### 2013-10-28 snippets/RMD_nRD.tex
snippets on RMD and helper packages

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{../helpers/myarticlestyledefault}

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

\title{Replicator-mutator and noise-perturbed replicator dynamics}
\author{}
\date{}

\begin{document}
\maketitle

\section{Preliminaries}

\paragraph{Notation.} Fix a signaling game with states $\States$,
messages $\Messgs$ and acts $\Acts$. % Let $\Util_{\sen,\rec} \mycolon
% \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ be the
% player's utility function.

Pure sender (receiver) strategies are functions $\Spure \in
\Messgs^\States$ ($\Rpure \in \Acts^\Messgs$). Mixed sender (receiver)
strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
($\Rmixed \in (\Acts^\Messgs)$). The latter give the relative
population frequencies of the former. We write $\Smixed_i$ for the
frequency $\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for
the receiver).

Let the sender's (receiver's) behavioral strategies be functions in
$\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
\Delta(\Acts)^\Messgs$). Every mixed strategy $\Smixed$ converts to a
unique behavioral strategy defined by:
\begin{align*}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align*} 
Let $G$ be this mapping from mixed to behavioral strategies. Notice
that $G$ is \emph{not} an injection, as many mixed strategies map onto
the same behavioral strategy.

For behavioral strategies, let $\EU(\messg, \state, \Rstrat)$ and
$\EU(\act, \messg, \Sstrat)$ be the players' expected utilities of
each action in each choice point as usual. For mixed strategies we
define the players' fitness in the usual manner. Let $F_i^{\Rmixed}$
be $\Spure_i$'s fitness given $\Rmixed$ and $F_i^{\Smixed}$ be
$\Rpure_i$'s fitness given $\Smixed$. Then $\Phi(\Smixed,\Rmixed) =
\sum_{k} \Smixed_k \cdot F_k^{\Rmixed}$ is the average fitness in the
sender population and $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k
\cdot F_k^{\Smixed}$ the average fitness in the receiver population.

\paragraph{Replicator dynamics in pure strategies.} The two-population
(non-payoff adjusted) continuous replicator dynamics in pure
strategies is defined as:
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j} \Smixed_j \cdot \left ( F_j^{\Rmixed} -
  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \sum_{j} \Rmixed_j \cdot \left ( F_j^{\Smixed} -
  \Phi(\Rmixed,\Smixed) \right ) \,.
\end{align*}
The discrete time version is given by: 
\begin{align*}
  \Smixed_i' & = \frac{\Smixed_i \cdot
  F_i^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} &     \Rmixed_i' & = \frac{\Rmixed_i \cdot
  F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
\end{align*}

\paragraph{Replicator-mutator equation.} Let $Q$ be a row-stochastic
\emph{mutation matrix} where $Q_{ji}$ gives the probability that pure
sender strategy $\Spure_j$ mutates into $\Spure_i$. Similarly, let $R$
be a row-stochastic \emph{mutation matrix} where $Q_{ji}$ gives the
probability that pure receiver strategy $\Rpure_j$ mutates into
$\Rpure_i$.

The two-population (non-payoff adjusted) continuous replicator-mutator
dynamics, is then given by::
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
\end{align*}
\begin{claim} A discrete time version of the above replicator mutator
  dynamics is:
  \begin{align*}
    \Smixed_i' & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
      F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \Rmixed_i' & = \sum_{j}
    R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
      \Phi(\Rmixed,\Smixed)}\,.
  \end{align*}
\end{claim}

\begin{proof}
  It suffices to check either sender or receiver part. Focusing on the
  former, we need to show that the continuous version is derivable
  from the discrete version if the discrete update steps get
  infinitely small so that:
  \begin{align*}
    \dot{\Smixed_i} & = \Smixed_i' - \Smixed_i = \sum_{j} Q_{ji}
    \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
    \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
        F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
    & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )
  \end{align*}
  As $\Phi(\Smixed,\Rmixed)$ is constant the rate of change we can
  drop it to derive the non-payoff adjusted continuous version above,
  since:
  \begin{align*}
    & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
      ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
        F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
        \Phi(\Smixed,\Rmixed) \\
       = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
        F_j^{\Rmixed} - \Smixed_i \cdot
        \Phi(\Smixed,\Rmixed) 
  \end{align*}
\end{proof}

The discrete time replicator-mutator dynamics have a nice sequential
update property: first we compute the fitness-driven change according
to the standard replicator dynamics; then we compute the perturbation
from mutation.

\section{Noise-perturbed replicator dynamics}

We look at signaling games with $\States = \Acts$ and fix a confusion
matrix $C$, which is a row-stochastic matrix whose elements $C_{ij}$
give the probability that $\state_i$ is perceived as
$\state_j$. Define player's average utility at each choice point given
the opponent's strategy as:
\begin{align*}
  \Phi(\state,\Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg,\state,\Rstrat) &
\Phi(\messg,\Sstrat) & = \sum_{\state} \Rstrat(\state \probbar \messg)
\cdot \EU(\state,\messg,\Sstrat)\,.
\end{align*}
The discrete noise-perturbed replicator dynamics on behavioral
strategies proposed by \citep{Correia2013:The-Bivalent-Tr} is:
\begin{align*}
  \Sstrat'(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
    \Rstrat'(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
\end{align*}
Notice that this update is also sequential: first we calculate an
update according to the standard replicator dynamics (in behavioral
strategies), then we compute the perturbation from perceptual noise.

\section{Exploring the relation}

We will show that the noise-perturbed replicator dynamics defined
above is the behavioral-strategy analogue to the replicator-mutator
dynamics when the only source of mutation is perceptual confusion of
states. Since, in general, the replicator dynamics in two-populations
is not equivalent in its formulations for behavioral and mixed
strategies,\todo{reference} we abstract from the dynamics and look
only at the effect of mutation and noise-perturbation. This is
justified given the above mentioned sequential nature of both
dynamics.

Let the mutation $Q(\Smixed)$ ($R(\Rmixed)$) of a mixed sender
(receiver) strategy be another mixed sender (receiver) strategy
defined by:
\begin{align*}
  Q(\Smixed)i & =  \sum_j  \Smixed_j \cdot
  Q_{ji} &   R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
  R_{ji} \,.
\end{align*}
We would like to compare this to the noise perturbation $C(\Sstrat)$
($C(\Rstrat)$) of behavioral strategy $\Sstrat$ ($\Rstrat)$, which is
another behavioral strategy given by:
\begin{align*}
  C(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
  \Sstrat(\messg \probbar \state_j) & C(\Rstrat)(\state_i \probbar
  \messg) & = \sum_{j} C_{ji} \cdot \Sstrat(\state_j \probbar \messg)
  \,.
\end{align*}

We hypothesize that a confusion matrix $C$ should give rise to a
unique mutation matrix $Q^C$ so that whenever $F(\Smixed) = \Sstrat$
we also have $F(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the
receiver.

\paragraph{Confusion-based mutations.} There are natural conversions
of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
we start with that.

The probability that $\Rpure$ mutates into $\Rpure'$ is the product
of the probabilities for each $\messg$ that $\Rpure(\messg)$ is
perceived as $\Rpure'(\messg)$. Abusing notation, by refer to the
index of $\Rpure(\messg)$ with $\Rpure(\messg)$ itself, we define:
\begin{align*}
  R^C_{ji} = \prod_{\messg} C_{\Rpure(\messg)\Rpure'(\messg)}\,.
\end{align*}

Now look at the sender. If $\Spure_j(\state_k)=\messg$ and
$\Spure_i(\state_k)=\messg'$, then the probability of confusion at
state $\state_k$ is given by the chance that $\state_k$ is perceived
as a state $\state_l$ which $\Spure_j$ would map onto $\messg'$. The
overal probability that $\Spure_j$ mutates into $\Sstrat_i$ is the
product of these chances for all states. So, define:
\begin{align*}
  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
\end{align*}

For example, consider a signaling game with two states and two
messages. Let the confusion matrix be:
\begin{align*}
  C=
  \begin{pmatrix}
    .8 & .2 \\
    .2 & .8 
  \end{pmatrix}
\end{align*}
The resulting mutation matrices are:
\begin{align*}
S^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & 1 & 0 & 0 & 0 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & 0 & 0 & 0 & 1 \cr}
&                    
  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & .64 & .16 & .16 & .04 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & .04 & .12 & .12 & .64 \cr}
\end{align*}
Here, a pair like $21$ refers to a pure strategy that plays message 2
and is states 1 and 2 respectively. Similarly for the receiver.

\medskip

\todo[inline]{necessary to prove that mutation matrices are row-stochastic}

\todo[inline]{maybe worthwhile to think about general properties of
  the mutation matrices that ensue from perceptual confusion?}

\medskip

\begin{theorem}
  \label{thm:sender-eq}
  If $G(\Smixed) = \Sstrat$, then $G(Q^C(\Smixed)) = C(\Sstrat)$.
\end{theorem}

\begin{proof}
  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
  the consequent:
  \begin{align*}
    C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
    \cdot \Sstrat(\messg_y \probbar \state_l) \\
    & =  \sum_{\state_l} C_{xl}
    \cdot  \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \\
    & = \sum_{\state_l}
    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
  \end{align*}
  Next consider the lhs of the consequent:
  \begin{align*}
    G(Q^C(\Smixed))(\messg_y \probbar \state_x) & =
    \sum_{\Spure_i(\state_x)=\messg_y} Q^C(\Smixed)(\Spure_i) \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed(\Spure_j) \cdot Q^C_{ji} \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed(\Spure_j) \cdot \prod_{\state_l} \sum_{\state_m \in
      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} \\
    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
  \end{align*}
  To simplify this further we look at a fixed $\Spure_j$ and consider
  the term: 
  \begin{align}
    \label{eq:term}
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
  \end{align}
  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
  over all pure strategies, we essentially look at each such
  mapping. Term (\ref{eq:term}) above quantifies over all pure
  strategies that map $\state_k$ onto $\messg_y$. The above term then
  sums over all products whose factors are tuples in $\times_{k>2}
  \set{y \setbar \exists l \mycolon y = Y_{kl}}$. So term
  (\ref{eq:term}) expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
  \begin{align*}
    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
    \dots \\
    & + (Y_{11} \cdot Y_{2d} \cdot
    Y_{3d} \cdot \ldots \cdot
    Y_{ed})\\
    = &    
  \end{align*}
  But since $Y$ is a row-stochastic matrix, this simplifies to
  $Y_{xy}$. Continuing the derivation with this:
  \begin{align*}
    G(Q^C(\Smixed))(\messg_y \probbar \state_x) 
    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed(\Spure_j) \cdot
     C_{xl} \\
     & = \sum_{\state_l}
    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
  \end{align*}

\end{proof}

\begin{theorem}
  If $G(\Rmixed) = \Rstrat$, then $G(R^C(\Rmixed)) = C(\Rstrat)$.
\end{theorem}

\todo[inline]{proof}

% \begin{proof}
%   The proof is almost entirely parallel to that of
%   Theorem~\ref{thm:sender-eq}. Fix $\Rmixed$ and $\Rstrat =
%   G(\Rmixed)$. The rhs of the consequent expands to:
%   \begin{align*}
%     C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{i}
%     \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}} \Rmixed_j \cdot C_{ix}\,.
%   \end{align*}
%   The rhs expands to:
%   \begin{align*}
%     G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{i}
%     \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
%         \state_x}} 
%     \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} \\
%     & = \sum_{i}
%     \Rmixed_i \cdot ( (C_{\Rpure_i(\messg_1)1} \cdot
%       C_{\Rpure_i(\messg_2)1} \cdot \ldots +
%       C_{\Rpure_i(\messg_d)1})  \\
%       & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
%       C_{\Rpure_i(\messg_2)1} \cdot \ldots \cdot
%       C_{\Rpure_i(\messg_d)2}) + \ldots  \\
%       & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
%       C_{\Rpure_i(\messg_2)e} \cdot \ldots \cdot
%       C_{\Rpure_i(\messg_d)e}) + \ldots ))
%   \end{align*}
%   If we assuming without loss of generality that $x=y=1$, then if $d =
%   \card{\Messgs}$, we see that:
%   \begin{align*}
%     \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}}
%     \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} = \sum_{i}
%     \Rmixed_i \cdot
%   \end{align*}
% \end{proof}

\printbibliography[heading=bibintoc]

\end{document}

### 2013-10-29 snippets/RMD_nRD.tex
RMD connection receiver part proof 1st half

@@
 \begin{align*}
-  \dot{\Smixed_i} & = \sum_{j} \Smixed_j \cdot \left ( F_j^{\Rmixed} -
-  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \sum_{j} \Rmixed_j \cdot \left ( F_j^{\Smixed} -
+  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
+  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
   \Phi(\Rmixed,\Smixed) \right ) \,.
@@
 \paragraph{Replicator-mutator equation.} Let $Q$ be a row-stochastic
-\emph{mutation matrix} where $Q_{ji}$ gives the probability that pure
-sender strategy $\Spure_j$ mutates into $\Spure_i$. Similarly, let $R$
-be a row-stochastic \emph{mutation matrix} where $Q_{ji}$ gives the
-probability that pure receiver strategy $\Rpure_j$ mutates into
-$\Rpure_i$.
+mutation matrix where $Q_{ji}$ gives the probability that pure sender
+strategy $\Spure_j$ mutates into $\Spure_i$. Similarly, let $R$ be a
+row-stochastic mutation matrix where $R_{ji}$ gives the probability
+that pure receiver strategy $\Rpure_j$ mutates into $\Rpure_i$.
 
@@
   \end{align*}
-  As $\Phi(\Smixed,\Rmixed)$ is constant the rate of change we can
-  drop it to derive the non-payoff adjusted continuous version above,
+  As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
+  it to derive the non-payoff adjusted continuous version above,
   since:
@@
 
-The discrete time replicator-mutator dynamics have a nice sequential
+The discrete time replicator-mutator dynamics has a nice sequential
 update property: first we compute the fitness-driven change according
@@
 give the probability that $\state_i$ is perceived as
-$\state_j$. Define player's average utility at each choice point given
+$\state_j$. Define players average utility at each choice point given
 the opponent's strategy as:
@@
 The discrete noise-perturbed replicator dynamics on behavioral
-strategies proposed by \citep{Correia2013:The-Bivalent-Tr} is:
+strategies proposed by \citet{Correia2013:The-Bivalent-Tr} is:
 \begin{align*}
@@
 is not equivalent in its formulations for behavioral and mixed
-strategies,\todo{reference} we abstract from the dynamics and look
-only at the effect of mutation and noise-perturbation. This is
-justified given the above mentioned sequential nature of both
-dynamics.
+strategies \citep{Cressman2003:Evolutionary-Dy}, we abstract from the
+dynamics and look only at the effect of mutation and
+noise-perturbation. This is justified given the above mentioned
+sequential nature of both dynamics.
 
@@
 \begin{align*}
-  Q(\Smixed)i & =  \sum_j  \Smixed_j \cdot
+  Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
   Q_{ji} &   R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
@@
   \Sstrat(\messg \probbar \state_j) & C(\Rstrat)(\state_i \probbar
-  \messg) & = \sum_{j} C_{ji} \cdot \Sstrat(\state_j \probbar \messg)
+  \messg) & = \sum_{j} C_{ji} \cdot \Rstrat(\state_j \probbar \messg)
   \,.
@@
 We hypothesize that a confusion matrix $C$ should give rise to a
-unique mutation matrix $Q^C$ so that whenever $F(\Smixed) = \Sstrat$
-we also have $F(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the
+unique mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$
+we also have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the
 receiver.
@@
 
-The probability that $\Rpure$ mutates into $\Rpure'$ is the product
-of the probabilities for each $\messg$ that $\Rpure(\messg)$ is
-perceived as $\Rpure'(\messg)$. Abusing notation, by refer to the
-index of $\Rpure(\messg)$ with $\Rpure(\messg)$ itself, we define:
+The probability that $\Rpure$ mutates into $\Rpure'$ is the product of
+the probabilities for each $\messg$ that $\Rpure(\messg)$ is perceived
+as $\Rpure'(\messg)$. Abusing notation, by refer to the index of
+$\Rpure(\messg)$ and $\Rpure'(\messg)$ with $\Rpure(\messg)$ and
+$\Rpure'(\messg)$ directly, we define:
 \begin{align*}
@@
     .2 & .8 
-  \end{pmatrix}
+  \end{pmatrix}\,.
 \end{align*}
@@
                       21 & .16 & .04 & .64 & .16 \cr
-                      22 & .04 & .12 & .12 & .64 \cr}
+                      22 & .04 & .12 & .12 & .64 \cr}\,.
 \end{align*}
-Here, a pair like $21$ refers to a pure strategy that plays message 2
-and is states 1 and 2 respectively. Similarly for the receiver.
+Here, a pair like $21$, for example, refers to a pure sender strategy
+with $\Spure(\state_1=\messg_2$ and $\Spure(\state_2) =
+\messg_1$. Similarly for the receiver.
 
@@
 
-\todo[inline]{necessary to prove that mutation matrices are row-stochastic}
+\todo[inline]{necessary to prove that mutation matrices are row-stochastic?}
 
@@
 
-\todo[inline]{proof}
-
-% \begin{proof}
-%   The proof is almost entirely parallel to that of
-%   Theorem~\ref{thm:sender-eq}. Fix $\Rmixed$ and $\Rstrat =
-%   G(\Rmixed)$. The rhs of the consequent expands to:
-%   \begin{align*}
-%     C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{i}
-%     \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}} \Rmixed_j \cdot C_{ix}\,.
-%   \end{align*}
-%   The rhs expands to:
-%   \begin{align*}
-%     G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{i}
-%     \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
-%         \state_x}} 
-%     \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} \\
-%     & = \sum_{i}
-%     \Rmixed_i \cdot ( (C_{\Rpure_i(\messg_1)1} \cdot
-%       C_{\Rpure_i(\messg_2)1} \cdot \ldots +
-%       C_{\Rpure_i(\messg_d)1})  \\
-%       & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
-%       C_{\Rpure_i(\messg_2)1} \cdot \ldots \cdot
-%       C_{\Rpure_i(\messg_d)2}) + \ldots  \\
-%       & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
-%       C_{\Rpure_i(\messg_2)e} \cdot \ldots \cdot
-%       C_{\Rpure_i(\messg_d)e}) + \ldots ))
-%   \end{align*}
-%   If we assuming without loss of generality that $x=y=1$, then if $d =
-%   \card{\Messgs}$, we see that:
-%   \begin{align*}
-%     \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}}
-%     \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} = \sum_{i}
-%     \Rmixed_i \cdot
-%   \end{align*}
-% \end{proof}
+\begin{proof}
+  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
+  expands to:
+  \begin{align*}
+    C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
+    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
+    \Rmixed_i \cdot C_{jx} \\
+    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
+    \Rmixed_i \cdot C_{jx} \\
+    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
+  \end{align*}
+  The rhs expands to:
+  \begin{align*}
+    G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{i}
+    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
+        \state_x}} 
+    \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} \\
+    & = \sum_{i}
+    \Rmixed_i \cdot ( (C_{\Rpure_i(\messg_1)1} \cdot
+      C_{\Rpure_i(\messg_2)1} \cdot \ldots +
+      C_{\Rpure_i(\messg_d)1})  \\
+      & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
+      C_{\Rpure_i(\messg_2)1} \cdot \ldots \cdot
+      C_{\Rpure_i(\messg_d)2}) + \ldots  \\
+      & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
+      C_{\Rpure_i(\messg_2)e} \cdot \ldots \cdot
+      C_{\Rpure_i(\messg_d)e}) + \ldots ))
+  \end{align*}
+  If we assuming without loss of generality that $x=y=1$, then if $d =
+  \card{\Messgs}$, we see that:
+  \begin{align*}
+    \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}}
+    \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} = \sum_{i}
+    \Rmixed_i \cdot
+  \end{align*}
+\end{proof}
 

### 2013-10-29 snippets/RMD_nRD.tex
RMD receiver proof

@@
   \begin{align*}
-    G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{i}
+    G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{\Rpure_i}
     \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
         \state_x}} 
-    \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} \\
+    \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}
+  \end{align*}
+  Without loss of generality, assume that $x=y=1$, and fix
+  $\card{M}=d$ and let $e$ be the number of pure receiver
+  strategies. Then the last term can be rewritten as:
+  \begin{align*}
     & = \sum_{i}
-    \Rmixed_i \cdot ( (C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)1} \cdot \ldots +
-      C_{\Rpure_i(\messg_d)1})  \\
-      & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)1} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)2}) + \ldots  \\
-      & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)e} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)e}) + \ldots ))
+    \Rmixed_i \cdot ( C_{\Rpure_i(\messg_1)1} \cdot
+      C_{\Rpure_i(\messg_2)\Rpure_1(\messg_2)} \cdot \ldots \cdot
+      C_{\Rpure_i(\messg_d)\Rpure_1(\messg_d)} + \ldots  \\
+      & \textcolor{white}{ = \sum_{i}
+    \Rmixed_i  (}  +  C_{\Rpure_i(\messg_1)1} \cdot
+      C_{\Rpure_i(\messg_2)\Rpure_2(\messg_2)} \cdot \ldots \cdot
+      C_{\Rpure_i(\messg_d)\Rpure_2(\messg_d)} + \ldots  \\
+      & \textcolor{white}{ = \sum_{i}
+    \Rmixed_i  (}  + C_{\Rpure_i(\messg_1)1} \cdot
+      C_{\Rpure_i(\messg_2)\Rpure_e(\messg_2)} \cdot \ldots \cdot
+      C_{\Rpure_i(\messg_d)\Rpure_e(\messg_d)}) )
   \end{align*}
-  If we assuming without loss of generality that $x=y=1$, then if $d =
-  \card{\Messgs}$, we see that:
+  Every for every messages $C_{\Rpure_i(\messg_l)}$ is a stochastic
+  vector. For $l>1$, all elements of these vectors appear equally
+  often. But that means that these cancel out. What remains is:
   \begin{align*}
-    \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}}
-    \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} = \sum_{i}
-    \Rmixed_i \cdot
+    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
   \end{align*}

### 2013-12-13 vague-voronoi-language.tex
Document describing conditions for a vague Voronoi language.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{todonotes}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\title{Vagueness games}
\author{José Pedro Correia \and Michael Franke}
\date{\today}

\begin{document}

\maketitle

\section{Vague Voronoi languages}

A \emph{vague Voronoi language} consists of a sender strategy $\sigma:T\rightarrow\Delta\left(M\right)$ and a receiver strategy $\rho:M\rightarrow \Delta\left(T\right)$ such that:

\begin{enumerate}
\item The receiver strategy assigns maximal probability to one state per message. The maximal element of a strategy for a message $m\in M$ can be understood as the prototype of that message:
\begin{equation}
\forall m\in M:\exists!t\in T:\forall t^{\prime}\in T:\rho\left(m,t\right)\geq\rho\left(m,t^{\prime}\right)
\end{equation}
Note that after we guarantee the uniqueness of message prototypes for a strategy $\rho$, we can define a prototype function $P_{\rho}:M\rightarrow T$
which associates to each message its prototype as $P_{\rho}\left(m\right)=\argmax_{t\in T}\rho\left(m,t\right)$

\item In the receiver strategy, for each message, the probability of a state being picked for that message increases with the state's similarity with the message's prototype:
\begin{equation}
\forall m\in M:\forall t,t^{\prime}\in T:s\left(t,P_{\rho}\left(m\right)\right)>s\left(t^{\prime},P_{\rho}\left(m\right)\right)\Rightarrow\rho\left(m,t\right)>\rho\left(m,t^{\prime}\right)
\end{equation}
\todo[inline]{Is this so far the same as defining $\rho\left(m\right)$ as a convez fuzzy set for every $m\in M$? Probably not...}

\item In the sender strategy, the more similar a state is to a message's prototype in comparison with other prototypes, the higher the probability that message will be used:
\begin{equation}
\forall t\in T:\forall m,m^{\prime}\in M:s\left(t,P_{\rho}\left(m\right)\right)>s\left(t,P_{\rho}\left(m^{\prime}\right)\right)\Rightarrow\sigma\left(t,m\right)>\sigma\left(t,m^{\prime}\right)
\end{equation}
\todo[inline]{A non-vague language Voronoi language should be a special case of this. We should show that and add a criterion of non-triviality, right?}
\end{enumerate}

\end{document}

### 2013-12-13 vagueness-games.py
Calculating "Voronoiness" for both speaker and hearer.

@@ -48,9 +48,10 @@ def plotStrategies(block=False):
     plt.plot(ExpectedUtilityHistory, label='$U(\\sigma,\\rho)$')
     plt.plot(EntropySpeakerHistory, label='$E(\\sigma)$', linestyle='--', color='green')
     plt.plot(ConvexitySpeakerHistory, label='$C(\\sigma)$', linestyle='-.', color='green')
+    plt.plot(VoronoinessSpeakerHistory, label='$V(\\sigma)$', linestyle=':', color='green')
     plt.plot(EntropyHearerHistory, label='$E(\\rho)$', linestyle='--', color='red')
 #    plt.plot(ConvexityHearerHistory, label='$C(\\rho)$', linestyle='-.', color='red')
-    plt.plot(PrototypicalityHearerHistory, label='$P(\\rho)$', linestyle='-.', color='red')
+    plt.plot(VoronoinessHearerHistory, label='$V(\\rho)$', linestyle=':', color='red')
     plt.ylim(ymin= -0.1, ymax=1.1)
     plt.legend(loc='best')
     plt.title('Measures')
@@ -99,17 +100,31 @@ def Convexity(Strategy):
     return sum(np.prod([ Strategy[c, s[c]] for c in xrange(Strategy.shape[0])])
                for s in ConvexPureStrategies)
     
-def Prototypicality(HearerStrategy, Similarity):
-    result = 0.0
+def Voronoiness(HearerStrategy, SpeakerStrategy, StateSimilarity):
+    Prototypes = [ np.argmax(HearerStrategy[m]) for m in xrange(HearerStrategy.shape[0]) ]
+
+    acc = 0.0
     for m in xrange(HearerStrategy.shape[0]):
-        p = np.argmax(HearerStrategy[m])
+        p = Prototypes[m]
         for t1 in xrange(HearerStrategy.shape[1]):
             for t2 in xrange(HearerStrategy.shape[1]):
-                if (Similarity[t1, p] > Similarity[t2, p] and HearerStrategy[m, t1] > HearerStrategy[m, t2]) \
-                   or (Similarity[t1, p] < Similarity[t2, p] and HearerStrategy[m, t1] < HearerStrategy[m, t2]):
-                    result += 1
-    return result / (HearerStrategy.shape[0] * HearerStrategy.shape[1] * (HearerStrategy.shape[1] - 1))
[…]

### 2013-12-13 run-batch.sh
Corrected headers.

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 DATE=$(date +%Y%m%d-%H%M)
-HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Expected utility,Iterations,Speaker strategy,Hearer strategy"
+HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Expected utility,Iterations,Speaker strategy,Hearer strategy"
 
 echo $HEADER > results-$DATE.csv
 for t in 6 10 14 18 22

### 2013-12-13 vagueness-games.py
Batch mode activated when there is at least one argument. Ran new batch of results.

@@ -142,7 +142,7 @@ OutputFile = sys.stdout
 
 # # Batch mode
  
-BatchMode = False
+BatchMode = len(sys.argv) > 1
 
 if BatchMode:
     if len(sys.argv) < 5:

### 2014-05-10 paper/paper.tex
Paper skeleton

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{myarticlestyledefault}

\title{Vagueness, Noise and Communication}
\author{Jos\'e Pedro Correia and Michael Franke}
\date{}

\begin{document}
\maketitle

\input{01_intro}

\input{02_noisyRMD}

\input{03_simulations}

\input{04_results}

\input{05_conclusions}

\printbibliography[heading=bibintoc]

\end{document}

### 2014-05-10 paper/02_noisyRMD.tex
paper skeleton first content

--- 2013-10-29 snippets/RMD_nRD.tex
@@
-\documentclass[fleqn,reqno,10pt]{article}
+\section{Noise-perturbed replicator dynamic}
 
-\usepackage{../helpers/myarticlestyledefault}
-
-\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
-\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}
-
-\title{Replicator-mutator and noise-perturbed replicator dynamics}
-\author{}
-\date{}
-
-\begin{document}
-\maketitle
-
-\section{Preliminaries}
+\subsection{Preliminaries}
 
@@
 
-\section{Noise-perturbed replicator dynamics}
+\subsection{Noise-perturbed replicator dynamics}
 
@@
 
-\section{Exploring the relation}
+\subsection{Exploring the relation}
 
@@
 
-\begin{proof}
-  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
-  the consequent:
-  \begin{align*}
-    C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
-    \cdot \Sstrat(\messg_y \probbar \state_l) \\
-    & =  \sum_{\state_l} C_{xl}
-    \cdot  \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \\
-    & = \sum_{\state_l}
-    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
-  \end{align*}
-  Next consider the lhs of the consequent:
-  \begin{align*}
-    G(Q^C(\Smixed))(\messg_y \probbar \state_x) & =
-    \sum_{\Spure_i(\state_x)=\messg_y} Q^C(\Smixed)(\Spure_i) \\
-    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
-    \Smixed(\Spure_j) \cdot Q^C_{ji} \\
-    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
-    \Smixed(\Spure_j) \cdot \prod_{\state_l} \sum_{\state_m \in
-      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} \\
-    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
-    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
-    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
-  \end{align*}
-  To simplify this further we look at a fixed $\Spure_j$ and consider
-  the term: 
-  \begin{align}
-    \label{eq:term}
-    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
-      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
-  \end{align}
-  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
-    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
-  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
-  over all pure strategies, we essentially look at each such
-  mapping. Term (\ref{eq:term}) above quantifies over all pure
-  strategies that map $\state_k$ onto $\messg_y$. The above term then
-  sums over all products whose factors are tuples in $\times_{k>2}
-  \set{y \setbar \exists l \mycolon y = Y_{kl}}$. So term
-  (\ref{eq:term}) expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
-  \begin{align*}
-    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
-    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
-    \dots \\
-    & + (Y_{11} \cdot Y_{2d} \cdot
-    Y_{3d} \cdot \ldots \cdot
-    Y_{ed})\\
-    = &    
-  \end{align*}
-  But since $Y$ is a row-stochastic matrix, this simplifies to
-  $Y_{xy}$. Continuing the derivation with this:
-  \begin{align*}
-    G(Q^C(\Smixed))(\messg_y \probbar \state_x) 
-    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
-    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
-    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed(\Spure_j) \cdot
-     C_{xl} \\
-     & = \sum_{\state_l}
-    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
-  \end{align*}
-
-\end{proof}
-
 \begin{theorem}
+  \label{thm:receiver-eq}
   If $G(\Rmixed) = \Rstrat$, then $G(R^C(\Rmixed)) = C(\Rstrat)$.
@@
 
-\begin{proof}
-  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
-  expands to:
-  \begin{align*}
-    C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
-    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
-    \Rmixed_i \cdot C_{jx} \\
-    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
-    \Rmixed_i \cdot C_{jx} \\
-    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
-  \end{align*}
-  The rhs expands to:
-  \begin{align*}
-    G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{i}
-    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
-        \state_x}} 
-    \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} \\
-    & = \sum_{i}
-    \Rmixed_i \cdot ( (C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)1} \cdot \ldots +
-      C_{\Rpure_i(\messg_d)1})  \\
-      & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)1} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)2}) + \ldots  \\
-      & + \quad (C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)e} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)e}) + \ldots ))
-  \end{align*}
-  If we assuming without loss of generality that $x=y=1$, then if $d =
-  \card{\Messgs}$, we see that:
-  \begin{align*}
-    \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) = \state_x}}
-    \prod_{\messg} C_{\Rstrat_i(\messg)\Rstrat_j(\messg)} = \sum_{i}
-    \Rmixed_i \cdot
-  \end{align*}
-\end{proof}
-
-\printbibliography[heading=bibintoc]
-
-\end{document}
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "paper"
+%%% TeX-PDF-mode: t
+%%% End:

### 2014-05-10 paper/03_simulations.tex
paper skeleton first content

--- 2014-05-10 paper/02_noisyRMD.tex
@@
-\section{Noise-perturbed replicator dynamic}
-
-\subsection{Preliminaries}
-
-\paragraph{Notation.} Fix a signaling game with states $\States$,
-messages $\Messgs$ and acts $\Acts$. % Let $\Util_{\sen,\rec} \mycolon
-% \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ be the
-% player's utility function.
-
-Pure sender (receiver) strategies are functions $\Spure \in
-\Messgs^\States$ ($\Rpure \in \Acts^\Messgs$). Mixed sender (receiver)
-strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
-($\Rmixed \in (\Acts^\Messgs)$). The latter give the relative
-population frequencies of the former. We write $\Smixed_i$ for the
-frequency $\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for
-the receiver).
-
-Let the sender's (receiver's) behavioral strategies be functions in
-$\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
-\Delta(\Acts)^\Messgs$). Every mixed strategy $\Smixed$ converts to a
-unique behavioral strategy defined by:
-\begin{align*}
-  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
-\end{align*} 
-Let $G$ be this mapping from mixed to behavioral strategies. Notice
-that $G$ is \emph{not} an injection, as many mixed strategies map onto
-the same behavioral strategy.
-
-For behavioral strategies, let $\EU(\messg, \state, \Rstrat)$ and
-$\EU(\act, \messg, \Sstrat)$ be the players' expected utilities of
-each action in each choice point as usual. For mixed strategies we
-define the players' fitness in the usual manner. Let $F_i^{\Rmixed}$
-be $\Spure_i$'s fitness given $\Rmixed$ and $F_i^{\Smixed}$ be
-$\Rpure_i$'s fitness given $\Smixed$. Then $\Phi(\Smixed,\Rmixed) =
-\sum_{k} \Smixed_k \cdot F_k^{\Rmixed}$ is the average fitness in the
-sender population and $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k
-\cdot F_k^{\Smixed}$ the average fitness in the receiver population.
-
-\paragraph{Replicator dynamics in pure strategies.} The two-population
-(non-payoff adjusted) continuous replicator dynamics in pure
-strategies is defined as:
-\begin{align*}
-  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
-  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
-  \Phi(\Rmixed,\Smixed) \right ) \,.
-\end{align*}
-The discrete time version is given by: 
-\begin{align*}
-  \Smixed_i' & = \frac{\Smixed_i \cdot
-  F_i^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} &     \Rmixed_i' & = \frac{\Rmixed_i \cdot
-  F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
-\end{align*}
-
-\paragraph{Replicator-mutator equation.} Let $Q$ be a row-stochastic
-mutation matrix where $Q_{ji}$ gives the probability that pure sender
-strategy $\Spure_j$ mutates into $\Spure_i$. Similarly, let $R$ be a
-row-stochastic mutation matrix where $R_{ji}$ gives the probability
-that pure receiver strategy $\Rpure_j$ mutates into $\Rpure_i$.
-
-The two-population (non-payoff adjusted) continuous replicator-mutator
-dynamics, is then given by::
-\begin{align*}
-  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
-    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
-    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
-    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
-\end{align*}
-\begin{claim} A discrete time version of the above replicator mutator
-  dynamics is:
-  \begin{align*}
-    \Smixed_i' & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
-      F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \Rmixed_i' & = \sum_{j}
-    R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
-      \Phi(\Rmixed,\Smixed)}\,.
-  \end{align*}
-\end{claim}
-
-\begin{proof}
-  It suffices to check either sender or receiver part. Focusing on the
-  former, we need to show that the continuous version is derivable
-  from the discrete version if the discrete update steps get
-  infinitely small so that:
-  \begin{align*}
-    \dot{\Smixed_i} & = \Smixed_i' - \Smixed_i = \sum_{j} Q_{ji}
-    \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
-    \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
-    & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )
-  \end{align*}
-  As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
-  it to derive the non-payoff adjusted continuous version above,
-  since:
-  \begin{align*}
-    & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
-      ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
-        F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
-        \Phi(\Smixed,\Rmixed) \\
-       = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot
-        \Phi(\Smixed,\Rmixed) 
-  \end{align*}
-\end{proof}
-
-The discrete time replicator-mutator dynamics has a nice sequential
-update property: first we compute the fitness-driven change according
-to the standard replicator dynamics; then we compute the perturbation
-from mutation.
-
-\subsection{Noise-perturbed replicator dynamics}
-
-We look at signaling games with $\States = \Acts$ and fix a confusion
-matrix $C$, which is a row-stochastic matrix whose elements $C_{ij}$
-give the probability that $\state_i$ is perceived as
-$\state_j$. Define players average utility at each choice point given
-the opponent's strategy as:
-\begin{align*}
-  \Phi(\state,\Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
-\EU(\messg,\state,\Rstrat) &
-\Phi(\messg,\Sstrat) & = \sum_{\state} \Rstrat(\state \probbar \messg)
-\cdot \EU(\state,\messg,\Sstrat)\,.
-\end{align*}
-The discrete noise-perturbed replicator dynamics on behavioral
-strategies proposed by \citet{Correia2013:The-Bivalent-Tr} is:
-\begin{align*}
-  \Sstrat'(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
-  \frac{\Sstrat(\messg \probbar \state_j) \cdot
-    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
-    \Rstrat'(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
-  \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
-\end{align*}
-Notice that this update is also sequential: first we calculate an
-update according to the standard replicator dynamics (in behavioral
-strategies), then we compute the perturbation from perceptual noise.
-
-\subsection{Exploring the relation}
-
-We will show that the noise-perturbed replicator dynamics defined
-above is the behavioral-strategy analogue to the replicator-mutator
-dynamics when the only source of mutation is perceptual confusion of
-states. Since, in general, the replicator dynamics in two-populations
-is not equivalent in its formulations for behavioral and mixed
-strategies \citep{Cressman2003:Evolutionary-Dy}, we abstract from the
-dynamics and look only at the effect of mutation and
-noise-perturbation. This is justified given the above mentioned
-sequential nature of both dynamics.
-
-Let the mutation $Q(\Smixed)$ ($R(\Rmixed)$) of a mixed sender
-(receiver) strategy be another mixed sender (receiver) strategy
-defined by:
-\begin{align*}
-  Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
-  Q_{ji} &   R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
-  R_{ji} \,.
-\end{align*}
-We would like to compare this to the noise perturbation $C(\Sstrat)$
-($C(\Rstrat)$) of behavioral strategy $\Sstrat$ ($\Rstrat)$, which is
-another behavioral strategy given by:
-\begin{align*}
-  C(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
-  \Sstrat(\messg \probbar \state_j) & C(\Rstrat)(\state_i \probbar
-  \messg) & = \sum_{j} C_{ji} \cdot \Rstrat(\state_j \probbar \messg)
-  \,.
-\end{align*}
-
-We hypothesize that a confusion matrix $C$ should give rise to a
-unique mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$
-we also have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the
-receiver.
-
-\paragraph{Confusion-based mutations.} There are natural conversions
-of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
-we start with that.
-
-The probability that $\Rpure$ mutates into $\Rpure'$ is the product of
-the probabilities for each $\messg$ that $\Rpure(\messg)$ is perceived
-as $\Rpure'(\messg)$. Abusing notation, by refer to the index of
-$\Rpure(\messg)$ and $\Rpure'(\messg)$ with $\Rpure(\messg)$ and
-$\Rpure'(\messg)$ directly, we define:
-\begin{align*}
-  R^C_{ji} = \prod_{\messg} C_{\Rpure(\messg)\Rpure'(\messg)}\,.
-\end{align*}
-
-Now look at the sender. If $\Spure_j(\state_k)=\messg$ and
-$\Spure_i(\state_k)=\messg'$, then the probability of confusion at
-state $\state_k$ is given by the chance that $\state_k$ is perceived
-as a state $\state_l$ which $\Spure_j$ would map onto $\messg'$. The
-overal probability that $\Spure_j$ mutates into $\Sstrat_i$ is the
-product of these chances for all states. So, define:
-\begin{align*}
-  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
-    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
-\end{align*}
-
-For example, consider a signaling game with two states and two
-messages. Let the confusion matrix be:
-\begin{align*}
-  C=
-  \begin{pmatrix}
-    .8 & .2 \\
-    .2 & .8 
-  \end{pmatrix}\,.
-\end{align*}
-The resulting mutation matrices are:
-\begin{align*}
-S^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
-                      11 & 1 & 0 & 0 & 0 \cr
-                      12 & .16 & .64 & .04 & .16 \cr
-                      21 & .16 & .04 & .64 & .16 \cr
-                      22 & 0 & 0 & 0 & 1 \cr}
-&                    
-  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
-                      11 & .64 & .16 & .16 & .04 \cr
-                      12 & .16 & .64 & .04 & .16 \cr
-                      21 & .16 & .04 & .64 & .16 \cr
-                      22 & .04 & .12 & .12 & .64 \cr}\,.
-\end{align*}
-Here, a pair like $21$, for example, refers to a pure sender strategy
-with $\Spure(\state_1=\messg_2$ and $\Spure(\state_2) =
-\messg_1$. Similarly for the receiver.
-
-\medskip
-
-\todo[inline]{necessary to prove that mutation matrices are row-stochastic?}
-
-\todo[inline]{maybe worthwhile to think about general properties of
-  the mutation matrices that ensue from perceptual confusion?}
-
-\medskip
-
-\begin{theorem}
-  \label{thm:sender-eq}
-  If $G(\Smixed) = \Sstrat$, then $G(Q^C(\Smixed)) = C(\Sstrat)$.
-\end{theorem}
-
-\begin{theorem}
-  \label{thm:receiver-eq}
-  If $G(\Rmixed) = \Rstrat$, then $G(R^C(\Rmixed)) = C(\Rstrat)$.
-\end{theorem}
+\section{Simulations}
 

### 2014-05-11 paper/01_intro.tex

\section{Introduction}

\begin{itemize}
\item introduce sim-max games
\item problem of vagueness \citep{Lipman2009:Why-is-Language}
\item \dots
\end{itemize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-PDF-mode: t
%%% End:

### 2014-05-11 paper/06_appendix.tex

--- 2013-10-29 snippets/RMD_nRD.tex
@@
-\documentclass[fleqn,reqno,10pt]{article}
+\appendix
 
-\usepackage{../helpers/myarticlestyledefault}
+\section{Proofs}
 
-\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
-\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}
-
-\title{Replicator-mutator and noise-perturbed replicator dynamics}
-\author{}
-\date{}
-
-\begin{document}
-\maketitle
-
-\section{Preliminaries}
-
-\paragraph{Notation.} Fix a signaling game with states $\States$,
-messages $\Messgs$ and acts $\Acts$. % Let $\Util_{\sen,\rec} \mycolon
-% \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ be the
-% player's utility function.
-
-Pure sender (receiver) strategies are functions $\Spure \in
-\Messgs^\States$ ($\Rpure \in \Acts^\Messgs$). Mixed sender (receiver)
-strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
-($\Rmixed \in (\Acts^\Messgs)$). The latter give the relative
-population frequencies of the former. We write $\Smixed_i$ for the
-frequency $\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for
-the receiver).
-
-Let the sender's (receiver's) behavioral strategies be functions in
-$\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
-\Delta(\Acts)^\Messgs$). Every mixed strategy $\Smixed$ converts to a
-unique behavioral strategy defined by:
-\begin{align*}
-  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
-\end{align*} 
-Let $G$ be this mapping from mixed to behavioral strategies. Notice
-that $G$ is \emph{not} an injection, as many mixed strategies map onto
-the same behavioral strategy.
-
-For behavioral strategies, let $\EU(\messg, \state, \Rstrat)$ and
-$\EU(\act, \messg, \Sstrat)$ be the players' expected utilities of
-each action in each choice point as usual. For mixed strategies we
-define the players' fitness in the usual manner. Let $F_i^{\Rmixed}$
-be $\Spure_i$'s fitness given $\Rmixed$ and $F_i^{\Smixed}$ be
-$\Rpure_i$'s fitness given $\Smixed$. Then $\Phi(\Smixed,\Rmixed) =
-\sum_{k} \Smixed_k \cdot F_k^{\Rmixed}$ is the average fitness in the
-sender population and $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k
-\cdot F_k^{\Smixed}$ the average fitness in the receiver population.
-
-\paragraph{Replicator dynamics in pure strategies.} The two-population
-(non-payoff adjusted) continuous replicator dynamics in pure
-strategies is defined as:
-\begin{align*}
-  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
-  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
-  \Phi(\Rmixed,\Smixed) \right ) \,.
-\end{align*}
-The discrete time version is given by: 
-\begin{align*}
-  \Smixed_i' & = \frac{\Smixed_i \cdot
-  F_i^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} &     \Rmixed_i' & = \frac{\Rmixed_i \cdot
-  F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
-\end{align*}
-
-\paragraph{Replicator-mutator equation.} Let $Q$ be a row-stochastic
-mutation matrix where $Q_{ji}$ gives the probability that pure sender
-strategy $\Spure_j$ mutates into $\Spure_i$. Similarly, let $R$ be a
-row-stochastic mutation matrix where $R_{ji}$ gives the probability
-that pure receiver strategy $\Rpure_j$ mutates into $\Rpure_i$.
-
-The two-population (non-payoff adjusted) continuous replicator-mutator
-dynamics, is then given by::
-\begin{align*}
-  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
-    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
-    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
-    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
-\end{align*}
-\begin{claim} A discrete time version of the above replicator mutator
-  dynamics is:
-  \begin{align*}
-    \Smixed_i' & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
-      F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \Rmixed_i' & = \sum_{j}
-    R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
-      \Phi(\Rmixed,\Smixed)}\,.
-  \end{align*}
-\end{claim}
-
-\begin{proof}
-  It suffices to check either sender or receiver part. Focusing on the
-  former, we need to show that the continuous version is derivable
-  from the discrete version if the discrete update steps get
-  infinitely small so that:
-  \begin{align*}
-    \dot{\Smixed_i} & = \Smixed_i' - \Smixed_i = \sum_{j} Q_{ji}
-    \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
-    \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
-    & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )
-  \end{align*}
-  As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
-  it to derive the non-payoff adjusted continuous version above,
-  since:
-  \begin{align*}
-    & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
-      ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
-        F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
-        \Phi(\Smixed,\Rmixed) \\
-       = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot
-        \Phi(\Smixed,\Rmixed) 
-  \end{align*}
-\end{proof}
-
-The discrete time replicator-mutator dynamics has a nice sequential
-update property: first we compute the fitness-driven change according
-to the standard replicator dynamics; then we compute the perturbation
-from mutation.
-
-\section{Noise-perturbed replicator dynamics}
-
-We look at signaling games with $\States = \Acts$ and fix a confusion
-matrix $C$, which is a row-stochastic matrix whose elements $C_{ij}$
-give the probability that $\state_i$ is perceived as
-$\state_j$. Define players average utility at each choice point given
-the opponent's strategy as:
-\begin{align*}
-  \Phi(\state,\Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
-\EU(\messg,\state,\Rstrat) &
-\Phi(\messg,\Sstrat) & = \sum_{\state} \Rstrat(\state \probbar \messg)
-\cdot \EU(\state,\messg,\Sstrat)\,.
-\end{align*}
-The discrete noise-perturbed replicator dynamics on behavioral
-strategies proposed by \citet{Correia2013:The-Bivalent-Tr} is:
-\begin{align*}
-  \Sstrat'(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
-  \frac{\Sstrat(\messg \probbar \state_j) \cdot
-    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
-    \Rstrat'(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
-  \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
-\end{align*}
-Notice that this update is also sequential: first we calculate an
-update according to the standard replicator dynamics (in behavioral
-strategies), then we compute the perturbation from perceptual noise.
-
-\section{Exploring the relation}
-
-We will show that the noise-perturbed replicator dynamics defined
-above is the behavioral-strategy analogue to the replicator-mutator
-dynamics when the only source of mutation is perceptual confusion of
-states. Since, in general, the replicator dynamics in two-populations
-is not equivalent in its formulations for behavioral and mixed
-strategies \citep{Cressman2003:Evolutionary-Dy}, we abstract from the
-dynamics and look only at the effect of mutation and
-noise-perturbation. This is justified given the above mentioned
-sequential nature of both dynamics.
-
-Let the mutation $Q(\Smixed)$ ($R(\Rmixed)$) of a mixed sender
-(receiver) strategy be another mixed sender (receiver) strategy
-defined by:
-\begin{align*}
-  Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
-  Q_{ji} &   R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
-  R_{ji} \,.
-\end{align*}
-We would like to compare this to the noise perturbation $C(\Sstrat)$
-($C(\Rstrat)$) of behavioral strategy $\Sstrat$ ($\Rstrat)$, which is
-another behavioral strategy given by:
-\begin{align*}
-  C(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
-  \Sstrat(\messg \probbar \state_j) & C(\Rstrat)(\state_i \probbar
-  \messg) & = \sum_{j} C_{ji} \cdot \Rstrat(\state_j \probbar \messg)
-  \,.
-\end{align*}
-
-We hypothesize that a confusion matrix $C$ should give rise to a
-unique mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$
-we also have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the
-receiver.
-
-\paragraph{Confusion-based mutations.} There are natural conversions
-of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
-we start with that.
-
-The probability that $\Rpure$ mutates into $\Rpure'$ is the product of
-the probabilities for each $\messg$ that $\Rpure(\messg)$ is perceived
-as $\Rpure'(\messg)$. Abusing notation, by refer to the index of
-$\Rpure(\messg)$ and $\Rpure'(\messg)$ with $\Rpure(\messg)$ and
-$\Rpure'(\messg)$ directly, we define:
-\begin{align*}
-  R^C_{ji} = \prod_{\messg} C_{\Rpure(\messg)\Rpure'(\messg)}\,.
-\end{align*}
-
-Now look at the sender. If $\Spure_j(\state_k)=\messg$ and
-$\Spure_i(\state_k)=\messg'$, then the probability of confusion at
-state $\state_k$ is given by the chance that $\state_k$ is perceived
-as a state $\state_l$ which $\Spure_j$ would map onto $\messg'$. The
-overal probability that $\Spure_j$ mutates into $\Sstrat_i$ is the
-product of these chances for all states. So, define:
-\begin{align*}
-  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
-    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
-\end{align*}
-
-For example, consider a signaling game with two states and two
-messages. Let the confusion matrix be:
-\begin{align*}
-  C=
-  \begin{pmatrix}
-    .8 & .2 \\
-    .2 & .8 
-  \end{pmatrix}\,.
-\end{align*}
-The resulting mutation matrices are:
-\begin{align*}
-S^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
-                      11 & 1 & 0 & 0 & 0 \cr
-                      12 & .16 & .64 & .04 & .16 \cr
-                      21 & .16 & .04 & .64 & .16 \cr
-                      22 & 0 & 0 & 0 & 1 \cr}
-&                    
-  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
-                      11 & .64 & .16 & .16 & .04 \cr
-                      12 & .16 & .64 & .04 & .16 \cr
-                      21 & .16 & .04 & .64 & .16 \cr
-                      22 & .04 & .12 & .12 & .64 \cr}\,.
-\end{align*}
-Here, a pair like $21$, for example, refers to a pure sender strategy
-with $\Spure(\state_1=\messg_2$ and $\Spure(\state_2) =
-\messg_1$. Similarly for the receiver.
-
-\medskip
-
-\todo[inline]{necessary to prove that mutation matrices are row-stochastic?}
-
-\todo[inline]{maybe worthwhile to think about general properties of
-  the mutation matrices that ensue from perceptual confusion?}
-
-\medskip
-
-\begin{theorem}
-  \label{thm:sender-eq}
-  If $G(\Smixed) = \Sstrat$, then $G(Q^C(\Smixed)) = C(\Sstrat)$.
-\end{theorem}
-
-\begin{proof}
+\begin{proof}[Proof of Theorem~\ref{thm:sender-eq}]
   Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
@@
 
-\begin{theorem}
-  If $G(\Rmixed) = \Rstrat$, then $G(R^C(\Rmixed)) = C(\Rstrat)$.
-\end{theorem}
-
-\begin{proof}
+\begin{proof}[Proof of \ref{thm:receiver-eq} ]
   Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
@@
 
-\printbibliography[heading=bibintoc]
-
-\end{document}
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "paper"
+%%% TeX-PDF-mode: t
+%%% End:

### 2014-05-20 data-analysis.R
Small fix.

@@ -39,6 +39,6 @@ plot.cluster <- function(language.names, data) {
     for (i in 1:length(language.names)) {
         plots[[i]] <- plot.language(data[language.names[i],'Speaker.strategy.file'], data[language.names[i],'Hearer.strategy.file'])
     }
-    do.call(arrangeGrob, plots)
+    do.call(grid.arrange, plots)
     
 }

### 2014-05-20 paper/03_simulations.tex
Added description of first metric: entropy.

@@
 \section{Simulations}
+
+In order to study the long-term behavior of the noise-perturbed sim-max games, we conducted computer simulations for different combinations of initial parameters.
+Namely, we were interested in investigating the impact of different values of impairment and its relation with the size of the state space.
+\mytodo{JPC}{Justify why size of state space is interesting}
+
+\subsection{Measures}
+To be able to quantitively analyze the results of our simulations, we decided to define a number of measures to capture different characteristics of sim-max languages.
+
+\paragraph{Entropy.}
+This classic information-theoretic measure captures the amount of uncertainty in a signaling strategy.
+The most natural way is to define it in terms of mixed strategies rather than behavioral strategies.
+Let us recalle that mixed sender (receiver) strategies are functions $\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).
+We can define the entropy of a sender strategy (henceforth $\Sstrat$-entropy) as:
+\begin{align*}
+  E(\Smixed) = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot \ln(\Smixed(\Spure)) \,.
+\end{align*} 
+and the entropy of a receiver strategy (henceforth $\Rstrat$-entropy) as:
+\begin{align*}
+  E(\Rmixed) = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \ln(\Rmixed(\Rpure)) \,.
+\end{align*} 
+These measures are computationally expensive to calculate, since the size of the domain over which the sum is computed grows exponentially with the number of choice points.
+Therefore, we converted\footnote{See Appendix~\ref{sec:proofs}.} the above definitions into equivalent measures defined in terms of behavioral strategies.
+These are as follows:
+\begin{align*}
+  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs} \Sstrat(\messg \probbar \state) \cdot \ln(\Sstrat(\messg \probbar \state)) \,.
+\end{align*} 
+\begin{align*}
+  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\act \probbar \messg) \cdot \ln(\Rstrat(\act \probbar \messg)) \,.
+\end{align*}
+The measures are lower bounded by $0$ and upper bounded by, respectively, $\ln(|\Messgs^\States|) = |\States| \cdot ln(|\Messgs|)$ and $\ln(|\Acts^\Messgs|) = |\Messgs| \cdot ln(|\Acts|)$, thus we can normalize them by dividing by these values.
 

### 2014-08-07 paper/03_simulations.tex
Added notions of informativity and Voronoiness.

--- 2014-05-10 paper/03_simulations.tex
@@
 \section{Simulations}
+
+In order to study the long-term behavior of the noise-perturbed sim-max games, we conducted computer simulations for different combinations of initial parameters.
+Namely, we were interested in investigating the impact of different values of impairment and its relation with the size of the state space.
+\mytodo{JPC}{Justify why size of state space is interesting}
+
+\subsection{Measures}
+To be able to quantitively analyze the results of our simulations, we define a number of measures to capture different characteristics of sim-max languages.
+
+\paragraph{Entropy.}
+This classic information-theoretic measure captures the amount of uncertainty in a signaling strategy.
+The most natural way to define it is in terms of mixed strategies rather than behavioral strategies.
+Let us recall that mixed sender (receiver) strategies are functions $\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).
+We can define the entropy of a sender strategy as
+\begin{align*}
+  E(\Smixed) = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot \log(\Smixed(\Spure))
+\end{align*} 
+and the entropy of a receiver strategy as
+\begin{align*}
+  E(\Rmixed) = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
+\end{align*} 
+These measures are computationally expensive to calculate, since the size of the domain over which the sum is computed grows exponentially with the number of choice points.
+Therefore, we converted\footnote{See Appendix~\ref{sec:proofs}.} the above definitions into equivalent measures defined in terms of behavioral strategies.
+These are as follows:
+\begin{align*}
+  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs} \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar \state))
+\end{align*} 
+\begin{align*}
+  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar \messg)) \,.
+\end{align*}
+The measures are lower bounded by $0$ and upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States| \cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot \log(|\Acts|)$, thus we can normalize them by dividing by these values.
+
+\paragraph{Informativity.}
+\mytodo{JPC}{Informativity is perhaps a bad name for the receiver...}
+The quantity of information in a signal is an old notion that also goes back to the start of information theory.
+Skyrms~\citet[ch.~3]{Skyrms2010} discusses its use in the context of signaling games.
+The main idea is that we can quantify the amount of information about a state $\state$ in a message $\messg$ via the relation between the probability of the state given the message $\Pr(\state \,|\, \messg)$ and the unconditional probability of the state $\Pr(\state)$.
+Following Bayes' theorem, we can express $\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg \,|\, \state)}{\Pr(\messg)}$.
+We then have $\Pr(\messg \,|\, \state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) = \sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)$.
+Finally, we equate sender infomativity with the average overall information about states in each signal.
+Based on the definition by Skyrms~\citet[p.~36]{Skyrms2010}, and the considerations above, we define sender informativity as follows:
+\begin{align*}
+  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\state \in \States} \frac{\Pr(\state) \cdot \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right) \,.
+\end{align*}
+Conversely, we can quantify the amount of information about an act in a message.
+We equate receiver informativity with the average overall information about acts in each signal.
+Based on Skyrms' definition~\citet[p.~39]{Skyrms2010}, we define receiver informativity as follows:
+\begin{align*}
+  I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
+\end{align*}
+Both measures are bounded between $0$ and $1$.
+
+\paragraph{Voronoiness.}
+This measure aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
+Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p.
+562).
+\mytodo{JPC}{Explain further?}
+Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive measure to characterize that in terms of degrees.
+With that in mind, and given $\Acts = \States$, we define Voronoiness as follows.
+Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
+\begin{multline*}
+  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state, \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then } 1 \textrm{ else } 0
+\end{multline*}
+is the Voronoiness of the sender strategy $\Sstrat$ given the receiver strategy $\Rstrat$ and
+\begin{align*}
+  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States} \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat, \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg, \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{ else } 0
+\end{align*}
+is the Voronoiness of the receiver strategy $\Rstrat$.
+\mytodo{JPC}{Give some intuitions on these?}
+The measures are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.
 

### 2014-08-09 paper/03_simulations.tex
Description of methodology for experiments.

@@
 Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive measure to characterize that in terms of degrees.
-With that in mind, and given $\Acts = \States$, we define Voronoiness as follows.
+With that in mind, and given that this measure only makes sense for $\Acts = \States$, we define Voronoiness as follows.
 Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
@@
 
+\subsection{Methodology}
+\newcommand{\impairment}{\alpha}
+In our simulations we focus on a concrete setup in line with \cite{Correia2013}.
+Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
+As similarity measure $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
+\begin{align*}
+  s(\state,\state^\prime) =
+    \begin{cases}
+    1 & \textrm{if } \impairment = 0 \textrm{ and } \state = \state^\prime \\
+    0 & \textrm{if } \impairment = 0 \textrm{ and } \state \neq \state^\prime \\
+    e^{-\frac{\lVert \state - \state^\prime \rVert ^2}{\impairment^2}} & \textrm{otherwise} \\
+    \end{cases}
+\end{align*}
+where $\impairment$ is an \emph{impairment} factor.
+Here we deviate from \cite{Correia2013} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
+Confusion is directly related with similarity, \emph{i.e.}~the more similar two states are, the more likely an agent is to perceive one as the other.
+For simplicity, we equate one with the other by defining $C_{ij} = s(\state_i,\state_j)$.
+
+We run the noise-perturbed replicator dynamics with random initial strategies $\Sstrat_0$ and $\Rstrat_0$ until a convergence criterion is met, namely that the total absolute change in both sender and receiver strategies does not amount to more than $0.01$, \emph{i.e.}:
+\begin{align*}
+  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left| \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| < 0.01
+\end{align*}
+and
+\begin{align*}
+  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.01 \,.
+\end{align*}
+
+We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
+For each parameter combination, we perform $100$ trials.
+
 %%% Local Variables: 

### 2014-08-09 data-analysis.R
Generated new plots, included them in the results section.

@@ -2,7 +2,7 @@ require(ggplot2)
 require(reshape)
 require(gridExtra)
 
-theme_set(theme_minimal())
+theme_set(theme_bw())
 
 plot.language <- function(speaker.filename, hearer.filename) {
 
@@ -42,3 +42,21 @@ plot.cluster <- function(language.names, data) {
     do.call(grid.arrange, plots)
     
 }
+
+x <- read.csv('results/20140121-141158.csv', colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))
+ggplot(x, aes(x=factor(Impairment), y=Speaker.entropy)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(E(sigma)))
+ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-entropy-', '20140121-141158.png'))
+ggplot(x, aes(x=factor(Impairment), y=Hearer.entropy)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(E(rho)))
+ggsave(filename=paste(sep='', 'paper/plots/', 'Hearer-entropy-', '20140121-141158.png'))
+ggplot(x, aes(x=factor(Impairment), y=Speaker.informativity)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(I(sigma)))
+ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-informativity-', '20140121-141158.png'))
+ggplot(x, aes(x=factor(Impairment), y=Hearer.informativity)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(I(rho)))
+ggsave(filename=paste(sep='', 'paper/plots/', 'Hearer-informativity-', '20140121-141158.png'))
+ggplot(x, aes(x=factor(Impairment), y=Speaker.Voronoiness)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(V(sigma)))
+ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-Voronoiness-', '20140121-141158.png'))
[…]

### 2014-08-11 data-analysis.R
Moved saving of plots to its own function to avoid rerunning that on every load.

@@ -29,7 +29,7 @@ plot.language <- function(speaker.filename, hearer.filename) {
         geom_line() + labs(title='Hearer',x='State',y='') + 
         theme(legend.position = 'none')
     
-    arrangeGrob(speaker.plot, hearer.plot)
+    return(arrangeGrob(speaker.plot, hearer.plot))
     
 }
 
@@ -43,20 +43,24 @@ plot.cluster <- function(language.names, data) {
     
 }
 
-x <- read.csv('results/20140121-141158.csv', colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))
-ggplot(x, aes(x=factor(Impairment), y=Speaker.entropy)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(E(sigma)))
-ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-entropy-', '20140121-141158.png'))
-ggplot(x, aes(x=factor(Impairment), y=Hearer.entropy)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(E(rho)))
-ggsave(filename=paste(sep='', 'paper/plots/', 'Hearer-entropy-', '20140121-141158.png'))
-ggplot(x, aes(x=factor(Impairment), y=Speaker.informativity)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(I(sigma)))
-ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-informativity-', '20140121-141158.png'))
-ggplot(x, aes(x=factor(Impairment), y=Hearer.informativity)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(I(rho)))
-ggsave(filename=paste(sep='', 'paper/plots/', 'Hearer-informativity-', '20140121-141158.png'))
-ggplot(x, aes(x=factor(Impairment), y=Speaker.Voronoiness)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(V(sigma)))
-ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-Voronoiness-', '20140121-141158.png'))
[…]

### 2014-08-11 paper/03_simulations.tex
Generated new plots, included them in the results section.
Added short description of results. Renamed results section to discussion.

@@
 \subsection{Measures}
+\label{sec:measures}
 To be able to quantitively analyze the results of our simulations, we define a number of measures to capture different characteristics of sim-max languages.
@@
 We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
-For each parameter combination, we perform $100$ trials.
+For each parameter combination, we perform $100$ trials and record the values of each of the measures introduced in Section~\ref{sec:measures} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
+Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.
+
+\subsection{Results}
+In order to visualize the impact of the impairment factor on the outcome of simulations, in Figure~\ref{fig:metrics-boxplots} we plot the distribution of each of the measurements per impairment value.
+%
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140121-141158}
+                \caption{Entropy of sender strategy.}
+%                \label{fig:gull}
+        \end{subfigure}
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140121-141158}
+                \caption{Entropy of receiver strategy.}
+        \end{subfigure}
+
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140121-141158}
+                \caption{Informativity of sender strategy.}
+        \end{subfigure}
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140121-141158}
+                \caption{Informativity of receiver strategy.}
+        \end{subfigure}
+
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140121-141158}
+                \caption{Voronoiness of sender strategy.}
+        \end{subfigure}
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140121-141158}
+                \caption{Voronoiness of receiver strategy.}
+        \end{subfigure}
+
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Expected-utility-20140121-141158}
+                \caption{Expected utility.}
+        \end{subfigure}
+        \begin{subfigure}{0.45\textwidth}
+                \includegraphics[width=\textwidth]{plots/Iterations-20140121-141158}
+                \caption{Number of iterations until convergence.}
+        \end{subfigure}
+        \caption{Boxplots of metrics per impairment value.}
+        \label{fig:metrics-boxplots}
+\end{figure}
+%
+Most trends are quite clear.
+Entropy seems to increase with higher impairment values for both sender and receiver.
+This is an expected outcome, since the higher the impairment value, the more likely each player is to confuse two states, thus the more uncertain he will be in unequivocally associating a state with a message, or vice-versa.
+The trend for informativity is also expected: more vague strategies convey less information, be it from the sender or the receiver perspective.
+
+Regarding Voronoiness, we observe that most simulation results approximate perfect Voronoi tessalations.
+This is not the case, however, for the lower values of impairment.
+We believe the explanation is that, with lower values of impairment, strategies can achieve a high expected utility without having perfect Voronoiness.
+Although the expected long term outcome in these situations is perfect Voronoiness (given J\"ager's analytical results~\cite{Jager2007}), actually reaching it involves minor refinements which at some point do not qualify as significant change given our convergence criterion.
+
+The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009}: more impairment necessarily leads to lower expected utility.
+Finally, the plot for number of iterations until convergence shows that a certain level of impairment enables faster convergence.
+We will discuss this further in more detail in Section~\ref{sec:discussion}.
+
+\mytodo{JPC}{Present and discuss results per size of state space? Don't forget O'Connor's claim that speed of convergence is more important for larger \#state/\#message ratio}
 

### 2014-08-11 paper/04_discussion.tex
Added short description of results. Renamed results section to discussion.

--- 2014-05-11 paper/06_appendix.tex
@@
-\appendix
-
-\section{Proofs}
-
-\begin{proof}[Proof of Theorem~\ref{thm:sender-eq}]
-  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
-  the consequent:
-  \begin{align*}
-    C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
-    \cdot \Sstrat(\messg_y \probbar \state_l) \\
-    & =  \sum_{\state_l} C_{xl}
-    \cdot  \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \\
-    & = \sum_{\state_l}
-    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
-  \end{align*}
-  Next consider the lhs of the consequent:
-  \begin{align*}
-    G(Q^C(\Smixed))(\messg_y \probbar \state_x) & =
-    \sum_{\Spure_i(\state_x)=\messg_y} Q^C(\Smixed)(\Spure_i) \\
-    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
-    \Smixed(\Spure_j) \cdot Q^C_{ji} \\
-    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
-    \Smixed(\Spure_j) \cdot \prod_{\state_l} \sum_{\state_m \in
-      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} \\
-    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
-    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
-    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
-  \end{align*}
-  To simplify this further we look at a fixed $\Spure_j$ and consider
-  the term: 
-  \begin{align}
-    \label{eq:term}
-    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
-      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
-  \end{align}
-  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
-    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
-  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
-  over all pure strategies, we essentially look at each such
-  mapping. Term (\ref{eq:term}) above quantifies over all pure
-  strategies that map $\state_k$ onto $\messg_y$. The above term then
-  sums over all products whose factors are tuples in $\times_{k>2}
-  \set{y \setbar \exists l \mycolon y = Y_{kl}}$. So term
-  (\ref{eq:term}) expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
-  \begin{align*}
-    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
-    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
-    \dots \\
-    & + (Y_{11} \cdot Y_{2d} \cdot
-    Y_{3d} \cdot \ldots \cdot
-    Y_{ed})\\
-    = &    
-  \end{align*}
-  But since $Y$ is a row-stochastic matrix, this simplifies to
-  $Y_{xy}$. Continuing the derivation with this:
-  \begin{align*}
-    G(Q^C(\Smixed))(\messg_y \probbar \state_x) 
-    & = \sum_{\Spure_j} \Smixed(\Spure_j) \cdot
-    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
-    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed(\Spure_j) \cdot
-     C_{xl} \\
-     & = \sum_{\state_l}
-    \sum_{\Spure(\state_l) = \messg_y} \Smixed(\Spure) \cdot C_{xl}\,.
-  \end{align*}
-
-\end{proof}
-
-\begin{proof}[Proof of \ref{thm:receiver-eq} ]
-  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
-  expands to:
-  \begin{align*}
-    C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
-    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
-    \Rmixed_i \cdot C_{jx} \\
-    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
-    \Rmixed_i \cdot C_{jx} \\
-    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
-  \end{align*}
-  The rhs expands to:
-  \begin{align*}
-    G(R^C(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{\Rpure_i}
-    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
-        \state_x}} 
-    \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}
-  \end{align*}
-  Without loss of generality, assume that $x=y=1$, and fix
-  $\card{M}=d$ and let $e$ be the number of pure receiver
-  strategies. Then the last term can be rewritten as:
-  \begin{align*}
-    & = \sum_{i}
-    \Rmixed_i \cdot ( C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)\Rpure_1(\messg_2)} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)\Rpure_1(\messg_d)} + \ldots  \\
-      & \textcolor{white}{ = \sum_{i}
-    \Rmixed_i  (}  +  C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)\Rpure_2(\messg_2)} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)\Rpure_2(\messg_d)} + \ldots  \\
-      & \textcolor{white}{ = \sum_{i}
-    \Rmixed_i  (}  + C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)\Rpure_e(\messg_2)} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)\Rpure_e(\messg_d)}) )
-  \end{align*}
-  Every for every messages $C_{\Rpure_i(\messg_l)}$ is a stochastic
-  vector. For $l>1$, all elements of these vectors appear equally
-  often. But that means that these cancel out. What remains is:
-  \begin{align*}
-    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
-  \end{align*}
-\end{proof}
+\section{Discussion}
+\label{sec:discussion}
 

### 2014-08-11 paper/paper.tex
paper skeleton first content
Added description of first metric: entropy.
Generated new plots, included them in the results section.
Added short description of results. Renamed results section to discussion.

@@
 
-\usepackage{myarticlestyledefault}
+%========================================
+% Packages
+%========================================
 
-\title{Vagueness, Noise and Communication}
+\usepackage[]{../helpers/mypackages}
+%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
+\bibliography{../helpers/MyRefGlobal}
+\usepackage{../helpers/myenvironments}
+\usepackage{../helpers/mycommands}
+\usepackage{todonotes}
+\usepackage{subcaption}
+
+%========================================
+% Standard Layout
+%========================================
+
+% Itemize
+\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
+\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
+\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
+\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
+% Description
+\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}
+
+% Figure Captions
+\usepackage{caption} % use corresponding myfiguresize!
+\setlength{\captionmargin}{20pt}
+\renewcommand{\captionfont}{\small}
+\setlength{\belowcaptionskip}{7pt} % standard is 0pt
+
+%========================================
+% Additional layout & commands
+%========================================
+
+\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
+\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}
+
+% Annotations
+\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
+\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
+\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}
+
+\title{Vagueness, Noise, and Communication}
 \author{Jos\'e Pedro Correia and Michael Franke}
@@
 \maketitle
+
+\begin{abstract}
+\mytodo{JPC}{This start needs to be adjusted depending on the audience}
+Signaling games have been productively used as models for studying natural language meaning.
+One important aspect that is, however, typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
+Recently, some models have been proposed to explicitly address this limitation.
+In this paper we explore one such models.
+Namely, we establish a link between the dynamics in the model and the replicator-mutator equations.
+Furthermore, we study the long term behavior of the model using computer simulations.
+\mytodo{JPC}{Kind of a big claim below\ldots}
+Our results indicate that a possible explanation of why vagueness is pervasive in natural language is that vague languages converge faster than non-vague ones, thus providing an evolutionary advantage in an changing environment.
+\end{abstract}
 
@@
 
-\input{04_results}
+\input{04_discussion}
 
 \input{05_conclusions}
+
+\input{06_appendix}
 

### 2014-08-22 data-analysis.R
Plotting languages in B&W. Function to process results is parameterizable.

@@ -22,11 +22,11 @@ plot.language <- function(speaker.filename, hearer.filename) {
     melted.speaker$State <- states
     melted.hearer$State <- states
     
-    speaker.plot <- ggplot(melted.speaker, aes(x=State, y=value, colour=variable)) +
-        geom_line() + labs(title='Speaker',x='State',y='') +
+    speaker.plot <- ggplot(melted.speaker, aes(x=State, y=value)) +
+        geom_line(aes(linetype=variable)) + labs(title='Speaker',x='State',y='') +
         theme(legend.position = 'none') + ylim(0,1)
-    hearer.plot <- ggplot(melted.hearer, aes(x=State, y=value, colour=variable)) +
-        geom_line() + labs(title='Hearer',x='State',y='') + 
+    hearer.plot <- ggplot(melted.hearer, aes(x=State, y=value)) +
+        geom_line(aes(linetype=variable)) + labs(title='Hearer',x='State',y='') + 
         theme(legend.position = 'none')
     
     return(arrangeGrob(speaker.plot, hearer.plot))
@@ -43,24 +43,24 @@ plot.cluster <- function(language.names, data) {
     
 }
 
-do.stuff <- function() {
-    x <- read.csv('results/20140121-141158.csv', colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))
+do.stuff <- function(timestamp) {
+    x <<- read.csv(paste(sep='','results/', timestamp,'.csv'), colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))
     width <- 8
     height <- 5
     ggplot(x, aes(x=factor(Impairment), y=Speaker.entropy)) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(E(sigma)))
-    ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-entropy-', '20140121-141158.png'), width=width, height=height)
+    ggsave(filename=paste(sep='', 'paper/plots/', 'Speaker-entropy-', timestamp, '.png'), width=width, height=height)
[…]

### 2014-08-22 run-batch.sh
Generated new set of results (100 per parameter combination)

@@ -4,16 +4,17 @@ DATE=$(date +%Y%m%d-%H%M%S)
 HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Speaker informativity,Hearer informativity,Expected utility,Iterations,Speaker strategy file,Hearer strategy file"
 
 echo $HEADER > results/$DATE.csv
-for t in 6 10 50 90
+for t in 10 50 90 130
 do
     echo $t
-    for i in 0 0.05 0.1 0.15 0.2 0.25 0.3 0.35 0.4
+    for i in 0 0.1 0.2 0.3 0.4 0.5
     do
         echo " $i"
-    	for n in {1..10}
+    	for n in {1..100}
     	do
-	        echo "  $n"
+	        echo -n "  $n"
         	python vagueness-games.py $t 2 $i results/$DATE.csv
     	done
+    	echo
     done
 done

### 2014-08-22 paper/03_simulations.tex
Using metrics not measures. Plotting new results.

@@
 
-\subsection{Measures}
-\label{sec:measures}
-To be able to quantitively analyze the results of our simulations, we define a number of measures to capture different characteristics of sim-max languages.
+\subsection{Metrics}
+\label{sec:metrics}
+To be able to quantitively analyze the results of our simulations, we define a number of metrics to capture different characteristics of sim-max languages.
 
 \paragraph{Entropy.}
-This classic information-theoretic measure captures the amount of uncertainty in a signaling strategy.
+This classic information-theoretic metric captures the amount of uncertainty in a signaling strategy.
 The most natural way to define it is in terms of mixed strategies rather than behavioral strategies.
@@
 \end{align*} 
-These measures are computationally expensive to calculate, since the size of the domain over which the sum is computed grows exponentially with the number of choice points.
-Therefore, we converted\footnote{See Appendix~\ref{sec:proofs}.} the above definitions into equivalent measures defined in terms of behavioral strategies.
+These metrics are computationally expensive to calculate, since the size of the domain over which the sum is computed grows exponentially with the number of choice points.
+Therefore, we converted\footnote{See Appendix~\ref{sec:proofs}.} the above definitions into equivalent metrics defined in terms of behavioral strategies.
 These are as follows:
@@
 \end{align*}
-The measures are lower bounded by $0$ and upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States| \cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot \log(|\Acts|)$, thus we can normalize them by dividing by these values.
+The metrics are lower bounded by $0$ and upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States| \cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot \log(|\Acts|)$, thus we can normalize them by dividing by these values.
 
@@
 \end{align*}
-Both measures are bounded between $0$ and $1$.
+Both metrics are bounded between $0$ and $1$.
 
 \paragraph{Voronoiness.}
-This measure aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
+This metric aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
 Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p.
@@
 \mytodo{JPC}{Explain further?}
-Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive measure to characterize that in terms of degrees.
-With that in mind, and given that this measure only makes sense for $\Acts = \States$, we define Voronoiness as follows.
+Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive metric to characterize that in terms of degrees.
+With that in mind, and given that this metric only makes sense for $\Acts = \States$, we define Voronoiness as follows.
 Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
@@
 \mytodo{JPC}{Give some intuitions on these?}
-The measures are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.
+The metrics are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.
 
@@
 Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
-As similarity measure $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
+As similarity metric $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
 \begin{align*}
@@
 We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
-For each parameter combination, we perform $100$ trials and record the values of each of the measures introduced in Section~\ref{sec:measures} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
+For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
 Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140813-194306}
                 \caption{Entropy of sender strategy.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140813-194306}
                 \caption{Entropy of receiver strategy.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140813-194306}
                 \caption{Informativity of sender strategy.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140813-194306}
                 \caption{Informativity of receiver strategy.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140813-194306}
                 \caption{Voronoiness of sender strategy.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140813-194306}
                 \caption{Voronoiness of receiver strategy.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Expected-utility-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Expected-utility-20140813-194306}
                 \caption{Expected utility.}
@@
         \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Iterations-20140121-141158}
+                \includegraphics[width=\textwidth]{plots/Iterations-20140813-194306}
                 \caption{Number of iterations until convergence.}

### 2014-08-22 paper/04_discussion.tex
Wrote part of discussion section on interpretation of the model.

@@
 \label{sec:discussion}
+
+\subsection{Model interpretation}
+The replicator-mutator equations are a mathematical model motivated by concepts from the theory of natural selection, namely natural variation, differential reproduction, and mutation.
+The most straightforward interpretation is thus a biological one: we can imagine signaling strategies as innate behaviorial tendencies of organisms, steps in the evolutionary process as successive generations, and selection as capturing the reproductive advantage of fitter individuals.
+In this interpretation, our proposed model is not an extremely interesting one, since the noise matrix simply corresponds to a restricted form of mutation.
+
+Another possibility is to take the equations as embodying cultural evolution.
+In this interpretation, what is subject to the evolutionary forces are not organisms but strategies themselves.
+This pressuposes that individuals can adopt strategies from other individuals.
+Fitness captures the success of strategies, which in the case of language can be thought of in terms of communicative success.
+Differential reproduction represents the tendency of more successful strategies to be more likely adopted by individuals, be it through imitation or some learning procedure within the population.
+Under this interpretation, the proposed noise perturbation becomes more interesting.
+Messages correspond to the overt data available to individuals, whereas states are interpreted as the more private aspect of the behavior.
+Futhermore, similarity-maximization games are especially interesting in situations where the number of states largely exceeds the number of messages available.
+When trying to adopt another individual's behavior, it is thus more likely that information about states is both less accessible and less practical to elicit, hence justifying a higher likelihood of mistakes and a more pressing need for a certain degree of generalization%
+\footnote{The consequence of perturbing a strategy with a noise matrix based on similarity can be seen as ``spreading'' the effects of the fitness of the behavior for a state to its more similar states, which can also be interpreted as generalization}%
+.
+
+Finally, the replicator dynamics are also related to some forms of reinforcement learning (\cite{Borgers1997,Hopkins2005,Beggs2005}), and can thus be interpreted in cognitive terms.
+
+\mytodo{JPC}{Discuss nature of noise}
+\mytodo{JPC}{Compare with O'Connor}
+\mytodo{JPC}{Compare with Franke2013}
 

### 2014-08-22 paper/paper.tex
New plots. Added separate bibliography.
Reshaped abstract.

@@
 \bibliography{../helpers/MyRefGlobal}
+\bibliography{paper}
 \usepackage{../helpers/myenvironments}
@@
 \begin{abstract}
-\mytodo{JPC}{This start needs to be adjusted depending on the audience}
-Signaling games have been productively used as models for studying natural language meaning.
-One important aspect that is, however, typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
+Signaling games have been garnering a lot of attention as models for studying natural language meaning.
+One important aspect that is nevertheless typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
 Recently, some models have been proposed to explicitly address this limitation.
-In this paper we explore one such models.
-Namely, we establish a link between the dynamics in the model and the replicator-mutator equations.
+In this paper we explore one such model and establish a link between the dynamics in the model and the replicator-mutator equations.
 Furthermore, we study the long term behavior of the model using computer simulations.
-\mytodo{JPC}{Kind of a big claim below\ldots}
-Our results indicate that a possible explanation of why vagueness is pervasive in natural language is that vague languages converge faster than non-vague ones, thus providing an evolutionary advantage in an changing environment.
+The results indicate that vague languages, despite having lower expected utility and informativity, are faster to converge to an equilibrium than non-vague ones.
+We argue that this could be an importante advantage that contributes to the pervasiveness of vagueness in natural language.
 \end{abstract}

### 2014-08-22 paper/03_simulations.tex
Added more comments on the results. Fixed two citations.

@@
 \subsection{Results}
-In order to visualize the impact of the impairment factor on the outcome of simulations, in Figure~\ref{fig:metrics-boxplots} we plot the distribution of each of the measurements per impairment value.
+In order to visualize the impact of the impairment factor on the outcome of simulations, we plot in Figure~\ref{fig:metrics-boxplots} the distribution of each of the measurements per impairment value.
 %
@@
 Regarding Voronoiness, we observe that most simulation results approximate perfect Voronoi tessalations.
-This is not the case, however, for the lower values of impairment.
-We believe the explanation is that, with lower values of impairment, strategies can achieve a high expected utility without having perfect Voronoiness.
+This is not the case, however, for impairment of $0$ (both sender and receiver) and $0.5$ (mostly sender only).
+\mytodo{JPC}{For impairment 0, the languages with the lowest sender voronoiness look like non-convex equilibria. Discuss in relation with Elliott's results?}
+For impairment of $0$,  the explanation is usually that strategies can achieve a high expected utility without having perfect Voronoiness.
 Although the expected long term outcome in these situations is perfect Voronoiness (given J\"ager's analytical results~\cite{Jager2007}), actually reaching it involves minor refinements which at some point do not qualify as significant change given our convergence criterion.
+The result is thus an artifact of stopping the simulations before they reach perfect convergence.
+For impairment of $0.5$ (and some cases with impairment of $0.4$), the reason is simply that the languages are representatives of something between babbling and pooling equilibria.
+These are equilibria where $\Sstrat(\state,\messg_1) \geq \Sstrat(\state,\messg_2)$ for all $\state \in \States$, but vary from situations where $\Sstrat(\state,\messg)$ is close to $0.5$ for both messages (babbling), or is significantly higher for one of them (pooling).
+\mytodo{JPC}{Plot examples}
 
-The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009}: more impairment necessarily leads to lower expected utility.
+The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009:Why-is-Language}: more impairment necessarily leads to lower expected utility.
 Finally, the plot for number of iterations until convergence shows that a certain level of impairment enables faster convergence.

### 2014-08-22 data-analysis.R

@@ -23,10 +23,10 @@ plot.language <- function(speaker.filename, hearer.filename) {
     melted.hearer$State <- states
     
     speaker.plot <- ggplot(melted.speaker, aes(x=State, y=value)) +
-        geom_line(aes(linetype=variable)) + labs(title='Speaker',x='State',y='') +
+        geom_line(aes(linetype=variable)) + labs(title='Sender',x='State',y='') +
         theme(legend.position = 'none') + ylim(0,1)
     hearer.plot <- ggplot(melted.hearer, aes(x=State, y=value)) +
-        geom_line(aes(linetype=variable)) + labs(title='Hearer',x='State',y='') + 
+        geom_line(aes(linetype=variable)) + labs(title='Receiver',x='State',y='') + 
         theme(legend.position = 'none')
     
     return(arrangeGrob(speaker.plot, hearer.plot))

### 2014-08-26 data_exploration_tmp.Rmd
data exploration MF

Data Exploration
================

load and pre-process the data

``` {r}
data = read.csv('results/20140813-194306.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))

ds = data[,c(1,4,5,7:16)] # only relevant data

attach(ds)
```

plot all the relevant variables against each other

``` {r warning = FALSE, fig.width=12, fig.height=12}

pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
        Speaker.convexity + Hearer.convexity + Speaker.Voronoiness + Hearer.Voronoiness +
        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
```

some properties are very clearly positively correlated:

* speaker informativity, hearer informativity & expected utility
* impairment & speaker entropy

these two "clusters" are also strongly negatively correlated

``` {r warning = FALSE, fig.width=7, fig.height=7}

pairs(~ Impairment + Speaker.entropy +
        Speaker.informativity + Hearer.informativity + Expected.utility, data=ds)
```

The benefit of impairment (some thoughts)
-------------------------

we want to argue that (small?) levels of impairment can be helpful (i.e., boost EU); but unfortunately, impairment and EU are strongly negatively correlated; so we should look for stages before final convergence

impairment does seem to speed up learning! that's different from Cailin's stuff, because there learning speed could be argued to be only due to more balls being added per learning step

also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment

### 2014-08-26 data_exploration_tmp.Rmd
data exploration MF

@@
 ``` {r}
+
+require('ellipse')
+require('ggplot2')
+
 data = read.csv('results/20140813-194306.csv', 
@@
 pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
-        Speaker.convexity + Hearer.convexity + Speaker.Voronoiness + Hearer.Voronoiness +
+        Speaker.convexity + Hearer.convexity +
         Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
@@
 
-pairs(~ Impairment + Speaker.entropy +
-        Speaker.informativity + Hearer.informativity + Expected.utility, data=ds)
+dss = ds[,c(2,4,10,11,12)]
+
+ctab <- cor(dss)
+
+plotcorr(ctab)
+
+# pairs(~ Impairment + Speaker.entropy +
+#         Speaker.informativity + Hearer.informativity + Expected.utility, data=dss)
 ```
@@
 
-also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment
+also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment, but we left this notion out [why what that again?]
+
+Faster learning by impairment
+-----------------------------
+
+if learning is faster for some impairment, we should see these effects also in terms of higher EU at earlier stages
+
+- plot EU at different stages in the nRD
+- show that early on slight impairment is beneficial to no impairment
+
+Next up
+-------
+- evaluate impact of number of states
+- replot Fig 1 with seperated numbers of states
+
+``` {r}
+
+x = data
+
+ggplot(x, aes(x=factor(Impairment), y=Expected.utility, color=factor(Number.of.states))) + geom_boxplot() + xlab('Impairment') + ylim(0, 1) + ylab(expression(EU))
+
+```

### 2014-08-26 paper/02_noisyRMD.tex
data exploration MF

@@
 
-Let the sender's (receiver's) behavioral strategies be functions in
+Let the sender's (receiver's) behavioral strategies be functions
 $\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
@@
 give the probability that $\state_i$ is perceived as
-$\state_j$. Define players average utility at each choice point given
+$\state_j$. Define the players' average utility at each choice point given
 the opponent's strategy as:

### 2014-08-26 paper/03_simulations.tex
data exploration MF

@@
 
-We perform a parameter sweep with size of the state space $n \in \{10, 50, 100, 200\}$ and impairment $\impairment \in \{0, 0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4\}$.
+We perform a parameter sweep with size of the state space $n \in \{10,
+50, 90, 130\}$ and impairment $\impairment \in \{0, 0.1, 0.2, 0.3,
+0.4, 0.5\}$.
 For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.

### 2014-08-27 data_exploration_tmp.Rmd
further data exploration, new plots by no. of states

--- 2014-08-26 data_exploration_tmp.Rmd
@@
 
-``` {r}
+``` {r warnings=FALSE, message=FALSE}
+
+require('ellipse')
+require('ggplot2')
+require('reshape2')
+source('~//Desktop//data//svn//programming//R//helpers//helpers.R')
+
 data = read.csv('results/20140813-194306.csv', 
-                colClasses=c('numeric','factor','numeric','numeric','numeric','factor','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','numeric','character','character'))
+                colClasses=c('numeric','factor','numeric','numeric','numeric',
+                             'factor','numeric','numeric','numeric','numeric',
+                             'numeric','numeric','numeric','numeric','numeric',
+                             'numeric','character','character'))
 
-ds = data[,c(1,4,5,7:16)] # only relevant data
+ds = data[,c(1,4,7,8,11:16)] # only relevant data
+dm = melt(ds,id.vars = c("Number.of.states", "Impairment")) # melted data
 
@@
 pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
-        Speaker.convexity + Hearer.convexity + Speaker.Voronoiness + Hearer.Voronoiness +
+        Speaker.Voronoiness + Hearer.Voronoiness +
         Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
@@
 
-pairs(~ Impairment + Speaker.entropy +
-        Speaker.informativity + Hearer.informativity + Expected.utility, data=ds)
+dss = ds[,c(2,3,7,8,9)]
+
+ctab <- cor(dss)
+
+plotcorr(ctab)
+
 ```
 
-The benefit of impairment (some thoughts)
--------------------------
+Replotting data for each combination of values of independent variables
+-----------------------------------------------------------------------
 
-we want to argue that (small?) levels of impairment can be helpful (i.e., boost EU); but unfortunately, impairment and EU are strongly negatively correlated; so we should look for stages before final convergence
+We have two independent variables: impairment & number of states.
+The following plots show the means of all 100 data points for each combination of values for independent variables, together with the estimated confidence intervals.
 
-impairment does seem to speed up learning! that's different from Cailin's stuff, because there learning speed could be argued to be only due to more balls being added per learning step
+We can exclude the informativity notions, because they are highly correlated with expected utility scores, and add no extra insight.
 
-also speaker and hearer voronoiness seems to be much higher for at least some bit of impairment
+```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+
+dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "variable"), measurevar="value")
+dsub = subset(dms, !grepl("informativity",dms$variable) )
+
+pd <- position_dodge(.05)
+sp = ggplot(dsub, aes(x=Impairment, y=value, colour=factor(Number.of.states))) + 
+     geom_point(position = pd) +
+     geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
+     geom_line(position=pd)
+
+sp + facet_wrap(~ variable, ncol=2, scales="free")
+
+```
+
+We are particularly interested in the interplay between learning speed and EU.
+Generally, the fewer states, the faster we reach convergence (unsurprising).
+But speed to convergence also depends on the impairment value.
+For values 0.2, 0.3 and 0.5 we see a greatly reduced time to convergence.
+However, it is also clear that this increase in speed comes with a decline in expected utility.
+[Notice that the confidence intervals are very tight here, so that we can be sure that there is an almost functional realtionship between convergence speed and EU for each pair of independent parameter values.]
+
+Faster learning by impairment
+-----------------------------
+
+we want to argue that (small?) levels of impairment can be helpful
+
+unfortunately, impairment and EU are strongly negatively correlated;
+
+however, impairment does seem to speed up learning;
+
+- that's different from Cailin's stuff, because there learning speed could be argued to be only due to more balls being added per learning step
+
+if learning is faster for some impairment, we should see these effects also in terms of higher EU at earlier stages
+- so we should look for stages before final convergence

### 2014-08-27 data_exploration_round_based.Rmd
round based data exploration

Data Exploration (round-based data)
===================================

load and pre-process the data

``` {r warnings=FALSE, message=FALSE}

require('ellipse')
require('ggplot2')
require('reshape2')
source('~//Desktop//data//svn//programming//R//helpers//helpers.R')

data = read.csv('results/20140827-121341.round_based.csv')

ds = data[,c(1,4,7:12)] # only relevant data
dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Iterations")) # melted data

attach(ds)
```

Plot EU by round
----------------

We plot EU at every fifth round, depending on number of states and impairments.

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Iterations", "variable"), measurevar="value")
dsub = subset(dms, dms$variable=="Expected.utility")

pd <- position_dodge(.05)
sp = ggplot(dsub, aes(x=Iterations, y=value, colour=factor(Impairment))) + 
     geom_point(position = pd) +
     geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
     geom_line(position=pd)

sp + facet_wrap(~ Number.of.states, ncol=2, scales="free")

```

Unfortunately, we see that lower impairment always trumps higher impairment.

### 2014-08-27 run-batch_round_based.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+DATE=$(date +%Y%m%d-%H%M%S)
+HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Expected utility,Iterations"
+
+echo $HEADER > results/$DATE.round_based.csv
+for t in 10 50 90 130
+do
+    echo $t
+    for i in 0 0.1 0.2 0.3 0.4 0.5
+    do
+        echo " $i"
+    	for n in {1..20}
+    	do
+	        echo -n "  $n"
+        	python vagueness-games_round_based.py $t 2 $i results/$DATE.round_based.csv
+    	done
+    	echo
+    done
+done

### 2014-08-27 vagueness-games_round_based.py

@@ -1,0 +1,211 @@
+import sys
+import copy
+import csv
+import math
+
+import numpy as np
+from numpy import random as random
+
+from scipy import stats as stats
+
+import matplotlib.pyplot as plt
+import itertools
+
+from datetime import datetime
+
+def normalize(Vector):
+    return Vector / np.max(Vector)
+
+def makePDF(Vector):
+    if np.sum(Vector) == 0:
+        Vector = np.ones(np.shape(Vector))
+    return Vector / np.sum(Vector)
+
+def makePDFPerRow(Matrix):
+    return np.array([ makePDF(Row) for Row in Matrix ])
+    
+def ExpectedUtility(Speaker, Hearer, Utility):
+    return np.sum(Speaker[t1, m] * Hearer[m, t2] * Utility[t1, t2]
+                  for t1 in xrange(Speaker.shape[0])
+                  for m in xrange(Speaker.shape[1])
+                  for t2 in xrange(Hearer.shape[1]))
+
+def NormalizedEntropy(Strategy):
+    return -sum(np.log(Strategy[c, a]) * Strategy[c, a] if Strategy[c, a] != 0 else 0
+                 for c in xrange(Strategy.shape[0]) for a in xrange(Strategy.shape[1])) / (Strategy.shape[0] * np.log(Strategy.shape[1]))
+
+def convex_sequences(options, length, current=None, acc=[]):
+    if length == 0:
+        return [acc]
+    else:
+        result = []
+        for x in options:
+            new_acc = list(acc)
+            new_acc.append(x)
+            new_options = list(options)
+            if x != current and current in new_options:
+                new_options.remove(current)
+            for seq in convex_sequences(new_options, length - 1, x, new_acc):
+                result.append(seq)
+        return result
+
+def Convexity(Strategy):
+    ConvexPureStrategies = convex_sequences(range(Strategy.shape[1]), Strategy.shape[0])
+    return sum(np.prod([ Strategy[c, s[c]] for c in xrange(Strategy.shape[0])])
+               for s in ConvexPureStrategies)
+    
+def Voronoiness(HearerStrategy, SpeakerStrategy, StateSimilarity):
[…]

### 2014-08-27 data_exploration_round_based.Rmd
data exploration round based

@@
 
-data = read.csv('results/20140827-121341.round_based.csv')
+data = read.csv('results/20140827-124230.round_based.csv')
 
-ds = data[,c(1,4,7:12)] # only relevant data
+ds = data[,c(1,4,7:8)] # only relevant data
 dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Iterations")) # melted data

### 2014-08-27 run-batch_round_based.sh
exploring inefficient equilibria with low state numbers

@@ -4,13 +4,13 @@ DATE=$(date +%Y%m%d-%H%M%S)
 HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Expected utility,Iterations"
 
 echo $HEADER > results/$DATE.round_based.csv
-for t in 10 50 90 130
+for t in 3 4 5 6 7
 do
     echo $t
-    for i in 0 0.1 0.2 0.3 0.4 0.5
+    for i in 0 0.025 0.05 0.075 0.1
     do
         echo " $i"
-    	for n in {1..20}
+    	for n in {1..50}
     	do
 	        echo -n "  $n"
         	python vagueness-games_round_based.py $t 2 $i results/$DATE.round_based.csv

### 2014-08-27 vagueness-games_round_based.py

@@ -171,7 +171,7 @@ else:
 OptimalExpectedUtility = ExpectedUtility(SpeakerOptimal, HearerOptimal, Utility)
 
 
-for i in xrange(101):
+for i in xrange(11):
 
     # # Dynamics
     
@@ -205,7 +205,7 @@ for i in xrange(101):
 
     Hearer = makePDFPerRow(Hearer)
 
-    if i % 5 == 0:
+    if i % 1 == 0:
 
         csv.writer(OutputFile).writerow([NStates, PriorDistributionType, NMessages, Impairment, Tolerance, Dynamics, \
             ExpectedUtility(Speaker, Hearer, Utility) / OptimalExpectedUtility, i])

### 2014-08-28 vagueness-games_round_based.py
sec com

@@ -171,7 +171,7 @@ else:
 OptimalExpectedUtility = ExpectedUtility(SpeakerOptimal, HearerOptimal, Utility)
 
 
-for i in xrange(11):
+for i in xrange(101):
 
     # # Dynamics
     
@@ -205,7 +205,7 @@ for i in xrange(11):
 
     Hearer = makePDFPerRow(Hearer)
 
-    if i % 1 == 0:
+    if i % 5 == 0:
 
         csv.writer(OutputFile).writerow([NStates, PriorDistributionType, NMessages, Impairment, Tolerance, Dynamics, \
             ExpectedUtility(Speaker, Hearer, Utility) / OptimalExpectedUtility, i])

### 2014-09-02 snippets/square_games.Rmd
exploring square games

``` {r warning=FALSE, message=FALSE, echo=FALSE}

require('gtools') # for dirichlet distribution
require('ggplot2')
require('reshape2')
require('fpc') # for function pamk(.) -> wrapper for kmeans!
source('~/Desktop/data/svn/programming/R/fuzzy-action/dev.r') # for basic sim-max operations (get.eu.s etc.)
```

Square Games
------------

A square game is a sim-max game with 4 states in a two dimensional similarity space. States are situated at coordinates (0,0), (0,x), (x,0) and (x,x) where x > 0. The distance between these states is given by the geometrical distance.

```{r}
x = 1
D = matrix(c(0,x,x,sqrt(2)*x,
             x,0,sqrt(2)*x,x,
             x,sqrt(2)*x,0,x,
             sqrt(2)*x,x,x,0),nrow=4, byrow=T)
show(D)
```

The (Nosofsky-)utility of these with parameter a is:

```{r}
a = 20
U = D
for (i in 1:4){
  for (j in 1:4){
    U[i,j] = exp(- (D[i,j]^2)/(a^2))
  }
}
show(U)
```

What matters to the utility table is the proportion between x and a, so we can better express the game with respect to a single parameter k=x/a. But we can then also think of x as fixed to x=1, so that a is our only parameter.

There are two messages in these games.

Equilibria
----------

There are two types of equilibria whose sender strategy partitions the state space.

* In *diagonal equilibria*, the sender uses one message for states 1 and 4, and another for 2 and 3. These are the states that are most dissimilar to each other. 
* In *convex equilibria*, the sender uses one message for two adjacent states, thereby partitioning the state space into convex regions.

Neither of these are strict, because the receiver can respond with different act to either message.

There are 4 diagonal equilibria, and 8 convex equilibria.

Convex equilibria are more efficient than diagonal equilibria.

We are interested in the basins of attraction of these under the replicator diffusion dynamic.

Replicator Dynamic
------------------

We run several trials of the (discrete-time) replicator dynamic to find out which attracting states exist.

```{r}
trials = 100
rounds = 100000
break.point = 0.0000001

ns = dim(U)[1]
nm = 2

priors = rep(1/ns,ns)
costs = matrix(rep(0,nm*ns),nrow=ns)

P = matrix(0,ncol=2*ns*nm+1,nrow=trials)
colnames(P) = c("S(t1,m1)", "S(t2,m1)", "S(t3,m1)", "S(t4,m1)",
                "S(t1,m2)", "S(t2,m2)", "S(t3,m2)", "S(t4,m2)",
                "R(m1,t1)", "R(m2,t1)", "R(m1,t2)", "R(m2,t2)",
                "R(m1,t3)", "R(m2,t3)", "R(m1,t4)", "R(m2,t4)",
                "EU")

for (t in 1:trials){

  setTxtProgressBar(pb, t)

  sen = rdirichlet(ns,rep(1/nm,nm))
  rec = rdirichlet(nm,rep(1/ns,ns))

  for (round in 1:rounds){

    s = sen
    r = rec

    eu.s = get.eu.s(rec ,U)
    eu.r = get.eu.r(sen,utils=U,priors=priors)

    sen = prop.table( sen * eu.s ,1)
    rec = prop.table( rec * eu.r ,1)

    if (sum(abs(sen-s)) <= break.point & sum(abs(rec-r)) <= break.point){
      break()
    }
  }
  eu = get.eu(priors,U,sen,rec,costs)
  P[t,] = c(as.vector(c(sen, rec)),eu)
}

P = P[,c(1:8,9,11,13,15,10,12,14,16,17)]

```

Show results:

```{r}
X = round(P,4)
Y = pamk(X,krange=2:22)
show(table(Y$pamobject$clustering))
show(Y$pamobject$medoids)
```

### 2014-09-03 snippets/square_games.Rmd
exploring square games

@@
 ```{r}
-a = 20
+a = 0.75
 U = D
@@
 
+Inner category coherence
+------------------------
+
+Get inner category coherence to classify results
+
+```{r}
+icc = function(D,sen,theta=0.01){
+  S = sen
+  S[sen>theta] = 1
+  S[sen<=theta] = 0
+  S = prop.table(S,2)
+  out = rep(0,dim(sen)[2])
+  for (i in 1:dim(sen)[2]) {
+    X = rowSums(D * S[,i])
+    Y = sum(S[,i] * X)
+    out[i] = Y
+  }
+  return(sum(out))
+}
+```
+
 Replicator Dynamic
@@
 ```{r}
-trials = 100
+trials = 50
 rounds = 100000
@@
 
-P = matrix(0,ncol=2*ns*nm+1,nrow=trials)
+P = matrix(0,ncol=2*ns*nm+2,nrow=trials)
 colnames(P) = c("S(t1,m1)", "S(t2,m1)", "S(t3,m1)", "S(t4,m1)",
@@
                 "R(m1,t3)", "R(m2,t3)", "R(m1,t4)", "R(m2,t4)",
-                "EU")
+                "EU", "ICC")
 
 for (t in 1:trials){
-
-  setTxtProgressBar(pb, t)
 
@@
   eu = get.eu(priors,U,sen,rec,costs)
-  P[t,] = c(as.vector(c(sen, rec)),eu)
+  icc.s = icc(D,sen)
+  P[t,] = c(as.vector(c(sen, rec)),eu,icc.s)
 }
 
-P = P[,c(1:8,9,11,13,15,10,12,14,16,17)]
+Q = P
+# P = P[,c(1:8,9,11,13,15,10,12,14,16,17,18,19)]
+P = P[,c(17,18)]
 
@@
 X = round(P,4)
-Y = pamk(X,krange=2:22)
+Y = pamk(X,krange=2:15)
 show(table(Y$pamobject$clustering))
 show(Y$pamobject$medoids)
+show(Q[Y$pamobject$id.med,1:8])
 ```

### 2014-09-12 snippets/iteration_diffusion.Rmd
investigating iterations of diffusion only

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(MCMCpack) # for Dirichlet distribution
source('~/Desktop/data/svn/programming/R/fuzzy-action/dev.r')
```

Iterated Diffusion
==================

The (discrete) Replicator Diffusion Dynamics first calculates the replicator dynamics and then applies diffusion, so to speak. To understand what diffusion does, we'd like to see it run without the replicator dynamics. We look at the sender and the receiver roles seperately.

Diffusion
---------

Diffusion is defined in terms of

```{r}

get.nbr = function(response,noise.matrix){
    if (dim(noise.matrix)[1] == dim(response)[2]){  # receiver part
        return(response %*% noise.matrix)
    }
    if (dim(noise.matrix)[2] == dim(response)[1]){  # sender part
        return(noise.matrix %*% response)
    }
}

```

Preliminaries & Parameters
--------------------------

```{r}

alpha= 0.1
rounds = 1000
ns = 6
nm = 3
states = seq(from=0,to=1,length.out = ns)
C = get.confusion.matrix(states,alpha)
```

Initialization
--------------

Create an arbitrary sender and receiver strategy.

```{r}

sen = rdirichlet(ns,rep(1,nm))
rec = rdirichlet(nm,rep(1,ns))

show(rec)

for (i in 1:rounds){
#   sen = get.nbr(sen,C)
  rec = get.nbr(rec,C)
}

show(rec)

```

### 2014-09-12 snippets/noisy_RD_and_RMD.r

@@ -1,0 +1,74 @@
+n.s = 2
+n.m = 2
+
+C = matrix(c(.8,.2,.2,.8),nrow=n.s)
+C = prop.table(C,1)
+rownames(C) = 1:n.s
+colnames(C) = 1:n.s
+
+pureS = rep("d",n.m^n.s)
+c = 1
+for (i in 1:n.s){
+  for (j in 1:n.m){
+    pureS[c] = paste(i,j,sep="")
+    c = c+1
+  }
+}
+
+Q = matrix(0,nrow=length(pureS), ncol=length(pureS))
+colnames(Q) = pureS
+rownames(Q) = pureS
+
+for (i in 1:dim(Q)[1]){
+  s1 = strsplit(rownames(Q)[i],split="")[[1]]
+  for (j in 1:dim(Q)[2]){
+    s2 = strsplit(colnames(Q)[j],split="")[[1]]
+    x = prod( sapply(1:n.s, function(t) sum(sapply(1:n.s, function(t2) ifelse(s2[t] == s1[t2], C[t,t2], 0 ) ) ) ) )
+    Q[i,j] = x
+  }
+}
+
+show(Q)
+
+mixed2behavioral = function(mixedS,pureS,n.s,n.m) {
+  b = matrix(0,n.s,n.m)
+  rownames(b)=1:n.s
+  colnames(b)=1:n.m
+  for (i in 1:n.s){
+    for (j in 1:n.m){
+      b[i,j] = sum( sapply(1:length(pureS), function(p) ifelse(strsplit(pureS[p],split="")[[1]][i]==j ,mixedS[p] ,0)) ) 
+    }
+  }
+  return(b)
+}
+
+mixedS = c(.24,.56,.06,.14)
+mixedS = c(1/3,1/3,1/3,0)
+mixedS = c(2/3,0,1/3,0)
+
+
+# show(mixed2behavioral(c(.24,.56,.06,.14), pureS,n.s,n.m))
+# show(mixed2behavioral(c(.2,.6,.1,.1), pureS,n.s,n.m))
+
+mutate = function(mixedS,Q){
+  m = rep(0,length(mixedS))
+  for (i in 1:length(mixedS)){
+    m[i] = sum(sapply(1:length(mixedS), function(j) Q[j,i] * mixedS[j]))
+  }
+  return(m)
+}
+
+mut = mutate(mixedS,Q) 
+show(mut)
[…]

### 2014-09-12 snippets/iteration_diffusion.Rmd
Iterated Diffusion

@@
 
+Preliminaries & Parameters
+--------------------------
+
+```{r}
+
+alpha= 0.2
+rounds = 1000
+ns = 7
+nm = 4
+states = seq(from=0,to=1,length.out = ns)
+C = get.confusion.matrix(states,alpha)
+```
+
 Diffusion
@@
 
-Diffusion is defined in terms of
+Diffusion is defined as simple matrix product for sender and receiver strategies.
 
@@
 
-Preliminaries & Parameters
---------------------------
+Iterated Diffusion
+------------------
 
-```{r}
-
-alpha= 0.1
-rounds = 1000
-ns = 6
-nm = 3
-states = seq(from=0,to=1,length.out = ns)
-C = get.confusion.matrix(states,alpha)
-```
-
-Initialization
---------------
-
-Create an arbitrary sender and receiver strategy.
+Create an arbitrary sender and receiver strategy, and iterate the diffusion process.
 
@@
 
-show(rec)
-
 for (i in 1:rounds){
-#   sen = get.nbr(sen,C)
+  sen = get.nbr(sen,C)
   rec = get.nbr(rec,C)
@@
 show(rec)
+show(sen)
 

### 2014-09-13 paper/01_intro.tex
introduction; iterated diffusion

@@
 
+\subsection{Situating our approach}
+
+A number of authors have recently tried to explain why vagueness
+evolved, based on considerations why a vague language might be useful
+\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,Jaegherde-JaegherRooijvan-Rooij2010:Strategic-Vague,BlumeBoard2013:Intentional-Vag}. In
+contrast to these, others have argued that vagueness if a natural
+byproduct of limitations in information processing
+\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
+paper makes a contribution to the latter.
+
+The view that vagueness is a natural concomitant of cognitive
+limitations of language users has been formalized in a number of
+ways. As in other applications of evolutionary game theory, we can
+distinguish micro- and macro-level approaches. Micro-level approaches
+look at adaptive behavior of individual agents. Usually, changes in
+the behavioral dispositions of agents occur after every single
+interaction. In contrast, macro-level approaches outline more
+abstract, aggregate population-based dynamics. Usually, the dynamic
+captures changes of frequencies of behavioral types in the population
+over time. Previous accounts have suggested both micro- and
+macro-level dynamics to show how vague signal use can arise in sim-max
+games.\todo{what about Ewan Klein's stuff?}
+
+As for micro-dynamics, \citet{FrankeJager2010:Vagueness-Signa} show
+how limited memory of past interactions can lead to vague signal use,
+when averaging over a single agent's behavior over time or over the
+momentary behavior of a population of several language
+users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
+reinforcement learning that includes a low-level form of stimulus
+generalization. Agents update their behavior after each round of play
+in such a way that similar states to the ones that actually occurred
+are subject to behavioral adjustment as well. (Alternative approaches
+are discussed in more detail in Section~XYZ, where we can compare them
+with the system introduced in this paper.)
+
+\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
+approach, using the notion of a \emph{quantal response}. A quantal
+response function is a probabilistic choice rule that formalizes the
+idea that agents make small mistakes when calculating the expected
+utility of choice options. In aggregation, these probabilistic
+mistakes lead to systematic ``trembles'' that produce vague signal
+use. The approach we take here is only superficially similar. For one,
+we not only look at equilibria, but take a dynamic perspective
+here. For another, we demonstrate in Section~XYZ that quantal
+responses can give rise to counterintuitive predictions in sim-max
+games, which suggests that the source of stochastic trembles should
+better not be seen in mistakes in calculating expected utility, but
+rather, as we assume here, as confusion of similar states.
+
+Against this background, we introduce here a new macro-level approach
+to the evolution of vague signal use in sim-max games, where the
+source of vagueness is the agents' natural disability to sharply
+distinguish perceptually distinct states. This applies to both
+production and comprehension, i.e., both to the sender as well as to
+the receiver role.
+
+We call our dynamic \textbf{replicator diffusion dynamic} (\rdd),
+because it is a special case of the replicator mutator dynamic
+(\rmd). It is a special case in the sense that the \rdd uses a special
+kind of mutation matrix that is based on the similarity of states (as
+defined in the sim-max game). More concretely, we start from the idea
+that similar states are more easily confused than less similar
+states. To formalize this idea, we assume that there is a confusion
+matrix $C$, where $C_{ij}$ is the probability that a state \mystate{i}
+is realized as a possibly different state \mystate{j}. (There is a
+slightly different sense of ``realization'' in the sender and the
+receiver side, and we will enlarge on this below.) Based on this, we
+show two things. For one, we show how the discrete-time formulation of
+the replicator dynamic can be easily adapted to integrate diffusion of
+behavior due to confusion of states. This yields the \rdd in its basic
+discrete-time formulation. For another, we also show that there exists
+a translation of the confusion matrix $C$ into a suitable mutation
+matrix $M_C$ such that the effect of $C$ and $M_C$ are ``equivalent''
+in a sense to be made precise. This supports the conclusion that the
+\rdd can indeed be thought of as a special case of the \rmd.
+
+To understand the workings of the \rdd, we study numerical simulations
+\dots
+
+The paper is structured as follows:
+\begin{enumerate}
+\item replicator diffusion dynamics
+\item relation to replicator mutator dynamics
+\item simulations of \rdd
+\item comparison with other approaches
+\item conclusions
+\end{enumerate}
+
 %%% Local Variables: 

### 2014-09-13 paper/03_simulations.tex
introduction; iterated diffusion

@@
 The quantity of information in a signal is an old notion that also goes back to the start of information theory.
-Skyrms~\citet[ch.~3]{Skyrms2010} discusses its use in the context of signaling games.
+Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of signaling games.
 The main idea is that we can quantify the amount of information about a state $\state$ in a message $\messg$ via the relation between the probability of the state given the message $\Pr(\state \,|\, \messg)$ and the unconditional probability of the state $\Pr(\state)$.
@@
 Finally, we equate sender infomativity with the average overall information about states in each signal.
-Based on the definition by Skyrms~\citet[p.~36]{Skyrms2010}, and the considerations above, we define sender informativity as follows:
+Based on the definition by Skyrms~\citet[p.~36]{Skyrms2010:Signals}, and the considerations above, we define sender informativity as follows:
 \begin{align*}
@@
 We equate receiver informativity with the average overall information about acts in each signal.
-Based on Skyrms' definition~\citet[p.~39]{Skyrms2010}, we define receiver informativity as follows:
+Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
 \begin{align*}
@@
 \newcommand{\impairment}{\alpha}
-In our simulations we focus on a concrete setup in line with \cite{Correia2013}.
+In our simulations we focus on a concrete setup in line with \cite{Correia2013:The-Bivalent-Tr}.
 Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
@@
 where $\impairment$ is an \emph{impairment} factor.
-Here we deviate from \cite{Correia2013} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
+Here we deviate from \cite{Correia2013:The-Bivalent-Tr} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
 Confusion is directly related with similarity, \emph{i.e.}~the more similar two states are, the more likely an agent is to perceive one as the other.

### 2014-09-13 snippets/iteration_diffusion.Rmd
introduction; iterated diffusion

@@
 
-Iterated Diffusion
+(Iterated) Diffusion
 ==================
 
-The (discrete) Replicator Diffusion Dynamics first calculates the replicator dynamics and then applies diffusion, so to speak.
+The (discrete) Replicator Diffusion Dynamic first calculates the replicator dynamic and then applies diffusion, so to speak.
 To understand what diffusion does, we'd like to see it run without the replicator dynamics.
-We look at the sender and the receiver roles seperately.
+
+Two things are interesting: (1) what does diffusion do in one step, (2) what does it do in iteration.
+
+We must look at the sender and the receiver roles seperately.
 
@@
 ```{r}
-
-alpha= 0.2
-rounds = 1000
-ns = 7
-nm = 4
+alpha= 0.1
+rounds = 10000
+ns = 11
+nm = 3
 states = seq(from=0,to=1,length.out = ns)
@@
 
+But that means that we can define iterated diffusion as either left or right multiplication with suitable powers of the confusion matrix. Therefore define a the n-th power of the confusion matrix.
+
+```{r}
+power = function(C,n){
+  # this could be improved for speed
+  if (n <=1){
+    return(C)
+  }
+  else{
+    if (n %% 2 ==0){
+      X = power(C,n/2)
+      return(X%*%X)
+    }
+    else{
+      return(C%*%power(C,n-1))
+    }
+  }
+}
+```
+
 Iterated Diffusion
@@
 
-for (i in 1:rounds){
-  sen = get.nbr(sen,C)
-  rec = get.nbr(rec,C)
-}
+CPower = power(C,rounds)
+show(CPower)
 
-show(rec)
+sen = CPower %*% sen
 show(sen)
 
+rec = rec %*% CPower  
+show(rec)
+
 ```

### 2014-09-15 paper/01_intro.tex
minor changes in introduction

@@
 \citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
-paper makes a contribution to the latter.
+paper makes a contribution to the latter line of thought.
 
@@
 interaction. In contrast, macro-level approaches outline more
-abstract, aggregate population-based dynamics. Usually, the dynamic
-captures changes of frequencies of behavioral types in the population
-over time. Previous accounts have suggested both micro- and
+abstract, aggregate population-based dynamics. Usually, a macro-level
+dynamic captures changes of frequencies of behavioral types in the
+population over time. Previous accounts have suggested both micro- and
 macro-level dynamics to show how vague signal use can arise in sim-max
-games.\todo{what about Ewan Klein's stuff?}
+games.
 
@@
 mistakes lead to systematic ``trembles'' that produce vague signal
-use. The approach we take here is only superficially similar. For one,
-we not only look at equilibria, but take a dynamic perspective
-here. For another, we demonstrate in Section~XYZ that quantal
-responses can give rise to counterintuitive predictions in sim-max
-games, which suggests that the source of stochastic trembles should
-better not be seen in mistakes in calculating expected utility, but
-rather, as we assume here, as confusion of similar states.
+use. The approach we take here is superficially similar, but there are
+crucial divergences. For one, we adopt a dynamic perspective by
+looking at the stable states of a suitable population dynamics. For
+another, we demonstrate in Section~XYZ that quantal responses can give
+rise to counterintuitive predictions. These counterintuitive examples
+suggest that vagueness in signaling use is not convincingly explained
+by appeal to mistakes in calculating expected utility, but rather, as
+we assume here, as the result of confusing similar states of the word.
 
-Against this background, we introduce here a new macro-level approach
+Against this background, we introduce a new macro-level approach
 to the evolution of vague signal use in sim-max games, where the
@@
 because it is a special case of the replicator mutator dynamic
-(\rmd). It is a special case in the sense that the \rdd uses a special
-kind of mutation matrix that is based on the similarity of states (as
+(\rmd). It is a special case in the sense that the \rdd incorporates a
+special kind of mutation that is based on the similarity of states (as
 defined in the sim-max game). More concretely, we start from the idea

### 2014-09-15 paper/03_simulations.tex
fixed RD in vagueness_games.py; adapted batch script for values where non-convexity shows for low impairment

@@
-\section{Simulations}
+\section{Exploring the RDD}
+
+\subsection{(Iterated) Diffusion}
+
+We have seen that the \rdd can be conceptualized as a two-step
+procedure. We first calculate where the replicator dynamic would take
+us, and then apply diffusion. Since the latter step is the novel
+addition, we should look at diffusion in isolation to understand what
+the \rdd does.
+
+\subsection{Simulations}
 
@@
 
-\subsection{Metrics}
+\subsubsection{Metrics}
 \label{sec:metrics}
@@
 
-\subsection{Methodology}
+\subsubsection{Methodology}
 \newcommand{\impairment}{\alpha}
@@
 
-\subsection{Results}
+\subsubsection{Results}
 In order to visualize the impact of the impairment factor on the outcome of simulations, we plot in Figure~\ref{fig:metrics-boxplots} the distribution of each of the measurements per impairment value.

### 2014-09-15 snippets/data_exploration/data_exploration_tmp.Rmd
fixed RD in vagueness_games.py; adapted batch script for values where non-convexity shows for low impairment

--- 2014-08-27 data_exploration_tmp.Rmd
@@
 require('reshape2')
-source('~//Desktop//data//svn//programming//R//helpers//helpers.R')
+source('~/Desktop/data/svn/programming/R/helpers/helpers.R')
 
-data = read.csv('results/20140813-194306.csv', 
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140915-190401.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',

### 2014-09-15 run-batch.sh

@@ -4,10 +4,10 @@ DATE=$(date +%Y%m%d-%H%M%S)
 HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Speaker informativity,Hearer informativity,Expected utility,Iterations,Speaker strategy file,Hearer strategy file"
 
 echo $HEADER > results/$DATE.csv
-for t in 10 50 90 130
+for t in 6 10
 do
     echo $t
-    for i in 0 0.1 0.2 0.3 0.4 0.5
+    for i in 0 0.025 0.05 0.1 0.2
     do
         echo " $i"
     	for n in {1..100}

### 2014-09-15 vagueness-games.py

@@ -73,7 +73,7 @@ def makePDF(Vector):
 
 def makePDFPerRow(Matrix):
     return np.array([ makePDF(Row) for Row in Matrix ])
-    
+
 def ExpectedUtility(Speaker, Hearer, Utility):
     return np.sum(Speaker[t1, m] * Hearer[m, t2] * Utility[t1, t2]
                   for t1 in xrange(Speaker.shape[0])
@@ -103,7 +103,7 @@ def Convexity(Strategy):
     ConvexPureStrategies = convex_sequences(range(Strategy.shape[1]), Strategy.shape[0])
     return sum(np.prod([ Strategy[c, s[c]] for c in xrange(Strategy.shape[0])])
                for s in ConvexPureStrategies)
-    
+
 def Voronoiness(HearerStrategy, SpeakerStrategy, StateSimilarity):
     Prototypes = [ np.argmax(HearerStrategy[m]) for m in xrange(HearerStrategy.shape[0]) ]
 
@@ -134,7 +134,7 @@ def InformationQuantity(SpeakerStrategy, HearerStrategy, Priors):
     NStates = SpeakerStrategy.shape[0]
     NMessages = SpeakerStrategy.shape[1]
     NActs = HearerStrategy.shape[1]
-    
+
     InformationOnStatesPerMessage = np.zeros(NMessages)
     for m in xrange(NMessages):
         MessageProbability = sum(Priors[t] * SpeakerStrategy[t, m] for t in xrange(NStates))
@@ -149,7 +149,7 @@ def InformationQuantity(SpeakerStrategy, HearerStrategy, Priors):
             ActProbability = sum(Priors[t] * SpeakerStrategy[t, m2] * HearerStrategy[m2, a] for t in xrange(NStates) for m2 in xrange(NMessages))
             InformationOnActsPerMessage[m] += HearerStrategy[m, a] * np.log(HearerStrategy[m, a] / ActProbability) if HearerStrategy[m, a] != 0 else 0
     InformationOnActs = sum(InformationOnActsPerMessage[m] for m in xrange(NMessages)) / NMessages
-    
+
     return (InformationOnStates, InformationOnActs)
 
 # # Settings
@@ -160,14 +160,17 @@ PriorDistributionType = 'uniform'
 NMessages = 2
 
 Dynamics = 'replicator dynamics'
-Impairment = 0.1
 
-Tolerance = Impairment
+Impairment = 0.2
+Tolerance = 0.025
+
[…]

### 2014-09-16 paper/02_noisyRMD.tex
improved structure Sec 02; data for convex vs. non-convex strats for small state spaces

@@
-\section{Noise-perturbed replicator dynamic}
+\section{Replicator diffusion dynamic}
 
-\subsection{Preliminaries}
+The replicator diffusion dynamic (\rdd) was first introduced by
+\citet{Correia2013:The-Bivalent-Tr} as a noise-perturbed variant of
+the replicator dynamic (\rd) in behavioral strategies. We begin by
+recapitulating the formulation of the \rd in behavioral
+strategies. Then we define the replicator diffusion dynamics in
+behavioral strategies, following \citet{Correia2013:The-Bivalent-Tr}.
+Finally we show that the \rdd is a special case of the \rmd.
 
-\paragraph{Notation.} Fix a signaling game with states $\States$,
-messages $\Messgs$ and acts $\Acts$. % Let $\Util_{\sen,\rec} \mycolon
-% \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ be the
-% player's utility function.
+\subsection{Replicator dynamic in behavioral strategies}
+
+Fix a signaling game with states $\States$, messages $\Messgs$ and
+acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
+distribution over states and $\Util_{\sen,\rec} \mycolon \States
+\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
+receiver's utility functions. The sender's behavioral strategies are
+functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
+functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
+utility of actions in each choice point are defined as usual:
+\begin{align*}
+  \EU(\messg \myts ; \myts \state, \Rstrat) & = \sum_{\act \in \Acts}
+  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
+  \EU(\act \myts ; \myts \messg, \Sstrat) & = \sum_{\state \in
+    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
+  U_\rec(\state, \messg, \act) \,.
+\end{align*}
+The \emph{fitness} of a behavioral strategy at a choice point is the
+frequency-weighted average of expected utilities of each choice, given
+the opponent's strategy:
+\begin{align*}
+  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
+\EU(\messg \myts ; \myts\state,\Rstrat) \\
+\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
+\cdot \EU(\act \myts ; \myts \messg,\Sstrat)\,.
+\end{align*}
+The discrete-time replicator dynamic maps current strategies $\Sstrat$
+and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$ in
+such a way that changes in frequency are proportional to expected
+utilities. For behavioral strategies, the changes take place locally
+at each of the agents' choice points:\todo{could be shortened by using
+  $\propto$ notation}
+\begin{align*}
+  \RD(\Sstrat)(\messg \probbar \state_i) & = \frac{\Sstrat(\messg \probbar \state_j) \cdot
+    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Sstrat, \Rstrat)} \\
+    \RD(\Rstrat)(\state_i \probbar \messg) & = \frac{\Rstrat(\state_j \probbar \messg) \cdot
+    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
+\end{align*}
+
+\todo[inline]{example? sim-max game 10 states, after so and so many rounds?}
+
+\subsection{Replicator diffusion dynamic in behavioral strategies}
+
+The replicator diffusion dynamic adds probabilistic confusion of
+similar states to the replicator dynamics. Fix a sim-max game with
+$\States = \Acts$ and a confusion matrix $C$. $C$ is a row-stochastic
+matrix whose elements $C_{ij}$ give the probability that $\state_i$ is
+realized as $\state_j$. The confusability of states affects senders
+and receivers alike, but in slightly different ways (see
+Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
+$C_{ij}$ is the probability that the actual states \mystate{i} is
+perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
+probability that \mystate{j} is the interpretation that is actually
+formed when \mystate{i} is the output of the receiver's strategy.
+
+\begin{figure}
+  \centering
+
+    \begin{tikzpicture}[node distance = 2cm, thick]
+
+      \begin{scope}
+
+      \node[rectangle, draw=mycol, fill=mycol!25, thick] (actual)
+      {actual state $\mystate{i}$};
+
+      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      actual] (perceived) {perceived state $\mystate{j}$};
+
+      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      perceived] (output) {chosen message $\messg$};
+
+      \node[rectangle, thick, below of = output, node distance=1cm]
+      (sender) {sender};
+
+      \draw[->] (actual) -> (perceived) node[midway,left] {noise
+        $C_{ij}$};
+
+      \draw[->] (perceived) -> (output) node[midway,left] (label)
+      {strategy $\Sstrat(\messg \probbar \mystate{j})$};
+
+      \begin{pgfonlayer}{background}
+        \node [draw=black!50, fill=black!20,fit=(actual) (label)
+        (output)] {};
+      \end{pgfonlayer}
+    \end{scope}
+
+      \begin{scope}[xshift=5cm]
+
+      \node[rectangle, draw=mycol, fill=mycol!25, thick] (message)
+      {observed message $\messg$};
+
+      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      message] (target) {target interpretation $\mystate{i}$};
+
+      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      target] (realized) {realized interpretation $\mystate{j}$};
+
+      \node[rectangle, thick, below of = realized, node distance=1cm]
+      (sender) {receiver};
+
+      \draw[->] (message) -> (target) node[midway,right] (bla) {strategy
+        $\Rstrat(\mystate{i} \probbar \messg)$};
+
+      \draw[->] (target) -> (realized) node[midway,right]  {noise $C_{ij}$};
+
+      \begin{pgfonlayer}{background}
+        \node [draw=black!50, fill=black!20,fit = (message) 
+        (realized) (bla)] {};
+      \end{pgfonlayer}
+    \end{scope}
+
+  \end{tikzpicture}
+
+  \caption{Effect of confusion of states on sender and receiver
+    choices.}
+  \label{fig:noise-perturbation-of-strategies}
+\end{figure}
+
+The aggregate effect of confusion of states on behavioral strategies
+can be captured in a function that maps behavioral strategies
+$\Sstrat$ and $\Rstrat$ on their noise-perturbed realizations
+$C(\Sstrat)$ and $C(\Rstrat)$. The idea is that $\Sstrat$ and
+$\Rstrat$ are what, on average, the idealized, confusion-free behavior
+would be, while $C(\Sstrat)$ and $C(\Rstrat)$ are behavioral
+strategies that describe the agents' actual noise-perturbed
+probabilistic behavior. If we conceive of behavioral strategies
+$\Sstrat$ and $\Rstrat$ as row-stochastic matrices, the effect of
+state confusability is easily captured by matrix multiplication as:
+\begin{align*}
+  C(\Sstrat) & = C \Sstrat &    C(\Rstrat) & = \Rstrat C\,.
+\end{align*}
+
+The discrete-time replicator diffusion dynamic takes the replicator
+dynamic as basic, but factors in the confusion of states at each
+update step in a sequential update:
+\begin{align*}
+  \RDD(\Sstrat) & = C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = C(\RD(\Rstrat)) 
+\end{align*}
+This is equivalent to the following definition, given by
+\citet{Correia2013:The-Bivalent-Tr}:
+\begin{align*}
+  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
+  \frac{\Sstrat(\messg \probbar \state_j) \cdot
+    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
+    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ij} \cdot
+  \frac{\Rstrat(\state_j \probbar \messg) \cdot
+    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
+\end{align*}
+
+The idea behind the sequential definition of the \rdd is that, at each
+time step, strategies are gradually optimized along the current
+fitness landscape, as described by the \rd, but the realization of
+optimized strategies is bound to be noisy, due to confusion of
+states. It is obvious that other (discrete-time) evolutionary dynamics
+can be subjected to state-confusability in an analogous fashion
+\citep{Correia2013:The-Bivalent-Tr}. In the case of the \rd, however,
+perturbation by state-confusion has a prominent close relative in the
+replicator mutator dynamic.
+
+\subsection{Relation with the replicator mutator dynamic}
+
+The replicator mutator dynamic (\rmd) has been proposed by Martin
+Nowak first in the context of signaling game models for the evolution
+of grammar
+\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. The
+relation of the \rmd to other prominent evolutionary dynamics is well
+understood \citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen
+further fruitful applications in the context of signaling games
+\citep{HutteggerSkyrms2010:Evolutionary-Dy}. It is therefore desirable
+to relate the \rdd that we propose here to the \rmd.
+
+A direct comparison between \rdd and the \rmd is not possible, because
+the usual formulation of the latter is in its continuous-time form,
+and based on mixed strategies, not behavioral strategies. We therefore
+give a discrete-time formulation of the \rmd in behavioral strategies,
+which is independently useful. The relation between \rdd and \rmd will
+then be plain to see.
+
+\subsubsection{Replicator mutator dynamic}
 
@@
 frequency $\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for
-the receiver).
+the receiver). For mixed strategies, we define the players' fitness in
+the usual manner. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
+$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
+$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
+F_k^{\Rmixed}$ is the average fitness in the sender population and
+$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
+average fitness in the receiver population.
 
-Let the sender's (receiver's) behavioral strategies be functions
-$\Sstrat \in \Delta(\Messgs)^\States$ ($\Rstrat \in
-\Delta(\Acts)^\Messgs$). Every mixed strategy $\Smixed$ converts to a
-unique behavioral strategy defined by:
+Every mixed strategy $\Smixed$ converts to a unique behavioral
+strategy defined by:
 \begin{align*}
@@
 that $G$ is \emph{not} an injection, as many mixed strategies map onto
-the same behavioral strategy.
-
-For behavioral strategies, let $\EU(\messg, \state, \Rstrat)$ and
-$\EU(\act, \messg, \Sstrat)$ be the players' expected utilities of
-each action in each choice point as usual. For mixed strategies we
-define the players' fitness in the usual manner. Let $F_i^{\Rmixed}$
-be $\Spure_i$'s fitness given $\Rmixed$ and $F_i^{\Smixed}$ be
-$\Rpure_i$'s fitness given $\Smixed$. Then $\Phi(\Smixed,\Rmixed) =
-\sum_{k} \Smixed_k \cdot F_k^{\Rmixed}$ is the average fitness in the
-sender population and $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k
-\cdot F_k^{\Smixed}$ the average fitness in the receiver population.
+the same behavioral strategy. \todo{this should not be here}
 

### 2014-09-16 snippets/data_exploration/data_exploration_convexity.Rmd
improved structure Sec 02; data for convex vs. non-convex strats for small state spaces

Data Exploration
================

``` {r warnings=FALSE, message=FALSE, echo=FALSE}

require('ellipse')
require('plyr')
require('ggplot2')
require('reshape2')
source('~/Desktop/data/svn/programming/R/helpers/helpers.R')

data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140916-134358.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric',
                             'factor','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','character','character'))

ddply(data, .(Impairment,Number.of.states), summarise, no.convex = sum(Speaker.Convex.Cat)) -> sumconv

qplot(Impairment,no.convex,data=sumconv, geom=c("point","line"),ylab="percentage of convex outcomes", color=factor(Number.of.states))

```

### 2014-09-16 snippets/data_exploration/data_exploration_tmp.Rmd
improved structure Sec 02; data for convex vs. non-convex strats for small state spaces

@@
 
-load and pre-process the data
-
-``` {r warnings=FALSE, message=FALSE}
+``` {r warnings=FALSE, message=FALSE, echo=FALSE}
 
@@
 
-data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140915-190401.csv', 
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140916-133324.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
@@
                              'numeric','numeric','numeric','numeric','numeric',
-                             'numeric','character','character'))
+                             'numeric','factor','factor','character','character'))
 
-ds = data[,c(1,4,7,8,11:16)] # only relevant data
+ds = data[,c(1,4,7,8,9,10,11:16)] # only relevant data
 dm = melt(ds,id.vars = c("Number.of.states", "Impairment")) # melted data
@@
 
-pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy +
+pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy + Speaker.convexity +
         Speaker.Voronoiness + Hearer.Voronoiness +
@@
 
-pd <- position_dodge(.05)
+pd <- position_dodge(.01)
 sp = ggplot(dsub, aes(x=Impairment, y=value, colour=factor(Number.of.states))) + 

### 2014-09-16 run-batch.sh

@@ -1,13 +1,13 @@
 #!/bin/bash
 
 DATE=$(date +%Y%m%d-%H%M%S)
-HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Speaker informativity,Hearer informativity,Expected utility,Iterations,Speaker strategy file,Hearer strategy file"
+HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Speaker informativity,Hearer informativity,Expected utility,Iterations,Speaker Convex Cat, Hearer Convex Cat, Speaker strategy file,Hearer strategy file"
 
 echo $HEADER > results/$DATE.csv
 for t in 6 10
 do
     echo $t
-    for i in 0 0.025 0.05 0.1 0.2
+    for i in 0 0.025 0.05 0.075 0.1 0.125 0.15
     do
         echo " $i"
     	for n in {1..100}

### 2014-09-16 vagueness-games.py

@@ -104,6 +104,11 @@ def Convexity(Strategy):
     return sum(np.prod([ Strategy[c, s[c]] for c in xrange(Strategy.shape[0])])
                for s in ConvexPureStrategies)
 
+def ConvexityCat(Strategy):
+    ConvexPureStrategies = convex_sequences(range(Strategy.shape[1]), Strategy.shape[0])
+    return sum(np.prod([ Strategy[c, s[c]] == max(Strategy[c]) for c in xrange(Strategy.shape[0])])
+               for s in ConvexPureStrategies)
+
 def Voronoiness(HearerStrategy, SpeakerStrategy, StateSimilarity):
     Prototypes = [ np.argmax(HearerStrategy[m]) for m in xrange(HearerStrategy.shape[0]) ]
 
@@ -329,4 +334,5 @@ csv.writer(OutputFile).writerow([NStates, PriorDistributionType, NMessages, Impa
     SpeakerVoronoiness, HearerVoronoiness, \
     InformativitySpeaker, InformativityHearer, \
     ExpectedUtility(Speaker, Hearer, Utility) / OptimalExpectedUtility, i, \
+    ConvexityCat(Speaker), ConvexityCat(Hearer), \
     SpeakerOutputFilename, HearerOutputFilename])

### 2014-09-17 paper/02_noisyRMD.tex
restructured and expanded Section 2

@@
 recapitulating the formulation of the \rd in behavioral
-strategies. Then we define the replicator diffusion dynamics in
+strategies. Then we define the replicator diffusion dynamic in
 behavioral strategies, following \citet{Correia2013:The-Bivalent-Tr}.
-Finally we show that the \rdd is a special case of the \rmd.
+Finally we show that the \rdd is a special case of the replicator
+mutator dynamic \citep[e.g.][]{Nowak2006:Evolutionary-Dy}.
 
@@
 
-Fix a signaling game with states $\States$, messages $\Messgs$ and
-acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
+Fix a signaling game with finite states $\States$, messages $\Messgs$
+and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
 distribution over states and $\Util_{\sen,\rec} \mycolon \States
@@
 functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
-utility of actions in each choice point are defined as usual:
+utility of actions in each choice point are as usual:
 \begin{align*}
@@
 \begin{align*}
-  \RD(\Sstrat)(\messg \probbar \state_i) & = \frac{\Sstrat(\messg \probbar \state_j) \cdot
-    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Sstrat, \Rstrat)} \\
-    \RD(\Rstrat)(\state_i \probbar \messg) & = \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
+  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
+    \EU(\messg \myts ; \myts \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
+    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
+    \EU(\act \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
 \end{align*}
@@
 The replicator diffusion dynamic adds probabilistic confusion of
-similar states to the replicator dynamics. Fix a sim-max game with
+similar states to the replicator dynamic. Fix a sim-max game with
 $\States = \Acts$ and a confusion matrix $C$. $C$ is a row-stochastic
@@
 can be captured in a function that maps behavioral strategies
-$\Sstrat$ and $\Rstrat$ on their noise-perturbed realizations
+$\Sstrat$ and $\Rstrat$ to their noise-perturbed realizations
 $C(\Sstrat)$ and $C(\Rstrat)$. The idea is that $\Sstrat$ and
@@
 \begin{align*}
-  \RDD(\Sstrat) & = C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = C(\RD(\Rstrat)) 
+  \RDD(\Sstrat) & = C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = C(\RD(\Rstrat)) \,.
 \end{align*}
-This is equivalent to the following definition, given by
-\citet{Correia2013:The-Bivalent-Tr}:
+This is equivalent to the following, perhaps more transparent
+formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
 \begin{align*}
@@
   \frac{\Sstrat(\messg \probbar \state_j) \cdot
-    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
+    \EU(\messg \myts ; \myts \state_j,\Rstrat)}
+  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
     \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ij} \cdot
   \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
+    \EU(\state_j \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
 \end{align*}
@@
 fitness landscape, as described by the \rd, but the realization of
-optimized strategies is bound to be noisy, due to confusion of
-states. It is obvious that other (discrete-time) evolutionary dynamics
-can be subjected to state-confusability in an analogous fashion
-\citep{Correia2013:The-Bivalent-Tr}. In the case of the \rd, however,
-perturbation by state-confusion has a prominent close relative in the
-replicator mutator dynamic.
+optimized strategies is noisy, due to confusion of states. It is
+obvious that other (discrete-time) evolutionary dynamics can be
+subjected to state-confusability in an analogous fashion. In the case
+of the \rd, however, perturbation by state-confusion has a prominent
+close relative in the replicator mutator dynamic.
 
-\subsection{Relation with the replicator mutator dynamic}
+\subsection{The replicator mutator dynamic}
 
@@
 further fruitful applications in the context of signaling games
-\citep{HutteggerSkyrms2010:Evolutionary-Dy}. It is therefore desirable
-to relate the \rdd that we propose here to the \rmd.
+\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy,Deo2014:The-Semantic-an}. It
+is therefore desirable to relate the \rdd that we propose here to the
+\rmd.
 
@@
 which is independently useful. The relation between \rdd and \rmd will
-then be plain to see.
+then be plain to see. We start with a delineation of mixed and
+behavioral strategies.
 
-\subsubsection{Replicator mutator dynamic}
+\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
+functions $\Spure \in \Messgs^\States$ ($\Rpure \in
+\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
+$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
+(\Acts^\Messgs)$). The latter give the relative population frequencies
+of the former. We write $\Smixed_i$ for the frequency
+$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for the
+receiver).
 
-Pure sender (receiver) strategies are functions $\Spure \in
-\Messgs^\States$ ($\Rpure \in \Acts^\Messgs$). Mixed sender (receiver)
-strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
-($\Rmixed \in (\Acts^\Messgs)$). The latter give the relative
-population frequencies of the former. We write $\Smixed_i$ for the
-frequency $\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for
-the receiver). For mixed strategies, we define the players' fitness in
-the usual manner. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
+Mixed strategies hold more information than behavioral
+strategies. This is because every mixed strategy $\Smixed$ converts to
+a unique behavioral strategy defined by:
+\begin{align*}
+  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
+\end{align*} 
+Let $G$ be this mapping from mixed to behavioral strategies. This $G$
+is \emph{not} an injection, as many mixed strategies map onto the same
+behavioral strategy. In this sense, mixed strategies hold more
+information about a distribution of pure strategies in a population.
+
+In the context of evolutionary dynamics, we can think of behavioral
+strategies as treating each choice point as an independent update site
+\citep[e.g.][]{Cressman2003:Evolutionary-Dy,Sandholm2013:Population-Game}. Dynamics
+for mixed strategies conceive of agents as adjusting their whole pure
+strategy globally. In this way, dependencies between different choices
+at different choice points can matter to the evolutionary path. In
+contrast, dynamics for behavioral strategies conceive of agents as
+adjusting their strategies locally at each choice point, regardless of
+what they are bound to do at other choice points. The latter
+perspective is then more coarse-grained, and not necessarily
+equivalent to the former. An advantage of dynamics in behavioral
+strategies is that they reduce complexity, which is especially helpful
+for computational simulations of larger games, like the ones we are
+interested in here.
+
+\paragraph{Replicator dynamic in mixed strategies.} To understand the
+replicator mutator dynamic in mixed strategies, we first introduce the
+replicator dynamic in mixed strategies as introduced by
+\citet{TaylorJonker1978:Evolutionary-St}.
+
+For mixed strategies, we define the players' fitness in the
+usual manner. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
 $\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
@@
 $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
-average fitness in the receiver population.
+average fitness in the receiver population.\todo{give actual
+  definition of fitness etc.?}
 
-Every mixed strategy $\Smixed$ converts to a unique behavioral
-strategy defined by:
-\begin{align*}
-  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
-\end{align*} 
-Let $G$ be this mapping from mixed to behavioral strategies. Notice
-that $G$ is \emph{not} an injection, as many mixed strategies map onto
-the same behavioral strategy. \todo{this should not be here}
-
-\paragraph{Replicator dynamics in pure strategies.} The two-population
-(non-payoff adjusted) continuous replicator dynamics in pure
-strategies is defined as:
+The continuous-time replicator dynamic defines the change of frequency
+of mixed strategies. Its so-called two-population version that takes
+sender and receiver roles into account separately is defined as:
 \begin{align*}
@@
 \end{align*}
-The discrete time version is given by: 
+Similar to the version for behavioral strategies, a discrete-time
+version of the \rd for mixed strategies can be conceptualized as an
+update function that maps mixed strategies $\Smixed$ and $\Rmixed$
+onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
+such that:
 \begin{align*}
-  \Smixed_i' & = \frac{\Smixed_i \cdot
-  F_i^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} &     \Rmixed_i' & = \frac{\Rmixed_i \cdot
+  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
+    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
     F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
@@
 
-\paragraph{Replicator-mutator equation.} Let $Q$ be a row-stochastic
-mutation matrix where $Q_{ji}$ gives the probability that pure sender
-strategy $\Spure_j$ mutates into $\Spure_i$. Similarly, let $R$ be a
-row-stochastic mutation matrix where $R_{ji}$ gives the probability
-that pure receiver strategy $\Rpure_j$ mutates into $\Rpure_i$.
+\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
+extends the \rd by adding probabilistic mutation. Let $Q$ be a
+row-stochastic mutation matrix where $Q_{ji}$ gives the probability
+that pure sender strategy $\Spure_j$ mutates into
+$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
+where $R_{ji}$ gives the probability that pure receiver strategy
+$\Rpure_j$ mutates into $\Rpure_i$.
 
-The two-population (non-payoff adjusted) continuous replicator-mutator
-dynamics, is then given by::
+The \rmd is usually given only in its continuous-time form. A
+two-population (non-payoff adjusted) formulation is then:
 \begin{align*}
@@
 \end{align*}
-\begin{claim} A discrete time version of the above replicator mutator
-  dynamics is:
+Although a discrete-time formulation of the dynamics is absent from the
+literature, as far as we know, it is easy to spell out, and clearly
+demonstrates that the sequential nature of the \rmd, parallel to the
+intuition that motivated the formulation of the \rdd.
+
+\begin{claim} A discrete time version of the replicator mutator
+  dynamic is:
   \begin{align*}
-    \Smixed_i' & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
-      F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \Rmixed_i' & = \sum_{j}
-    R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
+    \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
+      F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
+    \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
       \Phi(\Rmixed,\Smixed)}\,.
@@
 
-The discrete time replicator-mutator dynamics has a nice sequential
-update property: first we compute the fitness-driven change according
-to the standard replicator dynamics; then we compute the perturbation
-from mutation.
-
-\subsection{Noise-perturbed replicator dynamics}
-
-We look at signaling games with $\States = \Acts$ and fix a confusion
-matrix $C$, which is a row-stochastic matrix whose elements $C_{ij}$
-give the probability that $\state_i$ is perceived as
-$\state_j$. Define the players' average utility at each choice point given
-the opponent's strategy as:
-\begin{align*}
-  \Phi(\state,\Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
-\EU(\messg,\state,\Rstrat) &
-\Phi(\messg,\Sstrat) & = \sum_{\state} \Rstrat(\state \probbar \messg)
-\cdot \EU(\state,\messg,\Sstrat)\,.
-\end{align*}
-The discrete noise-perturbed replicator dynamics on behavioral
-strategies proposed by \citet{Correia2013:The-Bivalent-Tr} is:
-\begin{align*}
-  \Sstrat'(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
-  \frac{\Sstrat(\messg \probbar \state_j) \cdot
-    \EU(\messg,\state_j,\Rstrat)} {\Phi(\state_j,\Rstrat)} \\
-    \Rstrat'(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
-  \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j,\messg,\Sstrat)} {\Phi(\messg,\Sstrat)}  \,.
-\end{align*}
-Notice that this update is also sequential: first we calculate an
-update according to the standard replicator dynamics (in behavioral
-strategies), then we compute the perturbation from perceptual noise.
-
-\subsection{Exploring the relation}
-
-We will show that the noise-perturbed replicator dynamics defined
-above is the behavioral-strategy analogue to the replicator-mutator
-dynamics when the only source of mutation is perceptual confusion of
-states. Since, in general, the replicator dynamics in two-populations
-is not equivalent in its formulations for behavioral and mixed
-strategies \citep{Cressman2003:Evolutionary-Dy}, we abstract from the
-dynamics and look only at the effect of mutation and
-noise-perturbation. This is justified given the above mentioned
-sequential nature of both dynamics.
-
-Let the mutation $Q(\Smixed)$ ($R(\Rmixed)$) of a mixed sender
-(receiver) strategy be another mixed sender (receiver) strategy
-defined by:
+The discrete-time \rmd has the same sequential nature as the \rdd:
+first we compute the fitness-driven change according to the standard
+replicator dynamic; then we compute the perturbation from
+mutation. This becomes clear if we define an independent mutation
+function that maps mixed strategies onto the outcome of (one round of)
+mutation: 
 \begin{align*}
@@
 \end{align*}
-We would like to compare this to the noise perturbation $C(\Sstrat)$
-($C(\Rstrat)$) of behavioral strategy $\Sstrat$ ($\Rstrat)$, which is
-another behavioral strategy given by:
+The discrete-time \rmd given above can then be rewritten as:
 \begin{align*}
-  C(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ji} \cdot
-  \Sstrat(\messg \probbar \state_j) & C(\Rstrat)(\state_i \probbar
-  \messg) & = \sum_{j} C_{ji} \cdot \Rstrat(\state_j \probbar \messg)
-  \,.
+  \RDD(\Smixed) &= Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= Q(\RD(\Rmixed))\,. 
 \end{align*}
 
-We hypothesize that a confusion matrix $C$ should give rise to a
-unique mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$
-we also have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the
-receiver.
+\subsection{Diffusion as a special kind of mutation}
+
+The sequential reformulation of the \rmd in terms of a sequence of
+updates already closely resembles our initial formulation of the
+\rdd. However, there is a major difference between these. While the
+\rdd is a dynamic for behavioral strategies, the \rmd, as taken from
+the literature, is a dynamic for mixed strategies. Similarly, the
+confusability of states is an operation that is straightforwardly
+definable for behavioral strategies (see Equation~XYZ), while mutation
+in the \rmd is defined as mutation from one pure strategy to
+another. Nonetheless, it is possible to map each confusion matrix to a
+mutation matrix in such a way that the natural correspondence between
+strategies is conserved (see
+Figure~\ref{fig:correspondence-result}). This is the main result
+presented in this section, and it is our justification for regarding
+the \rdd as the behavioral-strategy analogue of the replicator-mutator
+dynamic when the only source of mutation is perceptual confusion of
+states.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[node distance = 2cm]
+
+    \begin{scope}
+      \node[] (mixed) {$\Smixed$};
+
+      \node[below of = mixed] (mutated) {$Q^C(\Smixed)$};
+
+      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};
+
+      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};
+
+      \draw[->, dotted, thick] (mixed) -> (mutated);
+
+      \begin{pgfonlayer}{background}
+        \node [draw=black!50, fill=black!20,fit = (dummyl) 
+        (mutated) (dummyr)] (mixedSpace) {};
+      \end{pgfonlayer}
+
+      \node[below of = mixedSpace, node distance = 1.7cm] {space of
+        mixed strategies};
+    \end{scope}
+
+    \begin{scope}[xshift=6cm]
+      \node[] (behavioral) {$\Sstrat$};
+
+      \node[below of = behavioral] (mutatedBeh) {$C(\Sstrat)$};
+
+      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};
+
+      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};
+
+      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);
+
+      \begin{pgfonlayer}{background}
+        \node [draw=black!50, fill=black!20,
+              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
+      \end{pgfonlayer}
+
+      \node[below of = behSpace, node distance = 1.7cm] {space of
+        behavioral strategies};
+    \end{scope}
+
+    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
+    above]{$G(\Smixed)$};
+
+    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(Q^C(\Smixed))$};
+
+  \end{tikzpicture}
+
+  \caption{Correspondence between state-confusion and mutation.}
+  \label{fig:correspondence-result}
+\end{figure}
+
+More concretely, since there is a non-injective mapping $G$ from
+behavioral to mixed strategies, we would like to show that there is a
+systematic translation from a confusion matrix $C$ to a mutation
+matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
+unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
+confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$ to the
+mutation-perturbation $Q^C(\Smixed)$ (see
+Figure~\ref{fig:correspondence-result}). In other words, we
+hypothesize that a confusion matrix $C$ should give rise to a unique
+mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
+have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver. This
+is the content of Theorems~\ref{thm:sender-eq} and
+~\ref{thm:receiver-eq}), the proofs of which are in
+Appendix~\ref{sec:proofs}. The remainder of this section introduces
+the relevant construction that maps arbitrary $C$ to suitable $Q^C$.
 
@@
 
-The probability that $\Rpure$ mutates into $\Rpure'$ is the product of
-the probabilities for each $\messg$ that $\Rpure(\messg)$ is perceived
-as $\Rpure'(\messg)$. Abusing notation, by refer to the index of
-$\Rpure(\messg)$ and $\Rpure'(\messg)$ with $\Rpure(\messg)$ and
-$\Rpure'(\messg)$ directly, we define:
-\begin{align*}
-  R^C_{ji} = \prod_{\messg} C_{\Rpure(\messg)\Rpure'(\messg)}\,.
-\end{align*}
+The probability that $\Rpure_i$ mutates into $\Rpure_j$ is the product of
+the probabilities for each $\messg$ that the state $\Rpure_i(\messg)$ is
+perceived as state $\Rpure_j(\messg)$. Abusing notation, by referring to
+the indices of states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
+$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
+\begin{align}
+  \label{eq:construction-rec}
+  R^C_{ij} = \prod_{\messg} C_{\Rpure(\messg)\Rpure'(\messg)}\,.
+\end{align}
 
@@
 as a state $\state_l$ which $\Spure_j$ would map onto $\messg'$. The
-overal probability that $\Spure_j$ mutates into $\Sstrat_i$ is the
+overal probability that $\Spure_j$ mutates into $\Spure_i$ is the
 product of these chances for all states. So, define:
-\begin{align*}
+\begin{align}
+  \label{eq:construction-rec}
   Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
     \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
-\end{align*}
+\end{align}
 
@@
 
-\medskip
-
-\todo[inline]{necessary to prove that mutation matrices are row-stochastic?}
-
-\todo[inline]{maybe worthwhile to think about general properties of
-  the mutation matrices that ensue from perceptual confusion?}
-
-\medskip
+Based on the translations from confusion matrices $C$ to mutation
+matrices $Q^C$ and $R^C$ in Equations~(\ref{eq:construction-rec}) and
+(\ref{eq:construction-rec}), we can finally formulate the
+desired correspondence result, that reveals the \rdd as a special case
+of the \rmd in behavioral strategies where the source of mutation is
+imperfect discriminability of states.
 
@@
 
+\noindent Proofs are given in Appendix~\ref{sec:proofs}.
+
 %%% Local Variables: 

### 2014-09-17 paper/02_noisyRMD.tex
section 2 proof read; iterated diffusion analysis

@@
 The discrete-time replicator dynamic maps current strategies $\Sstrat$
-and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$ in
-such a way that changes in frequency are proportional to expected
+and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
+in such a way that changes in frequency are proportional to expected
 utilities. For behavioral strategies, the changes take place locally
-at each of the agents' choice points:\todo{could be shortened by using
-  $\propto$ notation}
+at each of the agents' choice points:
 \begin{align*}
@@
 \end{align*}
-
-\todo[inline]{example? sim-max game 10 states, after so and so many rounds?}
 
@@
 state confusability is easily captured by matrix multiplication as:
-\begin{align*}
+\begin{align}
+  \label{eq:confusion-function}
   C(\Sstrat) & = C \Sstrat &    C(\Rstrat) & = \Rstrat C\,.
-\end{align*}
+\end{align}
 
@@
   {\Phi(\state_j,\Sstrat, \Rstrat)} \\
-    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ij} \cdot
+    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
   \frac{\Rstrat(\state_j \probbar \messg) \cdot
@@
 
-The replicator mutator dynamic (\rmd) has been proposed by Martin
-Nowak first in the context of signaling game models for the evolution
-of grammar
+The replicator mutator dynamic (\rmd) has been proposed first in the
+context of signaling game models for the evolution of grammar
 \citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. The
@@
 
-For mixed strategies, we define the players' fitness in the
-usual manner. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
-$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
-$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
-F_k^{\Rmixed}$ is the average fitness in the sender population and
+Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given $\Rmixed$ and
+$F_i^{\Smixed}$ be $\Rpure_i$'s fitness given $\Smixed$. Then
+$\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot F_k^{\Rmixed}$ is
+the average fitness in the sender population and
 $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
@@
 The continuous-time replicator dynamic defines the change of frequency
-of mixed strategies. Its so-called two-population version that takes
-sender and receiver roles into account separately is defined as:
+of mixed strategies. Its so-called two-population version, that takes
+sender and receiver roles into account separately, is defined as:
 \begin{align*}
@@
   \begin{align*}
-    \dot{\Smixed_i} & = \Smixed_i' - \Smixed_i = \sum_{j} Q_{ji}
+    \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
     \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
@@
     & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )
+        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
   \end{align*}
@@
 \rdd is a dynamic for behavioral strategies, the \rmd, as taken from
-the literature, is a dynamic for mixed strategies. Similarly, the
+the literature, applies to mixed strategies. Similarly, the
 confusability of states is an operation that is straightforwardly
-definable for behavioral strategies (see Equation~XYZ), while mutation
-in the \rmd is defined as mutation from one pure strategy to
-another. Nonetheless, it is possible to map each confusion matrix to a
-mutation matrix in such a way that the natural correspondence between
-strategies is conserved (see
-Figure~\ref{fig:correspondence-result}). This is the main result
+definable for behavioral strategies (see
+Equations~(\ref{eq:confusion-function})), while mutation in the \rmd is
+defined as mutation from one pure strategy to another. Nonetheless, it
+is possible to map each confusion matrix to a mutation matrix in such
+a way that the natural correspondence between strategies is conserved
+(see Figure~\ref{fig:correspondence-result}). This is the main result
 presented in this section, and it is our justification for regarding
@@
 unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
-confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$ to the
+confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
 mutation-perturbation $Q^C(\Smixed)$ (see
@@
 is the content of Theorems~\ref{thm:sender-eq} and
-~\ref{thm:receiver-eq}), the proofs of which are in
+~\ref{thm:receiver-eq}, the proofs of which are in
 Appendix~\ref{sec:proofs}. The remainder of this section introduces
@@
 the probabilities for each $\messg$ that the state $\Rpure_i(\messg)$ is
-perceived as state $\Rpure_j(\messg)$. Abusing notation, by referring to
+perceived as state $\Rpure_j(\messg)$. Abusing notation by referring to
 the indices of states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
@@
   \label{eq:construction-rec}
-  R^C_{ij} = \prod_{\messg} C_{\Rpure(\messg)\Rpure'(\messg)}\,.
+  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j\messg)}\,.
 \end{align}

### 2014-09-17 snippets/iteration_diffusion.Rmd
section 2 proof read; iterated diffusion analysis

@@
 library(MCMCpack) # for Dirichlet distribution
-source('~/Desktop/data/svn/programming/R/fuzzy-action/dev.r')
+library(ggplot2)
+library(reshape2)
+# source('~/Desktop/data/svn/programming/R/fuzzy-action/dev.r')
 ```
@@
 rounds = 10000
-ns = 11
-nm = 3
+ns = 60
+nm = 2
 states = seq(from=0,to=1,length.out = ns)
+
+get.confusion.matrix = function(states,alpha){
+  ns = length(states)
+  if (alpha <= 0){
+    return(diag(rep(1,ns)))
+  }
+  C = matrix(0,nrow=ns, ncol=ns)
+  for (i in 1:ns){
+    for (j in 1:ns){
+      C[i,j] = exp( - (states[i]-states[j])^2 / (alpha^2) )
+    }
+  }
+  return(prop.table(C,1))
+}
+
 C = get.confusion.matrix(states,alpha)
+
 ```
@@
 
-But that means that we can define iterated diffusion as either left or right multiplication with suitable powers of the confusion matrix. Therefore define a the n-th power of the confusion matrix.
+We can define iterated diffusion as either left or right multiplication with suitable powers of the confusion matrix. Therefore define a the n-th power of the confusion matrix.
 
@@
 }
+```
+
+Simple Diffusion
+----------------
+
+Let's plot an arbitrary sender and receiver strategy, and apply some noise.
+
+First a function to plot an arbitrary sender and receiver strategy
+
+```{r}
+
+plotSender = function(sen){
+  senDF = melt(sen,varnames = c("state.index","message"), value.name = "probability")
+  senDF$states = states[senDF[,1]]
+  senDF$message = factor(senDF$message)
+  ggplot(senDF, aes(x = states, y=probability, color=message)) + geom_point() + geom_line()
+}
+
+```
+
+```{r}
+
+sen = rdirichlet(ns,rep(1,nm))
+
+plotSender(sen)
+plotSender(get.nbr(sen,C))
+plotSender(get.nbr(get.nbr(sen,C),C))
+
+sys.exit()
+
+# rec = rdirichlet(nm,rep(1,ns))
+
 ```

### 2014-09-18 paper/03_simulations.tex
iterated diffusion; plots

@@
 the \rdd does.
+
+\begin{itemize}
+  \item which confusion matrices: normalized Nosofsky (positive matrices)
+  \item repeated applications
+  \item diffusion leads to a smoothing out of sender and receiver
+    behavior to neighboring states
+  \item how much depends on the level of confusion (see Figure~\ref{fig:diffusion-SenRec})
+  \item in the limit, this will lead sender and receiver to do the
+    same thing in each choice point (states for the sender, messages
+    for the receiver); this follows from the XYZ theorem
+\end{itemize}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\textwidth]{plots/confusion_sender.pdf}
+
+  \includegraphics[width=\textwidth]{plots/confusion_receiver.pdf}
+
+  \caption{Iterated diffusion for arbitraty sender (top) and receiver
+    strategy (bottom), with 50 states, 2 messages. Each row
+    corresponds to a different confusion level $\alpha$, each colum to
+    a different number of applications of iterated diffusion.}
+  \label{fig:confusion-SenRec}
+\end{figure}
 

### 2014-09-18 snippets/iteration_diffusion.Rmd
iterated diffusion; plots

@@
 nm = 2
-states = seq(from=0,to=1,length.out = ns)
+
+get.states = function(ns){
+  return(seq(from=0,to=1,length.out = ns))
+}
+
+states = get.states(ns)
 
@@
 
-Simple Diffusion
-----------------
+Single Diffusion Steps
+----------------------
 
-Let's plot an arbitrary sender and receiver strategy, and apply some noise.
+Let's plot an arbitrary sender and receiver strategy, and apply some noise, step by step until convergence.
 
-First a function to plot an arbitrary sender and receiver strategy
+Parameters
 
@@
 
-plotSender = function(sen){
-  senDF = melt(sen,varnames = c("state.index","message"), value.name = "probability")
-  senDF$states = states[senDF[,1]]
+ns = 50
+nm = 2
+rounds = c(1,2,10,50,1000)
+alphas = c(0.01,0.05,0.1,0.3)
+
+plotSender = function(ns,nm,alphas,rounds){
+  sen = rdirichlet(ns,rep(1,nm))
+  senDF = data.frame(state = 0,
+                     message = 0,
+                     round = 0,
+                     alpha = 0,
+                     probability = 0)
+
+  for (alpha in alphas){
+    C = get.confusion.matrix(get.states(ns),alpha)
+    newDF = data.frame(state = rep(get.states(ns),nm),
+                       message = rep(1:2,each=ns),
+                       round = rep(0,ns*nm),
+                       alpha = rep(alpha,ns*nm),
+                       probability = as.vector(sen))
+    for (r in rounds){
+      newC = power(C,r)
+      newSen = get.nbr(sen,newC)
+      newDF2 = data.frame(state = rep(get.states(ns),nm),
+                          message = rep(1:2,each=ns),
+                          round = rep(r,ns*nm),
+                          alpha = rep(alpha,ns*nm),
+                          probability = as.vector(newSen))
+      newDF = rbind(newDF,newDF2)                   
+      }
+    senDF = rbind(senDF,newDF)
+  }
+  senDF = senDF[-1,]
   senDF$message = factor(senDF$message)
-  ggplot(senDF, aes(x = states, y=probability, color=message)) + geom_point() + geom_line()
+  senDF$alpha = factor(senDF$alpha)
+  p = ggplot(senDF,aes(x = state, y = probability, color = message)) + 
+      geom_point(size=0.5) + geom_line() + facet_grid(alpha~round) +
+      scale_x_continuous(breaks=seq(0,1,1/2)) +
+      theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + 
+      scale_y_continuous(breaks=seq(0,1,1/2))
+
+  return(p)
 }
+
+plotReceiver = function(ns,nm,alphas,rounds){
+  rec = rdirichlet(nm,rep(1,ns))
+  recDF = data.frame(state = 0,
+                     message = 0,
+                     round = 0,
+                     alpha = 0,
+                     probability = 0)
+
+  for (alpha in alphas){
+    C = get.confusion.matrix(get.states(ns),alpha)
+    newDF = data.frame(state = rep(get.states(ns),nm),
+                       message = rep(1:2,each=ns),
+                       round = rep(0,ns*nm),
+                       alpha = rep(alpha,ns*nm),
+                       probability = as.vector(t(rec)))
+    for (r in rounds){
+      newC = power(C,r)
+      newRec = get.nbr(rec,newC)
+      newDF2 = data.frame(state = rep(get.states(ns),nm),
+                          message = rep(1:2,each=ns),
+                          round = rep(r,ns*nm),
+                          alpha = rep(alpha,ns*nm),
+                          probability = as.vector(t(newRec)))
+      newDF = rbind(newDF,newDF2)                   
+      }
+    recDF = rbind(recDF,newDF)
+  }
+  recDF = recDF[-1,]
+  recDF$message = factor(recDF$message)
+  recDF$alpha = factor(recDF$alpha)
+  p = ggplot(recDF,aes(x = state, y = probability, color = message)) + 
+      geom_point(size=0.5) + geom_line() + facet_grid(alpha~round) +
+      scale_x_continuous(breaks=seq(0,1,1/2)) +
+      theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + 
+      scale_y_continuous(breaks=seq(0,round(max(recDF$probability),2),round(max(recDF$probability),2)/2))
+  return(p)
+}
+
+pSen = plotSender(ns,nm,alphas,rounds)
+pRec = plotReceiver(ns,nm,alphas,rounds)
+show(pSen)
+show(pRec)
+
+pdf(file = "paper/plots/confusion_sender.pdf", width = 7, height = 4)
+show(pSen)
+dev.off()
+
+pdf(file = "paper/plots/confusion_receiver.pdf", width = 7, height = 4)
+show(pRec)
+dev.off()
 

### 2014-09-18 paper/03_simulations.tex
more on diffusion; Nosofsky similarity plots

@@
 the \rdd does.
+
+Confusion of states should intuitively be a function of their
+perceptual similarity. Here and in the following, we assume that the
+state space of the sim-max game consists of $n \ge 2$ states that are
+equally spaced across the unit interval, including $0$ and $1$. We
+assume here that the distance $\card{\mystate{i} - \mystate{j}}$ is
+the objective, physical similarity between two states $\mystate{i}$
+and $\mystate{j}$. Physical distance space feeds into a perceptual
+similarity function, as described by
+\citet{Nosofsky1986:Attention-Simil}:
+\begin{align}
+  \label{eq:NosofskySim}
+  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \alpha) = \expo \left ( -
+    \frac{\card{\mystate{i} - \mystate{j}}^s}{ \alpha^2} \right ) \,,
+\end{align}
+where $\alpha$ is an imprecision parameter. When $\alpha=0$ agents
+perfectly discriminate between states; when $\alpha \rightarrow
+\infty$ agents cannot discriminate states at
+all. Figure~\ref{fig:NosofskySim} gives an impression of
+Nosofsky-similarity for different parameter values. Other
+formalizations of perceptual similarity are possible, including ones
+that allow for different discriminability in different areas of the
+state space, but we stick with Nosofsky's similarity function for the
+time being, because it is a mathematically simple, yet established
+notion in mathematical psychology.
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\textwidth]{plots/NosofskySim.pdf}
+
+  \caption{Examples for Nosofsky-similarity (non-normalized) and
+    derived probability of state confusion (normalized) for different
+    values of $\alpha$ and $\card{\States} = 50$.}
+  \label{fig:NosofskySim}
+\end{figure}
+
+We further assume that the probability of confusing any two states
+$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
+similarity and therefore obtained by normalization:
+\begin{align*}
+  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
+  \alpha)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
+  \alpha)} \,.
+\end{align*}
+Confusion of states is then also a function of imprecision parameter
+$\alpha$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\alpha
+= 0$ the confusion matrix has $C_{ii} = 1$; for $\alpha > 0$ $C$ is
+positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\alpha
+\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
+$i$ and $j$.
 
@@
 \end{figure}
+
+\newpage
 

### 2014-09-18 snippets/Nosofsky_similarity.Rmd
more on diffusion; Nosofsky similarity plots

```{r warning=FALSE, echo=FALSE, message=FALSE}
library(ggplot2)
library(reshape2)
```

Nosofsky Similarity
===================

Nosofsky similarity is governed by a single parameter alpha.

```{r}
get.confusion.matrix = function(states,alpha){
  ns = length(states)
  if (alpha <= 0){
    return(diag(rep(1,ns)))
  }
  C = matrix(0,nrow=ns, ncol=ns)
  for (i in 1:ns){
    for (j in 1:ns){
      C[i,j] = exp( - (states[i]-states[j])^2 / (alpha^2) )
    }
  }
  return(C)
}
```

We can plot the similarity for different distances for different values of alpha.

```{r}
ns = 50
alphas = c(0.05, 0.1, 0.2, 0.3)

states = seq(0,1,length.out=ns)

out = matrix(0,nrow=length(alphas),ncol=ns)
outNorm = matrix(0,nrow=length(alphas),ncol=ns)

for (a in 1:length(alphas)){
  out[a,] = get.confusion.matrix(states,alphas[a])[1,]
  outNorm[a,] = prop.table(get.confusion.matrix(states,alphas[a]),1)[1,]
}

simDF = data.frame(distance = rep(states,2*length(alphas)),
                   alpha = factor(rep(rep(alphas, each=ns),2)),
                   normalized = factor(rep(c("non-normalized", "normalized"),each=ns*length(alphas))),
                   similarity = c(as.vector(t(out)),as.vector(t(outNorm)))  )

p = ggplot(simDF, aes(x = distance, y = similarity, color=alpha)) + geom_point() +
  geom_line() + facet_grid(normalized~. , scales = "free")

show(p)

pdf('paper/plots/NosofskySim.pdf', width=7, height=4)
show(p)
dev.off()
```

### 2014-09-18 paper/02_noisyRMD.tex
diffusion section ready

@@
 can be captured in a function that maps behavioral strategies
-$\Sstrat$ and $\Rstrat$ to their noise-perturbed realizations
-$C(\Sstrat)$ and $C(\Rstrat)$. The idea is that $\Sstrat$ and
-$\Rstrat$ are what, on average, the idealized, confusion-free behavior
-would be, while $C(\Sstrat)$ and $C(\Rstrat)$ are behavioral
-strategies that describe the agents' actual noise-perturbed
-probabilistic behavior. If we conceive of behavioral strategies
-$\Sstrat$ and $\Rstrat$ as row-stochastic matrices, the effect of
-state confusability is easily captured by matrix multiplication as:
+$\Sstrat$ and $\Rstrat$ to their diffusions $D(\Sstrat)$ and
+$D(\Rstrat)$. The idea is that $\Sstrat$ and $\Rstrat$ are what, on
+average, the idealized, noise-free behavior would be, while
+$D(\Sstrat)$ and $D(\Rstrat)$ are behavioral strategies that describe
+the agents' actual noise-perturbed probabilistic behavior. Since the
+effect of confusion of states is that behavior at one choice point
+percolates to behavior at similar choice points, we speak of
+\emph{diffusion of behavior under confusion of states} (see
+Section~\ref{sec:iterated-diffusion} for an investigation into this
+diffusion process). If we conceive of behavioral strategies $\Sstrat$
+and $\Rstrat$ as row-stochastic matrices, the diffusion effect of
+state confusability is easily captured by matrix multiplication as a
+diffusion function $\Diff_C$ that, based on confusion matrix $C$, maps
+behavioral strategies on their diffused realizations:
 \begin{align}
   \label{eq:confusion-function}
-  C(\Sstrat) & = C \Sstrat &    C(\Rstrat) & = \Rstrat C\,.
+  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
 \end{align}
@@
 \begin{align*}
-  \RDD(\Sstrat) & = C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = C(\RD(\Rstrat)) \,.
+  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
 \end{align*}
@@
 replicator dynamic; then we compute the perturbation from
-mutation. This becomes clear if we define an independent mutation
-function that maps mixed strategies onto the outcome of (one round of)
-mutation: 
+mutation. This becomes clear if we define independent mutation
+functions $\Mutate_{Q,R}$ that maps mixed strategies onto the outcome
+of (one round of) mutation:
 \begin{align*}
-  Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
-  Q_{ji} &   R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
+  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
+  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
   R_{ji} \,.
@@
 \begin{align*}
-  \RDD(\Smixed) &= Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= Q(\RD(\Rmixed))\,. 
+  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,. 
 \end{align*}
@@
 \begin{align*}
-S^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
+Q^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                       11 & 1 & 0 & 0 & 0 \cr
@@
   \label{thm:sender-eq}
-  If $G(\Smixed) = \Sstrat$, then $G(Q^C(\Smixed)) = C(\Sstrat)$.
+  If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) = \Diff_C(\Sstrat)$.
 \end{theorem}
@@
   \label{thm:receiver-eq}
-  If $G(\Rmixed) = \Rstrat$, then $G(R^C(\Rmixed)) = C(\Rstrat)$.
+  If $G(\Rmixed) = \Rstrat$, then $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
 \end{theorem}

### 2014-09-18 paper/03_simulations.tex
diffusion section ready

@@
 
+The previous section showed that the \rdd can be conceptualized as a
+two-step procedure. We first calculate where the replicator dynamic
+would take us, and then apply diffusion. Since the latter step is the
+novel addition, we should look at diffusion in isolation to understand
+what the \rdd does. This is what we will do in
+Section~\ref{sec:iterated-diffusion}. Section~\ref{sec:simulations}
+\dots does something else \dots
+
 \subsection{(Iterated) Diffusion}
-
-We have seen that the \rdd can be conceptualized as a two-step
-procedure. We first calculate where the replicator dynamic would take
-us, and then apply diffusion. Since the latter step is the novel
-addition, we should look at diffusion in isolation to understand what
-the \rdd does.
+\label{sec:iterated-diffusion}
 
@@
 the objective, physical similarity between two states $\mystate{i}$
-and $\mystate{j}$. Physical distance space feeds into a perceptual
+and $\mystate{j}$. Distance in physical space feeds into a perceptual
 similarity function, as described by
@@
 
-\begin{itemize}
-  \item which confusion matrices: normalized Nosofsky (positive matrices)
-  \item repeated applications
-  \item diffusion leads to a smoothing out of sender and receiver
-    behavior to neighboring states
-  \item how much depends on the level of confusion (see Figure~\ref{fig:diffusion-SenRec})
-  \item in the limit, this will lead sender and receiver to do the
-    same thing in each choice point (states for the sender, messages
-    for the receiver); this follows from the XYZ theorem
-\end{itemize}
+Difusion of behavior under confusion of states was defined in
+Equation~(\ref{eq:confusion-function}), repeated here:
+\begin{align*}
+  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
+\end{align*}
+If states are confused on average with a probability proportional to
+their similarity, the effect on noise-perturbed behavior is that
+behavioral dispositions gradually diffuse along a gradient of
+similarity of states as well. Consequently, iterated diffusion leads
+to a smoothing out and an eventual equalization of behavioral
+strategies, for both the sender and the
+receiver. Figure~\ref{fig:confusion-SenRec} shows an arbitrary sender
+and receiver strategy after one or several rounds of iterated
+application of diffusion, for different values of perceptual
+imprecision $\alpha$.
+
+The example in Figure~\ref{fig:confusion-SenRec} shows that iterated
+diffusion leads to sender and receiver behavior that is the same at
+all choice points. After $n$ steps of diffusion an initial sender
+strategy $\Sstrat$ will be $C^n \Sstrat$, and an initial receiver
+strategy will be $\Rstrat C^n$. If we assume that $\alpha > 0$ and so
+$C$ is a positive row-stochastic matrix (each state could in principle
+be confused as any other state, albeit with possibly a very low
+probability), it follows from the Perron-Frobenius theorem that the
+limit $C^\infty = \lim_{n \rightarrow \infty} C^n$ exists (it is the
+Perron-projection) and all of its rows are identical. That is why, in
+the limit of iterated diffusion, $C^\infty S$ has identical rows
+(messages are sent with the same probability in each state), and $R
+C^\infty$ likewise has identical rows (every message is interpreted
+equally). All of this is in line with the intuition that diffusion of
+behavior under confusion of states iteratively equalizes behavior for
+similar states; if all states can be confused for one another in
+principle, behavior smoothes out entirely in the limit.
 
@@
 
-  \caption{Iterated diffusion for arbitraty sender (top) and receiver
-    strategy (bottom), with 50 states, 2 messages. Each row
-    corresponds to a different confusion level $\alpha$, each colum to
-    a different number of applications of iterated diffusion.}
+  \caption{Iterated diffusion for one arbitrary sender (top) and one
+    receiver strategy (bottom), for a game with 50 states and 2
+    messages. Each row corresponds to a different confusion level
+    $\alpha$, each colum to a different number of applications of
+    iterated diffusion.}
   \label{fig:confusion-SenRec}
@@
 
-\newpage
+Now that we understand better what diffusion does, let us look at how
+diffusion interacts with optimization of behavior, as described by the
+replicator dynamics.
 
-\subsection{Simulations}
+\subsection{Exploring the replicator diffusion dynamics}
+\label{sec:simulations}
 

### 2014-09-18 paper/03_simulations.tex
more data exploration with varying tolerance and impairment

@@
   \label{eq:NosofskySim}
-  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \alpha) = \expo \left ( -
-    \frac{\card{\mystate{i} - \mystate{j}}^s}{ \alpha^2} \right ) \,,
+  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
+      \begin{cases}
+    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
+    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) \,,& \textrm{otherwise} \\
+    \end{cases}
 \end{align}
-where $\alpha$ is an imprecision parameter. When $\alpha=0$ agents
-perfectly discriminate between states; when $\alpha \rightarrow
+where $\impairment >= 0$ is an imprecision parameter. When $\impairment=0$ agents
+perfectly discriminate between states; when $\impairment \rightarrow
 \infty$ agents cannot discriminate states at
@@
     derived probability of state confusion (normalized) for different
-    values of $\alpha$ and $\card{\States} = 50$.}
+    values of $\impairment$ and $\card{\States} = 50$.}
   \label{fig:NosofskySim}
@@
   C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
-  \alpha)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
-  \alpha)} \,.
+  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
+  \impairment)} \,.
 \end{align*}
 Confusion of states is then also a function of imprecision parameter
-$\alpha$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\alpha
-= 0$ the confusion matrix has $C_{ii} = 1$; for $\alpha > 0$ $C$ is
-positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\alpha
+$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
+= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$ $C$ is
+positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
 \rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
@@
 application of diffusion, for different values of perceptual
-imprecision $\alpha$.
+imprecision $\impairment$.
 
@@
 strategy $\Sstrat$ will be $C^n \Sstrat$, and an initial receiver
-strategy will be $\Rstrat C^n$. If we assume that $\alpha > 0$ and so
+strategy will be $\Rstrat C^n$. If we assume that $\impairment > 0$ and so
 $C$ is a positive row-stochastic matrix (each state could in principle
@@
     messages. Each row corresponds to a different confusion level
-    $\alpha$, each colum to a different number of applications of
+    $\impairment$, each colum to a different number of applications of
     iterated diffusion.}
@@
 \label{sec:simulations}
+
+\begin{itemize}
+\item utility function: another Nosofsky-style function
+\item get data for different set sizes, impairment and utility parameters
+\item 
+\end{itemize}
+
+\newpage
 
@@
 \subsubsection{Methodology}
-\newcommand{\impairment}{\alpha}
+
 In our simulations we focus on a concrete setup in line with \cite{Correia2013:The-Bivalent-Tr}.

### 2014-09-18 run-batch.sh

@@ -4,17 +4,22 @@ DATE=$(date +%Y%m%d-%H%M%S)
 HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Speaker informativity,Hearer informativity,Expected utility,Iterations,Speaker Convex Cat, Hearer Convex Cat, Speaker strategy file,Hearer strategy file"
 
 echo $HEADER > results/$DATE.csv
-for t in 6 10
+for t in 6 10 50 90
 do
     echo $t
-    for i in 0 0.025 0.05 0.075 0.1 0.125 0.15
+    for i in 0 0.05 0.1 0.2 0.3
     do
         echo " $i"
-    	for n in {1..100}
-    	do
-	        echo -n "  $n"
-        	python vagueness-games.py $t 2 $i results/$DATE.csv
-    	done
+	for p in 0.05 0.1 0.2 0.3
+	do
+	    echo -n "   $p"
+    	    for n in {1..50}
+    	    do
+	        echo -n "    $n"
+        	python vagueness-games.py $t 2 $i $p results/$DATE.csv
+    	    done
+	    echo
+	done
     	echo
     done
 done

### 2014-09-18 vagueness-games.py

@@ -159,7 +159,7 @@ def InformationQuantity(SpeakerStrategy, HearerStrategy, Priors):
 
 # # Settings
 
-NStates = 6
+NStates = 10
 PriorDistributionType = 'uniform'
 
 NMessages = 2
@@ -180,13 +180,14 @@ BatchMode = len(sys.argv) > 1
 
 if BatchMode:
     if len(sys.argv) < 5:
-        print "Usage: python", sys.argv[0], "<number of states> <number of messages> <impairment> <output file>"
+        print "Usage: python", sys.argv[0], "<number of states> <number of messages> <impairment> <tolerance> <output file>"
         sys.exit(1)
     else:
         NStates = int(sys.argv[1])
         NMessages = int(sys.argv[2])
         Impairment = float(sys.argv[3])
-        OutputFile = open(sys.argv[4], 'ab')
+        Tolerance = float(sys.argv[4])
+        OutputFile = open(sys.argv[5], 'ab')
 
 # # Initialization
 

### 2014-09-18 snippets/data_exploration/data_exploration_tmp.Rmd
data exploration

@@
 
-data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140916-133324.csv', 
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
@@
                              'numeric','numeric','numeric','numeric','numeric',
-                             'numeric','factor','factor','character','character'))
+                             'numeric','numeric','numeric','character','character'))
 
-ds = data[,c(1,4,7,8,9,10,11:16)] # only relevant data
-dm = melt(ds,id.vars = c("Number.of.states", "Impairment")) # melted data
+ds = data[,c(1,4,5,7,8,11:17)] # only relevant data
+dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
 
@@
 
-pairs(~ Number.of.states + Impairment + Speaker.entropy + Hearer.entropy + Speaker.convexity +
+pairs(~ Number.of.states + Impairment + Tolerance + Speaker.entropy + Hearer.entropy + Speaker.convexity +
         Speaker.Voronoiness + Hearer.Voronoiness +
-        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations, data=ds)
+        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations +
+        Speaker.Convex.Cat, data=ds)
 ```
@@
 
-dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "variable"), measurevar="value")
+dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
 dsub = subset(dms, !grepl("informativity",dms$variable) )
@@
 
-sp + facet_wrap(~ variable, ncol=2, scales="free")
+sp + facet_grid(variable ~ Tolerance, scales="free")
 

### 2014-09-21 snippets/data_exploration/data_exploration_tmp.Rmd
data exploration with varying tolerance; some plotting functions added

@@
 require('reshape2')
+require('plyr')
 source('~/Desktop/data/svn/programming/R/helpers/helpers.R')
+source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
 
-data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.csv', 
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
@@
 attach(ds)
+
+plot.language.byindex = function(dataFrame,rowindex){
+  plot.language(dataFrame[rowindex,]$Speaker.strategy.file,
+                dataFrame[rowindex,]$Hearer.strategy.file)
+}
+
+plot.language.MinMaxProperty = function(dataFrame,colindex){
+  show(names(dataFrame)[colindex])
+  show(summary(dataFrame[,colindex]))
+  p = arrangeGrob(plot.language.byindex(dataFrame,which.min(dataFrame[,colindex])),
+                  plot.language.byindex(dataFrame,which.max(dataFrame[,colindex])))
+}
+
 ```
@@
 
-pairs(~ Number.of.states + Impairment + Tolerance + Speaker.entropy + Hearer.entropy + Speaker.convexity +
+pairs(~ Number.of.states + Impairment + Tolerance + Speaker.entropy + Hearer.entropy + 
         Speaker.Voronoiness + Hearer.Voronoiness +
@@
 dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
-dsub = subset(dms, !grepl("informativity",dms$variable) )
+dsub = dms
 
@@
 - so we should look for stages before final convergence
+
+Informativity
+-------------
+
+This is the notion from Skyrms. It seems very well-behaved. There is a clear positive correlations between speaker and hearer informativity.
+
+```{r}
+cor.test(data$Speaker.informativity, data$Hearer.informativity)
+qplot(data$Speaker.informativity, data$Hearer.informativity) + geom_smooth(method=lm)
+```
+
+There is a clear negative correlation between impairment and informativity. The remaining variance here must and can be accounted for in terms of Number.of.states (see main plot).
+
+```{r}
+cor.test(data$Impairment, data$Speaker.informativity)
+qplot(data$Impairment, data$Speaker.informativity) + geom_smooth(method=lm)
+```
+
+Let's look at the 50 state case, with impairment 0.1 and Tolerance 0.2.
+
+```{r}
+d = subset(data, Impairment == 0.05 & Number.of.states == 50 & Tolerance == 0.2)
+colindex = which(names(d) == "Speaker.informativity")
+show(plot.language.MinMaxProperty(d,colindex))
+```
+
+How much identity in outcomes?
+------------------------------
+
+I'm curious how much different properties actually differ within each triplet of dependent variables (tolerance, impairment, number of states).
+We can plot the difference between the maximmal and the minimal value of each property for each triplet of independent variables.
+
+```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+
+dm = melt(subset(ds,Iterations < 205)[,-which(names(ds) == "Iterations")],
+          id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
+
+MinMaxProperty = ddply(dm, .(Number.of.states, Impairment, Tolerance, variable), 
+                      summarise, 
+                      MaxMinDiff = max(value) - min(value),
+                      MinProperty = min(value), MaxProperty = max(value), MeanProperty = mean(value))
+
+pd <- position_dodge(.01)
+sp = ggplot(MinMaxProperty, aes(x=Impairment, y=MaxMinDiff, colour=factor(Number.of.states))) + 
+     geom_point(position = pd)
+
+sp + facet_grid(variable ~ Tolerance, scales="free")
+```
+
+We can then look at the maximal difference among all properties. (Removing Voronoiness here, because it screws up the y-axis dimensions.)
+
+```{r}
+diffTrials = ddply(subset(MinMaxProperty, variable != "Speaker.Voronoiness" & variable != "Hearer.Voronoiness"),
+                   .(Number.of.states, Impairment, Tolerance), 
+                  summarise, 
+                  diffTrials = max(MaxMinDiff) <=1e-5,
+                  maxDiffTrials = max(MaxMinDiff) )
+
+ggplot(diffTrials, aes(x=Impairment, y=maxDiffTrials, )) + 
+     geom_point() + facet_grid(Number.of.states ~ Tolerance, scales="free")
+
+```
+
+We see clearly that, all else equal (e.g.
+fixed tolerance and number of states), impairment 'unifies' evolutionary outcomes.
+The outcomes are all almost identical (in terms of the recorded properties).
+
+But we should check whether this 'uniformity' is different from just convexity vs. non-convexity.
+
+```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+
+d = subset(dsub, dsub$variable =="Speaker.Convex.Cat")
+
+pd <- position_dodge(.01)
+sp = ggplot(d, aes(x=Impairment, y=1-value )) + 
+     geom_point(position = pd)
+
+sp + facet_grid(Number.of.states ~ Tolerance, scales="free")
+
+```
+
+By visual inspection, there are differences in evolutionary outcomes that not due to convexity issues.

### 2014-09-21 data-analysis.R

@@ -6,6 +6,9 @@ theme_set(theme_bw())
 
 plot.language <- function(speaker.filename, hearer.filename) {
 
+    speaker.filename = paste("~/Desktop/data/svn/vagueness-games/", speaker.filename, sep="",col="")
+    hearer.filename = paste("~/Desktop/data/svn/vagueness-games/" , hearer.filename , sep="",col="")
+  
     speaker <- read.csv(speaker.filename, header=FALSE, colClasses='numeric')
     hearer <- read.csv(hearer.filename, header=FALSE, colClasses='numeric')
 

### 2014-09-21 run-batch.sh

@@ -4,7 +4,7 @@ DATE=$(date +%Y%m%d-%H%M%S)
 HEADER="Number of states,Prior distribution type,Number of messages,Impairment,Tolerance,Dynamics,Speaker entropy,Hearer entropy,Speaker convexity,Hearer convexity,Speaker Voronoiness,Hearer Voronoiness,Speaker informativity,Hearer informativity,Expected utility,Iterations,Speaker Convex Cat, Hearer Convex Cat, Speaker strategy file,Hearer strategy file"
 
 echo $HEADER > results/$DATE.csv
-for t in 6 10 50 90
+for t in 90
 do
     echo $t
     for i in 0 0.05 0.1 0.2 0.3

### 2014-09-22 snippets/data_exploration/data_exploration_tmp.Rmd
uniformity of evolutionary outcomes; first script

@@
                              'numeric','numeric','numeric','character','character'))
+
+# no idea why these numbers are messed up, but this fix seems to solve it
+data$Speaker.Voronoiness = ifelse(data$Speaker.Voronoiness > 1, 
+                                  yes = data$Speaker.Voronoiness / 1000, 
+                                  no = data$Speaker.Voronoiness)
+data$Hearer.Voronoiness = ifelse(data$Hearer.Voronoiness > 1, 
+                                  yes = data$Hearer.Voronoiness / 1000, 
+                                  no = data$Hearer.Voronoiness)
 
@@
 
-dss = ds[,c(2,3,7,8,9)]
+dss = ds[c("Impairment", "Speaker.informativity", "Hearer.informativity", "Expected.utility",
+           "Speaker.entropy")]
 
@@
 
-We have two independent variables: impairment & number of states.
-The following plots show the means of all 100 data points for each combination of values for independent variables, together with the estimated confidence intervals.
-
-We can exclude the informativity notions, because they are highly correlated with expected utility scores, and add no extra insight.
+We have three independent variables: impairment, tolerance & number of states.
+The following plots show the means of all 50 data points for each combination of values for independent variables, together with the estimated confidence intervals.
 
@@
 
-We are particularly interested in the interplay between learning speed and EU.
-Generally, the fewer states, the faster we reach convergence (unsurprising).
-But speed to convergence also depends on the impairment value.
-For values 0.2, 0.3 and 0.5 we see a greatly reduced time to convergence.
-However, it is also clear that this increase in speed comes with a decline in expected utility.
-[Notice that the confidence intervals are very tight here, so that we can be sure that there is an almost functional realtionship between convergence speed and EU for each pair of independent parameter values.]
-
-Faster learning by impairment
------------------------------
-
-we want to argue that (small?) levels of impairment can be helpful
-
-unfortunately, impairment and EU are strongly negatively correlated;
-
-however, impairment does seem to speed up learning;
-
-- that's different from Cailin's stuff, because there learning speed could be argued to be only due to more balls being added per learning step
-
-if learning is faster for some impairment, we should see these effects also in terms of higher EU at earlier stages
-- so we should look for stages before final convergence
-
 Informativity
@@
 
-This is the notion from Skyrms. It seems very well-behaved. There is a clear positive correlations between speaker and hearer informativity.
+This is the notion from Skyrms. It seems very well-behaved. There is a clear positive correlation between speaker and hearer informativity.
 
-```{r}
+```{r fig.width=5, fig.height=5}
 cor.test(data$Speaker.informativity, data$Hearer.informativity)
@@
 
-```{r}
+```{r fig.width=5, fig.height=5}
 cor.test(data$Impairment, data$Speaker.informativity)
 qplot(data$Impairment, data$Speaker.informativity) + geom_smooth(method=lm)
-```
-
-Let's look at the 50 state case, with impairment 0.1 and Tolerance 0.2.
-
-```{r}
-d = subset(data, Impairment == 0.05 & Number.of.states == 50 & Tolerance == 0.2)
-colindex = which(names(d) == "Speaker.informativity")
-show(plot.language.MinMaxProperty(d,colindex))
 ```
@@
 
-I'm curious how much different properties actually differ within each triplet of dependent variables (tolerance, impairment, number of states).
+It turns out that there is surprisingly little variance in the recorded values of dependent variables, at least for some triples of independent variables (tolerance, impairment, number of states).
 We can plot the difference between the maximmal and the minimal value of each property for each triplet of independent variables.
@@
 
-We can then look at the maximal difference among all properties. (Removing Voronoiness here, because it screws up the y-axis dimensions.)
+We can then look at the maximal difference among all properties. (This works because all relevant dependent measures scale inbetween 0 and 1.)
 
 ```{r}
-diffTrials = ddply(subset(MinMaxProperty, variable != "Speaker.Voronoiness" & variable != "Hearer.Voronoiness"),
+diffTrials = ddply(MinMaxProperty,
                    .(Number.of.states, Impairment, Tolerance), 
@@
 fixed tolerance and number of states), impairment 'unifies' evolutionary outcomes.
-The outcomes are all almost identical (in terms of the recorded properties).
+When impairment is at least 0.1, outcomes are all almost identical (in terms of the recorded properties).
 
-But we should check whether this 'uniformity' is different from just convexity vs. non-convexity.
+Notice that this 'uniformity' is different from just convexity vs. non-convexity!
 
-```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+```{r message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
 
@@
 
-By visual inspection, there are differences in evolutionary outcomes that not due to convexity issues.
+There are strategies whose speaker strategies are all convex (e.g.
+for ns = 6, tolerance = 0.3 and impairment = 0.05), but whose other properties differ.
+Here are two strategies from that class of independent variable values that differ maximally in terms of expected utility.
+
+```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
+
+d = subset(data, Number.of.states == 6 & Tolerance == 0.3 &
+             Impairment == 0.05 & Speaker.Convex.Cat == 1)
+
+show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))
+
+```
+
+In contrast, if we look at a "uniform" class of independent parameter values, such as with ns = 50, tolerance = 0.1, impairment = 0.1, then the strategies whose EU are minimal and or maximal in that class don't differ optically at all (modulo arbitrariness of message use).
+
+```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
+
+d = subset(data, Number.of.states == 50 & Tolerance == 0.1 &
+             Impairment == 0.1)
+
+show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))
+
+```
+
+What this suggests is that the differences in some of the groups might be due to non-evenly splitting the interval.
+Impairment leads to even splits and therefore to uniformity of evolutionary outcomes.

### 2014-09-22 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -1,0 +1,120 @@
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
+                colClasses=c('numeric','factor','numeric','numeric','numeric',
+                             'factor','numeric','numeric','numeric','numeric',
+                             'numeric','numeric','numeric','numeric','numeric',
+                             'numeric','numeric','numeric','character','character'))
+
+plot.language.byindex = function(dataFrame,rowindex){
+  plot.language(dataFrame[rowindex,]$Speaker.strategy.file,
+                dataFrame[rowindex,]$Hearer.strategy.file)
+}
+
+plot.language.MinMaxProperty = function(dataFrame,colindex){
+  show(names(dataFrame)[colindex])
+  show(summary(dataFrame[,colindex]))
+  p = arrangeGrob(plot.language.byindex(dataFrame,which.min(dataFrame[,colindex])),
+                  plot.language.byindex(dataFrame,which.max(dataFrame[,colindex])))
+}
+
+
+
+# Hellinger distance.
+
+HD = function(x, y){
+  return( sqrt(sum(sapply(1:length(x), function(i) (sqrt(x[i]) - sqrt(y[i]))^2  )) ) / (sqrt(2)) )
+}
+
+
+# Distance between two strategis as Hellinger distances at each choice point.
+
+strategy.distance = function(P,Q){
+  nc = dim(P)[1]
+  sum(sapply(1:nc, function(x) return(HD(P[x,], Q[x,]) ) ) ) / nc 
+}
+
+# Reading a strategy.
+
+read.speaker.strategy = function(dataFrame = data, rowIndex){
+  speaker.filename = paste("~/Desktop/data/svn/vagueness-games/", 
+                           dataFrame[rowIndex,]$Speaker.strategy.file, sep="",col="")
+  speaker <- prop.table(as.matrix(read.csv(speaker.filename, header=FALSE, colClasses='numeric')),1)
+  return(as.matrix(speaker))
+}
+
+read.speaker.strategy.byString = function(String){
+  speaker.filename = paste("~/Desktop/data/svn/vagueness-games/", 
+                           String, sep="",col="")
+  speaker <- prop.table(as.matrix(read.csv(speaker.filename, header=FALSE, colClasses='numeric')),1)
+  return(as.matrix(speaker))
+}
+
[…]

### 2014-09-23 snippets/exploring_uniformity_from_impairment/uniformity.R
in group EU

@@ -1,3 +1,5 @@
+computeD == FALSE
+
 data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
                              'factor','numeric','numeric','numeric','numeric',
@@ -16,14 +18,11 @@ plot.language.MinMaxProperty = function(dataFrame,colindex){
                   plot.language.byindex(dataFrame,which.max(dataFrame[,colindex])))
 }
 
-
-
 # Hellinger distance.
 
 HD = function(x, y){
   return( sqrt(sum(sapply(1:length(x), function(i) (sqrt(x[i]) - sqrt(y[i]))^2  )) ) / (sqrt(2)) )
 }
-
 
 # Distance between two strategis as Hellinger distances at each choice point.
 
@@ -116,5 +115,64 @@ ds = data[,c(1,4,5,19)] # only relevant "data" is speaker strategy file name
 # ds = subset(ds,Number.of.states == 50 & Impairment == 0 & Tolerance <= 0.3)
 dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
 
-D = ddply(dm, .(Number.of.states, Impairment, Tolerance), summarise, 
+if (computeD == TRUE){
+  D = ddply(dm, .(Number.of.states, Impairment, Tolerance), summarise, 
             inGroupDistance = speaker.vector.distance(value))
+}
+else{
+  D = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/D.csv")
+}
+
+pd <- position_dodge(.01)
+sp = ggplot(D, aes(x=Impairment, y=inGroupDistance)) + 
+  geom_point(position = pd) + geom_line()
+sp + facet_grid(Number.of.states ~ Tolerance, scales="free")
+
+# compute average EU between different strategies from the same parameter triple
+
+get.confusion.matrix = function(states,alpha){
+  ns = length(states)
+  if (alpha <= 0){
+    return(diag(rep(1,ns)))
+  }
+  C = matrix(0,nrow=ns, ncol=ns)
+  for (i in 1:ns){
+    for (j in 1:ns){
+      C[i,j] = exp( - (states[i]-states[j])^2 / (alpha^2) )
+    }
+  }
+  return(C)
+}
+
[…]

### 2014-09-24 snippets/data_exploration/data_exploration_tmp.Rmd
uniformity and EU

@@
 source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
+source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
 
@@
 attach(ds)
-
-plot.language.byindex = function(dataFrame,rowindex){
-  plot.language(dataFrame[rowindex,]$Speaker.strategy.file,
-                dataFrame[rowindex,]$Hearer.strategy.file)
-}
-
-plot.language.MinMaxProperty = function(dataFrame,colindex){
-  show(names(dataFrame)[colindex])
-  show(summary(dataFrame[,colindex]))
-  p = arrangeGrob(plot.language.byindex(dataFrame,which.min(dataFrame[,colindex])),
-                  plot.language.byindex(dataFrame,which.max(dataFrame[,colindex])))
-}
 

### 2014-09-24 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -1,22 +1,16 @@
-computeD == FALSE
+computeD = FALSE
+computeEU = TRUE
+
+require('plyr')
+require('reshape2')
+
+source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
 
 data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
                              'factor','numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','character','character'))
-
-plot.language.byindex = function(dataFrame,rowindex){
-  plot.language(dataFrame[rowindex,]$Speaker.strategy.file,
-                dataFrame[rowindex,]$Hearer.strategy.file)
-}
-
-plot.language.MinMaxProperty = function(dataFrame,colindex){
-  show(names(dataFrame)[colindex])
-  show(summary(dataFrame[,colindex]))
-  p = arrangeGrob(plot.language.byindex(dataFrame,which.min(dataFrame[,colindex])),
-                  plot.language.byindex(dataFrame,which.max(dataFrame[,colindex])))
-}
 
 # Hellinger distance.
 
@@ -40,19 +34,20 @@ read.speaker.strategy = function(dataFrame = data, rowIndex){
   return(as.matrix(speaker))
 }
 
-read.speaker.strategy.byString = function(String){
-  speaker.filename = paste("~/Desktop/data/svn/vagueness-games/", 
-                           String, sep="",col="")
-  speaker <- prop.table(as.matrix(read.csv(speaker.filename, header=FALSE, colClasses='numeric')),1)
-  return(as.matrix(speaker))
-}
-
 read.hearer.strategy = function(dataFrame = data, rowIndex){
   hearer.filename = paste("~/Desktop/data/svn/vagueness-games/", 
                            dataFrame[rowIndex,]$Hearer.strategy.file, sep="",col="")
   hearer <- prop.table(as.matrix(read.csv(hearer.filename, header=FALSE, colClasses='numeric')),1)
   return(as.matrix(hearer))
 }
+
+read.strategy.byString = function(String){
[…]

### 2014-09-24 vagueness-games.py

@@ -167,7 +167,7 @@ NMessages = 2
 Dynamics = 'replicator dynamics'
 
 Impairment = 0.2
-Tolerance = 0.025
+Tolerance = 0.05
 
 convThreshold = 0.001
 rounds = 200

### 2014-09-25 snippets/exploring_uniformity_from_impairment/uniformity.R
in Group EU results

@@ -1,5 +1,5 @@
 computeD = FALSE
-computeEU = TRUE
+computeEU = FALSE
 
 require('plyr')
 require('reshape2')
@@ -106,21 +106,21 @@ hearer.strategy.distance = function(data = data, i , j){
 # 
 # # compute distances for speaker strategies
 # 
-# ds = data[,c(1,4,5,19)] # only relevant "data" is speaker strategy file name
-# # ds = subset(ds,Number.of.states == 50 & Impairment == 0 & Tolerance <= 0.3)
-# dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
-# 
-# if (computeD == TRUE){
-#   D = ddply(dm, .(Number.of.states, Impairment, Tolerance), summarise, 
-#             inGroupDistance = speaker.vector.distance(value))
-# } else{
-#   D = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/D.csv")
-# }
-# 
-# pd <- position_dodge(.01)
-# sp = ggplot(D, aes(x=Impairment, y=inGroupDistance)) + 
-#   geom_point(position = pd) + geom_line()
-# sp + facet_grid(Number.of.states ~ Tolerance, scales="free")
+ds = data[,c(1,4,5,19)] # only relevant "data" is speaker strategy file name
+# ds = subset(ds,Number.of.states == 50 & Impairment == 0 & Tolerance <= 0.3)
+dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
+
+if (computeD == TRUE){
+  D = ddply(dm, .(Number.of.states, Impairment, Tolerance), summarise, 
+            inGroupDistance = speaker.vector.distance(value))
+} else{
+  D = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/D.csv")
+}
+
+pd <- position_dodge(.01)
+sp = ggplot(D, aes(x=Impairment, y=inGroupDistance)) + 
+  geom_point(position = pd) + geom_line()
+sp + facet_grid(Number.of.states ~ Tolerance)
 
 # compute average EU between different strategies from the same parameter triple
 
@@ -205,12 +205,12 @@ if (computeEU){
   }
   
   close(pb)
[…]

### 2014-09-25 snippets/data_exploration/data_exploration_tmp.Rmd
restructured snippets

@@
 source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
-
-data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
-                colClasses=c('numeric','factor','numeric','numeric','numeric',
-                             'factor','numeric','numeric','numeric','numeric',
-                             'numeric','numeric','numeric','numeric','numeric',
-                             'numeric','numeric','numeric','character','character'))
-
-# no idea why these numbers are messed up, but this fix seems to solve it
-data$Speaker.Voronoiness = ifelse(data$Speaker.Voronoiness > 1, 
-                                  yes = data$Speaker.Voronoiness / 1000, 
-                                  no = data$Speaker.Voronoiness)
-data$Hearer.Voronoiness = ifelse(data$Hearer.Voronoiness > 1, 
-                                  yes = data$Hearer.Voronoiness / 1000, 
-                                  no = data$Hearer.Voronoiness)
-
-ds = data[,c(1,4,5,7,8,11:17)] # only relevant data
-dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
-
-attach(ds)
 
@@
 ```
-
-How much identity in outcomes?
-------------------------------
-
-It turns out that there is surprisingly little variance in the recorded values of dependent variables, at least for some triples of independent variables (tolerance, impairment, number of states).
-We can plot the difference between the maximmal and the minimal value of each property for each triplet of independent variables.
-
-```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
-
-dm = melt(subset(ds,Iterations < 205)[,-which(names(ds) == "Iterations")],
-          id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
-
-MinMaxProperty = ddply(dm, .(Number.of.states, Impairment, Tolerance, variable), 
-                      summarise, 
-                      MaxMinDiff = max(value) - min(value),
-                      MinProperty = min(value), MaxProperty = max(value), MeanProperty = mean(value))
-
-pd <- position_dodge(.01)
-sp = ggplot(MinMaxProperty, aes(x=Impairment, y=MaxMinDiff, colour=factor(Number.of.states))) + 
-     geom_point(position = pd)
-
-sp + facet_grid(variable ~ Tolerance, scales="free")
-```
-
-We can then look at the maximal difference among all properties. (This works because all relevant dependent measures scale inbetween 0 and 1.)
-
-```{r}
-diffTrials = ddply(MinMaxProperty,
-                   .(Number.of.states, Impairment, Tolerance), 
-                  summarise, 
-                  diffTrials = max(MaxMinDiff) <=1e-5,
-                  maxDiffTrials = max(MaxMinDiff) )
-
-ggplot(diffTrials, aes(x=Impairment, y=maxDiffTrials, )) + 
-     geom_point() + facet_grid(Number.of.states ~ Tolerance, scales="free")
-
-```
-
-We see clearly that, all else equal (e.g.
-fixed tolerance and number of states), impairment 'unifies' evolutionary outcomes.
-When impairment is at least 0.1, outcomes are all almost identical (in terms of the recorded properties).
-
-Notice that this 'uniformity' is different from just convexity vs. non-convexity!
-
-```{r message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
-
-d = subset(dsub, dsub$variable =="Speaker.Convex.Cat")
-
-pd <- position_dodge(.01)
-sp = ggplot(d, aes(x=Impairment, y=1-value )) + 
-     geom_point(position = pd)
-
-sp + facet_grid(Number.of.states ~ Tolerance, scales="free")
-
-```
-
-There are strategies whose speaker strategies are all convex (e.g.
-for ns = 6, tolerance = 0.3 and impairment = 0.05), but whose other properties differ.
-Here are two strategies from that class of independent variable values that differ maximally in terms of expected utility.
-
-```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
-
-d = subset(data, Number.of.states == 6 & Tolerance == 0.3 &
-             Impairment == 0.05 & Speaker.Convex.Cat == 1)
-
-show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))
-
-```
-
-In contrast, if we look at a "uniform" class of independent parameter values, such as with ns = 50, tolerance = 0.1, impairment = 0.1, then the strategies whose EU are minimal and or maximal in that class don't differ optically at all (modulo arbitrariness of message use).
-
-```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
-
-d = subset(data, Number.of.states == 50 & Tolerance == 0.1 &
-             Impairment == 0.1)
-
-show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))
-
-```
-
-What this suggests is that the differences in some of the groups might be due to non-evenly splitting the interval.
-Impairment leads to even splits and therefore to uniformity of evolutionary outcomes.

### 2014-09-25 snippets/exploring_uniformity_from_impairment/uniformity.Rmd
restructured snippets

``` {r warnings=FALSE, message=FALSE, echo=FALSE}

require('ellipse')
require('ggplot2')
require('reshape2')
require('plyr')
source('~/Desktop/data/svn/programming/R/helpers/helpers.R')
source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
source('~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/uniformity.R')

data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
                colClasses=c('numeric','factor','numeric','numeric','numeric',
                             'factor','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','numeric','numeric',
                             'numeric','numeric','numeric','character','character'))

# no idea why these numbers are messed up, but this fix seems to solve it
data$Speaker.Voronoiness = ifelse(data$Speaker.Voronoiness > 1, 
                                  yes = data$Speaker.Voronoiness / 1000, 
                                  no = data$Speaker.Voronoiness)
data$Hearer.Voronoiness = ifelse(data$Hearer.Voronoiness > 1, 
                                 yes = data$Hearer.Voronoiness / 1000, 
                                 no = data$Hearer.Voronoiness)

ds = data[,c(1,4,5,7,8,11:15,17)] # only relevant data
dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
```

Uniformity and diversity in evolutionary outcomes
=================================================

Within each triplet of independent parameter values, how much variation do we see in the evolving lanugages? It turns out that a side-effect of impairment is that it unifies evolutionary outcomes. To see this, we can plot the difference between the maximmal and the minimal value of each property for each triplet of independent variables.

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
MinMaxProperty = ddply(dm, .(Number.of.states, Impairment, Tolerance, variable), 
                      summarise, 
                      MaxMinDiff = max(value) - min(value),
                      MinProperty = min(value), MaxProperty = max(value), MeanProperty = mean(value))

pd <- position_dodge(.01)
sp = ggplot(MinMaxProperty, aes(x=Impairment, y=MaxMinDiff, colour=factor(Number.of.states))) + 
     geom_point(position = pd)

sp + facet_grid(variable ~ Tolerance, scales="free")
```

We can then look at the maximal difference among all properties. (This works because all relevant dependent measures scale inbetween 0 and 1.)

```{r}
diffTrials = ddply(MinMaxProperty,
                   .(Number.of.states, Impairment, Tolerance), 
                  summarise, 
                  diffTrials = max(MaxMinDiff) <=1e-5,
                  maxDiffTrials = max(MaxMinDiff) )

ggplot(diffTrials, aes(x=Impairment, y=maxDiffTrials, )) + 
     geom_point() + facet_grid(Number.of.states ~ Tolerance, scales="free")
```

We see clearly that, all else equal (e.g. fixed tolerance and number of states), impairment 'unifies' evolutionary outcomes. When impairment is at least 0.1, outcomes are all almost identical (in terms of the recorded properties).

Notice that this 'uniformity' is different from just convexity vs. non-convexity!

```{r message=FALSE, warning=FALSE, fig.width=5, fig.height=5}

dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
d = subset(dms, dms$variable == "Speaker.Convex.Cat")

pd <- position_dodge(.01)
sp = ggplot(d, aes(x=Impairment, y=1-value )) + 
     geom_point(position = pd)

sp + facet_grid(Number.of.states ~ Tolerance, scales="free")

```

There are strategies whose speaker strategies are all convex (e.g. for ns = 6, tolerance = 0.3 and impairment = 0.05), but whose other properties differ. Here are two strategies from that class of independent variable values that differ maximally in terms of expected utility.

```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

d = subset(data, Number.of.states == 6 & Tolerance == 0.3 &
             Impairment == 0.05 & Speaker.Convex.Cat == 1)

show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))

```

In contrast, if we look at a "uniform" class of independent parameter values, such as with ns = 50, tolerance = 0.1, impairment = 0.1, then the strategies whose EU are minimal and or maximal in that class don't differ optically at all (modulo arbitrariness of message use).

```{r message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

d = subset(data, Number.of.states == 50 & Tolerance == 0.1 &
             Impairment == 0.1)

show(plot.language.MinMaxProperty(d, which(names(d) == "Expected.utility")))

```

What this suggests is that the differences in some of the groups might be due to non-evenly splitting the interval. Impairment leads to even splits and therefore to uniformity of evolutionary outcomes.

Mean distance between speaker strategies 
----------------------------------------

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
  show(sPlot)
```

Putting it all together 
-----------------------

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}
  show(combPlot)
```

### 2014-09-25 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -216,7 +216,9 @@ meanEUs = ddply(data,
                    .(Number.of.states, Impairment, Tolerance), 
                    summarise, 
                    Speaker.meanEU = mean(Speaker.meanEU),
-                   Hearer.meanEU = mean(Hearer.meanEU))
+                   Hearer.meanEU = mean(Hearer.meanEU),
+                    convexProportion = mean(Speaker.Convex.Cat),
+                    meanEU = mean(Expected.utility))
 
 # mean EUs of speaker and hearer strategies are pretty much perfectly aligned
 qplot(meanEUs$Speaker.meanEU, meanEUs$Hearer.meanEU)
@@ -227,17 +229,19 @@ sPlot = ggplot(meanEUs, aes(x=Impairment, y=Speaker.meanEU)) +
   geom_point() + facet_grid(Number.of.states ~ Tolerance) + geom_line() +
   ylab("average EU against group")
 
-show(sPlot)
+# show(sPlot)
 
 # Combine distance and average Group EU
 
-D$Speaker.meanEU = meanEUs$Speaker.meanEU
+D$withinGroupEU = meanEUs$Speaker.meanEU
+D$individualEU = meanEUs$meanEU
+D$convexProportion = meanEUs$convexProportion
 
 Dmelt = melt(D, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
 
 combPlot = ggplot(Dmelt, aes(x=Impairment, y=value, color = variable)) + 
   geom_point() + facet_grid(Number.of.states ~ Tolerance) + geom_line()
 
-show(sPlot)
+# show(combPlot)
 
 

### 2014-09-26 paper/03_simulations.tex
simulation results write up

@@
 The previous section showed that the \rdd can be conceptualized as a
-two-step procedure. We first calculate where the replicator dynamic
-would take us, and then apply diffusion. Since the latter step is the
+two-step procedure: we first calculate where the replicator dynamic
+would take us, then apply diffusion. Since the latter step is the
 novel addition, we should look at diffusion in isolation to understand
@@
 Section~\ref{sec:iterated-diffusion}. Section~\ref{sec:simulations}
-\dots does something else \dots
+ then explores how diffusion and fitness-based selection interact.
 
@@
 \end{align}
-where $\impairment >= 0$ is an imprecision parameter. When $\impairment=0$ agents
-perfectly discriminate between states; when $\impairment \rightarrow
-\infty$ agents cannot discriminate states at
-all. Figure~\ref{fig:NosofskySim} gives an impression of
-Nosofsky-similarity for different parameter values. Other
-formalizations of perceptual similarity are possible, including ones
-that allow for different discriminability in different areas of the
-state space, but we stick with Nosofsky's similarity function for the
-time being, because it is a mathematically simple, yet established
-notion in mathematical psychology.
+where $\impairment >= 0$ is an imprecision or indiscriminability
+parameter. When $\impairment=0$ agents perfectly discriminate between
+states; when $\impairment \rightarrow \infty$ agents cannot
+discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
+impression of Nosofsky-similarity for different parameter
+values. Other formalizations of perceptual similarity are possible,
+including ones that allow for different discriminability in different
+areas of the state space, but we stick with Nosofsky's similarity
+function for the time being, because it is a mathematically simple,
+yet established notion in mathematical psychology.
 
@@
   \caption{Examples for Nosofsky-similarity (non-normalized) and
-    derived probability of state confusion (normalized) for different
-    values of $\impairment$ and $\card{\States} = 50$.}
+    derived probability of state confusion (normalized) for
+    $\card{\States} = 50$ and different
+    values of $\impairment$.}
   \label{fig:NosofskySim}
@@
 
-The example in Figure~\ref{fig:confusion-SenRec} shows that iterated
-diffusion leads to sender and receiver behavior that is the same at
-all choice points. After $n$ steps of diffusion an initial sender
-strategy $\Sstrat$ will be $C^n \Sstrat$, and an initial receiver
-strategy will be $\Rstrat C^n$. If we assume that $\impairment > 0$ and so
-$C$ is a positive row-stochastic matrix (each state could in principle
-be confused as any other state, albeit with possibly a very low
-probability), it follows from the Perron-Frobenius theorem that the
-limit $C^\infty = \lim_{n \rightarrow \infty} C^n$ exists (it is the
-Perron-projection) and all of its rows are identical. That is why, in
-the limit of iterated diffusion, $C^\infty S$ has identical rows
-(messages are sent with the same probability in each state), and $R
-C^\infty$ likewise has identical rows (every message is interpreted
-equally). All of this is in line with the intuition that diffusion of
-behavior under confusion of states iteratively equalizes behavior for
-similar states; if all states can be confused for one another in
-principle, behavior smoothes out entirely in the limit.
+The example in Figure~\ref{fig:confusion-SenRec} suggests that
+iterated diffusion leads to sender and receiver behavior that is the
+same at all choice points. This holds in general. After $n$ steps of
+diffusion an initial sender strategy $\Sstrat$ will be $C^n \Sstrat$,
+and an initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If
+we assume that $\impairment > 0$, $C$ is a positive row-stochastic
+matrix (each state could in principle be confused as any other state,
+albeit with possibly a very low probability). It then follows from the
+Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
+  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
+all of its rows are identical. That is why, in the limit of iterated
+diffusion, $C^\infty S$ has identical rows (messages are sent with the
+same probability in each state), and $R C^\infty$ likewise has
+identical rows (every message is interpreted equally). All of this is
+in line with the intuition that diffusion of behavior under confusion
+of states iteratively equalizes behavior for similar states; if all
+states can be confused for one another in principle, behavior smoothes
+out entirely in the limit.
 
@@
 
+The main question we should address is whether diffusion and
+fitness-based replication reasonably interact, and if so, whether the
+inclusion of diffusion has any noteworthy effects on the evolving
+meaning of signals. Clearly, if $\impairment = 0$, the \rdd reduces to
+the \rd. If $\impairment \rightarrow \infty$, the diffusion component
+takes over and the \rdd reduces to the trivial iterated diffusion that
+we looked at in the previous section. For reasonable, in-between
+levels of imprecision $\impairment$, the \rdd leads to communicatively
+successful, yet vague signal meaning. Moreover, suitable levels of
+imprecision can have further accelerating and, therefore, unifying
+effects on meaning evolution.
+
+To explore the \rdd, we looked at numerical simulations in sim-max
+games with two signals whose states were evenly spaced elements of the
+unit interval. The state space always included 0 and 1. We assumed
+that priors were flat: $\Pr(\state) = \Pr(\state')$ for all $\state,
+\state'$. As for utility functions, we implemented another
+Nosofsky-style function:
+\begin{align}
+  \label{eq:NosofskyUtil}
+  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
+      \begin{cases}
+    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
+    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) \,,& \textrm{otherwise} \\
+    \end{cases}
+\end{align}
+The tolerance parameter $\toler$ models the amount of tolerable
+pragmatic slack or communicative imprecision. For This choice of
+utility function is governed partly by convenience in parallel to the
+similarity function, but also because it has, as we believe, the right
+general properties for a communicative payoff function. Unlike
+utilities that are, say, linearly or quadratically decreasing
+\citep[as used
+by][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
+utilities that are exponentially decreasing in negative quadratic
+distance can model situations where a small amount of imprecision in
+communication is tolerable, whereas similarly small differences in
+intolerably off interpretations matter very little, with a smooth
+transition between these regimes.
+
+\begin{figure}
+  \centering
+  \subfloat[][$\ns = 90$, $\toler = 0.2$, $\impairment = 0.1$]{
+    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-90_tol-2_imp1_ind3501.pdf}
+  }
+  \subfloat[][$\ns = 90$, $\toler = 0.05$, $\impairment = 0$]{
+    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-90_tol-005_imp0_ind3002.pdf}
+  }
+
+  \subfloat[][$\ns = 10$, $\toler = 0.05$, $\impairment = 0$]{
+    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
+  }
+  \subfloat[][$\ns = 6$, $\toler = 0.05$, $\impairment = 0.05$]{
+    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
+  }
+  \caption{Example strategies under \rdd at stopping time of our simulations.}
+  \label{fig:example_strats}
+\end{figure}
+
+We ran 50 trials of the \rdd, starting with randomly samples sender
+and receiver strategies, for each triplet of independent parameter
+values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
+\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
+  0.3}$. Each trial ran for a maximum of 200 update steps of the
+\rdd. This may not be enough to guarantee convergence to the eventual
+attracting state, but we were interested especially in whether the
+dynamics could lead to reasonable outcomes in reasonable time. A trial
+was considered converged if the total amount of change between
+strategies before and after the current \rdd step was smaller than
+0.001. Examples for the resulting strategy pairs at stopping time are
+given in Figure~\ref{fig:example_strats}.
+
 \begin{itemize}
-\item utility function: another Nosofsky-style function
-\item get data for different set sizes, impairment and utility parameters
-\item 
+\item we see some interesting stuff in these strategies
+\item now we look at the whole bunch
+\item introduce some measures
 \end{itemize}
@@
 %
-\begin{figure}
-        \centering
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140813-194306}
-                \caption{Entropy of sender strategy.}
-%                \label{fig:gull}
-        \end{subfigure}
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140813-194306}
-                \caption{Entropy of receiver strategy.}
-        \end{subfigure}
+% \begin{figure}
+%         \centering
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140813-194306}
+%                 \caption{Entropy of sender strategy.}
+% %                \label{fig:gull}
+%         \end{subfigure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140813-194306}
+%                 \caption{Entropy of receiver strategy.}
+%         \end{subfigure}
 
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140813-194306}
-                \caption{Informativity of sender strategy.}
-        \end{subfigure}
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140813-194306}
-                \caption{Informativity of receiver strategy.}
-        \end{subfigure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140813-194306}
+%                 \caption{Informativity of sender strategy.}
+%         \end{subfigure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140813-194306}
+%                 \caption{Informativity of receiver strategy.}
+%         \end{subfigure}
 
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140813-194306}
-                \caption{Voronoiness of sender strategy.}
-        \end{subfigure}
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140813-194306}
-                \caption{Voronoiness of receiver strategy.}
-        \end{subfigure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140813-194306}
+%                 \caption{Voronoiness of sender strategy.}
+%         \end{subfigure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140813-194306}
+%                 \caption{Voronoiness of receiver strategy.}
+%         \end{subfigure}
 
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Expected-utility-20140813-194306}
-                \caption{Expected utility.}
-        \end{subfigure}
-        \begin{subfigure}{0.45\textwidth}
-                \includegraphics[width=\textwidth]{plots/Iterations-20140813-194306}
-                \caption{Number of iterations until convergence.}
-        \end{subfigure}
-        \caption{Boxplots of metrics per impairment value.}
-        \label{fig:metrics-boxplots}
-\end{figure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Expected-utility-20140813-194306}
+%                 \caption{Expected utility.}
+%         \end{subfigure}
+%         \begin{subfigure}{0.45\textwidth}
+%                 \includegraphics[width=\textwidth]{plots/Iterations-20140813-194306}
+%                 \caption{Number of iterations until convergence.}
+%         \end{subfigure}
+%         \caption{Boxplots of metrics per impairment value.}
+%         \label{fig:metrics-boxplots}
+% \end{figure}
 %

### 2014-09-26 snippets/Nosofsky_similarity.Rmd
simulation results write up

@@
 simDF = data.frame(distance = rep(states,2*length(alphas)),
-                   alpha = factor(rep(rep(alphas, each=ns),2)),
+                   alpha_i = factor(rep(rep(alphas, each=ns),2)),
                    normalized = factor(rep(c("non-normalized", "normalized"),each=ns*length(alphas))),
@@
 
-p = ggplot(simDF, aes(x = distance, y = similarity, color=alpha)) + geom_point() +
-  geom_line() + facet_grid(normalized~. , scales = "free")
+p = ggplot(simDF, aes(x = distance, y = similarity, color=alpha_i)) + geom_point() +
+  geom_line() + facet_grid(normalized~. , scales = "free") + ylab("") + xlab("distance in physical space")
 

### 2014-09-26 snippets/example_strats/example_strats_plots.R

@@ -1,0 +1,21 @@
+require('plyr')
+require('reshape2')
+
+source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
+source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
+
+plot.and.save.language.by.index = function(data, index){
+  ns = data[index,]$Number.of.states
+  tolerance = data[index,]$Tolerance
+  impairment = data[index,]$Impairment
+  stratPlot = plot.language.byindex(data,index)
+  filename = paste("~/Desktop/data/svn/vagueness-games/paper/plots/strat_example_", "NS-", ns,                    "_tol-", tolerance, "_imp", impairment, "_ind", index, ".pdf", sep = "", collapse = "")
+  pdf(file = filename, height = 5, width = 5)
+  show(stratPlot)
+  dev.off()
+}
+
+plot.and.save.language.by.index(data,1201)
+plot.and.save.language.by.index(data,3002)
+plot.and.save.language.by.index(data,3501)
+plot.and.save.language.by.index(data,1001)

### 2014-09-26 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -25,29 +25,6 @@ strategy.distance = function(P,Q){
   sum(sapply(1:nc, function(x) return(HD(P[x,], Q[x,]) ) ) ) / nc 
 }
 
-# Reading a strategy.
-
-read.speaker.strategy = function(dataFrame = data, rowIndex){
-  speaker.filename = paste("~/Desktop/data/svn/vagueness-games/", 
-                           dataFrame[rowIndex,]$Speaker.strategy.file, sep="",col="")
-  speaker <- prop.table(as.matrix(read.csv(speaker.filename, header=FALSE, colClasses='numeric')),1)
-  return(as.matrix(speaker))
-}
-
-read.hearer.strategy = function(dataFrame = data, rowIndex){
-  hearer.filename = paste("~/Desktop/data/svn/vagueness-games/", 
-                           dataFrame[rowIndex,]$Hearer.strategy.file, sep="",col="")
-  hearer <- prop.table(as.matrix(read.csv(hearer.filename, header=FALSE, colClasses='numeric')),1)
-  return(as.matrix(hearer))
-}
-
-read.strategy.byString = function(String){
-  speaker.filename = paste("~/Desktop/data/svn/vagueness-games/", 
-                           String, sep="",col="")
-  speaker <- prop.table(as.matrix(read.csv(speaker.filename, header=FALSE, colClasses='numeric')),1)
-  return(as.matrix(speaker))
-}
-
 
 # Get HD-strategy distance between two strategies.
 

### 2014-09-28 paper/03_simulations.tex
results report

@@
 
-Now that we understand better what diffusion does, let us look at how
-diffusion interacts with optimization of behavior, as described by the
-replicator dynamics.
-
 \subsection{Exploring the replicator diffusion dynamics}
@@
 
-The main question we should address is whether diffusion and
-fitness-based replication reasonably interact, and if so, whether the
-inclusion of diffusion has any noteworthy effects on the evolving
-meaning of signals. Clearly, if $\impairment = 0$, the \rdd reduces to
-the \rd. If $\impairment \rightarrow \infty$, the diffusion component
-takes over and the \rdd reduces to the trivial iterated diffusion that
-we looked at in the previous section. For reasonable, in-between
-levels of imprecision $\impairment$, the \rdd leads to communicatively
-successful, yet vague signal meaning. Moreover, suitable levels of
-imprecision can have further accelerating and, therefore, unifying
-effects on meaning evolution.
+Now that we understand better what diffusion does, let us look at how
+diffusion interacts with optimization of behavior, as described by the
+replicator dynamics. The main question we should address is whether
+diffusion and fitness-based replication reasonably interact, and if
+so, whether the inclusion of diffusion has any noteworthy effects on
+the evolving meaning of signals. Clearly, if $\impairment = 0$, the
+\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
+diffusion component takes over and the \rdd reduces to the trivial
+iterated diffusion that we looked at in the previous section. For
+reasonable in-between levels of imprecision $\impairment$, the \rdd
+leads to communicatively successful, yet vague signal
+meaning. Moreover, suitable levels of imprecision can have further
+accelerating and, therefore, unifying effects on meaning evolution.
 
@@
 games with two signals whose states were evenly spaced elements of the
-unit interval. The state space always included 0 and 1. We assumed
-that priors were flat: $\Pr(\state) = \Pr(\state')$ for all $\state,
-\state'$. As for utility functions, we implemented another
-Nosofsky-style function:
+unit interval that always included 0 and 1. We assumed that priors
+were flat: $\Pr(\state) = \Pr(\state')$ for all $\state, \state'$. As
+for utility functions, we implemented another Nosofsky-style function:
 \begin{align}
@@
 The tolerance parameter $\toler$ models the amount of tolerable
-pragmatic slack or communicative imprecision. For This choice of
-utility function is governed partly by convenience in parallel to the
+pragmatic slack or communicative imprecision. This choice of utility
+function is governed partly by convenience in parallel to the
 similarity function, but also because it has, as we believe, the right
@@
 intolerably off interpretations matter very little, with a smooth
-transition between these regimes.
+transition between these regimes \citep{OConnor2013:The-Evolution-o}.
 
@@
   \centering
-  \subfloat[][$\ns = 90$, $\toler = 0.2$, $\impairment = 0.1$]{
-    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-90_tol-2_imp1_ind3501.pdf}
-  }
-  \subfloat[][$\ns = 90$, $\toler = 0.05$, $\impairment = 0$]{
-    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-90_tol-005_imp0_ind3002.pdf}
-  }
 
-  \subfloat[][$\ns = 10$, $\toler = 0.05$, $\impairment = 0$]{
-    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
-  }
-  \subfloat[][$\ns = 6$, $\toler = 0.05$, $\impairment = 0.05$]{
-    \includegraphics[width = 0.5\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
-  }
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp0_ind3098.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
+    \label{fig:example_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp01_ind3452.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
+    \label{fig:example_stratsB}
+  \end{subfigure}
+
+  % \begin{subfigure}[]{0.45\textwidth}
+  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
+  %   \label{fig:example_stratsC}
+  % \end{subfigure}
+  % \hfill
+  % \begin{subfigure}[]{0.45\textwidth}
+  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
+  %   \label{fig:example_stratsD}
+  % \end{subfigure}
+
   \caption{Example strategies under \rdd at stopping time of our simulations.}
@@
 
-We ran 50 trials of the \rdd, starting with randomly samples sender
+We ran 50 trials of the \rdd, starting with randomly sampled sender
 and receiver strategies, for each triplet of independent parameter
@@
 dynamics could lead to reasonable outcomes in reasonable time. A trial
-was considered converged if the total amount of change between
-strategies before and after the current \rdd step was smaller than
-0.001. Examples for the resulting strategy pairs at stopping time are
-given in Figure~\ref{fig:example_strats}.
+was considered converged and stopped before the maximum of 200 rounds
+if the total amount of change between strategies before and after the
+current \rdd step was smaller than 0.001, i.e., if both:\todo{say what
+the prime means}
+\begin{align*}
+  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
+    \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| <
+  0.001 & \text{\ \ and} &
+  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.001 \,.
+\end{align*}
+Representative examples for the resulting strategy pairs at stopping
+time are given in Figure~\ref{fig:example_strats}.
 
-\begin{itemize}
-\item we see some interesting stuff in these strategies
-\item now we look at the whole bunch
-\item introduce some measures
-\end{itemize}
+Figure~\ref{fig:example_stratsA} shows a strategy pair at stopping
+time with 90 states, tolerance $\toler = 0.1$ and imprecision
+$\impairment = 0$. The latter means that the trial was effectively a
+plain application of the \rd. Noteworthily, the given sender strategy
+approximate a pure sender strategy that crisply partitions the state
+space into non-convex sets. The irregular shape of the receiver
+strategy shows clearly that the pictured strategy pair has not yet
+reached a dynamically stable state under the \rd. Indeed, the trial
+did not converge in our technical sense, but was obtained by stopping
+after 200 rounds of iterated \rdd. In contrast, the outcome of a trial
+with identical parameters, except that imprecision was $\impairment =
+0.1$, is shown in Figure~\ref{fig:example_stratsB} had converged after
+98 rounds of iterated \rdd. The sender strategy shows a smooth
+blending from one ``category'' to the other, and the receiver's
+interpretations are rather extended curves, peaking at a central point
+in the relevant ``categories''. What these examples show is that (i)
+(speaker) strategies can approach non-convex strategies and that (ii)
+inclusion of imprecision can lead to seemingly well-behaved, yet vague
+strategies in the sense that we are after.
 
-\newpage
+To explore further what happened in all of our 50 trials for each
+parameter triple, we defined a small number of metrics that aim to
+capture numerically how vague, communicatively efficient and generally
+well-structured the resulting strategy pairs were.
 
-In order to study the long-term behavior of the noise-perturbed sim-max games, we conducted computer simulations for different combinations of initial parameters.
-Namely, we were interested in investigating the impact of different values of impairment and its relation with the size of the state space.
-\mytodo{JPC}{Justify why size of state space is interesting}
-
-\subsubsection{Metrics}
-\label{sec:metrics}
-To be able to quantitively analyze the results of our simulations, we define a number of metrics to capture different characteristics of sim-max languages.
-
-\paragraph{Entropy.}
-This classic information-theoretic metric captures the amount of uncertainty in a signaling strategy.
-The most natural way to define it is in terms of mixed strategies rather than behavioral strategies.
-Let us recall that mixed sender (receiver) strategies are functions $\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).
-We can define the entropy of a sender strategy as
+\paragraph{Entropy.} This classic information-theoretic metric captures
+the amount of uncertainty in a signaling strategy.  The most natural
+way to define it is in terms of mixed strategies rather than
+behavioral strategies.  Let us recall that mixed sender (receiver)
+strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
+($\Rmixed \in (\Acts^\Messgs)$).  We can define the entropy of a
+sender and receiver strategy as
 \begin{align*}
-  E(\Smixed) = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot \log(\Smixed(\Spure))
+  E(\Smixed) & = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot
+  \log(\Smixed(\Spure)) & 
+  E(\Rmixed) & = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
 \end{align*} 
-and the entropy of a receiver strategy as
+These metrics are computationally expensive to calculate, since the
+size of the domain over which the sum is computed grows exponentially
+with the number of choice points.  Therefore, we
+converted\footnote{See Appendix~\ref{sec:proofs}.}\todo{include this!?} the above
+definitions into equivalent metrics defined in terms of behavioral
+strategies.  These are as follows:
 \begin{align*}
-  E(\Rmixed) = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
-\end{align*} 
-These metrics are computationally expensive to calculate, since the size of the domain over which the sum is computed grows exponentially with the number of choice points.
-Therefore, we converted\footnote{See Appendix~\ref{sec:proofs}.} the above definitions into equivalent metrics defined in terms of behavioral strategies.
-These are as follows:
-\begin{align*}
-  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs} \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar \state))
+  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
+  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
+  \state)) 
 \end{align*} 
 \begin{align*}
-  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar \messg)) \,.
+  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts}
+  \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar
+  \messg)) \,. 
 \end{align*}
-The metrics are lower bounded by $0$ and upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States| \cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot \log(|\Acts|)$, thus we can normalize them by dividing by these values.
+The metrics are lower bounded by $0$ and upper bounded by,
+respectively, $\log(|\Messgs^\States|) = |\States| \cdot
+\log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot
+\log(|\Acts|)$, thus we can normalize them by dividing by these
+values. The reported values are these normalized values.
 
-\paragraph{Informativity.}
-\mytodo{JPC}{Informativity is perhaps a bad name for the receiver...}
-The quantity of information in a signal is an old notion that also goes back to the start of information theory.
-Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of signaling games.
-The main idea is that we can quantify the amount of information about a state $\state$ in a message $\messg$ via the relation between the probability of the state given the message $\Pr(\state \,|\, \messg)$ and the unconditional probability of the state $\Pr(\state)$.
-Following Bayes' theorem, we can express $\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg \,|\, \state)}{\Pr(\messg)}$.
-We then have $\Pr(\messg \,|\, \state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) = \sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)$.
-Finally, we equate sender infomativity with the average overall information about states in each signal.
-Based on the definition by Skyrms~\citet[p.~36]{Skyrms2010:Signals}, and the considerations above, we define sender informativity as follows:
+\paragraph{Informativity.} The quantity of information in a signal is
+an old notion that also goes back to the start of information theory.
+Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the
+context of signaling games.  The main idea is that we can quantify the
+amount of information about a state $\state$ in a message $\messg$ via
+the relation between the probability of the state given the message
+$\Pr(\state \,|\, \messg)$ and the unconditional probability of the
+state $\Pr(\state)$.  Following Bayes' theorem, we can express
+$\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg
+  \,|\, \state)}{\Pr(\messg)}$.  We then have $\Pr(\messg \,|\,
+\state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) =
+\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot
+\Sstrat(\state^\prime,\messg)$.  Finally, we equate sender
+infomativity with the average overall information about states in each
+signal.  Based on the definition by
+Skyrms~\citet[p.~36]{Skyrms2010:Signals}, and the considerations
+above, we define sender informativity as follows:
 \begin{align*}
@@
 We run the noise-perturbed replicator dynamics with random initial strategies $\Sstrat_0$ and $\Rstrat_0$ until a convergence criterion is met, namely that the total absolute change in both sender and receiver strategies does not amount to more than $0.01$, \emph{i.e.}:
-\begin{align*}
-  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left| \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| < 0.01
-\end{align*}
-and
-\begin{align*}
-  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.01 \,.
-\end{align*}
 

### 2014-09-28 snippets/example_strats/example_strats_plots.R

@@ -16,6 +16,6 @@ plot.and.save.language.by.index = function(data, index){
 }
 
 plot.and.save.language.by.index(data,1201)
-plot.and.save.language.by.index(data,3002)
-plot.and.save.language.by.index(data,3501)
 plot.and.save.language.by.index(data,1001)
+plot.and.save.language.by.index(data,3098)
+plot.and.save.language.by.index(data,3452)

### 2014-09-29 paper/03_simulations.tex
results section sec com

@@
 diffusion component takes over and the \rdd reduces to the trivial
-iterated diffusion that we looked at in the previous section. For
-reasonable in-between levels of imprecision $\impairment$, the \rdd
-leads to communicatively successful, yet vague signal
-meaning. Moreover, suitable levels of imprecision can have further
-accelerating and, therefore, unifying effects on meaning evolution.
+iterated diffusion that we looked at in the previous section. We will
+show presently that for reasonable in-between levels of imprecision
+$\impairment$, the \rdd leads to communicatively successful, yet vague
+signal meaning. Moreover, suitable levels of imprecision can have
+further accelerating and, therefore, unifying effects on meaning
+evolution.
 
@@
 \end{align}
-The tolerance parameter $\toler$ models the amount of tolerable
+The tolerance parameter $\toler > 0$ models the amount of tolerable
 pragmatic slack or communicative imprecision. This choice of utility
@@
 general properties for a communicative payoff function. Unlike
-utilities that are, say, linearly or quadratically decreasing
-\citep[as used
+utilities that are, say, linearly or quadratically decreasing in
+physical distance \citep[as used
 by][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
@@
 communication is tolerable, whereas similarly small differences in
-intolerably off interpretations matter very little, with a smooth
-transition between these regimes \citep{OConnor2013:The-Evolution-o}.
+intolerably far away interpretations matter very little, with a smooth
+transition between these regimes
+\citep[c.f.][]{OConnor2013:The-Evolution-o}.
 
@@
 if the total amount of change between strategies before and after the
-current \rdd step was smaller than 0.001, i.e., if both:\todo{say what
-the prime means}
+current \rdd step was smaller than 0.001, i.e., if:
 \begin{align*}
   \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
-    \Sstrat^\prime(\state,\messg) - \Sstrat(\state,\messg) \right| <
-  0.001 & \text{\ \ and} &
-  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left| \Rstrat^\prime(\messg,\state) - \Rstrat(\messg,\state) \right| < 0.001 \,.
+    \rdd(\Sstrat)(\state,\messg) - \Sstrat(\state,\messg) \right| <
+  0.001\,, & \text{\ \ and} \\
+  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left|
+    \rdd(\Rstrat)(\messg,\state) - \Rstrat(\messg,\state) \right| <
+  0.001 & \,.
 \end{align*}
-Representative examples for the resulting strategy pairs at stopping
-time are given in Figure~\ref{fig:example_strats}.
 
-Figure~\ref{fig:example_stratsA} shows a strategy pair at stopping
-time with 90 states, tolerance $\toler = 0.1$ and imprecision
-$\impairment = 0$. The latter means that the trial was effectively a
-plain application of the \rd. Noteworthily, the given sender strategy
-approximate a pure sender strategy that crisply partitions the state
-space into non-convex sets. The irregular shape of the receiver
-strategy shows clearly that the pictured strategy pair has not yet
-reached a dynamically stable state under the \rd. Indeed, the trial
-did not converge in our technical sense, but was obtained by stopping
-after 200 rounds of iterated \rdd. In contrast, the outcome of a trial
-with identical parameters, except that imprecision was $\impairment =
-0.1$, is shown in Figure~\ref{fig:example_stratsB} had converged after
-98 rounds of iterated \rdd. The sender strategy shows a smooth
-blending from one ``category'' to the other, and the receiver's
+Representative examples for resulting strategy pairs at stopping time
+are given in
+Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
+shows a strategy pair at stopping time with 90 states, tolerance
+$\toler = 0.1$ and imprecision $\impairment = 0$. The latter means
+that the trial was effectively a plain application of the
+\rd. Noteworthily, the given sender strategy approximates a pure
+sender strategy that crisply partitions the state space into
+non-convex sets. The irregular shape of the receiver strategy shows
+clearly that the pictured strategy pair has not yet reached a
+dynamically stable state under the \rd. Indeed, the trial did not
+converge in our technical sense, but was stopped after 200 rounds of
+iterated \rdd. In contrast, the outcome of a trial with identical
+parameters, except that imprecision was $\impairment = 0.1$, is shown
+in Figure~\ref{fig:example_stratsB} had converged (in our technical
+sense) after 98 rounds of iterated \rdd. The sender strategy shows a
+smooth blending from one ``category'' to the other, and the receiver's
 interpretations are rather extended curves, peaking at a central point
@@
 capture numerically how vague, communicatively efficient and generally
-well-structured the resulting strategy pairs were.
+well-structured the resulting strategy pairs were. Entropy and
+informativity capture (slightly different aspects of) the amount of
+systematicity or regularity in a language. Voronoiness and convexity
+are, respectively, gradient and categorical measures for the coherence
+of linguistically formed categories.
 
-\paragraph{Entropy.} This classic information-theoretic metric captures
-the amount of uncertainty in a signaling strategy.  The most natural
-way to define it is in terms of mixed strategies rather than
-behavioral strategies.  Let us recall that mixed sender (receiver)
-strategies are functions $\Smixed \in \Delta(\Messgs^\States)$
-($\Rmixed \in (\Acts^\Messgs)$).  We can define the entropy of a
-sender and receiver strategy as
+\paragraph{Entropy.} This classic information-theoretic metric
+captures the amount of uncertainty in a signaling strategy. Roughly
+put, entropy of a strategy captures inverse distance from a pure
+strategy in the present context. Vague languages will have higher
+entropy than crisp ones.
+
+The most natural way to define it is in terms of mixed strategies
+rather than behavioral strategies.  Let us recall that mixed sender
+(receiver) strategies are functions $\Smixed \in
+\Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).  We can
+define the entropy of a sender and receiver strategy as
 \begin{align*}
@@
 \end{align*} 
-These metrics are computationally expensive to calculate, since the
+These measures are computationally expensive to calculate, since the
 size of the domain over which the sum is computed grows exponentially
 with the number of choice points.  Therefore, we
-converted\footnote{See Appendix~\ref{sec:proofs}.}\todo{include this!?} the above
-definitions into equivalent metrics defined in terms of behavioral
-strategies.  These are as follows:
+converted\footnote{See Appendix~\ref{sec:proofs}.}\todo{include
+  this!?} the above definitions into equivalent metrics defined in
+terms of behavioral strategies.  These are as follows:
 \begin{align*}
@@
 \end{align*}
-The metrics are lower bounded by $0$ and upper bounded by,
-respectively, $\log(|\Messgs^\States|) = |\States| \cdot
-\log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot
-\log(|\Acts|)$, thus we can normalize them by dividing by these
-values. The reported values are these normalized values.
+Values obtained by these definitions are lower bounded by $0$ and
+upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
+\cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot
+\log(|\Acts|)$. We work with normalized values.
+
+The speaker strategies in Figures~\ref{fig:example_stratsA} and
+\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
+respectively. The receiver strategies have respective entropies $0.43$
+and $0.84$.
 
@@
 Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the
-context of signaling games.  The main idea is that we can quantify the
-amount of information about a state $\state$ in a message $\messg$ via
-the relation between the probability of the state given the message
+context of signaling games. Intuitively speaking, informativity in
+Skyrms' sense measures how revealing the message use is about which
+state is actual. It therefore measures, in our given set-up with flat
+priors, a sense of well-behavedness or regularity of a sender
+strategy. Vague languages will have lower informativity than crisp
+ones.
+
+The main idea behind Skyrm's notion is that we can quantify the amount
+of information about a state $\state$ in a message $\messg$ via the
+relation between the probability of the state given the message
 $\Pr(\state \,|\, \messg)$ and the unconditional probability of the
@@
 infomativity with the average overall information about states in each
-signal.  Based on the definition by
-Skyrms~\citet[p.~36]{Skyrms2010:Signals}, and the considerations
-above, we define sender informativity as follows:
+signal.  Following \citet[p.~36]{Skyrms2010:Signals}, and the
+considerations above, the informativity about states conveyed by a
+sender strategy is:
 \begin{align*}
-  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\state \in \States} \frac{\Pr(\state) \cdot \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right) \,.
+  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs}
+  \sum_{\state \in \States} \frac{\Pr(\state) \cdot
+    \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
+    \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log
+  \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
+      \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right)
+  \,. 
 \end{align*}
-Conversely, we can quantify the amount of information about an act in a message.
-We equate receiver informativity with the average overall information about acts in each signal.
-Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
-\begin{align*}
-  I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
-\end{align*}
-Both metrics are bounded between $0$ and $1$.
+This measure is bounded between $0$ and $1$. The speaker strategies in
+Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
+informativity $0.69$ and $0.54$, respectively.
+
+\citet[p.~39]{Skyrms2010:Signals} also offers a notion of
+informativity about the acts performed by the receiver, but it turns
+out that this is perfectly correlated with information about the
+states in the results of our simulations. We therefore only look at
+informativity about states.
+
+% Conversely, we can quantify the amount of information about an act in a message.
+% We equate receiver informativity with the average overall information about acts in each signal.
+% Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
+% \begin{align*}
+%   I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
+% \end{align*}
 
 \paragraph{Voronoiness.}
-This metric aims to quantify how close a strategy is to being a part of a Voronoi tessallation of the state space.
-Recalling the results by \cite{Jager2007}, the stable outcomes of a population of agents playing a similarity-maximization game using pure strategies and evolving according to the replicator dynamics are those where ``the sender strategy is consistent with the Voronoi tessallation that is induced by the image of the receiver strategy.''~(\cite{Jager2007}, p.
-562).
-\mytodo{JPC}{Explain further?}
-Given we are working with behavioral strategies, rather than pure strategies, and that we perform numerical simulations, rather than calculate analytical results, instead of looking into a binary criterion of whether a simulation result is such a Voronoi tessallation of the state space or not, we want a more progressive metric to characterize that in terms of degrees.
-With that in mind, and given that this metric only makes sense for $\Acts = \States$, we define Voronoiness as follows.
-Let $\pi(\Rstrat, \messg) = \argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype of a message $\messg \in \Messgs$, then
+This metric aims to quantify how close a strategy is to being a part
+of a Voronoi tessallation of the state space. Bearing in mind that
+this metric is specifically geared towards the sim-max games in our
+set-up, we define Voronoiness as follows. Let $\pi(\Rstrat, \messg) =
+\argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype
+of a message $\messg \in \Messgs$, then
 \begin{multline*}
-  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state, \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then } 1 \textrm{ else } 0
+  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in
+    \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state,
+  \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond
+  \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then
+  } 1 \textrm{ else } 0 
 \end{multline*}
-is the Voronoiness of the sender strategy $\Sstrat$ given the receiver strategy $\Rstrat$ and
+is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
+strategy $\Rstrat$ and
 \begin{align*}
-  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States} \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat, \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg, \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{ else } 0
+  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States}
+  \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat,
+  \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg,
+  \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{
+    else } 0
 \end{align*}
-is the Voronoiness of the receiver strategy $\Rstrat$.
-\mytodo{JPC}{Give some intuitions on these?}
-The metrics are lower bounded by $0$ and upper bounded by, respectively, $|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we can normalize them by dividing by these values.
+is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
+are lower bounded by $0$ and upper bounded by, respectively,
+$|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$,
+thus we can normalize them by dividing by these values.
 
-\subsubsection{Methodology}
+Despite being vague, a language can have maximal Voronoiness according
+to this gradient measure. For instance, the sender strategies in
+Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
+Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
+Voronoiness $0.78$ and $1$, respectively.
 
-In our simulations we focus on a concrete setup in line with \cite{Correia2013:The-Bivalent-Tr}.
-Namely, we define a similarity-maximization game with $\States = \Acts$, where $\States$ is a discrete set of equidistant points within the unit interval $[0,1]$, including the boundaries (thus we always use $n \geq 2$).
-As similarity metric $s:\States \times \States \rightarrow [0,1]$, we use a negative exponential of the of the square distance, \emph{i.e.}~for $\state,\state^\prime \in \States$:
+\paragraph{Convexity.} At least for speaker strategies, which converge
+faster, it also makes sense to define a categorical measure of
+convexity, that compensates for potential vagueness. To determine
+whether a sender strategy $\Sstrat$ was convex despite possibly being
+vague, we looked at the derived pure strategy $\Spure$ for which
+$\Spure(\state) = \messg$ only if $\Sstrat(\state,\messg) =
+\max_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$
+was convex, we also counted $\Sstrat$ as convex.
+
+\paragraph{Expected utility.} We also recorded the expected utility of
+a strategy pair:
 \begin{align*}
-  s(\state,\state^\prime) =
-    \begin{cases}
-    1 & \textrm{if } \impairment = 0 \textrm{ and } \state = \state^\prime \\
-    0 & \textrm{if } \impairment = 0 \textrm{ and } \state \neq \state^\prime \\
-    e^{-\frac{\lVert \state - \state^\prime \rVert ^2}{\impairment^2}} & \textrm{otherwise} \\
-    \end{cases}
+  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
+    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
+  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
+  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
 \end{align*}
-where $\impairment$ is an \emph{impairment} factor.
-Here we deviate from \cite{Correia2013:The-Bivalent-Tr} by working with impairment $\impairment$ rather than acuity $\theta$, since we are interested in exploring the case of having $0$ impairment, and acuity does not have a natural maximum as defined.
-Confusion is directly related with similarity, \emph{i.e.}~the more similar two states are, the more likely an agent is to perceive one as the other.
-For simplicity, we equate one with the other by defining $C_{ij} = s(\state_i,\state_j)$.
-
-We run the noise-perturbed replicator dynamics with random initial strategies $\Sstrat_0$ and $\Rstrat_0$ until a convergence criterion is met, namely that the total absolute change in both sender and receiver strategies does not amount to more than $0.01$, \emph{i.e.}:
-
-We perform a parameter sweep with size of the state space $n \in \{10,
-50, 90, 130\}$ and impairment $\impairment \in \{0, 0.1, 0.2, 0.3,
-0.4, 0.5\}$.
-For each parameter combination, we perform $100$ trials and record the values of each of the metrics introduced in Section~\ref{sec:metrics} after convergence, as well as two additional indicators: number of iterations taken until convergence, and the normalized expected utility.
-Expected utility is normalized by dividing by the expected utility of the theoretical optimum without impairment, \emph{i.e.}~a Voronoi tessalation of the state space.
+To make direct comparisons across different parameter settings, we
+normalized expected utility by the maximal amount of expected utility
+obtainable in the relevant game.
 
 \subsubsection{Results}
-In order to visualize the impact of the impairment factor on the outcome of simulations, we plot in Figure~\ref{fig:metrics-boxplots} the distribution of each of the measurements per impairment value.
-%
-% \begin{figure}
-%         \centering
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Speaker-entropy-20140813-194306}
-%                 \caption{Entropy of sender strategy.}
-% %                \label{fig:gull}
-%         \end{subfigure}
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Hearer-entropy-20140813-194306}
-%                 \caption{Entropy of receiver strategy.}
-%         \end{subfigure}
+Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
+together with their estimated 95\% confidence intervals, for each
+triple of independent variables. Most trends are quite clear. Entropy
+of sender and receiver strategies are increasing in imprecision
+$\impairment$, as can be expected. Entropy of speaker strategies
+hardly seems independent of tolerance levels and the number of states,
+while there seems to be an effect of both parameters on the entropy of
+the hearer strategy.
 
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Speaker-informativity-20140813-194306}
-%                 \caption{Informativity of sender strategy.}
-%         \end{subfigure}
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Hearer-informativity-20140813-194306}
-%                 \caption{Informativity of receiver strategy.}
-%         \end{subfigure}
+\begin{figure}
+  \centering
 
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Speaker-Voronoiness-20140813-194306}
-%                 \caption{Voronoiness of sender strategy.}
-%         \end{subfigure}
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Hearer-Voronoiness-20140813-194306}
-%                 \caption{Voronoiness of receiver strategy.}
-%         \end{subfigure}
+  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}
 
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Expected-utility-20140813-194306}
-%                 \caption{Expected utility.}
-%         \end{subfigure}
-%         \begin{subfigure}{0.45\textwidth}
-%                 \includegraphics[width=\textwidth]{plots/Iterations-20140813-194306}
-%                 \caption{Number of iterations until convergence.}
-%         \end{subfigure}
-%         \caption{Boxplots of metrics per impairment value.}
-%         \label{fig:metrics-boxplots}
-% \end{figure}
-%
-Most trends are quite clear.
-Entropy seems to increase with higher impairment values for both sender and receiver.
-This is an expected outcome, since the higher the impairment value, the more likely each player is to confuse two states, thus the more uncertain he will be in unequivocally associating a state with a message, or vice-versa.
-The trend for informativity is also expected: more vague strategies convey less information, be it from the sender or the receiver perspective.
+  \caption{Means of recorded metrics for each triple of independent variables.}
+  \label{fig:MeanMetrics}
+\end{figure}
 
-Regarding Voronoiness, we observe that most simulation results approximate perfect Voronoi tessalations.
-This is not the case, however, for impairment of $0$ (both sender and receiver) and $0.5$ (mostly sender only).
-\mytodo{JPC}{For impairment 0, the languages with the lowest sender voronoiness look like non-convex equilibria. Discuss in relation with Elliott's results?}
-For impairment of $0$,  the explanation is usually that strategies can achieve a high expected utility without having perfect Voronoiness.
-Although the expected long term outcome in these situations is perfect Voronoiness (given J\"ager's analytical results~\cite{Jager2007}), actually reaching it involves minor refinements which at some point do not qualify as significant change given our convergence criterion.
-The result is thus an artifact of stopping the simulations before they reach perfect convergence.
-For impairment of $0.5$ (and some cases with impairment of $0.4$), the reason is simply that the languages are representatives of something between babbling and pooling equilibria.
-These are equilibria where $\Sstrat(\state,\messg_1) \geq \Sstrat(\state,\messg_2)$ for all $\state \in \States$, but vary from situations where $\Sstrat(\state,\messg)$ is close to $0.5$ for both messages (babbling), or is significantly higher for one of them (pooling).
-\mytodo{JPC}{Plot examples}
+Speaker informativity and expected utility are decreasing under
+growing imprecision, again, as should be expected. There seem to be
+mild effects of the number of states in the game, but hardly any in
+terms of pragmatic tolerance.
 
-The trend for expected utility reflects Lipman's observation that vague languages are less optimal than non-vague ones~\cite{Lipman2009:Why-is-Language}: more impairment necessarily leads to lower expected utility.
-Finally, the plot for number of iterations until convergence shows that a certain level of impairment enables faster convergence.
-We will discuss this further in more detail in Section~\ref{sec:discussion}.
+Interestingly, voronoiness of both sender and receiver strategies are
+appear to be non-decreasing under increasing impairment and peaking at
+the maximal value. The only exception is the 6 states games and the
+transition from $\impairment = 0$ to $\impairment = 0.05$. This
+suggests that impairment can help evolve more regular, well-behaved
+languages.
 
-\mytodo{JPC}{Present and discuss results per size of state space? Don't forget O'Connor's claim that speed of convergence is more important for larger \#state/\#message ratio}
+This impression is also supported by the proportion of converged and
+convex trials, pictured in
+Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
+speaker strategies showed for several values of parameters in the
+absence of impairment. Increasing impairment always assured
+convergence and convexity.
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\textwidth]{plots/CategoricalMeasures.pdf}
+
+  \caption{Proportion of converged and convex trials for each triple
+    of independent variables.}
+  \label{fig:CategoricalMeasures}
+\end{figure}
+
+Summarizing so far, it seems that diffusion increases entropy and
+vagueness, decreases informativity and expected utility, but it also
+speeds up convergence to a convex and well-behaved outcome.
+
+Diffusion of strategies from confusion of states has another
+interesting regularizing effect on the evolution of signaling. If we
+look at the confidence intervals in Figure~\ref{fig:MeanMetrics}, we
+see that there is very little variation in the recorded metrics for
+evolved strategies, at least for higher values of impairment. On
+closer inspection, it turns out that variability in low-impairment
+conditions is not only due to non-convergence and/or associated
+non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more
+examples of strategy pairs at stopping time. Both are obtained for the
+same triple of parameters, both converged and are convex. However,
+they are not equally efficient. In fact, the pair in
+Figure~\ref{fig:example_stratsC} has a normalized expected utility of
+$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
+$0.89$.
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind41.pdf}
+    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
+    \label{fig:example_stratsC}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind23.pdf}
+    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
+    \label{fig:example_stratsD}
+  \end{subfigure}
+
+  \caption{More example strategies under \rdd at stopping time of our simulations.}
+  \label{fig:MoreExample_strats}
+\end{figure}
+
+This type of variability in evolutionary outcomes is also weeded out
+by mild increases in impairment.
+
+\begin{figure}
+  \centering
+
+    \includegraphics[width=\textwidth]{plots/AverageEUinGroups.pdf}
+
+   \caption{Average normalized expected utility between evolved strategies within each triplet of independent parameter values.}
+  \label{fig:AverageEUinGroups}
+\end{figure}
 

### 2014-09-29 snippets/data_exploration/data_exploration_tmp.Rmd
results section sec com

@@
 
-pairs(~ Number.of.states + Impairment + Tolerance + Speaker.entropy + Hearer.entropy + 
-        Speaker.Voronoiness + Hearer.Voronoiness +
-        Speaker.informativity + Hearer.informativity + Expected.utility + Iterations +
+pairs(~ Speaker.entropy + Hearer.entropy + Speaker.Voronoiness + Hearer.Voronoiness +
+        Speaker.informativity + Hearer.informativity + Expected.utility +
         Speaker.Convex.Cat, data=ds)
@@
 
+ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
+             "Speaker.entropy", "Hearer.entropy",
+             "Speaker.informativity",
+             "Speaker.Voronoiness", "Hearer.Voronoiness",
+             "Expected.utility")]
+
+dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
+
 dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
@@
 
-sp + facet_grid(variable ~ Tolerance, scales="free")
+sp = sp + facet_grid(variable ~ Tolerance, scales="free")
 
+show(sp)
+
+pdf("~/Desktop/data/svn/vagueness-games/paper/plots/MeanMetrics.pdf", width=8, height = 8)
+show(sp)
+dev.off()
+
+```
+
+Entropy
+-------
+
+```{r fig.width=5, fig.height=5}
+cor.test(data$Speaker.entropy, data$Hearer.entropy)
+qplot(data$Speaker.entropy, data$Hearer.entropy) + geom_smooth(method=lm)
 ```
@@
 ```
+
+Voronoiness
+-------
+
+```{r fig.width=5, fig.height=5}
+cor.test(data$Speaker.Voronoiness, data$Hearer.Voronoiness)
+qplot(data$Speaker.Voronoiness, data$Hearer.Voronoiness) + geom_smooth(method=lm)
+```
+
+Percentage of Converged and Convex strategies
+---------------------------------------------
+
+```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+
+data$Converged = data$Iterations < 201
+data$Convex = data$Speaker.Convex.Cat
+
+ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
+             "Converged", "Convex")]
+
+dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
+
+dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
+dsub = dms
+
+sp = ggplot(dms, aes(x=factor(Impairment), y = value)) 
+sp = sp + geom_bar(aes(fill = variable), color = "black", stat="identity", position= position_dodge(0.4), width = 0.3)
+sp = sp + facet_grid(Number.of.states ~ Tolerance, scales="free")
+
+show(sp)
+
+pdf("~/Desktop/data/svn/vagueness-games/paper/plots/CategoricalMeasures.pdf", width=8, height = 6)
+show(sp)
+dev.off()
+
+```

### 2014-09-29 snippets/exploring_uniformity_from_impairment/uniformity.Rmd
results section sec com

@@
   show(sPlot)
+
+pdf("~/Desktop/data/svn/vagueness-games/paper/plots/AverageEUinGroups.pdf", height = 6, width = 8)
+show(sPlot)
+dev.off()
+
 ```

### 2014-09-29 snippets/example_strats/example_strats_plots.R

@@ -15,7 +15,18 @@ plot.and.save.language.by.index = function(data, index){
   dev.off()
 }
 
-plot.and.save.language.by.index(data,1201)
-plot.and.save.language.by.index(data,1001)
-plot.and.save.language.by.index(data,3098)
-plot.and.save.language.by.index(data,3452)
+stratsToPlot = c(1201, 1001, 3098, 3452)
+
+# for (strat in stratsToPlot){
+#   plot.and.save.language.by.index(data,strat)
+# }
+
+
+show(ds[stratsToPlot,])
+
+d = subset(data, Number.of.states == 6 & Tolerance == 0.3 & Impairment == 0.05 & Speaker.Convex.Cat == 1)
+
+
+plot.and.save.language.by.index(d, which.min(d$Expected.utility))
+plot.and.save.language.by.index(d, which.max(d$Expected.utility))
+d[c(which.min(d$Expected.utility), which.max(d$Expected.utility)),]

### 2014-09-30 paper/03_simulations.tex
results write-up sec com

@@
 $\impairment$, the \rdd leads to communicatively successful, yet vague
-signal meaning. Moreover, suitable levels of imprecision can have
-further accelerating and, therefore, unifying effects on meaning
-evolution.
+signal meaning. Suitable levels of imprecision can have further
+accelerating and unifying effects on meaning evolution.
+
+\subsubsection{Experimental set-up}
 
@@
 in Figure~\ref{fig:example_stratsB} had converged (in our technical
-sense) after 98 rounds of iterated \rdd. The sender strategy shows a
+sense) after 99 rounds of iterated \rdd. The sender strategy shows a
 smooth blending from one ``category'' to the other, and the receiver's
 interpretations are rather extended curves, peaking at a central point
-in the relevant ``categories''. What these examples show is that (i)
-(speaker) strategies can approach non-convex strategies and that (ii)
-inclusion of imprecision can lead to seemingly well-behaved, yet vague
-strategies in the sense that we are after.
+in the relevant ``categories''. What these examples show is that (i),
+contrary to the limiting results of
+\citet{Jager2007:The-Evolution-o,JagerMetzger2011:Voronoi-Languag},
+(speaker) strategies can approach non-convex pure strategies and that
+(ii) inclusion of imprecision can lead to seemingly well-behaved, yet
+vague strategies in the sense that we are after.
+
+\subsubsection{Dependent measures}
 
@@
 are, respectively, gradient and categorical measures for the coherence
-of linguistically formed categories.
+of linguistically formed categories. Expected utility measures the
+communicative efficiency of evolved strategy pairs.
 
@@
 put, entropy of a strategy captures inverse distance from a pure
-strategy in the present context. Vague languages will have higher
-entropy than crisp ones.
-
-The most natural way to define it is in terms of mixed strategies
-rather than behavioral strategies.  Let us recall that mixed sender
-(receiver) strategies are functions $\Smixed \in
+strategy. The most natural way to define it is in terms of mixed
+strategies rather than behavioral strategies.  Let us recall that
+mixed sender (receiver) strategies are functions $\Smixed \in
 \Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).  We can
@@
 respectively. The receiver strategies have respective entropies $0.43$
-and $0.84$.
+and $0.84$. In general, we expect that vague languages will have
+higher entropy than crisp ones and that increasing imprecision will
+also increase entropy, all else being equal.
 
@@
 an old notion that also goes back to the start of information theory.
-Skyrms~\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the
-context of signaling games. Intuitively speaking, informativity in
-Skyrms' sense measures how revealing the message use is about which
-state is actual. It therefore measures, in our given set-up with flat
-priors, a sense of well-behavedness or regularity of a sender
-strategy. Vague languages will have lower informativity than crisp
-ones.
+\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of
+signaling games. Intuitively speaking, informativity in Skyrms' sense
+measures how revealing the sender's signal use is about which state is
+actual. It therefore measures, in our given set-up with flat priors, a
+sense of well-behavedness or regularity of a sender strategy.
 
@@
 Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
-informativity $0.69$ and $0.54$, respectively.
+informativity $0.69$ and $0.54$, respectively. In general, we expect
+that vague languages will have lower informativity than crisp
+ones and that increasing imprecision will decrease informativity.
 
@@
 Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
-Voronoiness $0.78$ and $1$, respectively.
+Voronoiness $0.78$ and $1$, respectively. We do not \emph{prima facie}
+expect a conceptual connection between vagueness and Voronoiness, and
+so do not predict a clear effect of imprecision on Voronoiness either.
 
@@
 vague, we looked at the derived pure strategy $\Spure$ for which
-$\Spure(\state) = \messg$ only if $\Sstrat(\state,\messg) =
-\max_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$
-was convex, we also counted $\Sstrat$ as convex.
+$\Spure(\state) = \messg$ only if $\messg =
+\argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$
+was convex, we also counted $\Sstrat$ as convex. The sender strategy
+in Figure~\ref{fig:example_stratsA} is not convex, while the one in
+Figure~\ref{fig:example_stratsB} is. Again, we do not expect a
+conceptual relation between vagueness and imprecision, on the one
+hand, and convexity on the other.
 
@@
 together with their estimated 95\% confidence intervals, for each
-triple of independent variables. Most trends are quite clear. Entropy
-of sender and receiver strategies are increasing in imprecision
-$\impairment$, as can be expected. Entropy of speaker strategies
-hardly seems independent of tolerance levels and the number of states,
-while there seems to be an effect of both parameters on the entropy of
-the hearer strategy.
+triple of independent variables. Most trends are quite clear, and not
+unexpected. Entropy of sender and receiver strategies are increasing
+in imprecision. Entropy of speaker strategies seems independent of
+tolerance levels and the number of states, while there seems to be an
+effect of both parameters on the entropy of the hearer
+strategy. Speaker informativity and expected utility are decreasing
+under growing imprecision. There seem to be mild effects of the number
+of states in the game, but hardly any in terms of pragmatic
+tolerance. Interestingly, Voronoiness of both sender and receiver
+strategies appear to be non-decreasing under increasing impairment and
+peaking at the maximal value. The only exception is the 6 states games
+and the transition from $\impairment = 0$ to $\impairment =
+0.05$. Expected utility of evolved languages mostly decreases in
+increasing imprecision, with the sole exception of a rise for $\toler
+= 0.2$ and $ns = 6$ at $\impairment = 0.1$. Still, normalized expected
+utility of evolved strategy pairs under mild values of imprecision are
+not necessarily much worse than their respective perfect precision
+\rd-analogues.
 
@@
 
-Speaker informativity and expected utility are decreasing under
-growing imprecision, again, as should be expected. There seem to be
-mild effects of the number of states in the game, but hardly any in
-terms of pragmatic tolerance.
-
-Interestingly, voronoiness of both sender and receiver strategies are
-appear to be non-decreasing under increasing impairment and peaking at
-the maximal value. The only exception is the 6 states games and the
-transition from $\impairment = 0$ to $\impairment = 0.05$. This
-suggests that impairment can help evolve more regular, well-behaved
-languages.
-
-This impression is also supported by the proportion of converged and
-convex trials, pictured in
+Taken together, these findings suggest the following effect of
+imprecise perception and diffusion of behavior. Mild forms of
+perceptual imprecision lead to slightly less communicative efficiency
+in the evolving strategy pairs, more vagueness and more regular,
+well-behaved languages. This impression is also supported by the
+proportion of converged and convex trials, pictured in
 Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
@@
 absence of impairment. Increasing impairment always assured
-convergence and convexity.
+convergence and convexity: perceptual imprecision speeds up
+convergence to a convex and regular language.
 
@@
 \end{figure}
-
-Summarizing so far, it seems that diffusion increases entropy and
-vagueness, decreases informativity and expected utility, but it also
-speeds up convergence to a convex and well-behaved outcome.
 
@@
 
-This type of variability in evolutionary outcomes is also weeded out
-by mild increases in impairment.
+Interestingly, this type of variability in evolutionary outcomes is
+also weeded out by mild increases in impairment. To see this, we
+calculated the average distance between evolved sender strategies
+within each group of trials that had identical independent
+parameters. We determined the distance between sender strategies
+$\Sstrat$ and $\Sstrat'$ as the average Hellinger distance between
+probability distributions $\Sstrat(\state)$ and $\Sstrat'(\state)$
+over messages at each choice point $\state$:
+\begin{align*}
+  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
+     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
+ \sqrt{\sum_{\messg \in  \Messgs}
+         \left ( \sqrt{\Sstrat(\state,\messg)} -
+         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
+      \\
+\end{align*}
+To compensate for the arbitrariness of message use, we set the
+distance between strategies $\Sstrat$ and $\Sstrat'$ to be the maximum
+of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
+where $\Sstrat^*$ is $Sstrat$ with reversed message indices. The
+average distances between sender strategies in this sense is plotted
+in Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
+increasing imprecision, the sender strategies evolving under \rdd were
+exactly alike, modulo which message was used for which part of the
+unit interval. In other words, perceptual imprecision clearly unifies
+evolutionary outcomes, and in fact guarantees sender strategies that
+are not only convex, but also maximally efficient in that they induce
+a vague category split exactly in the middle of the unit interval.
 
@@
 
-    \includegraphics[width=\textwidth]{plots/AverageEUinGroups.pdf}
+    \includegraphics[width=\textwidth]{plots/WithinGroupMeasures.pdf}
 
-   \caption{Average normalized expected utility between evolved strategies within each triplet of independent parameter values.}
+    \caption{Mean group measures as defined in the main text. For each
+      group, i.e., triplet of independent variables, the plot shows
+      (i) average normalized expected utility when playing against an
+      arbitrary other language from the group, (ii) the mean distance
+      between evolved strategies in the group.}
   \label{fig:AverageEUinGroups}
 \end{figure}
+
+This unifying property of perceptual imprecision could be considered
+an evolutionary beneficial side-effect. When we look at the normalized
+average expected utility that each evolved language (i.e., pair of
+sender and receiver strategy at stopping time) scored when playing
+against an arbitrary other language of the same group in
+Figure~\ref{fig:MoreExample_strats}, we see that higher imprecision
+can lead to higher such average expected utility. What this means is
+that, imprecision might decrease the communicative efficiency of
+individual languages, it increases the conceptual coherence and
+communicative success between independently evolving strategies. It is
+almost as if mere confusability of states imposes a regularity
+constraint on evolving linguistic meanings.
 

### 2014-09-30 snippets/example_strats/example_strats_plots.R

@@ -27,6 +27,6 @@ show(ds[stratsToPlot,])
 d = subset(data, Number.of.states == 6 & Tolerance == 0.3 & Impairment == 0.05 & Speaker.Convex.Cat == 1)
 
 
-plot.and.save.language.by.index(d, which.min(d$Expected.utility))
-plot.and.save.language.by.index(d, which.max(d$Expected.utility))
-d[c(which.min(d$Expected.utility), which.max(d$Expected.utility)),]
+# plot.and.save.language.by.index(d, which.min(d$Expected.utility))
+# plot.and.save.language.by.index(d, which.max(d$Expected.utility))
+# d[c(which.min(d$Expected.utility), which.max(d$Expected.utility)),]

### 2014-09-30 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -18,7 +18,7 @@ HD = function(x, y){
   return( sqrt(sum(sapply(1:length(x), function(i) (sqrt(x[i]) - sqrt(y[i]))^2  )) ) / (sqrt(2)) )
 }
 
-# Distance between two strategis as Hellinger distances at each choice point.
+# Distance between two strategis as mean Hellinger distances at each choice point.
 
 strategy.distance = function(P,Q){
   nc = dim(P)[1]
@@ -210,15 +210,20 @@ sPlot = ggplot(meanEUs, aes(x=Impairment, y=Speaker.meanEU)) +
 
 # Combine distance and average Group EU
 
-D$withinGroupEU = meanEUs$Speaker.meanEU
-D$individualEU = meanEUs$meanEU
-D$convexProportion = meanEUs$convexProportion
+D$meanEUGroup = meanEUs$Speaker.meanEU
+# D$meanEUIndiv = meanEUs$meanEU
+# D$convexProportion = meanEUs$convexProportion
 
 Dmelt = melt(D, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
 
 combPlot = ggplot(Dmelt, aes(x=Impairment, y=value, color = variable)) + 
-  geom_point() + facet_grid(Number.of.states ~ Tolerance) + geom_line()
+  geom_point(aes(shape = variable)) + facet_grid(Number.of.states ~ Tolerance) + geom_line()
 
-# show(combPlot)
+show(combPlot)
+
+pdf("~/Desktop/data/svn/vagueness-games/paper/plots/WithinGroupMeasures.pdf", height=6, width=8)
+show(combPlot)
+dev.off()
+
 
 

### 2014-09-30 paper/03_simulations.tex
comparison with QRE

@@
 
+The strategy pair in Figure~\ref{fig:example_stratsA} has a normalized
+expected utility of $0.992$, the pair in
+Figure~\ref{fig:example_stratsB} has $0.901$. Generally, vagueness and
+imprecision can be expected to decrease expected utility
+\citep{Lipman2009:Why-is-Language}. The crucial question is whether
+communicative success drops unacceptably fast with moderate levels of
+vagueness and imprecision.
+
 \subsubsection{Results}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}
+
+  \caption{Means of recorded metrics for each triple of independent variables.}
+  \label{fig:MeanMetrics}
+\end{figure}
+
 Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
@@
 of states in the game, but hardly any in terms of pragmatic
-tolerance. Interestingly, Voronoiness of both sender and receiver
-strategies appear to be non-decreasing under increasing impairment and
-peaking at the maximal value. The only exception is the 6 states games
-and the transition from $\impairment = 0$ to $\impairment =
-0.05$. Expected utility of evolved languages mostly decreases in
-increasing imprecision, with the sole exception of a rise for $\toler
-= 0.2$ and $ns = 6$ at $\impairment = 0.1$. Still, normalized expected
-utility of evolved strategy pairs under mild values of imprecision are
-not necessarily much worse than their respective perfect precision
-\rd-analogues.
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}
-
-  \caption{Means of recorded metrics for each triple of independent variables.}
-  \label{fig:MeanMetrics}
-\end{figure}
+tolerance. What was not expected \emph{a priori} was that Voronoiness
+of both sender and receiver strategies appear to be non-decreasing
+under increasing impairment and peaking at the maximal value. The only
+exception is the 6 states games and the transition from $\impairment =
+0$ to $\impairment = 0.05$. Expected utility of evolved languages
+mostly decreases in increasing imprecision, with the sole exception of
+a rise for $\toler = 0.2$ and $ns = 6$ at $\impairment = 0.1$. Still,
+normalized expected utility of evolved strategy pairs under mild
+values of imprecision are not necessarily much worse than their
+respective perfect precision \rd-analogues. This is especially so for
+lower numbers of states (which increase the chance of non-convergence
+and non-convexity; see Figure~\ref{fig:CategoricalMeasures} and the
+discussion below).
 
@@
 
-  \caption{Proportion of converged and convex trials for each triple
-    of independent variables.}
+  \caption{Proportion of converged and convex trials.}
   \label{fig:CategoricalMeasures}
@@
 
-Diffusion of strategies from confusion of states has another
-interesting regularizing effect on the evolution of signaling. If we
-look at the confidence intervals in Figure~\ref{fig:MeanMetrics}, we
-see that there is very little variation in the recorded metrics for
-evolved strategies, at least for higher values of impairment. On
-closer inspection, it turns out that variability in low-impairment
-conditions is not only due to non-convergence and/or associated
+Diffusion also has another interesting regularizing effect on the
+evolution of signaling. If we look at the confidence intervals in
+Figure~\ref{fig:MeanMetrics}, we see that there is very little
+variation in the recorded metrics for evolved strategies, at least for
+higher values of impairment. On closer inspection, it turns out that
+variability in low-impairment conditions is not only due to
+non-convergence and/or associated
 non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more
@@
 within each group of trials that had identical independent
-parameters. We determined the distance between sender strategies
+parameter values. We determined the distance between sender strategies
 $\Sstrat$ and $\Sstrat'$ as the average Hellinger distance between
@@
          \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
-      \\
 \end{align*}
@@
 of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
-where $\Sstrat^*$ is $Sstrat$ with reversed message indices. The
+where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
 average distances between sender strategies in this sense is plotted
@@
       arbitrary other language from the group, (ii) the mean distance
-      between evolved strategies in the group.}
+      between evolved sender strategies in the group.}
   \label{fig:AverageEUinGroups}

### 2014-09-30 paper/04_discussion.tex
simulation results write up
comparison with QRE

@@
 
+\subsection{Relation with previous accounts}
+
+\paragraph{Quantal response equilibria.}
+\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in
+which information processing limitations of signaling agents could
+lead to vague strategies. The model that is most clearly related to
+the present approach uses the notion of a quantal response, also known
+as a logit response or a soft-max function
+\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,McKelveyPalfrey1995:Quantal-Respons,McKelveyPalfrey1998:Quantal-Respons,GoereeHolt2008:Quantal-Respons}. A
+quantal response function is a paramterized generalization of the
+classic best response function. For example, if $U \mycolon \Acts
+\rightarrow \mathds{R}$ is the measure of expected utility over
+choices $\Acts$ of an agent, then a best response function would have
+the agent choose $\act$ only if $U(\act) = \max_{\act' \in \Acts}
+U(a)$. A quantal response function rather assumes that agents would
+choose $\act$ with a probability proportional to $\expo(\lambda \cdot
+U(\act))$, where $\lambda$ is a rationality parameter. If $\lambda
+\rightarrow \infty$ we retrieve the behavior of the best response
+choice function, but if it is positive but finite, any choice $\act$
+will receive a positive probability of being chosen, but acts with
+higher expected utility will be more
+likely. \citet{FrankeJager2010:Vagueness-Signa} show that quantal
+response equilibria of sim-max games, i.e., pairs of sender and
+receiver strategies such that the sender strategy is the quantal
+response to the expected utilities under the receiver strategy and
+vice versa, show the desired marks of vague
+signaling. Figure~\ref{fig:exampleQRE_stratsA} shows an example of a
+quantal response equilibrium for a sim-max game, as used in our set-up
+with $\toler = 0.5$. Sender and receiver strategies look very much
+like what evolves under \rdd with modest values of perceptual
+imprecision.
+
+There is a major conceptual difference in the source of vagueness
+between the quantal response approach and our favored approach in
+terms of perceptual noise. When we choose a lower level of pragmatic
+tolerance, like we used in our simulations, the predictions of the
+quantal response approach look
+different. Figure~\ref{fig:exampleQRE_stratsB} shows such an example
+that clearly shows that, unlike the evolving strategies under \rdd,
+sender strategies have vague boundaries also towards the end of the
+unit intervals. This is because quantal responses equalize message use
+far away from the ``prototypical'' interpretation. That this is
+conceptually odd shows even more clearly in a case where the state
+space is intuitively unbounded, as for instance for the property
+``tall''. If the usual interpretation of a ``tall man'' peaks at
+around, say, 195cm then when meeting a giant of $n$ meters speaker
+would, according to the quantal response approach, be ever more
+inclined to describe the giant as ``short''. This is because, as the
+distance from the prototype increases for larger $n$, the expected
+utilities of saying ``short'' or ``tall'' will both converge to zero,
+and the difference between the two will also converge to zero. Whence
+that the quantal response approach would predict that speakers would
+grow indifferent between choice of antonyms as $n$ grows, which seems
+weird. (Notice that this argument hinges on the choice of utility
+function. Still, to the extent that the chosen lower bounded utility
+functions are reasonable ---and we think they are very reasonable---,
+this is a conceptual oddity.)
+
+The \rdd approach does not have this peculiarity. \dots
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
+    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
+    \label{fig:exampleQRE_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
+    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
+    \label{fig:exampleQRE_stratsB}
+  \end{subfigure}
+
+  \caption{Examples of vague quantal response equilibria.}
+  \label{fig:exampleQREs}
+\end{figure}
+
 \mytodo{JPC}{Discuss nature of noise}

### 2014-09-30 snippets/data_exploration/data_exploration_tmp.Rmd
comparison with QRE

@@
 dsub = dms
+dsub$Number.of.states = factor(dsub$Number.of.states)
 
 pd <- position_dodge(.01)
-sp = ggplot(dsub, aes(x=Impairment, y=value, colour=factor(Number.of.states))) + 
+sp = ggplot(dsub, aes(x=Impairment, y=value, colour=Number.of.states)) + 
      geom_point(position = pd) +
      geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
-     geom_line(position=pd)
+     geom_line(position=pd) + xlab("imprecision")
 
@@
 
-pdf("~/Desktop/data/svn/vagueness-games/paper/plots/MeanMetrics.pdf", width=8, height = 8)
+pdf("~/Desktop/data/svn/vagueness-games/paper/plots/MeanMetrics.pdf", width=8, height = 6)
 show(sp)
@@
 sp = ggplot(dms, aes(x=factor(Impairment), y = value)) 
-sp = sp + geom_bar(aes(fill = variable), color = "black", stat="identity", position= position_dodge(0.4), width = 0.3)
-sp = sp + facet_grid(Number.of.states ~ Tolerance, scales="free")
+sp = sp + geom_bar(aes(fill = variable), color = "black", stat="identity", position= position_dodge(0.4), width = 0.35)
+sp = sp + facet_grid(Number.of.states ~ Tolerance, scales="free") + ylab("proportion") + xlab("imprecision")
 
@@
 
-pdf("~/Desktop/data/svn/vagueness-games/paper/plots/CategoricalMeasures.pdf", width=8, height = 6)
+pdf("~/Desktop/data/svn/vagueness-games/paper/plots/CategoricalMeasures.pdf", width=8, height = 5)
 show(sp)

### 2014-09-30 snippets/QRE.R

@@ -1,0 +1,39 @@
+library(MCMCpack) # for Dirichlet distribution
+library(ggplot2)
+library(reshape2)
+
+source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
+
+len = 1
+ns = 50
+lambda = 15
+tolerance = 0.05
+
+states = seq(0, len, length.out = ns)
+priors = rep(1/ns, ns)
+utils = get.confusion.matrix(states = states, alpha = tolerance)
+
+sen = rdirichlet(ns,rep(1,2))
+rec = rdirichlet(2, rep(1,ns))
+
+
+for (i in 1:100){
+  eu.s = get.eu.s(rec = rec, utils = utils)
+  senNext = get.qbr(utils = eu.s, lambda = lambda)
+  eu.r = get.eu.r(sen = sen, utils = utils, priors = priors)
+  recNext = get.qbr(utils = eu.r, lambda = lambda)
+  if (all(as.vector(sen) == as.vector(senNext))) {
+    show("converged")
+  }
+  sen = senNext
+  rec = recNext
+}
+
+outPlot = arrangeGrob(plot.sender.strat(sen,states),
+                      plot.receiver.strat(rec,states))
+
+show(outPlot)
+filename = paste("~/Desktop/data/svn/vagueness-games/paper/plots/exampleStratQRE_tolerance0", tolerance*10, ".pdf", sep = "", collapse = "")
+pdf(filename, height = 5, width = 6)
+show(outPlot)
+dev.off()

### 2014-09-30 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -125,21 +125,7 @@ expected.utility.byIndeces = function(data, i, j){
   return(max(eu1,eu2))
 }
 
-# Optimal sender & receiver strategy [
-#  works ONLY for 2 messages, ns %% 2 == 0 and ns/2 %% 2 == 1,  and flat priors only
 
-optimal.strategies = function(ns) {
-  if (!(ns %% 2 == 0 & (ns/2) %% 2 == 1)) {
-    stop("cannot get optimal strategies for this state number")
-  }
-  sen = matrix(0, nrow =ns, ncol = 2)
-  sen[1:(ns/2),1] = 1
-  sen[(ns/2+1):ns,2] = 1
-  rec = matrix(0, nrow = 2, ncol = ns)
-  rec[1,median(1:(ns/2))] = 1
-  rec[2,median((ns/2+1):ns)] = 1
-  return(list(sen,rec))
-}
 
 # utility of optimal sender & receiver strategy
 

### 2014-10-01 paper/04_discussion.tex
discussion section; oconnor

--- 2014-08-22 paper/04_discussion.tex
@@
 
-Finally, the replicator dynamics are also related to some forms of reinforcement learning (\cite{Borgers1997,Hopkins2005,Beggs2005}), and can thus be interpreted in cognitive terms.
+\subsection{Relation with previous accounts}
 
-\mytodo{JPC}{Discuss nature of noise}
-\mytodo{JPC}{Compare with O'Connor}
-\mytodo{JPC}{Compare with Franke2013}
+\paragraph{Generalized reinforcement.}
+\citet{OConnor2013:The-Evolution-o} introduces a generalization of
+Herrnstein reinforcement learning for sim-max games and showed that
+this not only leads to vague signaling patterns, but can speed-up
+evolution of efficient signaling strategies, especially in games with
+many states.
+
+Under plain Herrnstein reinforcement learning, sender and receiver
+play the game repeatedly and adjust their dispositions to act after
+each round of play, in such a way that the actual (non-negative)
+payoff gained in the current interaction is added to the
+non-normalized propensities for acting in exactly the way that they
+acted in the current round of play. For sim-max games, this means that
+when the sender chose $\messg$ in state $\state$, and this resulted in
+some non-negative payoff (which is guaranteed by our choice of utility
+function), the probability that the sender chooses $\messg$ again in
+$\state$ is increased, but nothing else changes. In particular, the
+speakers behavior in other choice points does not change. Generalized
+reinforcement learning is different here. When the use of $\messg$ in
+$\state$ gave positive payoff, then not only will its future use
+probability be promoted at $\state$ but also at other states,
+proportional to how similar these are to $\state$. Similar amendments
+take care of the way that the receiver updates his choice
+dispositions.
+
+\citet{OConnor2013:The-Evolution-o} shows that this extension not only
+leads to vague signaling of the appropriate kind, but also speeds up
+learning in such a way that, especially for games with higher numbers
+of states, higher levels of communicative success are reached in
+shorter learning periods than is possible without stimulus
+generalization. Technically, this result is partly due to the fact
+that signalers make bigger changes to their behavioral strategies
+after each round of play under generalized reinforcement than under
+the plain variety. But that only explains the speed of adaptation, not
+necessarily that generalization also leads to regularity and
+communicative efficiency.
+
+Diffusion of strategies in the \rdd can be conceived of as a form of
+generalization as well, and works in large part quite analogous to
+stimulus generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
+approach. This holds for the effect of diffusion and generalization,
+but not necessarily for the way that the effect is achieved. We also
+saw that diffusion in \rdd leads to more regular languages and
+speedier convergence. However, the \rdd is a more abstract framework
+than generalized reinforcement learning. The latter is foremost
+motivated as a learning dynamic that has two players adapt their
+individual strategies after each concrete round of play. In contrast,
+the \rdd describes a more abstract, average dynamical change in
+behavioral dispositions. Although the dynamics of (some forms of)
+reinforcement learning mirror those of the replicator dynamics (at
+some stage in time) (\cite{Borgers1997,Hopkins2005,Beggs2005}), this
+does not mean that \emph{generalized} reinforcement learning, in which
+stimulus generalization is best motivated at the level of a single
+agent's dispositional generalization after one round of play, is also
+directly a plausible high-level description of, say, generalized
+learning in a population of agents. Seen in this light, generalized
+reinforcement learning and the \rdd nicely complement each other, as
+similarly-minded accounts operating at different levels of
+abstraction.
+
+\paragraph{Quantal response equilibria.}
+\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in
+which information processing limitations of signaling agents could
+lead to vague strategies. The model that is most clearly related to
+the present approach uses the notion of a quantal response, also known
+as a logit response or a soft-max function
+\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,McKelveyPalfrey1995:Quantal-Respons,McKelveyPalfrey1998:Quantal-Respons,GoereeHolt2008:Quantal-Respons}. A
+quantal response function is a paramterized generalization of the
+classic best response function. For example, if $U \mycolon \Acts
+\rightarrow \mathds{R}$ is the measure of expected utility over
+choices $\Acts$ of an agent, then a best response function would have
+the agent choose $\act$ only if $U(\act) = \max_{\act' \in \Acts}
+U(a)$. A quantal response function rather assumes that agents would
+choose $\act$ with a probability proportional to $\expo(\lambda \cdot
+U(\act))$, where $\lambda$ is a rationality parameter. If $\lambda
+\rightarrow \infty$ we retrieve the behavior of the best response
+choice function, but if it is positive but finite, any choice $\act$
+will receive a positive probability, but acts with higher expected
+utility will be more likely. The underlying motivation for this choice
+rule is the assumption that there is noise in the computation of
+expected utilities and/or in maximization of expected
+utilities. Consequently, choices with almost equal expected utilities
+will be chosen with almost equal probability (for moderate values of
+$\lambda$).
+
+\citet{FrankeJager2010:Vagueness-Signa} show that quantal response
+equilibria of sim-max games, i.e., pairs of sender and receiver
+strategies such that the sender strategy is the quantal response to
+the expected utilities under the receiver strategy and vice versa, can
+show the desired marks of vague
+signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
+quantal response equilibrium for a sim-max game, as used in our set-up
+but with higher tolerance $\toler = 0.5$. Sender and receiver
+strategies look very much like what evolves under \rdd with modest
+values of perceptual imprecision.
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
+    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
+    \label{fig:exampleQRE_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
+    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
+    \label{fig:exampleQRE_stratsB}
+  \end{subfigure}
+
+  \caption{Examples of vague quantal response equilibria.}
+  \label{fig:exampleQREs}
+\end{figure}
+
+But not all quantal response equilibria are equally plausible
+explanations of vague language use, and the ones that are not suggest
+that it is less plausible to think of vagueness as arising because of
+errors in the computation and maximization of expected utility, as the
+quantal response approach assumes, than that it arises due to
+confusion of similar states, as the \rdd proposes. To see what the
+problem is, we can look at cases like given in
+Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
+equilibrium for a game with lower tolerance $\toler = 0.05$. Unlike
+what evolves under \rdd in this case, sender strategies have vague
+boundaries also towards the end of the unit intervals. Technically,
+this is because quantal responses equalize message use far away from
+the ``prototypical'' interpretation, not just in-between categories,
+so to speak. This, in turn, is because quantal responses introduce
+noise into the decision making at the level of computing or maximizing
+expected utility of choices.
+
+That this is conceptually odd shows even more clearly in a case where
+the state space is intuitively unbounded, as for instance for the
+property ``tall''. If the usual interpretation of a ``tall man'' peaks
+at around, say, 195cm then when meeting a giant of $n$ meters speakers
+would, according to the quantal response approach, be ever more
+inclined to describe the giant indifferently as either ``tall'' or
+``short'' the larger $n$ gets. This is because, as the distance from
+the prototype increases for larger $n$, the difference between the
+expected utilities of saying ``short'' or ``tall'' will converge to
+zero. Whence that the quantal response approach would predict that
+speakers would grow indifferent between choice of antonyms as $n$
+grows, which seems weird.
+
+Admittedly, this argument hinges on the choice of utility
+function. Still, to the extent that the chosen lower bounded utility
+functions are reasonable ---and we think they are very reasonable---,
+the case suggests that quantal responses are not a good model,
+intuitively speaking, for why linguistic categories are
+vague. Vagueness is more plausibly an effect of perceptual confusion
+of similar states, than of computational errors in maximizing expected
+utility.
 

### 2014-10-02 paper/01_intro.tex
intro; sec com

@@
 \section{Introduction}
+
+Many of our concepts and words are vague. A vague category knows clear
+cases that fall under it, clear cases that do not, and also so-called
+borderline cases in-between that neither clearly apply, nor clearly
+not apply, but might themselves not be equal in terms of how well they
+represent the category in question. Vagueness does not seem to
+dramatically affect the success of everyday communication, but it is
+troublesome for some of our best theories of language. Especially the
+logico-semantic tradition of Frege, Russell and the young Wittgenstein
+is challenged by vagueness and the paradoxes that it gives rise to. As
+can be expected, many proponents of this tradition have risen to the
+challenge \citep{Williamson1994:Vagueness}.
+
+There are other intriguing aspects about vagueness. Linguists, for
+example, wonder about the reasons why some expressions are less vague
+than others \citep[e.g.][]{Kennedy2007:Vagueness-and-G}. But the
+puzzle that we are concerned with here is how vagueness could arise
+and be maintained in the first place. This may not seem a particularly
+deep issue at first glance, but on closer inspection it is a serious
+worry to functionalist accounts that maintain that our concepts and
+language use evolved in order to be efficient. Since the existence of
+unclear borderline cases seems to entail inefficiency of
+categorization or communication, the challenge, succinctly put forward
+by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
+exist despite its obvious and demonstrable functional deficiency under
+evolutionary pressure to be optimally expressive.
+
+This problem, call it Lipman's problem, has a conceptual and a
+technical side to it. Conceptually, it asks for reasons and general
+mechanisms by which we could plausibly conceive of vagueness as
+persisting the pressure of evolutionary optimization towards
+precision. Such reasons and general mechanisms are, arguably, not hard
+to imagine. We will address some presently. But each putative
+explanation, no matter how plausible intuitively, must also be checked
+for internal consistency and its potential to account for the fact
+that we see vagueness as a part of a by-and-large efficient system of
+categorization and communication. In other words, not all first-shot
+rebuttals of Lipman's problem will do. Denying that there is any
+functional pressure on efficient communication, for instance, is
+dubious because it leaves the relative efficiency of our communication
+with vague words entirely unexplained. In sum, the technical answer to
+Lipman's problem is to give, ideally, a model of evolution of
+categorization or language use that shows how competing forces that
+realize both by-and-large efficient categories and vagueness.
+
+c
+
+\bigskip
 

### 2014-10-02 paper/01_intro.tex
introduction and background; sec com

@@
 can be expected, many proponents of this tradition have risen to the
-challenge \citep{Williamson1994:Vagueness}.
+challenge
+\citep[e.g.][]{Williamson1994:Vagueness,CobrerosEgre2012:Tolerant-Classi}.\todo{more
+  references}
 
-There are other intriguing aspects about vagueness. Linguists, for
-example, wonder about the reasons why some expressions are less vague
-than others \citep[e.g.][]{Kennedy2007:Vagueness-and-G}. But the
+Yet there are other intriguing aspects about vagueness. The
 puzzle that we are concerned with here is how vagueness could arise
@@
 mechanisms by which we could plausibly conceive of vagueness as
-persisting the pressure of evolutionary optimization towards
-precision. Such reasons and general mechanisms are, arguably, not hard
-to imagine. We will address some presently. But each putative
-explanation, no matter how plausible intuitively, must also be checked
-for internal consistency and its potential to account for the fact
-that we see vagueness as a part of a by-and-large efficient system of
+resisting the pressure of evolutionary selection for precision. Such
+reasons and general mechanisms are, arguably, not hard to imagine. We
+will address some presently. But each putative explanation, no matter
+how plausible intuitively, must also be checked for internal
+consistency and its potential to account for the fact that we see
+vagueness as a part of a by-and-large efficient system of
 categorization and communication. In other words, not all first-shot
 rebuttals of Lipman's problem will do. Denying that there is any
-functional pressure on efficient communication, for instance, is
-dubious because it leaves the relative efficiency of our communication
-with vague words entirely unexplained. In sum, the technical answer to
-Lipman's problem is to give, ideally, a model of evolution of
-categorization or language use that shows how competing forces that
-realize both by-and-large efficient categories and vagueness.
-
-c
-
-\bigskip
-
-\begin{itemize}
-\item introduce sim-max games
-\item problem of vagueness \citep{Lipman2009:Why-is-Language}
-\item \dots
-\end{itemize}
-
-\subsection{Situating our approach}
+functional pressure on efficient communication, for instance, is a
+non-starter, because it leaves the relative efficiency of our
+communication with vague words entirely unexplained. In sum, the
+technical answer to Lipman's problem is to give, ideally, a model of
+evolution of categorization or language use that shows how competing
+forces that realize both by-and-large efficient categories and
+vagueness.
 
@@
 \citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,Jaegherde-JaegherRooijvan-Rooij2010:Strategic-Vague,BlumeBoard2013:Intentional-Vag}. In
-contrast to these, others have argued that vagueness if a natural
+contrast to these, others have argued that vagueness is a natural
 byproduct of limitations in information processing
-\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
-paper makes a contribution to the latter line of thought.
+\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o,QingFranke2014:Gradable-Adject}. This
+paper makes a contribution to the latter line of thought.  Concretely,
+we introduce a novel variant of the replicator mutator dynamic that
+integrates stochastic noise on the differential confusability of
+similar stimuli. Our main contribution, therefore, is a technical
+answer to Lipman's problem in the sense introduced above. The dynamic
+proposed here nicely improves on and complements previous like-minded
+accounts. We show that that stochastic noise can not only lead to
+vague, yet communicative efficient signal use, but can also unify
+evolutionary outcomes and help avoid inefficient categorization.
+
+The next section introduces the background against which the work
+presented here can be appreciated. Section~\ref{sec:repl-diff-dynam},
+then, introduces the replicator diffusion dynamic, and elaborates its
+relation with the replicator mutator
+dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
+diffusion dynamic on the relevant class of generalized signaling games
+introduced in
+Section~\ref{sec:background}. Section~\ref{sec:discussion} reflects on
+the results and compares what has been achieved to related accounts in
+more detail.
+
+\section{Background}
+\label{sec:background}
 
 The view that vagueness is a natural concomitant of cognitive
-limitations of language users has been formalized in a number of
-ways. As in other applications of evolutionary game theory, we can
-distinguish micro- and macro-level approaches. Micro-level approaches
-look at adaptive behavior of individual agents. Usually, changes in
-the behavioral dispositions of agents occur after every single
-interaction. In contrast, macro-level approaches outline more
-abstract, aggregate population-based dynamics. Usually, a macro-level
-dynamic captures changes of frequencies of behavioral types in the
-population over time. Previous accounts have suggested both micro- and
-macro-level dynamics to show how vague signal use can arise in sim-max
-games.
+limitations of language users has been formalized in a number of ways,
+using evolutionary game theory and certain generalizations of
+signaling games, so called similarity-maximizing games or sim-max
+games, for short
+\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}. Our
+contribution is best seen in relation to these accounts, as it also
+relies on sim-max games. Let's introduce these first.
+
+Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
+a sender and a receiver. The sender knows the true state of the world,
+but the receiver does not. The sender can select a signal, or message,
+to reveal to the receiver, who then chooses an act. In Lewis' games,
+states are maximally distinct from each other and the receiver's acts
+are related to them one-to-one. If the receiver chooses the act that
+corresponds to the actual state, the play is a success, otherwise a
+failure. Similarity-maximizing games are generalizations of Lewis'
+games in that they allow different states to be more or less similar
+to one another, and, roughly put, to make success of communication a
+function of that similarity. Signaling games with utility-relevant
+similarities in the state space are fairly standard in economics
+\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
+but have received particular attention in a more philosophical and
+linguistic context by the more recent work of
+\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.
+
+More concretely, a sim-max game, in the relevant sense here, consists
+of a set of states $\States$, a set of messages $\Messgs$, a
+similarity metric on states: $\similarity \mycolon \States \times
+\States \rightarrow \mathds{R}$, and a utility function $\Utils
+\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
+the receiver's acts with the states of the world, so that the game is
+one of guessing the actual state, so to speak. We also assume that
+sender and receivers interests are alike, so we only have one utility
+function. We do not consider message costs, so utilities only depend
+on the actual state, and the receiver's response. The similarity function
+should satisfies the usual requirements for a metric:
+\begin{align*}
+  & \similarity(\mystate{1}, \mystate{2}) \ge 0 &&
+  \similarity(\mystate{1}, \mystate{2}) = 0 \Leftrightarrow
+  \mystate{1} = \mystate{2} \\
+  & \similarity(\mystate{1}, \mystate{2}) = \similarity(\mystate{2},
+  \mystate{1}) && \similarity(\mystate{1}, \mystate{2}) \le
+  \similarity(\mystate{1}, \mystate{3}) + \similarity(\mystate{3}, \mystate{2})\,.
+\end{align*}
+The utility function should be a monotonically decreasing function of
+similarity:
+\begin{align*}
+  \similarity(\mystate{1},\mystate{2}) \ge
+  \similarity(\mystate{1},\mystate{3}) \ \ \Rightarrow \ \ 
+  \Utils(\mystate{1},\mystate{2}) \ge
+  \Utils(\mystate{1},\mystate{3})\,.
+\end{align*}
+To keep matters simple, we only look at cases where $\States$ contains
+finitely many points from the unit interval, with similarity given by
+the Euclidean distance in this one-dimensional space.\todo{introduce
+  utilities here?}
+
+\begin{itemize}
+\item ESSs of sim-max games
+\item Voronoi tessallations
+\item conceptual spaces
+\item 
+\end{itemize}
+
+\bigskip
+
+As in other applications of
+evolutionary game theory, we can distinguish micro- and macro-level
+approaches. Micro-level approaches look at adaptive behavior of
+individual agents. Usually, changes in the behavioral dispositions of
+agents occur after every single interaction. In contrast, macro-level
+approaches outline more abstract, aggregate dynamics, happening in a
+huge population of agents, or otherwise abstracting from seemingly
+irrelevant detail. Usually, a macro-level dynamic captures changes of
+frequencies of behavioral types in the population over time. Previous
+accounts have suggested both micro- and macro-level dynamics to show
+how vague signal use can arise in sim-max games.
 
@@
 
-To understand the workings of the \rdd, we study numerical simulations
-\dots
-
-The paper is structured as follows:
-\begin{enumerate}
-\item replicator diffusion dynamics
-\item relation to replicator mutator dynamics
-\item simulations of \rdd
-\item comparison with other approaches
-\item conclusions
-\end{enumerate}
-
 %%% Local Variables: 

### 2014-10-02 paper/04_discussion.tex
introduction and background; sec com

@@
 reinforcement learning mirror those of the replicator dynamics (at
-some stage in time) (\cite{Borgers1997,Hopkins2005,Beggs2005}), this
-does not mean that \emph{generalized} reinforcement learning, in which
-stimulus generalization is best motivated at the level of a single
-agent's dispositional generalization after one round of play, is also
-directly a plausible high-level description of, say, generalized
-learning in a population of agents. Seen in this light, generalized
-reinforcement learning and the \rdd nicely complement each other, as
-similarly-minded accounts operating at different levels of
+some stage in time)
+(\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
+this does not mean that \emph{generalized} reinforcement learning, in
+which stimulus generalization is best motivated at the level of a
+single agent's dispositional generalization after one round of play,
+is also directly a plausible high-level description of, say,
+generalized learning in a population of agents. Seen in this light,
+generalized reinforcement learning and the \rdd nicely complement each
+other, as similarly-minded accounts operating at different levels of
 abstraction.
@@
 
+\paragraph{The conceptual spaces approach.}
+
+\begin{itemize}
+\item can only ``construct'' (fuzzy) tessellation's given the
+  prototypes (or prototype regions); but cannot explain how the
+  prototypes come about
+\end{itemize}
+
 %%% Local Variables: 

### 2014-10-02 paper/paper.tex
data exploration MF
introduction; iterated diffusion
improved structure Sec 02; data for convex vs. non-convex strats for small state spaces
restructured and expanded Section 2
iterated diffusion; plots
more on diffusion; Nosofsky similarity plots
diffusion section ready
more data exploration with varying tolerance and impairment
simulation results write up
results report
introduction and background; sec com

@@
 %\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
-\bibliography{../helpers/MyRefGlobal}
-\bibliography{paper}
+\bibliography{MyRefGlobal}
+% \bibliography{paper}
 \usepackage{../helpers/myenvironments}
@@
 
+\newcommand{\rd}{\acro{rd}} % replicator dynamic
+\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
+\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
+\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
+\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
+\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator
+
+\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
+\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation
+
+\newcommand{\impairment}{\ensuremath{\alpha_i}} % impairment
+\newcommand{\toler}{\ensuremath{\alpha_t}} % tolerance
+\newcommand{\ns}{\ensuremath{n_s}} % number of states
+
+\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function
+
 \title{Vagueness, Noise, and Communication}
-\author{Jos\'e Pedro Correia and Michael Franke}
+% \author{Jos\'e Pedro Correia and Michael Franke}
 \date{}
@@
 \begin{abstract}
-Signaling games have been garnering a lot of attention as models for studying natural language meaning.
-One important aspect that is nevertheless typically missing from most models in the literature is the possibility for vagueness as a feature of successful languages.
-Recently, some models have been proposed to explicitly address this limitation.
-In this paper we explore one such model and establish a link between the dynamics in the model and the replicator-mutator equations.
-Furthermore, we study the long term behavior of the model using computer simulations.
-The results indicate that vague languages, despite having lower expected utility and informativity, are faster to converge to an equilibrium than non-vague ones.
-We argue that this could be an importante advantage that contributes to the pervasiveness of vagueness in natural language.
+  Signaling games have attracted a lot of attention as models for
+  studying the evolution of meaning. One important aspect that is
+  nevertheless typically missing from most models in the literature is
+  the possibility for vagueness as a feature of successful signaling.
+  Complementing recent models that have been proposed to explicitly
+  address this limitation, we introduce the replicator diffusion
+  dynamic, provably a special case of the replicator mutator dynamic,
+  which implements a versatile and natural integration of stochastic
+  noise in the form of probabilistic confusion of similar stimuli, on
+  the one hand, and evolutionary pressure on optimal signal use, on
+  the other. Applying the replicator diffusion dynamic to a
+  generalization of David Lewis' signaling games, so-called
+  similarity-maximizing games, we show that stochastic noise can not
+  only lead to vague, yet communicative efficient signal use, but can
+  also unify evolutionary outcomes and help avoid inefficient
+  categorization.
 \end{abstract}

### 2014-10-02 paper/01_intro.tex
introduction and background; sec com

@@
 contribution is best seen in relation to these accounts, as it also
-relies on sim-max games. Let's introduce these first.
+relies on sim-max games. Let's introduce these first, and then zoom in
+on the problem of vagueness.
+
+\subsection{Sim-max games}
 
@@
 corresponds to the actual state, the play is a success, otherwise a
-failure. Similarity-maximizing games are generalizations of Lewis'
-games in that they allow different states to be more or less similar
-to one another, and, roughly put, to make success of communication a
-function of that similarity. Signaling games with utility-relevant
+failure. Certain regular combinations of sender signaling and receiver
+reaction make messages meaningful, in the sense that their use is
+correlated systematically to certain states or acts. To investigate
+the conditions under which such meaning-generating behavior can evolve
+is a highly interesting topic that we are only beginning to fully
+understand
+\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Zollman2005:Talking-to-Neig,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,Wagner2009:Communication-a,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals,HutteggerZollman2011:Signaling-Games}.
+
+Similarity-maximizing games are generalizations of Lewis' games in
+that they allow different states to be more or less similar to one
+another, and, roughly put, make success of communication a function of
+that similarity. Intuitively speaking, while Lewis' games treated
+communicative success as a matter of black and white, sim-max games
+allow for a gradient notion of communicative success: the more similar
+the receiver's interpretation is to the actual state, the more
+successful the play was. Signaling games with utility-relevant
 similarities in the state space are fairly standard in economics
@@
 but have received particular attention in a more philosophical and
-linguistic context by the more recent work of
-\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.
+linguistic context by the more recent work for reasons that will
+become clear presently
+\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.
 
 More concretely, a sim-max game, in the relevant sense here, consists
-of a set of states $\States$, a set of messages $\Messgs$, a
-similarity metric on states: $\similarity \mycolon \States \times
+of a set of states $\States$, a set of messages $\Messgs$ with much
+fever messages than states, a probability distribution $\Pr \in
+\Delta(\States)$ that gives the occurrence probabilities of states, a
+similarity metric on states $\similarity \mycolon \States \times
 \States \rightarrow \mathds{R}$, and a utility function $\Utils
@@
 one of guessing the actual state, so to speak. We also assume that
-sender and receivers interests are alike, so we only have one utility
-function. We do not consider message costs, so utilities only depend
-on the actual state, and the receiver's response. The similarity function
-should satisfies the usual requirements for a metric:
+sender's and receiver's interests are alike, so we only have one
+utility function. We do not consider message costs, so utilities only
+depend on the actual state, and the receiver's response. The
+similarity function should satisfy the usual requirements for a
+metric:
 \begin{align*}
@@
 To keep matters simple, we only look at cases where $\States$ contains
-finitely many points from the unit interval, with similarity given by
-the Euclidean distance in this one-dimensional space.\todo{introduce
-  utilities here?}
+finitely many points from the unit interval, all of which have equal
+probability, and with similarity given by the Euclidean distance in
+this one-dimensional space.\todo{introduce utilities here?}
+
+\citet{JagerMetzger2011:Voronoi-Languag} showed that the
+evolutionarily stable states of sim-max games are remarkably
+systematic. Their results were obtained for games with infinitely many
+states in $n$-dimensional Euclidean space $\States \subseteq
+\mathds{R}^n$ and a quadratic loss function for utilities
+$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
+\mystate{2})^2$. For these games, the evolutionarily stable states are
+demonstrably so-called Voronoi-languages. Intuitively speaking, a
+Voronoi language is a pair of sender and receiver strategies, such
+that the sender strategy (quasi-)partitions the state space into
+convex categories, while the receiver's interpretations are the
+central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
+an example for such a Voronoi-language for a sim-max game with
+$\States = [0,1]$, two messages, and a flat prior distribution that
+selects all states equally likely. The sender's strategy (a function
+from states to messages; more below) uses one message exclusively for
+all points in the lower interval and another for all points in the
+upper interval. The receiver's interpretations of messages are the
+central points in the respective intervals.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[scale=3]
+
+        \begin{scope}
+
+          \draw[-,thick] (0,0) -- (1.5,0);
+
+          \draw (0,0) node {\color{red}{\big{[}}};
+
+          \draw (.75,0) node {\color{red}{\big{)}}};
+
+          \draw (.76,0) node {\color{red}{\big{(}}};
+
+          \draw (1.5,0) node {\color{red}{\big{]}}};
+
+            \draw (.325,.15) node
+            {\color{red}{$\Spure(\mymessg{1}  \probbar \cdot)$}};
+
+            \draw (1.075,.15) node
+            {\color{red}{$\Spure(\mymessg{2} \probbar \cdot)$}};
+
+            \node at (.375,0) (node1) {\color{blue}{$\bullet$}};
+
+            \node at (.375,-.3) (label1)
+            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{1})$}};
+
+            \draw[-,color=blue] (node1) -- (label1);
+
+            \node at (1.125,0) (node2) {\color{blue}{$\bullet$}};
+
+            \node at (1.125,-.3) (label2)
+            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{2})$}};
+
+            \draw[-] (node2) -- (label2);
+
+            \draw (0,-.2) node {$0$};
+
+            \draw (.75,-.2) node {$0.5$};
+
+            \draw (1.5,-.2) node {$1$};
+
+          \end{scope}
+
+        \end{tikzpicture}
+
+  \caption{Example of a Voronoi-language on $\States = [0,1]$. The
+    sender strategy $\Spure$ uses one signal for the lower half, and
+    another for the upper half of the unit interval. The receiver
+    strategy $\Rpure$ interprets messages as the central elements in the
+    respective intervals.}
+  \label{fig:VoronoiLanguage}
+\end{figure}
+
+This result is particularly interesting, because it demonstrates that
+signaling can impose orderly categories on a metric space, without
+that being the ulterior purpose of it all. Sender and receiver can be
+distinct entities, whose purpose it is to communicate effectively
+about the actual state. In that case, evolving Voronoi-languages would
+explain why linguistic categories are well-behaved and orderly in the
+way they appear to be. More abstractly, sender and receiver can also
+be thought of as distinct modules in a single system, where the first
+module that must discretize the information that it is fed into a
+small sample of, suggestively, category labels to pass to a second
+system that tries to decode the original information. In this case,
+evolving Voronoi-languages would explain why conceptual categories are
+well-behaved and orderly in the way that they appear to be
+\citep[see][for more on this latter
+interpretation]{OConnor2013:Evolving-Percep}.
+
+Seen in this light, sim-max games appear to have the potential of
+providing a foundation to approaches in cognitive semantics that rely
+on the notion of conceptual spaces
+\citep[e.g.][]{Gardenfors2000:Conceptual-Spac}.
+
+\bigskip
 
@@
 \item conceptual spaces
-\item 
+\item desiderata on vagueness: higher order vagueness \dots
+\item no vagueness normally, but possible Franke, O'Connor ...
 \end{itemize}
@@
 
-As in other applications of
-evolutionary game theory, we can distinguish micro- and macro-level
-approaches. Micro-level approaches look at adaptive behavior of
-individual agents. Usually, changes in the behavioral dispositions of
-agents occur after every single interaction. In contrast, macro-level
-approaches outline more abstract, aggregate dynamics, happening in a
-huge population of agents, or otherwise abstracting from seemingly
-irrelevant detail. Usually, a macro-level dynamic captures changes of
-frequencies of behavioral types in the population over time. Previous
-accounts have suggested both micro- and macro-level dynamics to show
-how vague signal use can arise in sim-max games.
+\subsection{Vague signaling in sim-max games}
+
+As in other applications of evolutionary game theory, we can
+distinguish micro- and macro-level approaches. Micro-level approaches
+look at adaptive behavior of individual agents. Usually, changes in
+the behavioral dispositions of agents occur after every single
+interaction. In contrast, macro-level approaches outline more
+abstract, aggregate dynamics, happening in a huge population of
+agents, or otherwise abstracting from seemingly irrelevant
+detail. Usually, a macro-level dynamic captures changes of frequencies
+of behavioral types in the population over time. Previous accounts
+have suggested both micro- and macro-level dynamics to show how vague
+signal use can arise in sim-max games.
 

### 2014-10-02 paper/01_intro.tex
introduction and background; sec com

@@
 
-\subsection{Sim-max games}
+\subsection{Sim-max games \& conceptual spaces}
 
@@
 providing a foundation to approaches in cognitive semantics that rely
-on the notion of conceptual spaces
-\citep[e.g.][]{Gardenfors2000:Conceptual-Spac}.
-
-\bigskip
-
-\begin{itemize}
-\item ESSs of sim-max games
-\item Voronoi tessallations
-\item conceptual spaces
-\item desiderata on vagueness: higher order vagueness \dots
-\item no vagueness normally, but possible Franke, O'Connor ...
-\end{itemize}
-
-\bigskip
+on the notion of conceptual spaces.
+\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has
+prominently argued that natural categories are convex regions in a
+suitable conceptual space. If the conceptual space has a suitable
+metric, then it is possible to think of these convex categories as
+derived from a set of prototypes. Fixing a set of prototypes, we
+consider the category corresponding to each prototype $p$ as the set
+of points that are more similar to $p$ than to any other
+\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
+\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
+categorization system can be obtained: storing the prototypes lets us
+recover the categories without having to store each categories
+extension. However, what is left unexplained so far, is where the
+prototypes come from, and why we would not see just any distribution
+of prototypes as an equally efficient classification system. This is
+where sim-max games, might chip in a principled approach to deriving
+in an independent way, not only convex categories, but also
+prototypical exemplars belonging to them.
 
@@
 
-As in other applications of evolutionary game theory, we can
-distinguish micro- and macro-level approaches. Micro-level approaches
-look at adaptive behavior of individual agents. Usually, changes in
-the behavioral dispositions of agents occur after every single
-interaction. In contrast, macro-level approaches outline more
-abstract, aggregate dynamics, happening in a huge population of
-agents, or otherwise abstracting from seemingly irrelevant
-detail. Usually, a macro-level dynamic captures changes of frequencies
-of behavioral types in the population over time. Previous accounts
-have suggested both micro- and macro-level dynamics to show how vague
-signal use can arise in sim-max games.
+The brief outline of an approach to conceptual categorization using
+sim-max games leaves many problems unaddressed. One of them is that
+natural categories for continuously variable stimuli, like shades of
+color, pitch heights, spatial dimension and the like, do not have
+unique prototypes and clear category boundaries. We would like to
+account for the possibility of vagueness, in particular the following
+criteria
+\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re,Smith2008:Vagueness-and-D}:
+(i) clear positive examples of a vague category should show a
+gradient, perhaps smooth, transition to negative examples to
+accommodate also ``higher-order vagueness'' ; (ii) prototypes should
+likewise be gradient regions, peaking at the center of the vague
+category they represent.
+
+\citet{DouvenDecock2011:Vagueness:-A-Co} show that
+\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
+approach can be extended to account for the existence of borderline
+cases. From the assumption that prototypes are not unitary points, but
+extended, yet convex regions in conceptual space,
+\citeauthor{DouvenDecock2011:Vagueness:-A-Co} give a construction
+algorithm that yields ``collated Voronoi diagrams'' with thick,
+extended boundaries representing borderline regions. Building on this
+work, \citet{DecockDouven2012:What-is-Graded-} show further how it is
+possible to arrive at ``higher-order vagueness'' and degrees of
+category membership, by, essentially, weighing in the distance of
+different borderline examples to various prototypical regions. This
+accounts for the first of the two desiderata mentioned above, but
+still assumes that crisp non-gradient prototype regions must be
+pre-given.
+
+An alternative approach is taken by
+\citet{FrankeJager2010:Vagueness-Signa} and
+\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
+can be met by evolving strategies in sim-max games for certain types
+of solution concepts, that can be divided into micro- and macro-level
+approaches. Micro-level approaches look at adaptive behavior of
+individual agents. Usually, changes in the behavioral dispositions of
+agents occur after every single interaction. In contrast, macro-level
+approaches outline more abstract, aggregate dynamics, happening in a
+huge population of agents, or otherwise abstracting from seemingly
+irrelevant detail. Usually, a macro-level dynamic captures changes of
+frequencies of behavioral types in the population over time.
 
@@
 are subject to behavioral adjustment as well. (Alternative approaches
-are discussed in more detail in Section~XYZ, where we can compare them
-with the system introduced in this paper.)
+are discussed in more detail in Section~\ref{sec:discussion}, where we
+can compare them with the system introduced in this paper.)
 
@@
 use. The approach we take here is superficially similar, but there are
-crucial divergences. For one, we adopt a dynamic perspective by
-looking at the stable states of a suitable population dynamics. For
-another, we demonstrate in Section~XYZ that quantal responses can give
-rise to counterintuitive predictions. These counterintuitive examples
-suggest that vagueness in signaling use is not convincingly explained
-by appeal to mistakes in calculating expected utility, but rather, as
-we assume here, as the result of confusing similar states of the word.
+crucial differences. For one, we adopt a dynamic perspective by
+looking at limited-time outcomes of a dynamic process. For another, we
+demonstrate in Section~\ref{sec:discussion} that quantal responses can
+give rise to counterintuitive predictions. These counterintuitive
+examples suggest that vagueness in signaling use is not convincingly
+explained by appeal to mistakes in calculating expected utility, but
+rather, as we assume here, as the result of confusing similar states
+of the word.
 
-Against this background, we introduce a new macro-level approach
-to the evolution of vague signal use in sim-max games, where the
-source of vagueness is the agents' natural disability to sharply
-distinguish perceptually distinct states. This applies to both
-production and comprehension, i.e., both to the sender as well as to
-the receiver role.
-
-We call our dynamic \textbf{replicator diffusion dynamic} (\rdd),
-because it is a special case of the replicator mutator dynamic
-(\rmd). It is a special case in the sense that the \rdd incorporates a
-special kind of mutation that is based on the similarity of states (as
-defined in the sim-max game). More concretely, we start from the idea
-that similar states are more easily confused than less similar
-states. To formalize this idea, we assume that there is a confusion
-matrix $C$, where $C_{ij}$ is the probability that a state \mystate{i}
-is realized as a possibly different state \mystate{j}. (There is a
-slightly different sense of ``realization'' in the sender and the
-receiver side, and we will enlarge on this below.) Based on this, we
-show two things. For one, we show how the discrete-time formulation of
-the replicator dynamic can be easily adapted to integrate diffusion of
-behavior due to confusion of states. This yields the \rdd in its basic
+Against this background, we introduce a new macro-level approach to
+the evolution of vague signal use in sim-max games, where the source
+of vagueness is the agents' natural inability to sharply distinguish
+perceptually distinct states. This applies to the sender as well as to
+the receiver role (albeit in slightly different manner). We call our
+dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
+special case of the replicator mutator dynamic (\rmd). It is a special
+case in the sense that the \rdd incorporates a particular kind of
+mutation that is based on the similarity of states (as defined in the
+sim-max game). More concretely, we start from the idea that similar
+states are more easily confused than less similar states. To formalize
+this idea, we assume that there is a confusion matrix $C$, where
+$C_{ij}$ is the probability that a state \mystate{i} is realized as a
+possibly different state \mystate{j}. (There is a slightly different
+sense of ``realization'' in the sender and the receiver side, and we
+will enlarge on this below.) Based on this, we show two things. For
+one, we show how the discrete-time formulation of the replicator
+dynamic can be easily adapted to integrate diffusion of behavior due
+to confusion of states. This yields the \rdd in its basic
 discrete-time formulation. For another, we also show that there exists
@@
 
+Based on data from numerical simulations, we show that the inclusion
+of mild levels of such stochastic noise in the replicator dynamics
+does not disrupt the possibility of evolving communicative efficient
+signaling strategies at all. On the contrary, there might even be a
+higher-order benefit to the presence of perceptual imprecision, in
+that it accelerates converges, but also unifies and regularizes
+evolutionary outcomes in such a way that inefficient categorization is
+avoided. These results complement similar results by
+\citet{OConnor2013:The-Evolution-o}, obtained for a micro-dynamic
+extension of reinforcement learning. The \rdd adds to this a more
+general, abstract framework that is not tied to the specific
+assumptions of turn-based reinforcement.
+
 %%% Local Variables: 

### 2014-10-02 paper/01_intro.tex
discussion section; first almost complete version

@@
 probability, and with similarity given by the Euclidean distance in
-this one-dimensional space.\todo{introduce utilities here?}
+this one-dimensional space.
 

### 2014-10-02 paper/02_noisyRMD.tex
introduction and background; sec com
discussion section; first almost complete version

@@
 \section{Replicator diffusion dynamic}
+\label{sec:repl-diff-dynam}
 
@@
 \subsection{Replicator diffusion dynamic in behavioral strategies}
+\label{sec:repl-diff-dynam-1}
 
@@
 obvious that other (discrete-time) evolutionary dynamics can be
-subjected to state-confusability in an analogous fashion. In the case
-of the \rd, however, perturbation by state-confusion has a prominent
-close relative in the replicator mutator dynamic.
+subjected to state-confusability in an analogous fashion (see also
+Section~\ref{sec:model-interpretation}). In the case of the \rd,
+however, perturbation by state-confusion has a prominent close
+relative in the replicator mutator dynamic.
 

### 2014-10-02 paper/04_discussion.tex
discussion section; first almost complete version

@@
 
+We have introduced a new variant of the replicator mutator dynamics
+that implements specifically stochastic noise in the form of
+confusability of similar states. The probabilities of confusing nearby
+similar states is easily implemented in behavioral strategies, but can
+also be translated into a mutation matrix for pure strategies. This
+allows, in principle, several conceptual interpretations of the \rdd,
+on which we will briefly enlarge below in
+Section~\ref{sec:model-interpretation}.
+
+Next to this technical contribution, the results of our numerical
+simulations also advance our understanding of the possibility of
+evolving regular and efficient categories despite their (higher-order)
+vagueness. Section~\ref{sec:relat-with-prev} zooms in on the relation
+with closely related accounts once more, to delineate the present
+approach more precisely.
+
 \subsection{Model interpretation}
-The replicator-mutator equations are a mathematical model motivated by concepts from the theory of natural selection, namely natural variation, differential reproduction, and mutation.
-The most straightforward interpretation is thus a biological one: we can imagine signaling strategies as innate behaviorial tendencies of organisms, steps in the evolutionary process as successive generations, and selection as capturing the reproductive advantage of fitter individuals.
-In this interpretation, our proposed model is not an extremely interesting one, since the noise matrix simply corresponds to a restricted form of mutation.
+\label{sec:model-interpretation}
 
-Another possibility is to take the equations as embodying cultural evolution.
-In this interpretation, what is subject to the evolutionary forces are not organisms but strategies themselves.
-This pressuposes that individuals can adopt strategies from other individuals.
-Fitness captures the success of strategies, which in the case of language can be thought of in terms of communicative success.
-Differential reproduction represents the tendency of more successful strategies to be more likely adopted by individuals, be it through imitation or some learning procedure within the population.
-Under this interpretation, the proposed noise perturbation becomes more interesting.
-Messages correspond to the overt data available to individuals, whereas states are interpreted as the more private aspect of the behavior.
-Futhermore, similarity-maximization games are especially interesting in situations where the number of states largely exceeds the number of messages available.
-When trying to adopt another individual's behavior, it is thus more likely that information about states is both less accessible and less practical to elicit, hence justifying a higher likelihood of mistakes and a more pressing need for a certain degree of generalization%
-\footnote{The consequence of perturbing a strategy with a noise matrix based on similarity can be seen as ``spreading'' the effects of the fitness of the behavior for a state to its more similar states, which can also be interpreted as generalization}%
-.
+We mentioned in passing in Section~\ref{sec:repl-diff-dynam-1} that
+adding diffusion to other discrete-time evolutionary dynamics that
+operate on behavioral strategies is entirely straightforward. We
+could, for instance, easily diffuse the outcome of a best-response
+dynamic at each time step. That would make sense if we thought of
+agents as prone to confusing similar states that are otherwise
+rational optimizers of behavior. The reason that we chose the
+replicator dynamics to combine with diffusion is twofold. A minor
+reason is that it makes for a conceptually interesting link with the
+replicator mutator dynamics. A more important reason is that the
+replicator dynamic is especially versatile and non-committal about
+what the exact process of adaption is that is being modelled.
+
+Originally the \rd was introduced as mathematical model of evolution
+under asexual reproduction, motivated by concepts from the theory of
+natural selection \citep{TaylorJonker1978:Evolutionary-St}. The most
+conservative interpretation of the \rdd in our present context is thus
+a biological one: we can imagine signaling strategies as innate or
+fixed behaviorial tendencies of organisms, steps in the evolutionary
+process as successive generations, and selection as capturing the
+reproductive advantage of fitter individuals. This interpretation,
+strictly speaking, requires formalization in terms of mixed strategies
+via the \rmd, and possibly also a symmetrizing of the game, so that
+every agent is assumed to have a unique sender and receiver role at
+the same time the is bequeathed onto the next generation. Diffusion,
+in this context, could be either of two things. It could be
+differential mutation probabilities in line with other interpretation
+of mutation in the \rmd
+\citep[e.g.][]{NowakKomarova2001:Evolution-of-Un,KomarovaNiyogi2001:The-Evolutionar}:
+pure strategies are not faithfully reproduced, and mistakes in
+bequeathing pure strategies are more likely if they result from the
+confusion of similar states. Another possible biological
+interpretation of the \rdd, is that inheritance is faithful, but
+strategies are noisily realized. Strategies are not necessarily
+selected for what they are, but rather for how they are realized, once
+noise is factored in.
+
+But the replicator dynamic is not only a model of biological
+evolution. It can also be interpreted as a high-level description of
+the likely development of other behavioral adaptation processes, like
+differential imitation and cultural evolution in general
+\citep[see][for various derivations of the
+\rd]{Sandholm2013:Population-Game}. Under this interpretation, the \rd
+is consistent with the idea that what is subject to the evolutionary
+forces are not organisms but behavior: individuals can adapt their
+behavior to the perceived environment or adopt strategies from other
+individuals. Fitness captures the success of behavioral patterns,
+which in the case of language can be thought of in terms of
+communicative success.  Differential reproduction represents the
+tendency of more successful strategies to be more likely adopted by
+individuals, be it through imitation or some learning procedure within
+the population.
+
+Under this non-biological interpretation, we have again two options of
+picturing what diffusion is, similar to the biological cases
+before. One possibility is that diffusion is noise in the adoption of
+strategies, say, by conditional imitation, of strategies by other
+signalers. Another possibility is that diffusion is again noise on the
+realization of strategies: while behavior is optimized to be efficient
+(be it due to learning, introspection or imitation), realization of
+strategies is bound to be noisy due to confusability of similar
+states.
+
+We believe that all of the four mentioned interpretations are, on
+first approximations, feasible conceptualizations of the \rdd, and
+that it is a good thing to know of a working account of vague
+signaling that sketches where fitness-based selection under
+state-confusability will take is, abstracting away from the details of
+actually playing the game, inheriting, imitating or otherwise
+optimizing behavior in whatever particular way. It is a good thing to
+know this on the macro-level, especially since there are also
+micro-level accounts that nicely complement the picture. We turn to
+one such next.
 
 \subsection{Relation with previous accounts}
+\label{sec:relat-with-prev}
 
@@
 
-\paragraph{The conceptual spaces approach.}
-
-\begin{itemize}
-\item can only ``construct'' (fuzzy) tessellation's given the
-  prototypes (or prototype regions); but cannot explain how the
-  prototypes come about
-\end{itemize}
-
 %%% Local Variables: 

### 2014-10-07 paper/01_intro.tex
Small corrections and additions to abstract and introduction.

@@
 dramatically affect the success of everyday communication, but it is
-troublesome for some of our best theories of language. Especially the
+troublesome for some of our theories of language. This is especially the case for the
 logico-semantic tradition of Frege, Russell and the young Wittgenstein
-is challenged by vagueness and the paradoxes that it gives rise to. As
-can be expected, many proponents of this tradition have risen to the
-challenge
-\citep[e.g.][]{Williamson1994:Vagueness,CobrerosEgre2012:Tolerant-Classi}.\todo{more
-  references}
+which is challenged by the paradoxes vagueness gives rise to. As
+can be expected, many proponents of this tradition have tried to address the
+problem, ranging from suggesting that there is nothing wrong with the classical
+approach and the problem is of a different nature, as in Williamson's epistemic view~\citep*{Williamson1994:Vagueness},
+to proposing various kinds of modifications to either semantics, logic, or both, as in for example
+supervaluationism \citep{Mehlberg1958,Fine1975}, many-valued logic \citep{Zadeh1975,Machina1976,Edgington1997},
+or paraconsistent logic \citep{CobrerosEgre2012:Tolerant-Classi}.
 
@@
 answer to Lipman's problem in the sense introduced above. The dynamic
-proposed here nicely improves on and complements previous like-minded
-accounts. We show that that stochastic noise can not only lead to
+proposed here nicely generalizes on and complements previous like-minded
+accounts. We show that stochastic noise can not only lead to
 vague, yet communicative efficient signal use, but can also unify
@@
 The next section introduces the background against which the work
-presented here can be appreciated. Section~\ref{sec:repl-diff-dynam},
-then, introduces the replicator diffusion dynamic, and elaborates its
+presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
+introduces the replicator diffusion dynamic and elaborates on its
 relation with the replicator mutator

### 2014-10-07 paper/paper.tex
discussion section; first almost complete version
Small corrections and additions to abstract and introduction.

@@
 %\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
-\bibliography{MyRefGlobal}
-% \bibliography{paper}
+%\bibliography{MyRefGlobal}
+\bibliography{../helpers/MyRefGlobal}
+\bibliography{paper}
 \usepackage{../helpers/myenvironments}
@@
 \begin{document}
+
+\textbf{to do}
+
+\begin{itemize}
+\item check proofs in appendix; adapt to changed notation from main text
+\item make prettier and uniform plots
+\item write conclusion
+\end{itemize}
+
+\newpage
+
 \maketitle
@@
   address this limitation, we introduce the replicator diffusion
-  dynamic, provably a special case of the replicator mutator dynamic,
-  which implements a versatile and natural integration of stochastic
-  noise in the form of probabilistic confusion of similar stimuli, on
-  the one hand, and evolutionary pressure on optimal signal use, on
-  the other. Applying the replicator diffusion dynamic to a
+  dynamic. Provably a special case of the replicator mutator dynamic,
+  it implements a versatile and natural integration of, on the one hand, stochastic
+  noise in the form of probabilistic confusion of similar stimuli and, on
+  the other, evolutionary pressure on optimal signal use.
+  Applying the replicator diffusion dynamic to a
   generalization of David Lewis' signaling games, so-called

### 2014-10-07 paper/01_intro.tex
Small corrections to background section.

@@
 central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
-an example for such a Voronoi-language for a sim-max game with
+an example of such a Voronoi-language for a sim-max game with
 $\States = [0,1]$, two messages, and a flat prior distribution that
@@
 from states to messages; more below) uses one message exclusively for
-all points in the lower interval and another for all points in the
-upper interval. The receiver's interpretations of messages are the
+all points in the lower half of the interval and another for all points in the
+upper half. The receiver's interpretations of messages are the
 central points in the respective intervals.
@@
 that being the ulterior purpose of it all. Sender and receiver can be
-distinct entities, whose purpose it is to communicate effectively
+distinct entities, whose purpose is to communicate effectively
 about the actual state. In that case, evolving Voronoi-languages would
@@
 of prototypes as an equally efficient classification system. This is
-where sim-max games, might chip in a principled approach to deriving
-in an independent way, not only convex categories, but also
+where sim-max games might chip in a principled approach to deriving,
+in an independent way, not only convex categories but also
 prototypical exemplars belonging to them.
@@
 gradient, perhaps smooth, transition to negative examples to
-accommodate also ``higher-order vagueness'' ; (ii) prototypes should
+accommodate also higher-order vagueness; (ii) prototypes should
 likewise be gradient regions, peaking at the center of the vague
@@
 can be met by evolving strategies in sim-max games for certain types
-of solution concepts, that can be divided into micro- and macro-level
+of solution concepts which can be divided into micro- and macro-level
 approaches. Micro-level approaches look at adaptive behavior of

### 2014-10-07 paper/02_noisyRMD.tex
Small corrections to RDD section.

@@
 similar states to the replicator dynamic. Fix a sim-max game with
-$\States = \Acts$ and a confusion matrix $C$. $C$ is a row-stochastic
+$\States = \Acts$ and a confusion matrix $C : \States \times \States$. $C$ is a row-stochastic
 matrix whose elements $C_{ij}$ give the probability that $\state_i$ is
@@
 probability that \mystate{j} is the interpretation that is actually
-formed when \mystate{i} is the output of the receiver's strategy.
+formed when \mystate{i} is the intended interpretation.
 
@@
 
-Mixed strategies hold more information than behavioral
-strategies. This is because every mixed strategy $\Smixed$ converts to
-a unique behavioral strategy defined by:
+Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
 \begin{align*}
@@
 behavioral strategy. In this sense, mixed strategies hold more
-information about a distribution of pure strategies in a population.
+information than behavioral strategies about the distribution of pure strategies in a population.
 

### 2014-10-07 paper/03_simulations.tex
introduction and background; sec com
Small corrections to simulations section.

@@
 \section{Exploring the RDD}
+\label{sec:exploring-rdd}
 
@@
 
-The main idea behind Skyrm's notion is that we can quantify the amount
+The main idea behind Skyrms' notion is that we can quantify the amount
 of information about a state $\state$ in a message $\messg$ via the
@@
 convexity, that compensates for potential vagueness. To determine
-whether a sender strategy $\Sstrat$ was convex despite possibly being
-vague, we looked at the derived pure strategy $\Spure$ for which
+whether a sender strategy $\Sstrat$ is convex despite possibly being
+vague, we look at the derived pure strategy $\Spure$ for which
 $\Spure(\state) = \messg$ only if $\messg =
 \argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$
-was convex, we also counted $\Sstrat$ as convex. The sender strategy
+is convex, we also count $\Sstrat$ as convex. The sender strategy
 in Figure~\ref{fig:example_stratsA} is not convex, while the one in
@@
 
-\paragraph{Expected utility.} We also recorded the expected utility of
+\paragraph{Expected utility.} We also record the expected utility of
 a strategy pair:
@@
 To make direct comparisons across different parameter settings, we
-normalized expected utility by the maximal amount of expected utility
+normalize expected utility by the maximal amount of expected utility
 obtainable in the relevant game.
@@
 unexpected. Entropy of sender and receiver strategies are increasing
-in imprecision. Entropy of speaker strategies seems independent of
+with imprecision. Entropy of speaker strategies seems independent of
 tolerance levels and the number of states, while there seems to be an

### 2014-10-07 paper/03_simulations.tex
Speaker -> sender, hearer -> receiver

@@
 \citet{Jager2007:The-Evolution-o,JagerMetzger2011:Voronoi-Languag},
-(speaker) strategies can approach non-convex pure strategies and that
+(sender) strategies can approach non-convex pure strategies and that
 (ii) inclusion of imprecision can lead to seemingly well-behaved, yet
@@
 
-The speaker strategies in Figures~\ref{fig:example_stratsA} and
+The sender strategies in Figures~\ref{fig:example_stratsA} and
 \ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
@@
 \end{align*}
-This measure is bounded between $0$ and $1$. The speaker strategies in
+This measure is bounded between $0$ and $1$. The sender strategies in
 Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
@@
 
-\paragraph{Convexity.} At least for speaker strategies, which converge
+\paragraph{Convexity.} At least for sender strategies, which converge
 faster, it also makes sense to define a categorical measure of
@@
 unexpected. Entropy of sender and receiver strategies are increasing
-with imprecision. Entropy of speaker strategies seems independent of
+with imprecision. Entropy of sender strategies seems independent of
 tolerance levels and the number of states, while there seems to be an
-effect of both parameters on the entropy of the hearer
-strategy. Speaker informativity and expected utility are decreasing
+effect of both parameters on the entropy of the receiver
+strategy. Sender informativity and expected utility are decreasing
 under growing imprecision. There seem to be mild effects of the number
@@
 Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
-speaker strategies showed for several values of parameters in the
+sender strategies showed for several values of parameters in the
 absence of impairment. Increasing impairment always assured

### 2014-10-07 paper/04_discussion.tex
Speaker -> sender, hearer -> receiver

@@
 $\state$ is increased, but nothing else changes. In particular, the
-speakers behavior in other choice points does not change. Generalized
+senders behavior in other choice points does not change. Generalized
 reinforcement learning is different here. When the use of $\messg$ in
@@
 property ``tall''. If the usual interpretation of a ``tall man'' peaks
-at around, say, 195cm then when meeting a giant of $n$ meters speakers
+at around, say, 195cm then when meeting a giant of $n$ meters senders
 would, according to the quantal response approach, be ever more
@@
 zero. Whence that the quantal response approach would predict that
-speakers would grow indifferent between choice of antonyms as $n$
+senders would grow indifferent between choice of antonyms as $n$
 grows, which seems weird.

### 2014-10-07 paper/04_discussion.tex
Small corrections to discussion section.

@@
 allows, in principle, several conceptual interpretations of the \rdd,
-on which we will briefly enlarge below in
+on which we will briefly elaborate below in
 Section~\ref{sec:model-interpretation}.
@@
 \paragraph{Generalized reinforcement.}
-\citet{OConnor2013:The-Evolution-o} introduces a generalization of
-Herrnstein reinforcement learning for sim-max games and showed that
-this not only leads to vague signaling patterns, but can speed-up
-evolution of efficient signaling strategies, especially in games with
-many states.
+% Repeated below...
+%\citet{OConnor2013:The-Evolution-o} introduced a generalization of
+%Herrnstein reinforcement learning for sim-max games and showed that
+%this not only leads to vague signaling patterns, but can speed-up
+%evolution of efficient signaling strategies, especially in games with
+%many states.
 
@@
 expected utilities of saying ``short'' or ``tall'' will converge to
-zero. Whence that the quantal response approach would predict that
-senders would grow indifferent between choice of antonyms as $n$
-grows, which seems weird.
+zero. 
+%Whence that the quantal response approach would predict that
+%senders would grow indifferent between choice of antonyms as $n$
+%grows, which seems weird.
 

### 2014-10-16 paper/01_intro.tex
reworked Sections 1-3

@@
 cases that fall under it, clear cases that do not, and also so-called
-borderline cases in-between that neither clearly apply, nor clearly
-not apply, but might themselves not be equal in terms of how well they
-represent the category in question. Vagueness does not seem to
-dramatically affect the success of everyday communication, but it is
-troublesome for some of our theories of language. This is especially the case for the
-logico-semantic tradition of Frege, Russell and the young Wittgenstein
-which is challenged by the paradoxes vagueness gives rise to. As
-can be expected, many proponents of this tradition have tried to address the
-problem, ranging from suggesting that there is nothing wrong with the classical
-approach and the problem is of a different nature, as in Williamson's epistemic view~\citep*{Williamson1994:Vagueness},
-to proposing various kinds of modifications to either semantics, logic, or both, as in for example
-supervaluationism \citep{Mehlberg1958,Fine1975}, many-valued logic \citep{Zadeh1975,Machina1976,Edgington1997},
-or paraconsistent logic \citep{CobrerosEgre2012:Tolerant-Classi}.
+borderline cases. Borderline cases neither clearly apply, nor clearly
+not apply, and there may be differences between borderline cases in
+terms of how well they represent the category in question. Vagueness
+does not seem to dramatically affect the success of everyday
+communication, but it is troublesome for some of our theories of
+language. This is especially the case for the logico-semantic
+tradition of Frege, Russell and the young Wittgenstein which is
+challenged by the paradoxes vagueness gives rise to. As can be
+expected, many proponents of this tradition have tried to address the
+problem, ranging from suggesting that there is nothing wrong with the
+classical approach since the problem is of a different nature, as in
+Williamson's epistemic view~\citep*{Williamson1994:Vagueness}, to
+proposing various kinds of modifications to either semantics, logic,
+or both, as in for example supervaluationism
+\citep[e.g.][]{Mehlberg1958,Fine1975}, many-valued logic
+\citep[e.g.][]{Zadeh1975,Machina1976,Edgington1997}, or paraconsistent
+logic \citep[e.g.][]{CobrerosEgre2012:Tolerant-Classi}.
 
@@
 resisting the pressure of evolutionary selection for precision. Such
-reasons and general mechanisms are, arguably, not hard to imagine. We
-will address some presently. But each putative explanation, no matter
-how plausible intuitively, must also be checked for internal
-consistency and its potential to account for the fact that we see
-vagueness as a part of a by-and-large efficient system of
-categorization and communication. In other words, not all first-shot
-rebuttals of Lipman's problem will do. Denying that there is any
-functional pressure on efficient communication, for instance, is a
-non-starter, because it leaves the relative efficiency of our
-communication with vague words entirely unexplained. In sum, the
-technical answer to Lipman's problem is to give, ideally, a model of
-evolution of categorization or language use that shows how competing
-forces that realize both by-and-large efficient categories and
+reasons and general mechanisms are, arguably, not hard to imagine.
+But each putative explanation, no matter how plausible intuitively,
+must also be checked for internal consistency and its potential to
+account for the fact that we see vagueness as a part of a by-and-large
+efficient system of categorization and communication. In other words,
+not all first-shot rebuttals of Lipman's problem will do. Denying that
+there is any functional pressure on efficient communication, for
+instance, is a non-starter, because it leaves the relative efficiency
+of our communication with vague words entirely unexplained. In sum,
+the technical answer to Lipman's problem is to give, ideally, a
+conceptually sound model of evolution of categorization or language
+use that yields by-and-large efficient categories \emph{and}
 vagueness.
@@
 paper makes a contribution to the latter line of thought.  Concretely,
-we introduce a novel variant of the replicator mutator dynamic that
-integrates stochastic noise on the differential confusability of
-similar stimuli. Our main contribution, therefore, is a technical
-answer to Lipman's problem in the sense introduced above. The dynamic
-proposed here nicely generalizes on and complements previous like-minded
-accounts. We show that stochastic noise can not only lead to
-vague, yet communicative efficient signal use, but can also unify
-evolutionary outcomes and help avoid inefficient categorization.
+we introduce the replicator diffusion dynamics ---a novel variant of
+the replicator mutator dynamic--- that integrates stochastic noise on
+the differential confusability of similar stimuli. Our main
+contribution, therefore, is a technical answer to Lipman's problem in
+the sense introduced above. The dynamic proposed here generalizes and
+complements previous like-minded accounts. We show that stochastic
+noise can not only lead to vague, yet communicative efficient signal
+use, but can also unify evolutionary outcomes and help avoid
+suboptimal categorization.
 
@@
 using evolutionary game theory and certain generalizations of
-signaling games, so called similarity-maximizing games or sim-max
-games, for short
+signaling games, so called \emph{similarity-maximizing games}, or
+\emph{sim-max games}, for short
 \citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}. Our
@@
 the receiver's interpretation is to the actual state, the more
-successful the play was. Signaling games with utility-relevant
-similarities in the state space are fairly standard in economics
+successful the play is taken to be. Signaling games with
+utility-relevant similarities in the state space are fairly standard
+in economics
 \citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
 but have received particular attention in a more philosophical and
-linguistic context by the more recent work for reasons that will
-become clear presently
+linguistic context for reasons that will become clear presently
 \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.
 
-More concretely, a sim-max game, in the relevant sense here, consists
-of a set of states $\States$, a set of messages $\Messgs$ with much
-fever messages than states, a probability distribution $\Pr \in
+A sim-max game, in the relevant sense here, consists of a set of
+states $\States$, a set of messages $\Messgs$ with much fever messages
+than states, a prior probability distribution $\Pr \in
 \Delta(\States)$ that gives the occurrence probabilities of states, a
@@
 \end{align*}
-To keep matters simple, we only look at cases where $\States$ contains
-finitely many points from the unit interval, all of which have equal
-probability, and with similarity given by the Euclidean distance in
-this one-dimensional space.
+To keep matters simple, we here only look at cases where $\States$
+contains finitely many points from the unit interval, all of which
+have equal probability, and with similarity given by the Euclidean
+distance in this one-dimensional space.
 
@@
 central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
-an example of such a Voronoi-language for a sim-max game with
-$\States = [0,1]$, two messages, and a flat prior distribution that
-selects all states equally likely. The sender's strategy (a function
-from states to messages; more below) uses one message exclusively for
-all points in the lower half of the interval and another for all points in the
-upper half. The receiver's interpretations of messages are the
-central points in the respective intervals.
+an example of such a Voronoi-language for a sim-max game with $\States
+= [0,1]$, two messages, and a flat prior distribution over states. The
+sender uses one message exclusively for all points in the lower half
+of the interval and another for all points in the upper half. The
+receiver's interpretations of messages are the central points in the
+respective intervals.
 
@@
 
-  \caption{Example of a Voronoi-language on $\States = [0,1]$. The
-    sender strategy $\Spure$ uses one signal for the lower half, and
-    another for the upper half of the unit interval. The receiver
-    strategy $\Rpure$ interprets messages as the central elements in the
-    respective intervals.}
+        \caption{Example of a Voronoi-language on $\States =
+          [0,1]$. The (pure) sender strategy $\Spure \in
+          \Messgs^\States$ uses one signal for the lower half, and
+          another for the upper half of the unit interval. The (pure)
+          receiver strategy $\Rpure \in \States^\Messgs$ selects the
+          central elements in the respective intervals.}
   \label{fig:VoronoiLanguage}
@@
 
-This result is particularly interesting, because it demonstrates that
-signaling can impose orderly categories on a metric space, without
-that being the ulterior purpose of it all. Sender and receiver can be
-distinct entities, whose purpose is to communicate effectively
-about the actual state. In that case, evolving Voronoi-languages would
-explain why linguistic categories are well-behaved and orderly in the
-way they appear to be. More abstractly, sender and receiver can also
-be thought of as distinct modules in a single system, where the first
-module that must discretize the information that it is fed into a
-small sample of, suggestively, category labels to pass to a second
-system that tries to decode the original information. In this case,
-evolving Voronoi-languages would explain why conceptual categories are
+This result is interesting, because it demonstrates that signaling can
+impose orderly categories on a metric space, without that being the
+ulterior purpose of it all. Sender and receiver can be distinct
+entities, whose purpose is to communicate effectively about the actual
+state. In that case, evolving Voronoi-languages would explain why
+linguistic categories are well-behaved and orderly in the way they
+appear to be. More abstractly, sender and receiver can also be thought
+of as distinct modules in a single system, where the first module must
+discretize the information by selecting a small sample of,
+suggestively, category labels. These are passed to a second module
+that tries to decode the original information. In this case, evolving
+Voronoi-languages would explain why conceptual categories are
 well-behaved and orderly in the way that they appear to be
-\citep[see][for more on this latter
+\citep[e.g.][for more on this latter
 interpretation]{OConnor2013:Evolving-Percep}.
@@
 \citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has
-prominently argued that natural categories are convex regions in a
-suitable conceptual space. If the conceptual space has a suitable
-metric, then it is possible to think of these convex categories as
-derived from a set of prototypes. Fixing a set of prototypes, we
-consider the category corresponding to each prototype $p$ as the set
-of points that are more similar to $p$ than to any other
+prominently argued that natural categories are convex regions in
+conceptual space. If the conceptual space has a suitable metric, then
+it is possible to think of these convex categories as derived from a
+set of prototypes. Fixing a set of prototypes, we consider the
+category corresponding to each prototype $p$ as the set of points that
+are more similar to $p$ than to any other
 \citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
@@
 categorization system can be obtained: storing the prototypes lets us
-recover the categories without having to store each categories
+recover the categories without having to store each category's
 extension. However, what is left unexplained so far, is where the
@@
 of prototypes as an equally efficient classification system. This is
-where sim-max games might chip in a principled approach to deriving,
+where sim-max games can contribute a principled approach to deriving,
 in an independent way, not only convex categories but also
@@
 
-The brief outline of an approach to conceptual categorization using
+This brief outline of an approach to conceptual categorization using
 sim-max games leaves many problems unaddressed. One of them is that
@@
 color, pitch heights, spatial dimension and the like, do not have
-unique prototypes and clear category boundaries. We would like to
-account for the possibility of vagueness, in particular the following
-criteria
+unique, point-valued prototypes and clear category boundaries. We
+would like to account for the possibility of such vagueness, and that
+means in particular for the following criteria
 \citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re,Smith2008:Vagueness-and-D}:
 (i) clear positive examples of a vague category should show a
-gradient, perhaps smooth, transition to negative examples to
+gradient, perhaps smooth, transition to clear negative examples to
 accommodate also higher-order vagueness; (ii) prototypes should
@@
 work, \citet{DecockDouven2012:What-is-Graded-} show further how it is
-possible to arrive at ``higher-order vagueness'' and degrees of
-category membership, by, essentially, weighing in the distance of
-different borderline examples to various prototypical regions. This
-accounts for the first of the two desiderata mentioned above, but
-still assumes that crisp non-gradient prototype regions must be
-pre-given.
+possible to arrive at higher-order vagueness and degrees of category
+membership, by, essentially, weighing in the distance of different
+borderline cases to various prototypical regions. This accounts for
+the first of the two desiderata mentioned above, but still assumes
+that crisp non-gradient prototype regions must be pre-given.
 
@@
 \citet{OConnor2013:The-Evolution-o} who show how the above desiderata
-can be met by evolving strategies in sim-max games for certain types
-of solution concepts which can be divided into micro- and macro-level
+can be met by evolving strategies in sim-max games for various types
+of solution concepts. These can be divided into micro- and macro-level
 approaches. Micro-level approaches look at adaptive behavior of
@@
 approaches outline more abstract, aggregate dynamics, happening in a
-huge population of agents, or otherwise abstracting from seemingly
+population of agents, or otherwise abstracting from seemingly
 irrelevant detail. Usually, a macro-level dynamic captures changes of
@@
 users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
-reinforcement learning that includes a low-level form of stimulus
+reinforcement learning that entails a low-level form of stimulus
 generalization. Agents update their behavior after each round of play
-in such a way that similar states to the ones that actually occurred
-are subject to behavioral adjustment as well. (Alternative approaches
-are discussed in more detail in Section~\ref{sec:discussion}, where we
-can compare them with the system introduced in this paper.)
+in such a way that states similar to the one that actually occurred
+are subject to behavioral adjustment as well. (A more detailed
+discussion of alternative approaches is deferred until
+Section~\ref{sec:discussion}, where we can compare them better with the
+system introduced in this paper.)
 
@@
 give rise to counterintuitive predictions. These counterintuitive
-examples suggest that vagueness in signaling use is not convincingly
+examples suggest that vagueness in sim-max games is not convincingly
 explained by appeal to mistakes in calculating expected utility, but
-rather, as we assume here, as the result of confusing similar states
-of the word.
-
-Against this background, we introduce a new macro-level approach to
-the evolution of vague signal use in sim-max games, where the source
-of vagueness is the agents' natural inability to sharply distinguish
-perceptually distinct states. This applies to the sender as well as to
-the receiver role (albeit in slightly different manner). We call our
-dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
-special case of the replicator mutator dynamic (\rmd). It is a special
-case in the sense that the \rdd incorporates a particular kind of
-mutation that is based on the similarity of states (as defined in the
-sim-max game). More concretely, we start from the idea that similar
-states are more easily confused than less similar states. To formalize
-this idea, we assume that there is a confusion matrix $C$, where
-$C_{ij}$ is the probability that a state \mystate{i} is realized as a
-possibly different state \mystate{j}. (There is a slightly different
-sense of ``realization'' in the sender and the receiver side, and we
-will enlarge on this below.) Based on this, we show two things. For
-one, we show how the discrete-time formulation of the replicator
-dynamic can be easily adapted to integrate diffusion of behavior due
-to confusion of states. This yields the \rdd in its basic
-discrete-time formulation. For another, we also show that there exists
-a translation of the confusion matrix $C$ into a suitable mutation
-matrix $M_C$ such that the effect of $C$ and $M_C$ are ``equivalent''
-in a sense to be made precise. This supports the conclusion that the
-\rdd can indeed be thought of as a special case of the \rmd.
-
-Based on data from numerical simulations, we show that the inclusion
-of mild levels of such stochastic noise in the replicator dynamics
-does not disrupt the possibility of evolving communicative efficient
-signaling strategies at all. On the contrary, there might even be a
-higher-order benefit to the presence of perceptual imprecision, in
-that it accelerates converges, but also unifies and regularizes
-evolutionary outcomes in such a way that inefficient categorization is
-avoided. These results complement similar results by
-\citet{OConnor2013:The-Evolution-o}, obtained for a micro-dynamic
-extension of reinforcement learning. The \rdd adds to this a more
-general, abstract framework that is not tied to the specific
-assumptions of turn-based reinforcement.
+rather, as we assume here, as the result of confusing similar
+states.
 

### 2014-10-16 paper/02_noisyRMD.tex
reworked Sections 1-3

@@
 
-The replicator diffusion dynamic (\rdd) was first introduced by
-\citet{Correia2013:The-Bivalent-Tr} as a noise-perturbed variant of
-the replicator dynamic (\rd) in behavioral strategies. We begin by
-recapitulating the formulation of the \rd in behavioral
-strategies. Then we define the replicator diffusion dynamic in
-behavioral strategies, following \citet{Correia2013:The-Bivalent-Tr}.
-Finally we show that the \rdd is a special case of the replicator
-mutator dynamic \citep[e.g.][]{Nowak2006:Evolutionary-Dy}.
+Against this background, we introduce a new macro-level approach to
+the evolution of vague signal use in sim-max games, where the source
+of vagueness is the agents' natural inability to sharply distinguish
+similar states. This applies to the sender as well as to
+the receiver role (albeit in slightly different manner). We call our
+dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
+special case of the replicator mutator dynamic (\rmd).
+
+The \rdd was first introduced by \citet{Correia2013:The-Bivalent-Tr}
+as a noise-perturbed variant of the replicator dynamic (\rd) in
+behavioral strategies. We begin by recapitulating this formulation of
+the \rd in Section~\ref{sec:repl-dynam-behav}. Then we introduce the
+\rdd in Section~\ref{sec:repl-diff-dynam-1}.
+Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
+Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
+a special case of the \rdd.
 
 \subsection{Replicator dynamic in behavioral strategies}
+\label{sec:repl-dynam-behav}
 
@@
 functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
-utility of actions in each choice point are as usual:
+utility of choices in each choice point as usual:
 \begin{align*}
@@
 similar states to the replicator dynamic. Fix a sim-max game with
-$\States = \Acts$ and a confusion matrix $C : \States \times \States$. $C$ is a row-stochastic
-matrix whose elements $C_{ij}$ give the probability that $\state_i$ is
-realized as $\state_j$. The confusability of states affects senders
-and receivers alike, but in slightly different ways (see
+$\States = \Acts$ and a confusion matrix $C : \States \times
+\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
+the probability that $\state_i$ is realized as $\state_j$. The
+confusability of states affects senders and receivers alike, but in
+slightly different ways (see
 Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
-$C_{ij}$ is the probability that the actual states \mystate{i} is
+$C_{ij}$ is the probability that the actual state \mystate{i} is
 perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
@@
 percolates to behavior at similar choice points, we speak of
-\emph{diffusion of behavior under confusion of states} (see
-Section~\ref{sec:iterated-diffusion} for an investigation into this
-diffusion process). If we conceive of behavioral strategies $\Sstrat$
-and $\Rstrat$ as row-stochastic matrices, the diffusion effect of
-state confusability is easily captured by matrix multiplication as a
-diffusion function $\Diff_C$ that, based on confusion matrix $C$, maps
-behavioral strategies on their diffused realizations:
+\emph{diffusion of behavior under confusion of states}. If we
+conceive of behavioral strategies $\Sstrat$ and $\Rstrat$ as
+row-stochastic matrices, the diffusion effect of state confusability
+is easily captured by matrix multiplication as a diffusion function
+$\Diff_C$ that, based on confusion matrix $C$, maps behavioral
+strategies onto their diffused realizations:
 \begin{align}
@@
 fitness landscape, as described by the \rd, but the realization of
-optimized strategies is noisy, due to confusion of states. It is
+optimized strategies is noisy, due to similarity of states. It is
 obvious that other (discrete-time) evolutionary dynamics can be
@@
 \subsection{The replicator mutator dynamic}
+\label{sec:repl-mutat-dynam}
 
-The replicator mutator dynamic (\rmd) has been proposed first in the
-context of signaling game models for the evolution of grammar
-\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. The
-relation of the \rmd to other prominent evolutionary dynamics is well
-understood \citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen
-further fruitful applications in the context of signaling games
+The \rmd has been proposed first in the context of signaling game
+models for the evolution of grammar
+\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. It
+adds to the replicator dynamic stochastic mutation that captures, in
+the context of grammar evolution, the differential learning success of
+different dispositions in first language acquisition. The relation of
+the \rmd to other prominent evolutionary dynamics is well understood
+\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
+fruitful applications in the context of signaling games
 \citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy,Deo2014:The-Semantic-an}. It
@@
 of the former. We write $\Smixed_i$ for the frequency
-$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$ (also for the
-receiver).
+$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
+receiver.
 
@@
 adjusting their strategies locally at each choice point, regardless of
-what they are bound to do at other choice points. The latter
+what they are disposed to do at other choice points. The latter
 perspective is then more coarse-grained, and not necessarily
-equivalent to the former. An advantage of dynamics in behavioral
+equivalent to the former. But an advantage of dynamics in behavioral
 strategies is that they reduce complexity, which is especially helpful
@@
 \paragraph{Replicator dynamic in mixed strategies.} To understand the
-replicator mutator dynamic in mixed strategies, we first introduce the
-replicator dynamic in mixed strategies as introduced by
-\citet{TaylorJonker1978:Evolutionary-St}.
-
-Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given $\Rmixed$ and
-$F_i^{\Smixed}$ be $\Rpure_i$'s fitness given $\Smixed$. Then
-$\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot F_k^{\Rmixed}$ is
-the average fitness in the sender population and
+\rmd in mixed strategies, we first introduce the \rd in mixed
+strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
+$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
+$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
+F_k^{\Rmixed}$ is the average fitness in the sender population and
 $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
@@
 of mixed strategies. Its so-called two-population version, that takes
-sender and receiver roles into account separately, is defined as:
+sender and receiver roles as separate update sites:
 \begin{align*}
@@
 Similar to the version for behavioral strategies, a discrete-time
-version of the \rd for mixed strategies can be conceptualized as an
+formulation of the \rd for mixed strategies can be conceptualized as an
 update function that maps mixed strategies $\Smixed$ and $\Rmixed$
@@
 \end{align*}
-Although a discrete-time formulation of the dynamics is absent from the
-literature, as far as we know, it is easy to spell out, and clearly
-demonstrates that the sequential nature of the \rmd, parallel to the
-intuition that motivated the formulation of the \rdd.
+A discrete-time, two-population formulation of the \rmd is this 
+\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
 
-\begin{claim} A discrete time version of the replicator mutator
-  dynamic is:
 \begin{align*}
@@
 \end{align*}
-\end{claim}
 
-\begin{proof}
-  It suffices to check either sender or receiver part. Focusing on the
-  former, we need to show that the continuous version is derivable
-  from the discrete version if the discrete update steps get
-  infinitely small so that:
-  \begin{align*}
-    \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
-    \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
-    \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
-    & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
-  \end{align*}
-  As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
-  it to derive the non-payoff adjusted continuous version above,
-  since:
-  \begin{align*}
-    & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
-      ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
-        F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
-        \Phi(\Smixed,\Rmixed) \\
-       = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
-        F_j^{\Rmixed} - \Smixed_i \cdot
-        \Phi(\Smixed,\Rmixed) 
-  \end{align*}
-\end{proof}
+% \begin{proof}
+%   It suffices to check either sender or receiver part. Focusing on the
+%   former, we need to show that the continuous version is derivable
+%   from the discrete version if the discrete update steps get
+%   infinitely small so that:
+%   \begin{align*}
+%     \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
+%     \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
+%     \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
+%         F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
+%     & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
+%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
+%   \end{align*}
+%   As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
+%   it to derive the non-payoff adjusted continuous version above,
+%   since:
+%   \begin{align*}
+%     & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
+%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
+%       ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
+%         F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
+%         \Phi(\Smixed,\Rmixed) \\
+%        = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
+%         F_j^{\Rmixed} - \Smixed_i \cdot
+%         \Phi(\Smixed,\Rmixed) 
+%   \end{align*}
+% \end{proof}
 
@@
 mutation. This becomes clear if we define independent mutation
-functions $\Mutate_{Q,R}$ that maps mixed strategies onto the outcome
+functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
 of (one round of) mutation:
@@
 \subsection{Diffusion as a special kind of mutation}
+\label{sec:diffusion-as-special}
 
@@
 definable for behavioral strategies (see
-Equations~(\ref{eq:confusion-function})), while mutation in the \rmd is
-defined as mutation from one pure strategy to another. Nonetheless, it
-is possible to map each confusion matrix to a mutation matrix in such
-a way that the natural correspondence between strategies is conserved
-(see Figure~\ref{fig:correspondence-result}). This is the main result
-presented in this section, and it is our justification for regarding
-the \rdd as the behavioral-strategy analogue of the replicator-mutator
-dynamic when the only source of mutation is perceptual confusion of
-states.
+Equations~(\ref{eq:confusion-function})), while mutation in the \rmd
+is defined as mutation from one pure strategy to another. Nonetheless,
+it is possible to map each confusion matrix to a mutation matrix in a
+natural way so that the correspondence between mixed and behavioral
+strategies are conserved. This is the main result presented in this
+section, and it is our justification for regarding the \rdd as the
+behavioral-strategy analogue of the replicator-mutator dynamic when
+the only source of mutation is confusability of states.
 
@@
 have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver. This
-is the content of Theorems~\ref{thm:sender-eq} and
-~\ref{thm:receiver-eq}, the proofs of which are in
-Appendix~\ref{sec:proofs}. The remainder of this section introduces
-the relevant construction that maps arbitrary $C$ to suitable $Q^C$.
+is the content of Theorem~\ref{thm:Correspondence} given below, a
+proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
+section introduces the relevant construction that maps arbitrary $C$
+to suitable $Q^C$.
 
@@
 
-The probability that $\Rpure_i$ mutates into $\Rpure_j$ is the product of
-the probabilities for each $\messg$ that the state $\Rpure_i(\messg)$ is
-perceived as state $\Rpure_j(\messg)$. Abusing notation by referring to
-the indices of states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
+The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
+diffusion with $C$ is the product of the probabilities, for each
+$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
+$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
+states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
 $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
@@
   \label{eq:construction-rec}
-  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j\messg)}\,.
+  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
 \end{align}
 
-Now look at the sender. If $\Spure_j(\state_k)=\messg$ and
-$\Spure_i(\state_k)=\messg'$, then the probability of confusion at
-state $\state_k$ is given by the chance that $\state_k$ is perceived
-as a state $\state_l$ which $\Spure_j$ would map onto $\messg'$. The
-overal probability that $\Spure_j$ mutates into $\Spure_i$ is the
-product of these chances for all states. So, define:
+Now look at the sender. The probability that $\Spure_j$ is realized as
+$\Spure_i$ under diffusion with $C$ is the product of the
+probabilities, over all states $\mystate{k}$, that the message
+$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
+$\mystate{k}$ in the absence of noise, is produced by a noisy
+realization of $\Spure_j$, which is the probability $\sum_{\state_l
+  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
+$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
+would map unto $\Spure_i(\mystate{k})$. So, define:
 \begin{align}
-  \label{eq:construction-rec}
+  \label{eq:construction-sen}
   Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
@@
 Here, a pair like $21$, for example, refers to a pure sender strategy
-with $\Spure(\state_1=\messg_2$ and $\Spure(\state_2) =
+with $\Spure(\state_1)=\messg_2$ and $\Spure(\state_2) =
 \messg_1$. Similarly for the receiver.
@@
 matrices $Q^C$ and $R^C$ in Equations~(\ref{eq:construction-rec}) and
-(\ref{eq:construction-rec}), we can finally formulate the
+(\ref{eq:construction-sen}), we can finally formulate the
 desired correspondence result, that reveals the \rdd as a special case
-of the \rmd in behavioral strategies where the source of mutation is
+of the \rmd in behavioral strategies where the only source of mutation is
 imperfect discriminability of states.
@@
 \begin{theorem}
-  \label{thm:sender-eq}
-  If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) = \Diff_C(\Sstrat)$.
+  \label{thm:Correspondence}
+  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
+  \Diff_C(\Sstrat)$. And, (ii) if $G(\Rmixed) = \Rstrat$, then
+  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
 \end{theorem}
 
-\begin{theorem}
-  \label{thm:receiver-eq}
-  If $G(\Rmixed) = \Rstrat$, then $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
-\end{theorem}
-
-\noindent Proofs are given in Appendix~\ref{sec:proofs}.
+\noindent A proof is given in Appendix~\ref{sec:proofs}.
 

### 2014-10-16 paper/05_conclusions.tex
paper skeleton first content
introduction and background; sec com
reworked Sections 1-3

--- 2014-05-10 paper/03_simulations.tex
@@
-\section{Simulations}
+\section{Conclusions}
+\label{sec:conclusions}
+
+It is a special case in the sense that the \rdd incorporates a
+particular kind of mutation that is based on the similarity of states
+(as defined in the sim-max game). More concretely, we start from the
+idea that similar states are more easily confused than less similar
+states. To formalize this idea, we assume that there is a confusion
+matrix $C$, where $C_{ij}$ is the probability that a state \mystate{i}
+is realized as a possibly different state \mystate{j}. (There is a
+slightly different sense of ``realization'' in the sender and the
+receiver side, and we will enlarge on this below.) Based on this, we
+show two things. For one, we show how the discrete-time formulation of
+the replicator dynamic can be easily adapted to integrate diffusion of
+behavior due to confusion of states. This yields the \rdd in its basic
+discrete-time formulation. For another, we also show that there exists
+a translation of the confusion matrix $C$ into a suitable mutation
+matrix $M_C$ such that the effect of $C$ and $M_C$ are ``equivalent''
+in a sense to be made precise. This supports the conclusion that the
+\rdd can indeed be thought of as a special case of the \rmd.
+
+Based on data from numerical simulations, we show that the inclusion
+of mild levels of such stochastic noise in the replicator dynamics
+does not disrupt the possibility of evolving communicative efficient
+signaling strategies at all. On the contrary, there might even be a
+higher-order benefit to the presence of perceptual imprecision, in
+that it accelerates converges, but also unifies and regularizes
+evolutionary outcomes in such a way that inefficient categorization is
+avoided. These results complement similar results by
+\citet{OConnor2013:The-Evolution-o}, obtained for a micro-dynamic
+extension of reinforcement learning. The \rdd adds to this a more
+general, abstract framework that is not tied to the specific
+assumptions of turn-based reinforcement.
 

### 2014-10-16 paper/paper.tex
reworked Sections 1-3

@@
   only lead to vague, yet communicative efficient signal use, but can
-  also unify evolutionary outcomes and help avoid inefficient
+  also unify evolutionary outcomes and help avoid suboptimal
   categorization.

### 2014-10-16 paper/03_simulations.tex
revised sections 3 - 5

@@
 two-step procedure: we first calculate where the replicator dynamic
-would take us, then apply diffusion. Since the latter step is the
-novel addition, we should look at diffusion in isolation to understand
-what the \rdd does. This is what we will do in
-Section~\ref{sec:iterated-diffusion}. Section~\ref{sec:simulations}
- then explores how diffusion and fitness-based selection interact.
+would take us, then apply diffusion. To understand the \rdd
+better, Section~\ref{sec:iterated-diffusion} looks closer at the novel
+diffusion part, before Section~\ref{sec:simulations} explores how
+diffusion and fitness-based selection interact.
 
@@
 Confusion of states should intuitively be a function of their
-perceptual similarity. Here and in the following, we assume that the
+perceptual similarity. To make this concrete, let us assume that the
 state space of the sim-max game consists of $n \ge 2$ states that are
-equally spaced across the unit interval, including $0$ and $1$. We
-assume here that the distance $\card{\mystate{i} - \mystate{j}}$ is
-the objective, physical similarity between two states $\mystate{i}$
-and $\mystate{j}$. Distance in physical space feeds into a perceptual
+equally spaced across the unit interval, including $0$ and $1$. The
+distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical
+similarity between two states $\mystate{i}$ and
+$\mystate{j}$. Distance in physical space feeds into a perceptual
 similarity function, as described by
@@
 states; when $\impairment \rightarrow \infty$ agents cannot
-discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
-impression of Nosofsky-similarity for different parameter
-values. Other formalizations of perceptual similarity are possible,
-including ones that allow for different discriminability in different
-areas of the state space, but we stick with Nosofsky's similarity
-function for the time being, because it is a mathematically simple,
-yet established notion in mathematical psychology.
+discriminate states at all. The top part of
+Figure~\ref{fig:NosofskySim} gives an impression of
+Nosofsky-similarity for different parameter values. Other
+formalizations of perceptual similarity are possible, including ones
+that allow for different discriminability in different areas of the
+state space, but we stick with Nosofsky's similarity function for the
+time being, because it is a mathematically simple, yet established
+notion in mathematical psychology.
 
@@
 $\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
-= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$ $C$ is
+= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$ is
 positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
@@
 
-Difusion of behavior under confusion of states was defined in
+Diffusion of behavior was defined in
 Equation~(\ref{eq:confusion-function}), repeated here:
@@
 diffusion component takes over and the \rdd reduces to the trivial
-iterated diffusion that we looked at in the previous section. We will
-show presently that for reasonable in-between levels of imprecision
-$\impairment$, the \rdd leads to communicatively successful, yet vague
-signal meaning. Suitable levels of imprecision can have further
-accelerating and unifying effects on meaning evolution.
+iterated diffusion process that we looked at in the previous
+section. We will show presently that for reasonable in-between levels
+of imprecision $\impairment$, the \rdd leads to communicatively
+successful, yet vague signal meaning. Suitable levels of imprecision
+can have further accelerating and unifying effects on meaning
+evolution.
 
@@
 
-To explore the \rdd, we looked at numerical simulations in sim-max
-games with two signals whose states were evenly spaced elements of the
-unit interval that always included 0 and 1. We assumed that priors
-were flat: $\Pr(\state) = \Pr(\state')$ for all $\state, \state'$. As
-for utility functions, we implemented another Nosofsky-style function:
+To explore the \rdd, we turn to numerical simulations. Let states be
+evenly spaced elements of the unit interval that always include 0 and
+1, and let priors be flat: $\Pr(\state) = \Pr(\state')$ for all
+$\state, \state'$. As for utility functions, we take another
+Nosofsky-style function:
 \begin{align}
@@
 \end{align}
-The tolerance parameter $\toler > 0$ models the amount of tolerable
+The tolerance parameter $\toler \ge 0$ models the amount of tolerable
 pragmatic slack or communicative imprecision. This choice of utility
@@
 utilities that are, say, linearly or quadratically decreasing in
-physical distance \citep[as used
-by][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
+physical distance \citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
 utilities that are exponentially decreasing in negative quadratic
@@
 shows a strategy pair at stopping time with 90 states, tolerance
-$\toler = 0.1$ and imprecision $\impairment = 0$. The latter means
-that the trial was effectively a plain application of the
+$\toler = 0.1$ and imprecision $\impairment = 0$. Zero imprecision
+means that the trial was effectively an application of the plain
 \rd. Noteworthily, the given sender strategy approximates a pure
@@
 dynamically stable state under the \rd. Indeed, the trial did not
-converge in our technical sense, but was stopped after 200 rounds of
-iterated \rdd. In contrast, the outcome of a trial with identical
-parameters, except that imprecision was $\impairment = 0.1$, is shown
-in Figure~\ref{fig:example_stratsB} had converged (in our technical
-sense) after 99 rounds of iterated \rdd. The sender strategy shows a
-smooth blending from one ``category'' to the other, and the receiver's
-interpretations are rather extended curves, peaking at a central point
-in the relevant ``categories''. What these examples show is that (i),
-contrary to the limiting results of
-\citet{Jager2007:The-Evolution-o,JagerMetzger2011:Voronoi-Languag},
-(sender) strategies can approach non-convex pure strategies and that
-(ii) inclusion of imprecision can lead to seemingly well-behaved, yet
-vague strategies in the sense that we are after.
+converge in our technical sense, but was stopped after the maximum of
+200 rounds. In contrast, the outcome of a trial with identical
+parameters, except that imprecision was $\impairment = 0.1$, which is
+shown in Figure~\ref{fig:example_stratsB}, had converged (in our
+technical sense) after 99 rounds of iterated \rdd. The sender strategy
+shows a smooth blending from one ``category'' to the other, and the
+receiver's interpretations are rather extended curves, peaking at a
+central point in the relevant ``categories''.
+
+These examples already show two interesting things. Firstly, inclusion
+of imprecision can lead to seemingly well-behaved, yet vague
+strategies in the sense that we are after. The sender strategy in
+Figure~\ref{fig:example_stratsB} identifies clear positive and clear
+negative cases for each signal, with a smooth transition
+in-between. The receiver's interpretations of signals can be seen as
+smoothed out prototype regions.
+
+Secondly, in contrast to results concerning evolutionary stability in
+sim-max games
+\citep{Jager2007:The-Evolution-o,JagerMetzger2011:Voronoi-Languag},
+(sender) strategies can approach non-convex pure strategies under the
+replicator dynamic. We see this in our numerical simulations for the
+non-limiting cases, but this also holds, for some types of utility
+functions, for the limiting case. This was first observed by Elliott
+Wagner in unpublished work, and is further discussed by
+\citet{OConnor2013:Evolving-Percep}. Our simulation results do not
+speak directly to limiting results, but we will see shortly that
+diffusion from confusability of states clearly prevents evolutionary
+paths that meander for a long time in the vicinity of non-convex
+strategies.
 
@@
 capture numerically how vague, communicatively efficient and generally
-well-structured the resulting strategy pairs were. Entropy and
-informativity capture (slightly different aspects of) the amount of
-systematicity or regularity in a language. Voronoiness and convexity
-are, respectively, gradient and categorical measures for the coherence
-of linguistically formed categories. Expected utility measures the
-communicative efficiency of evolved strategy pairs.
+well-structured the resulting strategy pairs were. \emph{Entropy} and
+\emph{informativity} capture (slightly different aspects of) the
+amount of systematicity or regularity in signal
+use. \emph{Voronoiness} and \emph{convexity} are, respectively,
+gradient and categorical measures for the coherence of signal-induced
+categories. \emph{Expected utility} measures the communicative
+efficiency of evolved strategy pairs.
 
-\paragraph{Entropy.} This classic information-theoretic metric
-captures the amount of uncertainty in a signaling strategy. Roughly
-put, entropy of a strategy captures inverse distance from a pure
-strategy. The most natural way to define it is in terms of mixed
-strategies rather than behavioral strategies.  Let us recall that
-mixed sender (receiver) strategies are functions $\Smixed \in
-\Delta(\Messgs^\States)$ ($\Rmixed \in (\Acts^\Messgs)$).  We can
-define the entropy of a sender and receiver strategy as
+\paragraph{Entropy.} This classic information-theoretic notion
+captures the amount of uncertainty in a probability
+distribution. Roughly put, entropy of a signaling strategy captures
+inverse distance from a pure strategy. The usual definition of entropy
+applies directly to mixed strategies:
 \begin{align*}
@@
   \log(\Smixed(\Spure)) & 
-  E(\Rmixed) & = \sum_{\Rpure \in \Acts^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
+  E(\Rmixed) & = \sum_{\Rpure \in \States^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
 \end{align*} 
@@
 size of the domain over which the sum is computed grows exponentially
-with the number of choice points.  Therefore, we
-converted\footnote{See Appendix~\ref{sec:proofs}.}\todo{include
-  this!?} the above definitions into equivalent metrics defined in
-terms of behavioral strategies.  These are as follows:
+with the number of choice points. But provably equivalent metrics for behavioral
+strategies are ready to hand:
 \begin{align*}
@@
 \begin{align*}
-  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\act \in \Acts}
-  \Rstrat(\act \probbar \messg) \cdot \log(\Rstrat(\act \probbar
+  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
+  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
   \messg)) \,. 
@@
 upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
-\cdot \log(|\Messgs|)$ and $\log(|\Acts^\Messgs|) = |\Messgs| \cdot
-\log(|\Acts|)$. We work with normalized values.
+\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
+\log(|\States|)$. We work with normalized values.
 
@@
 of a message $\messg \in \Messgs$, then
-\begin{multline*}
-  V(\Sstrat, \Rstrat) = \sum_{\state \in \States} \sum_{\messg_1 \in
-    \Messgs} \sum_{\messg_2 \in \Messgs} \textrm{if } s(\state,
-  \pi(\Rstrat, \messg_1)) > s(\state, \pi(\Rstrat, \messg_2)) \bicond
-  \Sstrat(\state, \messg_1) > \Sstrat(\state, \messg_2) \textrm{ then
-  } 1 \textrm{ else } 0 
-\end{multline*}
+\begin{align*}
+  V(\Sstrat, \Rstrat) & = \sum_{\state \in \States} \sum_{\messg_1 \in
+    \Messgs} \sum_{\messg_2 \in \Messgs} F(\Sstrat, \Rstrat, \state,
+  \messg_1, \messg_2) \text{\ \ \ with } \\
+  F(\Sstrat, \Rstrat, \state,
+  \messg_1, \messg_2) & =
+  \begin{cases}
+    1 & \text{if } \similarity(\state, \pi(\Rstrat, \messg_1)) >
+    \similarity(\state, \pi(\Rstrat, \messg_2)) \bicond \\
+    &  \ \ \ \ \ \Sstrat(\state, \messg_1) >
+    \Sstrat(\state, \messg_2) \\
+    0 & \text{otherwise.}
+  \end{cases}
+\end{align*}
 is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
@@
 \begin{align*}
-  V(\Rstrat) = \sum_{\messg \in \Messgs} \sum_{\state_1 \in \States}
-  \sum_{\state_2 \in \States} \textrm{if } s(\state_1, \pi(\Rstrat,
-  \messg)) > s(\state_2, \pi(\Rstrat, \messg)) \bicond \Rstrat(\messg,
-  \state_1) > \Rstrat(\messg, \state_2) \textrm{ then } 1 \textrm{
-    else } 0
+  V(\Rstrat) & = \sum_{\messg \in
+    \Messgs} \sum_{\state_1 \in \States}  \sum_{\state_2 \in \States}
+  F(\Rstrat, \messg, \state_1, \state_2) \text{\ \ \ with } \\
+  F(\Rstrat, \messg, \state_1, \state_2) & =
+  \begin{cases}
+    1 & \text{if } \similarity(\state_1, \pi(\Rstrat,
+  \messg)) > \similarity(\state_2, \pi(\Rstrat, \messg)) \\
+  & \ \ \ \ \ \bicond \Rstrat(\messg,
+  \state_1) > \Rstrat(\messg, \state_2) \\
+    0 & \text{otherwise.}
+  \end{cases}
 \end{align*}
 is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
-are lower bounded by $0$ and upper bounded by, respectively,
-$|\States| \times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$,
-thus we can normalize them by dividing by these values.
+are lower bound by $0$ and upper bound by, respectively, $|\States|
+\times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we will
+normalize them by dividing by these values.
 
@@
 vague, we look at the derived pure strategy $\Spure$ for which
-$\Spure(\state) = \messg$ only if $\messg =
-\argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$
-is convex, we also count $\Sstrat$ as convex. The sender strategy
-in Figure~\ref{fig:example_stratsA} is not convex, while the one in
+$\Spure(\state) = \argmax_{\messg' \in \Messgs}
+\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
+$\Sstrat$ as convex. The sender strategy in
+Figure~\ref{fig:example_stratsA} is not convex, while the one in
 Figure~\ref{fig:example_stratsB} is. Again, we do not expect a
@@
 imprecision can be expected to decrease expected utility
-\citep{Lipman2009:Why-is-Language}. The crucial question is whether
+\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is whether
 communicative success drops unacceptably fast with moderate levels of
@@
 
-  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}
+  \includegraphics[width=0.9\textwidth]{plots/MeanMetrics.pdf}
 
-  \caption{Means of recorded metrics for each triple of independent variables.}
+  \caption{Means of recorded metrics for each triple of independent
+    variables. Notice that the plots also show 95\% confidence
+    intervals, but these are most often invisible, owing to the
+    uniformity of outcomes across trials.}
   \label{fig:MeanMetrics}
@@
 of both sender and receiver strategies appear to be non-decreasing
-under increasing impairment and peaking at the maximal value. The only
-exception is the 6 states games and the transition from $\impairment =
-0$ to $\impairment = 0.05$. Expected utility of evolved languages
-mostly decreases in increasing imprecision, with the sole exception of
-a rise for $\toler = 0.2$ and $ns = 6$ at $\impairment = 0.1$. Still,
-normalized expected utility of evolved strategy pairs under mild
-values of imprecision are not necessarily much worse than their
-respective perfect precision \rd-analogues. This is especially so for
-lower numbers of states (which increase the chance of non-convergence
-and non-convexity; see Figure~\ref{fig:CategoricalMeasures} and the
-discussion below).
+under increasing impairment, with the only exception the 6 states
+games and the transition from $\impairment = 0$ to $\impairment =
+0.05$. Expected utility of evolved languages mostly decreases in
+increasing imprecision, with the sole exception of a rise for $\toler
+= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized expected
+utility of evolved strategy pairs under mild values of imprecision are
+not necessarily much worse than their respective perfect precision
+\rd-analogues. This is especially so for lower numbers of states
+(which increase the chance of non-convergence and non-convexity; see
+Figure~\ref{fig:CategoricalMeasures} and the discussion below).
 
 Taken together, these findings suggest the following effect of
-imprecise perception and diffusion of behavior. Mild forms of
-perceptual imprecision lead to slightly less communicative efficiency
-in the evolving strategy pairs, more vagueness and more regular,
-well-behaved languages. This impression is also supported by the
-proportion of converged and convex trials, pictured in
+confusability of states and diffusion of behavior. Mild forms of
+perceptual imprecision lead to slightly less communicative efficiency,
+more vagueness and more regular, well-behaved languages. This
+impression is also supported by the proportion of converged and convex
+trials, pictured in
 Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
@@
 absence of impairment. Increasing impairment always assured
-convergence and convexity: perceptual imprecision speeds up
-convergence to a convex and regular language.
+convergence and convexity: diffusion speeds up convergence to a convex
+and regular language.
 
@@
 
-  \includegraphics[width=\textwidth]{plots/CategoricalMeasures.pdf}
+  \includegraphics[width=0.9\textwidth]{plots/CategoricalMeasures.pdf}
 
@@
 variability in low-impairment conditions is not only due to
-non-convergence and/or associated
-non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more
-examples of strategy pairs at stopping time. Both are obtained for the
-same triple of parameters, both converged and are convex. However,
-they are not equally efficient. In fact, the pair in
+non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
+gives two more examples of strategy pairs at stopping time. Both are
+obtained for the same triple of parameters, both converged and are
+convex. However, they are not equally efficient. In fact, the pair in
 Figure~\ref{fig:example_stratsC} has a normalized expected utility of
@@
 where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
-average distances between sender strategies in this sense is plotted
-in Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
+\emph{within group distances}, i.e., the average distances between all
+sender strategies in each parameter group, are plotted in
+Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
 increasing imprecision, the sender strategies evolving under \rdd were
@@
 
-    \caption{Mean group measures as defined in the main text. For each
-      group, i.e., triplet of independent variables, the plot shows
-      (i) average normalized expected utility when playing against an
-      arbitrary other language from the group, (ii) the mean distance
-      between evolved sender strategies in the group.}
+    \caption{For each group, i.e., triplet of independent variables,
+      the plots show (i) within group distances and (ii) within group
+      expected utilities, as defined in the main text.}
   \label{fig:AverageEUinGroups}
@@
 This unifying property of perceptual imprecision could be considered
-an evolutionary beneficial side-effect. When we look at the normalized
-average expected utility that each evolved language (i.e., pair of
-sender and receiver strategy at stopping time) scored when playing
-against an arbitrary other language of the same group in
-Figure~\ref{fig:MoreExample_strats}, we see that higher imprecision
-can lead to higher such average expected utility. What this means is
-that, imprecision might decrease the communicative efficiency of
-individual languages, it increases the conceptual coherence and
-communicative success between independently evolving strategies. It is
-almost as if mere confusability of states imposes a regularity
-constraint on evolving linguistic meanings.
+an evolutionary beneficial side-effect. Higher imprecision can lead to
+higher \emph{within group expected utility}, i.e., the average
+expected utility that each evolved language (i.e., pair of sender and
+receiver strategy at stopping time) scored when playing against an
+arbitrary other language of the same parameter group (see
+Figure~\ref{fig:AverageEUinGroups}). What this means is that, while
+imprecision might decrease the communicative efficiency of individual
+languages, it increases the conceptual coherence and communicative
+success between independently evolving strategies. It is almost as if
+mere confusability of states imposes a regularity constraint on
+evolving linguistic meanings.
 

### 2014-10-16 paper/04_discussion.tex
revised sections 3 - 5

@@
 
-We have introduced a new variant of the replicator mutator dynamics
-that implements specifically stochastic noise in the form of
-confusability of similar states. The probabilities of confusing nearby
-similar states is easily implemented in behavioral strategies, but can
-also be translated into a mutation matrix for pure strategies. This
-allows, in principle, several conceptual interpretations of the \rdd,
-on which we will briefly elaborate below in
-Section~\ref{sec:model-interpretation}.
+We have introduced the \rdd as a new variant of the \rmd that
+implements stochastic noise in the form of confusability of similar
+states. The probabilities of confusing similar states is easily
+implemented for behavioral strategies, but, as seen in
+Section~\ref{sec:repl-diff-dynam}, can also be translated into a
+mutation matrix for pure strategies. This allows, in principle,
+several conceptual interpretations of the \rdd, on which we will
+briefly elaborate below in Section~\ref{sec:model-interpretation}.
 
@@
 dynamic at each time step. That would make sense if we thought of
-agents as prone to confusing similar states that are otherwise
-rational optimizers of behavior. The reason that we chose the
-replicator dynamics to combine with diffusion is twofold. A minor
-reason is that it makes for a conceptually interesting link with the
-replicator mutator dynamics. A more important reason is that the
-replicator dynamic is especially versatile and non-committal about
-what the exact process of adaption is that is being modelled.
+agents as prone to confusing similar states, but otherwise rationally
+optimizing behavior. The reason that we chose the \rd to combine with
+diffusion is twofold. A minor reason is that it makes for a
+conceptually interesting link with the \rmd. A more important reason
+is that the \rd is especially versatile and non-committal about what
+the exact process of adaption is that is being modelled.
 
@@
 every agent is assumed to have a unique sender and receiver role at
-the same time the is bequeathed onto the next generation. Diffusion,
-in this context, could be either of two things. It could be
-differential mutation probabilities in line with other interpretation
-of mutation in the \rmd
+the same time, the pair of which is bequeathed onto the next
+generation. Diffusion, in this context, could be either of two
+things. It could be differential mutation probabilities in line with
+other interpretations of mutation in the \rmd
 \citep[e.g.][]{NowakKomarova2001:Evolution-of-Un,KomarovaNiyogi2001:The-Evolutionar}:
-pure strategies are not faithfully reproduced, and mistakes in
-bequeathing pure strategies are more likely if they result from the
+pure strategies are not faithfully reproduced, e.g. by imperfections
+of learning or underdetermination of input, and mistakes in
+inheriting pure strategies are more likely if they result from the
 confusion of similar states. Another possible biological
@@
 the likely development of other behavioral adaptation processes, like
-differential imitation and cultural evolution in general
-\citep[see][for various derivations of the
+differential imitation \citep[see][for various derivations of the
 \rd]{Sandholm2013:Population-Game}. Under this interpretation, the \rd
-is consistent with the idea that what is subject to the evolutionary
-forces are not organisms but behavior: individuals can adapt their
-behavior to the perceived environment or adopt strategies from other
-individuals. Fitness captures the success of behavioral patterns,
-which in the case of language can be thought of in terms of
+is consistent with the idea that what is subject to evolutionary
+optimization are not organisms but behavior: agents or systems can
+adapt their behavior to the perceived environment or agents can adopt
+strategies from others. Fitness captures the success of behavioral
+patterns, which in the case of language can be thought of in terms of
 communicative success.  Differential reproduction represents the
-tendency of more successful strategies to be more likely adopted by
-individuals, be it through imitation or some learning procedure within
-the population.
+tendency of more successful strategies to be more likely realized, be
+it through imitation or some solitary learning or optimization
+mechanism.
 
@@
 before. One possibility is that diffusion is noise in the adoption of
-strategies, say, by conditional imitation, of strategies by other
+strategies, e.g., by conditional imitation, of strategies by other
 signalers. Another possibility is that diffusion is again noise on the
@@
 
-Under plain Herrnstein reinforcement learning, sender and receiver
-play the game repeatedly and adjust their dispositions to act after
-each round of play, in such a way that the actual (non-negative)
+Under \emph{Herrnstein reinforcement learning}, sender and
+receiver play the game repeatedly and adjust their dispositions to act
+after each round of play, in such a way that the actual (non-negative)
 payoff gained in the current interaction is added to the
-non-normalized propensities for acting in exactly the way that they
-acted in the current round of play. For sim-max games, this means that
-when the sender chose $\messg$ in state $\state$, and this resulted in
-some non-negative payoff (which is guaranteed by our choice of utility
-function), the probability that the sender chooses $\messg$ again in
-$\state$ is increased, but nothing else changes. In particular, the
-senders behavior in other choice points does not change. Generalized
-reinforcement learning is different here. When the use of $\messg$ in
-$\state$ gave positive payoff, then not only will its future use
-probability be promoted at $\state$ but also at other states,
-proportional to how similar these are to $\state$. Similar amendments
-take care of the way that the receiver updates his choice
-dispositions.
+non-normalized propensities for repeating the same behavior. For
+sim-max games, this means that when the sender chose $\messg$ in state
+$\state$, and this resulted in some non-negative payoff (which is
+guaranteed by our choice of utility function), the probability that
+the sender chooses $\messg$ again in $\state$ is increased, but
+nothing else changes. In particular, the sender's behavior in other
+choice points does not change. \emph{Generalized reinforcement
+  learning} is different here. When the use of $\messg$ in $\state$
+gave positive payoff, then not only will its future use probability be
+promoted at $\state$ but also at other states, proportional to how
+similar these are to $\state$. Similar amendments take care of the way
+that the receiver updates his choice dispositions.
 
@@
 leads to vague signaling of the appropriate kind, but also speeds up
-learning in such a way that, especially for games with higher numbers
-of states, higher levels of communicative success are reached in
-shorter learning periods than is possible without stimulus
+learning in such a way that, especially for sim-max games with higher
+numbers of states, higher levels of communicative success are reached
+in shorter learning periods than is possible without stimulus
 generalization. Technically, this result is partly due to the fact
@@
 after each round of play under generalized reinforcement than under
-the plain variety. But that only explains the speed of adaptation, not
-necessarily that generalization also leads to regularity and
-communicative efficiency.
+the Herrnstein variety. But that only explains the speed of
+adaptation, not necessarily that generalization also leads to
+regularity and communicative efficiency.
 
-Diffusion of strategies in the \rdd can be conceived of as a form of
-generalization as well, and works in large part quite analogous to
-stimulus generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
-approach. This holds for the effect of diffusion and generalization,
-but not necessarily for the way that the effect is achieved. We also
-saw that diffusion in \rdd leads to more regular languages and
-speedier convergence. However, the \rdd is a more abstract framework
-than generalized reinforcement learning. The latter is foremost
-motivated as a learning dynamic that has two players adapt their
-individual strategies after each concrete round of play. In contrast,
-the \rdd describes a more abstract, average dynamical change in
-behavioral dispositions. Although the dynamics of (some forms of)
-reinforcement learning mirror those of the replicator dynamics (at
-some stage in time)
+Diffusion of strategies in the \rdd can also be conceived of as a form
+of generalization, and works in large part quite analogous to stimulus
+generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
+approach. But there are still
+differences. \citeauthor{OConnor2013:The-Evolution-o} showed that
+generalized reinforcement learning can speed up the development of
+efficient signaling, especially for higher numbers of
+states. Complementing this, we showed that diffusion regularizes
+evolutionary outcomes and prevents meandering around sub-optimal and
+non-convex strategies, also for low numbers of states.
+
+Conceptually, the \rdd is a more abstract framework than generalized
+reinforcement learning. The latter is foremost motivated as a learning
+dynamic that has two players adapt their individual strategies after
+each concrete round of play. In contrast, the \rdd describes a more
+abstract, average dynamical change in behavioral
+dispositions. Although the behavior of (some forms of) reinforcement
+learning mirror those of the \rd (at some stage in time)
 (\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
-this does not mean that \emph{generalized} reinforcement learning, in
-which stimulus generalization is best motivated at the level of a
-single agent's dispositional generalization after one round of play,
-is also directly a plausible high-level description of, say,
+this does not mean that \emph{generalized} reinforcement learning is
+also necessarily a plausible high-level description of, say,
 generalized learning in a population of agents. Seen in this light,
@@
 lead to vague strategies. The model that is most clearly related to
-the present approach uses the notion of a quantal response, also known
-as a logit response or a soft-max function
+the present approach uses the notion of a \emph{quantal response},
+also known as a \emph{logit response} or a \emph{soft-max function}
 \citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,McKelveyPalfrey1995:Quantal-Respons,McKelveyPalfrey1998:Quantal-Respons,GoereeHolt2008:Quantal-Respons}. A
@@
 \rightarrow \infty$ we retrieve the behavior of the best response
-choice function, but if it is positive but finite, any choice $\act$
-will receive a positive probability, but acts with higher expected
-utility will be more likely. The underlying motivation for this choice
-rule is the assumption that there is noise in the computation of
-expected utilities and/or in maximization of expected
-utilities. Consequently, choices with almost equal expected utilities
-will be chosen with almost equal probability (for moderate values of
-$\lambda$).
+function, but if it is positive but finite, any choice $\act$ will
+receive a positive probability, but acts with higher expected utility
+will be more likely. The underlying motivation for this choice rule is
+the assumption that there is noise in the computation of expected
+utilities and/or in maximization of expected utilities. Consequently,
+choices with almost equal expected utilities will be chosen with
+almost equal probability (for moderate values of $\lambda$).
 
@@
 
-But not all quantal response equilibria are equally plausible
-explanations of vague language use, and the ones that are not suggest
-that it is less plausible to think of vagueness as arising because of
-errors in the computation and maximization of expected utility, as the
-quantal response approach assumes, than that it arises due to
-confusion of similar states, as the \rdd proposes. To see what the
-problem is, we can look at cases like given in
+But not all quantal response equilibria are equally plausible, and the
+ones that are not suggest that it is less natural to think of
+vagueness as arising from of errors in the computation or maximization
+of expected utility than that it arises from confusion of similar
+states. To see what the problem is, we can look at cases like given in
 Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
@@
 property ``tall''. If the usual interpretation of a ``tall man'' peaks
-at around, say, 195cm then when meeting a giant of $n$ meters senders
+at around, say, 195cm then when meeting a giant of $n$ meters, senders
 would, according to the quantal response approach, be ever more

### 2014-10-17 paper/05_conclusions.tex
conclusion section

--- 2014-05-10 paper/03_simulations.tex
@@
-\section{Simulations}
+\section{Conclusions}
+\label{sec:conclusions}
+
+Vagueness is a pervasive but seemingly non-disruptive feature of
+natural communication and classification systems. From an evolutionary
+point of view, the challenge arises to explain how vagueness can
+persist under selective pressure for precision. While intuitive
+reasons for vagueness abound, this is foremost a technical challenge,
+probing for the possibility of integrating into a consistent model
+forces that lead to vagueness with forces that lead to efficient
+information transfer. This paper proposed one such model in the
+context of sim-max games and explored some of its consequences.
+
+We introduced the replicator diffusion dynamic as a special case of
+the established replicator mutator dynamic. It is a special case in
+the sense that the \rdd incorporates a particular kind of mutation
+that is based on the similarity of states (as defined in the sim-max
+games). More concretely, we started from the idea that similar states
+are more easily confused than less similar states. To formalize this
+idea, we assumed that there is a confusion matrix $C$, where $C_{ij}$
+is the probability that a state \mystate{i} is realized as a possibly
+different state \mystate{j}. Based on this, we showed two things. For
+one, we showed how the discrete-time formulation of the replicator
+dynamic can be easily adapted to integrate diffusion of behavior due
+to confusion of states. This yielded the \rdd in its basic
+discrete-time formulation. For another, we also showed that there
+exists a translation of the confusion matrix $C$ into a suitable
+mutation matrix $M_C$ such that the effect of $C$ and $M_C$ are
+effectively equivalent. This supports the conclusion that the \rdd can
+indeed be thought of as a special case of the \rmd.
+
+Based on data from numerical simulations, we showed that the inclusion
+of mild levels of stochastic confusability of states does not
+undermine the possibility of evolving communicatively successful
+signaling strategies at all. On the contrary, strategies that evolved
+under mild diffusion induce a highly regular and systematic category
+structure that shows the signs of vagueness as desired. Moreover, we
+noted that there might even be a higher-order benefit to the presence
+of imprecision, in that can accelerates convergence to optimal
+categorization, by preventing evolutionary paths to stay near
+inefficient non-convex strategies for a long time. Diffusion thus also
+unifies and regularizes evolutionary outcomes in such a way that
+sub-optimal categorization is avoided. These results complement
+similar results by \citet{OConnor2013:The-Evolution-o}, obtained for a
+micro-dynamic extension of reinforcement learning. The \rdd adds to
+this a more general, abstract framework that is not tied to the
+specific assumptions of turn-based reinforcement.
+
+Many interesting issues arise in this context that we have not yet
+explored. Firstly, we concentrated on the \rdd in behavioral
+strategies, because these are much less complex and unwieldy than
+mixed strategies or analyses of the symmetrized game. But we also
+noted that for especially biological interpretations of the dynamic,
+symmetrizing would be conceptually more plausible. It remains to be
+seen whether there are substantial differences between the symmetrized
+version of the \rdd and the behavioral strategy version that
+we have explored here.
+
+Secondly, we have only explored the effect of a single confusion
+matrix that affected both sender and receiver roles. Obviously, we
+could allow different roles to be affected differently by
+confusability of states. We might hypothesize that for the unifying
+and regularizing effects of diffusion, it is less important that the
+receiver's strategies are diffused, as long as the sender's are. But,
+again, this issues must wait for further exploration. Connected to
+that is the obvious extension that also allows for stochastic
+confusability of signals.
+
+Further promising extensions of the work presented here include
+investigating higher-dimensional state spaces, possibly with less
+uniform, more psychologically real similarity metrics, such as the
+three-dimensional color space. Also, to compare the work presented
+here better to other work on efficient categorization
+\citep[e.g.][]{Mohlin2014:Optimal-Categor}, it would be worthwhile
+exploring the possibility of signal innovation
+\citep[e.g.][]{McKenzie-AlexanderSkymrs2012:Inventing-New-S} also for
+sim-max games. This way, the game dynamics would ideally lead to an
+optimal number of categories, not only an optimal shape of a
+predefined number of categories.
 

### 2014-10-20 paper/05_conclusions.tex
conclusion revised

@@
 point of view, the challenge arises to explain how vagueness can
-persist under selective pressure for precision. While intuitive
-reasons for vagueness abound, this is foremost a technical challenge,
-probing for the possibility of integrating into a consistent model
-forces that lead to vagueness with forces that lead to efficient
-information transfer. This paper proposed one such model in the
-context of sim-max games and explored some of its consequences.
+persist under selective pressure for precision. This is foremost a
+technical challenge, probing for the possibility of integrating into a
+consistent model forces that lead to vagueness with forces that lead
+to efficient information transfer. This paper proposed one such model
+in the context of sim-max games and explored some of its consequences.
 
@@
 to confusion of states. This yielded the \rdd in its basic
-discrete-time formulation. For another, we also showed that there
-exists a translation of the confusion matrix $C$ into a suitable
-mutation matrix $M_C$ such that the effect of $C$ and $M_C$ are
-effectively equivalent. This supports the conclusion that the \rdd can
-indeed be thought of as a special case of the \rmd.
+discrete-time formulation for behavioral strategies. For another, we
+also showed that there exists a translation of the confusion matrix
+$C$ into a suitable mutation matrices $Q_C$ and $R_C$ such that the
+effect of former and latter are effectively equivalent, although $C$
+applies to behavioral strategies, while $Q_C$ and $R_C$ apply to mixed
+strategies. This shows that the \rdd may be thought of as a special
+case of the \rmd.
 
-Based on data from numerical simulations, we showed that the inclusion
-of mild levels of stochastic confusability of states does not
-undermine the possibility of evolving communicatively successful
+Based on data from numerical simulations, we demonstrated that the
+inclusion of mild levels of stochastic confusability of states does
+not undermine the possibility of evolving communicatively successful
 signaling strategies at all. On the contrary, strategies that evolved
 under mild diffusion induce a highly regular and systematic category
-structure that shows the signs of vagueness as desired. Moreover, we
-noted that there might even be a higher-order benefit to the presence
-of imprecision, in that can accelerates convergence to optimal
-categorization, by preventing evolutionary paths to stay near
-inefficient non-convex strategies for a long time. Diffusion thus also
-unifies and regularizes evolutionary outcomes in such a way that
-sub-optimal categorization is avoided. These results complement
-similar results by \citet{OConnor2013:The-Evolution-o}, obtained for a
-micro-dynamic extension of reinforcement learning. The \rdd adds to
-this a more general, abstract framework that is not tied to the
-specific assumptions of turn-based reinforcement.
+structure that shows the signs of vagueness as desired. There might
+even be a higher-order benefit to the presence of imprecision, in that
+it can accelerate convergence to optimal categorization, by preventing
+evolutionary paths to stay near inefficient non-convex strategies for
+a long time. Diffusion thus also unifies and regularizes evolutionary
+outcomes in such a way that sub-optimal categorization is
+avoided. These results complement similar findings by
+\citet{OConnor2013:The-Evolution-o}, obtained for a micro-dynamic
+extension of reinforcement learning. The \rdd adds to this a more
+general, abstract framework that is not tied to the specific
+assumptions of turn-based reinforcement.
 
@@
 explored. Firstly, we concentrated on the \rdd in behavioral
-strategies, because these are much less complex and unwieldy than
-mixed strategies or analyses of the symmetrized game. But we also
-noted that for especially biological interpretations of the dynamic,
-symmetrizing would be conceptually more plausible. It remains to be
-seen whether there are substantial differences between the symmetrized
-version of the \rdd and the behavioral strategy version that
-we have explored here.
+strategies, because numerical simulations of these are much less
+complex and unwieldy than of mixed strategies or analyses of the
+symmetrized game. But we also noted that for especially biological
+interpretations of the dynamic, symmetrizing would be conceptually
+more plausible. It remains to be seen whether there are substantial
+differences between the symmetrized version of the \rdd and the
+behavioral strategy version that we have explored here.
 

### 2014-10-20 paper/03_simulations.tex
plots reworked, almost all BW

@@
     $\card{\States} = 50$ and different
-    values of $\impairment$.}
+    values of imprecision $\impairment$.}
   \label{fig:NosofskySim}
@@
 
-  \includegraphics[width=\textwidth]{plots/confusion_sender.pdf}
+  \includegraphics[width=0.9\textwidth]{plots/confusion_sender.pdf}
 
-  \includegraphics[width=\textwidth]{plots/confusion_receiver.pdf}
+  \includegraphics[width=0.9\textwidth]{plots/confusion_receiver.pdf}
 
@@
 
-  \includegraphics[width=0.9\textwidth]{plots/MeanMetrics.pdf}
+  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}
 
@@
 
-  \includegraphics[width=0.9\textwidth]{plots/CategoricalMeasures.pdf}
+  \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}
 
@@
 
-    \includegraphics[width=\textwidth]{plots/WithinGroupMeasures.pdf}
+    \includegraphics[width=0.8\textwidth]{plots/WithinGroupMeasures.pdf}
 

### 2014-10-20 snippets/Nosofsky_similarity.Rmd
plots reworked, almost all BW

--- 2014-09-18 snippets/Nosofsky_similarity.Rmd
@@
 library(reshape2)
+theme_set(theme_bw())
 ```
@@
 simDF = data.frame(distance = rep(states,2*length(alphas)),
-                   alpha = factor(rep(rep(alphas, each=ns),2)),
-                   normalized = factor(rep(c("non-normalized", "normalized"),each=ns*length(alphas))),
+                   alpha_i = factor(rep(rep(alphas, each=ns),2)),
+                   normalized = factor(rep(c("Nosofsky similarity", "confusion probability"),each=ns*length(alphas)),
+                                       levels = c("Nosofsky similarity", "confusion probability")),
                    similarity = c(as.vector(t(out)),as.vector(t(outNorm)))  )
 
-p = ggplot(simDF, aes(x = distance, y = similarity, color=alpha)) + geom_point() +
-  geom_line() + facet_grid(normalized~. , scales = "free")
+p = ggplot(simDF, aes(x = distance, y = similarity, shape = alpha_i))+
+  geom_line(color='grey')  + geom_point()  + 
+  facet_grid(normalized~. , scales = "free") + ylab("") + 
+  xlab("distance in physical space") +
+  scale_shape_discrete(name="imprecision")
 
@@
 
-pdf('paper/plots/NosofskySim.pdf', width=7, height=4)
-show(p)
-dev.off()
+ggsave(p, file="paper/plots/NosofskySim.pdf", width=7, height=4)
+
 ```

### 2014-10-20 snippets/data_exploration/data_exploration_tmp.Rmd
plots reworked, almost all BW

@@
 
+myLabeller <- function(var, value){
+    value <- as.character(value)
+    if (var=="variable") { 
+        value[value=="Speaker.entropy"] <- "entropy sen"
+        value[value=="Hearer.entropy"] <- "entropy rec"
+        value[value=="Speaker.informativity"] <- "informativity"
+        value[value=="Speaker.Voronoiness"] <- "Voronoiness sen"
+        value[value=="Hearer.Voronoiness"] <- "Voronoiness rec"
+        value[value=="Expected.utility"] <- "expected utility"
+    }
+    return(value)
+}
+
 pd <- position_dodge(.01)
-sp = ggplot(dsub, aes(x=Impairment, y=value, colour=Number.of.states)) + 
-     geom_point(position = pd) +
+mysize = 1.65
+sp = ggplot(dsub, aes(x=Impairment, y=value, color = Number.of.states, shape=Number.of.states, linetype=Number.of.states)) + 
+    geom_line(color="gray", position = pd) + xlab("imprecision") +
+     geom_point(position = pd, size = mysize) +
      geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
-     geom_line(position=pd) + xlab("imprecision")
+    scale_shape_discrete(name = "# states") +
+    scale_linetype_discrete(name = "# states") +
+    scale_color_discrete(name = "# states")
 
-sp = sp + facet_grid(variable ~ Tolerance, scales="free")
+sp = sp + facet_grid(variable ~ Tolerance, scales="free", labeller=myLabeller)
 
+sp + facet_grid(. ~ sex, labeller=mf_labeller)
 show(sp)
 
-pdf("~/Desktop/data/svn/vagueness-games/paper/plots/MeanMetrics.pdf", width=8, height = 6)
-show(sp)
-dev.off()
+ggsave(sp, file = "paper/plots/MeanMetrics.pdf", width=8, height = 8)
 
@@
 
-sp = ggplot(dms, aes(x=factor(Impairment), y = value)) 
-sp = sp + geom_bar(aes(fill = variable), color = "black", stat="identity", position= position_dodge(0.4), width = 0.35)
-sp = sp + facet_grid(Number.of.states ~ Tolerance, scales="free") + ylab("proportion") + xlab("imprecision")
-
+pd <- position_dodge(.05)
+sp = ggplot(dms, aes(x=factor(Impairment), y = value, group = variable, shape = variable, linetype = variable)) 
+sp = sp + geom_line(position = pd, color = "grey") + geom_point(position = pd)
+sp = sp + facet_grid(Number.of.states ~ Tolerance, scales="free") + ylab("proportion") + xlab("imprecision") +
+  scale_shape_discrete(name = "property") +
+  scale_linetype_discrete(name = "property")
 show(sp)
-
-pdf("~/Desktop/data/svn/vagueness-games/paper/plots/CategoricalMeasures.pdf", width=8, height = 5)
-show(sp)
-dev.off()
+ggsave(sp, file = "paper/plots/CategoricalMeasures.pdf", width=8, height = 5)
 

### 2014-10-20 snippets/iteration_diffusion.Rmd
plots reworked, almost all BW

--- 2014-09-17 snippets/iteration_diffusion.Rmd
@@
 library(reshape2)
+theme_set(theme_bw())
+greyPalette = c("#999999", "#222222", "#999999", "#000000", "#999999", "#000000", "#999999", "#000000")
 # source('~/Desktop/data/svn/programming/R/fuzzy-action/dev.r')
@@
 
-plotSender = function(sen){
-  senDF = melt(sen,varnames = c("state.index","message"), value.name = "probability")
-  senDF$states = states[senDF[,1]]
+ns = 50
+nm = 2
+rounds = c(1,2,10,50,1000)
+alphas = c(0.01,0.05,0.1,0.3)
+
+plotSender = function(ns,nm,alphas,rounds){
+  sen = rdirichlet(ns,rep(1,nm))
+  senDF = data.frame(state = 0,
+                     message = 0,
+                     round = 0,
+                     alpha = 0,
+                     probability = 0)
+
+  for (alpha in alphas){
+    C = get.confusion.matrix(get.states(ns),alpha)
+    newDF = data.frame(state = rep(get.states(ns),nm),
+                       message = rep(1:2,each=ns),
+                       round = rep(0,ns*nm),
+                       alpha = rep(alpha,ns*nm),
+                       probability = as.vector(sen))
+    for (r in rounds){
+      newC = power(C,r)
+      newSen = get.nbr(sen,newC)
+      newDF2 = data.frame(state = rep(get.states(ns),nm),
+                          message = rep(1:2,each=ns),
+                          round = rep(r,ns*nm),
+                          alpha = rep(alpha,ns*nm),
+                          probability = as.vector(newSen))
+      newDF = rbind(newDF,newDF2)                   
+      }
+    senDF = rbind(senDF,newDF)
+  }
+  senDF = senDF[-1,]
   senDF$message = factor(senDF$message)
-  ggplot(senDF, aes(x = states, y=probability, color=message)) + geom_point() + geom_line()
+  senDF$alpha = factor(senDF$alpha)
+  p = ggplot(senDF,aes(x = state, y = probability, color = message)) + 
+      geom_line() + facet_grid(alpha~round) +
+                                          scale_colour_manual(values=greyPalette) +
+      ggtitle("sender") +
+      scale_x_continuous(breaks=seq(0,1,1/2)) +
+      theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + 
+      scale_y_continuous(breaks=seq(0,1,1/2))
+
+  return(p)
 }
+
+plotReceiver = function(ns,nm,alphas,rounds){
+  rec = rdirichlet(nm,rep(1,ns))
+  recDF = data.frame(state = 0,
+                     message = 0,
+                     round = 0,
+                     alpha = 0,
+                     probability = 0)
+
+  for (alpha in alphas){
+    C = get.confusion.matrix(get.states(ns),alpha)
+    newDF = data.frame(state = rep(get.states(ns),nm),
+                       message = rep(1:2,each=ns),
+                       round = rep(0,ns*nm),
+                       alpha = rep(alpha,ns*nm),
+                       probability = as.vector(t(rec)))
+    for (r in rounds){
+      newC = power(C,r)
+      newRec = get.nbr(rec,newC)
+      newDF2 = data.frame(state = rep(get.states(ns),nm),
+                          message = rep(1:2,each=ns),
+                          round = rep(r,ns*nm),
+                          alpha = rep(alpha,ns*nm),
+                          probability = as.vector(t(newRec)))
+      newDF = rbind(newDF,newDF2)                   
+      }
+    recDF = rbind(recDF,newDF)
+  }
+  recDF = recDF[-1,]
+  recDF$message = factor(recDF$message)
+  recDF$alpha = factor(recDF$alpha)
+  p = ggplot(recDF,aes(x = state, y = probability, color = message)) + 
+       geom_line() + facet_grid(alpha~round) +
+                                          scale_colour_manual(values=greyPalette) +
+      scale_x_continuous(breaks=seq(0,1,1/2)) +
+      ggtitle("receiver") +
+      theme(axis.text.x  = element_text(angle=90, vjust=0.5)) + 
+      scale_y_continuous(breaks=seq(0,round(max(recDF$probability),2),round(max(recDF$probability),2)/2))
+  return(p)
+}
+
+pSen = plotSender(ns,nm,alphas,rounds)
+pRec = plotReceiver(ns,nm,alphas,rounds)
+show(pSen)
+show(pRec)
+
+ggsave(pSen, file = "paper/plots/confusion_sender.pdf", width = 7, height = 4)
+
+ggsave(pRec, file = "paper/plots/confusion_receiver.pdf", width = 7, height = 4)
 

### 2014-10-20 snippets/QRE.R

@@ -31,9 +31,6 @@ for (i in 1:100){
 
 outPlot = arrangeGrob(plot.sender.strat(sen,states),
                       plot.receiver.strat(rec,states))
-
 show(outPlot)
-filename = paste("~/Desktop/data/svn/vagueness-games/paper/plots/exampleStratQRE_tolerance0", tolerance*10, ".pdf", sep = "", collapse = "")
-pdf(filename, height = 5, width = 6)
-show(outPlot)
-dev.off()
+filename = paste("paper/plots/exampleStratQRE_tolerance0", tolerance*10, ".pdf", sep = "", collapse = "")
+ggsave(outPlot, file = filename, height = 5, width = 5)

### 2014-10-20 snippets/example_strats/example_strats_plots.R

@@ -1,5 +1,6 @@
 require('plyr')
 require('reshape2')
+theme_set(theme_bw())
 
 source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
 source('~/Desktop/data/svn/vagueness-games/data-analysis.R')

### 2014-10-20 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -3,6 +3,7 @@ computeEU = FALSE
 
 require('plyr')
 require('reshape2')
+theme_set(theme_bw())
 
 source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
 
@@ -202,14 +203,15 @@ D$meanEUGroup = meanEUs$Speaker.meanEU
 
 Dmelt = melt(D, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
 
-combPlot = ggplot(Dmelt, aes(x=Impairment, y=value, color = variable)) + 
-  geom_point(aes(shape = variable)) + facet_grid(Number.of.states ~ Tolerance) + geom_line()
-
+combPlot = ggplot(Dmelt, aes(x=Impairment, y=value, group = variable, shape = variable, linetype = variable)) +  
+  geom_line( color = "grey") + geom_point() + facet_grid(Number.of.states ~ Tolerance) +
+  xlab("imprecision") +
+  scale_shape_discrete(name = "property", 
+                       labels = c("distance", "utility")) + 
+  scale_linetype_discrete(name = "property", 
+                          labels = c("distance", "utility"))
 show(combPlot)
-
-pdf("~/Desktop/data/svn/vagueness-games/paper/plots/WithinGroupMeasures.pdf", height=6, width=8)
-show(combPlot)
-dev.off()
+ggsave(combPlot, file = "~/Desktop/data/svn/vagueness-games/paper/plots/WithinGroupMeasures.pdf", height=5, width=8)
 
 
 

### 2014-10-20 helper_functions.R
clean up

@@ -1,0 +1,180 @@
+source('~/Desktop/data/svn/vagueness-games/data-analysis.R')
+
+theme_set(theme_bw())
+greyPalette = c("#999999", "#222222", "#999999", "#000000", "#999999", "#000000", "#999999", "#000000")
+
+
+##################
+## load data
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
+                colClasses=c('numeric','factor','numeric','numeric','numeric',
+                             'factor','numeric','numeric','numeric','numeric',
+                             'numeric','numeric','numeric','numeric','numeric',
+                             'numeric','numeric','numeric','character','character'))
+
+# no idea why these numbers are messed up, but this fix seems to solve it
+data$Speaker.Voronoiness = ifelse(data$Speaker.Voronoiness > 1, 
+                                  yes = data$Speaker.Voronoiness / 1000, 
+                                  no = data$Speaker.Voronoiness)
+data$Hearer.Voronoiness = ifelse(data$Hearer.Voronoiness > 1, 
+                                 yes = data$Hearer.Voronoiness / 1000, 
+                                 no = data$Hearer.Voronoiness)
+
+ds = data[,c(1,4,5,7,8,11:17)] # only relevant data
+dm = melt(ds,id.vars = c("Number.of.states", "Impairment", "Tolerance")) # melted data
+
+###################
+
+plot.language.byindex = function(dataFrame,rowindex){
+  plot.language(dataFrame[rowindex,]$Speaker.strategy.file,
+                dataFrame[rowindex,]$Hearer.strategy.file)
+}
+
+
+
+plot.language.MinMaxProperty = function(dataFrame,colindex){
+  show(names(dataFrame)[colindex])
+  show(summary(dataFrame[,colindex]))
+  p = arrangeGrob(plot.language.byindex(dataFrame,which.min(dataFrame[,colindex])),
+                  plot.language.byindex(dataFrame,which.max(dataFrame[,colindex])))
+}
+
+###################
+
+get.confusion.matrix = function(states,alpha){
+  ns = length(states)
+  if (alpha <= 0){
+    return(diag(rep(1,ns)))
+  }
+  C = matrix(0,nrow=ns, ncol=ns)
[…]

### 2014-10-20 paper/02_noisyRMD.tex
proof revised
removed todo note

@@
 Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
-\begin{align*}
+\begin{align}
+  \label{eq:CorrespondenceBehavioralMixed}
   \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
-\end{align*} 
+\end{align} 
 Let $G$ be this mapping from mixed to behavioral strategies. This $G$
@@
 $\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
-average fitness in the receiver population.\todo{give actual
-  definition of fitness etc.?}
+average fitness in the receiver population.
 
@@
 of (one round of) mutation:
-\begin{align*}
+\begin{align}
+  \label{eq:Mutation}
   \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
@@
   R_{ji} \,.
-\end{align*}
+\end{align}
 The discrete-time \rmd given above can then be rewritten as:

### 2014-10-21 paper/01_intro.tex
conclusion section
Some small changes.

@@
 byproduct of limitations in information processing
-\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o,QingFranke2014:Gradable-Adject}. This
+\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
 paper makes a contribution to the latter line of thought.  Concretely,
@@
 \emph{sim-max games}, for short
-\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}. Our
-contribution is best seen in relation to these accounts, as it also
+\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
+%\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
+Our contribution is best seen in relation to these accounts, as it also
 relies on sim-max games. Let's introduce these first, and then zoom in
@@
 
-Similarity-maximizing games are generalizations of Lewis' games in
-that they allow different states to be more or less similar to one
-another, and, roughly put, make success of communication a function of
+Similarity-maximizing games are generalizations of Lewis' games where
+different states are allowed to be more or less similar to one
+another, and, roughly put, success of communication is a function of
 that similarity. Intuitively speaking, while Lewis' games treated
@@
 \end{align*}
-To keep matters simple, we here only look at cases where $\States$
+To keep matters simple, in this paper we mostly focus on cases where $\States$
 contains finitely many points from the unit interval, all of which
@@
 \mystate{2})^2$. For these games, the evolutionarily stable states are
-demonstrably so-called Voronoi-languages. Intuitively speaking, a
+demonstrably so-called Voronoi languages. Intuitively speaking, a
 Voronoi language is a pair of sender and receiver strategies, such
@@
 central spots in each category. Figure~\ref{fig:VoronoiLanguage} gives
-an example of such a Voronoi-language for a sim-max game with $\States
+an example of such a Voronoi language for a sim-max game with $\States
 = [0,1]$, two messages, and a flat prior distribution over states. The
@@
 
-        \caption{Example of a Voronoi-language on $\States =
+        \caption{Example of a Voronoi language on $\States =
           [0,1]$. The (pure) sender strategy $\Spure \in
@@
 entities, whose purpose is to communicate effectively about the actual
-state. In that case, evolving Voronoi-languages would explain why
+state. In that case, evolving Voronoi languages would explain why
 linguistic categories are well-behaved and orderly in the way they
@@
 that tries to decode the original information. In this case, evolving
-Voronoi-languages would explain why conceptual categories are
+Voronoi languages would explain why conceptual categories are
 well-behaved and orderly in the way that they appear to be
@@
 in such a way that states similar to the one that actually occurred
-are subject to behavioral adjustment as well. (A more detailed
-discussion of alternative approaches is deferred until
-Section~\ref{sec:discussion}, where we can compare them better with the
-system introduced in this paper.)
+are subject to behavioral adjustment as well.
 
@@
 mistakes lead to systematic ``trembles'' that produce vague signal
-use. The approach we take here is superficially similar, but there are
+use.
+
+The approach we take here is superficially similar, but there are
 crucial differences. For one, we adopt a dynamic perspective by
@@
 rather, as we assume here, as the result of confusing similar
-states.
+states. A more in-depth discussion of alternative approaches is deferred until
+Section~\ref{sec:discussion}.
+Let us first look into our own proposal in more detail.
 

### 2014-10-21 paper/02_noisyRMD.tex
Some small changes.

@@
 the receiver role (albeit in slightly different manner). We call our
-dynamic \emph{replicator diffusion dynamic} (\rdd), because it is a
-special case of the replicator mutator dynamic (\rmd).
+dynamic \emph{replicator diffusion dynamic} (\rdd), because it is an
+extension of the replicator dynamic (\rd) that adds diffusion
+of behavior. It is also a special case of the replicator mutator dynamic (\rmd).
 
-The \rdd was first introduced by \citet{Correia2013:The-Bivalent-Tr}
-as a noise-perturbed variant of the replicator dynamic (\rd) in
-behavioral strategies. We begin by recapitulating this formulation of
-the \rd in Section~\ref{sec:repl-dynam-behav}. Then we introduce the
-\rdd in Section~\ref{sec:repl-diff-dynam-1}.
+We begin by recapitulating the formulation of the \rd in Section~\ref{sec:repl-dynam-behav}.
+Then we introduce the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
 Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
 Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
-a special case of the \rdd.
+a special case of the \rmd.
 
@@
 \begin{align*}
-  \EU(\messg \myts ; \myts \state, \Rstrat) & = \sum_{\act \in \Acts}
+  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
   \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
-  \EU(\act \myts ; \myts \messg, \Sstrat) & = \sum_{\state \in
+  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
     \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
@@
   \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
-\EU(\messg \myts ; \myts\state,\Rstrat) \\
+\EU(\messg, \state,\Rstrat) \\
 \Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
-\cdot \EU(\act \myts ; \myts \messg,\Sstrat)\,.
+\cdot \EU(\act, \messg,\Sstrat)\,.
 \end{align*}
@@
   \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
-    \EU(\messg \myts ; \myts \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
+    \EU(\messg, \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
     \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
-    \EU(\act \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
+    \EU(\act, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
 \end{align*}
@@
 
-The replicator diffusion dynamic adds probabilistic confusion of
-similar states to the replicator dynamic. Fix a sim-max game with
+The replicator diffusion dynamic was first introduced by \citet{Correia2013:The-Bivalent-Tr}
+as a noise-perturbed variant of the replicator dynamic (\rd) in
+behavioral strategies. Fix a sim-max game with
 $\States = \Acts$ and a confusion matrix $C : \States \times
@@
 
-      \begin{pgfonlayer}{background}
-        \node [draw=black!50, fill=black!20,fit=(actual) (label)
-        (output)] {};
-      \end{pgfonlayer}
+%      \begin{pgfonlayer}{background}
+%        \node [draw=black!50, fill=black!20,fit=(actual) (label)
+%        (output)] {};
+%      \end{pgfonlayer}
     \end{scope}
@@
 
-      \begin{pgfonlayer}{background}
-        \node [draw=black!50, fill=black!20,fit = (message) 
-        (realized) (bla)] {};
-      \end{pgfonlayer}
+%      \begin{pgfonlayer}{background}
+%        \node [draw=black!50, fill=black!20,fit = (message) 
+%        (realized) (bla)] {};
+%      \end{pgfonlayer}
     \end{scope}
@@
   \frac{\Sstrat(\messg \probbar \state_j) \cdot
-    \EU(\messg \myts ; \myts \state_j,\Rstrat)}
+    \EU(\messg, \state_j,\Rstrat)}
   {\Phi(\state_j,\Sstrat, \Rstrat)} \\
@@
   \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j \myts ; \myts \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
+    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
 \end{align*}
@@
 \citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un,Nowak2006:Evolutionary-Dy}. It
-adds to the replicator dynamic stochastic mutation that captures, in
+adds stochastic mutation to the replicator dynamic in order to capture, in
 the context of grammar evolution, the differential learning success of
@@
 \end{align} 
-Let $G$ be this mapping from mixed to behavioral strategies. This $G$
+Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
 is \emph{not} an injection, as many mixed strategies map onto the same
@@
 \end{align*}
-A discrete-time, two-population formulation of the \rmd is this 
+A discrete-time, two-population formulation of the \rmd can be defined as follows 
 \citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
@@
 mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
-have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver. This
-is the content of Theorem~\ref{thm:Correspondence} given below, a
-proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
-section introduces the relevant construction that maps arbitrary $C$
-to suitable $Q^C$.
+have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver.
+%This is the content of Theorem~\ref{thm:Correspondence} given below, a
+%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
+%section introduces the relevant construction that maps arbitrary $C$
+%to suitable $Q^C$.
 

### 2014-10-21 paper/03_simulations.tex
Some small changes.

@@
 would take us, then apply diffusion. To understand the \rdd
-better, Section~\ref{sec:iterated-diffusion} looks closer at the novel
-diffusion part, before Section~\ref{sec:simulations} explores how
+better, in Section~\ref{sec:iterated-diffusion} we look closer at the novel
+diffusion part, before exploring in Section~\ref{sec:simulations} how
 diffusion and fitness-based selection interact.
@@
 state space, but we stick with Nosofsky's similarity function for the
-time being, because it is a mathematically simple, yet established
+time being, because it is mathematically simple, yet an established
 notion in mathematical psychology.
@@
 
-  \caption{Examples for Nosofsky-similarity (non-normalized) and
+  \caption{Examples of Nosofsky-similarity (non-normalized) and
     derived probability of state confusion (normalized) for
@@
 200 rounds. In contrast, the outcome of a trial with identical
-parameters, except that imprecision was $\impairment = 0.1$, which is
+parameters, except with imprecision $\impairment = 0.1$, which is
 shown in Figure~\ref{fig:example_stratsB}, had converged (in our
@@
 
-To explore further what happened in all of our 50 trials for each
+To further explore what happened in all of our 50 trials for each
 parameter triple, we defined a small number of metrics that aim to
@@
 of both sender and receiver strategies appear to be non-decreasing
-under increasing impairment, with the only exception the 6 states
+under increasing impairment, with the only exception being the 6 states
 games and the transition from $\impairment = 0$ to $\impairment =

### 2014-10-21 paper/04_discussion.tex
Some small changes.

@@
 before. One possibility is that diffusion is noise in the adoption of
-strategies, e.g., by conditional imitation, of strategies by other
+strategies, e.g. by conditional imitation, of strategies by other
 signalers. Another possibility is that diffusion is again noise on the
@@
 function. Still, to the extent that the chosen lower bounded utility
-functions are reasonable ---and we think they are very reasonable---,
+functions are reasonable---and we think they are very reasonable---
 the case suggests that quantal responses are not a good model,
@@
 vague. Vagueness is more plausibly an effect of perceptual confusion
-of similar states, than of computational errors in maximizing expected
+of similar states than of computational errors in maximizing expected
 utility.

### 2014-10-21 paper/05_conclusions.tex
Some small changes.

@@
 also showed that there exists a translation of the confusion matrix
-$C$ into a suitable mutation matrices $Q_C$ and $R_C$ such that the
+$C$ into suitable mutation matrices $Q_C$ and $R_C$ such that the
 effect of former and latter are effectively equivalent, although $C$
@@
 receiver's strategies are diffused, as long as the sender's are. But,
-again, this issues must wait for further exploration. Connected to
+again, these issues must wait for further exploration. Connected to
 that is the obvious extension that also allows for stochastic

### 2014-11-04 paper/paper_short.tex
paper short version

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../helpers/myenvironments}
\usepackage{../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

% \linespread{1.5}

\title{Vagueness, Noise, and Signaling}
% \author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

\begin{document}

\maketitle

\begin{abstract}
  Signaling games are appealing models for studying the evolution of
  meaning, but typically do not explain vagueness as a feature of
  successful signaling.  Complementing recent like-minded models, we
  introduce the replicator diffusion dynamic as a special case of the
  replicator mutator dynamic that combines evolutionary pressure on
  optimal signal use with stochastic noise in the processing of
  similar stimuli. Applying this new dynamic to a generalization of
  Lewis' signaling games, we show that stochastic imprecision leads to
  vague, yet communicative efficient signal use, and, moreover, also
  unifies evolutionary outcomes and helps avoid suboptimal
  categorization.
\end{abstract}

Many of our concepts and words are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially the
case for the logico-semantic tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to. But there are other intriguing aspects about vagueness.

The puzzle that we are concerned with here is how vagueness could
arise and be maintained in the first place. This is a serious worry to
functionalist accounts that maintain that our concepts and language
use evolved in order to be efficient. Since the existence of unclear
borderline cases seems to entail inefficiency of categorization or
communication, the challenge, succinctly put forward by
\citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
exist despite its obvious and demonstrable functional deficiency under
evolutionary pressure to be optimally expressive. We address
\emph{Lipman's problem} as a technical challenge,
namely to find a conceptually sound and mathematically coherent formal
model of the evolution of categorization and language use that yields
by-and-large efficient categories \emph{and} vagueness.

A number of authors have recently tried to explain why vagueness
evolved as something that is itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of
limitations in information processing
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
paper contributes to the latter line of thought. Concretely, we
introduce the replicator diffusion dynamics ---a novel variant of the
replicator mutator dynamic--- that integrates stochastic noise on the
differential confusability of similar stimuli. The dynamic proposed
here generalizes and complements previous like-minded accounts. We
show that stochastic noise can not only lead to vague, yet
communicative efficient signal use, but can also unify evolutionary
outcomes and help avoid suboptimal categorization.

The next section introduces the background against which the work
presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
introduces the replicator diffusion dynamic and elaborates on its
relation with the replicator mutator
dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
diffusion dynamic. Section~\ref{sec:discussion} reflects and compares
our approach to others.

\section{Background}
\label{sec:background}

% The view that vagueness is a natural concomitant of cognitive
% limitations of language users has been formalized in a number of ways,
% using evolutionary game theory and certain generalizations of Lewis'
% signaling games \citep{Lewis_1969:Convention}, so called
% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
% Our contribution is best seen in relation to these accounts, as it also
% relies on sim-max games. Let's introduce these first, and then zoom in
% on the problem of vagueness.

\subsection{Sim-max games \& conceptual spaces}

Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
a sender and a receiver. The sender knows the true state of the world,
but the receiver does not. The sender can select a signal, or message,
to reveal to the receiver, who then chooses an act. In Lewis' games,
if the receiver chooses the act that corresponds to the actual state,
the play is a success, otherwise a failure. Certain regular
combinations of sender signaling and receiver reaction make messages
meaningful, in the sense that their use is correlated systematically
to certain states or acts. To investigate the conditions under which
such meaning-generating behavior can evolve is a highly interesting
topic that we are only beginning to fully understand
\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.

Similarity-maximizing games, or \emph{sim-max games} for short, are
generalizations of Lewis' games where different states are allowed to
be more or less similar to one another. While Lewis' games treated
communicative success as a matter of black and white, sim-max games
allow for a gradient notion of communicative success: the more similar
the receiver's interpretation is to the actual state, the more
successful the play is taken to be. Signaling games with
utility-relevant similarities in the state space are fairly standard
in economics
\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
but have received particular attention in a more philosophical and
linguistic context for reasons that will become clear presently
\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.

A sim-max game, in the relevant sense here, consists of a set of
states $\States$, a set of messages $\Messgs$ with much fever messages
than states, a prior probability distribution $\Pr \in
\Delta(\States)$ that gives the occurrence probabilities of states, a
similarity metric on states $\similarity \mycolon \States \times
\States \rightarrow \mathds{R}$, and a utility function $\Utils
\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
the receiver's acts with the states of the world, so that the game is
one of guessing the actual state, so to speak. We also assume that
sender's and receiver's interests are alike, so we only have one
utility function. We do not consider message costs, so utilities only
depend on the actual state, and the receiver's response. The
similarity function should satisfy the usual requirements for a metric
and the utility function should be a monotonically decreasing function
of similarity.

\citet{JagerMetzger2011:Voronoi-Languag} showed that the
evolutionarily stable states of sim-max games are remarkably
systematic. Their results were obtained for games with infinitely many
states in $n$-dimensional Euclidean space $\States \subseteq
\mathds{R}^n$ and a quadratic loss function for utilities
$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
\mystate{2})^2$. For these games, the evolutionarily stable states are
demonstrably so-called Voronoi languages. Roughly put, a Voronoi
language is a pair of sender and receiver strategies, such that the
sender strategy partitions the state space into convex categories,
while the receiver's interpretations are the central spots in each
category. For example, if \States is the unit interval and all states
are equiprobable, an \ESS of a sim-max game with two messages would
have the sender use one message exclusively for all points in the
lower half of the unit interval and another for all points in the upper
half; the receiver's interpretations of messages are the central
points in the respective intervals.

This result is interesting, because it demonstrates that signaling can
impose orderly categories on a metric space, without that being the
ulterior purpose of it all. Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information by selecting a small sample of,
suggestively, category labels. These are passed to a second module
that tries to decode the original information. In this case, evolving
Voronoi languages would explain why \emph{conceptual categories} are
well-behaved and orderly in the way that they appear to be.

Seen in this light, sim-max games may provide a foundation to
approaches in cognitive semantics that rely on the notion of
conceptual spaces.  \citet[][70--77]{Gardenfors2000:Conceptual-Spac},
for example, has prominently argued that natural categories are convex
regions in conceptual space. If the conceptual space has a suitable
metric, convex categories can be derived from a set of prototypes. The
category corresponding to prototype $p$ is the set of points that are
more similar to $p$ than to any other
\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
categorization system can be obtained: storing the prototypes lets us
recover the categories without having to store each category's
extension. However, what is left unexplained, is where the prototypes
come from, and why we would not see just any distribution of
prototypes as an equally efficient classification system. This is
where sim-max games can contribute a principled approach to deriving,
in an independent way, not only convex categories but also
prototypical exemplars belonging to them.

\subsection{Vague signaling in sim-max games}

This outline of an approach to categorization using sim-max games
leaves many problems unaddressed. One of them is that natural
categories for continuously variable stimuli usually do not have
unique, point-valued prototypes and clear category boundaries. We
would like to account for the possibility of such vagueness, in
particular for the following criteria: (i) clear positive examples of
a vague category should show a gradient transition to clear negative
examples to (so-called \emph{higher-order vagueness}); (ii) prototypes
should likewise be gradient regions, peaking at the center of the
vague category they represent
\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re}.

\citet{DouvenDecock2011:Vagueness:-A-Co} show that
\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
approach can be extended to account for the existence of borderline
cases. From the assumption that prototypes are extended, yet convex
regions in conceptual space, they give a construction algorithm that
yields ``collated Voronoi diagrams'' with thick boundaries
representing borderline
regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
is possible to arrive at higher-order vagueness, by weighing in the
distance of different borderline cases to various prototypical
regions. This accounts for the first of the two desiderata mentioned
above, but still assumes that crisp prototype regions must be
given.

An alternative approach is taken by
\citet{FrankeJager2010:Vagueness-Signa} and
\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
can be met by evolving strategies in sim-max games. We should
distinguish micro- and macro-level approaches. Micro-level approaches
look at adaptive behavior of individual agents. Usually, changes in
the behavioral dispositions of agents occur after every single
interaction. In contrast, macro-level approaches outline more
abstract, aggregate dynamics, happening in a population of agents, or
otherwise abstracting from seemingly irrelevant detail. Usually, a
macro-level dynamic captures changes of frequencies of behavioral
types in the population over time.

As for micro-level approaches, \citet{FrankeJager2010:Vagueness-Signa}
show how limited memory of past interactions can lead to vague signal
use, when averaging over a single agent's behavior over time or over
the momentary behavior of a population of several language
users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
reinforcement learning that entails a low-level form of stimulus
generalization. Agents update their behavior after each round of play
in such a way that states similar to the one that actually occurred
are subject to behavioral adjustment as well.

\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
approach, using the notion of a \emph{quantal response}. A quantal
response function is a probabilistic choice rule that formalizes the
idea that agents make small mistakes when calculating the expected
utility of choice options. In aggregation, these probabilistic
mistakes lead to systematic ``trembles'' that produce vague signal
use.

The approach we take here is superficially similar to the latter, but
there are crucial differences. For one, we adopt a dynamic perspective
by looking at limited-time outcomes of a dynamic process. For another,
we demonstrate in Section~\ref{sec:discussion} that quantal responses
can give rise to counterintuitive predictions. These counterintuitive
examples suggest that vagueness in sim-max games is not convincingly
explained by appeal to mistakes in calculating expected utility, but
rather, as we assume here, as the result of confusing similar
states. A more in-depth discussion of alternative approaches is
deferred until Section~\ref{sec:discussion}.  Let us first look into
our own proposal in more detail.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Replicator diffusion dynamic}
\label{sec:repl-diff-dynam}

Against this background, we introduce a new macro-level approach,
where the source of vagueness is the agents' natural inability to
sharply distinguish similar states. We call our dynamic
\emph{replicator diffusion dynamic} (\rdd), because it is an extension
of the replicator dynamic (\rd) that adds diffusion of behavior, and a
special case of the replicator mutator dynamic (\rmd).

We begin by recapitulating the formulation of the \rd in
Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
a special case of the \rmd.

\subsection{Replicator dynamic in behavioral strategies}
\label{sec:repl-dynam-behav}

Fix a signaling game with finite states $\States$, messages $\Messgs$
and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
distribution over states and $\Util_{\sen,\rec} \mycolon \States
\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
receiver's utility functions. The sender's behavioral strategies are
functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
utility of choices in each choice point as usual:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
The \emph{fitness} of a behavioral strategy at a choice point is the
frequency-weighted average of expected utilities of each choice, given
the opponent's strategy:
\begin{align*}
  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg, \state,\Rstrat) \\
\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
\cdot \EU(\act, \messg,\Sstrat)\,.
\end{align*}
The discrete-time replicator dynamic maps current strategies $\Sstrat$
and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
in such a way that changes in frequency are proportional to expected
utilities. For behavioral strategies, the changes take place locally
at each of the agents' choice points:
\begin{align*}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

\subsection{Replicator diffusion dynamic in behavioral strategies}
\label{sec:repl-diff-dynam-1}

The replicator diffusion dynamic was first introduced by \citet{Correia2013:The-Bivalent-Tr}
as a noise-perturbed variant of the replicator dynamic (\rd) in
behavioral strategies. Fix a sim-max game with
$\States = \Acts$ and a confusion matrix $C : \States \times
\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
the probability that $\state_i$ is realized as $\state_j$. The
confusability of states affects senders and receivers alike, but in
slightly different ways (see
Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
$C_{ij}$ is the probability that the actual state \mystate{i} is
perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
probability that \mystate{j} is the interpretation that is actually
formed when \mystate{i} is the intended interpretation.

\begin{figure}
  \centering

    \begin{tikzpicture}[node distance = 2cm, thick]

      \begin{scope}

      \node[rectangle, draw=mycol, fill=mycol!25, thick] (actual)
      {actual state $\mystate{i}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      actual] (perceived) {perceived state $\mystate{j}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      perceived] (output) {chosen message $\messg$};

      \node[rectangle, thick, below of = output, node distance=1cm]
      (sender) {sender};

      \draw[->] (actual) -> (perceived) node[midway,left] {noise
        $C_{ij}$};

      \draw[->] (perceived) -> (output) node[midway,left] (label)
      {strategy $\Sstrat(\messg \probbar \mystate{j})$};

%      \begin{pgfonlayer}{background}
%        \node [draw=black!50, fill=black!20,fit=(actual) (label)
%        (output)] {};
%      \end{pgfonlayer}
    \end{scope}

      \begin{scope}[xshift=5cm]

      \node[rectangle, draw=mycol, fill=mycol!25, thick] (message)
      {observed message $\messg$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      message] (target) {target interpretation $\mystate{i}$};

      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
      target] (realized) {realized interpretation $\mystate{j}$};

      \node[rectangle, thick, below of = realized, node distance=1cm]
      (sender) {receiver};

      \draw[->] (message) -> (target) node[midway,right] (bla) {strategy
        $\Rstrat(\mystate{i} \probbar \messg)$};

      \draw[->] (target) -> (realized) node[midway,right]  {noise $C_{ij}$};

%      \begin{pgfonlayer}{background}
%        \node [draw=black!50, fill=black!20,fit = (message) 
%        (realized) (bla)] {};
%      \end{pgfonlayer}
    \end{scope}

  \end{tikzpicture}

  \caption{Effect of confusion of states on sender and receiver
    choices.}
  \label{fig:noise-perturbation-of-strategies}
\end{figure}

The aggregate effect of confusion of states on behavioral strategies
can be captured in a function that maps behavioral strategies
$\Sstrat$ and $\Rstrat$ to their diffusions $D(\Sstrat)$ and
$D(\Rstrat)$. The idea is that $\Sstrat$ and $\Rstrat$ are what, on
average, the idealized, noise-free behavior would be, while
$D(\Sstrat)$ and $D(\Rstrat)$ are behavioral strategies that describe
the agents' actual noise-perturbed probabilistic behavior. Since the
effect of confusion of states is that behavior at one choice point
percolates to behavior at similar choice points, we speak of
\emph{diffusion of behavior under confusion of states}. If we
conceive of behavioral strategies $\Sstrat$ and $\Rstrat$ as
row-stochastic matrices, the diffusion effect of state confusability
is easily captured by matrix multiplication as a diffusion function
$\Diff_C$ that, based on confusion matrix $C$, maps behavioral
strategies onto their diffused realizations:
\begin{align}
  \label{eq:confusion-function}
  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
\end{align}

The discrete-time replicator diffusion dynamic takes the replicator
dynamic as basic, but factors in the confusion of states at each
update step in a sequential update:
\begin{align*}
  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
\end{align*}
This is equivalent to the following, perhaps more transparent
formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
\begin{align*}
  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
  \frac{\Sstrat(\messg \probbar \state_j) \cdot
    \EU(\messg, \state_j,\Rstrat)}
  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
  \frac{\Rstrat(\state_j \probbar \messg) \cdot
    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}

The idea behind the sequential definition of the \rdd is that, at each
time step, strategies are gradually optimized along the current
fitness landscape, as described by the \rd, but the realization of
optimized strategies is noisy, due to similarity of states. It is
obvious that other (discrete-time) evolutionary dynamics can be
subjected to state-confusability in an analogous fashion (see also
Section~\ref{sec:model-interpretation}). In the case of the \rd,
however, perturbation by state-confusion has a prominent close
relative in the replicator mutator dynamic.

\subsection{The replicator mutator dynamic}
\label{sec:repl-mutat-dynam}

The \rmd has been proposed first in the context of signaling game
models for the evolution of grammar
\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
adds stochastic mutation to the replicator dynamic in order to capture, in
the context of grammar evolution, the differential learning success of
different dispositions in first language acquisition. The relation of
the \rmd to other prominent evolutionary dynamics is well understood
\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
fruitful applications in the context of signaling games
\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}. It
is therefore desirable to relate the \rdd that we propose here to the
\rmd.

A direct comparison between \rdd and the \rmd is not possible, because
the usual formulation of the latter is in its continuous-time form,
and based on mixed strategies, not behavioral strategies. We therefore
give a discrete-time formulation of the \rmd in behavioral strategies,
which is independently useful. The relation between \rdd and \rmd will
then be plain to see. We start with a delineation of mixed and
behavioral strategies.

\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
functions $\Spure \in \Messgs^\States$ ($\Rpure \in
\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
(\Acts^\Messgs)$). The latter give the relative population frequencies
of the former. We write $\Smixed_i$ for the frequency
$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
receiver.

Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
\begin{align}
  \label{eq:CorrespondenceBehavioralMixed}
  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
\end{align} 
Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
is \emph{not} an injection, as many mixed strategies map onto the same
behavioral strategy. In this sense, mixed strategies hold more
information than behavioral strategies about the distribution of pure strategies in a population.

In the context of evolutionary dynamics, we can think of behavioral
strategies as treating each choice point as an independent update site
\citep[e.g.][]{Cressman2003:Evolutionary-Dy,Sandholm2013:Population-Game}. Dynamics
for mixed strategies conceive of agents as adjusting their whole pure
strategy globally. In this way, dependencies between different choices
at different choice points can matter to the evolutionary path. In
contrast, dynamics for behavioral strategies conceive of agents as
adjusting their strategies locally at each choice point, regardless of
what they are disposed to do at other choice points. The latter
perspective is then more coarse-grained, and not necessarily
equivalent to the former. But an advantage of dynamics in behavioral
strategies is that they reduce complexity, which is especially helpful
for computational simulations of larger games, like the ones we are
interested in here.

\paragraph{Replicator dynamic in mixed strategies.} To understand the
\rmd in mixed strategies, we first introduce the \rd in mixed
strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
F_k^{\Rmixed}$ is the average fitness in the sender population and
$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
average fitness in the receiver population.

The continuous-time replicator dynamic defines the change of frequency
of mixed strategies. Its so-called two-population version, that takes
sender and receiver roles as separate update sites:
\begin{align*}
  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
  \Phi(\Rmixed,\Smixed) \right ) \,.
\end{align*}
Similar to the version for behavioral strategies, a discrete-time
formulation of the \rd for mixed strategies can be conceptualized as an
update function that maps mixed strategies $\Smixed$ and $\Rmixed$
onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
such that:
\begin{align*}
  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
    F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
\end{align*}

\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
extends the \rd by adding probabilistic mutation. Let $Q$ be a
row-stochastic mutation matrix where $Q_{ji}$ gives the probability
that pure sender strategy $\Spure_j$ mutates into
$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
where $R_{ji}$ gives the probability that pure receiver strategy
$\Rpure_j$ mutates into $\Rpure_i$.

The \rmd is usually given only in its continuous-time form. A
two-population (non-payoff adjusted) formulation is then:
\begin{align*}
  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
\end{align*}
A discrete-time, two-population formulation of the \rmd can be defined as follows 
\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:

\begin{align*}
  \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
    F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
  \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
    \Phi(\Rmixed,\Smixed)}\,.
\end{align*}

% \begin{proof}
%   It suffices to check either sender or receiver part. Focusing on the
%   former, we need to show that the continuous version is derivable
%   from the discrete version if the discrete update steps get
%   infinitely small so that:
%   \begin{align*}
%     \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
%     \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
%     \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
%         F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
%     & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
%   \end{align*}
%   As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
%   it to derive the non-payoff adjusted continuous version above,
%   since:
%   \begin{align*}
%     & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
%       ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
%         F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
%         \Phi(\Smixed,\Rmixed) \\
%        = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
%         F_j^{\Rmixed} - \Smixed_i \cdot
%         \Phi(\Smixed,\Rmixed) 
%   \end{align*}
% \end{proof}

The discrete-time \rmd has the same sequential nature as the \rdd:
first we compute the fitness-driven change according to the standard
replicator dynamic; then we compute the perturbation from
mutation. This becomes clear if we define independent mutation
functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
of (one round of) mutation:
\begin{align}
  \label{eq:Mutation}
  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
  R_{ji} \,.
\end{align}
The discrete-time \rmd given above can then be rewritten as:
\begin{align*}
  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,. 
\end{align*}

\subsection{Diffusion as mutation}
\label{sec:diffusion-as-special}

The sequential reformulation of the \rmd in terms of a sequence of
updates already closely resembles our initial formulation of the
\rdd. However, there is a major difference between these. While the
\rdd is a dynamic for behavioral strategies, the \rmd, as taken from
the literature, applies to mixed strategies. Similarly, the
confusability of states is an operation that is straightforwardly
definable for behavioral strategies (see
Equations~(\ref{eq:confusion-function})), while mutation in the \rmd
is defined as mutation from one pure strategy to another. Nonetheless,
it is possible to map each confusion matrix to a mutation matrix in a
natural way so that the correspondence between mixed and behavioral
strategies are conserved. This is the main result presented in this
section, and it is our justification for regarding the \rdd as the
behavioral-strategy analogue of the replicator-mutator dynamic when
the only source of mutation is confusability of states.

\begin{figure}
  \centering

  \begin{tikzpicture}[node distance = 1cm]

    \begin{scope}
      \node[] (mixed) {$\Smixed$};

      \node[below of = mixed] (mutated) {$Q^C(\Smixed)$};

      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};

      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};

      \draw[->, dotted, thick] (mixed) -> (mutated);

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!10,fit = (dummyl) 
        (mutated) (dummyr)] (mixedSpace) {};
      \end{pgfonlayer}

      \node[below of = mixedSpace, node distance = 1.7cm] {space of
        mixed strategies};
    \end{scope}

    \begin{scope}[xshift=6cm]
      \node[] (behavioral) {$\Sstrat$};

      \node[below of = behavioral] (mutatedBeh) {$C(\Sstrat)$};

      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};

      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};

      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);

      \begin{pgfonlayer}{background}
        \node [draw=black!50, fill=black!10,
              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
      \end{pgfonlayer}

      \node[below of = behSpace, node distance = 1.7cm] {space of
        behavioral strategies};
    \end{scope}

    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
    above]{$G(\Smixed)$};

    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(Q^C(\Smixed))$};

  \end{tikzpicture}

  \caption{Correspondence between state-confusion and mutation.}
  \label{fig:correspondence-result}
\end{figure}

More concretely, since there is a non-injective mapping $G$ from
behavioral to mixed strategies, we would like to show that there is a
systematic translation from a confusion matrix $C$ to a mutation
matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
mutation-perturbation $Q^C(\Smixed)$ (see
Figure~\ref{fig:correspondence-result}). In other words, we
hypothesize that a confusion matrix $C$ should give rise to a unique
mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver.
%This is the content of Theorem~\ref{thm:Correspondence} given below, a
%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
%section introduces the relevant construction that maps arbitrary $C$
%to suitable $Q^C$.

\paragraph{Confusion-based mutations.} There are natural conversions
of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
we start with that.

The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
diffusion with $C$ is the product of the probabilities, for each
$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
\begin{align}
  \label{eq:construction-rec}
  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
\end{align}

Now look at the sender. The probability that $\Spure_j$ is realized as
$\Spure_i$ under diffusion with $C$ is the product of the
probabilities, over all states $\mystate{k}$, that the message
$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
$\mystate{k}$ in the absence of noise, is produced by a noisy
realization of $\Spure_j$, which is the probability $\sum_{\state_l
  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
would map unto $\Spure_i(\mystate{k})$. So, define:
\begin{align}
  \label{eq:construction-sen}
  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
\end{align}

For example, consider a signaling game with two states and two
messages. Let the confusion matrix be:
\begin{align*}
  C=
  \begin{pmatrix}
    .8 & .2 \\
    .2 & .8 
  \end{pmatrix}\,.
\end{align*}
The resulting mutation matrices are:
\begin{align*}
Q^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & 1 & 0 & 0 & 0 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & 0 & 0 & 0 & 1 \cr}
&                    
  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
                      11 & .64 & .16 & .16 & .04 \cr
                      12 & .16 & .64 & .04 & .16 \cr
                      21 & .16 & .04 & .64 & .16 \cr
                      22 & .04 & .12 & .12 & .64 \cr}\,.
\end{align*}
Here, a pair like $21$, for example, refers to a pure sender strategy
with $\Spure(\state_1)=\messg_2$ and $\Spure(\state_2) =
\messg_1$. Similarly for the receiver.

Based on the translations from confusion matrices $C$ to mutation
matrices $Q^C$ and $R^C$ in Equations~(\ref{eq:construction-rec}) and
(\ref{eq:construction-sen}), we can finally formulate the
desired correspondence result, that reveals the \rdd as a special case
of the \rmd in behavioral strategies where the only source of mutation is
imperfect discriminability of states.

\begin{theorem}
  \label{thm:Correspondence}
  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
  \Diff_C(\Sstrat)$. And, (ii) if $G(\Rmixed) = \Rstrat$, then
  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
\end{theorem}

\noindent A proof is given in Appendix~\ref{sec:proofs}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Exploring the RDD}
\label{sec:exploring-rdd}

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. To understand the \rdd
better, in Section~\ref{sec:iterated-diffusion} we look closer at the novel
diffusion part, before exploring in Section~\ref{sec:simulations} how
diffusion and fitness-based selection interact.

\subsection{(Iterated) Diffusion}
\label{sec:iterated-diffusion}

Confusion of states should intuitively be a function of their
perceptual similarity. To make this concrete, let us assume that the
state space of the sim-max game consists of $n \ge 2$ states that are
equally spaced across the unit interval, including $0$ and $1$. The
distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical
similarity between two states $\mystate{i}$ and
$\mystate{j}$. Distance in physical space feeds into a perceptual
similarity function, as described by
\citet{Nosofsky1986:Attention-Simil}:
\begin{align*}
  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) & \textrm{otherwise,} \\
    \end{cases}
\end{align*}
where $\impairment >= 0$ is an imprecision or indiscriminability
parameter. When $\impairment=0$ agents perfectly discriminate between
states; when $\impairment \rightarrow \infty$ agents cannot
discriminate states at all. The top part of
Figure~\ref{fig:NosofskySim} gives an impression of
Nosofsky-similarity for different parameter values. Other
formalizations of perceptual similarity are possible, including ones
that allow for different discriminability in different areas of the
state space, but we stick with Nosofsky's similarity function for the
time being, because it is mathematically simple, yet an established
notion in mathematical psychology.

\begin{figure}
  \centering

  \includegraphics[width=0.85\textwidth]{plots/NosofskySim.pdf}

  \caption{Examples of Nosofsky-similarity (non-normalized) and
    derived probability of state confusion (normalized) for
    $\card{\States} = 50$ and different
    values of imprecision $\impairment$.}
  \label{fig:NosofskySim}
\end{figure}

We further assume that the probability of confusing any two states
$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
similarity and therefore obtained by normalization:
\begin{align*}
  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
  \impairment)} \,.
\end{align*}
Confusion of states is then also a function of imprecision parameter
$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$ is
positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
$i$ and $j$.

If states are confused on average with a probability proportional to
their similarity, repeated applications of diffusion of behavior as
defined in Equation~(\ref{eq:confusion-function}) causes behavioral
dispositions gradually to diffuse along a gradient of similarity of
states as well. Consequently, iterated diffusion leads to a smoothing
out and an eventual equalization of behavioral strategies, for both
the sender and the receiver. After $n$ steps of iterated diffusion an
initial sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an
initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we
assume that $\impairment > 0$, $C$ is a positive row-stochastic matrix
(each state could in principle be confused as any other state, albeit
with possibly a very low probability). It then follows from the
Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
all of its rows are identical. That is why, in the limit of iterated
diffusion, $C^\infty S$ has identical rows (messages are sent with the
same probability in each state), and $R C^\infty$ likewise has
identical rows (every message is interpreted equally). All of this is
in line with the intuition that diffusion of behavior under confusion
of states iteratively equalizes behavior for similar states; if all
states can be confused for one another in principle, behavior smoothes
out entirely in the limit.

\subsection{Exploring the replicator diffusion dynamics}
\label{sec:simulations}

Now that we understand better what diffusion does, let us look at how
diffusion interacts with optimization of behavior, as described by the
replicator dynamics. The main question we should address is whether
diffusion and fitness-based replication reasonably interact, and if
so, whether the inclusion of diffusion has any noteworthy effects on
the evolving meaning of signals. Clearly, if $\impairment = 0$, the
\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
diffusion component takes over and the \rdd reduces to the trivial
iterated diffusion process that we looked at in the previous
section. We will show presently that for reasonable in-between levels
of imprecision $\impairment$, the \rdd leads to communicatively
successful, yet vague signal meaning. Suitable levels of imprecision
can have further accelerating and unifying effects on meaning
evolution.

\subsubsection{Experimental set-up}

To explore the \rdd, we turn to numerical simulations. Let states be
evenly spaced elements of the unit interval that always include 0 and
1, and let priors be flat: $\Pr(\state) = \Pr(\state')$ for all
$\state, \state'$. As for utility functions, we take another
Nosofsky-style function:
\begin{align*}
  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
      \begin{cases}
    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
 \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
    \end{cases}
\end{align*}
The tolerance parameter $\toler \ge 0$ models the amount of tolerable
pragmatic slack or communicative imprecision. This choice of utility
function is governed partly by convenience in parallel to the
similarity function, but also because it has, as we believe, the right
general properties for a communicative payoff function. Unlike
utilities that are, say, linearly or quadratically decreasing in
physical distance \citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
utilities that are exponentially decreasing in negative quadratic
distance can model situations where a small amount of imprecision in
communication is tolerable, whereas similarly small differences in
intolerably far away interpretations matter very little, with a smooth
transition between these regimes
\citep[c.f.][]{OConnor2013:The-Evolution-o}.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp0_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp01_ind3452.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
    \label{fig:example_stratsB}
  \end{subfigure}

  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
  %   \label{fig:example_stratsC}
  % \end{subfigure}
  % \hfill
  % \begin{subfigure}[]{0.45\textwidth}
  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
  %   \label{fig:example_stratsD}
  % \end{subfigure}

  \caption{Example strategies under \rdd at stopping time of our simulations.}
  \label{fig:example_strats}
\end{figure}

We ran 50 trials of the \rdd, starting with randomly sampled sender
and receiver strategies, for each triplet of independent parameter
values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
  0.3}$. Each trial ran for a maximum of 200 update steps of the
\rdd. This may not be enough to guarantee convergence to the eventual
attracting state, but we were interested especially in whether the
dynamics could lead to reasonable outcomes in reasonable time. A trial
was considered converged and stopped before the maximum of 200 rounds
if the total amount of change between strategies before and after the
current \rdd step was smaller than 0.001, i.e., if:
\begin{align*}
  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
    \rdd(\Sstrat)(\state,\messg) - \Sstrat(\state,\messg) \right| <
  0.001\,, & \text{\ \ and} \\
  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left|
    \rdd(\Rstrat)(\messg,\state) - \Rstrat(\messg,\state) \right| <
  0.001 & \,.
\end{align*}

Representative examples for resulting strategy pairs at stopping time
are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
shows a strategy pair at stopping time with 90 states, tolerance
$\toler = 0.1$ and imprecision $\impairment = 0$. Zero imprecision
means that the trial was effectively an application of the plain
\rd. Noteworthily, the given sender strategy approximates a pure
sender strategy that crisply partitions the state space into
non-convex sets. The irregular shape of the receiver strategy shows
clearly that the pictured strategy pair has not yet reached a
dynamically stable state under the \rd. Indeed, the trial did not
converge in our technical sense, but was stopped after the maximum of
200 rounds. In contrast, the outcome of a trial with identical
parameters, except with imprecision $\impairment = 0.1$, which is
shown in Figure~\ref{fig:example_stratsB}, had converged (in our
technical sense) after 99 rounds of iterated \rdd. The sender strategy
shows a smooth blending from one ``category'' to the other, and the
receiver's interpretations are rather extended curves, peaking at a
central point in the relevant ``categories.''

These examples already show two interesting things. Firstly, inclusion
of imprecision can lead to seemingly well-behaved, yet vague
strategies in the sense that we are after. The sender strategy in
Figure~\ref{fig:example_stratsB} identifies clear positive and clear
negative cases for each signal, with a smooth transition
in-between. The receiver's interpretations of signals can be seen as
smoothed-out prototype regions. Secondly, (sender) strategies can
approach non-convex pure strategies under the replicator dynamic. We
see this in our limited-time simulations, but this also holds, for
some types of utility functions, for the limiting case. This was first
observed by Elliott Wagner in unpublished work
\citep[c.f.][]{OConnor2013:Evolving-Percep}. A full analysis of the
dynamics of sim-max games is beyond the scope of this paper, but we
will see shortly that diffusion from confusability of states clearly
prevents evolutionary paths that meander for a long time in the
vicinity of non-convex strategies.

\subsubsection{Dependent measures}

To further explore our simulation results, we defined a small number
of metrics that aim to capture numerically how vague, communicatively
efficient and generally well-structured the resulting strategy pairs
were. \emph{Entropy} and \emph{informativity} capture (slightly
different aspects of) the amount of systematicity or regularity in
signal use. \emph{Voronoiness} and \emph{convexity} are, respectively,
gradient and categorical measures for the coherence of signal-induced
categories. \emph{Expected utility} measures the communicative
efficiency of evolved strategy pairs.

\paragraph{Entropy.} This classic information-theoretic notion
captures the amount of uncertainty in a probability
distribution. Roughly put, entropy of a signaling strategy captures
inverse distance from a pure strategy. The usual definition of entropy
applies directly to mixed strategies:
\begin{align*}
  E(\Smixed) & = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot
  \log(\Smixed(\Spure)) & 
  E(\Rmixed) & = \sum_{\Rpure \in \States^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
\end{align*} 
These measures are computationally expensive to calculate, since the
size of the domain over which the sum is computed grows exponentially
with the number of choice points. But provably equivalent metrics for behavioral
strategies are ready to hand:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) 
\end{align*} 
\begin{align*}
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
\log(|\States|)$. We work with normalized values.

The sender strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.84$. In general, we expect that vague languages will have
higher entropy than crisp ones and that increasing imprecision will
also increase entropy, all else being equal.

\paragraph{Informativity.} The quantity of information in a signal is
an old notion that also goes back to the start of information theory.
\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of
signaling games. Intuitively speaking, informativity in Skyrms' sense
measures how revealing the sender's signal use is about which state is
actual. It therefore measures, in our given set-up with flat priors, a
sense of well-behavedness or regularity of a sender strategy.

The main idea behind Skyrms' notion is that we can quantify the amount
of information about a state $\state$ in a message $\messg$ via the
relation between the probability of the state given the message
$\Pr(\state \,|\, \messg)$ and the unconditional probability of the
state $\Pr(\state)$.  Following Bayes' theorem, we can express
$\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg
  \,|\, \state)}{\Pr(\messg)}$.  We then have $\Pr(\messg \,|\,
\state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) =
\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot
\Sstrat(\state^\prime,\messg)$.  Finally, we equate sender
infomativity with the average overall information about states in each
signal.  Following \citet[p.~36]{Skyrms2010:Signals}, and the
considerations above, the informativity about states conveyed by a
sender strategy is:
\begin{align*}
  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs}
  \sum_{\state \in \States} \frac{\Pr(\state) \cdot
    \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
    \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log
  \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
      \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right)
  \,. 
\end{align*}
This measure is bounded between $0$ and $1$. The sender strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
informativity $0.69$ and $0.54$, respectively. In general, we expect
that vague languages will have lower informativity than crisp
ones and that increasing imprecision will decrease informativity.

\citet[p.~39]{Skyrms2010:Signals} also offers a notion of
informativity about the acts performed by the receiver, but it turns
out that this is perfectly correlated with information about the
states in the results of our simulations. We therefore only look at
informativity about states.

% Conversely, we can quantify the amount of information about an act in a message.
% We equate receiver informativity with the average overall information about acts in each signal.
% Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
% \begin{align*}
%   I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
% \end{align*}

\paragraph{Voronoiness.}
This metric aims to quantify how close a strategy is to being a part
of a Voronoi tessallation of the state space. Bearing in mind that
this metric is specifically geared towards the sim-max games in our
set-up, we define Voronoiness as follows. Let $\pi(\Rstrat, \messg) =
\argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype
of a message $\messg \in \Messgs$, then
\begin{align*}
  V(\Sstrat, \Rstrat) & = \sum_{\state \in \States} \sum_{\messg_1 \in
    \Messgs} \sum_{\messg_2 \in \Messgs} F(\Sstrat, \Rstrat, \state,
  \messg_1, \messg_2) \text{\ \ \ with } \\
  F(\Sstrat, \Rstrat, \state,
  \messg_1, \messg_2) & =
  \begin{cases}
    1 & \text{if } \similarity(\state, \pi(\Rstrat, \messg_1)) >
    \similarity(\state, \pi(\Rstrat, \messg_2)) \bicond \\
    &  \ \ \ \ \ \Sstrat(\state, \messg_1) >
    \Sstrat(\state, \messg_2) \\
    0 & \text{otherwise.}
  \end{cases}
\end{align*}
is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
strategy $\Rstrat$ and
\begin{align*}
  V(\Rstrat) & = \sum_{\messg \in
    \Messgs} \sum_{\state_1 \in \States}  \sum_{\state_2 \in \States}
  F(\Rstrat, \messg, \state_1, \state_2) \text{\ \ \ with } \\
  F(\Rstrat, \messg, \state_1, \state_2) & =
  \begin{cases}
    1 & \text{if } \similarity(\state_1, \pi(\Rstrat,
  \messg)) > \similarity(\state_2, \pi(\Rstrat, \messg)) \\
  & \ \ \ \ \ \bicond \Rstrat(\messg,
  \state_1) > \Rstrat(\messg, \state_2) \\
    0 & \text{otherwise.}
  \end{cases}
\end{align*}
is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
are lower bound by $0$ and upper bound by, respectively, $|\States|
\times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we will
normalize them by dividing by these values.

Despite being vague, a language can have maximal Voronoiness according
to this gradient measure. For instance, the sender strategies in
Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
Voronoiness $0.78$ and $1$, respectively. We do not \emph{prima facie}
expect a conceptual connection between vagueness and Voronoiness, and
so do not predict a clear effect of imprecision on Voronoiness either.

\paragraph{Convexity.} At least for sender strategies, which converge
faster, it also makes sense to define a categorical measure of
convexity, that compensates for potential vagueness. To determine
whether a sender strategy $\Sstrat$ is convex despite possibly being
vague, we look at the derived pure strategy $\Spure$ for which
$\Spure(\state) = \argmax_{\messg' \in \Messgs}
\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
$\Sstrat$ as convex. The sender strategy in
Figure~\ref{fig:example_stratsA} is not convex, while the one in
Figure~\ref{fig:example_stratsB} is. Again, we do not expect a
conceptual relation between vagueness and imprecision, on the one
hand, and convexity on the other.

\paragraph{Expected utility.} We also record the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game.

The strategy pair in Figure~\ref{fig:example_stratsA} has a normalized
expected utility of $0.992$, the pair in
Figure~\ref{fig:example_stratsB} has $0.901$. Generally, vagueness and
imprecision can be expected to decrease expected utility
\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is whether
communicative success drops unacceptably fast with moderate levels of
vagueness and imprecision.

\subsubsection{Results}

\begin{figure}
  \centering

  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}

  \caption{Means of recorded metrics for each triple of independent
    variables. Notice that the plots also show 95\% confidence
    intervals, but these are most often invisible, owing to the
    uniformity of outcomes across trials.}
  \label{fig:MeanMetrics}
\end{figure}

Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
together with their estimated 95\% confidence intervals, for each
triple of independent variables. Most trends are quite clear, and not
unexpected. Entropy of sender and receiver strategies are increasing
with imprecision. Entropy of sender strategies seems independent of
tolerance levels and the number of states, while there seems to be an
effect of both parameters on the entropy of the receiver
strategy. Sender informativity and expected utility are decreasing
under growing imprecision. There seem to be mild effects of the number
of states in the game, but hardly any in terms of pragmatic
tolerance. What was not expected \emph{a priori} was that Voronoiness
of both sender and receiver strategies appear to be non-decreasing
under increasing impairment, with the only exception being the 6 states
games and the transition from $\impairment = 0$ to $\impairment =
0.05$. Expected utility of evolved languages mostly decreases in
increasing imprecision, with the sole exception of a rise for $\toler
= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized expected
utility of evolved strategy pairs under mild values of imprecision are
not necessarily much worse than their respective perfect precision
\rd-analogues. This is especially so for lower numbers of states
(which increase the chance of non-convergence and non-convexity; see
Figure~\ref{fig:CategoricalMeasures} and the discussion below).

Taken together, these findings suggest the following effect of
confusability of states and diffusion of behavior. Mild forms of
perceptual imprecision lead to slightly less communicative efficiency,
more vagueness and more regular, well-behaved languages. This
impression is also supported by the proportion of converged and convex
trials, pictured in
Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
sender strategies showed for several values of parameters in the
absence of impairment. Increasing impairment always assured
convergence and convexity: diffusion speeds up convergence to a convex
and regular language.

\begin{figure}
  \centering

  \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}

  \caption{Proportion of converged and convex trials.}
  \label{fig:CategoricalMeasures}
\end{figure}

Diffusion also has another interesting regularizing effect on the
evolution of signaling. If we look at the confidence intervals in
Figure~\ref{fig:MeanMetrics}, we see that there is very little
variation in the recorded metrics for evolved strategies, at least for
higher values of impairment. On closer inspection, it turns out that
variability in low-impairment conditions is not only due to
non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
gives two more examples of strategy pairs at stopping time. Both are
obtained for the same triple of parameters, both converged and are
convex. However, they are not equally efficient. In fact, the pair in
Figure~\ref{fig:example_stratsC} has a normalized expected utility of
$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
$0.89$.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind41.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
    \label{fig:example_stratsC}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind23.pdf}
    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
    \label{fig:example_stratsD}
  \end{subfigure}

  \caption{More example strategies under \rdd at stopping time of our simulations.}
  \label{fig:MoreExample_strats}
\end{figure}

Interestingly, this type of variability in evolutionary outcomes is
also weeded out by mild increases in impairment. To see this, we
calculated the average distance between evolved sender strategies
within each group of trials that had identical independent
parameter values. We determined the distance between sender strategies
$\Sstrat$ and $\Sstrat'$ as the average Hellinger distance between
probability distributions $\Sstrat(\state)$ and $\Sstrat'(\state)$
over messages at each choice point $\state$:
\begin{align*}
  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
 \sqrt{\sum_{\messg \in  \Messgs}
         \left ( \sqrt{\Sstrat(\state,\messg)} -
         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
\end{align*}
To compensate for the arbitrariness of message use, we set the
distance between strategies $\Sstrat$ and $\Sstrat'$ to be the maximum
of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
\emph{within group distances}, i.e., the average distances between all
sender strategies in each parameter group, are plotted in
Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
increasing imprecision, the sender strategies evolving under \rdd were
exactly alike (modulo swapping of messages). In other words,
perceptual imprecision unifies evolutionary outcomes, and in fact
guarantees sender strategies that are not only convex, but also
maximally efficient in that they induce a vague category split exactly
in the middle of the unit interval.

\begin{figure}
  \centering

    \includegraphics[width=0.8\textwidth]{plots/WithinGroupMeasures.pdf}

    \caption{For each group, i.e., triplet of independent variables,
      the plots show (i) within group distances and (ii) within group
      expected utilities, as defined in the main text.}
  \label{fig:AverageEUinGroups}
\end{figure}

This unifying property of perceptual imprecision could be considered
an evolutionary beneficial side-effect. Higher imprecision can lead to
higher \emph{within group expected utility}, i.e., the average
expected utility that each evolved language (i.e., pair of sender and
receiver strategy at stopping time) scored when playing against an
arbitrary other language of the same parameter group (see
Figure~\ref{fig:AverageEUinGroups}). What this means is that, while
imprecision might decrease the communicative efficiency of individual
languages, it increases the conceptual coherence and communicative
success between independently evolving strategies. It is almost as if
mere confusability of states imposes a regularity constraint on
evolving categories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{sec:discussion}

\subsection{Model interpretation}
\label{sec:model-interpretation}

We mentioned in passing in Section~\ref{sec:repl-diff-dynam-1} that
adding diffusion to other discrete-time evolutionary dynamics that
operate on behavioral strategies is entirely straightforward. We
could, for instance, easily diffuse the outcome of a best-response
dynamic at each time step. That would make sense if we thought of
agents as prone to confusing similar states, but otherwise rationally
optimizing behavior. The reason that we chose the \rd to combine with
diffusion is twofold. A minor reason is that it makes for a
conceptually interesting link with the \rmd. A more important reason
is that the \rd is especially versatile and non-committal about what
the exact process of adaption is that is being modelled. Generally,
the \rd could be a biological or a cultural dynamic, and so could the
\rdd.

Under a biological interpretation of the \rd, signaling strategies
are innate behavioral dispositions of organisms, steps in the
evolutionary process are successive generations, and selection
captures the reproductive advantage of fitter individuals. This
interpretation, strictly speaking, requires formalization in terms of
mixed strategies via the \rmd, and possibly also a symmetrizing of the
game, so that every agent is assumed to have a unique sender and
receiver role at the same time, the pair of which is bequeathed onto
the next generation.

Diffusion, in this context, could be either of two things. It could be
differential mutation probabilities in line with other interpretations
of mutation in the \rmd
\citep[e.g.][]{NowakKomarova2001:Evolution-of-Un,KomarovaNiyogi2001:The-Evolutionar}:
pure strategies are not faithfully reproduced, e.g. by imperfections
of learning or underdetermination of input, and mistakes in inheriting
pure strategies are more likely if they result from the confusion of
similar states. Another possible biological interpretation of the
\rdd, is that inheritance is faithful, but strategies are noisily
realized. Strategies are not necessarily selected for what they are,
but rather for how they are realized, once noise is factored in.

As a cultural dynamic, the \rd can also be interpreted as a high-level
description of the likely development of other behavioral adaptation
processes, like differential imitation \citep[see][for various
derivations of the \rd]{Sandholm2013:Population-Game}. Under this
interpretation, what is subject to evolutionary optimization is
behavior, not genetic make-up. Fitness captures the success of
behavioral patterns.  Differential reproduction represents the
tendency of more successful ways of behavior to be more likely
realized, be it through imitation or some solitary learning or
optimization mechanism.

Under this non-biological interpretation, we have again two options of
picturing what diffusion is, similar to the biological cases
before. One possibility is that diffusion is noise in the adoption of
strategies, e.g. by conditional imitation, of strategies by other
signalers. Another possibility is, again, that diffusion is noise on
the realization of strategies: while behavior is optimized to be
efficient (be it due to learning, introspection or imitation),
realization of strategies is bound to be noisy due to confusability of
similar states.

We believe that all of the four mentioned interpretations are, on
first approximations, feasible conceptualizations of the \rdd, and
that it is a good thing to know of a working account of vague
signaling that sketches where fitness-based selection under
state-confusability will take us, abstracting away from the details of
actually playing the game, inheriting, imitating or otherwise
optimizing behavior in whatever particular way. It is a good thing to
know this on the macro-level, especially since there are also
micro-level accounts that nicely complement the picture. We turn to
one such next.

\subsection{Relation with previous accounts}
\label{sec:relat-with-prev}

\paragraph{Generalized reinforcement.}
% Repeated below...
%\citet{OConnor2013:The-Evolution-o} introduced a generalization of
%Herrnstein reinforcement learning for sim-max games and showed that
%this not only leads to vague signaling patterns, but can speed-up
%evolution of efficient signaling strategies, especially in games with
%many states.

Under \emph{Herrnstein reinforcement learning}, sender and receiver
adjust their dispositions to act after each round of play, in such a
way that the actual (non-negative) payoff gained in the previous
interaction is added to the non-normalized propensities for repeating
the same behavior. For sim-max games, this means that, when the sender
chose $\messg$ in state $\state$, and this resulted in some
non-negative payoff, the probability that the sender chooses $\messg$
again in $\state$ is increased, but nothing else changes. In
particular, the sender's behavior in other choice points does not
change. \emph{Generalized reinforcement learning} is different. When
the use of $\messg$ in $\state$ gave positive payoff, then not only
will its future use probability be promoted at $\state$ but also at
other states, proportional to how similar these are to
$\state$. Similar amendments take care of the way that the receiver
updates his choice dispositions.

\citet{OConnor2013:The-Evolution-o} shows that this extension not only
leads to vague signaling of the appropriate kind, but also speeds up
learning in such a way that, especially for sim-max games with higher
numbers of states, higher levels of communicative success are reached
in shorter learning periods. Technically, this result is partly due to
the fact that signalers make bigger changes to their behavioral
strategies after each round of play under generalized reinforcement
than under the Herrnstein variety. But that only explains the speed of
adaptation, not necessarily that generalization also leads to
regularity and communicative efficiency, but it does.

Diffusion of strategies in the \rdd can also be conceived of as a form
of generalization, and works in large part quite analogous to stimulus
generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
approach. But there are still
differences. \citeauthor{OConnor2013:The-Evolution-o} showed that
generalized reinforcement learning can speed up the development of
efficient signaling, especially for higher numbers of
states. Complementing this, we showed that diffusion regularizes
evolutionary outcomes and prevents meandering around sub-optimal and
non-convex strategies, also for low numbers of states.

Conceptually, the \rdd is a more abstract framework than generalized
reinforcement learning. The latter is foremost motivated as a learning
dynamic that has two players adapt their individual strategies after
each concrete round of play. In contrast, the \rdd describes a more
abstract, average dynamical change in behavioral
dispositions. Although the behavior of (some forms of) reinforcement
learning mirror those of the \rd (at some stage in time)
(\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
this does not mean that \emph{generalized} reinforcement learning is
also necessarily a plausible high-level description of, say,
generalized learning in a population of agents. Seen in this light,
generalized reinforcement learning and the \rdd nicely complement each
other, as similarly-minded accounts operating at different levels of
abstraction.

\paragraph{Quantal response equilibria.}
\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in
which information-processing limitations could lead to vague
strategies. The model that is most clearly related to the present
approach uses the notion of a \emph{quantal response}, also known as a
\emph{soft-max function}
\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. A
quantal response function is a paramterized generalization of the
classic best response function. For example, if $U \mycolon \Acts
\rightarrow \mathds{R}$ is the measure of expected utility over
choices $\Acts$ of an agent, then a best response function would have
the agent choose $\act$ only if $U(\act) = \max_{\act' \in \Acts}
U(a)$. A quantal response function rather assumes that agents would
choose $\act$ with a probability proportional to $\expo(\lambda \cdot
U(\act))$, where $\lambda$ is a rationality parameter. If $\lambda
\rightarrow \infty$ we retrieve the behavior of the best response
function, but if it is positive but finite, any choice $\act$ will
receive a positive probability, but acts with higher expected utility
will be more likely. The underlying motivation is the assumption that
there is noise in the computation of expected utilities and/or in
maximization of expected utilities. Consequently, choices with almost
equal expected utilities will be chosen with almost equal probability
(for moderate values of $\lambda$).

\citet{FrankeJager2010:Vagueness-Signa} show that quantal response
equilibria of sim-max games, i.e., pairs of sender and receiver
strategies such that the sender strategy is the quantal response to
the expected utilities under the receiver strategy and vice versa, can
show the desired marks of vague
signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
quantal response equilibrium for a sim-max game, as used in our set-up
but with higher tolerance $\toler = 0.5$. Sender and receiver
strategies look very much like what evolves under \rdd with modest
values of perceptual imprecision.

\begin{figure}
  \centering

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
    \label{fig:exampleQRE_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
    \label{fig:exampleQRE_stratsB}
  \end{subfigure}

  \caption{Examples of vague quantal response equilibria.}
  \label{fig:exampleQREs}
\end{figure}

But not all quantal response equilibria are equally plausible, and the
ones that are not suggest that it is less natural to think of
vagueness as arising from errors in the computation or maximization of
expected utility than that it arises from confusion of similar
states. To see what the problem is, we can look at cases like given in
Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
equilibrium for a game with lower tolerance $\toler = 0.05$. Unlike
what evolves under \rdd in this case, sender strategies have vague
boundaries also towards the end of the unit intervals. Technically,
this is because quantal responses equalize message use far away from
the ``prototypical'' interpretation, not just in-between categories,
so to speak. This, in turn, is because quantal responses introduce
noise into the decision making at the level of computing or maximizing
expected utility of choices.

That this is conceptually odd shows even more clearly in a case where
the state space is intuitively unbounded, as for instance for the
property ``tall''. If the usual interpretation of a ``tall man'' peaks
at around, say, 195cm then when meeting a giant of $n$ meters, senders
would, according to the quantal response approach, be ever more
inclined to describe the giant indifferently as either ``tall'' or
``short'' the larger $n$ gets. This is because, as the distance from
the prototype increases for larger $n$, the difference between the
expected utilities of saying ``short'' or ``tall'' will converge to
zero. 
%Whence that the quantal response approach would predict that
%senders would grow indifferent between choice of antonyms as $n$
%grows, which seems weird.

Admittedly, this argument hinges on the choice of utility
function. Still, to the extent that the chosen lower bounded utility
functions are reasonable---and we think they are very reasonable---
the case suggests that quantal responses are not a good model,
intuitively speaking, for why linguistic categories are
vague. Vagueness is more plausibly an effect of perceptual confusion
of similar states than of computational errors in maximizing expected
utility.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
\label{sec:conclusions}

Vagueness is a pervasive but seemingly non-disruptive feature of
natural communication and classification systems. From an evolutionary
point of view, the challenge arises to explain how vagueness can
persist under selective pressure for precision. This is foremost a
technical challenge, probing for the possibility of integrating into a
consistent model forces that lead to vagueness and forces that lead
to efficient information transfer. This paper proposed one such model
in the context of sim-max games and explored some of its consequences.

We introduced the replicator diffusion dynamic as a special case of
the replicator mutator dynamic. The \rdd incorporates a special kind
of stochastic mutation based on differential confusability of similar
states. Based on data from numerical simulations, we demonstrated that
the inclusion of mild levels of stochastic confusability of states
does not undermine the possibility of evolving communicatively
successful signaling strategies at all. On the contrary, strategies
that evolved under mild diffusion induce a highly regular and
systematic category structure that shows the signs of vagueness as
desired. There might even be a higher-order benefit to the presence of
imprecision, in that it can accelerate convergence to optimal
categorization, by preventing evolutionary paths to stay near
inefficient non-convex strategies for a long time. Diffusion thus also
unifies and regularizes evolutionary outcomes in such a way that
sub-optimal categorization is avoided.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Proof of Theorem}
\label{sec:proofs}

\begin{proof}[Part (i).]
  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
  the consequent:
  \begin{align*}
    \Diff_C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
    \cdot \Sstrat(\messg_y \probbar \state_l) && \text{(by Equation~(\ref{eq:confusion-function}))} \\
    & =  \sum_{\state_l} C_{xl}
    \cdot  \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i && \text{(by Equation~(\ref{eq:CorrespondenceBehavioralMixed}))} \\
    & = \sum_{\state_l}
    \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i \cdot C_{xl}\,.
  \end{align*}
  Next consider the lhs of the consequent:
  \begin{align*}
    G(\Mutate_{Q^C}(\Smixed))(\messg_y \probbar \state_x) & =
    \sum_{\Spure_i(\state_x)=\messg_y} \Mutate_{Q^C}(\Smixed_i) &&
    \text{(by Equation~(\ref{eq:CorrespondenceBehavioralMixed}))} \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed_j \cdot Q^C_{ji} &&
    \text{(by Equation~(\ref{eq:Mutation}))} \\
    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
    \Smixed_j \cdot \prod_{\state_l} \sum_{\state_m \in
      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} &&
    \text{(by Equation~(\ref{eq:construction-sen}))} \\
    & = \sum_{\Spure_j} \Smixed_j \cdot
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
  \end{align*}
  To simplify this further we look at a fixed $\Spure_j$ and consider
  the term: 
  \begin{align}
    \label{eq:term}
    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
  \end{align}
  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
  over all pure strategies, we essentially look at each such
  mapping. Term (\ref{eq:term}) above sums over all pure strategies
  that map $\state_k$ onto $\messg_y$. The above term then sums over
  all products whose factors are tuples in $\times_{k>2} \set{y
    \setbar \exists l \mycolon y = Y_{kl}}$. So term (\ref{eq:term})
  expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
  \begin{align*}
    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
    \dots \\
    & + (Y_{11} \cdot Y_{2d} \cdot
    Y_{3d} \cdot \ldots \cdot
    Y_{ed}) 
  \end{align*}
  But since $Y$ is a row-stochastic matrix, this simplifies to
  $Y_{xy}$. Continuing the derivation with this:
  \begin{align*}
    G(\Mutate_{Q^C}(\Smixed))(\messg_y \probbar \state_x) 
    & = \sum_{\Spure_j} \Smixed_j \cdot
    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed_j \cdot
     C_{xl} \\
     & = \sum_{\state_l}
    \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i \cdot C_{xl}\,.
  \end{align*}

\end{proof}

\begin{proof}[Part (ii).]
  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
  expands to:
  \begin{align*}
    \Diff_C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
    \Rmixed_i \cdot C_{jx} && \text{(by Equation~(\ref{eq:confusion-function}))} \\
    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
    \Rmixed_i \cdot C_{jx} \\
    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
  \end{align*}
  The rhs expands to (by Equations~(\ref{eq:CorrespondenceBehavioralMixed}),
  (\ref{eq:Mutation}) and (\ref{eq:construction-rec})):
  \begin{align*}
    G(\Mutate_{R^C}(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{\Rpure_i}
    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
        \state_x}} 
    \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)} \,.
  \end{align*}
  Without loss of generality, assume that $x=y=1$, and fix
  $\card{M}=d$ and let $e$ be the number of pure receiver
  strategies. Then the last term can be rewritten as:
  \begin{align*}
    & = \sum_{i}
    \Rmixed_i \cdot ( C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_1(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_1(\messg_d)} + \ldots  \\
      & \textcolor{white}{ = \sum_{i}
    \Rmixed_i  (}  +  C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_2(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_2(\messg_d)} + \ldots  \\
      & \textcolor{white}{ = \sum_{i}
    \Rmixed_i  (}  + C_{\Rpure_i(\messg_1)1} \cdot
      C_{\Rpure_i(\messg_2)\Rpure_e(\messg_2)} \cdot \ldots \cdot
      C_{\Rpure_i(\messg_d)\Rpure_e(\messg_d)}) ) \,.
  \end{align*}
  For every messages $\messg_l$, $C_{\Rpure_i(\messg_l)}$ is a stochastic
  vector. For $l>1$, all elements of these vectors appear equally
  often. But that means that these cancel out. What remains is:
  \begin{align*}
    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
  \end{align*}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\printbibliography[heading=bibintoc]

\end{document}

### 2014-11-05 paper/paper_short.tex
paper shortened, and plots resized/rearranged

@@
 \begin{abstract}
-  Signaling games are appealing models for studying the evolution of
-  meaning, but typically do not explain vagueness as a feature of
-  successful signaling.  Complementing recent like-minded models, we
-  introduce the replicator diffusion dynamic as a special case of the
-  replicator mutator dynamic that combines evolutionary pressure on
-  optimal signal use with stochastic noise in the processing of
-  similar stimuli. Applying this new dynamic to a generalization of
-  Lewis' signaling games, we show that stochastic imprecision leads to
-  vague, yet communicative efficient signal use, and, moreover, also
-  unifies evolutionary outcomes and helps avoid suboptimal
-  categorization.
+  Signaling games are popular models for studying the evolution of
+  meaning, but typical approaches do not allow for vagueness as a
+  feature of successful signaling.  Complementing recent like-minded
+  models, we introduce the replicator diffusion dynamic as a special
+  case of the replicator mutator dynamic that combines evolutionary
+  pressure on optimal signal use with stochastic noise in the
+  processing of similar stimuli. Applying this new dynamic to a
+  generalization of Lewis' signaling games, we show that stochastic
+  imprecision leads to vague, yet communicative efficient signal use,
+  and, moreover, also unifies evolutionary outcomes and helps avoid
+  suboptimal categorization.
 \end{abstract}
@@
 Wittgenstein which is challenged by the paradoxes vagueness gives rise
-to. But there are other intriguing aspects about vagueness.
-
-The puzzle that we are concerned with here is how vagueness could
-arise and be maintained in the first place. This is a serious worry to
+to.
+
+But there are other intriguing aspects about vagueness. The puzzle
+that we are concerned with here is how vagueness could arise and be
+maintained in the first place. This is a serious worry to
 functionalist accounts that maintain that our concepts and language
@@
 evolutionary pressure to be optimally expressive. We address
-\emph{Lipman's problem} as a technical challenge,
-namely to find a conceptually sound and mathematically coherent formal
-model of the evolution of categorization and language use that yields
-by-and-large efficient categories \emph{and} vagueness.
+\emph{Lipman's problem} as a technical challenge, namely to find a
+conceptually sound and mathematically coherent formal model of the
+evolution of categorization and language use that yields by-and-large
+efficient categories \emph{and} vagueness.
 
@@
 here generalizes and complements previous like-minded accounts. We
-show that stochastic noise can not only lead to vague, yet
-communicative efficient signal use, but can also unify evolutionary
-outcomes and help avoid suboptimal categorization.
+show that stochastic noise can not only lead to vague, yet efficient
+signal use, but can also unify evolutionary outcomes and help avoid
+suboptimal categorization.
 
@@
 utility function. We do not consider message costs, so utilities only
-depend on the actual state, and the receiver's response. The
+depend on the actual state and the receiver's response. The
 similarity function should satisfy the usual requirements for a metric
@@
 category. For example, if \States is the unit interval and all states
-are equiprobable, an \ESS of a sim-max game with two messages would
-have the sender use one message exclusively for all points in the
-lower half of the unit interval and another for all points in the upper
-half; the receiver's interpretations of messages are the central
-points in the respective intervals.
+are equiprobable, a Voronoi language with two messages would have the
+sender use one message exclusively for all points in the lower half of
+the unit interval and another for all points in the upper half; the
+receiver's interpretations of messages are the central points, .25 and
+.75, in the respective intervals.
 
@@
 thought of as distinct modules in a single system, where the first
-module must discretize the information by selecting a small sample of,
-suggestively, category labels. These are passed to a second module
-that tries to decode the original information. In this case, evolving
-Voronoi languages would explain why \emph{conceptual categories} are
-well-behaved and orderly in the way that they appear to be.
+module must discretize the information it is fed by selecting a small
+sample of, suggestively, category labels. These are passed to a second
+module that tries to decode the original information. In this case,
+evolving Voronoi languages would explain why \emph{conceptual
+  categories} are well-behaved and orderly in the way that they appear
+to be.
 
@@
 category corresponding to prototype $p$ is the set of points that are
-more similar to $p$ than to any other
-\citep[e.g.][]{OkabeBoots2000:Spatial-Tessell}. In this way,
+more similar to $p$ than to any other. In this way,
 \citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
@@
 would like to account for the possibility of such vagueness, in
-particular for the following criteria: (i) clear positive examples of
+particular: (i) clear positive examples of
 a vague category should show a gradient transition to clear negative
-examples to (so-called \emph{higher-order vagueness}); (ii) prototypes
+examples (so-called \emph{higher-order vagueness}); (ii) prototypes
 should likewise be gradient regions, peaking at the center of the
@@
 cases. From the assumption that prototypes are extended, yet convex
-regions in conceptual space, they give a construction algorithm that
+regions in conceptual space, a construction algorithm is available that
 yields ``collated Voronoi diagrams'' with thick boundaries
@@
 
-    \begin{tikzpicture}[node distance = 2cm, thick]
+    \begin{tikzpicture}[node distance = 1.6cm, thick]
 
@@
 
-      \node[rectangle, draw=mycol, fill=mycol!25, thick] (actual)
+      \node[rectangle, draw=black!50, fill=black!10, thick] (actual)
       {actual state $\mystate{i}$};
 
-      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
       actual] (perceived) {perceived state $\mystate{j}$};
 
-      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
       perceived] (output) {chosen message $\messg$};
 
-      \node[rectangle, thick, below of = output, node distance=1cm]
+      \node[rectangle, thick, above of = actual, node distance=0.75cm]
       (sender) {sender};
@@
 
-      \node[rectangle, draw=mycol, fill=mycol!25, thick] (message)
+      \node[rectangle, draw=black!50, fill=black!10, thick] (message)
       {observed message $\messg$};
 
-      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
       message] (target) {target interpretation $\mystate{i}$};
 
-      \node[rectangle, draw=mycol, fill=mycol!25, thick, below of =
+      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
       target] (realized) {realized interpretation $\mystate{j}$};
 
-      \node[rectangle, thick, below of = realized, node distance=1cm]
-      (sender) {receiver};
+      \node[rectangle, thick, above of = message, node distance=0.75cm]
+      (receiver) {receiver};
 
@@
 
-      \node[below of = mixedSpace, node distance = 1.7cm] {space of
+      \node[above of = mixedSpace, node distance = 1.4cm] {space of
         mixed strategies};
@@
 
-      \node[below of = behSpace, node distance = 1.7cm] {space of
+      \node[above of = behSpace, node distance = 1.4cm] {space of
         behavioral strategies};
@@
 
-For example, consider a signaling game with two states and two
-messages. Let the confusion matrix be:
-\begin{align*}
-  C=
-  \begin{pmatrix}
-    .8 & .2 \\
-    .2 & .8 
-  \end{pmatrix}\,.
-\end{align*}
-The resulting mutation matrices are:
-\begin{align*}
-Q^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
-                      11 & 1 & 0 & 0 & 0 \cr
-                      12 & .16 & .64 & .04 & .16 \cr
-                      21 & .16 & .04 & .64 & .16 \cr
-                      22 & 0 & 0 & 0 & 1 \cr}
-&                    
-  R^C & = \bordermatrix{ ~ & 11 & 12 & 21 & 22 \cr
-                      11 & .64 & .16 & .16 & .04 \cr
-                      12 & .16 & .64 & .04 & .16 \cr
-                      21 & .16 & .04 & .64 & .16 \cr
-                      22 & .04 & .12 & .12 & .64 \cr}\,.
-\end{align*}
-Here, a pair like $21$, for example, refers to a pure sender strategy
-with $\Spure(\state_1)=\messg_2$ and $\Spure(\state_2) =
-\messg_1$. Similarly for the receiver.
-
-Based on the translations from confusion matrices $C$ to mutation
-matrices $Q^C$ and $R^C$ in Equations~(\ref{eq:construction-rec}) and
-(\ref{eq:construction-sen}), we can finally formulate the
-desired correspondence result, that reveals the \rdd as a special case
-of the \rmd in behavioral strategies where the only source of mutation is
-imperfect discriminability of states.
+Based on Equations~(\ref{eq:construction-rec}) and
+(\ref{eq:construction-sen}), we can formulate the desired
+correspondence result. A proof is given in Appendix~\ref{sec:proofs}.
 
@@
 \end{theorem}
-
-\noindent A proof is given in Appendix~\ref{sec:proofs}.
 
@@
 
-  \includegraphics[width=0.85\textwidth]{plots/NosofskySim.pdf}
+  \includegraphics[width=0.7\textwidth]{plots/NosofskySim.pdf}
 
@@
   \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp0_ind3098.pdf}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
     \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
@@
   \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_NS-90_tol-01_imp01_ind3452.pdf}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind3452.pdf}
     \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
@@
 efficient and generally well-structured the resulting strategy pairs
-were. \emph{Entropy} and \emph{informativity} capture (slightly
-different aspects of) the amount of systematicity or regularity in
-signal use. \emph{Voronoiness} and \emph{convexity} are, respectively,
-gradient and categorical measures for the coherence of signal-induced
-categories. \emph{Expected utility} measures the communicative
-efficiency of evolved strategy pairs.
+were. \emph{Entropy} captures the amount of systematicity or
+regularity in signal use. \emph{Voronoiness} and \emph{convexity} are,
+respectively, gradient and categorical measures for the coherence of
+signal-induced categories. \emph{Expected utility} measures the
+communicative efficiency of evolved strategy pairs.
 
@@
 higher entropy than crisp ones and that increasing imprecision will
-also increase entropy, all else being equal.
-
-\paragraph{Informativity.} The quantity of information in a signal is
-an old notion that also goes back to the start of information theory.
-\citet[ch.~3]{Skyrms2010:Signals} discusses its use in the context of
-signaling games. Intuitively speaking, informativity in Skyrms' sense
-measures how revealing the sender's signal use is about which state is
-actual. It therefore measures, in our given set-up with flat priors, a
-sense of well-behavedness or regularity of a sender strategy.
-
-The main idea behind Skyrms' notion is that we can quantify the amount
-of information about a state $\state$ in a message $\messg$ via the
-relation between the probability of the state given the message
-$\Pr(\state \,|\, \messg)$ and the unconditional probability of the
-state $\Pr(\state)$.  Following Bayes' theorem, we can express
-$\Pr(\state \,|\, \messg)$ as $\frac{\Pr(\state) \cdot \Pr(\messg
-  \,|\, \state)}{\Pr(\messg)}$.  We then have $\Pr(\messg \,|\,
-\state) = \Sstrat(\state,\messg)$ and $\Pr(\messg) =
-\sum_{\state^\prime \in \States} \Pr(\state^\prime) \cdot
-\Sstrat(\state^\prime,\messg)$.  Finally, we equate sender
-infomativity with the average overall information about states in each
-signal.  Following \citet[p.~36]{Skyrms2010:Signals}, and the
-considerations above, the informativity about states conveyed by a
-sender strategy is:
-\begin{align*}
-  I(\Sstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs}
-  \sum_{\state \in \States} \frac{\Pr(\state) \cdot
-    \Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
-    \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \cdot \log
-  \left(\frac{\Sstrat(\state,\messg)}{\sum_{\state^\prime \in \States}
-      \Pr(\state^\prime) \cdot \Sstrat(\state^\prime,\messg)} \right)
-  \,. 
-\end{align*}
-This measure is bounded between $0$ and $1$. The sender strategies in
-Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
-informativity $0.69$ and $0.54$, respectively. In general, we expect
-that vague languages will have lower informativity than crisp
-ones and that increasing imprecision will decrease informativity.
-
-\citet[p.~39]{Skyrms2010:Signals} also offers a notion of
-informativity about the acts performed by the receiver, but it turns
-out that this is perfectly correlated with information about the
-states in the results of our simulations. We therefore only look at
-informativity about states.
-
-% Conversely, we can quantify the amount of information about an act in a message.
-% We equate receiver informativity with the average overall information about acts in each signal.
-% Based on Skyrms' definition~\citet[p.~39]{Skyrms2010:Signals}, we define receiver informativity as follows:
-% \begin{align*}
-%   I(\Sstrat,\Rstrat) = \frac{1}{|\Messgs|} \sum_{\messg \in \Messgs} \sum_{\act \in \Acts} \Rstrat(\messg,\act) \cdot \log \left(\frac{\Rstrat(\messg,\act)}{\sum_{\state \in \States} \Pr(\state) \cdot \sum_{\messg^\prime \in \Messgs} \Sstrat(\state,\messg^\prime) \cdot \Rstrat(\messg^\prime,\act)} \right) \,.
-% \end{align*}
+increase entropy, all else equal.
 
@@
     \Sstrat(\state, \messg_2) \\
-    0 & \text{otherwise.}
+    0 & \text{otherwise}
   \end{cases}
@@
   \state_1) > \Rstrat(\messg, \state_2) \\
-    0 & \text{otherwise.}
+    0 & \text{otherwise}
   \end{cases}
@@
 effect of both parameters on the entropy of the receiver
-strategy. Sender informativity and expected utility are decreasing
-under growing imprecision. There seem to be mild effects of the number
-of states in the game, but hardly any in terms of pragmatic
-tolerance. What was not expected \emph{a priori} was that Voronoiness
+strategy. What was not expected \emph{a priori} was that Voronoiness
 of both sender and receiver strategies appear to be non-decreasing
-under increasing impairment, with the only exception being the 6 states
-games and the transition from $\impairment = 0$ to $\impairment =
+under increasing impairment, with the only exception being games with
+6 states and the transition from $\impairment = 0$ to $\impairment =
 0.05$. Expected utility of evolved languages mostly decreases in
 increasing imprecision, with the sole exception of a rise for $\toler
-= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized expected
-utility of evolved strategy pairs under mild values of imprecision are
-not necessarily much worse than their respective perfect precision
-\rd-analogues. This is especially so for lower numbers of states
-(which increase the chance of non-convergence and non-convexity; see
-Figure~\ref{fig:CategoricalMeasures} and the discussion below).
+= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized
+expected utility of evolved strategy pairs under mild values of
+imprecision are not necessarily much worse than their respective
+perfect precision \rd-analogues. This is especially so for lower
+numbers of states (see discussion below).
 
@@
 trials, pictured in
-Figure~\ref{fig:CategoricalMeasures}. Non-converged and non-convex
+Figure~\ref{fig:MeanMetrics}. Non-converged and non-convex
 sender strategies showed for several values of parameters in the
@@
 
-\begin{figure}
-  \centering
-
-  \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}
-
-  \caption{Proportion of converged and convex trials.}
-  \label{fig:CategoricalMeasures}
-\end{figure}
+% \begin{figure}
+%   \centering
+
+%   \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}
+
+%   \caption{Proportion of converged and convex trials.}
+%   \label{fig:CategoricalMeasures}
+% \end{figure}
 
@@
   \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind41.pdf}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
     \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
@@
   \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_NS-6_tol-03_imp005_ind23.pdf}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
     \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
@@
 similar states. Another possible biological interpretation of the
-\rdd, is that inheritance is faithful, but strategies are noisily
+\rdd is that inheritance is faithful, but strategies are noisily
 realized. Strategies are not necessarily selected for what they are,
@@
 Under this non-biological interpretation, we have again two options of
-picturing what diffusion is, similar to the biological cases
-before. One possibility is that diffusion is noise in the adoption of
-strategies, e.g. by conditional imitation, of strategies by other
-signalers. Another possibility is, again, that diffusion is noise on
-the realization of strategies: while behavior is optimized to be
-efficient (be it due to learning, introspection or imitation),
-realization of strategies is bound to be noisy due to confusability of
-similar states.
+picturing what diffusion is. One possibility is that diffusion is
+noise in the adoption of strategies, e.g. by conditional imitation, of
+strategies by other signalers. Another possibility is, again, that
+diffusion is noise on the realization of strategies: while behavior is
+optimized to be efficient (be it due to learning, introspection or
+imitation), realization of strategies is bound to be noisy due to
+confusability of similar states.
 
@@
 the replicator mutator dynamic. The \rdd incorporates a special kind
-of stochastic mutation based on differential confusability of similar
+of stochastic mutation due to differential confusability of similar
 states. Based on data from numerical simulations, we demonstrated that

### 2014-11-05 snippets/Nosofsky_similarity.Rmd
paper shortened, and plots resized/rearranged

@@
 
-ggsave(p, file="paper/plots/NosofskySim.pdf", width=7, height=4)
+ggsave(p, file="paper/plots/NosofskySimConf.pdf", width=7, height=4)
+
+simDF = subset(simDF, simDF$normalized == "Nosofsky similarity")
+
+p = ggplot(simDF, aes(x = distance, y = similarity, shape = alpha_i))+
+  geom_line(color='grey')  + geom_point()  + 
+  ylab("") + 
+  xlab("distance in physical space") +
+  scale_shape_discrete(name="imprecision")
+
+show(p)
+
+ggsave(p, file="paper/plots/NosofskySim.pdf", width=6, height=2.25)
 

### 2014-11-05 data-analysis.R

@@ -25,15 +25,19 @@ plot.language <- function(speaker.filename, hearer.filename) {
     melted.speaker$State <- states
     melted.hearer$State <- states
     
-    speaker.plot <- ggplot(melted.speaker, aes(x=State, y=value)) +
-        geom_line(aes(linetype=variable)) + labs(title='Sender',x='State',y='') +
-        theme(legend.position = 'none') + ylim(0,1)
-    hearer.plot <- ggplot(melted.hearer, aes(x=State, y=value)) +
-        geom_line(aes(linetype=variable)) + labs(title='Receiver',x='State',y='') + 
-        theme(legend.position = 'none')
+    melted.speaker$role = "sender"
+    melted.hearer$role = "receiver"
     
-    return(arrangeGrob(speaker.plot, hearer.plot))
+    plotData = rbind(melted.speaker, melted.hearer)
+    plotData$role = factor(plotData$role, levels = c("sender", "receiver"))
     
+    p = ggplot(plotData, aes(x = State, y = value)) +
+      geom_line(aes(linetype=variable)) + labs(title='',x='State',y='') +
+      theme(legend.position = 'none') + 
+      facet_grid(role ~ ., scales = "free_y")
+    
+    return(p)
+      
 }
 
 plot.cluster <- function(language.names, data) {

### 2014-11-05 helper_functions.R

@@ -178,3 +178,20 @@ plot.receiver.strat = function(rec,states){
     geom_line(aes(linetype=message)) + labs(title='Receiver',x='State',y='') + 
     theme(legend.position = 'none')
 }
+
+plot.strats = function(sen, rec, states){
+  senMelt = melt(sen, varnames = c("state", "message"))
+  senMelt$state = rep(states, dim(sen)[2])
+  senMelt$message = factor(senMelt$message)
+  senMelt$role = "sender"
+  recMelt = melt(rec, varnames = c("message", "state"))
+  recMelt$state = rep(states, each = dim(rec)[1])
+  recMelt$message = factor(recMelt$message)
+  recMelt$role = "receiver"
+  plotData = rbind(senMelt, recMelt)
+  plotData$role = factor(plotData$role, levels = c("sender", "receiver"))
+  outplot = ggplot(data = plotData, aes(x = state, y = value)) + 
+    geom_line(aes(linetype=message)) + labs(title='',x='State',y='') +
+    theme(legend.position = 'none') + facet_grid(role ~ ., scale = "free_y")
+  return(outplot)
+}

### 2014-11-05 snippets/QRE.R

@@ -7,7 +7,7 @@ source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
 len = 1
 ns = 50
 lambda = 15
-tolerance = 0.05
+tolerance = 0.5
 
 states = seq(0, len, length.out = ns)
 priors = rep(1/ns, ns)
@@ -29,8 +29,8 @@ for (i in 1:100){
   rec = recNext
 }
 
-outPlot = arrangeGrob(plot.sender.strat(sen,states),
-                      plot.receiver.strat(rec,states))
+outPlot = plot.strats(sen,rec,states)
+
 show(outPlot)
 filename = paste("paper/plots/exampleStratQRE_tolerance0", tolerance*10, ".pdf", sep = "", collapse = "")
 ggsave(outPlot, file = filename, height = 5, width = 5)

### 2014-11-05 snippets/example_strats/example_strats_plots.R

@@ -10,24 +10,25 @@ plot.and.save.language.by.index = function(data, index){
   tolerance = data[index,]$Tolerance
   impairment = data[index,]$Impairment
   stratPlot = plot.language.byindex(data,index)
-  filename = paste("~/Desktop/data/svn/vagueness-games/paper/plots/strat_example_", "NS-", ns,                    "_tol-", tolerance, "_imp", impairment, "_ind", index, ".pdf", sep = "", collapse = "")
-  pdf(file = filename, height = 5, width = 5)
+  filename = paste("~/Desktop/data/svn/vagueness-games/paper/plots/strat_example_", "ind", index, ".pdf", sep = "", collapse = "")
+  pdf(file = filename, height = 2.75, width = 3)
   show(stratPlot)
   dev.off()
+  show(stratPlot)
 }
 
-stratsToPlot = c(1201, 1001, 3098, 3452)
+stratsToPlot = c(1201, 1001, 3098, 3452, 41, 23)
 
-# for (strat in stratsToPlot){
-#   plot.and.save.language.by.index(data,strat)
-# }
+for (strat in stratsToPlot){
+  plot.and.save.language.by.index(data,strat)
+}
 
 
 show(ds[stratsToPlot,])
 
-d = subset(data, Number.of.states == 6 & Tolerance == 0.3 & Impairment == 0.05 & Speaker.Convex.Cat == 1)
+d = subset(data, Number.of.states == 6 & Tolerance == 0.2 & Impairment == 0 & Speaker.Convex.Cat == 1)
 
 
-# plot.and.save.language.by.index(d, which.min(d$Expected.utility))
-# plot.and.save.language.by.index(d, which.max(d$Expected.utility))
-# d[c(which.min(d$Expected.utility), which.max(d$Expected.utility)),]
+plot.and.save.language.by.index(d, which.min(d$Expected.utility))
+plot.and.save.language.by.index(d, which.max(d$Expected.utility))
+d[c(which.min(d$Expected.utility), which.max(d$Expected.utility)),]

### 2014-11-05 paper/paper_short.tex
results graphs merged

@@
 
-  \includegraphics[width=\textwidth]{plots/MeanMetrics.pdf}
-
-  \caption{Means of recorded metrics for each triple of independent
-    variables. Notice that the plots also show 95\% confidence
-    intervals, but these are most often invisible, owing to the
-    uniformity of outcomes across trials.}
+  \includegraphics[width=\textwidth]{plots/MeanMetrics2.pdf}
+
+  \caption{Means of gradient and proportions of categorical metrics for
+    each triple of independent variables, with estimated 95\%
+    confidence intervals.}
   \label{fig:MeanMetrics}

### 2014-11-05 snippets/data_exploration/data_exploration_tmp.Rmd
results graphs merged

@@
 
+data$converged = data$Iterations < 201
+data$convex = data$Speaker.Convex.Cat
+
 ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
              "Speaker.entropy", "Hearer.entropy",
-             "Speaker.informativity",
              "Speaker.Voronoiness", "Hearer.Voronoiness",
-             "Expected.utility")]
+             "Expected.utility",
+             "convex", "converged"
+             )]
 
@@
         value[value=="Hearer.entropy"] <- "entropy rec"
-        value[value=="Speaker.informativity"] <- "informativity"
         value[value=="Speaker.Voronoiness"] <- "Voronoiness sen"
@@
      geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
-    scale_shape_discrete(name = "# states") +
-    scale_linetype_discrete(name = "# states") +
-    scale_color_discrete(name = "# states")
+    scale_shape_discrete(name = "# states", guide = guide_legend()) +
+    scale_linetype_discrete(name = "# states", guide = guide_legend()) +
+    scale_color_discrete(name = "# states", guide = guide_legend()) +
+    theme(legend.position="top")
 
@@
 
-ggsave(sp, file = "paper/plots/MeanMetrics.pdf", width=8, height = 8)
+ggsave(sp, file = "paper/plots/MeanMetrics2.pdf", width=8, height = 10)
 

### 2014-11-05 paper/paper_short.tex
cover letter

@@
 
-% \linespread{1.5}
+\doublespacing
 
@@
   Signaling games are popular models for studying the evolution of
-  meaning, but typical approaches do not allow for vagueness as a
+  meaning, but typical approaches do not incorporate vagueness as a
   feature of successful signaling.  Complementing recent like-minded
@@
   imprecision leads to vague, yet communicative efficient signal use,
-  and, moreover, also unifies evolutionary outcomes and helps avoid
+  and, moreover, unifies evolutionary outcomes and helps avoid
   suboptimal categorization.
@@
 
-Many of our concepts and words are vague. A vague category knows clear
+Many concepts and expressions are vague. A vague category knows clear
 cases that fall under it, clear cases that do not, and also so-called
@@
 of everyday communication, but it is troublesome for some of the most
-prominent theories of language and meaning. This is especially the
-case for the logico-semantic tradition of Frege, Russell and the young
+prominent theories of language and meaning. This is especially so for the logico-semantic tradition of Frege, Russell and the young
 Wittgenstein which is challenged by the paradoxes vagueness gives rise
@@
 
-The replicator diffusion dynamic was first introduced by \citet{Correia2013:The-Bivalent-Tr}
-as a noise-perturbed variant of the replicator dynamic (\rd) in
-behavioral strategies. Fix a sim-max game with
-$\States = \Acts$ and a confusion matrix $C : \States \times
+The replicator diffusion dynamic is a noise-perturbed variant of the
+replicator dynamic (\rd) in behavioral strategies. Fix a sim-max game
+with $\States = \Acts$ and a confusion matrix $C : \States \times
 \States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
@@
 This is equivalent to the following, perhaps more transparent
-formulation, given by \citet{Correia2013:The-Bivalent-Tr}:
+formulation:
 \begin{align*}

### 2015-01-05 paper/version_01/01_intro.tex
branched paper to version folder

--- 2014-10-21 paper/01_intro.tex
@@
 \end{align*}
-To keep matters simple, in this paper we mostly focus on cases where $\States$
-contains finitely many points from the unit interval, all of which
-have equal probability, and with similarity given by the Euclidean
-distance in this one-dimensional space.
+To keep matters simple, in this paper we mostly focus on cases where
+$\States$ contains finitely many points from the unit interval, all of
+which have equal probability.
 

### 2015-01-14 paper/version_02/paper.tex
interpretation of dynamics added to section 2

--- 2014-11-05 paper/paper_short.tex
@@
 
-\usepackage[]{../helpers/mypackages}
+\usepackage[]{../../helpers/mypackages}
 %\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
 %\bibliography{MyRefGlobal}
-\bibliography{../helpers/MyRefGlobal}
+\bibliography{../../helpers/MyRefGlobal}
 \bibliography{paper} 
-\usepackage{../helpers/myenvironments}
-\usepackage{../helpers/mycommands}
+\usepackage{../../helpers/myenvironments}
+\usepackage{../../helpers/mycommands}
 \usepackage{todonotes}
@@
   generalization of Lewis' signaling games, we show that stochastic
-  imprecision leads to vague, yet communicative efficient signal use,
-  and, moreover, unifies evolutionary outcomes and helps avoid
-  suboptimal categorization.
+  imprecision leads to vague, yet by-and-large communicative efficient
+  signal use, and, moreover, unifies evolutionary outcomes and helps
+  avoid suboptimal categorization.
 \end{abstract}
@@
 maintained in the first place. This is a serious worry to
-functionalist accounts that maintain that our concepts and language
-use evolved in order to be efficient. Since the existence of unclear
-borderline cases seems to entail inefficiency of categorization or
-communication, the challenge, succinctly put forward by
-\citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
-exist despite its obvious and demonstrable functional deficiency under
-evolutionary pressure to be optimally expressive. We address
-\emph{Lipman's problem} as a technical challenge, namely to find a
-conceptually sound and mathematically coherent formal model of the
-evolution of categorization and language use that yields by-and-large
-efficient categories \emph{and} vagueness.
+functionalist accounts that maintain that concepts and linguistic
+meanings evolved in order to be efficient. Since the existence of
+unclear borderline cases seems to entail inefficiency of
+categorization or communication, the challenge, succinctly put forward
+by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
+persist under evolutionary pressure to be optimally discriminative. We
+address \emph{Lipman's problem} as a technical challenge, namely to
+find a conceptually sound and mathematically coherent formal model of
+the evolution of categorization and language use that yields
+by-and-large efficient categories \emph{and} vagueness.
 
@@
 here generalizes and complements previous like-minded accounts. We
-show that stochastic noise can not only lead to vague, yet efficient
-signal use, but can also unify evolutionary outcomes and help avoid
-suboptimal categorization.
+show that stochastic noise can not only lead to vague, yet relatively
+efficient signal use, but can also unify evolutionary outcomes and
+help avoid suboptimal categorization.
 
@@
 \citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
-but have received particular attention in a more philosophical and
-linguistic context for reasons that will become clear presently
+but have received particular attention in a more philosophical context
+for reasons that will become clear presently
 \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.
 
-A sim-max game, in the relevant sense here, consists of a set of
-states $\States$, a set of messages $\Messgs$ with much fever messages
-than states, a prior probability distribution $\Pr \in
-\Delta(\States)$ that gives the occurrence probabilities of states, a
-similarity metric on states $\similarity \mycolon \States \times
-\States \rightarrow \mathds{R}$, and a utility function $\Utils
-\mycolon \States \times \States \rightarrow \mathds{R}$. We identify
-the receiver's acts with the states of the world, so that the game is
-one of guessing the actual state, so to speak. We also assume that
-sender's and receiver's interests are alike, so we only have one
-utility function. We do not consider message costs, so utilities only
-depend on the actual state and the receiver's response. The
-similarity function should satisfy the usual requirements for a metric
-and the utility function should be a monotonically decreasing function
-of similarity.
+A sim-max game consists of a set of states $\States$, a set of
+messages $\Messgs$ with much fever messages than states, a prior
+probability distribution $\Pr \in \Delta(\States)$ that gives the
+occurrence probabilities of states, a similarity metric on states
+$\similarity \mycolon \States \times \States \rightarrow \mathds{R}$,
+and a utility function $\Utils \mycolon \States \times \States
+\rightarrow \mathds{R}$. We identify the receiver's acts with the
+states of the world, so that the game is one of guessing the actual
+state, so to speak. We also assume that sender's and receiver's
+interests are alike, so we only have one utility function. We do not
+consider message costs, so utilities only depend on the actual state
+and the receiver's response. The utility function should be a
+monotonically decreasing function of similarity.
 
@@
 would like to account for the possibility of such vagueness, in
-particular: (i) clear positive examples of
-a vague category should show a gradient transition to clear negative
-examples (so-called \emph{higher-order vagueness}); (ii) prototypes
-should likewise be gradient regions, peaking at the center of the
-vague category they represent
-\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re}.
+particular: (i) clear positive examples of a vague category should
+show a gradient transition to clear negative examples (so-called
+\emph{higher-order vagueness});\todo{check usage of term} (ii)
+prototypes should likewise be gradient regions, peaking at the center
+of the vague category they represent
+\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re}.\todo{check
+  whether references are okay/appropriate in this context}
 
@@
 \citet{OConnor2013:The-Evolution-o} who show how the above desiderata
-can be met by evolving strategies in sim-max games. We should
-distinguish micro- and macro-level approaches. Micro-level approaches
-look at adaptive behavior of individual agents. Usually, changes in
-the behavioral dispositions of agents occur after every single
-interaction. In contrast, macro-level approaches outline more
-abstract, aggregate dynamics, happening in a population of agents, or
-otherwise abstracting from seemingly irrelevant detail. Usually, a
-macro-level dynamic captures changes of frequencies of behavioral
-types in the population over time.
-
-As for micro-level approaches, \citet{FrankeJager2010:Vagueness-Signa}
-show how limited memory of past interactions can lead to vague signal
-use, when averaging over a single agent's behavior over time or over
-the momentary behavior of a population of several language
-users. \citet{OConnor2013:The-Evolution-o} introduces a variant of
-reinforcement learning that entails a low-level form of stimulus
-generalization. Agents update their behavior after each round of play
-in such a way that states similar to the one that actually occurred
-are subject to behavioral adjustment as well.
-
-\citet{FrankeJager2010:Vagueness-Signa} also consider a macro-level
-approach, using the notion of a \emph{quantal response}. A quantal
-response function is a probabilistic choice rule that formalizes the
-idea that agents make small mistakes when calculating the expected
-utility of choice options. In aggregation, these probabilistic
-mistakes lead to systematic ``trembles'' that produce vague signal
-use.
-
-The approach we take here is superficially similar to the latter, but
-there are crucial differences. For one, we adopt a dynamic perspective
-by looking at limited-time outcomes of a dynamic process. For another,
-we demonstrate in Section~\ref{sec:discussion} that quantal responses
-can give rise to counterintuitive predictions. These counterintuitive
-examples suggest that vagueness in sim-max games is not convincingly
-explained by appeal to mistakes in calculating expected utility, but
-rather, as we assume here, as the result of confusing similar
-states. A more in-depth discussion of alternative approaches is
-deferred until Section~\ref{sec:discussion}.  Let us first look into
-our own proposal in more detail.
+can be met by evolving strategies in sim-max games. The approach we
+take here is similar in spirit, but different in relevant detail. We
+are argue here that \citeauthor{FrankeJager2010:Vagueness-Signa}'s
+approaches builds on a dubious assumption about the source of
+vagueness. While \citeauthor{OConnor2013:The-Evolution-o} gives an
+agent-level learning dynamic under which vague signaling arises, we
+complement the picture by giving a population-level dynamic that
+abstracts from the concrete behavior of signalers. A more in-depth
+comparison is deferred until Section~\ref{sec:discussion}. Let us
+first look into our own proposal in more detail.
 
@@
 
-Against this background, we introduce a new macro-level approach,
-where the source of vagueness is the agents' natural inability to
-sharply distinguish similar states. We call our dynamic
+We introduce a population-level dynamic in which communicatively more
+efficient strategies increase in proportion, as usual, but where the
+realization of signaling strategies is noisy because of the agents'
+inability to sharply distinguish similar states. We call our dynamic
 \emph{replicator diffusion dynamic} (\rdd), because it is an extension
-of the replicator dynamic (\rd) that adds diffusion of behavior, and a
-special case of the replicator mutator dynamic (\rmd).
+of the replicator dynamic (\rd) and a special case of the replicator
+mutator dynamic (\rmd).
+
+The formulation of the \rdd given and explored here is in terms of
+behavioral strategies, not in terms of mixed strategies (more on this
+below). Dynamics on behavioral strategies assume that agents can adapt
+their behavior locally, so to speak, i.e., independently at each
+choice point. This greatly reduces the complexity of the dynamic and
+simplifies numerical simulations. But it points towards an
+interpretation as a process of cultural evolution, rather than
+biological evolution: signaling strategies are not inherited as a
+whole, but adapted piece-meal, e.g., by conditional imitation (more
+below). Still, the approach we suggest here is compatible with a
+formulation in terms of mixed strategies and an interpretation as
+biological evolution as well. To show this, we translate our initial
+formulation of the \rdd into one in using mixed strategies. Doing so
+identifies the \rdd as a special case of the \rmd.
 
@@
 Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
-a special case of the \rmd.
+a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
+\ref{sec:diffusion-as-special} are conceptually important, but not
+strictly necessary to appreciate the main point of the paper.
 
@@
 \times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
-receiver's utility functions. The sender's behavioral strategies are
-functions $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are
-functions $\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected
-utility of choices in each choice point as usual:
+receiver's utility functions. A behavioral strategy is a function that
+maps an agent's choice points to a probability distribution over
+available choices. The sender's behavioral strategies are functions
+$\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are functions
+$\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected utility of
+choices in each choice point as usual:
 \begin{align*}
@@
 \end{align*}
+The \rd in behavioral strategies describes the most likely path of
+development of average behavior in a (huge and homogeneous) population
+of agents when behavioral adaptations occur independently at each
+choice point, e.g., when agents adopt the observed choice behavior
+\emph{at a given choice point} of another agent whose behavior
+\emph{at that choice point} has higher fitness (see
+\citeauthor{Sandholm2013:Population-Game}
+\citeyear{Sandholm2013:Population-Game} for agent-level
+interpretations of the replicator dynamics and
+\citeauthor{Cressman2003:Evolutionary-Dy}
+\citeyear{Cressman2003:Evolutionary-Dy} for applications to extensive
+form games and behavioral strategies).
 
@@
 The replicator diffusion dynamic is a noise-perturbed variant of the
-replicator dynamic (\rd) in behavioral strategies. Fix a sim-max game
-with $\States = \Acts$ and a confusion matrix $C : \States \times
-\States$. $C$ is a row-stochastic matrix whose elements $C_{ij}$ give
-the probability that $\state_i$ is realized as $\state_j$. The
-confusability of states affects senders and receivers alike, but in
-slightly different ways (see
+\rd in behavioral strategies. Fix a sim-max game with $\States =
+\Acts$ and a confusion matrix $C : \States \times \States$. $C$ is a
+row-stochastic matrix, i.e., each row is a probability
+distribution. Elements $C_{ij}$ give the probability that $\state_i$
+is realized as $\state_j$. The confusability of states affects senders
+and receivers alike, but in slightly different ways (see
 Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
@@
 $\Sstrat$ and $\Rstrat$ to their diffusions $D(\Sstrat)$ and
-$D(\Rstrat)$. The idea is that $\Sstrat$ and $\Rstrat$ are what, on
-average, the idealized, noise-free behavior would be, while
-$D(\Sstrat)$ and $D(\Rstrat)$ are behavioral strategies that describe
-the agents' actual noise-perturbed probabilistic behavior. Since the
-effect of confusion of states is that behavior at one choice point
-percolates to behavior at similar choice points, we speak of
-\emph{diffusion of behavior under confusion of states}. If we
+$D(\Rstrat)$. Suppose that $\Sstrat$ and $\Rstrat$ capture a
+population's average signaling behavior if there was no confusion of
+states. Then $D(\Sstrat)$ and $D(\Rstrat)$ give the expected aggregate
+population behavior under the actual noise-perturbed behavior of each
+agent. Since the effect of confusion of states is that behavior at one
+choice point percolates to behavior at similar choice points, we speak
+of \emph{diffusion of behavior under confusion of states}. If we
 conceive of behavioral strategies $\Sstrat$ and $\Rstrat$ as
-row-stochastic matrices, the diffusion effect of state confusability
-is easily captured by matrix multiplication as a diffusion function
-$\Diff_C$ that, based on confusion matrix $C$, maps behavioral
-strategies onto their diffused realizations:
+row-stochastic matrices (rows are choice points, columns are choices;
+each row is a probability distribution), the diffusion effect of
+confusion matrix $C$ is just matrix multiplication:
 \begin{align}
@@
 This is equivalent to the following, perhaps more transparent
-formulation:
-\begin{align*}
-  \RDD(\Sstrat)(\messg \probbar \state_i) & = \sum_{j} C_{ij} \cdot
-  \frac{\Sstrat(\messg \probbar \state_j) \cdot
+formulation:\footnote{To see this, start with $\RDD(\Sstrat)(\messg_k
+  \probbar \state_i)$. If $\RDD(\Sstrat)$ is considered as a matrix,
+  we can rewrite this as $(\RDD(\Sstrat))_{ik}$. By
+  Equation~(\ref{eq:confusion-function}), we get $(C \
+  \RD(\Sstrat))_{ik}$, which by definition of matrix product is
+  $\sum_j C_{ij} \cdot \RD(\Sstrat)_{jk}$. Substituting the definition
+  of $\RD(\Sstrat)$ completes the derivation. Similarly for the
+  receiver.}
+\begin{align*}
+  \RDD(\Sstrat)(\messg_k \probbar \state_i) & = \sum_{j} C_{ij} \cdot
+  \frac{\Sstrat(\messg_k \probbar \state_j) \cdot
     \EU(\messg, \state_j,\Rstrat)}
   {\Phi(\state_j,\Sstrat, \Rstrat)} \\
-    \RDD(\Rstrat)(\state_i \probbar \messg) & = \sum_{j} C_{ji} \cdot
-  \frac{\Rstrat(\state_j \probbar \messg) \cdot
-    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
-\end{align*}
-
-The idea behind the sequential definition of the \rdd is that, at each
-time step, strategies are gradually optimized along the current
-fitness landscape, as described by the \rd, but the realization of
-optimized strategies is noisy, due to similarity of states. It is
-obvious that other (discrete-time) evolutionary dynamics can be
-subjected to state-confusability in an analogous fashion (see also
-Section~\ref{sec:model-interpretation}). In the case of the \rd,
-however, perturbation by state-confusion has a prominent close
-relative in the replicator mutator dynamic.
+    \RDD(\Rstrat)(\state_i \probbar \messg_k) & = \sum_{j} C_{ji} \cdot
+  \frac{\Rstrat(\state_j \probbar \messg_k) \cdot
+    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg_k,\Sstrat, \Rstrat)}  \,.
+\end{align*}
+
+The most natural interpretation of this sequential definition is that
+the \rdd captures the most likely path of development of aggregate
+behavior in a (huge \& homogeneous) society, in which behavioral
+changes occur independently at each choice point, as described by the
+\rd for behavioral strategies. Whatever strategy an agent does
+actually hold, its realization is perturbed by imperfect
+discriminability of states in such a way that the expected overall
+aggregate behavior is the diffusion of the strategies that the agents
+would play in a noise-free world. Any agent-level interpretation of
+the \rd in behavioral strategies seems compatible with this
+interpretation.
+
+The sequential definition is advantageous also because it highlights
+that other (discrete-time) evolutionary dynamics can be subjected to
+state-confusability in an analogous fashion. It would be entirely
+straightforward, for instance, to perturbed a best-response dynamic in
+case we wanted to assume cognizing, rational agents with a noisy
+perceptual system. Our choice of the \rd as a carrier dynamic is
+motivated by its generality and austerity, being compatible with a
+variety of implicit agent-level behavior, without presupposing
+extravagant higher cognitive abilities. Moreover, using the \rd as a
+carrier for diffusion, we actually obtain a special case of a
+behavioral-strategy version of the replicator mutator dynamics. This
+is important to know because it enables similar applications of
+diffusion from perceptual confusion also in models of biological
+evolution.\todo{Do we need the continuous-time stuff? Give
+  discrete-time RMD immediately?}
 
@@
 \citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
-adds stochastic mutation to the replicator dynamic in order to capture, in
-the context of grammar evolution, the differential learning success of
+adds stochastic mutation to the replicator dynamic. The original
+motivation was to capture the differential learning success of
 different dispositions in first language acquisition. The relation of
@@
 fruitful applications in the context of signaling games
-\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}. It
-is therefore desirable to relate the \rdd that we propose here to the
-\rmd.
+\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}.
 
@@
 the usual formulation of the latter is in its continuous-time form,
-and based on mixed strategies, not behavioral strategies. We therefore
-give a discrete-time formulation of the \rmd in behavioral strategies,
-which is independently useful. The relation between \rdd and \rmd will
-then be plain to see. We start with a delineation of mixed and
-behavioral strategies.
+and based on mixed strategies, not behavioral strategies. But a
+discrete-time formulation of the \rmd in behavioral strategies,
+clearly identifies the \rdd to be a special case. From this, it
+follows that a mixed strategy version of the \rdd is just the \rmd
+with suitably defined mutation matrix. To see all this, we start with
+a delineation of mixed and behavioral strategies.
 
@@
 
-Every mixed strategy $\Smixed$ converts to a unique behavioral strategy defined by:
+Every mixed strategy $\Smixed$ converts to a unique behavioral
+strategy defined by:
 \begin{align}
@@
 strategies as treating each choice point as an independent update site
-\citep[e.g.][]{Cressman2003:Evolutionary-Dy,Sandholm2013:Population-Game}. Dynamics
-for mixed strategies conceive of agents as adjusting their whole pure
-strategy globally. In this way, dependencies between different choices
-at different choice points can matter to the evolutionary path. In
+\citep[e.g.][]{Cressman2003:Evolutionary-Dy}. Dynamics for mixed
+strategies conceive of agents as adjusting their whole pure strategy
+globally. In this way, dependencies between different choices at
+different choice points can matter to the evolutionary path. In
 contrast, dynamics for behavioral strategies conceive of agents as
 adjusting their strategies locally at each choice point, regardless of
-what they are disposed to do at other choice points. The latter
-perspective is then more coarse-grained, and not necessarily
-equivalent to the former. But an advantage of dynamics in behavioral
-strategies is that they reduce complexity, which is especially helpful
-for computational simulations of larger games, like the ones we are
-interested in here.
+what they are disposed to do at other choice points.
+
+This can matter to evolutionary dynamics. For illustration, take two
+sender strategies whose behavior is different only at two states. The
+first sends a message with a much higher expected utility in the state
+$A$ than the second, but a message with a slightly lower expected
+utility in state $B$. The expected utility of the first strategy is
+higher. Under evolutionary dynamics that work on entire strategies the
+first strategy will increase in proportion, relative to the
+second. So, a locally less efficient type of behavior can be demoted
+because it is correlated with efficient behavior at another choice
+point. Evolutionary dynamics working on behavioral strategies do not
+allow for that. Behavioral adaptations occur independently at
+different choice points.
+
+As a result, dynamics on behavioral strategies more coarse-grained and
+not necessarily equivalent to dynamics on mixed strategies. An
+advantage of dynamics in behavioral strategies is that they reduce
+complexity, which helps with numerical simulations. They might also be
+a more natural choice for cultural dynamics, e.g., when an agent
+adopts only part of another agent's strategy: not entire lexica are
+adopted, but only lexical entries, so to speak. On the other hand,
+biological evolution seems more naturally modeled in terms of full
+strategies: it is a full set of behavioral disposition that is
+bequeathed to the next generation.
 
@@
 of mixed strategies. Its so-called two-population version, that takes
-sender and receiver roles as separate update sites:
+sender and receiver roles as separate update sites is:
 \begin{align*}
@@
 
-      \node[below of = behavioral] (mutatedBeh) {$C(\Sstrat)$};
+      \node[below of = behavioral] (mutatedBeh) {$D_C(\Sstrat)$};
 
@@
 
-More concretely, since there is a non-injective mapping $G$ from
-behavioral to mixed strategies, we would like to show that there is a
-systematic translation from a confusion matrix $C$ to a mutation
-matrices $Q^C$ such that for each mixed sender strategy $\Smixed$, its
-unique corresponding behavioral strategy $\Sstrat = G(\Smixed)$ has a
+More concretely,given the non-injective mapping $G$ from mixed to
+behavioral strategies, we show that there is a systematic translation
+from a confusion matrix $C$ to a mutation matrices $Q^C$ such that for
+each mixed sender strategy $\Smixed$, its unique corresponding
+behavioral strategy $\Sstrat = G(\Smixed)$ has a
 confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
@@
 all of its rows are identical. That is why, in the limit of iterated
-diffusion, $C^\infty S$ has identical rows (messages are sent with the
-same probability in each state), and $R C^\infty$ likewise has
+diffusion, $C^\infty \Sstrat$ has identical rows (messages are sent with the
+same probability in each state), and $\Rstrat C^\infty$ likewise has
 identical rows (every message is interpreted equally). All of this is
@@
       \begin{cases}
-    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
-    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
+    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
+    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
  \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
@@
   0.3}$. Each trial ran for a maximum of 200 update steps of the
-\rdd. This may not be enough to guarantee convergence to the eventual
-attracting state, but we were interested especially in whether the
-dynamics could lead to reasonable outcomes in reasonable time. A trial
-was considered converged and stopped before the maximum of 200 rounds
-if the total amount of change between strategies before and after the
-current \rdd step was smaller than 0.001, i.e., if:
-\begin{align*}
-  \sum_{\state \in \States} \sum_{\messg \in \Messgs} \left|
-    \rdd(\Sstrat)(\state,\messg) - \Sstrat(\state,\messg) \right| <
-  0.001\,, & \text{\ \ and} \\
-  \sum_{\messg \in \Messgs} \sum_{\state \in \States} \left|
-    \rdd(\Rstrat)(\messg,\state) - \Rstrat(\messg,\state) \right| <
-  0.001 & \,.
-\end{align*}
+\rdd. A trial was considered converged and stopped before the maximum
+of 200 rounds if the total amount of change between strategies before
+and after the current \rdd step was smaller than 0.001. It is not
+guaranteed that strategies at halting time had converged to the
+eventual attracting state, whether they ran for 200 rounds or not. Our
+notion of convergence therefore is only a categorical measure for the
+speed of approaching some state in which force of evolutionary
+optimization has fallen under a (well-considered, but eventually
+arbitrary) threshold. In other words, we need to think of convergence
+in our sense as speed of evolution towards the first evolutionary
+bottleneck.\todo{check and improve this}
 
@@
 were. \emph{Entropy} captures the amount of systematicity or
-regularity in signal use. \emph{Voronoiness} and \emph{convexity} are,
-respectively, gradient and categorical measures for the coherence of
-signal-induced categories. \emph{Expected utility} measures the
-communicative efficiency of evolved strategy pairs.
+regularity in signal use. \emph{Convexity} captures whether a
+behavioral strategy would project onto a convex pure
+strategy. \emph{Expected utility} measures the communicative
+efficiency of evolved strategy pairs.\todo{take average of entropy}
 

### 2015-01-15 paper/version_02/paper.tex
reworked sections 1 & 2

@@
 Others have argued that vagueness is a natural byproduct of
-limitations in information processing
+limitations in information processing or particular learning
+strategies
 \citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
@@
 
-Similarity-maximizing games, or \emph{sim-max games} for short, are
-generalizations of Lewis' games where different states are allowed to
-be more or less similar to one another. While Lewis' games treated
-communicative success as a matter of black and white, sim-max games
-allow for a gradient notion of communicative success: the more similar
-the receiver's interpretation is to the actual state, the more
-successful the play is taken to be. Signaling games with
-utility-relevant similarities in the state space are fairly standard
-in economics
+Similarity-maximizing (short: sim-max) games are generalizations of
+Lewis' games where different states are allowed to be more or less
+similar to one another. While Lewis' games treated communicative
+success as a matter of black and white, sim-max games allow for a
+gradient notion: the more similar the receiver's interpretation is to
+the actual state, the better. Signaling games with utility-relevant
+similarities in the state space are fairly standard in economics
 \citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
@@
 while the receiver's interpretations are the central spots in each
-category. For example, if \States is the unit interval and all states
-are equiprobable, a Voronoi language with two messages would have the
-sender use one message exclusively for all points in the lower half of
-the unit interval and another for all points in the upper half; the
-receiver's interpretations of messages are the central points, .25 and
-.75, in the respective intervals.
+category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put,
+all points in $X$ are connected via a straight line that lies entirely
+in $X$; $X$ has no gaps or dents. For example, if \States is the unit
+interval and all states are equiprobable, a Voronoi language with two
+messages would have the sender use one message exclusively for all
+points in the lower half of the unit interval and another for all
+points in the upper half; the receiver's interpretations of messages
+are the central points, .25 and .75, in the respective intervals.
 
@@
 particular: (i) clear positive examples of a vague category should
-show a gradient transition to clear negative examples (so-called
-\emph{higher-order vagueness});\todo{check usage of term} (ii)
-prototypes should likewise be gradient regions, peaking at the center
-of the vague category they represent
+show a gradient transition to clear negative examples (towards
+so-called \emph{higher-order vagueness});\todo{check usage of term}
+(ii) prototypes should likewise be gradient regions, peaking at the
+center of the vague category they represent
 \citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re}.\todo{check
@@
 cases. From the assumption that prototypes are extended, yet convex
-regions in conceptual space, a construction algorithm is available that
-yields ``collated Voronoi diagrams'' with thick boundaries
+regions in conceptual space, a construction algorithm is available
+that yields ``collated Voronoi diagrams'' with thick boundaries
 representing borderline
 regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
-is possible to arrive at higher-order vagueness, by weighing in the
-distance of different borderline cases to various prototypical
-regions. This accounts for the first of the two desiderata mentioned
-above, but still assumes that crisp prototype regions must be
-given.
+is possible to arrive at a gradient transition between categories, by
+weighing in the distance of different borderline cases to various
+prototypical regions. This accounts for the first of the two
+desiderata mentioned above, but still assumes that crisp prototype
+regions must be given.
 
@@
 are argue here that \citeauthor{FrankeJager2010:Vagueness-Signa}'s
-approaches builds on a dubious assumption about the source of
+approach builds on a dubious assumption about the source of
 vagueness. While \citeauthor{OConnor2013:The-Evolution-o} gives an
@@
 choice point. This greatly reduces the complexity of the dynamic and
-simplifies numerical simulations. But it points towards an
+simplifies numerical simulations. But it also points towards an
 interpretation as a process of cultural evolution, rather than
@@
 
-We begin by recapitulating the formulation of the \rd in
-Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
+We begin by recapitulating the \rd in
+Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in
+Section~\ref{sec:repl-diff-dynam-1}.
 Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
@@
 a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
-\ref{sec:diffusion-as-special} are conceptually important, but not
-strictly necessary to appreciate the main point of the paper.
+\ref{sec:diffusion-as-special} are conceptually important, but rather
+technical and not strictly necessary to appreciate the main point of
+the paper.
 
@@
 $\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are functions
-$\Rstrat \in \Delta(\Acts)^\Messgs$. Define the expected utility of
-choices in each choice point as usual:
+$\Rstrat \in \Delta(\Acts)^\Messgs$. The expected utility of choices
+at each choice point is:
 \begin{align*}
@@
 of agents when behavioral adaptations occur independently at each
-choice point, e.g., when agents adopt the observed choice behavior
-\emph{at a given choice point} of another agent whose behavior
-\emph{at that choice point} has higher fitness (see
+choice point, e.g., when agents adopt the behavior of another agent
+\emph{at a given choice point} if that ``local'' behavior is
+``locally'' better than their own (see
 \citeauthor{Sandholm2013:Population-Game}
@@
 can be captured in a function that maps behavioral strategies
-$\Sstrat$ and $\Rstrat$ to their diffusions $D(\Sstrat)$ and
+$\Sstrat$ and $\Rstrat$ to their \emph{diffusions} $D(\Sstrat)$ and
 $D(\Rstrat)$. Suppose that $\Sstrat$ and $\Rstrat$ capture a
@@
 The discrete-time replicator diffusion dynamic takes the replicator
-dynamic as basic, but factors in the confusion of states at each
-update step in a sequential update:
+dynamic as basic, but factors in the confusion of states at each step
+in a sequential manner:
 \begin{align*}
@@
   \probbar \state_i)$. If $\RDD(\Sstrat)$ is considered as a matrix,
-  we can rewrite this as $(\RDD(\Sstrat))_{ik}$. By
+  we can rewrite this as $(RDD(\Sstrat)_{ik}$. By
   Equation~(\ref{eq:confusion-function}), we get $(C \
-  \RD(\Sstrat))_{ik}$, which by definition of matrix product is
+  \RD(\Sstrat))_{ik}$, which by definition of matrix multiplication is
   $\sum_j C_{ij} \cdot \RD(\Sstrat)_{jk}$. Substituting the definition
@@
   \frac{\Rstrat(\state_j \probbar \messg_k) \cdot
-    \EU(\state_j, \messg,\Sstrat)} {\Phi(\messg_k,\Sstrat, \Rstrat)}  \,.
+    \EU(\state_j, \messg_k,\Sstrat)} {\Phi(\messg_k,\Sstrat, \Rstrat)}  \,.
 \end{align*}
@@
 state-confusability in an analogous fashion. It would be entirely
-straightforward, for instance, to perturbed a best-response dynamic in
-case we wanted to assume cognizing, rational agents with a noisy
+straightforward, for instance, to perturb a best-response dynamic in
+case we wanted to model cognizing, rational agents with a noisy
 perceptual system. Our choice of the \rd as a carrier dynamic is
@@
 variety of implicit agent-level behavior, without presupposing
-extravagant higher cognitive abilities. Moreover, using the \rd as a
-carrier for diffusion, we actually obtain a special case of a
+extravagant higher cognitive abilities
+\citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, using the \rd
+as a carrier for diffusion, we actually obtain a special case of a
 behavioral-strategy version of the replicator mutator dynamics. This
@@
 diffusion from perceptual confusion also in models of biological
-evolution.\todo{Do we need the continuous-time stuff? Give
-  discrete-time RMD immediately?}
+evolution. The remainder of this section therefore works out the
+relation between the \rdd and the \rmd.
 
@@
 
-The \rmd has been proposed first in the context of signaling game
-models for the evolution of grammar
+The \rmd has first been proposed first as a tool for studying the
+evolution of grammar
 \citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
-adds stochastic mutation to the replicator dynamic. The original
-motivation was to capture the differential learning success of
-different dispositions in first language acquisition. The relation of
+adds stochastic mutation to the replicator dynamic. The relation of
 the \rmd to other prominent evolutionary dynamics is well understood
@@
 
-A direct comparison between \rdd and the \rmd is not possible, because
-the usual formulation of the latter is in its continuous-time form,
-and based on mixed strategies, not behavioral strategies. But a
-discrete-time formulation of the \rmd in behavioral strategies,
-clearly identifies the \rdd to be a special case. From this, it
-follows that a mixed strategy version of the \rdd is just the \rmd
-with suitably defined mutation matrix. To see all this, we start with
-a delineation of mixed and behavioral strategies.
+A direct comparison between the \rdd and the \rmd is not possible,
+because the usual formulation of the latter is based on mixed, not
+behavioral strategies. But a discrete-time formulation of the \rmd in
+behavioral strategies, clearly identifies the \rdd to be a special
+case. From this, it follows that a mixed strategy version of the \rdd
+is just the \rmd with suitably defined mutation matrix. From this,
+even a continuous-time version can easily be retrieved if desired. To
+show this, we start with a delineation of mixed and behavioral
+strategies.
 
@@
 
-In the context of evolutionary dynamics, we can think of behavioral
-strategies as treating each choice point as an independent update site
+In the context of evolutionary dynamics for signaling games (where
+each agent only chooses once), we can think of behavioral strategies
+as treating each choice point as an independent update site
 \citep[e.g.][]{Cressman2003:Evolutionary-Dy}. Dynamics for mixed
-strategies conceive of agents as adjusting their whole pure strategy
-globally. In this way, dependencies between different choices at
+strategies conceive of agents as adjusting their entire pure
+strategy. In this way, dependencies between different choices at
 different choice points can matter to the evolutionary path. In
@@
 This can matter to evolutionary dynamics. For illustration, take two
-sender strategies whose behavior is different only at two states. The
-first sends a message with a much higher expected utility in the state
-$A$ than the second, but a message with a slightly lower expected
-utility in state $B$. The expected utility of the first strategy is
+sender strategies whose behavior is different only at two
+states. Compared to the second strategy, the first scores a much
+higher expected utility at one state, but a slightly lower at the
+other, so that the total expected utility of the first strategy is
 higher. Under evolutionary dynamics that work on entire strategies the
 first strategy will increase in proportion, relative to the
-second. So, a locally less efficient type of behavior can be demoted
-because it is correlated with efficient behavior at another choice
-point. Evolutionary dynamics working on behavioral strategies do not
-allow for that. Behavioral adaptations occur independently at
+second. Consequently, a locally less efficient type of behavior can be
+demoted because it is correlated with efficient behavior at another
+choice point. Evolutionary dynamics working on behavioral strategies
+do not allow for that. Behavioral adaptations occur independently at
 different choice points.
@@
 a more natural choice for cultural dynamics, e.g., when an agent
-adopts only part of another agent's strategy: not entire lexica are
-adopted, but only lexical entries, so to speak. On the other hand,
-biological evolution seems more naturally modeled in terms of full
-strategies: it is a full set of behavioral disposition that is
-bequeathed to the next generation.
+observes and imitates another agent's strategy: it may be more
+plausible that agents adopt the use of a single word, not an entire
+lexicon, so to speak. On the other hand, biological evolution seems
+more naturally modeled in terms of full strategies: it is a full set
+of behavioral disposition that is bequeathed to the next generation.
 
@@
 
-The continuous-time replicator dynamic defines the change of frequency
-of mixed strategies. Its so-called two-population version, that takes
-sender and receiver roles as separate update sites is:
-\begin{align*}
-  \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
-  \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
-  \Phi(\Rmixed,\Smixed) \right ) \,.
-\end{align*}
+% The continuous-time replicator dynamic defines the change of frequency
+% of mixed strategies. Its so-called two-population version, that takes
+% sender and receiver roles as separate update sites is:
+% \begin{align*}
+%   \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
+%   \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
+%   \Phi(\Rmixed,\Smixed) \right ) \,.
+% \end{align*}
 Similar to the version for behavioral strategies, a discrete-time
@@
 
-The \rmd is usually given only in its continuous-time form. A
-two-population (non-payoff adjusted) formulation is then:
-\begin{align*}
-  \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
-    \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
-    \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
-    \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
-\end{align*}
-A discrete-time, two-population formulation of the \rmd can be defined as follows 
-\citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
-
+The \rmd is usually given only in its continuous-time form. % A
+% two-population (non-payoff adjusted) formulation is then:
+% \begin{align*}
+%   \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
+%     \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
+%     \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
+%     \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
+% \end{align*}
+But a discrete-time, two-population formulation of the \rmd is
+available too \citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
 \begin{align*}
@@
 \end{align*}
-
-% \begin{proof}
-%   It suffices to check either sender or receiver part. Focusing on the
-%   former, we need to show that the continuous version is derivable
-%   from the discrete version if the discrete update steps get
-%   infinitely small so that:
-%   \begin{align*}
-%     \dot{\Smixed_i} & = \RMD(\Smixed)_i - \Smixed_i = \sum_{j} Q_{ji}
-%     \frac{\Smixed_j \cdot F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} -
-%     \Smixed_i = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-%         F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} - \Smixed_i \right ) \\
-%     & = \sum_{j} \Smixed_j \left ( \frac{ Q_{ji} \cdot
-%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)}{ \Phi(\Smixed,\Rmixed)}  \right )\,.
-%   \end{align*}
-%   As $\Phi(\Smixed,\Rmixed)$ is a constant rate of change, we can drop
-%   it to derive the non-payoff adjusted continuous version above,
-%   since:
-%   \begin{align*}
-%     & \sum_{j} \Smixed_j \left ( Q_{ji} \cdot
-%         F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed)  \right
-%       ) =     \sum_{j} \Smixed_j  Q_{ji} \cdot
-%         F_j^{\Rmixed} - \sum_{j} \Smixed_j \cdot \Smixed_i \cdot
-%         \Phi(\Smixed,\Rmixed) \\
-%        = &    \sum_{j} \Smixed_j  Q_{ji} \cdot
-%         F_j^{\Rmixed} - \Smixed_i \cdot
-%         \Phi(\Smixed,\Rmixed) 
-%   \end{align*}
-% \end{proof}
-
 The discrete-time \rmd has the same sequential nature as the \rdd:
@@
 definable for behavioral strategies (see
-Equations~(\ref{eq:confusion-function})), while mutation in the \rmd
-is defined as mutation from one pure strategy to another. Nonetheless,
-it is possible to map each confusion matrix to a mutation matrix in a
+Equation~(\ref{eq:confusion-function})), while mutation in the \rmd is
+defined as mutation from one pure strategy to another. Nonetheless, it
+is possible to map each confusion matrix to a mutation matrix in a
 natural way so that the correspondence between mixed and behavioral
@@
 
-More concretely,given the non-injective mapping $G$ from mixed to
+More concretely, given the non-injective mapping $G$ from mixed to
 behavioral strategies, we show that there is a systematic translation
-from a confusion matrix $C$ to a mutation matrices $Q^C$ such that for
+from a confusion matrix $C$ to a mutation matriX $Q^C$ such that for
 each mixed sender strategy $\Smixed$, its unique corresponding
@@
   \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
-  \state)) 
-\end{align*} 
-\begin{align*}
+  \state)) \\
   E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}

### 2015-01-15 paper/version_02/paper.tex
rewrite Section 3

@@
 two-step procedure: we first calculate where the replicator dynamic
-would take us, then apply diffusion. To understand the \rdd
-better, in Section~\ref{sec:iterated-diffusion} we look closer at the novel
-diffusion part, before exploring in Section~\ref{sec:simulations} how
-diffusion and fitness-based selection interact.
+would take us, then apply diffusion. To understand the \rdd better, we
+look closer at the novel diffusion part in
+Section~\ref{sec:iterated-diffusion}, before exploring how diffusion
+and fitness-based selection interact in Section~\ref{sec:simulations}.
 
@@
 states; when $\impairment \rightarrow \infty$ agents cannot
-discriminate states at all. The top part of
-Figure~\ref{fig:NosofskySim} gives an impression of
-Nosofsky-similarity for different parameter values. Other
-formalizations of perceptual similarity are possible, including ones
-that allow for different discriminability in different areas of the
-state space, but we stick with Nosofsky's similarity function for the
-time being, because it is mathematically simple, yet an established
-notion in mathematical psychology.
+discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
+impression of Nosofsky-similarity for different parameter
+values. Other formalizations of perceptual similarity are possible,
+including ones that allow for different discriminability in different
+areas of the state space, but we stick with Nosofsky's similarity
+function for the time being, because it is mathematically simple, yet
+an established notion in mathematical psychology.
 
@@
 Confusion of states is then also a function of imprecision parameter
-$\impairment$ (see Figure~\ref{fig:NosofskySim}, bottom part). For $\impairment
-= 0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$ is
-positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
-\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$ for all
-$i$ and $j$.
+$\impairment$ (see Figure~\ref{fig:NosofskySim}). For $\impairment =
+0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$
+is positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
+\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$
+for all $i$ and $j$.
 
 If states are confused on average with a probability proportional to
-their similarity, repeated applications of diffusion of behavior as
-defined in Equation~(\ref{eq:confusion-function}) causes behavioral
-dispositions gradually to diffuse along a gradient of similarity of
-states as well. Consequently, iterated diffusion leads to a smoothing
-out and an eventual equalization of behavioral strategies, for both
-the sender and the receiver. After $n$ steps of iterated diffusion an
-initial sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an
-initial receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we
-assume that $\impairment > 0$, $C$ is a positive row-stochastic matrix
-(each state could in principle be confused as any other state, albeit
-with possibly a very low probability). It then follows from the
+their similarity, repeated application of the operation in
+Equation~(\ref{eq:confusion-function}) causes behavioral dispositions
+gradually to diffuse along a gradient of similarity of
+states. Consequently, iterated diffusion leads to a smoothing out and
+an eventual equalization of behavioral strategies, for both the sender
+and the receiver. After $n$ steps of iterated diffusion an initial
+sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an initial
+receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we assume that
+$\impairment > 0$, $C$ is a positive row-stochastic matrix (each state
+could in principle be confused as any other state, albeit with
+possibly a very low probability). It then follows from the
 Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
@@
 all of its rows are identical. That is why, in the limit of iterated
-diffusion, $C^\infty \Sstrat$ has identical rows (messages are sent with the
-same probability in each state), and $\Rstrat C^\infty$ likewise has
-identical rows (every message is interpreted equally). All of this is
-in line with the intuition that diffusion of behavior under confusion
-of states iteratively equalizes behavior for similar states; if all
-states can be confused for one another in principle, behavior smoothes
-out entirely in the limit.
+diffusion, $C^\infty \Sstrat$ has identical rows (messages are sent
+with the same probability in each state), and $\Rstrat C^\infty$
+likewise has identical rows (every message is interpreted
+equally). All of this is in line with the intuition that diffusion of
+behavior under confusion of states iteratively equalizes behavior for
+similar states; if all states can be confused for one another in
+principle, behavior smoothes out entirely in the limit.
 
@@
 section. We will show presently that for reasonable in-between levels
-of imprecision $\impairment$, the \rdd leads to communicatively
-successful, yet vague signal meaning. Suitable levels of imprecision
-can have further accelerating and unifying effects on meaning
-evolution.
+of imprecision $\impairment$, the \rdd can lead to reasonably
+efficient, yet vague signaling behavior. Suitable levels of
+imprecision can have further accelerating and, surprisingly, unifying
+effects on meaning evolution.
 
@@
 of 200 rounds if the total amount of change between strategies before
-and after the current \rdd step was smaller than 0.001. It is not
-guaranteed that strategies at halting time had converged to the
-eventual attracting state, whether they ran for 200 rounds or not. Our
-notion of convergence therefore is only a categorical measure for the
-speed of approaching some state in which force of evolutionary
-optimization has fallen under a (well-considered, but eventually
-arbitrary) threshold. In other words, we need to think of convergence
-in our sense as speed of evolution towards the first evolutionary
-bottleneck.\todo{check and improve this}
-
-Representative examples for resulting strategy pairs at stopping time
-are given in
+and after the current \rdd step was smaller than a suitably chosen
+threshold. It is not guaranteed that strategies at halting time had
+converged to the eventual attracting state, whether they ran for 200
+rounds or not. Our notion of convergence therefore is only a
+categorical measure for the speed of approaching some state in which
+the force of evolutionary optimization has fallen under a
+(well-considered, but eventually arbitrary) threshold. In other words,
+we need to think of convergence in our sense as speed of evolution
+towards the first evolutionary bottleneck.\todo{check and improve
+  this}
+
+Representative examples for resulting strategy pairs are given in
 Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
@@
 clearly that the pictured strategy pair has not yet reached a
-dynamically stable state under the \rd. Indeed, the trial did not
-converge in our technical sense, but was stopped after the maximum of
-200 rounds. In contrast, the outcome of a trial with identical
-parameters, except with imprecision $\impairment = 0.1$, which is
-shown in Figure~\ref{fig:example_stratsB}, had converged (in our
-technical sense) after 99 rounds of iterated \rdd. The sender strategy
-shows a smooth blending from one ``category'' to the other, and the
-receiver's interpretations are rather extended curves, peaking at a
-central point in the relevant ``categories.''
+dynamically stable state under the \rd. Indeed, the trial was stopped
+after the maximum of 200 rounds. In contrast, the outcome of a trial
+with identical parameters, except with imprecision $\impairment =
+0.1$, which is shown in Figure~\ref{fig:example_stratsB}, had
+converged (in our technical sense) after 99 rounds of iterated
+\rdd. The sender strategy shows a smooth blending from one
+``category'' to the other, and the receiver's interpretations are
+rather extended curves, peaking at a central point in the relevant
+``categories.''
 
@@
 
-To further explore our simulation results, we defined a small number
-of metrics that aim to capture numerically how vague, communicatively
-efficient and generally well-structured the resulting strategy pairs
-were. \emph{Entropy} captures the amount of systematicity or
-regularity in signal use. \emph{Convexity} captures whether a
-behavioral strategy would project onto a convex pure
+To further explore our simulation results, we track how vague,
+communicatively efficient and generally well-structured the recorded
+strategy pairs were. \emph{Entropy} captures the amount of
+systematicity or regularity in signal use. \emph{Convexity} captures
+whether a behavioral strategy would project onto a convex pure
 strategy. \emph{Expected utility} measures the communicative
@@
 respectively. The receiver strategies have respective entropies $0.43$
-and $0.84$. In general, we expect that vague languages will have
-higher entropy than crisp ones and that increasing imprecision will
+and $0.84$. In general, we expected that vague languages would have
+higher entropy than crisp ones and that increasing imprecision would
 increase entropy, all else equal.
-
-\paragraph{Voronoiness.}
-This metric aims to quantify how close a strategy is to being a part
-of a Voronoi tessallation of the state space. Bearing in mind that
-this metric is specifically geared towards the sim-max games in our
-set-up, we define Voronoiness as follows. Let $\pi(\Rstrat, \messg) =
-\argmax_{\state \in \States}\Rstrat(\messg,\state)$ be the prototype
-of a message $\messg \in \Messgs$, then
-\begin{align*}
-  V(\Sstrat, \Rstrat) & = \sum_{\state \in \States} \sum_{\messg_1 \in
-    \Messgs} \sum_{\messg_2 \in \Messgs} F(\Sstrat, \Rstrat, \state,
-  \messg_1, \messg_2) \text{\ \ \ with } \\
-  F(\Sstrat, \Rstrat, \state,
-  \messg_1, \messg_2) & =
-  \begin{cases}
-    1 & \text{if } \similarity(\state, \pi(\Rstrat, \messg_1)) >
-    \similarity(\state, \pi(\Rstrat, \messg_2)) \bicond \\
-    &  \ \ \ \ \ \Sstrat(\state, \messg_1) >
-    \Sstrat(\state, \messg_2) \\
-    0 & \text{otherwise}
-  \end{cases}
-\end{align*}
-is the Voronoiness of the sender strategy $\Sstrat$ given the receiver
-strategy $\Rstrat$ and
-\begin{align*}
-  V(\Rstrat) & = \sum_{\messg \in
-    \Messgs} \sum_{\state_1 \in \States}  \sum_{\state_2 \in \States}
-  F(\Rstrat, \messg, \state_1, \state_2) \text{\ \ \ with } \\
-  F(\Rstrat, \messg, \state_1, \state_2) & =
-  \begin{cases}
-    1 & \text{if } \similarity(\state_1, \pi(\Rstrat,
-  \messg)) > \similarity(\state_2, \pi(\Rstrat, \messg)) \\
-  & \ \ \ \ \ \bicond \Rstrat(\messg,
-  \state_1) > \Rstrat(\messg, \state_2) \\
-    0 & \text{otherwise}
-  \end{cases}
-\end{align*}
-is the Voronoiness of the receiver strategy $\Rstrat$.  The metrics
-are lower bound by $0$ and upper bound by, respectively, $|\States|
-\times |\Messgs|^2$ and $|\Messgs| \times |\States|^2$, thus we will
-normalize them by dividing by these values.
-
-Despite being vague, a language can have maximal Voronoiness according
-to this gradient measure. For instance, the sender strategies in
-Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have
-Voronoiness $0.91$ and $1$, respectively. The receiver strategies have
-Voronoiness $0.78$ and $1$, respectively. We do not \emph{prima facie}
-expect a conceptual connection between vagueness and Voronoiness, and
-so do not predict a clear effect of imprecision on Voronoiness either.
 
@@
 faster, it also makes sense to define a categorical measure of
-convexity, that compensates for potential vagueness. To determine
+convexity that compensates for potential vagueness. To determine
 whether a sender strategy $\Sstrat$ is convex despite possibly being
@@
 Figure~\ref{fig:example_stratsA} is not convex, while the one in
-Figure~\ref{fig:example_stratsB} is. Again, we do not expect a
-conceptual relation between vagueness and imprecision, on the one
+Figure~\ref{fig:example_stratsB} is.  One would perhaps not expect a
+systematic relation between vagueness and imprecision, on the one
 hand, and convexity on the other.
 
-\paragraph{Expected utility.} We also record the expected utility of
+\paragraph{Expected utility.} We also recorded the expected utility of
 a strategy pair:
@@
 
-  \caption{Means of gradient and proportions of categorical metrics for
-    each triple of independent variables, with estimated 95\%
-    confidence intervals.}
+  \caption{Means of gradient and proportions of categorical measures
+    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\impairment
+    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the mean of the
+    entropies for the sender and receiver strategy.}
   \label{fig:MeanMetrics}
@@
 
-Figure~\ref{fig:MeanMetrics} shows the means of the recorded metrics,
-together with their estimated 95\% confidence intervals, for each
-triple of independent variables. Most trends are quite clear, and not
-unexpected. Entropy of sender and receiver strategies are increasing
-with imprecision. Entropy of sender strategies seems independent of
-tolerance levels and the number of states, while there seems to be an
-effect of both parameters on the entropy of the receiver
-strategy. What was not expected \emph{a priori} was that Voronoiness
-of both sender and receiver strategies appear to be non-decreasing
-under increasing impairment, with the only exception being games with
-6 states and the transition from $\impairment = 0$ to $\impairment =
-0.05$. Expected utility of evolved languages mostly decreases in
-increasing imprecision, with the sole exception of a rise for $\toler
-= 0.2$ and $n_s = 6$ at $\impairment = 0.1$. Still, normalized
-expected utility of evolved strategy pairs under mild values of
-imprecision are not necessarily much worse than their respective
-perfect precision \rd-analogues. This is especially so for lower
-numbers of states (see discussion below).
-
-Taken together, these findings suggest the following effect of
-confusability of states and diffusion of behavior. Mild forms of
-perceptual imprecision lead to slightly less communicative efficiency,
-more vagueness and more regular, well-behaved languages. This
-impression is also supported by the proportion of converged and convex
-trials, pictured in
-Figure~\ref{fig:MeanMetrics}. Non-converged and non-convex
-sender strategies showed for several values of parameters in the
-absence of impairment. Increasing impairment always assured
-convergence and convexity: diffusion speeds up convergence to a convex
-and regular language.
-
-% \begin{figure}
-%   \centering
-
-%   \includegraphics[width=0.8\textwidth]{plots/CategoricalMeasures.pdf}
-
-%   \caption{Proportion of converged and convex trials.}
-%   \label{fig:CategoricalMeasures}
-% \end{figure}
-
-Diffusion also has another interesting regularizing effect on the
-evolution of signaling. If we look at the confidence intervals in
-Figure~\ref{fig:MeanMetrics}, we see that there is very little
-variation in the recorded metrics for evolved strategies, at least for
-higher values of impairment. On closer inspection, it turns out that
-variability in low-impairment conditions is not only due to
+Figure~\ref{fig:MeanMetrics} shows plots summarizing the results of
+our simulation. For perspicuity, we only plot results for one level of
+tolerance $\beta = 0.1$, and leave out the case of $n_s = 90$. Still,
+every qualitative trend we mention in the following applies to the
+whole set of results.
+
+As expected, increasing imprecision leads to higher entropy and lower
+expected utility. Importantly, however, imprecision does not
+necessarily lead to disastrous decline of communicative success. On
+the other hand, imprecision increased the number of outcomes with
+convex sender strategies, and led to higher proportions of
+convergence. In fact, sufficient impairment always assured convergence
+and convexity. It appears that perceptual imprecision leads to
+slightly less communicative efficiency, more vagueness but more
+regular, well-behaved languages in shorter time.
+
+Beyond promoting convexity, diffusion also has another interesting
+regularizing effect on the evolution of signaling. There was very
+little variation in the recorded metrics for evolved strategies, at
+least for higher values of impairment. On closer inspection, it turns
+out that variability in low-impairment conditions is not only due to
 non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
@@
 
-Interestingly, this type of variability in evolutionary outcomes is
-also weeded out by mild increases in impairment. To see this, we
-calculated the average distance between evolved sender strategies
-within each group of trials that had identical independent
-parameter values. We determined the distance between sender strategies
-$\Sstrat$ and $\Sstrat'$ as the average Hellinger distance between
-probability distributions $\Sstrat(\state)$ and $\Sstrat'(\state)$
-over messages at each choice point $\state$:
+Interestingly, this type of variability in evolutionary outcomes can
+be weeded out by impairment. To see this, we calculated the average
+distance between evolved sender strategies within each group of trials
+that had identical parameter values. We determined the distance
+between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
+distance between distributions $\Sstrat(\state)$ and
+$\Sstrat'(\state)$ at each choice point $\state$:
 \begin{align*}
@@
 \emph{within group distances}, i.e., the average distances between all
-sender strategies in each parameter group, are plotted in
-Figure~\ref{fig:AverageEUinGroups}. This clearly shows that with
-increasing imprecision, the sender strategies evolving under \rdd were
-exactly alike (modulo swapping of messages). In other words,
-perceptual imprecision unifies evolutionary outcomes, and in fact
-guarantees sender strategies that are not only convex, but also
-maximally efficient in that they induce a vague category split exactly
-in the middle of the unit interval.
+sender strategies in each parameter group, are plotted exemplarily in
+Figure~\ref{fig:AverageEUinGroups} for $\beta = 0.1$ and $n_s =
+10$. Despite minor quantitative differences the is the same for all
+other parameter settings that we tested: with increasing imprecision,
+the sender strategies evolving under \rdd were exactly alike (modulo
+swapping of messages). This means that perceptual imprecision unifies
+evolutionary outcomes and guarantees sender strategies that are not
+only convex, but also regular in that they induce a vague category
+split exactly in the middle of the unit interval.
 
@@
 an evolutionary beneficial side-effect. Higher imprecision can lead to
-higher \emph{within group expected utility}, i.e., the average
+higher \emph{within group expected utility}, defined as the average
 expected utility that each evolved language (i.e., pair of sender and
 receiver strategy at stopping time) scored when playing against an
-arbitrary other language of the same parameter group (see
-Figure~\ref{fig:AverageEUinGroups}). What this means is that, while
-imprecision might decrease the communicative efficiency of individual
-languages, it increases the conceptual coherence and communicative
-success between independently evolving strategies. It is almost as if
-mere confusability of states imposes a regularity constraint on
-evolving categories.
+arbitrary other language of the same parameter
+group. Figure~\ref{fig:AverageEUinGroups}) gives the representative
+example of $\toler = 0.1$ and $n_s = 10$. The observation repeats for
+other parameter values: while imprecision might decrease the
+communicative efficiency of individual languages, it increases the
+conceptual coherence and communicative success between independently
+evolving strategies. It is almost as if mere confusability of states
+imposes a regularity constraint on evolving categories.
 

### 2015-01-15 paper/version_02/paper.tex
revised section 3 with smaller plots

@@
 
-\begin{figure}
+\begin{figure}[t]
   \centering
 
-  \includegraphics[width=\textwidth]{plots/MeanMetrics2.pdf}
+  \includegraphics[width=0.75\textwidth]{plots/MeanMetrics3.pdf}
 
@@
 \emph{within group distances}, i.e., the average distances between all
-sender strategies in each parameter group, are plotted exemplarily in
-Figure~\ref{fig:AverageEUinGroups} for $\beta = 0.1$ and $n_s =
-10$. Despite minor quantitative differences the is the same for all
+sender strategies in each parameter group, are plotted exemplarily on
+the left in Figure~\ref{fig:GroupMeasures} for $\beta = 0.1$ and $n_s
+= 10$. Despite minor quantitative differences the is the same for all
 other parameter settings that we tested: with increasing imprecision,
@@
 
-    \includegraphics[width=0.8\textwidth]{plots/WithinGroupMeasures.pdf}
-
-    \caption{For each group, i.e., triplet of independent variables,
-      the plots show (i) within group distances and (ii) within group
-      expected utilities, as defined in the main text.}
-  \label{fig:AverageEUinGroups}
+    \includegraphics[width=0.475\textwidth]{plots/WithinGroupDistanceConcise.pdf}
+    \hfill
+    \includegraphics[width=0.475\textwidth]{plots/WithinGroupEUConcise.pdf}
+
+    \caption{Within group distance (left) and within group expected
+      utility (right) as a function of impairment for all runs with
+      $\toler = 0.1$ and $n_s = 10$.}
+  \label{fig:GroupMeasures}
 \end{figure}
@@
 receiver strategy at stopping time) scored when playing against an
-arbitrary other language of the same parameter
-group. Figure~\ref{fig:AverageEUinGroups}) gives the representative
-example of $\toler = 0.1$ and $n_s = 10$. The observation repeats for
-other parameter values: while imprecision might decrease the
-communicative efficiency of individual languages, it increases the
-conceptual coherence and communicative success between independently
-evolving strategies. It is almost as if mere confusability of states
-imposes a regularity constraint on evolving categories.
+arbitrary other language of the same parameter group. The right side
+of Figure~\ref{fig:GroupMeasures}) gives a representative example. The
+observation repeats for other parameter values: while imprecision
+might decrease the communicative efficiency of individual languages,
+it increases the conceptual coherence and communicative success
+between independently evolving strategies. It is almost as if mere
+confusability of states imposes a regularity constraint on evolving
+categories.
 

### 2015-01-15 snippets/data_exploration/data_exploration_tmp.Rmd
revised section 3 with smaller plots

@@
 
+More concise plot for second revision of paper
+
+```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+
+data$converged = data$Iterations < 201
+data$convex = data$Speaker.Convex.Cat
+data$entropy = (data$Speaker.entropy + data$Hearer.entropy) / 2
+
+ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
+             "entropy",
+             "Expected.utility",
+             "convex", "converged"
+             )]
+
+dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
+
+dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
+dsub = subset(dms, dms$Tolerance == 0.1 & dms$Number.of.states < 90)
+dsub$Number.of.states = factor(dsub$Number.of.states)
+
+myLabeller <- function(var, value){
+    value <- as.character(value)
+    if (var=="variable") { 
+        value[value=="Expected.utility"] <- "exp. utility"
+    }
+    return(value)
+}
+
+pd <- position_dodge(.01)
+mysize = 2.5
+sp = ggplot(dsub, aes(x=Impairment, y=value, color = Number.of.states, shape=Number.of.states, linetype=Number.of.states)) + 
+   geom_line(color="gray", position = pd) + xlab("imprecision") +
+   geom_point(position = pd, size = mysize) +
+    scale_shape_discrete(name = "# states", guide = guide_legend()) +
+    scale_linetype_discrete(name = "# states", guide = guide_legend()) +
+    scale_color_discrete(name = "# states", guide = guide_legend()) +
+    theme(legend.position="top")
+
+sp = sp + facet_grid(variable ~ ., scales="free", labeller=myLabeller)
+show(sp)
+
+ggsave(sp, file = "paper/version_02/plots/MeanMetrics3.pdf", width=5, height = 5.5)
+
+```
+
 Entropy

### 2015-01-15 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -211,7 +211,22 @@ combPlot = ggplot(Dmelt, aes(x=Impairment, y=value, group = variable, shape = va
   scale_linetype_discrete(name = "property", 
                           labels = c("distance", "utility"))
 show(combPlot)
-ggsave(combPlot, file = "~/Desktop/data/svn/vagueness-games/paper/plots/WithinGroupMeasures.pdf", height=5, width=8)
+# ggsave(combPlot, file = "~/Desktop/data/svn/vagueness-games/paper/plots/WithinGroupMeasures.pdf", height=5, width=8)
 
 
+## reduced plots for revision 2 of paper
 
+# focus on tolerance = 0.1 and 10 states:
+dsub = subset(D, D$Number.of.states == 10 & D$Tolerance == 0.1)
+wgdConcise = ggplot(dsub, aes(x = Impairment, y = inGroupDistance)) + 
+  geom_line( color = "grey") + geom_point() +
+  xlab("Impairment") + ylab("")
+show(wgdConcise)
+ggsave(wgdConcise, file = "~/Desktop/data/svn/vagueness-games/paper/version_02/plots/WithinGroupDistanceConcise.pdf", height=2, width=3) 
+
+eusub = subset(meanEUs, meanEUs$Number.of.states == 10 & meanEUs$Tolerance == 0.1)
+EUConcise = ggplot(eusub, aes(x = Impairment, y = Speaker.meanEU)) + 
+  geom_line( color = "grey") + geom_point() +
+  xlab("Impairment") + ylab("")
+show(EUConcise)
+ggsave(EUConcise, file = "~/Desktop/data/svn/vagueness-games/paper/version_02/plots/WithinGroupEUConcise.pdf", height=2, width=3) 

### 2015-01-16 paper/version_02/paper.tex
reworked sections 3 - 5

@@
 particular: (i) clear positive examples of a vague category should
-show a gradient transition to clear negative examples (towards
-so-called \emph{higher-order vagueness});\todo{check usage of term}
+show a gradient transition to clear negative examples;
 (ii) prototypes should likewise be gradient regions, peaking at the
@@
 biological evolution: signaling strategies are not inherited as a
-whole, but adapted piece-meal, e.g., by conditional imitation (more
-below). Still, the approach we suggest here is compatible with a
-formulation in terms of mixed strategies and an interpretation as
-biological evolution as well. To show this, we translate our initial
-formulation of the \rdd into one in using mixed strategies. Doing so
-identifies the \rdd as a special case of the \rmd.
+whole, but adapted piece-meal, e.g., by conditional imitation of
+``local'' behavior at a given choice point (more below). Still, the
+approach we suggest here is compatible with a formulation in terms of
+mixed strategies and an interpretation as biological evolution as
+well. To show this, we translate our initial formulation of the \rdd
+into one using mixed strategies. Doing so identifies the \rdd as a
+special case of the \rmd.
 
@@
 $\Sstrat$ and $\Rstrat$ to their \emph{diffusions} $D(\Sstrat)$ and
-$D(\Rstrat)$. Suppose that $\Sstrat$ and $\Rstrat$ capture a
+$D(\Rstrat)$. Suppose that $\Sstrat$ and $\Rstrat$ represent a
 population's average signaling behavior if there was no confusion of
@@
   \frac{\Sstrat(\messg_k \probbar \state_j) \cdot
-    \EU(\messg, \state_j,\Rstrat)}
+    \EU(\messg_k, \state_j,\Rstrat)}
   {\Phi(\state_j,\Sstrat, \Rstrat)} \\
@@
 
-The sequential definition is advantageous also because it highlights
-that other (discrete-time) evolutionary dynamics can be subjected to
-state-confusability in an analogous fashion. It would be entirely
-straightforward, for instance, to perturb a best-response dynamic in
-case we wanted to model cognizing, rational agents with a noisy
-perceptual system. Our choice of the \rd as a carrier dynamic is
-motivated by its generality and austerity, being compatible with a
-variety of implicit agent-level behavior, without presupposing
-extravagant higher cognitive abilities
-\citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, using the \rd
-as a carrier for diffusion, we actually obtain a special case of a
+The sequential definition highlights that other (discrete-time)
+evolutionary dynamics can be subjected to state-confusability in an
+analogous fashion. It would be entirely straightforward, for instance,
+to perturb a best-response dynamic in case we wanted to model
+cognizing, rational agents with a noisy perceptual system. Our choice
+of the \rd as a carrier dynamic is motivated by its generality and
+austerity, being compatible with a variety of implicit agent-level
+behavior, without presupposing extravagant higher cognitive abilities
+\citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, by using the
+\rd as a carrier for diffusion, we actually obtain a special case of a
 behavioral-strategy version of the replicator mutator dynamics. This
@@
 
-The \rmd has first been proposed first as a tool for studying the
-evolution of grammar
+The \rmd has first been proposed as a tool for studying the evolution
+of grammar
 \citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
@@
 
-A direct comparison between the \rdd and the \rmd is not possible,
-because the usual formulation of the latter is based on mixed, not
-behavioral strategies. But a discrete-time formulation of the \rmd in
-behavioral strategies, clearly identifies the \rdd to be a special
-case. From this, it follows that a mixed strategy version of the \rdd
-is just the \rmd with suitably defined mutation matrix. From this,
-even a continuous-time version can easily be retrieved if desired. To
-show this, we start with a delineation of mixed and behavioral
-strategies.
+The usual formulation of the \rmd is based on mixed, not behavioral
+strategies. But a discrete-time formulation of the \rmd in behavioral
+strategies clearly identifies the \rdd to be a special case. It
+follows that a mixed strategy version of the \rdd is just the \rmd
+with suitably defined mutation matrices. From this, even a
+continuous-time version can easily be retrieved if desired. To show
+this, we start with a delineation of mixed and behavioral strategies.
 
@@
 
-As a result, dynamics on behavioral strategies more coarse-grained and
-not necessarily equivalent to dynamics on mixed strategies. An
+As a result, dynamics on behavioral strategies are more coarse-grained
+and not necessarily equivalent to dynamics on mixed strategies. An
 advantage of dynamics in behavioral strategies is that they reduce
@@
 lexicon, so to speak. On the other hand, biological evolution seems
-more naturally modeled in terms of full strategies: it is a full set
+more naturally modeled in terms of mixed strategies: it is a full set
 of behavioral disposition that is bequeathed to the next generation.
@@
 
-The sequential reformulation of the \rmd in terms of a sequence of
-updates already closely resembles our initial formulation of the
-\rdd. However, there is a major difference between these. While the
-\rdd is a dynamic for behavioral strategies, the \rmd, as taken from
-the literature, applies to mixed strategies. Similarly, the
-confusability of states is an operation that is straightforwardly
-definable for behavioral strategies (see
+The sequential reformulation of the \rmd closely resembles our initial
+formulation of the \rdd. However, there is a major difference between
+these. While the \rdd is a dynamic for behavioral strategies, the
+\rmd, as taken from the literature, applies to mixed
+strategies. Similarly, the confusability of states is an operation
+that is straightforwardly definable for behavioral strategies (see
 Equation~(\ref{eq:confusion-function})), while mutation in the \rmd is
@@
 
-      \node[below of = mixed] (mutated) {$Q^C(\Smixed)$};
+      \node[below of = mixed] (mutated) {$\Mutate_{Q^C}(\Smixed)$};
 
@@
 
-      \node[below of = behavioral] (mutatedBeh) {$D_C(\Sstrat)$};
+      \node[below of = behavioral] (mutatedBeh) {$\Diff_C(\Sstrat)$};
 
@@
 
-    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(Q^C(\Smixed))$};
+    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(\Mutate_{Q^C}(\Smixed))$};
 
@@
 behavioral strategies, we show that there is a systematic translation
-from a confusion matrix $C$ to a mutation matriX $Q^C$ such that for
+from a confusion matrix $C$ to a mutation matrix $Q^C$ such that for
 each mixed sender strategy $\Smixed$, its unique corresponding
 behavioral strategy $\Sstrat = G(\Smixed)$ has a
-confusion-perturbation $C(\Sstrat)$ that corresponds, via $G$, to the
-mutation-perturbation $Q^C(\Smixed)$ (see
+confusion-perturbation $\Diff_C(\Sstrat)$ that corresponds, via $G$,
+to the mutation-perturbation $\Mutate_{Q^C}(\Smixed)$ (see
 Figure~\ref{fig:correspondence-result}). In other words, we
@@
 mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
-have $G(Q^C(\Smixed)) = C(\Sstrat)$. Similarly, for the receiver.
+have $G(\Mutate_{Q^C}(\Smixed)) = \Diff_C(\Sstrat)$. Similarly, for
+the receiver.
 %This is the content of Theorem~\ref{thm:Correspondence} given below, a
@@
 would take us, then apply diffusion. To understand the \rdd better, we
-look closer at the novel diffusion part in
+isolate the novel diffusion part in
 Section~\ref{sec:iterated-diffusion}, before exploring how diffusion
@@
 \end{align*}
-where $\impairment >= 0$ is an imprecision or indiscriminability
+where $\impairment \ge 0$ is an imprecision or indiscriminability
 parameter. When $\impairment=0$ agents perfectly discriminate between
@@
 we need to think of convergence in our sense as speed of evolution
-towards the first evolutionary bottleneck.\todo{check and improve
+towards the first evolutionary slowdown.\todo{check and improve
   this}
@@
 inverse distance from a pure strategy. The usual definition of entropy
-applies directly to mixed strategies:
-\begin{align*}
-  E(\Smixed) & = \sum_{\Spure \in \Messgs^\States} \Smixed(\Spure) \cdot
-  \log(\Smixed(\Spure)) & 
-  E(\Rmixed) & = \sum_{\Rpure \in \States^\Messgs} \Rmixed(\Rpure) \cdot \log(\Rmixed(\Rpure)) \,.
-\end{align*} 
-These measures are computationally expensive to calculate, since the
-size of the domain over which the sum is computed grows exponentially
-with the number of choice points. But provably equivalent metrics for behavioral
-strategies are ready to hand:
+applies directly to mixed strategies, but provably equivalent metrics
+for behavioral strategies are ready to hand:
 \begin{align*}
@@
 \cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
-\log(|\States|)$. We work with normalized values.
-
-The sender strategies in Figures~\ref{fig:example_stratsA} and
+\log(|\States|)$. We work with normalized values. The sender
+strategies in Figures~\ref{fig:example_stratsA} and
 \ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
@@
 normalize expected utility by the maximal amount of expected utility
-obtainable in the relevant game.
-
-The strategy pair in Figure~\ref{fig:example_stratsA} has a normalized
-expected utility of $0.992$, the pair in
-Figure~\ref{fig:example_stratsB} has $0.901$. Generally, vagueness and
-imprecision can be expected to decrease expected utility
-\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is whether
-communicative success drops unacceptably fast with moderate levels of
-vagueness and imprecision.
+obtainable in the relevant game. The strategy pair in
+Figure~\ref{fig:example_stratsA} has a normalized expected utility of
+$0.992$, the pair in Figure~\ref{fig:example_stratsB} has
+$0.901$. Generally, vagueness and imprecision can be expected to
+decrease expected utility
+\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
+whether communicative success drops unacceptably fast with moderate
+levels of vagueness and imprecision.
 
@@
     for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\impairment
-    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the mean of the
-    entropies for the sender and receiver strategy.}
+    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
+    the entropies for the sender and receiver strategy.}
   \label{fig:MeanMetrics}
@@
 gives two more examples of strategy pairs at stopping time. Both are
-obtained for the same triple of parameters, both converged and are
-convex. However, they are not equally efficient. In fact, the pair in
-Figure~\ref{fig:example_stratsC} has a normalized expected utility of
-$0.99$ while the one in Figure~\ref{fig:example_stratsD} only has
-$0.89$.
+obtained for the same triple of parameters, both converged and have
+convex sender strategies. However, they are not equally efficient. In
+fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized
+expected utility of $0.99$ while the one in
+Figure~\ref{fig:example_stratsD} only has $0.89$.
 
@@
 \emph{within group distances}, i.e., the average distances between all
-sender strategies in each parameter group, are plotted exemplarily on
-the left in Figure~\ref{fig:GroupMeasures} for $\beta = 0.1$ and $n_s
-= 10$. Despite minor quantitative differences the is the same for all
-other parameter settings that we tested: with increasing imprecision,
-the sender strategies evolving under \rdd were exactly alike (modulo
-swapping of messages). This means that perceptual imprecision unifies
-evolutionary outcomes and guarantees sender strategies that are not
-only convex, but also regular in that they induce a vague category
-split exactly in the middle of the unit interval.
+sender strategies obtained for the same parameter values, are plotted
+exemplarily on the left in Figure~\ref{fig:GroupMeasures} for $\beta =
+0.1$ and $n_s = 10$. Despite minor quantitative differences, the
+picture is the same for all other parameter settings that we tested:
+with increasing imprecision, the sender strategies evolving under \rdd
+were exactly alike (modulo swapping of messages). This means that
+perceptual imprecision unifies evolutionary outcomes and guarantees
+sender strategies that are not only convex, but also regular in that
+they induce a vague category split exactly in the middle of the unit
+interval.
 
@@
 receiver strategy at stopping time) scored when playing against an
-arbitrary other language of the same parameter group. The right side
-of Figure~\ref{fig:GroupMeasures}) gives a representative example. The
-observation repeats for other parameter values: while imprecision
-might decrease the communicative efficiency of individual languages,
-it increases the conceptual coherence and communicative success
-between independently evolving strategies. It is almost as if mere
-confusability of states imposes a regularity constraint on evolving
-categories.
+arbitrary other language obtained for the same parameter values. The
+right side of Figure~\ref{fig:GroupMeasures} gives a representative
+example. The observation repeats for other parameter values: while
+imprecision might decrease the communicative efficiency of individual
+languages, it increases the conceptual coherence and communicative
+success between independently evolving strategies. It is almost as if
+mere confusability of states imposes a regularity constraint on
+evolving categories.
 
@@
 
-\subsection{Model interpretation}
+\subsection{Interpretation of model and results}
 \label{sec:model-interpretation}
 
-We mentioned in passing in Section~\ref{sec:repl-diff-dynam-1} that
-adding diffusion to other discrete-time evolutionary dynamics that
-operate on behavioral strategies is entirely straightforward. We
-could, for instance, easily diffuse the outcome of a best-response
-dynamic at each time step. That would make sense if we thought of
-agents as prone to confusing similar states, but otherwise rationally
-optimizing behavior. The reason that we chose the \rd to combine with
-diffusion is twofold. A minor reason is that it makes for a
-conceptually interesting link with the \rmd. A more important reason
-is that the \rd is especially versatile and non-committal about what
-the exact process of adaption is that is being modelled. Generally,
-the \rd could be a biological or a cultural dynamic, and so could the
-\rdd.
-
-Under a biological interpretation of the \rd, signaling strategies
-are innate behavioral dispositions of organisms, steps in the
-evolutionary process are successive generations, and selection
-captures the reproductive advantage of fitter individuals. This
-interpretation, strictly speaking, requires formalization in terms of
-mixed strategies via the \rmd, and possibly also a symmetrizing of the
-game, so that every agent is assumed to have a unique sender and
-receiver role at the same time, the pair of which is bequeathed onto
-the next generation.
-
-Diffusion, in this context, could be either of two things. It could be
-differential mutation probabilities in line with other interpretations
-of mutation in the \rmd
-\citep[e.g.][]{NowakKomarova2001:Evolution-of-Un,KomarovaNiyogi2001:The-Evolutionar}:
-pure strategies are not faithfully reproduced, e.g. by imperfections
-of learning or underdetermination of input, and mistakes in inheriting
-pure strategies are more likely if they result from the confusion of
-similar states. Another possible biological interpretation of the
-\rdd is that inheritance is faithful, but strategies are noisily
-realized. Strategies are not necessarily selected for what they are,
-but rather for how they are realized, once noise is factored in.
-
-As a cultural dynamic, the \rd can also be interpreted as a high-level
-description of the likely development of other behavioral adaptation
-processes, like differential imitation \citep[see][for various
-derivations of the \rd]{Sandholm2013:Population-Game}. Under this
-interpretation, what is subject to evolutionary optimization is
-behavior, not genetic make-up. Fitness captures the success of
-behavioral patterns.  Differential reproduction represents the
-tendency of more successful ways of behavior to be more likely
-realized, be it through imitation or some solitary learning or
-optimization mechanism.
-
-Under this non-biological interpretation, we have again two options of
-picturing what diffusion is. One possibility is that diffusion is
-noise in the adoption of strategies, e.g. by conditional imitation, of
-strategies by other signalers. Another possibility is, again, that
-diffusion is noise on the realization of strategies: while behavior is
-optimized to be efficient (be it due to learning, introspection or
-imitation), realization of strategies is bound to be noisy due to
-confusability of similar states.
-
-We believe that all of the four mentioned interpretations are, on
-first approximations, feasible conceptualizations of the \rdd, and
-that it is a good thing to know of a working account of vague
-signaling that sketches where fitness-based selection under
-state-confusability will take us, abstracting away from the details of
-actually playing the game, inheriting, imitating or otherwise
-optimizing behavior in whatever particular way. It is a good thing to
-know this on the macro-level, especially since there are also
-micro-level accounts that nicely complement the picture. We turn to
-one such next.
+The \rdd adds expected confusion of similar states to the \rd in
+behavioral strategies. The result is an abstract evolutionary dynamic
+that describes the most likely path of development of aggregate
+population behavior in which any process of behavioral adaptation at
+the agent level that is consistent with the \rd in behavioral
+strategies is perturbed by systematic noise in the realization of the
+agents' strategies. Given the use of behavioral strategies, an
+interpretation as a process of cultural evolution is most salient,
+according to which agents adapt their choice dispositions locally at
+each choice point. But the \rdd is also extensible to biological
+evolution on mixed strategies, in which case it is a special case of
+the \rmd with mutation matrices derived from confusion matrices as
+shown in Section~\ref{sec:diffusion-as-special}.
+
+The \rdd is not only compatible with several interpretations of how,
+at an agent level, the population-level changes in strategy
+proportions come about. It is also compatible with several
+interpretations of diffusion. Diffusion can be pictured as a flaw or a
+virtue. If we think of perceptual imprecision as a hard limitation on
+information processing, vagueness is a natural byproduct of language
+evolution. Results from numerical simulations show that such
+imprecision leads to vagueness, but does not need to undermine the
+prospect of evolving regular and by-and-large efficient signaling
+behavior. On the other hand, there appears to be a possibly beneficial
+side-effect of state confusability: it speeds up convergence to
+regularly shaped and convex categories. This suggests that diffusion
+of behavior, as we implemented it here, could also be seen as a form
+of stimulus-generalization that is evolutionarily beneficial, because
+it accelerates the emergence of uniform category formation. This
+latter view is also suggested by a like-minded account to which we
+turn next.
 
@@
 Conceptually, the \rdd is a more abstract framework than generalized
-reinforcement learning. The latter is foremost motivated as a learning
-dynamic that has two players adapt their individual strategies after
-each concrete round of play. In contrast, the \rdd describes a more
-abstract, average dynamical change in behavioral
-dispositions. Although the behavior of (some forms of) reinforcement
-learning mirror those of the \rd (at some stage in time)
+reinforcement learning. The latter is most naturally seen as an
+agent-based learning dynamic that has two players adapt their
+individual strategies after each concrete round of play. In contrast,
+the \rdd describes a more abstract, average dynamical change in
+behavioral dispositions at the aggregate population level. Although
+the behavior of (some forms of) reinforcement learning mirror those of
+the \rd (at some stage in time)
 (\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
 this does not mean that \emph{generalized} reinforcement learning is
-also necessarily a plausible high-level description of, say,
+also necessarily a plausible population-level description of, say,
 generalized learning in a population of agents. Seen in this light,
@@
 \citet{FrankeJager2010:Vagueness-Signa} show that quantal response
-equilibria of sim-max games, i.e., pairs of sender and receiver
-strategies such that the sender strategy is the quantal response to
-the expected utilities under the receiver strategy and vice versa, can
-show the desired marks of vague
-signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
+equilibria of sim-max games can show the desired marks of vague
+signaling. A quantal response equilibrium is a pair of sender and
+receiver strategies such that the sender strategy is the quantal
+response to the expected utilities under the receiver strategy and
+vice versa. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
 quantal response equilibrium for a sim-max game, as used in our set-up
@@
 expected utility than that it arises from confusion of similar
-states. To see what the problem is, we can look at cases like given in
+states. To see what the problem is, consider
 Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
@@
 the state space is intuitively unbounded, as for instance for the
-property ``tall''. If the usual interpretation of a ``tall man'' peaks
+property ``tall.'' If the usual interpretation of a ``tall man'' peaks
 at around, say, 195cm then when meeting a giant of $n$ meters, senders
@@
 Admittedly, this argument hinges on the choice of utility
-function. Still, to the extent that the chosen lower bounded utility
-functions are reasonable---and we think they are very reasonable---
-the case suggests that quantal responses are not a good model,
-intuitively speaking, for why linguistic categories are
-vague. Vagueness is more plausibly an effect of perceptual confusion
-of similar states than of computational errors in maximizing expected
-utility.
+function. Still, to the extent that the utility functions we chose are
+reasonable---and we think they are very reasonable--- the case
+suggests that quantal responses are not a good model, intuitively
+speaking, for why linguistic categories are vague. Vagueness is more
+plausibly an effect of perceptual confusion of similar states than of
+computational errors in maximizing expected utility.
 
@@
 the inclusion of mild levels of stochastic confusability of states
-does not undermine the possibility of evolving communicatively
-successful signaling strategies at all. On the contrary, strategies
-that evolved under mild diffusion induce a highly regular and
-systematic category structure that shows the signs of vagueness as
-desired. There might even be a higher-order benefit to the presence of
-imprecision, in that it can accelerate convergence to optimal
-categorization, by preventing evolutionary paths to stay near
-inefficient non-convex strategies for a long time. Diffusion thus also
-unifies and regularizes evolutionary outcomes in such a way that
-sub-optimal categorization is avoided.
+does not undermine the possibility of evolving relatively successful
+signaling strategies. On the contrary, strategies that evolved under
+mild diffusion induce a highly regular and systematic category
+structure that shows the signs of vagueness as desired. There might
+even be a higher-order benefit to the presence of imprecision, in that
+it can accelerate convergence to optimal categorization, by preventing
+evolutionary paths to stay near inefficient non-convex strategies for
+a long time. Diffusion thus also unifies and regularizes evolutionary
+outcomes in such a way that sub-optimal categorization is avoided in
+short-term trajectories.
 

### 2015-01-16 paper/version_02/paper.tex
answer to reviewer comments

@@
 A sim-max game consists of a set of states $\States$, a set of
-messages $\Messgs$ with much fever messages than states, a prior
+messages $\Messgs$ with much fewer messages than states, a prior
 probability distribution $\Pr \in \Delta(\States)$ that gives the
@@
 convex sender strategies, and led to higher proportions of
-convergence. In fact, sufficient impairment always assured convergence
+convergence. In fact, sufficient impairment always ensured convergence
 and convexity. It appears that perceptual imprecision leads to

### 2015-01-17 paper/version_02/paper.tex
Small changes after first pass. Made introduction into a section. Changed general attitude towards quantal response.

@@
   imprecision leads to vague, yet by-and-large communicative efficient
-  signal use, and, moreover, unifies evolutionary outcomes and helps
-  avoid suboptimal categorization.
+  signal use, and, moreover, that it unifies evolutionary outcomes and helps
+  avoiding suboptimal categorization.
 \end{abstract}
+
+\section{Introduction}
+\label{sec:introduction}
 
@@
 that we are concerned with here is how vagueness could arise and be
-maintained in the first place. This is a serious worry to
+maintained in the first place. This is a serious issue for
 functionalist accounts that maintain that concepts and linguistic
-meanings evolved in order to be efficient. Since the existence of
+meanings evolved driven towards efficiency. Since the existence of
 unclear borderline cases seems to entail inefficiency of
@@
 the evolution of categorization and language use that yields
-by-and-large efficient categories \emph{and} vagueness.
+by-and-large efficient categorization \emph{and} vagueness.
 
@@
 paper contributes to the latter line of thought. Concretely, we
-introduce the replicator diffusion dynamics ---a novel variant of the
-replicator mutator dynamic--- that integrates stochastic noise on the
+introduce the replicator diffusion dynamics---a novel variant of the
+replicator mutator dynamic---that integrates stochastic noise on the
 differential confusability of similar stimuli. The dynamic proposed
@@
 show that stochastic noise can not only lead to vague, yet relatively
-efficient signal use, but can also unify evolutionary outcomes and
+efficient signal use, but also unify evolutionary outcomes and
 help avoid suboptimal categorization.
@@
 take here is similar in spirit, but different in relevant detail. We
-are argue here that \citeauthor{FrankeJager2010:Vagueness-Signa}'s
-approach builds on a dubious assumption about the source of
-vagueness. While \citeauthor{OConnor2013:The-Evolution-o} gives an
-agent-level learning dynamic under which vague signaling arises, we
-complement the picture by giving a population-level dynamic that
+argue here that \citeauthor{FrankeJager2010:Vagueness-Signa}'s
+approach gives rise to problematic predictions under certain conditions
+which are avoided in our approach. However, the two are not necessarily
+competing explanations since they are motivated differently and it is
+certainly possible that both the mechanisms proposed by them, as the
+ones proposed here play a role in the whole phenomenon.
+Regarding the work by \citeauthor{OConnor2013:The-Evolution-o}, while she
+gives an agent-level learning dynamic under which vague signaling arises,
+we complement the picture by giving a population-level dynamic that
 abstracts from the concrete behavior of signalers. A more in-depth
@@
 simplifies numerical simulations. But it also points towards an
-interpretation as a process of cultural evolution, rather than
-biological evolution: signaling strategies are not inherited as a
+interpretation of the process as cultural evolution, rather than
+biological evolution. Signaling strategies are not inherited as a
 whole, but adapted piece-meal, e.g., by conditional imitation of
@@
 The \rd in behavioral strategies describes the most likely path of
-development of average behavior in a (huge and homogeneous) population
+development of average behavior in a population
 of agents when behavioral adaptations occur independently at each
@@
 the \rdd captures the most likely path of development of aggregate
-behavior in a (huge \& homogeneous) society, in which behavioral
+behavior in a society, in which behavioral
 changes occur independently at each choice point, as described by the
@@
 (\ref{eq:construction-sen}), we can formulate the desired
-correspondence result. A proof is given in Appendix~\ref{sec:proofs}.
+correspondence result, a proof for which is given in Appendix~\ref{sec:proofs}.
 
@@
   (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
-  \Diff_C(\Sstrat)$. And, (ii) if $G(\Rmixed) = \Rstrat$, then
+  \Diff_C(\Sstrat)$. (ii) If $G(\Rmixed) = \Rstrat$, then
   $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
@@
 
-\subsection{(Iterated) Diffusion}
+\subsection{Iterated diffusion}
 \label{sec:iterated-diffusion}
@@
 
-To further explore our simulation results, we track how vague,
-communicatively efficient and generally well-structured the recorded
+To further explore our simulation results, we calculated
+metrics that aim to numerically capture how vague, generally
+well-structured, and communicatively efficient the recorded
 strategy pairs were. \emph{Entropy} captures the amount of
@@
 higher entropy than crisp ones and that increasing imprecision would
-increase entropy, all else equal.
+lead to increased entropy, all else being equal.
 
@@
 Figure~\ref{fig:example_stratsA} is not convex, while the one in
-Figure~\ref{fig:example_stratsB} is.  One would perhaps not expect a
-systematic relation between vagueness and imprecision, on the one
-hand, and convexity on the other.
+Figure~\ref{fig:example_stratsB} is.  One would a priori not expect a
+systematic relation between imprecision and convexity.
 
@@
 expected utility. Importantly, however, imprecision does not
-necessarily lead to disastrous decline of communicative success. On
-the other hand, imprecision increased the number of outcomes with
-convex sender strategies, and led to higher proportions of
+necessarily lead to disastrous decline of communicative success. What was
+not expected, on the other hand, was a relation between imprecision and
+convexity. However, the corresponding plot clearly shows that
+higher imprecision led to a higher number of outcomes with
+convex sender strategies. Moreover, it also led to higher likelihood of
 convergence. In fact, sufficient impairment always ensured convergence
 and convexity. It appears that perceptual imprecision leads to
-slightly less communicative efficiency, more vagueness but more
+more entropy, slightly less communicative efficiency, but more
 regular, well-behaved languages in shorter time.
 
-Beyond promoting convexity, diffusion also has another interesting
+Beyond promoting convexity and convergence, diffusion also has another interesting
 regularizing effect on the evolution of signaling. There was very
@@
 gives two more examples of strategy pairs at stopping time. Both are
-obtained for the same triple of parameters, both converged and have
+obtained for the same triple of parameters, both converged before the
+maximum number of rounds, and both have
 convex sender strategies. However, they are not equally efficient. In
@@
 Interestingly, this type of variability in evolutionary outcomes can
-be weeded out by impairment. To see this, we calculated the average
+be weeded out by impairment. To investigate this, we calculated the average
 distance between evolved sender strategies within each group of trials
@@
 of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
-where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. The
+where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. An example of the
 \emph{within group distances}, i.e., the average distances between all
-sender strategies obtained for the same parameter values, are plotted
-exemplarily on the left in Figure~\ref{fig:GroupMeasures} for $\beta =
-0.1$ and $n_s = 10$. Despite minor quantitative differences, the
-picture is the same for all other parameter settings that we tested:
+sender strategies obtained for the same parameter values, is plotted
+on the left in Figure~\ref{fig:GroupMeasures} for $\beta =
+0.1$ and $n_s = 10$. Despite some quantitative differences, the
+general trend is the same for all other parameter settings that we tested:
 with increasing imprecision, the sender strategies evolving under \rdd
-were exactly alike (modulo swapping of messages). This means that
+were much more alike (modulo swapping of messages). This means that
 perceptual imprecision unifies evolutionary outcomes and guarantees
@@
 regularly shaped and convex categories. This suggests that diffusion
-of behavior, as we implemented it here, could also be seen as a form
+of behavior, as we implemented here, could also be seen as a form
 of stimulus-generalization that is evolutionarily beneficial, because
@@
 choices $\Acts$ of an agent, then a best response function would have
-the agent choose $\act$ only if $U(\act) = \max_{\act' \in \Acts}
-U(a)$. A quantal response function rather assumes that agents would
+the agent choose $\act$ only if $U(\act) = \max_{\act^\prime \in \Acts}
+U(\act^\prime)$. A quantal response function rather assumes that agents would
 choose $\act$ with a probability proportional to $\expo(\lambda \cdot
@@
 
-But not all quantal response equilibria are equally plausible, and the
-ones that are not suggest that it is less natural to think of
-vagueness as arising from errors in the computation or maximization of
-expected utility than that it arises from confusion of similar
-states. To see what the problem is, consider
+But not all quantal response equilibria are equally plausible.
+%, and the
+%ones that are not suggest that it is less natural to think of
+%vagueness as arising from errors in the computation or maximization of
+%expected utility than that it arises from confusion of similar states
+To see what the problem is, consider
 Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
@@
 
-That this is conceptually odd shows even more clearly in a case where
+That this is problematic shows even more clearly in a case where
 the state space is intuitively unbounded, as for instance for the
@@
 function. Still, to the extent that the utility functions we chose are
-reasonable---and we think they are very reasonable--- the case
-suggests that quantal responses are not a good model, intuitively
-speaking, for why linguistic categories are vague. Vagueness is more
-plausibly an effect of perceptual confusion of similar states than of
-computational errors in maximizing expected utility.
+reasonable---and we think they are very reasonable---the case
+suggests that quantal responses do not paint a complete picture. Naturally,
+we cannot claim our approach provides valid predictions for all cases.
+Given the complexity of the phenomenon, and the fact that
+the quantal response approach is built on very reasonable assumptions,
+it is likely that both proposals are relevant pieces of the puzzle.
+%Vagueness is more
+%plausibly an effect of perceptual confusion of similar states than of
+%computational errors in maximizing expected utility.
 

### 2015-01-18 paper/version_02/paper.tex
Moved diffusion as mutation to its own section

@@
 presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
-introduces the replicator diffusion dynamic and elaborates on its
-relation with the replicator mutator
+introduces the replicator diffusion dynamic and Section~\ref{sec:diffusion-as-special}
+elaborates on its relation with the replicator mutator
 dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
@@
 The formulation of the \rdd given and explored here is in terms of
-behavioral strategies, not in terms of mixed strategies (more on this
-below). Dynamics on behavioral strategies assume that agents can adapt
+behavioral strategies, not in terms of mixed strategies. Dynamics on
+behavioral strategies assume that agents can adapt
 their behavior locally, so to speak, i.e., independently at each
@@
 mixed strategies and an interpretation as biological evolution as
-well. To show this, we translate our initial formulation of the \rdd
+well. In Section~\ref{sec:diffusion-as-special}, we
+translate our initial formulation of the \rdd
 into one using mixed strategies. Doing so identifies the \rdd as a
-special case of the \rmd.
+special case of the \rmd. For now, let us first focus on the
+formulation of the dynamic in terms of behavioral strategies alone.
 
@@
 Section~\ref{sec:repl-diff-dynam-1}.
-Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
-Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
-a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
-\ref{sec:diffusion-as-special} are conceptually important, but rather
-technical and not strictly necessary to appreciate the main point of
-the paper.
+%Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
+%Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
+%a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
+%\ref{sec:diffusion-as-special} are conceptually important, but rather
+%technical and not strictly necessary to appreciate the main point of
+%the paper.
 
@@
 diffusion from perceptual confusion also in models of biological
-evolution. The remainder of this section therefore works out the
-relation between the \rdd and the \rmd.
+evolution.
+
+The following section works out the relation between the \rdd and the \rmd.
+Although this section is conceptually important, it is also rather
+technical and not strictly necessary to appreciate the main point of
+the paper.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Diffusion as mutation}
+\label{sec:diffusion-as-special}
 
@@
 
-\subsection{Diffusion as mutation}
-\label{sec:diffusion-as-special}
+\subsection{Relating the dynamics}
 
@@
 
-The previous section showed that the \rdd can be conceptualized as a
+The previous sections showed that the \rdd can be conceptualized as a
 two-step procedure: we first calculate where the replicator dynamic
@@
 evolution on mixed strategies, in which case it is a special case of
-the \rmd with mutation matrices derived from confusion matrices as
+the \rmd with mutation matrices derived from confusion matrices, as
 shown in Section~\ref{sec:diffusion-as-special}.

### 2015-01-18 paper/version_02/paper.tex
Removed reference to Sainsbury and Keefe.

@@
 (ii) prototypes should likewise be gradient regions, peaking at the
-center of the vague category they represent
-\citep[e.g.][]{Sainsbury1991:Is-There-Higher,KeefeSmith1997:Vagueness:-A-Re}.\todo{check
-  whether references are okay/appropriate in this context}
+center of the vague category they represent.
 

### 2015-01-18 paper/version_02/paper.tex
Added motivation for the fixed maximum number of iterations.

@@
   0.3}$. Each trial ran for a maximum of 200 update steps of the
-\rdd. A trial was considered converged and stopped before the maximum
-of 200 rounds if the total amount of change between strategies before
-and after the current \rdd step was smaller than a suitably chosen
+\rdd. A trial was considered converged, and thus stopped before the maximum
+of 200 rounds, if the total amount of change between strategies before
+and after an \rdd step was smaller than a suitably chosen
 threshold. It is not guaranteed that strategies at halting time had
 converged to the eventual attracting state, whether they ran for 200
-rounds or not. Our notion of convergence therefore is only a
-categorical measure for the speed of approaching some state in which
-the force of evolutionary optimization has fallen under a
-(well-considered, but eventually arbitrary) threshold. In other words,
-we need to think of convergence in our sense as speed of evolution
-towards the first evolutionary slowdown.\todo{check and improve
-  this}
+rounds or not. Our notion of convergence is therefore only a
+categorical measure for reaching a certain (well-considered, but
+eventually arbitrary) degree of stability. In principle, rather than
+setting a hard limit on the maximum number of rounds and analyzing
+convergence categorically, one could also run each trial until reaching
+our criterion for convergence and analyze the number of rounds it took to get there.
+Our choice was initially motivated by pragmatic concerns regarding length of
+simulation time, but it is also justifiable if we reflect on the
+nature of language evolution as a continuous process.
+Language adapts to an ever-changing reality; it is rare, if not impossible,
+that a lexicon has an infinite amount of time to neatly converge to a stable
+state. Therefore, investigating speed of convergence within a limited
+time frame can be seen as better reflecting the natural constraints of
+language evolution.
 

### 2015-01-26 paper/version_02/paper.tex
Renamed section 5.2.
paper revision 2 some (almost) final edits

@@
   signal use, and, moreover, that it unifies evolutionary outcomes and helps
-  avoiding suboptimal categorization.
+  avoid suboptimal categorization.
 \end{abstract}
@@
 argue here that \citeauthor{FrankeJager2010:Vagueness-Signa}'s
-approach gives rise to problematic predictions under certain conditions
-which are avoided in our approach. However, the two are not necessarily
-competing explanations since they are motivated differently and it is
-certainly possible that both the mechanisms proposed by them, as the
-ones proposed here play a role in the whole phenomenon.
-Regarding the work by \citeauthor{OConnor2013:The-Evolution-o}, while she
-gives an agent-level learning dynamic under which vague signaling arises,
-we complement the picture by giving a population-level dynamic that
-abstracts from the concrete behavior of signalers. A more in-depth
-comparison is deferred until Section~\ref{sec:discussion}. Let us
-first look into our own proposal in more detail.
+approach gives rise to problematic predictions under certain
+conditions which are avoided in our approach. While
+\citeauthor{OConnor2013:The-Evolution-o} gives an agent-level learning
+dynamic under which vague signaling arises, we complement the picture
+by giving a population-level dynamic that abstracts from the concrete
+behavior of signalers. A more in-depth comparison is deferred until
+Section~\ref{sec:discussion}. Let us first look into our own proposal
+in more detail.
 
@@
 
-The most natural interpretation of this sequential definition is that
-the \rdd captures the most likely path of development of aggregate
-behavior in a society, in which behavioral
-changes occur independently at each choice point, as described by the
-\rd for behavioral strategies. Whatever strategy an agent does
-actually hold, its realization is perturbed by imperfect
-discriminability of states in such a way that the expected overall
-aggregate behavior is the diffusion of the strategies that the agents
-would play in a noise-free world. Any agent-level interpretation of
-the \rd in behavioral strategies seems compatible with this
+The most straightforward interpretation of this sequential definition
+is that the \rdd captures the most likely path of development of
+aggregate behavior in a society, in which behavioral changes occur
+independently at each choice point, as described by the \rd for
+behavioral strategies. Whatever strategy an agent does actually hold,
+its realization is perturbed by imperfect discriminability of states
+in such a way that the expected overall aggregate behavior is the
+diffusion of the strategies that the agents would play in a noise-free
+world. Any agent-level interpretation of the \rd in behavioral
+strategies seems \emph{prima facie} compatible with this
 interpretation.
@@
 evolutionary dynamics can be subjected to state-confusability in an
-analogous fashion. It would be entirely straightforward, for instance,
-to perturb a best-response dynamic in case we wanted to model
-cognizing, rational agents with a noisy perceptual system. Our choice
-of the \rd as a carrier dynamic is motivated by its generality and
-austerity, being compatible with a variety of implicit agent-level
-behavior, without presupposing extravagant higher cognitive abilities
+analogous fashion. It is easy, for instance, to perturb a
+best-response dynamic in case we wanted to model cognizing, rational
+agents with a noisy perceptual system. Our choice of the \rd as a
+carrier dynamic is motivated by its generality and austerity, being
+compatible with a variety of implicit agent-level behavior, without
+presupposing extravagant higher cognitive abilities
 \citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, by using the
@@
 second. Consequently, a locally less efficient type of behavior can be
-demoted because it is correlated with efficient behavior at another
+promoted because it is correlated with efficient behavior at another
 choice point. Evolutionary dynamics working on behavioral strategies
@@
 lexicon, so to speak. On the other hand, biological evolution seems
-more naturally modeled in terms of mixed strategies: it is a full set
-of behavioral disposition that is bequeathed to the next generation.
+more naturally modeled in terms of mixed strategies, when it is a full
+set of behavioral disposition that is bequeathed to the next
+generation.
 
@@
 
-  \caption{Examples of Nosofsky-similarity (non-normalized) and
-    derived probability of state confusion (normalized) for
-    $\card{\States} = 50$ and different
-    values of imprecision $\impairment$.}
+  \caption{Examples of Nosofsky-similarity for different values of
+    imprecision.}
   \label{fig:NosofskySim}
@@
 Confusion of states is then also a function of imprecision parameter
-$\impairment$ (see Figure~\ref{fig:NosofskySim}). For $\impairment =
-0$ the confusion matrix has $C_{ii} = 1$; for $\impairment > 0$, $C$
-is positive, i.e., $C_{ij} >0$ for all $i$ and $j$; for $\impairment
-\rightarrow \infty$ we find $C_{ij} = \nicefrac{1}{\card{\States}}$
-for all $i$ and $j$.
+$\impairment$. For $\impairment = 0$, the confusion matrix has $C_{ii}
+= 1$; for $\impairment > 0$, $C$ is positive, i.e., $C_{ij} >0$ for
+all $i$ and $j$; for $\impairment \rightarrow \infty$, we find $C_{ij}
+= \nicefrac{1}{\card{\States}}$ for all $i$ and $j$.
 
@@
 
-\subsection{Exploring the replicator diffusion dynamics}
+\subsection{Numerical simulations}
 \label{sec:simulations}
@@
 smoothed-out prototype regions. Secondly, (sender) strategies can
-approach non-convex pure strategies under the replicator dynamic. We
-see this in our limited-time simulations, but this also holds, for
-some types of utility functions, for the limiting case. This was first
+approach non-convex pure strategies under the replicator dynamic and
+linger there for vast amounts of time, possibly indefinitely. We see
+this in our limited-time simulations, but this also holds, for some
+types of utility functions, for the limiting case. This was first
 observed by Elliott Wagner in unpublished work
@@
 
-To further explore our simulation results, we calculated
-metrics that aim to numerically capture how vague, generally
-well-structured, and communicatively efficient the recorded
-strategy pairs were. \emph{Entropy} captures the amount of
-systematicity or regularity in signal use. \emph{Convexity} captures
-whether a behavioral strategy would project onto a convex pure
+To further explore our simulation results, we calculated metrics that
+aim to numerically capture how vague, generally well-structured, and
+communicatively efficient the recorded strategy pairs
+were. \emph{Entropy} captures the amount of systematicity or
+regularity in signal use. \emph{Convexity} captures whether a
+behavioral strategy would project onto a convex pure
 strategy. \emph{Expected utility} measures the communicative
-efficiency of evolved strategy pairs.\todo{take average of entropy}
+efficiency of evolved strategy pairs.
 
@@
 
-\paragraph{Convexity.} At least for sender strategies, which converge
-faster, it also makes sense to define a categorical measure of
-convexity that compensates for potential vagueness. To determine
-whether a sender strategy $\Sstrat$ is convex despite possibly being
-vague, we look at the derived pure strategy $\Spure$ for which
-$\Spure(\state) = \argmax_{\messg' \in \Messgs}
+\paragraph{Convexity.} At least for sender strategies, which develop
+faster than receiver strategies, it also makes sense to define a
+categorical measure of convexity that compensates for potential
+vagueness. To determine whether a sender strategy $\Sstrat$ is convex
+despite possibly being vague, we look at the derived pure strategy
+$\Spure$ for which $\Spure(\state) = \argmax_{\messg' \in \Messgs}
 \Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
@@
 expected utility. Importantly, however, imprecision does not
-necessarily lead to disastrous decline of communicative success. What was
-not expected, on the other hand, was a relation between imprecision and
-convexity. However, the corresponding plot clearly shows that
-higher imprecision led to a higher number of outcomes with
-convex sender strategies. Moreover, it also led to higher likelihood of
-convergence. In fact, sufficient impairment always ensured convergence
-and convexity. It appears that perceptual imprecision leads to
-more entropy, slightly less communicative efficiency, but more
-regular, well-behaved languages in shorter time.
+necessarily lead to disastrous decline of communicative success. What
+was not expected, on the other hand, was a relation between
+imprecision and convexity. However, the corresponding plot clearly
+shows that higher imprecision led to a higher number of outcomes with
+convex sender strategies. Moreover, it also led to higher likelihood
+of convergence. In fact, sufficient impairment always ensured
+convergence and convexity. It appears that perceptual imprecision
+leads to more vagueness, slightly less communicative efficiency, but
+more regular, well-behaved languages in shorter time.
 
@@
 where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. An example of the
-\emph{within group distances}, i.e., the average distances between all
+\emph{within group distance}, i.e., the average distances between all
 sender strategies obtained for the same parameter values, is plotted
@@
 function. Still, to the extent that the utility functions we chose are
-reasonable---and we think they are very reasonable---the case
-suggests that quantal responses do not paint a complete picture. Naturally,
-we cannot claim our approach provides valid predictions for all cases.
-Given the complexity of the phenomenon, and the fact that
-the quantal response approach is built on very reasonable assumptions,
-it is likely that both proposals are relevant pieces of the puzzle.
-%Vagueness is more
-%plausibly an effect of perceptual confusion of similar states than of
-%computational errors in maximizing expected utility.
+reasonable---and we think they are very reasonable--- the case
+suggests that quantal responses alone are not a good model,
+intuitively speaking, for why linguistic categories are
+vague. Vagueness is more plausibly an effect of perceptual confusion
+of similar states than of computational errors in maximizing expected
+utility. Even more plausibly, though, perceptual confusion \emph{and}
+computational errors, together with even further factors such as
+limited observations, imperfect memory and more, are needed for a
+complete understanding of why vague categorization and linguistic
+meaning as a natural concomitant of a distinguished cognitive
+make-up.
 

### 2015-01-26 paper/version_02/paper.tex
Small fixes. Rephrased  end of Section 6.

@@
 distribution over states and $\Util_{\sen,\rec} \mycolon \States
-\times \Messgs \times \Acts \rightarrow \mathds{R}$ the senders's and
+\times \Messgs \times \Acts \rightarrow \mathds{R}$ the sender's and
 receiver's utility functions. A behavioral strategy is a function that
@@
 lexicon, so to speak. On the other hand, biological evolution seems
-more naturally modeled in terms of mixed strategies, when it is a full
+more naturally modeled in terms of mixed strategies, since it is a full
 set of behavioral disposition that is bequeathed to the next
@@
 \citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. A
-quantal response function is a paramterized generalization of the
+quantal response function is a parameterized generalization of the
 classic best response function. For example, if $U \mycolon \Acts
@@
 function. Still, to the extent that the utility functions we chose are
-reasonable---and we think they are very reasonable--- the case
-suggests that quantal responses alone are not a good model,
-intuitively speaking, for why linguistic categories are
-vague. Vagueness is more plausibly an effect of perceptual confusion
-of similar states than of computational errors in maximizing expected
-utility. Even more plausibly, though, perceptual confusion \emph{and}
-computational errors, together with even further factors such as
-limited observations, imperfect memory and more, are needed for a
-complete understanding of why vague categorization and linguistic
-meaning as a natural concomitant of a distinguished cognitive
-make-up.
+reasonable---and we think they are very reasonable---the case
+suggests that quantal responses alone are not a good model for why
+linguistic categories are vague. With respect to this issue, perceptual
+confusion of similar states produces more intuitive predictions. What
+is more plausible, though, is that both perceptual confusion \emph{and}
+computational errors in maximizing expected utility play a role in
+making vague categorization a pervasive feature of natural language
+meaning. Perhaps even further factors, such as limited observations,
+imperfect memory, and more, should be considered in order to form
+a more complete understanding of the phenomenon of vagueness.
 

### 2015-01-27 paper/version_02/paper.tex
Revision 1, final sent

@@
 lexicon, so to speak. On the other hand, biological evolution seems
-more naturally modeled in terms of mixed strategies, since it is a full
-set of behavioral disposition that is bequeathed to the next
+more naturally modeled in terms of mixed strategies whenever it is a
+full set of behavioral disposition that is bequeathed to the next
 generation.
@@
 function. Still, to the extent that the utility functions we chose are
-reasonable---and we think they are very reasonable---the case
-suggests that quantal responses alone are not a good model for why
-linguistic categories are vague. With respect to this issue, perceptual
-confusion of similar states produces more intuitive predictions. What
-is more plausible, though, is that both perceptual confusion \emph{and}
+reasonable---and we think they are very reasonable---the case suggests
+that quantal responses alone are not a good model for why linguistic
+categories are vague. With respect to this issue, perceptual confusion
+of similar states produces more intuitive predictions. What is more
+plausible, though, is that both perceptual confusion \emph{and}
 computational errors in maximizing expected utility play a role in
@@
 meaning. Perhaps even further factors, such as limited observations,
-imperfect memory, and more, should be considered in order to form
-a more complete understanding of the phenomenon of vagueness.
+imperfect memory or uncertainty about the exact prior distribution of
+states, should be considered in order to form a more complete
+understanding of the phenomenon of vagueness.
 

### 2015-04-03 paper/derivation_RDD/derivation_RDD.tex
derivation of RDD from conditional imitiation - first shot

\documentclass[fleqn,reqno,10pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\doublespacing

\begin{document}

\section*{Derivation of the RDD by imitation-based revision}

We derive the proposed \rdd as the mean field of an imitative revision
protocol, following the example 4.3.1 of
\citet{Sandholm2010:Population-Game}. Our derivation relies on the
usual assumptions about conditional imitation at the agent level, but
must add special assumptions about how noise affects revision
probabilities. In keeping with the idea that agents in evolutionary
models are largely myopic, we assume here that noise is real but that
agents are unaware of it. When agents perceive the actual expected
utility of their choice and of another agent's choice (as is usual in
the motivation of imitation-based revision protocols), we assume
additionally that agents confuse their possibly non-veridical
subjective representations of states with the actual ones (details
below).

\paragraph{Notation.} As previously in this paper, let
$\Sstrat(\mymessg{i} \probbar \mystate{o})$ be the probability that a
speaker randomly picked from the population chooses $\mymessg{i}$ if
the actual/objective state is $\mystate{o}$. This factors in the
possibility that the sampled speaker perceives the objective state
$\mystate{o}$ as some subjective state $\mystate{s}$. In contrast, let
$\Sstrat_s(\mymessg{i} \probbar \mystate{s})$ be the probability that
a randomly sampled speaker chooses $\mymessg{i}$ if she experiences
some objective state $\mystate{o}$ as $\mystate{s}$. This is then the
hypothetical behavior of the speaker population in a (possibly
counterfactual) world without perceptual confusion. It is also what
the speakers actually do from their subjective point of view, and it
is this that changes in behavior (e.g., by imitation) will change in
the first place. Changes to $\Sstrat_s$ will, of course, also imply
changes in $\Sstrat$, as the latter is the result of what actually
happens when agents would like to do behave as in $\Sstrat_s$ in a
noisy world, as described by Equation~XYZ.

Similarly, $\Rstrat(\mystate{o} \probbar \messg)$ is the probability
that a randomly sampled receiver actually realizes objective state
$\mystate{o}$. $\Rstrat_s(\mystate{s} \probbar \messg)$ is the
probability of a subjectively intended interpretation $\mystate{s}$
that a randomly sampled receiver would pick in a noise-free
world. Again, behavior changes cause changes in $\Rstrat_s$, which in
turn influence $\Rstrat$.

Let $\EU(\messg)$ be short for $\EU(\messg,\state,\Rstrat)$ whenever
$\state$ and $\Rstrat$ are implicit, similarly for $\EU(\state)$ as a
shorthand for $\EU(\state,\messg,\Sstrat)$.

\paragraph{Standard assumptions.} Let's assume that we are in an
entirely noise-free world so that $\Sstrat = \Sstrat_s$ and $\Rstrat =
\Rstrat_s$. The standard assumptions for imitation-based revisions are
these. Agents get revision opportunities for all actual choice points
with equal probability (independently of how likely these would occur
in actual play). Take the case of a sender (receiver side is
parallel). If agent $A$ is given a revision opportunity for actual
state $\state$, she consults her own choice, which is $\mymessg{i}$
with probability $\Sstrat(\mymessg{i} \probbar \state)$, and observes
the behavior of one randomly picked other agent $B$, which is
$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
\state)$. $A$ also observes the expected utility of both choices (at
$\state$) and switches to $B$'s choice with probability $r_{ij} = [
\EU(\mymessg{j}) - \EU(\mymessg{i}) ]_+$, so that the overall
switching probability of an agent like $A$ who samples $\mymessg{i}$
for $\state$ is $p_{ij} = \Sstrat(\mymessg{j} \probbar \state) \
r_{ij}$.

\paragraph{Imitation under perceptual noise.} There are many
conceivable ways in which perceptual noise might influence the above
revision protocol. Our choice here is guided by theoretical and
practical considerations alike. We would like maintain the assumption
that agents in evolutionary models are not necessarily fully aware of
their surroundings. In particular, we assume here that agents are not
themselves aware of perceptual noise. Switching probabilities are
therefore to be defined in terms of actual expected utility (and not
subjective expectations about actual expected utility that take noise
into consideration). Moreover, we assume that agents who have
(unconscious) access to expected utilities for actual states
$\mystate{o}$ will confuse whatever subjective representation
$\mystate{s}$ they may have had during their revision opportunity for
the actual state $\mystate{o}$.

Concretely, suppose agent $A$ has a revision opportunity for objective
state $\mystate{o}$, and chooses $\mymessg{i}$ with probability
$\Sstrat(\mymessg{i} \probbar \mystate{o})$. Similarly, $B$ chooses
$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
\mystate{o})$. It may be that $A$ perceived $\mystate{o}$ as
$\mystate{s_A}$ and that $B$ perceived $\mystate{s_B}$, but if $A$ is
unaware of noise he would see expect and see no difference. A observes
(as usual) the actual expected utilities of $\mymessg{i}$ and
$\mymessg{j}$ for the actual state $\mystate{o}$. Consequently, he
updates his subjective behavior for $\mystate{o}$ (even if he has
actually chosen $\mymessg{i}$ since he perceived
$\mystate{s_A}$). This myopic revision behavior gives rise to expected
changes in the average subjective/noise-free behavior $\Sstrat_s$ as
follows:
\begin{align*}
  \dot{\Sstrat}_s(\mymessg{i} \probbar \state) & = \sum_j
  \Sstrat(\mymessg{j} \probbar \state) \cdot p_{ji} -  
  \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j p_{ij} \\
  & = \sum_j  \Sstrat(\mymessg{j} \probbar \state) \cdot
    \Sstrat(\mymessg{i} \probbar \state) \cdot [\EU(\mymessg{i}) -
    \EU(\mymessg{j})]_+ \\ 
   & \ \ \ \  - \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j
   \Sstrat(\mymessg{j} \probbar \state) \cdot [\EU(\mymessg{j}) -
    \EU(\mymessg{i})]_+ \\
   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left ( \sum_j
     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{i}) - \sum_j
     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{j})
   \right ) \\
   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left (
     \EU(\mymessg{i}) - \text{average EU at $\state$} \right )
\end{align*}
This is the (continuous-time) replicator dynamics for behavioral
strategies.

Similar arguments let us derive the receiver part. Suppose that agent
$A$ has a revision opportunity for choice point $\messg$. He and a
random other agent $B$ actually choose $\mystate{o_A}$ and
$\mystate{o_B}$ with probability $\Rstrat(\mystate{o_{A,B}} \probbar
\messg)$, even though they may have intended to realize different
interpretations $\mystate{s_A}$ and $\mystate{s_B}$ coming from
$\Rstrat_s(\mystate{o_{A,B}} \probbar \messg)$, but then
noise-perturbed. Agent $A$ observes the actual expected utility for
the actually realized interpretations and, by assumptions, believes
(myopically) that he had actually wanted to realize $\mystate{o_A}$
and that $B$ had wanted to realize $\mystate{o_B}$. Under these
assumptions, the derivation of the replicator dynamic for subjectively
intended (noise-free) receiver behavior $\Rstrat_s$ proceeds in
parallel with the above sender case.

At each time step, the noise-free behavior of agents is, of course,
perturbed by noise. This is why, after each update step, the \rdd
defined in Equation~XYZ perturbes the outcome of the replicator step
with perceptual noise.

\printbibliography[heading=bibintoc]

\end{document}

### 2015-04-03 paper/derivation_RDD/derivation_RDD.tex
derivation RDD minor edits

@@
 
-We derive the proposed \rdd as the mean field of an imitative revision
-protocol, following the example 4.3.1 of
-\citet{Sandholm2010:Population-Game}. Our derivation relies on the
-usual assumptions about conditional imitation at the agent level, but
-must add special assumptions about how noise affects revision
-probabilities. In keeping with the idea that agents in evolutionary
-models are largely myopic, we assume here that noise is real but that
-agents are unaware of it. When agents perceive the actual expected
-utility of their choice and of another agent's choice (as is usual in
-the motivation of imitation-based revision protocols), we assume
-additionally that agents confuse their possibly non-veridical
-subjective representations of states with the actual ones (details
+We derive the proposed \rdd as a mean dynamic of an imitative revision
+protocol. For expository reasons we follow closely the example 4.3.1
+of \citet{Sandholm2010:Population-Game}, but other possibilities
+exist. Our derivation relies on the usual assumptions about
+conditional imitation at the agent level, but must add special
+assumptions about how noise affects revision probabilities. In keeping
+with the idea that agents are of possibly limited cognitive
+capacities, we assume here that noise is real but that agents are
+unaware of it. When agents perceive the actual expected utility of
+their choice and of another agent's choice (as is usual in the
+motivation of imitation-based revision protocols), we assume
+additionally that agents mistake their possibly non-veridical
+subjective representations of states for the actual ones (details
 below).
@@
 a randomly sampled speaker chooses $\mymessg{i}$ if she experiences
-some objective state $\mystate{o}$ as $\mystate{s}$. This is then the
-hypothetical behavior of the speaker population in a (possibly
-counterfactual) world without perceptual confusion. It is also what
-the speakers actually do from their subjective point of view, and it
-is this that changes in behavior (e.g., by imitation) will change in
-the first place. Changes to $\Sstrat_s$ will, of course, also imply
-changes in $\Sstrat$, as the latter is the result of what actually
-happens when agents would like to do behave as in $\Sstrat_s$ in a
-noisy world, as described by Equation~XYZ.
+some objective state $\mystate{o}$ as $\mystate{s}$. This is then also
+the average sender behavior in a (possibly counterfactual) world
+without perceptual confusion. It is also where behavioral changes
+first take effect, to then percolate to $\Sstrat$ as well.
 
 Similarly, $\Rstrat(\mystate{o} \probbar \messg)$ is the probability
-that a randomly sampled receiver actually realizes objective state
+that a randomly sampled receiver chooses interpretation
 $\mystate{o}$. $\Rstrat_s(\mystate{s} \probbar \messg)$ is the
-probability of a subjectively intended interpretation $\mystate{s}$
-that a randomly sampled receiver would pick in a noise-free
-world. Again, behavior changes cause changes in $\Rstrat_s$, which in
-turn influence $\Rstrat$.
+probability of a subjectively intended interpretation
+$\mystate{s}$. Again, behavioral changes first target $\Rstrat_s$ but
+cause changes in $\Rstrat$ as well.
 
@@
 revision protocol. Our choice here is guided by theoretical and
-practical considerations alike. We would like maintain the assumption
-that agents in evolutionary models are not necessarily fully aware of
-their surroundings. In particular, we assume here that agents are not
-themselves aware of perceptual noise. Switching probabilities are
-therefore to be defined in terms of actual expected utility (and not
-subjective expectations about actual expected utility that take noise
-into consideration). Moreover, we assume that agents who have
-(unconscious) access to expected utilities for actual states
-$\mystate{o}$ will confuse whatever subjective representation
-$\mystate{s}$ they may have had during their revision opportunity for
-the actual state $\mystate{o}$.
+practical considerations alike. We would like maintain the common
+assumption that agents are ideal optimizers and not necessarily fully
+aware of their surroundings. In particular, we assume here that agents
+are not aware of perceptual noise. Switching probabilities are
+therefore defined in terms of actual expected utility (and not
+subjective expectations that take noise into consideration). Moreover,
+we assume that agents who have (unconscious) access to expected
+utilities for actual states $\mystate{o}$ will confuse whatever
+subjective representation $\mystate{s}$ they may have had during their
+revision opportunity for the actual state $\mystate{o}$.
 
@@
 \mystate{o})$. It may be that $A$ perceived $\mystate{o}$ as
-$\mystate{s_A}$ and that $B$ perceived $\mystate{s_B}$, but if $A$ is
-unaware of noise he would see expect and see no difference. A observes
-(as usual) the actual expected utilities of $\mymessg{i}$ and
-$\mymessg{j}$ for the actual state $\mystate{o}$. Consequently, he
-updates his subjective behavior for $\mystate{o}$ (even if he has
-actually chosen $\mymessg{i}$ since he perceived
-$\mystate{s_A}$). This myopic revision behavior gives rise to expected
-changes in the average subjective/noise-free behavior $\Sstrat_s$ as
-follows:
+$\mystate{s_A}$ and that $B$ perceived $\mystate{s_B}$, but since $A$
+is unaware of noise he expects no difference. $A$ observes (as usual)
+the actual expected utilities of $\mymessg{i}$ and $\mymessg{j}$ for
+$\mystate{o}$. Consequently, he updates his subjective behavior for
+$\mystate{o}$ (not being aware that the perceived $\mystate{s_A}$
+might have been different). This myopic revision behavior gives rise
+to expected changes in the subjective/noise-free choice probabilities
+$\Sstrat_s(\mymessg{i} \probbar \state)$ as follows:
 \begin{align*}
@@
 
-Similar arguments let us derive the receiver part. Suppose that agent
+Similar considerations motivate the receiver part. Suppose that agent
 $A$ has a revision opportunity for choice point $\messg$. He and a
@@
 $\mystate{o_B}$ with probability $\Rstrat(\mystate{o_{A,B}} \probbar
-\messg)$, even though they may have intended to realize different
-interpretations $\mystate{s_A}$ and $\mystate{s_B}$ coming from
-$\Rstrat_s(\mystate{o_{A,B}} \probbar \messg)$, but then
-noise-perturbed. Agent $A$ observes the actual expected utility for
-the actually realized interpretations and, by assumptions, believes
-(myopically) that he had actually wanted to realize $\mystate{o_A}$
-and that $B$ had wanted to realize $\mystate{o_B}$. Under these
-assumptions, the derivation of the replicator dynamic for subjectively
-intended (noise-free) receiver behavior $\Rstrat_s$ proceeds in
-parallel with the above sender case.
+\messg)$, even though they may have intended interpretations
+$\mystate{s_A}$ and $\mystate{s_B}$. Agent $A$ observes the actual
+expected utility and, by assumption, believes (myopically) that he had
+wanted to realize $\mystate{o_A}$ and that $B$ had wanted to realize
+$\mystate{o_B}$. Under these assumptions, the derivation of the
+replicator dynamic for subjectively intended (noise-free) receiver
+behavior $\Rstrat_s$ proceeds in parallel with the above sender case.
 

### 2015-04-08 paper/version_03/paper.tex
Moved Section 4 to appendix.

--- 2015-01-27 paper/version_02/paper.tex
@@
 presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
-introduces the replicator diffusion dynamic and Section~\ref{sec:diffusion-as-special}
-elaborates on its relation with the replicator mutator
-dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
+introduces the replicator diffusion dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
 diffusion dynamic. Section~\ref{sec:discussion} reflects and compares
 our approach to others.
+In Appendix~\ref{sec:diffusion-as-special} we give a formal elaboration on the relation between the replicator diffusion dynamic and the replicator mutator dynamic.
+This finding is not necessary to understand the main points of this paper regarding vagueness, hence its inclusion in an appendix.
+We decided to nevertheless include it since it is related, conceptually interesting, and potentially valuable for future work related to the model introduced in this paper.
 
@@
 mixed strategies and an interpretation as biological evolution as
-well. In Section~\ref{sec:diffusion-as-special}, we
+well. In Appendix~\ref{sec:diffusion-as-special}, we
 translate our initial formulation of the \rdd
@@
 
-The following section works out the relation between the \rdd and the \rmd.
-Although this section is conceptually important, it is also rather
-technical and not strictly necessary to appreciate the main point of
-the paper.
+%The following section works out the relation between the \rdd and the \rmd.
+%Although this section is conceptually important, it is also rather
+%technical and not strictly necessary to appreciate the main point of
+%the paper.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Diffusion as mutation}
-\label{sec:diffusion-as-special}
-
-\subsection{The replicator mutator dynamic}
-\label{sec:repl-mutat-dynam}
-
-The \rmd has first been proposed as a tool for studying the evolution
-of grammar
-\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
-adds stochastic mutation to the replicator dynamic. The relation of
-the \rmd to other prominent evolutionary dynamics is well understood
-\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
-fruitful applications in the context of signaling games
-\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}.
-
-The usual formulation of the \rmd is based on mixed, not behavioral
-strategies. But a discrete-time formulation of the \rmd in behavioral
-strategies clearly identifies the \rdd to be a special case. It
-follows that a mixed strategy version of the \rdd is just the \rmd
-with suitably defined mutation matrices. From this, even a
-continuous-time version can easily be retrieved if desired. To show
-this, we start with a delineation of mixed and behavioral strategies.
-
-\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
-functions $\Spure \in \Messgs^\States$ ($\Rpure \in
-\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
-$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
-(\Acts^\Messgs)$). The latter give the relative population frequencies
-of the former. We write $\Smixed_i$ for the frequency
-$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
-receiver.
-
-Every mixed strategy $\Smixed$ converts to a unique behavioral
-strategy defined by:
-\begin{align}
-  \label{eq:CorrespondenceBehavioralMixed}
-  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
-\end{align} 
-Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
-is \emph{not} an injection, as many mixed strategies map onto the same
-behavioral strategy. In this sense, mixed strategies hold more
-information than behavioral strategies about the distribution of pure strategies in a population.
-
-In the context of evolutionary dynamics for signaling games (where
-each agent only chooses once), we can think of behavioral strategies
-as treating each choice point as an independent update site
-\citep[e.g.][]{Cressman2003:Evolutionary-Dy}. Dynamics for mixed
-strategies conceive of agents as adjusting their entire pure
-strategy. In this way, dependencies between different choices at
-different choice points can matter to the evolutionary path. In
-contrast, dynamics for behavioral strategies conceive of agents as
-adjusting their strategies locally at each choice point, regardless of
-what they are disposed to do at other choice points.
-
-This can matter to evolutionary dynamics. For illustration, take two
-sender strategies whose behavior is different only at two
-states. Compared to the second strategy, the first scores a much
-higher expected utility at one state, but a slightly lower at the
-other, so that the total expected utility of the first strategy is
-higher. Under evolutionary dynamics that work on entire strategies the
-first strategy will increase in proportion, relative to the
-second. Consequently, a locally less efficient type of behavior can be
-promoted because it is correlated with efficient behavior at another
-choice point. Evolutionary dynamics working on behavioral strategies
-do not allow for that. Behavioral adaptations occur independently at
-different choice points.
-
-As a result, dynamics on behavioral strategies are more coarse-grained
-and not necessarily equivalent to dynamics on mixed strategies. An
-advantage of dynamics in behavioral strategies is that they reduce
-complexity, which helps with numerical simulations. They might also be
-a more natural choice for cultural dynamics, e.g., when an agent
-observes and imitates another agent's strategy: it may be more
-plausible that agents adopt the use of a single word, not an entire
-lexicon, so to speak. On the other hand, biological evolution seems
-more naturally modeled in terms of mixed strategies whenever it is a
-full set of behavioral disposition that is bequeathed to the next
-generation.
-
-\paragraph{Replicator dynamic in mixed strategies.} To understand the
-\rmd in mixed strategies, we first introduce the \rd in mixed
-strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
-$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
-$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
-F_k^{\Rmixed}$ is the average fitness in the sender population and
-$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
-average fitness in the receiver population.
-
-% The continuous-time replicator dynamic defines the change of frequency
-% of mixed strategies. Its so-called two-population version, that takes
-% sender and receiver roles as separate update sites is:
-% \begin{align*}
-%   \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
-%   \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
-%   \Phi(\Rmixed,\Smixed) \right ) \,.
-% \end{align*}
-Similar to the version for behavioral strategies, a discrete-time
-formulation of the \rd for mixed strategies can be conceptualized as an
-update function that maps mixed strategies $\Smixed$ and $\Rmixed$
-onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
-such that:
-\begin{align*}
-  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
-    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
-    F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
-\end{align*}
-
-\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
-extends the \rd by adding probabilistic mutation. Let $Q$ be a
-row-stochastic mutation matrix where $Q_{ji}$ gives the probability
-that pure sender strategy $\Spure_j$ mutates into
-$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
-where $R_{ji}$ gives the probability that pure receiver strategy
-$\Rpure_j$ mutates into $\Rpure_i$.
-
-The \rmd is usually given only in its continuous-time form. % A
-% two-population (non-payoff adjusted) formulation is then:
-% \begin{align*}
-%   \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
-%     \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
-%     \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
-%     \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
-% \end{align*}
-But a discrete-time, two-population formulation of the \rmd is
-available too \citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
-\begin{align*}
-  \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
-    F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
-  \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
-    \Phi(\Rmixed,\Smixed)}\,.
-\end{align*}
-The discrete-time \rmd has the same sequential nature as the \rdd:
-first we compute the fitness-driven change according to the standard
-replicator dynamic; then we compute the perturbation from
-mutation. This becomes clear if we define independent mutation
-functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
-of (one round of) mutation:
-\begin{align}
-  \label{eq:Mutation}
-  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
-  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
-  R_{ji} \,.
-\end{align}
-The discrete-time \rmd given above can then be rewritten as:
-\begin{align*}
-  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,. 
-\end{align*}
-
-\subsection{Relating the dynamics}
-
-The sequential reformulation of the \rmd closely resembles our initial
-formulation of the \rdd. However, there is a major difference between
-these. While the \rdd is a dynamic for behavioral strategies, the
-\rmd, as taken from the literature, applies to mixed
-strategies. Similarly, the confusability of states is an operation
-that is straightforwardly definable for behavioral strategies (see
-Equation~(\ref{eq:confusion-function})), while mutation in the \rmd is
-defined as mutation from one pure strategy to another. Nonetheless, it
-is possible to map each confusion matrix to a mutation matrix in a
-natural way so that the correspondence between mixed and behavioral
-strategies are conserved. This is the main result presented in this
-section, and it is our justification for regarding the \rdd as the
-behavioral-strategy analogue of the replicator-mutator dynamic when
-the only source of mutation is confusability of states.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[node distance = 1cm]
-
-    \begin{scope}
-      \node[] (mixed) {$\Smixed$};
-
-      \node[below of = mixed] (mutated) {$\Mutate_{Q^C}(\Smixed)$};
-
-      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};
-
-      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};
-
-      \draw[->, dotted, thick] (mixed) -> (mutated);
-
-      \begin{pgfonlayer}{background}
-        \node [draw=black!50, fill=black!10,fit = (dummyl) 
-        (mutated) (dummyr)] (mixedSpace) {};
-      \end{pgfonlayer}
-
-      \node[above of = mixedSpace, node distance = 1.4cm] {space of
-        mixed strategies};
-    \end{scope}
-
-    \begin{scope}[xshift=6cm]
-      \node[] (behavioral) {$\Sstrat$};
-
-      \node[below of = behavioral] (mutatedBeh) {$\Diff_C(\Sstrat)$};
-
-      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};
-
-      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};
-
-      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);
-
-      \begin{pgfonlayer}{background}
-        \node [draw=black!50, fill=black!10,
-              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
-      \end{pgfonlayer}
-
-      \node[above of = behSpace, node distance = 1.4cm] {space of
-        behavioral strategies};
-    \end{scope}
-
-    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
-    above]{$G(\Smixed)$};
-
-    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(\Mutate_{Q^C}(\Smixed))$};
-
-  \end{tikzpicture}
-
-  \caption{Correspondence between state-confusion and mutation.}
-  \label{fig:correspondence-result}
-\end{figure}
-
-More concretely, given the non-injective mapping $G$ from mixed to
-behavioral strategies, we show that there is a systematic translation
-from a confusion matrix $C$ to a mutation matrix $Q^C$ such that for
-each mixed sender strategy $\Smixed$, its unique corresponding
-behavioral strategy $\Sstrat = G(\Smixed)$ has a
-confusion-perturbation $\Diff_C(\Sstrat)$ that corresponds, via $G$,
-to the mutation-perturbation $\Mutate_{Q^C}(\Smixed)$ (see
-Figure~\ref{fig:correspondence-result}). In other words, we
-hypothesize that a confusion matrix $C$ should give rise to a unique
-mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
-have $G(\Mutate_{Q^C}(\Smixed)) = \Diff_C(\Sstrat)$. Similarly, for
-the receiver.
-%This is the content of Theorem~\ref{thm:Correspondence} given below, a
-%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
-%section introduces the relevant construction that maps arbitrary $C$
-%to suitable $Q^C$.
-
-\paragraph{Confusion-based mutations.} There are natural conversions
-of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
-we start with that.
-
-The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
-diffusion with $C$ is the product of the probabilities, for each
-$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
-$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
-states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
-$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
-\begin{align}
-  \label{eq:construction-rec}
-  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
-\end{align}
-
-Now look at the sender. The probability that $\Spure_j$ is realized as
-$\Spure_i$ under diffusion with $C$ is the product of the
-probabilities, over all states $\mystate{k}$, that the message
-$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
-$\mystate{k}$ in the absence of noise, is produced by a noisy
-realization of $\Spure_j$, which is the probability $\sum_{\state_l
-  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
-$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
-would map unto $\Spure_i(\mystate{k})$. So, define:
-\begin{align}
-  \label{eq:construction-sen}
-  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
-    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
-\end{align}
-
-Based on Equations~(\ref{eq:construction-rec}) and
-(\ref{eq:construction-sen}), we can formulate the desired
-correspondence result, a proof for which is given in Appendix~\ref{sec:proofs}.
-
-\begin{theorem}
-  \label{thm:Correspondence}
-  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
-  \Diff_C(\Sstrat)$. (ii) If $G(\Rmixed) = \Rstrat$, then
-  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
-\end{theorem}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The previous sections showed that the \rdd can be conceptualized as a
+The previous section argued that the \rdd can be conceptualized as a
 two-step procedure: we first calculate where the replicator dynamic
@@
 the \rmd with mutation matrices derived from confusion matrices, as
-shown in Section~\ref{sec:diffusion-as-special}.
+shown in Appendix~\ref{sec:diffusion-as-special}.
 
@@
 
-\section{Proof of Theorem}
-\label{sec:proofs}
+\section{Diffusion as mutation}
+\label{sec:diffusion-as-special}
+
+\subsection{The replicator mutator dynamic}
+\label{sec:repl-mutat-dynam}
+
+The \rmd has first been proposed as a tool for studying the evolution
+of grammar
+\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
+adds stochastic mutation to the replicator dynamic. The relation of
+the \rmd to other prominent evolutionary dynamics is well understood
+\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
+fruitful applications in the context of signaling games
+\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}.
+
+The usual formulation of the \rmd is based on mixed, not behavioral
+strategies. But a discrete-time formulation of the \rmd in behavioral
+strategies clearly identifies the \rdd to be a special case. It
+follows that a mixed strategy version of the \rdd is just the \rmd
+with suitably defined mutation matrices. From this, even a
+continuous-time version can easily be retrieved if desired. To show
+this, we start with a delineation of mixed and behavioral strategies.
+
+\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
+functions $\Spure \in \Messgs^\States$ ($\Rpure \in
+\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
+$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
+(\Acts^\Messgs)$). The latter give the relative population frequencies
+of the former. We write $\Smixed_i$ for the frequency
+$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
+receiver.
+
+Every mixed strategy $\Smixed$ converts to a unique behavioral
+strategy defined by:
+\begin{align}
+  \label{eq:CorrespondenceBehavioralMixed}
+  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
+\end{align}
+Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
+is \emph{not} an injection, as many mixed strategies map onto the same
+behavioral strategy. In this sense, mixed strategies hold more
+information than behavioral strategies about the distribution of pure strategies in a population.
+
+In the context of evolutionary dynamics for signaling games (where
+each agent only chooses once), we can think of behavioral strategies
+as treating each choice point as an independent update site
+\citep[e.g.][]{Cressman2003:Evolutionary-Dy}. Dynamics for mixed
+strategies conceive of agents as adjusting their entire pure
+strategy. In this way, dependencies between different choices at
+different choice points can matter to the evolutionary path. In
+contrast, dynamics for behavioral strategies conceive of agents as
+adjusting their strategies locally at each choice point, regardless of
+what they are disposed to do at other choice points.
+
+This can matter to evolutionary dynamics. For illustration, take two
+sender strategies whose behavior is different only at two
+states. Compared to the second strategy, the first scores a much
+higher expected utility at one state, but a slightly lower at the
+other, so that the total expected utility of the first strategy is
+higher. Under evolutionary dynamics that work on entire strategies the
+first strategy will increase in proportion, relative to the
+second. Consequently, a locally less efficient type of behavior can be
+promoted because it is correlated with efficient behavior at another
+choice point. Evolutionary dynamics working on behavioral strategies
+do not allow for that. Behavioral adaptations occur independently at
+different choice points.
+
+As a result, dynamics on behavioral strategies are more coarse-grained
+and not necessarily equivalent to dynamics on mixed strategies. An
+advantage of dynamics in behavioral strategies is that they reduce
+complexity, which helps with numerical simulations. They might also be
+a more natural choice for cultural dynamics, e.g., when an agent
+observes and imitates another agent's strategy: it may be more
+plausible that agents adopt the use of a single word, not an entire
+lexicon, so to speak. On the other hand, biological evolution seems
+more naturally modeled in terms of mixed strategies whenever it is a
+full set of behavioral disposition that is bequeathed to the next
+generation.
+
+\paragraph{Replicator dynamic in mixed strategies.} To understand the
+\rmd in mixed strategies, we first introduce the \rd in mixed
+strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
+$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
+$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
+F_k^{\Rmixed}$ is the average fitness in the sender population and
+$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
+average fitness in the receiver population.
+
+% The continuous-time replicator dynamic defines the change of frequency
+% of mixed strategies. Its so-called two-population version, that takes
+% sender and receiver roles as separate update sites is:
+% \begin{align*}
+%   \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
+%   \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
+%   \Phi(\Rmixed,\Smixed) \right ) \,.
+% \end{align*}
+Similar to the version for behavioral strategies, a discrete-time
+formulation of the \rd for mixed strategies can be conceptualized as an
+update function that maps mixed strategies $\Smixed$ and $\Rmixed$
+onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
+such that:
+\begin{align*}
+  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
+    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
+    F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
+\end{align*}
+
+\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
+extends the \rd by adding probabilistic mutation. Let $Q$ be a
+row-stochastic mutation matrix where $Q_{ji}$ gives the probability
+that pure sender strategy $\Spure_j$ mutates into
+$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
+where $R_{ji}$ gives the probability that pure receiver strategy
+$\Rpure_j$ mutates into $\Rpure_i$.
+
+The \rmd is usually given only in its continuous-time form. % A
+% two-population (non-payoff adjusted) formulation is then:
+% \begin{align*}
+%   \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
+%     \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
+%     \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
+%     \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
+% \end{align*}
+But a discrete-time, two-population formulation of the \rmd is
+available too \citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
+\begin{align*}
+  \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
+    F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
+  \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
+    \Phi(\Rmixed,\Smixed)}\,.
+\end{align*}
+The discrete-time \rmd has the same sequential nature as the \rdd:
+first we compute the fitness-driven change according to the standard
+replicator dynamic; then we compute the perturbation from
+mutation. This becomes clear if we define independent mutation
+functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
+of (one round of) mutation:
+\begin{align}
+  \label{eq:Mutation}
+  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
+  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
+  R_{ji} \,.
+\end{align}
+The discrete-time \rmd given above can then be rewritten as:
+\begin{align*}
+  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,.
+\end{align*}
+
+\subsection{Relating the dynamics}
+
+The sequential reformulation of the \rmd closely resembles our initial
+formulation of the \rdd. However, there is a major difference between
+these. While the \rdd is a dynamic for behavioral strategies, the
+\rmd, as taken from the literature, applies to mixed
+strategies. Similarly, the confusability of states is an operation
+that is straightforwardly definable for behavioral strategies (see
+Equation~(\ref{eq:confusion-function})), while mutation in the \rmd is
+defined as mutation from one pure strategy to another. Nonetheless, it
+is possible to map each confusion matrix to a mutation matrix in a
+natural way so that the correspondence between mixed and behavioral
+strategies are conserved. This is the main result presented in this
+section, and it is our justification for regarding the \rdd as the
+behavioral-strategy analogue of the replicator-mutator dynamic when
+the only source of mutation is confusability of states.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[node distance = 1cm]
+
+    \begin{scope}
+      \node[] (mixed) {$\Smixed$};
+
+      \node[below of = mixed] (mutated) {$\Mutate_{Q^C}(\Smixed)$};
+
+      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};
+
+      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};
+
+      \draw[->, dotted, thick] (mixed) -> (mutated);
+
+      \begin{pgfonlayer}{background}
+        \node [draw=black!50, fill=black!10,fit = (dummyl)
+        (mutated) (dummyr)] (mixedSpace) {};
+      \end{pgfonlayer}
+
+      \node[above of = mixedSpace, node distance = 1.4cm] {space of
+        mixed strategies};
+    \end{scope}
+
+    \begin{scope}[xshift=6cm]
+      \node[] (behavioral) {$\Sstrat$};
+
+      \node[below of = behavioral] (mutatedBeh) {$\Diff_C(\Sstrat)$};
+
+      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};
+
+      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};
+
+      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);
+
+      \begin{pgfonlayer}{background}
+        \node [draw=black!50, fill=black!10,
+              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
+      \end{pgfonlayer}
+
+      \node[above of = behSpace, node distance = 1.4cm] {space of
+        behavioral strategies};
+    \end{scope}
+
+    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
+    above]{$G(\Smixed)$};
+
+    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(\Mutate_{Q^C}(\Smixed))$};
+
+  \end{tikzpicture}
+
+  \caption{Correspondence between state-confusion and mutation.}
+  \label{fig:correspondence-result}
+\end{figure}
+
+More concretely, given the non-injective mapping $G$ from mixed to
+behavioral strategies, we show that there is a systematic translation
+from a confusion matrix $C$ to a mutation matrix $Q^C$ such that for
+each mixed sender strategy $\Smixed$, its unique corresponding
+behavioral strategy $\Sstrat = G(\Smixed)$ has a
+confusion-perturbation $\Diff_C(\Sstrat)$ that corresponds, via $G$,
+to the mutation-perturbation $\Mutate_{Q^C}(\Smixed)$ (see
+Figure~\ref{fig:correspondence-result}). In other words, we
+hypothesize that a confusion matrix $C$ should give rise to a unique
+mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
+have $G(\Mutate_{Q^C}(\Smixed)) = \Diff_C(\Sstrat)$. Similarly, for
+the receiver.
+%This is the content of Theorem~\ref{thm:Correspondence} given below, a
+%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
+%section introduces the relevant construction that maps arbitrary $C$
+%to suitable $Q^C$.
+
+\paragraph{Confusion-based mutations.} There are natural conversions
+of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
+we start with that.
+
+The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
+diffusion with $C$ is the product of the probabilities, for each
+$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
+$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
+states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
+$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
+\begin{align}
+  \label{eq:construction-rec}
+  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
+\end{align}
+
+Now look at the sender. The probability that $\Spure_j$ is realized as
+$\Spure_i$ under diffusion with $C$ is the product of the
+probabilities, over all states $\mystate{k}$, that the message
+$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
+$\mystate{k}$ in the absence of noise, is produced by a noisy
+realization of $\Spure_j$, which is the probability $\sum_{\state_l
+  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
+$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
+would map unto $\Spure_i(\mystate{k})$. So, define:
+\begin{align}
+  \label{eq:construction-sen}
+  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
+    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
+\end{align}
+
+Based on Equations~(\ref{eq:construction-rec}) and
+(\ref{eq:construction-sen}), we can formulate the desired
+correspondence result:
+
+\begin{theorem}
+  \label{thm:Correspondence}
+  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
+  \Diff_C(\Sstrat)$. (ii) If $G(\Rmixed) = \Rstrat$, then
+  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
+\end{theorem}
+
+We can prove this theorem as follows.
+%\section{Proof of Theorem}
+%\label{sec:proofs}
 

### 2015-04-16 paper/version_03/paper.tex
Added placeholder figure for example of Voronoi language. Added footnote to alert user that "imprecision" and "impairment" are the same thing and the terms are used interchangeably.

@@
 points in the upper half; the receiver's interpretations of messages
-are the central points, .25 and .75, in the respective intervals.
+are the central points, .25 and .75, in the respective intervals
+(see Figure~\ref{fig:Voronoi-language-example} for an illustration).
+
+\begin{figure}
+  \centering
+  \missingfigure{Example of Voronoi language for unit interval}
+  \caption{}
+  \label{fig:Voronoi-language-example}
+\end{figure}
 
@@
 \end{align*}
-where $\impairment \ge 0$ is an imprecision or indiscriminability
-parameter. When $\impairment=0$ agents perfectly discriminate between
+where $\impairment \ge 0$ is an impairment, imprecision or indiscriminability
+parameter\footnote{We use these terms interchangeably throughout the paper.}.
+When $\impairment=0$ agents perfectly discriminate between
 states; when $\impairment \rightarrow \infty$ agents cannot

### 2015-04-22 paper/version_03/paper.tex
Added reaction and comment in paper about exploring a connection between RDD and generalized reinforcement learning.

@@
 the use of $\messg$ in $\state$ gave positive payoff, then not only
-will its future use probability be promoted at $\state$ but also at
+will the probability of its future use be promoted at $\state$ but also at
 other states, proportional to how similar these are to
@@
 strategies after each round of play under generalized reinforcement
-than under the Herrnstein variety. But that only explains the speed of
+than under the Herrnstein variety. That only explains the speed of
 adaptation, not necessarily that generalization also leads to
@@
 Diffusion of strategies in the \rdd can also be conceived of as a form
-of generalization, and works in large part quite analogous to stimulus
+of generalization, and works in large part quite analogously to stimulus
 generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
@@
 the behavior of (some forms of) reinforcement learning mirror those of
-the \rd (at some stage in time)
+the \rd at some stage in time
 (\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
@@
 abstraction.
+Although it would be interesting to explore the possibility of a mathematical connection between the two, this is beyond the scope of this paper and thus left as future work.
 

### 2015-05-03 paper/version_03/paper.tex
Added reaction to suggestion for more pictures. Updated reference to "Evolving perceptual categories" and updated the text accordingly, addressing point v) of first reviewer.

@@
 but have received particular attention in a more philosophical context
-for reasons that will become clear presently
-\citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag}.
+for reasons that will become clear presently.
 
@@
 points in the upper half; the receiver's interpretations of messages
-are the central points, .25 and .75, in the respective intervals
-(see Figure~\ref{fig:Voronoi-language-example} for an illustration).
+are the central points, .25 and .75, in the respective intervals.
+See Figure~\ref{fig:Voronoi-language-example} for an illustration.
 
@@
   \centering
-  \missingfigure{Example of Voronoi language for unit interval}
-  \caption{}
+  \missingfigure{Please include here something like Figure 4 in your paper with Elliot.}
+  \caption{Example of Voronoi language for the unit interval.}
   \label{fig:Voronoi-language-example}
@@
 prototypical exemplars belonging to them.
+Development of these ideas, and more, were carried out by \citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC}.
 
@@
 types of utility functions, for the limiting case. This was first
-observed by Elliott Wagner in unpublished work
-\citep[c.f.][]{OConnor2013:Evolving-Percep}. A full analysis of the
-dynamics of sim-max games is beyond the scope of this paper, but we
-will see shortly that diffusion from confusability of states clearly
+observed by Elliott Wagner, as mentioned by \citet{OConnor2014-OCOEPC}.
+A full analysis of the dynamics of sim-max games is beyond the scope of this paper, but we
+will shortly see that diffusion from confusability of states clearly
 prevents evolutionary paths that meander for a long time in the

### 2015-05-09 paper/version_03/paper.tex
Some minor tweaks. Added descriptions for sender/receiver strategies and expected utility function.

@@
 presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
-introduces the replicator diffusion dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
-diffusion dynamic. Section~\ref{sec:discussion} reflects and compares
+introduces the replicator diffusion dynamic, which is then explored in Section~\ref{sec:exploring-rdd}.
+Section~\ref{sec:discussion} reflects and compares
 our approach to others.
@@
 This finding is not necessary to understand the main points of this paper regarding vagueness, hence its inclusion in an appendix.
-We decided to nevertheless include it since it is related, conceptually interesting, and potentially valuable for future work related to the model introduced in this paper.
+We nevertheless include it since it is related, conceptually interesting, and potentially valuable for future work related to the model introduced here.
 
@@
 
-Similarity-maximizing (short: sim-max) games are generalizations of
-Lewis' games where different states are allowed to be more or less
+Similarity-maximizing (short: sim-max) games are a variation of
+Lewis' games where the receiver's actions are equated with the state
+space (one can think of the actions as choosing states) and different states are allowed to be more or less
 similar to one another. While Lewis' games treated communicative
@@
 prototypical exemplars belonging to them.
-Development of these ideas, and more, were carried out by \citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC}.
+These ideas, and more, are developed further by \citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC}, among others.
 
@@
 into one using mixed strategies. Doing so identifies the \rdd as a
-special case of the \rmd. For now, let us first focus on the
+special case of the \rmd.
+
+For now, let us first focus on the
 formulation of the dynamic in terms of behavioral strategies alone.
-
 We begin by recapitulating the \rd in
@@
 available choices. The sender's behavioral strategies are functions
-$\Sstrat \in \Delta(\Messgs)^\States$; the receiver's are functions
-$\Rstrat \in \Delta(\Acts)^\Messgs$. The expected utility of choices
+$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
+$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a probability of each state $\state \in \States$ being interpreted by the receiver given $\messg$.
+The expected utility of choices
 at each choice point is:
@@
 \end{align*}
+This means that the expected utility for the sender of choosing message $\messg \in \Messgs$, given a state $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability the receiver will choose act $\act$ given message $\messg$ times the actual utility of using act $\act$ for state $\state$.
+Conversely, the expected utility for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior probability of state $\state$ occurring, times the probability the sender will choose message $\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.
+
 The \emph{fitness} of a behavioral strategy at a choice point is the

### 2015-05-13 paper/version_03/paper.tex
Some more minor tweaks. Added descriptions for definitions of diffusion. Labelling within group measure figure using subfigures.

@@
 \rd in behavioral strategies. Fix a sim-max game with $\States =
-\Acts$ and a confusion matrix $C : \States \times \States$. $C$ is a
-row-stochastic matrix, i.e., each row is a probability
+\Acts$ and a confusion matrix $C : \States \times \States$, where $C$ is
+row-stochastic, i.e., each row is a probability
 distribution. Elements $C_{ij}$ give the probability that $\state_i$
 is realized as $\state_j$. The confusability of states affects senders
-and receivers alike, but in slightly different ways (see
-Figure~\ref{fig:noise-perturbation-of-strategies}). For the sender,
+and receivers alike, but in slightly different ways. For the sender,
 $C_{ij}$ is the probability that the actual state \mystate{i} is
@@
 formed when \mystate{i} is the intended interpretation.
+See Figure~\ref{fig:noise-perturbation-of-strategies} for an illustration.
 
@@
 row-stochastic matrices (rows are choice points, columns are choices;
-each row is a probability distribution), the diffusion effect of
-confusion matrix $C$ is just matrix multiplication:
+each row is a probability distribution), we can define the diffusion effect of
+confusion matrix $C$ as matrix multiplication:
 \begin{align}
@@
 \end{align}
+For example, diffusion thus defined determines that, on the diffused sender strategy, the probability that message $\messg$ is chosen given a state $\state_i$ is given by $\sum_{j} C_{ij} \cdot \Sstrat(\messg \probbar \state_j)$.
+This means that the original probability given by $\Sstrat$ is increased by the original probability of this message being chosen in other states, proportionally to how likely it is that they are mistaken for $\state_i$.
 
@@
 evolution.
-
-%The following section works out the relation between the \rdd and the \rmd.
-%Although this section is conceptually important, it is also rather
-%technical and not strictly necessary to appreciate the main point of
-%the paper.
+We worked out the relation between the \rdd and the \rmd, but given that it is rather technical and not strictly necessary to appreciate the main point of the paper, we defer it to Appendix~\ref{sec:diffusion-as-special}.
+In the following section we explore the long-term effects of using the \rdd.
 
@@
 Confusion of states should intuitively be a function of their
-perceptual similarity. To make this concrete, let us assume that the
+perceptual similarity: the more similar two states are, the more likely they could be mistaken for each other.
+To make this concrete, let us assume that the
 state space of the sim-max game consists of $n \ge 2$ states that are
@@
 $\impairment$. For $\impairment = 0$, the confusion matrix has $C_{ii}
-= 1$; for $\impairment > 0$, $C$ is positive, i.e., $C_{ij} >0$ for
-all $i$ and $j$; for $\impairment \rightarrow \infty$, we find $C_{ij}
-= \nicefrac{1}{\card{\States}}$ for all $i$ and $j$.
+= 1$ and $C_{ij} = 0$ for $i \neq j$. For $\impairment > 0$, $C$ is positive, i.e., $C_{ij} >0$ for
+all $i$ and $j$. For $\impairment \rightarrow \infty$, we find $C_{ij}
+\rightarrow \nicefrac{1}{\card{\States}}$ for all $i$ and $j$.
 
@@
 \citep[c.f.][]{OConnor2013:The-Evolution-o}.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
-    \label{fig:example_stratsA}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind3452.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
-    \label{fig:example_stratsB}
-  \end{subfigure}
-
-  % \begin{subfigure}[]{0.45\textwidth}
-  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
-  %   \label{fig:example_stratsC}
-  % \end{subfigure}
-  % \hfill
-  % \begin{subfigure}[]{0.45\textwidth}
-  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
-  %   \label{fig:example_stratsD}
-  % \end{subfigure}
-
-  \caption{Example strategies under \rdd at stopping time of our simulations.}
-  \label{fig:example_strats}
-\end{figure}
 
@@
 
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
+    \label{fig:example_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind3452.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
+    \label{fig:example_stratsB}
+  \end{subfigure}
+
+  % \begin{subfigure}[]{0.45\textwidth}
+  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
+  %   \label{fig:example_stratsC}
+  % \end{subfigure}
+  % \hfill
+  % \begin{subfigure}[]{0.45\textwidth}
+  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
+  %   \label{fig:example_stratsD}
+  % \end{subfigure}
+
+  \caption{Example strategies under \rdd at stopping time of our simulations.}
+  \label{fig:example_strats}
+\end{figure}
+
 Representative examples for resulting strategy pairs are given in
@@
 dynamically stable state under the \rd. Indeed, the trial was stopped
-after the maximum of 200 rounds. In contrast, the outcome of a trial
+for reaching the maximum of 200 rounds. In contrast, the outcome of a trial
 with identical parameters, except with imprecision $\impairment =
@@
 
-Figure~\ref{fig:MeanMetrics} shows plots summarizing the results of
-our simulation. For perspicuity, we only plot results for one level of
+Figure~\ref{fig:MeanMetrics} shows plots summarizing some results.
+%of our simulation.
+For perspicuity, we only plot results for one level of
 tolerance $\beta = 0.1$, and leave out the case of $n_s = 90$. Still,
-every qualitative trend we mention in the following applies to the
+every qualitative trend mentioned in the following applies to the
 whole set of results.
@@
 \emph{within group distance}, i.e., the average distances between all
-sender strategies obtained for the same parameter values, is plotted
-on the left in Figure~\ref{fig:GroupMeasures} for $\beta =
+sender strategies obtained for the same parameter values, is plotted in Figure~\ref{fig:GroupMeasuresA} for $\beta =
 0.1$ and $n_s = 10$. Despite some quantitative differences, the
@@
 
-    \includegraphics[width=0.475\textwidth]{plots/WithinGroupDistanceConcise.pdf}
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/WithinGroupDistanceConcise.pdf}
+    \caption{Within group distance}
+    \label{fig:GroupMeasuresA}
+  \end{subfigure}
     \hfill
-    \includegraphics[width=0.475\textwidth]{plots/WithinGroupEUConcise.pdf}
-
-    \caption{Within group distance (left) and within group expected
-      utility (right) as a function of impairment for all runs with
-      $\toler = 0.1$ and $n_s = 10$.}
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/WithinGroupEUConcise.pdf}
+    \caption{Within group expected utility}
+    \label{fig:GroupMeasuresB}
+  \end{subfigure}
+
+    \caption{Within group measures for all runs with $\toler = 0.1$ and $n_s = 10$.}
   \label{fig:GroupMeasures}
@@
 higher \emph{within group expected utility}, defined as the average
-expected utility that each evolved language (i.e., pair of sender and
-receiver strategy at stopping time) scored when playing against an
-arbitrary other language obtained for the same parameter values. The
-right side of Figure~\ref{fig:GroupMeasures} gives a representative
+expected utility that each evolved language scored when playing against an
+arbitrary other language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a representative
 example. The observation repeats for other parameter values: while

### 2015-05-14 paper/version_03/paper.tex
PHoS paper Derivation RDD added; major edits and shortening relation to other accounts; figure Voronoi-language added

@@
 
-\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
+\newcommand{\imprecision}{\ensuremath{\alpha}} % imprecision
 \newcommand{\toler}{\ensuremath{\beta}} % tolerance
@@
 \begin{abstract}
-  Signaling games are popular models for studying the evolution of
-  meaning, but typical approaches do not incorporate vagueness as a
-  feature of successful signaling.  Complementing recent like-minded
-  models, we introduce the replicator diffusion dynamic as a special
-  case of the replicator mutator dynamic that combines evolutionary
-  pressure on optimal signal use with stochastic noise in the
-  processing of similar stimuli. Applying this new dynamic to a
-  generalization of Lewis' signaling games, we show that stochastic
-  imprecision leads to vague, yet by-and-large communicative efficient
-  signal use, and, moreover, that it unifies evolutionary outcomes and helps
-  avoid suboptimal categorization.
+  Signaling games are popular models for studying the evolution of meaning, but typical
+  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
+  recent like-minded models, we introduce the replicator diffusion dynamic that combines
+  evolutionary pressure on optimal signal use with stochastic noise in the processing of
+  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
+  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
+  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal categorization.
 \end{abstract}
@@
 
-A number of authors have recently tried to explain why vagueness
-evolved as something that is itself useful
+A number of authors have recently tried to explain why vagueness evolved as something that is
+itself useful
 \citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
-Others have argued that vagueness is a natural byproduct of
-limitations in information processing or particular learning
-strategies
-\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This
-paper contributes to the latter line of thought. Concretely, we
-introduce the replicator diffusion dynamics---a novel variant of the
-replicator mutator dynamic---that integrates stochastic noise on the
-differential confusability of similar stimuli. The dynamic proposed
-here generalizes and complements previous like-minded accounts. We
-show that stochastic noise can not only lead to vague, yet relatively
-efficient signal use, but also unify evolutionary outcomes and
-help avoid suboptimal categorization.
-
-The next section introduces the background against which the work
-presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
-introduces the replicator diffusion dynamic, which is then explored in Section~\ref{sec:exploring-rdd}.
-Section~\ref{sec:discussion} reflects and compares
-our approach to others.
-In Appendix~\ref{sec:diffusion-as-special} we give a formal elaboration on the relation between the replicator diffusion dynamic and the replicator mutator dynamic.
-This finding is not necessary to understand the main points of this paper regarding vagueness, hence its inclusion in an appendix.
-We nevertheless include it since it is related, conceptually interesting, and potentially valuable for future work related to the model introduced here.
+Others have argued that vagueness is a natural byproduct of limitations in information
+processing or particular learning strategies
+\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This paper
+contributes to the latter line of thought. Concretely, we introduce the replicator diffusion
+dynamics that integrates stochastic noise on the differential confusability of similar
+stimuli. The dynamic proposed here generalizes and complements previous like-minded
+accounts. We show that stochastic noise can not only lead to vague, yet relatively efficient
+signal use, but can also unify evolutionary outcomes and help avoid suboptimal categorization.
+
+The next section introduces the background against which the work presented here can be
+appreciated. Section~\ref{sec:repl-diff-dynam} introduces the replicator diffusion dynamic,
+which is then explored in Section~\ref{sec:exploring-rdd}.  Section~\ref{sec:discussion}
+reflects and compares our approach to others. Two appendices cover technical material that is
+not strictly necessary to appreciate the main conceptual points of the paper.
+Appendix~\ref{sec:repl-mutat-dynam} shows that out replicator diffusion dynamic is a special
+case of the well-known replicator mutator dynamics. Appendix~\ref{sec:deriv-rdd-imit} gives a
+derivation of our dynamic as the most-likely path of a population whose agents imitate
+successful behavior without being aware of perceptual noise themselves.
 
@@
 
-A sim-max game consists of a set of states $\States$, a set of
-messages $\Messgs$ with much fewer messages than states, a prior
-probability distribution $\Pr \in \Delta(\States)$ that gives the
-occurrence probabilities of states, a similarity metric on states
-$\similarity \mycolon \States \times \States \rightarrow \mathds{R}$,
-and a utility function $\Utils \mycolon \States \times \States
-\rightarrow \mathds{R}$. We identify the receiver's acts with the
-states of the world, so that the game is one of guessing the actual
-state, so to speak. We also assume that sender's and receiver's
-interests are alike, so we only have one utility function. We do not
-consider message costs, so utilities only depend on the actual state
-and the receiver's response. The utility function should be a
-monotonically decreasing function of similarity.
+A sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ with much
+fewer messages than states, a prior probability distribution $\Pr \in \Delta(\States)$ such
+that $\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity
+metric\footnote{A metric is a function of distance between any two points in a given space. It
+  should satisfy certain axioms that one would intuitively expect from the words ``similarity''
+  and ``distance'' alone, but these details do not matter for the purposes of this paper.}
+  % : similarity should always be non-negative, zero only for the distance
+  % between an element and itself, symmetric, and satisfy the triangle inequality that
+  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
+on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
+$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
+and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
+\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
+for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
+identify the receiver's acts with the states of the world, so that the game is one of guessing
+the actual state, so to speak. The utility function should be a monotonically decreasing
+function of similarity.
 
@@
   \centering
-  \missingfigure{Please include here something like Figure 4 in your paper with Elliot.}
-  \caption{Example of Voronoi language for the unit interval.}
+
+  \begin{tikzpicture}[scale=3]
+
+        \begin{scope}
+
+          \draw[-,thick] (0,0) -- (1.5,0);
+
+          \draw (0,0) node {\color{red}{\big{[}}};
+
+          \draw (.75,0) node {\color{red}{\big{)}}};
+
+          \draw (.76,0) node {\color{red}{\big{(}}};
+
+          \draw (1.5,0) node {\color{red}{\big{]}}};
+
+            \draw (.325,.15) node
+            {\color{red}{$\Spure(\mymessg{1}  \probbar \cdot)$}};
+
+            \draw (1.075,.15) node
+            {\color{red}{$\Spure(\mymessg{2} \probbar \cdot)$}};
+
+            \node at (.375,0) (node1) {\color{blue}{$\bullet$}};
+
+            \node at (.375,-.3) (label1)
+            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{1})$}};
+
+            \draw[-,color=blue] (node1) -- (label1);
+
+            \node at (1.125,0) (node2) {\color{blue}{$\bullet$}};
+
+            \node at (1.125,-.3) (label2)
+            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{2})$}};
+
+            \draw[-] (node2) -- (label2);
+
+            \draw (0,-.2) node {$0$};
+
+            \draw (.75,-.2) node {$0.5$};
+
+            \draw (1.5,-.2) node {$1$};
+
+          \end{scope}
+
+        \end{tikzpicture}
+
+        \caption{Example of a Voronoi language on $\States = [0,1]$. The (pure) sender strategy
+          $\Spure \in \Messgs^\States$ uses one signal for the lower half, and another for the
+          upper half of the unit interval. The (pure) receiver strategy $\Rpure \in
+          \States^\Messgs$ selects the central elements in the respective intervals. See
+          \citet{JagerMetzger2011:Voronoi-Languag} for further details.}
   \label{fig:Voronoi-language-example}
@@
 
-The formulation of the \rdd given and explored here is in terms of
-behavioral strategies, not in terms of mixed strategies. Dynamics on
-behavioral strategies assume that agents can adapt
-their behavior locally, so to speak, i.e., independently at each
-choice point. This greatly reduces the complexity of the dynamic and
-simplifies numerical simulations. But it also points towards an
-interpretation of the process as cultural evolution, rather than
-biological evolution. Signaling strategies are not inherited as a
-whole, but adapted piece-meal, e.g., by conditional imitation of
-``local'' behavior at a given choice point (more below). Still, the
-approach we suggest here is compatible with a formulation in terms of
-mixed strategies and an interpretation as biological evolution as
-well. In Appendix~\ref{sec:diffusion-as-special}, we
-translate our initial formulation of the \rdd
-into one using mixed strategies. Doing so identifies the \rdd as a
-special case of the \rmd.
-
-For now, let us first focus on the
-formulation of the dynamic in terms of behavioral strategies alone.
-We begin by recapitulating the \rd in
-Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in
-Section~\ref{sec:repl-diff-dynam-1}.
+The formulation of the \rdd given and explored here is in terms of behavioral strategies, not
+in terms of mixed strategies. Dynamics on behavioral strategies assume that agents can adapt
+their behavior locally, so to speak, i.e., independently at each choice
+point. Section~\ref{sec:repl-mutat-dynam} elaborates on the
+difference. Section~\ref{sec:relating-dynamics} shows how to translate our initial formulation
+of the \rdd into one using mixed strategies. Doing so identifies the \rdd as a special case of
+the \rmd.
+
+Our focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
+numerical simulations. But it also points towards an interpretation of the process as cultural
+evolution, rather than biological evolution. Signaling strategies are not inherited as a whole,
+but adapted piece-meal, e.g., by conditional imitation of ``local'' behavior at a given choice
+point. Appendix~\ref{sec:deriv-rdd-imit} gives one possible derivation of our dynamic as the
+most likely path in a population of agents that imitate successful behavior while being
+entirely unaware of perceptual imprecision themselves. Still, the possibility to formulate the
+\rdd as a special case of the \rmd (Appendix~\ref{sec:diffusion-as-special}) also suggests that
+an interpretation as a process of biological evolution may be feasible.
+
+For now, let us first focus on the formulation of the dynamic in terms of behavioral strategies
+alone.  We begin by recapitulating the \rd in Section~\ref{sec:repl-dynam-behav}, and introduce
+the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
 %Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
@@
 \end{align*}
-This means that the expected utility for the sender of choosing message $\messg \in \Messgs$, given a state $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability the receiver will choose act $\act$ given message $\messg$ times the actual utility of using act $\act$ for state $\state$.
-Conversely, the expected utility for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior probability of state $\state$ occurring, times the probability the sender will choose message $\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.
+This means that the expected utility for the sender of choosing message $\messg \in \Messgs$,
+given tate $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act
+$\act \in \Acts$, of the probability that the receiver will choose act $\act$ given message
+$\messg$ times the actual utility of using act $\act$ for state $\state$.  The expected utility
+for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a
+sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior
+probability of state $\state$ occurring, times the probability the sender will choose message
+$\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.
 
@@
 \end{align*}
-The \rd in behavioral strategies describes the most likely path of
-development of average behavior in a population
-of agents when behavioral adaptations occur independently at each
-choice point, e.g., when agents adopt the behavior of another agent
-\emph{at a given choice point} if that ``local'' behavior is
-``locally'' better than their own (see
-\citeauthor{Sandholm2013:Population-Game}
-\citeyear{Sandholm2013:Population-Game} for agent-level
-interpretations of the replicator dynamics and
-\citeauthor{Cressman2003:Evolutionary-Dy}
-\citeyear{Cressman2003:Evolutionary-Dy} for applications to extensive
-form games and behavioral strategies).
+The \rd in behavioral strategies describes the most likely path of development of average
+behavior in a population of agents when behavioral adaptations occur independently at each
+choice point, e.g., when agents adopt the behavior of another agent \emph{at a given choice
+  point} if that ``local'' behavior is ``locally'' better than their own (see
+\cite{Sandholm2010:Population-Game} for agent-level interpretations of the replicator dynamics
+and \citeauthor{Cressman2003:Evolutionary-Dy} \citeyear{Cressman2003:Evolutionary-Dy} for
+applications to extensive form games and behavioral strategies).
 
@@
 in a sequential manner:
-\begin{align*}
+\begin{align}
+  \label{eq:RDD-Def}
   \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
-\end{align*}
+\end{align}
 This is equivalent to the following, perhaps more transparent
@@
 
-The most straightforward interpretation of this sequential definition
-is that the \rdd captures the most likely path of development of
-aggregate behavior in a society, in which behavioral changes occur
-independently at each choice point, as described by the \rd for
-behavioral strategies. Whatever strategy an agent does actually hold,
-its realization is perturbed by imperfect discriminability of states
-in such a way that the expected overall aggregate behavior is the
-diffusion of the strategies that the agents would play in a noise-free
-world. Any agent-level interpretation of the \rd in behavioral
-strategies seems \emph{prima facie} compatible with this
-interpretation.
-
-The sequential definition highlights that other (discrete-time)
-evolutionary dynamics can be subjected to state-confusability in an
-analogous fashion. It is easy, for instance, to perturb a
-best-response dynamic in case we wanted to model cognizing, rational
-agents with a noisy perceptual system. Our choice of the \rd as a
-carrier dynamic is motivated by its generality and austerity, being
-compatible with a variety of implicit agent-level behavior, without
-presupposing extravagant higher cognitive abilities
-\citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, by using the
-\rd as a carrier for diffusion, we actually obtain a special case of a
-behavioral-strategy version of the replicator mutator dynamics. This
-is important to know because it enables similar applications of
-diffusion from perceptual confusion also in models of biological
-evolution.
-We worked out the relation between the \rdd and the \rmd, but given that it is rather technical and not strictly necessary to appreciate the main point of the paper, we defer it to Appendix~\ref{sec:diffusion-as-special}.
-In the following section we explore the long-term effects of using the \rdd.
+The most straightforward interpretation of this sequential definition is that the \rdd captures
+the most likely path of development of aggregate behavior in a society, in which behavioral
+changes occur independently at each choice point, as described by the \rd for behavioral
+strategies. Perceptual noise affects the realization of behavior, but not the fitness-based
+selection thereof. Alternative noise-perturbed versions of the replicator dynamics in which the
+fitness-based selection of strategies takes noise into account are conceivable, but we must put
+the exploration of alternatives aside here. We suggest that our approach is practical,
+conservative and plausible: it is practical because it is easy to implement; it is conservative
+because it connects the \rdd with the \rmd (see Appendix~\ref{sec:diffusion-as-special}); it is
+also plausible because the \rdd can be derived from a myopic update protocol in which agents
+themselves are unaware of perceptual noise (see Appendix~\ref{sec:deriv-rdd-imit}).
 
@@
 
-The previous section argued that the \rdd can be conceptualized as a
-two-step procedure: we first calculate where the replicator dynamic
-would take us, then apply diffusion. To understand the \rdd better, we
-isolate the novel diffusion part in
-Section~\ref{sec:iterated-diffusion}, before exploring how diffusion
-and fitness-based selection interact in Section~\ref{sec:simulations}.
+The \rdd is as a two-step procedure: first calculate where the replicator dynamic takes us,
+then apply diffusion. To understand the \rdd better, we isolate the novel diffusion part in
+Section~\ref{sec:iterated-diffusion}, before exploring how diffusion and fitness-based
+selection interact in Section~\ref{sec:simulations}.
 
@@
 \begin{align*}
-  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \impairment) =
+  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
       \begin{cases}
-    1 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
-    0 & \textrm{if } \impairment = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \impairment^2} \right ) & \textrm{otherwise,} \\
+    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
+    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
     \end{cases}
 \end{align*}
-where $\impairment \ge 0$ is an impairment, imprecision or indiscriminability
-parameter\footnote{We use these terms interchangeably throughout the paper.}.
-When $\impairment=0$ agents perfectly discriminate between
-states; when $\impairment \rightarrow \infty$ agents cannot
-discriminate states at all. Figure~\ref{fig:NosofskySim} gives an
-impression of Nosofsky-similarity for different parameter
-values. Other formalizations of perceptual similarity are possible,
-including ones that allow for different discriminability in different
-areas of the state space, but we stick with Nosofsky's similarity
-function for the time being, because it is mathematically simple, yet
-an established notion in mathematical psychology.
+where $\imprecision \ge 0$ is a imprecision parameter. When $\imprecision=0$ agents perfectly
+discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
+states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
+different parameter values. Other formalizations of perceptual similarity are possible,
+including ones that allow for different discriminability in different areas of the state space,
+but we stick with Nosofsky's similarity function for the time being, because it is
+mathematically simple, yet an established notion in mathematical psychology.
 
@@
   C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
-  \impairment)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
-  \impairment)} \,.
+  \imprecision)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
+  \imprecision)} \,.
 \end{align*}
 Confusion of states is then also a function of imprecision parameter
-$\impairment$. For $\impairment = 0$, the confusion matrix has $C_{ii}
-= 1$ and $C_{ij} = 0$ for $i \neq j$. For $\impairment > 0$, $C$ is positive, i.e., $C_{ij} >0$ for
-all $i$ and $j$. For $\impairment \rightarrow \infty$, we find $C_{ij}
+$\imprecision$. For $\imprecision = 0$, the confusion matrix has $C_{ii}
+= 1$ and $C_{ij} = 0$ for $i \neq j$. For $\imprecision > 0$, $C$ is positive, i.e., $C_{ij} >0$ for
+all $i$ and $j$. For $\imprecision \rightarrow \infty$, we find $C_{ij}
 \rightarrow \nicefrac{1}{\card{\States}}$ for all $i$ and $j$.
@@
 receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we assume that
-$\impairment > 0$, $C$ is a positive row-stochastic matrix (each state
+$\imprecision > 0$, $C$ is a positive row-stochastic matrix (each state
 could in principle be confused as any other state, albeit with
@@
 so, whether the inclusion of diffusion has any noteworthy effects on
-the evolving meaning of signals. Clearly, if $\impairment = 0$, the
-\rdd reduces to the \rd. If $\impairment \rightarrow \infty$, the
+the evolving meaning of signals. Clearly, if $\imprecision = 0$, the
+\rdd reduces to the \rd. If $\imprecision \rightarrow \infty$, the
 diffusion component takes over and the \rdd reduces to the trivial
@@
 section. We will show presently that for reasonable in-between levels
-of imprecision $\impairment$, the \rdd can lead to reasonably
+of imprecision $\imprecision$, the \rdd can lead to reasonably
 efficient, yet vague signaling behavior. Suitable levels of
@@
 and receiver strategies, for each triplet of independent parameter
-values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\impairment \in
+values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\imprecision \in
 \set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
@@
     \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0$}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
     \label{fig:example_stratsA}
@@
     \includegraphics[width=\textwidth]{plots/strat_example_ind3452.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\impairment = 0.1$}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.1$}
     \label{fig:example_stratsB}
@@
   %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0$}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0$}
   %   \label{fig:example_stratsC}
@@
   %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\impairment = 0.05$}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0.05$}
   %   \label{fig:example_stratsD}
@@
 Representative examples for resulting strategy pairs are given in
-Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA}
-shows a strategy pair at stopping time with 90 states, tolerance
-$\toler = 0.1$ and imprecision $\impairment = 0$. Zero imprecision
-means that the trial was effectively an application of the plain
-\rd. Noteworthily, the given sender strategy approximates a pure
-sender strategy that crisply partitions the state space into
-non-convex sets. The irregular shape of the receiver strategy shows
-clearly that the pictured strategy pair has not yet reached a
-dynamically stable state under the \rd. Indeed, the trial was stopped
-for reaching the maximum of 200 rounds. In contrast, the outcome of a trial
-with identical parameters, except with imprecision $\impairment =
-0.1$, which is shown in Figure~\ref{fig:example_stratsB}, had
-converged (in our technical sense) after 99 rounds of iterated
-\rdd. The sender strategy shows a smooth blending from one
-``category'' to the other, and the receiver's interpretations are
-rather extended curves, peaking at a central point in the relevant
-``categories.''
-
-These examples already show two interesting things. Firstly, inclusion
-of imprecision can lead to seemingly well-behaved, yet vague
-strategies in the sense that we are after. The sender strategy in
-Figure~\ref{fig:example_stratsB} identifies clear positive and clear
-negative cases for each signal, with a smooth transition
-in-between. The receiver's interpretations of signals can be seen as
-smoothed-out prototype regions. Secondly, (sender) strategies can
-approach non-convex pure strategies under the replicator dynamic and
-linger there for vast amounts of time, possibly indefinitely. We see
-this in our limited-time simulations, but this also holds, for some
-types of utility functions, for the limiting case. This was first
-observed by Elliott Wagner, as mentioned by \citet{OConnor2014-OCOEPC}.
-A full analysis of the dynamics of sim-max games is beyond the scope of this paper, but we
-will shortly see that diffusion from confusability of states clearly
-prevents evolutionary paths that meander for a long time in the
-vicinity of non-convex strategies.
+Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
+stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
+imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
+the given sender strategy approximates a pure sender strategy that crisply partitions the state
+space into non-convex sets. The irregular shape of the receiver strategy suggests that the
+pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
+the trial was stopped for reaching the maximum of 200 rounds. In contrast, the outcome of a
+trial with identical parameters, except with imprecision $\imprecision = 0.1$, which is shown
+in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99 rounds of
+iterated \rdd. The sender strategy shows a smooth blending from one ``category'' to the other,
+and the receiver's interpretations are rather extended curves, peaking at a central point in
+the relevant ``categories.''
+
+These examples already show two interesting things. Firstly, inclusion of imprecision can lead
+to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
+strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
+for each signal, with a smooth transition in-between. The receiver's interpretations of signals
+can be seen as smoothed-out prototype regions. Secondly, (sender) strategies can approach
+non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
+time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
+Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
+for the limiting case. This was first observed by Elliott Wagner, as mentioned by
+\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
+scope of this paper, but we will shortly see that diffusion from confusability of states
+clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
+strategies.
 
@@
   \caption{Means of gradient and proportions of categorical measures
-    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\impairment
+    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\imprecision
     \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
@@
 
-Figure~\ref{fig:MeanMetrics} shows plots summarizing some results.
-%of our simulation.
-For perspicuity, we only plot results for one level of
-tolerance $\beta = 0.1$, and leave out the case of $n_s = 90$. Still,
-every qualitative trend mentioned in the following applies to the
-whole set of results.
+Figure~\ref{fig:MeanMetrics} shows plots summarizing some results.  For perspicuity, we only
+plot results for one level of tolerance $\beta = 0.1$, and leave out the case of $n_s =
+90$. Still, every qualitative trend mentioned in the following applies to the whole set of
+results.
 
@@
 convex sender strategies. Moreover, it also led to higher likelihood
-of convergence. In fact, sufficient impairment always ensured
+of convergence. In fact, sufficient imprecision always ensured
 convergence and convexity. It appears that perceptual imprecision
@@
 little variation in the recorded metrics for evolved strategies, at
-least for higher values of impairment. On closer inspection, it turns
-out that variability in low-impairment conditions is not only due to
+least for higher values of imprecision. On closer inspection, it turns
+out that variability in low-imprecision conditions is not only due to
 non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
@@
     \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
-    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
+    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
     \label{fig:example_stratsC}
@@
     \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
-    \caption{$\ns = 6$, $\toler = 0.2$, $\impairment = 0$}
+    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
     \label{fig:example_stratsD}
@@
 Interestingly, this type of variability in evolutionary outcomes can
-be weeded out by impairment. To investigate this, we calculated the average
+be weeded out by imprecision. To investigate this, we calculated the average
 distance between evolved sender strategies within each group of trials
@@
 \end{align*}
-To compensate for the arbitrariness of message use, we set the
-distance between strategies $\Sstrat$ and $\Sstrat'$ to be the maximum
-of $\text{HD}(\Sstrat,\Sstrat')$ and $\text{HD}(\Sstrat^*,\Sstrat')$
-where $\Sstrat^*$ is $\Sstrat$ with reversed message indices. An example of the
-\emph{within group distance}, i.e., the average distances between all
-sender strategies obtained for the same parameter values, is plotted in Figure~\ref{fig:GroupMeasuresA} for $\beta =
-0.1$ and $n_s = 10$. Despite some quantitative differences, the
-general trend is the same for all other parameter settings that we tested:
-with increasing imprecision, the sender strategies evolving under \rdd
-were much more alike (modulo swapping of messages). This means that
-perceptual imprecision unifies evolutionary outcomes and guarantees
-sender strategies that are not only convex, but also regular in that
-they induce a vague category split exactly in the middle of the unit
-interval.
+To compensate for the arbitrariness of message use, we set the distance between strategies
+$\Sstrat$ and $\Sstrat'$ to be the maximum of $\text{HD}(\Sstrat,\Sstrat')$ and
+$\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
+indices. An example of the \emph{within group distance}, i.e., the average distances between
+all sender strategies obtained for the same parameter values, is plotted in
+Figure~\ref{fig:GroupMeasuresA} for $\beta = 0.1$ and $n_s = 10$. Despite some quantitative
+differences, the general trend is the same for all other parameter settings that we tested:
+with increasing imprecision, the sender strategies evolving under the \rdd were much more alike
+(modulo swapping of messages). This means that perceptual imprecision unifies evolutionary
+outcomes and guarantees sender strategies that are not only convex, but also regular in that
+they induce a vague category split exactly in the middle of the unit interval.
 
@@
 \paragraph{Generalized reinforcement.}
-% Repeated below...
-%\citet{OConnor2013:The-Evolution-o} introduced a generalization of
-%Herrnstein reinforcement learning for sim-max games and showed that
-%this not only leads to vague signaling patterns, but can speed-up
-%evolution of efficient signaling strategies, especially in games with
-%many states.
-
-Under \emph{Herrnstein reinforcement learning}, sender and receiver
-adjust their dispositions to act after each round of play, in such a
-way that the actual (non-negative) payoff gained in the previous
-interaction is added to the non-normalized propensities for repeating
-the same behavior. For sim-max games, this means that, when the sender
-chose $\messg$ in state $\state$, and this resulted in some
-non-negative payoff, the probability that the sender chooses $\messg$
-again in $\state$ is increased, but nothing else changes. In
-particular, the sender's behavior in other choice points does not
-change. \emph{Generalized reinforcement learning} is different. When
-the use of $\messg$ in $\state$ gave positive payoff, then not only
-will the probability of its future use be promoted at $\state$ but also at
-other states, proportional to how similar these are to
-$\state$. Similar amendments take care of the way that the receiver
-updates his choice dispositions.
-
-\citet{OConnor2013:The-Evolution-o} shows that this extension not only
-leads to vague signaling of the appropriate kind, but also speeds up
-learning in such a way that, especially for sim-max games with higher
-numbers of states, higher levels of communicative success are reached
-in shorter learning periods. Technically, this result is partly due to
-the fact that signalers make bigger changes to their behavioral
-strategies after each round of play under generalized reinforcement
-than under the Herrnstein variety. That only explains the speed of
-adaptation, not necessarily that generalization also leads to
-regularity and communicative efficiency, but it does.
+
+% Under \emph{Herrnstein reinforcement learning}, sender and receiver
+% adjust their dispositions to act after each round of play, in such a
+% way that the actual (non-negative) payoff gained in the previous
+% interaction is added to the non-normalized propensities for repeating
+% the same behavior. For sim-max games, this means that, when the sender
+% chose $\messg$ in state $\state$, and this resulted in some
+% non-negative payoff, the probability that the sender chooses $\messg$
+% again in $\state$ is increased, but nothing else changes. In
+% particular, the sender's behavior in other choice points does not
+% change. \emph{Generalized reinforcement learning} is different. When
+% the use of $\messg$ in $\state$ gave positive payoff, then not only
+% will the probability of its future use be promoted at $\state$ but also at
+% other states, proportional to how similar these are to
+% $\state$. Similar amendments take care of the way that the receiver
+% updates his choice dispositions.
+
+\citet{OConnor2013:The-Evolution-o} introduces a generalization of classical Herrnstein
+reinforcement learning for sim-max games, in which successful play leads to reinforcement of
+acts also for states similar to the one that actually occurred. This generalization not only
+leads to vague signaling of the appropriate kind, but also speeds up learning in such a way
+that, especially for sim-max games with higher numbers of states, higher levels of
+communicative success are reached in shorter learning periods. % Technically, this result is
+% partly due to the fact that signalers make bigger changes to their behavioral strategies
+% after each round of play under generalized reinforcement than under the Herrnstein
+% variety. That only explains the speed of adaptation, not necessarily that generalization also
+% leads to regularity and communicative efficiency, but it does.
 
@@
 \paragraph{Quantal response equilibria.}
-\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in
-which information-processing limitations could lead to vague
-strategies. The model that is most clearly related to the present
-approach uses the notion of a \emph{quantal response}, also known as a
-\emph{soft-max function}
+\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
+information-processing limitations could lead to vague strategies. The model that is most
+clearly related to the present approach uses the notion of a \emph{quantal response}, also
+known as a \emph{soft-max function}
 \citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. A
-quantal response function is a parameterized generalization of the
-classic best response function. For example, if $U \mycolon \Acts
-\rightarrow \mathds{R}$ is the measure of expected utility over
-choices $\Acts$ of an agent, then a best response function would have
-the agent choose $\act$ only if $U(\act) = \max_{\act^\prime \in \Acts}
-U(\act^\prime)$. A quantal response function rather assumes that agents would
-choose $\act$ with a probability proportional to $\expo(\lambda \cdot
-U(\act))$, where $\lambda$ is a rationality parameter. If $\lambda
-\rightarrow \infty$ we retrieve the behavior of the best response
-function, but if it is positive but finite, any choice $\act$ will
-receive a positive probability, but acts with higher expected utility
-will be more likely. The underlying motivation is the assumption that
-there is noise in the computation of expected utilities and/or in
-maximization of expected utilities. Consequently, choices with almost
-equal expected utilities will be chosen with almost equal probability
-(for moderate values of $\lambda$).
-
-\citet{FrankeJager2010:Vagueness-Signa} show that quantal response
-equilibria of sim-max games can show the desired marks of vague
-signaling. A quantal response equilibrium is a pair of sender and
-receiver strategies such that the sender strategy is the quantal
-response to the expected utilities under the receiver strategy and
-vice versa. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a
-quantal response equilibrium for a sim-max game, as used in our set-up
-but with higher tolerance $\toler = 0.5$. Sender and receiver
-strategies look very much like what evolves under \rdd with modest
-values of perceptual imprecision.
+quantal response is a stochastic generalization of the best response function (i.e., the
+deterministic choice of an action that maximizes expected utility). Quantal responses factor in
+stochastic trembles at the level of the computation or maximization of expected utilities:
+agents may make mistakes in calculating how good given choice options really are for
+them. \citet{FrankeJager2010:Vagueness-Signa} show that quantal response equilibria of sim-max
+games can show the desired marks of vague signaling. Figure~\ref{fig:exampleQRE_stratsA} gives
+an example of a quantal response equilibrium for a sim-max game, as used in our set-up but with
+higher tolerance $\toler = 0.5$. Sender and receiver strategies look very much like what
+evolves under \rdd with modest values of perceptual imprecision.
 
@@
 
-  \caption{Examples of vague quantal response equilibria.}
+  \caption{Examples of vague quantal response equilibria. Here, $\lambda$ is an imprecision
+    parameter measuring the extent of noise in the calculation of expected utilities
+    \citep[see][for details]{FrankeJager2010:Vagueness-Signa}.}
   \label{fig:exampleQREs}
@@
 
-But not all quantal response equilibria are equally plausible.
-%, and the
-%ones that are not suggest that it is less natural to think of
-%vagueness as arising from errors in the computation or maximization of
-%expected utility than that it arises from confusion of similar states
-To see what the problem is, consider
-Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response
-equilibrium for a game with lower tolerance $\toler = 0.05$. Unlike
-what evolves under \rdd in this case, sender strategies have vague
-boundaries also towards the end of the unit intervals. Technically,
-this is because quantal responses equalize message use far away from
-the ``prototypical'' interpretation, not just in-between categories,
-so to speak. This, in turn, is because quantal responses introduce
-noise into the decision making at the level of computing or maximizing
-expected utility of choices.
-
-That this is problematic shows even more clearly in a case where
-the state space is intuitively unbounded, as for instance for the
-property ``tall.'' If the usual interpretation of a ``tall man'' peaks
-at around, say, 195cm then when meeting a giant of $n$ meters, senders
-would, according to the quantal response approach, be ever more
-inclined to describe the giant indifferently as either ``tall'' or
-``short'' the larger $n$ gets. This is because, as the distance from
-the prototype increases for larger $n$, the difference between the
-expected utilities of saying ``short'' or ``tall'' will converge to
-zero. 
-%Whence that the quantal response approach would predict that
-%senders would grow indifferent between choice of antonyms as $n$
-%grows, which seems weird.
-
-Admittedly, this argument hinges on the choice of utility
-function. Still, to the extent that the utility functions we chose are
-reasonable---and we think they are very reasonable---the case suggests
-that quantal responses alone are not a good model for why linguistic
-categories are vague. With respect to this issue, perceptual confusion
-of similar states produces more intuitive predictions. What is more
-plausible, though, is that both perceptual confusion \emph{and}
-computational errors in maximizing expected utility play a role in
-making vague categorization a pervasive feature of natural language
-meaning. Perhaps even further factors, such as limited observations,
-imperfect memory or uncertainty about the exact prior distribution of
-states, should be considered in order to form a more complete
-understanding of the phenomenon of vagueness.
+But not all quantal response equilibria are equally plausible.  To see what the problem is,
+consider Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response equilibrium for a
+game with lower tolerance $\toler = 0.05$. Unlike what evolves under \rdd in this case, sender
+strategies have vague boundaries also towards the end of the unit intervals. Technically, this
+is because quantal responses equalize message use far away from the ``prototypical''
+interpretation, not just in-between categories, so to speak. Conceptually, this is because
+quantal responses introduce noise at the level of computing or maximizing expected utility of
+choices.
+
+In contrast, the \rdd implements noise at the level of the perception of different
+states. Unintuitive outcomes like in Figure~\ref{fig:exampleQRE_stratsB} are thereby
+avoided. Still, we believe that it would be most plausible to assume that both perceptual
+confusion \emph{and} computational errors in maximizing expected utility play a role in making
+vague categorization a pervasive feature of natural language meaning. Perhaps even further
+factors, such as limited observations, imperfect memory or uncertainty about the exact prior
+distribution of states, should be considered in order to form a more complete understanding of
+the phenomenon of vagueness.
 
@@
 
-We introduced the replicator diffusion dynamic as a special case of
-the replicator mutator dynamic. The \rdd incorporates a special kind
-of stochastic mutation due to differential confusability of similar
-states. Based on data from numerical simulations, we demonstrated that
-the inclusion of mild levels of stochastic confusability of states
-does not undermine the possibility of evolving relatively successful
-signaling strategies. On the contrary, strategies that evolved under
-mild diffusion induce a highly regular and systematic category
-structure that shows the signs of vagueness as desired. There might
-even be a higher-order benefit to the presence of imprecision, in that
-it can accelerate convergence to optimal categorization, by preventing
-evolutionary paths to stay near inefficient non-convex strategies for
-a long time. Diffusion thus also unifies and regularizes evolutionary
-outcomes in such a way that sub-optimal categorization is avoided in
+We introduced the replicator diffusion dynamic and identified it as a special case of the
+replicator mutator dynamic (Appendix~\ref{sec:diffusion-as-special}). The \rdd incorporates a
+special kind of stochastic mutation due to differential confusability of similar states. It can
+be thought of as the aggregate population-level behavior process that results from imitation
+among agents that are unaware of noise (Appendix~\ref{sec:deriv-rdd-imit}). Based on data from
+numerical simulations, we demonstrated that the inclusion of mild levels of stochastic
+confusability of states does not undermine the possibility of evolving relatively successful
+signaling strategies. On the contrary, strategies that evolved under mild diffusion induce a
+highly regular and systematic category structure that shows the signs of vagueness as
+desired. There might even be a higher-order benefit to the presence of imprecision, in that it
+can accelerate convergence to optimal categorization, by preventing evolutionary paths to stay
+near inefficient non-convex strategies for a long time. Diffusion thus also unifies and
+regularizes evolutionary outcomes in such a way that sub-optimal categorization is avoided in
 short-term trajectories.
@@
 \subsection{Relating the dynamics}
+\label{sec:relating-dynamics}
 
@@
 
+\section{Derivation of the RDD by imitation-based revision}
+\label{sec:deriv-rdd-imit}
+
+We can derive the proposed \rdd as a mean dynamic of an imitative revision protocol. For
+expository reasons we follow closely the example 4.3.1 of \citet{Sandholm2010:Population-Game},
+but other possibilities exist. Our derivation relies on the usual assumptions about conditional
+imitation at the agent level, but must add special assumptions about how noise affects revision
+probabilities. In keeping with the idea that agents are of possibly limited cognitive
+capacities, we assume here that noise is real but that agents are unaware of it. When agents
+perceive the actual expected utility of their choice and of another agent's choice (as is usual
+in the motivation of imitation-based revision protocols), we assume additionally that agents
+mistake their possibly non-veridical subjective representations of states for the actual ones
+(details below).
+
+\paragraph{Notation.} As before, let $\Sstrat(\mymessg{i} \probbar \mystate{o})$ be the
+probability that a speaker randomly picked from the population chooses $\mymessg{i}$ if the
+actual/objective state is $\mystate{o}$. This speaker may perceive the objective state
+$\mystate{o}$ subjectively as $\mystate{s}$. In contrast, let $\Sstrat_s(\mymessg{i} \probbar
+\mystate{s})$ be the probability that a randomly sampled speaker chooses $\mymessg{i}$ if she
+experiences some objective state $\mystate{o}$ as $\mystate{s}$. The latter is the average
+sender behavior in a (possibly counterfactual) world without perceptual confusion. It is also
+where behavioral changes first take effect, to then percolate to $\Sstrat$ as well.
+
+Similarly, $\Rstrat(\mystate{o} \probbar \messg)$ is the probability that a randomly sampled
+receiver actually chooses interpretation $\mystate{o}$. $\Rstrat_s(\mystate{s} \probbar
+\messg)$ is the probability of a subjectively intended interpretation $\mystate{s}$. Again,
+behavioral changes first target $\Rstrat_s$ but cause changes in $\Rstrat$ as well.
+
+Let $\EU(\messg)$ be short for $\EU(\messg,\state,\Rstrat)$ whenever
+$\state$ and $\Rstrat$ are implicit, similarly for $\EU(\state)$ as a
+shorthand for $\EU(\state,\messg,\Sstrat)$.
+
+\paragraph{Standard assumptions.} Let's assume that we are in an
+entirely noise-free world so that $\Sstrat = \Sstrat_s$ and $\Rstrat =
+\Rstrat_s$. The standard assumptions for imitation-based revisions are
+these. Agents get revision opportunities for all actual choice points
+with equal probability (independently of how likely these would occur
+in actual play). Take the case of a sender (receiver side is
+parallel). If agent $A$ is given a revision opportunity for actual
+state $\state$, she consults her own choice, which is $\mymessg{i}$
+with probability $\Sstrat(\mymessg{i} \probbar \state)$, and observes
+the behavior of one randomly picked other agent $B$, which is
+$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
+\state)$. $A$ also observes the expected utility of both choices (at
+$\state$) and switches to $B$'s choice with probability $r_{ij} = [
+\EU(\mymessg{j}) - \EU(\mymessg{i}) ]_+$, so that the overall
+switching probability of an agent like $A$ who samples $\mymessg{i}$
+for $\state$ is $p_{ij} = \Sstrat(\mymessg{j} \probbar \state) \
+r_{ij}$.
+
+\paragraph{Imitation under perceptual noise.} There are many conceivable ways in which
+perceptual noise might influence the above revision protocol. Our choice here is guided by
+theoretical and practical considerations alike. We would like maintain the common assumption
+that agents might not be ideally rational and not necessarily fully aware of their
+surroundings. In particular, we assume here that agents are not aware of perceptual
+noise. Switching probabilities are therefore defined in terms of actual expected utility (and
+not subjective expectations that take noise into consideration). Moreover, we assume that
+agents who have (unconscious) access to expected utilities for actual states $\mystate{o}$ will
+also (mistakenly) treat whatever subjective state $\mystate{s}$ they may have experienced
+during their revision opportunity as if it had been the actual state $\mystate{o}$.
+
+Concretely, suppose agent $A$ has a revision opportunity for objective
+state $\mystate{o}$, and chooses $\mymessg{i}$ with probability
+$\Sstrat(\mymessg{i} \probbar \mystate{o})$. Similarly, $B$ chooses
+$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
+\mystate{o})$. It may be that $A$ perceived $\mystate{o}$ as
+$\mystate{s_A}$ and that $B$ perceived $\mystate{s_B}$, but since $A$
+is unaware of noise he expects no difference. $A$ observes (as usual)
+the actual expected utilities of $\mymessg{i}$ and $\mymessg{j}$ for
+$\mystate{o}$. Consequently, he updates his subjective behavior for
+$\mystate{o}$ (not being aware that the perceived $\mystate{s_A}$
+might have been different). This myopic revision behavior gives rise
+to expected changes in the subjective/noise-free choice probabilities
+$\Sstrat_s(\mymessg{i} \probbar \state)$ as follows:
+\begin{align*}
+  \dot{\Sstrat}_s(\mymessg{i} \probbar \state) & = \sum_j
+  \Sstrat(\mymessg{j} \probbar \state) \cdot p_{ji} -  
+  \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j p_{ij} \\
+  & = \sum_j  \Sstrat(\mymessg{j} \probbar \state) \cdot
+    \Sstrat(\mymessg{i} \probbar \state) \cdot [\EU(\mymessg{i}) -
+    \EU(\mymessg{j})]_+ \\ 
+   & \ \ \ \  - \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j
+   \Sstrat(\mymessg{j} \probbar \state) \cdot [\EU(\mymessg{j}) -
+    \EU(\mymessg{i})]_+ \\
+   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left ( \sum_j
+     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{i}) - \sum_j
+     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{j})
+   \right ) \\
+   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left (
+     \EU(\mymessg{i}) - \text{average EU at $\state$} \right )
+\end{align*}
+This is the (continuous-time) replicator dynamics for behavioral
+strategies.
+
+Similar considerations motivate the receiver part. Suppose that agent
+$A$ has a revision opportunity for choice point $\messg$. He and a
+random other agent $B$ actually choose $\mystate{o_A}$ and
+$\mystate{o_B}$ with probability $\Rstrat(\mystate{o_{A,B}} \probbar
+\messg)$, even though they may have intended interpretations
+$\mystate{s_A}$ and $\mystate{s_B}$. Agent $A$ observes the actual
+expected utility and, by assumption, believes (myopically) that he had
+wanted to realize $\mystate{o_A}$ and that $B$ had wanted to realize
+$\mystate{o_B}$. Under these assumptions, the derivation of the
+replicator dynamic for subjectively intended (noise-free) receiver
+behavior $\Rstrat_s$ proceeds in parallel with the above sender case.
+
+At each time step, the noise-free behavior of agents is, of course, perturbed by noise. This is
+why, after each update step, the \rdd defined in Equation~(\ref{eq:RDD-Def}) perturbes the
+outcome of the replicator step with perceptual noise.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-05-14 snippets/exploring_uniformity_from_impairment/uniformity.R
changed plot caption to 'imprecision'

@@ -220,13 +220,13 @@ show(combPlot)
 dsub = subset(D, D$Number.of.states == 10 & D$Tolerance == 0.1)
 wgdConcise = ggplot(dsub, aes(x = Impairment, y = inGroupDistance)) + 
   geom_line( color = "grey") + geom_point() +
-  xlab("Impairment") + ylab("")
+  xlab("Imprecision") + ylab("")
 show(wgdConcise)
-ggsave(wgdConcise, file = "~/Desktop/data/svn/vagueness-games/paper/version_02/plots/WithinGroupDistanceConcise.pdf", height=2, width=3) 
+ggsave(wgdConcise, file = "~/Desktop/data/svn/vagueness-games/paper/version_03/plots/WithinGroupDistanceConcise.pdf", height=2, width=3) 
 
 eusub = subset(meanEUs, meanEUs$Number.of.states == 10 & meanEUs$Tolerance == 0.1)
 EUConcise = ggplot(eusub, aes(x = Impairment, y = Speaker.meanEU)) + 
   geom_line( color = "grey") + geom_point() +
-  xlab("Impairment") + ylab("")
+  xlab("Imprecision") + ylab("")
 show(EUConcise)
-ggsave(EUConcise, file = "~/Desktop/data/svn/vagueness-games/paper/version_02/plots/WithinGroupEUConcise.pdf", height=2, width=3) 
+ggsave(EUConcise, file = "~/Desktop/data/svn/vagueness-games/paper/version_03/plots/WithinGroupEUConcise.pdf", height=2, width=3) 

### 2015-05-15 paper/version_03/paper.tex
revision version 3; this version sent

@@
 contributes to the latter line of thought. Concretely, we introduce the replicator diffusion
-dynamics that integrates stochastic noise on the differential confusability of similar
-stimuli. The dynamic proposed here generalizes and complements previous like-minded
-accounts. We show that stochastic noise can not only lead to vague, yet relatively efficient
-signal use, but can also unify evolutionary outcomes and help avoid suboptimal categorization.
+dynamic that integrates confusability of similar stimuli. The dynamic proposed here generalizes
+and complements previous like-minded accounts. We show that stochastic noise can not only lead
+to vague, yet relatively efficient signal use, but can also unify evolutionary outcomes and
+help avoid suboptimal categorization.
 
@@
 not strictly necessary to appreciate the main conceptual points of the paper.
-Appendix~\ref{sec:repl-mutat-dynam} shows that out replicator diffusion dynamic is a special
-case of the well-known replicator mutator dynamics. Appendix~\ref{sec:deriv-rdd-imit} gives a
-derivation of our dynamic as the most-likely path of a population whose agents imitate
+Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
+special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
+gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
 successful behavior without being aware of perceptual noise themselves.
@@
 A sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ with much
-fewer messages than states, a prior probability distribution $\Pr \in \Delta(\States)$ such
-that $\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity
-metric\footnote{A metric is a function of distance between any two points in a given space. It
-  should satisfy certain axioms that one would intuitively expect from the words ``similarity''
-  and ``distance'' alone, but these details do not matter for the purposes of this paper.}
+fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
+$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
+  metric is a function of distance between any two points in a given space. It should satisfy
+  certain axioms that ensure behavior that one would intuitively expect from the words
+  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
+  this paper.}
   % : similarity should always be non-negative, zero only for the distance
@@
 
-\citet{JagerMetzger2011:Voronoi-Languag} showed that the
-evolutionarily stable states of sim-max games are remarkably
-systematic. Their results were obtained for games with infinitely many
-states in $n$-dimensional Euclidean space $\States \subseteq
-\mathds{R}^n$ and a quadratic loss function for utilities
-$\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} -
-\mystate{2})^2$. For these games, the evolutionarily stable states are
-demonstrably so-called Voronoi languages. Roughly put, a Voronoi
-language is a pair of sender and receiver strategies, such that the
-sender strategy partitions the state space into convex categories,
-while the receiver's interpretations are the central spots in each
-category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put,
-all points in $X$ are connected via a straight line that lies entirely
-in $X$; $X$ has no gaps or dents. For example, if \States is the unit
-interval and all states are equiprobable, a Voronoi language with two
-messages would have the sender use one message exclusively for all
-points in the lower half of the unit interval and another for all
-points in the upper half; the receiver's interpretations of messages
-are the central points, .25 and .75, in the respective intervals.
-See Figure~\ref{fig:Voronoi-language-example} for an illustration.
+\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of
+sim-max games are remarkably systematic: the evolutionarily stable states are demonstrably
+so-called Voronoi languages.\footnote{Their results were obtained for games with infinitely
+  many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a
+  quadratic loss function for utilities
+  $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$.} Roughly put, a Voronoi
+language is a pair of sender and receiver strategies, such that the sender strategy partitions
+the state space into convex categories, while the receiver's interpretations are the central
+spots in each category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put, all points
+in $X$ are connected via a straight line that lies entirely in $X$; $X$ has no gaps or
+dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
+language with two messages would have the sender use one message exclusively for all points in
+the lower half of the unit interval and another for all points in the upper half; the
+receiver's interpretations of messages are the central points, .25 and .75, in the respective
+intervals.  See Figure~\ref{fig:Voronoi-language-example} for an illustration.
 
@@
 
-This result is interesting, because it demonstrates that signaling can
+This result is interesting because it demonstrates that signaling can
 impose orderly categories on a metric space, without that being the
@@
 
-An alternative approach is taken by
-\citet{FrankeJager2010:Vagueness-Signa} and
-\citet{OConnor2013:The-Evolution-o} who show how the above desiderata
-can be met by evolving strategies in sim-max games. The approach we
-take here is similar in spirit, but different in relevant detail. We
-argue here that \citeauthor{FrankeJager2010:Vagueness-Signa}'s
-approach gives rise to problematic predictions under certain
-conditions which are avoided in our approach. While
-\citeauthor{OConnor2013:The-Evolution-o} gives an agent-level learning
-dynamic under which vague signaling arises, we complement the picture
-by giving a population-level dynamic that abstracts from the concrete
-behavior of signalers. A more in-depth comparison is deferred until
-Section~\ref{sec:discussion}. Let us first look into our own proposal
-in more detail.
+An alternative approach is taken by \citet{FrankeJager2010:Vagueness-Signa} and
+\citet{OConnor2013:The-Evolution-o} who show how the above desiderata can be met by evolving
+strategies in sim-max games. The approach we take here is similar in spirit, but different in
+relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
+us first look into our own proposal in more detail.
 
@@
 inability to sharply distinguish similar states. We call our dynamic
-\emph{replicator diffusion dynamic} (\rdd), because it is an extension
+\emph{replicator diffusion dynamic} (\rdd) because it is an extension
 of the replicator dynamic (\rd) and a special case of the replicator
@@
 their behavior locally, so to speak, i.e., independently at each choice
-point. Section~\ref{sec:repl-mutat-dynam} elaborates on the
-difference. Section~\ref{sec:relating-dynamics} shows how to translate our initial formulation
+point. Appendix~\ref{sec:repl-mutat-dynam} elaborates on the
+difference. Appendix~\ref{sec:relating-dynamics} shows how to translate our initial formulation
 of the \rdd into one using mixed strategies. Doing so identifies the \rdd as a special case of
@@
 point. Appendix~\ref{sec:deriv-rdd-imit} gives one possible derivation of our dynamic as the
-most likely path in a population of agents that imitate successful behavior while being
-entirely unaware of perceptual imprecision themselves. Still, the possibility to formulate the
-\rdd as a special case of the \rmd (Appendix~\ref{sec:diffusion-as-special}) also suggests that
-an interpretation as a process of biological evolution may be feasible.
-
-For now, let us first focus on the formulation of the dynamic in terms of behavioral strategies
-alone.  We begin by recapitulating the \rd in Section~\ref{sec:repl-dynam-behav}, and introduce
+most likely path of development in a population of agents that imitate successful behavior
+while being entirely unaware of perceptual imprecision themselves. Still, the possibility to
+formulate the \rdd as a special case of the \rmd (Appendix~\ref{sec:diffusion-as-special}) also
+suggests that an interpretation as a process of biological evolution may be feasible.
+
+We begin by recapitulating the \rd in Section~\ref{sec:repl-dynam-behav}, and introduce
 the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
@@
   point} if that ``local'' behavior is ``locally'' better than their own (see
-\cite{Sandholm2010:Population-Game} for agent-level interpretations of the replicator dynamics
+\cite{Sandholm2010:Population-Game} for agent-level interpretations of the replicator dynamic
 and \citeauthor{Cressman2003:Evolutionary-Dy} \citeyear{Cressman2003:Evolutionary-Dy} for
@@
 \end{align}
-For example, diffusion thus defined determines that, on the diffused sender strategy, the probability that message $\messg$ is chosen given a state $\state_i$ is given by $\sum_{j} C_{ij} \cdot \Sstrat(\messg \probbar \state_j)$.
-This means that the original probability given by $\Sstrat$ is increased by the original probability of this message being chosen in other states, proportionally to how likely it is that they are mistaken for $\state_i$.
+For example, diffusion thus defined determines that, on the diffused sender strategy, the
+probability that message $\messg$ is chosen given a state $\mystate{i}$ is given by
+$\sum_{j} C_{ij} \cdot \Sstrat(\messg \probbar \mystate{j})$.  This means that the original
+probability given by $\Sstrat$ is increased by the original probability of this message being
+chosen in other states $\mystate{j}$, proportionally to how likely it is that $\mystate{i}$ is
+(mistakenly) perceived as $\mystate{j}$.
 
@@
 strategies. Perceptual noise affects the realization of behavior, but not the fitness-based
-selection thereof. Alternative noise-perturbed versions of the replicator dynamics in which the
+selection thereof. Alternative noise-perturbed versions of the replicator dynamic in which the
 fitness-based selection of strategies takes noise into account are conceivable, but we must put
@@
 
-The \rdd is as a two-step procedure: first calculate where the replicator dynamic takes us,
-then apply diffusion. To understand the \rdd better, we isolate the novel diffusion part in
+The \rdd is a two-step procedure: first calculate where the replicator dynamic takes us, then
+apply diffusion. To understand the \rdd better, we isolate the novel diffusion part in
 Section~\ref{sec:iterated-diffusion}, before exploring how diffusion and fitness-based
@@
 diffusion interacts with optimization of behavior, as described by the
-replicator dynamics. The main question we should address is whether
+replicator dynamic. The main question we should address is whether
 diffusion and fitness-based replication reasonably interact, and if
@@
 
-The \rdd adds expected confusion of similar states to the \rd in
-behavioral strategies. The result is an abstract evolutionary dynamic
-that describes the most likely path of development of aggregate
-population behavior in which any process of behavioral adaptation at
-the agent level that is consistent with the \rd in behavioral
-strategies is perturbed by systematic noise in the realization of the
-agents' strategies. Given the use of behavioral strategies, an
-interpretation as a process of cultural evolution is most salient,
-according to which agents adapt their choice dispositions locally at
-each choice point. But the \rdd is also extensible to biological
-evolution on mixed strategies, in which case it is a special case of
-the \rmd with mutation matrices derived from confusion matrices, as
-shown in Appendix~\ref{sec:diffusion-as-special}.
+The \rdd adds expected confusion of similar states to the \rd in behavioral strategies. The
+result is an abstract evolutionary dynamic that describes the most likely path of development
+of aggregate population behavior that is consistent with the idea that agents imitate other
+agents' behavior (Appendix~\ref{sec:deriv-rdd-imit}). This would be a process of cultural
+evolution operating on behavioral strategies. But we speculate that the \rdd is also extensible
+to biological evolution on mixed strategies, in which case it is a special case of the \rmd
+with mutation matrices derived from confusion matrices, as shown in
+Appendix~\ref{sec:diffusion-as-special}.
 
@@
 
-\citet{OConnor2013:The-Evolution-o} introduces a generalization of classical Herrnstein
-reinforcement learning for sim-max games, in which successful play leads to reinforcement of
-acts also for states similar to the one that actually occurred. This generalization not only
-leads to vague signaling of the appropriate kind, but also speeds up learning in such a way
-that, especially for sim-max games with higher numbers of states, higher levels of
-communicative success are reached in shorter learning periods. % Technically, this result is
+\citet{OConnor2013:The-Evolution-o} introduces a variant of classical Herrnstein reinforcement
+learning for sim-max games, in which successful play leads to reinforcement of acts also for
+states similar to the one that actually occurred. This is a form of generalization to unseen
+contingencies. It not only leads to vague signaling of the appropriate kind, but also speeds up
+learning in such a way that, especially for sim-max games with higher numbers of states, higher
+levels of communicative success are reached in shorter learning
+periods. % Technically, this result is
 % partly due to the fact that signalers make bigger changes to their behavioral strategies
@@
 
-Diffusion of strategies in the \rdd can also be conceived of as a form
-of generalization, and works in large part quite analogously to stimulus
-generalization in \citeauthor{OConnor2013:The-Evolution-o}'s
-approach. But there are still
-differences. \citeauthor{OConnor2013:The-Evolution-o} showed that
-generalized reinforcement learning can speed up the development of
-efficient signaling, especially for higher numbers of
-states. Complementing this, we showed that diffusion regularizes
-evolutionary outcomes and prevents meandering around sub-optimal and
-non-convex strategies, also for low numbers of states.
-
-Conceptually, the \rdd is a more abstract framework than generalized
-reinforcement learning. The latter is most naturally seen as an
-agent-based learning dynamic that has two players adapt their
-individual strategies after each concrete round of play. In contrast,
-the \rdd describes a more abstract, average dynamical change in
-behavioral dispositions at the aggregate population level. Although
-the behavior of (some forms of) reinforcement learning mirror those of
-the \rd at some stage in time
+Diffusion of strategies in the \rdd can also be conceived of as a form of generalization, and
+works in large part quite analogously to stimulus generalization in
+\citeauthor{OConnor2013:The-Evolution-o}'s approach. But there are still
+differences. \citeauthor{OConnor2013:The-Evolution-o} showed that her generalized reinforcement
+learning can speed up the development of efficient signaling, especially for higher numbers of
+states. Complementing this, we showed that diffusion regularizes evolutionary outcomes and
+prevents meandering around sub-optimal and non-convex strategies, also for low numbers of
+states.
+
+Conceptually, the \rdd is a more abstract framework than generalized reinforcement
+learning. The latter is most naturally seen as an agent-based learning dynamic that has two
+players adapt their individual strategies after each concrete round of play. In contrast, the
+\rdd describes a more abstract, average dynamical change in behavioral dispositions at the
+aggregate population level. Although the behavior of (some forms of) reinforcement learning
+mirror those of the \rd at some stage in time
 (\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
-this does not mean that \emph{generalized} reinforcement learning is
-also necessarily a plausible population-level description of, say,
-generalized learning in a population of agents. Seen in this light,
-generalized reinforcement learning and the \rdd nicely complement each
-other, as similarly-minded accounts operating at different levels of
-abstraction.
-Although it would be interesting to explore the possibility of a mathematical connection between the two, this is beyond the scope of this paper and thus left as future work.
+this does not mean that \emph{generalized} reinforcement learning is also necessarily a
+plausible population-level description of, say, generalized learning in a population of
+agents. Seen in this light, generalized reinforcement learning and the \rdd nicely complement
+each other, as similarly-minded accounts operating at different levels of abstraction.
+Although it would be interesting to explore the possibility of a mathematical connection
+between the two, this is beyond the scope of this paper.
 
@@
 agents may make mistakes in calculating how good given choice options really are for
-them. \citet{FrankeJager2010:Vagueness-Signa} show that quantal response equilibria of sim-max
-games can show the desired marks of vague signaling. Figure~\ref{fig:exampleQRE_stratsA} gives
-an example of a quantal response equilibrium for a sim-max game, as used in our set-up but with
-higher tolerance $\toler = 0.5$. Sender and receiver strategies look very much like what
-evolves under \rdd with modest values of perceptual imprecision.
+them. \citet{FrankeJager2010:Vagueness-Signa} demonstrate that quantal response equilibria of
+sim-max games can show the desired marks of vague
+signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a quantal response
+equilibrium for a sim-max game, as used in our set-up but with higher tolerance $\toler =
+0.5$.
+Sender and receiver strategies look very much like what evolves under \rdd with modest values
+of perceptual imprecision.
 
@@
 be thought of as the aggregate population-level behavior process that results from imitation
-among agents that are unaware of noise (Appendix~\ref{sec:deriv-rdd-imit}). Based on data from
-numerical simulations, we demonstrated that the inclusion of mild levels of stochastic
-confusability of states does not undermine the possibility of evolving relatively successful
-signaling strategies. On the contrary, strategies that evolved under mild diffusion induce a
-highly regular and systematic category structure that shows the signs of vagueness as
-desired. There might even be a higher-order benefit to the presence of imprecision, in that it
-can accelerate convergence to optimal categorization, by preventing evolutionary paths to stay
-near inefficient non-convex strategies for a long time. Diffusion thus also unifies and
-regularizes evolutionary outcomes in such a way that sub-optimal categorization is avoided in
-short-term trajectories.
+among agents that are unaware of perceptual confusion
+(Appendix~\ref{sec:deriv-rdd-imit}). Based on data from numerical simulations, we demonstrated
+that the inclusion of mild levels of perceptual imprecision does not undermine the possibility
+of evolving relatively successful signaling strategies. On the contrary, strategies that
+evolved under mild imprecision induce a highly regular and systematic category structure that
+shows the signs of vagueness as desired. There might even be a higher-order benefit to the
+presence of imprecision, in that it can accelerate convergence to optimal categorization, by
+preventing evolutionary paths to stay near inefficient non-convex strategies for a long time. A
+little imprecision thus also unifies and regularizes evolutionary outcomes in such a way that
+sub-optimal categorization is avoided in short-term trajectories.
 
@@
 
-The usual formulation of the \rmd is based on mixed, not behavioral
-strategies. But a discrete-time formulation of the \rmd in behavioral
-strategies clearly identifies the \rdd to be a special case. It
-follows that a mixed strategy version of the \rdd is just the \rmd
-with suitably defined mutation matrices. From this, even a
-continuous-time version can easily be retrieved if desired. To show
-this, we start with a delineation of mixed and behavioral strategies.
+The usual formulation of the \rmd is based on mixed, not behavioral strategies. But a
+discrete-time formulation of the \rmd in behavioral strategies clearly identifies the \rdd to
+be a special case. It follows that a mixed-strategy version of the \rdd is just the \rmd with
+suitably defined mutation matrices. From this, even a continuous-time version can easily be
+retrieved if desired. To show this, we start with a delineation of mixed and behavioral
+strategies.
 
@@
 
-We can derive the proposed \rdd as a mean dynamic of an imitative revision protocol. For
-expository reasons we follow closely the example 4.3.1 of \citet{Sandholm2010:Population-Game},
-but other possibilities exist. Our derivation relies on the usual assumptions about conditional
-imitation at the agent level, but must add special assumptions about how noise affects revision
-probabilities. In keeping with the idea that agents are of possibly limited cognitive
-capacities, we assume here that noise is real but that agents are unaware of it. When agents
-perceive the actual expected utility of their choice and of another agent's choice (as is usual
-in the motivation of imitation-based revision protocols), we assume additionally that agents
-mistake their possibly non-veridical subjective representations of states for the actual ones
-(details below).
+We can derive the proposed \rdd as a mean dynamic of an imitative revision protocol, following
+closely the example 4.3.1 of \citet{Sandholm2010:Population-Game}. Our derivation relies on the
+usual assumptions about conditional imitation at the agent level, but must add special
+assumptions about how noise affects revision probabilities. In keeping with the idea that
+agents are of possibly limited cognitive capacities, we assume here that noise is real but that
+agents are unaware of it. When agents perceive the actual expected utility of their choice and
+of another agent's choice (as is usual in the motivation of imitation-based revision
+protocols), we assume additionally that agents mistake their possibly non-veridical subjective
+representations of states for the actual ones.
 
@@
 perceptual noise might influence the above revision protocol. Our choice here is guided by
-theoretical and practical considerations alike. We would like maintain the common assumption
+theoretical and practical considerations alike. We would like to maintain the common assumption
 that agents might not be ideally rational and not necessarily fully aware of their
@@
 \end{align*}
-This is the (continuous-time) replicator dynamics for behavioral
+This is the (continuous-time) replicator dynamic for behavioral
 strategies.

### 2015-07-23 paper/derivation_deepNoise/derivation_deepNoise.tex
notes on deep noise imitatoin

\documentclass[fleqn,reqno,11pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{../../helpers/mypackages}
%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
%\bibliography{MyRefGlobal}
\bibliography{../../helpers/MyRefGlobal}
\bibliography{paper} 
\usepackage{../../helpers/myenvironments}
\usepackage{../../helpers/mycommands}
\usepackage{todonotes}
\usepackage{subcaption}

%========================================
% Standard Layout
%========================================

% Itemize
\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
% Description
\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}

% Figure Captions
\usepackage{caption} % use corresponding myfiguresize!
\setlength{\captionmargin}{20pt}
\renewcommand{\captionfont}{\small}
\setlength{\belowcaptionskip}{7pt} % standard is 0pt

%========================================
% Additional layout & commands
%========================================

\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}

% Annotations
\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation

\newcommand{\impairment}{\ensuremath{\alpha}} % impairment
\newcommand{\toler}{\ensuremath{\beta}} % tolerance
\newcommand{\ns}{\ensuremath{n_s}} % number of states

\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function

\usepackage{fullpage}

% \doublespacing

\begin{document}

Our goal is to derive a simple and practical dynamic for behavioral adaptations in a sim-max
game given noisy perception and realization of states. We want to take noise into account at
all levels: the realization of strategies and imperfections in perceiving other agents'
strategy choices (for imitation). The focus here in on adaptations triggered by imitation of
behavior, so that we look at adaptations of local behavior, in terms of behavioral strategies.

Fix a sim-max game and fix behavioral strategies $\Sstrat$ and $\Rstrat$ that represent the
actual behavior of agents without noise or perturbations. As usual, we assume that agents carry
pure strategies and that behavioral strategies express population frequencies. So:
$\Sstrat(\messg \probbar \mystate{o})$ is the probability with which a randomly drawn sender
from the population plays $\messg$ when she perceives the state $\mystate{o}$;
$\Rstrat(\mystate{i} \probbar \messg)$ is the probability that a receiver, randomly drawn from
the population will intend to play $\mystate{i}$ when hearing message $\messg$. There could be
various sources of noise that perturb: (i) what agents observe (conditional probability
distribution $P_o$, see below) and (ii) how receivers realize intended state (interpretation)
choices (conditional probability distribution $P_r$, see below). Together with the basic priors
over states, these are the basic probabilities from which everything else will be derived. The
replicator dynamic is the limiting case where $P_o$ and $P_r$ are trivial.

Concretely, we think of the process of playing the sim-max game as subject to several pieces of
stochasticity:
\begin{enumerate}
\item the actual state $\mystate{a}$ occurs with $P(\mystate{a}$)
\item any randomly sampled sender would observes state $\mystate{o}$ with probability
  $P_o(\mystate{o} \probbar \mystate{a})$
  \begin{itemize}
  \item given observed $\mystate{o}$, the probability of actual state $\mystate{a}$ is
    $P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P(\mystate{a}) P_o(\mystate{o}
    \probbar \mystate{a})$
  \end{itemize}
\item a randomly sampled sender will play $\messg$ when observing $\mystate{o}$ with
  probability $\Sstrat(\messg \probbar \mystate{o})$
  \begin{itemize}
  \item when the true state is $\mystate{a}$ the probability that a randomly sampled sender
    plays $\messg$ is $P_\Sstrat(\messg \probbar \mystate{a}) = \sum_{\mystate{o}}
    P_o(\mystate{o} \probbar \mystate{a}) \Sstrat(\messg \probbar \mystate{o})$
  \end{itemize}
\item given $\messg$ the probability that a randomly drawn receiver will intend to interpret
  with $\mystate{i}$ is $\Rstrat(\mystate{i} \probbar \messg)$
\item if a receiver intends to realize $\mystate{i}$, the probability that he realizes
  $\mystate{r}$ is $P_r(\mystate{r} \probbar \mystate{i})$
\item $P_\Rstrat(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)$ is the probability that a randomly sampled
  receiver realizes $\mystate{r}$ when observing $\messg$
\end{enumerate}

We want to define a protocol with which agents update their behavior locally, for each choice
point. Given that we have many aspects of noise in the system, revision protocols will be
subject to this noise as well. As a consequence, it appears that \textbf{different revision
  protocol that would all give rise to the replicator dynamic in a noise-free setting, can give
  rise to different mean field dynamics if we take noise seriously}. Our aim here is to
consider one revision protocol, namely ``imitation of success'' (Sandholm) that is plausible
enough and yields a practical and elegant mean field dynamic that we can study/implement.

Revision by ``imitation of success'' generally proceeds as follows. Call the possibly
updating/switching agent ``learner'' and the possibly to-be-imitated agent ``teacher''. A
random learner is drawn from the population and given a chance to change her behavior at choice
point $k$. Suppose, she currently plays $c$ at $k$. She observes what a randomly sampled
teacher would do at choice point $k$, say $c'$ and switches from $c$ to $c'$ with switch
probability $\text{SWITCH}(c \rightarrow c' , k)$. For the imitation of success protocol, this
latter switching probability is just the expected utility of $c'$ at choice point $k$ (if that
is always positive, like in our case). The mean change induced by this is:
\begin{align*}
  \text{mean change of frequency $c$, at $k$} & = \text{probability of switchers $c'
    \rightarrow c$ at $k$ (inflow)} \\
& - \text{probability of switchers $c \rightarrow c'$ at $k$ (outflow)} \\
\end{align*}
with
\begin{align*}
  \text{inflow} & = \sum_{c'} \text{probability learner plays $c'$ at $k$} \\
  & \ \ \ \times \text{probability teacher plays $c$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c' \rightarrow c , k) \\
  \text{outflow} & = \sum_{c'} \text{probability learner plays $c$ at $k$} \\
  & \ \ \ \times \text{probability teacher plays $c'$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k) \\
\end{align*}

Given noise in a sim-max game, as described, this update protocol must be adapted so as to take
into account that learners may not always perfectly perceive what the teacher is doing, and
that expected utilities (and hence switching probabilities) are also subject to noisy
realizations of strategies. This gives us:

\begin{align*}
  \text{inflow} & = \sum_{c'} \text{probability learner actually plays $c'$ at perceived $k$} \\
  & \ \ \ \times \text{probability learner observes teacher play $c$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c' \rightarrow c , k) \\
  \text{outflow} & = \sum_{c'} \text{probability learner actually plays $c$ at perceived $k$} \\
  & \ \ \ \times \text{probability learner observes teacher play $c'$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k) \\
\end{align*}

Define:
\begin{itemize}
\item $P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})$ is the probability that
  a learner with revision opportunity for $\mystate{o}$ perceives a random teacher produce $\messg$
\item $P_o(\mystate{o} \probbar \messg) =  \sum_{\mystate{r}} P_o(\mystate{o} \probbar
  \mystate{r}) \myts P_{\Rstrat}(\mystate{r} \probbar \messg)$ is the probability that a
  learner with a revision opportunity for $\messg$ perceives a random teacher produce
  $\mystate{o}$\footnote{We assume that the noise that applies to the sender's observations
    also applies to the receiver's observations when trying to imitate other agents'
    strategies}
\item $\EU(\messg , \mystate{o}, \Rstrat) = \sum_{\mystate{a}} \sum_{\mystate{r}}
  P_{\overline{o}}(\mystate{a} \probbar \mystate{o})  P_\Rstrat(\mystate{r} \probbar
  \messg) \utils(\mystate{a}, \messg, \mystate{r})$ is the actual noise-aware utility of
  sending message $\messg$ when perceiving state $\mystate{o}$ given that the receiver
  population plays a noisy realization of $\Rstrat$
\item $P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P(\mystate{a})
  P_{\Sstrat}(\messg \probbar \mystate{a})$ is the probability that the real state is
  $\mystate{a}$ given that a random sender produced $\messg$
\item $\EU(\mystate{i} , \messg, \Sstrat) = \sum_{\mystate{a}}
  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})$ is the actual noise-aware utility of
  intending to realize interpretation $\mystate{i}$ given message $\messg$ when the sender
  population plays a noisy realization of $\Sstrat$
\end{itemize}

Given this we can derive the continuous time expected change in behavioral sender strategies as
follows:
\begin{align*}
  \Sstrat(\messg \probbar \state) & = \sum_{\messg'} \Sstrat(\messg' \probbar \state) \myts
  P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \state) \myts
  P_o(\messg' \probbar \state) \myts \EU(\messg', \state, \Rstrat) \\
  & = P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \Sstrat(\messg \probbar \state) \myts \sum_{\messg'} 
  P_o(\messg' \probbar \state) \myts \EU(\messg', \state, \Rstrat)
\end{align*}
which has a discrete-time solution that is easily implementable:
\begin{align*}
  \Sstrat'(\messg \probbar \state) \propto P_o(\messg \probbar \state) \myts \EU(\messg,
  \state, \Rstrat)
\end{align*}
The case of the receiver is analogous:
\begin{align*}
  \Rstrat(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg) \myts
  P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'} \Rstrat(\state \probbar \messg) \myts
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \\
  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar \messg) \myts \sum_{\state'} 
  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat)\\
  \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg) \myts \EU(\state,
  \messg, \Sstrat)
\end{align*}

\printbibliography[heading=bibintoc]

\end{document}

### 2015-07-24 paper/derivation_deepNoise/derivation_deepNoise.tex
notes minor edits

@@
 Our goal is to derive a simple and practical dynamic for behavioral adaptations in a sim-max
-game given noisy perception and realization of states. We want to take noise into account at
-all levels: the realization of strategies and imperfections in perceiving other agents'
-strategy choices (for imitation). The focus here in on adaptations triggered by imitation of
-behavior, so that we look at adaptations of local behavior, in terms of behavioral strategies.
+game given noisy perception and realization of states. The focus here in on adaptations
+triggered by imitation of behavior, so that we look at adaptations of local behavior, in terms
+of behavioral strategies. We want to take noise into account (i) during the realization of
+strategies and (ii) in perceiving other agents' strategy choices (for imitation).
 
@@
 distribution $P_o$, see below) and (ii) how receivers realize intended state (interpretation)
-choices (conditional probability distribution $P_r$, see below). Together with the basic priors
-over states, these are the basic probabilities from which everything else will be derived. The
-replicator dynamic is the limiting case where $P_o$ and $P_r$ are trivial.
+choices (conditional probability distribution $P_r$, see below). Together with priors over
+states, these are the basic probabilities from which everything else will be derived. (The
+replicator dynamic is the limiting case where $P_o$ and $P_r$ are trivial.)
 
@@
 subject to this noise as well. As a consequence, it appears that \textbf{different revision
-  protocol that would all give rise to the replicator dynamic in a noise-free setting, can give
-  rise to different mean field dynamics if we take noise seriously}. Our aim here is to
-consider one revision protocol, namely ``imitation of success'' (Sandholm) that is plausible
-enough and yields a practical and elegant mean field dynamic that we can study/implement.
+  protocols that would all give rise to the replicator dynamic in a noise-free setting, can
+  give rise to different mean field dynamics under noise}. Our aim here is to consider one
+revision protocol, namely ``imitation of success'' (Sandholm) that is plausible enough and
+yields a practical and elegant mean field dynamic that we can study easily through a
+discrete-time implementation.
 
@@
     \rightarrow c$ at $k$ (inflow)} \\
-& - \text{probability of switchers $c \rightarrow c'$ at $k$ (outflow)} \\
+& - \text{probability of switchers $c \rightarrow c'$ at $k$ (outflow)}
 \end{align*}
-with
-\begin{align*}
-  \text{inflow} & = \sum_{c'} \text{probability learner plays $c'$ at $k$} \\
-  & \ \ \ \times \text{probability teacher plays $c$ at $k$} \\
-  & \ \ \ \times \text{SWITCH}(c' \rightarrow c , k) \\
-  \text{outflow} & = \sum_{c'} \text{probability learner plays $c$ at $k$} \\
-  & \ \ \ \times \text{probability teacher plays $c'$ at $k$} \\
-  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k) \\
-\end{align*}
-
-Given noise in a sim-max game, as described, this update protocol must be adapted so as to take
-into account that learners may not always perfectly perceive what the teacher is doing, and
-that expected utilities (and hence switching probabilities) are also subject to noisy
-realizations of strategies. This gives us:
-
+with:
 \begin{align*}
@@
   & \ \ \ \times \text{probability learner observes teacher play $c'$ at $k$} \\
-  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k) \\
+  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k)
 \end{align*}
+
+Given noise in a sim-max game, as described, this update protocol must take into account that
+learners may not always perfectly perceive what the teacher is doing, and that expected
+utilities (and hence switching probabilities) are also subject to noisy realizations of
+strategies.
 
@@
   a learner with revision opportunity for $\mystate{o}$ perceives a random teacher produce $\messg$
-\item $P_o(\mystate{o} \probbar \messg) =  \sum_{\mystate{r}} P_o(\mystate{o} \probbar
-  \mystate{r}) \myts P_{\Rstrat}(\mystate{r} \probbar \messg)$ is the probability that a
-  learner with a revision opportunity for $\messg$ perceives a random teacher produce
-  $\mystate{o}$\footnote{We assume that the noise that applies to the sender's observations
-    also applies to the receiver's observations when trying to imitate other agents'
-    strategies}
-\item $\EU(\messg , \mystate{o}, \Rstrat) = \sum_{\mystate{a}} \sum_{\mystate{r}}
-  P_{\overline{o}}(\mystate{a} \probbar \mystate{o})  P_\Rstrat(\mystate{r} \probbar
+\item
+  $P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
+  \myts P_{\Rstrat}(\mystate{r} \probbar \messg)$
+  is the probability that a learner with a revision opportunity for $\messg$ perceives a random
+  teacher produce $\mystate{o}$\footnote{We assume that the noise
+    $P_o(\mystate{o} \probbar \mystate{a})$ that applies to the sender's observations also
+    applies to the receiver's observations when trying to imitate other agents' strategies. We
+    could easily make the model more complex and introduce another measure of noise for the
+    perceptual state confusability during receiver learning. But I don't think that this is
+    needed.}
+\item $\EU(\messg , \mystate{o}, \Rstrat) = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
+  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
   \messg) \utils(\mystate{a}, \messg, \mystate{r})$ is the actual noise-aware utility of

### 2015-08-11 paper/version_04/paper.tex
started on version 4

--- 2015-05-15 paper/version_03/paper.tex
@@
 \newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator
+\newcommand{\icid}{\acro{icid}} % imprecise conditional imitation dynamic
+\newcommand{\ICID}{\ensuremath{\mathrm{ICID}}} % imprecise conditional imitation dynamic
 
@@
 
+\newcommand{\myred}[1]{\textcolor{red}{#1}}
+
 \doublespacing
@@
   approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
-  recent like-minded models, we introduce the replicator diffusion dynamic that combines
-  evolutionary pressure on optimal signal use with stochastic noise in the processing of
-  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
-  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
-  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal categorization.
+  recent like-minded models, we describe an aggregate population-level dynamic that describes a
+  process of imitation of successful signaling behavior under imprecise perception and
+  realization of similar stimuli. Applying this new dynamic to a generalization of Lewis'
+  signaling games, we show that stochastic imprecision leads to vague, yet by-and-large
+  efficient signal use, and, moreover, that it unifies evolutionary outcomes and helps avoid
+  suboptimal categorization. \myred{The upshot of this is that generalization emerges, at an
+    aggregate level, without agents actually generalizing.}
 \end{abstract}
@@
 Others have argued that vagueness is a natural byproduct of limitations in information
-processing or particular learning strategies
-\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This paper
-contributes to the latter line of thought. Concretely, we introduce the replicator diffusion
-dynamic that integrates confusability of similar stimuli. The dynamic proposed here generalizes
-and complements previous like-minded accounts. We show that stochastic noise can not only lead
-to vague, yet relatively efficient signal use, but can also unify evolutionary outcomes and
-help avoid suboptimal categorization.
+processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or the result of particular
+low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
+contributes to the latter line of thought. Concretely, we introduce an dynamic that describes
+average behavior resulting from imitation under confusability of similar stimuli. The dynamic
+proposed here complements previous like-minded accounts. We show that stochastic noise can not
+only lead to vague, yet relatively efficient signal use, but can also unify evolutionary
+outcomes and help avoid suboptimal categorization. Moreover, the system shows generalization
+behavior without a single generalizing agent.
 
 The next section introduces the background against which the work presented here can be
-appreciated. Section~\ref{sec:repl-diff-dynam} introduces the replicator diffusion dynamic,
+appreciated. Section~\ref{sec:repl-diff-dynam} introduces our noise-affected imitation dynamic,
 which is then explored in Section~\ref{sec:exploring-rdd}.  Section~\ref{sec:discussion}
-reflects and compares our approach to others. Two appendices cover technical material that is
-not strictly necessary to appreciate the main conceptual points of the paper.
-Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
-special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
-gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
-successful behavior without being aware of perceptual noise themselves.
+reflects and compares our approach to others. % Two appendices cover technical material that is
+% not strictly necessary to appreciate the main conceptual points of the paper.
+% Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
+% special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
+% gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
+% successful behavior without being aware of perceptual noise themselves.
 
@@
 
-Seen in this light, sim-max games may provide a foundation to
-approaches in cognitive semantics that rely on the notion of
-conceptual spaces.  \citet[][70--77]{Gardenfors2000:Conceptual-Spac},
-for example, has prominently argued that natural categories are convex
-regions in conceptual space. If the conceptual space has a suitable
-metric, convex categories can be derived from a set of prototypes. The
-category corresponding to prototype $p$ is the set of points that are
-more similar to $p$ than to any other. In this way,
-\citet{Gardenfors2000:Conceptual-Spac} argues, an efficient
-categorization system can be obtained: storing the prototypes lets us
-recover the categories without having to store each category's
-extension. However, what is left unexplained, is where the prototypes
-come from, and why we would not see just any distribution of
-prototypes as an equally efficient classification system. This is
-where sim-max games can contribute a principled approach to deriving,
-in an independent way, not only convex categories but also
-prototypical exemplars belonging to them.
-These ideas, and more, are developed further by \citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC}, among others.
+Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
+that rely on the notion of conceptual spaces.
+\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
+natural categories are convex regions in conceptual space. If the conceptual space has a
+suitable metric, convex categories can be derived from a set of prototypes. The category
+corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
+other. In this way, \citet{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
+system can be obtained: storing the prototypes lets us recover the categories without having to
+store each category's extension. However, what is left unexplained, is where the prototypes
+come from, and why we would not see just any distribution of prototypes as an equally efficient
+classification system. This is where sim-max games can contribute a principled approach to
+deriving, in an independent way, not only convex categories but also prototypical exemplars
+belonging to them.  These ideas, and more, are developed further by
+\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
+among others.
 
@@
 
-\section{Replicator diffusion dynamic}
+\section{Imprecise conditional imitation}
 \label{sec:repl-diff-dynam}
 
-We introduce a population-level dynamic in which communicatively more
-efficient strategies increase in proportion, as usual, but where the
-realization of signaling strategies is noisy because of the agents'
-inability to sharply distinguish similar states. We call our dynamic
-\emph{replicator diffusion dynamic} (\rdd) because it is an extension
-of the replicator dynamic (\rd) and a special case of the replicator
-mutator dynamic (\rmd).
-
-The formulation of the \rdd given and explored here is in terms of behavioral strategies, not
-in terms of mixed strategies. Dynamics on behavioral strategies assume that agents can adapt
-their behavior locally, so to speak, i.e., independently at each choice
-point. Appendix~\ref{sec:repl-mutat-dynam} elaborates on the
-difference. Appendix~\ref{sec:relating-dynamics} shows how to translate our initial formulation
-of the \rdd into one using mixed strategies. Doing so identifies the \rdd as a special case of
-the \rmd.
-
-Our focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
-numerical simulations. But it also points towards an interpretation of the process as cultural
-evolution, rather than biological evolution. Signaling strategies are not inherited as a whole,
-but adapted piece-meal, e.g., by conditional imitation of ``local'' behavior at a given choice
-point. Appendix~\ref{sec:deriv-rdd-imit} gives one possible derivation of our dynamic as the
-most likely path of development in a population of agents that imitate successful behavior
-while being entirely unaware of perceptual imprecision themselves. Still, the possibility to
-formulate the \rdd as a special case of the \rmd (Appendix~\ref{sec:diffusion-as-special}) also
-suggests that an interpretation as a process of biological evolution may be feasible.
-
-We begin by recapitulating the \rd in Section~\ref{sec:repl-dynam-behav}, and introduce
-the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
-%Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
-%Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
-%a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
-%\ref{sec:diffusion-as-special} are conceptually important, but rather
-%technical and not strictly necessary to appreciate the main point of
-%the paper.
+Signaling agents can adapt their dispositions to act, given some feedback about their past
+success, in multiple ways. Usually, we would assume that changes in behavior should, at least
+on average, lean towards increasing chances of communicative success. Such a population
+of signaling agents and their behavioral adaptations can be described at different levels of
+abstraction. At the level of individual agents, we can describe a more or less idealized
+process of how each agent adapts dispositions for future actions based on detailed descriptions
+of information available to the agent, e.g., about the game, the other players' behavior and
+the like. More abstractly, at the level of a population of agents, we can describe how
+average behavioral dispositions will evolve. The latter population-level perspective abstracts
+over petty stochastic fluctuations and zooms in on the general tendency or direction of
+evolution that ensues from some (often: simplified) behavior at the agent-level.
+
+We are interested here in exploring how a tendency to favor communicatively successful
+strategies is affected by imprecise perception and realization of states. We therefore look at
+a population-level dynamic that describes the most likely evolutionary path of a population of
+signaling agents who imitate other agents' behavior (proportional to how successful it is) but
+may occasionally confuse states for one another (proportional to how similar they are). We call
+our dynamic \emph{imprecise conditional imitation dynamic} (\icid), because it derives from an
+update protocol based on conditional imitation and allows for imprecise perception and
+realization of states. In the special limiting case where imprecision vanishes, the process is
+just the well-known replicator dynamic. To appreciate the \icid and its motivation, we first
+briefly review the replicator dynamic and its derivation from conditional imitation of
+successes.
 
@@
 
-Fix a signaling game with finite states $\States$, messages $\Messgs$
-and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
-distribution over states and $\Util_{\sen,\rec} \mycolon \States
-\times \Messgs \times \Acts \rightarrow \mathds{R}$ the sender's and
-receiver's utility functions. A behavioral strategy is a function that
-maps an agent's choice points to a probability distribution over
-available choices. The sender's behavioral strategies are functions
-$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
-$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a probability of each state $\state \in \States$ being interpreted by the receiver given $\messg$.
-The expected utility of choices
-at each choice point is:
+Fix a signaling game with finite states $\States$, messages $\Messgs$ and acts $\Acts$. Let
+$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
+$\Util_{\sen,\rec} \mycolon \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ the
+sender's and receiver's utility functions. A \emph{behavioral strategy} is a function that maps
+an agent's choice points to a probability distribution over available choices.\footnote{Our
+  focus is on behavioral strategies, not \emph{mixed strategies}, i.e., probability
+  distributions over functions from each choice point to an act, such as
+  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
+  adapt their behavior locally, so to speak, i.e., independently at each choice point. Our
+  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
+  numerical simulations. But it also seems the more plausible choice for imitation-based update
+  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
+  particular situation, not how that agent would behave in \emph{all} relevant choice
+  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
+  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
+  behavioral strategies.)} The sender's behavioral strategies are functions
+$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
+probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
+$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a
+probability of each state $\state \in \States$ being interpreted by the receiver given
+$\messg$. Although behavioral strategies are probabilistic, we may conveniently imagine that
+every individual agent has a non-probabilistic strategy nonetheless. If the population is large
+enough, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability
+that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$.
+
+The \emph{expected utility} of choices at each choice point is:
 \begin{align*}
@@
 
-The \emph{fitness} of a behavioral strategy at a choice point is the
-frequency-weighted average of expected utilities of each choice, given
-the opponent's strategy:
-\begin{align*}
-  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
-\EU(\messg, \state,\Rstrat) \\
-\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
-\cdot \EU(\act, \messg,\Sstrat)\,.
-\end{align*}
-The discrete-time replicator dynamic maps current strategies $\Sstrat$
-and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
-in such a way that changes in frequency are proportional to expected
-utilities. For behavioral strategies, the changes take place locally
-at each of the agents' choice points:
-\begin{align*}
+% The \emph{fitness} of a behavioral strategy at a choice point is the
+% frequency-weighted average of expected utilities of each choice, given
+% the opponent's strategy:
+% \begin{align*}
+%   \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
+% \EU(\messg, \state,\Rstrat) \\
+% \Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
+% \cdot \EU(\act, \messg,\Sstrat)\,.
+% \end{align*}
+
+The discrete-time replicator dynamic maps current strategies $\Sstrat$ and $\Rstrat$ to future
+strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$ in such a way that changes in frequency are
+proportional to expected utilities:
+\begin{align}
+  \label{eq:RDDiscrete}
   \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
-    \EU(\messg, \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
+    \EU(\messg, \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
+   \EU(\messg', \state,\Rstrat)} \\
     \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
-    \EU(\act, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
-\end{align*}
-The \rd in behavioral strategies describes the most likely path of development of average
-behavior in a population of agents when behavioral adaptations occur independently at each
-choice point, e.g., when agents adopt the behavior of another agent \emph{at a given choice
-  point} if that ``local'' behavior is ``locally'' better than their own (see
-\cite{Sandholm2010:Population-Game} for agent-level interpretations of the replicator dynamic
-and \citeauthor{Cressman2003:Evolutionary-Dy} \citeyear{Cressman2003:Evolutionary-Dy} for
-applications to extensive form games and behavioral strategies).
+    \EU(\act, \messg,\Sstrat)} {\sum_{\act'} \Rstrat(\act' \probbar \messg)
+    \cdot \EU(\act', \messg,\Sstrat)}  \,. \nonumber
+\end{align}
+Intuitively put, this definition updates the average frequency with which each choice (message
+or act) will occur, independently at each choice point (state or message). For a given choice
+point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
+the population \emph{before} the update is $\Sstrat{\messg \probbar \state}$. \emph{After} the
+update it will be proportional to the frequency before the update times the expected utility of
+$\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
+obtain probabilities that sum to one from this. In particular, the denominator in the
+definition above captures the average expected utility of all choices at the relevant choice
+point.
+
+\subsection{Conditional imitation of success}
+\label{sec:cond-imit-succ}
+
+The replicator dynamic is an abstract population-level dynamic that describes the mean,
+to-be-expected change in behavioral dispositions in a population of signalers. There are
+several ways of deriving the replicator dynamic from agent-level processes of behavioral
+adaptation. We focus here on one of the simplest: \emph{conditional imitation of success}
+\citep[see][for this and many other ways of deriving the replicator
+dynamics]{Sandholm2010:Population-Game}. We choose this agent-level derivation, because it is
+more convenient than its alternatives when we want to add imprecision in the following. The
+main idea is that agents imitate the behavior of another agent at some choice point with a
+probability that is proportional to the expected utility of the latter agent's act for that
+choice point. Since the sim-max games that we will be looking at have all expected utilities
+conveniently lie between 0 and 1, we may even identify the switching probability with the
+expected utility directly.
+
+Let's consider sender strategies, as the receiver case is parallel (as long as there is no
+noise or imprecision). Revision by imitation of success proceeds as follows. Call
+(misleadingly!) the agent who gets a chance to change behavior ``learner'' and the possibly
+to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
+chance to change behavior at choice point $\state$. The probability that that learner plays
+$\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
+teacher does at $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar
+\state)$. The learner then starts using $\messg'$ instead of $\messg$ with probability
+$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)
+
+If agents get repeated updating chances for their choice points, the \emph{expected change} of
+frequency for $\messg$ at $\state$ becomes:
+\begin{align}
+  \label{eq:MeanChange}
+  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
+\end{align}
+where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
+$\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
+that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
+population, these probabilities can be given as:
+\begin{align*}
+  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
+    \state)}_{\text{learner plays $\messg'$}} \cdot
+  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
+  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
+  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
+    \state)}_{\text{learner plays $\messg$}} \cdot
+  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
+  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
+\end{align*}
+
+From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
+to:
+\begin{align*}
+  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
+  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
+      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
+& = \underbrace{\Sstrat(\messg \probbar \state)}_{\text{frequency of $\messg$ at $\state$}}  ( \underbrace{\EU(\messg,
+  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
+      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
+      EU at choice point $\state$}}  ) 
+\end{align*}
+This is the continuous-time formulation of the replicator dynamics for behavioral
+strategies. The discrete time formulation in Equation (\ref{eq:RDDiscrete}) solves this
+continuous-time formulation by assuming that discrete update steps are infinitesimally
+small:\footnote{Notice that we drop the denominator here, to obtain the non-payoff corrected
+  replicator dynamic.}
+\begin{align*}
+  \dot{\Sstrat}(\messg \probbar \state) & = \RD(\Sstrat)(\messg \probbar \state) -
+  \Sstrat(\messg \probbar \state) \\
+  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat)} {\sum_{\messg'}
+    \Sstrat(\messg' \probbar \state) \cdot
+    \EU(\messg', \state,\Rstrat)} - \Sstrat(\messg \probbar \state) \\
+  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat) - \Sstrat(\messg
+    \probbar \state) \cdot \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
+    \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
+    \EU(\messg', \state,\Rstrat)}  \\
+  & \propto \Sstrat(\messg \probbar \state) \cdot \left ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
+  \state,\Rstrat) \right )
+\end{align*}
+
+\bigskip
+
+ Suppose, she currently plays $c$ at $k$. She
+observes what a randomly sampled teacher would do at choice point $k$, say $c'$ and switches
+from $c$ to $c'$ with switch probability $\text{SWITCH}(c \rightarrow c' , k)$, which we can
+here identify with the expected utility $\EU(c',k)$ of $c'$ at $k$. The mean change induced by
+offering all agents imitation chances for $k$ is:
+\begin{align*}
+  \text{mean change of frequency $c$, at $k$} & = \text{probability of switchers $c'
+    \rightarrow c$ at $k$ \mygray{(inflow)}} \\
+& - \text{probability of switchers $c \rightarrow c'$ at $k$ \mygray{(outflow)}}
+\end{align*}
+We should spell out the inflow and outflow here with some care, so as to prepare ourselves for
+the subsequent addition of imprecise perception and realization of stimuli. So in general, we
+would define:
+\begin{align*}
+  \text{inflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c'$} \\
+  & \ \ \ \times \text{prob.~learner observes $c$ by random teacher} \\
+  & \ \ \ \times \EU(c', k) \\
+  \text{outflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c$} \\
+  & \ \ \ \times \text{prob.~learner observes random teacher play $c'$ at $k$} \\
+  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k)
+\end{align*}
+
+Suffice it to look at the case of a sender.
 
@@
   \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind3452.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.1$}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.0.5$}
     \label{fig:example_stratsB}
@@
 the trial was stopped for reaching the maximum of 200 rounds. In contrast, the outcome of a
-trial with identical parameters, except with imprecision $\imprecision = 0.1$, which is shown
+trial with identical parameters, except with imprecision $\imprecision = 0.0.5$, which is shown
 in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99 rounds of
@@
 strategies in Figures~\ref{fig:example_stratsA} and
-\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
+\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
 respectively. The receiver strategies have respective entropies $0.43$
-and $0.84$. In general, we expected that vague languages would have
+and $0.81$. In general, we expected that vague languages would have
 higher entropy than crisp ones and that increasing imprecision would
@@
 Figure~\ref{fig:example_stratsA} has a normalized expected utility of
-$0.992$, the pair in Figure~\ref{fig:example_stratsB} has
-$0.901$. Generally, vagueness and imprecision can be expected to
+$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
+$0.95$. Generally, vagueness and imprecision can be expected to
 decrease expected utility

### 2015-08-11 snippets/data_exploration/data_exploration_tmp.Rmd
started on version 4

@@
 
-```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
-
-data$converged = data$Iterations < 201
-data$convex = data$Speaker.Convex.Cat
-
-ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
-             "Speaker.entropy", "Hearer.entropy",
-             "Speaker.Voronoiness", "Hearer.Voronoiness",
-             "Expected.utility",
-             "convex", "converged"
-             )]
-
-dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
-
-dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
-dsub = dms
-dsub$Number.of.states = factor(dsub$Number.of.states)
-
-myLabeller <- function(var, value){
-    value <- as.character(value)
-    if (var=="variable") { 
-        value[value=="Speaker.entropy"] <- "entropy sen"
-        value[value=="Hearer.entropy"] <- "entropy rec"
-        value[value=="Speaker.Voronoiness"] <- "Voronoiness sen"
-        value[value=="Hearer.Voronoiness"] <- "Voronoiness rec"
-        value[value=="Expected.utility"] <- "expected utility"
-    }
-    return(value)
-}
-
-pd <- position_dodge(.01)
-mysize = 1.65
-sp = ggplot(dsub, aes(x=Impairment, y=value, color = Number.of.states, shape=Number.of.states, linetype=Number.of.states)) + 
-    geom_line(color="gray", position = pd) + xlab("imprecision") +
-     geom_point(position = pd, size = mysize) +
-     geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
-    scale_shape_discrete(name = "# states", guide = guide_legend()) +
-    scale_linetype_discrete(name = "# states", guide = guide_legend()) +
-    scale_color_discrete(name = "# states", guide = guide_legend()) +
-    theme(legend.position="top")
-
-sp = sp + facet_grid(variable ~ Tolerance, scales="free", labeller=myLabeller)
-
-sp + facet_grid(. ~ sex, labeller=mf_labeller)
-show(sp)
-
-ggsave(sp, file = "paper/plots/MeanMetrics2.pdf", width=8, height = 10)
-
-```
+# ```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
+# 
+# data$converged = data$Iterations < 201
+# data$convex = data$Speaker.Convex.Cat
+# 
+# ds = data[,c("Number.of.states", "Impairment", "Tolerance", 
+#              "Speaker.entropy", "Hearer.entropy",
+#              "Speaker.Voronoiness", "Hearer.Voronoiness",
+#              "Expected.utility",
+#              "convex", "converged"
+#              )]
+# 
+# dm = melt(ds, id.vars = c("Number.of.states", "Impairment", "Tolerance"))
+# 
+# dms = summarySE(dm, groupvars = c("Number.of.states", "Impairment", "Tolerance", "variable"), measurevar="value")
+# dsub = dms
+# dsub$Number.of.states = factor(dsub$Number.of.states)
+# 
+# myLabeller <- function(var, value){
+#     value <- as.character(value)
+#     if (var=="variable") { 
+#         value[value=="Speaker.entropy"] <- "entropy sen"
+#         value[value=="Hearer.entropy"] <- "entropy rec"
+#         value[value=="Speaker.Voronoiness"] <- "Voronoiness sen"
+#         value[value=="Hearer.Voronoiness"] <- "Voronoiness rec"
+#         value[value=="Expected.utility"] <- "expected utility"
+#     }
+#     return(value)
+# }
+# 
+# 
+# pd <- position_dodge(.01)
+# mysize = 1.65
+# sp = ggplot(dsub, aes(x=Impairment, y=value, color = Number.of.states, shape=Number.of.states, linetype=Number.of.states)) + 
+#     geom_line(color="gray", position = pd) + xlab("imprecision") +
+#      geom_point(position = pd, size = mysize) +
+#      geom_errorbar(aes(ymin=value-ci, ymax=value+ci), width=0, position=pd) +
+#     scale_shape_discrete(name = "# states", guide = guide_legend()) +
+#     scale_linetype_discrete(name = "# states", guide = guide_legend()) +
+#     scale_color_discrete(name = "# states", guide = guide_legend()) +
+#     theme(legend.position="top")
+# 
+# sp = sp + facet_grid(variable ~ Tolerance, scales="free", labeller=myLabeller)
+# 
+# sp + facet_grid(. ~ sex, labeller=mf_labeller)
+# show(sp)
+# 
+# ggsave(sp, file = "paper/version_04/plots/MeanMetrics2.pdf", width=8, height = 10)
+# 
+# ```
 
@@
 
-ggsave(sp, file = "paper/version_02/plots/MeanMetrics3.pdf", width=5, height = 5.5)
+ggsave(sp, file = "paper/version_04/plots/MeanMetrics3.pdf", width=5, height = 5.5)
 
@@
 show(sp)
-ggsave(sp, file = "paper/plots/CategoricalMeasures.pdf", width=8, height = 5)
+ggsave(sp, file = "paper/version_04/plots/CategoricalMeasures.pdf", width=8, height = 5)
 

### 2015-08-11 helper_functions.R

@@ -6,7 +6,8 @@ greyPalette = c("#999999", "#222222", "#999999", "#000000", "#999999", "#000000"
 
 ##################
 ## load data
-data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
+# data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20150729-104740.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
                              'factor','numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','numeric','numeric',

### 2015-08-11 newCode/vagueness-games.py

@@ -160,7 +160,7 @@ def InformationQuantity(SpeakerStrategy, HearerStrategy, Priors):
 
 # # Settings
 
-NStates = 90
+NStates = 5
 PriorDistributionType = 'uniform'
 
 NMessages = 2
@@ -292,24 +292,23 @@ while not converged:
 
     # # Speaker strategy
 
-    PoT = makePDFPerRow(Priors * np.transpose(Confusion))
-    PSigma = np.dot(Confusion, Speaker)
-    PoSender = np.dot(PoT, PSigma)
-    PRho = np.dot(Hearer, ConfusionR)
+    PoT = makePDFPerRow(Priors * np.transpose(Confusion)) # prob actual state (col) given observed state (row)
+    PSigma = np.dot(Confusion, Speaker) # prob mess (col) given actual state (row)
+    PoSender = np.dot(PoT, PSigma) # P_o(m|t_o) prob of seeing agent j do mess (row) given agent i sees t_o
+    PRho = np.dot(Hearer, ConfusionR) # prob receiver plays act (row) given message (column)
     ExpUS = np.array([
         [np.sum([PoT[to, ta] * PRho[m, tr] * Utility[ta, tr] for ta in xrange(NStates) for tr in xrange(NStates)])
-         for m in xrange(NMessages)]
-         for to in xrange(NStates)])
+          for m in xrange(NMessages)]  for to in xrange(NStates)])
     Speaker = makePDFPerRow(PoSender * ExpUS)
 
     # # Hearer strategy
 
-    PoReceiver = np.dot(PRho, Confusion)
+    PoReceiver = np.dot(PRho, Confusion) # P_o(t_o|m)
     PSigmaInverse = makePDFPerRow(Priors * np.transpose(PSigma))
     ExpUR = np.array([
-        [np.sum([PSigmaInverse[m, ta] * ConfusionR[ti, tr] * Utility[ta, tr] for ta in xrange(NStates) for tr in xrange(NStates)])
-         for ti in xrange(NStates)]
-         for m in xrange(NMessages)])
+        [np.sum([PSigmaInverse[m, ta] * ConfusionR[ti, tr] * Utility[ta, tr] for ta in xrange(NStates) for tr in
+                 xrange(NStates)])
+         for ti in xrange(NStates)] for m in xrange(NMessages)])
     Hearer = makePDFPerRow(PoReceiver * ExpUR)
 
[…]

### 2015-08-11 snippets/example_strats/example_strats_plots.R

@@ -1,5 +1,6 @@
 require('plyr')
 require('reshape2')
+require('dplyr')
 theme_set(theme_bw())
 
 source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
@@ -10,7 +11,7 @@ plot.and.save.language.by.index = function(data, index){
   tolerance = data[index,]$Tolerance
   impairment = data[index,]$Impairment
   stratPlot = plot.language.byindex(data,index)
-  filename = paste("~/Desktop/data/svn/vagueness-games/paper/plots/strat_example_", "ind", index, ".pdf", sep = "", collapse = "")
+  filename = paste("~/Desktop/data/svn/vagueness-games/paper/version_04/plots/strat_example_", "ind", index, ".pdf", sep = "", collapse = "")
   pdf(file = filename, height = 2.75, width = 3)
   show(stratPlot)
   dev.off()
@@ -27,7 +28,7 @@ for (strat in stratsToPlot){
 show(ds[stratsToPlot,])
 
 d = subset(data, Number.of.states == 6 & Tolerance == 0.2 & Impairment == 0 & Speaker.Convex.Cat == 1)
-
+# d = filter(data, Number.of.states == 90, Tolerance == 0.1, Impairment == 0.05)
 
 plot.and.save.language.by.index(d, which.min(d$Expected.utility))
 plot.and.save.language.by.index(d, which.max(d$Expected.utility))

### 2015-08-11 snippets/exploring_uniformity_from_impairment/uniformity.R

@@ -7,11 +7,13 @@ theme_set(theme_bw())
 
 source('~/Desktop/data/svn/vagueness-games/helper_functions.R')
 
-data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
+#data = read.csv('~/Desktop/data/svn/vagueness-games/results/20140918-142408.varTolerance.csv', 
+data = read.csv('~/Desktop/data/svn/vagueness-games/results/20150729-104740.csv', 
                 colClasses=c('numeric','factor','numeric','numeric','numeric',
                              'factor','numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','numeric','numeric',
                              'numeric','numeric','numeric','character','character'))
+data = subset(data, data$Number.of.states < 90)
 
 # Hellinger distance.
 
@@ -92,7 +94,7 @@ if (computeD == TRUE){
   D = ddply(dm, .(Number.of.states, Impairment, Tolerance), summarise, 
             inGroupDistance = speaker.vector.distance(value))
 } else{
-  D = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/D.csv")
+  D = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/DNew.csv")
 }
 
 pd <- position_dodge(.01)
@@ -169,9 +171,9 @@ if (computeEU){
   }
   
   close(pb)
-  write.csv(data, paste("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/dataSec",Sys.time() , ".csv", sep = "", collapse = ""))
+  write.csv(data, paste("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/dataSecNew",Sys.time() , ".csv", sep = "", collapse = ""))
 } else{
-  data = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/dataSec.csv")
+  data = read.csv("~/Desktop/data/svn/vagueness-games/snippets/exploring_uniformity_from_impairment/dataSecNew.csv")
 }
 
 
@@ -222,11 +224,11 @@ wgdConcise = ggplot(dsub, aes(x = Impairment, y = inGroupDistance)) +
[…]

### 2015-08-12 paper/version_04/paper.tex
explain new dynamics

@@
 success, in multiple ways. Usually, we would assume that changes in behavior should, at least
-on average, lean towards increasing chances of communicative success. Such a population
-of signaling agents and their behavioral adaptations can be described at different levels of
-abstraction. At the level of individual agents, we can describe a more or less idealized
-process of how each agent adapts dispositions for future actions based on detailed descriptions
-of information available to the agent, e.g., about the game, the other players' behavior and
-the like. More abstractly, at the level of a population of agents, we can describe how
-average behavioral dispositions will evolve. The latter population-level perspective abstracts
-over petty stochastic fluctuations and zooms in on the general tendency or direction of
-evolution that ensues from some (often: simplified) behavior at the agent-level.
-
-We are interested here in exploring how a tendency to favor communicatively successful
-strategies is affected by imprecise perception and realization of states. We therefore look at
-a population-level dynamic that describes the most likely evolutionary path of a population of
-signaling agents who imitate other agents' behavior (proportional to how successful it is) but
-may occasionally confuse states for one another (proportional to how similar they are). We call
-our dynamic \emph{imprecise conditional imitation dynamic} (\icid), because it derives from an
-update protocol based on conditional imitation and allows for imprecise perception and
-realization of states. In the special limiting case where imprecision vanishes, the process is
-just the well-known replicator dynamic. To appreciate the \icid and its motivation, we first
-briefly review the replicator dynamic and its derivation from conditional imitation of
+on average, lean towards increasing chances of communicative success. Such a population of
+signaling agents and their behavioral adaptations can be described at different levels of
+abstraction. At the level of individual agents, we can picture a more or less idealized process
+of how each agent adapts dispositions for future actions based on various pieces of information
+available to the agent, e.g., about the game, the other players' behavior and the like. More
+abstractly, at the level of a population of agents, we can describe how average behavioral
+dispositions will evolve. The population-level perspective abstracts over petty stochastic
+fluctuations and zooms in on the general tendency or direction of evolution that ensues from
+some (often: simplified) behavior at the agent-level.
+
+To better understand the interaction of selective pressure successful communication and
+stochastic noise, we look at a population-level dynamic that describes the most likely
+evolutionary path of a population of signaling agents who imitate other agents' behavior but
+may occasionally confuse states for one another. We call our dynamic \emph{imprecise
+  conditional imitation dynamic} (\icid), because it derives from an update protocol based on
+conditional imitation and allows for imprecise perception and realization of states. In the
+special limiting case where imprecision vanishes, the process is just the well-known replicator
+dynamic \citep{TaylorJonker1978:Evolutionary-St}. To appreciate the \icid and its motivation,
+we first briefly review the replicator dynamic and its derivation from conditional imitation of
 successes.
@@
 $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
-probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
-$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a
-probability of each state $\state \in \States$ being interpreted by the receiver given
-$\messg$. Although behavioral strategies are probabilistic, we may conveniently imagine that
-every individual agent has a non-probabilistic strategy nonetheless. If the population is large
-enough, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability
-that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$.
+probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's are
+functions $\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$
+to a probability of each act $\act \in \Acts$ being chosen in response to $\messg$. Although
+behavioral strategies are probabilistic, we may conveniently imagine that every individual
+agent has a non-probabilistic strategy nonetheless. If the population is large enough, the
+number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability that a randomly
+sampled sender would have a strategy that responds with $\messg$ to $\state$.
 
@@
   \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
-    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
+    \States} \frac{\Pr(\state) \cdot \Sstrat(\messg \probbar \state)}{\sum_{\state'} \Pr(\state') \cdot \Sstrat(\messg \probbar \state')} \cdot
   U_\rec(\state, \messg, \act) \,.
 \end{align*}
-This means that the expected utility for the sender of choosing message $\messg \in \Messgs$,
-given tate $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act
-$\act \in \Acts$, of the probability that the receiver will choose act $\act$ given message
-$\messg$ times the actual utility of using act $\act$ for state $\state$.  The expected utility
-for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a
-sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior
-probability of state $\state$ occurring, times the probability the sender will choose message
-$\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.
-
-% The \emph{fitness} of a behavioral strategy at a choice point is the
-% frequency-weighted average of expected utilities of each choice, given
-% the opponent's strategy:
-% \begin{align*}
-%   \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
-% \EU(\messg, \state,\Rstrat) \\
-% \Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
-% \cdot \EU(\act, \messg,\Sstrat)\,.
-% \end{align*}
-
-The discrete-time replicator dynamic maps current strategies $\Sstrat$ and $\Rstrat$ to future
-strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$ in such a way that changes in frequency are
-proportional to expected utilities:
+Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
+outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
+sender of choosing message $\messg \in \Messgs$, given tate $\state \in \States$ and a receiver
+strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability that the
+receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
+$\act$ for state $\state$, given $\state$ and $\messg$.
+
+The continuous-time replicator dynamic for behavioral strategies describes the mean rate of
+change in the proportion of each choice, independently for each choice point:
+\begin{align}
+  \label{eq:RDContinuous}
+  \dot{\Sstrat}(\messg \probbar \state) & = \underbrace{\Sstrat(\messg \probbar
+    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
+  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
+      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
+      EU at choice point $\state$}} \Big ) \\
+  \dot{\Rstrat}(\act \probbar \messg) & = \Rstrat(\act \probbar \messg) \thinspace \Big ( \EU(\act,
+  \messg, \Sstrat) -  \sum_{\act'} \Rstrat(\act' \probbar \messg) \cdot \EU(\act',
+  \messg,\Sstrat)  \Big ) \nonumber
+\end{align}
+The rate of change in proportion of $\messg$ choices at choice point $\state$, for instance, is
+given by the difference of $\messg$'s expected utility at $\state$ and the average expected
+utility at $\state$, times the probability of seeing $\messg$ at $\state$ in the first place.
+
+For numerical simulations, it is often more convenient to work with a discrete-time
+formulation, which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\RD(\Sstrat)$
+and $\RD(\Rstrat)$ that describe the population at the next discrete time step. Like the
+continuous-time version, the discrete-time replicator dynamics tracks changes in frequency of
+choices as proportional to their expected utilities
+\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
+  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
+  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
+  utilities are always positive.}
 \begin{align}
@@
 point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
-the population \emph{before} the update is $\Sstrat{\messg \probbar \state}$. \emph{After} the
+the population \emph{before} the update is $\Sstrat(\messg \probbar \state)$. \emph{After} the
 update it will be proportional to the frequency before the update times the expected utility of
 $\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
-obtain probabilities that sum to one from this. In particular, the denominator in the
+obtain probabilities that sum to one. In particular, the denominator in the
 definition above captures the average expected utility of all choices at the relevant choice
@@
 
-The replicator dynamic is an abstract population-level dynamic that describes the mean,
-to-be-expected change in behavioral dispositions in a population of signalers. There are
-several ways of deriving the replicator dynamic from agent-level processes of behavioral
-adaptation. We focus here on one of the simplest: \emph{conditional imitation of success}
-\citep[see][for this and many other ways of deriving the replicator
-dynamics]{Sandholm2010:Population-Game}. We choose this agent-level derivation, because it is
-more convenient than its alternatives when we want to add imprecision in the following. The
-main idea is that agents imitate the behavior of another agent at some choice point with a
-probability that is proportional to the expected utility of the latter agent's act for that
-choice point. Since the sim-max games that we will be looking at have all expected utilities
-conveniently lie between 0 and 1, we may even identify the switching probability with the
-expected utility directly.
+The replicator dynamic is an abstract population-level dynamic that describes the mean expected
+change of behavioral dispositions in a population of signalers. There are several ways of
+deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
+here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
+many other derivations]{Sandholm2010:Population-Game}. We choose this agent-level derivation,
+because it is more convenient than its alternatives when we want to add imprecision in the
+following. The main idea is that agents imitate the behavior of another agent at some choice
+point with a probability that is proportional to the expected utility of the latter agent's act
+for that choice point. Since the sim-max games that we will be looking at in
+Section~\ref{sec:exploring-rdd} have positive utilities, we may even identify the switching
+probability with the expected utility directly.
 
@@
 to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
-chance to change behavior at choice point $\state$. The probability that that learner plays
+chance to change behavior at choice point $\state$. The probability that our learner plays
 $\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
@@
       \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
-& = \underbrace{\Sstrat(\messg \probbar \state)}_{\text{frequency of $\messg$ at $\state$}}  ( \underbrace{\EU(\messg,
-  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
-      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
-      EU at choice point $\state$}}  ) 
-\end{align*}
-This is the continuous-time formulation of the replicator dynamics for behavioral
-strategies. The discrete time formulation in Equation (\ref{eq:RDDiscrete}) solves this
-continuous-time formulation by assuming that discrete update steps are infinitesimally
-small:\footnote{Notice that we drop the denominator here, to obtain the non-payoff corrected
-  replicator dynamic.}
-\begin{align*}
-  \dot{\Sstrat}(\messg \probbar \state) & = \RD(\Sstrat)(\messg \probbar \state) -
-  \Sstrat(\messg \probbar \state) \\
-  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat)} {\sum_{\messg'}
-    \Sstrat(\messg' \probbar \state) \cdot
-    \EU(\messg', \state,\Rstrat)} - \Sstrat(\messg \probbar \state) \\
-  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat) - \Sstrat(\messg
-    \probbar \state) \cdot \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
-    \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
-    \EU(\messg', \state,\Rstrat)}  \\
-  & \propto \Sstrat(\messg \probbar \state) \cdot \left ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
-  \state,\Rstrat) \right )
-\end{align*}
-
-\bigskip
-
- Suppose, she currently plays $c$ at $k$. She
-observes what a randomly sampled teacher would do at choice point $k$, say $c'$ and switches
-from $c$ to $c'$ with switch probability $\text{SWITCH}(c \rightarrow c' , k)$, which we can
-here identify with the expected utility $\EU(c',k)$ of $c'$ at $k$. The mean change induced by
-offering all agents imitation chances for $k$ is:
-\begin{align*}
-  \text{mean change of frequency $c$, at $k$} & = \text{probability of switchers $c'
-    \rightarrow c$ at $k$ \mygray{(inflow)}} \\
-& - \text{probability of switchers $c \rightarrow c'$ at $k$ \mygray{(outflow)}}
-\end{align*}
-We should spell out the inflow and outflow here with some care, so as to prepare ourselves for
-the subsequent addition of imprecise perception and realization of stimuli. So in general, we
-would define:
-\begin{align*}
-  \text{inflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c'$} \\
-  & \ \ \ \times \text{prob.~learner observes $c$ by random teacher} \\
-  & \ \ \ \times \EU(c', k) \\
-  \text{outflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c$} \\
-  & \ \ \ \times \text{prob.~learner observes random teacher play $c'$ at $k$} \\
-  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k)
-\end{align*}
-
-Suffice it to look at the case of a sender.
+& = \Sstrat(\messg \probbar \state) \cdot \Big ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
+  \state,\Rstrat) \Big )\,,
+\end{align*}
+which is the continuous-time formulation of the replicator dynamic in (\ref{eq:RDContinuous}).
+
+\subsection{Imitation of success with noise}
+\label{sec:imit-succ-with}
+
+The above derivation of the replicator dynamic assumes that agents can discriminate choices and
+choice points perfectly. Let's dispense with that assumption. With an eye towards sim-max
+games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
+any two states may have a chance of being confused for one another, proportional to how similar
+they are. This will affect the way that agents realize the (pure) strategies that they carry
+and also how learners observe teacher choices and the probability with which they switch to the
+observed teacher choice.
+
+Agents carry pure dispositions to act, but noise can affect their realization. As a sender,
+every agent maps states to messages: these are subjectively perceived states, and no longer
+necessarily also the actually occurring state. As a receiver, every agent maps messages to
+state interpretations: these are intended interpretations that need not always be faithfully
+realized. This means that behavioral strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and
+$\Rstrat \in (\Delta(\States))^\Messgs$ represent the average proportions of \emph{actual}
+behavioral dispositions in the population, the realization and observation of which can be
+distorted by confusion of similar states.
+
+If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
+probability that a given agent observes state $\mystate{o}$. States $\mystate{a}$ and
+$\mystate{o}$ may be identical. In the limiting case, $P_o(\mystate{o} \probbar \mystate{a})$
+could be one only if $\mystate{a}$ and $\mystate{o}$ are identical. For the more general case,
+we should treat $P_o(\mystate{o} \probbar \mystate{a})$ as a primitive of the model. Similarly,
+if a given receiver intends to select interpretation $\mystate{i}$, we will assume that the
+probability with which state $\mystate{r}$ is realized is
+$P_r(\mystate{r} \probbar \mystate{i})$. Section~\ref{sec:exploring-rdd} gives a concrete
+specification of $P_o$ and $P_r$ in terms of similarity of states.
+
+A single round of play is then governed by five pieces of stochastic information, where
+previously there were only three. At the outset, the actual state is chosen with probability
+$P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with probability
+$P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
+$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
+state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
+to realize interpretation $\mystate{i}$ with probability
+$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
+$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.
+
+Expected utilities of choices at choice points should take into account as well that actual
+states need not be observed states, and intended interpretations need not be realized
+interpretations. First, note that 
+\begin{align*}
+  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) = \frac{P_a(\mystate{a}) P_o(\mystate{o}
+    \probbar \mystate{a})}{\sum_{\state}P_a(\state) P_o(\mystate{o}
+    \probbar \state)}
+\end{align*}
+is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed. The probability
+that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
+\begin{align*}
+  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
+  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
+\end{align*}
+The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
+\begin{align*}
+  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) = \frac{P_a(\mystate{a})
+  P_{\Sstrat}(\messg \probbar \mystate{a})}{\sum_{\state} P_a(\state)
+  P_{\Sstrat}(\messg \probbar \state)}\,. 
+\end{align*}
+The probability that $\mystate{r}$ is realized by a random receiver in response to
+message $\messg$ is:
+\begin{align*}
+  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
+  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
+\end{align*}
+This lets us capture the expected utilities for observed states (sender) and intended
+interpretations (receiver) by taking into consideration what the likely actual states and
+realized interpretations will be:
+\begin{align*}
+  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
+  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
+  \messg) \utils(\mystate{a}, \messg, \mystate{r})\,, \\
+  \EU(\mystate{i} , \messg, \Sstrat) & = \sum_{\mystate{a}}
+  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
+  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})\,.
+\end{align*}
+If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
+observability and realizability of non-identical states, above definitions reduce to the
+previous definitions of expected utilities. This also legitimates the overload of notation.
+
+Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
+the dynamic that ensues from imitation of successes. Since imprecision works differently on
+senders and receivers (the former confuse choice points, the latter confuse choices), we need
+to look separately at each case.
+
+As before, suppose that senders receive a chance to change their behavior independently for a
+given choice point. In the present case, this would be a chance to change how to respond to a
+subjectively perceived state $\mystate{o}$, which need not be the actual one. We must therefore
+take into account what the probability is that, given that the learner observed $\mystate{o}$,
+he will observe a randomly sampled teacher play some $\messg$ or other. The probability
+$P_o(\messg \probbar \mystate{o})$ that some teacher plays $\messg$ in the same situation in
+which some learner observes $\mystate{o}$ is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as
+defined above):
+\begin{align*}
+  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
+  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
+\end{align*}
+The probability $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches
+from $\messg'$ to $\messg$ in subjectively perceived state $\mystate{o}$ is therefore:
+\begin{align*}
+  P(\messg' \rightarrow \messg, \mystate{o}) = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
+    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
+  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
+  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
+\end{align*}
+The mean change to the proportion of $\messg$ choices at state $\state$ are then as before:
+\begin{align*}
+  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
+  \rightarrow m', \mystate{o}) \\
+  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
+  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \\
+  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)
+\end{align*}
+This has an elegant and practical discrete-time solution:\footnote{To see how the discrete-time formulation
+  gives rise to the continuous-time formulation, let's assume that update steps are
+  infinitesimally small, so that
+  \begin{align*}
+    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
+    \probbar \state)\\ 
+    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
+      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
+      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
+      \EU(\messg', \state, \Rstrat)}\,.
+  \end{align*}
+By dropping the nominator, which is constant for all $\messg$ for fixed
+$\state$, we obtain the above continuous-time formulation.}
+\begin{align*}
+  \Sstrat'(\messg \probbar \state) = \frac{P_o(\messg \probbar \state) \myts \EU(\messg,
+  \state, \Rstrat)}{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg',
+  \state, \Rstrat)}\,.
+\end{align*}
+
+The case of the receiver is mostly analogous. Presented with an update opportunity for choice
+point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
+with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
+  that applies to the sender's observations also applies to the receiver's observations when
+  trying to imitate other agents' strategies. We could easily make the model more complex and
+  introduce another measure of noise for state confusability during receiver attempts to
+  imitate. We refrain from it here, because we see no immediate theoretical gain.}
+\begin{align*}
+  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
+  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
+  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
+  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
+\end{align*}
+The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
+from using $\state'$ to $\state$ in response to $\messg$ is therefore:
+\begin{align*}
+  P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
+    \messg)}_{\text{learner plays $\state'$}} \cdot
+  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
+  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
+\end{align*}
+This gives rise to continuous-time and discrete-time formulations of resulting dynamics,
+parallel to those for the sender case:
+\begin{align*}
+  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg) \myts
+  P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'} \Rstrat(\state \probbar \messg) \myts
+  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \\
+  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar \messg) \myts \sum_{\state'} 
+  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat)\\
+  \Rstrat'(\state \probbar \messg) & = \frac{P_o(\state \probbar \messg) \myts \EU(\state,
+  \messg, \Sstrat)}{\sum_{\state'} P_o(\state' \probbar \messg) \myts \EU(\state',
+  \messg, \Sstrat)}
+\end{align*}
+
+\newpage
 

### 2015-08-12 paper/version_04/paper.tex
explain new dynamics
dynamics write up

@@
 success, in multiple ways. Usually, we would assume that changes in behavior should, at least
-on average, lean towards increasing chances of communicative success. Such a population of
-signaling agents and their behavioral adaptations can be described at different levels of
-abstraction. At the level of individual agents, we can picture a more or less idealized process
-of how each agent adapts dispositions for future actions based on various pieces of information
-available to the agent, e.g., about the game, the other players' behavior and the like. More
-abstractly, at the level of a population of agents, we can describe how average behavioral
-dispositions will evolve. The population-level perspective abstracts over petty stochastic
-fluctuations and zooms in on the general tendency or direction of evolution that ensues from
-some (often: simplified) behavior at the agent-level.
-
-To better understand the interaction of selective pressure successful communication and
-stochastic noise, we look at a population-level dynamic that describes the most likely
+on average, lean towards increasing chances of communicative success. Such behavioral
+adaptations can be described at different levels of abstraction. At the level of individual
+agents, we can picture a more or less idealized process of how each agent adapts dispositions
+for future actions based on various pieces of information available to the agent, e.g., about
+the game, the other players' behavior and the like. More abstractly, at the level of a
+population of agents, we can describe how average behavioral dispositions will evolve. The
+population-level perspective abstracts over petty stochastic fluctuations and zooms in on the
+general tendency or direction of evolution that ensues from some (often: simplified) behavior
+at the agent-level.
+
+To better understand the interaction of confusability of stimuli and selective pressure towards
+successful communication, we look at a population-level dynamic that describes the most likely
 evolutionary path of a population of signaling agents who imitate other agents' behavior but
@@
 receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
-$\act$ for state $\state$, given $\state$ and $\messg$.
+$\act$, given $\state$ and $\messg$.
 
@@
 
-For numerical simulations, it is often more convenient to work with a discrete-time
-formulation, which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\RD(\Sstrat)$
-and $\RD(\Rstrat)$ that describe the population at the next discrete time step. Like the
-continuous-time version, the discrete-time replicator dynamics tracks changes in frequency of
-choices as proportional to their expected utilities
-\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
-  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
-  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
-  utilities are always positive.}
+For numerical simulations, it can be more convenient to work with a discrete-time formulation,
+which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\Sstrat'$ and $\Rstrat'$
+that describe the population at the next discrete time step. Like the continuous-time version,
+the discrete-time replicator dynamics tracks changes in frequency of choices as proportional to
+their expected utilities \citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The
+  formulation given in Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one
+  we will be looking at in Section~\ref{sec:exploring-rdd}, where utilities are always
+  non-negative and expected utilities are always positive.}\textsuperscript{,}\footnote{The
+  notation $\propto$, for ``proportional to'', means that the right-hand side must possibly yet
+  be normalized so as to have probabilities sum to one. In general, writing $P(x) \propto f(x)$
+  for any function $f \mycolon X \rightarrow \mathds{R}$ is shorthand for
+  $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. The implicit normalizing denominators for the
+  replicator equations are average expected utilities, e.g., for the sender:
+  $\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg', \state,\Rstrat)$.}
 \begin{align}
   \label{eq:RDDiscrete}
-  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
-    \EU(\messg, \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
-   \EU(\messg', \state,\Rstrat)} \\
-    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
-    \EU(\act, \messg,\Sstrat)} {\sum_{\act'} \Rstrat(\act' \probbar \messg)
-    \cdot \EU(\act', \messg,\Sstrat)}  \,. \nonumber
+  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \cdot
+    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \cdot
+    \EU(\act, \messg,\Sstrat)\,.   
 \end{align}
-Intuitively put, this definition updates the average frequency with which each choice (message
-or act) will occur, independently at each choice point (state or message). For a given choice
-point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
-the population \emph{before} the update is $\Sstrat(\messg \probbar \state)$. \emph{After} the
-update it will be proportional to the frequency before the update times the expected utility of
-$\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
-obtain probabilities that sum to one. In particular, the denominator in the
-definition above captures the average expected utility of all choices at the relevant choice
-point.
+For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
+average agent in the population \emph{after} the update is proportional to the probability of
+seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
+expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
+by a gradient of current frequency and a measure of how good they are.
 
@@
 here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
-many other derivations]{Sandholm2010:Population-Game}. We choose this agent-level derivation,
-because it is more convenient than its alternatives when we want to add imprecision in the
-following. The main idea is that agents imitate the behavior of another agent at some choice
-point with a probability that is proportional to the expected utility of the latter agent's act
-for that choice point. Since the sim-max games that we will be looking at in
-Section~\ref{sec:exploring-rdd} have positive utilities, we may even identify the switching
-probability with the expected utility directly.
-
-Let's consider sender strategies, as the receiver case is parallel (as long as there is no
-noise or imprecision). Revision by imitation of success proceeds as follows. Call
-(misleadingly!) the agent who gets a chance to change behavior ``learner'' and the possibly
-to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
-chance to change behavior at choice point $\state$. The probability that our learner plays
-$\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
-teacher does at $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar
-\state)$. The learner then starts using $\messg'$ instead of $\messg$ with probability
-$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)
-
-If agents get repeated updating chances for their choice points, the \emph{expected change} of
+many other derivations]{Sandholm2010:Population-Game}. The main idea is that agents imitate the
+behavior of another agent at some choice point with a probability that is proportional to the
+expected utility of the latter agent's act for that choice point. Since the sim-max games that
+we will be looking at in Section~\ref{sec:exploring-rdd} have positive utilities, we may
+identify the switching probability with the expected utility.
+
+Let's consider sender strategies, as the receiver case is parallel. Revision by imitation of
+success proceeds as follows. Call (misleadingly!) the agent who gets a chance to change
+behavior ``learner'' and the possibly to-be-imitated agent ``teacher.'' A random learner is
+drawn from the population and given a chance to change behavior at choice point $\state$. The
+probability that our learner plays $\messg$ is $\Sstrat(\messg \probbar \state)$. The learner
+observes what a randomly sampled teacher does at $\state$. That would be $\messg'$ with
+probability $\Sstrat(\messg' \probbar \state)$. The learner then starts using $\messg'$ instead
+of $\messg$ with probability $\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and
+$\messg$ could be the same.)
+
+If agents get repeated update chances for their choice points, the \emph{expected change} of
 frequency for $\messg$ at $\state$ becomes:
@@
 that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
-population, these probabilities can be given as:
+population, the probability $P(\messg' \rightarrow m, \state)$ of switching from $\messg'$ to
+$\messg$ at $\state$ can be written as:
 \begin{align*}
@@
   \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
-  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
-  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
-    \state)}_{\text{learner plays $\messg$}} \cdot
-  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
-  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
-\end{align*}
-
+  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} 
+  % P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
+  %   \state)}_{\text{learner plays $\messg$}} \cdot
+  % \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
+  % \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
+\end{align*}
 From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
@@
 games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
-any two states may have a chance of being confused for one another, proportional to how similar
-they are. This will affect the way that agents realize the (pure) strategies that they carry
-and also how learners observe teacher choices and the probability with which they switch to the
-observed teacher choice.
-
-Agents carry pure dispositions to act, but noise can affect their realization. As a sender,
-every agent maps states to messages: these are subjectively perceived states, and no longer
-necessarily also the actually occurring state. As a receiver, every agent maps messages to
-state interpretations: these are intended interpretations that need not always be faithfully
-realized. This means that behavioral strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and
-$\Rstrat \in (\Delta(\States))^\Messgs$ represent the average proportions of \emph{actual}
-behavioral dispositions in the population, the realization and observation of which can be
-distorted by confusion of similar states.
+any two states, but not messages, may be confused for one another.\footnote{It is relatively
+  straightforward to also incorporate confusability of messages, but this is irrelevant to our
+  present purposes.} Confusability of states will affect how agents behave (on average), how
+they perceive others that they might want to imitate, and how good any behavioral disposition
+actually is.
+
+To keep matters simple, let us assume that agents carry pure dispositions to act, but noise can
+affect their realization. As a sender, every agent maps states to messages: these are
+subjectively perceived states, and no longer necessarily also the actually occurring state. As
+a receiver, every agent maps messages to state interpretations: these are intended
+interpretations that need not always be faithfully realized. This means that behavioral
+strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and $\Rstrat \in (\Delta(\States))^\Messgs$
+represent the average proportions of \emph{actual} behavioral dispositions in the population,
+the realization and observation of which can be distorted by confusion of similar states.
 
 If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
-probability that a given agent observes state $\mystate{o}$. States $\mystate{a}$ and
-$\mystate{o}$ may be identical. In the limiting case, $P_o(\mystate{o} \probbar \mystate{a})$
-could be one only if $\mystate{a}$ and $\mystate{o}$ are identical. For the more general case,
-we should treat $P_o(\mystate{o} \probbar \mystate{a})$ as a primitive of the model. Similarly,
-if a given receiver intends to select interpretation $\mystate{i}$, we will assume that the
-probability with which state $\mystate{r}$ is realized is
-$P_r(\mystate{r} \probbar \mystate{i})$. Section~\ref{sec:exploring-rdd} gives a concrete
-specification of $P_o$ and $P_r$ in terms of similarity of states.
-
-A single round of play is then governed by five pieces of stochastic information, where
-previously there were only three. At the outset, the actual state is chosen with probability
-$P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with probability
-$P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
+probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
+intends to select interpretation $\mystate{i}$, we will assume that the probability with which
+state $\mystate{r}$ is realized is $P_r(\mystate{r} \probbar
+\mystate{i})$.
+Section~\ref{sec:exploring-rdd} gives a concrete specification of $P_o$ and $P_r$ in terms of
+similarity of states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary
+conditional confusion probabilities.
+
+A single round of play of a sim-max game is then governed by five pieces of stochastic
+information, where previously there were only three. At the outset, the actual state is chosen
+with probability $P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with
+probability $P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
 $\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
@@
 
-Expected utilities of choices at choice points should take into account as well that actual
-states need not be observed states, and intended interpretations need not be realized
+Expected utilities of choices at choice points should also take into account that actual states
+need not be observed states, and intended interpretations need not be realized
 interpretations. First, note that
 \begin{align*}
-  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) = \frac{P_a(\mystate{a}) P_o(\mystate{o}
-    \probbar \mystate{a})}{\sum_{\state}P_a(\state) P_o(\mystate{o}
-    \probbar \state)}
+  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
+    \probbar \mystate{a})
 \end{align*}
@@
 \begin{align*}
-  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) = \frac{P_a(\mystate{a})
-  P_{\Sstrat}(\messg \probbar \mystate{a})}{\sum_{\state} P_a(\state)
-  P_{\Sstrat}(\messg \probbar \state)}\,. 
+  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
+  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
 \end{align*}
@@
 given choice point. In the present case, this would be a chance to change how to respond to a
-subjectively perceived state $\mystate{o}$, which need not be the actual one. We must therefore
-take into account what the probability is that, given that the learner observed $\mystate{o}$,
-he will observe a randomly sampled teacher play some $\messg$ or other. The probability
-$P_o(\messg \probbar \mystate{o})$ that some teacher plays $\messg$ in the same situation in
-which some learner observes $\mystate{o}$ is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as
-defined above):
+subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
+consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
+observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
+$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
 \begin{align*}
@@
 \end{align*}
-The mean change to the proportion of $\messg$ choices at state $\state$ are then as before:
-\begin{align*}
+The mean change to the proportion of $\messg$ choices at state $\state$ are then:
+\begin{align}
+  \label{eq:nRDSender}
   \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
-  \rightarrow m', \mystate{o}) \\
-  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
-  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \\
-  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)
-\end{align*}
-This has an elegant and practical discrete-time solution:\footnote{To see how the discrete-time formulation
-  gives rise to the continuous-time formulation, let's assume that update steps are
-  infinitesimally small, so that
-  \begin{align*}
-    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
-    \probbar \state)\\ 
-    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
-      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
-      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
-      \EU(\messg', \state, \Rstrat)}\,.
-  \end{align*}
-By dropping the nominator, which is constant for all $\messg$ for fixed
-$\state$, we obtain the above continuous-time formulation.}
-\begin{align*}
-  \Sstrat'(\messg \probbar \state) = \frac{P_o(\messg \probbar \state) \myts \EU(\messg,
-  \state, \Rstrat)}{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg',
-  \state, \Rstrat)}\,.
-\end{align*}
+  \rightarrow m', \mystate{o}) \nonumber \\
+  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
+  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
+  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
+  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
+\end{align}
 
@@
 \end{align*}
-This gives rise to continuous-time and discrete-time formulations of resulting dynamics,
-parallel to those for the sender case:
-\begin{align*}
-  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg) \myts
-  P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'} \Rstrat(\state \probbar \messg) \myts
-  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \\
-  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar \messg) \myts \sum_{\state'} 
-  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat)\\
-  \Rstrat'(\state \probbar \messg) & = \frac{P_o(\state \probbar \messg) \myts \EU(\state,
-  \messg, \Sstrat)}{\sum_{\state'} P_o(\state' \probbar \messg) \myts \EU(\state',
-  \messg, \Sstrat)}
-\end{align*}
-
-\newpage
-
-\subsection{Replicator diffusion dynamic in behavioral strategies}
-\label{sec:repl-diff-dynam-1}
-
-The replicator diffusion dynamic is a noise-perturbed variant of the
-\rd in behavioral strategies. Fix a sim-max game with $\States =
-\Acts$ and a confusion matrix $C : \States \times \States$, where $C$ is
-row-stochastic, i.e., each row is a probability
-distribution. Elements $C_{ij}$ give the probability that $\state_i$
-is realized as $\state_j$. The confusability of states affects senders
-and receivers alike, but in slightly different ways. For the sender,
-$C_{ij}$ is the probability that the actual state \mystate{i} is
-perceived as \mystate{j}. For the receiver, $C_{ij}$ is the
-probability that \mystate{j} is the interpretation that is actually
-formed when \mystate{i} is the intended interpretation.
-See Figure~\ref{fig:noise-perturbation-of-strategies} for an illustration.
-
-\begin{figure}
-  \centering
-
-    \begin{tikzpicture}[node distance = 1.6cm, thick]
-
-      \begin{scope}
-
-      \node[rectangle, draw=black!50, fill=black!10, thick] (actual)
-      {actual state $\mystate{i}$};
-
-      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
-      actual] (perceived) {perceived state $\mystate{j}$};
-
-      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
-      perceived] (output) {chosen message $\messg$};
-
-      \node[rectangle, thick, above of = actual, node distance=0.75cm]
-      (sender) {sender};
-
-      \draw[->] (actual) -> (perceived) node[midway,left] {noise
-        $C_{ij}$};
-
-      \draw[->] (perceived) -> (output) node[midway,left] (label)
-      {strategy $\Sstrat(\messg \probbar \mystate{j})$};
-
-%      \begin{pgfonlayer}{background}
-%        \node [draw=black!50, fill=black!20,fit=(actual) (label)
-%        (output)] {};
-%      \end{pgfonlayer}
-    \end{scope}
-
-      \begin{scope}[xshift=5cm]
-
-      \node[rectangle, draw=black!50, fill=black!10, thick] (message)
-      {observed message $\messg$};
-
-      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
-      message] (target) {target interpretation $\mystate{i}$};
-
-      \node[rectangle, draw=black!50, fill=black!10, thick, below of =
-      target] (realized) {realized interpretation $\mystate{j}$};
-
-      \node[rectangle, thick, above of = message, node distance=0.75cm]
-      (receiver) {receiver};
-
-      \draw[->] (message) -> (target) node[midway,right] (bla) {strategy
-        $\Rstrat(\mystate{i} \probbar \messg)$};
-
-      \draw[->] (target) -> (realized) node[midway,right]  {noise $C_{ij}$};
-
-%      \begin{pgfonlayer}{background}
-%        \node [draw=black!50, fill=black!20,fit = (message) 
-%        (realized) (bla)] {};
-%      \end{pgfonlayer}
-    \end{scope}
-
-  \end{tikzpicture}
-
-  \caption{Effect of confusion of states on sender and receiver
-    choices.}
-  \label{fig:noise-perturbation-of-strategies}
-\end{figure}
-
-The aggregate effect of confusion of states on behavioral strategies
-can be captured in a function that maps behavioral strategies
-$\Sstrat$ and $\Rstrat$ to their \emph{diffusions} $D(\Sstrat)$ and
-$D(\Rstrat)$. Suppose that $\Sstrat$ and $\Rstrat$ represent a
-population's average signaling behavior if there was no confusion of
-states. Then $D(\Sstrat)$ and $D(\Rstrat)$ give the expected aggregate
-population behavior under the actual noise-perturbed behavior of each
-agent. Since the effect of confusion of states is that behavior at one
-choice point percolates to behavior at similar choice points, we speak
-of \emph{diffusion of behavior under confusion of states}. If we
-conceive of behavioral strategies $\Sstrat$ and $\Rstrat$ as
-row-stochastic matrices (rows are choice points, columns are choices;
-each row is a probability distribution), we can define the diffusion effect of
-confusion matrix $C$ as matrix multiplication:
+Parallel to those for the sender case, this gives rise to:
 \begin{align}
-  \label{eq:confusion-function}
-  \Diff_C(\Sstrat) & = C \Sstrat &    \Diff_C(\Rstrat) & = \Rstrat C\,.
+  \label{eq:nRDReceiver}
+  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
+  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
+  \Rstrat(\state \probbar \messg) \myts
+  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
+  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
+  \messg) \myts \sum_{\state'}
+  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
 \end{align}
-For example, diffusion thus defined determines that, on the diffused sender strategy, the
-probability that message $\messg$ is chosen given a state $\mystate{i}$ is given by
-$\sum_{j} C_{ij} \cdot \Sstrat(\messg \probbar \mystate{j})$.  This means that the original
-probability given by $\Sstrat$ is increased by the original probability of this message being
-chosen in other states $\mystate{j}$, proportionally to how likely it is that $\mystate{i}$ is
-(mistakenly) perceived as $\mystate{j}$.
-
-The discrete-time replicator diffusion dynamic takes the replicator
-dynamic as basic, but factors in the confusion of states at each step
-in a sequential manner:
+
+The continuous-time formulations in Equations~(\ref{eq:nRDSender}) and (\ref{eq:nRDReceiver})
+have elegant and practical discrete-time solutions in:\footnote{To see how the discrete-time
+  formulation gives rise to the continuous-time formulation, let's assume that update steps are
+  infinitesimally small, so that
+  \begin{align*}
+    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
+    \probbar \state)\\ 
+    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
+      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
+      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
+      \EU(\messg', \state, \Rstrat)}\,.
+  \end{align*}
+By dropping the denominator, which is constant for all $\messg$ for fixed
+$\state$, we obtain the above continuous-time formulation.}
 \begin{align}
-  \label{eq:RDD-Def}
-  \RDD(\Sstrat) & = \Diff_C(\RD(\Sstrat)) &   \RDD(\Rstrat) & = \Diff_C(\RD(\Rstrat)) \,.
+  \label{eq:nRD}
+  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
+  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
+  \myts \EU(\state, \messg, \Sstrat)\,.
 \end{align}
-This is equivalent to the following, perhaps more transparent
-formulation:\footnote{To see this, start with $\RDD(\Sstrat)(\messg_k
-  \probbar \state_i)$. If $\RDD(\Sstrat)$ is considered as a matrix,
-  we can rewrite this as $(RDD(\Sstrat)_{ik}$. By
-  Equation~(\ref{eq:confusion-function}), we get $(C \
-  \RD(\Sstrat))_{ik}$, which by definition of matrix multiplication is
-  $\sum_j C_{ij} \cdot \RD(\Sstrat)_{jk}$. Substituting the definition
-  of $\RD(\Sstrat)$ completes the derivation. Similarly for the
-  receiver.}
-\begin{align*}
-  \RDD(\Sstrat)(\messg_k \probbar \state_i) & = \sum_{j} C_{ij} \cdot
-  \frac{\Sstrat(\messg_k \probbar \state_j) \cdot
-    \EU(\messg_k, \state_j,\Rstrat)}
-  {\Phi(\state_j,\Sstrat, \Rstrat)} \\
-    \RDD(\Rstrat)(\state_i \probbar \messg_k) & = \sum_{j} C_{ji} \cdot
-  \frac{\Rstrat(\state_j \probbar \messg_k) \cdot
-    \EU(\state_j, \messg_k,\Sstrat)} {\Phi(\messg_k,\Sstrat, \Rstrat)}  \,.
-\end{align*}
-
-The most straightforward interpretation of this sequential definition is that the \rdd captures
-the most likely path of development of aggregate behavior in a society, in which behavioral
-changes occur independently at each choice point, as described by the \rd for behavioral
-strategies. Perceptual noise affects the realization of behavior, but not the fitness-based
-selection thereof. Alternative noise-perturbed versions of the replicator dynamic in which the
-fitness-based selection of strategies takes noise into account are conceivable, but we must put
-the exploration of alternatives aside here. We suggest that our approach is practical,
-conservative and plausible: it is practical because it is easy to implement; it is conservative
-because it connects the \rdd with the \rmd (see Appendix~\ref{sec:diffusion-as-special}); it is
-also plausible because the \rdd can be derived from a myopic update protocol in which agents
-themselves are unaware of perceptual noise (see Appendix~\ref{sec:deriv-rdd-imit}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Exploring the RDD}
+This looks very much like the discrete-time formulation of the standard replicator dynamic in
+(\ref{eq:RDDiscrete}). A minor difference is that expected utility here takes stochastic
+confusability of states into account. The main difference is that here we are not looking at,
+say, the probability $\Sstrat(\messg \probbar \state)$ with which $\messg$ is be played in
+response to $\state$ on average in the population, but at the probability $P_o(\messg \probbar
+\state)$ of seeing one agent pick $\messg$ given that another agent has observed
+$\state$. Similarly for the receiver, we are not looking at occurrence probability
+$\Sstrat(\state \probbar \messg)$ but at the probability $P_o(\state \probbar \messg)$ with
+which an agent who observes $\messg$ observes another agent play $\state$. It is easy to check
+formally and intuitive to appreciate that this major difference vanishes if confusion
+probabilities $P_o$ and $P_r$ are trivial, i.e., assigning zero probability to all pairs of
+non-identical states.
+
+\section{Exploring noisy imitation}
 \label{sec:exploring-rdd}
 
-The \rdd is a two-step procedure: first calculate where the replicator dynamic takes us, then
-apply diffusion. To understand the \rdd better, we isolate the novel diffusion part in
-Section~\ref{sec:iterated-diffusion}, before exploring how diffusion and fitness-based
-selection interact in Section~\ref{sec:simulations}.
-
-\subsection{Iterated diffusion}
+How does a tendency to confuse similar states interact with a drive towards more efficient
+signaling strategies under the imprecise conditional imitation dynamic? To be able to properly
+address this question, it pays to be more precise about similarity of states, the conditional
+confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max games.
+
+\subsection{Confusion \& utilities}
 \label{sec:iterated-diffusion}
 
-Confusion of states should intuitively be a function of their
-perceptual similarity: the more similar two states are, the more likely they could be mistaken for each other.
-To make this concrete, let us assume that the
-state space of the sim-max game consists of $n \ge 2$ states that are
-equally spaced across the unit interval, including $0$ and $1$. The
-distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical
-similarity between two states $\mystate{i}$ and
-$\mystate{j}$. Distance in physical space feeds into a perceptual
-similarity function, as described by
-\citet{Nosofsky1986:Attention-Simil}:
+Confusion of states should intuitively be a function of their perceptual similarity: the more
+similar two states are, the more likely they could be mistaken for each other.  To make this
+concrete, let us assume that the state space of the sim-max game consists of $n \ge 2$ states
+that are equally spaced across the unit interval, including $0$ and $1$. The distance
+$\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between two states
+$\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual similarity
+function, as described by \citet{Nosofsky1986:Attention-Simil}:
 \begin{align*}
@@
 \end{align*}
-where $\imprecision \ge 0$ is a imprecision parameter. When $\imprecision=0$ agents perfectly
+where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
 discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
@@
 
-We further assume that the probability of confusing any two states
-$\mystate{i}$ and $\mystate{j}$ is proportional to their perceived
-similarity and therefore obtained by normalization:
-\begin{align*}
-  C_{ij} = \frac{\similarity(\mystate{i}, \mystate{j} \myts ; \myts
-  \imprecision)} {\sum_j \similarity(\mystate{i}, \mystate{j} \myts ; \myts
-  \imprecision)} \,.
-\end{align*}
-Confusion of states is then also a function of imprecision parameter
-$\imprecision$. For $\imprecision = 0$, the confusion matrix has $C_{ii}
-= 1$ and $C_{ij} = 0$ for $i \neq j$. For $\imprecision > 0$, $C$ is positive, i.e., $C_{ij} >0$ for
-all $i$ and $j$. For $\imprecision \rightarrow \infty$, we find $C_{ij}
-\rightarrow \nicefrac{1}{\card{\States}}$ for all $i$ and $j$.
-
-If states are confused on average with a probability proportional to
-their similarity, repeated application of the operation in
-Equation~(\ref{eq:confusion-function}) causes behavioral dispositions
-gradually to diffuse along a gradient of similarity of
-states. Consequently, iterated diffusion leads to a smoothing out and
-an eventual equalization of behavioral strategies, for both the sender
-and the receiver. After $n$ steps of iterated diffusion an initial
-sender strategy $\Sstrat$ will be $C^n \Sstrat$, and an initial
-receiver strategy $\Rstrat$ will be $\Rstrat C^n$. If we assume that
-$\imprecision > 0$, $C$ is a positive row-stochastic matrix (each state
-could in principle be confused as any other state, albeit with
-possibly a very low probability). It then follows from the
-Perron-Frobenius theorem that the limit $C^\infty = \lim_{n
-  \rightarrow \infty} C^n$ exists (it is the Perron-projection) and
-all of its rows are identical. That is why, in the limit of iterated
-diffusion, $C^\infty \Sstrat$ has identical rows (messages are sent
-with the same probability in each state), and $\Rstrat C^\infty$
-likewise has identical rows (every message is interpreted
-equally). All of this is in line with the intuition that diffusion of
-behavior under confusion of states iteratively equalizes behavior for
-similar states; if all states can be confused for one another in
-principle, behavior smoothes out entirely in the limit.
-
-\subsection{Numerical simulations}
-\label{sec:simulations}
-
-Now that we understand better what diffusion does, let us look at how
-diffusion interacts with optimization of behavior, as described by the
-replicator dynamic. The main question we should address is whether
-diffusion and fitness-based replication reasonably interact, and if
-so, whether the inclusion of diffusion has any noteworthy effects on
-the evolving meaning of signals. Clearly, if $\imprecision = 0$, the
-\rdd reduces to the \rd. If $\imprecision \rightarrow \infty$, the
-diffusion component takes over and the \rdd reduces to the trivial
-iterated diffusion process that we looked at in the previous
-section. We will show presently that for reasonable in-between levels
-of imprecision $\imprecision$, the \rdd can lead to reasonably
-efficient, yet vague signaling behavior. Suitable levels of
-imprecision can have further accelerating and, surprisingly, unifying
-effects on meaning evolution.
-
-\subsubsection{Experimental set-up}
-
-To explore the \rdd, we turn to numerical simulations. Let states be
-evenly spaced elements of the unit interval that always include 0 and
-1, and let priors be flat: $\Pr(\state) = \Pr(\state')$ for all
-$\state, \state'$. As for utility functions, we take another
-Nosofsky-style function:
+We further assume that the probability of confusing any two states $\mystate{i}$ and
+$\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
+observational confusability $P_o$ is simply the same as realizational confusability $P_r$ and
+that both are governed by the same imprecision parameter $\alpha$:
+\begin{align*}
+  P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
+  \imprecision) &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
+  \imprecision) \,.
+\end{align*}
+For $\imprecision = 0$, we obtain trivial confusion probabilities, i.e., confusion probability
+zero between non-identical states and our dynamic reduces to the replicator dynamics. For
+$\imprecision > 0$, any state can be confused for any other state with some positive
+probability. For the uninteresting case of $\imprecision \rightarrow \infty$, confusion is
+maximal and every state can be perceived or realized as any other state with probability
+$\nicefrac{1}{\card{\States}}$.
+
+As for utility functions, we take another Nosofsky-style function:
 \begin{align*}
@@
 \end{align*}
-The tolerance parameter $\toler \ge 0$ models the amount of tolerable
-pragmatic slack or communicative imprecision. This choice of utility
-function is governed partly by convenience in parallel to the
-similarity function, but also because it has, as we believe, the right
-general properties for a communicative payoff function. Unlike
-utilities that are, say, linearly or quadratically decreasing in
-physical distance \citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa},
-utilities that are exponentially decreasing in negative quadratic
-distance can model situations where a small amount of imprecision in
-communication is tolerable, whereas similarly small differences in
-intolerably far away interpretations matter very little, with a smooth
-transition between these regimes
-\citep[c.f.][]{OConnor2013:The-Evolution-o}.
-
-We ran 50 trials of the \rdd, starting with randomly sampled sender
-and receiver strategies, for each triplet of independent parameter
-values: $\ns = \card{T} \in \set{6, 10, 50, 90}$, $\imprecision \in
-\set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2,
-  0.3}$. Each trial ran for a maximum of 200 update steps of the
-\rdd. A trial was considered converged, and thus stopped before the maximum
-of 200 rounds, if the total amount of change between strategies before
-and after an \rdd step was smaller than a suitably chosen
-threshold. It is not guaranteed that strategies at halting time had
-converged to the eventual attracting state, whether they ran for 200
-rounds or not. Our notion of convergence is therefore only a
-categorical measure for reaching a certain (well-considered, but
-eventually arbitrary) degree of stability. In principle, rather than
-setting a hard limit on the maximum number of rounds and analyzing
-convergence categorically, one could also run each trial until reaching
-our criterion for convergence and analyze the number of rounds it took to get there.
-Our choice was initially motivated by pragmatic concerns regarding length of
-simulation time, but it is also justifiable if we reflect on the
-nature of language evolution as a continuous process.
-Language adapts to an ever-changing reality; it is rare, if not impossible,
-that a lexicon has an infinite amount of time to neatly converge to a stable
-state. Therefore, investigating speed of convergence within a limited
-time frame can be seen as better reflecting the natural constraints of
-language evolution.
+The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
+communicative imprecision. This choice of utility function is governed partly by convenience in
+parallel to the similarity function, but also because it has, as we believe, the right general
+properties for a communicative payoff function. Unlike utilities that are, say, linearly or
+quadratically decreasing in physical distance
+\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
+that are exponentially decreasing in negative quadratic distance can model situations where a
+small amount of imprecision in communication is tolerable, whereas similarly small differences
+in intolerably far away interpretations matter very little, with a smooth transition between
+these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.
+
+In conclusion, to structure thinking about the behavior of the \icid, the system is governed by
+two parameters: imprecision $\imprecision$ and tolerance $\toler$, which can vary
+independently.
+
+\subsection{Numerical simulations}
+\label{sec:simulations}
+
+The main question we should address is whether state confusability and utility-driven imitation
+reasonably interact, and if so, whether the inclusion of confusability has any noteworthy
+effects on the evolving meaning of signals. It transpires that moderate levels of imprecision
+$\imprecision$ can lead to reasonably efficient, yet vague signaling behavior. Moreover,
+imprecision can have further accelerating and, surprisingly, unifying effects on meaning
+evolution. It appears that state confusability has an effect on aggregate signaling behavior
+that can be described as generalization or systematization, as if ``conclusions'' about what
+works for one state are ``carried over'' to similar states. This, however, is merely an
+emergent phenomenon in the sense no single agent genuinely generalizes over stimuli or reasons
+about what a more systematic signaling strategy would be.
+
+For all simulations, priors were flat: $P_a(\state) = P_a(\state')$ for all
+$\state, \state'$. We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting
+with randomly sampled sender and receiver strategies, for each triplet of independent parameter
+values: $\ns = \card{T} \in \set{6, 10, 50, 90}$,
+$\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each
+trial ran for a maximum of 200 update steps of the \rdd. A trial was considered converged, and
+thus stopped before the maximum of 200 rounds, if the total amount of change between strategies
+before and after an \rdd step was smaller than a suitably chosen threshold. It is not
+guaranteed that strategies at halting time had converged to the eventual attracting state,
+whether they ran for 200 rounds or not. Our notion of convergence is therefore only a
+categorical measure for reaching a certain (well-considered, but eventually arbitrary) degree
+of stability. In principle, rather than setting a hard limit on the maximum number of rounds
+and analyzing convergence categorically, one could also run each trial until reaching our
+criterion for convergence and analyze the number of rounds it took to get there.  Our choice
+was initially motivated by pragmatic concerns regarding length of simulation time, but it is
+also justifiable if we reflect on the nature of language evolution as a continuous process.
+Language adapts to an ever-changing reality; it is rare, if not impossible, that a lexicon has
+an infinite amount of time to neatly converge to a stable state. Therefore, investigating speed
+of convergence within a limited time frame can be seen as better reflecting the natural
+constraints of language evolution.
 
@@
     \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.0.5$}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.05$}
     \label{fig:example_stratsB}
@@
 
-\appendix
-
-\section{Diffusion as mutation}
-\label{sec:diffusion-as-special}
-
-\subsection{The replicator mutator dynamic}
-\label{sec:repl-mutat-dynam}
-
-The \rmd has first been proposed as a tool for studying the evolution
-of grammar
-\citep[e.g.][]{KomarovaNiyogi2001:The-Evolutionar,NowakKomarova2001:Evolution-of-Un}. It
-adds stochastic mutation to the replicator dynamic. The relation of
-the \rmd to other prominent evolutionary dynamics is well understood
-\citep{PageNowak2002:Unifying-Evolut}. The \rmd has seen further
-fruitful applications in the context of signaling games
-\citep[e.g.][]{HutteggerSkyrms2010:Evolutionary-Dy}.
-
-The usual formulation of the \rmd is based on mixed, not behavioral strategies. But a
-discrete-time formulation of the \rmd in behavioral strategies clearly identifies the \rdd to
-be a special case. It follows that a mixed-strategy version of the \rdd is just the \rmd with
-suitably defined mutation matrices. From this, even a continuous-time version can easily be
-retrieved if desired. To show this, we start with a delineation of mixed and behavioral
-strategies.
-
-\paragraph{Mixed strategies.} Pure sender (receiver) strategies are
-functions $\Spure \in \Messgs^\States$ ($\Rpure \in
-\Acts^\Messgs$). Mixed sender (receiver) strategies are functions
-$\Smixed \in \Delta(\Messgs^\States)$ ($\Rmixed \in
-(\Acts^\Messgs)$). The latter give the relative population frequencies
-of the former. We write $\Smixed_i$ for the frequency
-$\Smixed(\Spure_i)$ of pure strategy $\Spure_i$. Likewise for the
-receiver.
-
-Every mixed strategy $\Smixed$ converts to a unique behavioral
-strategy defined by:
-\begin{align}
-  \label{eq:CorrespondenceBehavioralMixed}
-  \Sstrat(\messg \probbar \state) = \sum_{\Spure(\state) = \messg} \Smixed(\Spure)\,.
-\end{align}
-Let $G$ be this mapping from mixed to behavioral strategies. Note that $G$
-is \emph{not} an injection, as many mixed strategies map onto the same
-behavioral strategy. In this sense, mixed strategies hold more
-information than behavioral strategies about the distribution of pure strategies in a population.
-
-In the context of evolutionary dynamics for signaling games (where
-each agent only chooses once), we can think of behavioral strategies
-as treating each choice point as an independent update site
-\citep[e.g.][]{Cressman2003:Evolutionary-Dy}. Dynamics for mixed
-strategies conceive of agents as adjusting their entire pure
-strategy. In this way, dependencies between different choices at
-different choice points can matter to the evolutionary path. In
-contrast, dynamics for behavioral strategies conceive of agents as
-adjusting their strategies locally at each choice point, regardless of
-what they are disposed to do at other choice points.
-
-This can matter to evolutionary dynamics. For illustration, take two
-sender strategies whose behavior is different only at two
-states. Compared to the second strategy, the first scores a much
-higher expected utility at one state, but a slightly lower at the
-other, so that the total expected utility of the first strategy is
-higher. Under evolutionary dynamics that work on entire strategies the
-first strategy will increase in proportion, relative to the
-second. Consequently, a locally less efficient type of behavior can be
-promoted because it is correlated with efficient behavior at another
-choice point. Evolutionary dynamics working on behavioral strategies
-do not allow for that. Behavioral adaptations occur independently at
-different choice points.
-
-As a result, dynamics on behavioral strategies are more coarse-grained
-and not necessarily equivalent to dynamics on mixed strategies. An
-advantage of dynamics in behavioral strategies is that they reduce
-complexity, which helps with numerical simulations. They might also be
-a more natural choice for cultural dynamics, e.g., when an agent
-observes and imitates another agent's strategy: it may be more
-plausible that agents adopt the use of a single word, not an entire
-lexicon, so to speak. On the other hand, biological evolution seems
-more naturally modeled in terms of mixed strategies whenever it is a
-full set of behavioral disposition that is bequeathed to the next
-generation.
-
-\paragraph{Replicator dynamic in mixed strategies.} To understand the
-\rmd in mixed strategies, we first introduce the \rd in mixed
-strategies. Let $F_i^{\Rmixed}$ be $\Spure_i$'s fitness given
-$\Rmixed$ and $F_i^{\Smixed}$ be $\Rpure_i$'s fitness given
-$\Smixed$. Then $\Phi(\Smixed,\Rmixed) = \sum_{k} \Smixed_k \cdot
-F_k^{\Rmixed}$ is the average fitness in the sender population and
-$\Phi(\Rmixed,\Smixed) = \sum_{k} \Rmixed_k \cdot F_k^{\Smixed}$ the
-average fitness in the receiver population.
-
-% The continuous-time replicator dynamic defines the change of frequency
-% of mixed strategies. Its so-called two-population version, that takes
-% sender and receiver roles as separate update sites is:
-% \begin{align*}
-%   \dot{\Smixed_i} & = \Smixed_i \cdot \left ( F_i^{\Rmixed} -
-%   \Phi(\Smixed,\Rmixed) \right ) &   \dot{\Rmixed_i} &  = \Rmixed_i \cdot \left ( F_i^{\Smixed} -
-%   \Phi(\Rmixed,\Smixed) \right ) \,.
-% \end{align*}
-Similar to the version for behavioral strategies, a discrete-time
-formulation of the \rd for mixed strategies can be conceptualized as an
-update function that maps mixed strategies $\Smixed$ and $\Rmixed$
-onto mixed strategies $\RD(\Smixed)$ and $\RD(\Rmixed)$ respectively,
-such that:
-\begin{align*}
-  \RD(\Smixed)_i & = \frac{\Smixed_i \cdot F_i^{\Rmixed}}{
-    \Phi(\Smixed,\Rmixed)} & \RD(\Rmixed)_i & = \frac{\Rmixed_i \cdot
-    F_i^{\Smixed}}{ \Phi(\Rmixed,\Smixed)} \,.
-\end{align*}
-
-\paragraph{Replicator mutator dynamic.} The replicator mutator dynamic
-extends the \rd by adding probabilistic mutation. Let $Q$ be a
-row-stochastic mutation matrix where $Q_{ji}$ gives the probability
-that pure sender strategy $\Spure_j$ mutates into
-$\Spure_i$. Similarly, let $R$ be a row-stochastic mutation matrix
-where $R_{ji}$ gives the probability that pure receiver strategy
-$\Rpure_j$ mutates into $\Rpure_i$.
-
-The \rmd is usually given only in its continuous-time form. % A
-% two-population (non-payoff adjusted) formulation is then:
-% \begin{align*}
-%   \dot{\Smixed_i} & = \sum_{j}  Q_{ji} \cdot \Smixed_j
-%     \cdot F_j^{\Rmixed} - \Smixed_i \cdot \Phi(\Smixed,\Rmixed) &
-%     \dot{\Rmixed_i} & = \sum_{j}  R_{ji} \cdot \Rmixed_j
-%     \cdot F_j^{\Smixed} - \Rmixed_i \cdot \Phi(\Rmixed,\Smixed) \,.
-% \end{align*}
-But a discrete-time, two-population formulation of the \rmd is
-available too \citep[c.f.][97]{PageNowak2002:Unifying-Evolut}:
-\begin{align*}
-  \RMD(\Smixed)_i & = \sum_{j} Q_{ji} \frac{\Smixed_j \cdot
-    F_j^{\Rmixed}}{ \Phi(\Smixed,\Rmixed)} & \RDD(\Rmixed)_i & =
-  \sum_{j} R_{ji} \frac{\Rmixed_j \cdot F_j^{\Smixed}}{
-    \Phi(\Rmixed,\Smixed)}\,.
-\end{align*}
-The discrete-time \rmd has the same sequential nature as the \rdd:
-first we compute the fitness-driven change according to the standard
-replicator dynamic; then we compute the perturbation from
-mutation. This becomes clear if we define independent mutation
-functions $\Mutate_{Q,R}$ that map mixed strategies onto the outcome
-of (one round of) mutation:
-\begin{align}
-  \label{eq:Mutation}
-  \Mutate_Q(\Smixed)_i & =  \sum_j  \Smixed_j \cdot
-  Q_{ji} &   \Mutate_R(\Rmixed)_i & =  \sum_{j}  \Rmixed_j \cdot
-  R_{ji} \,.
-\end{align}
-The discrete-time \rmd given above can then be rewritten as:
-\begin{align*}
-  \RDD(\Smixed) &= \Mutate_Q(\RD(\Smixed)) &   \RDD(\Rmixed) &= \Mutate_R(\RD(\Rmixed))\,.
-\end{align*}
-
-\subsection{Relating the dynamics}
-\label{sec:relating-dynamics}
-
-The sequential reformulation of the \rmd closely resembles our initial
-formulation of the \rdd. However, there is a major difference between
-these. While the \rdd is a dynamic for behavioral strategies, the
-\rmd, as taken from the literature, applies to mixed
-strategies. Similarly, the confusability of states is an operation
-that is straightforwardly definable for behavioral strategies (see
-Equation~(\ref{eq:confusion-function})), while mutation in the \rmd is
-defined as mutation from one pure strategy to another. Nonetheless, it
-is possible to map each confusion matrix to a mutation matrix in a
-natural way so that the correspondence between mixed and behavioral
-strategies are conserved. This is the main result presented in this
-section, and it is our justification for regarding the \rdd as the
-behavioral-strategy analogue of the replicator-mutator dynamic when
-the only source of mutation is confusability of states.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[node distance = 1cm]
-
-    \begin{scope}
-      \node[] (mixed) {$\Smixed$};
-
-      \node[below of = mixed] (mutated) {$\Mutate_{Q^C}(\Smixed)$};
-
-      \node[left of = mixed, node distance = 1.5cm] (dummyl) {};
-
-      \node[right of = mixed, node distance = 1.5cm] (dummyr) {};
-
-      \draw[->, dotted, thick] (mixed) -> (mutated);
-
-      \begin{pgfonlayer}{background}
-        \node [draw=black!50, fill=black!10,fit = (dummyl)
-        (mutated) (dummyr)] (mixedSpace) {};
-      \end{pgfonlayer}
-
-      \node[above of = mixedSpace, node distance = 1.4cm] {space of
-        mixed strategies};
-    \end{scope}
-
-    \begin{scope}[xshift=6cm]
-      \node[] (behavioral) {$\Sstrat$};
-
-      \node[below of = behavioral] (mutatedBeh) {$\Diff_C(\Sstrat)$};
-
-      \node[left of = behavioral, node distance = 1.5cm] (dummylB) {};
-
-      \node[right of = behavioral, node distance = 1.5cm] (dummyrB) {};
-
-      \draw[->, dotted, thick] (behavioral) -> (mutatedBeh);
-
-      \begin{pgfonlayer}{background}
-        \node [draw=black!50, fill=black!10,
-              fit = (dummylB) (mutatedBeh) (dummyrB)] (behSpace) {};
-      \end{pgfonlayer}
-
-      \node[above of = behSpace, node distance = 1.4cm] {space of
-        behavioral strategies};
-    \end{scope}
-
-    \draw[->,thick,dotted] (mixed) -> (behavioral) node[midway,
-    above]{$G(\Smixed)$};
-
-    \draw[->,thick,dotted] (mutated) -> (mutatedBeh) node[midway, below]{$G(\Mutate_{Q^C}(\Smixed))$};
-
-  \end{tikzpicture}
-
-  \caption{Correspondence between state-confusion and mutation.}
-  \label{fig:correspondence-result}
-\end{figure}
-
-More concretely, given the non-injective mapping $G$ from mixed to
-behavioral strategies, we show that there is a systematic translation
-from a confusion matrix $C$ to a mutation matrix $Q^C$ such that for
-each mixed sender strategy $\Smixed$, its unique corresponding
-behavioral strategy $\Sstrat = G(\Smixed)$ has a
-confusion-perturbation $\Diff_C(\Sstrat)$ that corresponds, via $G$,
-to the mutation-perturbation $\Mutate_{Q^C}(\Smixed)$ (see
-Figure~\ref{fig:correspondence-result}). In other words, we
-hypothesize that a confusion matrix $C$ should give rise to a unique
-mutation matrix $Q^C$ so that whenever $G(\Smixed) = \Sstrat$ we also
-have $G(\Mutate_{Q^C}(\Smixed)) = \Diff_C(\Sstrat)$. Similarly, for
-the receiver.
-%This is the content of Theorem~\ref{thm:Correspondence} given below, a
-%proof for which is in Appendix~\ref{sec:proofs}. The remainder of this
-%section introduces the relevant construction that maps arbitrary $C$
-%to suitable $Q^C$.
-
-\paragraph{Confusion-based mutations.} There are natural conversions
-of $C$ into $Q^C$ and $R^C$. The case for the receiver is easier, so
-we start with that.
-
-The probability that $\Rpure_i$ is realized as $\Rpure_j$ under
-diffusion with $C$ is the product of the probabilities, for each
-$\messg$, that the state $\Rpure_i(\messg)$ is perceived as state
-$\Rpure_j(\messg)$. Abusing notation by referring to the indices of
-states $\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ with
-$\Rpure_i(\messg)$ and $\Rpure_j(\messg)$ directly, we define:
-\begin{align}
-  \label{eq:construction-rec}
-  R^C_{ij} = \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)}\,.
-\end{align}
-
-Now look at the sender. The probability that $\Spure_j$ is realized as
-$\Spure_i$ under diffusion with $C$ is the product of the
-probabilities, over all states $\mystate{k}$, that the message
-$\Spure_i(\mystate{k})$, that $\Spure_i$ would produce at state
-$\mystate{k}$ in the absence of noise, is produced by a noisy
-realization of $\Spure_j$, which is the probability $\sum_{\state_l
-  \in \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}$ that the state
-$\mystate{k}$ is realized as a state $\mystate{l}$ which $\Spure_j$
-would map unto $\Spure_i(\mystate{k})$. So, define:
-\begin{align}
-  \label{eq:construction-sen}
-  Q^C_{ji} = \prod_{\state_k} \sum_{\state_l \in
-    \Spure_j^{-1}(\Spure_i(\state_k))} C_{kl}\,.
-\end{align}
-
-Based on Equations~(\ref{eq:construction-rec}) and
-(\ref{eq:construction-sen}), we can formulate the desired
-correspondence result:
-
-\begin{theorem}
-  \label{thm:Correspondence}
-  (i) If $G(\Smixed) = \Sstrat$, then $G(\Mutate_{Q^C}(\Smixed)) =
-  \Diff_C(\Sstrat)$. (ii) If $G(\Rmixed) = \Rstrat$, then
-  $G(\Mutate_{R^C}(\Rmixed)) = \Diff_C(\Rstrat)$.
-\end{theorem}
-
-We can prove this theorem as follows.
-%\section{Proof of Theorem}
-%\label{sec:proofs}
-
-\begin{proof}[Part (i).]
-  Fix $\Smixed$ and $\Sstrat = G(\Smixed)$. Look first at the rhs of
-  the consequent:
-  \begin{align*}
-    \Diff_C(\Sstrat)(\messg_y \probbar \state_x) & =  \sum_{\state_l} C_{xl}
-    \cdot \Sstrat(\messg_y \probbar \state_l) && \text{(by Equation~(\ref{eq:confusion-function}))} \\
-    & =  \sum_{\state_l} C_{xl}
-    \cdot  \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i && \text{(by Equation~(\ref{eq:CorrespondenceBehavioralMixed}))} \\
-    & = \sum_{\state_l}
-    \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i \cdot C_{xl}\,.
-  \end{align*}
-  Next consider the lhs of the consequent:
-  \begin{align*}
-    G(\Mutate_{Q^C}(\Smixed))(\messg_y \probbar \state_x) & =
-    \sum_{\Spure_i(\state_x)=\messg_y} \Mutate_{Q^C}(\Smixed_i) &&
-    \text{(by Equation~(\ref{eq:CorrespondenceBehavioralMixed}))} \\
-    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
-    \Smixed_j \cdot Q^C_{ji} &&
-    \text{(by Equation~(\ref{eq:Mutation}))} \\
-    & = \sum_{\Spure_i(\state_x)=\messg_y} \sum_{\Spure_j}
-    \Smixed_j \cdot \prod_{\state_l} \sum_{\state_m \in
-      \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm} &&
-    \text{(by Equation~(\ref{eq:construction-sen}))} \\
-    & = \sum_{\Spure_j} \Smixed_j \cdot
-    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l}
-    \sum_{\state_m \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}
-  \end{align*}
-  To simplify this further we look at a fixed $\Spure_j$ and consider
-  the term: 
-  \begin{align}
-    \label{eq:term}
-    \sum_{\Spure_i(\state_x)=\messg_y} \prod_{\state_l} \sum_{\state_m
-      \in \Spure_j^{-1}(\Spure_i(\state_l))} C_{lm}\,.
-  \end{align}
-  Let $Y$ be the row-stochastic matrix with $Y_{kl} = \sum_{\state_m
-    \in \Spure_j^{-1}(\messg_l)} C_{km}$. Every pure sender strategy
-  maps each state $\state_k$ onto exactly one $Y_{kl}$. If we quantify
-  over all pure strategies, we essentially look at each such
-  mapping. Term (\ref{eq:term}) above sums over all pure strategies
-  that map $\state_k$ onto $\messg_y$. The above term then sums over
-  all products whose factors are tuples in $\times_{k>2} \set{y
-    \setbar \exists l \mycolon y = Y_{kl}}$. So term (\ref{eq:term})
-  expands to (where $e = \card{\States}$ and $d=\card{\Messgs}$):
-  \begin{align*}
-    & (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e1}) +
-    (Y_{11} \cdot Y_{21} \cdot Y_{31} \cdot \ldots \cdot Y_{e2}) + 
-    \dots \\
-    & + (Y_{11} \cdot Y_{2d} \cdot
-    Y_{3d} \cdot \ldots \cdot
-    Y_{ed}) 
-  \end{align*}
-  But since $Y$ is a row-stochastic matrix, this simplifies to
-  $Y_{xy}$. Continuing the derivation with this:
-  \begin{align*}
-    G(\Mutate_{Q^C}(\Smixed))(\messg_y \probbar \state_x) 
-    & = \sum_{\Spure_j} \Smixed_j \cdot
-    \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} C_{xl} \\
-    & = \sum_{\Spure_j} \sum_{\state_l \in \Spure_j^{-1}(\messg_y)} \Smixed_j \cdot
-     C_{xl} \\
-     & = \sum_{\state_l}
-    \sum_{\Spure_i(\state_l) = \messg_y} \Smixed_i \cdot C_{xl}\,.
-  \end{align*}
-
-\end{proof}
-
-\begin{proof}[Part (ii).]
-  Fix $\Rmixed$ and $\Rstrat = G(\Rmixed)$. The rhs of the consequent
-  expands to:
-  \begin{align*}
-    \Diff_C(\Rstrat)(\state_x \probbar \messg_y) & = \sum_{\state_j}
-    \sum_{i \in \set{k \setbar \Rpure_k(\messg_y) = \state_x}}
-    \Rmixed_i \cdot C_{jx} && \text{(by Equation~(\ref{eq:confusion-function}))} \\
-    & = \sum_{\Rpure_i} \sum_{j \in \set{k \setbar \Rpure_i(\messg_y) = \state_j}}
-    \Rmixed_i \cdot C_{jx} \\
-    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
-  \end{align*}
-  The rhs expands to (by Equations~(\ref{eq:CorrespondenceBehavioralMixed}),
-  (\ref{eq:Mutation}) and (\ref{eq:construction-rec})):
-  \begin{align*}
-    G(\Mutate_{R^C}(\Rmixed))(\state_x \probbar \messg_y) & = \sum_{\Rpure_i}
-    \Rmixed_i \cdot \sum_{j \in \set{j \setbar \Rpure_k(\messg_y) =
-        \state_x}} 
-    \prod_{\messg} C_{\Rpure_i(\messg)\Rpure_j(\messg)} \,.
-  \end{align*}
-  Without loss of generality, assume that $x=y=1$, and fix
-  $\card{M}=d$ and let $e$ be the number of pure receiver
-  strategies. Then the last term can be rewritten as:
-  \begin{align*}
-    & = \sum_{i}
-    \Rmixed_i \cdot ( C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)\Rpure_1(\messg_2)} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)\Rpure_1(\messg_d)} + \ldots  \\
-      & \textcolor{white}{ = \sum_{i}
-    \Rmixed_i  (}  +  C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)\Rpure_2(\messg_2)} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)\Rpure_2(\messg_d)} + \ldots  \\
-      & \textcolor{white}{ = \sum_{i}
-    \Rmixed_i  (}  + C_{\Rpure_i(\messg_1)1} \cdot
-      C_{\Rpure_i(\messg_2)\Rpure_e(\messg_2)} \cdot \ldots \cdot
-      C_{\Rpure_i(\messg_d)\Rpure_e(\messg_d)}) ) \,.
-  \end{align*}
-  For every messages $\messg_l$, $C_{\Rpure_i(\messg_l)}$ is a stochastic
-  vector. For $l>1$, all elements of these vectors appear equally
-  often. But that means that these cancel out. What remains is:
-  \begin{align*}
-    & = \sum_{\Rpure_i} \Rmixed_i \cdot C_{\Rpure_i(\messg_y)x} \,.
-  \end{align*}
-\end{proof}
-
-\section{Derivation of the RDD by imitation-based revision}
-\label{sec:deriv-rdd-imit}
-
-We can derive the proposed \rdd as a mean dynamic of an imitative revision protocol, following
-closely the example 4.3.1 of \citet{Sandholm2010:Population-Game}. Our derivation relies on the
-usual assumptions about conditional imitation at the agent level, but must add special
-assumptions about how noise affects revision probabilities. In keeping with the idea that
-agents are of possibly limited cognitive capacities, we assume here that noise is real but that
-agents are unaware of it. When agents perceive the actual expected utility of their choice and
-of another agent's choice (as is usual in the motivation of imitation-based revision
-protocols), we assume additionally that agents mistake their possibly non-veridical subjective
-representations of states for the actual ones.
-
-\paragraph{Notation.} As before, let $\Sstrat(\mymessg{i} \probbar \mystate{o})$ be the
-probability that a speaker randomly picked from the population chooses $\mymessg{i}$ if the
-actual/objective state is $\mystate{o}$. This speaker may perceive the objective state
-$\mystate{o}$ subjectively as $\mystate{s}$. In contrast, let $\Sstrat_s(\mymessg{i} \probbar
-\mystate{s})$ be the probability that a randomly sampled speaker chooses $\mymessg{i}$ if she
-experiences some objective state $\mystate{o}$ as $\mystate{s}$. The latter is the average
-sender behavior in a (possibly counterfactual) world without perceptual confusion. It is also
-where behavioral changes first take effect, to then percolate to $\Sstrat$ as well.
-
-Similarly, $\Rstrat(\mystate{o} \probbar \messg)$ is the probability that a randomly sampled
-receiver actually chooses interpretation $\mystate{o}$. $\Rstrat_s(\mystate{s} \probbar
-\messg)$ is the probability of a subjectively intended interpretation $\mystate{s}$. Again,
-behavioral changes first target $\Rstrat_s$ but cause changes in $\Rstrat$ as well.
-
-Let $\EU(\messg)$ be short for $\EU(\messg,\state,\Rstrat)$ whenever
-$\state$ and $\Rstrat$ are implicit, similarly for $\EU(\state)$ as a
-shorthand for $\EU(\state,\messg,\Sstrat)$.
-
-\paragraph{Standard assumptions.} Let's assume that we are in an
-entirely noise-free world so that $\Sstrat = \Sstrat_s$ and $\Rstrat =
-\Rstrat_s$. The standard assumptions for imitation-based revisions are
-these. Agents get revision opportunities for all actual choice points
-with equal probability (independently of how likely these would occur
-in actual play). Take the case of a sender (receiver side is
-parallel). If agent $A$ is given a revision opportunity for actual
-state $\state$, she consults her own choice, which is $\mymessg{i}$
-with probability $\Sstrat(\mymessg{i} \probbar \state)$, and observes
-the behavior of one randomly picked other agent $B$, which is
-$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
-\state)$. $A$ also observes the expected utility of both choices (at
-$\state$) and switches to $B$'s choice with probability $r_{ij} = [
-\EU(\mymessg{j}) - \EU(\mymessg{i}) ]_+$, so that the overall
-switching probability of an agent like $A$ who samples $\mymessg{i}$
-for $\state$ is $p_{ij} = \Sstrat(\mymessg{j} \probbar \state) \
-r_{ij}$.
-
-\paragraph{Imitation under perceptual noise.} There are many conceivable ways in which
-perceptual noise might influence the above revision protocol. Our choice here is guided by
-theoretical and practical considerations alike. We would like to maintain the common assumption
-that agents might not be ideally rational and not necessarily fully aware of their
-surroundings. In particular, we assume here that agents are not aware of perceptual
-noise. Switching probabilities are therefore defined in terms of actual expected utility (and
-not subjective expectations that take noise into consideration). Moreover, we assume that
-agents who have (unconscious) access to expected utilities for actual states $\mystate{o}$ will
-also (mistakenly) treat whatever subjective state $\mystate{s}$ they may have experienced
-during their revision opportunity as if it had been the actual state $\mystate{o}$.
-
-Concretely, suppose agent $A$ has a revision opportunity for objective
-state $\mystate{o}$, and chooses $\mymessg{i}$ with probability
-$\Sstrat(\mymessg{i} \probbar \mystate{o})$. Similarly, $B$ chooses
-$\mymessg{j}$ with probability $\Sstrat(\mymessg{j} \probbar
-\mystate{o})$. It may be that $A$ perceived $\mystate{o}$ as
-$\mystate{s_A}$ and that $B$ perceived $\mystate{s_B}$, but since $A$
-is unaware of noise he expects no difference. $A$ observes (as usual)
-the actual expected utilities of $\mymessg{i}$ and $\mymessg{j}$ for
-$\mystate{o}$. Consequently, he updates his subjective behavior for
-$\mystate{o}$ (not being aware that the perceived $\mystate{s_A}$
-might have been different). This myopic revision behavior gives rise
-to expected changes in the subjective/noise-free choice probabilities
-$\Sstrat_s(\mymessg{i} \probbar \state)$ as follows:
-\begin{align*}
-  \dot{\Sstrat}_s(\mymessg{i} \probbar \state) & = \sum_j
-  \Sstrat(\mymessg{j} \probbar \state) \cdot p_{ji} -  
-  \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j p_{ij} \\
-  & = \sum_j  \Sstrat(\mymessg{j} \probbar \state) \cdot
-    \Sstrat(\mymessg{i} \probbar \state) \cdot [\EU(\mymessg{i}) -
-    \EU(\mymessg{j})]_+ \\ 
-   & \ \ \ \  - \Sstrat(\mymessg{i} \probbar \state) \cdot \sum_j
-   \Sstrat(\mymessg{j} \probbar \state) \cdot [\EU(\mymessg{j}) -
-    \EU(\mymessg{i})]_+ \\
-   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left ( \sum_j
-     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{i}) - \sum_j
-     \Sstrat(\mymessg{j} \probbar \state) \cdot \EU(\mymessg{j})
-   \right ) \\
-   & = \Sstrat(\mymessg{i} \probbar \state) \cdot \left (
-     \EU(\mymessg{i}) - \text{average EU at $\state$} \right )
-\end{align*}
-This is the (continuous-time) replicator dynamic for behavioral
-strategies.
-
-Similar considerations motivate the receiver part. Suppose that agent
-$A$ has a revision opportunity for choice point $\messg$. He and a
-random other agent $B$ actually choose $\mystate{o_A}$ and
-$\mystate{o_B}$ with probability $\Rstrat(\mystate{o_{A,B}} \probbar
-\messg)$, even though they may have intended interpretations
-$\mystate{s_A}$ and $\mystate{s_B}$. Agent $A$ observes the actual
-expected utility and, by assumption, believes (myopically) that he had
-wanted to realize $\mystate{o_A}$ and that $B$ had wanted to realize
-$\mystate{o_B}$. Under these assumptions, the derivation of the
-replicator dynamic for subjectively intended (noise-free) receiver
-behavior $\Rstrat_s$ proceeds in parallel with the above sender case.
-
-At each time step, the noise-free behavior of agents is, of course, perturbed by noise. This is
-why, after each update step, the \rdd defined in Equation~(\ref{eq:RDD-Def}) perturbes the
-outcome of the replicator step with perceptual noise.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \printbibliography[heading=bibintoc]

### 2015-08-13 paper/version_04/paper.tex
edits section 3

@@
 \newcommand{\myred}[1]{\textcolor{red}{#1}} 
-
-\doublespacing
+\renewcommand{\Pr}{\ensuremath{P}}
+
+% \doublespacing
 
@@
 agents, we can picture a more or less idealized process of how each agent adapts dispositions
-for future actions based on various pieces of information available to the agent, e.g., about
-the game, the other players' behavior and the like. More abstractly, at the level of a
-population of agents, we can describe how average behavioral dispositions will evolve. The
-population-level perspective abstracts over petty stochastic fluctuations and zooms in on the
-general tendency or direction of evolution that ensues from some (often: simplified) behavior
-at the agent-level.
+for future actions based on various pieces of information available to the agent. More
+abstractly, at the level of a population of agents, we can describe how average behavioral
+dispositions will evolve. The population-level perspective abstracts over petty stochastic
+fluctuations and zooms in on the general tendency or direction of evolution that ensues from
+behavior at the agent-level.
 
@@
 evolutionary path of a population of signaling agents who imitate other agents' behavior but
-may occasionally confuse states for one another. We call our dynamic \emph{imprecise
-  conditional imitation dynamic} (\icid), because it derives from an update protocol based on
-conditional imitation and allows for imprecise perception and realization of states. In the
-special limiting case where imprecision vanishes, the process is just the well-known replicator
-dynamic \citep{TaylorJonker1978:Evolutionary-St}. To appreciate the \icid and its motivation,
-we first briefly review the replicator dynamic and its derivation from conditional imitation of
-successes.
+are liable to confuse states for one another. In the special limiting case where state
+confusability vanishes, the process is just the well-known replicator dynamic
+\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.
 
@@
 
-The \emph{expected utility} of choices at each choice point is:
+The \emph{expected utility} of choices at each choice point is:\footnote{The notation
+  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
+  side must possibly yet be normalized so as to have probabilities sum to one. In general,
+  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
+  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
 \begin{align*}
@@
   \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
-    \States} \frac{\Pr(\state) \cdot \Sstrat(\messg \probbar \state)}{\sum_{\state'} \Pr(\state') \cdot \Sstrat(\messg \probbar \state')} \cdot
-  U_\rec(\state, \messg, \act) \,.
+    \States} P(\state \probbar \messg)  \cdot
+  U_\rec(\state, \messg, \act)\,, \text{\ \  where } &
+P(\state \probbar \messg) & \propto \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \,.
 \end{align*}
@@
 outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
-sender of choosing message $\messg \in \Messgs$, given tate $\state \in \States$ and a receiver
-strategy $\Rstrat$ is the sum, for each act $\act \in \Acts$, of the probability that the
-receiver will choose act $\act$ given message $\messg$ times the actual utility of using act
-$\act$, given $\state$ and $\messg$.
+sender of choosing message $\messg$, given tate $\state$ and a receiver
+strategy $\Rstrat$ is the sum, for each act $\act$, of the probability that the
+receiver will choose $\act$ given $\messg$ times the actual utility of triple
+$\state$, $\messg$ and $\act$.
 
@@
   we will be looking at in Section~\ref{sec:exploring-rdd}, where utilities are always
-  non-negative and expected utilities are always positive.}\textsuperscript{,}\footnote{The
-  notation $\propto$, for ``proportional to'', means that the right-hand side must possibly yet
-  be normalized so as to have probabilities sum to one. In general, writing $P(x) \propto f(x)$
-  for any function $f \mycolon X \rightarrow \mathds{R}$ is shorthand for
-  $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. The implicit normalizing denominators for the
-  replicator equations are average expected utilities, e.g., for the sender:
-  $\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg', \state,\Rstrat)$.}
+  non-negative and expected utilities are always positive.}
 \begin{align}
@@
 deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
-here on one of the simplest: \emph{conditional imitation of success} \citep[see][for this and
-many other derivations]{Sandholm2010:Population-Game}. The main idea is that agents imitate the
-behavior of another agent at some choice point with a probability that is proportional to the
-expected utility of the latter agent's act for that choice point. Since the sim-max games that
-we will be looking at in Section~\ref{sec:exploring-rdd} have positive utilities, we may
-identify the switching probability with the expected utility.
-
-Let's consider sender strategies, as the receiver case is parallel. Revision by imitation of
-success proceeds as follows. Call (misleadingly!) the agent who gets a chance to change
-behavior ``learner'' and the possibly to-be-imitated agent ``teacher.'' A random learner is
-drawn from the population and given a chance to change behavior at choice point $\state$. The
-probability that our learner plays $\messg$ is $\Sstrat(\messg \probbar \state)$. The learner
-observes what a randomly sampled teacher does at $\state$. That would be $\messg'$ with
-probability $\Sstrat(\messg' \probbar \state)$. The learner then starts using $\messg'$ instead
-of $\messg$ with probability $\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and
-$\messg$ could be the same.)
+here on one of the simplest: \emph{conditional imitation of success}
+\citep[see][]{Sandholm2010:Population-Game}. The main idea is that agents imitate the behavior
+of other agents at some choice point with a probability that is proportional to the expected
+utility of the latter agents' choice. Since the sim-max games that we will be looking at in
+Section~\ref{sec:exploring-rdd} have positive utilities, we can identify the switching
+probability with the expected utility.
+
+Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
+agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
+``teacher.'' A random learner is drawn from the population and given a chance to change
+behavior at choice point $\state$. The probability that our learner plays $\messg$ is
+$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
+$\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar \state)$. The
+learner then starts using $\messg'$ instead of $\messg$ with probability
+$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)
 
@@
 
-\subsection{Imitation of success with noise}
+\subsection{Imitation of success with imprecision}
 \label{sec:imit-succ-with}
@@
   straightforward to also incorporate confusability of messages, but this is irrelevant to our
-  present purposes.} Confusability of states will affect how agents behave (on average), how
-they perceive others that they might want to imitate, and how good any behavioral disposition
-actually is.
-
-To keep matters simple, let us assume that agents carry pure dispositions to act, but noise can
-affect their realization. As a sender, every agent maps states to messages: these are
-subjectively perceived states, and no longer necessarily also the actually occurring state. As
-a receiver, every agent maps messages to state interpretations: these are intended
+  present purposes.} Confusability of states will affect how agents behave, how they perceive
+the behavior of others, and the expected utilities of behavioral dispositions.
+
+To keep matters simple, let us assume that agents carry pure dispositions to act. Noise can
+affect the realization of these strategies. As a sender, every agent maps states to messages:
+these are subjectively perceived states, and no longer necessarily also the actually occurring
+states. As a receiver, every agent maps messages to state interpretations: these are intended
 interpretations that need not always be faithfully realized. This means that behavioral
-strategies $\Sstrat \in (\Delta(\Messgs))^\States$ and $\Rstrat \in (\Delta(\States))^\Messgs$
-represent the average proportions of \emph{actual} behavioral dispositions in the population,
-the realization and observation of which can be distorted by confusion of similar states.
+strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
+behavioral dispositions in the population, the realization and observation of which can be
+distorted by agents' confusion of similar states.
 
@@
 probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
-intends to select interpretation $\mystate{i}$, we will assume that the probability with which
-state $\mystate{r}$ is realized is $P_r(\mystate{r} \probbar
-\mystate{i})$.
-Section~\ref{sec:exploring-rdd} gives a concrete specification of $P_o$ and $P_r$ in terms of
-similarity of states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary
-conditional confusion probabilities.
+intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
+the probability with which state $\mystate{r}$ is realized.  Section~\ref{sec:exploring-rdd}
+gives a concrete (parameterized) specification of $P_o$ and $P_r$ in terms of similarity of
+states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary conditional
+confusion probabilities.
 
@@
 
-Expected utilities of choices at choice points should also take into account that actual states
+Expected utilities of choices at choice points should likewise take into account that actual states
 need not be observed states, and intended interpretations need not be realized
@@
 \end{align*}
-is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed. The probability
-that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
+is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed by an agent. The
+probability that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
 \begin{align*}
@@
 If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
-observability and realizability of non-identical states, above definitions reduce to the
+confusability of non-identical states, above definitions reduce to the
 previous definitions of expected utilities. This also legitimates the overload of notation.
@@
 Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
-the dynamic that ensues from imitation of successes. Since imprecision works differently on
-senders and receivers (the former confuse choice points, the latter confuse choices), we need
-to look separately at each case.
+the dynamic that ensues from imitation of successes. Since imprecision works slightly
+differently on senders and receivers (the former confuse choice points, the latter confuse
+choices), we need to look separately at each case.
 
@@
 have elegant and practical discrete-time solutions in:\footnote{To see how the discrete-time
-  formulation gives rise to the continuous-time formulation, let's assume that update steps are
-  infinitesimally small, so that
+  formulations gives rise to the continuous-time formulations, let's assume that update steps are
+  infinitesimally small, so that, for the sender case:
   \begin{align*}
@@
 confusability of states into account. The main difference is that here we are not looking at,
-say, the probability $\Sstrat(\messg \probbar \state)$ with which $\messg$ is be played in
+say, the probability $\Sstrat(\messg \probbar \state)$ with which $\messg$ is played in
 response to $\state$ on average in the population, but at the probability $P_o(\messg \probbar
@@
 $\state$. Similarly for the receiver, we are not looking at occurrence probability
-$\Sstrat(\state \probbar \messg)$ but at the probability $P_o(\state \probbar \messg)$ with
+$\Rstrat(\state \probbar \messg)$ but at the probability $P_o(\state \probbar \messg)$ with
 which an agent who observes $\messg$ observes another agent play $\state$. It is easy to check
@@
 probabilities $P_o$ and $P_r$ are trivial, i.e., assigning zero probability to all pairs of
-non-identical states.
+non-identical states. The noisy replicator dynamic in (\ref{eq:nRD}) conservatively extends the
+classic case in (\ref{eq:RDDiscrete}).
 

### 2015-08-13 paper/version_04/paper.tex
first pre-final version 04

@@
   efficient signal use, and, moreover, that it unifies evolutionary outcomes and helps avoid
-  suboptimal categorization. \myred{The upshot of this is that generalization emerges, at an
-    aggregate level, without agents actually generalizing.}
+  suboptimal categorization. The upshot of this is that we see \emph{as-if}-generalization at
+  an aggregate level, without agents actually generalizing.
 \end{abstract}
@@
 low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
-contributes to the latter line of thought. Concretely, we introduce an dynamic that describes
-average behavior resulting from imitation under confusability of similar stimuli. The dynamic
-proposed here complements previous like-minded accounts. We show that stochastic noise can not
-only lead to vague, yet relatively efficient signal use, but can also unify evolutionary
-outcomes and help avoid suboptimal categorization. Moreover, the system shows generalization
-behavior without a single generalizing agent.
+contributes to the latter line of thought. Concretely, we introduce a dynamic that describes
+average behavior resulting from imitation under confusability of similar stimuli. We show that
+stochastic noise can not only lead to vague, yet relatively efficient signal use, but can also
+unify evolutionary outcomes and help avoid suboptimal categorization: regularities in
+confusability of stimuli also regularize signaling behavior and evolving categories. The system
+gives rise to generalization behavior without agents actually generalizing.
 
@@
 
-How does a tendency to confuse similar states interact with a drive towards more efficient
-signaling strategies under the imprecise conditional imitation dynamic? To be able to properly
-address this question, it pays to be more precise about similarity of states, the conditional
-confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max games.
-
-\subsection{Confusion \& utilities}
-\label{sec:iterated-diffusion}
-
-Confusion of states should intuitively be a function of their perceptual similarity: the more
-similar two states are, the more likely they could be mistaken for each other.  To make this
-concrete, let us assume that the state space of the sim-max game consists of $n \ge 2$ states
-that are equally spaced across the unit interval, including $0$ and $1$. The distance
-$\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between two states
-$\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual similarity
-function, as described by \citet{Nosofsky1986:Attention-Simil}:
+How does a tendency to confuse similar states interact with selective pressure towards more
+efficient signaling strategies under the imprecise conditional imitation dynamic? If confusion
+probabilities are moderate and regular, in that they track similarity of states, we might
+expect a regularizing effect also on evolving signaling strategies. Indeed, we hypothesize that
+confusion of states can give rise to population-level aggregate behavior that looks \emph{as
+  if} information about what is good for one state percolates also to similar states. In other
+words, imprecise imitation may make signaling behavior on average look like agents generalize
+across similar states, even if no agent actually generalizes. To explore whether confusion of
+states can have this effect, we turn to numerical simulation.
+
+\subsection{Setting the stage}
+\label{sec:setting-stage}
+
+To obtain concrete results, we must fix how to represent states, similarity of states, the
+conditional confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max
+games. Confusion probabilities between states should be a function of perceptual similarity:
+the more similar two states are, the more likely they could be mistaken for each other.
+
+Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
+interval, including $0$ and $1$. All states occur, for simplicity, with the same probability.
+The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
+two states $\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual
+similarity function, as described by \citet{Nosofsky1986:Attention-Simil}:
 \begin{align*}
@@
 \end{align*}
-For $\imprecision = 0$, we obtain trivial confusion probabilities, i.e., confusion probability
-zero between non-identical states and our dynamic reduces to the replicator dynamics. For
-$\imprecision > 0$, any state can be confused for any other state with some positive
-probability. For the uninteresting case of $\imprecision \rightarrow \infty$, confusion is
-maximal and every state can be perceived or realized as any other state with probability
-$\nicefrac{1}{\card{\States}}$.
+For $\imprecision = 0$, we obtain trivial confusion probabilities for which everything is reduced
+to the replicator dynamic. For $\imprecision > 0$, any state can be confused for any other
+state with some positive probability. For the uninteresting case of
+$\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
+realized as any other state with probability $\nicefrac{1}{\card{\States}}$.
 
@@
 
-In conclusion, to structure thinking about the behavior of the \icid, the system is governed by
-two parameters: imprecision $\imprecision$ and tolerance $\toler$, which can vary
-independently.
-
-\subsection{Numerical simulations}
+In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
+is governed by three parameters: the number $\ns = \card{T}s$ of states in the state space,
+imprecision $\imprecision$ and tolerance $\toler$.
+
+\subsection{Experimental set-up}
 \label{sec:simulations}
 
-The main question we should address is whether state confusability and utility-driven imitation
-reasonably interact, and if so, whether the inclusion of confusability has any noteworthy
-effects on the evolving meaning of signals. It transpires that moderate levels of imprecision
-$\imprecision$ can lead to reasonably efficient, yet vague signaling behavior. Moreover,
-imprecision can have further accelerating and, surprisingly, unifying effects on meaning
-evolution. It appears that state confusability has an effect on aggregate signaling behavior
-that can be described as generalization or systematization, as if ``conclusions'' about what
-works for one state are ``carried over'' to similar states. This, however, is merely an
-emergent phenomenon in the sense no single agent genuinely generalizes over stimuli or reasons
-about what a more systematic signaling strategy would be.
-
-For all simulations, priors were flat: $P_a(\state) = P_a(\state')$ for all
-$\state, \state'$. We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting
-with randomly sampled sender and receiver strategies, for each triplet of independent parameter
-values: $\ns = \card{T} \in \set{6, 10, 50, 90}$,
-$\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$, $\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each
-trial ran for a maximum of 200 update steps of the \rdd. A trial was considered converged, and
-thus stopped before the maximum of 200 rounds, if the total amount of change between strategies
-before and after an \rdd step was smaller than a suitably chosen threshold. It is not
-guaranteed that strategies at halting time had converged to the eventual attracting state,
-whether they ran for 200 rounds or not. Our notion of convergence is therefore only a
-categorical measure for reaching a certain (well-considered, but eventually arbitrary) degree
-of stability. In principle, rather than setting a hard limit on the maximum number of rounds
-and analyzing convergence categorically, one could also run each trial until reaching our
-criterion for convergence and analyze the number of rounds it took to get there.  Our choice
-was initially motivated by pragmatic concerns regarding length of simulation time, but it is
-also justifiable if we reflect on the nature of language evolution as a continuous process.
-Language adapts to an ever-changing reality; it is rare, if not impossible, that a lexicon has
-an infinite amount of time to neatly converge to a stable state. Therefore, investigating speed
-of convergence within a limited time frame can be seen as better reflecting the natural
-constraints of language evolution.
+We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
+sender and receiver strategies, for each triplet of independent parameter values:
+$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
+$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps of the
+\rdd. A trial was considered converged, and thus stopped before the maximum of 200 rounds, if
+the total amount of change between strategies before and after an update step was smaller than
+a suitably chosen threshold. It is not guaranteed that strategies at halting time had converged
+to the eventual attracting state, whether they ran for 200 rounds or not. Our notion of
+convergence is therefore only a categorical measure for reaching a certain (well-considered,
+but eventually arbitrary) degree of stability. In other words, out notion of ``convergence'' is
+a measure of relative speed: it is true that the system reached a state in which evolutionary
+adaptations had slowed down almost to a halt before 200 update steps? This is motivated by
+pragmatic concerns regarding length of simulation time, but also theoretically justified,
+because we hypothesize that confusability of states leads to regularization of evolving
+strategies, which would show exactly in an increased speed of evolutionary trajectories towards
+well-behaved and regular signaling behavior.
 
@@
 pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
-the trial was stopped for reaching the maximum of 200 rounds. In contrast, the outcome of a
-trial with identical parameters, except with imprecision $\imprecision = 0.0.5$, which is shown
-in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99 rounds of
-iterated \rdd. The sender strategy shows a smooth blending from one ``category'' to the other,
-and the receiver's interpretations are rather extended curves, peaking at a central point in
-the relevant ``categories.''
+the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
+trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
+in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
+rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
+receiver's interpretations are rather extended curves, peaking at a central point in the
+relevant ``categories.''
 
@@
 
-\subsubsection{Dependent measures}
-
-To further explore our simulation results, we calculated metrics that
-aim to numerically capture how vague, generally well-structured, and
-communicatively efficient the recorded strategy pairs
-were. \emph{Entropy} captures the amount of systematicity or
-regularity in signal use. \emph{Convexity} captures whether a
-behavioral strategy would project onto a convex pure
-strategy. \emph{Expected utility} measures the communicative
-efficiency of evolved strategy pairs.
+\subsection{Dependent measures}
+
+To further explore our simulation results, we calculated metrics that aim to numerically
+capture how vague, generally well-structured, and communicatively efficient the recorded
+strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
+signal use. \emph{Convexity} captures whether a behavioral strategy would project onto a convex
+pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
+strategy pairs.
 
@@
 
-\paragraph{Convexity.} At least for sender strategies, which develop
-faster than receiver strategies, it also makes sense to define a
-categorical measure of convexity that compensates for potential
-vagueness. To determine whether a sender strategy $\Sstrat$ is convex
-despite possibly being vague, we look at the derived pure strategy
-$\Spure$ for which $\Spure(\state) = \argmax_{\messg' \in \Messgs}
-\Sstrat(\state,\messg')$. If that $\Spure$ is convex, we also count
-$\Sstrat$ as convex. The sender strategy in
+\paragraph{Convexity.} At least for sender strategies, which develop faster than receiver
+strategies, it also makes sense to define a categorical measure of convexity that compensates
+for potential vagueness. To determine whether a sender strategy $\Sstrat$ is convex despite
+possibly being vague, we look at the derived pure strategy $\Spure$ for which
+$\Spure(\state) = \argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$ is
+convex, we also count $\Sstrat$ as convex. The sender strategy in
 Figure~\ref{fig:example_stratsA} is not convex, while the one in
-Figure~\ref{fig:example_stratsB} is.  One would a priori not expect a
-systematic relation between imprecision and convexity.
+Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
+strategies, like we hypothesized, we should see more convexity with increasing impairment all
+else equal.
 
@@
 
-\subsubsection{Results}
+\subsection{Results}
 
@@
 
-Figure~\ref{fig:MeanMetrics} shows plots summarizing some results.  For perspicuity, we only
-plot results for one level of tolerance $\beta = 0.1$, and leave out the case of $n_s =
-90$. Still, every qualitative trend mentioned in the following applies to the whole set of
-results.
-
-As expected, increasing imprecision leads to higher entropy and lower
-expected utility. Importantly, however, imprecision does not
-necessarily lead to disastrous decline of communicative success. What
-was not expected, on the other hand, was a relation between
-imprecision and convexity. However, the corresponding plot clearly
-shows that higher imprecision led to a higher number of outcomes with
-convex sender strategies. Moreover, it also led to higher likelihood
-of convergence. In fact, sufficient imprecision always ensured
-convergence and convexity. It appears that perceptual imprecision
-leads to more vagueness, slightly less communicative efficiency, but
-more regular, well-behaved languages in shorter time.
+Figure~\ref{fig:MeanMetrics} shows plots summarizing a selected part of our findings.  For
+perspicuity, we only plot results for one level of tolerance $\beta = 0.1$, and leave out the
+case of $n_s = 90$. Still, every qualitative trend mentioned in the following applies to the
+whole set of results.
+
+As expected, increasing imprecision leads to higher entropy and lower expected
+utility. Importantly, however, imprecision does not necessarily lead to disastrous decline of
+communicative success. What is more, in line with our hypothesis that mere imprecision in
+imitation behavior can lead to behavior that looks like agents generalize across similar
+stimuli, higher imprecision led to a higher number of outcomes with convex sender
+strategies. It also led to higher rates of convergence. In fact, sufficient imprecision always
+ensured convergence and convexity. It appears that perceptual imprecision leads to more
+vagueness, slightly less communicative efficiency, but more regular, well-behaved languages in
+shorter time.
 
@@
 differences, the general trend is the same for all other parameter settings that we tested:
-with increasing imprecision, the sender strategies evolving under the \rdd were much more alike
-(modulo swapping of messages). This means that perceptual imprecision unifies evolutionary
-outcomes and guarantees sender strategies that are not only convex, but also regular in that
-they induce a vague category split exactly in the middle of the unit interval.
+with increasing imprecision, the resulting sender strategies were much more alike (modulo
+swapping of messages). This means that perceptual imprecision can speed up and unify
+evolutionary outcomes. It can amplify the emergence of sender strategies that are not only
+convex, but also regular in that they induce a vague category split exactly in the middle of
+the unit interval.
 
@@
 
-This unifying property of perceptual imprecision could be considered
-an evolutionary beneficial side-effect. Higher imprecision can lead to
-higher \emph{within group expected utility}, defined as the average
-expected utility that each evolved language scored when playing against an
-arbitrary other language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a representative
-example. The observation repeats for other parameter values: while
-imprecision might decrease the communicative efficiency of individual
-languages, it increases the conceptual coherence and communicative
-success between independently evolving strategies. It is almost as if
-mere confusability of states imposes a regularity constraint on
-evolving categories.
+This unifying property of perceptual imprecision could be considered an evolutionary beneficial
+side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
+defined as the average expected utility that each evolved language scored when playing against
+an arbitrary other language obtained for the same parameter
+values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
+for other parameter values: while imprecision might decrease the communicative efficiency of
+individual languages, it increases the conceptual coherence and communicative success between
+independently evolving strategies. It is as if mere confusability of states imposes a
+regularity constraint on evolving categories.
 
@@
 
-\section{Discussion}
+\section{Conclusion}
 \label{sec:discussion}
 
-\subsection{Interpretation of model and results}
-\label{sec:model-interpretation}
-
-The \rdd adds expected confusion of similar states to the \rd in behavioral strategies. The
-result is an abstract evolutionary dynamic that describes the most likely path of development
-of aggregate population behavior that is consistent with the idea that agents imitate other
-agents' behavior (Appendix~\ref{sec:deriv-rdd-imit}). This would be a process of cultural
-evolution operating on behavioral strategies. But we speculate that the \rdd is also extensible
-to biological evolution on mixed strategies, in which case it is a special case of the \rmd
-with mutation matrices derived from confusion matrices, as shown in
-Appendix~\ref{sec:diffusion-as-special}.
-
-The \rdd is not only compatible with several interpretations of how,
-at an agent level, the population-level changes in strategy
-proportions come about. It is also compatible with several
-interpretations of diffusion. Diffusion can be pictured as a flaw or a
-virtue. If we think of perceptual imprecision as a hard limitation on
-information processing, vagueness is a natural byproduct of language
-evolution. Results from numerical simulations show that such
-imprecision leads to vagueness, but does not need to undermine the
-prospect of evolving regular and by-and-large efficient signaling
-behavior. On the other hand, there appears to be a possibly beneficial
-side-effect of state confusability: it speeds up convergence to
-regularly shaped and convex categories. This suggests that diffusion
-of behavior, as we implemented here, could also be seen as a form
-of stimulus-generalization that is evolutionarily beneficial, because
-it accelerates the emergence of uniform category formation. This
-latter view is also suggested by a like-minded account to which we
-turn next.
-
-\subsection{Relation with previous accounts}
-\label{sec:relat-with-prev}
-
-\paragraph{Generalized reinforcement.}
-
-% Under \emph{Herrnstein reinforcement learning}, sender and receiver
-% adjust their dispositions to act after each round of play, in such a
-% way that the actual (non-negative) payoff gained in the previous
-% interaction is added to the non-normalized propensities for repeating
-% the same behavior. For sim-max games, this means that, when the sender
-% chose $\messg$ in state $\state$, and this resulted in some
-% non-negative payoff, the probability that the sender chooses $\messg$
-% again in $\state$ is increased, but nothing else changes. In
-% particular, the sender's behavior in other choice points does not
-% change. \emph{Generalized reinforcement learning} is different. When
-% the use of $\messg$ in $\state$ gave positive payoff, then not only
-% will the probability of its future use be promoted at $\state$ but also at
-% other states, proportional to how similar these are to
-% $\state$. Similar amendments take care of the way that the receiver
-% updates his choice dispositions.
-
-\citet{OConnor2013:The-Evolution-o} introduces a variant of classical Herrnstein reinforcement
-learning for sim-max games, in which successful play leads to reinforcement of acts also for
-states similar to the one that actually occurred. This is a form of generalization to unseen
-contingencies. It not only leads to vague signaling of the appropriate kind, but also speeds up
-learning in such a way that, especially for sim-max games with higher numbers of states, higher
-levels of communicative success are reached in shorter learning
-periods. % Technically, this result is
-% partly due to the fact that signalers make bigger changes to their behavioral strategies
-% after each round of play under generalized reinforcement than under the Herrnstein
-% variety. That only explains the speed of adaptation, not necessarily that generalization also
-% leads to regularity and communicative efficiency, but it does.
-
-Diffusion of strategies in the \rdd can also be conceived of as a form of generalization, and
-works in large part quite analogously to stimulus generalization in
-\citeauthor{OConnor2013:The-Evolution-o}'s approach. But there are still
-differences. \citeauthor{OConnor2013:The-Evolution-o} showed that her generalized reinforcement
-learning can speed up the development of efficient signaling, especially for higher numbers of
-states. Complementing this, we showed that diffusion regularizes evolutionary outcomes and
-prevents meandering around sub-optimal and non-convex strategies, also for low numbers of
-states.
-
-Conceptually, the \rdd is a more abstract framework than generalized reinforcement
-learning. The latter is most naturally seen as an agent-based learning dynamic that has two
-players adapt their individual strategies after each concrete round of play. In contrast, the
-\rdd describes a more abstract, average dynamical change in behavioral dispositions at the
-aggregate population level. Although the behavior of (some forms of) reinforcement learning
-mirror those of the \rd at some stage in time
-(\cite{BorgersSarin997:Learning-Throug,HopkinsPosch2005:Attainability-o,Beggs2005:On-the-Converge}),
-this does not mean that \emph{generalized} reinforcement learning is also necessarily a
-plausible population-level description of, say, generalized learning in a population of
-agents. Seen in this light, generalized reinforcement learning and the \rdd nicely complement
-each other, as similarly-minded accounts operating at different levels of abstraction.
-Although it would be interesting to explore the possibility of a mathematical connection
-between the two, this is beyond the scope of this paper.
-
-\paragraph{Quantal response equilibria.}
+We set out to meet a technical challenge posed by Lipman's problem
+\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
+formal model that shows how vague language can evolve under selective pressure for efficient
+communication? On the assumption that agents are likely to mistake states for one another,
+especially when they are perceptually similar, derived a generalization of the replicator
+dynamic in terms of imitation of successes. The resulting population-level dynamic produced
+signaling behavior that is at the same time regular and by-and-large communicatively efficient,
+while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
+by-product of an arguably natural limitation on the discriminatory power of signaling agents.
+
+But the inclusion of confusability of similar states also had noteworthy effects on the
+evolving meaning of signals. It transpired that imprecision can have further accelerating and,
+surprisingly, unifying effects on meaning evolution. State confusability has an effect on
+aggregate signaling behavior that can be described as generalization without generalizers: the
+dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
+``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
+single agent genuinely generalizes over stimuli or reasons about what a more systematic
+signaling strategy would be.
+
+In sum, inability to sharply discriminate similar stimuli may lead to vagueness and bring about
+a (slight) decrease in communicative efficiency. But there may also be an advantage, not of
+vagueness itself, but of its reason: systematicity in the confusability of states (which may be
+a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
+population-level without having to assume that agents themselves have any generalization power.
+
+\citet{OConnor2013:The-Evolution-o} makes a related case based on a version of reinforcement
+learning for sim-max games, in which successful play leads to reinforcement of choice options
+also for states similar to the ones that actually occurred. This not only leads to vague
+signaling of the appropriate kind, but also speeds up learning in such a way that, especially
+for sim-max games with higher numbers of states, higher levels of communicative success are
+reached in shorter learning periods. Our results complement and extend
+\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
+similar regularizing effects also for cases with low numbers of states and do not assume that
+agents have any kind of generalizing capacity in and of themselves, even if that is only
+implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
+
 \citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
@@
 known as a \emph{soft-max function}
-\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. A
-quantal response is a stochastic generalization of the best response function (i.e., the
-deterministic choice of an action that maximizes expected utility). Quantal responses factor in
-stochastic trembles at the level of the computation or maximization of expected utilities:
-agents may make mistakes in calculating how good given choice options really are for
-them. \citet{FrankeJager2010:Vagueness-Signa} demonstrate that quantal response equilibria of
-sim-max games can show the desired marks of vague
-signaling. Figure~\ref{fig:exampleQRE_stratsA} gives an example of a quantal response
-equilibrium for a sim-max game, as used in our set-up but with higher tolerance $\toler =
-0.5$.
-Sender and receiver strategies look very much like what evolves under \rdd with modest values
-of perceptual imprecision.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance05.pdf}
-    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.5$}
-    \label{fig:exampleQRE_stratsA}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/exampleStratQRE_tolerance005.pdf}
-    \caption{$\ns = 90$, $\lambda = 15$, $\toler = 0.05$}
-    \label{fig:exampleQRE_stratsB}
-  \end{subfigure}
-
-  \caption{Examples of vague quantal response equilibria. Here, $\lambda$ is an imprecision
-    parameter measuring the extent of noise in the calculation of expected utilities
-    \citep[see][for details]{FrankeJager2010:Vagueness-Signa}.}
-  \label{fig:exampleQREs}
-\end{figure}
-
-But not all quantal response equilibria are equally plausible.  To see what the problem is,
-consider Figure~\ref{fig:exampleQRE_stratsB}, which is the quantal response equilibrium for a
-game with lower tolerance $\toler = 0.05$. Unlike what evolves under \rdd in this case, sender
-strategies have vague boundaries also towards the end of the unit intervals. Technically, this
-is because quantal responses equalize message use far away from the ``prototypical''
-interpretation, not just in-between categories, so to speak. Conceptually, this is because
-quantal responses introduce noise at the level of computing or maximizing expected utility of
-choices.
-
-In contrast, the \rdd implements noise at the level of the perception of different
-states. Unintuitive outcomes like in Figure~\ref{fig:exampleQRE_stratsB} are thereby
-avoided. Still, we believe that it would be most plausible to assume that both perceptual
-confusion \emph{and} computational errors in maximizing expected utility play a role in making
-vague categorization a pervasive feature of natural language meaning. Perhaps even further
-factors, such as limited observations, imperfect memory or uncertainty about the exact prior
-distribution of states, should be considered in order to form a more complete understanding of
-the phenomenon of vagueness.
+\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
+main difference between this and our present approach is in where stochastic noise is assumed
+to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
+in case of imprecise imitation, it is perception and realization of similar states. There are
+cases, then, where evolving signaling behavior, as predicted by these two approaches, is quite
+different. Intuitively speaking, for a case with two messages, the further we venture away from
+a prototypical interpretation of either message, the less discriminative a signaling strategy
+would be when the source of ``trembles'' is the computation of expected utilities: to wit,
+since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
+response dynamics predict that speaker's would be almost indifferent. Speaker behavior that
+evolves under confusion of states does not have this puzzling property.
 
@@
 
-\section{Conclusions}
-\label{sec:conclusions}
-
-Vagueness is a pervasive but seemingly non-disruptive feature of
-natural communication and classification systems. From an evolutionary
-point of view, the challenge arises to explain how vagueness can
-persist under selective pressure for precision. This is foremost a
-technical challenge, probing for the possibility of integrating into a
-consistent model forces that lead to vagueness and forces that lead
-to efficient information transfer. This paper proposed one such model
-in the context of sim-max games and explored some of its consequences.
-
-We introduced the replicator diffusion dynamic and identified it as a special case of the
-replicator mutator dynamic (Appendix~\ref{sec:diffusion-as-special}). The \rdd incorporates a
-special kind of stochastic mutation due to differential confusability of similar states. It can
-be thought of as the aggregate population-level behavior process that results from imitation
-among agents that are unaware of perceptual confusion
-(Appendix~\ref{sec:deriv-rdd-imit}). Based on data from numerical simulations, we demonstrated
-that the inclusion of mild levels of perceptual imprecision does not undermine the possibility
-of evolving relatively successful signaling strategies. On the contrary, strategies that
-evolved under mild imprecision induce a highly regular and systematic category structure that
-shows the signs of vagueness as desired. There might even be a higher-order benefit to the
-presence of imprecision, in that it can accelerate convergence to optimal categorization, by
-preventing evolutionary paths to stay near inefficient non-convex strategies for a long time. A
-little imprecision thus also unifies and regularizes evolutionary outcomes in such a way that
-sub-optimal categorization is avoided in short-term trajectories.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \printbibliography[heading=bibintoc]

### 2015-09-10 paper/version_04/paper.tex
Created separate Discussion section. Added argument and citations for group selection.

@@
 
-This unifying property of perceptual imprecision could be considered an evolutionary beneficial
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Vagueness and evolution}
+The inclusion of confusability of similar states has noteworthy effects on the
+evolving meaning of signals. It transpired from our results that imprecision can have further accelerating and,
+surprisingly, unifying effects on meaning evolution. State confusability has an effect on
+aggregate signaling behavior that can be described as generalization without generalizers: the
+dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
+``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
+single agent genuinely generalizes over stimuli or reasons about what a more systematic
+signaling strategy would be.
+
+The unifying property of perceptual imprecision could be considered an evolutionary beneficial
 side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusion}
-\label{sec:discussion}
-
-We set out to meet a technical challenge posed by Lipman's problem
-\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
-formal model that shows how vague language can evolve under selective pressure for efficient
-communication? On the assumption that agents are likely to mistake states for one another,
-especially when they are perceptually similar, derived a generalization of the replicator
-dynamic in terms of imitation of successes. The resulting population-level dynamic produced
-signaling behavior that is at the same time regular and by-and-large communicatively efficient,
-while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
-by-product of an arguably natural limitation on the discriminatory power of signaling agents.
-
-But the inclusion of confusability of similar states also had noteworthy effects on the
-evolving meaning of signals. It transpired that imprecision can have further accelerating and,
-surprisingly, unifying effects on meaning evolution. State confusability has an effect on
-aggregate signaling behavior that can be described as generalization without generalizers: the
-dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
-``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
-single agent genuinely generalizes over stimuli or reasons about what a more systematic
-signaling strategy would be.
-
-In sum, inability to sharply discriminate similar stimuli may lead to vagueness and bring about
-a (slight) decrease in communicative efficiency. But there may also be an advantage, not of
-vagueness itself, but of its reason: systematicity in the confusability of states (which may be
-a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
-population-level without having to assume that agents themselves have any generalization power.
-
+Although in our model the imprecision parameter was fixed for each simulation run, and thus not subject to evolutionary selection.
+However, given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
+This is for now purely speculative, but it makes sense especially if we consider it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.
+
+\subsection{Related work}
 \citet{OConnor2013:The-Evolution-o} makes a related case based on a version of reinforcement
@@
 
+\section{Conclusion}
+\label{sec:conclusion}
+
+We set out to meet a technical challenge posed by Lipman's problem
+\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
+formal model that shows how vague language can evolve under selective pressure for efficient
+communication? On the assumption that agents are likely to mistake states for one another,
+especially when they are perceptually similar, derived a generalization of the replicator
+dynamic in terms of imitation of successes. The resulting population-level dynamic produced
+signaling behavior that is at the same time regular and by-and-large communicatively efficient,
+while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
+by-product of an arguably natural limitation on the discriminatory power of signaling agents.
+Although inability to sharply discriminate similar stimuli may lead to vagueness and bring about
+a (slight) decrease in communicative efficiency, there may also be an advantage, not of
+vagueness itself, but of its reason: systematicity in the confusability of states (which may be
+a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
+population-level without having to assume that agents themselves have any generalization power.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \printbibliography[heading=bibintoc]

### 2015-09-20 paper/version_04/paper.tex
Added discussion point on levels of vagueness (population vs individual). Small further corrections to discussion section.

@@
 
-\subsection{Vagueness and evolution}
-The inclusion of confusability of similar states has noteworthy effects on the
-evolving meaning of signals. It transpired from our results that imprecision can have further accelerating and,
-surprisingly, unifying effects on meaning evolution. State confusability has an effect on
-aggregate signaling behavior that can be described as generalization without generalizers: the
-dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
-``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
-single agent genuinely generalizes over stimuli or reasons about what a more systematic
-signaling strategy would be.
-
+\subsection{Levels of vagueness}
+The main motivation and intuitions given in Section~\ref{sec:repl-diff-dynam} describe probabilities in sender and receiver strategies as capturing the overall distribution of pure strategies in a population: $\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the receiver).
+This is in line with the standard interpretation of the replicator equation, but we should consider its philosophical implications.
+Based on this interpretation and our model, vagueness is seemingly being characterized as a strictly population-level phenomenon since it arises in a signaling system from the inability of individual agents to fully align their (non-vague) strategies because of imprecision.
+This is, we believe, a plausible mechanism that can already explain the existence of vagueness in a language even if we assume each agent commands a non-vague idiolect.
+
+We do not, however, want to commit to the idea that vagueness does not exist at the level of individual agents.
+The fact is that the model can be interpreted in alternative ways.
+The probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will be used by a sender when observing state $\state$.
+Whether the uncertainty represents variations in a population of agents playing pure strategies, a single sender using a non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is up for interpretation.
+The first option supported what we believe to be a plausible motivation for the mechanism of imitation of success with imprecision, but other motivations could be imagined to better fit an interpretation that includes non-deterministic agents.
+We thus believe that the model introduced in this paper does not exclude vagueness at the individual level, but to explore all possible interpretations would go beyond our current scope.
+
+\subsection{Evolutionary benefits of imprecision}
+The inclusion of confusability of similar states has noteworthy effects on the evolving meaning of signals.
+It transpired from our results that imprecision can have further accelerating and, surprisingly, unifying effects on meaning evolution.
 The unifying property of perceptual imprecision could be considered an evolutionary beneficial
@@
 
-Although in our model the imprecision parameter was fixed for each simulation run, and thus not subject to evolutionary selection.
-However, given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
-This is for now purely speculative, but it makes sense especially if we consider it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.
+The phenomenon could potentially not just be a side effect.
+Given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
+This could be achieved by letting the imprecision parameter be subject to the evolutionary dynamics.
+The idea is for now mostly speculative, but we consider it an interesting avenue for future research.
+There is additional motivation to consider its potential if we see it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.
 
 \subsection{Related work}
-\citet{OConnor2013:The-Evolution-o} makes a related case based on a version of reinforcement
+\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of reinforcement
 learning for sim-max games, in which successful play leads to reinforcement of choice options
@@
 implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
+State confusability has an effect on
+aggregate signaling behavior that can be described as generalization without generalizers: the
+dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
+``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
+single agent genuinely generalizes over stimuli or reasons about what a more systematic
+signaling strategy would be.
 
 \citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
-information-processing limitations could lead to vague strategies. The model that is most
+in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
 clearly related to the present approach uses the notion of a \emph{quantal response}, also
@@
 since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
-response dynamics predict that speaker's would be almost indifferent. Speaker behavior that
+response dynamics predict that senders would be almost indifferent. Sender behavior that
 evolves under confusion of states does not have this puzzling property.

### 2015-09-23 paper/version_04/paper.tex
Added a couple of diagrams to illustrate a round of play in normal sim-max games versus when imprecision is added.

@@
 
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
+  \node (3) [right=of 2] {$\messg = \Spure(\mystate{1}) \in \Messgs$};
+  \node (4) [right=of 3] {$\mystate{2} = \Rpure(\messg) \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {S} (3)
+    (3) edge node {R} (4);
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg = \Spure(\mystate{1})$ determined by his pure strategy; a receiver (R) selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy.
+Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
+   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
+  \label{fig:sim-max-illustration}
+\end{figure}
+
 Similarity-maximizing (short: sim-max) games are a variation of
@@
 
-A sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ with much
+Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
 fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
@@
 function of similarity.
+In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
+A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.
 
@@
 confusion probabilities.
-
 A single round of play of a sim-max game is then governed by five pieces of stochastic
-information, where previously there were only three. At the outset, the actual state is chosen
-with probability $P_a(\mystate{a})$. A randomly sampled sender observes $\mystate{o}$ with
-probability $P_o(\mystate{o} \probbar \mystate{a})$ and so we expect $\messg$ with probability
-$\Sstrat(\messg \probbar \mystate{o})$. (Senders respond to observations, not the actual
-state!) As perception of messages is flawless, we expect a randomly sampled receiver to intend
-to realize interpretation $\mystate{i}$ with probability
-$\Rstrat(\mystate{i} \probbar \messg)$.  The actually realized interpretation will be
-$\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.
+information, where previously there were only three.
+An illustration is given in Figure~\ref{fig:imprecise-sim-max-illustration}.
+Note that, and this is crucial, senders respond to observations not actual states, receiver strategies determine intentions not realized states, but payoff is calculated based on actual and realized states.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$P_a(\mystate{a})$] {$\mystate{a} \in \States$};
+  \node (2') [right=of 2,label=below:$P_o(\mystate{o} \probbar \mystate{a})$] {$\mystate{o} \in \States$};
+  \node (3) [right=of 2',label=below:$\Sstrat(\messg \probbar \mystate{o})$] {$\messg \in \Messgs$};
+  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{i} \probbar \messg)$] {$\mystate{i} \in \States$};
+  \node (4') [right=of 4,label=below:$P_r(\mystate{r} \probbar \mystate{i})$] {$\mystate{r} \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {} (2')
+    (2') edge node {S} (3)
+    (3) edge node {R} (4)
+    (4) edge node {} (4');
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game with imprecise imitation: Nature (N) chooses an actual state with probability $P_a(\mystate{a})$; a randomly sampled sender (S) observes $\mystate{o}$ with probability $P_o(\mystate{o} \probbar \mystate{a})$ and subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.
+Payoff for both sender and receiver is given by $\Utils(\mystate{a}, \messg, \mystate{r})$.}
+  %which is typically proportional to state similarity $\similarity(\mystate{a}, \mystate{r})$.}
+  \label{fig:imprecise-sim-max-illustration}
+\end{figure}
 

### 2015-09-29 paper/version_04/paper.tex
First diagram with behavioral strategies, better for comparing with the other diagram. (cherry picked from commit 9c9e0bb)

@@
 \usepackage{subcaption}
+
+\usepackage{a4wide}
 
@@
 The next section introduces the background against which the work presented here can be
-appreciated. Section~\ref{sec:repl-diff-dynam} introduces our noise-affected imitation dynamic,
-which is then explored in Section~\ref{sec:exploring-rdd}.  Section~\ref{sec:discussion}
-reflects and compares our approach to others. % Two appendices cover technical material that is
-% not strictly necessary to appreciate the main conceptual points of the paper.
-% Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
-% special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
-% gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
-% successful behavior without being aware of perceptual noise themselves.
+appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
+dynamic which is derived from the idea that agents imitate other agents' behavior while
+possibly confusing similar states. This imprecise imitation dynamic is explored in
+Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
+approach to others. Appendix~\ref{sec:repl-diff-dynam} provides formal detail.
 
@@
 \citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
-  \node (3) [right=of 2] {$\messg = \Spure(\mystate{1}) \in \Messgs$};
-  \node (4) [right=of 3] {$\mystate{2} = \Rpure(\messg) \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {S} (3)
-    (3) edge node {R} (4);
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg = \Spure(\mystate{1})$ determined by his pure strategy; a receiver (R) selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy.
-Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
-   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
-  \label{fig:sim-max-illustration}
-\end{figure}
 
@@
 function of similarity.
-In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
-A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.
+A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the \emph{probability} that a given message would be used for each state.
+The analogous happens for the receiver's behavior.
+A play in such a game with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
+  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
+  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {S} (3)
+    (3) edge node {R} (4);
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a receiver (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
+Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
+   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
+  \label{fig:sim-max-illustration}
+\end{figure}
 
@@
 dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
-language with two messages would have the sender use one message exclusively for all points in
+language with two messages could have the sender use one message exclusively for all points in
 the lower half of the unit interval and another for all points in the upper half; the
@@
         \caption{Example of a Voronoi language on $\States = [0,1]$. The (pure) sender strategy
-          $\Spure \in \Messgs^\States$ uses one signal for the lower half, and another for the
-          upper half of the unit interval. The (pure) receiver strategy $\Rpure \in
-          \States^\Messgs$ selects the central elements in the respective intervals. See
-          \citet{JagerMetzger2011:Voronoi-Languag} for further details.}
+          $\Spure$ uses one signal for the lower half, and another for the upper half of the
+          unit interval. The (pure) receiver strategy $\Rpure$ selects the central elements in
+          the respective intervals. See \citet{JagerMetzger2011:Voronoi-Languag} for further
+          details.}
   \label{fig:Voronoi-language-example}
@@
 
-\section{Imprecise conditional imitation}
+\section{Imprecise imitation}
 \label{sec:repl-diff-dynam}
@@
 
-Fix a signaling game with finite states $\States$, messages $\Messgs$ and acts $\Acts$. Let
+Fix a sim-max game with finite states $\States$ and messages $\Messgs$. As usual, we assume
+that the receiver chooses states in $\States$ in response to messages. Let
 $\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
-$\Util_{\sen,\rec} \mycolon \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ the
-sender's and receiver's utility functions. A \emph{behavioral strategy} is a function that maps
-an agent's choice points to a probability distribution over available choices.\footnote{Our
-  focus is on behavioral strategies, not \emph{mixed strategies}, i.e., probability
-  distributions over functions from each choice point to an act, such as
+$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility
+function shared by senders and receivers in the population. A \emph{behavioral strategy} is a
+function that maps an agent's choice points to a probability distribution over available
+choices.\footnote{Our focus is on behavioral strategies, not \emph{mixed strategies}, i.e.,
+  probability distributions over functions from each choice point to an act, such as
   $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
-  adapt their behavior locally, so to speak, i.e., independently at each choice point. Our
-  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
-  numerical simulations. But it also seems the more plausible choice for imitation-based update
-  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
-  particular situation, not how that agent would behave in \emph{all} relevant choice
-  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
-  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
-  behavioral strategies.)} The sender's behavioral strategies are functions
-$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
-probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's are
-functions $\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$
-to a probability of each act $\act \in \Acts$ being chosen in response to $\messg$. Although
-behavioral strategies are probabilistic, we may conveniently imagine that every individual
-agent has a non-probabilistic strategy nonetheless. If the population is large enough, the
-number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability that a randomly
-sampled sender would have a strategy that responds with $\messg$ to $\state$.
+  adapt their behavior locally, i.e., independently at each choice point. Our focus on
+  behavioral strategies greatly reduces the complexity of the dynamic and simplifies numerical
+  simulations. But it also seems the more plausible choice for imitation-based update protocols
+  of the kind we consider here: agents only observe how, on some occasion, some other agent
+  behaved in \emph{one} particular situation, not how that agent would behave in \emph{all}
+  relevant choice situations; they imitate the use of a single word, so to speak, not a whole
+  lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference between
+  dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
+functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
+to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
+are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
+$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
+in response to $\messg$. Although behavioral strategies are probabilistic, evolutionary
+modeling usually imagines that every individual agent has a non-probabilistic
+strategy. Behavioral strategies then capture average population behavior. Assuming a virtually
+infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
+probability that a randomly sampled sender would send message $\messg$ if the actual state was
+$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
+randomly sampled receiver interprets $\messg$ as $\state$.
 
@@
 \begin{align*}
-  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
-  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
-  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
-    \States} P(\state \probbar \messg)  \cdot
-  U_\rec(\state, \messg, \act)\,, \text{\ \  where } &
-P(\state \probbar \messg) & \propto \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \,.
+  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
+  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
+  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
+    \States} P(\state \probbar \messg)  \myts
+  U(\state, \state')\,, \text{\ \  where } &
+P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
 \end{align*}
 Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
-outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
-sender of choosing message $\messg$, given tate $\state$ and a receiver
-strategy $\Rstrat$ is the sum, for each act $\act$, of the probability that the
-receiver will choose $\act$ given $\messg$ times the actual utility of triple
-$\state$, $\messg$ and $\act$.
-
-The continuous-time replicator dynamic for behavioral strategies describes the mean rate of
-change in the proportion of each choice, independently for each choice point:
-\begin{align}
-  \label{eq:RDContinuous}
-  \dot{\Sstrat}(\messg \probbar \state) & = \underbrace{\Sstrat(\messg \probbar
-    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
-  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
-      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
-      EU at choice point $\state$}} \Big ) \\
-  \dot{\Rstrat}(\act \probbar \messg) & = \Rstrat(\act \probbar \messg) \thinspace \Big ( \EU(\act,
-  \messg, \Sstrat) -  \sum_{\act'} \Rstrat(\act' \probbar \messg) \cdot \EU(\act',
-  \messg,\Sstrat)  \Big ) \nonumber
-\end{align}
-The rate of change in proportion of $\messg$ choices at choice point $\state$, for instance, is
-given by the difference of $\messg$'s expected utility at $\state$ and the average expected
-utility at $\state$, times the probability of seeing $\messg$ at $\state$ in the first place.
-
-For numerical simulations, it can be more convenient to work with a discrete-time formulation,
-which maps current strategies $\Sstrat$ and $\Rstrat$ to strategies $\Sstrat'$ and $\Rstrat'$
-that describe the population at the next discrete time step. Like the continuous-time version,
-the discrete-time replicator dynamics tracks changes in frequency of choices as proportional to
-their expected utilities \citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The
-  formulation given in Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one
-  we will be looking at in Section~\ref{sec:exploring-rdd}, where utilities are always
-  non-negative and expected utilities are always positive.}
+outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
+of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
+sum, for each interpretation $\state'$, of the probability that some random receiver will
+choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
+$\state'$.
+
+The discrete-time replicator dynamic tracks changes in frequency of choices in the population
+as proportional to their expected utilities
+\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
+  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
+  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
+  utilities are always positive.}
 \begin{align}
   \label{eq:RDDiscrete}
-  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \cdot
-    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \cdot
+  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
+    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \myts
     \EU(\act, \messg,\Sstrat)\,.   
@@
 
-\subsection{Conditional imitation of success}
-\label{sec:cond-imit-succ}
-
 The replicator dynamic is an abstract population-level dynamic that describes the mean expected
@@
 deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
-here on one of the simplest: \emph{conditional imitation of success}
-\citep[see][]{Sandholm2010:Population-Game}. The main idea is that agents imitate the behavior
-of other agents at some choice point with a probability that is proportional to the expected
-utility of the latter agents' choice. Since the sim-max games that we will be looking at in
-Section~\ref{sec:exploring-rdd} have positive utilities, we can identify the switching
-probability with the expected utility.
-
-Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
-agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
-``teacher.'' A random learner is drawn from the population and given a chance to change
-behavior at choice point $\state$. The probability that our learner plays $\messg$ is
-$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
-$\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar \state)$. The
-learner then starts using $\messg'$ instead of $\messg$ with probability
-$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)
-
-If agents get repeated update chances for their choice points, the \emph{expected change} of
-frequency for $\messg$ at $\state$ becomes:
-\begin{align}
-  \label{eq:MeanChange}
-  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
-\end{align}
-where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
-$\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
-that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
-population, the probability $P(\messg' \rightarrow m, \state)$ of switching from $\messg'$ to
-$\messg$ at $\state$ can be written as:
-\begin{align*}
-  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
-    \state)}_{\text{learner plays $\messg'$}} \cdot
-  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
-  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} 
-  % P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
-  %   \state)}_{\text{learner plays $\messg$}} \cdot
-  % \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
-  % \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
-\end{align*}
-From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
-to:
-\begin{align*}
-  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
-  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
-      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
-& = \Sstrat(\messg \probbar \state) \cdot \Big ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
-  \state,\Rstrat) \Big )\,,
-\end{align*}
-which is the continuous-time formulation of the replicator dynamic in (\ref{eq:RDContinuous}).
-
-\subsection{Imitation of success with imprecision}
-\label{sec:imit-succ-with}
-
-The above derivation of the replicator dynamic assumes that agents can discriminate choices and
-choice points perfectly. Let's dispense with that assumption. With an eye towards sim-max
-games, we will assume that receivers choose interpretation actions $\Acts = \States$ and that
-any two states, but not messages, may be confused for one another.\footnote{It is relatively
-  straightforward to also incorporate confusability of messages, but this is irrelevant to our
-  present purposes.} Confusability of states will affect how agents behave, how they perceive
-the behavior of others, and the expected utilities of behavioral dispositions.
-
-To keep matters simple, let us assume that agents carry pure dispositions to act. Noise can
-affect the realization of these strategies. As a sender, every agent maps states to messages:
-these are subjectively perceived states, and no longer necessarily also the actually occurring
-states. As a receiver, every agent maps messages to state interpretations: these are intended
-interpretations that need not always be faithfully realized. This means that behavioral
-strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
-behavioral dispositions in the population, the realization and observation of which can be
-distorted by agents' confusion of similar states.
-
-If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
-probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
-intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
-the probability with which state $\mystate{r}$ is realized.  Section~\ref{sec:exploring-rdd}
-gives a concrete (parameterized) specification of $P_o$ and $P_r$ in terms of similarity of
-states. For the time being, it is best to treat $P_o$ and $P_r$ as arbitrary conditional
-confusion probabilities.
-A single round of play of a sim-max game is then governed by five pieces of stochastic
-information, where previously there were only three.
-An illustration is given in Figure~\ref{fig:imprecise-sim-max-illustration}.
-Note that, and this is crucial, senders respond to observations not actual states, receiver strategies determine intentions not realized states, but payoff is calculated based on actual and realized states.
+here on one of the simplest: \emph{imitation of success}
+\citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
+Appendix~XYZ for a derivation of the standard replicator dynamic from this update scheme).
+Every now and then a random agent gets a chance to alter his behavior for one of his choice
+points. Say, this is a sender who gets to ``reconsider'' his choice of message for state
+$\state$. The revising agent then observes what some random agent does at $\state$, say
+$\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
+expected utility of $\messg$ for state $\state$.
+
+\subsection{Noisy replicator dynamic}
+\label{sec:noisy-repl-dynam}
+
+Imitation of success, as described above, presupposes that agents make no mistakes when
+observing states, or choosing interpretations. This may not always be an appropriate
+assumption, especially when some states can be fairly similar and therefore likely to be
+confused for one another. Therefore, imitation dynamics could be affected by at least two
+sources of probabilistic noise:
+
+\begin{enumerate}
+\item \emph{observation noise:} whenever a state $\mystate{a}$ actually occurs, the probability
+  that an agent observes it as $\mystate{o}$ is $P_o(\mystate{o} \probbar \mystate{a})$;
+\item \emph{realization noise:} whenever an agent intends to realize interpretation
+  $\mystate{i}$, the probability that $\mystate{r}$ is realized is
+  $P_r(\mystate{r} \probbar \mystate{o})$.
+\end{enumerate}
+
+\noindent A round of play of a sim-max game with these two sources of confusability of states
+is pictured in Figure~\ref{fig:imprecise-sim-max-illustration}. Note that, and this is crucial,
+senders respond to observations not actual states, receiver strategies determine intentions not
+realized states, but payoff is calculated based on actual and realized states. Behavioral
+strategies thus encode what agents actually do from their subjective point of view; or, put
+differently, what they would do in a noise-free world. The noise-free situation depicted in
+Figure~\ref{fig:sim-max-illustration} is then the special case where
+$P_o(\mystate{o} \probbar \mystate{a}) = 1$ iff $\mystate{a} = \mystate{o}$ and
+$P_r(\mystate{r} \probbar \mystate{i}) = 1$ iff $\mystate{i} = \mystate{r}$.
 
@@
 
-  \caption{A round of play in a sim-max game with imprecise imitation: Nature (N) chooses an actual state with probability $P_a(\mystate{a})$; a randomly sampled sender (S) observes $\mystate{o}$ with probability $P_o(\mystate{o} \probbar \mystate{a})$ and subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$.
-Payoff for both sender and receiver is given by $\Utils(\mystate{a}, \messg, \mystate{r})$.}
+  \caption{A round of play in a sim-max game with probabilistic confusability of states: Nature
+    (N) chooses an actual state with probability $P_a(\mystate{a})$; a randomly sampled sender
+    (S) observes $\mystate{o}$ with probability $P_o(\mystate{o} \probbar \mystate{a})$ and
+    subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a
+    randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with
+    probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation
+    $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$. Payoff for both
+    sender and receiver is given by $\Utils(\mystate{a}, \mystate{r})$.}
   %which is typically proportional to state similarity $\similarity(\mystate{a}, \mystate{r})$.}
@@
 
-Expected utilities of choices at choice points should likewise take into account that actual states
-need not be observed states, and intended interpretations need not be realized
-interpretations. First, note that
-\begin{align*}
-  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
-    \probbar \mystate{a})
-\end{align*}
-is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed by an agent. The
-probability that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
-\begin{align*}
-  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
-  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
-\end{align*}
-The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
-\begin{align*}
-  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
-  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
-\end{align*}
-The probability that $\mystate{r}$ is realized by a random receiver in response to
-message $\messg$ is:
-\begin{align*}
-  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
-  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
-\end{align*}
-This lets us capture the expected utilities for observed states (sender) and intended
-interpretations (receiver) by taking into consideration what the likely actual states and
-realized interpretations will be:
-\begin{align*}
-  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}}  P_{\overline{o}}(\mystate{a}
-  \probbar \mystate{o}) \sum_{\mystate{r}}  P_\Rstrat(\mystate{r} \probbar
-  \messg) \utils(\mystate{a}, \messg, \mystate{r})\,, \\
-  \EU(\mystate{i} , \messg, \Sstrat) & = \sum_{\mystate{a}}
-  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \sum_{\mystate{r}}
-  P_{r}(\mystate{r} \probbar \mystate{i})  \utils(\mystate{a}, \messg, \mystate{r})\,.
-\end{align*}
-If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
-confusability of non-identical states, above definitions reduce to the
-previous definitions of expected utilities. This also legitimates the overload of notation.
-
-Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
-the dynamic that ensues from imitation of successes. Since imprecision works slightly
-differently on senders and receivers (the former confuse choice points, the latter confuse
-choices), we need to look separately at each case.
-
-As before, suppose that senders receive a chance to change their behavior independently for a
-given choice point. In the present case, this would be a chance to change how to respond to a
-subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
-consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
-observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
-$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
-\begin{align*}
-  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
-  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
-\end{align*}
-The probability $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches
-from $\messg'$ to $\messg$ in subjectively perceived state $\mystate{o}$ is therefore:
-\begin{align*}
-  P(\messg' \rightarrow \messg, \mystate{o}) = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
-    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
-  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
-  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
-\end{align*}
-The mean change to the proportion of $\messg$ choices at state $\state$ are then:
-\begin{align}
-  \label{eq:nRDSender}
-  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
-  \rightarrow m', \mystate{o}) \nonumber \\
-  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
-  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
-  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
-  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
-\end{align}
-
-The case of the receiver is mostly analogous. Presented with an update opportunity for choice
-point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
-with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
-  that applies to the sender's observations also applies to the receiver's observations when
-  trying to imitate other agents' strategies. We could easily make the model more complex and
-  introduce another measure of noise for state confusability during receiver attempts to
-  imitate. We refrain from it here, because we see no immediate theoretical gain.}
-\begin{align*}
-  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
-  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
-  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
-  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
-\end{align*}
-The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
-from using $\state'$ to $\state$ in response to $\messg$ is therefore:
-\begin{align*}
-  P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
-    \messg)}_{\text{learner plays $\state'$}} \cdot
-  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
-  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
-\end{align*}
-Parallel to those for the sender case, this gives rise to:
-\begin{align}
-  \label{eq:nRDReceiver}
-  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
-  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
-  \Rstrat(\state \probbar \messg) \myts
-  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
-  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
-  \messg) \myts \sum_{\state'}
-  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
-\end{align}
-
-The continuous-time formulations in Equations~(\ref{eq:nRDSender}) and (\ref{eq:nRDReceiver})
-have elegant and practical discrete-time solutions in:\footnote{To see how the discrete-time
-  formulations gives rise to the continuous-time formulations, let's assume that update steps are
-  infinitesimally small, so that, for the sender case:
-  \begin{align*}
-    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
-    \probbar \state)\\ 
-    &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat)
-      - \Sstrat(\messg \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts
-      \EU(\messg', \state, \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts
-      \EU(\messg', \state, \Rstrat)}\,.
-  \end{align*}
-By dropping the denominator, which is constant for all $\messg$ for fixed
-$\state$, we obtain the above continuous-time formulation.}
+The presence of observation and realization noise also affects imitation of successes. Here, we
+focus on the main ideas. Formal detail is provided by Appendix~XYZ. Take a sender who gets to
+revise behavior for state $\state$. What agents can plausibly revise by imitation is their pure
+strategy, which maps perceived states onto messages. We should therefore assume that when an
+agent gets to revise his strategy for the perceived state $\state$, this need not necessarily
+be the actual state. Moreover, when that agent observes $\state$, another possibly
+to-be-imitated agent may yet perceive a completely different state altogether. Given what that
+latter agent perceives, his actual (pure) strategy will determine what he plays. In sum, to
+describe how likely the potential imitator observes a message choice $\messg$, we are
+interested in the conditional probability $P_o(\messg \probbar \state)$ that some other random
+agent realizes $\messg$ when the first agent perceived $\state$. This
+$P_o(\messg \probbar \state)$ is derived from the prior probability of states, the current
+sender population behavior $\Sstrat$ and the given observation noise (see
+Appendix~XYZ). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
+probability given by the expected utility of sending $\messg$ when perceiving state
+$\state$. Expected utility should, of course, take the probabilistic confusability of states
+into account as well.
+
+Similar considerations apply to the receiver side. If an agent who gets a chance to change his
+intended interpretation of $\messg$, we need to look at the conditional probability
+$P_o(\state \probbar \messg)$ of observing another agent realize interpretation $\state$ given
+that the first agent (and therefore the second as well) perceived $\messg$. This depends on
+observation and realization noise, as well as on the current receiver population behavior
+$\Rstrat$.
+
+Appendix~XYZ shows how imprecise imitation of this sort leads to mean changes in population
+frequencies of choices that can be covered by the following discrete-time formulation:
 \begin{align}
@@
 \end{align}
-This looks very much like the discrete-time formulation of the standard replicator dynamic in
-(\ref{eq:RDDiscrete}). A minor difference is that expected utility here takes stochastic
-confusability of states into account. The main difference is that here we are not looking at,
-say, the probability $\Sstrat(\messg \probbar \state)$ with which $\messg$ is played in
-response to $\state$ on average in the population, but at the probability $P_o(\messg \probbar
-\state)$ of seeing one agent pick $\messg$ given that another agent has observed
-$\state$. Similarly for the receiver, we are not looking at occurrence probability
-$\Rstrat(\state \probbar \messg)$ but at the probability $P_o(\state \probbar \messg)$ with
-which an agent who observes $\messg$ observes another agent play $\state$. It is easy to check
-formally and intuitive to appreciate that this major difference vanishes if confusion
-probabilities $P_o$ and $P_r$ are trivial, i.e., assigning zero probability to all pairs of
-non-identical states. The noisy replicator dynamic in (\ref{eq:nRD}) conservatively extends the
-classic case in (\ref{eq:RDDiscrete}).
+\noindent This looks very much like the discrete-time formulation of the standard replicator
+dynamic in (\ref{eq:RDDiscrete}), but there are, of course, the aforementioned
+differences. Firstly, expected utility here takes stochastic confusability of states into
+account. Secondly, where the standard replicator dynamic had probabilities
+$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
+$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. If there is no
+observation or realization noise, $P_o(\messg \probbar \state)$ reduces to
+$\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
+$\Rstrat(\state \probbar \messg)$. The noisy replicator dynamic in (\ref{eq:nRD})
+conservatively extends the classic case in (\ref{eq:RDDiscrete}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 $\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
-observational confusability $P_o$ is simply the same as realizational confusability $P_r$ and
+observation noise $P_o$ is simply the same as realizational confusability $P_r$ and
 that both are governed by the same imprecision parameter $\alpha$:
@@
 \end{align*}
-For $\imprecision = 0$, we obtain trivial confusion probabilities for which everything is reduced
-to the replicator dynamic. For $\imprecision > 0$, any state can be confused for any other
-state with some positive probability. For the uninteresting case of
+For $\imprecision = 0$, we obtain trivial confusion probabilities: everything is reduced to
+perfect imitation and the replicator dynamic. For $\imprecision > 0$, any state can be confused
+for any other state with some positive probability. For the uninteresting case of
 $\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
@@
 $\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
-$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps of the
-\rdd. A trial was considered converged, and thus stopped before the maximum of 200 rounds, if
-the total amount of change between strategies before and after an update step was smaller than
-a suitably chosen threshold. It is not guaranteed that strategies at halting time had converged
-to the eventual attracting state, whether they ran for 200 rounds or not. Our notion of
-convergence is therefore only a categorical measure for reaching a certain (well-considered,
-but eventually arbitrary) degree of stability. In other words, out notion of ``convergence'' is
-a measure of relative speed: it is true that the system reached a state in which evolutionary
-adaptations had slowed down almost to a halt before 200 update steps? This is motivated by
-pragmatic concerns regarding length of simulation time, but also theoretically justified,
-because we hypothesize that confusability of states leads to regularization of evolving
-strategies, which would show exactly in an increased speed of evolutionary trajectories towards
-well-behaved and regular signaling behavior.
+$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps. A
+trial was considered converged, and thus stopped before the maximum of 200 rounds, if the total
+amount of change between strategies before and after an update step was smaller than a suitably
+chosen threshold. It is not guaranteed that strategies at halting time had converged to the
+eventual attracting state, whether they ran for 200 rounds or not. Our notion of convergence is
+therefore only a categorical measure for reaching a certain (well-considered, but eventually
+arbitrary) degree of stability. In other words, out notion of ``convergence'' is a measure of
+relative speed: it is true that the system reached a state in which evolutionary adaptations
+had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
+concerns regarding length of simulation time, but also theoretically justified, because we
+hypothesize that confusability of states leads to regularization of evolving strategies, which
+would show exactly in an increased speed of evolutionary trajectories towards well-behaved and
+regular signaling behavior.
 
@@
 
+Our imprecise imitation dynamic leads to by-and-large successful signaling behavior, even in
+the presence of noise, that show the hallmarks of vagueness as desired. It also gives rise to
+population-level behavior that looks as if agents are generalizing across similar
+stimuli. Here, we would like to reflect briefly on some further conceptually relevant points
+and compare our approach to related work.
+
 \subsection{Levels of vagueness}
-The main motivation and intuitions given in Section~\ref{sec:repl-diff-dynam} describe probabilities in sender and receiver strategies as capturing the overall distribution of pure strategies in a population: $\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the receiver).
-This is in line with the standard interpretation of the replicator equation, but we should consider its philosophical implications.
-Based on this interpretation and our model, vagueness is seemingly being characterized as a strictly population-level phenomenon since it arises in a signaling system from the inability of individual agents to fully align their (non-vague) strategies because of imprecision.
-This is, we believe, a plausible mechanism that can already explain the existence of vagueness in a language even if we assume each agent commands a non-vague idiolect.
-
-We do not, however, want to commit to the idea that vagueness does not exist at the level of individual agents.
-The fact is that the model can be interpreted in alternative ways.
-The probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will be used by a sender when observing state $\state$.
-Whether the uncertainty represents variations in a population of agents playing pure strategies, a single sender using a non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is up for interpretation.
-The first option supported what we believe to be a plausible motivation for the mechanism of imitation of success with imprecision, but other motivations could be imagined to better fit an interpretation that includes non-deterministic agents.
-We thus believe that the model introduced in this paper does not exclude vagueness at the individual level, but to explore all possible interpretations would go beyond our current scope.
+The imprecise imitation dynamic was introduced in Section~\ref{sec:repl-diff-dynam} as tracing
+changes in the overall distribution of pure strategies in a population:
+$\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled
+sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the
+receiver). This is in line with the standard interpretation of the replicator equation, but we
+should consider its philosophical implications. Based on this picture, vagueness in signal use
+would seem to be characterized as a strictly population-level phenomenon since it arises in a
+signaling system from the inability of individual agents to fully align their (non-vague)
+strategies because of imprecision. This is, we believe, a plausible mechanism that can already
+explain the existence of vagueness in a language even if we assume each agent commands a
+non-vague idiolect.
+
+We would not, however, want to commit to the idea that vagueness does not exist at the level of
+individual agents. True, our derivation of the imprecise imitation dynamic assumed that agents
+carry and revise pure strategies. But that was an assumption of convenience, not of
+conviction. Moreover, even if individual agents command a non-vague pure strategy, the
+realization of that pure strategy, according to our model, is bound to be vague (e.g., the same
+agent could signal differently in repeated exposure to the same state). We have used the term
+``observation noise'' here, but it may well be that this could equally well be perceived as an
+inseparable component of an agent's ``signaling faculty.'' In this sense, then, the model might
+not be entirely incompatible with a picture of agents who have internalized a vague signaling
+strategy. It would need to be seen, however, how revision of non-deterministic individual-level
+behavior must be spelled out rigorously and whether the resulting population-level dynamic
+would be equivalent to our present proposal in all respects.
+
+% Moreover, we believe that the model may be interpreted in yet alternative ways. The
+% probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will
+% be used by a sender when observing state $\state$.  Whether the uncertainty represents
+% variations in a population of agents playing pure strategies, a single sender using a
+% non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is
+% up for interpretation.  The first option supported what we believe to be a plausible motivation
+% for the mechanism of imitation of success with imprecision, but other motivations could be
+% imagined to better fit an interpretation that includes non-deterministic agents.  We thus
+% believe that the model introduced in this paper does not exclude vagueness at the individual
+% level, but to explore all possible interpretations would go beyond our current scope.
 
@@
 
-The phenomenon could potentially not just be a side effect.
-Given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
-This could be achieved by letting the imprecision parameter be subject to the evolutionary dynamics.
-The idea is for now mostly speculative, but we consider it an interesting avenue for future research.
-There is additional motivation to consider its potential if we see it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.
+The phenomenon could potentially not just be a side effect.  Given the benefit of a certain
+amount of imprecision we observe when comparing within group expected utility, it would be
+interesting to study whether, under certain conditions, this group level advantage could trump
+the individual level disadvantage of a vague language, and thus actually select for a certain
+amount of imprecision.  This could be achieved by letting the imprecision parameter be an
+evolving part in the dynamics as well. The idea is for now mostly speculative, but we consider
+it an interesting avenue for future research.  There is additional motivation to consider its
+potential if we see it in the eyes of multilevel selection theory
+\citep{Wilson1994,OGorman2008}.
 
@@
 response dynamics predict that senders would be almost indifferent. Sender behavior that
-evolves under confusion of states does not have this puzzling property.
+evolves under confusion of states does not have this puzzling property, because a giant would
+not likely be confused for a dwarf.
 
@@
 communication? On the assumption that agents are likely to mistake states for one another,
-especially when they are perceptually similar, derived a generalization of the replicator
-dynamic in terms of imitation of successes. The resulting population-level dynamic produced
-signaling behavior that is at the same time regular and by-and-large communicatively efficient,
-while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
-by-product of an arguably natural limitation on the discriminatory power of signaling agents.
-Although inability to sharply discriminate similar stimuli may lead to vagueness and bring about
-a (slight) decrease in communicative efficiency, there may also be an advantage, not of
-vagueness itself, but of its reason: systematicity in the confusability of states (which may be
-a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
-population-level without having to assume that agents themselves have any generalization power.
+especially when they are perceptually similar, we derived a generalization of the replicator
+dynamic from an agent-level process of imprecise imitation. The resulting population-level
+dynamic produced signaling behavior that is at the same time regular and by-and-large
+communicatively efficient, while also showing the crucial marks of vagueness. In a sense, the
+model derives vagueness as a by-product of an arguably natural limitation on the discriminatory
+power of signaling agents.  Although inability to sharply discriminate similar stimuli may lead
+to vagueness and bring about a (slight) decrease in communicative efficiency, there may also be
+an advantage, not of vagueness itself, but of its reason: systematicity in the confusability of
+states (which may be a natural by-product of the perceptual system) supports
+\emph{as-if}-generalization at the population-level without having to assume that agents
+themselves have any generalization power.
 
@@
 
+\appendix
+
+\section{Imprecise conditional imitation}
+\label{sec:formal-stuff}
+
+The goal of this section is to provide technical details for
+Section~\ref{sec:repl-diff-dynam}. We first show, in Section~\ref{sec:cond-imit-succ}, how to
+derive the standard replicator dynamic from noise-free imitation of success. The, in
+Section~\ref{sec:imit-succ-with}, we derive the imprecise imitation dynamic by a parallel
+chain of arguments.
+
+\subsection{Deriving the replicator dynamic from imitation of success}
+\label{sec:cond-imit-succ}
+
+The main idea behind \emph{imitation of success} is that agents imitate the behavior of other
+agents at some choice point with a probability that is proportional to the expected utility of
+the latter agents' choice. Since the sim-max games that we are looking at here have positive
+utilities upper bound by 1, we can identify the switching probability with the expected
+utility.
+
+Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
+agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
+``teacher.'' A random learner is drawn from the population and given a chance to change
+behavior at choice point $\state$. The probability that our learner plays $\messg$ is
+$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
+$\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar \state)$. The
+learner then starts using $\messg'$ instead of $\messg$ with probability
+$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same; the
+learner could even be the teacher as well, by random sampling.)
+
+If agents get repeated update chances for their choice points, the \emph{expected change} of
+frequency of $\messg$-choices at $\state$ becomes:
+\begin{align}
+  \label{eq:MeanChange}
+  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
+\end{align}
+where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
+any $\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
+that agents switch from $\messg$ to any $\messg'$. Since we are dealing with expectations in a
+huge population, these can be spelled out as:
+\begin{align*}
+  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
+    \state)}_{\text{learner plays $\messg'$}} \cdot
+  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
+  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
+  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
+    \state)}_{\text{learner plays $\messg$}} \cdot
+  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
+  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
+\end{align*}
+From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
+to:
+\begin{align*}
+  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
+  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
+      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
+& = \underbrace{\Sstrat(\messg \probbar
+    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
+  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
+      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
+      EU at choice point $\state$}} \Big )\,.
+\end{align*}
+This latter formulation is the continuous-time formulation of the replicator
+dynamic. We obtain the discrete-time formulation from it by assuming that discrete update steps
+are infinitesimally small, so that:
+ \begin{align*}
+    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
+    \probbar \state)\\ 
+    & = \frac{\Sstrat(\messg \probbar
+    \state)  \myts \EU(\messg,\state,\Rstrat)}{\sum_{\messg'}\Sstrat(\messg' \probbar
+    \state)  \myts \EU(\messg,\state,\Rstrat)} - \sigma(\messg
+      \probbar \state) \\
+    & = \frac{\Sstrat(\messg \probbar
+    \state)  \myts \EU(\messg,\state,\Rstrat) - \Sstrat(\messg \probbar \state) \sum_{\messg'}\Sstrat(\messg' \probbar
+    \state)  \myts \EU(\messg,\state,\Rstrat) }{\sum_{\messg'}\Sstrat(\messg' \probbar
+    \state)  \myts \EU(\messg,\state,\Rstrat)}
+      \,.
+  \end{align*}
+By dropping the denominator, which is constant for all $\messg$ for fixed
+$\state$, we obtain the above continuous-time formulation.
+
+\subsection{Imitation of success with imprecision}
+\label{sec:imit-succ-with}
+
+The above derivation of the replicator dynamic assumes that agents can discriminate choices and
+choice points perfectly. Let's dispense with that assumption. With an eye toward sim-max games,
+we will assume that states, but not messages, may be confused for one another.\footnote{It is
+  relatively straightforward to also incorporate confusability of messages, but this is
+  irrelevant to our present purposes.} Confusability of states will affect how agents behave,
+how they perceive the behavior of others, and the expected utilities of behavioral
+dispositions.
+
+To keep matters simple, let us assume that agents carry pure dispositions to act. Noise can
+affect the realization of these strategies. As a sender, every agent maps states to messages:
+these are subjectively perceived states, and no longer necessarily also the actually occurring
+states. As a receiver, every agent maps messages to state interpretations: these are intended
+interpretations that need not always be faithfully realized. This means that behavioral
+strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
+behavioral dispositions in the population, the realization and observation of which can be
+distorted by agents' confusion of similar states.
+
+If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
+probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
+intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
+the probability with which state $\mystate{r}$ is realized. A single round of play of a sim-max
+game is then governed by five pieces of stochastic information, where previously there were
+only three (see Figures~\ref{fig:sim-max-illustration} and
+\ref{fig:imprecise-sim-max-illustration}).
+
+Expected utilities of choices at choice points should likewise take into account that actual states
+need not be observed states, and intended interpretations need not be realized
+interpretations. First, note that
+\begin{align*}
+  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
+    \probbar \mystate{a})
+\end{align*}
+is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed by an agent. The
+probability that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
+\begin{align*}
+  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
+  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
+\end{align*}
+The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
+\begin{align*}
+  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
+  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
+\end{align*}
+The probability that $\mystate{r}$ is realized by a random receiver in response to
+message $\messg$ is:
+\begin{align*}
+  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
+  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
+\end{align*}
+This lets us capture the expected utilities for observed states (sender) and intended
+interpretations (receiver) by taking into consideration what the likely actual states and
+realized interpretations will be:
+\begin{align*}
+  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
+  \probbar \mystate{o}) \sum_{\mystate{r}} P_\Rstrat(\mystate{r} \probbar
+  \messg) \utils(\mystate{a}, \mystate{r})\,, \\
+  \EU(\mystate{i}, \messg, \Sstrat) & = \sum_{\mystate{a}} P_{\overline{\Sstrat}}(\mystate{a}
+  \probbar \messg) \sum_{\mystate{r}} P_{r}(\mystate{r} \probbar \mystate{i})
+  \utils(\mystate{a}, \mystate{r})\,.
+\end{align*}
+If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
+confusability of non-identical states, above definitions reduce to the
+previous definitions of expected utilities. This also legitimates the overload of notation.
+
+Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
+the dynamic that ensues from imitation of successes. Since imprecision works slightly
+differently on senders and receivers (the former confuse choice points, the latter confuse
+choices), we need to look separately at each case.
+
+As before, suppose that senders receive a chance to change their behavior independently for a
+given choice point. In the present case, this would be a chance to change how to respond to a
+subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
+consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
+observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
+$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
+\begin{align*}
+  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
+  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
+\end{align*}
+The ``inflow'' and ``outflow'' probabilities $P(\messg' \rightarrow \messg, \state)$ and
+$P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches from any
+$\messg'$ to $\messg$ or from $\messg$ to any $\messg'$ in subjectively perceived state
+$\mystate{o}$ are therefore:
+\begin{align*}
+  P(\messg' \rightarrow \messg, \mystate{o}) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
+    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
+  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
+  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
+  \\
+  P(\messg \rightarrow \messg', \mystate{o}) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
+    \mystate{o})}_{\text{learner plays $\messg$ at $\mystate{o}$}} \cdot
+  \underbrace{P_{o}(\messg' \probbar \mystate{o})}_{\text{observe teacher play $\messg'$}} \cdot
+  \underbrace{\EU(\messg', \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
+\end{align*}
+The mean change to the proportion of $\messg$ choices at state $\state$ are then:
+\begin{align}
+  \label{eq:nRDSender}
+  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
+  \rightarrow m', \mystate{o}) \nonumber \\
+  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
+  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
+  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
+  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
+\end{align}
+
+The case of the receiver is mostly analogous. Presented with an update opportunity for choice
+point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
+with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
+  that applies to the sender's observations also applies to the receiver's observations when
+  trying to imitate other agents' strategies. We could easily make the model more complex and
+  introduce another measure of noise for state confusability during receiver attempts to
+  imitate. We refrain from it here, because we see no immediate theoretical gain.}
+\begin{align*}
+  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
+  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
+  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
+  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
+\end{align*}
+% The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
+% from using $\state'$ to $\state$ in response to $\messg$ is therefore:
+% \begin{align*}
+%   P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
+%     \messg)}_{\text{learner plays $\state'$}} \cdot
+%   \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
+%   \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
+% \end{align*}
+Parallel to the sender case, this gives rise to:
+\begin{align}
+  \label{eq:nRDReceiver}
+  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
+  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
+  \Rstrat(\state \probbar \messg) \myts
+  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
+  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
+  \messg) \myts \sum_{\state'}
+  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
+\end{align}
+
+The continuous-time formulations in Equations~(\ref{eq:nRDSender}) and (\ref{eq:nRDReceiver})
+have elegant and practical discrete-time solutions in:
+\begin{align*}
+  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
+  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
+  \myts \EU(\state, \messg, \Sstrat)\,,
+\end{align*}
+which is the discrete-time formulation of the imprecise imitation dynamic given in
+Section~\ref{sec:repl-diff-dynam} in Equation~(\ref{eq:nRD}). To see how the discrete-time
+formulation gives rise to the continuous-time formulations above, let's assume that update
+steps are infinitesimally small, so that, for the sender case:
+\begin{align*}
+  \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
+  \probbar \state)\\
+  &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \Sstrat(\messg
+    \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg', \state,
+    \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg', \state,
+    \Rstrat)}\,.
+\end{align*}
+As before, we drop the denominator, which is constant for all $\messg$ for fixed $\state$, and
+obtain the above continuous-time formulation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \printbibliography[heading=bibintoc]

### 2015-09-29 paper/version_04/paper.tex
Higher similarity, higher utility, right?
Higher similarity, higher utility, right? (cherry picked from commit e273325)
Small fixes. Changed 'teacher' to 'model', since it doesn't have the connotation that the agent is actively passing on knowledge.

@@
 of everyday communication, but it is troublesome for some of the most
-prominent theories of language and meaning. This is especially so for the logico-semantic tradition of Frege, Russell and the young
+prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell and the young
 Wittgenstein which is challenged by the paradoxes vagueness gives rise
@@
 identify the receiver's acts with the states of the world, so that the game is one of guessing
-the actual state, so to speak. The utility function should be a monotonically decreasing
+the actual state, so to speak. The utility function should be a monotonically increasing
 function of similarity.
-A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the \emph{probability} that a given message would be used for each state.
+A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the probability that a given message would be used for each state.
 The analogous happens for the receiver's behavior.
@@
 corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
-other. In this way, \citet{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
+other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
 system can be obtained: storing the prototypes lets us recover the categories without having to
-store each category's extension. However, what is left unexplained, is where the prototypes
+store each category's extension. However, what is left unexplained is where the prototypes
 come from, and why we would not see just any distribution of prototypes as an equally efficient
@@
 
-An alternative approach is taken by \citet{FrankeJager2010:Vagueness-Signa} and
-\citet{OConnor2013:The-Evolution-o} who show how the above desiderata can be met by evolving
+Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
+\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be met by evolving
 strategies in sim-max games. The approach we take here is similar in spirit, but different in
@@
 abstractly, at the level of a population of agents, we can describe how average behavioral
-dispositions will evolve. The population-level perspective abstracts over petty stochastic
+dispositions will evolve. The population-level perspective abstracts over small stochastic
 fluctuations and zooms in on the general tendency or direction of evolution that ensues from
@@
   $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
-  side must possibly yet be normalized so as to have probabilities sum to one. In general,
+  side might still need to be normalized so as to have probabilities sum to one. In general,
   writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
@@
 outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
-sender of choosing message $\messg$, given tate $\state$ and a receiver
+sender of choosing message $\messg$, given state $\state$ and a receiver
 strategy $\Rstrat$ is the sum, for each act $\act$, of the probability that the
@@
 of other agents at some choice point with a probability that is proportional to the expected
-utility of the latter agents' choice. Since the sim-max games that we will be looking at in
+utility of the latter agents' choice. Since the sim-max games we will be looking at in
 Section~\ref{sec:exploring-rdd} have positive utilities, we can identify the switching
@@
 
-Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
+Let's consider sender strategies, as the receiver case is parallel. Call the
 agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
-``teacher.'' A random learner is drawn from the population and given a chance to change
+``model.'' A random learner is drawn from the population and given a chance to change
 behavior at choice point $\state$. The probability that our learner plays $\messg$ is
-$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
+$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled model does at
 $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar \state)$. The
@@
     \state)}_{\text{learner plays $\messg'$}} \cdot
-  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
-  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} 
+  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{model plays $\messg$}} \cdot
+  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU model choice}}
   % P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
   %   \state)}_{\text{learner plays $\messg$}} \cdot
-  % \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
-  % \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
+  % \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{model plays $\messg'$}} \cdot
+  % \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU model choice}}
 \end{align*}
@@
 Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
-the dynamic that ensues from imitation of successes. Since imprecision works slightly
+the dynamic that ensues from imitation of success. Since imprecision works slightly
 differently on senders and receivers (the former confuse choice points, the latter confuse
@@
 consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
-observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
+observed $\mystate{o}$, he will simultaneously observe a randomly sampled model play
 $\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
@@
     \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
-  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
-  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
+  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe model play $\messg$}} \cdot
+  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU model choice in learner's view}}
 \end{align*}
@@
 The case of the receiver is mostly analogous. Presented with an update opportunity for choice
-point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
+point $\messg$, a learner will observe a random model choose interpretation $\mystate{o}$
 with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
@@
     \messg)}_{\text{learner plays $\state'$}} \cdot
-  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
-  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
+  \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe model play $\state$}} \cdot
+  \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU model choice}}
 \end{align*}
@@
 Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
-interval, including $0$ and $1$. All states occur, for simplicity, with the same probability.
+interval, including $0$ and $1$. All states occur, for simplicity, with the same probability (\emph{i.e.}~$P_a$ is uniform).
 The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
@@
     0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
     \end{cases}
@@
     0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \toler^2} \right ) & \textrm{otherwise.} \\
     \end{cases}
@@
 convergence is therefore only a categorical measure for reaching a certain (well-considered,
-but eventually arbitrary) degree of stability. In other words, out notion of ``convergence'' is
-a measure of relative speed: it is true that the system reached a state in which evolutionary
-adaptations had slowed down almost to a halt before 200 update steps? This is motivated by
+but eventually arbitrary) degree of stability. In other words, it is
+a measure of relative speed capturing whether the system reached a state in which evolutionary
+adaptations had slowed down almost to a halt before 200 update steps. This is motivated by
 pragmatic concerns regarding length of simulation time, but also theoretically justified,
@@
 
-  \caption{Example strategies under \rdd at stopping time of our simulations.}
+  \caption{Example strategies under \rdd at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
   \label{fig:example_strats}
@@
 The unifying property of perceptual imprecision could be considered an evolutionary beneficial
-side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
+side-effect. A certain degree of imprecision can lead to higher \emph{within group expected utility},
 defined as the average expected utility that each evolved language scored when playing against
@@
 especially when they are perceptually similar, derived a generalization of the replicator
-dynamic in terms of imitation of successes. The resulting population-level dynamic produced
+dynamic in terms of imitation of success. The resulting population-level dynamic produced
 signaling behavior that is at the same time regular and by-and-large communicatively efficient,

### 2015-09-29 paper/version_04/paper.tex
Small fixes (cherry picked from commit 7d7fc88)

--- 2015-09-29 paper/version_04/paper.tex
@@
 Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
-approach to others. Appendix~\ref{sec:repl-diff-dynam} provides formal detail.
+approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
 
@@
 \citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
-Appendix~XYZ for a derivation of the standard replicator dynamic from this update scheme).
+Appendix~\ref{sec:formal-stuff} for a derivation of the standard replicator dynamic from this update scheme).
 Every now and then a random agent gets a chance to alter his behavior for one of his choice
@@
 The presence of observation and realization noise also affects imitation of successes. Here, we
-focus on the main ideas. Formal detail is provided by Appendix~XYZ. Take a sender who gets to
+focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:formal-stuff}. Take a sender who gets to
 revise behavior for state $\state$. What agents can plausibly revise by imitation is their pure
@@
 sender population behavior $\Sstrat$ and the given observation noise (see
-Appendix~XYZ). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
+Appendix~\ref{sec:formal-stuff}). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
 probability given by the expected utility of sending $\messg$ when perceiving state
@@
 
-Appendix~XYZ shows how imprecise imitation of this sort leads to mean changes in population
+Appendix~\ref{sec:formal-stuff} shows how imprecise imitation of this sort leads to mean changes in population
 frequencies of choices that can be covered by the following discrete-time formulation:

### 2015-09-29 paper/version_04/paper.tex
shortened section 3 to move technicalities into appendix; first stages

--- 2015-09-29 paper/version_04/paper.tex
@@
 \citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
+  \node (3) [right=of 2] {$\messg = \Spure(\mystate{1}) \in \Messgs$};
+  \node (4) [right=of 3] {$\mystate{2} = \Rpure(\messg) \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {S} (3)
+    (3) edge node {R} (4);
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg = \Spure(\mystate{1})$ determined by his pure strategy; a receiver (R) selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy.
+Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
+   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
+  \label{fig:sim-max-illustration}
+\end{figure}
 
@@
 function of similarity.
-A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the \emph{probability} that a given message would be used for each state.
-The analogous happens for the receiver's behavior.
-A play in such a game with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
-  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
-  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {S} (3)
-    (3) edge node {R} (4);
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a receiver (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
-Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
-   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
-  \label{fig:sim-max-illustration}
-\end{figure}
+In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
+A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Imprecise conditional imitation}
+\label{sec:repl-diff-dynam}
+
+Signaling agents can adapt their dispositions to act, given some feedback about their past
+success, in multiple ways. Usually, we would assume that changes in behavior should, at least
+on average, lean towards increasing chances of communicative success. Such behavioral
+adaptations can be described at different levels of abstraction. At the level of individual
+agents, we can picture a more or less idealized process of how each agent adapts dispositions
+for future actions based on various pieces of information available to the agent. More
+abstractly, at the level of a population of agents, we can describe how average behavioral
+dispositions will evolve. The population-level perspective abstracts over petty stochastic
+fluctuations and zooms in on the general tendency or direction of evolution that ensues from
+behavior at the agent-level.
+
+To better understand the interaction of confusability of stimuli and selective pressure towards
+successful communication, we look at a population-level dynamic that describes the most likely
+evolutionary path of a population of signaling agents who imitate other agents' behavior but
+are liable to confuse states for one another. In the special limiting case where state
+confusability vanishes, the process is just the well-known replicator dynamic
+\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.
+
+\subsection{Replicator dynamic in behavioral strategies}
+\label{sec:repl-dynam-behav}
+
+Fix a signaling game with finite states $\States$ and messages $\Messgs$. We consider an
+interpretation game in which the receiver chooses states in $\States$ in response to
+messages. Let $\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
+$\Util \mycolon \States \times \Messgs \times \States \rightarrow \mathds{R}$ the utility
+function shared by senders and receivers in the population. A \emph{behavioral strategy} is a
+function that maps an agent's choice points to a probability distribution over available
+choices.\footnote{Our focus is on behavioral strategies, not \emph{mixed strategies}, i.e.,
+  probability distributions over functions from each choice point to an act, such as
+  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
+  adapt their behavior locally, i.e., independently at each choice point. Our
+  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
+  numerical simulations. But it also seems the more plausible choice for imitation-based update
+  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
+  particular situation, not how that agent would behave in \emph{all} relevant choice
+  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
+  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
+  behavioral strategies.)} The sender's behavioral strategies are functions
+$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
+probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's are
+functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message $\messg \in \Messgs$
+to a probability of each interpretation $\state \in \States$ being chosen in response to
+$\messg$. Although behavioral strategies are probabilistic, evolutionary modeling usually
+imagines that every individual agent has a non-probabilistic strategy. Behavioral strategies
+then capture average population behavior. In a virtually infinite population, the number
+$\Sstrat(\messg \probbar \state)$, for instance, is then the probability that a randomly
+sampled sender would send message $\messg$ if the actual state was $\state$. Similarly,
+$\Rstrat(\state \probbar \messg)$ is then the probability with which a randomly sampled
+receiver interprets $\messg$ as $\state$.
+
+The \emph{expected utility} of choices at each choice point is:\footnote{The notation
+  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
+  side must possibly yet be normalized so as to have probabilities sum to one. In general,
+  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
+  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
+\begin{align*}
+  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
+  \Rstrat(\state' \probbar \messg) \myts U(\state, \messg, \state') \\
+  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
+    \States} P(\state \probbar \messg)  \myts
+  U(\state, \messg, \state')\,, \text{\ \  where } &
+P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
+\end{align*}
+Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
+outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
+sender of choosing message $\messg$, given tate $\state$ and a receiver strategy $\Rstrat$ is
+the sum, for each act $\act$, of the probability that some random receiver will choose
+$\act$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and $\act$.
+
+The discrete-time replicator dynamic tracks changes in frequency of choices in the population
+as proportional to their expected utilities
+\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
+  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
+  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
+  utilities are always positive.}
+\begin{align}
+  \label{eq:RDDiscrete}
+  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \cdot
+    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \cdot
+    \EU(\act, \messg,\Sstrat)\,.   
+\end{align}
+For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
+average agent in the population \emph{after} the update is proportional to the probability of
+seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
+expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
+by a gradient of current frequency and a measure of how good they are.
+
+\subsection{A noisy replicator dynamic}
+\label{sec:noisy-repl-dynam}
+
+The replicator dynamic is an abstract population-level dynamic that describes the mean expected
+change of behavioral dispositions in a population of signalers. There are several ways of
+deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
+here on one of the simplest: \emph{probabilistic imitation of success}
+\citep[see][]{Sandholm2010:Population-Game}: agents imitate the behavior of some other agent,
+which they happen to observe, with a probability that is proportional to the expected utility
+of the latter agents' choice (see Appendix~XYZ for a derivation of the standard replicator
+dynamic from this update scheme).
+
+Probabilistic imitation of success presupposes that agents make no mistakes when observing
+states. This may not always be an appropriate assumption. Especially for sim-max games, where,
+intuitively, some states can be fairly similar and therefore likely to be confused for one
+another. Appendix~XYZ shows how to derive a generalization of the replicator dynamic from the
+assumption that probabilistic imitation of success is affected by two sources of probabilistic
+confusion of states:
+
+\begin{enumerate}
+\item \emph{observation noise:} whenever a state $\mystate{a}$ actually occurs, the
+  probability that is observed as $\mystate{o}$ is $P_o(\mystate{o} \probbar \mystate{a})$;
+\item \emph{realization noise:} whenever a receiver intends to realize interpretation
+  $\mystate{i}$, the probability that $\mystate{r}$ is realized is $P_r(\mystate{r} \probbar \mystate{o})$.
+\end{enumerate}
+
+The mean changes in the population frequencies of choices under noise-affected imitation of
+success can be covered by the following discrete-time formulation:
+
+\begin{align}
+  \label{eq:nRD}
+  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
+  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
+  \myts \EU(\state, \messg, \Sstrat)\,.
+\end{align}
+
+This looks very much like the discrete-time formulation of the standard replicator dynamic in
+(\ref{eq:RDDiscrete}), but there are differences (the Appendix~XYZ for full definitions and
+formal detail). Firstly, expected utility here takes stochastic confusability of states into
+account in a natural way. Secondly, where the standard replicator dynamic had probabilities
+$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
+$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. This is the major
+difference, so we should dwell on it for a brief moment.
+
+Look at the sender case first, where $\Sstrat(\messg \probbar \state)$ is the probability with
+which $\messg$ is played in response to actual state $\state$ on average in the population. If
+agents may not perceive the actual state, this should also affect what they observe when trying
+to imitate another agents behavior. $P_o(\messg \probbar \state)$ is therefore the probability,
+given that a potential imitator has observed $\state$, that he would observe a random agent
+play $\messg$, which he may or may not choose to imitate. Similarly for the receiver, we are
+not looking at occurrence probability $\Rstrat(\state \probbar \messg)$ but at the probability
+$P_o(\state \probbar \messg)$ with which an agent who observes $\messg$ observes another agent
+choose interpretation $\state$.
+
+Obviously, $P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ are to be derived
+from, among others, probabilistic confusion of states $P_o(\mystate{o} \probbar \mystate{a})$
+and $P_r(\mystate{r} \probbar \mystate{o})$. Details are given in Appendix~XYZ. If there is no
+noise, $P_o(\messg \probbar \state)$ reduces to $\Sstrat(\messg \probbar \state)$ and
+$P_o(\state \probbar \messg)$ reduces to $\Rstrat(\state \probbar \messg)$. But if there is,
+imitation probabilities must take confusability of states into account. In this sense, the
+noisy replicator dynamic in (\ref{eq:nRD}) conservatively extends the classic case in
+(\ref{eq:RDDiscrete}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Exploring noisy imitation}
+\label{sec:exploring-rdd}
+
+How does a tendency to confuse similar states interact with selective pressure towards more
+efficient signaling strategies under the imprecise conditional imitation dynamic? If confusion
+probabilities are moderate and regular, in that they track similarity of states, we might
+expect a regularizing effect also on evolving signaling strategies. Indeed, we hypothesize that
+confusion of states can give rise to population-level aggregate behavior that looks \emph{as
+  if} information about what is good for one state percolates also to similar states. In other
+words, imprecise imitation may make signaling behavior on average look like agents generalize
+across similar states, even if no agent actually generalizes. To explore whether confusion of
+states can have this effect, we turn to numerical simulation.
+
+\subsection{Setting the stage}
+\label{sec:setting-stage}
+
+To obtain concrete results, we must fix how to represent states, similarity of states, the
+conditional confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max
+games. Confusion probabilities between states should be a function of perceptual similarity:
+the more similar two states are, the more likely they could be mistaken for each other.
+
+Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
+interval, including $0$ and $1$. All states occur, for simplicity, with the same probability.
+The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
+two states $\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual
+similarity function, as described by \citet{Nosofsky1986:Attention-Simil}:
+\begin{align*}
+  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
+      \begin{cases}
+    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
+    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
+    \end{cases}
+\end{align*}
+where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
+discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
+states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
+different parameter values. Other formalizations of perceptual similarity are possible,
+including ones that allow for different discriminability in different areas of the state space,
+but we stick with Nosofsky's similarity function for the time being, because it is
+mathematically simple, yet an established notion in mathematical psychology.
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=0.7\textwidth]{plots/NosofskySim.pdf}
+
+  \caption{Examples of Nosofsky-similarity for different values of
+    imprecision.}
+  \label{fig:NosofskySim}
+\end{figure}
+
+We further assume that the probability of confusing any two states $\mystate{i}$ and
+$\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
+observational confusability $P_o$ is simply the same as realizational confusability $P_r$ and
+that both are governed by the same imprecision parameter $\alpha$:
+\begin{align*}
+  P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
+  \imprecision) &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
+  \imprecision) \,.
+\end{align*}
+For $\imprecision = 0$, we obtain trivial confusion probabilities for which everything is reduced
+to the replicator dynamic. For $\imprecision > 0$, any state can be confused for any other
+state with some positive probability. For the uninteresting case of
+$\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
+realized as any other state with probability $\nicefrac{1}{\card{\States}}$.
+
+As for utility functions, we take another Nosofsky-style function:
+\begin{align*}
+  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
+      \begin{cases}
+    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
+    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
+    \end{cases}
+\end{align*}
+The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
+communicative imprecision. This choice of utility function is governed partly by convenience in
+parallel to the similarity function, but also because it has, as we believe, the right general
+properties for a communicative payoff function. Unlike utilities that are, say, linearly or
+quadratically decreasing in physical distance
+\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
+that are exponentially decreasing in negative quadratic distance can model situations where a
+small amount of imprecision in communication is tolerable, whereas similarly small differences
+in intolerably far away interpretations matter very little, with a smooth transition between
+these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.
+
+In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
+is governed by three parameters: the number $\ns = \card{T}s$ of states in the state space,
+imprecision $\imprecision$ and tolerance $\toler$.
+
+\subsection{Experimental set-up}
+\label{sec:simulations}
+
+We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
+sender and receiver strategies, for each triplet of independent parameter values:
+$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
+$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps of the
+\rdd. A trial was considered converged, and thus stopped before the maximum of 200 rounds, if
+the total amount of change between strategies before and after an update step was smaller than
+a suitably chosen threshold. It is not guaranteed that strategies at halting time had converged
+to the eventual attracting state, whether they ran for 200 rounds or not. Our notion of
+convergence is therefore only a categorical measure for reaching a certain (well-considered,
+but eventually arbitrary) degree of stability. In other words, out notion of ``convergence'' is
+a measure of relative speed: it is true that the system reached a state in which evolutionary
+adaptations had slowed down almost to a halt before 200 update steps? This is motivated by
+pragmatic concerns regarding length of simulation time, but also theoretically justified,
+because we hypothesize that confusability of states leads to regularization of evolving
+strategies, which would show exactly in an increased speed of evolutionary trajectories towards
+well-behaved and regular signaling behavior.
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
+    \label{fig:example_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
+    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.05$}
+    \label{fig:example_stratsB}
+  \end{subfigure}
+
+  % \begin{subfigure}[]{0.45\textwidth}
+  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0$}
+  %   \label{fig:example_stratsC}
+  % \end{subfigure}
+  % \hfill
+  % \begin{subfigure}[]{0.45\textwidth}
+  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
+  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0.05$}
+  %   \label{fig:example_stratsD}
+  % \end{subfigure}
+
+  \caption{Example strategies under \rdd at stopping time of our simulations.}
+  \label{fig:example_strats}
+\end{figure}
+
+Representative examples for resulting strategy pairs are given in
+Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
+stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
+imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
+the given sender strategy approximates a pure sender strategy that crisply partitions the state
+space into non-convex sets. The irregular shape of the receiver strategy suggests that the
+pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
+the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
+trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
+in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
+rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
+receiver's interpretations are rather extended curves, peaking at a central point in the
+relevant ``categories.''
+
+These examples already show two interesting things. Firstly, inclusion of imprecision can lead
+to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
+strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
+for each signal, with a smooth transition in-between. The receiver's interpretations of signals
+can be seen as smoothed-out prototype regions. Secondly, (sender) strategies can approach
+non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
+time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
+Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
+for the limiting case. This was first observed by Elliott Wagner, as mentioned by
+\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
+scope of this paper, but we will shortly see that diffusion from confusability of states
+clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
+strategies.
+
+\subsection{Dependent measures}
+
+To further explore our simulation results, we calculated metrics that aim to numerically
+capture how vague, generally well-structured, and communicatively efficient the recorded
+strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
+signal use. \emph{Convexity} captures whether a behavioral strategy would project onto a convex
+pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
+strategy pairs.
+
+\paragraph{Entropy.} This classic information-theoretic notion
+captures the amount of uncertainty in a probability
+distribution. Roughly put, entropy of a signaling strategy captures
+inverse distance from a pure strategy. The usual definition of entropy
+applies directly to mixed strategies, but provably equivalent metrics
+for behavioral strategies are ready to hand:
+\begin{align*}
+  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
+  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
+  \state)) \\
+  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
+  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
+  \messg)) \,. 
+\end{align*}
+Values obtained by these definitions are lower bounded by $0$ and
+upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
+\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
+\log(|\States|)$. We work with normalized values. The sender
+strategies in Figures~\ref{fig:example_stratsA} and
+\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
+respectively. The receiver strategies have respective entropies $0.43$
+and $0.81$. In general, we expected that vague languages would have
+higher entropy than crisp ones and that increasing imprecision would
+lead to increased entropy, all else being equal.
+
+\paragraph{Convexity.} At least for sender strategies, which develop faster than receiver
+strategies, it also makes sense to define a categorical measure of convexity that compensates
+for potential vagueness. To determine whether a sender strategy $\Sstrat$ is convex despite
+possibly being vague, we look at the derived pure strategy $\Spure$ for which
+$\Spure(\state) = \argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$ is
+convex, we also count $\Sstrat$ as convex. The sender strategy in
+Figure~\ref{fig:example_stratsA} is not convex, while the one in
+Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
+strategies, like we hypothesized, we should see more convexity with increasing impairment all
+else equal.
+
+\paragraph{Expected utility.} We also recorded the expected utility of
+a strategy pair:
+\begin{align*}
+  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
+    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
+  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
+  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
+\end{align*}
+To make direct comparisons across different parameter settings, we
+normalize expected utility by the maximal amount of expected utility
+obtainable in the relevant game. The strategy pair in
+Figure~\ref{fig:example_stratsA} has a normalized expected utility of
+$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
+$0.95$. Generally, vagueness and imprecision can be expected to
+decrease expected utility
+\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
+whether communicative success drops unacceptably fast with moderate
+levels of vagueness and imprecision.
+
+\subsection{Results}
+
+\begin{figure}[t]
+  \centering
+
+  \includegraphics[width=0.75\textwidth]{plots/MeanMetrics3.pdf}
+
+  \caption{Means of gradient and proportions of categorical measures
+    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\imprecision
+    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
+    the entropies for the sender and receiver strategy.}
+  \label{fig:MeanMetrics}
+\end{figure}
+
+Figure~\ref{fig:MeanMetrics} shows plots summarizing a selected part of our findings.  For
+perspicuity, we only plot results for one level of tolerance $\beta = 0.1$, and leave out the
+case of $n_s = 90$. Still, every qualitative trend mentioned in the following applies to the
+whole set of results.
+
+As expected, increasing imprecision leads to higher entropy and lower expected
+utility. Importantly, however, imprecision does not necessarily lead to disastrous decline of
+communicative success. What is more, in line with our hypothesis that mere imprecision in
+imitation behavior can lead to behavior that looks like agents generalize across similar
+stimuli, higher imprecision led to a higher number of outcomes with convex sender
+strategies. It also led to higher rates of convergence. In fact, sufficient imprecision always
+ensured convergence and convexity. It appears that perceptual imprecision leads to more
+vagueness, slightly less communicative efficiency, but more regular, well-behaved languages in
+shorter time.
+
+Beyond promoting convexity and convergence, diffusion also has another interesting
+regularizing effect on the evolution of signaling. There was very
+little variation in the recorded metrics for evolved strategies, at
+least for higher values of imprecision. On closer inspection, it turns
+out that variability in low-imprecision conditions is not only due to
+non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
+gives two more examples of strategy pairs at stopping time. Both are
+obtained for the same triple of parameters, both converged before the
+maximum number of rounds, and both have
+convex sender strategies. However, they are not equally efficient. In
+fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized
+expected utility of $0.99$ while the one in
+Figure~\ref{fig:example_stratsD} only has $0.89$.
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
+    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
+    \label{fig:example_stratsC}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
+    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
+    \label{fig:example_stratsD}
+  \end{subfigure}
+
+  \caption{More example strategies under \rdd at stopping time of our simulations.}
+  \label{fig:MoreExample_strats}
+\end{figure}
+
+Interestingly, this type of variability in evolutionary outcomes can
+be weeded out by imprecision. To investigate this, we calculated the average
+distance between evolved sender strategies within each group of trials
+that had identical parameter values. We determined the distance
+between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
+distance between distributions $\Sstrat(\state)$ and
+$\Sstrat'(\state)$ at each choice point $\state$:
+\begin{align*}
+  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
+     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
+ \sqrt{\sum_{\messg \in  \Messgs}
+         \left ( \sqrt{\Sstrat(\state,\messg)} -
+         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
+\end{align*}
+To compensate for the arbitrariness of message use, we set the distance between strategies
+$\Sstrat$ and $\Sstrat'$ to be the maximum of $\text{HD}(\Sstrat,\Sstrat')$ and
+$\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
+indices. An example of the \emph{within group distance}, i.e., the average distances between
+all sender strategies obtained for the same parameter values, is plotted in
+Figure~\ref{fig:GroupMeasuresA} for $\beta = 0.1$ and $n_s = 10$. Despite some quantitative
+differences, the general trend is the same for all other parameter settings that we tested:
+with increasing imprecision, the resulting sender strategies were much more alike (modulo
+swapping of messages). This means that perceptual imprecision can speed up and unify
+evolutionary outcomes. It can amplify the emergence of sender strategies that are not only
+convex, but also regular in that they induce a vague category split exactly in the middle of
+the unit interval.
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/WithinGroupDistanceConcise.pdf}
+    \caption{Within group distance}
+    \label{fig:GroupMeasuresA}
+  \end{subfigure}
+    \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{plots/WithinGroupEUConcise.pdf}
+    \caption{Within group expected utility}
+    \label{fig:GroupMeasuresB}
+  \end{subfigure}
+
+    \caption{Within group measures for all runs with $\toler = 0.1$ and $n_s = 10$.}
+  \label{fig:GroupMeasures}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Levels of vagueness}
+The main motivation and intuitions given in Section~\ref{sec:repl-diff-dynam} describe probabilities in sender and receiver strategies as capturing the overall distribution of pure strategies in a population: $\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the receiver).
+This is in line with the standard interpretation of the replicator equation, but we should consider its philosophical implications.
+Based on this interpretation and our model, vagueness is seemingly being characterized as a strictly population-level phenomenon since it arises in a signaling system from the inability of individual agents to fully align their (non-vague) strategies because of imprecision.
+This is, we believe, a plausible mechanism that can already explain the existence of vagueness in a language even if we assume each agent commands a non-vague idiolect.
+
+We do not, however, want to commit to the idea that vagueness does not exist at the level of individual agents.
+The fact is that the model can be interpreted in alternative ways.
+The probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will be used by a sender when observing state $\state$.
+Whether the uncertainty represents variations in a population of agents playing pure strategies, a single sender using a non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is up for interpretation.
+The first option supported what we believe to be a plausible motivation for the mechanism of imitation of success with imprecision, but other motivations could be imagined to better fit an interpretation that includes non-deterministic agents.
+We thus believe that the model introduced in this paper does not exclude vagueness at the individual level, but to explore all possible interpretations would go beyond our current scope.
+
+\subsection{Evolutionary benefits of imprecision}
+The inclusion of confusability of similar states has noteworthy effects on the evolving meaning of signals.
+It transpired from our results that imprecision can have further accelerating and, surprisingly, unifying effects on meaning evolution.
+The unifying property of perceptual imprecision could be considered an evolutionary beneficial
+side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
+defined as the average expected utility that each evolved language scored when playing against
+an arbitrary other language obtained for the same parameter
+values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
+for other parameter values: while imprecision might decrease the communicative efficiency of
+individual languages, it increases the conceptual coherence and communicative success between
+independently evolving strategies. It is as if mere confusability of states imposes a
+regularity constraint on evolving categories.
+
+The phenomenon could potentially not just be a side effect.
+Given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
+This could be achieved by letting the imprecision parameter be subject to the evolutionary dynamics.
+The idea is for now mostly speculative, but we consider it an interesting avenue for future research.
+There is additional motivation to consider its potential if we see it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.
+
+\subsection{Related work}
+\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of reinforcement
+learning for sim-max games, in which successful play leads to reinforcement of choice options
+also for states similar to the ones that actually occurred. This not only leads to vague
+signaling of the appropriate kind, but also speeds up learning in such a way that, especially
+for sim-max games with higher numbers of states, higher levels of communicative success are
+reached in shorter learning periods. Our results complement and extend
+\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
+similar regularizing effects also for cases with low numbers of states and do not assume that
+agents have any kind of generalizing capacity in and of themselves, even if that is only
+implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
+State confusability has an effect on
+aggregate signaling behavior that can be described as generalization without generalizers: the
+dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
+``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
+single agent genuinely generalizes over stimuli or reasons about what a more systematic
+signaling strategy would be.
+
+\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
+in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
+clearly related to the present approach uses the notion of a \emph{quantal response}, also
+known as a \emph{soft-max function}
+\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
+main difference between this and our present approach is in where stochastic noise is assumed
+to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
+in case of imprecise imitation, it is perception and realization of similar states. There are
+cases, then, where evolving signaling behavior, as predicted by these two approaches, is quite
+different. Intuitively speaking, for a case with two messages, the further we venture away from
+a prototypical interpretation of either message, the less discriminative a signaling strategy
+would be when the source of ``trembles'' is the computation of expected utilities: to wit,
+since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
+response dynamics predict that senders would be almost indifferent. Sender behavior that
+evolves under confusion of states does not have this puzzling property.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+We set out to meet a technical challenge posed by Lipman's problem
+\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
+formal model that shows how vague language can evolve under selective pressure for efficient
+communication? On the assumption that agents are likely to mistake states for one another,
+especially when they are perceptually similar, derived a generalization of the replicator
+dynamic in terms of imitation of successes. The resulting population-level dynamic produced
+signaling behavior that is at the same time regular and by-and-large communicatively efficient,
+while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
+by-product of an arguably natural limitation on the discriminatory power of signaling agents.
+Although inability to sharply discriminate similar stimuli may lead to vagueness and bring about
+a (slight) decrease in communicative efficiency, there may also be an advantage, not of
+vagueness itself, but of its reason: systematicity in the confusability of states (which may be
+a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
+population-level without having to assume that agents themselves have any generalization power.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\appendix
 
@@
 
-\section{Exploring noisy imitation}
-\label{sec:exploring-rdd}
-
-How does a tendency to confuse similar states interact with selective pressure towards more
-efficient signaling strategies under the imprecise conditional imitation dynamic? If confusion
-probabilities are moderate and regular, in that they track similarity of states, we might
-expect a regularizing effect also on evolving signaling strategies. Indeed, we hypothesize that
-confusion of states can give rise to population-level aggregate behavior that looks \emph{as
-  if} information about what is good for one state percolates also to similar states. In other
-words, imprecise imitation may make signaling behavior on average look like agents generalize
-across similar states, even if no agent actually generalizes. To explore whether confusion of
-states can have this effect, we turn to numerical simulation.
-
-\subsection{Setting the stage}
-\label{sec:setting-stage}
-
-To obtain concrete results, we must fix how to represent states, similarity of states, the
-conditional confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max
-games. Confusion probabilities between states should be a function of perceptual similarity:
-the more similar two states are, the more likely they could be mistaken for each other.
-
-Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
-interval, including $0$ and $1$. All states occur, for simplicity, with the same probability.
-The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
-two states $\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual
-similarity function, as described by \citet{Nosofsky1986:Attention-Simil}:
-\begin{align*}
-  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
-      \begin{cases}
-    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
-    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
-    \end{cases}
-\end{align*}
-where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
-discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
-states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
-different parameter values. Other formalizations of perceptual similarity are possible,
-including ones that allow for different discriminability in different areas of the state space,
-but we stick with Nosofsky's similarity function for the time being, because it is
-mathematically simple, yet an established notion in mathematical psychology.
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=0.7\textwidth]{plots/NosofskySim.pdf}
-
-  \caption{Examples of Nosofsky-similarity for different values of
-    imprecision.}
-  \label{fig:NosofskySim}
-\end{figure}
-
-We further assume that the probability of confusing any two states $\mystate{i}$ and
-$\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
-observational confusability $P_o$ is simply the same as realizational confusability $P_r$ and
-that both are governed by the same imprecision parameter $\alpha$:
-\begin{align*}
-  P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
-  \imprecision) &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
-  \imprecision) \,.
-\end{align*}
-For $\imprecision = 0$, we obtain trivial confusion probabilities for which everything is reduced
-to the replicator dynamic. For $\imprecision > 0$, any state can be confused for any other
-state with some positive probability. For the uninteresting case of
-$\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
-realized as any other state with probability $\nicefrac{1}{\card{\States}}$.
-
-As for utility functions, we take another Nosofsky-style function:
-\begin{align*}
-  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
-      \begin{cases}
-    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
-    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
-    \end{cases}
-\end{align*}
-The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
-communicative imprecision. This choice of utility function is governed partly by convenience in
-parallel to the similarity function, but also because it has, as we believe, the right general
-properties for a communicative payoff function. Unlike utilities that are, say, linearly or
-quadratically decreasing in physical distance
-\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
-that are exponentially decreasing in negative quadratic distance can model situations where a
-small amount of imprecision in communication is tolerable, whereas similarly small differences
-in intolerably far away interpretations matter very little, with a smooth transition between
-these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.
-
-In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
-is governed by three parameters: the number $\ns = \card{T}s$ of states in the state space,
-imprecision $\imprecision$ and tolerance $\toler$.
-
-\subsection{Experimental set-up}
-\label{sec:simulations}
-
-We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
-sender and receiver strategies, for each triplet of independent parameter values:
-$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
-$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps of the
-\rdd. A trial was considered converged, and thus stopped before the maximum of 200 rounds, if
-the total amount of change between strategies before and after an update step was smaller than
-a suitably chosen threshold. It is not guaranteed that strategies at halting time had converged
-to the eventual attracting state, whether they ran for 200 rounds or not. Our notion of
-convergence is therefore only a categorical measure for reaching a certain (well-considered,
-but eventually arbitrary) degree of stability. In other words, out notion of ``convergence'' is
-a measure of relative speed: it is true that the system reached a state in which evolutionary
-adaptations had slowed down almost to a halt before 200 update steps? This is motivated by
-pragmatic concerns regarding length of simulation time, but also theoretically justified,
-because we hypothesize that confusability of states leads to regularization of evolving
-strategies, which would show exactly in an increased speed of evolutionary trajectories towards
-well-behaved and regular signaling behavior.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
-    \label{fig:example_stratsA}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.05$}
-    \label{fig:example_stratsB}
-  \end{subfigure}
-
-  % \begin{subfigure}[]{0.45\textwidth}
-  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0$}
-  %   \label{fig:example_stratsC}
-  % \end{subfigure}
-  % \hfill
-  % \begin{subfigure}[]{0.45\textwidth}
-  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0.05$}
-  %   \label{fig:example_stratsD}
-  % \end{subfigure}
-
-  \caption{Example strategies under \rdd at stopping time of our simulations.}
-  \label{fig:example_strats}
-\end{figure}
-
-Representative examples for resulting strategy pairs are given in
-Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
-stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
-imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
-the given sender strategy approximates a pure sender strategy that crisply partitions the state
-space into non-convex sets. The irregular shape of the receiver strategy suggests that the
-pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
-the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
-trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
-in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
-rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
-receiver's interpretations are rather extended curves, peaking at a central point in the
-relevant ``categories.''
-
-These examples already show two interesting things. Firstly, inclusion of imprecision can lead
-to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
-strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
-for each signal, with a smooth transition in-between. The receiver's interpretations of signals
-can be seen as smoothed-out prototype regions. Secondly, (sender) strategies can approach
-non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
-time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
-Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
-for the limiting case. This was first observed by Elliott Wagner, as mentioned by
-\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
-scope of this paper, but we will shortly see that diffusion from confusability of states
-clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
-strategies.
-
-\subsection{Dependent measures}
-
-To further explore our simulation results, we calculated metrics that aim to numerically
-capture how vague, generally well-structured, and communicatively efficient the recorded
-strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
-signal use. \emph{Convexity} captures whether a behavioral strategy would project onto a convex
-pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
-strategy pairs.
-
-\paragraph{Entropy.} This classic information-theoretic notion
-captures the amount of uncertainty in a probability
-distribution. Roughly put, entropy of a signaling strategy captures
-inverse distance from a pure strategy. The usual definition of entropy
-applies directly to mixed strategies, but provably equivalent metrics
-for behavioral strategies are ready to hand:
-\begin{align*}
-  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
-  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
-  \state)) \\
-  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
-  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
-  \messg)) \,. 
-\end{align*}
-Values obtained by these definitions are lower bounded by $0$ and
-upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
-\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
-\log(|\States|)$. We work with normalized values. The sender
-strategies in Figures~\ref{fig:example_stratsA} and
-\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
-respectively. The receiver strategies have respective entropies $0.43$
-and $0.81$. In general, we expected that vague languages would have
-higher entropy than crisp ones and that increasing imprecision would
-lead to increased entropy, all else being equal.
-
-\paragraph{Convexity.} At least for sender strategies, which develop faster than receiver
-strategies, it also makes sense to define a categorical measure of convexity that compensates
-for potential vagueness. To determine whether a sender strategy $\Sstrat$ is convex despite
-possibly being vague, we look at the derived pure strategy $\Spure$ for which
-$\Spure(\state) = \argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$ is
-convex, we also count $\Sstrat$ as convex. The sender strategy in
-Figure~\ref{fig:example_stratsA} is not convex, while the one in
-Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
-strategies, like we hypothesized, we should see more convexity with increasing impairment all
-else equal.
-
-\paragraph{Expected utility.} We also recorded the expected utility of
-a strategy pair:
-\begin{align*}
-  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
-    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
-  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
-  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
-\end{align*}
-To make direct comparisons across different parameter settings, we
-normalize expected utility by the maximal amount of expected utility
-obtainable in the relevant game. The strategy pair in
-Figure~\ref{fig:example_stratsA} has a normalized expected utility of
-$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
-$0.95$. Generally, vagueness and imprecision can be expected to
-decrease expected utility
-\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
-whether communicative success drops unacceptably fast with moderate
-levels of vagueness and imprecision.
-
-\subsection{Results}
-
-\begin{figure}[t]
-  \centering
-
-  \includegraphics[width=0.75\textwidth]{plots/MeanMetrics3.pdf}
-
-  \caption{Means of gradient and proportions of categorical measures
-    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\imprecision
-    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
-    the entropies for the sender and receiver strategy.}
-  \label{fig:MeanMetrics}
-\end{figure}
-
-Figure~\ref{fig:MeanMetrics} shows plots summarizing a selected part of our findings.  For
-perspicuity, we only plot results for one level of tolerance $\beta = 0.1$, and leave out the
-case of $n_s = 90$. Still, every qualitative trend mentioned in the following applies to the
-whole set of results.
-
-As expected, increasing imprecision leads to higher entropy and lower expected
-utility. Importantly, however, imprecision does not necessarily lead to disastrous decline of
-communicative success. What is more, in line with our hypothesis that mere imprecision in
-imitation behavior can lead to behavior that looks like agents generalize across similar
-stimuli, higher imprecision led to a higher number of outcomes with convex sender
-strategies. It also led to higher rates of convergence. In fact, sufficient imprecision always
-ensured convergence and convexity. It appears that perceptual imprecision leads to more
-vagueness, slightly less communicative efficiency, but more regular, well-behaved languages in
-shorter time.
-
-Beyond promoting convexity and convergence, diffusion also has another interesting
-regularizing effect on the evolution of signaling. There was very
-little variation in the recorded metrics for evolved strategies, at
-least for higher values of imprecision. On closer inspection, it turns
-out that variability in low-imprecision conditions is not only due to
-non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
-gives two more examples of strategy pairs at stopping time. Both are
-obtained for the same triple of parameters, both converged before the
-maximum number of rounds, and both have
-convex sender strategies. However, they are not equally efficient. In
-fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized
-expected utility of $0.99$ while the one in
-Figure~\ref{fig:example_stratsD} only has $0.89$.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
-    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
-    \label{fig:example_stratsC}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
-    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
-    \label{fig:example_stratsD}
-  \end{subfigure}
-
-  \caption{More example strategies under \rdd at stopping time of our simulations.}
-  \label{fig:MoreExample_strats}
-\end{figure}
-
-Interestingly, this type of variability in evolutionary outcomes can
-be weeded out by imprecision. To investigate this, we calculated the average
-distance between evolved sender strategies within each group of trials
-that had identical parameter values. We determined the distance
-between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
-distance between distributions $\Sstrat(\state)$ and
-$\Sstrat'(\state)$ at each choice point $\state$:
-\begin{align*}
-  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
-     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
- \sqrt{\sum_{\messg \in  \Messgs}
-         \left ( \sqrt{\Sstrat(\state,\messg)} -
-         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
-\end{align*}
-To compensate for the arbitrariness of message use, we set the distance between strategies
-$\Sstrat$ and $\Sstrat'$ to be the maximum of $\text{HD}(\Sstrat,\Sstrat')$ and
-$\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
-indices. An example of the \emph{within group distance}, i.e., the average distances between
-all sender strategies obtained for the same parameter values, is plotted in
-Figure~\ref{fig:GroupMeasuresA} for $\beta = 0.1$ and $n_s = 10$. Despite some quantitative
-differences, the general trend is the same for all other parameter settings that we tested:
-with increasing imprecision, the resulting sender strategies were much more alike (modulo
-swapping of messages). This means that perceptual imprecision can speed up and unify
-evolutionary outcomes. It can amplify the emergence of sender strategies that are not only
-convex, but also regular in that they induce a vague category split exactly in the middle of
-the unit interval.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/WithinGroupDistanceConcise.pdf}
-    \caption{Within group distance}
-    \label{fig:GroupMeasuresA}
-  \end{subfigure}
-    \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/WithinGroupEUConcise.pdf}
-    \caption{Within group expected utility}
-    \label{fig:GroupMeasuresB}
-  \end{subfigure}
-
-    \caption{Within group measures for all runs with $\toler = 0.1$ and $n_s = 10$.}
-  \label{fig:GroupMeasures}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Discussion}
-\label{sec:discussion}
-
-\subsection{Levels of vagueness}
-The main motivation and intuitions given in Section~\ref{sec:repl-diff-dynam} describe probabilities in sender and receiver strategies as capturing the overall distribution of pure strategies in a population: $\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the receiver).
-This is in line with the standard interpretation of the replicator equation, but we should consider its philosophical implications.
-Based on this interpretation and our model, vagueness is seemingly being characterized as a strictly population-level phenomenon since it arises in a signaling system from the inability of individual agents to fully align their (non-vague) strategies because of imprecision.
-This is, we believe, a plausible mechanism that can already explain the existence of vagueness in a language even if we assume each agent commands a non-vague idiolect.
-
-We do not, however, want to commit to the idea that vagueness does not exist at the level of individual agents.
-The fact is that the model can be interpreted in alternative ways.
-The probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will be used by a sender when observing state $\state$.
-Whether the uncertainty represents variations in a population of agents playing pure strategies, a single sender using a non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is up for interpretation.
-The first option supported what we believe to be a plausible motivation for the mechanism of imitation of success with imprecision, but other motivations could be imagined to better fit an interpretation that includes non-deterministic agents.
-We thus believe that the model introduced in this paper does not exclude vagueness at the individual level, but to explore all possible interpretations would go beyond our current scope.
-
-\subsection{Evolutionary benefits of imprecision}
-The inclusion of confusability of similar states has noteworthy effects on the evolving meaning of signals.
-It transpired from our results that imprecision can have further accelerating and, surprisingly, unifying effects on meaning evolution.
-The unifying property of perceptual imprecision could be considered an evolutionary beneficial
-side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
-defined as the average expected utility that each evolved language scored when playing against
-an arbitrary other language obtained for the same parameter
-values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
-for other parameter values: while imprecision might decrease the communicative efficiency of
-individual languages, it increases the conceptual coherence and communicative success between
-independently evolving strategies. It is as if mere confusability of states imposes a
-regularity constraint on evolving categories.
-
-The phenomenon could potentially not just be a side effect.
-Given the benefit of a certain amount of imprecision we observe when comparing within group expected utility, it would be interesting to study whether, under certain conditions, this group level advantage could trump the individual level disadvantage of a vague language, and thus actually select for a certain amount of imprecision.
-This could be achieved by letting the imprecision parameter be subject to the evolutionary dynamics.
-The idea is for now mostly speculative, but we consider it an interesting avenue for future research.
-There is additional motivation to consider its potential if we see it in the eyes of multilevel selection theory \citep{Wilson1994,OGorman2008}.
-
-\subsection{Related work}
-\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of reinforcement
-learning for sim-max games, in which successful play leads to reinforcement of choice options
-also for states similar to the ones that actually occurred. This not only leads to vague
-signaling of the appropriate kind, but also speeds up learning in such a way that, especially
-for sim-max games with higher numbers of states, higher levels of communicative success are
-reached in shorter learning periods. Our results complement and extend
-\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
-similar regularizing effects also for cases with low numbers of states and do not assume that
-agents have any kind of generalizing capacity in and of themselves, even if that is only
-implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
-State confusability has an effect on
-aggregate signaling behavior that can be described as generalization without generalizers: the
-dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
-``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
-single agent genuinely generalizes over stimuli or reasons about what a more systematic
-signaling strategy would be.
-
-\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
-in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
-clearly related to the present approach uses the notion of a \emph{quantal response}, also
-known as a \emph{soft-max function}
-\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
-main difference between this and our present approach is in where stochastic noise is assumed
-to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
-in case of imprecise imitation, it is perception and realization of similar states. There are
-cases, then, where evolving signaling behavior, as predicted by these two approaches, is quite
-different. Intuitively speaking, for a case with two messages, the further we venture away from
-a prototypical interpretation of either message, the less discriminative a signaling strategy
-would be when the source of ``trembles'' is the computation of expected utilities: to wit,
-since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
-response dynamics predict that senders would be almost indifferent. Sender behavior that
-evolves under confusion of states does not have this puzzling property.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-We set out to meet a technical challenge posed by Lipman's problem
-\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
-formal model that shows how vague language can evolve under selective pressure for efficient
-communication? On the assumption that agents are likely to mistake states for one another,
-especially when they are perceptually similar, derived a generalization of the replicator
-dynamic in terms of imitation of successes. The resulting population-level dynamic produced
-signaling behavior that is at the same time regular and by-and-large communicatively efficient,
-while also showing the crucial marks of vagueness. In a sense, the model derives vagueness as a
-by-product of an arguably natural limitation on the discriminatory power of signaling agents.
-Although inability to sharply discriminate similar stimuli may lead to vagueness and bring about
-a (slight) decrease in communicative efficiency, there may also be an advantage, not of
-vagueness itself, but of its reason: systematicity in the confusability of states (which may be
-a natural by-product of the perceptual system) supports \emph{as-if}-generalization at the
-population-level without having to assume that agents themselves have any generalization power.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \printbibliography[heading=bibintoc]

### 2015-09-30 paper/version_04/paper.tex
section 3 edits

@@
 
-\section{Imprecise conditional imitation}
+\section{Imitation with confusability of states}
 \label{sec:repl-diff-dynam}
@@
   $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
-  adapt their behavior locally, i.e., independently at each choice point. Our
-  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
-  numerical simulations. But it also seems the more plausible choice for imitation-based update
-  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
-  particular situation, not how that agent would behave in \emph{all} relevant choice
-  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
-  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
-  behavioral strategies.)} The sender's behavioral strategies are functions
-$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
-probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's are
-functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message $\messg \in \Messgs$
-to a probability of each interpretation $\state \in \States$ being chosen in response to
-$\messg$. Although behavioral strategies are probabilistic, evolutionary modeling usually
-imagines that every individual agent has a non-probabilistic strategy. Behavioral strategies
-then capture average population behavior. In a virtually infinite population, the number
-$\Sstrat(\messg \probbar \state)$, for instance, is then the probability that a randomly
-sampled sender would send message $\messg$ if the actual state was $\state$. Similarly,
-$\Rstrat(\state \probbar \messg)$ is then the probability with which a randomly sampled
-receiver interprets $\messg$ as $\state$.
+  adapt their behavior locally, i.e., independently at each choice point. Our focus on
+  behavioral strategies greatly reduces the complexity of the dynamic and simplifies numerical
+  simulations. But it also seems the more plausible choice for imitation-based update protocols
+  of the kind we consider here: agents only observe how, on some occasion, some other agent
+  behaved in \emph{one} particular situation, not how that agent would behave in \emph{all}
+  relevant choice situations; they imitate the use of a single word, so to speak, not a whole
+  lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference between
+  dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
+functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
+to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
+are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
+$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
+in response to $\messg$. Although behavioral strategies are probabilistic, evolutionary
+modeling usually imagines that every individual agent has a non-probabilistic
+strategy. Behavioral strategies then capture average population behavior. Assuming a virtually
+infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
+probability that a randomly sampled sender would send message $\messg$ if the actual state was
+$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
+randomly sampled receiver interprets $\messg$ as $\state$.
 
@@
 Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
-outcomes, weighted by how likely they are to occur. For instance, the expected utility for the
-sender of choosing message $\messg$, given tate $\state$ and a receiver strategy $\Rstrat$ is
-the sum, for each act $\act$, of the probability that some random receiver will choose
-$\act$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and $\act$.
+outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
+of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
+sum, for each interpretation $\state'$, of the probability that some random receiver will
+choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
+$\state'$.
 
@@
   \label{eq:RDDiscrete}
-  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \cdot
-    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \cdot
+  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
+    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \myts
     \EU(\act, \messg,\Sstrat)\,.   
@@
 
-\subsection{A noisy replicator dynamic}
+\subsection{Noisy replicator dynamic}
 \label{sec:noisy-repl-dynam}
@@
 Probabilistic imitation of success presupposes that agents make no mistakes when observing
-states. This may not always be an appropriate assumption. Especially for sim-max games, where,
-intuitively, some states can be fairly similar and therefore likely to be confused for one
+states. This may not always be an appropriate assumption, especially for sim-max games, where,
+some states can be fairly similar and therefore likely to be confused for one
 another. Appendix~XYZ shows how to derive a generalization of the replicator dynamic from the
-assumption that probabilistic imitation of success is affected by two sources of probabilistic
-confusion of states:
+assumption that imitation of success is affected by two sources of probabilistic noise:
 
@@
 
-The mean changes in the population frequencies of choices under noise-affected imitation of
-success can be covered by the following discrete-time formulation:
+\noindent Noise-free imitation of success is the special case where $P_o(\mystate{o} \probbar
+\mystate{a}) = 1$ iff $\mystate{a} = \mystate{o}$ and $P_r(\mystate{r} \probbar
+\mystate{i}) = 1$ iff $\mystate{i} = \mystate{r}$. In that case, the population-level impact of
+imitation is described by the replicator equation in (\ref{eq:RDDiscrete}).
+
+If noise is possibly non-trivial, the mean changes in population frequencies of choices
+under noise-affected imitation of success can be covered by the following discrete-time
+formulation:
 
@@
 
-This looks very much like the discrete-time formulation of the standard replicator dynamic in
-(\ref{eq:RDDiscrete}), but there are differences (the Appendix~XYZ for full definitions and
-formal detail). Firstly, expected utility here takes stochastic confusability of states into
-account in a natural way. Secondly, where the standard replicator dynamic had probabilities
-$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
-$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. This is the major
-difference, so we should dwell on it for a brief moment.
-
-Look at the sender case first, where $\Sstrat(\messg \probbar \state)$ is the probability with
-which $\messg$ is played in response to actual state $\state$ on average in the population. If
-agents may not perceive the actual state, this should also affect what they observe when trying
-to imitate another agents behavior. $P_o(\messg \probbar \state)$ is therefore the probability,
-given that a potential imitator has observed $\state$, that he would observe a random agent
-play $\messg$, which he may or may not choose to imitate. Similarly for the receiver, we are
-not looking at occurrence probability $\Rstrat(\state \probbar \messg)$ but at the probability
+\noindent This looks very much like the discrete-time formulation of the standard replicator
+dynamic in (\ref{eq:RDDiscrete}), but there are differences (the Appendix~XYZ for full
+definitions and formal detail). Firstly, expected utility here takes stochastic confusability
+of states into account in a natural way. Secondly, where the standard replicator dynamic had
+probabilities $\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now
+have $P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. This is the
+major difference, so we should dwell on it for a brief moment.
+
+Look at the sender case first. Is some agent $X$ gets a chance to revise his behavior, he
+observes a state $\state$ and, at the same time, that some agent $Y$ chose $\messg$. But if $X$
+perceives $\state$ that does not necessarily mean that the actual state is $\state$, and
+similarly, that agent $Y$ also perceived state $\state$. Hence, where the replicator dynamic in
+(\ref{eq:RDDiscrete}), derived from noise-free imitation, has the probability
+$\Sstrat(\messg \probbar \state)$ with which $\messg$ is played in response to actual state
+$\state$ on average in the population, we substitute $P_o(\messg \probbar \state)$: the
+probability, given that a potential imitator has observed $\state$, that he would observe a
+random agent play $\messg$. Similar considerations apply to the receiver case: we are not
+looking at occurrence probability $\Rstrat(\state \probbar \messg)$ but at the probability
 $P_o(\state \probbar \messg)$ with which an agent who observes $\messg$ observes another agent
@@
 Obviously, $P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ are to be derived
-from, among others, probabilistic confusion of states $P_o(\mystate{o} \probbar \mystate{a})$
-and $P_r(\mystate{r} \probbar \mystate{o})$. Details are given in Appendix~XYZ. If there is no
-noise, $P_o(\messg \probbar \state)$ reduces to $\Sstrat(\messg \probbar \state)$ and
-$P_o(\state \probbar \messg)$ reduces to $\Rstrat(\state \probbar \messg)$. But if there is,
-imitation probabilities must take confusability of states into account. In this sense, the
-noisy replicator dynamic in (\ref{eq:nRD}) conservatively extends the classic case in
-(\ref{eq:RDDiscrete}).
+from, among others, the assumed probabilistic confusion rates
+$P_o(\mystate{o} \probbar \mystate{a})$ and $P_r(\mystate{r} \probbar \mystate{o})$. Details
+are given in Appendix~XYZ. If there is no noise, $P_o(\messg \probbar \state)$ reduces to
+$\Sstrat(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
+$\Rstrat(\state \probbar \messg)$. But if there is, we must take
+confusability of states into account. In this sense, the noisy replicator dynamic in
+(\ref{eq:nRD}) conservatively extends the classic case in (\ref{eq:RDDiscrete}).
 

### 2015-10-02 paper/version_04/paper.tex
section 3 edits
Section 3 edits and appendix

--- 2015-09-29 paper/version_04/paper.tex
@@
 of everyday communication, but it is troublesome for some of the most
-prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell and the young
+prominent theories of language and meaning. This is especially so for the logico-semantic tradition of Frege, Russell and the young
 Wittgenstein which is challenged by the paradoxes vagueness gives rise
@@
 Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
-approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
+approach to others. Appendix~\ref{sec:repl-diff-dynam} provides formal detail.
 
@@
 \citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
+  \node (3) [right=of 2] {$\messg = \Spure(\mystate{1}) \in \Messgs$};
+  \node (4) [right=of 3] {$\mystate{2} = \Rpure(\messg) \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {S} (3)
+    (3) edge node {R} (4);
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with
+    probability $\Pr(\mystate{1})$; a sender (S) selects a message
+    $\messg = \Spure(\mystate{1})$ determined by his pure strategy $\Spure$; a receiver (R)
+    selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy
+    $\Rpure$. Payoff for both sender and receiver is given by
+    $\Utils(\mystate{1}, \mystate{2})$.}
+  \label{fig:sim-max-illustration}
+\end{figure}
 
@@
 identify the receiver's acts with the states of the world, so that the game is one of guessing
-the actual state, so to speak. The utility function should be a monotonically increasing
+the actual state, so to speak. The utility function should be a monotonically decreasing
 function of similarity.
-A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the probability that a given message would be used for each state.
-The analogous happens for the receiver's behavior.
-A play in such a game with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
-  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
-  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {S} (3)
-    (3) edge node {R} (4);
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a receiver (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
-Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
-   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
-  \label{fig:sim-max-illustration}
-\end{figure}
+In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
+A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.
 
@@
 corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
-other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
+other. In this way, \citet{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
 system can be obtained: storing the prototypes lets us recover the categories without having to
-store each category's extension. However, what is left unexplained is where the prototypes
+store each category's extension. However, what is left unexplained, is where the prototypes
 come from, and why we would not see just any distribution of prototypes as an equally efficient
@@
 
-Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
-\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be met by evolving
+An alternative approach is taken by \citet{FrankeJager2010:Vagueness-Signa} and
+\citet{OConnor2013:The-Evolution-o} who show how the above desiderata can be met by evolving
 strategies in sim-max games. The approach we take here is similar in spirit, but different in
@@
 abstractly, at the level of a population of agents, we can describe how average behavioral
-dispositions will evolve. The population-level perspective abstracts over small stochastic
+dispositions will evolve. The population-level perspective abstracts over petty stochastic
 fluctuations and zooms in on the general tendency or direction of evolution that ensues from
@@
   $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
-  side might still need to be normalized so as to have probabilities sum to one. In general,
+  side must possibly yet be normalized so as to have probabilities sum to one. In general,
   writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
@@
 \citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
-Appendix~\ref{sec:formal-stuff} for a derivation of the standard replicator dynamic from this update scheme).
+Appendix~XYZ for a derivation of the standard replicator dynamic from this update scheme).
 Every now and then a random agent gets a chance to alter his behavior for one of his choice
@@
 The presence of observation and realization noise also affects imitation of successes. Here, we
-focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:formal-stuff}. Take a sender who gets to
+focus on the main ideas. Formal detail is provided by Appendix~XYZ. Take a sender who gets to
 revise behavior for state $\state$. What agents can plausibly revise by imitation is their pure
@@
 sender population behavior $\Sstrat$ and the given observation noise (see
-Appendix~\ref{sec:formal-stuff}). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
+Appendix~XYZ). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
 probability given by the expected utility of sending $\messg$ when perceiving state
@@
 
-Appendix~\ref{sec:formal-stuff} shows how imprecise imitation of this sort leads to mean changes in population
+Appendix~XYZ shows how imprecise imitation of this sort leads to mean changes in population
 frequencies of choices that can be covered by the following discrete-time formulation:
@@
 Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
-interval, including $0$ and $1$. All states occur, for simplicity, with the same probability (\emph{i.e.}~$P_a$ is uniform).
+interval, including $0$ and $1$. All states occur, for simplicity, with the same probability.
 The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
@@
     0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
     \end{cases}
@@
     0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \toler^2} \right ) & \textrm{otherwise.} \\
+ \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
     \end{cases}
@@
 
-  \caption{Example strategies under \rdd at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
+  \caption{Example strategies under \rdd at stopping time of our simulations.}
   \label{fig:example_strats}
@@
 The unifying property of perceptual imprecision could be considered an evolutionary beneficial
-side-effect. A certain degree of imprecision can lead to higher \emph{within group expected utility},
+side-effect. Higher imprecision can lead to higher \emph{within group expected utility},
 defined as the average expected utility that each evolved language scored when playing against

### 2015-10-05 paper/version_04/paper.tex
Fixed references to Appendix.

@@
 \citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
-  \node (3) [right=of 2] {$\messg = \Spure(\mystate{1}) \in \Messgs$};
-  \node (4) [right=of 3] {$\mystate{2} = \Rpure(\messg) \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {S} (3)
-    (3) edge node {R} (4);
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game: Nature (N) chooses a state $\mystate{1}$ with
-    probability $\Pr(\mystate{1})$; a sender (S) selects a message
-    $\messg = \Spure(\mystate{1})$ determined by his pure strategy $\Spure$; a receiver (R)
-    selects a state $\mystate{2} = \Rpure(\messg)$ determined by her pure strategy
-    $\Rpure$. Payoff for both sender and receiver is given by
-    $\Utils(\mystate{1}, \mystate{2})$.}
-  \label{fig:sim-max-illustration}
-\end{figure}
 
@@
 identify the receiver's acts with the states of the world, so that the game is one of guessing
-the actual state, so to speak. The utility function should be a monotonically decreasing
+the actual state, so to speak. The utility function should be a monotonically increasing
 function of similarity.
-In its original formulation, a sender's behavior is represented by a pure strategy $\Spure \in \Messgs^\States$ which associates states to messages (given a state $\state \in \States$ which message $\messg \in \Messgs$ would be sent), and a receiver's behavior by a pure strategy $\Rpure \in \States^\Messgs$ which does the opposite.
-A play in such a game is illustrated in Figure~\ref{fig:sim-max-illustration}.
+A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the \emph{probability} that a given message would be used for each state.
+The analogous happens for the receiver's behavior.
+A play in such a game with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
+  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
+  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {S} (3)
+    (3) edge node {R} (4);
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a receiver (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
+Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
+   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
+  \label{fig:sim-max-illustration}
+\end{figure}
 

### 2015-10-06 paper/version_04/paper.tex
Defining utility directly in terms of similarity.

@@
 function of similarity.
-A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the \emph{probability} that a given message would be used for each state.
+A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the probability that a given message would be used for each state.
 The analogous happens for the receiver's behavior.
@@
 
-As for utility functions, we take another Nosofsky-style function:
+As for utility, we define it in terms of similarity and introduce another free parameter, $\toler \ge 0$.
+ The intention is for this parameter to model the amount of tolerable pragmatic slack, which we felt should be allowed to vary separately from perceptual imprecision:
 \begin{align*}
   \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
-      \begin{cases}
-    1 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
-    0 & \textrm{if } \toler = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^s}{ \toler^2} \right ) & \textrm{otherwise.} \\
-    \end{cases}
-\end{align*}
-The tolerance parameter $\toler \ge 0$ models the amount of tolerable pragmatic slack or
-communicative imprecision. This choice of utility function is governed partly by convenience in
-parallel to the similarity function, but also because it has, as we believe, the right general
-properties for a communicative payoff function. Unlike utilities that are, say, linearly or
+      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)
+\end{align*}
+This choice of utility function is governed partly by convenience, but also because we believe it has the right general properties for a communicative payoff function.
+Unlike utilities that are, say, linearly or
 quadratically decreasing in physical distance

### 2015-10-15 paper/version_04/paper.tex
minor cosmetics; draft email

@@
   recent like-minded models, we describe an aggregate population-level dynamic that describes a
-  process of imitation of successful signaling behavior under imprecise perception and
-  realization of similar stimuli. Applying this new dynamic to a generalization of Lewis'
-  signaling games, we show that stochastic imprecision leads to vague, yet by-and-large
-  efficient signal use, and, moreover, that it unifies evolutionary outcomes and helps avoid
-  suboptimal categorization. The upshot of this is that we see \emph{as-if}-generalization at
-  an aggregate level, without agents actually generalizing.
+  process of imitation of successful behavior under imprecise perception and realization of
+  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
+  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
+  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal
+  categorization. The upshot of this is that we see \emph{as-if}-generalization at an aggregate
+  level, without agents actually generalizing.
 \end{abstract}
@@
 contributes to the latter line of thought. Concretely, we introduce a dynamic that describes
-average behavior resulting from imitation under confusability of similar stimuli. We show that
-stochastic noise can not only lead to vague, yet relatively efficient signal use, but can also
-unify evolutionary outcomes and help avoid suboptimal categorization: regularities in
-confusability of stimuli also regularize signaling behavior and evolving categories. The system
-gives rise to generalization behavior without agents actually generalizing.
+changes in aggregate population behavior as the result of imitation with probabilistic
+confusability of similar stimuli. We show that stochastic noise can not only lead to vague, yet
+relatively efficient signal use, but can also unify evolutionary outcomes and help avoid
+suboptimal categorization: regularities in confusability of stimuli also regularize signaling
+behavior and evolving categories. The system gives rise to generalization behavior without
+agents actually generalizing.
 
@@
 $\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
-and $\mystate{2}$, and a utility function $\Utils \mycolon \States \times \States \rightarrow
-\mathds{R}$ such that $\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver
-for a play with actual state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We
-identify the receiver's acts with the states of the world, so that the game is one of guessing
-the actual state, so to speak. The utility function should be a monotonically increasing
-function of similarity.
-A sender's behavior can be represented by a pure strategy $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for each state, a mixed strategy $\Smixed \in \Delta(\Messgs^\States)$, which characterizes a probability distribution over pure strategies, or a behavioral strategy $\Sstrat \in \Delta(\Messgs)^\States$, which defines the probability that a given message would be used for each state.
-The analogous happens for the receiver's behavior.
-A play in such a game with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
-  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
-  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {S} (3)
-    (3) edge node {R} (4);
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a sender (S) selects a message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a receiver (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
-Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
-   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
-  \label{fig:sim-max-illustration}
-\end{figure}
+and $\mystate{2}$, and a utility function
+$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
+$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
+state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
+with the states of the world, so that the game is one of guessing the actual state, so to
+speak. The utility function should be a monotonically increasing function of similarity.
+Non-probabilistic sender behavior can be represented by a pure strategy
+$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
+each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.
 
@@
 
-        \caption{Example of a Voronoi language on $\States = [0,1]$. The (pure) sender strategy
+        \caption{Example of a Voronoi language on $\States = [0,1]$. The pure sender strategy
           $\Spure$ uses one signal for the lower half, and another for the upper half of the
-          unit interval. The (pure) receiver strategy $\Rpure$ selects the central elements in
+          unit interval. The pure receiver strategy $\Rpure$ selects the central elements in
           the respective intervals. See \citet{JagerMetzger2011:Voronoi-Languag} for further
@@
 $\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
-$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility
-function shared by senders and receivers in the population. A \emph{behavioral strategy} is a
-function that maps an agent's choice points to a probability distribution over available
-choices.\footnote{Our focus is on behavioral strategies, not \emph{mixed strategies}, i.e.,
-  probability distributions over functions from each choice point to an act, such as
-  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
-  adapt their behavior locally, i.e., independently at each choice point. Our focus on
-  behavioral strategies greatly reduces the complexity of the dynamic and simplifies numerical
-  simulations. But it also seems the more plausible choice for imitation-based update protocols
-  of the kind we consider here: agents only observe how, on some occasion, some other agent
-  behaved in \emph{one} particular situation, not how that agent would behave in \emph{all}
-  relevant choice situations; they imitate the use of a single word, so to speak, not a whole
-  lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference between
-  dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
+$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
+senders and receivers in the population. A \emph{behavioral strategy} is a function that maps
+an agent's choice points to a probability distribution over available
+choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioral strategies, not \emph{mixed
+    strategies}, i.e., probability distributions over functions from each choice point to an
+  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume
+  that agents can adapt their behavior locally, i.e., independently at each choice point. Our
+  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
+  numerical simulations. But it also seems the more plausible choice for imitation-based update
+  protocols of the kind we consider here: agents only observe how, on some occasion, some other
+  agent behaved in \emph{one} particular situation, not how that agent would behave in
+  \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
+  not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
+  between dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
 functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
@@
 $\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
-randomly sampled receiver interprets $\messg$ as $\state$.
+randomly sampled receiver interprets $\messg$ as $\state$. A play of single evolutionary game
+with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
+
+\begin{figure}
+  \centering
+
+  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
+
+  \node (1) {};
+  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
+  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
+  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
+
+  \path
+    (1) edge node {N} (2)
+    (2) edge node {S} (3)
+    (3) edge node {R} (4);
+
+  \end{tikzpicture}
+
+  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a
+    state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a random sender (S) selects a
+    message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a random receiver
+    (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
+    Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
+   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
+  \label{fig:sim-max-illustration}
+\end{figure}
 
@@
   \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
-    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\act \probbar \messg) & \propto \Rstrat(\act \probbar \messg) \myts
-    \EU(\act, \messg,\Sstrat)\,.   
+    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
+    \EU(\state, \messg,\Sstrat)\,.   
 \end{align}
@@
 \citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
-Appendix~\ref{sec:formal-stuff} for a derivation of the standard replicator dynamic from this update scheme).
-Every now and then a random agent gets a chance to alter his behavior for one of his choice
-points. Say, this is a sender who gets to ``reconsider'' his choice of message for state
-$\state$. The revising agent then observes what some random agent does at $\state$, say
-$\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
+Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
+update scheme).  Every now and then a random agent gets a chance to alter his behavior for one
+of his choice points. Say, this is a sender who gets to ``reconsider'' his choice of message
+for state $\state$. The revising agent then observes what some random agent does at $\state$,
+say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
 expected utility of $\messg$ for state $\state$.
@@
 observing states, or choosing interpretations. This may not always be an appropriate
-assumption, especially when some states can be fairly similar and therefore likely to be
+assumption, especially when some states can be perceptually similar and therefore likely to be
 confused for one another. Therefore, imitation dynamics could be affected by at least two
@@
 The presence of observation and realization noise also affects imitation of successes. Here, we
-focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:formal-stuff}. Take a sender who gets to
-revise behavior for state $\state$. What agents can plausibly revise by imitation is their pure
-strategy, which maps perceived states onto messages. We should therefore assume that when an
-agent gets to revise his strategy for the perceived state $\state$, this need not necessarily
-be the actual state. Moreover, when that agent observes $\state$, another possibly
-to-be-imitated agent may yet perceive a completely different state altogether. Given what that
-latter agent perceives, his actual (pure) strategy will determine what he plays. In sum, to
-describe how likely the potential imitator observes a message choice $\messg$, we are
-interested in the conditional probability $P_o(\messg \probbar \state)$ that some other random
-agent realizes $\messg$ when the first agent perceived $\state$. This
-$P_o(\messg \probbar \state)$ is derived from the prior probability of states, the current
-sender population behavior $\Sstrat$ and the given observation noise (see
-Appendix~\ref{sec:formal-stuff}). Eventually, the imitating agent adopts $\messg$ as his choice for $\state$ with
-probability given by the expected utility of sending $\messg$ when perceiving state
-$\state$. Expected utility should, of course, take the probabilistic confusability of states
-into account as well.
-
-Similar considerations apply to the receiver side. If an agent who gets a chance to change his
+focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:imit-succ-with}. Take a
+sender who gets to revise behavior for state $\state$. What agents can plausibly revise by
+imitation is their pure strategy, which maps perceived states onto messages. When an agent gets
+to revise his strategy for the perceived state $\state$, this need not necessarily be the
+actual state. Moreover, when that agent observes $\state$, another agent may perceive yet a
+different state. Given what that latter agent perceives, his actual (pure) strategy will
+determine what he plays. In sum, to describe how likely the potential imitator observes a
+message choice $\messg$, we are interested in the conditional probability
+$P_o(\messg \probbar \state)$ that some other random agent selects $\messg$ when the first
+agent perceives $\state$. This $P_o(\messg \probbar \state)$ is derived from the prior
+probability of states, the current sender population behavior $\Sstrat$ and the given
+observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
+$\messg$ as his choice for $\state$ with probability given by the expected utility of sending
+$\messg$ when perceiving state $\state$. Expected utility should, of course, take the
+probabilistic confusability of states into account as well.
+
+Similar considerations apply to the receiver side. If an agent gets a chance to change his
 intended interpretation of $\messg$, we need to look at the conditional probability
@@
 
-Appendix~\ref{sec:formal-stuff} shows how imprecise imitation of this sort leads to mean changes in population
+Appendix~\ref{sec:imit-succ-with} shows how imprecise imitation of this sort leads to mean changes in population
 frequencies of choices that can be covered by the following discrete-time formulation:
@@
 Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
-interval, including $0$ and $1$. All states occur, for simplicity, with the same probability (\emph{i.e.}~$P_a$ is uniform).
-The distance $\card{\mystate{i} - \mystate{j}}$ is the objective, physical similarity between
-two states $\mystate{i}$ and $\mystate{j}$. Distance in physical space feeds into a perceptual
-similarity function, as described by \citet{Nosofsky1986:Attention-Simil}:
+interval, including $0$ and $1$. All states occur, for simplicity, with the same probability
+(i.e., $P_a$ is uniform).  The distance $\card{\mystate{i} - \mystate{j}}$ is the objective,
+physical similarity between two states $\mystate{i}$ and $\mystate{j}$. Distance in physical
+space feeds into a perceptual similarity function, as described by
+\citet{Nosofsky1986:Attention-Simil}:
 \begin{align*}
@@
   P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
-  \imprecision) &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
+  \imprecision)\,, &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
   \imprecision) \,.
@@
   \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
-      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)
+      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)\,.
 \end{align*}
@@
 therefore only a categorical measure for reaching a certain (well-considered, but eventually
-arbitrary) degree of stability. In other words, out notion of ``convergence'' is a measure of
-relative speed: it is true that the system reached a state in which evolutionary adaptations
+arbitrary) degree of stability. In other words, our notion of ``convergence'' is a measure of
+relative speed: is it true that the system reached a state in which evolutionary adaptations
 had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
@@
 space into non-convex sets. The irregular shape of the receiver strategy suggests that the
-pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
+pictured strategy pair has not yet reached a stable state. Indeed,
 the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
@@
 
-\paragraph{Entropy.} This classic information-theoretic notion
-captures the amount of uncertainty in a probability
-distribution. Roughly put, entropy of a signaling strategy captures
-inverse distance from a pure strategy. The usual definition of entropy
-applies directly to mixed strategies, but provably equivalent metrics
-for behavioral strategies are ready to hand:
+\paragraph{Entropy.} This classic information-theoretic notion captures the amount of
+uncertainty in a probability distribution. Roughly put, entropy of a signaling strategy
+captures inverse distance from a pure strategy. The usual definition of entropy applies
+directly to mixed strategies (see Footnote~\ref{fn:mixedStrats}), but provably equivalent
+metrics for behavioral strategies are ready to hand:
 \begin{align*}
@@
   \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
-  \state)) \\
+  \state))\,, \\
   E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
@@
 communicative success. What is more, in line with our hypothesis that mere imprecision in
-imitation behavior can lead to behavior that looks like agents generalize across similar
+imitation behavior can lead to behavior that looks as if agents generalize across similar
 stimuli, higher imprecision led to a higher number of outcomes with convex sender
@@
 
-Beyond promoting convexity and convergence, diffusion also has another interesting
-regularizing effect on the evolution of signaling. There was very
-little variation in the recorded metrics for evolved strategies, at
-least for higher values of imprecision. On closer inspection, it turns
-out that variability in low-imprecision conditions is not only due to
-non-convergence or non-convexity. Figure~\ref{fig:MoreExample_strats}
-gives two more examples of strategy pairs at stopping time. Both are
-obtained for the same triple of parameters, both converged before the
-maximum number of rounds, and both have
-convex sender strategies. However, they are not equally efficient. In
-fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized
-expected utility of $0.99$ while the one in
-Figure~\ref{fig:example_stratsD} only has $0.89$.
+Beyond promoting convexity and convergence, diffusion also has another interesting regularizing
+effect on the evolution of signaling. There is very little variation in the recorded metrics
+for evolved strategies, at least for higher values of imprecision. On closer inspection, it
+turns out that variability in low-imprecision conditions is not only due to non-convergence or
+non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more examples of strategy pairs at
+stopping time. Both are obtained for the same triple of parameters, both converged before the
+maximum number of rounds, and both have convex sender strategies. However, they are not equally
+efficient. In fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized expected
+utility of $0.99$ while the one in Figure~\ref{fig:example_stratsD} only has $0.89$.
 
@@
 Our imprecise imitation dynamic leads to by-and-large successful signaling behavior, even in
-the presence of noise, that show the hallmarks of vagueness as desired. It also gives rise to
+the presence of noise, that shows the hallmarks of vagueness as desired. It also gives rise to
 population-level behavior that looks as if agents are generalizing across similar
@@
 strategies because of imprecision. This is, we believe, a plausible mechanism that can already
-explain the existence of vagueness in a language even if we assume each agent commands a
+explain the existence of vagueness in a language even if we assume that each agent commands a
 non-vague idiolect.
@@
 behavior must be spelled out rigorously and whether the resulting population-level dynamic
-would be equivalent to our present proposal in all respects.
+would be equivalent to our present proposal in all relevant respects.
 
@@
 \subsection{Evolutionary benefits of imprecision}
-The inclusion of confusability of similar states has noteworthy effects on the evolving meaning of signals.
-It transpired from our results that imprecision can have further accelerating and, surprisingly, unifying effects on meaning evolution.
-The unifying property of perceptual imprecision could be considered an evolutionary beneficial
-side-effect. A certain degree of imprecision can lead to higher \emph{within group expected utility},
-defined as the average expected utility that each evolved language scored when playing against
-an arbitrary other language obtained for the same parameter
-values. Figure~\ref{fig:GroupMeasuresB} gives a representative example. The observation repeats
-for other parameter values: while imprecision might decrease the communicative efficiency of
-individual languages, it increases the conceptual coherence and communicative success between
-independently evolving strategies. It is as if mere confusability of states imposes a
-regularity constraint on evolving categories.
+The inclusion of confusability of similar states has noteworthy effects on the evolving meaning
+of signals.  It transpired from our results that imprecision can have further accelerating and,
+surprisingly, unifying effects on meaning evolution.  The unifying property of perceptual
+imprecision could be considered an evolutionarily beneficial side-effect. A certain degree of
+imprecision can lead to higher \emph{within group expected utility}, defined as the average
+expected utility that each evolved language scored when playing against an arbitrary other
+language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a
+representative example. The observation repeats for other parameter values: while imprecision
+might decrease the communicative efficiency of individual languages, it increases the
+conceptual coherence and communicative success between independently evolving strategies. It is
+as if mere confusability of states imposes a regularity constraint on evolving categories.
 
@@
 \subsection{Related work}
-\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of reinforcement
-learning for sim-max games, in which successful play leads to reinforcement of choice options
-also for states similar to the ones that actually occurred. This not only leads to vague
-signaling of the appropriate kind, but also speeds up learning in such a way that, especially
-for sim-max games with higher numbers of states, higher levels of communicative success are
-reached in shorter learning periods. Our results complement and extend
+\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of
+reinforcement learning for sim-max games, in which successful play leads to reinforcement of
+choice options also for states similar to the ones that actually occurred. This not only leads
+to vague signaling of the appropriate kind, but also speeds up learning in such a way that,
+especially for sim-max games with higher numbers of states, higher levels of communicative
+success are reached in shorter learning periods. Our results complement and extend
 \citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
@@
 implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
-State confusability has an effect on
-aggregate signaling behavior that can be described as generalization without generalizers: the
-dynamics of imprecise imitation look as if ``conclusions'' about what works for one state are
-``carried over'' to similar states. This, however, is merely an epiphenomenon in the sense no
-single agent genuinely generalizes over stimuli or reasons about what a more systematic
-signaling strategy would be.
+State confusability has an effect on aggregate signaling behavior that can be described as
+generalization without generalizers: the dynamics of imprecise imitation look as if
+``conclusions'' about what works for one state are ``carried over'' to similar states. This,
+however, is merely an epiphenomenon in the sense that no single agent genuinely generalizes
+over stimuli or reasons about what a more systematic signaling strategy would be.
 
@@
 Section~\ref{sec:repl-diff-dynam}. We first show, in Section~\ref{sec:cond-imit-succ}, how to
-derive the standard replicator dynamic from noise-free imitation of success. The, in
+derive the standard replicator dynamic from noise-free imitation of success. Then, in
 Section~\ref{sec:imit-succ-with}, we derive the imprecise imitation dynamic by a parallel
@@
 \end{align*}
-This latter formulation is the continuous-time formulation of the replicator
-dynamic. We obtain the discrete-time formulation from it by assuming that discrete update steps
-are infinitesimally small, so that:
+This latter formulation is the continuous-time version of the replicator dynamic. We obtain a
+discrete-time formulation from it by assuming that discrete update steps are infinitesimally
+small, so that:
  \begin{align*}
@@
   P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
-  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
-  \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
-  \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg) \,.
+  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) % = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
+  % \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
+  % \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg)
+  \,.
 \end{align*}
@@
 which is the discrete-time formulation of the imprecise imitation dynamic given in
-Section~\ref{sec:repl-diff-dynam} in Equation~(\ref{eq:nRD}). To see how the discrete-time
-formulation gives rise to the continuous-time formulations above, let's assume that update
-steps are infinitesimally small, so that, for the sender case:
+Equation~(\ref{eq:nRD}). To see how the discrete-time formulation gives rise to the
+continuous-time formulations above, let's assume that update steps are infinitesimally small,
+so that, for the sender case:
 \begin{align*}

### 2015-10-23 paper/version_04/title_page.tex
Title page for submission to BPoS.

--- 2015-10-15 paper/version_04/paper.tex
@@
-\documentclass[fleqn,reqno,10pt]{article}
+\documentclass{article}
+\usepackage[utf8]{inputenc}
 
-%========================================
-% Packages
-%========================================
+\title{Vagueness and \emph{as-if}-generalization from imprecise imitation in signaling games}
 
-\usepackage[]{../../helpers/mypackages}
-%\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
-%\bibliography{MyRefGlobal}
-\bibliography{../../helpers/MyRefGlobal}
-\bibliography{paper} 
-\usepackage{../../helpers/myenvironments}
-\usepackage{../../helpers/mycommands}
-\usepackage{todonotes}
-\usepackage{subcaption}
+\author{
+Michael Franke%
+\thanks{Department of General Linguistics, Faculty of Humanities, University of Tübingen}\\
+\texttt{mchfranke@gmail.com}
+\and
+Jos\'e Pedro Correia\\
+\texttt{zepedro.correia@gmail.com}
+}
 
-\usepackage{a4wide}
-
-%========================================
-% Standard Layout
-%========================================
-
-% Itemize
-\renewcommand{\labelitemi}{\large{$\mathbf{\cdot}$}}    % itemize symbols
-\renewcommand{\labelitemii}{\large{$\mathbf{\cdot}$}}
-\renewcommand{\labelitemiii}{\large{$\mathbf{\cdot}$}}
-\renewcommand{\labelitemiv}{\large{$\mathbf{\cdot}$}}
-% Description
-\renewcommand{\descriptionlabel}[1]{\hspace\labelsep\textsc{#1}}
-
-% Figure Captions
-\usepackage{caption} % use corresponding myfiguresize!
-\setlength{\captionmargin}{20pt}
-\renewcommand{\captionfont}{\small}
-\setlength{\belowcaptionskip}{7pt} % standard is 0pt
-
-%========================================
-% Additional layout & commands
-%========================================
-
-\renewcommand{\Smixed}{\ensuremath{\mathrm{\mathbf{s}}}}
-\renewcommand{\Rmixed}{\ensuremath{\mathrm{\mathbf{r}}}}
-
-% Annotations
-\newcommand{\mytodo}[2]{\todo[inline,color=yellow,author=#1]{#2}}
-\newcommand{\question}[2]{\todo[inline,color=blue,author=#1]{#2}}
-\newcommand{\answer}[2]{\todo[inline,color=green,author=#1]{#2}}
-
-\newcommand{\rd}{\acro{rd}} % replicator dynamic
-\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
-\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
-\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
-\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
-\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator
-\newcommand{\icid}{\acro{icid}} % imprecise conditional imitation dynamic
-\newcommand{\ICID}{\ensuremath{\mathrm{ICID}}} % imprecise conditional imitation dynamic
-
-\newcommand{\Diff}{\ensuremath{\mathrm{D}}} % Difusion 
-\newcommand{\Mutate}{\ensuremath{\mathrm{M}}} % Mutation
-
-\newcommand{\imprecision}{\ensuremath{\alpha}} % imprecision
-\newcommand{\toler}{\ensuremath{\beta}} % tolerance
-\newcommand{\ns}{\ensuremath{n_s}} % number of states
-
-\newcommand{\similarity}{\ensuremath{\mathrm{Sim}}} % similarity function
-
-\newcommand{\myred}[1]{\textcolor{red}{#1}} 
-\renewcommand{\Pr}{\ensuremath{P}}
-
-% \doublespacing
-
-\title{Vagueness, Noise, and Signaling}
-% \author{Michael Franke \and Jos\'e Pedro Correia}
 \date{}
@@
 
-\section{Introduction}
-\label{sec:introduction}
-
-Many concepts and expressions are vague. A vague category knows clear
-cases that fall under it, clear cases that do not, and also so-called
-borderline cases. Borderline cases do not clearly apply, nor do they
-clearly not apply, and there may be differences between borderline
-cases in terms of how well they represent the category in
-question. Vagueness does not seem to dramatically affect the success
-of everyday communication, but it is troublesome for some of the most
-prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell and the young
-Wittgenstein which is challenged by the paradoxes vagueness gives rise
-to.
-
-But there are other intriguing aspects about vagueness. The puzzle
-that we are concerned with here is how vagueness could arise and be
-maintained in the first place. This is a serious issue for
-functionalist accounts that maintain that concepts and linguistic
-meanings evolved driven towards efficiency. Since the existence of
-unclear borderline cases seems to entail inefficiency of
-categorization or communication, the challenge, succinctly put forward
-by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
-persist under evolutionary pressure to be optimally discriminative. We
-address \emph{Lipman's problem} as a technical challenge, namely to
-find a conceptually sound and mathematically coherent formal model of
-the evolution of categorization and language use that yields
-by-and-large efficient categorization \emph{and} vagueness.
-
-A number of authors have recently tried to explain why vagueness evolved as something that is
-itself useful
-\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
-Others have argued that vagueness is a natural byproduct of limitations in information
-processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or the result of particular
-low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
-contributes to the latter line of thought. Concretely, we introduce a dynamic that describes
-changes in aggregate population behavior as the result of imitation with probabilistic
-confusability of similar stimuli. We show that stochastic noise can not only lead to vague, yet
-relatively efficient signal use, but can also unify evolutionary outcomes and help avoid
-suboptimal categorization: regularities in confusability of stimuli also regularize signaling
-behavior and evolving categories. The system gives rise to generalization behavior without
-agents actually generalizing.
-
-The next section introduces the background against which the work presented here can be
-appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
-dynamic which is derived from the idea that agents imitate other agents' behavior while
-possibly confusing similar states. This imprecise imitation dynamic is explored in
-Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
-approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
-
-\section{Background}
-\label{sec:background}
-
-% The view that vagueness is a natural concomitant of cognitive
-% limitations of language users has been formalized in a number of ways,
-% using evolutionary game theory and certain generalizations of Lewis'
-% signaling games \citep{Lewis_1969:Convention}, so called
-% \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
-% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
-% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
-% Our contribution is best seen in relation to these accounts, as it also
-% relies on sim-max games. Let's introduce these first, and then zoom in
-% on the problem of vagueness.
-
-\subsection{Sim-max games \& conceptual spaces}
-
-Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
-a sender and a receiver. The sender knows the true state of the world,
-but the receiver does not. The sender can select a signal, or message,
-to reveal to the receiver, who then chooses an act. In Lewis' games,
-if the receiver chooses the act that corresponds to the actual state,
-the play is a success, otherwise a failure. Certain regular
-combinations of sender signaling and receiver reaction make messages
-meaningful, in the sense that their use is correlated systematically
-to certain states or acts. To investigate the conditions under which
-such meaning-generating behavior can evolve is a highly interesting
-topic that we are only beginning to fully understand
-\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.
-
-Similarity-maximizing (short: sim-max) games are a variation of
-Lewis' games where the receiver's actions are equated with the state
-space (one can think of the actions as choosing states) and different states are allowed to be more or less
-similar to one another. While Lewis' games treated communicative
-success as a matter of black and white, sim-max games allow for a
-gradient notion: the more similar the receiver's interpretation is to
-the actual state, the better. Signaling games with utility-relevant
-similarities in the state space are fairly standard in economics
-\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
-but have received particular attention in a more philosophical context
-for reasons that will become clear presently.
-
-Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
-fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
-$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
-  metric is a function of distance between any two points in a given space. It should satisfy
-  certain axioms that ensure behavior that one would intuitively expect from the words
-  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
-  this paper.}
-  % : similarity should always be non-negative, zero only for the distance
-  % between an element and itself, symmetric, and satisfy the triangle inequality that
-  % $\similarity(x,z) \le \similarity(x,y) + \similarity(y,z)$
-on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
-$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
-and $\mystate{2}$, and a utility function
-$\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
-$\Utils(\mystate{1}, \mystate{2})$ is the payoff for sender and receiver for a play with actual
-state $\mystate{1}$ and receiver interpretation $\mystate{2}$. We identify the receiver's acts
-with the states of the world, so that the game is one of guessing the actual state, so to
-speak. The utility function should be a monotonically increasing function of similarity.
-Non-probabilistic sender behavior can be represented by a pure strategy
-$\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
-each state. Similarly, a pure receiver strategy is a function $\Rpure \in \States^\Messgs$.
-
-\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of
-sim-max games are remarkably systematic: the evolutionarily stable states are demonstrably
-so-called Voronoi languages.\footnote{Their results were obtained for games with infinitely
-  many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a
-  quadratic loss function for utilities
-  $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$.} Roughly put, a Voronoi
-language is a pair of sender and receiver strategies, such that the sender strategy partitions
-the state space into convex categories, while the receiver's interpretations are the central
-spots in each category. A subset $X$ of $\mathds{R}^n$ is convex if, informally put, all points
-in $X$ are connected via a straight line that lies entirely in $X$; $X$ has no gaps or
-dents. For example, if \States is the unit interval and all states are equiprobable, a Voronoi
-language with two messages could have the sender use one message exclusively for all points in
-the lower half of the unit interval and another for all points in the upper half; the
-receiver's interpretations of messages are the central points, .25 and .75, in the respective
-intervals.  See Figure~\ref{fig:Voronoi-language-example} for an illustration.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[scale=3]
-
-        \begin{scope}
-
-          \draw[-,thick] (0,0) -- (1.5,0);
-
-          \draw (0,0) node {\color{red}{\big{[}}};
-
-          \draw (.75,0) node {\color{red}{\big{)}}};
-
-          \draw (.76,0) node {\color{red}{\big{(}}};
-
-          \draw (1.5,0) node {\color{red}{\big{]}}};
-
-            \draw (.325,.15) node
-            {\color{red}{$\Spure(\mymessg{1}  \probbar \cdot)$}};
-
-            \draw (1.075,.15) node
-            {\color{red}{$\Spure(\mymessg{2} \probbar \cdot)$}};
-
-            \node at (.375,0) (node1) {\color{blue}{$\bullet$}};
-
-            \node at (.375,-.3) (label1)
-            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{1})$}};
-
-            \draw[-,color=blue] (node1) -- (label1);
-
-            \node at (1.125,0) (node2) {\color{blue}{$\bullet$}};
-
-            \node at (1.125,-.3) (label2)
-            {\color{blue}{${\Rpure}(\cdot \probbar \mymessg{2})$}};
-
-            \draw[-] (node2) -- (label2);
-
-            \draw (0,-.2) node {$0$};
-
-            \draw (.75,-.2) node {$0.5$};
-
-            \draw (1.5,-.2) node {$1$};
-
-          \end{scope}
-
-        \end{tikzpicture}
-
-        \caption{Example of a Voronoi language on $\States = [0,1]$. The pure sender strategy
-          $\Spure$ uses one signal for the lower half, and another for the upper half of the
-          unit interval. The pure receiver strategy $\Rpure$ selects the central elements in
-          the respective intervals. See \citet{JagerMetzger2011:Voronoi-Languag} for further
-          details.}
-  \label{fig:Voronoi-language-example}
-\end{figure}
-
-This result is interesting because it demonstrates that signaling can
-impose orderly categories on a metric space, without that being the
-ulterior purpose of it all. Sender and receiver can be distinct
-entities, whose purpose is to communicate effectively about the actual
-state. In that case, evolving Voronoi languages would explain why
-\emph{linguistic categories} are well-behaved and orderly in the way
-they appear to be. More abstractly, sender and receiver can also be
-thought of as distinct modules in a single system, where the first
-module must discretize the information it is fed by selecting a small
-sample of, suggestively, category labels. These are passed to a second
-module that tries to decode the original information. In this case,
-evolving Voronoi languages would explain why \emph{conceptual
-  categories} are well-behaved and orderly in the way that they appear
-to be.
-
-Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
-that rely on the notion of conceptual spaces.
-\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
-natural categories are convex regions in conceptual space. If the conceptual space has a
-suitable metric, convex categories can be derived from a set of prototypes. The category
-corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
-other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
-system can be obtained: storing the prototypes lets us recover the categories without having to
-store each category's extension. However, what is left unexplained is where the prototypes
-come from, and why we would not see just any distribution of prototypes as an equally efficient
-classification system. This is where sim-max games can contribute a principled approach to
-deriving, in an independent way, not only convex categories but also prototypical exemplars
-belonging to them.  These ideas, and more, are developed further by
-\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
-among others.
-
-\subsection{Vague signaling in sim-max games}
-
-This outline of an approach to categorization using sim-max games
-leaves many problems unaddressed. One of them is that natural
-categories for continuously variable stimuli usually do not have
-unique, point-valued prototypes and clear category boundaries. We
-would like to account for the possibility of such vagueness, in
-particular: (i) clear positive examples of a vague category should
-show a gradient transition to clear negative examples;
-(ii) prototypes should likewise be gradient regions, peaking at the
-center of the vague category they represent.
-
-\citet{DouvenDecock2011:Vagueness:-A-Co} show that
-\citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
-approach can be extended to account for the existence of borderline
-cases. From the assumption that prototypes are extended, yet convex
-regions in conceptual space, a construction algorithm is available
-that yields ``collated Voronoi diagrams'' with thick boundaries
-representing borderline
-regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
-is possible to arrive at a gradient transition between categories, by
-weighing in the distance of different borderline cases to various
-prototypical regions. This accounts for the first of the two
-desiderata mentioned above, but still assumes that crisp prototype
-regions must be given.
-
-Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
-\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be met by evolving
-strategies in sim-max games. The approach we take here is similar in spirit, but different in
-relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
-us first look into our own proposal in more detail.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Imprecise imitation}
-\label{sec:repl-diff-dynam}
-
-Signaling agents can adapt their dispositions to act, given some feedback about their past
-success, in multiple ways. Usually, we would assume that changes in behavior should, at least
-on average, lean towards increasing chances of communicative success. Such behavioral
-adaptations can be described at different levels of abstraction. At the level of individual
-agents, we can picture a more or less idealized process of how each agent adapts dispositions
-for future actions based on various pieces of information available to the agent. More
-abstractly, at the level of a population of agents, we can describe how average behavioral
-dispositions will evolve. The population-level perspective abstracts over small stochastic
-fluctuations and zooms in on the general tendency or direction of evolution that ensues from
-behavior at the agent-level.
-
-To better understand the interaction of confusability of stimuli and selective pressure towards
-successful communication, we look at a population-level dynamic that describes the most likely
-evolutionary path of a population of signaling agents who imitate other agents' behavior but
-are liable to confuse states for one another. In the special limiting case where state
-confusability vanishes, the process is just the well-known replicator dynamic
-\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.
-
-\subsection{Replicator dynamic in behavioral strategies}
-\label{sec:repl-dynam-behav}
-
-Fix a sim-max game with finite states $\States$ and messages $\Messgs$. As usual, we assume
-that the receiver chooses states in $\States$ in response to messages. Let
-$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
-$\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
-senders and receivers in the population. A \emph{behavioral strategy} is a function that maps
-an agent's choice points to a probability distribution over available
-choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioral strategies, not \emph{mixed
-    strategies}, i.e., probability distributions over functions from each choice point to an
-  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume
-  that agents can adapt their behavior locally, i.e., independently at each choice point. Our
-  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
-  numerical simulations. But it also seems the more plausible choice for imitation-based update
-  protocols of the kind we consider here: agents only observe how, on some occasion, some other
-  agent behaved in \emph{one} particular situation, not how that agent would behave in
-  \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
-  not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
-  between dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
-functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
-to a probability of each message $\messg \in \Messgs$ being sent in $\state$; the receiver's
-are functions $\Rstrat \in \Delta(\States)^\Messgs$, thus mapping each message
-$\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
-in response to $\messg$. Although behavioral strategies are probabilistic, evolutionary
-modeling usually imagines that every individual agent has a non-probabilistic
-strategy. Behavioral strategies then capture average population behavior. Assuming a virtually
-infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
-probability that a randomly sampled sender would send message $\messg$ if the actual state was
-$\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
-randomly sampled receiver interprets $\messg$ as $\state$. A play of single evolutionary game
-with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$\Pr(\mystate{1})$] {$\mystate{1} \in \States$};
-  \node (3) [right=of 2,label=below:$\Sstrat(\messg \probbar \mystate{1})$] {$\messg \in \Messgs$};
-  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{2} \probbar \messg)$] {$\mystate{2} \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {S} (3)
-    (3) edge node {R} (4);
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a
-    state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a random sender (S) selects a
-    message $\messg$ with probability $\Sstrat(\messg \probbar \mystate{1})$; a random receiver
-    (R) selects a state $\mystate{2}$ with probability $\Rstrat(\mystate{2} \probbar \messg)$.
-    Payoff for both sender and receiver is given by $\Utils(\mystate{1}, \mystate{2})$.}
-   %which is typically proportional to state similarity $\similarity(\mystate{1}, \mystate{2})$.}
-  \label{fig:sim-max-illustration}
-\end{figure}
-
-The \emph{expected utility} of choices at each choice point is:\footnote{The notation
-  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
-  side might still need to be normalized so as to have probabilities sum to one. In general,
-  writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
-  shorthand for $P(x) = \frac{f(x)}{\sum_{x' \in X} f(x')}$. }
-\begin{align*}
-  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
-  \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
-  \EU(\state', \messg, \Sstrat) & = \sum_{\state \in
-    \States} P(\state \probbar \messg)  \myts
-  U(\state, \state')\,, \text{\ \  where } &
-P(\state \probbar \messg) & \propto \Pr(\state) \myts \Sstrat(\messg \probbar \state) \,.
-\end{align*}
-Expected utilities are sums, over all possible concrete outcomes, of the utilities of these
-outcomes, weighted by how likely they occur. For instance, the expected utility for the sender
-of choosing message $\messg$, given state $\state$ and a receiver strategy $\Rstrat$ is the
-sum, for each interpretation $\state'$, of the probability that some random receiver will
-choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
-$\state'$.
-
-The discrete-time replicator dynamic tracks changes in frequency of choices in the population
-as proportional to their expected utilities
-\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
-  Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
-  in Section~\ref{sec:exploring-rdd}, where utilities are always non-negative and expected
-  utilities are always positive.}
-\begin{align}
-  \label{eq:RDDiscrete}
-  \Sstrat'(\messg \probbar \state) & \propto \Sstrat(\messg \probbar \state) \myts
-    \EU(\messg, \state,\Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto \Rstrat(\state \probbar \messg) \myts
-    \EU(\state, \messg,\Sstrat)\,.   
-\end{align}
-For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
-average agent in the population \emph{after} the update is proportional to the probability of
-seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
-expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
-by a gradient of current frequency and a measure of how good they are.
-
-The replicator dynamic is an abstract population-level dynamic that describes the mean expected
-change of behavioral dispositions in a population of signalers. There are several ways of
-deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
-here on one of the simplest: \emph{imitation of success}
-\citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
-Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
-update scheme).  Every now and then a random agent gets a chance to alter his behavior for one
-of his choice points. Say, this is a sender who gets to ``reconsider'' his choice of message
-for state $\state$. The revising agent then observes what some random agent does at $\state$,
-say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
-expected utility of $\messg$ for state $\state$.
-
-\subsection{Noisy replicator dynamic}
-\label{sec:noisy-repl-dynam}
-
-Imitation of success, as described above, presupposes that agents make no mistakes when
-observing states, or choosing interpretations. This may not always be an appropriate
-assumption, especially when some states can be perceptually similar and therefore likely to be
-confused for one another. Therefore, imitation dynamics could be affected by at least two
-sources of probabilistic noise:
-
-\begin{enumerate}
-\item \emph{observation noise:} whenever a state $\mystate{a}$ actually occurs, the probability
-  that an agent observes it as $\mystate{o}$ is $P_o(\mystate{o} \probbar \mystate{a})$;
-\item \emph{realization noise:} whenever an agent intends to realize interpretation
-  $\mystate{i}$, the probability that $\mystate{r}$ is realized is
-  $P_r(\mystate{r} \probbar \mystate{o})$.
-\end{enumerate}
-
-\noindent A round of play of a sim-max game with these two sources of confusability of states
-is pictured in Figure~\ref{fig:imprecise-sim-max-illustration}. Note that, and this is crucial,
-senders respond to observations not actual states, receiver strategies determine intentions not
-realized states, but payoff is calculated based on actual and realized states. Behavioral
-strategies thus encode what agents actually do from their subjective point of view; or, put
-differently, what they would do in a noise-free world. The noise-free situation depicted in
-Figure~\ref{fig:sim-max-illustration} is then the special case where
-$P_o(\mystate{o} \probbar \mystate{a}) = 1$ iff $\mystate{a} = \mystate{o}$ and
-$P_r(\mystate{r} \probbar \mystate{i}) = 1$ iff $\mystate{i} = \mystate{r}$.
-
-\begin{figure}
-  \centering
-
-  \begin{tikzpicture}[->,auto,semithick,>=stealth',every label/.style={gray},node distance=0.9cm]
-
-  \node (1) {};
-  \node (2) [right=of 1,label=below:$P_a(\mystate{a})$] {$\mystate{a} \in \States$};
-  \node (2') [right=of 2,label=below:$P_o(\mystate{o} \probbar \mystate{a})$] {$\mystate{o} \in \States$};
-  \node (3) [right=of 2',label=below:$\Sstrat(\messg \probbar \mystate{o})$] {$\messg \in \Messgs$};
-  \node (4) [right=of 3,label=below:$\Rstrat(\mystate{i} \probbar \messg)$] {$\mystate{i} \in \States$};
-  \node (4') [right=of 4,label=below:$P_r(\mystate{r} \probbar \mystate{i})$] {$\mystate{r} \in \States$};
-
-  \path
-    (1) edge node {N} (2)
-    (2) edge node {} (2')
-    (2') edge node {S} (3)
-    (3) edge node {R} (4)
-    (4) edge node {} (4');
-
-  \end{tikzpicture}
-
-  \caption{A round of play in a sim-max game with probabilistic confusability of states: Nature
-    (N) chooses an actual state with probability $P_a(\mystate{a})$; a randomly sampled sender
-    (S) observes $\mystate{o}$ with probability $P_o(\mystate{o} \probbar \mystate{a})$ and
-    subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a
-    randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with
-    probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation
-    $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$. Payoff for both
-    sender and receiver is given by $\Utils(\mystate{a}, \mystate{r})$.}
-  %which is typically proportional to state similarity $\similarity(\mystate{a}, \mystate{r})$.}
-  \label{fig:imprecise-sim-max-illustration}
-\end{figure}
-
-The presence of observation and realization noise also affects imitation of successes. Here, we
-focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:imit-succ-with}. Take a
-sender who gets to revise behavior for state $\state$. What agents can plausibly revise by
-imitation is their pure strategy, which maps perceived states onto messages. When an agent gets
-to revise his strategy for the perceived state $\state$, this need not necessarily be the
-actual state. Moreover, when that agent observes $\state$, another agent may perceive yet a
-different state. Given what that latter agent perceives, his actual (pure) strategy will
-determine what he plays. In sum, to describe how likely the potential imitator observes a
-message choice $\messg$, we are interested in the conditional probability
-$P_o(\messg \probbar \state)$ that some other random agent selects $\messg$ when the first
-agent perceives $\state$. This $P_o(\messg \probbar \state)$ is derived from the prior
-probability of states, the current sender population behavior $\Sstrat$ and the given
-observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
-$\messg$ as his choice for $\state$ with probability given by the expected utility of sending
-$\messg$ when perceiving state $\state$. Expected utility should, of course, take the
-probabilistic confusability of states into account as well.
-
-Similar considerations apply to the receiver side. If an agent gets a chance to change his
-intended interpretation of $\messg$, we need to look at the conditional probability
-$P_o(\state \probbar \messg)$ of observing another agent realize interpretation $\state$ given
-that the first agent (and therefore the second as well) perceived $\messg$. This depends on
-observation and realization noise, as well as on the current receiver population behavior
-$\Rstrat$.
-
-Appendix~\ref{sec:imit-succ-with} shows how imprecise imitation of this sort leads to mean changes in population
-frequencies of choices that can be covered by the following discrete-time formulation:
-\begin{align}
-  \label{eq:nRD}
-  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
-  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
-  \myts \EU(\state, \messg, \Sstrat)\,.
-\end{align}
-\noindent This looks very much like the discrete-time formulation of the standard replicator
-dynamic in (\ref{eq:RDDiscrete}), but there are, of course, the aforementioned
-differences. Firstly, expected utility here takes stochastic confusability of states into
-account. Secondly, where the standard replicator dynamic had probabilities
-$\Sstrat(\messg \probbar \state)$ and $\Rstrat(\state \probbar \messg)$, we now have
-$P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. If there is no
-observation or realization noise, $P_o(\messg \probbar \state)$ reduces to
-$\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
-$\Rstrat(\state \probbar \messg)$. The noisy replicator dynamic in (\ref{eq:nRD})
-conservatively extends the classic case in (\ref{eq:RDDiscrete}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Exploring noisy imitation}
-\label{sec:exploring-rdd}
-
-How does a tendency to confuse similar states interact with selective pressure towards more
-efficient signaling strategies under the imprecise conditional imitation dynamic? If confusion
-probabilities are moderate and regular, in that they track similarity of states, we might
-expect a regularizing effect also on evolving signaling strategies. Indeed, we hypothesize that
-confusion of states can give rise to population-level aggregate behavior that looks \emph{as
-  if} information about what is good for one state percolates also to similar states. In other
-words, imprecise imitation may make signaling behavior on average look like agents generalize
-across similar states, even if no agent actually generalizes. To explore whether confusion of
-states can have this effect, we turn to numerical simulation.
-
-\subsection{Setting the stage}
-\label{sec:setting-stage}
-
-To obtain concrete results, we must fix how to represent states, similarity of states, the
-conditional confusion probabilities $P_o$ and $P_r$ and the utilities of our sim-max
-games. Confusion probabilities between states should be a function of perceptual similarity:
-the more similar two states are, the more likely they could be mistaken for each other.
-
-Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
-interval, including $0$ and $1$. All states occur, for simplicity, with the same probability
-(i.e., $P_a$ is uniform).  The distance $\card{\mystate{i} - \mystate{j}}$ is the objective,
-physical similarity between two states $\mystate{i}$ and $\mystate{j}$. Distance in physical
-space feeds into a perceptual similarity function, as described by
-\citet{Nosofsky1986:Attention-Simil}:
-\begin{align*}
-  \similarity(\mystate{i}, \mystate{j} \myts ; \myts \imprecision) =
-      \begin{cases}
-    1 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} = \mystate{j} \\
-    0 & \textrm{if } \imprecision = 0 \textrm{ and } \mystate{i} \neq \mystate{j} \\
- \expo \left ( -  \frac{\card{\mystate{i} - \mystate{j}}^2}{ \imprecision^2} \right ) & \textrm{otherwise,} \\
-    \end{cases}
-\end{align*}
-where $\imprecision \ge 0$ is an imprecision parameter. When $\imprecision=0$ agents perfectly
-discriminate between states; when $\imprecision \rightarrow \infty$ agents cannot discriminate
-states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
-different parameter values. Other formalizations of perceptual similarity are possible,
-including ones that allow for different discriminability in different areas of the state space,
-but we stick with Nosofsky's similarity function for the time being, because it is
-mathematically simple, yet an established notion in mathematical psychology.
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=0.7\textwidth]{plots/NosofskySim.pdf}
-
-  \caption{Examples of Nosofsky-similarity for different values of
-    imprecision.}
-  \label{fig:NosofskySim}
-\end{figure}
-
-We further assume that the probability of confusing any two states $\mystate{i}$ and
-$\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
-observation noise $P_o$ is simply the same as realizational confusability $P_r$ and
-that both are governed by the same imprecision parameter $\alpha$:
-\begin{align*}
-  P_o(\mystate{o} \probbar \mystate{a}) & \propto \similarity(\mystate{o}, \mystate{a} \myts ; \myts
-  \imprecision)\,, &   P_r(\mystate{r} \probbar \mystate{i}) & \propto \similarity(\mystate{r}, \mystate{i} \myts ; \myts
-  \imprecision) \,.
-\end{align*}
-For $\imprecision = 0$, we obtain trivial confusion probabilities: everything is reduced to
-perfect imitation and the replicator dynamic. For $\imprecision > 0$, any state can be confused
-for any other state with some positive probability. For the uninteresting case of
-$\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
-realized as any other state with probability $\nicefrac{1}{\card{\States}}$.
-
-As for utility, we define it in terms of similarity and introduce another free parameter, $\toler \ge 0$.
- The intention is for this parameter to model the amount of tolerable pragmatic slack, which we felt should be allowed to vary separately from perceptual imprecision:
-\begin{align*}
-  \util(\mystate{i}, \mystate{j} \myts ; \myts \toler) =
-      \similarity(\mystate{i}, \mystate{j} \myts ; \myts \toler)\,.
-\end{align*}
-This choice of utility function is governed partly by convenience, but also because we believe it has the right general properties for a communicative payoff function.
-Unlike utilities that are, say, linearly or
-quadratically decreasing in physical distance
-\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
-that are exponentially decreasing in negative quadratic distance can model situations where a
-small amount of imprecision in communication is tolerable, whereas similarly small differences
-in intolerably far away interpretations matter very little, with a smooth transition between
-these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.
-
-In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
-is governed by three parameters: the number $\ns = \card{T}s$ of states in the state space,
-imprecision $\imprecision$ and tolerance $\toler$.
-
-\subsection{Experimental set-up}
-\label{sec:simulations}
-
-We ran 50 trials of the discrete-time dynamic in (\ref{eq:nRD}), starting with randomly sampled
-sender and receiver strategies, for each triplet of independent parameter values:
-$\ns \in \set{6, 10, 50, 90}$, $\imprecision \in \set{0, 0.05, 0.1, 0.2, 0.3}$,
-$\toler \in \set{0.05, 0.1, 0.2, 0.3}$. Each trial ran for a maximum of 200 update steps. A
-trial was considered converged, and thus stopped before the maximum of 200 rounds, if the total
-amount of change between strategies before and after an update step was smaller than a suitably
-chosen threshold. It is not guaranteed that strategies at halting time had converged to the
-eventual attracting state, whether they ran for 200 rounds or not. Our notion of convergence is
-therefore only a categorical measure for reaching a certain (well-considered, but eventually
-arbitrary) degree of stability. In other words, our notion of ``convergence'' is a measure of
-relative speed: is it true that the system reached a state in which evolutionary adaptations
-had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
-concerns regarding length of simulation time, but also theoretically justified, because we
-hypothesize that confusability of states leads to regularization of evolving strategies, which
-would show exactly in an increased speed of evolutionary trajectories towards well-behaved and
-regular signaling behavior.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
-    \label{fig:example_stratsA}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
-    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.05$}
-    \label{fig:example_stratsB}
-  \end{subfigure}
-
-  % \begin{subfigure}[]{0.45\textwidth}
-  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp0_ind1001.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0$}
-  %   \label{fig:example_stratsC}
-  % \end{subfigure}
-  % \hfill
-  % \begin{subfigure}[]{0.45\textwidth}
-  %   \includegraphics[width=\textwidth]{plots/strat_example_NS-10_tol-005_imp005_ind1201.pdf}
-  %   \caption{$\ns = 10$, $\toler = 0.1$, $\imprecision = 0.05$}
-  %   \label{fig:example_stratsD}
-  % \end{subfigure}
-
-  \caption{Example strategies under \rdd at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
-  \label{fig:example_strats}
-\end{figure}
-
-Representative examples for resulting strategy pairs are given in
-Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
-stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
-imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
-the given sender strategy approximates a pure sender strategy that crisply partitions the state
-space into non-convex sets. The irregular shape of the receiver strategy suggests that the
-pictured strategy pair has not yet reached a stable state. Indeed,
-the trial was stopped when reaching the maximum of 200 rounds. In contrast, the outcome of a
-trial with identical parameters, except with imprecision $\imprecision = 0.05$, which is shown
-in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
-rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
-receiver's interpretations are rather extended curves, peaking at a central point in the
-relevant ``categories.''
-
-These examples already show two interesting things. Firstly, inclusion of imprecision can lead
-to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
-strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
-for each signal, with a smooth transition in-between. The receiver's interpretations of signals
-can be seen as smoothed-out prototype regions. Secondly, (sender) strategies can approach
-non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
-time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
-Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
-for the limiting case. This was first observed by Elliott Wagner, as mentioned by
-\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
-scope of this paper, but we will shortly see that diffusion from confusability of states
-clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
-strategies.
-
-\subsection{Dependent measures}
-
-To further explore our simulation results, we calculated metrics that aim to numerically
-capture how vague, generally well-structured, and communicatively efficient the recorded
-strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
-signal use. \emph{Convexity} captures whether a behavioral strategy would project onto a convex
-pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
-strategy pairs.
-
-\paragraph{Entropy.} This classic information-theoretic notion captures the amount of
-uncertainty in a probability distribution. Roughly put, entropy of a signaling strategy
-captures inverse distance from a pure strategy. The usual definition of entropy applies
-directly to mixed strategies (see Footnote~\ref{fn:mixedStrats}), but provably equivalent
-metrics for behavioral strategies are ready to hand:
-\begin{align*}
-  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
-  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
-  \state))\,, \\
-  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
-  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
-  \messg)) \,. 
-\end{align*}
-Values obtained by these definitions are lower bounded by $0$ and
-upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
-\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
-\log(|\States|)$. We work with normalized values. The sender
-strategies in Figures~\ref{fig:example_stratsA} and
-\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
-respectively. The receiver strategies have respective entropies $0.43$
-and $0.81$. In general, we expected that vague languages would have
-higher entropy than crisp ones and that increasing imprecision would
-lead to increased entropy, all else being equal.
-
-\paragraph{Convexity.} At least for sender strategies, which develop faster than receiver
-strategies, it also makes sense to define a categorical measure of convexity that compensates
-for potential vagueness. To determine whether a sender strategy $\Sstrat$ is convex despite
-possibly being vague, we look at the derived pure strategy $\Spure$ for which
-$\Spure(\state) = \argmax_{\messg' \in \Messgs} \Sstrat(\state,\messg')$. If that $\Spure$ is
-convex, we also count $\Sstrat$ as convex. The sender strategy in
-Figure~\ref{fig:example_stratsA} is not convex, while the one in
-Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
-strategies, like we hypothesized, we should see more convexity with increasing impairment all
-else equal.
-
-\paragraph{Expected utility.} We also recorded the expected utility of
-a strategy pair:
-\begin{align*}
-  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
-    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
-  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
-  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
-\end{align*}
-To make direct comparisons across different parameter settings, we
-normalize expected utility by the maximal amount of expected utility
-obtainable in the relevant game. The strategy pair in
-Figure~\ref{fig:example_stratsA} has a normalized expected utility of
-$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
-$0.95$. Generally, vagueness and imprecision can be expected to
-decrease expected utility
-\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
-whether communicative success drops unacceptably fast with moderate
-levels of vagueness and imprecision.
-
-\subsection{Results}
-
-\begin{figure}[t]
-  \centering
-
-  \includegraphics[width=0.75\textwidth]{plots/MeanMetrics3.pdf}
-
-  \caption{Means of gradient and proportions of categorical measures
-    for $\toler = 0.1$, $\ns \in \set{6, 10, 50}$, and $\imprecision
-    \in \set{0, 0.05, 0.1, 0.2, 0.3}$. The plot shows the average of
-    the entropies for the sender and receiver strategy.}
-  \label{fig:MeanMetrics}
-\end{figure}
-
-Figure~\ref{fig:MeanMetrics} shows plots summarizing a selected part of our findings.  For
-perspicuity, we only plot results for one level of tolerance $\beta = 0.1$, and leave out the
-case of $n_s = 90$. Still, every qualitative trend mentioned in the following applies to the
-whole set of results.
-
-As expected, increasing imprecision leads to higher entropy and lower expected
-utility. Importantly, however, imprecision does not necessarily lead to disastrous decline of
-communicative success. What is more, in line with our hypothesis that mere imprecision in
-imitation behavior can lead to behavior that looks as if agents generalize across similar
-stimuli, higher imprecision led to a higher number of outcomes with convex sender
-strategies. It also led to higher rates of convergence. In fact, sufficient imprecision always
-ensured convergence and convexity. It appears that perceptual imprecision leads to more
-vagueness, slightly less communicative efficiency, but more regular, well-behaved languages in
-shorter time.
-
-Beyond promoting convexity and convergence, diffusion also has another interesting regularizing
-effect on the evolution of signaling. There is very little variation in the recorded metrics
-for evolved strategies, at least for higher values of imprecision. On closer inspection, it
-turns out that variability in low-imprecision conditions is not only due to non-convergence or
-non-convexity. Figure~\ref{fig:MoreExample_strats} gives two more examples of strategy pairs at
-stopping time. Both are obtained for the same triple of parameters, both converged before the
-maximum number of rounds, and both have convex sender strategies. However, they are not equally
-efficient. In fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized expected
-utility of $0.99$ while the one in Figure~\ref{fig:example_stratsD} only has $0.89$.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind21.pdf}
-    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
-    \label{fig:example_stratsC}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/strat_example_ind23.pdf}
-    \caption{$\ns = 6$, $\toler = 0.2$, $\imprecision = 0$}
-    \label{fig:example_stratsD}
-  \end{subfigure}
-
-  \caption{More example strategies under \rdd at stopping time of our simulations.}
-  \label{fig:MoreExample_strats}
-\end{figure}
-
-Interestingly, this type of variability in evolutionary outcomes can
-be weeded out by imprecision. To investigate this, we calculated the average
-distance between evolved sender strategies within each group of trials
-that had identical parameter values. We determined the distance
-between strategies $\Sstrat$ and $\Sstrat'$ as the average Hellinger
-distance between distributions $\Sstrat(\state)$ and
-$\Sstrat'(\state)$ at each choice point $\state$:
-\begin{align*}
-  \text{HD}(\Sstrat,\Sstrat') & = \frac{1}{{\card{\States} \cdot
-     \sqrt{2}}} \cdot  \sum_{\state \in \States} 
- \sqrt{\sum_{\messg \in  \Messgs}
-         \left ( \sqrt{\Sstrat(\state,\messg)} -
-         \sqrt{\Sstrat'(\state,\messg)} \right )^2 }\,.
-\end{align*}
-To compensate for the arbitrariness of message use, we set the distance between strategies
-$\Sstrat$ and $\Sstrat'$ to be the maximum of $\text{HD}(\Sstrat,\Sstrat')$ and
-$\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
-indices. An example of the \emph{within group distance}, i.e., the average distances between
-all sender strategies obtained for the same parameter values, is plotted in
-Figure~\ref{fig:GroupMeasuresA} for $\beta = 0.1$ and $n_s = 10$. Despite some quantitative
-differences, the general trend is the same for all other parameter settings that we tested:
-with increasing imprecision, the resulting sender strategies were much more alike (modulo
-swapping of messages). This means that perceptual imprecision can speed up and unify
-evolutionary outcomes. It can amplify the emergence of sender strategies that are not only
-convex, but also regular in that they induce a vague category split exactly in the middle of
-the unit interval.
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/WithinGroupDistanceConcise.pdf}
-    \caption{Within group distance}
-    \label{fig:GroupMeasuresA}
-  \end{subfigure}
-    \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{plots/WithinGroupEUConcise.pdf}
-    \caption{Within group expected utility}
-    \label{fig:GroupMeasuresB}
-  \end{subfigure}
-
-    \caption{Within group measures for all runs with $\toler = 0.1$ and $n_s = 10$.}
-  \label{fig:GroupMeasures}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Discussion}
-\label{sec:discussion}
-
-Our imprecise imitation dynamic leads to by-and-large successful signaling behavior, even in
-the presence of noise, that shows the hallmarks of vagueness as desired. It also gives rise to
-population-level behavior that looks as if agents are generalizing across similar
-stimuli. Here, we would like to reflect briefly on some further conceptually relevant points
-and compare our approach to related work.
-
-\subsection{Levels of vagueness}
-The imprecise imitation dynamic was introduced in Section~\ref{sec:repl-diff-dynam} as tracing
-changes in the overall distribution of pure strategies in a population:
-$\Sstrat(\messg \probbar \state)$ was said to represent the probability that a randomly sampled
-sender would have a strategy that responds with $\messg$ to $\state$ (likewise for the
-receiver). This is in line with the standard interpretation of the replicator equation, but we
-should consider its philosophical implications. Based on this picture, vagueness in signal use
-would seem to be characterized as a strictly population-level phenomenon since it arises in a
-signaling system from the inability of individual agents to fully align their (non-vague)
-strategies because of imprecision. This is, we believe, a plausible mechanism that can already
-explain the existence of vagueness in a language even if we assume that each agent commands a
-non-vague idiolect.
-
-We would not, however, want to commit to the idea that vagueness does not exist at the level of
-individual agents. True, our derivation of the imprecise imitation dynamic assumed that agents
-carry and revise pure strategies. But that was an assumption of convenience, not of
-conviction. Moreover, even if individual agents command a non-vague pure strategy, the
-realization of that pure strategy, according to our model, is bound to be vague (e.g., the same
-agent could signal differently in repeated exposure to the same state). We have used the term
-``observation noise'' here, but it may well be that this could equally well be perceived as an
-inseparable component of an agent's ``signaling faculty.'' In this sense, then, the model might
-not be entirely incompatible with a picture of agents who have internalized a vague signaling
-strategy. It would need to be seen, however, how revision of non-deterministic individual-level
-behavior must be spelled out rigorously and whether the resulting population-level dynamic
-would be equivalent to our present proposal in all relevant respects.
-
-% Moreover, we believe that the model may be interpreted in yet alternative ways. The
-% probability $\Sstrat(\messg \probbar \state)$ merely gives us the likelihood that $\messg$ will
-% be used by a sender when observing state $\state$.  Whether the uncertainty represents
-% variations in a population of agents playing pure strategies, a single sender using a
-% non-deterministic strategy, or even a heterogeneous population of non-deterministic senders, is
-% up for interpretation.  The first option supported what we believe to be a plausible motivation
-% for the mechanism of imitation of success with imprecision, but other motivations could be
-% imagined to better fit an interpretation that includes non-deterministic agents.  We thus
-% believe that the model introduced in this paper does not exclude vagueness at the individual
-% level, but to explore all possible interpretations would go beyond our current scope.
-
-\subsection{Evolutionary benefits of imprecision}
-The inclusion of confusability of similar states has noteworthy effects on the evolving meaning
-of signals.  It transpired from our results that imprecision can have further accelerating and,
-surprisingly, unifying effects on meaning evolution.  The unifying property of perceptual
-imprecision could be considered an evolutionarily beneficial side-effect. A certain degree of
-imprecision can lead to higher \emph{within group expected utility}, defined as the average
-expected utility that each evolved language scored when playing against an arbitrary other
-language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a
-representative example. The observation repeats for other parameter values: while imprecision
-might decrease the communicative efficiency of individual languages, it increases the
-conceptual coherence and communicative success between independently evolving strategies. It is
-as if mere confusability of states imposes a regularity constraint on evolving categories.
-
-The phenomenon could potentially not just be a side effect.  Given the benefit of a certain
-amount of imprecision we observe when comparing within group expected utility, it would be
-interesting to study whether, under certain conditions, this group level advantage could trump
-the individual level disadvantage of a vague language, and thus actually select for a certain
-amount of imprecision.  This could be achieved by letting the imprecision parameter be an
-evolving part in the dynamics as well. The idea is for now mostly speculative, but we consider
-it an interesting avenue for future research.  There is additional motivation to consider its
-potential if we see it in the eyes of multilevel selection theory
-\citep{Wilson1994,OGorman2008}.
-
-\subsection{Related work}
-\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of
-reinforcement learning for sim-max games, in which successful play leads to reinforcement of
-choice options also for states similar to the ones that actually occurred. This not only leads
-to vague signaling of the appropriate kind, but also speeds up learning in such a way that,
-especially for sim-max games with higher numbers of states, higher levels of communicative
-success are reached in shorter learning periods. Our results complement and extend
-\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
-similar regularizing effects also for cases with low numbers of states and do not assume that
-agents have any kind of generalizing capacity in and of themselves, even if that is only
-implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
-State confusability has an effect on aggregate signaling behavior that can be described as
-generalization without generalizers: the dynamics of imprecise imitation look as if
-``conclusions'' about what works for one state are ``carried over'' to similar states. This,
-however, is merely an epiphenomenon in the sense that no single agent genuinely generalizes
-over stimuli or reasons about what a more systematic signaling strategy would be.
-
-\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
-in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
-clearly related to the present approach uses the notion of a \emph{quantal response}, also
-known as a \emph{soft-max function}
-\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
-main difference between this and our present approach is in where stochastic noise is assumed
-to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
-in case of imprecise imitation, it is perception and realization of similar states. There are
-cases, then, where evolving signaling behavior, as predicted by these two approaches, is quite
-different. Intuitively speaking, for a case with two messages, the further we venture away from
-a prototypical interpretation of either message, the less discriminative a signaling strategy
-would be when the source of ``trembles'' is the computation of expected utilities: to wit,
-since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
-response dynamics predict that senders would be almost indifferent. Sender behavior that
-evolves under confusion of states does not have this puzzling property, because a giant would
-not likely be confused for a dwarf.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-We set out to meet a technical challenge posed by Lipman's problem
-\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
-formal model that shows how vague language can evolve under selective pressure for efficient
-communication? On the assumption that agents are likely to mistake states for one another,
-especially when they are perceptually similar, we derived a generalization of the replicator
-dynamic from an agent-level process of imprecise imitation. The resulting population-level
-dynamic produced signaling behavior that is at the same time regular and by-and-large
-communicatively efficient, while also showing the crucial marks of vagueness. In a sense, the
-model derives vagueness as a by-product of an arguably natural limitation on the discriminatory
-power of signaling agents.  Although inability to sharply discriminate similar stimuli may lead
-to vagueness and bring about a (slight) decrease in communicative efficiency, there may also be
-an advantage, not of vagueness itself, but of its reason: systematicity in the confusability of
-states (which may be a natural by-product of the perceptual system) supports
-\emph{as-if}-generalization at the population-level without having to assume that agents
-themselves have any generalization power.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Imprecise conditional imitation}
-\label{sec:formal-stuff}
-
-The goal of this section is to provide technical details for
-Section~\ref{sec:repl-diff-dynam}. We first show, in Section~\ref{sec:cond-imit-succ}, how to
-derive the standard replicator dynamic from noise-free imitation of success. Then, in
-Section~\ref{sec:imit-succ-with}, we derive the imprecise imitation dynamic by a parallel
-chain of arguments.
-
-\subsection{Deriving the replicator dynamic from imitation of success}
-\label{sec:cond-imit-succ}
-
-The main idea behind \emph{imitation of success} is that agents imitate the behavior of other
-agents at some choice point with a probability that is proportional to the expected utility of
-the latter agents' choice. Since the sim-max games that we are looking at here have positive
-utilities upper bound by 1, we can identify the switching probability with the expected
-utility.
-
-Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
-agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
-``teacher.'' A random learner is drawn from the population and given a chance to change
-behavior at choice point $\state$. The probability that our learner plays $\messg$ is
-$\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
-$\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar \state)$. The
-learner then starts using $\messg'$ instead of $\messg$ with probability
-$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same; the
-learner could even be the teacher as well, by random sampling.)
-
-If agents get repeated update chances for their choice points, the \emph{expected change} of
-frequency of $\messg$-choices at $\state$ becomes:
-\begin{align}
-  \label{eq:MeanChange}
-  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
-\end{align}
-where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
-any $\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
-that agents switch from $\messg$ to any $\messg'$. Since we are dealing with expectations in a
-huge population, these can be spelled out as:
-\begin{align*}
-  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
-    \state)}_{\text{learner plays $\messg'$}} \cdot
-  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
-  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
-  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
-    \state)}_{\text{learner plays $\messg$}} \cdot
-  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
-  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
-\end{align*}
-From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
-to:
-\begin{align*}
-  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
-  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
-      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
-& = \underbrace{\Sstrat(\messg \probbar
-    \state)}_{\text{frequency of $\messg$ at $\state$}}  \Big ( \underbrace{\EU(\messg,
-  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
-      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
-      EU at choice point $\state$}} \Big )\,.
-\end{align*}
-This latter formulation is the continuous-time version of the replicator dynamic. We obtain a
-discrete-time formulation from it by assuming that discrete update steps are infinitesimally
-small, so that:
- \begin{align*}
-    \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
-    \probbar \state)\\ 
-    & = \frac{\Sstrat(\messg \probbar
-    \state)  \myts \EU(\messg,\state,\Rstrat)}{\sum_{\messg'}\Sstrat(\messg' \probbar
-    \state)  \myts \EU(\messg,\state,\Rstrat)} - \sigma(\messg
-      \probbar \state) \\
-    & = \frac{\Sstrat(\messg \probbar
-    \state)  \myts \EU(\messg,\state,\Rstrat) - \Sstrat(\messg \probbar \state) \sum_{\messg'}\Sstrat(\messg' \probbar
-    \state)  \myts \EU(\messg,\state,\Rstrat) }{\sum_{\messg'}\Sstrat(\messg' \probbar
-    \state)  \myts \EU(\messg,\state,\Rstrat)}
-      \,.
-  \end{align*}
-By dropping the denominator, which is constant for all $\messg$ for fixed
-$\state$, we obtain the above continuous-time formulation.
-
-\subsection{Imitation of success with imprecision}
-\label{sec:imit-succ-with}
-
-The above derivation of the replicator dynamic assumes that agents can discriminate choices and
-choice points perfectly. Let's dispense with that assumption. With an eye toward sim-max games,
-we will assume that states, but not messages, may be confused for one another.\footnote{It is
-  relatively straightforward to also incorporate confusability of messages, but this is
-  irrelevant to our present purposes.} Confusability of states will affect how agents behave,
-how they perceive the behavior of others, and the expected utilities of behavioral
-dispositions.
-
-To keep matters simple, let us assume that agents carry pure dispositions to act. Noise can
-affect the realization of these strategies. As a sender, every agent maps states to messages:
-these are subjectively perceived states, and no longer necessarily also the actually occurring
-states. As a receiver, every agent maps messages to state interpretations: these are intended
-interpretations that need not always be faithfully realized. This means that behavioral
-strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
-behavioral dispositions in the population, the realization and observation of which can be
-distorted by agents' confusion of similar states.
-
-If $\mystate{a}$ is the actual state, let $P_o(\mystate{o} \probbar \mystate{a})$ be the
-probability that a given agent observes state $\mystate{o}$. Similarly, if a given receiver
-intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
-the probability with which state $\mystate{r}$ is realized. A single round of play of a sim-max
-game is then governed by five pieces of stochastic information, where previously there were
-only three (see Figures~\ref{fig:sim-max-illustration} and
-\ref{fig:imprecise-sim-max-illustration}).
-
-Expected utilities of choices at choice points should likewise take into account that actual states
-need not be observed states, and intended interpretations need not be realized
-interpretations. First, note that
-\begin{align*}
-  P_{\overline{o}}(\mystate{a} \probbar \mystate{o}) \propto P_a(\mystate{a}) P_o(\mystate{o}
-    \probbar \mystate{a})
-\end{align*}
-is the probability that $\mystate{a}$ is actual if $\mystate{o}$ is observed by an agent. The
-probability that a random sender produces $\messg$ when the actual state is $\mystate{a}$ is:
-\begin{align*}
-  P_{\Sstrat}(\messg \probbar \mystate{a}) = \sum_{\mystate{o}} P_o(\mystate{o} \probbar
-  \mystate{a}) \Sstrat(\messg \probbar \mystate{o})\,. 
-\end{align*}
-The probability that the actual state is $\mystate{a}$ if a random sender produced $\messg$ is:
-\begin{align*}
-  P_{\overline{\Sstrat}}(\mystate{a} \probbar \messg) \propto P_a(\mystate{a})
-  P_{\Sstrat}(\messg \probbar \mystate{a})\,. 
-\end{align*}
-The probability that $\mystate{r}$ is realized by a random receiver in response to
-message $\messg$ is:
-\begin{align*}
-  P_{\Rstrat}(\mystate{r} \probbar \messg) = \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
-  \mystate{i}) \Rstrat(\mystate{i} \probbar \messg)\,.
-\end{align*}
-This lets us capture the expected utilities for observed states (sender) and intended
-interpretations (receiver) by taking into consideration what the likely actual states and
-realized interpretations will be:
-\begin{align*}
-  \EU(\messg , \mystate{o}, \Rstrat) & = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
-  \probbar \mystate{o}) \sum_{\mystate{r}} P_\Rstrat(\mystate{r} \probbar
-  \messg) \utils(\mystate{a}, \mystate{r})\,, \\
-  \EU(\mystate{i}, \messg, \Sstrat) & = \sum_{\mystate{a}} P_{\overline{\Sstrat}}(\mystate{a}
-  \probbar \messg) \sum_{\mystate{r}} P_{r}(\mystate{r} \probbar \mystate{i})
-  \utils(\mystate{a}, \mystate{r})\,.
-\end{align*}
-If the conditional probabilities $P_o$ and $P_r$ are trivial, i.e., assign probability 0 to the
-confusability of non-identical states, above definitions reduce to the
-previous definitions of expected utilities. This also legitimates the overload of notation.
-
-Presence of potential imprecision in the form of non-trivial $P_o$ and $P_r$ will also affect
-the dynamic that ensues from imitation of successes. Since imprecision works slightly
-differently on senders and receivers (the former confuse choice points, the latter confuse
-choices), we need to look separately at each case.
-
-As before, suppose that senders receive a chance to change their behavior independently for a
-given choice point. In the present case, this would be a chance to change how to respond to a
-subjectively perceived state $\mystate{o}$, which need not be the actual one. We must then
-consult the probability $P_o(\messg \probbar \mystate{o})$ that, given that the learner
-observed $\mystate{o}$, he will simultaneously observe a randomly sampled teacher play
-$\messg$. This is (with $P_{\overline{o}}$ and $P_{\Sstrat}$ as defined above):
-\begin{align*}
-  P_o(\messg \probbar \mystate{o}) = \sum_{\mystate{a}} P_{\overline{o}}(\mystate{a}
-  \probbar \mystate{o}) \myts P_{\Sstrat}(\messg \probbar \mystate{a})\,.
-\end{align*}
-The ``inflow'' and ``outflow'' probabilities $P(\messg' \rightarrow \messg, \state)$ and
-$P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches from any
-$\messg'$ to $\messg$ or from $\messg$ to any $\messg'$ in subjectively perceived state
-$\mystate{o}$ are therefore:
-\begin{align*}
-  P(\messg' \rightarrow \messg, \mystate{o}) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
-    \mystate{o})}_{\text{learner plays $\messg'$ at $\mystate{o}$}} \cdot
-  \underbrace{P_{o}(\messg \probbar \mystate{o})}_{\text{observe teacher play $\messg$}} \cdot
-  \underbrace{\EU(\messg, \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
-  \\
-  P(\messg \rightarrow \messg', \mystate{o}) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
-    \mystate{o})}_{\text{learner plays $\messg$ at $\mystate{o}$}} \cdot
-  \underbrace{P_{o}(\messg' \probbar \mystate{o})}_{\text{observe teacher play $\messg'$}} \cdot
-  \underbrace{\EU(\messg', \mystate{o}, \Rstrat)}_{\text{EU teacher choice in learner's view}}
-\end{align*}
-The mean change to the proportion of $\messg$ choices at state $\state$ are then:
-\begin{align}
-  \label{eq:nRDSender}
-  \dot{\Sstrat}(\messg \probbar \mystate{o}) & = P(\messg' \rightarrow m, \mystate{o}) - P(\messg
-  \rightarrow m', \mystate{o}) \nonumber \\
-  & = \sum_{\messg'} \Sstrat(\messg' \probbar \mystate{o}) \myts P_o(\messg \probbar
-  \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \sum_{\messg'} \Sstrat(\messg \probbar \mystate{o}) \myts
-  P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat) \nonumber \\
-  & = P_o(\messg \probbar \mystate{o}) \myts \EU(\messg, \mystate{o}, \Rstrat) - \Sstrat(\messg \probbar \mystate{o}) \myts \sum_{\messg'} P_o(\messg' \probbar \mystate{o}) \myts \EU(\messg', \mystate{o}, \Rstrat)\,.
-\end{align}
-
-The case of the receiver is mostly analogous. Presented with an update opportunity for choice
-point $\messg$, a learner will observe a random teacher choose interpretation $\mystate{o}$
-with probability:\footnote{We assume that the noise $P_o(\mystate{o} \probbar \mystate{a})$
-  that applies to the sender's observations also applies to the receiver's observations when
-  trying to imitate other agents' strategies. We could easily make the model more complex and
-  introduce another measure of noise for state confusability during receiver attempts to
-  imitate. We refrain from it here, because we see no immediate theoretical gain.}
-\begin{align*}
-  P_o(\mystate{o} \probbar \messg) = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
-  \myts P_{\Rstrat}(\mystate{r} \probbar \messg) % = \sum_{\mystate{r}} P_o(\mystate{o} \probbar \mystate{r})
-  % \myts \sum_{\mystate{i}} P_r(\mystate{r} \ \probbar
-  % \mystate{i}) \myts \Rstrat(\mystate{i} \probbar \messg)
-  \,.
-\end{align*}
-% The probability $P(\state' \rightarrow \state, \messg)$ that a randomly sampled learner switches
-% from using $\state'$ to $\state$ in response to $\messg$ is therefore:
-% \begin{align*}
-%   P(\state' \rightarrow \state, \messg) = \sum_{\state'} \underbrace{\Rstrat(\state' \probbar
-%     \messg)}_{\text{learner plays $\state'$}} \cdot
-%   \underbrace{P_{o}(\state \probbar \messg)}_{\text{observe teacher play $\state$}} \cdot
-%   \underbrace{\EU(\state, \messg, \Rstrat)}_{\text{EU teacher choice}}
-% \end{align*}
-Parallel to the sender case, this gives rise to:
-\begin{align}
-  \label{eq:nRDReceiver}
-  \dot{\Rstrat}(\state \probbar \messg) & = \sum_{\state'} \Rstrat(\state' \probbar \messg)
-  \myts P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \sum_{\state'}
-  \Rstrat(\state \probbar \messg) \myts
-  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) \nonumber \\
-  & = P_o(\state \probbar \messg) \myts \EU(\state, \messg, \Sstrat) - \Rstrat(\state \probbar
-  \messg) \myts \sum_{\state'}
-  P_o(\state' \probbar \messg) \myts \EU(\state', \messg, \Sstrat) 
-\end{align}
-
-The continuous-time formulations in Equations~(\ref{eq:nRDSender}) and (\ref{eq:nRDReceiver})
-have elegant and practical discrete-time solutions in:
-\begin{align*}
-  \Sstrat'(\messg \probbar \state) & \propto P_o(\messg \probbar \state) \myts \EU(\messg,
-  \state, \Rstrat)\,, & \Rstrat'(\state \probbar \messg) & \propto P_o(\state \probbar \messg)
-  \myts \EU(\state, \messg, \Sstrat)\,,
-\end{align*}
-which is the discrete-time formulation of the imprecise imitation dynamic given in
-Equation~(\ref{eq:nRD}). To see how the discrete-time formulation gives rise to the
-continuous-time formulations above, let's assume that update steps are infinitesimally small,
-so that, for the sender case:
-\begin{align*}
-  \dot{\sigma}(\messg \probbar \state) & = \sigma'(\messg \probbar \state) - \sigma(\messg
-  \probbar \state)\\
-  &= \frac{P_o(\messg \probbar \state) \myts \EU(\messg, \state, \Rstrat) - \Sstrat(\messg
-    \probbar \state) \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg', \state,
-    \Rstrat) }{ \sum_{\messg'} P_o(\messg' \probbar \state) \myts \EU(\messg', \state,
-    \Rstrat)}\,.
-\end{align*}
-As before, we drop the denominator, which is constant for all $\messg$ for fixed $\state$, and
-obtain the above continuous-time formulation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\printbibliography[heading=bibintoc]
-
 \end{document}

### 2015-11-17 presentations/01_toulouse.Rmd
slides Toulouse

---
title: "Imprecise imitation in signaling games"
author: "Michael Franke & Jose Pedro Correia"
runtime: shiny
output:
  ioslides_presentation:
    css: mistyle.css
    smaller: yes
    transition: faster
---
```{r setup, include=FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dev.args = list(bg = 'transparent'))
require('ggplot2')
require('reshape2')
require('coda')
require('ggmcmc')
require('grid')
require('rjags')
require('runjags')
```

### 2016-01-15 paper/version_04/title_page.tex
Updating title.

@@
 
-\title{Vagueness and \emph{as-if}-generalization from imprecise imitation in signaling games}
+\title{Vagueness and imprecise imitation in signaling games}
 
@@
 Michael Franke%
-\thanks{Department of General Linguistics, Faculty of Humanities, University of Tübingen}\\
+\thanks{Department of Linguistics, Faculty of Humanities, University of Tübingen, Germany}\\
 \texttt{mchfranke@gmail.com}
 \and
-Jos\'e Pedro Correia\\
+Jos\'e Pedro Correia%
+\thanks{Department of Philosophy, University of Porto, Portugal}\\
 \texttt{zepedro.correia@gmail.com}
 }
+
+% \usepackage{a4wide}
 
@@
 
-\maketitle
+\noindent \textbf{Title:}
 
-\begin{abstract}
-  Signaling games are popular models for studying the evolution of meaning, but typical
+\medskip
+
+\noindent Vagueness and imprecise imitation in signaling games
+
+\medskip
+\medskip
+\medskip
+
+\noindent \textbf{Authors:}
+
+\medskip
+
+\begin{minipage}{0.4\linewidth}
+  Michael Franke \\
+  Wilhelmstra\ss e 19\\
+  Department of Linguistics\\
+  University of T\"ubingen\\
+  72072 T\"u bingen \\
+  Germany \\
+\end{minipage}
+\hfill
+\begin{minipage}{0.5\linewidth}
+  Jos\'e Pedro Correia\\
+  Praça Rainha D. Amélia 268 4 HAB3\\
+  4000-075 Porto\\
+  Portugal \\
+\end{minipage}
+
+\medskip
+\medskip
+\medskip
+
+\noindent \textbf{Abstract:}
+
+\medskip
+
+\noindent  Signaling games are popular models for studying the evolution of meaning, but typical
   approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
@@
   level, without agents actually generalizing.
-\end{abstract}
+
+\medskip
+\medskip
+\medskip
+
+\noindent \textbf{Acknowledgements:}
+
+\medskip
+
+\noindent We would like to thank Gerhard J\"ager, Robert van Rooij and Martin Stokhof for
+inspiring discussion and comments. MF gratefully acknowledges financial support during the
+completion of this work by NWO-VENI grant 275-80-004, the Institutional Strategy of the
+University of T\"ubingen (Deutsche Forschungsgemeinschaft, ZUK 63) and the Priority Program
+XPrag.de (DFG Schwerpunktprogramm 1727).
 

### 2016-01-20 paper/version_05/paper.tex
Copied BPoS version to new folder to make changes. Added Markdown file for reactions to the reviewers.
Made some of the easier changes.

--- 2015-10-15 paper/version_04/paper.tex
@@
 
-\title{Vagueness, Noise, and Signaling}
+\title{Vagueness and imprecise imitation in signaling games}
 % \author{Michael Franke \and Jos\'e Pedro Correia}
@@
 
-\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of
-sim-max games are remarkably systematic: the evolutionarily stable states are demonstrably
-so-called Voronoi languages.\footnote{Their results were obtained for games with infinitely
-  many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a
-  quadratic loss function for utilities
-  $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$.} Roughly put, a Voronoi
+\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of sim-max games with infinitely many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a quadratic loss function for utilities $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$ are remarkably systematic: the evolutionarily stable states are demonstrably so-called Voronoi languages.
+Roughly put, a Voronoi
 language is a pair of sender and receiver strategies, such that the sender strategy partitions
@@
 
-\section{Exploring noisy imitation}
+\section{Exploring imprecise imitation}
 \label{sec:exploring-rdd}
@@
 
-Let the state space consist of $n \ge 2$ states that are equally spaced across the unit
+Let the state space consist of $\ns \ge 2$ states that are equally spaced across the unit
 interval, including $0$ and $1$. All states occur, for simplicity, with the same probability
@@
 
+In order to simplify the analysis, we focus on games with two messages, \emph{i.e.}~we fix $\card{M} = 2$.
 In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
-is governed by three parameters: the number $\ns = \card{T}s$ of states in the state space,
+is governed by three parameters: the number $\ns = \card{T}$ of states in the state space,
 imprecision $\imprecision$ and tolerance $\toler$.
 
-\subsection{Experimental set-up}
+\subsection{Simulation set-up}
 \label{sec:simulations}

### 2016-02-15 paper/version_05/paper.tex
Added some motivation for talking about orderly categories.

@@
 of everyday communication, but it is troublesome for some of the most
-prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell and the young
-Wittgenstein which is challenged by the paradoxes vagueness gives rise
+prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell, and early
+Wittgenstein, which is challenged by the paradoxes vagueness gives rise
 to.
@@
 
-This result is interesting because it demonstrates that signaling can
-impose orderly categories on a metric space, without that being the
-ulterior purpose of it all. Sender and receiver can be distinct
+It is intuitive to think that linguistic and conceptual categories are orderly in such a manner.
+For example, if you consider two people, one with a height of 2m and another with a height of 2.2m, both `tall', it would be difficult to defend considering a person with a height of 2.1m `short'.
+Another example is that of color categorization.
+The World Color Survey project~\citep{WCS} collected color naming data for 110 unwritten languages of 45 language families.
+For the great majority of these languages a pattern can be observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}.
+It is premature to argue that these observations can be extended to \emph{all} cases of categorization.
+However, for the cases where it does apply, the result of \citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling can impose this kind of orderly categories on a metric space without that being the ulterior purpose of it all.
+
+There are at least two potential ways of interpreting the signaling setup.
+Sender and receiver can be distinct
 entities, whose purpose is to communicate effectively about the actual
@@
 to be.
-
 Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics

### 2016-02-17 paper/version_05/paper.tex
Added some references, including to paper in preparation.

@@
 Another example is that of color categorization.
-The World Color Survey project~\citep{WCS} collected color naming data for 110 unwritten languages of 45 language families.
+The World Color Survey project~\citep{CookKay2005:The-World-Color,WCS} collected color naming data for 110 unwritten languages of 45 language families.
 For the great majority of these languages a pattern can be observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}.
 It is premature to argue that these observations can be extended to \emph{all} cases of categorization.
-However, for the cases where it does apply, the result of \citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling can impose this kind of orderly categories on a metric space without that being the ulterior purpose of it all.
+However, for the cases where it does apply, the result of \citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling can impose this kind of orderly categories on a metric space without that being the ulterior purpose of it all.%
+\footnote{For the concrete case of color categorization, see also \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
 

### 2016-02-22 paper/version_05/paper.tex

@@
 assumption, especially when some states can be perceptually similar and therefore likely to be
-confused for one another. Therefore, imitation dynamics could be affected by at least two
-sources of probabilistic noise:
+confused for one another.
+Human performance in these situations has been studied by a number of authors in experimental psychology~\citep{takane_structures_1992}.
+In a stimulus identification experiment~\citep{luce_detection_1963}, subjects are presented in each trial with a stimulus to be identified out of a fixed set.
+The more similar the stimuli are to each other, the larger the number of errors subjects make.
+For example, in Robert Nosofsky's experiments~\citep{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the center to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
+In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
+
+In the context of imitation of success in a game where states have a degree of similarity between them, the possibility of agents mistaking one state for another is thus something that should be taken into account.
+Imitation dynamics could be affected by at least two sources of probabilistic noise:
 

### 2016-02-22 paper/version_05/paper.tex
introduction extended

@@
 \renewcommand{\Pr}{\ensuremath{P}}
+
+\definecolor{Red}{RGB}{178,34,34}
+\newcommand{\mf}[1]{\textcolor{Red}{[mf: #1]}}
 
@@
 
-Many concepts and expressions are vague. A vague category knows clear
-cases that fall under it, clear cases that do not, and also so-called
-borderline cases. Borderline cases do not clearly apply, nor do they
-clearly not apply, and there may be differences between borderline
-cases in terms of how well they represent the category in
-question. Vagueness does not seem to dramatically affect the success
-of everyday communication, but it is troublesome for some of the most
-prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell, and early
-Wittgenstein, which is challenged by the paradoxes vagueness gives rise
-to.
-
-But there are other intriguing aspects about vagueness. The puzzle
-that we are concerned with here is how vagueness could arise and be
-maintained in the first place. This is a serious issue for
-functionalist accounts that maintain that concepts and linguistic
-meanings evolved driven towards efficiency. Since the existence of
-unclear borderline cases seems to entail inefficiency of
-categorization or communication, the challenge, succinctly put forward
-by \citet{Lipman2009:Why-is-Language}, is to explain how vagueness can
-persist under evolutionary pressure to be optimally discriminative. We
-address \emph{Lipman's problem} as a technical challenge, namely to
-find a conceptually sound and mathematically coherent formal model of
-the evolution of categorization and language use that yields
-by-and-large efficient categorization \emph{and} vagueness.
-
-A number of authors have recently tried to explain why vagueness evolved as something that is
-itself useful
+Many concepts and expressions are vague. A vague category knows clear cases that fall under it,
+clear cases that do not, and also so-called borderline cases. Borderline cases do not clearly
+apply, nor do they clearly not apply, and there may be differences between borderline cases in
+terms of how well they represent the category in question. Vagueness does not seem to
+dramatically affect the success of everyday communication, but it is troublesome for some of
+the most prominent theories of language and meaning. This is especially so for the
+logico-positivist tradition of Frege, Russell, and early Wittgenstein, which is challenged by
+the paradoxes vagueness gives rise to.
+
+But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
+could arise and be maintained in the first place. This is an apparent puzzle for functionalist
+accounts that maintain that concepts and linguistic meanings evolved driven towards
+efficiency. Since the existence of unclear borderline cases seems to entail inefficiency of
+categorization or communication, the challenge, succinctly put forward by
+\citet{Lipman2009:Why-is-Language}, is to explain two things at the same time: (i) how
+vagueness can persist under evolutionary pressure to be optimally discriminative, (ii) without
+jeopardizing the possibility of evolving, learning and communicating with a meaningful
+language. A number of authors have consequently tried to explain why vagueness evolved as
+something that is itself useful
 \citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
 Others have argued that vagueness is a natural byproduct of limitations in information
-processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or the result of particular
-low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
-contributes to the latter line of thought. Concretely, we introduce a dynamic that describes
-changes in aggregate population behavior as the result of imitation with probabilistic
-confusability of similar stimuli. We show that stochastic noise can not only lead to vague, yet
-relatively efficient signal use, but can also unify evolutionary outcomes and help avoid
-suboptimal categorization: regularities in confusability of stimuli also regularize signaling
-behavior and evolving categories. The system gives rise to generalization behavior without
-agents actually generalizing.
+processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalization in low-level
+learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
+latter line of thought. We believe that vagueness in language and thought may have many
+reasons, not just one. We focus here on one possibility that has not received due attention,
+namely imprecise perception of similar stimuli and imprecise realization of similar
+responses.
+
+Our conceptual starting point is the widely shared conviction that language is shaped by at
+least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
+functional pressure towards efficient communication. On the other hand, there is systematic
+error, noise or imprecision in the transmission of linguistic behavior, knowledge or traits. As
+an example for the latter, consider a child learning syntactic rules from a parent
+generation. The child must infer these unobservable rules from observable speech. Inductive
+biases of the learner may influence which syntactic rules are likely to be inferred from
+(finite) parental input. Over the course of many generations, the effects of such transmission
+biases can, in a manner of speaking, ``accumulate'' and lead to surprising results, such as the
+evolution of compositional form-meaning mappings or the use of regular recursive syntactic
+structure. This can happen when the effects of tranmission biases are isolated, as in iterated
+learning models
+\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
+or when they interact with functional pressure towards efficient communication, e.g., as
+formalized in the replicator mutator dynamic
+\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.
+
+The emphasis of previous models that studied the effects of transmission infidelity on the
+evolution of language has been on inductive biases and the systematicity, compression and
+regularization that they can introduce. Here, we would like to show that transmission noise of
+a different kind can lead to regularization as well and also give rise to vague
+meaning. Towards this end, we formalize the expected change in the behavior of a population of
+agents that try to imitate other agents' signaling behavior. We assume, however, that both
+realization of behavior, as well as observability of others' behavior, is molested by noise, in
+particular the probabilistic confusability of similar stimuli. The resulting population-level
+dynamic generalizes the replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its
+interpretation as a cultural evolutionary dynamic based on imitation
+\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
+confusability of stimuli does not undermine the possibility of evolving communicative signaling
+behavior. Instead, it leads to the evolution of vague meanings. It moreover unifies and
+regularizes evolutionary outcomes, making it appear as if agents were applying inductive biases
+or generalizing over partial observations, when in truth this is the sole effect of confusion
+of similar stimuli.
+
+% The replicator mutator dynamic is neat because it keeps the effects of fitness-based selection
+% (replication) and transmission biases (mutation) nicely apart. This is possible because it
+% adheres to a picture of (asexual) biological reproduction: the number of children a (single)
+% parent has is higher, the higher the communicative success of the parent's language; all
+% children of this parent will learn from this parent alone, but may make mistakes during
+% language acquisition.
 
 The next section introduces the background against which the work presented here can be
-appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
+appreciated. Section~\ref{sec:repl-diff-dynam} introduces our generalization of the replicator
 dynamic which is derived from the idea that agents imitate other agents' behavior while
@@
 approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
+
+\mf{intro misses a mentioning of signaling games}
 

### 2016-02-23 paper/version_05/paper.tex
some improvements on intro, references added for semantic maps

@@
 learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
-latter line of thought. We believe that vagueness in language and thought may have many
-reasons, not just one. We focus here on one possibility that has not received due attention,
-namely imprecise perception of similar stimuli and imprecise realization of similar
-responses.
+latter line of thought.
+
+We believe that vagueness in language and thought may have many reasons, not just one. We focus
+here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
+idea is that at least some meanings are vague, in part because of imprecision in the perception
+of similar stimuli and in the realization of similar responses.\footnote{On this view,
+  vagueness in language may be seen as a necessary sub-optimality due to limitations in another
+  domain of cognition, first and foremost perception. The whole system, e.g., perception and
+  language together, may be an almost optimal adaptation in a larger frame of reference, e.g.,
+  if we take into account the metabolic costs for increased accuracy.} While this idea may be
+\emph{prima facie} plausible, it is not immediately clear what the consequences of
+confusability of stimuli are on a process of meaning evolution. In other words, in order to
+address Lipman's challenge seriously, and not just hand-wave it away, it needs to be spelled
+out how exactly confusability of stimuli could come into play in a process of meaning evolution
+and how it could lead to vague but by-and-large informative signal meaning. The formal model
+that this paper introduces does exactly that. And more. % We find that confusability of stimuli
+% can regularize and systematize evolving meaning, and may thereby imply an advantageous
+% side-effect that may counter some of the disadvantages that imprecise and vague communication
+% may have. Here is some elaboration on this point.
 
@@
 generation. The child must infer these unobservable rules from observable speech. Inductive
-biases of the learner may influence which syntactic rules are likely to be inferred from
-(finite) parental input. Over the course of many generations, the effects of such transmission
-biases can, in a manner of speaking, ``accumulate'' and lead to surprising results, such as the
-evolution of compositional form-meaning mappings or the use of regular recursive syntactic
-structure. This can happen when the effects of tranmission biases are isolated, as in iterated
-learning models
+biases may influence which syntactic rules are likely to be inferred from (finite) parental
+input. Over the course of many generations, the effects of such biases can, in a manner of
+speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
+form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
+the effects of tranmission biases are isolated, as in iterated learning models
 \citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
@@
 confusability of stimuli does not undermine the possibility of evolving communicative signaling
-behavior. Instead, it leads to the evolution of vague meanings. It moreover unifies and
-regularizes evolutionary outcomes, making it appear as if agents were applying inductive biases
-or generalizing over partial observations, when in truth this is the sole effect of confusion
-of similar stimuli.
+behavior. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
+emergence of communicative signaling because it unifies and regularizes evolutionary outcomes,
+making it appear as if agents were applying inductive biases or generalizing over partial
+observations, when in truth this is the sole effect of confusion of similar stimuli.
 
@@
 
-Similarity-maximizing (short: sim-max) games are a variation of
-Lewis' games where the receiver's actions are equated with the state
-space (one can think of the actions as choosing states) and different states are allowed to be more or less
-similar to one another. While Lewis' games treated communicative
-success as a matter of black and white, sim-max games allow for a
-gradient notion: the more similar the receiver's interpretation is to
-the actual state, the better. Signaling games with utility-relevant
-similarities in the state space are fairly standard in economics
-\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor},
-but have received particular attention in a more philosophical context
-for reasons that will become clear presently.
+Similarity-maximizing (short: sim-max) games are a variation of Lewis' games where the
+receiver's actions are equated with the state space (one can think of the actions as choosing
+states) and different states are allowed to be more or less similar to one another. While
+Lewis' games treated communicative success as a matter of black and white, sim-max games allow
+for shades of gray: the more similar the receiver's interpretation is to the actual state, the
+better. Signaling games with utility-relevant similarities in the state space are fairly
+standard in economics
+\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor}, but have received
+particular attention in a more philosophical context for reasons that will become clear
+presently.
 
@@
 
-It is intuitive to think that linguistic and conceptual categories are orderly in such a manner.
-For example, if you consider two people, one with a height of 2m and another with a height of 2.2m, both `tall', it would be difficult to defend considering a person with a height of 2.1m `short'.
-Another example is that of color categorization.
-The World Color Survey project~\citep{CookKay2005:The-World-Color,WCS} collected color naming data for 110 unwritten languages of 45 language families.
-For the great majority of these languages a pattern can be observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}.
-It is premature to argue that these observations can be extended to \emph{all} cases of categorization.
-However, for the cases where it does apply, the result of \citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling can impose this kind of orderly categories on a metric space without that being the ulterior purpose of it all.%
-\footnote{For the concrete case of color categorization, see also \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
+It is intuitive to think that linguistic and conceptual categories are orderly in such a
+manner.  For example, if you consider two people, one with a height of 2m and another with a
+height of 2.2m, both `tall', it would be difficult to defend considering a person with a height
+of 2.1m `short'.  Another example is that of color categorization. The World Color Survey
+project~\citep{CookKay2005:The-World-Color,WCS} collected color naming data for 110 unwritten
+languages of 45 language families.  For the great majority of these languages a pattern can be
+observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}. A prominent
+school of thought in comparative linguistics also assumes that more abstract conceptual domains
+---e.g., spatial-topological relations, temporal reference, or the meanings expressible by
+indefinite pronouns--- are preferably carved up by the languages of the world in such a way
+that meanings are connected regions on a so-called \emph{semantic map}
+\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}. It
+is premature to argue that these observations can be extended to \emph{all} cases of
+categorization. However, for the cases where it does apply, the result of
+\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling
+can impose this kind of orderly categories on a metric space without that being the ulterior
+purpose of it all.%
+\footnote{For the concrete case of color categorization, see also
+  \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
 

### 2016-02-24 paper/version_05/paper.tex
Added reference to other identification experiments that confirm the pervasiveness of confusion and the relation with similarity.

@@
 confused for one another.
-Human performance in these situations has been studied by a number of authors in experimental psychology~\citep{takane_structures_1992}.
+Human performance in these situations has been studied by a number of authors in experimental psychology. %~\citep{takane_structures_1992}.
 In a stimulus identification experiment~\citep{luce_detection_1963}, subjects are presented in each trial with a stimulus to be identified out of a fixed set.
@@
 In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
+Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colors, and magnitude of lines and areas~\citep{donkin_why_2015}.
+The empirical data confirms not only the pervasiveness of variation in the subjects' ability to correctly identify stimuli, but also the relation between similarity and likelihood of stimulus confusion, and further suggests that the phenomenon might extend to all types of perception.
 

### 2016-02-25 paper/version_05/paper.tex
Rephrased Lipman's reference for clarity. Small changes in second to last paragraph of Section 2.1. Wrote some possible replies to the reviewers.

@@
 accounts that maintain that concepts and linguistic meanings evolved driven towards
-efficiency. Since the existence of unclear borderline cases seems to entail inefficiency of
-categorization or communication, the challenge, succinctly put forward by
-\citet{Lipman2009:Why-is-Language}, is to explain two things at the same time: (i) how
-vagueness can persist under evolutionary pressure to be optimally discriminative, (ii) without
+efficiency. \citet{Lipman2009:Why-is-Language} argues that, technically, the existence of unclear borderline cases always entails inefficiency of
+categorization or communication. The challenge is then to explain how
+vagueness can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii) without
 jeopardizing the possibility of evolving, learning and communicating with a meaningful
@@
 It is intuitive to think that linguistic and conceptual categories are orderly in such a
-manner.  For example, if you consider two people, one with a height of 2m and another with a
-height of 2.2m, both `tall', it would be difficult to defend considering a person with a height
-of 2.1m `short'.  Another example is that of color categorization. The World Color Survey
+manner.  For example, if you consider two people `tall', one with a height of 2m and another with a
+height of 2.2m, it would be difficult to defend not considering a person with a height
+of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of color categorization. The World Color Survey
 project~\citep{CookKay2005:The-World-Color,WCS} collected color naming data for 110 unwritten
 languages of 45 language families.  For the great majority of these languages a pattern can be
-observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}. A prominent
-school of thought in comparative linguistics also assumes that more abstract conceptual domains
----e.g., spatial-topological relations, temporal reference, or the meanings expressible by
-indefinite pronouns--- are preferably carved up by the languages of the world in such a way
-that meanings are connected regions on a so-called \emph{semantic map}
-\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}. It
+observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}. It
 is premature to argue that these observations can be extended to \emph{all} cases of
@@
   \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
+Finally, these considerations are in line with a prominent
+school of thought in comparative linguistics which also assumes that more abstract conceptual domains
+(e.g., spatial-topological relations, temporal reference, or the meanings expressible by
+indefinite pronouns) are preferably carved up by the languages of the world in such a way
+that meanings are connected regions on a so-called \emph{semantic map}
+\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}.
 

### 2016-02-27 paper/version_05/paper.tex
sections 1-2 reshuffled; partial pooling paragraph added

@@
 vagueness can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii) without
-jeopardizing the possibility of evolving, learning and communicating with a meaningful
+undermining the possibility of evolving, learning and communicating with a meaningful
 language. A number of authors have consequently tried to explain why vagueness evolved as
@@
 here on one \emph{a priori} plausible reason for why vagueness is natural and pervasive. The
-idea is that at least some meanings are vague, in part because of imprecision in the perception
-of similar stimuli and in the realization of similar responses.\footnote{On this view,
-  vagueness in language may be seen as a necessary sub-optimality due to limitations in another
-  domain of cognition, first and foremost perception. The whole system, e.g., perception and
-  language together, may be an almost optimal adaptation in a larger frame of reference, e.g.,
-  if we take into account the metabolic costs for increased accuracy.} While this idea may be
-\emph{prima facie} plausible, it is not immediately clear what the consequences of
-confusability of stimuli are on a process of meaning evolution. In other words, in order to
-address Lipman's challenge seriously, and not just hand-wave it away, it needs to be spelled
-out how exactly confusability of stimuli could come into play in a process of meaning evolution
-and how it could lead to vague but by-and-large informative signal meaning. The formal model
-that this paper introduces does exactly that. And more. % We find that confusability of stimuli
-% can regularize and systematize evolving meaning, and may thereby imply an advantageous
-% side-effect that may counter some of the disadvantages that imprecise and vague communication
-% may have. Here is some elaboration on this point.
-
-Our conceptual starting point is the widely shared conviction that language is shaped by at
-least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
-functional pressure towards efficient communication. On the other hand, there is systematic
-error, noise or imprecision in the transmission of linguistic behavior, knowledge or traits. As
-an example for the latter, consider a child learning syntactic rules from a parent
-generation. The child must infer these unobservable rules from observable speech. Inductive
-biases may influence which syntactic rules are likely to be inferred from (finite) parental
-input. Over the course of many generations, the effects of such biases can, in a manner of
-speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
-form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
-the effects of tranmission biases are isolated, as in iterated learning models
-\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
-or when they interact with functional pressure towards efficient communication, e.g., as
-formalized in the replicator mutator dynamic
-\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.
-
-The emphasis of previous models that studied the effects of transmission infidelity on the
-evolution of language has been on inductive biases and the systematicity, compression and
-regularization that they can introduce. Here, we would like to show that transmission noise of
-a different kind can lead to regularization as well and also give rise to vague
-meaning. Towards this end, we formalize the expected change in the behavior of a population of
-agents that try to imitate other agents' signaling behavior. We assume, however, that both
-realization of behavior, as well as observability of others' behavior, is molested by noise, in
-particular the probabilistic confusability of similar stimuli. The resulting population-level
-dynamic generalizes the replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its
-interpretation as a cultural evolutionary dynamic based on imitation
-\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
-confusability of stimuli does not undermine the possibility of evolving communicative signaling
-behavior. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
-emergence of communicative signaling because it unifies and regularizes evolutionary outcomes,
-making it appear as if agents were applying inductive biases or generalizing over partial
-observations, when in truth this is the sole effect of confusion of similar stimuli.
+idea is that vagueness is, at least in part, due to imprecision in the perception of similar
+stimuli and imprecision in the realization of similar responses. On this view, vagueness in
+language may be seen as a necessary sub-optimality due to limitations in another domain of
+cognition, first and foremost perception. It could then be speculated that the whole system,
+e.g., perception and language together, may be an almost optimal adaptation in a larger frame
+of reference, e.g., if we take into account the metabolic costs for increased perceptual
+accuracy. We will not engage in such speculation here. Rather we will explore the consequences
+of perceptual limitations on processes of meaning evolution in a suitable formal framework. In
+other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
+by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
+of stimuli could come into play in a process of meaning evolution and how it could lead to
+vague but by-and-large informative signal meaning. The formal model that this paper introduces
+does exactly that. But it also does more. We find that confusability of stimuli can regularize
+and systematize evolving meaning. This suggests a possibly advantageous side-effect that a
+natural cause of vagueness may have and that might counter some of the disadvantages that
+vagueness may have for communicative efficiency \citep[c.f.][]{OConnor2013:The-Evolution-o}.
 
@@
 The next section introduces the background against which the work presented here can be
-appreciated. Section~\ref{sec:repl-diff-dynam} introduces our generalization of the replicator
+appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
 dynamic which is derived from the idea that agents imitate other agents' behavior while
@@
 approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
-
-\mf{intro misses a mentioning of signaling games}
 
@@
 
+To illustrate, let us briefly consider a sim-max game with six equiprobable states and two
+messages, its equilibria and what vagueness intuitively would look like \citep[see][for a more
+thorough discussion]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that
+utilities are strictly decreasing with decreasing
+similarity. Figure~\ref{fig:EquilibriaVagueness} shows three pairs of sender and receiver
+strategies. States and acts are arranged according to their similarity: the closer they are to
+each other, the more similar they are. The pair in Figure~\ref{fig:ineffNonEq} is not an
+equilibrium, because the sender's non-convex use of signals is suboptimal given the receiver's
+behavior. In contrast, Figure~\ref{fig:efficientEq}Z shows a maximally efficient
+equilibrium. This is a partial pooling equilibrium in the sense that the speaker uses the same
+message for several states. This would be inefficient if there were more messages, but it is
+maximally efficient for the two-message case. So, while partial pooling may entail
+inefficiency, and while partial pooling can hamper the evolution of maximally efficient signal
+use
+\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
+this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
+would look like in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets close, so
+to speak. It shows smooth transitions across similar states at the boundaries of categories and
+across acts around the most prototypical instances (as indicated by decreasing thickness of
+arrows in Figure~\ref{fig:vagueUse}).
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.475\textwidth}
+    \centering
+    \begin{tikzpicture}
+
+      \draw[] (0,2) node (t1) {$\mystate{1}$};
+
+      \draw[] (1,2) node (t2) {$\mystate{2}$};
+
+      \draw[] (2,2) node (t3) {$\mystate{3}$};
+
+      \draw[] (3,2) node (t4) {$\mystate{4}$};
+
+      \draw[] (4,2) node (t5) {$\mystate{5}$};
+
+      \draw[] (5,2) node (t6) {$\mystate{6}$};
+
+      \draw[] (1,1) node (m1) {$\mymessg{1}$};
+
+      \draw[] (4,1) node (m2) {$\mymessg{2}$};
+
+      \draw[] (0,0) node (a1) {$\myact{1}$};
+
+      \draw[] (1,0) node (a2) {$\myact{2}$};
+
+      \draw[] (2,0) node (a3) {$\myact{3}$};
+
+      \draw[] (3,0) node (a4) {$\myact{4}$};
+
+      \draw[] (4,0) node (a5) {$\myact{5}$};
+
+      \draw[] (5,0) node (a6) {$\myact{6}$};
+
+      % arrows
+
+      \draw[->, thick] (t1) -- (m1);
+
+      \draw[->, thick] (t2) -- (m1);
+
+      \draw[->, thick] (t3) -- (m2);
+
+      \draw[->, thick] (t4) -- (m1);
+
+      \draw[->, thick] (t5) -- (m2);
+
+      \draw[->, thick] (t6) -- (m2);
+
+      \draw[->, thick] (m1) -- (a2);
+
+      \draw[->, thick] (m2) -- (a5);
+
+    \end{tikzpicture}
+
+    \caption{inefficient, non-convex, non-equilibrium}
+    \label{fig:ineffNonEq}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.475\textwidth}
+    \centering
+
+    \begin{tikzpicture}
+
+      \draw[] (0,2) node (t1) {$\mystate{1}$};
+
+      \draw[] (1,2) node (t2) {$\mystate{2}$};
+
+      \draw[] (2,2) node (t3) {$\mystate{3}$};
+
+      \draw[] (3,2) node (t4) {$\mystate{4}$};
+
+      \draw[] (4,2) node (t5) {$\mystate{5}$};
+
+      \draw[] (5,2) node (t6) {$\mystate{6}$};
+
+      \draw[] (1,1) node (m1) {$\mymessg{1}$};
+
+      \draw[] (4,1) node (m2) {$\mymessg{2}$};
+
+      \draw[] (0,0) node (a1) {$\myact{1}$};
+
+      \draw[] (1,0) node (a2) {$\myact{2}$};
+
+      \draw[] (2,0) node (a3) {$\myact{3}$};
+
+      \draw[] (3,0) node (a4) {$\myact{4}$};
+
+      \draw[] (4,0) node (a5) {$\myact{5}$};
+
+      \draw[] (5,0) node (a6) {$\myact{6}$};
+
+      % arrows
+
+      \draw[->, thick] (t1) -- (m1);
+
+      \draw[->, thick] (t2) -- (m1);
+
+      \draw[->, thick] (t3) -- (m1);
+
+      \draw[->, thick] (t4) -- (m2);
+
+      \draw[->, thick] (t5) -- (m2);
+
+      \draw[->, thick] (t6) -- (m2);
+
+      \draw[->, thick] (m1) -- (a2);
+
+      \draw[->, thick] (m2) -- (a5);
+
+    \end{tikzpicture}
+
+    \caption{maximally efficient equilibrium, partial pooling}
+    \label{fig:efficientEq}
+  \end{subfigure}
+
+\begin{subfigure}[]{0.475\textwidth}
+    \centering
+
+    \begin{tikzpicture}
+
+      \draw[] (0,2) node (t1) {$\mystate{1}$};
+
+      \draw[] (1,2) node (t2) {$\mystate{2}$};
+
+      \draw[] (2,2) node (t3) {$\mystate{3}$};
+
+      \draw[] (3,2) node (t4) {$\mystate{4}$};
+
+      \draw[] (4,2) node (t5) {$\mystate{5}$};
+
+      \draw[] (5,2) node (t6) {$\mystate{6}$};
+
+      \draw[] (1,1) node (m1) {$\mymessg{1}$};
+
+      \draw[] (4,1) node (m2) {$\mymessg{2}$};
+
+      \draw[] (0,0) node (a1) {$\myact{1}$};
+
+      \draw[] (1,0) node (a2) {$\myact{2}$};
+
+      \draw[] (2,0) node (a3) {$\myact{3}$};
+
+      \draw[] (3,0) node (a4) {$\myact{4}$};
+
+      \draw[] (4,0) node (a5) {$\myact{5}$};
+
+      \draw[] (5,0) node (a6) {$\myact{6}$};
+
+      % arrows
+
+      \draw[->, thick] (t1) -- (m1);
+
+      \draw[->, thick] (t2) -- (m1);
+
+      \draw[->, thick, dashed] (t3) -- (m1);
+
+      \draw[->, thick, dotted] (t3) -- (m2);
+
+      \draw[->, thick, dotted] (t4) -- (m1);
+
+      \draw[->, thick, dashed] (t4) -- (m2);
+
+      \draw[->, thick] (t5) -- (m2);
+
+      \draw[->, thick] (t6) -- (m2);
+
+      \draw[->, thick, dotted] (m1) -- (a1);
+
+      \draw[->, thick, dashed] (m1) -- (a2);
+
+      \draw[->, thick, dotted] (m1) -- (a3);
+
+      \draw[->, dotted] (m1) -- (a4);
+
+      \draw[->, dotted] (m2) -- (a3);
+
+      \draw[->, thick, dotted] (m2) -- (a4);
+
+      \draw[->, thick, dashed] (m2) -- (a5);
+
+      \draw[->, thick, dotted] (m2) -- (a6);
+
+    \end{tikzpicture}
+
+    \caption{almost efficient, probabilistic, intuitively vague}
+    \label{fig:vagueUse}
+  \end{subfigure}
+
+  \caption{Examples of strategy pairs in sim-max games.}
+  \label{fig:EquilibriaVagueness}
+\end{figure}
+
 \citet{DouvenDecock2011:Vagueness:-A-Co} show that
@@
 us first look into our own proposal in more detail.
+
+Our conceptual starting point is the widely shared conviction that language is shaped by at
+least two forces, which may, on occasion, pull in opposite direction. On the one hand, there is
+functional pressure towards efficient communication. On the other hand, there is systematic
+error, noise or imprecision in the transmission of linguistic behavior, knowledge or traits. As
+an example for the latter, consider a child learning syntactic rules from a parent
+generation. The child must infer these unobservable rules from observable speech. Inductive
+biases may influence which syntactic rules are likely to be inferred from (finite) parental
+input. Over the course of many generations, the effects of such biases can, in a manner of
+speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
+form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
+the effects of tranmission biases are isolated, as in iterated learning models
+\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
+or when they interact with functional pressure towards efficient communication, e.g., as
+formalized in the replicator mutator dynamic
+\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.
+
+The emphasis of previous models that studied the effects of transmission infidelity on the
+evolution of language has been on inductive biases and the systematicity, compression and
+regularization that they can introduce. Here, we would like to show that transmission noise of
+a different kind can lead to regularization as well and also give rise to vague
+meaning. Towards this end, we formalize the expected change in the behavior of a population of
+agents that try to imitate other agents' signaling behavior. We assume, however, that both
+realization of behavior, as well as observability of others' behavior, is systematically
+perturbed by noise. The resulting population-level dynamic generalizes the replicator dynamic
+\citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural evolutionary
+dynamic based on imitation
+\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
+confusability of stimuli does not undermine the possibility of evolving communicative signaling
+behavior. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
+emergence of communicative signaling because it unifies and regularizes evolutionary outcomes,
+making it appear as if agents were applying inductive biases or generalizing over partial
+observations, when in truth this is the sole effect of confusion of similar stimuli.
 

### 2016-02-28 paper/version_05/paper.tex
Tiny fixes.

@@
 cognition, first and foremost perception. It could then be speculated that the whole system,
-e.g., perception and language together, may be an almost optimal adaptation in a larger frame
+perception and language together, may be an almost optimal adaptation in a larger frame
 of reference, e.g., if we take into account the metabolic costs for increased perceptual
@@
 and systematize evolving meaning. This suggests a possibly advantageous side-effect that a
-natural cause of vagueness may have and that might counter some of the disadvantages that
+natural cause of vagueness may have and that might compensate some of the disadvantages that
 vagueness may have for communicative efficiency \citep[c.f.][]{OConnor2013:The-Evolution-o}.
@@
 equilibrium, because the sender's non-convex use of signals is suboptimal given the receiver's
-behavior. In contrast, Figure~\ref{fig:efficientEq}Z shows a maximally efficient
+behavior. In contrast, Figure~\ref{fig:efficientEq} shows a maximally efficient
 equilibrium. This is a partial pooling equilibrium in the sense that the speaker uses the same
@@
 this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
-would look like in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets close, so
+would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets close, so
 to speak. It shows smooth transitions across similar states at the boundaries of categories and

### 2016-03-01 paper/version_05/paper.tex
text changes final

@@
 accounts that maintain that concepts and linguistic meanings evolved driven towards
-efficiency. \citet{Lipman2009:Why-is-Language} argues that, technically, the existence of unclear borderline cases always entails inefficiency of
-categorization or communication. The challenge is then to explain how
-vagueness can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii) without
-undermining the possibility of evolving, learning and communicating with a meaningful
+efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signaling
+situations, the existence of unclear borderline cases entails inefficiency of categorization or
+communication, or at least no advantage. The challenge is then to explain how vagueness
+can persist both: (i) under evolutionary pressure to be optimally discriminative, and (ii)
+without undermining the possibility of evolving, learning and communicating with a meaningful
 language. A number of authors have consequently tried to explain why vagueness evolved as
@@
 
-Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$ typically with much
-fewer messages than states, a probability distribution $\Pr \in \Delta(\States)$ such that
-$\Pr(\state)$ gives the probability that state $\state$ occurs, a similarity metric\footnote{A
-  metric is a function of distance between any two points in a given space. It should satisfy
-  certain axioms that ensure behavior that one would intuitively expect from the words
-  ``similarity'' and ``distance'' alone, but these details do not matter for the purposes of
-  this paper.}
+Formally, a sim-max game consists of a set of states $\States$, a set of messages $\Messgs$
+typically with much fewer messages than states, a probability distribution
+$\Pr \in \Delta(\States)$ such that $\Pr(\state)$ gives the probability that state $\state$
+occurs, a similarity metric\footnote{A metric is a function of distance between any two points
+  in a given space. It should satisfy certain axioms that ensure behavior that one would
+  intuitively expect from the words ``similarity'' and ``distance'' alone, but these details do
+  not matter for the purposes of this paper. The assumption that state similarity forms a
+  metric is conceptually loaded, but we follow the literature here and conceive of it as a
+  first and pragmatic simplification, possibly to be dispensed with later.}
   % : similarity should always be non-negative, zero only for the distance
@@
 with the states of the world, so that the game is one of guessing the actual state, so to
-speak. The utility function should be a monotonically increasing function of similarity.
+speak. The utility function should be a monotonically increasing function of
+similarity.\footnote{Section~\ref{sec:setting-stage} motivates a particular choice of utility
+  function that we will explore in more detail. But since the general discussion also applies
+  and makes reference to other choices, we prefer to remain general for the time being.}
 Non-probabilistic sender behavior can be represented by a pure strategy
@@
 
-\subsection{Vague signaling in sim-max games}
+\subsection{Vagueness in sim-max games and conceptual spaces}
 
@@
 center of the vague category they represent.
-
-To illustrate, let us briefly consider a sim-max game with six equiprobable states and two
-messages, its equilibria and what vagueness intuitively would look like \citep[see][for a more
-thorough discussion]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that
-utilities are strictly decreasing with decreasing
-similarity. Figure~\ref{fig:EquilibriaVagueness} shows three pairs of sender and receiver
-strategies. States and acts are arranged according to their similarity: the closer they are to
-each other, the more similar they are. The pair in Figure~\ref{fig:ineffNonEq} is not an
-equilibrium, because the sender's non-convex use of signals is suboptimal given the receiver's
-behavior. In contrast, Figure~\ref{fig:efficientEq} shows a maximally efficient
-equilibrium. This is a partial pooling equilibrium in the sense that the speaker uses the same
-message for several states. This would be inefficient if there were more messages, but it is
-maximally efficient for the two-message case. So, while partial pooling may entail
-inefficiency, and while partial pooling can hamper the evolution of maximally efficient signal
-use
-\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
-this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
-would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets close, so
-to speak. It shows smooth transitions across similar states at the boundaries of categories and
-across acts around the most prototypical instances (as indicated by decreasing thickness of
-arrows in Figure~\ref{fig:vagueUse}).
-
-\begin{figure}
-  \centering
-
-  \begin{subfigure}[]{0.475\textwidth}
-    \centering
-    \begin{tikzpicture}
-
-      \draw[] (0,2) node (t1) {$\mystate{1}$};
-
-      \draw[] (1,2) node (t2) {$\mystate{2}$};
-
-      \draw[] (2,2) node (t3) {$\mystate{3}$};
-
-      \draw[] (3,2) node (t4) {$\mystate{4}$};
-
-      \draw[] (4,2) node (t5) {$\mystate{5}$};
-
-      \draw[] (5,2) node (t6) {$\mystate{6}$};
-
-      \draw[] (1,1) node (m1) {$\mymessg{1}$};
-
-      \draw[] (4,1) node (m2) {$\mymessg{2}$};
-
-      \draw[] (0,0) node (a1) {$\myact{1}$};
-
-      \draw[] (1,0) node (a2) {$\myact{2}$};
-
-      \draw[] (2,0) node (a3) {$\myact{3}$};
-
-      \draw[] (3,0) node (a4) {$\myact{4}$};
-
-      \draw[] (4,0) node (a5) {$\myact{5}$};
-
-      \draw[] (5,0) node (a6) {$\myact{6}$};
-
-      % arrows
-
-      \draw[->, thick] (t1) -- (m1);
-
-      \draw[->, thick] (t2) -- (m1);
-
-      \draw[->, thick] (t3) -- (m2);
-
-      \draw[->, thick] (t4) -- (m1);
-
-      \draw[->, thick] (t5) -- (m2);
-
-      \draw[->, thick] (t6) -- (m2);
-
-      \draw[->, thick] (m1) -- (a2);
-
-      \draw[->, thick] (m2) -- (a5);
-
-    \end{tikzpicture}
-
-    \caption{inefficient, non-convex, non-equilibrium}
-    \label{fig:ineffNonEq}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[]{0.475\textwidth}
-    \centering
-
-    \begin{tikzpicture}
-
-      \draw[] (0,2) node (t1) {$\mystate{1}$};
-
-      \draw[] (1,2) node (t2) {$\mystate{2}$};
-
-      \draw[] (2,2) node (t3) {$\mystate{3}$};
-
-      \draw[] (3,2) node (t4) {$\mystate{4}$};
-
-      \draw[] (4,2) node (t5) {$\mystate{5}$};
-
-      \draw[] (5,2) node (t6) {$\mystate{6}$};
-
-      \draw[] (1,1) node (m1) {$\mymessg{1}$};
-
-      \draw[] (4,1) node (m2) {$\mymessg{2}$};
-
-      \draw[] (0,0) node (a1) {$\myact{1}$};
-
-      \draw[] (1,0) node (a2) {$\myact{2}$};
-
-      \draw[] (2,0) node (a3) {$\myact{3}$};
-
-      \draw[] (3,0) node (a4) {$\myact{4}$};
-
-      \draw[] (4,0) node (a5) {$\myact{5}$};
-
-      \draw[] (5,0) node (a6) {$\myact{6}$};
-
-      % arrows
-
-      \draw[->, thick] (t1) -- (m1);
-
-      \draw[->, thick] (t2) -- (m1);
-
-      \draw[->, thick] (t3) -- (m1);
-
-      \draw[->, thick] (t4) -- (m2);
-
-      \draw[->, thick] (t5) -- (m2);
-
-      \draw[->, thick] (t6) -- (m2);
-
-      \draw[->, thick] (m1) -- (a2);
-
-      \draw[->, thick] (m2) -- (a5);
-
-    \end{tikzpicture}
-
-    \caption{maximally efficient equilibrium, partial pooling}
-    \label{fig:efficientEq}
-  \end{subfigure}
-
-\begin{subfigure}[]{0.475\textwidth}
-    \centering
-
-    \begin{tikzpicture}
-
-      \draw[] (0,2) node (t1) {$\mystate{1}$};
-
-      \draw[] (1,2) node (t2) {$\mystate{2}$};
-
-      \draw[] (2,2) node (t3) {$\mystate{3}$};
-
-      \draw[] (3,2) node (t4) {$\mystate{4}$};
-
-      \draw[] (4,2) node (t5) {$\mystate{5}$};
-
-      \draw[] (5,2) node (t6) {$\mystate{6}$};
-
-      \draw[] (1,1) node (m1) {$\mymessg{1}$};
-
-      \draw[] (4,1) node (m2) {$\mymessg{2}$};
-
-      \draw[] (0,0) node (a1) {$\myact{1}$};
-
-      \draw[] (1,0) node (a2) {$\myact{2}$};
-
-      \draw[] (2,0) node (a3) {$\myact{3}$};
-
-      \draw[] (3,0) node (a4) {$\myact{4}$};
-
-      \draw[] (4,0) node (a5) {$\myact{5}$};
-
-      \draw[] (5,0) node (a6) {$\myact{6}$};
-
-      % arrows
-
-      \draw[->, thick] (t1) -- (m1);
-
-      \draw[->, thick] (t2) -- (m1);
-
-      \draw[->, thick, dashed] (t3) -- (m1);
-
-      \draw[->, thick, dotted] (t3) -- (m2);
-
-      \draw[->, thick, dotted] (t4) -- (m1);
-
-      \draw[->, thick, dashed] (t4) -- (m2);
-
-      \draw[->, thick] (t5) -- (m2);
-
-      \draw[->, thick] (t6) -- (m2);
-
-      \draw[->, thick, dotted] (m1) -- (a1);
-
-      \draw[->, thick, dashed] (m1) -- (a2);
-
-      \draw[->, thick, dotted] (m1) -- (a3);
-
-      \draw[->, dotted] (m1) -- (a4);
-
-      \draw[->, dotted] (m2) -- (a3);
-
-      \draw[->, thick, dotted] (m2) -- (a4);
-
-      \draw[->, thick, dashed] (m2) -- (a5);
-
-      \draw[->, thick, dotted] (m2) -- (a6);
-
-    \end{tikzpicture}
-
-    \caption{almost efficient, probabilistic, intuitively vague}
-    \label{fig:vagueUse}
-  \end{subfigure}
-
-  \caption{Examples of strategy pairs in sim-max games.}
-  \label{fig:EquilibriaVagueness}
-\end{figure}
 
@@
 Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
-\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be met by evolving
-strategies in sim-max games. The approach we take here is similar in spirit, but different in
-relevant detail. A more in-depth comparison is deferred until Section~\ref{sec:discussion}. Let
-us first look into our own proposal in more detail.
+\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
+met by evolving strategies in sim-max games. To illustrate what vague signaling would look
+like, let us briefly consider a sim-max game with six equiprobable states and two messages and
+what its equilibria would be like \citep[see][for a more thorough discussion of equilibria of
+sim-max games]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that utilities
+are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
+three pairs of sender and receiver strategies. States are arranged according to their
+similarity: the closer they are to each other, the more similar they are. The pair in
+Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
+signals is suboptimal given the receiver's behavior. Namely, if \mymessg{1} is interpreted as
+\mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
+sending \mymessg{1} in \mystate{3} than from sending \mymessg{2}, because (by assumption)
+\mystate{3} is more similar to \mystate{2} than \mystate{5} is. In contrast,
+Figure~\ref{fig:efficientEq} shows a maximally efficient equilibrium. This is a partial pooling
+equilibrium in the sense that the speaker uses the same message for several states. Partial
+pooling equilibria can be less inefficient than other non-pooling equilibria, but that requires
+enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
+the two-message case. So, while partial pooling may entail inefficiency in some sense, and
+while partial pooling can hamper the evolution of maximally efficient signal use
+\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
+this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
+would look like the pair in Figure~\ref{fig:vagueUse}. This is not an equilibrium, but it gets
+close, so to speak. It shows smooth transitions across similar states at the boundaries of
+categories and across acts around the most prototypical instances (as indicated by decreasing
+thickness of arrows in Figure~\ref{fig:vagueUse}).
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.475\textwidth}
+    \centering
+    \begin{tikzpicture}
+
+      \draw[] (0,2) node (t1) {$\mystate{1}$};
+
+      \draw[] (1,2) node (t2) {$\mystate{2}$};
+
+      \draw[] (2,2) node (t3) {$\mystate{3}$};
+
+      \draw[] (3,2) node (t4) {$\mystate{4}$};
+
+      \draw[] (4,2) node (t5) {$\mystate{5}$};
+
+      \draw[] (5,2) node (t6) {$\mystate{6}$};
+
+      \draw[] (1,1) node (m1) {$\mymessg{1}$};
+
+      \draw[] (4,1) node (m2) {$\mymessg{2}$};
+
+      \draw[] (0,0) node (a1) {$\mystate{1}$};
+
+      \draw[] (1,0) node (a2) {$\mystate{2}$};
+
+      \draw[] (2,0) node (a3) {$\mystate{3}$};
+
+      \draw[] (3,0) node (a4) {$\mystate{4}$};
+
+      \draw[] (4,0) node (a5) {$\mystate{5}$};
+
+      \draw[] (5,0) node (a6) {$\mystate{6}$};
+
+      % arrows
+
+      \draw[->, thick] (t1) -- (m1);
+
+      \draw[->, thick] (t2) -- (m1);
+
+      \draw[->, thick] (t3) -- (m2);
+
+      \draw[->, thick] (t4) -- (m1);
+
+      \draw[->, thick] (t5) -- (m2);
+
+      \draw[->, thick] (t6) -- (m2);
+
+      \draw[->, thick] (m1) -- (a2);
+
+      \draw[->, thick] (m2) -- (a5);
+
+    \end{tikzpicture}
+
+    \caption{inefficient, non-convex, non-equilibrium}
+    \label{fig:ineffNonEq}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.475\textwidth}
+    \centering
+
+    \begin{tikzpicture}
+
+      \draw[] (0,2) node (t1) {$\mystate{1}$};
+
+      \draw[] (1,2) node (t2) {$\mystate{2}$};
+
+      \draw[] (2,2) node (t3) {$\mystate{3}$};
+
+      \draw[] (3,2) node (t4) {$\mystate{4}$};
+
+      \draw[] (4,2) node (t5) {$\mystate{5}$};
+
+      \draw[] (5,2) node (t6) {$\mystate{6}$};
+
+      \draw[] (1,1) node (m1) {$\mymessg{1}$};
+
+      \draw[] (4,1) node (m2) {$\mymessg{2}$};
+
+      \draw[] (0,0) node (a1) {$\mystate{1}$};
+
+      \draw[] (1,0) node (a2) {$\mystate{2}$};
+
+      \draw[] (2,0) node (a3) {$\mystate{3}$};
+
+      \draw[] (3,0) node (a4) {$\mystate{4}$};
+
+      \draw[] (4,0) node (a5) {$\mystate{5}$};
+
+      \draw[] (5,0) node (a6) {$\mystate{6}$};
+
+      % arrows
+
+      \draw[->, thick] (t1) -- (m1);
+
+      \draw[->, thick] (t2) -- (m1);
+
+      \draw[->, thick] (t3) -- (m1);
+
+      \draw[->, thick] (t4) -- (m2);
+
+      \draw[->, thick] (t5) -- (m2);
+
+      \draw[->, thick] (t6) -- (m2);
+
+      \draw[->, thick] (m1) -- (a2);
+
+      \draw[->, thick] (m2) -- (a5);
+
+    \end{tikzpicture}
+
+    \caption{maximally efficient equilibrium, partial pooling}
+    \label{fig:efficientEq}
+  \end{subfigure}
+
+\begin{subfigure}[]{0.475\textwidth}
+    \centering
+
+    \begin{tikzpicture}
+
+      \draw[] (0,2) node (t1) {$\mystate{1}$};
+
+      \draw[] (1,2) node (t2) {$\mystate{2}$};
+
+      \draw[] (2,2) node (t3) {$\mystate{3}$};
+
+      \draw[] (3,2) node (t4) {$\mystate{4}$};
+
+      \draw[] (4,2) node (t5) {$\mystate{5}$};
+
+      \draw[] (5,2) node (t6) {$\mystate{6}$};
+
+      \draw[] (1,1) node (m1) {$\mymessg{1}$};
+
+      \draw[] (4,1) node (m2) {$\mymessg{2}$};
+
+      \draw[] (0,0) node (a1) {$\mystate{1}$};
+
+      \draw[] (1,0) node (a2) {$\mystate{2}$};
+
+      \draw[] (2,0) node (a3) {$\mystate{3}$};
+
+      \draw[] (3,0) node (a4) {$\mystate{4}$};
+
+      \draw[] (4,0) node (a5) {$\mystate{5}$};
+
+      \draw[] (5,0) node (a6) {$\mystate{6}$};
+
+      % arrows
+
+      \draw[->, thick] (t1) -- (m1);
+
+      \draw[->, thick] (t2) -- (m1);
+
+      \draw[->, thick, dashed] (t3) -- (m1);
+
+      \draw[->, thick, dotted] (t3) -- (m2);
+
+      \draw[->, thick, dotted] (t4) -- (m1);
+
+      \draw[->, thick, dashed] (t4) -- (m2);
+
+      \draw[->, thick] (t5) -- (m2);
+
+      \draw[->, thick] (t6) -- (m2);
+
+      \draw[->, thick, dotted] (m1) -- (a1);
+
+      \draw[->, thick, dashed] (m1) -- (a2);
+
+      \draw[->, thick, dotted] (m1) -- (a3);
+
+      \draw[->, dotted] (m1) -- (a4);
+
+      \draw[->, dotted] (m2) -- (a3);
+
+      \draw[->, thick, dotted] (m2) -- (a4);
+
+      \draw[->, thick, dashed] (m2) -- (a5);
+
+      \draw[->, thick, dotted] (m2) -- (a6);
+
+    \end{tikzpicture}
+
+    \caption{almost efficient, probabilistic, intuitively vague}
+    \label{fig:vagueUse}
+  \end{subfigure}
+
+  \caption{Examples of strategy pairs in sim-max games. Sender strategies map states onto
+    messages (top two rows); receiver strategies map messages onto states (bottom two rows). In
+    Figure~\ref{fig:EquilibriaVagueness}, the thickness of arrows indicates the probability of
+    a choice in a probabilistic strategy.}
+  \label{fig:EquilibriaVagueness}
+\end{figure}
+
+\subsection{Vagueness, functional pressure \& transmission biases}
+\label{sec:vagu-funct-press}
+
+The approach we take here is similar in spirit to that of
+\citet{FrankeJager2010:Vagueness-Signa} and \citet{OConnor2013:The-Evolution-o}, but different
+in relevant detail. A more in-depth comparison is deferred until
+Section~\ref{sec:discussion}. Let us first motivate our approach here, and spell it out in more
+detail in the following section.
 
@@
 regularization that they can introduce. Here, we would like to show that transmission noise of
-a different kind can lead to regularization as well and also give rise to vague
-meaning. Towards this end, we formalize the expected change in the behavior of a population of
-agents that try to imitate other agents' signaling behavior. We assume, however, that both
-realization of behavior, as well as observability of others' behavior, is systematically
-perturbed by noise. The resulting population-level dynamic generalizes the replicator dynamic
+a different kind can lead to regularization as well and also give rise to vague meaning. While
+perhaps \emph{prima facie} unexpected, we find that shared perceptual biases that perturb the
+transmission of successful signaling behavior can regularize, facilitate and accelerate the
+evolution of meaning conventions, albeit at the cost of vagueness. To show this, we formalize
+the expected change in the behavior of a population of agents that try to imitate other agents'
+signaling behavior. We assume, however, that both realization of behavior, as well as
+observability of others' behavior, is systematically perturbed by noise. The resulting
+population-level dynamic generalizes the replicator dynamic
 \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural evolutionary
@@
   not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
-  between dynamics on mixed or behavioral strategies.)} The sender's behavioral strategies are
+  between dynamics on mixed or behavioral strategies.) It is in this respect that the cultural
+  evolutionary dynamic introduced here differs most visibly from the replicator mutator dynamic
+  \citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}, which
+  operates on mixed strategies and is motivated by assumptions of (asexual) biological
+  inheritance with transmission infidelity.} The sender's behavioral strategies are
 functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
@@
 $\state$. Similarly, $\Rstrat(\state \probbar \messg)$ is then the probability with which a
-randomly sampled receiver interprets $\messg$ as $\state$. A play of single evolutionary game
+randomly sampled receiver interprets $\messg$ as $\state$. A play of a single evolutionary game
 with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
@@
 sum, for each interpretation $\state'$, of the probability that some random receiver will
-choose $\state'$ given $\messg$ times the actual utility of triple $\state$, $\messg$ and
+choose $\state'$ given $\messg$ times the actual utility of pair $\state$ and
 $\state'$.
@@
 
-\subsection{Noisy replicator dynamic}
+\subsection{Noise perturbed conditional imitation}
 \label{sec:noisy-repl-dynam}
@@
   $\mystate{i}$, the probability that $\mystate{r}$ is realized is
-  $P_r(\mystate{r} \probbar \mystate{o})$.
+  $P_r(\mystate{r} \probbar \mystate{i})$.
 \end{enumerate}
@@
 observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
-$\messg$ as his choice for $\state$ with probability given by the expected utility of sending
+$\messg$ as his choice for $\state$ with a probability given by the expected utility of sending
 $\messg$ when perceiving state $\state$. Expected utility should, of course, take the
@@
   if} information about what is good for one state percolates also to similar states. In other
-words, imprecise imitation may make signaling behavior on average look like agents generalize
+words, imprecise imitation may make signaling behavior on average look as if agents generalize
 across similar states, even if no agent actually generalizes. To explore whether confusion of
@@
 
-As for utility, we define it in terms of similarity and introduce another free parameter, $\toler \ge 0$.
- The intention is for this parameter to model the amount of tolerable pragmatic slack, which we felt should be allowed to vary separately from perceptual imprecision:
+As for utility, we define it in terms of similarity and introduce another free parameter,
+$\toler \ge 0$.  The intention is for this parameter to model the amount of tolerable pragmatic
+slack, which should be allowed to vary separately from perceptual imprecision:
 \begin{align*}
@@
 
-In order to simplify the analysis, we focus on games with two messages, \emph{i.e.}~we fix $\card{M} = 2$.
+In order to simplify the analysis, we focus on games with two messages, i.e., we fix $\card{M} = 2$.
 In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
@@
 
-  \caption{Example strategies under \rdd at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
+  \caption{Example strategies at stopping time. Each line corresponds to a message and plots, for each state, the probability that the message is used.}
   \label{fig:example_strats}
@@
 stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
-imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
+imprecision means that the trial was effectively an application of the standard replicator dynamic. Noteworthily,
 the given sender strategy approximates a pure sender strategy that crisply partitions the state
@@
 \citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
-scope of this paper, but we will shortly see that diffusion from confusability of states
+scope of this paper, but we will see shortly that diffusion from confusability of states
 clearly prevents evolutionary paths that meander for a long time in the vicinity of non-convex
@@
 
-\subsection{Dependent measures}
+\subsection{Measures of interest}
 
@@
 \begin{align*}
-  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
+  E(\Sstrat) & = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
   \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
-  \state))\,, \\
-  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
+  \state))\,, &
+  E(\Rstrat) & = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
   \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
@@
 \end{align*}
-Values obtained by these definitions are lower bounded by $0$ and
-upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
-\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
-\log(|\States|)$. We work with normalized values. The sender
-strategies in Figures~\ref{fig:example_stratsA} and
-\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
-respectively. The receiver strategies have respective entropies $0.43$
-and $0.81$. In general, we expected that vague languages would have
-higher entropy than crisp ones and that increasing imprecision would
-lead to increased entropy, all else being equal.
+Values obtained by these definitions have lower bound of zero and an upper bound of,
+respectively, $\log(|\Messgs^\States|) = |\States| \cdot \log(|\Messgs|)$ and
+$\log(|\States^\Messgs|) = |\Messgs| \cdot \log(|\States|)$. We work with values rescaled to
+lie in $[0;1]$ for cross-comparability. The sender strategies in
+Figures~\ref{fig:example_stratsA} and \ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and
+$0.08$, respectively. The receiver strategies have respective entropies $0.43$ and $0.81$. In
+general, we expected that vague languages would have higher entropy than crisp ones and that
+increasing imprecision would lead to increased entropy, all else being equal.
 
@@
 Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
-strategies, like we hypothesized, we should see more convexity with increasing impairment all
+strategies, like we hypothesized, we should see more convexity with increasing imprecision all
 else equal.
@@
 
-  \caption{More example strategies under \rdd at stopping time of our simulations.}
+  \caption{More example strategies at stopping time of our simulations.}
   \label{fig:MoreExample_strats}
@@
 inseparable component of an agent's ``signaling faculty.'' In this sense, then, the model might
-not be entirely incompatible with a picture of agents who have internalized a vague signaling
+be compatible with a picture of agents who have internalized a vague signaling
 strategy. It would need to be seen, however, how revision of non-deterministic individual-level
@@
 amount of imprecision we observe when comparing within group expected utility, it would be
-interesting to study whether, under certain conditions, this group level advantage could trump
-the individual level disadvantage of a vague language, and thus actually select for a certain
+interesting to study whether, under certain conditions, this group-level advantage could trump
+the individual-level disadvantage of a vague language, and thus actually select for a certain
 amount of imprecision.  This could be achieved by letting the imprecision parameter be an
@@
 clearly related to the present approach uses the notion of a \emph{quantal response}, also
-known as a \emph{soft-max function}
+known as a \emph{soft-max response function}
 \citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
@@
 formal model that shows how vague language can evolve under selective pressure for efficient
-communication? On the assumption that agents are likely to mistake states for one another,
-especially when they are perceptually similar, we derived a generalization of the replicator
-dynamic from an agent-level process of imprecise imitation. The resulting population-level
-dynamic produced signaling behavior that is at the same time regular and by-and-large
-communicatively efficient, while also showing the crucial marks of vagueness. In a sense, the
-model derives vagueness as a by-product of an arguably natural limitation on the discriminatory
-power of signaling agents.  Although inability to sharply discriminate similar stimuli may lead
-to vagueness and bring about a (slight) decrease in communicative efficiency, there may also be
-an advantage, not of vagueness itself, but of its reason: systematicity in the confusability of
-states (which may be a natural by-product of the perceptual system) supports
-\emph{as-if}-generalization at the population-level without having to assume that agents
-themselves have any generalization power.
+communication if agents tend to confuse similar stimuli? To address this, we derived a
+generalization of the replicator dynamic from an agent-level process of imprecise
+imitation. The resulting population-level dynamic produced signaling behavior that is at the
+same time regular and by-and-large communicatively efficient, while also showing the crucial
+marks of vagueness. In a sense, the model derives vagueness as a by-product of an arguably
+natural limitation on the discriminatory power of signaling agents.  Although inability to
+sharply discriminate similar stimuli may lead to vagueness and bring about a (slight) decrease
+in communicative efficiency, there may also be an advantage, not of vagueness itself, but of
+its reason: systematicity in the confusability of states (which may be a natural by-product of
+the perceptual system) supports \emph{as-if}-generalization at the population-level without
+having to assume that agents themselves have any generalization power. In this way, the
+presented model extends research into the effects of transmission biases on processes of
+meaning evolution. While most previous models have focused on inductive biases of language
+learners and the regularization that these may effect
+\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un,KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
+we have shown here that shared perceptual biases, of which such an effect was not necessarily
+expected, can also regularize, facilitate and accelerate the evolution of meaning conventions.
 

### 2016-03-03 paper/version_05/paper.tex
replies to reviewers
final tweaks responses and paper

@@
 on states $\similarity \mycolon \States \times \States \rightarrow \mathds{R}$ such that
-$\similarity(\mystate{1}, \mystate{2})$ is the (perceptual) similarity between $\mystate{1}$
-and $\mystate{2}$, and a utility function
+$\similarity(\mystate{1}, \mystate{2})$ is the (physical) similarity between $\mystate{1}$ and
+$\mystate{2}$, and a utility function
 $\Utils \mycolon \States \times \States \rightarrow \mathds{R}$ such that
@@
 with the states of the world, so that the game is one of guessing the actual state, so to
-speak. The utility function should be a monotonically increasing function of
-similarity.\footnote{Section~\ref{sec:setting-stage} motivates a particular choice of utility
-  function that we will explore in more detail. But since the general discussion also applies
-  and makes reference to other choices, we prefer to remain general for the time being.}
+speak. For the modeling purposes of this paper, we make the simplifying assumption that
+$\States$ is a set of points in Euclidean space, whose closeness to each other tracks physical
+similarity. Perceived similarity, where it is necessary, would be a monotonic function of
+physical similarity. Likewise, the utility function should be a monotonically increasing
+function of physical similarity.\footnote{Section~\ref{sec:setting-stage} motivates particular
+  choices of similarity and utility functions that we will explore in more detail.}
 Non-probabilistic sender behavior can be represented by a pure strategy
@@
 equilibrium in the sense that the speaker uses the same message for several states. Partial
-pooling equilibria can be less inefficient than other non-pooling equilibria, but that requires
+pooling equilibria can be less inefficient than other non-pooling equilibria, if there are
 enough messages. The strategy pair of Figure~\ref{fig:efficientEq} is maximally efficient for
@@
 regularization that they can introduce. Here, we would like to show that transmission noise of
-a different kind can lead to regularization as well and also give rise to vague meaning. While
-perhaps \emph{prima facie} unexpected, we find that shared perceptual biases that perturb the
-transmission of successful signaling behavior can regularize, facilitate and accelerate the
-evolution of meaning conventions, albeit at the cost of vagueness. To show this, we formalize
-the expected change in the behavior of a population of agents that try to imitate other agents'
-signaling behavior. We assume, however, that both realization of behavior, as well as
-observability of others' behavior, is systematically perturbed by noise. The resulting
-population-level dynamic generalizes the replicator dynamic
-\citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural evolutionary
-dynamic based on imitation
+a different kind can lead to regularization as well and also give rise to vague meaning. We
+find that shared perceptual biases that perturb the transmission of successful signaling
+behavior can regularize, facilitate and accelerate the evolution of meaning conventions, albeit
+at the cost of vagueness. Concretely, we formalize the expected change in the behavior of a
+population of agents that try to imitate other agents' signaling behavior. We assume, however,
+that both realization of behavior, as well as observability of others' behavior, is
+systematically perturbed by noise. The resulting population-level dynamic generalizes the
+replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
+evolutionary dynamic based on imitation
 \citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
@@
 $\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
-$\Rstrat(\state \probbar \messg)$. The noisy replicator dynamic in (\ref{eq:nRD})
+$\Rstrat(\state \probbar \messg)$. The imprecise imitation dynamic in (\ref{eq:nRD})
 conservatively extends the classic case in (\ref{eq:RDDiscrete}).

### 2016-05-26 paper/version_final/paper.tex
Added authors academic addresses.

--- 2016-03-03 paper/version_05/paper.tex
@@
 
+\begin{flushright}
+Michael Franke \\
+Wilhelmstra\ss e 19 \\
+Department of Linguistics \\
+University of T\"ubingen \\
+72072 T\"u bingen \\
+Germany \\
+\texttt{mchfranke@gmail.com} \\
+\bigskip
+Jos\'e Pedro Correia \\
+Via Panorâmica s/n \\
+Institute of Philosophy \\
+Faculty of Arts \\
+University of Porto \\
+4150-564 Porto \\
+Portugal \\
+\texttt{zepedro.correia@gmail.com} \\
+\end{flushright}
+
 \printbibliography[heading=bibintoc]

### 2016-05-26 paper/version_final/paper.tex
Ran aspell spell-checker.

@@
 
-\title{Vagueness and imprecise imitation in signaling games}
+\title{Vagueness and imprecise imitation in signalling games}
 % \author{Michael Franke \and Jos\'e Pedro Correia}
@@
 \begin{abstract}
-  Signaling games are popular models for studying the evolution of meaning, but typical
-  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
+  Signalling games are popular models for studying the evolution of meaning, but typical
+  approaches do not incorporate vagueness as a feature of successful signalling.  Complementing
   recent like-minded models, we describe an aggregate population-level dynamic that describes a
-  process of imitation of successful behavior under imprecise perception and realization of
-  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
+  process of imitation of successful behaviour under imprecise perception and realisation of
+  similar stimuli. Applying this new dynamic to a generalisation of Lewis' signalling games, we
   show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
-  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal
-  categorization. The upshot of this is that we see \emph{as-if}-generalization at an aggregate
-  level, without agents actually generalizing.
+  moreover, that it unifies evolutionary outcomes and helps avoid sub-optimal
+  categorisation. The upshot of this is that we see \emph{as-if}-generalisation at an aggregate
+  level, without agents actually generalising.
 \end{abstract}
@@
 accounts that maintain that concepts and linguistic meanings evolved driven towards
-efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signaling
-situations, the existence of unclear borderline cases entails inefficiency of categorization or
+efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signalling
+situations, the existence of unclear borderline cases entails inefficiency of categorisation or
 communication, or at least no advantage. The challenge is then to explain how vagueness
@@
 Others have argued that vagueness is a natural byproduct of limitations in information
-processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalization in low-level
+processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
 learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
@@
 idea is that vagueness is, at least in part, due to imprecision in the perception of similar
-stimuli and imprecision in the realization of similar responses. On this view, vagueness in
+stimuli and imprecision in the realisation of similar responses. On this view, vagueness in
 language may be seen as a necessary sub-optimality due to limitations in another domain of
@@
 vague but by-and-large informative signal meaning. The formal model that this paper introduces
-does exactly that. But it also does more. We find that confusability of stimuli can regularize
-and systematize evolving meaning. This suggests a possibly advantageous side-effect that a
+does exactly that. But it also does more. We find that confusability of stimuli can regularise
+and systematise evolving meaning. This suggests a possibly advantageous side-effect that a
 natural cause of vagueness may have and that might compensate some of the disadvantages that
@@
 The next section introduces the background against which the work presented here can be
-appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalization of the replicator
-dynamic which is derived from the idea that agents imitate other agents' behavior while
+appreciated. Section~\ref{sec:repl-diff-dynam} introduces a generalisation of the replicator
+dynamic which is derived from the idea that agents imitate other agents' behaviour while
 possibly confusing similar states. This imprecise imitation dynamic is explored in
@@
 
-Signaling games, as introduced by \citet{Lewis_1969:Convention}, have
+Signalling games, as introduced by \citet{Lewis_1969:Convention}, have
 a sender and a receiver. The sender knows the true state of the world,
@@
 the play is a success, otherwise a failure. Certain regular
-combinations of sender signaling and receiver reaction make messages
+combinations of sender signalling and receiver reaction make messages
 meaningful, in the sense that their use is correlated systematically
 to certain states or acts. To investigate the conditions under which
-such meaning-generating behavior can evolve is a highly interesting
+such meaning-generating behaviour can evolve is a highly interesting
 topic that we are only beginning to fully understand
@@
 
-Similarity-maximizing (short: sim-max) games are a variation of Lewis' games where the
+Similarity-maximising (short: sim-max) games are a variation of Lewis' games where the
 receiver's actions are equated with the state space (one can think of the actions as choosing
@@
 Lewis' games treated communicative success as a matter of black and white, sim-max games allow
-for shades of gray: the more similar the receiver's interpretation is to the actual state, the
-better. Signaling games with utility-relevant similarities in the state space are fairly
+for shades of grey: the more similar the receiver's interpretation is to the actual state, the
+better. Signalling games with utility-relevant similarities in the state space are fairly
 standard in economics
@@
 occurs, a similarity metric\footnote{A metric is a function of distance between any two points
-  in a given space. It should satisfy certain axioms that ensure behavior that one would
+  in a given space. It should satisfy certain axioms that ensure behaviour that one would
   intuitively expect from the words ``similarity'' and ``distance'' alone, but these details do
@@
 with the states of the world, so that the game is one of guessing the actual state, so to
-speak. For the modeling purposes of this paper, we make the simplifying assumption that
+speak. For the modelling purposes of this paper, we make the simplifying assumption that
 $\States$ is a set of points in Euclidean space, whose closeness to each other tracks physical
@@
   choices of similarity and utility functions that we will explore in more detail.}
-Non-probabilistic sender behavior can be represented by a pure strategy
+Non-probabilistic sender behaviour can be represented by a pure strategy
 $\Spure \in \Messgs^\States$, which deterministically defines which message would be used for
@@
 height of 2.2m, it would be difficult to defend not considering a person with a height
-of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of color categorization. The World Color Survey
-project~\citep{CookKay2005:The-World-Color,WCS} collected color naming data for 110 unwritten
+of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
+project~\citep{CookKay2005:The-World-Color,WCS} collected colour naming data for 110 unwritten
 languages of 45 language families.  For the great majority of these languages a pattern can be
-observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}. It
+observed: basic colour terms are by and large convex~\citep{Regier07,Jager2010}. It
 is premature to argue that these observations can be extended to \emph{all} cases of
-categorization. However, for the cases where it does apply, the result of
-\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling
+categorisation. However, for the cases where it does apply, the result of
+\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
 can impose this kind of orderly categories on a metric space without that being the ulterior
 purpose of it all.%
-\footnote{For the concrete case of color categorization, see also
+\footnote{For the concrete case of colour categorisation, see also
   \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
@@
 
-There are at least two potential ways of interpreting the signaling setup.
+There are at least two potential ways of interpreting the signalling setup.
 Sender and receiver can be distinct
@@
 thought of as distinct modules in a single system, where the first
-module must discretize the information it is fed by selecting a small
+module must discretise the information it is fed by selecting a small
 sample of, suggestively, category labels. These are passed to a second
@@
 corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
-other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
+other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorisation
 system can be obtained: storing the prototypes lets us recover the categories without having to
@@
 
-This outline of an approach to categorization using sim-max games
+This outline of an approach to categorisation using sim-max games
 leaves many problems unaddressed. One of them is that natural
@@
 (ii) prototypes should likewise be gradient regions, peaking at the
-center of the vague category they represent.
+centre of the vague category they represent.
 
@@
 \citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
-met by evolving strategies in sim-max games. To illustrate what vague signaling would look
+met by evolving strategies in sim-max games. To illustrate what vague signalling would look
 like, let us briefly consider a sim-max game with six equiprobable states and two messages and
@@
 Figure~\ref{fig:ineffNonEq} is not an equilibrium, because the sender's non-convex use of
-signals is suboptimal given the receiver's behavior. Namely, if \mymessg{1} is interpreted as
+signals is sub-optimal given the receiver's behaviour. Namely, if \mymessg{1} is interpreted as
 \mystate{2} and \mymessg{2} as \mystate{4}, then the sender would get a higher payoff from
@@
 functional pressure towards efficient communication. On the other hand, there is systematic
-error, noise or imprecision in the transmission of linguistic behavior, knowledge or traits. As
+error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
 an example for the latter, consider a child learning syntactic rules from a parent
@@
 form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
-the effects of tranmission biases are isolated, as in iterated learning models
+the effects of transmission biases are isolated, as in iterated learning models
 \citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
 or when they interact with functional pressure towards efficient communication, e.g., as
-formalized in the replicator mutator dynamic
+formalised in the replicator mutator dynamic
 \citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.
@@
 evolution of language has been on inductive biases and the systematicity, compression and
-regularization that they can introduce. Here, we would like to show that transmission noise of
-a different kind can lead to regularization as well and also give rise to vague meaning. We
-find that shared perceptual biases that perturb the transmission of successful signaling
-behavior can regularize, facilitate and accelerate the evolution of meaning conventions, albeit
-at the cost of vagueness. Concretely, we formalize the expected change in the behavior of a
-population of agents that try to imitate other agents' signaling behavior. We assume, however,
-that both realization of behavior, as well as observability of others' behavior, is
-systematically perturbed by noise. The resulting population-level dynamic generalizes the
+regularisation that they can introduce. Here, we would like to show that transmission noise of
+a different kind can lead to regularisation as well and also give rise to vague meaning. We
+find that shared perceptual biases that perturb the transmission of successful signalling
+behaviour can regularise, facilitate and accelerate the evolution of meaning conventions, albeit
+at the cost of vagueness. Concretely, we formalise the expected change in the behaviour of a
+population of agents that try to imitate other agents' signalling behaviour. We assume, however,
+that both realisation of behaviour, as well as observability of others' behaviour, is
+systematically perturbed by noise. The resulting population-level dynamic generalises the
 replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
@@
 \citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
-confusability of stimuli does not undermine the possibility of evolving communicative signaling
-behavior. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
-emergence of communicative signaling because it unifies and regularizes evolutionary outcomes,
-making it appear as if agents were applying inductive biases or generalizing over partial
+confusability of stimuli does not undermine the possibility of evolving communicative signalling
+behaviour. Instead, it leads to the evolution of vague meanings. It moreover accelerates the
+emergence of communicative signalling because it unifies and regularises evolutionary outcomes,
+making it appear as if agents were applying inductive biases or generalising over partial
 observations, when in truth this is the sole effect of confusion of similar stimuli.
@@
 
-Signaling agents can adapt their dispositions to act, given some feedback about their past
-success, in multiple ways. Usually, we would assume that changes in behavior should, at least
-on average, lean towards increasing chances of communicative success. Such behavioral
+Signalling agents can adapt their dispositions to act, given some feedback about their past
+success, in multiple ways. Usually, we would assume that changes in behaviour should, at least
+on average, lean towards increasing chances of communicative success. Such behavioural
 adaptations can be described at different levels of abstraction. At the level of individual
-agents, we can picture a more or less idealized process of how each agent adapts dispositions
+agents, we can picture a more or less idealised process of how each agent adapts dispositions
 for future actions based on various pieces of information available to the agent. More
-abstractly, at the level of a population of agents, we can describe how average behavioral
+abstractly, at the level of a population of agents, we can describe how average behavioural
 dispositions will evolve. The population-level perspective abstracts over small stochastic
 fluctuations and zooms in on the general tendency or direction of evolution that ensues from
-behavior at the agent-level.
+behaviour at the agent-level.
 
@@
 successful communication, we look at a population-level dynamic that describes the most likely
-evolutionary path of a population of signaling agents who imitate other agents' behavior but
+evolutionary path of a population of signalling agents who imitate other agents' behaviour but
 are liable to confuse states for one another. In the special limiting case where state
@@
 
-\subsection{Replicator dynamic in behavioral strategies}
+\subsection{Replicator dynamic in behavioural strategies}
 \label{sec:repl-dynam-behav}
@@
 $\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
-senders and receivers in the population. A \emph{behavioral strategy} is a function that maps
+senders and receivers in the population. A \emph{behavioural strategy} is a function that maps
 an agent's choice points to a probability distribution over available
-choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioral strategies, not \emph{mixed
+choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioural strategies, not \emph{mixed
     strategies}, i.e., probability distributions over functions from each choice point to an
-  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume
-  that agents can adapt their behavior locally, i.e., independently at each choice point. Our
-  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
+  act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioural strategies assume
+  that agents can adapt their behaviour locally, i.e., independently at each choice point. Our
+  focus on behavioural strategies greatly reduces the complexity of the dynamic and simplifies
   numerical simulations. But it also seems the more plausible choice for imitation-based update
@@
   not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
-  between dynamics on mixed or behavioral strategies.) It is in this respect that the cultural
+  between dynamics on mixed or behavioural strategies.) It is in this respect that the cultural
   evolutionary dynamic introduced here differs most visibly from the replicator mutator dynamic
@@
   operates on mixed strategies and is motivated by assumptions of (asexual) biological
-  inheritance with transmission infidelity.} The sender's behavioral strategies are
+  inheritance with transmission infidelity.} The sender's behavioural strategies are
 functions $\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$
@@
 $\messg \in \Messgs$ to a probability of each interpretation $\state \in \States$ being chosen
-in response to $\messg$. Although behavioral strategies are probabilistic, evolutionary
-modeling usually imagines that every individual agent has a non-probabilistic
-strategy. Behavioral strategies then capture average population behavior. Assuming a virtually
+in response to $\messg$. Although behavioural strategies are probabilistic, evolutionary
+modelling usually imagines that every individual agent has a non-probabilistic
+strategy. Behavioural strategies then capture average population behaviour. Assuming a virtually
 infinite population, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the
@@
 randomly sampled receiver interprets $\messg$ as $\state$. A play of a single evolutionary game
-with behavioral strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
+with behavioural strategies is illustrated in Figure~\ref{fig:sim-max-illustration}.
 
@@
 
-  \caption{A round of play in a sim-max game with behavioral strategies: Nature (N) chooses a
+  \caption{A round of play in a sim-max game with behavioural strategies: Nature (N) chooses a
     state $\mystate{1}$ with probability $\Pr(\mystate{1})$; a random sender (S) selects a
@@
   $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
-  side might still need to be normalized so as to have probabilities sum to one. In general,
+  side might still need to be normalised so as to have probabilities sum to one. In general,
   writing $P(x) \propto f(x)$ for any function $f \mycolon X \rightarrow \mathds{R}$ is
@@
 The replicator dynamic is an abstract population-level dynamic that describes the mean expected
-change of behavioral dispositions in a population of signalers. There are several ways of
-deriving the replicator dynamic from agent-level processes of behavioral adaptation. We focus
+change of behavioural dispositions in a population of signallers. There are several ways of
+deriving the replicator dynamic from agent-level processes of behavioural adaptation. We focus
 here on one of the simplest: \emph{imitation of success}
@@
 Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
-update scheme).  Every now and then a random agent gets a chance to alter his behavior for one
+update scheme).  Every now and then a random agent gets a chance to alter his behaviour for one
 of his choice points. Say, this is a sender who gets to ``reconsider'' his choice of message
@@
 The more similar the stimuli are to each other, the larger the number of errors subjects make.
-For example, in Robert Nosofsky's experiments~\citep{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the center to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
+For example, in Robert Nosofsky's experiments~\citep{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
 In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
-Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colors, and magnitude of lines and areas~\citep{donkin_why_2015}.
+Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citep{donkin_why_2015}.
 The empirical data confirms not only the pervasiveness of variation in the subjects' ability to correctly identify stimuli, but also the relation between similarity and likelihood of stimulus confusion, and further suggests that the phenomenon might extend to all types of perception.
@@
   that an agent observes it as $\mystate{o}$ is $P_o(\mystate{o} \probbar \mystate{a})$;
-\item \emph{realization noise:} whenever an agent intends to realize interpretation
-  $\mystate{i}$, the probability that $\mystate{r}$ is realized is
+\item \emph{realisation noise:} whenever an agent intends to realise interpretation
+  $\mystate{i}$, the probability that $\mystate{r}$ is realised is
   $P_r(\mystate{r} \probbar \mystate{i})$.
@@
 senders respond to observations not actual states, receiver strategies determine intentions not
-realized states, but payoff is calculated based on actual and realized states. Behavioral
+realised states, but payoff is calculated based on actual and realised states. Behavioural
 strategies thus encode what agents actually do from their subjective point of view; or, put
@@
     subsequently selects $\messg$ with probability $\Sstrat(\messg \probbar \mystate{o})$; a
-    randomly sampled receiver (R) intends to realize interpretation $\mystate{i}$ with
-    probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realizes interpretation
+    randomly sampled receiver (R) intends to realise interpretation $\mystate{i}$ with
+    probability $\Rstrat(\mystate{i} \probbar \messg)$ but actually realises interpretation
     $\mystate{r}$ with probability $P_r(\mystate{r} \probbar \mystate{i})$. Payoff for both
@@
 
-The presence of observation and realization noise also affects imitation of successes. Here, we
+The presence of observation and realisation noise also affects imitation of successes. Here, we
 focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:imit-succ-with}. Take a
-sender who gets to revise behavior for state $\state$. What agents can plausibly revise by
+sender who gets to revise behaviour for state $\state$. What agents can plausibly revise by
 imitation is their pure strategy, which maps perceived states onto messages. When an agent gets
@@
 agent perceives $\state$. This $P_o(\messg \probbar \state)$ is derived from the prior
-probability of states, the current sender population behavior $\Sstrat$ and the given
+probability of states, the current sender population behaviour $\Sstrat$ and the given
 observation noise (see Appendix~\ref{sec:imit-succ-with}). Eventually, the imitating agent adopts
@@
 intended interpretation of $\messg$, we need to look at the conditional probability
-$P_o(\state \probbar \messg)$ of observing another agent realize interpretation $\state$ given
+$P_o(\state \probbar \messg)$ of observing another agent realise interpretation $\state$ given
 that the first agent (and therefore the second as well) perceived $\messg$. This depends on
-observation and realization noise, as well as on the current receiver population behavior
+observation and realisation noise, as well as on the current receiver population behaviour
 $\Rstrat$.
@@
 $P_o(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ respectively. If there is no
-observation or realization noise, $P_o(\messg \probbar \state)$ reduces to
+observation or realisation noise, $P_o(\messg \probbar \state)$ reduces to
 $\sigma(\messg \probbar \state)$ and $P_o(\state \probbar \messg)$ reduces to
@@
 How does a tendency to confuse similar states interact with selective pressure towards more
-efficient signaling strategies under the imprecise conditional imitation dynamic? If confusion
+efficient signalling strategies under the imprecise conditional imitation dynamic? If confusion
 probabilities are moderate and regular, in that they track similarity of states, we might
-expect a regularizing effect also on evolving signaling strategies. Indeed, we hypothesize that
-confusion of states can give rise to population-level aggregate behavior that looks \emph{as
+expect a regularising effect also on evolving signalling strategies. Indeed, we hypothesise that
+confusion of states can give rise to population-level aggregate behaviour that looks \emph{as
   if} information about what is good for one state percolates also to similar states. In other
-words, imprecise imitation may make signaling behavior on average look as if agents generalize
-across similar states, even if no agent actually generalizes. To explore whether confusion of
+words, imprecise imitation may make signalling behaviour on average look as if agents generalise
+across similar states, even if no agent actually generalises. To explore whether confusion of
 states can have this effect, we turn to numerical simulation.
@@
 states at all. Figure~\ref{fig:NosofskySim} gives an impression of Nosofsky-similarity for
-different parameter values. Other formalizations of perceptual similarity are possible,
+different parameter values. Other formalisations of perceptual similarity are possible,
 including ones that allow for different discriminability in different areas of the state space,
@@
 $\mystate{j}$ is proportional to their perceived similarity and, to keep matters simple, that
-observation noise $P_o$ is simply the same as realizational confusability $P_r$ and
+observation noise $P_o$ is simply the same as realisational confusability $P_r$ and
 that both are governed by the same imprecision parameter $\alpha$:
@@
 $\imprecision \rightarrow \infty$, confusion is maximal and every state can be perceived or
-realized as any other state with probability $\nicefrac{1}{\card{\States}}$.
+realised as any other state with probability $\nicefrac{1}{\card{\States}}$.
 
@@
 In order to simplify the analysis, we focus on games with two messages, i.e., we fix $\card{M} = 2$.
-In sum, to structure thinking about the behavior of our imprecise imitation dynamic, the system
+In sum, to structure thinking about the behaviour of our imprecise imitation dynamic, the system
 is governed by three parameters: the number $\ns = \card{T}$ of states in the state space,
@@
 concerns regarding length of simulation time, but also theoretically justified, because we
-hypothesize that confusability of states leads to regularization of evolving strategies, which
+hypothesise that confusability of states leads to regularisation of evolving strategies, which
 would show exactly in an increased speed of evolutionary trajectories towards well-behaved and
-regular signaling behavior.
+regular signalling behaviour.
 
@@
 strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
-signal use. \emph{Convexity} captures whether a behavioral strategy would project onto a convex
+signal use. \emph{Convexity} captures whether a behavioural strategy would project onto a convex
 pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
@@
 \paragraph{Entropy.} This classic information-theoretic notion captures the amount of
-uncertainty in a probability distribution. Roughly put, entropy of a signaling strategy
+uncertainty in a probability distribution. Roughly put, entropy of a signalling strategy
 captures inverse distance from a pure strategy. The usual definition of entropy applies
 directly to mixed strategies (see Footnote~\ref{fn:mixedStrats}), but provably equivalent
-metrics for behavioral strategies are ready to hand:
+metrics for behavioural strategies are ready to hand:
 \begin{align*}
@@
 Figure~\ref{fig:example_stratsA} is not convex, while the one in
-Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularize signaling
-strategies, like we hypothesized, we should see more convexity with increasing imprecision all
+Figure~\ref{fig:example_stratsB} is.  If confusion of states can regularise signalling
+strategies, like we hypothesised, we should see more convexity with increasing imprecision all
 else equal.
@@
 To make direct comparisons across different parameter settings, we
-normalize expected utility by the maximal amount of expected utility
+normalise expected utility by the maximal amount of expected utility
 obtainable in the relevant game. The strategy pair in
-Figure~\ref{fig:example_stratsA} has a normalized expected utility of
+Figure~\ref{fig:example_stratsA} has a normalised expected utility of
 $0.99$, the pair in Figure~\ref{fig:example_stratsB} has
@@
 
-Figure~\ref{fig:MeanMetrics} shows plots summarizing a selected part of our findings.  For
+Figure~\ref{fig:MeanMetrics} shows plots summarising a selected part of our findings.  For
 perspicuity, we only plot results for one level of tolerance $\beta = 0.1$, and leave out the
@@
 communicative success. What is more, in line with our hypothesis that mere imprecision in
-imitation behavior can lead to behavior that looks as if agents generalize across similar
+imitation behaviour can lead to behaviour that looks as if agents generalise across similar
 stimuli, higher imprecision led to a higher number of outcomes with convex sender
@@
 
-Beyond promoting convexity and convergence, diffusion also has another interesting regularizing
-effect on the evolution of signaling. There is very little variation in the recorded metrics
+Beyond promoting convexity and convergence, diffusion also has another interesting regularising
+effect on the evolution of signalling. There is very little variation in the recorded metrics
 for evolved strategies, at least for higher values of imprecision. On closer inspection, it
@@
 maximum number of rounds, and both have convex sender strategies. However, they are not equally
-efficient. In fact, the pair in Figure~\ref{fig:example_stratsC} has a normalized expected
+efficient. In fact, the pair in Figure~\ref{fig:example_stratsC} has a normalised expected
 utility of $0.99$ while the one in Figure~\ref{fig:example_stratsD} only has $0.89$.
@@
 
-Our imprecise imitation dynamic leads to by-and-large successful signaling behavior, even in
+Our imprecise imitation dynamic leads to by-and-large successful signalling behaviour, even in
 the presence of noise, that shows the hallmarks of vagueness as desired. It also gives rise to
-population-level behavior that looks as if agents are generalizing across similar
+population-level behaviour that looks as if agents are generalising across similar
 stimuli. Here, we would like to reflect briefly on some further conceptually relevant points
@@
 should consider its philosophical implications. Based on this picture, vagueness in signal use
-would seem to be characterized as a strictly population-level phenomenon since it arises in a
-signaling system from the inability of individual agents to fully align their (non-vague)
+would seem to be characterised as a strictly population-level phenomenon since it arises in a
+signalling system from the inability of individual agents to fully align their (non-vague)
 strategies because of imprecision. This is, we believe, a plausible mechanism that can already
@@
 conviction. Moreover, even if individual agents command a non-vague pure strategy, the
-realization of that pure strategy, according to our model, is bound to be vague (e.g., the same
+realisation of that pure strategy, according to our model, is bound to be vague (e.g., the same
 agent could signal differently in repeated exposure to the same state). We have used the term
 ``observation noise'' here, but it may well be that this could equally well be perceived as an
-inseparable component of an agent's ``signaling faculty.'' In this sense, then, the model might
-be compatible with a picture of agents who have internalized a vague signaling
+inseparable component of an agent's ``signalling faculty.'' In this sense, then, the model might
+be compatible with a picture of agents who have internalised a vague signalling
 strategy. It would need to be seen, however, how revision of non-deterministic individual-level
-behavior must be spelled out rigorously and whether the resulting population-level dynamic
+behaviour must be spelled out rigorously and whether the resulting population-level dynamic
 would be equivalent to our present proposal in all relevant respects.
@@
 choice options also for states similar to the ones that actually occurred. This not only leads
-to vague signaling of the appropriate kind, but also speeds up learning in such a way that,
+to vague signalling of the appropriate kind, but also speeds up learning in such a way that,
 especially for sim-max games with higher numbers of states, higher levels of communicative
@@
 \citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
-similar regularizing effects also for cases with low numbers of states and do not assume that
-agents have any kind of generalizing capacity in and of themselves, even if that is only
-implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalized reinforcement learning.
-State confusability has an effect on aggregate signaling behavior that can be described as
-generalization without generalizers: the dynamics of imprecise imitation look as if
+similar regularising effects also for cases with low numbers of states and do not assume that
+agents have any kind of generalising capacity in and of themselves, even if that is only
+implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalised reinforcement learning.
+State confusability has an effect on aggregate signalling behaviour that can be described as
+generalisation without generalisers: the dynamics of imprecise imitation look as if
 ``conclusions'' about what works for one state are ``carried over'' to similar states. This,
-however, is merely an epiphenomenon in the sense that no single agent genuinely generalizes
-over stimuli or reasons about what a more systematic signaling strategy would be.
+however, is merely an epiphenomenon in the sense that no single agent genuinely generalises
+over stimuli or reasons about what a more systematic signalling strategy would be.
 
@@
 to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
-in case of imprecise imitation, it is perception and realization of similar states. There are
-cases, then, where evolving signaling behavior, as predicted by these two approaches, is quite
+in case of imprecise imitation, it is perception and realisation of similar states. There are
+cases, then, where evolving signalling behaviour, as predicted by these two approaches, is quite
 different. Intuitively speaking, for a case with two messages, the further we venture away from
-a prototypical interpretation of either message, the less discriminative a signaling strategy
+a prototypical interpretation of either message, the less discriminative a signalling strategy
 would be when the source of ``trembles'' is the computation of expected utilities: to wit,
 since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
-response dynamics predict that senders would be almost indifferent. Sender behavior that
+response dynamics predict that senders would be almost indifferent. Sender behaviour that
 evolves under confusion of states does not have this puzzling property, because a giant would
@@
 communication if agents tend to confuse similar stimuli? To address this, we derived a
-generalization of the replicator dynamic from an agent-level process of imprecise
-imitation. The resulting population-level dynamic produced signaling behavior that is at the
+generalisation of the replicator dynamic from an agent-level process of imprecise
+imitation. The resulting population-level dynamic produced signalling behaviour that is at the
 same time regular and by-and-large communicatively efficient, while also showing the crucial
 marks of vagueness. In a sense, the model derives vagueness as a by-product of an arguably
-natural limitation on the discriminatory power of signaling agents.  Although inability to
+natural limitation on the discriminatory power of signalling agents.  Although inability to
 sharply discriminate similar stimuli may lead to vagueness and bring about a (slight) decrease
@@
 its reason: systematicity in the confusability of states (which may be a natural by-product of
-the perceptual system) supports \emph{as-if}-generalization at the population-level without
-having to assume that agents themselves have any generalization power. In this way, the
+the perceptual system) supports \emph{as-if}-generalisation at the population-level without
+having to assume that agents themselves have any generalisation power. In this way, the
 presented model extends research into the effects of transmission biases on processes of
 meaning evolution. While most previous models have focused on inductive biases of language
-learners and the regularization that these may effect
+learners and the regularisation that these may effect
 \citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un,KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
 we have shown here that shared perceptual biases, of which such an effect was not necessarily
-expected, can also regularize, facilitate and accelerate the evolution of meaning conventions.
+expected, can also regularise, facilitate and accelerate the evolution of meaning conventions.
 
@@
 
-The main idea behind \emph{imitation of success} is that agents imitate the behavior of other
+The main idea behind \emph{imitation of success} is that agents imitate the behaviour of other
 agents at some choice point with a probability that is proportional to the expected utility of
@@
 Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
-agent who gets a chance to change behavior ``learner'' and the possibly to-be-imitated agent
+agent who gets a chance to change behaviour ``learner'' and the possibly to-be-imitated agent
 ``teacher.'' A random learner is drawn from the population and given a chance to change
-behavior at choice point $\state$. The probability that our learner plays $\messg$ is
+behaviour at choice point $\state$. The probability that our learner plays $\messg$ is
 $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled teacher does at
@@
   irrelevant to our present purposes.} Confusability of states will affect how agents behave,
-how they perceive the behavior of others, and the expected utilities of behavioral
+how they perceive the behaviour of others, and the expected utilities of behavioural
 dispositions.
@@
 To keep matters simple, let us assume that agents carry pure dispositions to act. Noise can
-affect the realization of these strategies. As a sender, every agent maps states to messages:
+affect the realisation of these strategies. As a sender, every agent maps states to messages:
 these are subjectively perceived states, and no longer necessarily also the actually occurring
 states. As a receiver, every agent maps messages to state interpretations: these are intended
-interpretations that need not always be faithfully realized. This means that behavioral
+interpretations that need not always be faithfully realised. This means that behavioural
 strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
-behavioral dispositions in the population, the realization and observation of which can be
+behavioural dispositions in the population, the realisation and observation of which can be
 distorted by agents' confusion of similar states.
@@
 intends to select interpretation $\mystate{i}$, let $P_r(\mystate{r} \probbar \mystate{i})$ be
-the probability with which state $\mystate{r}$ is realized. A single round of play of a sim-max
+the probability with which state $\mystate{r}$ is realised. A single round of play of a sim-max
 game is then governed by five pieces of stochastic information, where previously there were
@@
 Expected utilities of choices at choice points should likewise take into account that actual states
-need not be observed states, and intended interpretations need not be realized
+need not be observed states, and intended interpretations need not be realised
 interpretations. First, note that
@@
 \end{align*}
-The probability that $\mystate{r}$ is realized by a random receiver in response to
+The probability that $\mystate{r}$ is realised by a random receiver in response to
 message $\messg$ is:
@@
 interpretations (receiver) by taking into consideration what the likely actual states and
-realized interpretations will be:
+realised interpretations will be:
 \begin{align*}
@@
 
-As before, suppose that senders receive a chance to change their behavior independently for a
+As before, suppose that senders receive a chance to change their behaviour independently for a
 given choice point. In the present case, this would be a chance to change how to respond to a

### 2016-05-31 paper/version_final/paper.tex
Adapted paper to BJPS template and bib style (using bibtex8 from now on for bib management).

@@
-\documentclass[fleqn,reqno,10pt]{article}
+\documentclass[12pt,english]{article}
+\usepackage{etex}
 
@@
 %\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
-%\bibliography{MyRefGlobal}
-\bibliography{../../helpers/MyRefGlobal}
-\bibliography{paper} 
 \usepackage{../../helpers/myenvironments}
 \usepackage{../../helpers/mycommands}
-\usepackage{todonotes}
+%\usepackage{todonotes}
 \usepackage{subcaption}
@@
 
-\title{Vagueness and imprecise imitation in signalling games}
-% \author{Michael Franke \and Jos\'e Pedro Correia}
+% paper size, margins, etc:
+\usepackage{geometry}
+\geometry{verbose,a4paper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
+
+\usepackage[titletoc,title]{appendix}
+
+% spacing
+\usepackage{setspace}
+\doublespacing
+
+% left justification:
+\raggedright
+
+% indentation
+\setlength\parindent{12pt}
+
+% section headings
+\usepackage{sectsty}
+\allsectionsfont{\normalsize\raggedright\centering}
+
+% table of contents stuff
+\setcounter{tocdepth}{2}
+\usepackage[]{tocloft}
+\addtocontents{toc}{\cftpagenumbersoff{section}}
+\addtocontents{toc}{\cftpagenumbersoff{subsection}}
+\renewcommand{\cftsecfont}{\itshape}
+\renewcommand{\cftsubsecfont}{\itshape}
+\makeatletter
+\renewcommand{\@cftmaketoctitle}{}
+\makeatother
+\renewcommand{\cftdot}{}
+
+% footnotes
+\usepackage[]{footmisc}
+\renewcommand\footnotelayout{\normalsize\doublespacing\raggedright}
+
+% for bibliography
+\usepackage{natbib}
+\usepackage{url}
+\makeatletter
+\def\url@leostyle{%
+  %\@ifundefined{selectfont}{
+    \def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}
+\makeatother
+\urlstyle{leo}
+
+\bibpunct
+{(} %the opening bracket symbol
+{)} %the closing bracket symbol
+{,} %the punctuation between multiple citations
+{a} %the letter `n' for numerical style, or `s' for numerical superscript style, any other letter for author-year
+{} %the punctuation that comes between the author names and the year
+{;} %the punctuation that comes between years or numbers when common author lists are suppressed
+
+% headers
+\pagestyle{plain}
+% no page numbers:
+%\pagestyle{empty}
+
+% for author affiliation
+\usepackage{authblk}
+
+% misc
+\usepackage{graphicx}
+\usepackage[english]{babel}
+
+% math packages
+\usepackage{amsmath}
+\usepackage{amsfonts}
+
+% frames
+%\usepackage{framed}
+
+\makeatletter
+\makeatother
+
+% restart equation numbering after every section:
+\numberwithin{equation}{section}
+
+% FONTS:
+%-------
+% for times roman fonts:
+\usepackage{txfonts}
+% recommended if output will have accented characters:
+\usepackage[T1]{fontenc}
+% recommended if inputting accented directly from the keyboard:
+%\usepackage[latin1]{inputenc}
+
+% BJPS specific cite commands:
+\newcommand{\citetbjps}[1]{\citeauthor{#1} ([\citeyear{#1}])}
+\newcommand{\citetbjpspost}[2]{\citeauthor{#2} ([\citeyear{#2}], #1)}
+\newcommand{\citealtbjps}[1]{\citeauthor{#1} [\citeyear{#1}]}
+\newcommand{\citepbjps}[1]{(\citeauthor{#1} [\citeyear{#1}])}
+\newcommand{\citepbjpspre}[2]{(#1~\citeauthor{#2} [\citeyear{#2}])}
+\newcommand{\citeyearbjps}[1]{[\citeyear{#1}]}
+\newcommand{\citeyearparbjps}[1]{([\citeyear{#1}])}
+
+\begin{document}
+
+\title{{\normalsize \textbf{Vagueness and imprecise imitation in signalling games}}}
+\author{{\normalsize \textbf{Michael Franke}} \and {\normalsize \textbf{Jos\'e Pedro Correia}}}
 \date{}
 
-\begin{document}
-
 \maketitle
+
+%% no headers on the first page:
+\thispagestyle{empty}
 
@@
 
+\mbox{} \\
+
+\tableofcontents
+
+\mbox{} \\
+
 \section{Introduction}
@@
 accounts that maintain that concepts and linguistic meanings evolved driven towards
-efficiency. \citet{Lipman2009:Why-is-Language} argues that, in common interest signalling
+efficiency. \citetbjps{Lipman2009:Why-is-Language} argues that, in common interest signalling
 situations, the existence of unclear borderline cases entails inefficiency of categorisation or
@@
 something that is itself useful
-\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
+(e.g.~\citealtbjps{Jaegherde-Jaegher2003:A-Game-Theoreti}; \citealtbjps{Deemter2009:Utility-and-Lan}; \citealtbjps{BlumeBoard2013:Intentional-Vag}).
 Others have argued that vagueness is a natural byproduct of limitations in information
-processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
-learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper contributes to the
+processing \citepbjpspre{e.g.}{FrankeJager2010:Vagueness-Signa} or of generalisation in low-level
+learning strategies \citepbjpspre{e.g.}{OConnor2013:The-Evolution-o}. This paper contributes to the
 latter line of thought.
@@
 natural cause of vagueness may have and that might compensate some of the disadvantages that
-vagueness may have for communicative efficiency \citep[c.f.][]{OConnor2013:The-Evolution-o}.
+vagueness may have for communicative efficiency \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.
 
@@
 % using evolutionary game theory and certain generalizations of Lewis'
-% signaling games \citep{Lewis_1969:Convention}, so called
+% signaling games \citepbjps{Lewis_1969:Convention}, so called
 % \emph{similarity-maximizing games}, or \emph{sim-max games}, for short
-% \citep{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
-% %\citep{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
+% \citepbjps{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct}.
+% %\citepbjps{FrankeJager2010:Vagueness-Signa,OConnor2013:Evolving-Percep,OConnor2013:The-Evolution-o}.
 % Our contribution is best seen in relation to these accounts, as it also
@@
 
-Signalling games, as introduced by \citet{Lewis_1969:Convention}, have
+Signalling games, as introduced by \citetbjps{Lewis_1969:Convention}, have
 a sender and a receiver. The sender knows the true state of the world,
@@
 topic that we are only beginning to fully understand
-\citep[e.g.][]{Warneryd1993:Cheap-Talk-Coor,BlumeKim1993:Evolutionary-St,Huttegger2007:Evolution-and-t,Pawlowitsch2008:Why-Evolution-d,Barrett2009:The-Evolution-o,HutteggerSkyrms2010:Evolutionary-Dy,Skyrms2010:Signals}.
+(e.g.~\citealtbjps{Warneryd1993:Cheap-Talk-Coor}; \citealtbjps{BlumeKim1993:Evolutionary-St}; \citealtbjps{Huttegger2007:Evolution-and-t}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{Barrett2009:The-Evolution-o}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}; \citealtbjps{Skyrms2010:Signals}).
 
@@
 standard in economics
-\citep[e.g.][]{Spence1973:Job-market-sign,CrawfordSobel1982:Strategic-Infor}, but have received
+(e.g.~\citealtbjps{Spence1973:Job-market-sign}; \citealtbjps{CrawfordSobel1982:Strategic-Infor}), but have received
 particular attention in a more philosophical context for reasons that will become clear
@@
 
-\citet{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of sim-max games with infinitely many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a quadratic loss function for utilities $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$ are remarkably systematic: the evolutionarily stable states are demonstrably so-called Voronoi languages.
+\citetbjps{JagerMetzger2011:Voronoi-Languag} showed that the evolutionarily stable states of sim-max games with infinitely many states in $n$-dimensional Euclidean space $\States \subseteq \mathds{R}^n$ and a quadratic loss function for utilities $\Utils(\mystate{1}, \mystate{2}) = - (\mystate{1} - \mystate{2})^2$ are remarkably systematic: the evolutionarily stable states are demonstrably so-called Voronoi languages.
 Roughly put, a Voronoi
@@
           unit interval. The pure receiver strategy $\Rpure$ selects the central elements in
-          the respective intervals. See \citet{JagerMetzger2011:Voronoi-Languag} for further
+          the respective intervals. See \citetbjps{JagerMetzger2011:Voronoi-Languag} for further
           details.}
@@
 of 2.1m `tall' as well. Another example where this intuition is additionally supported by empirical data is that of colour categorisation. The World Color Survey
-project~\citep{CookKay2005:The-World-Color,WCS} collected colour naming data for 110 unwritten
+project~(\citealtbjps{CookKay2005:The-World-Color}; \citealtbjps{WCS}) collected colour naming data for 110 unwritten
 languages of 45 language families.  For the great majority of these languages a pattern can be
-observed: basic colour terms are by and large convex~\citep{Regier07,Jager2010}. It
+observed: basic colour terms are by and large convex~(\citealtbjps{Regier07}; \citealtbjps{Jager2010}). It
 is premature to argue that these observations can be extended to \emph{all} cases of
 categorisation. However, for the cases where it does apply, the result of
-\citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
+\citetbjps{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signalling
 can impose this kind of orderly categories on a metric space without that being the ulterior
@@
 \footnote{For the concrete case of colour categorisation, see also
-  \citet{JagerRooijvan-Rooij2007:Language-Struct} and \citet{Correia2015}}
+  \citetbjps{JagerRooijvan-Rooij2007:Language-Struct} and \citetbjps{Correia2015}}
 Finally, these considerations are in line with a prominent
@@
 that meanings are connected regions on a so-called \emph{semantic map}
-\citep[e.g.][]{Croft2003:Typology-and-Un,Haspelmath2003:The-geometry-of,LevinsonMeira2003:Natural-concept}.
+(e.g.~\citealtbjps{Croft2003:Typology-and-Un}; \citealtbjps{Haspelmath2003:The-geometry-of}; \citealtbjps{LevinsonMeira2003:Natural-concept}).
 
@@
 that rely on the notion of conceptual spaces.
-\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
+\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
 natural categories are convex regions in conceptual space. If the conceptual space has a
@@
 deriving, in an independent way, not only convex categories but also prototypical exemplars
-belonging to them.  These ideas, and more, are developed further by
-\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
+belonging to them.  These ideas, and more, are developed further in
+(\citealtbjps{Jager2007:The-Evolution-o}; \citealtbjps{JagerRooijvan-Rooij2007:Language-Struct}; \citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}),
 among others.
@@
 
-\citet{DouvenDecock2011:Vagueness:-A-Co} show that
+\citetbjps{DouvenDecock2011:Vagueness:-A-Co} show that
 \citeauthor{Gardenfors2000:Conceptual-Spac}'s conceptual spaces
@@
 representing borderline
-regions. \citet{DecockDouven2012:What-is-Graded-} show further how it
+regions. \citetbjps{DecockDouven2012:What-is-Graded-} show further how it
 is possible to arrive at a gradient transition between categories, by
@@
 
-Alternative approaches are taken by, for example, \citet{FrankeJager2010:Vagueness-Signa} and
-\citet{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
+Alternative approaches are taken by, for example, \citetbjps{FrankeJager2010:Vagueness-Signa} and
+\citetbjps{OConnor2013:The-Evolution-o}, who show different ways how the above desiderata can be
 met by evolving strategies in sim-max games. To illustrate what vague signalling would look
 like, let us briefly consider a sim-max game with six equiprobable states and two messages and
-what its equilibria would be like \citep[see][for a more thorough discussion of equilibria of
-sim-max games]{Jager2007:The-Evolution-o,OConnor2013:The-Evolution-o}. We assume that utilities
+what its equilibria would be like (see \citealtbjps{Jager2007:The-Evolution-o} and \citealtbjps{OConnor2013:The-Evolution-o} for a more thorough discussion of equilibria of
+sim-max games). We assume that utilities
 are linearly decreasing with decreasing similarity. Figure~\ref{fig:EquilibriaVagueness} shows
@@
 while partial pooling can hamper the evolution of maximally efficient signal use
-\citep[e.g.][]{Hutteger:2007_Evol_Indicatives_Imperatives,Pawlowitsch2008:Why-Evolution-d,HutteggerSkyrms2010:Evolutionary-Dy},
+(e.g.~\citealtbjps{Hutteger:2007-Evol-Indicatives-Imperatives}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}),
 this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
@@
 The approach we take here is similar in spirit to that of
-\citet{FrankeJager2010:Vagueness-Signa} and \citet{OConnor2013:The-Evolution-o}, but different
+\citetbjps{FrankeJager2010:Vagueness-Signa} and \citetbjps{OConnor2013:The-Evolution-o}, but different
 in relevant detail. A more in-depth comparison is deferred until
@@
 the effects of transmission biases are isolated, as in iterated learning models
-\citep{KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
+(\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
 or when they interact with functional pressure towards efficient communication, e.g., as
 formalised in the replicator mutator dynamic
-\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}.
+(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}).
 
@@
 systematically perturbed by noise. The resulting population-level dynamic generalises the
-replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
+replicator dynamic \citepbjps{TaylorJonker1978:Evolutionary-St} in its interpretation as a cultural
 evolutionary dynamic based on imitation
-\citep[e.g.][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and}. The inclusion of
+(e.g.~\citealtbjps{Helbing1996:A-Stochastic-Be}; \citealtbjps{Schlag1998:Why-Imitate-and}). The inclusion of
 confusability of stimuli does not undermine the possibility of evolving communicative signalling
@@
 confusability vanishes, the process is just the well-known replicator dynamic
-\citep{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.
+\citepbjps{TaylorJonker1978:Evolutionary-St}, which should therefore be briefly reviewed first.
 
@@
   \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
-  not a whole lexicon. (See \citet{Cressman2003:Evolutionary-Dy} for more on the difference
+  not a whole lexicon. (See \citealtbjps{Cressman2003:Evolutionary-Dy} for more on the difference
   between dynamics on mixed or behavioural strategies.) It is in this respect that the cultural
   evolutionary dynamic introduced here differs most visibly from the replicator mutator dynamic
-  \citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un}, which
+  (e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}), which
   operates on mixed strategies and is motivated by assumptions of (asexual) biological
@@
 as proportional to their expected utilities
-\citep[e.g.][]{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
+\citepbjpspre{e.g.}{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
   Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
@@
 here on one of the simplest: \emph{imitation of success}
-\citep[see][]{Sandholm2010:Population-Game}. The intuitive idea is the following (see
+\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following (see
 Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
@@
 confused for one another.
-Human performance in these situations has been studied by a number of authors in experimental psychology. %~\citep{takane_structures_1992}.
-In a stimulus identification experiment~\citep{luce_detection_1963}, subjects are presented in each trial with a stimulus to be identified out of a fixed set.
+Human performance in these situations has been studied by a number of authors in experimental psychology. %~\citepbjps{takane_structures_1992}.
+In a stimulus identification experiment~\citepbjps{luce-detection-1963}, subjects are presented in each trial with a stimulus to be identified out of a fixed set.
 The more similar the stimuli are to each other, the larger the number of errors subjects make.
-For example, in Robert Nosofsky's experiments~\citep{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
+For example, in Robert Nosofsky's experiments~\citepbjps{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
 In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
-Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citep{donkin_why_2015}.
+Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citepbjps{donkin-why-2015}.
 The empirical data confirms not only the pervasiveness of variation in the subjects' ability to correctly identify stimuli, but also the relation between similarity and likelihood of stimulus confusion, and further suggests that the phenomenon might extend to all types of perception.
@@
 space feeds into a perceptual similarity function, as described by
-\citet{Nosofsky1986:Attention-Simil}:
+\citetbjps{Nosofsky1986:Attention-Simil}:
 \begin{align*}
@@
 quadratically decreasing in physical distance
-\citep[c.f.][]{JagerMetzger2011:Voronoi-Languag,FrankeJager2010:Vagueness-Signa}, utilities
+(c.f.~\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{FrankeJager2010:Vagueness-Signa}), utilities
 that are exponentially decreasing in negative quadratic distance can model situations where a
@@
 in intolerably far away interpretations matter very little, with a smooth transition between
-these regimes \citep[c.f.][]{OConnor2013:The-Evolution-o}.
+these regimes \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.
 
@@
 for the limiting case. This was first observed by Elliott Wagner, as mentioned by
-\citet{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
+\citetbjps{OConnor2014-OCOEPC}.  A full analysis of the dynamics of sim-max games is beyond the
 scope of this paper, but we will see shortly that diffusion from confusability of states
@@
 decrease expected utility
-\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
+\citepbjpspre{c.f.}{Lipman2009:Why-is-Language}. The crucial question is
 whether communicative success drops unacceptably fast with moderate
@@
 potential if we see it in the eyes of multilevel selection theory
-\citep{Wilson1994,OGorman2008}.
+(\citealtbjps{Wilson1994}; \citealtbjps{OGorman2008}).
 
 \subsection{Related work}
-\citet{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of
+\citetbjps{OConnor2013:The-Evolution-o} makes a proposal related to ours based on a version of
 reinforcement learning for sim-max games, in which successful play leads to reinforcement of
@@
 
-\citet{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
+\citetbjps{FrankeJager2010:Vagueness-Signa} suggested a number of ways in which
 in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
@@
 known as a \emph{soft-max response function}
-\citep[e.g.][]{Luce1959:Individual-Choi,McFadden1976:Quantal-Choice-,GoereeHolt2008:Quantal-Respons}. The
+(e.g.~\citealtbjps{Luce1959:Individual-Choi}; \citealtbjps{McFadden1976:Quantal-Choice-}; \citealtbjps{GoereeHolt2008:Quantal-Respons}). The
 main difference between this and our present approach is in where stochastic noise is assumed
@@
 We set out to meet a technical challenge posed by Lipman's problem
-\citep{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
+\citepbjps{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
 formal model that shows how vague language can evolve under selective pressure for efficient
@@
 learners and the regularisation that these may effect
-\citep[e.g.][]{NowakPlotkin2000:The-Evolution-o,NowakKomarova2001:Evolution-of-Un,KirbyHurford2002:The-Emergence-o,SmithKirby2003:Iterated-Learni,GriffithsKalish2007:Language-Evolut,KirbyGriffith2014:Iterated-Learni},
+(e.g.~\citealtbjps{NowakPlotkin2000:The-Evolution-o}; \citealtbjps{NowakKomarova2001:Evolution-of-Un}; \citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
 we have shown here that shared perceptual biases, of which such an effect was not necessarily
@@
 
-\appendix
+\begin{appendices}
 
@@
 
+\end{appendices}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\newpage
 \begin{flushright}
@@
 
-\printbibliography[heading=bibintoc]
+\bibliographystyle{bjps}
+\bibliography{../../helpers/MyRefGlobal,paper}
 

### 2016-06-05 paper/version_final/paper.tex
added acknowledgements, deleted \emph, swapped double- for single quotes

@@
   moreover, that it unifies evolutionary outcomes and helps avoid sub-optimal
-  categorisation. The upshot of this is that we see \emph{as-if}-generalisation at an aggregate
+  categorisation. The upshot of this is that we see `as-if'-generalisation at an aggregate
   level, without agents actually generalising.
@@
   in a given space. It should satisfy certain axioms that ensure behaviour that one would
-  intuitively expect from the words ``similarity'' and ``distance'' alone, but these details do
+  intuitively expect from the words `similarity' and `distance' alone, but these details do
   not matter for the purposes of this paper. The assumption that state similarity forms a
@@
 observed: basic colour terms are by and large convex~(\citealtbjps{Regier07}; \citealtbjps{Jager2010}). It
-is premature to argue that these observations can be extended to \emph{all} cases of
+is premature to argue that these observations can be extended to all cases of
 categorisation. However, for the cases where it does apply, the result of
@@
 indefinite pronouns) are preferably carved up by the languages of the world in such a way
-that meanings are connected regions on a so-called \emph{semantic map}
+that meanings are connected regions on a so-called `semantic map'
 (e.g.~\citealtbjps{Croft2003:Typology-and-Un}; \citealtbjps{Haspelmath2003:The-geometry-of}; \citealtbjps{LevinsonMeira2003:Natural-concept}).
 
-There are at least two potential ways of interpreting the signalling setup.
-Sender and receiver can be distinct
-entities, whose purpose is to communicate effectively about the actual
-state. In that case, evolving Voronoi languages would explain why
-\emph{linguistic categories} are well-behaved and orderly in the way
-they appear to be. More abstractly, sender and receiver can also be
-thought of as distinct modules in a single system, where the first
-module must discretise the information it is fed by selecting a small
-sample of, suggestively, category labels. These are passed to a second
-module that tries to decode the original information. In this case,
-evolving Voronoi languages would explain why \emph{conceptual
-  categories} are well-behaved and orderly in the way that they appear
-to be.
-Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
-that rely on the notion of conceptual spaces.
-\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
-natural categories are convex regions in conceptual space. If the conceptual space has a
+There are at least two potential ways of interpreting the signalling setup.  Sender and
+receiver can be distinct entities, whose purpose is to communicate effectively about the actual
+state. In that case, evolving Voronoi languages would explain why linguistic categories are
+well-behaved and orderly in the way they appear to be. More abstractly, sender and receiver can
+also be thought of as distinct modules in a single system, where the first module must
+discretise the information it is fed by selecting a small sample of, suggestively, category
+labels. These are passed to a second module that tries to decode the original information. In
+this case, evolving Voronoi languages would explain why conceptual categories are well-behaved
+and orderly in the way that they appear to be.  Seen in this light, sim-max games may provide a
+foundation to approaches in cognitive semantics that rely on the notion of conceptual spaces.
+\citetbjpspost{pp.~70--77}{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued
+that natural categories are convex regions in conceptual space. If the conceptual space has a
 suitable metric, convex categories can be derived from a set of prototypes. The category
 corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
-other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorisation
-system can be obtained: storing the prototypes lets us recover the categories without having to
-store each category's extension. However, what is left unexplained is where the prototypes
-come from, and why we would not see just any distribution of prototypes as an equally efficient
-classification system. This is where sim-max games can contribute a principled approach to
-deriving, in an independent way, not only convex categories but also prototypical exemplars
-belonging to them.  These ideas, and more, are developed further in
-(\citealtbjps{Jager2007:The-Evolution-o}; \citealtbjps{JagerRooijvan-Rooij2007:Language-Struct}; \citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}),
-among others.
+other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient
+categorisation system can be obtained: storing the prototypes lets us recover the categories
+without having to store each category's extension. However, what is left unexplained is where
+the prototypes come from, and why we would not see just any distribution of prototypes as an
+equally efficient classification system. This is where sim-max games can contribute a
+principled approach to deriving, in an independent way, not only convex categories but also
+prototypical exemplars belonging to them.  These ideas, and more, are developed further in
+(\citealtbjps{Jager2007:The-Evolution-o};
+\citealtbjps{JagerRooijvan-Rooij2007:Language-Struct};
+\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}), among
+others.
 
@@
 regions in conceptual space, a construction algorithm is available
-that yields ``collated Voronoi diagrams'' with thick boundaries
+that yields `collated Voronoi diagrams' with thick boundaries
 representing borderline
@@
 input. Over the course of many generations, the effects of such biases can, in a manner of
-speaking, ``accumulate'' and lead to surprising results, such as the evolution of compositional
+speaking, `accumulate' and lead to surprising results, such as the evolution of compositional
 form-meaning mappings or the use of regular recursive syntactic structure. This can happen when
@@
 $\Util \mycolon \States \times \States \rightarrow \mathds{R}$ the utility function shared by
-senders and receivers in the population. A \emph{behavioural strategy} is a function that maps
+senders and receivers in the population. A behavioural strategy is a function that maps
 an agent's choice points to a probability distribution over available
-choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioural strategies, not \emph{mixed
-    strategies}, i.e., probability distributions over functions from each choice point to an
+choices.\footnote{\label{fn:mixedStrats}Our focus is on behavioural strategies, not mixed
+    strategies, i.e., probability distributions over functions from each choice point to an
   act, such as $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioural strategies assume
@@
   protocols of the kind we consider here: agents only observe how, on some occasion, some other
-  agent behaved in \emph{one} particular situation, not how that agent would behave in
-  \emph{all} relevant choice situations; they imitate the use of a single word, so to speak,
+  agent behaved in one particular situation, not how that agent would behave in
+  all relevant choice situations; they imitate the use of a single word, so to speak,
   not a whole lexicon. (See \citealtbjps{Cressman2003:Evolutionary-Dy} for more on the difference
@@
 
-The \emph{expected utility} of choices at each choice point is:\footnote{The notation
-  $\propto$, for ``proportional to'', that is used here and hereafter means that the right-hand
+The expected utility of choices at each choice point is:\footnote{The notation
+  $\propto$, for `proportional to', that is used here and hereafter means that the right-hand
   side might still need to be normalised so as to have probabilities sum to one. In general,
@@
 For a given choice point, say a state $\state$, the probability of seeing $\messg$ played by an
-average agent in the population \emph{after} the update is proportional to the probability of
-seeing it \emph{before} the update, which is $\Sstrat(\messg \probbar \state)$, times the
+average agent in the population after the update is proportional to the probability of
+seeing it before the update, which is $\Sstrat(\messg \probbar \state)$, times the
 expected utility of $\messg$ at state $\state$. Intuitively put, frequencies of choices change
@@
 deriving the replicator dynamic from agent-level processes of behavioural adaptation. We focus
-here on one of the simplest: \emph{imitation of success}
+here on one of the simplest: imitation of success
 \citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following (see
@@
 update scheme).  Every now and then a random agent gets a chance to alter his behaviour for one
-of his choice points. Say, this is a sender who gets to ``reconsider'' his choice of message
+of his choice points. Say, this is a sender who gets to `reconsider' his choice of message
 for state $\state$. The revising agent then observes what some random agent does at $\state$,
@@
 probabilities are moderate and regular, in that they track similarity of states, we might
-expect a regularising effect also on evolving signalling strategies. Indeed, we hypothesise that
-confusion of states can give rise to population-level aggregate behaviour that looks \emph{as
-  if} information about what is good for one state percolates also to similar states. In other
-words, imprecise imitation may make signalling behaviour on average look as if agents generalise
-across similar states, even if no agent actually generalises. To explore whether confusion of
-states can have this effect, we turn to numerical simulation.
+expect a regularising effect also on evolving signalling strategies. Indeed, we hypothesise
+that confusion of states can give rise to population-level aggregate behaviour that looks as
+if information about what is good for one state percolates also to similar states. In other
+words, imprecise imitation may make signalling behaviour on average look as if agents
+generalise across similar states, even if no agent actually generalises. To explore whether
+confusion of states can have this effect, we turn to numerical simulation.
 
@@
 \end{align*}
-This choice of utility function is governed partly by convenience, but also because we believe it has the right general properties for a communicative payoff function.
-Unlike utilities that are, say, linearly or
-quadratically decreasing in physical distance
-(c.f.~\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{FrankeJager2010:Vagueness-Signa}), utilities
-that are exponentially decreasing in negative quadratic distance can model situations where a
-small amount of imprecision in communication is tolerable, whereas similarly small differences
-in intolerably far away interpretations matter very little, with a smooth transition between
-these regimes \citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.
+With $\toler = 0$, we regain the case of Lewis' games where only a perfect match of actual
+state $\mystate{i}$ with receiver interpretation $\mystate{j}$ leads to a positive payoff. The
+higher $\toler$ the more acceptable a wider environment of interpretations $\mystate{j}$ around
+$\mystate{i}$ is (see Figure~\ref{fig:NosofskySim}).  This choice of utility function is
+governed partly by convenience, but also because we believe it has the right general properties
+for a communicative payoff function.  Unlike utilities that are, say, linearly or quadratically
+decreasing in physical distance (c.f.~\citealtbjps{JagerMetzger2011:Voronoi-Languag};
+\citealtbjps{FrankeJager2010:Vagueness-Signa}), utilities that are exponentially decreasing in
+negative quadratic distance can model situations where a small amount of imprecision in
+communication is tolerable, whereas similarly small differences in intolerably far away
+interpretations matter very little, with a smooth transition between these regimes
+\citepbjpspre{c.f.}{OConnor2013:The-Evolution-o}.
 
@@
 therefore only a categorical measure for reaching a certain (well-considered, but eventually
-arbitrary) degree of stability. In other words, our notion of ``convergence'' is a measure of
+arbitrary) degree of stability. In other words, our notion of `convergence' is a measure of
 relative speed: is it true that the system reached a state in which evolutionary adaptations
@@
 in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99
-rounds. The sender strategy shows a smooth blending from one ``category'' to the other, and the
+rounds. The sender strategy shows a smooth blending from one `category' to the other, and the
 receiver's interpretations are rather extended curves, peaking at a central point in the
-relevant ``categories.''
+relevant `categories.'
 
@@
 capture how vague, generally well-structured, and communicatively efficient the recorded
-strategy pairs were. \emph{Entropy} captures the amount of systematicity or regularity in
-signal use. \emph{Convexity} captures whether a behavioural strategy would project onto a convex
-pure strategy. \emph{Expected utility} measures the communicative efficiency of evolved
-strategy pairs.
+strategy pairs were. Entropy captures the amount of systematicity or regularity in signal
+use. Convexity captures whether a behavioural strategy would project onto a convex pure
+strategy. Expected utility measures the communicative efficiency of evolved strategy pairs.
 
@@
 $\text{HD}(\Sstrat^*,\Sstrat')$ where $\Sstrat^*$ is $\Sstrat$ with reversed message
-indices. An example of the \emph{within group distance}, i.e., the average distances between
+indices. An example of the `within group distance,' i.e., the average distances between
 all sender strategies obtained for the same parameter values, is plotted in
@@
 agent could signal differently in repeated exposure to the same state). We have used the term
-``observation noise'' here, but it may well be that this could equally well be perceived as an
-inseparable component of an agent's ``signalling faculty.'' In this sense, then, the model might
+`observation noise' here, but it may well be that this could equally well be perceived as an
+inseparable component of an agent's `signalling faculty.' In this sense, then, the model might
 be compatible with a picture of agents who have internalised a vague signalling
@@
 imprecision could be considered an evolutionarily beneficial side-effect. A certain degree of
-imprecision can lead to higher \emph{within group expected utility}, defined as the average
+imprecision can lead to higher `within group expected utility', defined as the average
 expected utility that each evolved language scored when playing against an arbitrary other
@@
 generalisation without generalisers: the dynamics of imprecise imitation look as if
-``conclusions'' about what works for one state are ``carried over'' to similar states. This,
+`conclusions' about what works for one state are `carried over' to similar states. This,
 however, is merely an epiphenomenon in the sense that no single agent genuinely generalises
@@
 in\-for\-ma\-tion-processing limitations could lead to vague strategies. The model that is most
-clearly related to the present approach uses the notion of a \emph{quantal response}, also
-known as a \emph{soft-max response function}
+clearly related to the present approach uses the notion of a quantal response, also
+known as a soft-max response function
 (e.g.~\citealtbjps{Luce1959:Individual-Choi}; \citealtbjps{McFadden1976:Quantal-Choice-}; \citealtbjps{GoereeHolt2008:Quantal-Respons}). The
@@
 a prototypical interpretation of either message, the less discriminative a signalling strategy
-would be when the source of ``trembles'' is the computation of expected utilities: to wit,
-since both ``tall'' and ``short'' are almost equally bad descriptions for a giant, quantal
+would be when the source of `trembles' is the computation of expected utilities: to wit,
+since both `tall' and `short' are almost equally bad descriptions for a giant, quantal
 response dynamics predict that senders would be almost indifferent. Sender behaviour that
@@
 its reason: systematicity in the confusability of states (which may be a natural by-product of
-the perceptual system) supports \emph{as-if}-generalisation at the population-level without
+the perceptual system) supports `as-if'-generalisation at the population-level without
 having to assume that agents themselves have any generalisation power. In this way, the
@@
 
-The main idea behind \emph{imitation of success} is that agents imitate the behaviour of other
+The main idea behind imitation of success is that agents imitate the behaviour of other
 agents at some choice point with a probability that is proportional to the expected utility of
@@
 Let's consider sender strategies, as the receiver case is parallel. Call (misleadingly!) the
-agent who gets a chance to change behaviour ``learner'' and the possibly to-be-imitated agent
-``teacher.'' A random learner is drawn from the population and given a chance to change
+agent who gets a chance to change behaviour `learner' and the possibly to-be-imitated agent
+`teacher.' A random learner is drawn from the population and given a chance to change
 behaviour at choice point $\state$. The probability that our learner plays $\messg$ is
@@
 
-If agents get repeated update chances for their choice points, the \emph{expected change} of
+If agents get repeated update chances for their choice points, the expected change of
 frequency of $\messg$-choices at $\state$ becomes:
@@
 \end{align}
-where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
-any $\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
+where $P(\messg' \rightarrow m, \state)$ is the `inflow' probability that agents switch from
+any $\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the `outflow' probability
 that agents switch from $\messg$ to any $\messg'$. Since we are dealing with expectations in a
@@
 interpretations that need not always be faithfully realised. This means that behavioural
-strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of \emph{actual}
+strategies $\Sstrat$ and $\Rstrat$ represent the average proportions of actual
 behavioural dispositions in the population, the realisation and observation of which can be
@@
 \end{align*}
-The ``inflow'' and ``outflow'' probabilities $P(\messg' \rightarrow \messg, \state)$ and
+The `inflow' and `outflow' probabilities $P(\messg' \rightarrow \messg, \state)$ and
 $P(\messg' \rightarrow \messg, \state)$ that a randomly sampled learner switches from any
@@
 \end{appendices}
+
+\section*{Acknowledgements}
+
+We would like to thank Gerhard J\"{a}ger, \textcolor{red}{\dots who else to thank? - Pedro
+  thesis?} and several anonymous reviewers for their insightful comments and suggestions on
+earlier versions of this paper. Funding for this research was provided by the Dutch Research
+Council (NWO-VENI grant 275-80-004 to Michael Franke), the Institutional Strategy of the
+University of T\"{u}bingen (German Research Foundation ZUK 63 to Michael Franke), the Priority
+Program XPrag.de (German Research Foundation Schwerpunktprogramm 1727 to Michael Franke), and
+\dots \textcolor{red}{Pedro funding?}.
 

### 2016-06-05 paper/version_final/paper.tex
added footnote and references on stochastic game theory

@@
 dynamic which is derived from the idea that agents imitate other agents' behaviour while
-possibly confusing similar states. This imprecise imitation dynamic is explored in
-Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares our
-approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
+possibly confusing similar states.\footnote{Much previous work has investigated the interplay
+  of adaptive dynamics, including imitation-based update protocols, on the one hand, and noise
+  or mutation, on the other hand (e.g., \citealtbjps{FosterYoung1990:Stochastic-Evol};
+  \citealtbjps{FudenbergHarris1992:Evolutionary-Dy};
+  \citealtbjps{KandoriMailath1993:Learning-Mutati}; \citealtbjps{Young1993:The-Evolution-o};
+  \citealtbjps{FudenbergImhof2006:Imitation-Proce}). While this line of research often looks at
+  finite populations and the extreme long-term behavior of the system under generic randomness,
+  the focus here is on a quite particular source of stochastic noise and its quite particular
+  role in the evolution of meanings through signalling.}  This imprecise imitation dynamic is
+explored in Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares
+our approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
 

### 2016-06-07 paper/version_final/paper.tex
moved floats to end (really?)
Small changes after final reading. Fixed some bibliography issues.

@@
 \usepackage{a4wide}
+
+\usepackage[nomarkers, nolists]{endfloat}
 
@@
 clear cases that do not, and also so-called borderline cases. Borderline cases do not clearly
-apply, nor do they clearly not apply, and there may be differences between borderline cases in
+belong to the category, nor do they clearly not belong, and there may be differences between borderline cases in
 terms of how well they represent the category in question. Vagueness does not seem to
@@
 
-But there are other intriguing aspects about vagueness. One perplexing issue is how vagueness
+There are other intriguing aspects about vagueness. One perplexing issue is how vagueness
 could arise and be maintained in the first place. This is an apparent puzzle for functionalist
@@
 perception and language together, may be an almost optimal adaptation in a larger frame
-of reference, e.g., if we take into account the metabolic costs for increased perceptual
+of reference, for example if we take into account the metabolic costs for increased perceptual
 accuracy. We will not engage in such speculation here. Rather we will explore the consequences
@@
 other words, in order to address Lipman's challenge seriously, and not just hand-wave it away
-by appeal to the naturalness of vagueness, it needs to be spelled out how exactly confusability
-of stimuli could come into play in a process of meaning evolution and how it could lead to
+by appeal to the naturalness of vagueness, one needs to spell out how exactly could confusability
+of stimuli come into play in a process of meaning evolution and how it could lead to
 vague but by-and-large informative signal meaning. The formal model that this paper introduces
@@
   of adaptive dynamics, including imitation-based update protocols, on the one hand, and noise
-  or mutation, on the other hand (e.g., \citealtbjps{FosterYoung1990:Stochastic-Evol};
+  or mutation, on the other hand (e.g.~\citealtbjps{FosterYoung1990:Stochastic-Evol};
   \citealtbjps{FudenbergHarris1992:Evolutionary-Dy};
@@
 explored in Section~\ref{sec:exploring-rdd}. Section~\ref{sec:discussion} reflects and compares
-our approach to others. Appendix~\ref{sec:formal-stuff} provides formal detail.
+our approach to that of others. Appendix~\ref{sec:formal-stuff} provides formal detail.
 
@@
 to certain states or acts. To investigate the conditions under which
-such meaning-generating behaviour can evolve is a highly interesting
+such meaning-generating behaviour can evolve is a
 topic that we are only beginning to fully understand
@@
 principled approach to deriving, in an independent way, not only convex categories but also
-prototypical exemplars belonging to them.  These ideas, and more, are developed further in
-(\citealtbjps{Jager2007:The-Evolution-o};
-\citealtbjps{JagerRooijvan-Rooij2007:Language-Struct};
-\citealtbjps{JagerMetzger2011:Voronoi-Languag}; \citealtbjps{OConnor2014-OCOEPC}), among
+prototypical exemplars belonging to them.  These ideas, and more, are developed further by
+\citetbjps{Jager2007:The-Evolution-o}, \citetbjps{JagerRooijvan-Rooij2007:Language-Struct}, \citetbjps{JagerMetzger2011:Voronoi-Languag}, \citetbjps{OConnor2014-OCOEPC}, among
 others.
@@
 \subsection{Vagueness in sim-max games and conceptual spaces}
+\label{sec:vagueness-in-sim-max-games}
 
 This outline of an approach to categorisation using sim-max games
-leaves many problems unaddressed. One of them is that natural
-categories for continuously variable stimuli usually do not have
-unique, point-valued prototypes and clear category boundaries. We
+leaves some problems unaddressed. One of them is that, usually, natural
+categories for continuously variable stimuli neither have
+clear boundaries, nor do they have unique, point-valued prototypes. We
 would like to account for the possibility of such vagueness, in
@@
 while partial pooling can hamper the evolution of maximally efficient signal use
-(e.g.~\citealtbjps{Hutteger:2007-Evol-Indicatives-Imperatives}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}),
+(\citealtbjps{Hutteger:2007-Evol-Indicatives-Imperatives}; \citealtbjps{Pawlowitsch2008:Why-Evolution-d}; \citealtbjps{HutteggerSkyrms2010:Evolutionary-Dy}),
 this is orthogonal to our concerns about vagueness. A regular and natural yet vague signal use
@@
 error, noise or imprecision in the transmission of linguistic behaviour, knowledge or traits. As
-an example for the latter, consider a child learning syntactic rules from a parent
+an example of the latter, consider a child learning syntactic rules from a parent
 generation. The child must infer these unobservable rules from observable speech. Inductive
@@
 (\citealtbjps{KirbyHurford2002:The-Emergence-o}; \citealtbjps{SmithKirby2003:Iterated-Learni}; \citealtbjps{GriffithsKalish2007:Language-Evolut}; \citealtbjps{KirbyGriffith2014:Iterated-Learni}),
-or when they interact with functional pressure towards efficient communication, e.g., as
+or when they interact with functional pressure towards efficient communication, for example as
 formalised in the replicator mutator dynamic
@@
 population of agents that try to imitate other agents' signalling behaviour. We assume, however,
-that both realisation of behaviour, as well as observability of others' behaviour, is
+that both observation of others' behaviour, as well as realisation of behaviour, are both
 systematically perturbed by noise. The resulting population-level dynamic generalises the
@@
 making it appear as if agents were applying inductive biases or generalising over partial
-observations, when in truth this is the sole effect of confusion of similar stimuli.
+observations, when this is actually the sole effect of confusion of similar stimuli.
 
@@
 \begin{align*}
-  \EU(\messg, \state, \Rstrat) & = \sum_{\state \in \States}
+  \EU(\messg, \state, \Rstrat) & = \sum_{\state' \in \States}
   \Rstrat(\state' \probbar \messg) \myts U(\state, \state') \\
@@
 as proportional to their expected utilities
-\citepbjpspre{e.g.}{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
+\citepbjps{HofbauerSigmund1998:Evolutionary-Ga}:\footnote{The formulation given in
   Equation~(\ref{eq:RDDiscrete}) is adequate only for cases like the one we will be looking at
@@
 here on one of the simplest: imitation of success
-\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following (see
+\citepbjpspre{see}{Sandholm2010:Population-Game}. The intuitive idea is the following: every now and then a random agent gets a chance to alter his behaviour for one
+of his choice points (say this is a sender who gets to `reconsider' his choice of message
+for state $\state$); the revising agent then observes what some random agent does at $\state$,
+say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
+expected utility of $\messg$ for state $\state$. See
 Appendix~\ref{sec:cond-imit-succ} for a derivation of the standard replicator dynamic from this
-update scheme).  Every now and then a random agent gets a chance to alter his behaviour for one
-of his choice points. Say, this is a sender who gets to `reconsider' his choice of message
-for state $\state$. The revising agent then observes what some random agent does at $\state$,
-say $\messg$, and will henceforth play $\messg$ in $\state$ with a probability given by the
-expected utility of $\messg$ for state $\state$.
+update scheme.
 
@@
 The more similar the stimuli are to each other, the larger the number of errors subjects make.
-For example, in Robert Nosofsky's experiments~\citepbjps{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
+For example, in Robert Nosofsky's experiments~\citeyearparbjps{Nosofsky1986:Attention-Simil}, stimuli consisted of 16 semicircles with a radial line from the centre to the rim, varying in size (radius of 0.478, 0.500, 0.522, or 0.544 cm) and angle of the line (50º, 53º, 56º, or 59º).
 In over 9000 trials, the two subjects identified the stimulus correctly only approximately 44\% and 35\% of the time.
-Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citepbjps{donkin-why-2015}.
+Similar experiments have been conducted with other identification tasks, for example, for frequency and intensity of tones, taste, hue of colours, and magnitude of lines and areas~\citepbjpspre{see}{donkin-why-2015}.
 The empirical data confirms not only the pervasiveness of variation in the subjects' ability to correctly identify stimuli, but also the relation between similarity and likelihood of stimulus confusion, and further suggests that the phenomenon might extend to all types of perception.
@@
 The presence of observation and realisation noise also affects imitation of successes. Here, we
-focus on the main ideas. Formal detail is provided by Appendix~\ref{sec:imit-succ-with}. Take a
+focus on the main ideas, formal detail is provided in Appendix~\ref{sec:imit-succ-with}. Take a
 sender who gets to revise behaviour for state $\state$. What agents can plausibly revise by
@@
 but we stick with Nosofsky's similarity function for the time being, because it is
-mathematically simple, yet an established notion in mathematical psychology.
+mathematically simple, and an established notion in mathematical psychology.
 
@@
 relative speed: is it true that the system reached a state in which evolutionary adaptations
-had slowed down almost to a halt before 200 update steps? This is motivated by pragmatic
-concerns regarding length of simulation time, but also theoretically justified, because we
+had slowed down almost to a halt before 200 update steps? This is motivated by practical
+concerns regarding length of simulation time, but also theoretically justifiable, because we
 hypothesise that confusability of states leads to regularisation of evolving strategies, which
@@
 These examples already show two interesting things. Firstly, inclusion of imprecision can lead
-to seemingly well-behaved, yet vague strategies in the sense that we are after. The sender
+to seemingly well-behaved, yet vague strategies in the sense that we are after (see again Section~\ref{sec:vagueness-in-sim-max-games}). The sender
 strategy in Figure~\ref{fig:example_stratsB} identifies clear positive and clear negative cases
@@
 non-convex pure strategies under the replicator dynamic and linger there for vast amounts of
-time, possibly indefinitely. We see this in our limited-time simulations (e.g.,
-Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
+time, possibly indefinitely. We see this in our limited-time simulations (e.g.~Figure~\ref{fig:example_stratsA}), but this also holds, for some types of utility functions,
 for the limiting case. This was first observed by Elliott Wagner, as mentioned by
@@
 the unit interval.
+This is then reflected in the `within group expected utility', defined as the average
+expected utility that each evolved language scored when playing against an arbitrary other
+language obtained for the same parameter values. Figure~\ref{fig:GroupMeasuresB} gives a
+representative example.
 
@@
 conviction. Moreover, even if individual agents command a non-vague pure strategy, the
-realisation of that pure strategy, according to our model, is bound to be vague (e.g., the same
-agent could signal differently in repeated exposure to the same state). We have used the term
-`observation noise' here, but it may well be that this could equally well be perceived as an
+realisation of that pure strategy, according to our model, is bound to be vague: the same
+agent could signal differently in repeated exposure to the same state because of the non-deterministic nature of observation noise. We have used the term
+`observation noise' here, but this could equally well be interpreted as an
 inseparable component of an agent's `signalling faculty.' In this sense, then, the model might
@@
 
-The phenomenon could potentially not just be a side effect.  Given the benefit of a certain
+The phenomenon could potentially be more than just a side effect.  Given the benefit of a certain
 amount of imprecision we observe when comparing within group expected utility, it would be
@@
 success are reached in shorter learning periods. Our results complement and extend
-\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that we obtain
-similar regularising effects also for cases with low numbers of states and do not assume that
+\citeauthor{OConnor2013:The-Evolution-o}'s. The most important differences are that (i) we obtain
+similar regularising effects also for cases with low numbers of states and (ii) we do not assume that
 agents have any kind of generalising capacity in and of themselves, even if that is only
-implicit, as in \citeauthor{OConnor2013:The-Evolution-o}'s generalised reinforcement learning.
+implicit in \citeauthor{OConnor2013:The-Evolution-o}'s generalised reinforcement learning.
 State confusability has an effect on aggregate signalling behaviour that can be described as
@@
 main difference between this and our present approach is in where stochastic noise is assumed
-to reside. In case of a quantal response dynamic, it is the computation of expected utilities;
-in case of imprecise imitation, it is perception and realisation of similar states. There are
+to reside. In case of a quantal response dynamic, it resides in the computation of expected utilities;
+in case of imprecise imitation, it resides in perception and realisation of similar states. There are
 cases, then, where evolving signalling behaviour, as predicted by these two approaches, is quite
@@
 We set out to meet a technical challenge posed by Lipman's problem
-\citepbjps{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
+\citeyearparbjps{Lipman2009:Why-is-Language}: is there a conceptually sound and mathematically coherent
 formal model that shows how vague language can evolve under selective pressure for efficient
@@
 in communicative efficiency, there may also be an advantage, not of vagueness itself, but of
-its reason: systematicity in the confusability of states (which may be a natural by-product of
+its cause. Systematicity in the confusability of states (which may be a natural by-product of
 the perceptual system) supports `as-if'-generalisation at the population-level without
@@
 
-We would like to thank Gerhard J\"{a}ger, \textcolor{red}{\dots who else to thank? - Pedro
-  thesis?} and several anonymous reviewers for their insightful comments and suggestions on
+We would like to thank Gerhard J\"{a}ger and several anonymous reviewers for their insightful comments and suggestions on
 earlier versions of this paper. Funding for this research was provided by the Dutch Research
 Council (NWO-VENI grant 275-80-004 to Michael Franke), the Institutional Strategy of the
-University of T\"{u}bingen (German Research Foundation ZUK 63 to Michael Franke), the Priority
-Program XPrag.de (German Research Foundation Schwerpunktprogramm 1727 to Michael Franke), and
-\dots \textcolor{red}{Pedro funding?}.
+University of T\"{u}bingen (German Research Foundation ZUK 63 to Michael Franke), and the Priority
+Program XPrag.de (German Research Foundation Schwerpunktprogramm 1727 to Michael Franke).
 

### 2016-06-09 paper/version_final/paper.tex
Added reference to PhD grant.

@@
 Council (NWO-VENI grant 275-80-004 to Michael Franke), the Institutional Strategy of the
-University of T\"{u}bingen (German Research Foundation ZUK 63 to Michael Franke), and the Priority
-Program XPrag.de (German Research Foundation Schwerpunktprogramm 1727 to Michael Franke).
+University of T\"{u}bingen (German Research Foundation ZUK 63 to Michael Franke), the Priority
+Program XPrag.de (German Research Foundation Schwerpunktprogramm 1727 to Michael Franke), and the Fundação para a Ciência	e a Tecnologia (PhD studentship SFRH/BD/100437/2014 to José Pedro Correia).
 

### 2016-06-10 paper/version_final/paper_clean_format.tex
final version

--- 2016-06-09 paper/version_final/paper.tex
@@
-\documentclass[12pt,english]{article}
+\documentclass[11pt,english]{article}
 \usepackage{etex}
@@
 
-\usepackage[]{../../helpers/mypackages}
+\usepackage[]{mypackages}
 %\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false]{biblatex}
-\usepackage{../../helpers/myenvironments}
-\usepackage{../../helpers/mycommands}
+\usepackage{myenvironments}
+\usepackage{mycommands}
 %\usepackage{todonotes}
@@
 
-\usepackage[nomarkers, nolists]{endfloat}
+% \usepackage[nomarkers, nolists]{endfloat}
 
@@
 % spacing
-\usepackage{setspace}
-\doublespacing
+% \usepackage{setspace}
+% \doublespacing
 
 % left justification:
-\raggedright
-
-% indentation
-\setlength\parindent{12pt}
-
-% section headings
-\usepackage{sectsty}
-\allsectionsfont{\normalsize\raggedright\centering}
-
-% table of contents stuff
-\setcounter{tocdepth}{2}
-\usepackage[]{tocloft}
-\addtocontents{toc}{\cftpagenumbersoff{section}}
-\addtocontents{toc}{\cftpagenumbersoff{subsection}}
-\renewcommand{\cftsecfont}{\itshape}
-\renewcommand{\cftsubsecfont}{\itshape}
-\makeatletter
-\renewcommand{\@cftmaketoctitle}{}
-\makeatother
-\renewcommand{\cftdot}{}
-
-% footnotes
-\usepackage[]{footmisc}
-\renewcommand\footnotelayout{\normalsize\doublespacing\raggedright}
+
+% % indentation
+% \setlength\parindent{12pt}
+
+% % section headings
+% \usepackage{sectsty}
+% \allsectionsfont{\normalsize\raggedright\centering}
+
+% % table of contents stuff
+% \setcounter{tocdepth}{2}
+% \usepackage[]{tocloft}
+% \addtocontents{toc}{\cftpagenumbersoff{section}}
+% \addtocontents{toc}{\cftpagenumbersoff{subsection}}
+% \renewcommand{\cftsecfont}{\itshape}
+% \renewcommand{\cftsubsecfont}{\itshape}
+% \makeatletter
+% \renewcommand{\@cftmaketoctitle}{}
+% \makeatother
+% \renewcommand{\cftdot}{}
+
+% % footnotes
+% \usepackage[]{footmisc}
+% \renewcommand\footnotelayout{\normalsize\doublespacing\raggedright}
 
@@
 % for times roman fonts:
-\usepackage{txfonts}
+% \usepackage{txfonts}
 % recommended if output will have accented characters:
-\usepackage[T1]{fontenc}
+% \usepackage[T1]{fontenc}
 % recommended if inputting accented directly from the keyboard:
@@
 % BJPS specific cite commands:
-\newcommand{\citetbjps}[1]{\citeauthor{#1} ([\citeyear{#1}])}
-\newcommand{\citetbjpspost}[2]{\citeauthor{#2} ([\citeyear{#2}], #1)}
-\newcommand{\citealtbjps}[1]{\citeauthor{#1} [\citeyear{#1}]}
-\newcommand{\citepbjps}[1]{(\citeauthor{#1} [\citeyear{#1}])}
-\newcommand{\citepbjpspre}[2]{(#1~\citeauthor{#2} [\citeyear{#2}])}
-\newcommand{\citeyearbjps}[1]{[\citeyear{#1}]}
-\newcommand{\citeyearparbjps}[1]{([\citeyear{#1}])}
+\newcommand{\citetbjps}[1]{\citeauthor{#1} (\citeyear{#1})}
+\newcommand{\citetbjpspost}[2]{\citeauthor{#2} (\citeyear{#2}, #1)}
+\newcommand{\citealtbjps}[1]{\citeauthor{#1} \citeyear{#1}}
+\newcommand{\citepbjps}[1]{(\citeauthor{#1} \citeyear{#1})}
+\newcommand{\citepbjpspre}[2]{(#1~\citeauthor{#2} \citeyear{#2})}
+\newcommand{\citeyearbjps}[1]{\citeyear{#1}}
+\newcommand{\citeyearparbjps}[1]{(\citeyear{#1})}
 
@@
 \title{{\normalsize \textbf{Vagueness and imprecise imitation in signalling games}}}
-\author{{\normalsize \textbf{Michael Franke}} \and {\normalsize \textbf{Jos\'e Pedro Correia}}}
+\author{{\normalsize {Michael Franke \& Jos\'e Pedro Correia}}}
 \date{}
@@
 \end{abstract}
-
-\mbox{} \\
-
-\tableofcontents
-
-\mbox{} \\
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newpage
-\begin{flushright}
-Michael Franke \\
-Wilhelmstra\ss e 19 \\
-Department of Linguistics \\
-University of T\"ubingen \\
-72072 T\"u bingen \\
-Germany \\
-\texttt{mchfranke@gmail.com} \\
-\bigskip
-Jos\'e Pedro Correia \\
-Via Panorâmica s/n \\
-Institute of Philosophy \\
-Faculty of Arts \\
-University of Porto \\
-4150-564 Porto \\
-Portugal \\
-\texttt{zepedro.correia@gmail.com} \\
-\end{flushright}
-
-\bibliographystyle{bjps}
-\bibliography{../../helpers/MyRefGlobal,paper}
+
+\bibliographystyle{apalike}
+\bibliography{bibexport}
 

### 2017-01-21 book_chapter/paper.tex
First version, only with abstract and some chapter titles.

\documentclass{article}
\usepackage{todonotes}
\usepackage{a4wide}
\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false]{biblatex}
\bibliography{paper}

\begin{document}

\title{Natural sources of vagueness and their implications}
\author{Michael Franke \and Jos\'e Pedro Correia}
\date{}

\maketitle

\begin{abstract}
A vexing puzzle about vagueness, rationality and evolution runs, in crude abbreviation, as follows: vague language use is demonstrably suboptimal if the goal is efficient precise and cooperative information transmission; hence rational deliberation or evolutionary selection should, under this assumed goal, eradicate vagueness from language use.
Since vagueness is persistent in all human languages, something has to give.
In this paper, we investigate a number of reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary optimal signaling.
We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness, and how, given vagueness, natural linguistic practices are likely to create more reason for vagueness.
We explore the consequences of these reasons and mechanism for a notion of meaning and a notion of language.
\end{abstract}

\tableofcontents

\section{Vaguenes}
\label{sec:vagueness}

\section{Paradigm shift}
\label{sec:Wittgenstein-and-signaling}

\section{Signaling models of vagueness}
\label{sec:vague-signaling}

\printbibliography

\end{document}

### 2017-01-23 book_chapter/paper.tex
Classical problem statement.

@@
-\documentclass{article}
+\documentclass[a4paper]{article}
+
+\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false]{biblatex}
+\bibliography{paper}
+
 \usepackage{todonotes}
 \usepackage{a4wide}
-\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false]{biblatex}
-\bibliography{paper}
+\usepackage{amssymb}
+\usepackage{syllogism}
 
@@
 
-\section{Vaguenes}
+\section{Vagueness}
 \label{sec:vagueness}
+
+The classical philosophical problem of vagueness is most starkly embodied by the Sorites paradox.
+The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\citep{sorensen_sorites_2009}, and uses the example of a heap of sand: if one million grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then by repeated applications of this reasoning we are lead to acknowledge that one single grain of sand also forms a heap.
+This paradox can be stated in a very general way.
+We start with an $x$ to which a certain predicate $P$ clearly applies.
+We observe that a certain transformation $f$ has no impact on whether something is or is not $P$.
+Thus we are lead to acknowledge that the predicate $P$ also applies to $y = f(x)$.
+Moreover, we can apply the same reasoning subsequently to $f(x)$ and repeat that as many times as the transformation is still applicable.
+The paradox arises when $n$ repeated applications of the transformation $f$ lead us from $x$ to $z = f^n(x)$, where $z$ is something for which $P$ clearly does not apply.
+However, the original premisses seem to force us to accept that $P$ does apply to $x$.
+The paradox with the heap can thus be seen as an instance of this general formulation, if $x$ is the group of grains of sand, the transformation is removing a grain, and $z$ the state of one or zero grains.
+It can also be reversed: it seems reasonable to accept that piling up a grain of sand to something that is not a heap will not turn it into a heap of sand, thus 
+by repeated applications of that transformation one can pile up as many grains of sand as one likes and will never obtain a heap.
+
+Predicates for which one can find a suitable instance of the general formulation of the Sorites paradox are called \emph{vague}.
+Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\citep{Keefe1997}.
+An important intuition regarding these predicates that enables the Sorites paradox to work is that, for a certain level of granularity, there is no clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
+Intuitively, a person with a height of 2 meters is tall, and one with a height of 1.5 meters is not, but there is no value of height in between where one would draw the line; saying, for example, that a person with a height of 1.80 meters is tall, but a person measuring 1.79 meters is not, just seems preposterous.
+Neither drawing a sharp boundary seems reasonable, nor do the consequences of not doing so, and in there lies the truely paradoxical issue of vagueness.
+How widespread is the problem?
+Mereological nihilists argue that instances of the Sorites can be designed for any material object that can be decomposed into small enough parts.
+If we subscribe to the scientific picture of material objects as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\citep{Unger1979}.
+Any heap of molecules that we have a name for is potentially at the mercy of being hypotheticall chipped away into nothingness while still keeping its name.
+The problem thus seems to be a serious one.
+In face of such a powerful paradox like the Sorites, something has got to give.
 

### 2017-01-25 book_chapter/paper.tex
Wrote paragraph on epistemic view.

@@
 The classical philosophical problem of vagueness is most starkly embodied by the Sorites paradox.
-The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\citep{sorensen_sorites_2009}, and uses the example of a heap of sand: if one million grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then by repeated applications of this reasoning we are lead to acknowledge that one single grain of sand also forms a heap.
+The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if one million grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then by repeated applications of this reasoning we are lead to acknowledge that one single grain of sand also forms a heap.
 This paradox can be stated in a very general way.
@@
 Predicates for which one can find a suitable instance of the general formulation of the Sorites paradox are called \emph{vague}.
-Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\citep{Keefe1997}.
+Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\parencite{Keefe1997}.
 An important intuition regarding these predicates that enables the Sorites paradox to work is that, for a certain level of granularity, there is no clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
@@
 Mereological nihilists argue that instances of the Sorites can be designed for any material object that can be decomposed into small enough parts.
-If we subscribe to the scientific picture of material objects as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\citep{Unger1979}.
+If we subscribe to the scientific picture of material objects as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
 Any heap of molecules that we have a name for is potentially at the mercy of being hypotheticall chipped away into nothingness while still keeping its name.
@@
 In face of such a powerful paradox like the Sorites, something has got to give.
+
+Vagueness is typically seen as a challenge to a certain classical conception of language and meaning.
+This is the picture that our words stand in a direct or mediated correspondence to their meanings; sentences are combinations of such words bound by logical rules; knowledge of the meanings of words and of how the rules combine them allows us to know the meaning of sentences and determine whether they are true or false.
+With variations, this picture could be said to underlie theories of meaning of authors life Gottlob Frege, Bertrand Russell, Ludwig Wittgenstein (his early work), Alfred Tarski, Richard Montague, and many others, especially those working in the analytic tradition.
+Vagueness can be seen as a challenge to such a view in different ways.
+Vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to anything at all?
+The classical idea that words like `table' stand in a relation to a set of entities for which it is true to say that they are tables is problematic in light of a Sorites paradox that stresses the boundaries of such a hypothetical set of entities to show that there is no way of defining one.
+Is this a problem of sticking to a bivalent notion of truth?
+Can we solve the issue by adopting more truth values besides `true' and `false'?
+What other alternatives are there?
+
+\todo[inline]{Should we include the following detailed arguments?}
+Some argue that we should stick to the classical picture outlined above.
+Timothy Williamson, for example, claims that ``[c]lassical logic and semantics are vastly superior to the alternatives in simplicity, power, past success, and integration with theories in other domains.''~\parencite*[162]{williamson_vagueness_1992}
+In the face of vagueness, Williamson defends the so-called \emph{epistemic view}: apparently vague terms do determine precise boundaries, we just don't have the ability to know where they are, \emph{i.e.}~vagueness is a kind of ignorance.
+There is thus a precise height above which people are tall, and below which they are not, we just cannot know what that value is.
+The epistemic approach is curious in that in its attempt to save truth, it renders it sterile as an explanatory notion.
+If we can never ascertain truth when it comes to vague terms, how can truth be relevant for explaining how we understand or produce language?
+If the concept of truth is crucially tied to meaning, but we have no access to the former, how could we ever have learned the latter?
+Williamson's reason to ignore these problems is a reasonable one: he believes it helps us, as philosophers or linguists, to better understand language.
+But unless we strongly agree with this assessment of the success of the classical picture, we would do better to search for a more intuitive alternative.
 

### 2017-01-26 book_chapter/paper.tex
Wrote paragraphs on supervaluationism, many-valued logics and degree theories, and why these things don't work.

@@
 
-The classical philosophical problem of vagueness is most starkly embodied by the Sorites paradox.
+The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
 The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if one million grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then by repeated applications of this reasoning we are lead to acknowledge that one single grain of sand also forms a heap.
@@
 
-Predicates for which one can find a suitable instance of the general formulation of the Sorites paradox are called \emph{vague}.
+Predicates for which one can find a suitable instance of the general formulation of the sorites paradox are called \emph{vague}.
 Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\parencite{Keefe1997}.
-An important intuition regarding these predicates that enables the Sorites paradox to work is that, for a certain level of granularity, there is no clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
+An important intuition regarding these predicates that enables the sorites paradox to work is that, for a certain level of granularity, there is no clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
 Intuitively, a person with a height of 2 meters is tall, and one with a height of 1.5 meters is not, but there is no value of height in between where one would draw the line; saying, for example, that a person with a height of 1.80 meters is tall, but a person measuring 1.79 meters is not, just seems preposterous.
@@
 How widespread is the problem?
-Mereological nihilists argue that instances of the Sorites can be designed for any material object that can be decomposed into small enough parts.
+Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
 If we subscribe to the scientific picture of material objects as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
@@
 The problem thus seems to be a serious one.
-In face of such a powerful paradox like the Sorites, something has got to give.
+In face of such a powerful paradox like the sorites, something has got to give.
 
@@
 Vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to anything at all?
-The classical idea that words like `table' stand in a relation to a set of entities for which it is true to say that they are tables is problematic in light of a Sorites paradox that stresses the boundaries of such a hypothetical set of entities to show that there is no way of defining one.
+The classical idea that words like `table' stand in a relation to a set of entities for which it is true to say that they are tables is problematic in light of a sorites paradox that stresses the boundaries of such a hypothetical set of entities to show that there is no way of defining one.
 Is this a problem of sticking to a bivalent notion of truth?
@@
 
+Another suggested theory that attempts to retain as much as possible from classical logic by making minimal modifications to the classical picture of semantics is \emph{supervaluationism}.
+Championed by Kit Fine~\parencite*{Fine1975}, the idea is that vague predicates can be made precise, even arbitrarily.
+Setting a boundary between tall and not tall people at 1.80 meters would be an admissible%
+\footnote{Only precisifications that meet certain constraints should be considered. We refer the reader to Fine's article~\parencite*{Fine1975} for the details.}
+\emph{precisification} of the word `tall'.
+What supervaluationism defends is that, when considering the truth value of a sentence with vague terms, we should take all admissible precisifications of those terms into account: the sentence is true if true under all of them, false if false under all of them, and neither true nor false otherwise.
+The intuition is that a person with a height of 2 meters would be considered tall under all admissible precisifications, thus we could say it is true that that person is tall; for someone measuring 1.79 this would not be the case, hypothetically being true under some precisifications and false under others.
+%In defending this approach, supervaluationism either assumes or postulates the potential existence of a boundary between the sentences for which all admissible precisifications of the vague terms therein make them true, and those for which at least one precisification makes them false.
+Although supervaluationism claims to retain classical logics, it seems to implicitly require a third truth value to account for sentences for which not all precisifications are in agreement.
+Other responses to the problem of vagueness develop alternative logics that explicitly break away from the bivalent assumption that a sentence is either true or false, and introduce additional truth values.
+These range from three-valued logics~\parencite[\emph{e.g.}][]{tye_sorites_1994} to infinite-valued degree theories~\parencite[\emph{e.g.}][]{machina_truth_1976}.
+
+Mark Sainsbury~\parencite*{sainsbury_concepts_1999} argues that all of these solutions fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
+Supervaluationism, many-valued logics, and degree theories all end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
+But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
+Further introducing borderline borderline cases would lead into an infinite regress.
+The point that these classical approaches to vagueness miss is aptly described as follows:
+\begin{quote}
+But to what in our actual use of language does this division correspond?
+It looks as if, as before, it should correspond to the sentences true beyond the shadow of vagueness, those in some kind of borderline position, and those false beyond the shadow.
+But [\ldots] we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly.
+Hence they cannot be included in a correct description of our language.%
+~\parencite[256]{sainsbury_concepts_1999}
+\end{quote}
+By trying to cling as much as possible to the classical picture of logic and semantics, these approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.} are able to recognize the logical inconsistency but do not have a good answer to overcome it.
+Even more importantly, we apparently do not need to solve the inconsistency in order to continue using natural language productively.
+Nobody ever stopped using the word `tall` after being confronted with a sorites series to deconstruct it.
+Why should we develop theories of meaning that are impervious to the paradox?
+In attempting to do so, we are imposing a requirement that is contrary to our observations, and by refusing to give that requirement up we are producing unintuitve, unrealistic theories of meaning that, worst of all, do not even seem to resolve the paradox but merely obfuscate the problem.
+Ludwig Wittgenstein put it better:
+\begin{quote}
+The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
+(For the crystalline purity of logic was, of course, not something I had \emph{discovered}: it was a requirement.)
+The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
+-- We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
+We want to walk: so we need \emph{friction}.
+Back to the rough ground!%
+~\parencite[\S 107]{wittgenstein_philosophical_1953}
+\end{quote}
+
 \section{Paradigm shift}

### 2017-01-30 book_chapter/paper.tex
Wrote on Wittgenstein, reformulated some other text, wrapped up the argument of the introduction.

--- 2017-01-25 book_chapter/paper.tex
@@
 
-\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false]{biblatex}
+\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
@@
 The problem thus seems to be a serious one.
-In face of such a powerful paradox like the Sorites, something has got to give.
+In face of such a powerful paradox like the sorites, something has got to give.
 
-Vagueness is typically seen as a challenge to a certain classical conception of language and meaning.
+Vagueness is typically seen as a challenge to a classical conception of language and meaning.
 This is the picture that our words stand in a direct or mediated correspondence to their meanings; sentences are combinations of such words bound by logical rules; knowledge of the meanings of words and of how the rules combine them allows us to know the meaning of sentences and determine whether they are true or false.
 With variations, this picture could be said to underlie theories of meaning of authors life Gottlob Frege, Bertrand Russell, Ludwig Wittgenstein (his early work), Alfred Tarski, Richard Montague, and many others, especially those working in the analytic tradition.
-Vagueness can be seen as a challenge to such a view in different ways.
-Vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to anything at all?
-The classical idea that words like `table' stand in a relation to a set of entities for which it is true to say that they are tables is problematic in light of a Sorites paradox that stresses the boundaries of such a hypothetical set of entities to show that there is no way of defining one.
-Is this a problem of sticking to a bivalent notion of truth?
-Can we solve the issue by adopting more truth values besides `true' and `false'?
-What other alternatives are there?
+The problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to anything at all?
+And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding this person?
 
-\todo[inline]{Should we include the following detailed arguments?}
-Some argue that we should stick to the classical picture outlined above.
+Some argue that the problems caused by vagueness is not necessarily enough to throw away the picture of meaning outlined above.
 Timothy Williamson, for example, claims that ``[c]lassical logic and semantics are vastly superior to the alternatives in simplicity, power, past success, and integration with theories in other domains.''~\parencite*[162]{williamson_vagueness_1992}
-In the face of vagueness, Williamson defends the so-called \emph{epistemic view}: apparently vague terms do determine precise boundaries, we just don't have the ability to know where they are, \emph{i.e.}~vagueness is a kind of ignorance.
-There is thus a precise height above which people are tall, and below which they are not, we just cannot know what that value is.
-The epistemic approach is curious in that in its attempt to save truth, it renders it sterile as an explanatory notion.
+In the face of vagueness, Williamson defends the so-called \emph{epistemic view}: apparently vague terms do determine precise boundaries, we just don't have the ability to know where they are; vagueness is thus a kind of ignorance.
+There is a precise height above which people are tall, and below which they are not, we just cannot know what that value is.\todo{Talk about Frazee, J., \& Beaver, D. (2010)?}
+The epistemic approach is curious in that in its attempt to save truth, it seems to render it sterile as an explanatory notion.
 If we can never ascertain truth when it comes to vague terms, how can truth be relevant for explaining how we understand or produce language?
 If the concept of truth is crucially tied to meaning, but we have no access to the former, how could we ever have learned the latter?
-Williamson's reason to ignore these problems is a reasonable one: he believes it helps us, as philosophers or linguists, to better understand language.
-But unless we strongly agree with this assessment of the success of the classical picture, we would do better to search for a more intuitive alternative.
+Williamson's reason to ignore these problems is a reasonable one: he believes it helps us, from philosophers of language to theoretical linguists, to better understand language.
+But this is tightly connected with this assessment of the success of the classical picture.
 
-\section{Paradigm shift}
-\label{sec:Wittgenstein-and-signaling}
+Criticism of the adequacy of this picture comes from various quarters.
+One of the strongest can be found in the later work of philosopher Ludwig Wittgenstein\footnote{To the extent that substantial views can be said to be defended by the author.
+Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} or Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}, most importantly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}.
+Undermining several assumptions behind the classical picture and the implications thereof, Wittgenstein pushes for a paradigm shift:
+\begin{quote}
+The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
+(For the crystalline purity of logic was, of course, not something I had \emph{discovered}: it was a requirement.)
+The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
+-- We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
+We want to walk: so we need \emph{friction}.
+Back to the rough ground!%
+~\parencite[\S 107]{wittgenstein_philosophical_1953}
+\end{quote}
+The ideal of exactness that is tied to the notions of truth and logic dazzles and leads to misunderstandings~\parencite*[\S 100]{wittgenstein_philosophical_1953}; it creates a haze around the workings of language~\parencite*[\S 5]{wittgenstein_philosophical_1953} that keeps us from seeing what is right in front of us.
+We can only escape this by ``turning our whole inquiry around''~\parencite*[\S 108]{wittgenstein_philosophical_1953}.
+
+%%% This discussion is perhaps too detailed for the scope of this paper
+% Another suggested theory that attempts to retain as much as possible from classical logic by making minimal modifications to the classical picture of semantics is \emph{supervaluationism}.
+% Championed by Kit Fine~\parencite*{Fine1975}, the idea is that vague predicates can be made precise, even arbitrarily.
+% Setting a boundary between tall and not tall people at 1.80 meters would be an admissible%
+% \footnote{Only precisifications that meet certain constraints should be considered. We refer the reader to Fine's article~\parencite*{Fine1975} for the details.}
+% \emph{precisification} of the word `tall'.
+% What supervaluationism defends is that, when considering the truth value of a sentence with vague terms, we should take all admissible precisifications of those terms into account: the sentence is true if true under all of them, false if false under all of them, and neither true nor false otherwise.
+% The intuition is that a person with a height of 2 meters would be considered tall under all admissible precisifications, thus we could say it is true that that person is tall; for someone measuring 1.79 this would not be the case, hypothetically being true under some precisifications and false under others.
+% %In defending this approach, supervaluationism either assumes or postulates the potential existence of a boundary between the sentences for which all admissible precisifications of the vague terms therein make them true, and those for which at least one precisification makes them false.
+% Although supervaluationism claims to retain classical logics, it seems to implicitly require a third truth value to account for sentences for which not all precisifications are in agreement.
+% Other responses to the problem of vagueness develop alternative logics that explicitly break away from the bivalent assumption that a sentence is either true or false, and introduce additional truth values.
+% These range from three-valued logics~\parencite[\emph{e.g.}][]{tye_sorites_1994} to infinite-valued degree theories~\parencite[\emph{e.g.}][]{machina_truth_1976}.
+
+The way in how the ideal of exactness leads into misunderstandings is somewhat patent in the standard approaches to dealing with vagueness that try to hang on to as much as possible of the classical picture.
+Supervaluationism, many-valued logics, and degree theories all propose changes to it in order to accommodate for vague predicates.
+The proposals are, however, still within the general molds of the classical picture.
+Mark Sainsbury~\parencite*{sainsbury_concepts_1999} argues that, because of that, they all fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
+All the aforementioned proposals end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
+But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
+Further introducing borderline borderline cases would lead into an infinite regress.
+Because of their attachment to the classical picture, the standard approaches to vagueness fail to see an important lesson:
+\begin{quote}
+But to what in our actual use of language does this division correspond?
+It looks as if, as before, it should correspond to the sentences true beyond the shadow of vagueness, those in some kind of borderline position, and those false beyond the shadow.
+But [\ldots] we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly.
+Hence they cannot be included in a correct description of our language.%
+~\parencite[256]{sainsbury_concepts_1999}
+\end{quote}
+By trying to cling as much as possible to the classical picture of logic and semantics, these approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~are able to recognize the logical inconsistency but do not have a good answer to overcome it.
+Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
+Nobody ever stopped using the word `tall` after being confronted with a sorites series to deconstruct it.
+Why should we develop theories of meaning that are impervious to the paradox?
+
+The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear that, by doing so, we would need to give up important associated notions like rationality\todo{Cite Rorty (PMN)?}.
+Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
+The fear could thus be that, if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground that rationality stands on.
+But giving up the ideal of truth and logic as relevant explanatory notions to understand natural language does not mean giving up on rationality altogether.
+To say that language does not follow strict logical rules, or that truth is not the ideal the logicians make it out to be, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
+Vagueness can shed light on features of our language, but only after we step out of the classical picture and start looking at language under a different paradigm.
+
+From Wittgenstein we learn that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}.
+This introduces the notion of language-games as methodological tools: one can set up a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and reflect on the assumptions that underlie it and their implications and on how the scenario would play according to them.
+Jos\'e Pedro Correia~\parencite*{correia_bivalent_2013} argues that the framework of signaling games, introduced for the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that, while reaping benefits from mathematical formalization and computer simulation.
+In the following, we will be looking at signaling models of vagueness in order to provide an overview of proposals of the natural sources of vagueness, what that tells us about rationality, and the implications of these ideas.
+But first, let us look in more detail into how Lewis-Skyrms signaling games work, and how vagueness shows up again as a different kind of problem for these models.
+%With regards to vagueness and exactness, it is also important to note the following:
+%\begin{quote}
+%``Inexact'' is really a reproach, and ``exact'' is praise.
+%And that is to say that what is inexact attains its goal less perfectly than does what is more exact.
+%So it all depends on what we call ``the goal''.
+%\end{quote}
+
+%%% This is perhaps too philosophical and off-topic
+% The reluctance to give up truth and logic as fundamental concepts to understand meaning is perhaps related to a fear of losing a grip on rationality, and more importantly, on the supposed necessity of these concepts to underwrite normative statements and thus condition discourse.
+% Richard Rorty describes this as follows:
+% \begin{quote}
+% The traditional view is that there is a core self which can look at, decide among, use, and express itself by means of, [\ldots] belief and desires.
+% Further, these beliefs and desires are criticizable not simply by reference to their ability to cohere with one another, but by reference to something exterior to the network within which they are strands.
+% Beliefs are, on this account, criticizable because they fail to correspond to reality.
+% Desires are criticizable because they fail to correspond to the essential nature of the human self -- because they are ``irrational'' or ``unnatural''.
+% \end{quote}
+% Since the linguistic turn, language is seen as the medium that puts the self in contact with either reality or the nature of the self.
+% Rationality is, in this picture, dependent on truth as its objective and logic as the means to attain it.
+% The latter are embodied in language.
+% The fear is thus that if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground that rationality stands on, and thus lose the ability to make value judgments on sentences from a supposedly objective standpoint.
+% Full-fledged relativism thus ensues.
+
+\section{Signaling games and another problem of vagueness}
+\label{sec:signaling-and-Lipman}
 

### 2017-02-02 book_chapter/paper.tex
Made some adjustments to the first section. Added something more about the relation between truth and rationality. Added some more text around Wittgenstein, language-games and their relation with signaling games.

@@
 The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
-The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if one million grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then by repeated applications of this reasoning we are lead to acknowledge that one single grain of sand also forms a heap.
-This paradox can be stated in a very general way.
-We start with an $x$ to which a certain predicate $P$ clearly applies.
+The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if $1 000 000$ grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then if we remove one grain, the remaining $999 999$ also clearly form a heap; by repeated applications of this reasoning we can remove almost all grains of sand one by one and are lead to acknowledge that $1$ single grain of sand also forms a heap.
+This paradox is interesting because it can be made general and re-applied to many other words besides `heap'.
+In the general formulation, we start with an $x$ to which a certain predicate $P$ clearly applies.
 We observe that a certain transformation $f$ has no impact on whether something is or is not $P$.
@@
 Moreover, we can apply the same reasoning subsequently to $f(x)$ and repeat that as many times as the transformation is still applicable.
-The paradox arises when $n$ repeated applications of the transformation $f$ lead us from $x$ to $z = f^n(x)$, where $z$ is something for which $P$ clearly does not apply.
-However, the original premisses seem to force us to accept that $P$ does apply to $x$.
-The paradox with the heap can thus be seen as an instance of this general formulation, if $x$ is the group of grains of sand, the transformation is removing a grain, and $z$ the state of one or zero grains.
+The paradox arises when $n$ repeated applications of the transformation $f$ lead us from $x$ to $z = f^n(x)$, where $z$ is something for which originally we would say $P$ clearly does not apply, but since we see it as difficult to reject either the premisses or the reasoning, we seem to be forced to accept that $P$ \emph{does in fact apply} to $x$.
+The paradox with the heap can thus be seen as an instance of this general formulation, if $x$ is the group of $1 000 000$ grains of sand, the transformation is removing a grain, and $z$ the state of one (or even zero) grains.
 It can also be reversed: it seems reasonable to accept that piling up a grain of sand to something that is not a heap will not turn it into a heap of sand, thus 
@@
 Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\parencite{Keefe1997}.
-An important intuition regarding these predicates that enables the sorites paradox to work is that, for a certain level of granularity, there is no clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
+An important intuition regarding these predicates that enables the sorites paradox to go through is that there does not seem to be any clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
 Intuitively, a person with a height of 2 meters is tall, and one with a height of 1.5 meters is not, but there is no value of height in between where one would draw the line; saying, for example, that a person with a height of 1.80 meters is tall, but a person measuring 1.79 meters is not, just seems preposterous.
-Neither drawing a sharp boundary seems reasonable, nor do the consequences of not doing so, and in there lies the truely paradoxical issue of vagueness.
+Neither drawing a sharp boundary seems reasonable, nor do the logical consequences of not doing so, and in there lies the truely paradoxical issue of vagueness.
 How widespread is the problem?
+It is easy to come up with more examples of predicates that are based on more finely grained properties, like `tall' is intuitively based on height, for which constructing a sorites paradox would be easy.
 Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
-If we subscribe to the scientific picture of material objects as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
+If we subscribe to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
 Any heap of molecules that we have a name for is potentially at the mercy of being hypotheticall chipped away into nothingness while still keeping its name.
 The problem thus seems to be a serious one.
+Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
 In face of such a powerful paradox like the sorites, something has got to give.
@@
 Vagueness is typically seen as a challenge to a classical conception of language and meaning.
-This is the picture that our words stand in a direct or mediated correspondence to their meanings; sentences are combinations of such words bound by logical rules; knowledge of the meanings of words and of how the rules combine them allows us to know the meaning of sentences and determine whether they are true or false.
+We can summarize this classical picture as such: words stand in a direct or mediated correspondence to their meanings; sentences are combinations of such words bound by logical rules; knowledge of the meanings of words and of how the rules combine them allows us to know the meaning of sentences and determine whether they are true or false.
 With variations, this picture could be said to underlie theories of meaning of authors life Gottlob Frege, Bertrand Russell, Ludwig Wittgenstein (his early work), Alfred Tarski, Richard Montague, and many others, especially those working in the analytic tradition.
-The problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to anything at all?
+The problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to a well defined set of people?
 And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding this person?
 
-Some argue that the problems caused by vagueness is not necessarily enough to throw away the picture of meaning outlined above.
-Timothy Williamson, for example, claims that ``[c]lassical logic and semantics are vastly superior to the alternatives in simplicity, power, past success, and integration with theories in other domains.''~\parencite*[162]{williamson_vagueness_1992}
-In the face of vagueness, Williamson defends the so-called \emph{epistemic view}: apparently vague terms do determine precise boundaries, we just don't have the ability to know where they are; vagueness is thus a kind of ignorance.
-There is a precise height above which people are tall, and below which they are not, we just cannot know what that value is.\todo{Talk about Frazee, J., \& Beaver, D. (2010)?}
-The epistemic approach is curious in that in its attempt to save truth, it seems to render it sterile as an explanatory notion.
+Some argue that the problems caused by vagueness are not serious enough for us to throw away the picture of meaning outlined above.
+Timothy Williamson, for example, claims that we should stick with the classical picture because ``[c]lassical logic and semantics are vastly superior to the alternatives in simplicity, power, past success, and integration with theories in other domains.''~\parencite*[162]{williamson_vagueness_1992}
+In the face of vagueness, Williamson defends the so-called \emph{epistemic view}: vague terms do determine precise boundaries, we just don't have the ability to know where they are; there is a precise height above which people are tall, and below which they are not, we are just ignorant of what this value is.
+%%% Could talk about Frazee, J., and Beaver, D. (2010) here
+The epistemic approach is curious in that, in its attempt to save truth, it seems to actually render it sterile as an explanatory notion.
 If we can never ascertain truth when it comes to vague terms, how can truth be relevant for explaining how we understand or produce language?
 If the concept of truth is crucially tied to meaning, but we have no access to the former, how could we ever have learned the latter?
-Williamson's reason to ignore these problems is a reasonable one: he believes it helps us, from philosophers of language to theoretical linguists, to better understand language.
-But this is tightly connected with this assessment of the success of the classical picture.
+Williamson's reason to ignore these problems is a pragmatic one: he believes it helps us, philosophers of language and theoretical linguists, to better understand natural language.
+But this reason holds only insofar as we agree with this assessment of the success of the classical picture.
 
-Criticism of the adequacy of this picture comes from various quarters.
+Criticism of the adequacy of this picture, however, comes from various quarters.
 One of the strongest can be found in the later work of philosopher Ludwig Wittgenstein\footnote{To the extent that substantial views can be said to be defended by the author.
-Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} or Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}, most importantly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}.
+Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}, most importantly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}.
 Undermining several assumptions behind the classical picture and the implications thereof, Wittgenstein pushes for a paradigm shift:
@@
 \end{quote}
-The ideal of exactness that is tied to the notions of truth and logic dazzles and leads to misunderstandings~\parencite*[\S 100]{wittgenstein_philosophical_1953}; it creates a haze around the workings of language~\parencite*[\S 5]{wittgenstein_philosophical_1953} that keeps us from seeing what is right in front of us.
-We can only escape this by ``turning our whole inquiry around''~\parencite*[\S 108]{wittgenstein_philosophical_1953}.
+He argues that the ideal of exactness that is tied to the notions of truth and logic dazzles and leads to misunderstandings~\parencite*[\S 100]{wittgenstein_philosophical_1953}; it creates a haze around the workings of language~\parencite*[\S 5]{wittgenstein_philosophical_1953} that keeps us from seeing what is right in front of us.
+We can only escape this predicament, not by bargaining exactness out of logic, but by turning our whole inquiry around~\parencite*[\S 108]{wittgenstein_philosophical_1953}, \emph{i.e.}~by thinking of language in terms of a different paradigm.
 
@@
 Further introducing borderline borderline cases would lead into an infinite regress.
-Because of their attachment to the classical picture, the standard approaches to vagueness fail to see an important lesson:
-\begin{quote}
-But to what in our actual use of language does this division correspond?
-It looks as if, as before, it should correspond to the sentences true beyond the shadow of vagueness, those in some kind of borderline position, and those false beyond the shadow.
-But [\ldots] we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly.
-Hence they cannot be included in a correct description of our language.%
-~\parencite[256]{sainsbury_concepts_1999}
-\end{quote}
-By trying to cling as much as possible to the classical picture of logic and semantics, these approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~are able to recognize the logical inconsistency but do not have a good answer to overcome it.
+Because of their attachment to the classical picture, the standard approaches to vagueness fail to see an important lesson, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
+% \begin{quote}
+% But to what in our actual use of language does this division correspond?
+% It looks as if, as before, it should correspond to the sentences true beyond the shadow of vagueness, those in some kind of borderline position, and those false beyond the shadow.
+% But [\ldots] we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly.
+% Hence they cannot be included in a correct description of our language.%
+% ~\parencite[256]{sainsbury_concepts_1999}
+% \end{quote}
+By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~are able to recognize the logical inconsistency but do not have a good answer to overcome it.
 Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
@@
 
-The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear that, by doing so, we would need to give up important associated notions like rationality\todo{Cite Rorty (PMN)?}.
+The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also consequently need to be abandoned down the line.
+Talking about philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
+\begin{quote}
+In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
+[\ldots]
+The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
+~\parencite*[1]{rorty_response_2000-1}
+\end{quote}
 Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
-The fear could thus be that, if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground that rationality stands on.
-But giving up the ideal of truth and logic as relevant explanatory notions to understand natural language does not mean giving up on rationality altogether.
-To say that language does not follow strict logical rules, or that truth is not the ideal the logicians make it out to be, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
-Vagueness can shed light on features of our language, but only after we step out of the classical picture and start looking at language under a different paradigm.
+The fear could be that, if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground on which rationality stands, and the whole edifice would collapse.
+But giving up the ideal of truth and logic as relevant explanatory notions to understand natural language does not mean giving up on rationality, or any of the other notions, altogether.
+To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
+Vagueness can shed light on features of our language and our rationality, but only after we step out of the classical picture and start looking at these concepts in terms of a different paradigm.
 
-From Wittgenstein we learn that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}.
-This introduces the notion of language-games as methodological tools: one can set up a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and reflect on the assumptions that underlie it and their implications and on how the scenario would play according to them.
-Jos\'e Pedro Correia~\parencite*{correia_bivalent_2013} argues that the framework of signaling games, introduced for the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that, while reaping benefits from mathematical formalization and computer simulation.
-In the following, we will be looking at signaling models of vagueness in order to provide an overview of proposals of the natural sources of vagueness, what that tells us about rationality, and the implications of these ideas.
+What can such a paradigm be?
+The Wittgensteinian picture of language is one of multiplicity, heterogeneity, and change.
+Languages are thought of as patchworks of various language-games, with new ones continuously being added and old ones falling out of use.
+These language-games, in turn, are situated uses of words in certain activities.
+The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
+Meaning is thus contingent to the language-games we play, and dynamic.
+The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
+Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up such a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play through according to those assumptions.
+Jos\'e Pedro Correia~\parencite*{correia_bivalent_2013} argues that the framework of signaling games, introduced for the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that, while reaping the benefits of a mathematical formalism and the further ability to conduct computer simulations.
+From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting computer simulations is like contemplating how the game would play through according to the implications of the assumptions one built into the model.
+
+In the following, we will be looking at signaling models of vagueness in order to provide an overview of proposals of natural sources of vagueness, what that tells us about rationality, and the implications of these ideas.
 But first, let us look in more detail into how Lewis-Skyrms signaling games work, and how vagueness shows up again as a different kind of problem for these models.

### 2017-02-15 book_chapter/paper.tex
Added introduction to signaling games.

@@
 \usepackage{a4wide}
+\usepackage{amsmath}
 \usepackage{amssymb}
@@
 
+Signaling games were first introduced as models of communication by David Lewis~\parencite*{lewis_convention_1969}.
+He wanted to address an argument raised by W.~V.~Quine~\parencite*{quine_truth_1936} and others against the possibility of language having had started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
+To this, Lewis retorts:
+\begin{quote}
+I offer this rejoinder: an agreement sufficient to create a convention need not be a transaction involving language or any other conventional activity.
+All it takes is an exchange of manifestations of a propensity to conform to a regularity.%
+~\parencite*[87--88]{lewis_convention_1969}
+\end{quote}
+In order to support this claim, Lewis studies coordination problems formalized in terms of game theory.
+
+Coordination problems are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
+In game theory terms, the agents interested in the coordination are the players, the game involves each player making an independent choice from his set of available actions, a payoff is what is attributed to each player based on the choices of both.
+The setup of the game, \emph{i.e.}~the available actions and the relative interests of the players, can be represented in a payoff matrix, where rows show one player's available actions, columns the other player's available actions, and each cell gives the payoff for each of the players based on its row and column combination.
+For example, consider the following matrix:
+\begin{center}
+\begin{tabular}{r|c|c|}
+\multicolumn{1}{r}{}
+ & \multicolumn{1}{c}{$b_1$}
+ & \multicolumn{1}{c}{$b_2$} \\ \cline{2-3}
+   $a_1$ & $1,1$ & $0,0$ \\ \cline{2-3}
+   $a_2$ & $0,0$ & $1,1$ \\ \cline{2-3}
+\end{tabular}
+\end{center}
+We can see this as representing the following game: one player has actions $A = \lbrace a_1, a_2 \rbrace$ available, the other $B = \lbrace b_1, b_2 \rbrace$; they prefer to coordinate $a_1$ with $b_1$ or $a_2$ with $b_2$, thus if this is achieved each gets a payoff of $1$, otherwise they each get $0$.
+Formally, all we need to define in order to characterize such a situation are the sets of actions $A$ and $B$, and a utility function $U : A \times B \rightarrow \mathbb{R}^2$ that specifies the payoff for each player given the choices of actions.
+
+The above example is a very simple case, but it serves to illustrate Lewis' notion of convention.
+The two pairs of choices $(a_1,b_1)$ and $(a_2,b_2)$ are stable coordination equilibria, because in such a scenario no player has an incentive to unilaterally change his choice: if the first player is going to choose action $a_1$, the second player would get a payoff of $0$ for switching to $b_2$, instead of $1$ from sticking to $b_1$; the same reasoning applies, mutatis mutandis, to the other player and the other equilibrium.
+However, neither player has a preference between those two pairs.
+Furthermore, if the same coordination problem arises again, Lewis says, we can expect precedence to induce a kind of regularity: if the players manage to coordinate on one of the two equilibria, they should be expected to repeat the choices of actions that lead to that success, thus remaining in the same equilibrium.
+Not only can we expect the players in such a repeated interaction to conform to a regularity, but also to expect others to do the same.
+Their preference is to remain in a certain equilibrium given that others do too.
+This summarizes Lewis's notion of a convention.
+
+In order to extend this notion to \emph{linguistic} conventions, Lewis considers situations where the actions available involve sending and receiving signals or messages.
+Thus, we could think of two players with different roles.
+The first player, which Lewis calls the communicator~\parencite*[130]{lewis_convention_1969} and we will here call the sender, has knowledge about which of a number of possible states of affairs obtains, and depending on this information chooses a signal to send.
+The second player, which Lewis calls the audience~\parencite*[130]{lewis_convention_1969} and we will here call the receiver, has knowledge about which signal the sender chose, and based on this information chooses one of several possible responses.
+A preference relation exists between responses and states of affairs, and a payoff is attributed to each player based on the choices of both.
+Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
+Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.
+
+Despite the added dimension of the signal exchange, these so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
+A communicator's contingency plan, or sender strategy as we will call it, is a specification of a choice of message for each possible state of affairs.
+It thus describes the sender's behavior conditional on the state of affairs that obtains.
+An audience's contingency plan, or receiver strategy as we will call it, analogously specifies a choice of action for each possible message.
+Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
+The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function between states of affairs and actions as a sum of the payoffs for all cases, \emph{i.e.}:
+$$
+EU(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
+$$
+As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, and the following utility matrix:
+\begin{center}
+\begin{tabular}{r|c|c|}
+\multicolumn{1}{r}{}
+ & \multicolumn{1}{c}{$a_1$}
+ & \multicolumn{1}{c}{$a_2$} \\ \cline{2-3}
+   $t_1$ & $1,1$ & $0,0$ \\ \cline{2-3}
+   $t_2$ & $0,0$ & $1,1$ \\ \cline{2-3}
+\end{tabular}
+\end{center}
+Possible sender and receiver strategies could be, for example, $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
+According to our formula above, these would have an expected utility of $2$ for both sender and receiver.
+% the following expected utility:
+% \begin{equation}
+% \begin{split}
+% EU(\sigma, \rho) & = \sum_{t \in T} U(t, \rho(\sigma(t))) \\
+%                  & = U(t_1, \rho(\sigma(t_1))) + U(t_2, \rho(\sigma(t_2))) \\
+%                  & = U(t_1, \rho(m_2)) + U(t_2, \rho(m_1)) \\
+%                  & = U(t_1, t_1) + U(t_2, t_1) \\
+%                  & = (1,1) + (1,1) \\
+%                  & = (2,2)
+% \end{split}
+% \end{equation}
+% Based on this, we could also create a matrix of expected utilities for this example as follows:
+% \begin{center}
+% \begin{tabular}{r|c|c|c|c|}
+% \multicolumn{1}{r}{}
+%  & \multicolumn{1}{c}{$m_1 \mapsto a_1$}
+%  & \multicolumn{1}{c}{$m_1 \mapsto a_1$}
+%  & \multicolumn{1}{c}{$m_1 \mapsto a_2$}
+%  & \multicolumn{1}{c}{$m_1 \mapsto a_2$} \\
+% \multicolumn{1}{r}{}
+%  & \multicolumn{1}{c}{$m_2 \mapsto a_1$}
+%  & \multicolumn{1}{c}{$m_2 \mapsto a_2$}
+%  & \multicolumn{1}{c}{$m_2 \mapsto a_1$}
+%  & \multicolumn{1}{c}{$m_2 \mapsto a_2$} \\ \cline{2-5}
+%    $t_1 \mapsto m_1, t_2 \mapsto m_1$ & $1,1$ & $1,1$ & $1,1$ & $1,1$ \\ \cline{2-5}
+%    $t_1 \mapsto m_1, t_2 \mapsto m_2$ & $1,1$ & $2,2$ & $0,0$ & $1,1$ \\ \cline{2-5}
+%    $t_1 \mapsto m_2, t_2 \mapsto m_1$ & $1,1$ & $0,0$ & $2,2$ & $1,1$ \\ \cline{2-5}
+%    $t_1 \mapsto m_2, t_2 \mapsto m_2$ & $1,1$ & $1,1$ & $1,1$ & $1,1$ \\ \cline{2-5}
+% \end{tabular}
+% \end{center}
+They also represent one of the two stable conventions in this game, the other being the pair of strategies $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
+Conventions of this kind in a signaling problem are what Lewis calls \emph{signaling systems}.
+
+Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
+Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be the required common knowledge between the players involved.
+Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
+These requirements can seem even more excessive if we consider how simple signaling systems are, compared to human languages.
+The purpose of studying them was to tell a story about how language as a conventional system could get off the ground without any prior sort of agreement.
+It is implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
+Also, communication through simple message exchange is something that almost all animals do, from monkeys to birds to bees.
+A plausible account of the origin of language should first explain how signaling systems like those could get started without a great deal of rationality from the part of the agents involved.
+
+% One interesting aspect of these signaling systems is that, if we were to observe the repeated interactions of sender and receiver according to those strategies from a bird's eye view, we would be inclined to say that the signals are meaningful.
+% For example, if we consider the latter example, a sender would always send $m_1$ when state of affairs $t_1$ obtains to which the receiver would always respond with action $a_1$ which maximizes both players' payoff.
+% It is as if $m_1$ means either ``$t_1$ obtains'' or ``do $a_1$''.
+%
+
 \section{Signaling models of vagueness}

### 2017-02-20 book_chapter/paper.tex
Wrote about Skyrms and the use of evolutionary game theory for signaling games.

@@
 
-Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
+The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like the classical picture.
+However, we can already see it attempts to address the study of meaning from a very different angle.
+There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes.
+There is also no appeal to truth as a guide for our inquiry.
+In this scenario, agents are assumed to possess and exercise rationality.
+It is not, however, the same rationality as we find in the classical picture, \emph{i.e.}~the willingness to abide by truth; it is rather a more pragmatic notion of rationality as utility maximization.
+One advantage is that we are no longer tied to the need to explain how language hooks on to the world, since this is not relevant on this account.
+We are merely trying to see how agents can use signals to cope with the world and achieve their purposes.
+On the way, we will better understand meaning, not by trying to say what meaning is, but by seeing how language works.
+
+Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996}, however, identifies some problems with the story so far.
 Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be the required common knowledge between the players involved.
 Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
-These requirements can seem even more excessive if we consider how simple signaling systems are, compared to human languages.
-The purpose of studying them was to tell a story about how language as a conventional system could get off the ground without any prior sort of agreement.
-It is implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
-Also, communication through simple message exchange is something that almost all animals do, from monkeys to birds to bees.
+These requirements can seem excessive, even more so if we consider how simple signaling systems are, compared to human languages.
+The models were introduced in order to help explain how language as a conventional system could get off the ground without any prior sort of agreement.
+If we think of the origins of language in the history of mankind, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
+Also, communication through simple message exchange is something that almost all animals do, from monkeys with their calls, to birds with their songs, to bees with their dances, to ants with their pheromone trails.
 A plausible account of the origin of language should first explain how signaling systems like those could get started without a great deal of rationality from the part of the agents involved.
+
+In order to address this problem, Skyrms proposes we study signaling problems in evolutionary terms.
+Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
+Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
+Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they appear to be.
+The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
+The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
+Using it, one can compute which strategies evolve under which conditions.
+
+The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
+Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits}.
+In a simple case like the example discussed above, an evolutionary process always drives the system into a state where one signaling system takes over the whole population; which one of the possible two will depend on their relative proportions in the original population.
+More complex signaling problems may have different evolutionary outcomes, sometimes surprising ones.
+Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, explanations of deception).
+
+Skyrms' evolutionary game theory approach not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change, in that it moves most of the theory and mathematical formalisms to the descriptive side of the investigation.
+Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
+Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
+Focus is put on understanding how various ingredients to the model interact, and which results they produce, not on metaphysical concerns.
+While the general framework manages to abstract some details away from the formalization, it does leave room for them, especially when it comes to the dynamics.
+We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired on learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players.\todo{give examples and references?}
+This range of options goes hand in hand with assumptions of rationality, from no assumptions in a biologically-inspired setting, to a certain degree of rationality captured by a learning method, to higher level rationality and even game-theoretical reasoning.
+
+\todo[inline]{How about vagueness?}
 

### 2017-03-07 book_chapter/paper.tex
Made some corrections to written text. Finished Section 2. Wrote introduction to Section 3.

--- 2017-02-15 book_chapter/paper.tex
@@
 \usepackage{syllogism}
+\usepackage{subcaption}
 
@@
 Signaling games were first introduced as models of communication by David Lewis~\parencite*{lewis_convention_1969}.
-He wanted to address an argument raised by W.~V.~Quine~\parencite*{quine_truth_1936} and others against the possibility of language having had started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
+His original objective was to address an argument raised by W.~V.~Quine~\parencite*{quine_truth_1936} and others against the possibility of language having started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
 To this, Lewis retorts:
@@
 In order to support this claim, Lewis studies coordination problems formalized in terms of game theory.
+These are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
+In game theory terms, the agents interested in the coordination are the players, the game involves each player making an independent choice from his set of available actions, a payoff is what is attributed to each player based on the choices of both.
+Adapting one of Lewis' examples, say Alice and Bob want to get together.
+They usually meet at either Caf\'e One or Bistro Two.
+Imagine there is no way for them to make any explicit agreement about it.
+They are thus left to independently decide to either go to Caf\'e One or go to Bistro Two and hope for the other to show up there.
+Neither has any preference for either place, but they do want to meet.
+Each thus prefers to go to one of the places only if the other also decides to go to that particular place.
 
-Coordination problems are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
-In game theory terms, the agents interested in the coordination are the players, the game involves each player making an independent choice from his set of available actions, a payoff is what is attributed to each player based on the choices of both.
-The setup of the game, \emph{i.e.}~the available actions and the relative interests of the players, can be represented in a payoff matrix, where rows show one player's available actions, columns the other player's available actions, and each cell gives the payoff for each of the players based on its row and column combination.
+The setup of a coordination problem, \emph{i.e.}~the available actions and the relative interests of the players, can be represented in a payoff matrix, where rows show one player's available actions, columns the other player's available actions, and each cell gives the payoff for each of the players based on its row and column combination.
 For example, consider the following matrix:
@@
 We can see this as representing the following game: one player has actions $A = \lbrace a_1, a_2 \rbrace$ available, the other $B = \lbrace b_1, b_2 \rbrace$; they prefer to coordinate $a_1$ with $b_1$ or $a_2$ with $b_2$, thus if this is achieved each gets a payoff of $1$, otherwise they each get $0$.
+Connecting with the example laid out above, think of $a_1$ as Alice going to Caf\'e One, $a_2$ as Alice going to Bistro Two, $b_1$ as Bob going to Caf\'e One, and finally $b_2$ as Bob going to Bistro Two; their payoff is $1$ for both if they coordinate on the place to meet (independently of which one), and $0$ if they miscoordinate.
 Formally, all we need to define in order to characterize such a situation are the sets of actions $A$ and $B$, and a utility function $U : A \times B \rightarrow \mathbb{R}^2$ that specifies the payoff for each player given the choices of actions.
@@
 However, neither player has a preference between those two pairs.
-Furthermore, if the same coordination problem arises again, Lewis says, we can expect precedence to induce a kind of regularity: if the players manage to coordinate on one of the two equilibria, they should be expected to repeat the choices of actions that lead to that success, thus remaining in the same equilibrium.
+If the same coordination problem arises again, Lewis says, we can expect precedence to induce a kind of regularity: if the players manage to coordinate on one of the two equilibria, they should be expected to repeat the choices of actions that lead to that success, thus remaining in the same equilibrium.
 Not only can we expect the players in such a repeated interaction to conform to a regularity, but also to expect others to do the same.
@@
 Thus, we could think of two players with different roles.
-The first player, which Lewis calls the communicator~\parencite*[130]{lewis_convention_1969} and we will here call the sender, has knowledge about which of a number of possible states of affairs obtains, and depending on this information chooses a signal to send.
-The second player, which Lewis calls the audience~\parencite*[130]{lewis_convention_1969} and we will here call the receiver, has knowledge about which signal the sender chose, and based on this information chooses one of several possible responses.
+The first player, which Lewis calls the communicator~\parencite*[130]{lewis_convention_1969} and we will here call the sender, has knowledge about which of a number of possible states of affairs obtains and, depending on this information, chooses a signal to send.
+The second player, which Lewis calls the audience~\parencite*[130]{lewis_convention_1969} and we will here call the receiver, has knowledge about which signal the sender chose and, based on this information, chooses one of several possible responses.
 A preference relation exists between responses and states of affairs, and a payoff is attributed to each player based on the choices of both.
@@
 Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
-The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function between states of affairs and actions as a sum of the payoffs for all cases, \emph{i.e.}:
-$$
-EU(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
-$$
+The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function between states of affairs and actions as a sum of the payoffs for all cases.
+%, \emph{i.e.}:
+% $$
+% EU(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
+% $$
 As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, and the following utility matrix:
@@
 \end{center}
-Possible sender and receiver strategies could be, for example, $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
-According to our formula above, these would have an expected utility of $2$ for both sender and receiver.
+Possible sender and receiver strategies are, for example, $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
+These would have an expected utility of $2$ for both sender and receiver, since when $t_1$ obtains the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, when $t_2$ obtains the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
 % the following expected utility:
@@
 Conventions of this kind in a signaling problem are what Lewis calls \emph{signaling systems}.
+An example of complete miscoordination would be $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
+Partial coordination is achieved, for example, by $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
 
+The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
+However, we can already see how it attempts to address the study of meaning from a very different angle.
+There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes.
+There is also no appeal to truth as a guide for our inquiry.
+In the examples discussed so far, agents are assumed to possess and exercise rationality.
+It is not, however, the same rationality as we find in the classical picture, \emph{i.e.}~the willingness to abide by truth; it is rather a more pragmatic notion of rationality as utility maximization.
+One advantage of such a paradigm shift is that we are no longer tied to the need to explain how language hooks on to the world; this question is no longer relevant on this account.
+We can merely focus on trying to see how agents can use signals to cope with the world and achieve their purposes.
+On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
+
+Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this use of his contribution.
+In order to get there, some aspects of the approach need to be refined.
 Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
-Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be the required common knowledge between the players involved.
+Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be a certain degree of required common knowledge between the players.
 Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
-These requirements can seem even more excessive if we consider how simple signaling systems are, compared to human languages.
-The purpose of studying them was to tell a story about how language as a conventional system could get off the ground without any prior sort of agreement.
-It is implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems that (hypothetically) evolved into languages.
-Also, communication through simple message exchange is something that almost all animals do, from monkeys to birds to bees.
-A plausible account of the origin of language should first explain how signaling systems like those could get started without a great deal of rationality from the part of the agents involved.
+These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
+The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
+However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
+Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
+A plausible account of the origin of language should first explain how signaling systems like those could get started, without assuming a great deal of rationality from the part of the agents involved.
+
+In order to address this problem, Skyrms proposes we study signaling problems in evolutionary terms.
+Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
+Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
+Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be.
+The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
+The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
+Using it, one can compute which strategies evolve under which conditions.
+
+Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalisms to the descriptive side of the investigation.
+Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
+Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
+Focus is put on understanding how various ingredients to the model interact, and which results they produce, not on metaphysical concerns.
+While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
+We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players\todo{give examples and references?}.
+This range of options goes hand in hand with assumptions of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
+Each of these can be utilized depending on the problem that one is interested in characterizing.
+
+The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
+Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?}.
+In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
+%; which one of the possible two will depend on their relative proportions in the original population.
+More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
+Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
+Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?}.
+
+So what about vagueness?
+In the context of game-theoretic approaches to language, vagueness presents a new challenge.
+Barton Lipman~\parencite*{lipman_why_2009} gives a detailed account of the problem.
+Very succinctly, it can be put as follows.
+In standard game-theoretic models of communication, vague usage of messages is less efficient than precise use.
+Therefore, given that the dynamics (be it natural selection, cultural evolution, or rational choice) maximize utility, vagueness should be weeded out by these forces, giving rise to only precise languages.
+However, vagueness is pervasive in our natural languages, and there is no reason to believe it is going away.
+Lipman concludes then that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}
+In the following section we provide an overview of models that try to account for vagueness in a signaling games framework, and what the implications of each account are for how we view rationality.
 
@@
 
+Before we discuss the particular suggestions of how to account for vagueness in a signaling games framework, we need to discuss what useful changes could we make to the classic setup.
+First, in order to work, the sorites paradox requires us to assume a relation between the vague term and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
+Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
+In terms of signaling games, we can take the height to be the state space and the terms the message space.
+Because of the difference in granularity, we will typically be interested in cases where the state space is much larger than the message space.
+For the structure of the state space, we can model this by defining a distance or similarity function between every element in the set, effectively making it a metric space.
+Second, another important ingredient of the paradox is the acknowledgment of a certain degree of tolerance with respect to whether a certain term applies or not.
+This tolerance decreases with distance: assuming a 180cm person is tall, one would easily tolerate the use of the term for a person measuring 179cm, less so for someone who is 170cm, and much less so for 160cm.
+This can be modeled by using a utility function that is continuous rather than discrete and that is monotonously decreasing with distance in a metric space, \emph{i.e.}~success is not a matter of black and white, right or wrong, but a matter of degree, of how close the agent got to the optimal.
+
+The simplest type of game to study in this scenario is one where the state space and the action space are the same.
+We can imagine this as a game of guessing the state of affairs: the sender has knowledge of it, sends a message to the receiver, who in turn has to guess it; their payoff, as discussed above, is proportional to how close he got.
+Other games are possible, depending on what situation of use we want to model.
+For example, imagine I have two editions of a book, one of them is red the other one is green.
+Wanting to read that book, I ask you to fetch the red book for me.
+If you bring me a different red book, whose hue is very similar to my prototypical notion of red, I would still be disappointed.
+If, however, you bring me my other edition of that book, despite it being far from red, I would be satisfied.
+This example is just to illustrate that there is more to the use of vague terms in natural language than what the setup we just described can cover.
+But, we believe, the setup captures one possible scenario where vagueness is relevant that would be difficult to capture in the classic formulations of signaling discussed before.
+
+These games, called similarity-maximization or sim-max games for short, were first introduced by Gerhard J{\"a}ger and Robert van~Rooij~\parencite{jager_language_2007,Jager2007} and further studied by~\textcite{jager_voronoi_2011}.
+What these authors find about this setup is that the evolutionary stable states are what they call Voronoi languages.
+Roughly, these are situations where the sender uses messages in a way that can be seen as partitioning the state space into convex regions and the receiver responds with the central element of those regions.
+For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}.
+Given two possible messages, the optimal sender strategy is to use one for all values from 40 up to 160, and the other for all values from 160 up to 280.
+The optimal receiver strategy is to guess with 100 if given the first message and with 220 given the second message.
+Confirming Lipman's argument, there is no vagueness in such an optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, associating only one guess with each message.
+
+\begin{figure}
+  \centering
+
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{Rplot-example-strict.pdf}
+    \caption{Crisp}
+    \label{fig:example_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{Rplot-example-vague.pdf}
+    \caption{Vague}
+    \label{fig:example_stratsB}
+  \end{subfigure}
+
+  \caption{Example strategies for a state space with 50 states.
+Each line corresponds to a message.
+For the sender, it plots for each state the probability that the message is used.
+For the receiver, it plots for each state the probability that the response is that state, given the message.}
+  \label{fig:example_strats}
+\end{figure}
+
+In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}.
+The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}.
+While in the former the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message, at a given point, in the latter these transitions are smooth.
+This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is almost certain.
+Similarly for the receiver strategy, in the first example for each message there is an almost certain response, whereas in the second example this is a degree of uncertainty.
+
+The interpretation of this uncertainty will be different depending on the interpretation of the model.
+If we see it as an explanatory model of how two agents play the game, we can see it as randomization.
+If we interpret the model as descriptive, it simply represents expected behavior in a manner agnostic to the underlying mechanisms.
+A third option is to see probabilities as capturing relative numbers in a population of agents.
+For example, if the sender strategy assigns a probability of $0.4$ of message $m$ being sent for state $t$, this would mean that $40\%$ of the population uses that message for that state, while $60\%$ uses the other.
+This latter option leaves open the possible interpretation that each agent commands a crisp language and vagueness is only a population-level effect.
+However, given the level of abstraction of the description so far, none of this is necessarily implied by the model.
+
+Our question is thus what additional modifications to sim-max games are sufficient for the optimal languages to be more like Figure~\ref{fig:example_stratsB}, rather than like Figure~\ref{fig:example_stratsA}.
+Let's look at some proposals in the literature.
+
+\subsection{Bounded rationality}
+\parencite{franke_vagueness_2010}
+
+\subsection{Generalized reinforcement}
+\parencite{oconnor_evolution_2014}
+
+\subsection{Noisy perception}
+\parencite{correia_bivalent_2013,franke_vagueness_2017}
+
+\subsection{Preference misalignment}
+\todo[inline]{Test CS~\parencite{crawford_strategic_1982} results in our model by using bias in utility function. READ Förster, M., and Riedel, F. (2011). Distorted Voronoi languages (working Paper).}
+
+\subsection{Noisy communication}
+\todo[inline]{Test if noise in communication channel as modeled by Blume~\parencite{blume_intentional_2014} induces vague strategies as well}
+
 \printbibliography

### 2017-03-09 book_chapter/paper.tex
Added reference to Huttegger's more recent overviews of signaling games literature. Summarized LMF and QRE.

@@
 Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
-Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?}.
+Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
+less examples?}.
+More recent overviews are given by \textcites{huttegger_how_2014} and \textcite{huttegger_dynamics_2014}.
 
@@
 \subsection{Bounded rationality}
-\parencite{franke_vagueness_2010}
+\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can arise as a consequence of limitations in the cognitive capacities of rational agents.
+The first proposal is called limited memory fictitious play and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
+For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
+Agents are assumed to be rational, thus each plays a best response (given the utility function) to their estimate of the other's strategy.
+In order to observe the evolution of strategies in repeated interaction, this approach involves modeling several individual agents in actual play.
+What the authors observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
+However, each agent still commands a crisp language, which is inadequate if the intention is to capture how vagueness presents itself in human languages.
+
+In order to overcome this limitation, they make another proposal using the notion of a quantal response equilibrium.
+The idea, inspired by experimental psychology, is to model the choice of best response as stochastic rather than deterministic, deviating from the optimal according to an exponentially decaying distribution (where small deviations are very likely, and large ones very rare).
+The approach is not explanatory but descriptive: it is assumed that agents are not always able to make the absolute best choice, but the mechanism behind this limitation is left open to interpretation.
+The degree to which agents deviate from the optimal behavior can be characterized by a parameter.
+The authors find that for low values of this parameter, only babbling equilibria (where sender and receiver simply randomize, respectively, message and response uniformly) where no communication actually occurs.
+Above a certain value of the parameter, other equilibria of the kind described in the beginning of this section arise, where agents communicate successfully, though not perfectly, using fuzzy strategies similar to the ones depicted in Figure~\ref{fig:example_stratsB}.
 
@@
 \subsection{Preference misalignment}
-\todo[inline]{Test CS~\parencite{crawford_strategic_1982} results in our model by using bias in utility function. READ Förster, M., and Riedel, F. (2011). Distorted Voronoi languages (working Paper).}
+\todo[inline]{Test CS~\parencite{crawford_strategic_1982} results in our model by using bias in utility function.
+READ Förster, M., and Riedel, F.
+(2011).
+Distorted Voronoi languages (working Paper).
+-> distorted Voronoi languages, i.e.
+shifted according to bias, but still crisp}
 

### 2017-03-18 book_chapter/paper.tex
Section on reinforcement learning.

@@
 \textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can arise as a consequence of limitations in the cognitive capacities of rational agents.
-The first proposal is called limited memory fictitious play and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
+The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
 For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
@@
 
-In order to overcome this limitation, they make another proposal using the notion of a quantal response equilibrium.
+In order to overcome this limitation, they make another proposal using the notion of a quantal response equilibrium (QRE).
 The idea, inspired by experimental psychology, is to model the choice of best response as stochastic rather than deterministic, deviating from the optimal according to an exponentially decaying distribution (where small deviations are very likely, and large ones very rare).
@@
 
-\subsection{Generalized reinforcement}
-\parencite{oconnor_evolution_2014}
+\subsection{Generalized reinforcement learning}
+Cailin O'Connor~\parencite*{oconnor_evolution_2014} proposes a way in which vagueness could be expected to evolve as a side-effect of a particular type of learning process.
+She studies sim-max games driven not by rational choice dynamics, but by generalized reinforcement learning (GRL), a variant of Herrnstein reinforcement learning (HRL)~\parencite{roth_learning_1995}.
+In HRL, agents learn to play a signaling game by strengthening particular choices (of messages for the sender, of responses for the receiver) proportionally to how successful that choice proves to be in an interaction.
+O'Connor's proposal is to model generalization as the propagation of reinforcement to nearby states, where ``nearby'' is defined in terms of distance in state space.
+For example, if a sender was successful in using message $m$ for state $t$, he will not only positively reinforce that choice of message for $t$, but also for states similar to $t$.
+This is done in a degree that is proportional to the degree of similarity between $t$ and other states.
+The dynamics also gives rise to vague signaling of the kind we are looking for.
+
+But why should agents evolve to generalize?
+%First, it is not argued that vagueness in itself is beneficial, but that it is a by-product of a learning mechanism.
+O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though not obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
+Considerations of optimality of strategies are typically made in terms of hypothetical atemporal comparisons of expected utility.
+However, generalized reinforcement learning has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a smaller period of time.
+In a naturalized evolutionary setting this temporary advantage should also be taken into account.
+Imagine an initial population of agents with random strategies, some using GRL and others using classical Herrnstein reinforcement learning to adapt to each other.
+Although the latter type of agent can hypothetically develop a precise and more efficient signaling system, agents using GRL would coordinate on vague signaling strategies with high (though not optimal) expected utility sooner than agents using HRL.
+In such a scenario, they could drive the other agents to extinction before the latter had time to achieve coordination and reap the benefits of a more precise signaling system.
+
+But would such a language be stable?
+In terms of evolutionary game theory, one could argue that a small enough group of mutants playing a precise signaling system could later invade a population of vague signalers.
+There is, an additional advantage to the generalized learning that is relevant here, namely the ability to adapt to a changing environment.
+Analysis of a game in terms of evolutionary stable states works well if we assume that the environment where the agents interact remains static.
+However, in a more realistic setting this is not the case.
+In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utiliry function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of their short-term gains.
 

### 2017-03-18 book_chapter/paper.tex
Shortened introduction.

@@
 The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
-The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if $1 000 000$ grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then if we remove one grain, the remaining $999 999$ also clearly form a heap; by repeated applications of this reasoning we can remove almost all grains of sand one by one and are lead to acknowledge that $1$ single grain of sand also forms a heap.
-This paradox is interesting because it can be made general and re-applied to many other words besides `heap'.
-In the general formulation, we start with an $x$ to which a certain predicate $P$ clearly applies.
-We observe that a certain transformation $f$ has no impact on whether something is or is not $P$.
-Thus we are lead to acknowledge that the predicate $P$ also applies to $y = f(x)$.
-Moreover, we can apply the same reasoning subsequently to $f(x)$ and repeat that as many times as the transformation is still applicable.
-The paradox arises when $n$ repeated applications of the transformation $f$ lead us from $x$ to $z = f^n(x)$, where $z$ is something for which originally we would say $P$ clearly does not apply, but since we see it as difficult to reject either the premisses or the reasoning, we seem to be forced to accept that $P$ \emph{does in fact apply} to $x$.
-The paradox with the heap can thus be seen as an instance of this general formulation, if $x$ is the group of $1 000 000$ grains of sand, the transformation is removing a grain, and $z$ the state of one (or even zero) grains.
-It can also be reversed: it seems reasonable to accept that piling up a grain of sand to something that is not a heap will not turn it into a heap of sand, thus 
-by repeated applications of that transformation one can pile up as many grains of sand as one likes and will never obtain a heap.
-
+The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, we can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, we would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
+Neither choice is, however, intuitively satisfying.
+The paradox is interesting because it can be made general and re-applied to many other words besides `heap'.
 Predicates for which one can find a suitable instance of the general formulation of the sorites paradox are called \emph{vague}.
 Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\parencite{Keefe1997}.
-An important intuition regarding these predicates that enables the sorites paradox to go through is that there does not seem to be any clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
-Intuitively, a person with a height of 2 meters is tall, and one with a height of 1.5 meters is not, but there is no value of height in between where one would draw the line; saying, for example, that a person with a height of 1.80 meters is tall, but a person measuring 1.79 meters is not, just seems preposterous.
-Neither drawing a sharp boundary seems reasonable, nor do the logical consequences of not doing so, and in there lies the truely paradoxical issue of vagueness.
 How widespread is the problem?
-It is easy to come up with more examples of predicates that are based on more finely grained properties, like `tall' is intuitively based on height, for which constructing a sorites paradox would be easy.
+It is easy to come up with more examples of predicates based on more finely grained properties---as `tall' is intuitively based on height---for which constructing a sorites paradox would be easy.
 Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
 If we subscribe to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
-Any heap of molecules that we have a name for is potentially at the mercy of being hypotheticall chipped away into nothingness while still keeping its name.
+Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
 The problem thus seems to be a serious one.
-Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
-In face of such a powerful paradox like the sorites, something has got to give.
+But it does not obviously undermine our everyday practices, so the issue must be with the conceptions we have of the mechanisms that underlie them.
 
-Vagueness is typically seen as a challenge to a classical conception of language and meaning.
-We can summarize this classical picture as such: words stand in a direct or mediated correspondence to their meanings; sentences are combinations of such words bound by logical rules; knowledge of the meanings of words and of how the rules combine them allows us to know the meaning of sentences and determine whether they are true or false.
-With variations, this picture could be said to underlie theories of meaning of authors life Gottlob Frege, Bertrand Russell, Ludwig Wittgenstein (his early work), Alfred Tarski, Richard Montague, and many others, especially those working in the analytic tradition.
-The problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to a well defined set of people?
-And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding this person?
-
-Some argue that the problems caused by vagueness are not serious enough for us to throw away the picture of meaning outlined above.
-Timothy Williamson, for example, claims that we should stick with the classical picture because ``[c]lassical logic and semantics are vastly superior to the alternatives in simplicity, power, past success, and integration with theories in other domains.''~\parencite*[162]{williamson_vagueness_1992}
-In the face of vagueness, Williamson defends the so-called \emph{epistemic view}: vague terms do determine precise boundaries, we just don't have the ability to know where they are; there is a precise height above which people are tall, and below which they are not, we are just ignorant of what this value is.
-%%% Could talk about Frazee, J., and Beaver, D. (2010) here
-The epistemic approach is curious in that, in its attempt to save truth, it seems to actually render it sterile as an explanatory notion.
-If we can never ascertain truth when it comes to vague terms, how can truth be relevant for explaining how we understand or produce language?
-If the concept of truth is crucially tied to meaning, but we have no access to the former, how could we ever have learned the latter?
-Williamson's reason to ignore these problems is a pragmatic one: he believes it helps us, philosophers of language and theoretical linguists, to better understand natural language.
-But this reason holds only insofar as we agree with this assessment of the success of the classical picture.
-
-Criticism of the adequacy of this picture, however, comes from various quarters.
-One of the strongest can be found in the later work of philosopher Ludwig Wittgenstein\footnote{To the extent that substantial views can be said to be defended by the author.
-Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}, most importantly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}.
-Undermining several assumptions behind the classical picture and the implications thereof, Wittgenstein pushes for a paradigm shift:
-\begin{quote}
-The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
-(For the crystalline purity of logic was, of course, not something I had \emph{discovered}: it was a requirement.)
-The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
--- We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
-We want to walk: so we need \emph{friction}.
-Back to the rough ground!%
-~\parencite[\S 107]{wittgenstein_philosophical_1953}
-\end{quote}
-He argues that the ideal of exactness that is tied to the notions of truth and logic dazzles and leads to misunderstandings~\parencite*[\S 100]{wittgenstein_philosophical_1953}; it creates a haze around the workings of language~\parencite*[\S 5]{wittgenstein_philosophical_1953} that keeps us from seeing what is right in front of us.
-We can only escape this predicament, not by bargaining exactness out of logic, but by turning our whole inquiry around~\parencite*[\S 108]{wittgenstein_philosophical_1953}, \emph{i.e.}~by thinking of language in terms of a different paradigm.
-
-%%% This discussion is perhaps too detailed for the scope of this paper
-% Another suggested theory that attempts to retain as much as possible from classical logic by making minimal modifications to the classical picture of semantics is \emph{supervaluationism}.
-% Championed by Kit Fine~\parencite*{Fine1975}, the idea is that vague predicates can be made precise, even arbitrarily.
-% Setting a boundary between tall and not tall people at 1.80 meters would be an admissible%
-% \footnote{Only precisifications that meet certain constraints should be considered. We refer the reader to Fine's article~\parencite*{Fine1975} for the details.}
-% \emph{precisification} of the word `tall'.
-% What supervaluationism defends is that, when considering the truth value of a sentence with vague terms, we should take all admissible precisifications of those terms into account: the sentence is true if true under all of them, false if false under all of them, and neither true nor false otherwise.
-% The intuition is that a person with a height of 2 meters would be considered tall under all admissible precisifications, thus we could say it is true that that person is tall; for someone measuring 1.79 this would not be the case, hypothetically being true under some precisifications and false under others.
-% %In defending this approach, supervaluationism either assumes or postulates the potential existence of a boundary between the sentences for which all admissible precisifications of the vague terms therein make them true, and those for which at least one precisification makes them false.
-% Although supervaluationism claims to retain classical logics, it seems to implicitly require a third truth value to account for sentences for which not all precisifications are in agreement.
-% Other responses to the problem of vagueness develop alternative logics that explicitly break away from the bivalent assumption that a sentence is either true or false, and introduce additional truth values.
-% These range from three-valued logics~\parencite[\emph{e.g.}][]{tye_sorites_1994} to infinite-valued degree theories~\parencite[\emph{e.g.}][]{machina_truth_1976}.
-
-The way in how the ideal of exactness leads into misunderstandings is somewhat patent in the standard approaches to dealing with vagueness that try to hang on to as much as possible of the classical picture.
-Supervaluationism, many-valued logics, and degree theories all propose changes to it in order to accommodate for vague predicates.
-The proposals are, however, still within the general molds of the classical picture.
+Vagueness is typically seen as a challenge to the classical conception of language and meaning.
+One problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to something like a well defined set of tall people?
+And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding these borderline cases?
+Supervaluationism, many-valued logics, and degree theories all propose changes to the classical picture in order to accommodate for vague predicates.
+They are, however, still strongly committed to retaining as much as possible of it, particularly the core notions of truth and reference.
 Mark Sainsbury~\parencite*{sainsbury_concepts_1999} argues that, because of that, they all fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
@@
 The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also consequently need to be abandoned down the line.
-Talking about philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
+One notion that seems to quickly be in peril in that of rationality.
+In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
 \begin{quote}
@@
 To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
-Vagueness can shed light on features of our language and our rationality, but only after we step out of the classical picture and start looking at these concepts in terms of a different paradigm.
+What if we try to step out of the classical picture and start looking at meaning and rationality in terms of a different paradigm?
 
 What can such a paradigm be?
-The Wittgensteinian picture of language is one of multiplicity, heterogeneity, and change.
-Languages are thought of as patchworks of various language-games, with new ones continuously being added and old ones falling out of use.
-These language-games, in turn, are situated uses of words in certain activities.
+Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
+\footnote{To the extent that substantial views can be said to be defended by the author.
+Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
+Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} we can find a picture of language is terms of multiplicity, heterogeneity, and change.
+Languages are seen as patchworks of various language-games, with new ones continuously being added and old ones falling out of use.
+These language-games, in turn, can be thought of language intermingled with a practice.
 The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
-Meaning is thus contingent to the language-games we play, and dynamic.
+There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
 The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
-Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up such a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play through according to those assumptions.
-Jos\'e Pedro Correia~\parencite*{correia_bivalent_2013} argues that the framework of signaling games, introduced for the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that, while reaping the benefits of a mathematical formalism and the further ability to conduct computer simulations.
-From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting computer simulations is like contemplating how the game would play through according to the implications of the assumptions one built into the model.
+Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play through according to those assumptions.
+It has been argued~\parencite{correia_bivalent_2013} that the framework of signaling games, introduced for the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that, while reaping the benefits of mathematical formalism and computer simulation to explore the implications of complex hypothetical language-games in depth.
+From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting computer simulations is like contemplating how the game would play through according to the assumptions one built into the model.
 

### 2017-04-05 book_chapter/paper.tex
Shortened intro to signaling games

@@
 Signaling games were first introduced as models of communication by David Lewis~\parencite*{lewis_convention_1969}.
-His original objective was to address an argument raised by W.~V.~Quine~\parencite*{quine_truth_1936} and others against the possibility of language having started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
-To this, Lewis retorts:
-\begin{quote}
-I offer this rejoinder: an agreement sufficient to create a convention need not be a transaction involving language or any other conventional activity.
-All it takes is an exchange of manifestations of a propensity to conform to a regularity.%
-~\parencite*[87--88]{lewis_convention_1969}
-\end{quote}
-In order to support this claim, Lewis studies coordination problems formalized in terms of game theory.
+His original objective was to address arguments raised against the possibility of language having started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
+In order to support the idea that conventions can arise without any prior conventional activity, Lewis studies coordination problems formalized in terms of game theory.
 These are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
@@
 
-The setup of a coordination problem, \emph{i.e.}~the available actions and the relative interests of the players, can be represented in a payoff matrix, where rows show one player's available actions, columns the other player's available actions, and each cell gives the payoff for each of the players based on its row and column combination.
-For example, consider the following matrix:
-\begin{center}
-\begin{tabular}{r|c|c|}
-\multicolumn{1}{r}{}
- & \multicolumn{1}{c}{$b_1$}
- & \multicolumn{1}{c}{$b_2$} \\ \cline{2-3}
-   $a_1$ & $1,1$ & $0,0$ \\ \cline{2-3}
-   $a_2$ & $0,0$ & $1,1$ \\ \cline{2-3}
-\end{tabular}
-\end{center}
-We can see this as representing the following game: one player has actions $A = \lbrace a_1, a_2 \rbrace$ available, the other $B = \lbrace b_1, b_2 \rbrace$; they prefer to coordinate $a_1$ with $b_1$ or $a_2$ with $b_2$, thus if this is achieved each gets a payoff of $1$, otherwise they each get $0$.
-Connecting with the example laid out above, think of $a_1$ as Alice going to Caf\'e One, $a_2$ as Alice going to Bistro Two, $b_1$ as Bob going to Caf\'e One, and finally $b_2$ as Bob going to Bistro Two; their payoff is $1$ for both if they coordinate on the place to meet (independently of which one), and $0$ if they miscoordinate.
-Formally, all we need to define in order to characterize such a situation are the sets of actions $A$ and $B$, and a utility function $U : A \times B \rightarrow \mathbb{R}^2$ that specifies the payoff for each player given the choices of actions.
-
-The above example is a very simple case, but it serves to illustrate Lewis' notion of convention.
-The two pairs of choices $(a_1,b_1)$ and $(a_2,b_2)$ are stable coordination equilibria, because in such a scenario no player has an incentive to unilaterally change his choice: if the first player is going to choose action $a_1$, the second player would get a payoff of $0$ for switching to $b_2$, instead of $1$ from sticking to $b_1$; the same reasoning applies, mutatis mutandis, to the other player and the other equilibrium.
-However, neither player has a preference between those two pairs.
-If the same coordination problem arises again, Lewis says, we can expect precedence to induce a kind of regularity: if the players manage to coordinate on one of the two equilibria, they should be expected to repeat the choices of actions that lead to that success, thus remaining in the same equilibrium.
-Not only can we expect the players in such a repeated interaction to conform to a regularity, but also to expect others to do the same.
-Their preference is to remain in a certain equilibrium given that others do too.
-This summarizes Lewis's notion of a convention.
-
-In order to extend this notion to \emph{linguistic} conventions, Lewis considers situations where the actions available involve sending and receiving signals or messages.
+In order to extend his notion of convention to \emph{linguistic} exchanges, Lewis considers situations where the actions available involve sending and receiving signals or messages.
 Thus, we could think of two players with different roles.
-The first player, which Lewis calls the communicator~\parencite*[130]{lewis_convention_1969} and we will here call the sender, has knowledge about which of a number of possible states of affairs obtains and, depending on this information, chooses a signal to send.
-The second player, which Lewis calls the audience~\parencite*[130]{lewis_convention_1969} and we will here call the receiver, has knowledge about which signal the sender chose and, based on this information, chooses one of several possible responses.
+The first player, the sender, has knowledge about which of a number of possible states of affairs obtains and, depending on this information, chooses a signal to send.
+The second player, the receiver, has knowledge about which signal the sender chose and, based on this information, chooses one of several possible responses.
 A preference relation exists between responses and states of affairs, and a payoff is attributed to each player based on the choices of both.
@@
 These would have an expected utility of $2$ for both sender and receiver, since when $t_1$ obtains the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, when $t_2$ obtains the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
-% the following expected utility:
-% \begin{equation}
-% \begin{split}
-% EU(\sigma, \rho) & = \sum_{t \in T} U(t, \rho(\sigma(t))) \\
-%                  & = U(t_1, \rho(\sigma(t_1))) + U(t_2, \rho(\sigma(t_2))) \\
-%                  & = U(t_1, \rho(m_2)) + U(t_2, \rho(m_1)) \\
-%                  & = U(t_1, t_1) + U(t_2, t_1) \\
-%                  & = (1,1) + (1,1) \\
-%                  & = (2,2)
-% \end{split}
-% \end{equation}
-% Based on this, we could also create a matrix of expected utilities for this example as follows:
-% \begin{center}
-% \begin{tabular}{r|c|c|c|c|}
-% \multicolumn{1}{r}{}
-%  & \multicolumn{1}{c}{$m_1 \mapsto a_1$}
-%  & \multicolumn{1}{c}{$m_1 \mapsto a_1$}
-%  & \multicolumn{1}{c}{$m_1 \mapsto a_2$}
-%  & \multicolumn{1}{c}{$m_1 \mapsto a_2$} \\
-% \multicolumn{1}{r}{}
-%  & \multicolumn{1}{c}{$m_2 \mapsto a_1$}
-%  & \multicolumn{1}{c}{$m_2 \mapsto a_2$}
-%  & \multicolumn{1}{c}{$m_2 \mapsto a_1$}
-%  & \multicolumn{1}{c}{$m_2 \mapsto a_2$} \\ \cline{2-5}
-%    $t_1 \mapsto m_1, t_2 \mapsto m_1$ & $1,1$ & $1,1$ & $1,1$ & $1,1$ \\ \cline{2-5}
-%    $t_1 \mapsto m_1, t_2 \mapsto m_2$ & $1,1$ & $2,2$ & $0,0$ & $1,1$ \\ \cline{2-5}
-%    $t_1 \mapsto m_2, t_2 \mapsto m_1$ & $1,1$ & $0,0$ & $2,2$ & $1,1$ \\ \cline{2-5}
-%    $t_1 \mapsto m_2, t_2 \mapsto m_2$ & $1,1$ & $1,1$ & $1,1$ & $1,1$ \\ \cline{2-5}
-% \end{tabular}
-% \end{center}
 They also represent one of the two stable conventions in this game, the other being the pair of strategies $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
@@
 
-Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this use of his contribution.
+Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this.
 In order to get there, some aspects of the approach need to be refined.
@@
 Each of these can be utilized depending on the problem that one is interested in characterizing.
+Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.
 

### 2017-04-10 book_chapter/paper.tex
Merged introduction on vagueness with introduction on signaling games

@@
 From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting computer simulations is like contemplating how the game would play through according to the assumptions one built into the model.
-
-In the following, we will be looking at signaling models of vagueness in order to provide an overview of proposals of natural sources of vagueness, what that tells us about rationality, and the implications of these ideas.
-But first, let us look in more detail into how Lewis-Skyrms signaling games work, and how vagueness shows up again as a different kind of problem for these models.
-%With regards to vagueness and exactness, it is also important to note the following:
-%\begin{quote}
-%``Inexact'' is really a reproach, and ``exact'' is praise.
-%And that is to say that what is inexact attains its goal less perfectly than does what is more exact.
-%So it all depends on what we call ``the goal''.
-%\end{quote}
-
-%%% This is perhaps too philosophical and off-topic
-% The reluctance to give up truth and logic as fundamental concepts to understand meaning is perhaps related to a fear of losing a grip on rationality, and more importantly, on the supposed necessity of these concepts to underwrite normative statements and thus condition discourse.
-% Richard Rorty describes this as follows:
-% \begin{quote}
-% The traditional view is that there is a core self which can look at, decide among, use, and express itself by means of, [\ldots] belief and desires.
-% Further, these beliefs and desires are criticizable not simply by reference to their ability to cohere with one another, but by reference to something exterior to the network within which they are strands.
-% Beliefs are, on this account, criticizable because they fail to correspond to reality.
-% Desires are criticizable because they fail to correspond to the essential nature of the human self -- because they are ``irrational'' or ``unnatural''.
-% \end{quote}
-% Since the linguistic turn, language is seen as the medium that puts the self in contact with either reality or the nature of the self.
-% Rationality is, in this picture, dependent on truth as its objective and logic as the means to attain it.
-% The latter are embodied in language.
-% The fear is thus that if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground that rationality stands on, and thus lose the ability to make value judgments on sentences from a supposedly objective standpoint.
-% Full-fledged relativism thus ensues.
-
-\section{Signaling games and another problem of vagueness}
-\label{sec:signaling-and-Lipman}
 

### 2017-04-11 book_chapter/paper.tex
Shortened the intro on signaling games a bit more. Tried to better fit discussions of rationality.

@@
 Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
+Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
+Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
+This type of rationality is one that supposedly demarcates humans from other animals.
 The fear could be that, if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground on which rationality stands, and the whole edifice would collapse.
@@
 To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
-What if we try to step out of the classical picture and start looking at meaning and rationality in terms of a different paradigm?
 
+The aforementioned classical picture of language and meaning is thus focused on an aspect of rationality that we can call \emph{procedural}.
+There are, however, other aspects that are also important.
+We can think of rationality as \emph{instrumental}, \emph{i.e.} as the efficient and effective achieving of goals (see Brown~\parencite*[???]{brown_rationality_1990}).
+An agent is rational if he employs his available means in the best way that helps him achieve a given objective.
+This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
+Other animals, however, do this all the time.
+The difference with humans, if we want to establish one, can reveal yet another aspect of rationality.
+Jon Elster~\parencite*{elster_ulysses_1979} suggests that our rationality is characterized by its ability to be more \emph{strategic}, \emph{i.e.}~it can take into account more than the expected immediate results of an action, like anticipation of other players' actions, or consideration of medium/long term gains.
+All of these aspects of rationality (procedural, instrumental, strategic) are probably intertwined in our understanding of rationality.
+
+Do we get different pictures of rationality if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
 What can such a paradigm be?
@@
 Signaling games were first introduced as models of communication by David Lewis~\parencite*{lewis_convention_1969}.
-His original objective was to address arguments raised against the possibility of language having started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
-In order to support the idea that conventions can arise without any prior conventional activity, Lewis studies coordination problems formalized in terms of game theory.
-These are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
-In game theory terms, the agents interested in the coordination are the players, the game involves each player making an independent choice from his set of available actions, a payoff is what is attributed to each player based on the choices of both.
-Adapting one of Lewis' examples, say Alice and Bob want to get together.
-They usually meet at either Caf\'e One or Bistro Two.
-Imagine there is no way for them to make any explicit agreement about it.
-They are thus left to independently decide to either go to Caf\'e One or go to Bistro Two and hope for the other to show up there.
-Neither has any preference for either place, but they do want to meet.
-Each thus prefers to go to one of the places only if the other also decides to go to that particular place.
+% His original objective was to address arguments raised against the possibility of language having started as a conventional system: if language is a convention, it had to be originally established by an agreement; in order to establish an agreement, a convention-governed system of communication would have to already have been in place; thus, although some languages could have been established by agreement if another convention was already in place, not all of them could.
+% In order to support the idea that conventions can arise without any prior conventional activity, Lewis studies coordination problems formalized in terms of game theory.
+% These are ``situations of interdependent decision by two or more agents in which coincidence of interest predominates and in which there are two or more proper coordination equilibria''~\parencite*[24]{lewis_convention_1969}.
+% In game theory terms, the agents interested in the coordination are the players, the game involves each player making an independent choice from his set of available actions, a payoff is what is attributed to each player based on the choices of both.
+% Adapting one of Lewis' examples, say Alice and Bob want to get together.
+% They usually meet at either Caf\'e One or Bistro Two.
+% Imagine there is no way for them to make any explicit agreement about it.
+% They are thus left to independently decide to either go to Caf\'e One or go to Bistro Two and hope for the other to show up there.
+% Neither has any preference for either place, but they do want to meet.
+% Each thus prefers to go to one of the places only if the other also decides to go to that particular place.
 
-In order to extend his notion of convention to \emph{linguistic} exchanges, Lewis considers situations where the actions available involve sending and receiving signals or messages.
+% In order to extend his notion of convention to \emph{linguistic} exchanges
+In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where the actions available involve sending and receiving signals or messages.
 Thus, we could think of two players with different roles.
@@
 Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.
-
-Despite the added dimension of the signal exchange, these so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
+These so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
 A communicator's contingency plan, or sender strategy as we will call it, is a specification of a choice of message for each possible state of affairs.
@@
 There is also no appeal to truth as a guide for our inquiry.
-In the examples discussed so far, agents are assumed to possess and exercise rationality.
-It is not, however, the same rationality as we find in the classical picture, \emph{i.e.}~the willingness to abide by truth; it is rather a more pragmatic notion of rationality as utility maximization.
 One advantage of such a paradigm shift is that we are no longer tied to the need to explain how language hooks on to the world; this question is no longer relevant on this account.
@@
 On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
+In the examples discussed so far, agents are assumed to possess and exercise rationality.
+It is not, however, the procedural notion of rationality of the classical picture, but rather one that is instrumental and heavily strategic.
 
-Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this.
-In order to get there, some aspects of the approach need to be refined.
-Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
+% Lewis signaling games can be seen as an alternative to the classical picture, even though the author himself might not have agreed with this.
+% In order to get there, some aspects of the approach need to be refined.
+% Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
 Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be a certain degree of required common knowledge between the players.
-Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, rationality, mutual ascription of rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
+Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, strategic rationality, mutual ascription of strategic rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
 These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
 The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
-However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
+However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of strategic rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
 Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
-A plausible account of the origin of language should first explain how signaling systems like those could get started, without assuming a great deal of rationality from the part of the agents involved.
+A plausible account of the origin of language should first explain how signaling systems like those could get started, without assuming a great deal of strategic rationality from the part of the agents involved.
 
-In order to address this problem, Skyrms proposes we study signaling problems in evolutionary terms.
+In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes we study signaling problems in evolutionary terms.
 Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
@@
 We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players\todo{give examples and references?}.
-This range of options goes hand in hand with assumptions of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
+This range of options goes hand in hand with assumptions of rationality, from nothing more than a basic instrumental rationality that aims for survival in a biologically-inspired setting, to a certain degree of strategic rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
 Each of these can be utilized depending on the problem that one is interested in characterizing.
-Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.
+Thus, although highly strategic rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where it can play a role.
 
-The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
-Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?}.
-In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
+% The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
+% Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?}.
+% In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
 %; which one of the possible two will depend on their relative proportions in the original population.
-More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
-Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
-Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
+% More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
+% Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
+% Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
 less examples?}.
-More recent overviews are given by \textcites{huttegger_how_2014} and \textcite{huttegger_dynamics_2014}.
+% More recent overviews are given by \textcites{huttegger_how_2014} and \textcite{huttegger_dynamics_2014}.
 
-So what about vagueness?
+%So what about vagueness?
 In the context of game-theoretic approaches to language, vagueness presents a new challenge.

### 2017-04-19 book_chapter/mf_notes.tex
note taking MF

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{myarticlestyledefault}

\title{Notes on vagueness \& rationality}
\author{Michael Franke}
\date{April 19 2017}

\begin{document}
\maketitle

\section*{Where this is going}

\begin{itemize}
\item The main question to be addressed is:
  \begin{quote}
    How is vagueness in language compatible with rationality and/or evolutionary selection for
    efficient communication?
  \end{quote}
\item The answer I will try to give is something like:
  \begin{quote}
    If whatever we want to call ``language'' is not vague, it is likely quite a useless notion.
  \end{quote}
  With ``quite useless notion'' I do mean ``philosophically'' or ``theoretically useless'', much in
  the vein of Wittgenstein's later views on logical analysis of language: yes, we can do most
  of this, but it shows us little (supply proper references).
\item The way to give this answer is by a more careful analysis of rational or evolutionary
  efficient language use. We look at signaling games and models of rational agency /
  evolutionary selection. It is easy to identify multiple reasons why this idealized picture
  breaks by giving up a number of assumptions, none of which is necessarily that of idealized
  rationality or evolutionary optimization.
\end{itemize}

\section*{Where do we find the meaning of a language?}

\begin{itemize}
\item There are different places where ``a language's meaning'' might reside:
  \begin{enumerate}
  \item at the \textbf{population-level}; or
  \item at the \textbf{agent-level}, where we might be interested in either:
  \begin{enumerate}
    \item an agent's behavior (language use \& interpretation across contexts); or
    \item her beliefs (about language use, or semantic meaning)
  \end{enumerate}
\end{enumerate}
\item it may seem natural to assume that population-level meaning supervenes on something at
  the agent-level (either behavior, beliefs or both) of individual agents;
  \begin{itemize}
  \item this is putting aside the option that meanings are innate concepts or any such
    esoteric moves
  \end{itemize}
\item We note that there is no generally accepted (formal) criterion for how exactly a
  language's meaning arises from anything more downstream (agent-level stuff, ways of
  interaction, etc.). We have to make do in this vacuum.
  \begin{itemize}
  \item Some attempts to fill the vacuum:
    \citet{Harms2004:Information-and,Harms2010:Determining-tru,Franke2013:An-adaptationis}
  \end{itemize}

\end{itemize}

\section*{A closer look at the agent's beliefs and behavior}

\begin{itemize}
\item game theoretic models of signaling agents define agent behavior as possibly probabilistic
  functions:
  \begin{itemize}
  \item sender behavior $P_S(\messg \mid \state)$ assigns a probability to the choice of message
    $\messg$ to each state $\state$
  \item receiver behavior $P_R(\act \mid \messg)$ assigns a probability to the choice of
    reaction $\act$ after observing message $\messg$
  \end{itemize}
\item When we fix a context (think: sim-max signaling game), we can define which choices are
  rational. We see that, indeed, no vagueness shows.
  \begin{itemize}
  \item \textcolor{gray}{MF2JPC: I'll skip this here; you did most of this already and its
      clear to us anyway}
  \end{itemize}
\item This is all rather impoverished. You, I, she and he don't just play one game. Utilities
  differ in different situations. We also don't always play sim-max games, but may play a
  reference game. Speakers and listeners may have partial knowledge of the situation and of
  each other. They may also have beliefs about meaning, and their choices would depend on
  these. Recent probabilistic models of pragmatic language use explicitly models agents
  (beliefs) about each other's mental lexica
  \citep{BergenLevy2012:Thats-what-she-,BergenLevy2014:Pragmatic-Reaso,PottsLassiter2016:Embedded-implic}.
  \begin{itemize}
  \item More on dynamic adaptivity to speaker idiolect (Yildirim et al., \dots
    \citet{Davidson_1986:Derangement})
  \end{itemize}
\item The picture of language users which we should adopt minimally adopt is this:
  \begin{itemize}
  \item speaker behavior is a probabilistic function $P_S(\messg \mid \state, O^S_c, B^S_l)$
    specifying a message choice probability for a meaning/state $\state$, a (possibly partial)
    observation of the context $O_c$ and a belief $B_l$ about the language's meaning
  \item listener interpretation behavior is similarly a probabilistic function
    $P_L(\state \mid \messg, O^L_c, B^L_l)$
    \begin{itemize}
    \item speaker and listener beliefs about context and meaning need not be the same; in fact,
      almost always they will be different
    \item the beliefs about contextual parameters $O_c$ which may influence choice behavior are
      bountiful:
      \begin{itemize}
      \item think: shape and parameters of the utility function in a sim-max game
      \item imprecise perception of the referent set in a reference game
      \item different beliefs about anything related to the interlocutor
      \item \dots
      \end{itemize}
    \item the beliefs $B_l$ about the (lexical \& compositional) meanings of expressions could,
      for simplicity, be captured as a distribution over all possible Boolean lexica, where a
      Boolean lexicon $l$ assigns a subset set of states to each message: $\den{\messg}_l
      \subseteq \States$
    \end{itemize}
  \end{itemize}
\item Nothing has been said so far how beliefs are acquired, whether these beliefs are
  rational, how choice are made based on beliefs. There are many options. We cannot look at
  them all. We will browse a few and note how they all naturally give rise to something like
  ``vagueness'' without abandoning rationality.

\end{itemize}

\newpage

Sources of vagueness:

- uncertainty about threshold:
   - learning from finite examples
   - rich language makes decisions about borderline cases obsolete; enhances problem of learning data scarcity
   - imperfect memory
- uncertainty about QUD, utility, interlocutor's beliefs

=> why not just have a prior and integrate and have a fixed threshold

1. Priors are non-introspective; sampling based assessment

2. Choice is probabilistic -> utilities are noisy

Why not adopt a fixed threshold? Why is an agent with uncertainty in the lexicon better off? => revision cost? surprisal cost?

If a single agent's semantic representation is the outcome of rational learning and possibly an internal optimization towards efficient communication.

Finiteness of life is not irrational.

Convince by number of examples of natural sources for vagueness.

Any precise threshold must be a function of priors about properties, QUDs, utilities, and interlocutor beliefs. But these are not stable over time. Vagueness might be optimal to compensate mistakes in a changing environment. Assumption here: revision cost high when too low prob on observable events. (van Deemter example: DECENT behavior)

If speakers' uncertain beliefs come from natural probabilistic evidence, they will not be sharp step-functions. If we measure success of communication by KL-divergence or similar, the update of a literal listener with a vague threshold would give higher utility. But what about pragmatic listeners?

Suppose agents play repeated reference games with different naturals kinds. Each kind has a mean and variance for each of several open and closed scale features. Learners entertain different hypotheses about lexical meaning (McNally?) They would like to assign a fixed degree to an expression because that is economical. They can do that for closed scale natural kinds unless their mean and variance is such that exemplars are frequently away from the end points. Think wine glasses! They cannot do this for open scales. They must consider a more difficult hypothesis, namely to use the prior after all.

Good point by van Deemter: reliable metrics may be absent; the goal of communication may be expression of sentiment, and the real value may be assessed in a goal-oriented way (BIG when playing soccer not equal to big when painting).

Languages supervene on agents' beliefs about lexica. Lexica are uncertain. They are still useful. Lexical uncertainty might even be rational in the light of belief revision in changing environments.

Languages are optimized for communication and social interaction. Not necessarily for logical consistency.

Plurality of language games forces uncertainty. Wittgenstein.

So then logic breaks, so what? The possibility of visual illusions does not prove the visual system ill adapted or suboptimal.

\printbibliography[heading=bibintoc]

\end{document}

### 2017-04-20 book_chapter/paper.tex
minor edits main paper

--- 2017-04-05 book_chapter/paper.tex
@@
 
+\usepackage{xcolor}
+\definecolor{Red}{RGB}{178,34,34}
+\newcommand{\mf}[1]{\textcolor{Red}{[mf: #1]}}
+
 \begin{document}
@@
 \title{Natural sources of vagueness and their implications}
-\author{Michael Franke \and Jos\'e Pedro Correia}
+\author{Jos\'e Pedro Correia \and Michael Franke}
 \date{}
@@
 
-The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
+The classic philosophical problem of vagueness is most starkly embodied by the sorites paradox.
 The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, we can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, we would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
@@
 Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
-Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} we can find a picture of language is terms of multiplicity, heterogeneity, and change.
+Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} we can find a picture of language in terms of multiplicity, heterogeneity, and change.
 Languages are seen as patchworks of various language-games, with new ones continuously being added and old ones falling out of use.
@@
 Neither has any preference for either place, but they do want to meet.
-Each thus prefers to go to one of the places only if the other also decides to go to that particular place.
+Each thus prefers to go to one of the places only if the other also decides to go to that
+particular place. \mf{shorten this paragraph? and the next?}
 
@@
 Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
-Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.
+Formally, in order to describe the setup all we need is to specify a set of possible states of
+affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and
+the utility function $U : T \times A \rightarrow \mathbb{R}^2$. \mf{explain sender and receiver
+payoffs}
 
@@
 
-In order to address this problem, Skyrms proposes we study signaling problems in evolutionary terms.
-Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
-Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
-Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be.
-The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
-The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
-Using it, one can compute which strategies evolve under which conditions.
+In order to address this problem, Skyrms proposes we study signaling problems in evolutionary
+terms.  Rather than imagining, as Lewis does, rational agents making conscious decisions in
+possession of knowledge of the game and expectations of the behavior of other agents, we can
+imagine a simpler scenario inspired by biological evolution: there is a population of agents
+with biologically hardwired behaviors for engaging in interactions characteristic of a
+signaling problem; utility does not represent preference, but rather fitness for survival and
+reproduction; the make-up of the population evolves based on the relative fitness of the
+strategies represented in the population.  Such a setup attempts to capture the main features
+of natural selection: in a diverse population, agents with more successful strategies thrive,
+while agents with less fit strategies die off.  Although the inspiration for this scenario is
+biological evolution, similar things could be
+said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread
+in a population of agents who can adopt or abandon them depending on how successful they prove
+to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-}.  The
+principles can be captured in a formal model that abstracts away from the interpretations: the
+replicator dynamics.  The only thing relevant to this equation are the relative proportions of
+strategies in a given population and the utility function.  Using it, one can compute which
+strategies evolve under which conditions.
 
-Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalisms to the descriptive side of the investigation.
-Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
-Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
-Focus is put on understanding how various ingredients to the model interact, and which results they produce, not on metaphysical concerns.
-While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
-We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players\todo{give examples and references?}.
-This range of options goes hand in hand with assumptions of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
-Each of these can be utilized depending on the problem that one is interested in characterizing.
-Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.
+Skyrms' evolutionary game theory approach to signaling games not only gives more plausible
+grounds to support Lewis' discussion of convention, but it also accomplishes an important
+conceptual change: it moves most of the theory and mathematical formalisms to the descriptive
+side of the investigation.  Utility represents how the modeler views the signaling problem and
+understands the relative advantages or disadvantages of different possible strategy
+combinations.  Dynamics describe how strategies can evolve when driven by mechanisms of utility
+maximization.  Focus is put on understanding how various ingredients to the model interact, and
+which results they produce, not on metaphysical concerns.  While the general framework manages
+to abstract quite some detail away from the formalization, it does leave room for them,
+especially when it comes to the dynamics.  We already mentioned the replicator equation that
+can be seen as representing biological or cultural evolution, but one can also use dynamics
+inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game
+and other players~\parencite[\emph{e.g.}][]{Muhlenbernd2011:Learning-with-N}.  This range of
+options goes hand in hand with assumptions of rationality, from nothing more than survival of
+the fittest in a biologically-inspired setting, to a certain degree of rationality in the case
+of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of
+rationality and even game-theoretical reasoning.  Each of these can be utilized depending on
+the problem that one is interested in characterizing.  Thus, although Skyrms shows that
+rationality is not necessary for signaling conventions to evolve, the framework does leave room
+for the study of linguistic interactions where rationality can play a role.
 
-The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
-Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?}.
+The characterization of signaling problems in terms of evolutionary game theory allow us to
+explain why certain equilibria come to be and how. \mf{explain/introduce equilibrium and/or ESS?}
+Not only can we better understand why signaling systems are stable even without any assumptions
+of rationality, but we can also map out which initial conditions drive the system towards which
+equilibria and which don't\todo{mention something about phase portraits? \mf{too much detail,
+    rather not!}}.
 In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
@@
 Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
-Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
-less examples?}.
-More recent overviews are given by \textcites{huttegger_how_2014} and \textcite{huttegger_dynamics_2014}.
+Other uses of signaling games include discussions of
+categorization~\parencite[\emph{e.g.}][]{jager_language_2007},
+compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009},
+incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a
+few\todo{more examples? less examples? \mf{good as it is}}.
+More recent overviews are given by \textcites{huttegger_how_2014},
+\textcite{huttegger_dynamics_2014} or \citet{FrankeWagner2014:Game-Theory-and}.
 
@@
 However, vagueness is pervasive in our natural languages, and there is no reason to believe it is going away.
-Lipman concludes then that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}
+Lipman concludes then that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}.
 In the following section we provide an overview of models that try to account for vagueness in a signaling games framework, and what the implications of each account are for how we view rationality.
@@
 
-Before we discuss the particular suggestions of how to account for vagueness in a signaling games framework, we need to discuss what useful changes could we make to the classic setup.
+Before we discuss the particular suggestions of how to account for vagueness in a signaling
+games framework, we need to discuss what useful changes we could make to the classic setup.
 First, in order to work, the sorites paradox requires us to assume a relation between the vague term and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
@@
 This example is just to illustrate that there is more to the use of vague terms in natural language than what the setup we just described can cover.
-But, we believe, the setup captures one possible scenario where vagueness is relevant that would be difficult to capture in the classic formulations of signaling discussed before.
+But, we believe, the setup captures one possible scenario where vagueness is relevant that
+would be difficult to capture in the classic formulations of signaling discussed before. \mf{I
+  suggest to enlarge on the ``other games are possible'' idea later; much in the general spirit
+  of Wittgenstein: consider multiple language games and what all of these suggest about
+  language in conjunction (perspicuous presentation)}
 

### 2017-04-20 book_chapter/mf_notes.tex

@@
   the vein of Wittgenstein's later views on logical analysis of language: yes, we can do most
-  of this, but it shows us little (supply proper references).
+  of this, but it shows us little.
 \item The way to give this answer is by a more careful analysis of rational or evolutionary
@@
   rationality or evolutionary optimization.
+\item The main reason for vagueness in language argued for here is this:
+  \begin{itemize}
+  \item what counts as optimal crisp language use crucially depends on certain
+    statistical properties of the world
+    \begin{itemize}
+    \item language must be optimal in different language games (see below)
+      \begin{enumerate}
+      \item sim-max games
+      \item reference games without displacement
+      \item reference games with displacement
+      \end{enumerate}
+    \item optimal behavior for all of these language games depends on specifics of the game
+      situation, most importantly the utilities and the priors about what happens in the real
+      world
+    \end{itemize}
+  \item neither rationality, nor evolutionary pressure can eradicate uncertainty about a
+    complex and dynamically changing world
+    \begin{itemize}
+    \item the main implicit premisse here is that it is not irrational not to be omniscient;
+      it's not irrational to have remaining uncertainty after any finite number of
+      observations; finiteness of life is not to blame on irrationality either
+    \end{itemize}
+  \item neither rationality nor evolutionary selection could possibly guarantee an optimal
+    crisp language
+  \end{itemize}
+\item Despite this uncertainty and the vagueness it gives rise to, there seems to be no need to
+  postulate an idealized language as an explanatory ingredient of either agents beliefs (about
+  language), nor about their actual linguistic practices.
 \end{itemize}
@@
   \end{itemize}
-
+\item Although speculative and provocative, we here suggest that everything of observational
+  relevance (agents linguistic behavior, possibly their beliefs) can be explained well (i)
+  without postulating an idealized non-vague language, and (ii) without abandoning the ideas of
+  idealized individual-level rationality or idealized long-run evolutionary selection.
 \end{itemize}
 
-\section*{A closer look at the agent's beliefs and behavior}
+\section*{A closer look at an agent's beliefs and behavior}
 
@@
   each other. They may also have beliefs about meaning, and their choices would depend on
-  these. Recent probabilistic models of pragmatic language use explicitly models agents
-  (beliefs) about each other's mental lexica
+  these. Recent probabilistic models of pragmatic use explicitly model agents' (beliefs) about
+  each other's mental lexica
   \citep{BergenLevy2012:Thats-what-she-,BergenLevy2014:Pragmatic-Reaso,PottsLassiter2016:Embedded-implic}.
@@
   \end{itemize}
-\item The picture of language users which we should adopt minimally adopt is this:
+\item The picture of language users which we should minimally adopt is this:
   \begin{itemize}
-  \item speaker behavior is a probabilistic function $P_S(\messg \mid \state, O^S_c, B^S_l)$
-    specifying a message choice probability for a meaning/state $\state$, a (possibly partial)
-    observation of the context $O_c$ and a belief $B_l$ about the language's meaning
+  \item speaker behavior is a probabilistic function $P_S(\messg \mid \state, B^S_c, B^S_l)$
+    specifying a message choice probability for a meaning/state $\state$, a belief after a
+    (possibly partial) observation of the context $B_c$ and a belief $B_l$ about the language's
+    meaning
   \item listener interpretation behavior is similarly a probabilistic function
-    $P_L(\state \mid \messg, O^L_c, B^L_l)$
+    $P_L(\state \mid \messg, B^L_c, B^L_l)$
     \begin{itemize}
@@
       almost always they will be different
-    \item the beliefs about contextual parameters $O_c$ which may influence choice behavior are
+    \item the beliefs about contextual parameters $B_c$ which may influence choice behavior are
       bountiful:
       \begin{itemize}
-      \item think: shape and parameters of the utility function in a sim-max game
+      \item think: shape and parameters of the utility function in a sim-max game; most
+        importantly: prior distribution of states (more on this below)
       \item imprecise perception of the referent set in a reference game
-      \item different beliefs about anything related to the interlocutor
       \item \dots
@@
       \subseteq \States$
+      \begin{itemize}
+      \item even if we do not want to have agents with ``fancy'' beliefs about meanings, a
+        definition of rational choice would require a belief about opponent behavior; in a
+        large population, this behavioral belief is good enough for our purposes: it will
+        likewise be subject to heavy uncertainty, being obtained from finite (partial)
+        observations about a dynamically changing world
+      \end{itemize}
+
     \end{itemize}
   \end{itemize}
+\end{itemize}
+
+\section*{Different language games and natural uncertainty}
+
+\begin{itemize}
 \item Nothing has been said so far how beliefs are acquired, whether these beliefs are
@@
   ``vagueness'' without abandoning rationality.
+\item We do this by noting that language games are plenty. We do not just ever engage in a
+  single sim-max games (with everlastingly fixed priors, utilities and co-players). 
+\item Let us distinguish at least three relevant kinds of games in which gradable properties
+  could be used for immediate communication:
+  \begin{enumerate}
+  \item sim-max games
+    \begin{itemize}
+    \item clear how they look like
+    \item a point which has so far not been stressed (at all? or enough?): the optimal strategy
+      is a function of the prior; what counts as ``tall'' depends on (beliefs about) the
+      distribution of states/degrees
+    \item if priors track, say the actual frequency with which someone is, say, 180cm tall that
+      real world frequency is impossible to know precisely
+    \item as a result even perfectly rational agents would not be certain about the optimal
+      strategy and will (given different life histories) have different strategies
+    \item being rational, they would also know that they don't know for certain what the prior
+      is and that there is no commonly known and temporally stable prior to begin with
+    \end{itemize}
+  \item reference games without displacement
+    \begin{itemize}
+    \item in each trial nature selects a variable number of referents (objects) as the
+      \emph{context}; she also selects a designated object for the speaker to refer to; speaker
+      and listener both observe the context, but only the speaker observes the designated
+      object; speaker and listener obtain a positive payoff if they coordinate on the
+      designated referent \citep{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}
+    \item crucially, these could be considered as an infinite collection of games, all
+      conditioned on the current context
+    \item optimal behavior for these games could be defined as a function of the actual
+      context; the problem for even a rational agent is to predict the co-player's behavior,
+      for in almost any new trial there will not be a preceding trial in which the exact same
+      context has been observed (adding perceptual noise and knowledge of perceptual
+      limitations would make this even more pressing)
+    \item as a result, even rational agents will be highly uncertain (and aware of their
+      uncertainty) as to how language is used in these situations
+    \item a rational response will take care of this uncertainty by preferring, if possible,
+      expressions whose use conditions are clearer over those that are less clear: borderline
+      case uses will be avoided (especially when the language is rich) and underdetermination
+      by use creates further uncertainty creates further underdetermination etc.
+    \end{itemize}
+  \item reference games with displacement
+    \begin{itemize}
+    \item similar to the above; these games have been considered (informally) by
+      \citet{Lipman2009:Why-is-Language} and \citet{Deemter2009:Utility-and-Lan}
+    \item the sender observes a referent and describes it for the listener; but unlike before,
+      the listener will have to pick out the designated referent in a future context (not known
+      by the sender): e.g., picking up a person from the airport, identifying a robber or (from
+      the main paper draft) picking up a book from the next room based on the color
+    \item as the sender does not know the future context in which the designated referent will
+      occur (the other objects co-present in the environment), the sender's optimal choice
+      depends crucially on a conjecture about the likelihood of seeing particular contexts
+    \item again the sender must needs know the statistical properties of a changing world for
+      genuinely optimal crisp language use
+    \end{itemize}
+  \end{enumerate}
+\end{itemize}
 
-\end{itemize}
+\printbibliography[heading=bibintoc]
 
@@
 
-Sources of vagueness:
+\section*{Snippets}
 
-- uncertainty about threshold:
-   - learning from finite examples
-   - rich language makes decisions about borderline cases obsolete; enhances problem of learning data scarcity
-   - imperfect memory
-- uncertainty about QUD, utility, interlocutor's beliefs
+\begin{itemize}
+\item - uncertainty about threshold: - learning from finite examples - rich language makes
+  decisions about borderline cases obsolete; enhances problem of learning data scarcity -
+  imperfect memory - uncertainty about QUD, utility, interlocutor's beliefs
 
-=> why not just have a prior and integrate and have a fixed threshold
+\item   => why not just have a prior and integrate and have a fixed threshold
 
-1. Priors are non-introspective; sampling based assessment
+\item   1. Priors are non-introspective; sampling based assessment
 
-2. Choice is probabilistic -> utilities are noisy
+\item   2. Choice is probabilistic -> utilities are noisy
 
-Why not adopt a fixed threshold? Why is an agent with uncertainty in the lexicon better off? => revision cost? surprisal cost?
+\item   Why not adopt a fixed threshold? Why is an agent with uncertainty in the lexicon better off?
+  => revision cost? surprisal cost?
 
-If a single agent's semantic representation is the outcome of rational learning and possibly an internal optimization towards efficient communication.
+\item   If a single agent's semantic representation is the outcome of rational learning and possibly
+  an internal optimization towards efficient communication.
 
-Finiteness of life is not irrational.
+\item   Finiteness of life is not irrational.
 
-Convince by number of examples of natural sources for vagueness.
+\item   Convince by number of examples of natural sources for vagueness.
 
-Any precise threshold must be a function of priors about properties, QUDs, utilities, and interlocutor beliefs.
-But these are not stable over time.
-Vagueness might be optimal to compensate mistakes in a changing environment.
-Assumption here: revision cost high when too low prob on observable events.
-(van Deemter example: DECENT behavior)
+\item   Any precise threshold must be a function of priors about properties, QUDs, utilities, and
+  interlocutor beliefs. But these are not stable over time. Vagueness might be optimal to
+  compensate mistakes in a changing environment. Assumption here: revision cost high when too
+  low prob on observable events. (van Deemter example: DECENT behavior)
 
-If speakers' uncertain beliefs come from natural probabilistic evidence, they will not be sharp step-functions.
-If we measure success of communication by KL-divergence or similar, the update of a literal listener with a vague threshold would give higher utility.
-But what about pragmatic listeners?
+\item   If speakers' uncertain beliefs come from natural probabilistic evidence, they will not be
+  sharp step-functions. If we measure success of communication by KL-divergence or similar, the
+  update of a literal listener with a vague threshold would give higher utility. But what about
+  pragmatic listeners?
 
-Suppose agents play repeated reference games with different naturals kinds.
-Each kind has a mean and variance for each of several open and closed scale features.
-Learners entertain different hypotheses about lexical meaning (McNally?) They would like to assign a fixed degree to an expression because that is economical.
-They can do that for closed scale natural kinds unless their mean and variance is such that exemplars are frequently away from the end points.
-Think wine glasses!
-They cannot do this for open scales.
-They must consider a more difficult hypothesis, namely to use the prior after all.
+\item   Suppose agents play repeated reference games with different naturals kinds. Each kind has a
+  mean and variance for each of several open and closed scale features. Learners entertain
+  different hypotheses about lexical meaning (McNally?) They would like to assign a fixed
+  degree to an expression because that is economical. They can do that for closed scale natural
+  kinds unless their mean and variance is such that exemplars are frequently away from the end
+  points. Think wine glasses! They cannot do this for open scales. They must consider a more
+  difficult hypothesis, namely to use the prior after all.
 
-Good point by van Deemter: reliable metrics may be absent; the goal of communication may be expression of sentiment, and the real value may be assessed in a goal-oriented way (BIG when playing soccer not equal to big when painting).
+\item   Good point by van Deemter: reliable metrics may be absent; the goal of communication may be
+  expression of sentiment, and the real value may be assessed in a goal-oriented way (BIG when
+  playing soccer not equal to big when painting).
 
-Languages supervene on agents' beliefs about lexica. Lexica are uncertain. They are still useful. Lexical uncertainty might even be rational in the light of belief revision in changing environments.
+\item   Languages supervene on agents' beliefs about lexica. Lexica are uncertain. They are still
+  useful. Lexical uncertainty might even be rational in the light of belief revision in
+  changing environments.
 
-Languages are optimized for communication and social interaction. Not necessarily for logical consistency.
+\item   Languages are optimized for communication and social interaction. Not necessarily for logical
+  consistency.
 
-Plurality of language games forces uncertainty. Wittgenstein.
+\item   Plurality of language games forces uncertainty. Wittgenstein.
 
-So then logic breaks, so what? The possibility of visual illusions does not prove the visual system ill adapted or suboptimal.
-
-\printbibliography[heading=bibintoc]
+\item   So then logic breaks, so what? The possibility of visual illusions does not prove the visual
+  system ill adapted or suboptimal.
+\end{itemize}
 

### 2017-04-20 book_chapter/paper.tex

--- 2017-04-11 book_chapter/paper.tex
@@
 Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
-Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.
+Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.\mf{explain sender and receiver
+payoffs}
 These so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
@@
 Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
-Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be.
+Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-}.
 The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
@@
 While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
-We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players\todo{give examples and references?}.
-This range of options goes hand in hand with assumptions of rationality, from nothing more than a basic instrumental rationality that aims for survival in a biologically-inspired setting, to a certain degree of strategic rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
+We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{Muhlenbernd2011:Learning-with-N}..
+This range of options goes hand in hand with assumptions of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
 Each of these can be utilized depending on the problem that one is interested in characterizing.
-Thus, although highly strategic rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where it can play a role.
+Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.
 
-% The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
-% Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?}.
-% In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
+The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how. \mf{explain/introduce equilibrium and/or ESS?}
+Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?
+\mf{too much detail,
+    rather not!}}.
+In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
 %; which one of the possible two will depend on their relative proportions in the original population.
-% More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
-% Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
-% Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
-less examples?}.
-% More recent overviews are given by \textcites{huttegger_how_2014} and \textcite{huttegger_dynamics_2014}.
+More complex signaling problems may have different evolutionary outcomes, sometimes unexpected ones.
+Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
+Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
+less examples?
+\mf{good as it is}}.
+More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.
 

### 2017-04-25 book_chapter/paper.tex
Added summary of Lawry and James. Added section with the argument for rationality of strategies with uncertainty.

@@
 
-\title{Natural sources of vagueness and their implications}
+\title{Natural sources of uncertainty and their implications}
 \author{Jos\'e Pedro Correia \and Michael Franke}
@@
 % Each thus prefers to go to one of the places only if the other also decides to go to that particular place.
-
 % In order to extend his notion of convention to \emph{linguistic} exchanges
@@
 The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how. \mf{explain/introduce equilibrium and/or ESS?}
-Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't\todo{mention something about phase portraits?
-\mf{too much detail,
-    rather not!}}.
+Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't.
 In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
@@
 Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
-Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few\todo{more examples?
-less examples?
-\mf{good as it is}}.
+Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few.
 More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.
@@
 
-\section{Signaling models of vagueness}
-\label{sec:vague-signaling}
+\section{Vagueness in sim-max games}
+\label{sec:sim-max-vagueness}
 
@@
 However, given the level of abstraction of the description so far, none of this is necessarily implied by the model.
+Our question is thus what additional modifications to sim-max games are sufficient for the optimal languages to be more like Figure~\ref{fig:example_stratsB}, rather than like Figure~\ref{fig:example_stratsA}.
+%Let's look at some proposals in the literature.
 
-Our question is thus what additional modifications to sim-max games are sufficient for the optimal languages to be more like Figure~\ref{fig:example_stratsB}, rather than like Figure~\ref{fig:example_stratsA}.
-Let's look at some proposals in the literature.
-
-\subsection{Bounded rationality}
+% \subsection{Bounded rationality}
 \textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can arise as a consequence of limitations in the cognitive capacities of rational agents.
@@
 
-\subsection{Generalized reinforcement learning}
+% \subsection{Generalized reinforcement learning}
 Cailin O'Connor~\parencite*{oconnor_evolution_2014} proposes a way in which vagueness could be expected to evolve as a side-effect of a particular type of learning process.
@@
 
-\subsection{Noisy perception}
-\parencite{correia_bivalent_2013,franke_vagueness_2017}
+% \subsection{Noisy perception}
+\todo[inline]{talk about \textcite{correia_bivalent_2013,franke_vagueness_2017}}
 
-\subsection{Preference misalignment}
-\todo[inline]{Test CS~\parencite{crawford_strategic_1982} results in our model by using bias in utility function.
+% \subsection{Aggregation from multiple senders}
+\textcite{lawry_vagueness_2017} explore the potential benefits of vagueness in variations of a sim-max game with multiple senders and one receiver.
+Although the models considered are not explicitly identified as sim-max games, they are equivalent in the relevant aspects: they are signaling games where senders have private knowledge of a given state $x$, each sends a message to the receiver, which then selects an estimate $y$ of the original state; although the authors do not define a utility function, their analysis evaluates the models in terms of the expected error between the estimate and the original state, defined as $(x-y)^2$; better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse; the game could then be seen as a sim-max game with multiple senders and negative squared difference as similarity function.
+Unlike the models discussed so far that introduce minimal constraints to a dynamic and study the strategies that emerge out of an evolutionary process, these authors' approach is mostly static and analytical.
+They consider several possible sender strategies with uncertainty and without and calculate if and when sender and receiver achieve better transmission accuracy, \emph{i.e.}~smaller expected error between original state and receiver estimate.
+Senders without uncertainty have a strategy that boils down to what is represented in Figure~\ref{fig:example_stratsA}: they split the state space according to a fixed thresholds, switching from using one message to another abruptly.
+The model for the other type of sender is similar, but assumes that there is uncertainty regarding the value of the thresholds, representing them as uniformly distributed random variables.
+What results from this in terms of sender strategy is also a smooth transition between the use of two messages, but unlike what we see in Figure~\ref{fig:example_stratsB}, it is a linear transition\footnote{This is certainly connected with the choice of modeling the uncertain threshold as uniformly distributed.
+We speculate that another choice of distribution, like for example a normal distribution, would yield something similar.
+What is important to characterize a vague use of messages, however, is that the transition is smooth and monotonous, not its particular shape.}.
+The receiver strategy is not inherently vague for either type of sender, it merely aggregates the messages received to provide an estimate of the original state using a simple arithmetic mean.
+
+The authors analyze a number of variations of this model.
+For a model with one sender, results are consistent with Lipman's~\parencite{lipman_why_2009} in that strategies without uncertainty lead to more accurate communication.
+For models with two messages and more than one sender, the expected error is a strictly decreasing function of the number of senders, and above a certain number of senders, strategies with uncertainty actually achieve lower expected error than sharp ones.
+This minimal number of senders from which vagueness surpasses sharpness is a decreasing function of the number of messages, \emph{i.e.}~whereas at least 8 senders are needed if the number of messages available is 2, this number lowers to 5 for 3 messages, 4 for and 5 messages, and 3 for more than 6 messages.
+If random errors are introduced between the choice of the sender and the message that reaches the receiver, increasing the number of senders can compensate for the noise, to a certain extent.
+Results are mixed when dropping the assumption of uniform priors for the state space.
+
+These results are interesting since they reveal scenarios where vague strategies have an advantage over strict ones, even in a setting of higher rationality.
+The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the idealist\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
+In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (or better memory) would lead to more precision.
+\citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{ellison_learning_1993}\todo{is this the canonical reference?}.
+A combination of both would hypothetically lead to a better outcome in the scenario of \citeauthor{franke_vagueness_2017}.
+However, \citeauthor{lawry_vagueness_2017} give us something that holds even if the idealist is unwilling to acknowledge the constraints that permeate our finite existence.
+Whether the scenario of multiple senders is representative of the language games where vagueness comes about, and whether the strategies proposed by the authors would be the ones to evolve\footnote{Their analysis is a static one.
+Despite the advantage of the proposed vague strategies against strict ones in certain settings, one cannot know whether other strategies of higher accuracy would not be possible in those same settings without studying some dynamics in an evolutionary context.}, is something that is, however, open for debate.
+
+\section{Vagueness in referential games}
+\label{sec:referential-vagueness}
+
+% \section{Vagueness in related models}
+% \label{sec:other-vagueness}
+%
+% \subsection{Preference misalignment}
+% \todo[inline]{Test CS~\parencite{crawford_strategic_1982} results in our model by using bias in utility function.
 READ Förster, M., and Riedel, F.
@@
 shifted according to bias, but still crisp}
+%
+% \subsection{Noisy communication}
+% \todo[inline]{Test if noise in communication channel as modeled by Blume~\parencite{blume_intentional_2014} induces vague strategies as well}
 
-\subsection{Noisy communication}
-\todo[inline]{Test if noise in communication channel as modeled by Blume~\parencite{blume_intentional_2014} induces vague strategies as well}
+\section{How could language not be vague?}
+\label{sec:finite-experience}
+The research surveyed here suggests various mechanisms that can lead to vague use of messages in certain signaling games, such as cognitive limitations like a finite memory or the inability to always calculate the absolute optimum choice to make, learning mechanisms that compensate for the finiteness of experience through generalization based on similarity, imperfect senses that allow for the possibility of stimulae to be confused, etc\todo{fit Lawry and James plus referential games stuff}.
+In order to account for vagueness in natural language, they all need to peel away from an idealized model of agents, language, and the world, and bring some assumptions down to earth.
+They point to some ways, but by all means not all, in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world.
+Consider the following:
+\begin{enumerate}
+\item Our existence is temporally finite; language does not have an infinite amount of time to evolve or take a long time to be learned. The faster a language can start to be useful, the better;
+\item Language learning through experience has to rely on a limited number of observations.
+Not only is the state space typically much larger than what one can survey in sufficient time, but it is also potentially infinite and constantly changing;
+\item A corollary of the former is that there will always be heterogeneity in a population of language learners, at least of prior experience since each agent will have relied on a different set of observations.
+Furthermore, this information is not directly or fully accessible to others;
+\item Given that an agent is almost always learning and using language in such a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice;
+\item We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another;
+\item These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time.
+\end{enumerate}
+It is almost inevitable that uncertainty gets introduced at many levels of language learning and language use.
+In the context of a signaling game, we can expect uncertainty about priors, observations, utilities, behavior of other players, and more.
+This is true both from the perspective of the modeler, as from the perspective of the agents involved.
+Arguments, of the kind presented by Lipman~\parencite*{lipman_why_2009}, that a vague language is suboptimal compared to a crisp one depend on ignoring all of these aspects of the world, language, and experience.
+
+We learned from \citeauthor{oconnor_evolution_2014} and \citeauthor{franke_vagueness_2017} that vague languages are quicker to converge and more adaptable, which is valuable given the finite and dynamic character of our experience (point 1).
+\citeauthor{oconnor_evolving_2015} also showed how generalization, an invaluable feature of any procedure for learning from a limited number of observations (point 2), leads to vagueness.
+We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism thats leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
+\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).\todo{add something for 5 and 6, maybe referential games}
+Trying to answer the question of ``Why is language vague?'', is important to force us to recognize those aspects of language and the world that are easily forgotten when idealizing the subject matter.
+The finiteness of human experience, the dynamic nature of the world, and the communitarian context of language use and development, these are things we need to keep in mind lest we end up with inadequate models of language and meaning.
+% Wittgenstein put it better:
+% \begin{quote}
+% The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
+(For the crystalline purity of logic was, of course, not something I had discovered: it was a requirement.) The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
+– We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
+We want to walk: so we need friction.
+Back to the rough ground!%
+% ~\parencite*[\S 107]{wittgenstein_philosophical_1953}
+% \end{quote}
+
+What do these observations tell us about rationality?
+An agent with minimal instrumental rationality, merely adapting its behavior based on reinforcement, will be directly conditioned by the limitations imposed by its environment.
+Therefore, given all the natural sources of uncertainty discussed so far, even the most effective of such agents (or population thereof) would develop a signaling behavior that inevitably makes vague use of messages.
+The agent would still be driving for the highest possible payoff it can achieve, by definition, and the presence of vagueness could not be attributed to any lack of rationality.
+For a more complex agent, the story is perhaps more nuanced.
+Consider an agent that has awareness that something like a game is being played, some estimate of the potential payoff involved, ability to gauge other players' expected behaviors, and the capacity to think strategically.
+Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
+If he is \emph{not} aware of the natural sources of uncertainty discussed so far, one cannot charge him with irrationality since by adopting a crisp strategy he is choosing the optimal strategy for his assumptions.
+However, if he \emph{is} aware, then it would be less rational to use a crisp strategy than to use a strategy that incorporates at least some source of uncertainty.
+A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
+
+\todo[inline]{precision is possible though, but not absolute precision, and vagueness comes in degrees, figuring out how much vagueness to use is the real challenge, and this needs to be studied}
 

### 2017-04-27 book_chapter/paper.tex
minor edits sections 1 and 2

@@
 
-\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false,isbn=false]{biblatex}
+\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
@@
 How widespread is the problem?
-It is easy to come up with more examples of predicates based on more finely grained properties---as `tall' is intuitively based on height---for which constructing a sorites paradox would be easy.
+It is easy to find more examples of predicates based on more finely grained properties---as `tall' is intuitively based on height---for which constructing a sorites paradox would be easy.
 Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
@@
 Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
-Nobody ever stopped using the word `tall` after being confronted with a sorites series to deconstruct it.
+Nobody ever stopped using the word `tall` after being confronted with a sorites series.
 Why should we develop theories of meaning that are impervious to the paradox?
@@
 The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also consequently need to be abandoned down the line.
-One notion that seems to quickly be in peril in that of rationality.
+One notion that seems to quickly be in peril is that of rationality.
 In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
@@
 Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
-Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-}.
+Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}.
 The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
@@
 While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
-We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{Muhlenbernd2011:Learning-with-N}..
+We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
 This range of options goes hand in hand with assumptions of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
@@
 
-The simplest type of game to study in this scenario is one where the state space and the action space are the same.
+Although we will look at more complex cases below, the simplest type of game to study in this scenario is one where the state space and the action space are the same. 
 We can imagine this as a game of guessing the state of affairs: the sender has knowledge of it, sends a message to the receiver, who in turn has to guess it; their payoff, as discussed above, is proportional to how close he got.
-Other games are possible, depending on what situation of use we want to model.
-For example, imagine I have two editions of a book, one of them is red the other one is green.
-Wanting to read that book, I ask you to fetch the red book for me.
-If you bring me a different red book, whose hue is very similar to my prototypical notion of red, I would still be disappointed.
-If, however, you bring me my other edition of that book, despite it being far from red, I would be satisfied.
-This example is just to illustrate that there is more to the use of vague terms in natural language than what the setup we just described can cover.
-But, we believe, the setup captures one possible scenario where vagueness is relevant that
-would be difficult to capture in the classic formulations of signaling discussed before. \mf{I
-  suggest to enlarge on the ``other games are possible'' idea later; much in the general spirit
-  of Wittgenstein: consider multiple language games and what all of these suggest about
-  language in conjunction (perspicuous presentation)}
-
 These games, called similarity-maximization or sim-max games for short, were first introduced by Gerhard J{\"a}ger and Robert van~Rooij~\parencite{jager_language_2007,Jager2007} and further studied by~\textcite{jager_voronoi_2011}.
@@
 Roughly, these are situations where the sender uses messages in a way that can be seen as partitioning the state space into convex regions and the receiver responds with the central element of those regions.
-For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}.
+For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280.\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}
 Given two possible messages, the optimal sender strategy is to use one for all values from 40 up to 160, and the other for all values from 160 up to 280.
@@
 In order to overcome this limitation, they make another proposal using the notion of a quantal response equilibrium (QRE).
-The idea, inspired by experimental psychology, is to model the choice of best response as stochastic rather than deterministic, deviating from the optimal according to an exponentially decaying distribution (where small deviations are very likely, and large ones very rare).
-The approach is not explanatory but descriptive: it is assumed that agents are not always able to make the absolute best choice, but the mechanism behind this limitation is left open to interpretation.
-The degree to which agents deviate from the optimal behavior can be characterized by a parameter.
-The authors find that for low values of this parameter, only babbling equilibria (where sender and receiver simply randomize, respectively, message and response uniformly) where no communication actually occurs.
+The idea, inspired by experimental psychology, is to model the choice of best response as stochastic rather than deterministic.\footnote{Probabilistic choice rules are also the source of vagueness in recent accounts by \textcite{LassiterGoodman2015:Adjectival-vagu} and \textcite{QingFranke2014:Gradable-Adject}.}
+A prominent explanation for such soft-max or quantal choice behavior is that agents make small mistakes in the calculation of their expected utilities \parencite{Train2009:Discrete-Choice}. 
+They still choose rationally the option with the highest expected utility, but each assessment of the expected utilities is noise perturbed.
+This, in turn, may actually be optimal behavior if the calculation of expected utilities relies on assessing stochastic uncertainty. 
+Choice based on a few samples from a distribution can be optimal if taking more samples or other means of better approximating degrees of uncertainty is resource costly \parencite[\emph{e.g.}][]{VulGoodman2014:One-and-Done-Op,SanbornChater2016:Bayesian-Brains}.
+The degree to which agents tremble in the calculation of expected utilities and therefore deviate from the optimal behavior can be characterized by a parameter.
+\textcite{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria (where sender and receiver simply randomize, respectively, message and response uniformly) where no communication actually occurs.
 Above a certain value of the parameter, other equilibria of the kind described in the beginning of this section arise, where agents communicate successfully, though not perfectly, using fuzzy strategies similar to the ones depicted in Figure~\ref{fig:example_stratsB}.
+However, it is not clear that soft-max choices model the right stochastic trembles in decision making as they would arise under natural sources of uncertainty about the context \parencite[see][]{franke_vagueness_2017}.
 

### 2017-04-27 book_chapter/paper.tex
first lines for reference games

@@
 
+Linguistic interaction, as captured by a sim-max game, has a rather narrow scope.
+There is a fixed prior and fixed utility function and a fixed similarity measure between states.
+If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t$.
+Here is a natural example of  a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue $t$; the closer Bubu's color choice matches Kiki's idea, the better.
+We could say that Kiki wants to describe the color for Bubu to pick out.
+It's a natural thing to want to do with language.
+But a lot of communication with vague and contiguously variable properties is rather different.
+In the general spirit of Wittgenstein's language games, we should look at some more cases of language use side by side.
+Let us therefore consider a continuum of cases of referential communication.
+
+A \emph{reference game} is a model of a situation in which the speaker has a particular object in mind and uses a description in order for the receiver to pick this object and no other object.
+One crucial difference to descriptive communication and sim-max games is that close will not be close enough: if Kiki needs a long nail, a nail a little longer or shorter may be good enough; but if Bubu needs the long, slightly bend nail with the flat top (which she previously disassembled from her art installation), no other nail will do.
+
+Interestingly, referential communication can take place under a plethora of different epistemic conditions.
+A lot of modeling work considers \emph{reference games without displacement} where sender and receiver communicate about a context, i.e.
+a set of objects, which is immediately accessible in their direct perceptual environment \parencite{BaronchelliGong2010:Modeling-the-Em,Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.\footnote{add more references later!} Agents may make mistakes about observing exactly how blue or long etc.~a particular object is, but there is no further fundamental uncertainty about the context.
+This, of course, need not be so.
+\textcite{lipman_why_2009} and \textcite{Deemter2009:Utility-and-Lan}, for example, consider \emph{reference games with displacement}.
+Kiki asks Bubu to fetch the red book from the library.
+No other book than the one Kiki needs will serve.
+Yet when Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki.
+Yet other situation are imaginable.
+In \emph{reference games with partial uncertainty}, for instance, interlocutors may know that they have shared and immediate access only to a subset of the objects in the context, but be uncertain about whether the co-player sees any other objects.
+This is the set-up for the classic director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
+
 % \section{Vagueness in related models}

### 2017-04-28 book_chapter/paper.tex
readjusted to biber engine
Added summary of Franke and Correia

@@
 
-\usepackage[natbib=true,style=authoryear-comp,backend=bibtex,doi=false,url=false,isbn=false]{biblatex}
+\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
@@
 % \subsection{Noisy perception}
-\todo[inline]{talk about \textcite{correia_bivalent_2013,franke_vagueness_2017}}
+\textcite{franke_vagueness_2017} study a variant of the replicator dynamics that is motivated by perceptual limitations.
+Assuming that agents do not have perfect perception, there will always be a possibility that senders confuse states and receivers mix up responses.
+Furthermore, it seems reasonable to assume that this state confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is they will be mistaken for each other.
+Incorporating thiese considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of that dynamic that also induces vague signal use of the kind we expect here.
+The consequence is very similar to that of the GRL model discussed above, in that the way the behavior for a given state is updated takes into account the behavior of similar states, proportionally to their similarity.
+Given the known relation between reinforcement learning and the replicator dynamics~\parencite{Beggs2005}, it is actually quite plausible that the two are tightly related, but this would need to be formally demonstrated.
+The account is, however, interpretable at a lower level of rationality since the replicator equation is suitable to represent biological processes of natural selection.
+
+The motivation underlying this model of vague signaling is still one of inevitability.
+A vague strategy is still not claimed to have higher expected utility than a crisp one.
+However, the authors observe an effect similar to that pointed out by~\citeauthor{oconnor_evolution_2014}: signaling converges faster in scenarios where there is some degree of state confusability.
+Furthermore, they observe one additional potentially beneficial property.
+By running several rounds of simulation for each parameter set, they can measure how close resulting strategies are to each other within each group, and how they would fair playing against each other by calculating a within group expected utility.
+The results show that the within group distances between strategies is smaller with growing confusability, and the within group expected utility is actually higher for strategies evolving under a certain degree of state confusion.
+Thus, some amount of uncertainty seems to promote more homogeneous populations of signalers that are better at achieving cooperation within a group of such signalers
 

### 2017-04-28 book_chapter/paper.tex
reference game section edits

@@
 
-\section{Vagueness in referential games}
+\section{Vagueness in multiple reference games}
 \label{sec:referential-vagueness}
@@
 Linguistic interaction, as captured by a sim-max game, has a rather narrow scope.
-There is a fixed prior and fixed utility function and a fixed similarity measure between states.
+There is a fixed prior, a fixed utility function and a fixed similarity measure between states.
 If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t$.
 Here is a natural example of  a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue $t$; the closer Bubu's color choice matches Kiki's idea, the better.
-We could say that Kiki wants to describe the color for Bubu to pick out.
+We could say that Kiki wants to \emph{describe} the color for Bubu to pick out.
 It's a natural thing to want to do with language.
 But a lot of communication with vague and contiguously variable properties is rather different.
-In the general spirit of Wittgenstein's language games, we should look at some more cases of language use side by side.
+In the general spirit of Wittgenstein's method of arranging multiple language games side by side for a more encompassing picture, we should look at some more cases of language use as well.
 Let us therefore consider a continuum of cases of referential communication.
@@
 Yet when Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki.
-Yet other situation are imaginable.
+Yet other situations are imaginable.
 In \emph{reference games with partial uncertainty}, for instance, interlocutors may know that they have shared and immediate access only to a subset of the objects in the context, but be uncertain about whether the co-player sees any other objects.
 This is the set-up for the classic director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
+
+What would a rational agent do in repeated reference games?
+Take a universe in which agents play repeated reference games without displacement.
+A single game has sender and receiver both flawlessly perceive a context of $k$ objects (with $k$ possibly randomly drawn for each encounter); each object has its individual features, e.g., drawn from a rich multidimensional feature space of continuously variable properties \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.
+The sender chooses a message $m$ for the target object $t$, but she can condition her choice on the current context $c$.
+Likewise, the receiver chooses a referent $t'$ from the available objects in $c$, again conditional on $c$.
+But if agents play many such reference games with different contexts $c$ and $c'$, and if strategic rationality means to signal and interpret so as to coordinate on the correct referent, strategic rationality of local choices in $c$ and $c'$ is not enough to guarantee that the agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
+In other words, local strategic rationality in a fixed context does not entail a globally consistent behavior across multiple contexts.
+The situation only gets worse if we want to have agents use the same expressions in a consistent way also across different types of reference games, with possibly different epistemic conditions.
+
+The picture sketched in the last paragraph neglects a rational agent's ability to learn from previous experience and to generalize from prior experience.
+Which strategies for generalization are best depends on the stochastic properties of the environment in which learning and decision making takes place.
+The problem for language is that it what works best for on agent depends on what the other agent does.
+Communication problems are games in the formal sense of delivering payoffs only as a function of what all players do.
+Consequently, the problem of how best to adapt behavior from past experience in language games also depends on how other agents categorize, generalize and behave.
+Nothing gets off the ground with hyper-rationality alone.
+There needs to be learning, evolution and circumstances.
+As with sim-max games above, biases and other quirks of general cognition are likely only helpful in solving this massive coordination problem.
+
+\begin{itemize}
+\item uncertainty
+\end{itemize}
+
+What about meaning and vagueness in a universe of multiple language games and heavy uncertainty?
+We could look for meaning at several levels.
+Firstly, as before, there is the actual behavior of a single agent in each actualized language game.
+Like with sim-max games, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
+Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used, conditional on a given context.
+Either one of these two levels, the agent's behavior or linguistic generalizations, could be the foundation for a notion of emerging abstract meaning as constituted by the practices of a population at a particular time.
+The question is: for which of these three options of situating meaning would we like to have a theory of so that being vague would be disastrous for that theory?
 

### 2017-04-30 book_chapter/paper.tex
biber rollback
reference games section further edits

@@
 
-Linguistic interaction, as captured by a sim-max game, has a rather narrow scope.
-There is a fixed prior, a fixed utility function and a fixed similarity measure between states.
+Linguistic interaction, as captured by a sim-max game, has a rather narrow scope. There is a fixed prior, a fixed utility function and a fixed similarity measure between states. 
 If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t$. 
 Here is a natural example of a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue $t$; the closer Bubu's color choice matches Kiki's idea, the better.
-We could say that Kiki wants to \emph{describe} the color for Bubu to pick out.
+Here Kiki wants to \emph{describe} the color for Bubu to pick out or realize from the whole set of possibilities.
 It's a natural thing to want to do with language. 
@@
 A lot of modeling work considers \emph{reference games without displacement} where sender and receiver communicate about a context, i.e.
-a set of objects, which is immediately accessible in their direct perceptual environment \parencite{BaronchelliGong2010:Modeling-the-Em,Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.\footnote{add more references later!} Agents may make mistakes about observing exactly how blue or long etc.~a particular object is, but there is no further fundamental uncertainty about the context.
-This, of course, need not be so.
-\textcite{lipman_why_2009} and \textcite{Deemter2009:Utility-and-Lan}, for example, consider \emph{reference games with displacement}.
+a set of objects, which is immediately accessible in their direct perceptual environment \parencite[\emph{e.g.}][]{BaronchelliGong2010:Modeling-the-Em,Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.
+Agents may make mistakes about observing exactly how blue or long etc.~a particular object is, but there is no further fundamental uncertainty about the context. 
+This, of course, need not be so. \textcite{lipman_why_2009} and \textcite{Deemter2009:Utility-and-Lan}, for example, consider \emph{reference games with displacement}. 
 Kiki asks Bubu to fetch the red book from the library. 
 No other book than the one Kiki needs will serve. 
-Yet when Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki.
+When Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki at the time of speaking. 
 Yet other situations are imaginable. 
@@
 Likewise, the receiver chooses a referent $t'$ from the available objects in $c$, again conditional on $c$. 
-But if agents play many such reference games with different contexts $c$ and $c'$, and if strategic rationality means to signal and interpret so as to coordinate on the correct referent, strategic rationality of local choices in $c$ and $c'$ is not enough to guarantee that the agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
-In other words, local strategic rationality in a fixed context does not entail a globally consistent behavior across multiple contexts.
+But if agents play many such reference games with different contexts $c$ and $c'$, and if strategic rationality means to signal and interpret so as to coordinate on the correct referent, strategic rationality of local choices in $c$ and $c'$ is not enough to guarantee that agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
+An agent could use a certain signal to denote, say, the tallest object in every context with between five and twelve objects or when the number of objects is a prime number, but use the same signal for the darkest object in all other contexts.
+This could even be rational if most agents do it this way.
+It could evolve as a convention under selection for communicative success as well.
+In other words, neither local strategic rationality in a fixed context nor evolutionary adaptation to each single context entails a natural, globally consistent behavior across multiple contexts. 
 The situation only gets worse if we want to have agents use the same expressions in a consistent way also across different types of reference games, with possibly different epistemic conditions.
 
-The picture sketched in the last paragraph neglects a rational agent's ability to learn from previous experience and to generalize from prior experience.
+The picture sketched in the last paragraph neglects one crucial thing: learning from observation. 
+For rationality, we need to consider a rational agent's ability to learn from previous experience and to generalize from prior experience. 
+For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice (\mf{insert refs for iterated learning}).
+But even if every language learner in history had been perfectly rational, it is unlikely that anyone ever succeeded in reducing all uncertainty about language use in context and say with confidence: this is the threshold that truly separates heaps from non-heaps (\mf{ref / footnote Williamson?}).
+
  Which strategies for generalization are best depends on the stochastic properties of the environment in which learning and decision making takes place.
@@
 As with sim-max games above, biases and other quirks of general cognition are likely only helpful in solving this massive coordination problem.
+Maybe our world is a place where particular learning biases are needed for any regular and shared communicative practice to get off the ground.
 

### 2017-04-30 book_chapter/paper.tex
finished first write up of ref games section

@@
 
-The picture sketched in the last paragraph neglects one crucial thing: learning from observation. 
-For rationality, we need to consider a rational agent's ability to learn from previous experience and to generalize from prior experience. 
-For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice (\mf{insert refs for iterated learning}).
-But even if every language learner in history had been perfectly rational, it is unlikely that anyone ever succeeded in reducing all uncertainty about language use in context and say with confidence: this is the threshold that truly separates heaps from non-heaps (\mf{ref / footnote Williamson?}).
-
- Which strategies for generalization are best depends on the stochastic properties of the environment in which learning and decision making takes place.
-The problem for language is that it what works best for on agent depends on what the other agent does.
-Communication problems are games in the formal sense of delivering payoffs only as a function of what all players do.
-Consequently, the problem of how best to adapt behavior from past experience in language games also depends on how other agents categorize, generalize and behave.
-Nothing gets off the ground with hyper-rationality alone.
-There needs to be learning, evolution and circumstances.
-As with sim-max games above, biases and other quirks of general cognition are likely only helpful in solving this massive coordination problem.
-Maybe our world is a place where particular learning biases are needed for any regular and shared communicative practice to get off the ground.
-
-\begin{itemize}
-\item uncertainty
-\end{itemize}
+The picture sketched in the last paragraph neglects learning from observation and generalization across contexts. 
+For rationality, we need to consider a rational agent's ability to learn from previous experience and to generalize from  the past to the future. 
+For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
+But even if every language learner in history had been perfectly rational, it is extremely unlikely that anyone ever succeeded in reducing all uncertainty about language use and say with confidence: this is the threshold that truly separates heaps from non-heaps no matter what.
+For one, there may simply not be enough time in the lifespan of a rational learner to observe enough instances of language use for certainty about crisp denotations. 
+For another, what needs to be learned likely changes over time. 
+The optimal use for gradable properties hinges crucially on the prior over states \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie,QingFranke2014:Gradable-Adject,LassiterGoodman2015:Adjectival-vagu}.
+If the prior captures real world attributes, such as the probability of an individual's body height, these attributes will not only be hard to know for certain, but they will also vary over time. 
+They may also depend on location: do we use our language exclusively in this village or do we travel beyond? 
+If so, the priors that determine the goodness of a convention for one agent need not be the same as those for another. 
+Going beyond the priors for property degrees, the probability of seeing a particular language game, e.g.~a particular context $c$ in some kind of reference game, is also crucial for determining globally efficient linguistic behavior.
+But here it is even more inconceivable how a rational learner could acquire a fairly confident estimate over how often a particular language game will be played. 
+The list of sources of natural and seemingly insurmountable uncertainty about the parameters necessary to determine locally or globally optimal strategies is much longer than hinted at here. 
+Still, it should be clear enough that a finite existence, a sparse learning input and the inherent fuzziness of the linguistic context itself (its temporal, local and social variability) is not a faul of rationality or the mechanism of selection of successful linguistic practices.
 
@@
 Firstly, as before, there is the actual behavior of a single agent in each actualized language game. 
-Like with sim-max games, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
+As mentioned above, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
 Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used, conditional on a given context.
+Notice that a single agent's rational beliefs about linguistic practices or linguistics meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
 Either one of these two levels, the agent's behavior or linguistic generalizations, could be the foundation for a notion of emerging abstract meaning as constituted by the practices of a population at a particular time.
-The question is: for which of these three options of situating meaning would we like to have a theory of so that being vague would be disastrous for that theory?
+If both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp? 
+Maybe there is a reason.
+The main question to be answered is even more general: given a universe of (rational) language users who learn from finite experience in a changing and stochastically complex arrangement of multiple language games, is a notion of linguistic meaning a theoretically useful and empirically explanatory concept in this picture in such a way that if that notion entailed vague meanings the whole view would have disastrously little explanatory bite?
+
+Taken together, when we look at the picture of multiple language games side by side, the question of what rational or optimal behavior is in any one of them, or the question of why human language users do not more closely approximate rational or optimal local behavior more closely seem entirely misplaced.
+What seems more pressing is to answer how the interaction of individual learning, individual local optimization and general pressure for successful communication at a cross-situational level lead to the shaping of linguistic practices in an open-ended landscape of language games.
+How to attain locally sub-optimal play in a single language game is not a theoretical puzzle we should take serious, if what really matters is a global multi-game perspective.
+Adopting this perspective does not \emph{solve} the problem of how vagueness is compatible with rationality. 
+But it suggest that it might be an inflated issue that we need not or should not worry about, because it will not matter for a more general and pressing issue in a more realistic and encompassing picture of language, its use and evolution.
+The following section will expand on this.
 

### 2017-05-06 book_chapter/paper.tex
Trying to make assumptions about rationality a bit more explicit. Developed argument about the consequences of the natural sources of vagueness a bit more.

@@
 
-\title{Natural sources of uncertainty and their implications}
+\title{Natural sources of vagueness and their implications}
 \author{Jos\'e Pedro Correia \and Michael Franke}
@@
 The fear could be that, if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground on which rationality stands, and the whole edifice would collapse.
-But giving up the ideal of truth and logic as relevant explanatory notions to understand natural language does not mean giving up on rationality, or any of the other notions, altogether.
+But, we want to defend here, giving up the ideal of truth and logic as relevant explanatory notions to understand natural language does not mean giving up on rationality, or any of the other notions, altogether.
 To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
 
-The aforementioned classical picture of language and meaning is thus focused on an aspect of rationality that we can call \emph{procedural}.
-There are, however, other aspects that are also important.
-We can think of rationality as \emph{instrumental}, \emph{i.e.} as the efficient and effective achieving of goals (see Brown~\parencite*[???]{brown_rationality_1990}).
-An agent is rational if he employs his available means in the best way that helps him achieve a given objective.
-This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
-Other animals, however, do this all the time.
-The difference with humans, if we want to establish one, can reveal yet another aspect of rationality.
-Jon Elster~\parencite*{elster_ulysses_1979} suggests that our rationality is characterized by its ability to be more \emph{strategic}, \emph{i.e.}~it can take into account more than the expected immediate results of an action, like anticipation of other players' actions, or consideration of medium/long term gains.
-All of these aspects of rationality (procedural, instrumental, strategic) are probably intertwined in our understanding of rationality.
+Although frequently debated (especially in philosophy, psychology, and economics), rationality is a notion for which clear definitions are hard to find.
+Discussions around it usually rely on common connotations and implicit assumptions.
+The aforementioned classical picture of language and meaning is focused on an aspect of rationality that we can call \emph{procedural}.
+It attempts to characterize a particular choice of an agent as rational based on the procedures used to select it.
+We can also focus instead on the consequences, rather than the procedures.
+\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if the available means were employed in the best way that helps him achieve a given objective.
+%This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
+The notion is linked to David Hume~\parencite*{hume_treatise_1738} and his assertion that ``Reason is, and ought to only be the slave of the passions.''
+Another assumption that is important to qualify when talking about rationality of an agent's choice is the amount of information that the agent has access to, when making his decision.
+Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices will be different.
+This is difficult to characterize in general, but, in the context of a given model, we can describe agents with respect to how much knowledge they have of the ingredients of the model.
+An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
+Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account
+Thus we can say that an agent is more or less \emph{strategic} depending on to what extent he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
+The classical picture is, in these terms, tied to a procedural account of rationality, where agents are pretty much omniscient.
+Questions of how strategic agents are do not really come to the fore, since the approach is typically both atemporal and individualistic.
 
-Do we get different pictures of rationality if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
-What can such a paradigm be?
+What kind of picture of rationality do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
+What could such a paradigm be?
 Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
@@
 \label{sec:finite-experience}
-The research surveyed here suggests various mechanisms that can lead to vague use of messages in certain signaling games, such as cognitive limitations like a finite memory or the inability to always calculate the absolute optimum choice to make, learning mechanisms that compensate for the finiteness of experience through generalization based on similarity, imperfect senses that allow for the possibility of stimulae to be confused, etc\todo{fit Lawry and James plus referential games stuff}.
-In order to account for vagueness in natural language, they all need to peel away from an idealized model of agents, language, and the world, and bring some assumptions down to earth.
-They point to some ways, but by all means not all, in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world.
-Consider the following:
+Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that a vague language is suboptimal compared to a crisp one, work under a number of various assumptions.
+Together, they form a picture of the context where agents learn and use language that is highly idealized.
+This picture has been criticized for several decades, at least in the field of theoretical economics.
+Herbert A.~Simon describes it as such:
+\begin{quote}
+Traditional economic theory postulates an ``economic man,'' who, in the course of being ``economic'' is also ``rational.''
+This man is assumed to have knowledge of the relevant aspects of his environment which, if not absolutely complete, is at least impressively clear and voluminous.
+He is assumed also to have a well-organized and stable system of preferences, and a skill in computation that enables him to calculate, for the alternative courses of action that are available to him, which of these will permit him to reach the highest attainable point on his preference scale.%
+~\parencite[99]{simon_behavioral_1955}
+\end{quote}
+Perhaps this could be seen as the setup of a straw man argument, but we believe it captures some of the crucial assumptions that still underlie some game theoretic models of language.
+
+In fact, in order to account for vagueness in natural language in the context of these models, the proposals surveyed here all need to peel away from this idealized model of agents, language, and the world, and bring some of these assumptions down to earth.
+They suggest various mechanisms that can lead to vague use of messages in certain signaling games, such as cognitive limitations like a finite memory or the inability to always calculate the absolute optimum choice to make, learning mechanisms that compensate for the finiteness of experience through generalization based on similarity, imperfect senses that allow for the possibility of stimulae to be confused, etc\todo{fit Lawry and James plus referential games stuff}.
+In the process, we realize ways in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world:
 \begin{enumerate}
@@
 \end{enumerate}
-It is almost inevitable that uncertainty gets introduced at many levels of language learning and language use.
-In the context of a signaling game, we can expect uncertainty about priors, observations, utilities, behavior of other players, and more.
-This is true both from the perspective of the modeler, as from the perspective of the agents involved.
-Arguments, of the kind presented by Lipman~\parencite*{lipman_why_2009}, that a vague language is suboptimal compared to a crisp one depend on ignoring all of these aspects of the world, language, and experience.
+All of these observations support the weakening of our modeling assumptions.
+% If we take these points seriously, we have to concede that it is almost inevitable that uncertainty gets introduced at almost all levels of language learning and language use.
+% In the context of a signaling game, we can expect uncertainty about priors, observations, utilities, behavior of other players, and more.
+% This is true both from the perspective of the modeler, as from the perspective of the agents involved, supporting the motivation to take it into account in our modeling efforts.
 
+The research surveyed here shows us some examples of assumptions which, when weakened, make vague signal use a natural outcome of certain evolutionary dynamics.
+But it gives us even more.
 We learned from \citeauthor{oconnor_evolution_2014} and \citeauthor{franke_vagueness_2017} that vague languages are quicker to converge and more adaptable, which is valuable given the finite and dynamic character of our experience (point 1).
@@
 We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism thats leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
-\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).\todo{add something for 5 and 6, maybe referential games}
-Trying to answer the question of ``Why is language vague?'', is important to force us to recognize those aspects of language and the world that are easily forgotten when idealizing the subject matter.
-The finiteness of human experience, the dynamic nature of the world, and the communitarian context of language use and development, these are things we need to keep in mind lest we end up with inadequate models of language and meaning.
+\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).
+\todo[inline]{add something for 5 and 6, maybe referential games}
+%Trying to answer the question of ``Why is language vague?'', is important to force us to recognize those aspects of language and the world that are easily forgotten when idealizing the subject matter.
+%The finiteness of human experience, the dynamic nature of the world, and the communitarian context of language use and development, these are things we need to keep in mind lest we end up with inadequate models of language and meaning.
 % Wittgenstein put it better:
@@
 What do these observations tell us about rationality?
-An agent with minimal instrumental rationality, merely adapting its behavior based on reinforcement, will be directly conditioned by the limitations imposed by its environment.
-Therefore, given all the natural sources of uncertainty discussed so far, even the most effective of such agents (or population thereof) would develop a signaling behavior that inevitably makes vague use of messages.
-The agent would still be driving for the highest possible payoff it can achieve, by definition, and the presence of vagueness could not be attributed to any lack of rationality.
-For a more complex agent, the story is perhaps more nuanced.
+% The replicator equation lends itself to an interpretation where agents could be said to be devoid of rationality, procedural or instrumental.
+% This is the case where we see the equation as purely descriptive of a population of agents with hardwired behaviors, where the most successful strive and the least successful die off.
+% In such an interpretation there is no room for questions regarding the rationality of vagueness.
+% Another possible interpretation is that of a population of agents learning by imitation.
+The LMF dynamics of~\textcite{franke_vagueness_2011}, generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, all paint a picture of agents with a basic level of instrumental rationality, very limited awareness of the game, and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
+Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
+All approaches introduce constraints into the dynamics that prevent the evolution of crisp signal use.
+But a crisp language would still have a higher expected utility than the evolved strategies.
+Agents in those models seem to be rational only to the extent of their capabilities.
+Despite the plausibility of the mechanisms proposed (limited memory, generalization, state confusability), this feels like a bittersweet conclusion because of this hypothetical possibility of a better, hence more rational, strategy, that seems to be only artificially barred from the agents.
+But this is so only if rationality is equated with maximization of expected utility.
+The mechanisms that lead to vague signal use, as~\textcite{oconnor_evolution_2014} and~\textcite{franke_vagueness_2017} stress, have the aforementioned important advantages of faster speed of convergence, higher flexibility, and homogenization.
+These advantages matter, especially in the complex, dynamic context of real language use.
+One would thus like to say that using a strategy that promotes those characteristics without a significant loss of expected utility is certainly a rational choice.
+But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond maximization of expected utility.
+
+For more complex agents, the story is potentially more nuanced.
 Consider an agent that has awareness that something like a game is being played, some estimate of the potential payoff involved, ability to gauge other players' expected behaviors, and the capacity to think strategically.
 Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
-If he is \emph{not} aware of the natural sources of uncertainty discussed so far, one cannot charge him with irrationality since by adopting a crisp strategy he is choosing the optimal strategy for his assumptions.
-However, if he \emph{is} aware, then it would be less rational to use a crisp strategy than to use a strategy that incorporates at least some source of uncertainty.
+In a scenario where the agent ability to make the optimal choice is bounded, such as the QRE model of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
+\textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
+The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as utility maximization.
+But even in other scenarios, we could again argue that, if we broaden our notion of rationality, it would be less rational to use a crisp strategy rather than one that incorporates at least some source of uncertainty.
 A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
 
-\todo[inline]{precision is possible though, but not absolute precision, and vagueness comes in degrees, figuring out how much vagueness to use is the real challenge, and this needs to be studied}
+Thus, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.
+Furthermore, if we broaden our notion of rationality beyond mere maximization of expected utility, we also find reasons to welcome the positive aspects of these processes and to see their acceptance as a perfectly rational choice.
+The investigation doesn't end here, though.
+The research discussed here suggests that although vagueness can be beneficial, that is the case only in moderation: too much can lead to babbling equilibria where no information is transmitted.
+Thus, for agents capable of adjusting the level of vagueness in their strategies, questions remain regarding how this can be modeled and what factors influence the amount of vagueness that is optimal for a given situation.
+Other challenges include how to incorporate the intuitions about the relevance of the aspects enumerated above into our models.
+
+\section{Conclusion}
+\label{sec:conclusion}
 

### 2017-05-07 book_chapter/paper.tex
Separated introduction of vagueness from introduction of signaling games. Made some small changes on the former.

@@
 
-\section{Vagueness}
+\section{Vagueness and the classical picture}
 \label{sec:vagueness}
@@
 Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
-The problem thus seems to be a serious one.
-But it does not obviously undermine our everyday practices, so the issue must be with the conceptions we have of the mechanisms that underlie them.
+The problem thus seems to be a wide-ranging one, but it does not obviously undermine our everyday linguistic practices.
+Therefore, the issue must lie with the conceptions we have of the forces or mechanisms that underlie those practices.
 
-Vagueness is typically seen as a challenge to the classical conception of language and meaning.
+Vagueness is typically seen as a challenge to a classical conception of language that relies on reference and truth as two core notions for understanding meaning.
 One problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to something like a well defined set of tall people?
@@
 Supervaluationism, many-valued logics, and degree theories all propose changes to the classical picture in order to accommodate for vague predicates.
-They are, however, still strongly committed to retaining as much as possible of it, particularly the core notions of truth and reference.
-Mark Sainsbury~\parencite*{sainsbury_concepts_1999} argues that, because of that, they all fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
+They are, however, still strongly committed to its two core notions.
+As argued by Mark Sainsbury~\parencite*{sainsbury_concepts_1999}, because of that they fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
 All the aforementioned proposals end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
 But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
-Further introducing borderline borderline cases would lead into an infinite regress.
-Because of their attachment to the classical picture, the standard approaches to vagueness fail to see an important lesson, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
+Further introducing borderline borderline cases would lead to an infinite regress.
+Because of their attachment to the classical picture, the standard approaches to vagueness also fail to draw an important lesson from the sorites paradox, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
 % \begin{quote}
@@
 \end{quote}
+By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
 Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
@@
 This type of rationality is one that supposedly demarcates humans from other animals.
-The fear could be that, if we drop the picture of meaning as intimately tied to truth and logic, we lose the ground on which rationality stands, and the whole edifice would collapse.
-But, we want to defend here, giving up the ideal of truth and logic as relevant explanatory notions to understand natural language does not mean giving up on rationality, or any of the other notions, altogether.
+We want to argue here that giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
 To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
+What we need is to revisit the notion of rationality tied to the classical picture.
 
-Although frequently debated (especially in philosophy, psychology, and economics), rationality is a notion for which clear definitions are hard to find.
+Although frequently debated, especially in areas like philosophy, psychology, and economics, rationality is a notion for which clear definitions are hard to come by.
 Discussions around it usually rely on common connotations and implicit assumptions.
@@
 We can also focus instead on the consequences, rather than the procedures.
-\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if the available means were employed in the best way that helps him achieve a given objective.
+\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the means available to him were employed in the best way possible that enables him achieve a certain goal.
 %This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
@@
 Another assumption that is important to qualify when talking about rationality of an agent's choice is the amount of information that the agent has access to, when making his decision.
-Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices will be different.
+Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices can vary.
 This is difficult to characterize in general, but, in the context of a given model, we can describe agents with respect to how much knowledge they have of the ingredients of the model.
@@
 Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account
-Thus we can say that an agent is more or less \emph{strategic} depending on to what extent he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
+Thus we can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
 The classical picture is, in these terms, tied to a procedural account of rationality, where agents are pretty much omniscient.
-Questions of how strategic agents are do not really come to the fore, since the approach is typically both atemporal and individualistic.
+Questions of how strategic they are do not really come to the fore, since the analysis is typically both atemporal and individualistic.
 
-What kind of picture of rationality do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
-What could such a paradigm be?
+What kind of rationality do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
+What could even such a paradigm be?
 Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
 \footnote{To the extent that substantial views can be said to be defended by the author.
-Skipping over the debate (see \cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
-Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} we can find a picture of language in terms of multiplicity, heterogeneity, and change.
-Languages are seen as patchworks of various language-games, with new ones continuously being added and old ones falling out of use.
-These language-games, in turn, can be thought of language intermingled with a practice.
-The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
+Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
+Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}, we can find a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
+Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
+These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
+%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
 There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
 The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
-Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play through according to those assumptions.
-It has been argued~\parencite{correia_bivalent_2013} that the framework of signaling games, introduced for the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that, while reaping the benefits of mathematical formalism and computer simulation to explore the implications of complex hypothetical language-games in depth.
-From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting computer simulations is like contemplating how the game would play through according to the assumptions one built into the model.
+Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
+It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
+From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.
+
+Following this idea, we here survey proposals for addressing vagueness in the context of signaling games.
+The following section introduces the framework of signaling games.
+Then in sections~\ref{sec:sim-max-vagueness} and~\ref{sec:referential-vagueness} we discuss two particular types of games, namely similarity maximization and referential games, and which suggestions there are with regards to how vagueness can be accounted for in those models.
+We bring the insights together in section~\ref{sec:finite-experience} and reflect on the consequences of the proposed natural sources of vagueness for notions of rationality.
+
+\section{Signaling games}
+\label{sec:signaling-games}
 

### 2017-05-07 book_chapter/paper.tex
Fine-tuning the introduction to signaling games.

@@
 % In order to extend his notion of convention to \emph{linguistic} exchanges
-In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where the actions available involve sending and receiving signals or messages.
+In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where agents' choices involve sending and receiving signals or messages.
 Thus, we could think of two players with different roles.
@@
 Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
-Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function $U : T \times A \rightarrow \mathbb{R}^2$.\mf{explain sender and receiver
-payoffs}
+Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function%
+\footnote{This could also be represented with two separate functions.
+In cases of perfectly aligned preferences, the function can be simplified to yield only one value, to be equally attributed to each player.}
+$U : T \times A \rightarrow \mathbb{R}^2$ which yields a pair of payoff values, one for the sender, one for the receiver, for each possible pairing of state and action.
 These so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
-A communicator's contingency plan, or sender strategy as we will call it, is a specification of a choice of message for each possible state of affairs.
+A sender strategy is a specification of a choice of message for each possible state of affairs.
 It thus describes the sender's behavior conditional on the state of affairs that obtains.
-An audience's contingency plan, or receiver strategy as we will call it, analogously specifies a choice of action for each possible message.
+A receiver strategy analogously specifies a choice of action for each possible message.
 Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
-The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function between states of affairs and actions as a sum of the payoffs for all cases.
+The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions.
 %, \emph{i.e.}:
@@
 Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, strategic rationality, mutual ascription of strategic rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
+Agents are thus envisioned as omniscient of the game and highly strategic.
 These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
 The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
-However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of strategic rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
+However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
 Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
-A plausible account of the origin of language should first explain how signaling systems like those could get started, without assuming a great deal of strategic rationality from the part of the agents involved.
+A plausible account of the origin of language should first explain how signaling systems like those could get started, without requiring high standards of rationality of the agents involved.
 
@@
 Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
-Although the inspiration for this scenario is biological evolution, similar things could be said~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985} about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}.
+Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985}.
 The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
@@
 
-Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalisms to the descriptive side of the investigation.
+Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
 Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
 Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
-Focus is put on understanding how various ingredients to the model interact, and which results they produce, not on metaphysical concerns.
+Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
 While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
-We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
-This range of options goes hand in hand with assumptions of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of rationality in the case of dynamics that attempt to mimic mechanisms of iterative learning, to higher levels of rationality and even game-theoretical reasoning.
+We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
+This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even game-theoretical reasoning.
 Each of these can be utilized depending on the problem that one is interested in characterizing.
@@
 
-The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how. \mf{explain/introduce equilibrium and/or ESS?}
-Not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't.
+The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
+A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
+With these tools, not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't.
 In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
@@
 Very succinctly, it can be put as follows.
-In standard game-theoretic models of communication, vague usage of messages is less efficient than precise use.
+In standard game-theoretic models of communication, vague signal use yields a lower expected utility than crisp use.
 Therefore, given that the dynamics (be it natural selection, cultural evolution, or rational choice) maximize utility, vagueness should be weeded out by these forces, giving rise to only precise languages.
 However, vagueness is pervasive in our natural languages, and there is no reason to believe it is going away.
-Lipman concludes then that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}.
-In the following section we provide an overview of models that try to account for vagueness in a signaling games framework, and what the implications of each account are for how we view rationality.
+Lipman concludes that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}.
+What proposals have been put forward to address this issue?
 
@@
 
-Before we discuss the particular suggestions of how to account for vagueness in a signaling
-games framework, we need to discuss what useful changes we could make to the classic setup.
-First, in order to work, the sorites paradox requires us to assume a relation between the vague term and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
+In order to work, the sorites paradox requires us to assume a relation between the vague term and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
 Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
@@
 In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (or better memory) would lead to more precision.
-\citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{ellison_learning_1993}\todo{is this the canonical reference?}.
+\citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{gilboa_social_1991}.
 A combination of both would hypothetically lead to a better outcome in the scenario of \citeauthor{franke_vagueness_2017}.

### 2017-05-07 book_chapter/paper.tex
Fine-tuning section on vagueness in sim-max games.

@@
 
-In order to work, the sorites paradox requires us to assume a relation between the vague term and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
+In order to work, the sorites paradox requires us to assume a relation between the vague terms and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
 Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
-In terms of signaling games, we can take the height to be the state space and the terms the message space.
+In terms of signaling games, we can take the state space to be constituted of values of the underlying dimension, and the message space as constituted by the terms in question.
 Because of the difference in granularity, we will typically be interested in cases where the state space is much larger than the message space.
-For the structure of the state space, we can model this by defining a distance or similarity function between every element in the set, effectively making it a metric space.
-Second, another important ingredient of the paradox is the acknowledgment of a certain degree of tolerance with respect to whether a certain term applies or not.
-This tolerance decreases with distance: assuming a 180cm person is tall, one would easily tolerate the use of the term for a person measuring 179cm, less so for someone who is 170cm, and much less so for 160cm.
-This can be modeled by using a utility function that is continuous rather than discrete and that is monotonously decreasing with distance in a metric space, \emph{i.e.}~success is not a matter of black and white, right or wrong, but a matter of degree, of how close the agent got to the optimal.
+We can model the structure of the state space by defining a distance or similarity function between every value, effectively making it a metric space.
+Another important ingredient of the paradox is the acknowledgment of a certain degree of tolerance with respect to whether a certain term applies or not.
+This tolerance decreases with distance in state space: assuming a 180cm person is tall, one would easily tolerate the use of the term for a person measuring 179cm, less so for someone who is 170cm, and much less so for 160cm.
+This can be modeled by using a utility function that is continuous rather than discrete and that is monotonously decreasing with distance, \emph{i.e.}~success is not a matter of black and white, right or wrong, but a matter of degree, of how close the receiver got to the optimal response to the sender's perceived state.
 
-Although we will look at more complex cases below, the simplest type of game to study in this scenario is one where the state space and the action space are the same. 
+The simplest type of game to study in this scenario is one where the state space and the action space are the same. 
 We can imagine this as a game of guessing the state of affairs: the sender has knowledge of it, sends a message to the receiver, who in turn has to guess it; their payoff, as discussed above, is proportional to how close he got.
@@
 Roughly, these are situations where the sender uses messages in a way that can be seen as partitioning the state space into convex regions and the receiver responds with the central element of those regions.
-For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280.\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}
+For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280.%
+\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}
 Given two possible messages, the optimal sender strategy is to use one for all values from 40 up to 160, and the other for all values from 160 up to 280.
 The optimal receiver strategy is to guess with 100 if given the first message and with 220 given the second message.
-Confirming Lipman's argument, there is no vagueness in such an optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, associating only one guess with each message.
+Confirming Lipman's argument, there is no vagueness in such an optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, univocally associating one guess with each message.
 
@@
 
-In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}.
-The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}.
-While in the former the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message, at a given point, in the latter these transitions are smooth.
+In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space.
+The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where use of one message over the other by the sender is characterized by a smooth transition, and the receiver strategy assigns for each message a positive probability to more than one state.
+The particular shapes of these strategies are not important to characterize a vague signal use.
+What is important is that the transition is smooth and monotonous for both players.
 This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is almost certain.
-Similarly for the receiver strategy, in the first example for each message there is an almost certain response, whereas in the second example this is a degree of uncertainty.
 
@@
 % \subsection{Bounded rationality}
-\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can arise as a consequence of limitations in the cognitive capacities of rational agents.
+\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can arise as a consequence of limitations in the cognitive capacities of instrumentally rational agents.
 The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
 For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
-Agents are assumed to be rational, thus each plays a best response (given the utility function) to their estimate of the other's strategy.
-In order to observe the evolution of strategies in repeated interaction, this approach involves modeling several individual agents in actual play.
+Each agent plays a best response (given the utility function) to their estimate of the other's strategy.
+In order to observe the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
 What the authors observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
@@
 The degree to which agents tremble in the calculation of expected utilities and therefore deviate from the optimal behavior can be characterized by a parameter.
-\textcite{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria (where sender and receiver simply randomize, respectively, message and response uniformly) where no communication actually occurs.
+\citeauthor{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria are possible, where sender and receiver simply randomize, respectively, message and response uniformly.
 Above a certain value of the parameter, other equilibria of the kind described in the beginning of this section arise, where agents communicate successfully, though not perfectly, using fuzzy strategies similar to the ones depicted in Figure~\ref{fig:example_stratsB}.
@@
 %First, it is not argued that vagueness in itself is beneficial, but that it is a by-product of a learning mechanism.
-O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though not obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
-Considerations of optimality of strategies are typically made in terms of hypothetical atemporal comparisons of expected utility.
-However, generalized reinforcement learning has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a smaller period of time.
+O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though less obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
+Considerations of optimality of strategies, for example by considering the evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
+However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a smaller period of time.
 In a naturalized evolutionary setting this temporary advantage should also be taken into account.
-Imagine an initial population of agents with random strategies, some using GRL and others using classical Herrnstein reinforcement learning to adapt to each other.
+Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.
 Although the latter type of agent can hypothetically develop a precise and more efficient signaling system, agents using GRL would coordinate on vague signaling strategies with high (though not optimal) expected utility sooner than agents using HRL.
@@
 In terms of evolutionary game theory, one could argue that a small enough group of mutants playing a precise signaling system could later invade a population of vague signalers.
-There is, an additional advantage to the generalized learning that is relevant here, namely the ability to adapt to a changing environment.
+There is, however, an additional advantage to the generalized learning that is relevant here, namely the ability to adapt to a changing environment.
 Analysis of a game in terms of evolutionary stable states works well if we assume that the environment where the agents interact remains static.
 However, in a more realistic setting this is not the case.
-In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utiliry function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of their short-term gains.
+In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utiliry function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of its short-term gains.
 
@@
 Furthermore, it seems reasonable to assume that this state confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is they will be mistaken for each other.
-Incorporating thiese considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of that dynamic that also induces vague signal use of the kind we expect here.
+Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of that dynamic that also induces vague signal use of the kind we expect here.
 The consequence is very similar to that of the GRL model discussed above, in that the way the behavior for a given state is updated takes into account the behavior of similar states, proportionally to their similarity.
-Given the known relation between reinforcement learning and the replicator dynamics~\parencite{Beggs2005}, it is actually quite plausible that the two are tightly related, but this would need to be formally demonstrated.
-The account is, however, interpretable at a lower level of rationality since the replicator equation is suitable to represent biological processes of natural selection.
+Given the known relation between reinforcement learning and the replicator dynamics~\parencite{Beggs2005}, it is actually quite plausible that the two are tightly related (although this would need to be formally demonstrated).
+The account is, furthermore, interpretable at a lower level of rationality, given that the replicator equation is suitable to represent biological processes of natural selection.
 
 The motivation underlying this model of vague signaling is still one of inevitability.
-A vague strategy is still not claimed to have higher expected utility than a crisp one.
+A vague strategy is not claimed to have higher expected utility than a crisp one.
 However, the authors observe an effect similar to that pointed out by~\citeauthor{oconnor_evolution_2014}: signaling converges faster in scenarios where there is some degree of state confusability.
 Furthermore, they observe one additional potentially beneficial property.
-By running several rounds of simulation for each parameter set, they can measure how close resulting strategies are to each other within each group, and how they would fair playing against each other by calculating a within group expected utility.
-The results show that the within group distances between strategies is smaller with growing confusability, and the within group expected utility is actually higher for strategies evolving under a certain degree of state confusion.
-Thus, some amount of uncertainty seems to promote more homogeneous populations of signalers that are better at achieving cooperation within a group of such signalers
+By running several rounds of simulation for each parameter set, they can measure how close resulting strategies are within each group of results, and how they would fair playing against each other.
+The results show that the within group distance between strategies becomes smaller with growing confusability, and the within group expected utility is actually higher for strategies evolving under a certain degree of state confusion.
+Thus, some amount of uncertainty seems to promote more homogeneous populations of signalers that are better at achieving cooperation within a group.
 
@@
 \textcite{lawry_vagueness_2017} explore the potential benefits of vagueness in variations of a sim-max game with multiple senders and one receiver.
-Although the models considered are not explicitly identified as sim-max games, they are equivalent in the relevant aspects: they are signaling games where senders have private knowledge of a given state $x$, each sends a message to the receiver, which then selects an estimate $y$ of the original state; although the authors do not define a utility function, their analysis evaluates the models in terms of the expected error between the estimate and the original state, defined as $(x-y)^2$; better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse; the game could then be seen as a sim-max game with multiple senders and negative squared difference as similarity function.
+Although the models considered are not explicitly identified as sim-max games, they are equivalent in the relevant aspects.
+They are signaling games where senders have private knowledge of a given state $x$, each sends a message to the receiver, which then selects an estimate $y$ of the original state.
+Although the authors do not define a utility function, their analysis evaluates the models in terms of the expected error between the estimate and the original state, defined as $(x-y)^2$.
+Better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse
+The game could then be seen as a sim-max game with multiple senders and negative squared difference as similarity function.
+
 Unlike the models discussed so far that introduce minimal constraints to a dynamic and study the strategies that emerge out of an evolutionary process, these authors' approach is mostly static and analytical.
-They consider several possible sender strategies with uncertainty and without and calculate if and when sender and receiver achieve better transmission accuracy, \emph{i.e.}~smaller expected error between original state and receiver estimate.
-Senders without uncertainty have a strategy that boils down to what is represented in Figure~\ref{fig:example_stratsA}: they split the state space according to a fixed thresholds, switching from using one message to another abruptly.
-The model for the other type of sender is similar, but assumes that there is uncertainty regarding the value of the thresholds, representing them as uniformly distributed random variables.
-What results from this in terms of sender strategy is also a smooth transition between the use of two messages, but unlike what we see in Figure~\ref{fig:example_stratsB}, it is a linear transition\footnote{This is certainly connected with the choice of modeling the uncertain threshold as uniformly distributed.
-We speculate that another choice of distribution, like for example a normal distribution, would yield something similar.
-What is important to characterize a vague use of messages, however, is that the transition is smooth and monotonous, not its particular shape.}.
+They consider several possible sender strategies, with and without uncertainty, and calculate if and when sender and receiver achieve better transmission accuracy, \emph{i.e.}~smaller expected error between original state and receiver estimate.
+Senders without uncertainty have a strategy that boils down to what is represented in Figure~\ref{fig:example_stratsA}: they split the state space according to a fixed threshold, switching from using one message to another abruptly.
+The model for the other type of sender is similar, but assumes that there is uncertainty regarding the value of the threshold, representing it as a uniformly distributed random variable.
+What results from this in terms of sender strategy is also a smooth transition between the use of two messages%
+\footnote{Unlike what we see in Figure~\ref{fig:example_stratsB}, it is a linear transition.
+This is certainly connected with the choice of modeling the uncertain threshold as uniformly distributed.
+We speculate that another choice of distribution, such as for example a normal distribution, could yield something similar to Figure~\ref{fig:example_stratsB}.}%
+.
 The receiver strategy is not inherently vague for either type of sender, it merely aggregates the messages received to provide an estimate of the original state using a simple arithmetic mean.
@@
 For a model with one sender, results are consistent with Lipman's~\parencite{lipman_why_2009} in that strategies without uncertainty lead to more accurate communication.
-For models with two messages and more than one sender, the expected error is a strictly decreasing function of the number of senders, and above a certain number of senders, strategies with uncertainty actually achieve lower expected error than sharp ones.
+For models with two messages and more than one sender, the expected error is a strictly decreasing function of the number of senders, and above a certain number of senders, strategies with uncertainty actually achieve lower expected error than crisp ones.
 This minimal number of senders from which vagueness surpasses sharpness is a decreasing function of the number of messages, \emph{i.e.}~whereas at least 8 senders are needed if the number of messages available is 2, this number lowers to 5 for 3 messages, 4 for and 5 messages, and 3 for more than 6 messages.
@@
 
-These results are interesting since they reveal scenarios where vague strategies have an advantage over strict ones, even in a setting of higher rationality.
+These results are interesting since they reveal scenarios where vague strategies have an advantage over strict ones, even in a setting of game-omniscient strategic agents.
 The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the idealist\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
@@
 A combination of both would hypothetically lead to a better outcome in the scenario of \citeauthor{franke_vagueness_2017}.
-However, \citeauthor{lawry_vagueness_2017} give us something that holds even if the idealist is unwilling to acknowledge the constraints that permeate our finite existence.
-Whether the scenario of multiple senders is representative of the language games where vagueness comes about, and whether the strategies proposed by the authors would be the ones to evolve\footnote{Their analysis is a static one.
+However, \citeauthor{lawry_vagueness_2017} give us something that holds even if the idealist is unwilling to acknowledge the constraints that pervade our finite existence.
+Whether the scenario of multiple senders is representative of the language games where vagueness comes about, and whether the strategies proposed by the authors would be the ones to evolve\footnote{The authors' analysis is a static one.
 Despite the advantage of the proposed vague strategies against strict ones in certain settings, one cannot know whether other strategies of higher accuracy would not be possible in those same settings without studying some dynamics in an evolutionary context.}, is something that is, however, open for debate.

### 2017-05-08 book_chapter/paper.tex
Moved section on referential games. Fine-tuned discussion on uncertainty and rationality. Moved last part of that discussion to conclusion section.

@@
 
+\section{How could language not be vague?}
+\label{sec:finite-experience}
+
+Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that a vague language is suboptimal compared to a crisp one, work under a number of various assumptions.
+Together, these assumptions form a picture of the context where agents learn and use language that is highly idealized and probably derives, via game theory, from the conception of rationality of traditional theoretical economics.
+%This picture has been criticized for several decades, at least in the field of theoretical economics.
+Herbert A.~Simon describes this picture as follows:
+\begin{quote}
+Traditional economic theory postulates an ``economic man,'' who, in the course of being ``economic'' is also ``rational.''
+This man is assumed to have knowledge of the relevant aspects of his environment which, if not absolutely complete, is at least impressively clear and voluminous.
+He is assumed also to have a well-organized and stable system of preferences, and a skill in computation that enables him to calculate, for the alternative courses of action that are available to him, which of these will permit him to reach the highest attainable point on his preference scale.%
+~\parencite[99]{simon_behavioral_1955}
+\end{quote}
+This ``economic man'', a game-omniscient and highly strategic rational agent, seems to have survived almost unscathed to now populate most game theoretic models of language and signaling.
+Both Simon and Lipman call for this picture to be revisited.
+
+In order to account for vagueness in natural language in the context of these models, the proposals surveyed here all need to peel away from this idealized picture and bring some of these assumptions down to earth.
+In the process, they point to ways in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world:
+\begin{enumerate}
+\item Our existence is temporally finite; language does not have an infinite amount of time to evolve, nor can it take an infinite time to be learned.
+The faster a language can start being useful, the better;
+\item Language learning through experience has to rely on a limited number of observations.
+Not only is the state space typically much larger than what one can survey in sufficient time, it is even potentially infinite and constantly changing;
+\item A corollary of the former is that there will always be heterogeneity in a population of language learners, at the very least in their prior experience, since each agent will have relied on a different set of observations.
+Furthermore, this information is not directly or fully accessible to others;
+\item Given that an agent is almost always learning and using language in a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice;
+\end{enumerate}
+All of these observations support the weakening of our modeling assumptions.
+% If we take these points seriously, we have to concede that it is almost inevitable that uncertainty gets introduced at almost all levels of language learning and language use.
+% In the context of a signaling game, we can expect uncertainty about priors, observations, utilities, behavior of other players, and more.
+% This is true both from the perspective of the modeler, as from the perspective of the agents involved, supporting the motivation to take it into account in our modeling efforts.
+The research surveyed here shows us some examples of assumptions which, when weakened, make vague signal use a natural outcome of certain evolutionary dynamics.
+% That includes considering various mechanisms that can lead to vague use of messages in certain signaling games: cognitive limitations like a finite memory or the inability to always calculate the absolute optimum choice to make; learning mechanisms that compensate for the finiteness of experience through generalization based on similarity; and imperfect senses that allow for the possibility of stimuli to be confused, etc\todo{fit Lawry and James plus referential games stuff}.
+But it gives us even more.
+It suggests ways in which the mechanisms that lead to vagueness can have positive effects that are extremely important in the context of the points just enumerated.
+We learned from \citeauthor{oconnor_evolution_2014} and \citeauthor{franke_vagueness_2017} that vague languages are quicker to converge and adapt, which is valuable given the finite and dynamic character of our experience (point 1).
+\citeauthor{oconnor_evolving_2015} also showed how generalization, an invaluable feature of any procedure for learning from a limited number of observations (point 2), leads to vagueness.
+We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism thats leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
+\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).
+%Trying to answer the question of ``Why is language vague?'', is important to force us to recognize those aspects of language and the world that are easily forgotten when idealizing the subject matter.
+%The finiteness of human experience, the dynamic nature of the world, and the communitarian context of language use and development, these are things we need to keep in mind lest we end up with inadequate models of language and meaning.
+% Wittgenstein put it better:
+
+What do these observations tell us about rationality?
+% The replicator equation lends itself to an interpretation where agents could be said to be devoid of rationality, procedural or instrumental.
+% This is the case where we see the equation as purely descriptive of a population of agents with hardwired behaviors, where the most successful strive and the least successful die off.
+% In such an interpretation there is no room for questions regarding the rationality of vagueness.
+% Another possible interpretation is that of a population of agents learning by imitation.
+The LMF dynamics of~\textcite{franke_vagueness_2011}, generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, all paint a picture of agents with a basic level of instrumental rationality, very limited awareness of the game, and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
+Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
+All approaches introduce constraints into the dynamics that prevent the evolution of crisp signal use.
+But a crisp language would still have a higher expected utility than the evolved strategies.
+Agents in those models seem to be rational only to the extent of their capabilities.
+Despite the plausibility of the mechanisms proposed (limited memory, generalization, state confusability), this feels like a bittersweet conclusion because of this hypothetical possibility of an ideal strategy that seems to be only artificially barred from the agents.
+But this is so only if rationality is equated with maximization of expected utility.
+The mechanisms that lead to vague signal use, as~\textcite{oconnor_evolution_2014} and~\textcite{franke_vagueness_2017} stress, have the aforementioned important advantages of faster speed of convergence, higher flexibility, and homogenization.
+These advantages matter, especially in the complex, dynamic context of real language use.
+One would thus like to say that using a strategy that promotes those characteristics without a significant loss of expected utility is certainly a rational choice.
+But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond maximization of expected utility, or a notion of expected utility that incorporates other dimensions.
+
+For more complex agents, the story is potentially more nuanced.
+Consider an agent that has awareness that something like a game is being played, some estimate of the potential payoff involved, ability to gauge other players' expected behaviors, and the capacity to think strategically.
+Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
+In a scenario where the agent's ability to make the optimal choice is bounded, such as the QRE model of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
+\textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
+The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as utility maximization.
+But even in other scenarios, we could again argue that, if we broaden our notion of rationality, it would be less rational to use a crisp strategy rather than one that incorporates at least some source of uncertainty.
+A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
+
+We believe these considerations are, again, very much in line with Wittgenstein's later philosophy:
+\begin{quote}
+The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
+(For the crystalline purity of logic was, of course, not something I had discovered: it was a requirement.)
+The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
+– We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
+We want to walk: so we need friction.
+Back to the rough ground!%
+~\parencite*[\S 107]{wittgenstein_philosophical_1953}
+\end{quote}
+This seems like a very clear call for what we are arguing here.
+But are we exposing ourselves to ``philosophical disease'' by following a ``one-sided diet''~\parencite*[\S 593]{wittgenstein_philosophical_1953} when looking only at sim-max games?
+In the spirit of Wittgenstein's pluralistic picture of language, we would like to reflect on vagueness in one more type of signaling game and speculate on the impact of plurality for vagueness.
+
 \section{Vagueness in multiple reference games}
 \label{sec:referential-vagueness}
+
+% \item We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another;
+% \item These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time.
 
@@
 
-% \section{Vagueness in related models}
-% \label{sec:other-vagueness}
-%
-% \subsection{Preference misalignment}
-% \todo[inline]{Test CS~\parencite{crawford_strategic_1982} results in our model by using bias in utility function.
-READ Förster, M., and Riedel, F.
-(2011).
-Distorted Voronoi languages (working Paper).
--> distorted Voronoi languages, i.e.
-shifted according to bias, but still crisp}
-%
-% \subsection{Noisy communication}
-% \todo[inline]{Test if noise in communication channel as modeled by Blume~\parencite{blume_intentional_2014} induces vague strategies as well}
+\section{Conclusion}
+\label{sec:conclusion}
 
-\section{How could language not be vague?}
-\label{sec:finite-experience}
-Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that a vague language is suboptimal compared to a crisp one, work under a number of various assumptions.
-Together, they form a picture of the context where agents learn and use language that is highly idealized.
-This picture has been criticized for several decades, at least in the field of theoretical economics.
-Herbert A.~Simon describes it as such:
-\begin{quote}
-Traditional economic theory postulates an ``economic man,'' who, in the course of being ``economic'' is also ``rational.''
-This man is assumed to have knowledge of the relevant aspects of his environment which, if not absolutely complete, is at least impressively clear and voluminous.
-He is assumed also to have a well-organized and stable system of preferences, and a skill in computation that enables him to calculate, for the alternative courses of action that are available to him, which of these will permit him to reach the highest attainable point on his preference scale.%
-~\parencite[99]{simon_behavioral_1955}
-\end{quote}
-Perhaps this could be seen as the setup of a straw man argument, but we believe it captures some of the crucial assumptions that still underlie some game theoretic models of language.
-
-In fact, in order to account for vagueness in natural language in the context of these models, the proposals surveyed here all need to peel away from this idealized model of agents, language, and the world, and bring some of these assumptions down to earth.
-They suggest various mechanisms that can lead to vague use of messages in certain signaling games, such as cognitive limitations like a finite memory or the inability to always calculate the absolute optimum choice to make, learning mechanisms that compensate for the finiteness of experience through generalization based on similarity, imperfect senses that allow for the possibility of stimulae to be confused, etc\todo{fit Lawry and James plus referential games stuff}.
-In the process, we realize ways in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world:
-\begin{enumerate}
-\item Our existence is temporally finite; language does not have an infinite amount of time to evolve or take a long time to be learned. The faster a language can start to be useful, the better;
-\item Language learning through experience has to rely on a limited number of observations.
-Not only is the state space typically much larger than what one can survey in sufficient time, but it is also potentially infinite and constantly changing;
-\item A corollary of the former is that there will always be heterogeneity in a population of language learners, at least of prior experience since each agent will have relied on a different set of observations.
-Furthermore, this information is not directly or fully accessible to others;
-\item Given that an agent is almost always learning and using language in such a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice;
-\item We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another;
-\item These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time.
-\end{enumerate}
-All of these observations support the weakening of our modeling assumptions.
-% If we take these points seriously, we have to concede that it is almost inevitable that uncertainty gets introduced at almost all levels of language learning and language use.
-% In the context of a signaling game, we can expect uncertainty about priors, observations, utilities, behavior of other players, and more.
-% This is true both from the perspective of the modeler, as from the perspective of the agents involved, supporting the motivation to take it into account in our modeling efforts.
-
-The research surveyed here shows us some examples of assumptions which, when weakened, make vague signal use a natural outcome of certain evolutionary dynamics.
-But it gives us even more.
-We learned from \citeauthor{oconnor_evolution_2014} and \citeauthor{franke_vagueness_2017} that vague languages are quicker to converge and more adaptable, which is valuable given the finite and dynamic character of our experience (point 1).
-\citeauthor{oconnor_evolving_2015} also showed how generalization, an invaluable feature of any procedure for learning from a limited number of observations (point 2), leads to vagueness.
-We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism thats leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
-\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).
-\todo[inline]{add something for 5 and 6, maybe referential games}
-%Trying to answer the question of ``Why is language vague?'', is important to force us to recognize those aspects of language and the world that are easily forgotten when idealizing the subject matter.
-%The finiteness of human experience, the dynamic nature of the world, and the communitarian context of language use and development, these are things we need to keep in mind lest we end up with inadequate models of language and meaning.
-% Wittgenstein put it better:
-% \begin{quote}
-% The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
-(For the crystalline purity of logic was, of course, not something I had discovered: it was a requirement.) The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
-– We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
-We want to walk: so we need friction.
-Back to the rough ground!%
-% ~\parencite*[\S 107]{wittgenstein_philosophical_1953}
-% \end{quote}
-
-What do these observations tell us about rationality?
-% The replicator equation lends itself to an interpretation where agents could be said to be devoid of rationality, procedural or instrumental.
-% This is the case where we see the equation as purely descriptive of a population of agents with hardwired behaviors, where the most successful strive and the least successful die off.
-% In such an interpretation there is no room for questions regarding the rationality of vagueness.
-% Another possible interpretation is that of a population of agents learning by imitation.
-The LMF dynamics of~\textcite{franke_vagueness_2011}, generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, all paint a picture of agents with a basic level of instrumental rationality, very limited awareness of the game, and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
-Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
-All approaches introduce constraints into the dynamics that prevent the evolution of crisp signal use.
-But a crisp language would still have a higher expected utility than the evolved strategies.
-Agents in those models seem to be rational only to the extent of their capabilities.
-Despite the plausibility of the mechanisms proposed (limited memory, generalization, state confusability), this feels like a bittersweet conclusion because of this hypothetical possibility of a better, hence more rational, strategy, that seems to be only artificially barred from the agents.
-But this is so only if rationality is equated with maximization of expected utility.
-The mechanisms that lead to vague signal use, as~\textcite{oconnor_evolution_2014} and~\textcite{franke_vagueness_2017} stress, have the aforementioned important advantages of faster speed of convergence, higher flexibility, and homogenization.
-These advantages matter, especially in the complex, dynamic context of real language use.
-One would thus like to say that using a strategy that promotes those characteristics without a significant loss of expected utility is certainly a rational choice.
-But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond maximization of expected utility.
-
-For more complex agents, the story is potentially more nuanced.
-Consider an agent that has awareness that something like a game is being played, some estimate of the potential payoff involved, ability to gauge other players' expected behaviors, and the capacity to think strategically.
-Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
-In a scenario where the agent ability to make the optimal choice is bounded, such as the QRE model of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
-\textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
-The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as utility maximization.
-But even in other scenarios, we could again argue that, if we broaden our notion of rationality, it would be less rational to use a crisp strategy rather than one that incorporates at least some source of uncertainty.
-A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
-
-Thus, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.
+We tried to show that, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.
 Furthermore, if we broaden our notion of rationality beyond mere maximization of expected utility, we also find reasons to welcome the positive aspects of these processes and to see their acceptance as a perfectly rational choice.
+Irrational would be to ignore the finite 
 The investigation doesn't end here, though.
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
-
 \printbibliography

### 2017-05-08 book_chapter/paper.tex
Fine-tuned reference games section. Finished conclusion.

@@
 
-% \item We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another;
-% \item These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time.
-
-Linguistic interaction, as captured by a sim-max game, has a rather narrow scope. There is a fixed prior, a fixed utility function and a fixed similarity measure between states. 
-If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t$. 
-Here is a natural example of a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue $t$; the closer Bubu's color choice matches Kiki's idea, the better.
+The type of linguistic interaction captured by a sim-max game has its own particular scope.
+There is a fixed prior, a fixed utility function and a fixed similarity measure between states. 
+If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t'$.
+Here is a natural example of a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue; the closer Bubu's color choice matches Kiki's idea, the better.
 Here Kiki wants to \emph{describe} the color for Bubu to pick out or realize from the whole set of possibilities.
 It's a natural thing to want to do with language. 
-But a lot of communication with vague and contiguously variable properties is rather different. 
-In the general spirit of Wittgenstein's method of arranging multiple language games side by side for a more encompassing picture, we should look at some more cases of language use as well. 
-Let us therefore consider a continuum of cases of referential communication.
-
+But other types of communication with vague and contiguously variable properties are possible. 
+%In the general spirit of Wittgenstein's method of arranging multiple language games side by side for a more encompassing picture, we should look at some more cases of language use as well. 
+%Let us therefore consider a continuum of cases of referential communication.
 A \emph{reference game} is a model of a situation in which the speaker has a particular object in mind and uses a description in order for the receiver to pick this object and no other object. 
@@
 
-The picture sketched in the last paragraph neglects learning from observation and generalization across contexts. 
-For rationality, we need to consider a rational agent's ability to learn from previous experience and to generalize from  the past to the future. 
+The picture just sketched neglects learning from observation, as well as generalization across contexts. 
+For rationality\todo{clarify}, we need to consider a rational agent's ability to learn from previous experience and to generalize from  the past to the future. 
 For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
-But even if every language learner in history had been perfectly rational, it is extremely unlikely that anyone ever succeeded in reducing all uncertainty about language use and say with confidence: this is the threshold that truly separates heaps from non-heaps no matter what.
+But even if every language learner in history had been perfectly rational, it is extremely unlikely, as was discussed in the previous section, that anyone ever succeeded in reducing all uncertainty about language use and say with confidence: this is the threshold that truly separates heaps from non-heaps no matter what.
 For one, there may simply not be enough time in the lifespan of a rational learner to observe enough instances of language use for certainty about crisp denotations. 
@@
 If so, the priors that determine the goodness of a convention for one agent need not be the same as those for another. 
-Going beyond the priors for property degrees, the probability of seeing a particular language game, e.g.~a particular context $c$ in some kind of reference game, is also crucial for determining globally efficient linguistic behavior.
-But here it is even more inconceivable how a rational learner could acquire a fairly confident estimate over how often a particular language game will be played. 
+Going beyond the priors for property degrees, the probability of seeing a particular context $c$ in some kind of reference game, is also crucial for determining globally efficient linguistic behavior.
+But here it is even more inconceivable how a rational learner could acquire a fairly confident estimate over how often the upcoming reference games will have a particular context or another. 
 The list of sources of natural and seemingly insurmountable uncertainty about the parameters necessary to determine locally or globally optimal strategies is much longer than hinted at here. 
-Still, it should be clear enough that a finite existence, a sparse learning input and the inherent fuzziness of the linguistic context itself (its temporal, local and social variability) is not a faul of rationality or the mechanism of selection of successful linguistic practices.
+Still, it should be clear enough that a finite existence, a sparse learning input and the inherent fuzziness of the linguistic context itself (its temporal, local and social variability) is not a fault of rationality on the part of the agent, or the mechanism of selection of successful linguistic practices.
 
 What about meaning and vagueness in a universe of multiple language games and heavy uncertainty?
+We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
+These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time. 
 We could look for meaning at several levels. 
 Firstly, as before, there is the actual behavior of a single agent in each actualized language game. 
-As mentioned above, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
+As mentioned above\todo{where?}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
 Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used, conditional on a given context.
-Notice that a single agent's rational beliefs about linguistic practices or linguistics meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
+Notice that a single agent's rational beliefs about linguistic practices or linguistic meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
 Either one of these two levels, the agent's behavior or linguistic generalizations, could be the foundation for a notion of emerging abstract meaning as constituted by the practices of a population at a particular time.
@@
 Maybe there is a reason.
-The main question to be answered is even more general: given a universe of (rational) language users who learn from finite experience in a changing and stochastically complex arrangement of multiple language games, is a notion of linguistic meaning a theoretically useful and empirically explanatory concept in this picture in such a way that if that notion entailed vague meanings the whole view would have disastrously little explanatory bite?
+The main question\todo{needs more argumentation, if we even want to go there} to be answered is even more general: given a universe of (rational) language users who learn from finite experience in a changing and stochastically complex arrangement of multiple language games, is a notion of linguistic meaning a theoretically useful and empirically explanatory concept in this picture in such a way that if that notion entailed vague meanings the whole view would have disastrously little explanatory bite?
 
@@
 But it suggest that it might be an inflated issue that we need not or should not worry about, because it will not matter for a more general and pressing issue in a more realistic and encompassing picture of language, its use and evolution.
-The following section will expand on this.
 
@@
 We tried to show that, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.
-Furthermore, if we broaden our notion of rationality beyond mere maximization of expected utility, we also find reasons to welcome the positive aspects of these processes and to see their acceptance as a perfectly rational choice.
-Irrational would be to ignore the finite 
+We supported this claim by giving an overview of recent research that tries to accommodate for vagueness in the context of sim-max games, and by discussing, in more general terms, how these considerations are relevant for reference games, and in the broader context of language as a multiplicity of language-games.
+Furthermore, we argued that if we broaden our notion of rationality beyond mere maximization of expected utility, we also find reasons to welcome the positive aspects of these processes and to see their acceptance as a perfectly rational choice.
+Irrational would be to ignore the finite nature of our existence and the constraints that condition language learning and use.
+Our challenge as researchers is of course not only to draw attention to the possibilities, but also to characterize which particular constraints give rise to vague signal use and how.
+
 The investigation doesn't end here, though.
@@
 Thus, for agents capable of adjusting the level of vagueness in their strategies, questions remain regarding how this can be modeled and what factors influence the amount of vagueness that is optimal for a given situation.
-Other challenges include how to incorporate the intuitions about the relevance of the aspects enumerated above into our models.
+Other challenges include how to incorporate into our models the intuitions about the importance of speed of convergence, adaptability, and the promotion of group homogeneity.
+Finally, perhaps the biggest of them all is how to be able to focus on one language-game without losing sight of the ecosystem of games that need to be taken into account as well.
+Whatever solutions we find to these challenges, we believe that a notion of rationality cannot be strongly tied to mere utility maximization in a particular game, but has to consider an agents' need to optimize his behavior to navigate the plurality of uncertainties that besiege him.
 

### 2017-05-10 book_chapter/paper.tex
typos, minor polishing and some inline comments on more juicy matters

@@
 We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness, and how, given vagueness, natural linguistic practices are likely to create more reason for vagueness.
-We explore the consequences of these reasons and mechanism for a notion of meaning and a notion of language.
+We explore the consequences of these reasons and mechanisms for a notion of meaning and a notion of language.
 \end{abstract}
@@
 
-The classic philosophical problem of vagueness is most starkly embodied by the sorites paradox.
+The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
 The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, we can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, we would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
@@
 The problem thus seems to be a wide-ranging one, but it does not obviously undermine our everyday linguistic practices.
+Most of the time we seem to communicate just fine.
 Therefore, the issue must lie with the conceptions we have of the forces or mechanisms that underlie those practices.
@@
 % \end{quote}
-By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~are able to recognize the logical inconsistency but do not have a good answer to overcome it.
+By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~they are able to recognize the logical inconsistency but do not have a good answer as to how to overcome it.
 Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
@@
 
-The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also consequently need to be abandoned down the line.
+The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also need to be abandoned as a consequence.
 One notion that seems to quickly be in peril is that of rationality.
@@
 We want to argue here that giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
-To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore completely irrational.
+To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
 What we need is to revisit the notion of rationality tied to the classical picture.
@@
 We can also focus instead on the consequences, rather than the procedures.
-\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the means available to him were employed in the best way possible that enables him achieve a certain goal.
+\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the available means were employed in the best way possible to achieve a certain goal.
 %This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
@@
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
-Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account
+Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account.
 Thus we can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
@@
 Questions of how strategic they are do not really come to the fore, since the analysis is typically both atemporal and individualistic.
+\mf{I suggest to rewrite this paragraph, keeping as much as possible, but to include a formal definition of ``instrumental rationality'' as maximization of expected utilities, and to also include notions ``ecological rationality'' and ``bounded rationality'' with special-purpose and clearly spelled out definitions, namely roughly like this: where ``instrumental rationality'' requires a single choice to maximize EU in a single local context / a single stage-game, \emph{ecological rationality} requires a general way of choosing (call it, e.g., a \emph{choice mechanism}) to be optimal (max EU) across a whole class of different contexts / games / decision-making situations; a choice mechanism is \emph{bounded rationality} if it is ecologically rational if we also take into account the processing costs necessary to reach a decision (so a choice mechanism that is fast-and-cheap can be better (under bounded rationality) than a highly precise one on average because it saves processing costs, while it would be strictly outperformed in any game/context/decision situation by a more precise mechanism)}
 
-What kind of rationality do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
-What could even such a paradigm be?
+What implications for a notion of rationality of language users do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
+What could such a paradigm even be?
 Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
@@
 The following section introduces the framework of signaling games.
-Then in sections~\ref{sec:sim-max-vagueness} and~\ref{sec:referential-vagueness} we discuss two particular types of games, namely similarity maximization and referential games, and which suggestions there are with regards to how vagueness can be accounted for in those models.
-We bring the insights together in section~\ref{sec:finite-experience} and reflect on the consequences of the proposed natural sources of vagueness for notions of rationality.
+Section~\ref{sec:sim-max-vagueness} looks at explanations for vagueness in a special kind of signaling game, a so-called sim-max game.
+Section~\ref{sec:finite-experience} reflects on the consequences of these explanations for notions of rationality.
+Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game. 
+\mf{Please check if this overview paragraph is appropriate or if you would like to go into a little more detail.}
 
@@
 Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
-Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a set of available signals or messages $M$, a set of responses or actions $A$, and the utility function%
-\footnote{This could also be represented with two separate functions.
-In cases of perfectly aligned preferences, the function can be simplified to yield only one value, to be equally attributed to each player.}
-$U : T \times A \rightarrow \mathbb{R}^2$ which yields a pair of payoff values, one for the sender, one for the receiver, for each possible pairing of state and action.
+Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a probability measure $P$ such that $P(t)$ is the probability of frequency with which $t \in T$ occurs, a set of available signals or messages $M$, a set of responses or actions $A$, and a pair of utility functions $U_{S,R} : T \times A \rightarrow \mathbb{R}$, one for the sender and one for the receiver, each of which yields a payoff value for each possible pairing of state and action.
 These so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
@@
 Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
-The expected utility $EU$ of a pair of strategies $(\sigma,\rho)$ can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions.
-%, \emph{i.e.}:
+The expected utility $EU$ of a strategy can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
+Concretely, the expected utility of $\sigma$ given $\rho$ is $EU_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $EU_S(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
+The expected u %, \emph{i.e.}:
 % $$
@@
 Possible sender and receiver strategies are, for example, $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
-These would have an expected utility of $2$ for both sender and receiver, since when $t_1$ obtains the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, when $t_2$ obtains the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
+These would have an expected utility of $1$ for both sender and receiver, since when $t_1$ obtains with probability $P(t_1)$ the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, and when $t_2$ obtains with probability $P(t_2) = 1-P(t_1)$ the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
 They also represent one of the two stable conventions in this game, the other being the pair of strategies $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
@@
 However, we can already see how it attempts to address the study of meaning from a very different angle.
-There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes.
+There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes (but see \cite{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis} for characterizations of signal meaning in terms of some correspondence with states or acts).
 There is also no appeal to truth as a guide for our inquiry.
-One advantage of such a paradigm shift is that we are no longer tied to the need to explain how language hooks on to the world; this question is no longer relevant on this account.
+One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
 We can merely focus on trying to see how agents can use signals to cope with the world and achieve their purposes.
@@
 
-In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes we study signaling problems in evolutionary terms.
+In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes to study signaling problems in evolutionary terms.
 Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
@@
 Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985}.
-The principles can be captured in a formal model that abstracts away from the interpretations: the replicator dynamics.
-The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
+The principles can be captured in formal models that abstracts away from details of single interactions and individual agent's behavior, \emph{e.g.}~the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
+The only thing relevant to the replicator dynamics are the relative proportions of strategies in a given population and the utility function.
 Using it, one can compute which strategies evolve under which conditions.
@@
 We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
-This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even game-theoretical reasoning.
+This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even recursive strategic reasoning about the co-players' beliefs and choices.
 Each of these can be utilized depending on the problem that one is interested in characterizing.
@@
 A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
-With these tools, not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which don't.
+With these tools, not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which do not.
 In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
@@
 \footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}
-Given two possible messages, the optimal sender strategy is to use one for all values from 40 up to 160, and the other for all values from 160 up to 280.
-The optimal receiver strategy is to guess with 100 if given the first message and with 220 given the second message.
-Confirming Lipman's argument, there is no vagueness in such an optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, univocally associating one guess with each message.
+Given two possible messages, the optimal sender strategy could be to use one message for all values from 40 up to 160, and the other for all values from 160 up to 280; the optimal receiver strategy could be to guess 100 if given the first message and 220 given the second message.
+These precise numerical values give mutually optimal behavior for a case where the prior probability is the same for each state and utility is a linear or quadratic function of the distance between the actual state and the receiver's guess.
+In general, at which point the sender switches the use of messages and which guesses of the receiver are optimal or rational critically hinges on the priors over states and the utility function.
+This will be important later.
+Still, confirming Lipman's argument, there is no vagueness in any such optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, univocally associating one guess with each message.
 
@@
 What is important is that the transition is smooth and monotonous for both players.
-This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is almost certain.
+This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is as good as certain.
 
@@
 For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
-Each agent plays a best response (given the utility function) to their estimate of the other's strategy.
+Each agent plays an instrumentally rational best response (given the utility function) to their estimate of the other's strategy.
 In order to observe the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
@@
 
-In order to overcome this limitation, they make another proposal using the notion of a quantal response equilibrium (QRE).
+In order to overcome this limitation, \citeauthor{franke_vagueness_2011} make another proposal using the notion of a quantal response equilibrium (QRE).
 The idea, inspired by experimental psychology, is to model the choice of best response as stochastic rather than deterministic.\footnote{Probabilistic choice rules are also the source of vagueness in recent accounts by \textcite{LassiterGoodman2015:Adjectival-vagu} and \textcite{QingFranke2014:Gradable-Adject}.}
 A prominent explanation for such soft-max or quantal choice behavior is that agents make small mistakes in the calculation of their expected utilities \parencite{Train2009:Discrete-Choice}. 
-They still choose rationally the option with the highest expected utility, but each assessment of the expected utilities is noise perturbed.
+They still choose by instrumental rationality the option with the highest expected utility, but each assessment of the expected utilities is noise perturbed.
 This, in turn, may actually be optimal behavior if the calculation of expected utilities relies on assessing stochastic uncertainty. 
@@
 The degree to which agents tremble in the calculation of expected utilities and therefore deviate from the optimal behavior can be characterized by a parameter.
-\citeauthor{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria are possible, where sender and receiver simply randomize, respectively, message and response uniformly.
+\citeauthor{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria are possible, where sender and receiver simply randomize, respectively, message and interpretation choice uniformly.
 Above a certain value of the parameter, other equilibria of the kind described in the beginning of this section arise, where agents communicate successfully, though not perfectly, using fuzzy strategies similar to the ones depicted in Figure~\ref{fig:example_stratsB}.
-However, it is not clear that soft-max choices model the right stochastic trembles in decision making as they would arise under natural sources of uncertainty about the context \parencite[see][]{franke_vagueness_2017}.
+However, it is not clear whether soft-max choices capture the right stochastic trembles in decision making as they would arise under natural sources of uncertainty about the context \parencite[see][]{franke_vagueness_2017}.
+\mf{use concept ``bounded rationality'' in this paragraph}
 
@@
 For example, if a sender was successful in using message $m$ for state $t$, he will not only positively reinforce that choice of message for $t$, but also for states similar to $t$.
-This is done in a degree that is proportional to the degree of similarity between $t$ and other states.
+This is done to a degree that is proportional to the similarity between $t$ and other states.
 The dynamics also gives rise to vague signaling of the kind we are looking for.
@@
 In such a scenario, they could drive the other agents to extinction before the latter had time to achieve coordination and reap the benefits of a more precise signaling system.
+\mf{use concept ``ecological rationality'' in this paragraph}
 
@@
 However, in a more realistic setting this is not the case.
-In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utiliry function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of its short-term gains.
+In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utility function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of its short-term gains.
 
@@
 Assuming that agents do not have perfect perception, there will always be a possibility that senders confuse states and receivers mix up responses.
-Furthermore, it seems reasonable to assume that this state confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is they will be mistaken for each other.
+Furthermore, it seems reasonable to assume that this state confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is that they will be mistaken for each other.
 Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of that dynamic that also induces vague signal use of the kind we expect here.
@@
 Furthermore, they observe one additional potentially beneficial property.
-By running several rounds of simulation for each parameter set, they can measure how close resulting strategies are within each group of results, and how they would fair playing against each other.
+By running several rounds of simulation for each parameter set, they can measure how close resulting strategies are within each group of results, and how they would fare playing against each other.
 The results show that the within group distance between strategies becomes smaller with growing confusability, and the within group expected utility is actually higher for strategies evolving under a certain degree of state confusion.
@@
 Better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse
-The game could then be seen as a sim-max game with multiple senders and negative squared difference as similarity function.
+The game could then be seen as a sim-max game with multiple senders and negative squared difference as utility function.
 
-Unlike the models discussed so far that introduce minimal constraints to a dynamic and study the strategies that emerge out of an evolutionary process, these authors' approach is mostly static and analytical.
+Unlike the models discussed so far that introduce minimal constraints to a dynamic and study the strategies that emerge out of an evolutionary process, \citeauthor{lawry_vagueness_2017}' approach is mostly static and analytical.
 They consider several possible sender strategies, with and without uncertainty, and calculate if and when sender and receiver achieve better transmission accuracy, \emph{i.e.}~smaller expected error between original state and receiver estimate.
@@
 The model for the other type of sender is similar, but assumes that there is uncertainty regarding the value of the threshold, representing it as a uniformly distributed random variable.
-What results from this in terms of sender strategy is also a smooth transition between the use of two messages%
+What results from this in terms of sender strategy is also a smooth transition between the use of two messages.%
 \footnote{Unlike what we see in Figure~\ref{fig:example_stratsB}, it is a linear transition.
@@
 We speculate that another choice of distribution, such as for example a normal distribution, could yield something similar to Figure~\ref{fig:example_stratsB}.}%
-.
 The receiver strategy is not inherently vague for either type of sender, it merely aggregates the messages received to provide an estimate of the original state using a simple arithmetic mean.
@@
 The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the idealist\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
-In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (or better memory) would lead to more precision.
+In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (better memory, or more reliable computation of expected utilities) would lead to more precision.
 \citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{gilboa_social_1991}.
@@
 \item Our existence is temporally finite; language does not have an infinite amount of time to evolve, nor can it take an infinite time to be learned.
-The faster a language can start being useful, the better;
+The faster a language can start being useful, the better.
 \item Language learning through experience has to rely on a limited number of observations.
-Not only is the state space typically much larger than what one can survey in sufficient time, it is even potentially infinite and constantly changing;
+Not only is the state space typically much larger than what one can survey in sufficient time, it is even potentially infinite and constantly changing.
 \item A corollary of the former is that there will always be heterogeneity in a population of language learners, at the very least in their prior experience, since each agent will have relied on a different set of observations.
-Furthermore, this information is not directly or fully accessible to others;
-\item Given that an agent is almost always learning and using language in a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice;
+Furthermore, this information is not directly or fully accessible to others.
+\item Given that an agent is almost always learning and using language in a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice.
 \end{enumerate}
@@
 But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond maximization of expected utility, or a notion of expected utility that incorporates other dimensions.
+\mf{reformulate last bit of paragraph using notion ``ecological rationality''?}
 
@@
 But other types of communication with vague and contiguously variable properties are possible. 
-%In the general spirit of Wittgenstein's method of arranging multiple language games side by side for a more encompassing picture, we should look at some more cases of language use as well. 
-%Let us therefore consider a continuum of cases of referential communication.
 A \emph{reference game} is a model of a situation in which the speaker has a particular object in mind and uses a description in order for the receiver to pick this object and no other object. 
-One crucial difference to descriptive communication and sim-max games is that close will not be close enough: if Kiki needs a long nail, a nail a little longer or shorter may be good enough; but if Bubu needs the long, slightly bend nail with the flat top (which she previously disassembled from her art installation), no other nail will do.
+One crucial difference with sim-max games is that close will not be close enough: if Kiki needs a long nail, a nail a little longer or shorter may be good enough; but if Bubu needs the long, slightly bend nail with the flat top (which she previously disassembled from her art installation), no other nail will do.
 
@@
 In \emph{reference games with partial uncertainty}, for instance, interlocutors may know that they have shared and immediate access only to a subset of the objects in the context, but be uncertain about whether the co-player sees any other objects.
-This is the set-up for the classic director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
+This is the set-up for the widely-used director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
 
@@
 Likewise, the receiver chooses a referent $t'$ from the available objects in $c$, again conditional on $c$. 
-But if agents play many such reference games with different contexts $c$ and $c'$, and if strategic rationality means to signal and interpret so as to coordinate on the correct referent, strategic rationality of local choices in $c$ and $c'$ is not enough to guarantee that agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
+But if agents play many such reference games with different contexts $c$ and $c'$, strategic and instrumental rationality of local choices in $c$ and $c'$ is not enough to guarantee that agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
 An agent could use a certain signal to denote, say, the tallest object in every context with between five and twelve objects or when the number of objects is a prime number, but use the same signal for the darkest object in all other contexts.
-This could even be rational if most agents do it this way.
-It could evolve as a convention under selection for communicative success as well.
-In other words, neither local strategic rationality in a fixed context nor evolutionary adaptation to each single context entails a natural, globally consistent behavior across multiple contexts. 
+In other words, neither rationality of choice in a fixed context nor evolutionary adaptation to each single context entails a natural, globally consistent behavior across multiple contexts. 
 The situation only gets worse if we want to have agents use the same expressions in a consistent way also across different types of reference games, with possibly different epistemic conditions.
@@
 The picture just sketched neglects learning from observation, as well as generalization across contexts. 
-For rationality\todo{clarify}, we need to consider a rational agent's ability to learn from previous experience and to generalize from  the past to the future. 
+For strategically and instrumentally rational agents, we need to consider also the ability to learn from previous experience and to generalize from the past to the future. 
 For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
@@
 If the prior captures real world attributes, such as the probability of an individual's body height, these attributes will not only be hard to know for certain, but they will also vary over time. 
-They may also depend on location: do we use our language exclusively in this village or do we travel beyond? 
+They may also depend on location: do we use our language exclusively in this village (where all heaps are extraordinarily small and people are extraordinarily tall) or do we travel beyond? 
 If so, the priors that determine the goodness of a convention for one agent need not be the same as those for another. 
@@
 These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time. 
-We could look for meaning at several levels. 
+We could look for meaning at several levels in this pluralistic picture. 
 Firstly, as before, there is the actual behavior of a single agent in each actualized language game. 
-As mentioned above\todo{where?}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic (with the caveat that in the real world we would never know because we cannot ever elicit a production in the exact same circumstances in a constantly changing environment including the mind of a speaker).
-Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used, conditional on a given context.
+As mentioned above in connection the soft-max choice function used by \textcite{franke_vagueness_2011}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic.
+\mf{mention ``bounded rationality'' here}
+Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
 Notice that a single agent's rational beliefs about linguistic practices or linguistic meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
-Either one of these two levels, the agent's behavior or linguistic generalizations, could be the foundation for a notion of emerging abstract meaning as constituted by the practices of a population at a particular time.
-If both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp? 
-Maybe there is a reason.
-The main question\todo{needs more argumentation, if we even want to go there} to be answered is even more general: given a universe of (rational) language users who learn from finite experience in a changing and stochastically complex arrangement of multiple language games, is a notion of linguistic meaning a theoretically useful and empirically explanatory concept in this picture in such a way that if that notion entailed vague meanings the whole view would have disastrously little explanatory bite?
+In sum, both at the level of behavior and at the level of beliefs about use or meaning, we should expect to find vagueness.
+Still, despite the natural vagueness, there does not seem to anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
 
-Taken together, when we look at the picture of multiple language games side by side, the question of what rational or optimal behavior is in any one of them, or the question of why human language users do not more closely approximate rational or optimal local behavior more closely seem entirely misplaced.
+Finally, we might be interested in a notion of meaning of language as an abstraction over their users and its use.
+We might not care about individual agents' behavior and beliefs as much as we care about the meaning of words, expressions and sentences.
+The questions to address then are probably: what explanatory purpose is a notion of meaning of an abstracted language to play?; if both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp?; what can be explained only by a notion of meaning that is not vague?
+\mf{I leave this paragraph in for you to read and ponder, but would recommend to leave it out entirely}
+
+Taken together, when we look at the picture of multiple language games side by side, the question of what rational or optimal behavior is in any one of them, or the question of why human language users do not more closely approximate instrumentally rational or optimal local behavior more closely seem entirely misplaced.
 What seems more pressing is to answer how the interaction of individual learning, individual local optimization and general pressure for successful communication at a cross-situational level lead to the shaping of linguistic practices in an open-ended landscape of language games.
@@
 Adopting this perspective does not \emph{solve} the problem of how vagueness is compatible with rationality. 
-But it suggest that it might be an inflated issue that we need not or should not worry about, because it will not matter for a more general and pressing issue in a more realistic and encompassing picture of language, its use and evolution.
+But it suggest that it might be an inflated issue that we need not or should not worry about, because it will likely not matter in a more realistic and encompassing picture of language, its use and evolution as driven by an appropriate notion of rational choice and evolutionary selection.
 

### 2017-05-11 book_chapter/paper.tex
extension paragraph on rationality

@@
 
-Although frequently debated, especially in areas like philosophy, psychology, and economics, rationality is a notion for which clear definitions are hard to come by.
+Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
 Discussions around it usually rely on common connotations and implicit assumptions.
@@
 \emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the available means were employed in the best way possible to achieve a certain goal.
-%This is close to the notion of rationality that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility.
 The notion is linked to David Hume~\parencite*{hume_treatise_1738} and his assertion that ``Reason is, and ought to only be the slave of the passions.''
-Another assumption that is important to qualify when talking about rationality of an agent's choice is the amount of information that the agent has access to, when making his decision.
+This is close to a notion of rational choice that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility. 
+Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.
+
+Choice rationality is not all there is to rationality though.
+An agent's mental representation of the decision situation is also subject to considerations of rationality. 
 Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices can vary.
@@
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
-Finally, the third assumption we will draw attention to, for now, regards the ability of an agent to take predictions into account.
-Thus we can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate and consider either the actions of other agents, or considerations of medium/long term gains.
+In many cases, agents may impossibly have all decision relevant information.
+This can happen when an agent has made only a rather small number of observations, e.g., about how a particular word is used.
+But even with \emph{partial information}, an agent can be required to have rational beliefs, i.e., to use all of the information available in the best possible way.
+This leads back to procedural rationality, logical inference and inductive learning.
+It is very hard to say what exactly rational learning is.
+It is easier to gesture at what it is not: ignoring observational evidence; failing to draw simple and obvious conclusions; failing to make simple and obvious generalizations etc.
+A further aspect of rational belief formation is very important for our present concerns, namely the ability of an agent to make accurate predictions about how other agents behave.
+We can say that an agent is more or less \emph{strategically rational} depending on the extent to which he is able to anticipate and consider either the actions and beliefs of other agents, or considerations of medium/long term gains.
+Strategic rationality is important for language use, because it is natural to assume that an agent's believes about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use a said expression.
+
+We draw attention to two additional aspects of rationality which will be helpful in the following.
+Where the above definition of instrumental rationality is usually understood as requiring a single choice of act $a$ to maximize the expected utility in a single concrete decision situation, we might also be interested more generally not in specific single local choices, so to speak, but in general global choice mechanisms.
+A choice mechanism in thes sense would be a general way of behaving, \emph{e.g.} a function from choice situations or contexts to single choices in each context.
+A notion of \emph{ecological rationality} requires a \emph{choice mechanism} to be optimal in the sense of expected utility maximization, across a whole class of different decision situations or contexts, weighed by their probability of frequency of occurrence.
+To make this a bit more transparent, consider the following rough formalization.
+If a context $c$ from some set $C$ gives us a decision problem with action set $A_c$, states $T_c$, probabilities of states $P_c \in \Delta(T)$ and utility function $u_c$, then a choice mechanism would be a function $CM \colon C \rightarrow \bigcup_{c \in C} A_c$.
+A choice mechanism $CM$ would be ecologically rational if it maximizes the
+
 The classical picture is, in these terms, tied to a procedural account of rationality, where agents are pretty much omniscient.
 Questions of how strategic they are do not really come to the fore, since the analysis is typically both atemporal and individualistic.
+
 \mf{I suggest to rewrite this paragraph, keeping as much as possible, but to include a formal definition of ``instrumental rationality'' as maximization of expected utilities, and to also include notions ``ecological rationality'' and ``bounded rationality'' with special-purpose and clearly spelled out definitions, namely roughly like this: where ``instrumental rationality'' requires a single choice to maximize EU in a single local context / a single stage-game, \emph{ecological rationality} requires a general way of choosing (call it, e.g., a \emph{choice mechanism}) to be optimal (max EU) across a whole class of different contexts / games / decision-making situations; a choice mechanism is \emph{bounded rationality} if it is ecologically rational if we also take into account the processing costs necessary to reach a decision (so a choice mechanism that is fast-and-cheap can be better (under bounded rationality) than a highly precise one on average because it saves processing costs, while it would be strictly outperformed in any game/context/decision situation by a more precise mechanism)}

### 2017-05-11 book_chapter/paper.tex
section on rationality

@@
 In this paper, we investigate a number of reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary optimal signaling.
-We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness, and how, given vagueness, natural linguistic practices are likely to create more reason for vagueness.
-We explore the consequences of these reasons and mechanisms for a notion of meaning and a notion of language.
+We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness.
+We explore the consequences of these reasons and mechanisms for a notion of meaning and a notion of rationality.
 \end{abstract}
@@
 This type of rationality is one that supposedly demarcates humans from other animals.
+
 We want to argue here that giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
+It also does not necessarily eradicate all hope for a (formal) explication of meaning.
 To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
-What we need is to revisit the notion of rationality tied to the classical picture.
+To support this view, we need to investigate different aspects of rationality (which is what the next section will do) and see how these relate to different ways of thinking about language, use and meaning (which is what the rest of the paper will do).
+
+\section{Facets of rationality and a different picture}
+\label{sec:facets-rationality}
 
@@
 It attempts to characterize a particular choice of an agent as rational based on the procedures used to select it.
+It is intimately related to logically valid deductive inference.  
 We can also focus instead on the consequences, rather than the procedures.
@@
 Choice rationality is not all there is to rationality though.
-An agent's mental representation of the decision situation is also subject to considerations of rationality. 
+An agent's mental representation of the decision situation can also be subject to considerations of rationality. 
 Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices can vary.
@@
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
-In many cases, agents may impossibly have all decision relevant information.
+In many cases, it is simply impossible for an agent to own all decision relevant information.
 This can happen when an agent has made only a rather small number of observations, e.g., about how a particular word is used.
-But even with \emph{partial information}, an agent can be required to have rational beliefs, i.e., to use all of the information available in the best possible way.
+Having only partial information is not irrational.
+But even with partial information, an agent can be required to have rational beliefs, i.e., to use all of the information available in the best possible way.
 This leads back to procedural rationality, logical inference and inductive learning.
@@
 A further aspect of rational belief formation is very important for our present concerns, namely the ability of an agent to make accurate predictions about how other agents behave.
-We can say that an agent is more or less \emph{strategically rational} depending on the extent to which he is able to anticipate and consider either the actions and beliefs of other agents, or considerations of medium/long term gains.
-Strategic rationality is important for language use, because it is natural to assume that an agent's believes about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use a said expression.
+We can say that an agent is more or less \emph{strategically rational} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
+Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.
 
 We draw attention to two additional aspects of rationality which will be helpful in the following.
-Where the above definition of instrumental rationality is usually understood as requiring a single choice of act $a$ to maximize the expected utility in a single concrete decision situation, we might also be interested more generally not in specific single local choices, so to speak, but in general global choice mechanisms.
-A choice mechanism in thes sense would be a general way of behaving, \emph{e.g.} a function from choice situations or contexts to single choices in each context.
-A notion of \emph{ecological rationality} requires a \emph{choice mechanism} to be optimal in the sense of expected utility maximization, across a whole class of different decision situations or contexts, weighed by their probability of frequency of occurrence.
-To make this a bit more transparent, consider the following rough formalization.
-If a context $c$ from some set $C$ gives us a decision problem with action set $A_c$, states $T_c$, probabilities of states $P_c \in \Delta(T)$ and utility function $u_c$, then a choice mechanism would be a function $CM \colon C \rightarrow \bigcup_{c \in C} A_c$.
-A choice mechanism $CM$ would be ecologically rational if it maximizes the
+Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general \emph{choice mechanisms} \parencite[\emph{e.g.}]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
+A choice mechanism is a general way of behaving.
+Think of it as a function from contexts (choice situations, decision problems, different language games) to action choices in each context.
+A notion of \emph{ecological rationality} requires a choice mechanism to be optimal, in the sense of expected utility maximization, across the whole class $C$ of potential contexts, weighed by their probability or frequency of occurrence $P \in \Delta(C)$.
+To make this more transparent, consider the following rough formalization.
+A context $c \in C$ gives us a decision problem with action set $A_c$, states $T_c$, probabilities of states $P_c \in \Delta(T_c)$ and utility function $u_c$.
+A choice mechanism is a function $CM \colon c \mapsto CM(c) \in A_c$. 
+Choice mechanism $CM$ is ecologically rational if it maximizes the expected utility $EU(CM) = \int \sum_{t \in T_c} P_c(t) \, u_c(CM(c) , t) \ \text{d} P(c)$.
+A $CM$ is ecologically rational in this sense only if it delivers instrumentally rational choices in each local context.
+But ecological rationality trades off with requirements on rationality of beliefs (captured here roughly by $P \in \Delta(C)$ and $P_c \in \Delta(T_c)$).
+Especially when agents do not face a static, unchanging environment, such as when decision relevant beliefs hinge on the behavior of other agents, it may simply be impossible to achieve perfect ecological rationality.
+Moreover, not every choice mechanism is as natural as another. 
+Some may be easier to learn and apply; some may rely on imprecise but useful generalizations across contexts, others may not.
+This leads to a related notion of rationality, which we will call \emph{bounded rationality} here.
+This notion takes into account that the application of a choice mechanism may be more or less costly, \emph{e.g.}~how much time and effort does the agent invest into approximating or precisely calculating decision relevant beliefs and preferences?
+In this way, it is conceivable that a given choice mechanism returns many good but still sub-optimal action choices in many contexts from the point of view of locally instrumental rationality.
+At the same time, this choice mechanism, though locally sub-optimal, may outperform a strictly ecologically rational mechanism if the former is easier to compute, especially with an eye to generalization to novel contexts where precise computation of all decision relevant factors might be difficult.
+In a nutshel, bounded rationality integrates performance factors into what should count as optimal or rational behavior. 
+It is an alternative to instrumental and ecological rationality.
+It becomes particularly relevant when we consider an open class of decision situations, like we do for ecological rationality, and also consider rationality of beliefs and strategic multi-player considerations in a non-static environment.
 
-The classical picture is, in these terms, tied to a procedural account of rationality, where agents are pretty much omniscient.
-Questions of how strategic they are do not really come to the fore, since the analysis is typically both atemporal and individualistic.
-
-\mf{I suggest to rewrite this paragraph, keeping as much as possible, but to include a formal definition of ``instrumental rationality'' as maximization of expected utilities, and to also include notions ``ecological rationality'' and ``bounded rationality'' with special-purpose and clearly spelled out definitions, namely roughly like this: where ``instrumental rationality'' requires a single choice to maximize EU in a single local context / a single stage-game, \emph{ecological rationality} requires a general way of choosing (call it, e.g., a \emph{choice mechanism}) to be optimal (max EU) across a whole class of different contexts / games / decision-making situations; a choice mechanism is \emph{bounded rationality} if it is ecologically rational if we also take into account the processing costs necessary to reach a decision (so a choice mechanism that is fast-and-cheap can be better (under bounded rationality) than a highly precise one on average because it saves processing costs, while it would be strictly outperformed in any game/context/decision situation by a more precise mechanism)}
-
-What implications for a notion of rationality of language users do we get if we try to step out of the classical picture and start looking at meaning in terms of a different paradigm?
+Within this landscape of aspects of rationality, it seems that the classical picture of meaning is intimately related to a procedural account of rationality, where agents are pretty much omniscient.
+Logically valid inference is important.
+The sorites paradox is a real threat.
+Questions of how language is used or in which sense it is used rationally do not really come to the fore, since the analysis is typically both atemporal and abstracts away from individual language users.
+But, what other aspects of rationality of language users are important when we step out of the classical picture and start looking at meaning in terms of a different paradigm?
 What could such a paradigm even be?
@@
 Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
-\mf{Please check if this overview paragraph is appropriate or if you would like to go into a little more detail.}
 
@@
 However, we can already see how it attempts to address the study of meaning from a very different angle.
-There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes (but see \cite{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis} for characterizations of signal meaning in terms of some correspondence with states or acts).
+There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes \parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
 There is also no appeal to truth as a guide for our inquiry.
@@
 % \subsection{Bounded rationality}
-\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can arise as a consequence of limitations in the cognitive capacities of instrumentally rational agents.
+\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can be boundedly rational, \emph{i.e.}~how vagueness could arise as a consequence of cost-saving limitations in the cognitive capacities of instrumentally rational agents.
 The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
@@
 A prominent explanation for such soft-max or quantal choice behavior is that agents make small mistakes in the calculation of their expected utilities \parencite{Train2009:Discrete-Choice}. 
-They still choose by instrumental rationality the option with the highest expected utility, but each assessment of the expected utilities is noise perturbed.
-This, in turn, may actually be optimal behavior if the calculation of expected utilities relies on assessing stochastic uncertainty. 
+They still choose the option with the highest expected utility, but each assessment of the expected utilities is noise perturbed.
+This, in turn, may actually be boundedly rational since the calculation of expected utilities relies on assessing stochastic uncertainty, which in turn may be costly to calculate precisely. 
 Choice based on a few samples from a distribution can be optimal if taking more samples or other means of better approximating degrees of uncertainty is resource costly \parencite[\emph{e.g.}][]{VulGoodman2014:One-and-Done-Op,SanbornChater2016:Bayesian-Brains}.
-The degree to which agents tremble in the calculation of expected utilities and therefore deviate from the optimal behavior can be characterized by a parameter.
+The degree to which agents tremble in the calculation of expected utilities and therefore deviate from the instrumentally rational behavior can be characterized by a parameter.
 \citeauthor{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria are possible, where sender and receiver simply randomize, respectively, message and interpretation choice uniformly.
@@
 However, it is not clear whether soft-max choices capture the right stochastic trembles in decision making as they would arise under natural sources of uncertainty about the context \parencite[see][]{franke_vagueness_2017}.
-\mf{use concept ``bounded rationality'' in this paragraph}
 
@@
 However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a smaller period of time.
-In a naturalized evolutionary setting this temporary advantage should also be taken into account.
+In a naturalized evolutionary setting, where we care about bounded rationality in a multi-context environment, learning speed is an advantage which should also be taken into account.
 Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.
@@
 In such a scenario, they could drive the other agents to extinction before the latter had time to achieve coordination and reap the benefits of a more precise signaling system.
-\mf{use concept ``ecological rationality'' in this paragraph}
 
@@
 Thus, some amount of uncertainty seems to promote more homogeneous populations of signalers that are better at achieving cooperation within a group.
+In this picture, vagueness is the natural and unavoidable consequence of perceptual limitations, but may still contribute to the overall boundedly rational character of language if we take multiple contexts of learning into account.
 
@@
 \end{quote}
-This ``economic man'', a game-omniscient and highly strategic rational agent, seems to have survived almost unscathed to now populate most game theoretic models of language and signaling.
+This ``economic man'', a game-omniscient and highly strategic, instrumentally rational agent, seems to have survived almost unscathed to now populate most game theoretic models of language and signaling.
 Both Simon and Lipman call for this picture to be revisited.
@@
 % Another possible interpretation is that of a population of agents learning by imitation.
-The LMF dynamics of~\textcite{franke_vagueness_2011}, generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, all paint a picture of agents with a basic level of instrumental rationality, very limited awareness of the game, and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
+The models of~\textcite{franke_vagueness_2011}, generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, all paint a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
 Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
-All approaches introduce constraints into the dynamics that prevent the evolution of crisp signal use.
+All approaches introduce constraints on agent behavior or information processing that prevent the evolution of crisp signal use.
 But a crisp language would still have a higher expected utility than the evolved strategies.
 Agents in those models seem to be rational only to the extent of their capabilities.
-Despite the plausibility of the mechanisms proposed (limited memory, generalization, state confusability), this feels like a bittersweet conclusion because of this hypothetical possibility of an ideal strategy that seems to be only artificially barred from the agents.
+Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), this feels like a bittersweet conclusion because of this hypothetical possibility of an ideal strategy that seems to be only artificially barred from the agents.
 But this is so only if rationality is equated with maximization of expected utility.
@@
 These advantages matter, especially in the complex, dynamic context of real language use.
+Limits on memory and precision in the calculation of complex probabilistic beliefs may have a reduced metabolic processing cost.
 One would thus like to say that using a strategy that promotes those characteristics without a significant loss of expected utility is certainly a rational choice.
 But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond maximization of expected utility, or a notion of expected utility that incorporates other dimensions.
-\mf{reformulate last bit of paragraph using notion ``ecological rationality''?}
+It is in this sense that these explanations tie in with a notion of bounded ecological rationality.
 

### 2017-05-11 book_chapter/paper.tex
section on rationality

@@
 
-Finally, we might be interested in a notion of meaning of language as an abstraction over their users and its use.
-We might not care about individual agents' behavior and beliefs as much as we care about the meaning of words, expressions and sentences.
-The questions to address then are probably: what explanatory purpose is a notion of meaning of an abstracted language to play?; if both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp?; what can be explained only by a notion of meaning that is not vague?
-\mf{I leave this paragraph in for you to read and ponder, but would recommend to leave it out entirely}
+% Finally, we might be interested in a notion of meaning of language as an abstraction over their users and its use.
+% We might not care about individual agents' behavior and beliefs as much as we care about the meaning of words, expressions and sentences.
+% The questions to address then are probably: what explanatory purpose is a notion of meaning of an abstracted language to play?; if both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp?; what can be explained only by a notion of meaning that is not vague?
+% \mf{I leave this paragraph in for you to read and ponder, but would recommend to leave it out entirely}
 

### 2017-05-12 book_chapter/paper.tex
Moved talk of Wittgenstein and signaling games to Section 1.

@@
 
-We want to argue here that giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
+We want to argue here that to say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
+Giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
 It also does not necessarily eradicate all hope for a (formal) explication of meaning.
-To say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
-To support this view, we need to investigate different aspects of rationality (which is what the next section will do) and see how these relate to different ways of thinking about language, use and meaning (which is what the rest of the paper will do).
+In order to make this argument, we need a different paradigm of language and meaning.
+But what could such a paradigm even be?
 
-\section{Facets of rationality and a different picture}
+We suggest that inspiration can be drawn from the later work of philosopher Ludwig Wittgenstein%
+\footnote{To the extent that substantial views can be said to be defended by the author.
+Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
+His \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} adumbrate a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
+Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
+These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
+%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
+There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
+The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
+Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
+It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
+From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.
+
+Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
+We start, in Section~\ref{sec:facets-rationality}, by clarifying some concepts that can help characterize different notions of rationality.
+In Section~\ref{sec:signaling-games} we introduce the framework of signaling games and in Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a special kind of signaling game, a so-called sim-max game.
+Section~\ref{sec:finite-experience} reflects on the implications of these explanations for notions of rationality.
+Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
+
+\section{Facets of rationality}
 \label{sec:facets-rationality}
@@
 Questions of how language is used or in which sense it is used rationally do not really come to the fore, since the analysis is typically both atemporal and abstracts away from individual language users.
-But, what other aspects of rationality of language users are important when we step out of the classical picture and start looking at meaning in terms of a different paradigm?
-What could such a paradigm even be?
-Perhaps we can draw inspiration from the later work of philosopher Ludwig Wittgenstein%
-\footnote{To the extent that substantial views can be said to be defended by the author.
-Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
-Most starkly in the \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953}, we can find a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
-Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
-These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
-%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
-There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
-The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
-Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
-It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
-From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.
-
-Following this idea, we here survey proposals for addressing vagueness in the context of signaling games.
-The following section introduces the framework of signaling games.
-Section~\ref{sec:sim-max-vagueness} looks at explanations for vagueness in a special kind of signaling game, a so-called sim-max game.
-Section~\ref{sec:finite-experience} reflects on the consequences of these explanations for notions of rationality.
-Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
+Let us know look into signaling games as another framework for the study of language use and meaning and see what account of rationality is characteristic of that approach.
 

### 2017-05-12 book_chapter/paper.tex
Changed Section 2 to keep discussion at an abstract level. Commented out parts that I think fit better later in the paper.

@@
 Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
-Discussions around it usually rely on common connotations and implicit assumptions.
-The aforementioned classical picture of language and meaning is focused on an aspect of rationality that we can call \emph{procedural}.
-It attempts to characterize a particular choice of an agent as rational based on the procedures used to select it.
-It is intimately related to logically valid deductive inference.  
-We can also focus instead on the consequences, rather than the procedures.
-\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if, in making that choice, the available means were employed in the best way possible to achieve a certain goal.
-The notion is linked to David Hume~\parencite*{hume_treatise_1738} and his assertion that ``Reason is, and ought to only be the slave of the passions.''
-This is close to a notion of rational choice that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility. 
-Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.
+Discussions around it, although touching on different aspects of the concept, often rely on a number of implicit assumptions.
+We can start by considering what is deemed to be most important when characterizing the rationality of a given choice.
+The aforementioned classical picture of language and meaning is focused on procedures.
+Thus, a choice is deemed rational if it was arrived at by means of following proper rules of reasoning, like for example a logically valid deductive inference.
+A \emph{procedural} account of rationality focuses on the means, rather than the ends.
+One can also do the opposite and focus on the consequences of the choice instead.
+\emph{Instrumental} rationality, as it is typically called\todo{references}, characterizes an agent's choice as rational if it maximizes the possibility of achieving a desired goal, regardless of the means.
+The notion is linked to David Hume~\parencite*{hume_treatise_1738} and epitomized in his assertion that ``Reason is, and ought to only be the slave of the passions.''
+This is close to a notion of rational choice that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility%
+\footnote{\todo[inline]{discuss other options like minimax and such}}%
+.
+A more in-depth discussion of the opposition in the context of theoretical economics can be found, for example, in the work of Herbert A.~Simon~\parencite*[\emph{e.g.}][]{simon_rationality_1986}%
+\footnote{Simon uses the term substantive instead of instrumental rationality, but the characterization is basically the same~\parencite[210-212]{simon_rationality_1986}.}%
+.
+%Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}
 
-Choice rationality is not all there is to rationality though.
-An agent's mental representation of the decision situation can also be subject to considerations of rationality. 
-Depending on how accurate and complete the agent's knowledge of its environment, the objectives to be achieved, the choices available, the relation between those choices and the objectives, etc, the verdict over the rationality of his choices can vary.
-This is difficult to characterize in general, but, in the context of a given model, we can describe agents with respect to how much knowledge they have of the ingredients of the model.
+When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between whatever agents are involved and the elements of the model can come to bear on considerations of rationality.
+Depending on how accurate and complete the agent's knowledge of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, etc, the verdict over the rationality of certain choices can vary.
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
-In many cases, it is simply impossible for an agent to own all decision relevant information.
-This can happen when an agent has made only a rather small number of observations, e.g., about how a particular word is used.
-Having only partial information is not irrational.
-But even with partial information, an agent can be required to have rational beliefs, i.e., to use all of the information available in the best possible way.
-This leads back to procedural rationality, logical inference and inductive learning.
-It is very hard to say what exactly rational learning is.
-It is easier to gesture at what it is not: ignoring observational evidence; failing to draw simple and obvious conclusions; failing to make simple and obvious generalizations etc.
-A further aspect of rational belief formation is very important for our present concerns, namely the ability of an agent to make accurate predictions about how other agents behave.
-We can say that an agent is more or less \emph{strategically rational} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
-Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.
+Models working within the classical picture typically do not make a distinction between modeler and agent, thus there is no room to express any epistemic gaps.
+It follows that agents can only be assumed to be omniscient of the details involved.
+By abstracting away from language users, these models cannot represent potential interactions between them, let alone repeated interaction and language change.
+In other types of models, however, a further aspect of rational belief formation needs to be considered, namely the ability of agents to make accurate predictions about how other agents behave.
+We can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
+%Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.\todo{too soon}
 
-We draw attention to two additional aspects of rationality which will be helpful in the following.
-Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general \emph{choice mechanisms} \parencite[\emph{e.g.}]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
-A choice mechanism is a general way of behaving.
-Think of it as a function from contexts (choice situations, decision problems, different language games) to action choices in each context.
-A notion of \emph{ecological rationality} requires a choice mechanism to be optimal, in the sense of expected utility maximization, across the whole class $C$ of potential contexts, weighed by their probability or frequency of occurrence $P \in \Delta(C)$.
-To make this more transparent, consider the following rough formalization.
-A context $c \in C$ gives us a decision problem with action set $A_c$, states $T_c$, probabilities of states $P_c \in \Delta(T_c)$ and utility function $u_c$.
-A choice mechanism is a function $CM \colon c \mapsto CM(c) \in A_c$. 
-Choice mechanism $CM$ is ecologically rational if it maximizes the expected utility $EU(CM) = \int \sum_{t \in T_c} P_c(t) \, u_c(CM(c) , t) \ \text{d} P(c)$.
-A $CM$ is ecologically rational in this sense only if it delivers instrumentally rational choices in each local context.
-But ecological rationality trades off with requirements on rationality of beliefs (captured here roughly by $P \in \Delta(C)$ and $P_c \in \Delta(T_c)$).
-Especially when agents do not face a static, unchanging environment, such as when decision relevant beliefs hinge on the behavior of other agents, it may simply be impossible to achieve perfect ecological rationality.
-Moreover, not every choice mechanism is as natural as another. 
-Some may be easier to learn and apply; some may rely on imprecise but useful generalizations across contexts, others may not.
-This leads to a related notion of rationality, which we will call \emph{bounded rationality} here.
-This notion takes into account that the application of a choice mechanism may be more or less costly, \emph{e.g.}~how much time and effort does the agent invest into approximating or precisely calculating decision relevant beliefs and preferences?
-In this way, it is conceivable that a given choice mechanism returns many good but still sub-optimal action choices in many contexts from the point of view of locally instrumental rationality.
-At the same time, this choice mechanism, though locally sub-optimal, may outperform a strictly ecologically rational mechanism if the former is easier to compute, especially with an eye to generalization to novel contexts where precise computation of all decision relevant factors might be difficult.
-In a nutshel, bounded rationality integrates performance factors into what should count as optimal or rational behavior. 
-It is an alternative to instrumental and ecological rationality.
-It becomes particularly relevant when we consider an open class of decision situations, like we do for ecological rationality, and also consider rationality of beliefs and strategic multi-player considerations in a non-static environment.
+% \todo[inline]{Following discussion is already drawing conclusions, and is too closely tied to notions that haven't been introduced yet}
+% In many cases, it is simply impossible for an agent to own all decision relevant information.
+% This can happen when an agent has made only a rather small number of observations, e.g., about how a particular word is used.
+% Having only partial information is not irrational.
+% But even with partial information, an agent can be required to have rational beliefs, i.e., to use all of the information available in the best possible way.
+% This leads back to procedural rationality, logical inference and inductive learning.
+% It is very hard to say what exactly rational learning is.
+% It is easier to gesture at what it is not: ignoring observational evidence; failing to draw simple and obvious conclusions; failing to make simple and obvious generalizations etc.
 
-Within this landscape of aspects of rationality, it seems that the classical picture of meaning is intimately related to a procedural account of rationality, where agents are pretty much omniscient.
-Logically valid inference is important.
-The sorites paradox is a real threat.
-Questions of how language is used or in which sense it is used rationally do not really come to the fore, since the analysis is typically both atemporal and abstracts away from individual language users.
-Let us know look into signaling games as another framework for the study of language use and meaning and see what account of rationality is characteristic of that approach.
+%We draw attention to two additional aspects of rationality which will be helpful in the following.
+Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
+A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
+When considering only a single situation, rationality can only be \emph{local}.
+If we take into account the possibility of the agent's choice or choice mechanism hinging on multiple situations, we can also talk about \emph{global} or \emph{ecological} rationality.
+This is important because, hypothetically, there could be choice mechanisms that are sub-optimal at a local level (for each situation) but are actually perfectly rational at a global level.
+
+% \todo[inline]{Formalization narrows down the concepts too much}
+% Think of it as a function from contexts (choice situations, decision problems, different language games) to action choices in each context.
+% A notion of \emph{ecological rationality} requires a choice mechanism to be optimal, in the sense of expected utility maximization, across the whole class $C$ of potential contexts, weighed by their probability or frequency of occurrence $P \in \Delta(C)$.
+% To make this more transparent, consider the following rough formalization.
+% A context $c \in C$ gives us a decision problem with action set $A_c$, states $T_c$, probabilities of states $P_c \in \Delta(T_c)$ and utility function $u_c$.
+% A choice mechanism is a function $CM \colon c \mapsto CM(c) \in A_c$. 
+% Choice mechanism $CM$ is ecologically rational if it maximizes the expected utility $EU(CM) = \int \sum_{t \in T_c} P_c(t) \, u_c(CM(c) , t) \ \text{d} P(c)$.
+% A $CM$ is ecologically rational in this sense only if it delivers instrumentally rational choices in each local context.
+% But ecological rationality trades off with requirements on rationality of beliefs (captured here roughly by $P \in \Delta(C)$ and $P_c \in \Delta(T_c)$).
+% Especially when agents do not face a static, unchanging environment, such as when decision relevant beliefs hinge on the behavior of other agents, it may simply be impossible to achieve perfect ecological rationality.
+% Moreover, not every choice mechanism is as natural as another. 
+% Some may be easier to learn and apply; some may rely on imprecise but useful generalizations across contexts, others may not.
+
+Note that we consider that, in theory, most combinations of these aspects are possible.
+Although we have been pinning logic to the classical picture, this is only with the most traditional logical systems in mind.
+Dynamic, epistemic, fuzzy, paraconsistent, and other types of logic\todo{cite Modal Logic for Open Minds?}, can potentially capture procedural rationality with limited non-strategic agents.
+Game-theoretical models of language can, on the other hand, also assume local instrumental rationality with omnisciente highly strategic agents.
+Our objective here is not to survey all the possibilities.
+We want to focus on signaling games as a framework for the study of language use and meaning.
+The vocabulary introduced in this section will, we hope, help inform the discussion that follows.
+
+% This leads to a related notion of rationality, which we will call \emph{bounded rationality} here.
+% This notion takes into account that the application of a choice mechanism may be more or less costly, \emph{e.g.}~how much time and effort does the agent invest into approximating or precisely calculating decision relevant beliefs and preferences?
+% In this way, it is conceivable that a given choice mechanism returns many good but still sub-optimal action choices in many contexts from the point of view of locally instrumental rationality.
+% At the same time, this choice mechanism, though locally sub-optimal, may outperform a strictly ecologically rational mechanism if the former is easier to compute, especially with an eye to generalization to novel contexts where precise computation of all decision relevant factors might be difficult.
+% In a nutshel, bounded rationality integrates performance factors into what should count as optimal or rational behavior. 
+% It is an alternative to instrumental and ecological rationality.
+% It becomes particularly relevant when we consider an open class of decision situations, like we do for ecological rationality, and also consider rationality of beliefs and strategic multi-player considerations in a non-static environment.
+%
 

### 2017-05-16 book_chapter/paper.tex
Moved Section 2 back into the intro. Some other minor changes.

@@
 Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
-Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
+Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions.
+% (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
 Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
@@
 Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
-It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996,skyrms_signals_2010}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
+It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
 From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.
@@
 Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
-We start, in Section~\ref{sec:facets-rationality}, by clarifying some concepts that can help characterize different notions of rationality.
-In Section~\ref{sec:signaling-games} we introduce the framework of signaling games and in Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a special kind of signaling game, a so-called sim-max game.
-Section~\ref{sec:finite-experience} reflects on the implications of these explanations for notions of rationality.
-Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
-
-\section{Facets of rationality}
-\label{sec:facets-rationality}
-
+Before we start, we want establish some vocabulary that will be useful later on.
+%
+%
+%\section{Facets of rationality}
+%\label{sec:facets-rationality}
+%
 Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
-Discussions around it, although touching on different aspects of the concept, often rely on a number of implicit assumptions.
+Discussions around it, although touching on different aspects of the concept, often fail to clearly demarcate them.
 We can start by considering what is deemed to be most important when characterizing the rationality of a given choice.
-The aforementioned classical picture of language and meaning is focused on procedures.
-Thus, a choice is deemed rational if it was arrived at by means of following proper rules of reasoning, like for example a logically valid deductive inference.
-A \emph{procedural} account of rationality focuses on the means, rather than the ends.
-One can also do the opposite and focus on the consequences of the choice instead.
-\emph{Instrumental} rationality, as it is typically called\todo{references}, characterizes an agent's choice as rational if it maximizes the possibility of achieving a desired goal, regardless of the means.
-The notion is linked to David Hume~\parencite*{hume_treatise_1738} and epitomized in his assertion that ``Reason is, and ought to only be the slave of the passions.''
-This is close to a notion of rational choice that is used in economics and game theory: agents are rational if and only if they make decisions that maximize their expected utility%
-\footnote{\todo[inline]{discuss other options like minimax and such}}%
-.
+The aforementioned classical picture of language and meaning is focused on dichotomies like true and false, meaningful and meaningless, correct and incorrect.
+A sentence can be true or false only if it is meaningful, and it is meaningful if it is constructed according to correct rules.
+Rationality is intimately connected with the ability to follow good procedures, not only of sentence production, but ultimately of sentence combination and reasoning (think of what is required for making a logically valid deductive inference).
+A \emph{procedural} account of rationality thus focuses on the means, rather than the ends.
+One can also do the opposite and focus on the consequences instead.
+\emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if it maximizes the possibility of achieving a desired goal, regardless of the means.
+The notion is linked to David Hume~\parencite*{hume_treatise_1738}, epitomized in his assertion that ``Reason is, and ought to only be the slave of the passions.''
+Instrumental rationality is close to a notion of rational choice that is used in economics and game theory%
+\footnote{In fact, it has been argued~\parencite{vanderschraaf_informal_1998} that Hume's whole account of convention is very closely in line with modern game theory.}%
+: agents are rational if and only if they make decisions that maximize their expected utility.
 A more in-depth discussion of the opposition in the context of theoretical economics can be found, for example, in the work of Herbert A.~Simon~\parencite*[\emph{e.g.}][]{simon_rationality_1986}%
@@
 %Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}
+% The dichotomy between procedural and instrumental views of rationality is not an exclusive one, since usually both means and ends are taking into account, but it helps us describe where the emphasis is placed on.
 
@@
 %Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.\todo{too soon}
+Lack of perfect knowledge of the situation or lack of strategicness can be caused by many possible factors.
+These include, but are not limited to, limitations in handling information (receiving, storing, retrieving, transmitting) and limited computational resources to solve complex problems.
+We can talk about \emph{bounded} rationality to characterize the choices of such agents.
 
@@
 If we take into account the possibility of the agent's choice or choice mechanism hinging on multiple situations, we can also talk about \emph{global} or \emph{ecological} rationality.
-This is important because, hypothetically, there could be choice mechanisms that are sub-optimal at a local level (for each situation) but are actually perfectly rational at a global level.
+This can be important because, hypothetically, there could be choice mechanisms that are sub-optimal at a local level (for each situation) but are actually perfectly rational at a global level.
 
@@
 Although we have been pinning logic to the classical picture, this is only with the most traditional logical systems in mind.
-Dynamic, epistemic, fuzzy, paraconsistent, and other types of logic\todo{cite Modal Logic for Open Minds?}, can potentially capture procedural rationality with limited non-strategic agents.
-Game-theoretical models of language can, on the other hand, also assume local instrumental rationality with omnisciente highly strategic agents.
+We are not denying that advances in dynamic, epistemic, fuzzy, paraconsistent, and other types of logic could potentially enable one to capture procedural rationality with different characteristics.
+Game-theoretical models of language can, on the other hand, also assume local instrumental rationality with omniscient highly strategic agents.
 Our objective here is not to survey all the possibilities.
 We want to focus on signaling games as a framework for the study of language use and meaning.
-The vocabulary introduced in this section will, we hope, help inform the discussion that follows.
+The vocabulary just introduced will, we hope, help inform the discussion that follows.
+We proceed by introducing the framework of signaling games in Section~\ref{sec:signaling-games}.
+In Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a particular kind of signaling game.
+Section~\ref{sec:finite-experience} reflects on the implications of these explanations for considerations of rationality.
+Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
 
@@
 However, we can already see how it attempts to address the study of meaning from a very different angle.
-There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific purposes \parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
+There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific goals~\parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
 There is also no appeal to truth as a guide for our inquiry.
 One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
-We can merely focus on trying to see how agents can use signals to cope with the world and achieve their purposes.
+We can merely focus on trying to see how agents can use signals to cope with the world and achieve their goals.
 On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
 In the examples discussed so far, agents are assumed to possess and exercise rationality.
-It is not, however, the procedural notion of rationality of the classical picture, but rather one that is instrumental and heavily strategic.
+Unlike in the classical picture, it is an instrumental, rather than procedural, rationality, but one where agents are still omniscient and highly strategic.
 
@@
 The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
-However, if we consider the origins of language from a historical perspective, it seems implausible to assume a high degree of rationality of the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages.
+However, if we consider the origins of language from a historical perspective, it seems implausible that the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages possessed such advanced rationality.
 Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
@@
 Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
-Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite[\emph{e.g.}][]{dawkins_selfish_1978,boyd_culture_1985}.
-The principles can be captured in formal models that abstracts away from details of single interactions and individual agent's behavior, \emph{e.g.}~the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
+Although the inspiration for this scenario is biological evolution, similar things could be said about how ideas spread in a population of agents who can adopt or abandon them depending on how successful they prove to be~\parencite[\emph{e.g.}][]{BenzJager2006:Game-Theory-and,Pagel2009:Human-Language-,ThompsonKirby2016:Culture-Shapes-}, \emph{i.e.}~we can interpret these notions in terms of cultural evolution~\parencite{dawkins_selfish_1978,boyd_culture_1985}.
+The principles can be captured in formal models that abstract away from details of single interactions and behavior of individual agents, for example in the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
 The only thing relevant to the replicator dynamics are the relative proportions of strategies in a given population and the utility function.
@@
 Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
-Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
-While the general framework manages to abstract quite some detail away from the formalization, it does leave room for them, especially when it comes to the dynamics.
+The shift in perspective allows interpretations that accommodate for limited non-strategic agents.
+% Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
+While the general framework manages to abstract quite some details away from the formalization, it nevertheless leaves room for them, especially when it comes to the dynamics.
 We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
@@
 Each of these can be utilized depending on the problem that one is interested in characterizing.
-Thus, although Skyrms shows that rationality is not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions where rationality can play a role.
+Thus, although Skyrms shows that high requirements of rationality are not necessary for signaling conventions to evolve, the framework does leave room for the study of linguistic interactions between highly strategic agents.
 
@@
 A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
-With these tools, not only can we better understand why signaling systems are stable even without any assumptions of rationality, but we can also map out which initial conditions drive the system towards which equilibria and which do not.
+With these tools, we can better understand why signaling systems are stable even without any strong assumptions of rationality.
+We can also map out which initial conditions drive the system towards which equilibria and which do not.
 In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
@@
 Each agent plays an instrumentally rational best response (given the utility function) to their estimate of the other's strategy.
-In order to observe the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
-What the authors observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
+In order to study the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
+What they observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
 However, each agent still commands a crisp language, which is inadequate if the intention is to capture how vagueness presents itself in human languages.
@@
 Considerations of optimality of strategies, for example by considering the evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
-However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a smaller period of time.
+However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a shorter period of time.
 In a naturalized evolutionary setting, where we care about bounded rationality in a multi-context environment, learning speed is an advantage which should also be taken into account.
@@
 Furthermore, they observe one additional potentially beneficial property.
-By running several rounds of simulation for each parameter set, they can measure how close resulting strategies are within each group of results, and how they would fare playing against each other.
+By running several rounds of simulation for each parameter set, they can measure for each group of results how close resulting strategies are to each other, and how they would fare playing against one another.
 The results show that the within group distance between strategies becomes smaller with growing confusability, and the within group expected utility is actually higher for strategies evolving under a certain degree of state confusion.
@@
 Although the authors do not define a utility function, their analysis evaluates the models in terms of the expected error between the estimate and the original state, defined as $(x-y)^2$.
-Better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse
+Better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse.
 The game could then be seen as a sim-max game with multiple senders and negative squared difference as utility function.
@@
 This ``economic man'', a game-omniscient and highly strategic, instrumentally rational agent, seems to have survived almost unscathed to now populate most game theoretic models of language and signaling.
-Both Simon and Lipman call for this picture to be revisited.
+Both Simon and Lipman call for this picture to be revised.
 
@@
 \end{enumerate}
-All of these observations support the weakening of our modeling assumptions.
+All of these observations support the weakening of the modeling assumptions.
 % If we take these points seriously, we have to concede that it is almost inevitable that uncertainty gets introduced at almost all levels of language learning and language use.
@@
 % Another possible interpretation is that of a population of agents learning by imitation.
-The models of~\textcite{franke_vagueness_2011}, generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, all paint a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
+Generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, both paint a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
 Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
@@
 Agents in those models seem to be rational only to the extent of their capabilities.
-Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), this feels like a bittersweet conclusion because of this hypothetical possibility of an ideal strategy that seems to be only artificially barred from the agents.
-But this is so only if rationality is equated with maximization of expected utility.
+Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), this feels like a bittersweet conclusion because of the hypothetical possibility of an ideal strategy that seems to be only artificially barred from the agents.
+But this is so only if rationality is equated with maximization of expected utility on a pure local level.
 The mechanisms that lead to vague signal use, as~\textcite{oconnor_evolution_2014} and~\textcite{franke_vagueness_2017} stress, have the aforementioned important advantages of faster speed of convergence, higher flexibility, and homogenization.
@@
 Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
-In a scenario where the agent's ability to make the optimal choice is bounded, such as the QRE model of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
+In a scenario where the agent's ability to make the optimal choice is bounded, such as in the models of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
 \textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
-The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as utility maximization.
+The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as local utility maximization.
 But even in other scenarios, we could again argue that, if we broaden our notion of rationality, it would be less rational to use a crisp strategy rather than one that incorporates at least some source of uncertainty.
 A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
+Though potentially a locally rational choice, it might thus not be globally optimal.
 
@@
 The picture just sketched neglects learning from observation, as well as generalization across contexts. 
-For strategically and instrumentally rational agents, we need to consider also the ability to learn from previous experience and to generalize from the past to the future. 
+For strategic and instrumentally rational agents, we need to consider also the ability to learn from previous experience and to generalize from the past to the future. 
 For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
@@
 We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
-These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time. 
+These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[\S 23]{wittgenstein_philosophical_1953}.
 We could look for meaning at several levels in this pluralistic picture. 
@@
 As mentioned above in connection the soft-max choice function used by \textcite{franke_vagueness_2011}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic.
-\mf{mention ``bounded rationality'' here}
+% \mf{mention ``bounded rationality'' here}
 Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
@@
 In sum, both at the level of behavior and at the level of beliefs about use or meaning, we should expect to find vagueness.
-Still, despite the natural vagueness, there does not seem to anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
+Still, despite the natural vagueness, there does not seem to be anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
 
@@
 
-The investigation doesn't end here, though.
+The investigation doesn't end here, then.
 The research discussed here suggests that although vagueness can be beneficial, that is the case only in moderation: too much can lead to babbling equilibria where no information is transmitted.
@@
 Whatever solutions we find to these challenges, we believe that a notion of rationality cannot be strongly tied to mere utility maximization in a particular game, but has to consider an agents' need to optimize his behavior to navigate the plurality of uncertainties that besiege him.
+We need a clear notion of ecological rationality in order to see vagueness in the right light.
 

### 2017-05-17 book_chapter/paper.tex
minor text edits

@@
 By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~they are able to recognize the logical inconsistency but do not have a good answer as to how to overcome it.
-Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language productively.
+Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language to achieve their daily goals.
 Nobody ever stopped using the word `tall` after being confronted with a sorites series.
@@
 Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
-Before we start, we want establish some vocabulary that will be useful later on.
+Before we start, we want to establish some vocabulary that will be useful later on.
 %
@@
 
-When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between whatever agents are involved and the elements of the model can come to bear on considerations of rationality.
-Depending on how accurate and complete the agent's knowledge of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, etc, the verdict over the rationality of certain choices can vary.
+When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between agents and their environment can come to bear on considerations of rationality.
+Depending on how accurate and complete an agent's knowledge of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, etc, the verdict over the rationality of certain choices can vary.
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.

### 2017-08-16 book_chapter/meta-dynamics.tex
Wrote down first sketch of a formalization of meta-dynamics

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{bm}

\title{Meta-dynamics: a sketch}
\author{José Pedro Correia}
\date{\today}

\begin{document}
\maketitle

Say that we want to test different dynamics against each other in one population of agents.
Each dynamic is used by a certain percentage of the population.
Each sub-population is characterized by a sender (receiver) strategy which is evolved according to its dynamic, but they interact with each other, so each strategy's expected utility is calculated against the whole population.
I will here assume that senders and receivers are separate sub-populations.

We can denote the percentage of sender population using dynamic $i \in I$ by $S_i$ and the percentage of receiver population using the same dynamic by $R_i$.
Strategies are also indexed by the dynamic, so we have $\sigma_i$ and $\rho_i$ for each $i \in I$.
Dynamics are functions $D_i$ that calculate a new generation of strategies based on the existing ones (this could benefit from a more formal characterization).

The calculation of a new generation now applies to both the strategies themselves, as to the dynamics.
Sender strategy $\sigma_i$ gets updated according to its dynamic $D_i$:
$$
\sigma_i^\prime(m \mid t) = D_i(\sigma_i, EU(m \mid t))
$$

For example, if $D_3$ is the replicator dynamic, we would have the following:
$$
\sigma_3^\prime(m \mid t) = \sigma_3(m \mid t) \cdot \frac{EU(m \mid t)}{\sum_{m^\prime \in M} EU(m^\prime \mid t)}
$$

One important point is that expected utility for the sender strategy of one dynamic is calculated against the receiver strategies of all other dynamics:
$$
EU(m \mid t) = \sum_{i \in I} R_i \cdot \sum_{t^\prime \in T} \rho_i(t^\prime \mid m) \cdot U(t, t^\prime)
$$

Then we also update the percentage of senders using a certain dynamic.
In this case, using the replicator equation as meta-dynamic, we get the following:
$$
S_i^\prime = S_i \cdot \frac{EU(\sigma_i)}{\sum_{i^\prime \in I} EU(\sigma_{i^\prime})}
$$
where the expected utility of a whole sender strategy of one dynamic is also calculated against the receiver strategies of all other dynamics:
$$
EU(\sigma) = \sum_{i \in I} R_i \cdot \sum_{t, t^\prime \in T} \sum_{m \in M} P(t) \cdot \sigma(m \mid t) \cdot \rho_i(t^\prime \mid m) \cdot U(t, t^\prime)
$$

We can make similar formulations, \emph{mutatis mutandis}, for the receiver populations.

\end{document}

### 2017-10-25 book_chapter/simulation/vagueness-games.py
Fixing a big mistake. Should be storing final strategies per impairment value.

@@ -429,10 +429,10 @@ with open(MeasurementsOutputFilename, 'w') as csvfile:
 
 for i in xrange(NPopulations):
     SpeakerOutputFilename = ResultsDirectory + '/strategies/' + \
-                            SimulationID + '-speaker.csv'
+                            SimulationID + '-speaker-' + str(Impairments[i]) + '.csv'
     np.savetxt(SpeakerOutputFilename, Speakers[i], delimiter=',')
     HearerOutputFilename = ResultsDirectory + '/strategies/' + \
-                           SimulationID + '-hearer.csv'
+                           SimulationID + '-hearer-' + str(Impairments[i]) + '.csv'
     np.savetxt(HearerOutputFilename, Hearers[i], delimiter=',')
 
 # (SpeakerVoronoiness, HearerVoronoiness) = Voronoiness(Hearer, Speaker, Similarity)

### 2017-10-26 book_chapter/simulation/run-batch.sh
Regenerated results, because strategies for certain impairment values were missing from previous run.

@@ -1,6 +1,16 @@
 #!/bin/bash
 
-for a in "0.0" "0.0 0.05" "0.0 0.05 0.1"
+for n in {1..25}
+do
+    echo -n "  $n "
+    python vagueness-games.py --batch 30 1 0.05 weakest 0.0
+    if [ $? -ne 0 ]; then
+        exit
+    fi
+done
+echo
+
+for a in "0.0 0.05" "0.0 0.05 0.1"
 do
     echo $a
     for b in "strong" "weak" "weakest"

### 2017-10-26 book_chapter/paper.tex
references fixed
Wrote up on tight population interaction model. Generated some plots.

@@
 
-\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false,isbn=false]{biblatex}
+\usepackage[natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
@@
 
-We believe these considerations are, again, very much in line with Wittgenstein's later philosophy:
-\begin{quote}
-The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
-(For the crystalline purity of logic was, of course, not something I had discovered: it was a requirement.)
-The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
-– We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
-We want to walk: so we need friction.
-Back to the rough ground!%
-~\parencite*[\S 107]{wittgenstein_philosophical_1953}
-\end{quote}
-This seems like a very clear call for what we are arguing here.
-But are we exposing ourselves to ``philosophical disease'' by following a ``one-sided diet''~\parencite*[\S 593]{wittgenstein_philosophical_1953} when looking only at sim-max games?
-In the spirit of Wittgenstein's pluralistic picture of language, we would like to reflect on vagueness in one more type of signaling game and speculate on the impact of plurality for vagueness.
+In order to explore these considerations further, in the remainder of the paper we propose three variations of a multi-population model of imprecise imitation dynamics.
+The idea is to investigate how populations with different levels of imprecision interact, and to what extent does the hypothesized temporary advantage of faster convergence helps vagueness to thrive in such an environment.
 
-\section{Vagueness in multiple reference games}
-\label{sec:referential-vagueness}
+%We believe these considerations are, again, very much in line with Wittgenstein's later philosophy:
+%\begin{quote}
+%The more closely we examine actual language, the greater becomes the conflict between it and our requirement.
+%(For the crystalline purity of logic was, of course, not something I had discovered: it was a requirement.)
+%The conflict becomes intolerable; the requirement is now in danger of becoming vacuous.
+%– We have got on to slippery ice where there is no friction, and so, in a certain sense, the conditions are ideal; but also, just because of that, we are unable to walk.
+%We want to walk: so we need friction.
+%Back to the rough ground!%
+%~\parencite*[\S 107]{wittgenstein_philosophical_1953}
+%\end{quote}
+%This seems like a very clear call for what we are arguing here.
+%But are we exposing ourselves to ``philosophical disease'' by following a ``one-sided diet''~\parencite*[\S 593]{wittgenstein_philosophical_1953} when looking only at sim-max games?
+%In the spirit of Wittgenstein's pluralistic picture of language, we would like to reflect on vagueness in one more type of signaling game and speculate on the impact of plurality for vagueness.
 
-The type of linguistic interaction captured by a sim-max game has its own particular scope.
-There is a fixed prior, a fixed utility function and a fixed similarity measure between states. 
-If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t'$.
-Here is a natural example of a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue; the closer Bubu's color choice matches Kiki's idea, the better.
+%\section{Vagueness in multiple reference games}
+%\label{sec:referential-vagueness}
+%
+%The type of linguistic interaction captured by a sim-max game has its own particular scope.
+%There is a fixed prior, a fixed utility function and a fixed similarity measure between states.
+%If the sender observes a state $t$ and the receiver guesses state $t'$ the utility of that interaction is proportional to the similarity between $t$ and $t'$.
+%Here is a natural example of a situation captured by a sim-max model: Kiki wants Bubu to draw her dissertation cover in a particular shade of blue; the closer Bubu's color choice matches Kiki's idea, the better.
 Here Kiki wants to \emph{describe} the color for Bubu to pick out or realize from the whole set of possibilities.
-It's a natural thing to want to do with language. 
-But other types of communication with vague and contiguously variable properties are possible. 
-A \emph{reference game} is a model of a situation in which the speaker has a particular object in mind and uses a description in order for the receiver to pick this object and no other object. 
-One crucial difference with sim-max games is that close will not be close enough: if Kiki needs a long nail, a nail a little longer or shorter may be good enough; but if Bubu needs the long, slightly bend nail with the flat top (which she previously disassembled from her art installation), no other nail will do.
+%It's a natural thing to want to do with language.
+%But other types of communication with vague and contiguously variable properties are possible.
+%A \emph{reference game} is a model of a situation in which the speaker has a particular object in mind and uses a description in order for the receiver to pick this object and no other object.
+%One crucial difference with sim-max games is that close will not be close enough: if Kiki needs a long nail, a nail a little longer or shorter may be good enough; but if Bubu needs the long, slightly bend nail with the flat top (which she previously disassembled from her art installation), no other nail will do.
+%
+%Interestingly, referential communication can take place under a plethora of different epistemic conditions.
+%A lot of modeling work considers \emph{reference games without displacement} where sender and receiver communicate about a context, i.e.
+a set of objects, which is immediately accessible in their direct perceptual environment \parencite[\emph{e.g.}][]{BaronchelliGong2010:Modeling-the-Em,Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.
+%Agents may make mistakes about observing exactly how blue or long etc.~a particular object is, but there is no further fundamental uncertainty about the context.
+%This, of course, need not be so. \textcite{lipman_why_2009} and \textcite{Deemter2009:Utility-and-Lan}, for example, consider \emph{reference games with displacement}.
+%Kiki asks Bubu to fetch the red book from the library.
+%No other book than the one Kiki needs will serve.
+%When Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki at the time of speaking.
+%Yet other situations are imaginable.
+%In \emph{reference games with partial uncertainty}, for instance, interlocutors may know that they have shared and immediate access only to a subset of the objects in the context, but be uncertain about whether the co-player sees any other objects.
+%This is the set-up for the widely-used director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
+%
+%What would a rational agent do in repeated reference games?
+%Take a universe in which agents play repeated reference games without displacement.
+%A single game has sender and receiver both flawlessly perceive a context of $k$ objects (with $k$ possibly randomly drawn for each encounter); each object has its individual features, e.g., drawn from a rich multidimensional feature space of continuously variable properties \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.
+%The sender chooses a message $m$ for the target object $t$, but she can condition her choice on the current context $c$.
+%Likewise, the receiver chooses a referent $t'$ from the available objects in $c$, again conditional on $c$.
+%But if agents play many such reference games with different contexts $c$ and $c'$, strategic and instrumental rationality of local choices in $c$ and $c'$ is not enough to guarantee that agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
+%An agent could use a certain signal to denote, say, the tallest object in every context with between five and twelve objects or when the number of objects is a prime number, but use the same signal for the darkest object in all other contexts.
+%In other words, neither rationality of choice in a fixed context nor evolutionary adaptation to each single context entails a natural, globally consistent behavior across multiple contexts.
+%The situation only gets worse if we want to have agents use the same expressions in a consistent way also across different types of reference games, with possibly different epistemic conditions.
+%
+%The picture just sketched neglects learning from observation, as well as generalization across contexts.
+%For strategic and instrumentally rational agents, we need to consider also the ability to learn from previous experience and to generalize from the past to the future.
+%For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
+%But even if every language learner in history had been perfectly rational, it is extremely unlikely, as was discussed in the previous section, that anyone ever succeeded in reducing all uncertainty about language use and say with confidence: this is the threshold that truly separates heaps from non-heaps no matter what.
+%For one, there may simply not be enough time in the lifespan of a rational learner to observe enough instances of language use for certainty about crisp denotations.
+%For another, what needs to be learned likely changes over time.
+%The optimal use for gradable properties hinges crucially on the prior over states \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie,QingFranke2014:Gradable-Adject,LassiterGoodman2015:Adjectival-vagu}.
+%If the prior captures real world attributes, such as the probability of an individual's body height, these attributes will not only be hard to know for certain, but they will also vary over time.
+%They may also depend on location: do we use our language exclusively in this village (where all heaps are extraordinarily small and people are extraordinarily tall) or do we travel beyond?
+%If so, the priors that determine the goodness of a convention for one agent need not be the same as those for another.
+%Going beyond the priors for property degrees, the probability of seeing a particular context $c$ in some kind of reference game, is also crucial for determining globally efficient linguistic behavior.
+%But here it is even more inconceivable how a rational learner could acquire a fairly confident estimate over how often the upcoming reference games will have a particular context or another.
+%The list of sources of natural and seemingly insurmountable uncertainty about the parameters necessary to determine locally or globally optimal strategies is much longer than hinted at here.
+%Still, it should be clear enough that a finite existence, a sparse learning input and the inherent fuzziness of the linguistic context itself (its temporal, local and social variability) is not a fault of rationality on the part of the agent, or the mechanism of selection of successful linguistic practices.
+%
+%What about meaning and vagueness in a universe of multiple language games and heavy uncertainty?
+%We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
+%These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[\S 23]{wittgenstein_philosophical_1953}.
+%We could look for meaning at several levels in this pluralistic picture.
+%Firstly, as before, there is the actual behavior of a single agent in each actualized language game.
+%As mentioned above in connection the soft-max choice function used by \textcite{franke_vagueness_2011}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic.
+%% \mf{mention ``bounded rationality'' here}
+%Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
+%Notice that a single agent's rational beliefs about linguistic practices or linguistic meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
+%In sum, both at the level of behavior and at the level of beliefs about use or meaning, we should expect to find vagueness.
+%Still, despite the natural vagueness, there does not seem to be anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
+%
+%% Finally, we might be interested in a notion of meaning of language as an abstraction over their users and its use.
+%% We might not care about individual agents' behavior and beliefs as much as we care about the meaning of words, expressions and sentences.
+%% The questions to address then are probably: what explanatory purpose is a notion of meaning of an abstracted language to play?; if both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp?; what can be explained only by a notion of meaning that is not vague?
+%% \mf{I leave this paragraph in for you to read and ponder, but would recommend to leave it out entirely}
+%
+%Taken together, when we look at the picture of multiple language games side by side, the question of what rational or optimal behavior is in any one of them, or the question of why human language users do not more closely approximate instrumentally rational or optimal local behavior more closely seem entirely misplaced.
+%What seems more pressing is to answer how the interaction of individual learning, individual local optimization and general pressure for successful communication at a cross-situational level lead to the shaping of linguistic practices in an open-ended landscape of language games.
+%How to attain locally sub-optimal play in a single language game is not a theoretical puzzle we should take serious, if what really matters is a global multi-game perspective.
+%Adopting this perspective does not \emph{solve} the problem of how vagueness is compatible with rationality.
+%But it suggest that it might be an inflated issue that we need not or should not worry about, because it will likely not matter in a more realistic and encompassing picture of language, its use and evolution as driven by an appropriate notion of rational choice and evolutionary selection.
 
-Interestingly, referential communication can take place under a plethora of different epistemic conditions. 
-A lot of modeling work considers \emph{reference games without displacement} where sender and receiver communicate about a context, i.e.
-a set of objects, which is immediately accessible in their direct perceptual environment \parencite[\emph{e.g.}][]{BaronchelliGong2010:Modeling-the-Em,Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.
-Agents may make mistakes about observing exactly how blue or long etc.~a particular object is, but there is no further fundamental uncertainty about the context. 
-This, of course, need not be so. \textcite{lipman_why_2009} and \textcite{Deemter2009:Utility-and-Lan}, for example, consider \emph{reference games with displacement}. 
-Kiki asks Bubu to fetch the red book from the library. 
-No other book than the one Kiki needs will serve. 
-When Kiki describes the book, the perceptual context in which Bubu must make her choice is not co-present to Kiki at the time of speaking. 
-Yet other situations are imaginable. 
-In \emph{reference games with partial uncertainty}, for instance, interlocutors may know that they have shared and immediate access only to a subset of the objects in the context, but be uncertain about whether the co-player sees any other objects.
-This is the set-up for the widely-used director task in psycholinguistics \parencite[\emph{e.g.}][]{KraussGlucksberg1977:Social-and-nons,Keysar2000,KeyzarLin2003:Limits-on-Theor}.
+\section{Two multi-population models of imprecise imitation}
+\label{sec:multi-population-model}
 
-What would a rational agent do in repeated reference games? 
-Take a universe in which agents play repeated reference games without displacement. 
-A single game has sender and receiver both flawlessly perceive a context of $k$ objects (with $k$ possibly randomly drawn for each encounter); each object has its individual features, e.g., drawn from a rich multidimensional feature space of continuously variable properties \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie}.
-The sender chooses a message $m$ for the target object $t$, but she can condition her choice on the current context $c$. 
-Likewise, the receiver chooses a referent $t'$ from the available objects in $c$, again conditional on $c$. 
-But if agents play many such reference games with different contexts $c$ and $c'$, strategic and instrumental rationality of local choices in $c$ and $c'$ is not enough to guarantee that agents' beliefs and behavior in $c$ have any relation to their beliefs and behavior in $c'$.
-An agent could use a certain signal to denote, say, the tallest object in every context with between five and twelve objects or when the number of objects is a prime number, but use the same signal for the darkest object in all other contexts.
-In other words, neither rationality of choice in a fixed context nor evolutionary adaptation to each single context entails a natural, globally consistent behavior across multiple contexts. 
-The situation only gets worse if we want to have agents use the same expressions in a consistent way also across different types of reference games, with possibly different epistemic conditions.
+We build upon the model of~\textcite{franke_vagueness_2017} by adding the possibility for more than one population, where each population has its own imprecision parameter.
+The dynamic needs to be adapted to allow different types of interactions between these populations, namely regarding which agents will be imitated, and which agents will be expected to play against.
+Finally, in order to see the impact of the success of different imprecision levels, we add another process of selection.
+Imagine the following: agents are born with varying perceptual abilities (modeled as populations with different imprecision values), they learn a signaling strategy by imitation of other agents (either within or across populations), depending on the success of the strategies they develop, they will be more or less likely to survive and reproduce, thus the proportion of each population will change depending on their fitness compared to other populations.
 
-The picture just sketched neglects learning from observation, as well as generalization across contexts. 
-For strategic and instrumentally rational agents, we need to consider also the ability to learn from previous experience and to generalize from the past to the future. 
-For evolutionary adaptation, the story so far misses the key fact that languages and practices have to be learnable in appropriate time from the actual observables, lest what is learned and further transmitted will simply be a different language or practice \parencite[\emph{e.g.}][]{KirbyGriffith2014:Iterated-Learni}.
-But even if every language learner in history had been perfectly rational, it is extremely unlikely, as was discussed in the previous section, that anyone ever succeeded in reducing all uncertainty about language use and say with confidence: this is the threshold that truly separates heaps from non-heaps no matter what.
-For one, there may simply not be enough time in the lifespan of a rational learner to observe enough instances of language use for certainty about crisp denotations. 
-For another, what needs to be learned likely changes over time. 
-The optimal use for gradable properties hinges crucially on the prior over states \parencite[\emph{e.g.}][]{Franke2012:Scales-Salience,Franke2012:On-Scales-Salie,QingFranke2014:Gradable-Adject,LassiterGoodman2015:Adjectival-vagu}.
-If the prior captures real world attributes, such as the probability of an individual's body height, these attributes will not only be hard to know for certain, but they will also vary over time. 
-They may also depend on location: do we use our language exclusively in this village (where all heaps are extraordinarily small and people are extraordinarily tall) or do we travel beyond? 
-If so, the priors that determine the goodness of a convention for one agent need not be the same as those for another. 
-Going beyond the priors for property degrees, the probability of seeing a particular context $c$ in some kind of reference game, is also crucial for determining globally efficient linguistic behavior.
-But here it is even more inconceivable how a rational learner could acquire a fairly confident estimate over how often the upcoming reference games will have a particular context or another. 
-The list of sources of natural and seemingly insurmountable uncertainty about the parameters necessary to determine locally or globally optimal strategies is much longer than hinted at here. 
-Still, it should be clear enough that a finite existence, a sparse learning input and the inherent fuzziness of the linguistic context itself (its temporal, local and social variability) is not a fault of rationality on the part of the agent, or the mechanism of selection of successful linguistic practices.
+Let $A$ be the set of imprecision values considered.
+For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$ such that $\sum_{\alpha \in A} P(\alpha) = 1$.
+The probability that a given agent with imprecision $\alpha$ (or of type $\alpha$) observes $t_o$ when the actual state is $t_a$ is given by $P_{o}^{\alpha}(t_{o}|t_{a})$.
+If the same agent intends to realize $t_i$, the probability that he actually realizes $t_r$ instead is given by $P_{r}^{\alpha}(t_{r}|t_{i})$.
+Following~\textcite[26]{franke_vagueness_2017}, we can define the following values.
+Probability that $t_a$ is actual if $t_o$ is observed by an agent of type $\alpha$:
+$$
+P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\propto P_{a}(t_{a})P_{o}^{\alpha}(t_{o}|t_{a})
+$$
+Probability that a random sender of type $\alpha$ produces $m$ when the actual state is $t_a$:
+$$
+P_{\sigma}^{\alpha}(m|t_{a})=\sum_{t_{o}}P_{o}^{\alpha}(t_{o}|t_{a})\sigma^{\alpha}(m|t_{o})
+$$
+Probability that the actual state is $t_a$ if a random sender of type $\alpha$ produced $m$:
+$$
+P_{\rho}^{\alpha}(t_{r}|m)=\sum_{t_{i}}P_{r}^{\alpha}(t_{r}|t_{i})\rho^{\alpha}(t_{i}|m)
+$$
+Probability that $t_r$ is realized by a random receiver of type $\alpha$ in response to message $m$:
+$$
+P_{\bar{\sigma}}^{\alpha}(t_{a}|m)\propto P_{a}(t_{a})P_{\sigma}^{\alpha}(m|t_{a})
+$$
 
-What about meaning and vagueness in a universe of multiple language games and heavy uncertainty?
-We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
-These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[\S 23]{wittgenstein_philosophical_1953}.
-We could look for meaning at several levels in this pluralistic picture. 
-Firstly, as before, there is the actual behavior of a single agent in each actualized language game. 
-As mentioned above in connection the soft-max choice function used by \textcite{franke_vagueness_2011}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic.
-% \mf{mention ``bounded rationality'' here}
-Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
-Notice that a single agent's rational beliefs about linguistic practices or linguistic meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
-In sum, both at the level of behavior and at the level of beliefs about use or meaning, we should expect to find vagueness.
-Still, despite the natural vagueness, there does not seem to be anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
+Given these probabilities, we can compute expected utilities and update probabilities for each type.
+We have a number of possible types of interaction between populations.
 
-% Finally, we might be interested in a notion of meaning of language as an abstraction over their users and its use.
-% We might not care about individual agents' behavior and beliefs as much as we care about the meaning of words, expressions and sentences.
-% The questions to address then are probably: what explanatory purpose is a notion of meaning of an abstracted language to play?; if both agents' rational behavior and rational beliefs are plausibly fuzzy, why should a concept of supervening meaning be crisp?; what can be explained only by a notion of meaning that is not vague?
-% \mf{I leave this paragraph in for you to read and ponder, but would recommend to leave it out entirely}
+\subsection{Tight population interaction}
+\label{sec:tight-interaction-model}
+The first we will consider is a tight one, where each agent plays with, observes, and potentially imitates any other agent, regardless of their type.
+Expected utilities for this case are given, for a sender of type $\alpha$, by:
+$$
+\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})\sum_{t_{r}}P_{\rho}^{\alpha^{\prime}}(t_{r}|m)U(t_{a},t_{r})
+$$
+and, for a receiver of type $\alpha$, by:
+$$
+\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})=\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})\sum_{t_{a}}P_{\bar{\sigma}}^{\alpha^{\prime}}(t_{a}|m)\sum_{t_{r}}P_{r}^{\alpha}(t_{r}|t_{i})U(t_{a},t_{r})
+$$
 
-Taken together, when we look at the picture of multiple language games side by side, the question of what rational or optimal behavior is in any one of them, or the question of why human language users do not more closely approximate instrumentally rational or optimal local behavior more closely seem entirely misplaced.
-What seems more pressing is to answer how the interaction of individual learning, individual local optimization and general pressure for successful communication at a cross-situational level lead to the shaping of linguistic practices in an open-ended landscape of language games.
-How to attain locally sub-optimal play in a single language game is not a theoretical puzzle we should take serious, if what really matters is a global multi-game perspective.
-Adopting this perspective does not \emph{solve} the problem of how vagueness is compatible with rationality. 
-But it suggest that it might be an inflated issue that we need not or should not worry about, because it will likely not matter in a more realistic and encompassing picture of language, its use and evolution as driven by an appropriate notion of rational choice and evolutionary selection.
+Furthermore, we can define the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
+$$
+P_{o}^{\alpha}(m|t_{o})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})P_{\sigma}^{\alpha^{\prime}}(m|t_{a})
+$$
+and the probability that a receiver of type $\alpha$ observes a randomly sampled agent choose interpretation $t_o$ given message $m$ as:
+$$
+P_{o}^{\alpha}(t_{o}|m)=\sum_{t_{r}}P_{o}^{\alpha}(t_{o}|t_{r})\sum_{\alpha^{\prime}\in A}P(\alpha^{\prime})P_{\rho}^{\alpha^{\prime}}(t_{r}|m)
+$$
+
+Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ is given by:
+$$
+\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)EU_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
+$$
+and similarly for a receiver strategy of type $\alpha$ by:
+$$
+\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)EU_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
+$$
+
+The formulations so far cover the evolution of the particular strategies of agents.
+We can think of this as the level of cultural evolution: agents are born with a certain level of imprecision and learn strategies based on the behavior of others.
+Alongside this process, we can imagine another level of selection, where agents die and new agents are born.
+More successful agents will have a higher likelihood of surviving and reproducing, giving rise to more agents with their level of imprecision.
+Levels of imprecision are thus subject to an evolutionary dynamic that is indirectly influenced by the cultural dynamic.
+\todo{genotype vs phenotype?}
+Importantly, only the level of imprecision is passed on to new generations under this dynamic, not the actual strategies developed by the agents at the cultural level.
+
+We model this process by changing the proportion of each type $P(\alpha)$ according to the replicator dynamic.
+A population of type $\alpha$ consists of agents employing both a sender and a receiver strategy, thus the overall fitness of the population must involve the expected utilities of each.
+We could imagine this process happening at a different speed than the cultural process, in which case we could have different time scales.
+For the sake of simplicity, we choose to have them both happen at each time step, but calculate the changes occurring at this level of selection after the other level.
+We define the proportion of type $\alpha$ at time step $i+1$ as:
+$$
+P_{i+1}(\alpha) \propto P_i(\alpha)(EU_{\sigma_{i+1}}^{\alpha}(m,t,\rho_{i}^{\star}) + EU_{\rho_{i+1}}^{\alpha}(t,m,\sigma_{i}^{\star}))
+$$
+In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma}$ and $\tilde{\rho}$ according to a parameter $\gamma$ defining the birth rate within the population.
+This can be described, for a certain time step $i+1$, in the following formulas:
+$$
+\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}
+$$
+$$
+\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}
+$$
+
+We performed 25 simulation runs of this model for each of three population scenarios: only one population with $\alpha = 0$ (for reference), two populations with $\alpha \in \{0, 0.05\}$, and three populations with $\alpha \in \{0, 0.05\}$.
+For each scenario, starting proportions $P(\alpha)$ were equal for each value of $\alpha$.
+Given the observations by~\textcite{franke_vagueness_2017} that state space size and tolerance parameter $\beta$ do not result in important qualitative difference, we fixed these values at $n_S = 30$ and $\beta = 0.1$.
+We also used a fixed uniform distribution for the priors and a message space with 2 messages.
+Regarding the duration of the simulations, due to the mixing in of random strategies at each step in order to simulate new individuals being introduced into the population (birth rate was fixed at $\gamma = 0.05$), the convergence criteria is not applicable.
+Therefore, all simulation runs were stopped after 200 iterations.
+
+The first thing to observe from the simulation results is that the population with no imprecision ($\alpha = 0$) always dominated the other populations in every run.
+\begin{figure}
+  \centering
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-example-0-005-strong.png}
+%    \caption{Crisp}
+%    \label{fig:example_stratsA}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-example-0-005-01-strong.png}
+%    \caption{Vague}
+%    \label{fig:example_stratsB}
+  \end{subfigure}
+  \caption{\todo[inline]{description}}
+  \label{fig:examples-proportion-tight-interaction}
+\end{figure}
+In Figure~\ref{fig:examples-proportion-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for one example trial.
+There were no qualitative differences across trials.
+As the plot shows, the population with no imprecision steadily increases its proportion against the others.
+In the three population scenario, the proportion of the population with $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
+
+These observations go against the expectation of~\textcite{franke_vagueness_2017} that faster convergence to a convex strategy by populations with a certain level of imprecision could give them a temporary advantage to take over and eliminate other types.
+What happens in this setup is nevertheless easy to explain.
+Because the populations interact very tightly, they develop strategies in close tandem with each other.
+One of the consequences of this is that the population with no imprecision reaps some benefits from evolving together with a population with a certain level of imprecision.
+Namely, they reach convexity faster and develop more homogeneous strategies
+\begin{figure}
+  \centering
+  \begin{subfigure}[]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-percentage-all-strong.png}
+%    \caption{Crisp}
+    \label{fig:convexity-tight-interaction}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-sender-all-strong.png}
+%    \caption{Vague}
+    \label{fig:entropy-sender-tight-interaction}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.3\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-receiver-all-strong.png}
+%    \caption{Vague}
+    \label{fig:entropy-receiver-tight-interaction}
+  \end{subfigure}
+  \caption{\todo[inline]{description}}
+  \label{fig:metrics-tight-interaction}
+\end{figure}
+We can see the effect on convexity by looking to Figure~\ref{fig:convexity-tight-interaction} where we plot, for each scenario, the percentage of strategies that are convex at a given iteration, including the scenario with a population with no imprecision evolving on its own.
+These effects preclude the hypothesized temporary advantage to even manifest itself.
+
+The flipside of this tight connection between populations is that strategies evolved by populations with no imprecision are also more vague (in the sense defined in Section~\ref{sec:sim-max-vagueness}).
+This is visible by looking at entropy values, namely sender strategy entropy, plotted in Figure~\ref{fig:entropy-sender-tight-interaction}, and receiver strategy entropy, plotted in Figure~\ref{fig:entropy-receiver-tight-interaction}.
+The values for the population with no imprecision are clearly higher in the scenarios where it evolves together with populations with imprecision than in the scenario where it evolves on its own.
+Given the trends in population proportions, one expects this to eventually be eliminated when the population with no imprecision finally takes over the others, but it is interesting to observe that while populations with vague strategies persist, the population with no imprecision takes much longer to evolve a crisp strategy.
+
+Given the tight interaction between populations precluding the manifestation of temporary advantages of vagueness, we decided to test a variation of the model where populations interact more loosely.
 

### 2017-10-27 book_chapter/paper.tex
Added discussion of results for loose interaction model. Cleaned up some plots.

@@
 .
-%Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $EU(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}
+%Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $\text{EU}(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}
 % The dichotomy between procedural and instrumental views of rationality is not an exclusive one, since usually both means and ends are taking into account, but it helps us describe where the emphasis is placed on.
@@
 Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
-The expected utility $EU$ of a strategy can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
-Concretely, the expected utility of $\sigma$ given $\rho$ is $EU_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $EU_S(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
+The expected utility $\text{EU}$ of a strategy can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
+Concretely, the expected utility of $\sigma$ given $\rho$ is $\text{EU}_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $\text{EU}_S(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
 The expected u %, \emph{i.e.}:
 % $$
-% EU(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
+% \text{EU}(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
 % $$
@@
 
-\section{How could language not be vague?}
+\section{The ecology of vagueness}
 \label{sec:finite-experience}
@@
 
-In order to explore these considerations further, in the remainder of the paper we propose three variations of a multi-population model of imprecise imitation dynamics.
+In order to explore these considerations further, in the remainder of the paper we propose two variations of a multi-population model of imprecise imitation dynamics.
 The idea is to investigate how populations with different levels of imprecision interact, and to what extent does the hypothesized temporary advantage of faster convergence helps vagueness to thrive in such an environment.
@@
 $$
-\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)EU_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
+\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
 $$
@@
 $$
-\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)EU_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
+\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
 $$
@@
 Levels of imprecision are thus subject to an evolutionary dynamic that is indirectly influenced by the cultural dynamic.
-\todo{genotype vs phenotype?}
+%\todo{genotype vs phenotype?}
 Importantly, only the level of imprecision is passed on to new generations under this dynamic, not the actual strategies developed by the agents at the cultural level.
@@
 $$
-P_{i+1}(\alpha) \propto P_i(\alpha)(EU_{\sigma_{i+1}}^{\alpha}(m,t,\rho_{i}^{\star}) + EU_{\rho_{i+1}}^{\alpha}(t,m,\sigma_{i}^{\star}))
+P_{i+1}(\alpha) \propto P_i(\alpha)(\text{EU}_{\sigma_{i+1}}^{\alpha}(m,t,\rho_{i}^{\star}) + \text{EU}_{\rho_{i+1}}^{\alpha}(t,m,\sigma_{i}^{\star}))
 $$
@@
   \centering
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-example-0-005-strong.png}
-%    \caption{Crisp}
-%    \label{fig:example_stratsA}
+  \begin{subfigure}[]{\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-strong.png}
+    \caption{Two populations scenario.}
+    \label{fig:proportion-cases-two-tight-interaction}
   \end{subfigure}
   \hfill
-  \begin{subfigure}[]{0.45\textwidth}
-    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-example-0-005-01-strong.png}
-%    \caption{Vague}
-%    \label{fig:example_stratsB}
+  \begin{subfigure}[]{\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-strong.png}
+    \caption{Three populations scenario.}
+    \label{fig:proportion-cases-three-tight-interaction}
   \end{subfigure}
-  \caption{\todo[inline]{description}}
-  \label{fig:examples-proportion-tight-interaction}
+  \caption{Evolution of population proportions through time for each simulation trial of the tight interaction model.}
+  \label{fig:proportion-cases-tight-interaction}
 \end{figure}
-In Figure~\ref{fig:examples-proportion-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for one example trial.
-There were no qualitative differences across trials.
-As the plot shows, the population with no imprecision steadily increases its proportion against the others.
+In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for all trials.
+As the plot shows, for every trial, the population with no imprecision steadily increases its proportion against the others.
 In the three population scenario, the proportion of the population with $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
@@
 These observations go against the expectation of~\textcite{franke_vagueness_2017} that faster convergence to a convex strategy by populations with a certain level of imprecision could give them a temporary advantage to take over and eliminate other types.
-What happens in this setup is nevertheless easy to explain.
-Because the populations interact very tightly, they develop strategies in close tandem with each other.
-One of the consequences of this is that the population with no imprecision reaps some benefits from evolving together with a population with a certain level of imprecision.
-Namely, they reach convexity faster and develop more homogeneous strategies
+The reason for this is interesting in itself.
+What happens is that, because of the tight interaction between populations, the strategies of each type evolve in close tandem with each other.
+One of the consequences of this is that the population with no imprecision reaches convexity faster and develops more homogeneous strategies because of the interaction with the populations with imprecision.
+We can see the effect on convexity by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: population with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
 \begin{figure}
   \centering
-  \begin{subfigure}[]{0.3\textwidth}
+  \begin{subfigure}[]{0.32\textwidth}
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-percentage-all-strong.png}
-%    \caption{Crisp}
+    \caption{Percentage of convex sender strategies.}
     \label{fig:convexity-tight-interaction}
@@
   \hfill
-  \begin{subfigure}[]{0.3\textwidth}
+  \begin{subfigure}[]{0.32\textwidth}
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-sender-all-strong.png}
-%    \caption{Vague}
+    \caption{Mean entropy of sender strategies.}
     \label{fig:entropy-sender-tight-interaction}
@@
   \hfill
-  \begin{subfigure}[]{0.3\textwidth}
+  \begin{subfigure}[]{0.32\textwidth}
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-receiver-all-strong.png}
-%    \caption{Vague}
+    \caption{Mean entropy of receiver strategies.}
     \label{fig:entropy-receiver-tight-interaction}
   \end{subfigure}
-  \caption{\todo[inline]{description}}
+  \caption{Development of some metrics through time for the $\alpha = 0$ population in each of three scenarios: evolving alone (`0'), with an $\alpha = 0.05$ population (`0/0.05'), and with an additional $\alpha = 0.1$ population (`0/0.05/0.1').}
   \label{fig:metrics-tight-interaction}
 \end{figure}
-We can see the effect on convexity by looking to Figure~\ref{fig:convexity-tight-interaction} where we plot, for each scenario, the percentage of strategies that are convex at a given iteration, including the scenario with a population with no imprecision evolving on its own.
-These effects preclude the hypothesized temporary advantage to even manifest itself.
+This is plotted in Figure~\ref{fig:convexity-tight-interaction}.
+What we see is more trials reaching convexity earlier for the multi-population scenarios when compared with the single-poulation scenario.
+This effect precludes the hypothesized temporary advantage of imprecision to even manifest itself, but it can be seen as a positive influence on the population with no imprecision.
 
@@
 
+\subsection{Loose population interaction}
+\label{sec:loose-interaction-model}
+In order to model a different type of interaction between populations in our multi-population model, we need to make changes to some formulations.
+Any formula that is not redefined in this section should be assumed to be the same as used for the model in Section~\ref{sec:tight-interaction-model}.
+The dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of the same type.
+First, if agents learn only within their population, the imitation dynamic needs to consider only the expected utility against agents of that population.
+We thus define the following expected utilities for a sender of type $\alpha$:
+$$
+\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\alpha})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})\sum_{t_{r}}P_{\rho}^{\alpha}(t_{r}|m)U(t_{a},t_{r})
+$$
+and for a receiver of type $\alpha$:
+$$
+\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\alpha})=\sum_{t_{a}}P_{\bar{\sigma}}^{\alpha}(t_{a}|m)\sum_{t_{r}}P_{r}^{\alpha}(t_{r}|t_{i})U(t_{a},t_{r})
+$$
+The main difference is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type.
+
+We can redefine the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
+$$
+P_{o}^{\alpha}(m|t_{o})=\sum_{t_{a}}P_{\bar{o}}^{\alpha}(t_{a}|t_{o})P_{\sigma}^{\alpha}(m|t_{a})
+$$
+and the probability that a receiver of type $\alpha$ observes a randomly sampled agent choose interpretation $t_o$ given message $m$ as:
+$$
+P_{o}^{\alpha}(t_{o}|m)=\sum_{t_{r}}P_{o}^{\alpha}(t_{o}|t_{r})P_{\rho}^{\alpha}(t_{r}|m)
+$$
+Again, the main difference is that agents make observations within their own population, so to model a randomly sampled agent one needs only to take into account the same type.
+
+Update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ is then given by:
+$$
+\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\alpha})
+$$
+and similarly for a receiver strategy of type $\alpha$ by:
+$$
+\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\alpha})
+$$
+
+The selection process between different populations still follows the same motivation: agents of a certain type that evolve successful strategies (with respect to other types) will be more likely to survive and reproduce, benifitting the proportion of their population.
+The formulation of the dynamic of the proportion of each type $P(\alpha)$ thus stays the same.
+In order to avoid confusion, we want to stress that this means that these calculations thus still rely on the calculation of expected utilities across populations (\emph{i.e.}~$\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})$) and not the newly introduced $\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\alpha})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\alpha})$.
+
+We ran the same number of simulation trials under the same conditions as for the model with tight population interaction.
+\begin{figure}
+  \centering
+  \begin{subfigure}[]{\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-weakest.png}
+    \caption{Two populations scenario.}
+    \label{fig:proportion-cases-two-loose-interaction}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-weakest.png}
+    \caption{Three populations scenario.}
+    \label{fig:proportion-cases-three-loose-interaction}
+  \end{subfigure}
+  \caption{Evolution of population proportions through time for each simulation trial of the loose interaction model.}
+  \label{fig:proportion-cases-loose-interaction}
+\end{figure}
+In Figure~\ref{fig:proportion-cases-loose-interaction} we plot the evolution of population proportions for all trials of the multi-population scenarios.
+The first thing to observe is that proportions evolve faster than in the model with tight interaction.
+Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in the one population fully dominating the others.
+In the two-population scenario, some population reached 99\% proportion in 24 out of 25 trials.
+In the three-population scenario, this happened in 18 out of 25 trials.
+Given that strategies no longer develop in close tandem with each other, differences in expected utility can be more salient resulting in some population reaching a more clear advantage than the other.
+
+More interestingly for our investigation, some trials actually resulted in a population with some degree of imprecision dominating the others.
+For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials, a point from which it is technically impossible for the other population to recover.
+In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
+In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then start losing ground to $\alpha = 0$.
+For the three population scenario, $\alpha = 0.05$ fully reached 100\% proportion in only 1 trial and $\alpha = 0.1$ never reached more than 59\%.
+In 5 other trials we see a positive trend for $\alpha = 0.05$, but given the reversals that sometimes occur (\emph{e.g.}~trial 171), one cannot be sure whether full dominance would be achieved if the simulation was left to run for longer time.
+
 \section{Conclusion}
 \label{sec:conclusion}
-
+\todo[inline]{Extend/Rewrite}
 We tried to show that, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.

### 2017-10-27 book_chapter/simulation/analysis.R

@@ -1,0 +1,110 @@
+library(tidyverse)
+
+read.data <- function() {
+    measurement_data <<- tibble()
+    for (fname in dir('.', pattern = '*.csv')) {
+        x <- read_csv(fname)
+        x$sender.converged <- parse_logical(toupper(x$sender.converged))
+        x$receiver.converged <- parse_logical(toupper(x$receiver.converged))
+        x$file <- fname
+        x$populations <- as.character(summarise(x, paste(unique(impairment), collapse='/')))
+        x$number.of.populations <- as.integer(summarise(x, length(unique(impairment))))
+        measurement_data <<- rbind(measurement_data, x)
+    }
+    measurement_data$sim.id <<- as.integer(factor(measurement_data$file))
+}
+
+plot.proportion <- function(population.scenario, interaction.type, simulation.id=NULL) {
+    dt <- filter(measurement_data, populations == population.scenario, interaction == interaction.type)
+    if (!is.null(simulation.id)) {
+        dt <- filter(dt, sim.id == simulation.id)
+    }
+    p <- ggplot(dt, aes(x=iteration, y=proportion)) +
+        geom_line(aes(linetype=factor(impairment))) +
+        labs(x=expression(i), y=expression(P(alpha)), linetype=expression(alpha))
+    if (is.null(simulation.id)) {
+        p <- p + facet_grid(. ~ sim.id) +
+            scale_x_continuous(breaks=c(100,200)) + theme(axis.text.x = element_text(angle=45))
+    }
+    return(p)
+}
+
+plot.convex.percentage <- function(interaction.type) {
+    dt <- filter(measurement_data, interaction == interaction.type | populations == '0', impairment == 0)
+    p <- ggplot(dt, aes(x=iteration)) +
+        stat_summary(fun.y='mean', geom='line', aes(y=sender.convex, linetype=factor(populations))) +
+        labs(x=expression(i), y='Convex percentage', linetype='Scenario') + theme_bw()
+    return(p)
+}
+
+plot.sender.entropy <- function(interaction.type) {
+    dt <- filter(measurement_data, interaction == interaction.type | populations == '0', impairment == 0)
+    p <- ggplot(dt, aes(x=iteration)) +
+        # geom_point(alpha=0.1) +
+        stat_summary(fun.y='mean', geom='line', aes(y=sender.entropy, linetype=factor(populations))) +
[…]

### 2017-10-30 book_chapter/paper.tex
Reworked parts of section 4.

@@
 
-\usepackage[natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
+\usepackage[backend=biber,natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
 
+\usepackage[utf8]{inputenc}
 \usepackage{todonotes}
@@
 %We draw attention to two additional aspects of rationality which will be helpful in the following.
-Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
+Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
 A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
@@
 The vocabulary just introduced will, we hope, help inform the discussion that follows.
+\todo[inline]{Review}
 We proceed by introducing the framework of signaling games in Section~\ref{sec:signaling-games}.
@@
 Section~\ref{sec:finite-experience} reflects on the implications of these explanations for considerations of rationality.
-Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
+%Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
 
@@
 
-Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that a vague language is suboptimal compared to a crisp one, work under a number of various assumptions.
-Together, these assumptions form a picture of the context where agents learn and use language that is highly idealized and probably derives, via game theory, from the conception of rationality of traditional theoretical economics.
-%This picture has been criticized for several decades, at least in the field of theoretical economics.
+Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that vague signal use is suboptimal when compared to a crisp one, work under a number of various assumptions.
+Part of the picture formed by these assumptions is a highly idealized conception of the agents involved and of the context in which they develop and use signals.
+These idealizations probably originate, via game theory, from the conception of rationality of traditional theoretical economics.
 Herbert A.~Simon describes this picture as follows:
@@
 \end{quote}
-This ``economic man'', a game-omniscient and highly strategic, instrumentally rational agent, seems to have survived almost unscathed to now populate most game theoretic models of language and signaling.
-Both Simon and Lipman call for this picture to be revised.
-
-In order to account for vagueness in natural language in the context of these models, the proposals surveyed here all need to peel away from this idealized picture and bring some of these assumptions down to earth.
+%This ``economic man'', a game-omniscient and highly strategic, instrumentally rational agent, seems to have survived almost unscathed to now populate most game theoretic models of language and signaling.
+Both Simon and Lipman call for this picture to be revised, and this is what the proposals surveyed here all do.
+In order to account for vagueness in natural language in the context of these models, they peel away from this idealized picture and bring some of these assumptions down to earth.
 In the process, they point to ways in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world:
@@
 Generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, both paint a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
-Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
-All approaches introduce constraints on agent behavior or information processing that prevent the evolution of crisp signal use.
+%Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
+These approaches introduce constraints on agent behavior or information processing that prevent the evolution of crisp signal use.
 But a crisp language would still have a higher expected utility than the evolved strategies.
 Agents in those models seem to be rational only to the extent of their capabilities.
-Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), this feels like a bittersweet conclusion because of the hypothetical possibility of an ideal strategy that seems to be only artificially barred from the agents.
-But this is so only if rationality is equated with maximization of expected utility on a pure local level.
+Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), this feels like a bittersweet conclusion because of the hypothetical possibility of an ideal strategy, seemingly barred from the agents in an artificial manner.
+Couldn't a more rational agent evolve and drive the system into crisp signal use?
+Aren't we, human beings, that kind of agent?
+
+Perhaps a deeper understanding of vagueness and the reasons for its pervasiveness in natural language are to be found only when we broaden the scope of the models employed.
+We call this the \emph{ecology} of vagueness.
+All the models discussed so far explore evolutionary dynamics for one homogeneous population playing one game.
+Different types of agents and different game setups are considered, but each of these different possible scenarios has so far been tested separately.
+We see at least two ways in which one could have a more ecological perspective.
+The first is to think about meaning and vagueness in a universe of multiple language games.
+We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
+These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[\S 23]{wittgenstein_philosophical_1953}.
+We could look for meaning at several levels in this pluralistic picture.
+Firstly, as before, there is the actual behavior of a single agent in each actualized language game.
+As mentioned above in connection the soft-max choice function used by \textcite{franke_vagueness_2011}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic.
+%% \mf{mention ``bounded rationality'' here}
+Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
+Notice that a single agent's rational beliefs about linguistic practices or linguistic meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
+In sum, both at the level of behavior and at the level of beliefs about use or meaning, we should expect to find vagueness.
+Still, despite the natural vagueness, there does not seem to be anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
+
+Another way to go beyond locality is to work with more heterogeneous population models.
 The mechanisms that lead to vague signal use, as~\textcite{oconnor_evolution_2014} and~\textcite{franke_vagueness_2017} stress, have the aforementioned important advantages of faster speed of convergence, higher flexibility, and homogenization.
-These advantages matter, especially in the complex, dynamic context of real language use.
-Limits on memory and precision in the calculation of complex probabilistic beliefs may have a reduced metabolic processing cost.
-One would thus like to say that using a strategy that promotes those characteristics without a significant loss of expected utility is certainly a rational choice.
-But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond maximization of expected utility, or a notion of expected utility that incorporates other dimensions.
-It is in this sense that these explanations tie in with a notion of bounded ecological rationality.
-
-For more complex agents, the story is potentially more nuanced.
-Consider an agent that has awareness that something like a game is being played, some estimate of the potential payoff involved, ability to gauge other players' expected behaviors, and the capacity to think strategically.
-Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
-In a scenario where the agent's ability to make the optimal choice is bounded, such as in the models of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
-\textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
-The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as local utility maximization.
-But even in other scenarios, we could again argue that, if we broaden our notion of rationality, it would be less rational to use a crisp strategy rather than one that incorporates at least some source of uncertainty.
-A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
-Though potentially a locally rational choice, it might thus not be globally optimal.
-
-In order to explore these considerations further, in the remainder of the paper we propose two variations of a multi-population model of imprecise imitation dynamics.
-The idea is to investigate how populations with different levels of imprecision interact, and to what extent does the hypothesized temporary advantage of faster convergence helps vagueness to thrive in such an environment.
+The argument goes that these side-effects, by temporarily enabling a higher expected utility, could allow a population using some generalization (or affected by some imprecision) to take over.
+However, despite the intuitiveness of the argument, it is based on comparing isolated runs of different dynamics.
+The models do not allow the hypothesis to be tested, because they do not accommodate different populations evolving together.
+%These advantages matter, especially in the complex, dynamic context of real language use.
+%Limits on memory and precision in the calculation of complex probabilistic beliefs may have a reduced metabolic processing cost.
+%One would thus like to say that using a strategy that promotes those characteristics without a significant loss of expected utility is certainly a rational choice.
+%But in order to be able to say that in the context of the model, we need either a notion of rationality that goes beyond local maximization of expected utility, or a notion of expected utility that incorporates other dimensions.
+%It is in this sense that these explanations tie in with a notion of bounded ecological rationality.
+%
+%For more complex agents, the story is potentially more nuanced.
+%Consider an agent that has awareness that something like a game is being played, some estimate of the potential payoff involved, ability to gauge other players' expected behaviors, and the capacity to think strategically.
+%Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
+%In a scenario where the agent's ability to make the optimal choice is bounded, such as in the models of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
+%\textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
+%The authors' analysis shows that such an approach would be irrational in a number of different scenarios, even in the narrow interpretation of rationality as local utility maximization.
+%But even in other scenarios, we could again argue that, if we broaden our notion of rationality, it would be less rational to use a crisp strategy rather than one that incorporates at least some source of uncertainty.
+%A crisp strategy for a particular game would, in most realistic scenarios, take more time to achieve convergence, be less adaptable to changes, less easy to reuse in other similar games, and less able to foster cooperation with a bigger group of agents.
+%Though potentially a locally rational choice, it might thus not be globally optimal.
+%
+%In the context of signaling games, thinking about vagueness from a more ecological perspective means broadening our models to incorporate more diversity and allow the study of additional interactions within the model.
+%There are many ways in which this can be done.
+In the remainder of the paper we propose a way to do this for the model of~\textcite{franke_vagueness_2017}.
+We introduce two variations of a multi-population model of the imprecise imitation dynamics, where populations characterized by different imprecision values interact and evolve together.
+Using this model, we can better see under which conditions the hypothesized advantages of some imprecision can lead to the evolution of vagueness.
 

### 2017-10-30 book_chapter/paper.tex
Trying to improve the text in section 5.

@@
 In order to account for vagueness in natural language in the context of these models, they peel away from this idealized picture and bring some of these assumptions down to earth.
-In the process, they point to ways in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic world:
+In the process, they point to ways in which we, as language learners and language users, are finite beings finding ways to cope with a highly complex and dynamic environment:
 \begin{enumerate}
 \item Our existence is temporally finite; language does not have an infinite amount of time to evolve, nor can it take an infinite time to be learned.
-The faster a language can start being useful, the better.
+The faster a language can start being useful, the better;
 \item Language learning through experience has to rely on a limited number of observations.
-Not only is the state space typically much larger than what one can survey in sufficient time, it is even potentially infinite and constantly changing.
+Not only is the state space typically much larger than what one can survey in sufficient time, it is even potentially infinite and constantly changing;
 \item A corollary of the former is that there will always be heterogeneity in a population of language learners, at the very least in their prior experience, since each agent will have relied on a different set of observations.
-Furthermore, this information is not directly or fully accessible to others.
+Furthermore, this information is not directly or fully accessible to others;
 \item Given that an agent is almost always learning and using language in a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice.
@@
 \label{sec:multi-population-model}
+We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations, with different imprecision parameters, evolving together.
+In order to do this, first we consider that each population evolves according to the imprecise imitation dynamics.
+However, we want to model populations that are able to interact with each other.
+Therefore, the dynamic needs to be adapted in order to reflect the fact that agents can hypothetically imitate and play against agents with different imprecision levels.
+Second, in order to see the impact of the success of different imprecision levels, we need to add another process of selection.
+The motivation is as follows: agents are born with varying perceptual abilities (modeled as populations with different imprecision values), they learn a signaling strategy by imitation of other agents (either within or across populations), depending on the success of the strategies they develop, they will be more or less likely to survive and reproduce, thus the proportion of each population changes depending on its fitness compared to other populations.
 
-We build upon the model of~\textcite{franke_vagueness_2017} by adding the possibility for more than one population, where each population has its own imprecision parameter.
-The dynamic needs to be adapted to allow different types of interactions between these populations, namely regarding which agents will be imitated, and which agents will be expected to play against.
-Finally, in order to see the impact of the success of different imprecision levels, we add another process of selection.
-Imagine the following: agents are born with varying perceptual abilities (modeled as populations with different imprecision values), they learn a signaling strategy by imitation of other agents (either within or across populations), depending on the success of the strategies they develop, they will be more or less likely to survive and reproduce, thus the proportion of each population will change depending on their fitness compared to other populations.
-
-Let $A$ be the set of imprecision values considered.
-For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$ such that $\sum_{\alpha \in A} P(\alpha) = 1$.
+Formally, let $A$ be the set of imprecision values considered.
+For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$, such that $\sum_{\alpha \in A} P(\alpha) = 1$.
 The probability that a given agent with imprecision $\alpha$ (or of type $\alpha$) observes $t_o$ when the actual state is $t_a$ is given by $P_{o}^{\alpha}(t_{o}|t_{a})$.
 If the same agent intends to realize $t_i$, the probability that he actually realizes $t_r$ instead is given by $P_{r}^{\alpha}(t_{r}|t_{i})$.
-Following~\textcite[26]{franke_vagueness_2017}, we can define the following values.
+Following~\textcite[26]{franke_vagueness_2017}, we can then define the following values.
 Probability that $t_a$ is actual if $t_o$ is observed by an agent of type $\alpha$:
@@
 $$
-P_{\rho}^{\alpha}(t_{r}|m)=\sum_{t_{i}}P_{r}^{\alpha}(t_{r}|t_{i})\rho^{\alpha}(t_{i}|m)
+P_{\bar{\sigma}}^{\alpha}(t_{a}|m)\propto P_{a}(t_{a})P_{\sigma}^{\alpha}(m|t_{a})
 $$
@@
 $$
-P_{\bar{\sigma}}^{\alpha}(t_{a}|m)\propto P_{a}(t_{a})P_{\sigma}^{\alpha}(m|t_{a})
+P_{\rho}^{\alpha}(t_{r}|m)=\sum_{t_{i}}P_{r}^{\alpha}(t_{r}|t_{i})\rho^{\alpha}(t_{i}|m)
 $$
-
-Given these probabilities, we can compute expected utilities and update probabilities for each type.
-We have a number of possible types of interaction between populations.
+These formulations are merely parameterized versions of the single-population model.
+They encapsulate calculations that we can then use to compute expected utilities and update probabilities for each type.
+The latter depend on the types of interaction that we imagine occurring between populations.
+In the following sections, we consider two different types.
 
@@
 \label{sec:tight-interaction-model}
-The first we will consider is a tight one, where each agent plays with, observes, and potentially imitates any other agent, regardless of their type.
-Expected utilities for this case are given, for a sender of type $\alpha$, by:
+In a multi-population model with tight interaction between populations, each agent plays with, observes, and potentially imitates any other agent, regardless of their type.
+This has an impact on the expected utilities of sender and receiver strategies of each type, and on the update probabilities for those strategies.
+Let's start with the expected utilities.
+For a sender of type $\alpha$, the expected utility of its strategy $\sigma^\alpha$ against all other receiver strategies $\rho^\star$, is given by:
 $$
@@
 $$
-and, for a receiver of type $\alpha$, by:
+and, for a receiver of type $\alpha$, the expected utility of its strategy $\rho^\alpha$ against all other senderstrategies $\sigma^\star$, is given by:
 $$
@@
 $$
+Expected utilities thus take into account the existence of strategies of other types, and weigh the relevance of each type $\alpha^\prime$ according to its relative proportion $P(\alpha^\prime)$.
 
-Furthermore, we can define the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
+Another important value to calculate has to do with the types which agents observe and imitate.
+In a model with tight interaction, we imagine that this occurring across populations.
+Therefore, we can define the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
 $$
@@
 $$
+Thus, these calculation also incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
 
-Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ is given by:
+Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ are given by:
 $$
@@
 
-The formulations so far cover the evolution of the particular strategies of agents.
+These formulations cover the evolution of the particular strategies of each type.
 We can think of this as the level of cultural evolution: agents are born with a certain level of imprecision and learn strategies based on the behavior of others.
 Alongside this process, we can imagine another level of selection, where agents die and new agents are born.
-More successful agents will have a higher likelihood of surviving and reproducing, giving rise to more agents with their level of imprecision.
+More successful agents have a higher likelihood of surviving and reproducing, giving rise to more agents with their level of imprecision.
 Levels of imprecision are thus subject to an evolutionary dynamic that is indirectly influenced by the cultural dynamic.
-%\todo{genotype vs phenotype?}
 Importantly, only the level of imprecision is passed on to new generations under this dynamic, not the actual strategies developed by the agents at the cultural level.
@@
 We model this process by changing the proportion of each type $P(\alpha)$ according to the replicator dynamic.
-A population of type $\alpha$ consists of agents employing both a sender and a receiver strategy, thus the overall fitness of the population must involve the expected utilities of each.
+A population of type $\alpha$ consists of agents employing both a sender and a receiver strategy, thus the overall fitness of the population must include the expected utilities of both.
 We could imagine this process happening at a different speed than the cultural process, in which case we could have different time scales.
-For the sake of simplicity, we choose to have them both happen at each time step, but calculate the changes occurring at this level of selection after the other level.
+For the sake of simplicity, we choose to have them both happen at each time step, but calculate the changes occurring at this level of selection after the calculation for the other level.
 We define the proportion of type $\alpha$ at time step $i+1$ as:

### 2017-11-03 book_chapter/paper.tex
Shortened and adapted introduction.

@@
 
-\title{Natural sources of vagueness and their implications}
+\title{Towards an ecology of vagueness}
 \author{Jos\'e Pedro Correia \and Michael Franke}
@@
 Since vagueness is persistent in all human languages, something has to give.
-In this paper, we investigate a number of reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary optimal signaling.
-We show how uncertainty about not only the linguistic practices of others, but also about the world itself can lead to vagueness.
-We explore the consequences of these reasons and mechanisms for a notion of meaning and a notion of rationality.
+In this paper, we focus on this problem in the context of signaling games.
+We provide an overview of a number of proposed reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary signaling.
+Most of them argue that vague signal use is simply the best we can get, given certain factors.
+Despite the plausibility of the proposals, we argue that a deeper understanding of the benefits of vagueness needs a more ecological perspective, namely one that goes beyond local optimization of signaling strategies of a homogeneous population.
+As an example of one possible way to expand upon our current models, we propose two variants of a novel multi-population dynamic of imprecise imitation where, under certain conditions, populations with a certain level of imprecision dominate over precise ones.
 \end{abstract}
@@
 Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
-The problem thus seems to be a wide-ranging one, but it does not obviously undermine our everyday linguistic practices.
+If we think of language as governed by logical rules, and of rationality as the ability to follow those rules, the sorites paradox seems \emph{prima facie} to show that vagueness in our language can undermine our rationality.
+
+But we can think of language in different terms.
+A possibility is to think of signs as tools agents use to coordinate actions.
+An example of a way to formally study language along those lines is by using the framework of signaling games~\parencite{lewis_convention_1969}.
+In such a context, rationality becomes the ability to choose the use of signs that is optimal to achieve the desired coordination.
+%So what about vagueness?
+However, in the context of these game-theoretic approaches to language, vagueness presents a new challenge.
+Barton Lipman~\parencite*{lipman_why_2009} gives a detailed account of the problem.
+Very succinctly, it can be put as follows.
+In standard game-theoretic models of communication, vague signal use yields a lower expected utility than crisp use.
+Therefore, given that the dynamics (be it natural selection, cultural evolution, or rational choice) maximize utility, vagueness should be weeded out by these forces, giving rise to only precise languages.
+Thus it would be irrational to stick to vague language use when we could (theoretically) switch to a better system.
+But as we said before, vagueness is pervasive in our natural languages, and there is no reason to believe it is going away.
+The problem seems to be a theoretically serious one, but it does not obviously undermine our everyday linguistic practices.
 Most of the time we seem to communicate just fine.
 Therefore, the issue must lie with the conceptions we have of the forces or mechanisms that underlie those practices.
+Lipman concludes that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}.
 
-Vagueness is typically seen as a challenge to a classical conception of language that relies on reference and truth as two core notions for understanding meaning.
-One problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to something like a well defined set of tall people?
-And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding these borderline cases?
-Supervaluationism, many-valued logics, and degree theories all propose changes to the classical picture in order to accommodate for vague predicates.
-They are, however, still strongly committed to its two core notions.
-As argued by Mark Sainsbury~\parencite*{sainsbury_concepts_1999}, because of that they fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
-All the aforementioned proposals end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
-But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
-Further introducing borderline borderline cases would lead to an infinite regress.
-Because of their attachment to the classical picture, the standard approaches to vagueness also fail to draw an important lesson from the sorites paradox, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
+%Vagueness is typically seen as a challenge to a classical conception of language that relies on reference and truth as two core notions for understanding meaning.
+%One problem is that vague predicates seem to lack precise boundaries; if that is the case, how can words like `tall' stand in correspondence to something like a well defined set of tall people?
+%And if we cannot know exactly whether a certain person is `tall' or not, as with borderline cases, how are we to determine the truth value of sentences that involve statements of tallness regarding these borderline cases?
+%Supervaluationism, many-valued logics, and degree theories all propose changes to the classical picture in order to accommodate for vague predicates.
+%They are, however, still strongly committed to its two core notions.
+%As argued by Mark Sainsbury~\parencite*{sainsbury_concepts_1999}, because of that they fail to address an important characteristic of vague predicates: \emph{higher order vagueness}.
+%All the aforementioned proposals end up being committed to new artificial demarcating boundaries (\emph{e.g.}~true-under-all-precisifications versus neither true nor false versus false-under-all-precisifications, true versus indefinite versus false, true to degree 1 versus true to degree 0 versus the rest).
+%But a vague predicate not only fails to demarcate between the cases where it clearly applies and the ones where it clearly doesn't, it also fails to establish a boundary between the cases where it clearly applies and the borderline cases, as well as between the borderline cases and the cases where it clearly doesn't apply.
+%Further introducing borderline borderline cases would lead to an infinite regress.
+%Because of their attachment to the classical picture, the standard approaches to vagueness also fail to draw an important lesson from the sorites paradox, namely that ``we do not know, cannot know, and do not need to know these supposed boundaries to use language correctly''~\parencite*[256]{sainsbury_concepts_1999}.
 % \begin{quote}
@@
 % \end{quote}
-By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~they are able to recognize the logical inconsistency but do not have a good answer as to how to overcome it.
-Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language to achieve their daily goals.
-Nobody ever stopped using the word `tall` after being confronted with a sorites series.
-Why should we develop theories of meaning that are impervious to the paradox?
+%By trying to cling as much as possible to the classical picture of logic and semantics, these standard approaches are ignoring a simple observation: natural language users are sensitive to the sorites paradox, \emph{i.e.}~they are able to recognize the logical inconsistency but do not have a good answer as to how to overcome it.
+%Even more importantly, they apparently do not need to solve the inconsistency in order to continue using natural language to achieve their daily goals.
+%Nobody ever stopped using the word `tall` after being confronted with a sorites series.
+%Why should we develop theories of meaning that are impervious to the paradox?
+%
+%The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also need to be abandoned as a consequence.
+%One notion that seems to quickly be in peril is that of rationality.
+%In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
+%\begin{quote}
+%In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
+%[\ldots]
+%The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
+%~\parencite*[1]{rorty_response_2000-1}
+%\end{quote}
+%By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
+%Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
+%Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions.
+% (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
+%Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
+%This type of rationality is one that supposedly demarcates humans from other animals.
 
-The reluctance to give up truth and logic as valuable notions to explain meaning is perhaps associated with the fear of what would also need to be abandoned as a consequence.
-One notion that seems to quickly be in peril is that of rationality.
-In reference to philosophers who defend the desirability of a classical notion of truth, Richard Rorty says:
-\begin{quote}
-In the past, such philosophers have typically conjoined the claim that there is universal human agreement on the supreme desirability of truth with two further premises: that truth is correspondence to reality, and that reality has an intrinsic nature (that there is, in Nelson Goodman's terms, a Way the World Is).
-[\ldots]
-The rise of relatively democratic, relatively tolerant, societies in the last few hundred years is said to be due to the increased rationality of modern times, where `rationality' denotes the employment of an innate truth-oriented faculty.%
-~\parencite*[1]{rorty_response_2000-1}
-\end{quote}
-By tying all of these notions tightly together, the fear could be that by dropping the picture of meaning as intimately tied to truth and logic we necessarily lose the ground on which rationality stands, and the whole edifice could collapse.
-Rationality, in this picture, has truth as its guiding light and logic as the means to attain it.
-Being rational is about following universally valid rules of reasoning that, given the correct inputs, necessarily lead us to the correct conclusions.
-% (see, for example, Harold Brown~\parencite*[19]{brown_rationality_1990} for a characterization).
-Although humans are not necessarily consciously aware of the rules, they are considered to be nevertheless unconsciously following them when they act rationally.
-This type of rationality is one that supposedly demarcates humans from other animals.
-
-We want to argue here that to say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
-Giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
-It also does not necessarily eradicate all hope for a (formal) explication of meaning.
-In order to make this argument, we need a different paradigm of language and meaning.
-But what could such a paradigm even be?
-
-We suggest that inspiration can be drawn from the later work of philosopher Ludwig Wittgenstein%
-\footnote{To the extent that substantial views can be said to be defended by the author.
+%We want to argue here that to say that language does not follow strict logical rules, or that truth is not a useful notion to guide our inquiries into meaning, is not to say that language is unstructured, meaningless, or unusable; neither is it to say that we, language users, are therefore irrational.
+%Giving up the ideal of truth and logic as relevant explanatory notions for understanding meaning in natural language does not imply giving up on rationality altogether.
+%It also does not necessarily eradicate all hope for a (formal) explication of meaning.
+%In order to make this argument, we need a different paradigm of language and meaning.
+%But what could such a paradigm even be?
+%
+%We suggest that inspiration can be drawn from the later work of philosopher Ludwig Wittgenstein%
+%\footnote{To the extent that substantial views can be said to be defended by the author.
 Skipping over the debate (but see~\cite{kahane_wittgenstein_2007} for more details), we are here assuming an interpretation of Wittgenstein as a kind of pragmatist, along the lines of the readings of Hilary Putnam~\parencite*{putnam_pragmatism_1994} and Richard Rorty~\parencite*{rorty_wittgenstein_2007}.}.
-His \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} adumbrate a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
-Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
-These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
-%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
-There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
-The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
-Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
-It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
-From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.
+%His \emph{Philosophical Investigations}~\parencite*{wittgenstein_philosophical_1953} adumbrate a characterization of language that departs from the classical picture by highlighting its diversity, heterogeneity, and dynamism.
+%Language is seen as a patchwork of various language-games, with new ones continuously being added and old ones falling out of use.
+%These language-games, in turn, can be thought of as the use of signs in the context of and intermingled with a practice.
+%%The metaphor emphasizes a concept of meaning as strongly linked to the use that words or signs are put to.
+%There is no meaning in an abstract atemporal world, it is rather created dynamically and in a way that is contingent to the practices we engage in, to the language-games we play.
+%The notion of language-game is also used to characterize one of Wittgenstein's methodological tools.
+%Following the idea that ``[i]t disperses the fog if we study the phenomena of language in primitive kinds of use in which one can clearly survey the purpose and functioning of the words''~\parencite*[\S 5]{wittgenstein_philosophical_1953}, the method is to set up a language-game as a hypothetical scenario, or thought experiment, where language is used in a certain type of activity, and then reflect on the assumptions that underlie our interpretation of this set-up, as well as on how the scenario would play out according to those assumptions.
+%It has been argued~\parencite{correia_bivalent_2013} that the framework of \emph{signaling games}, introduced to the study of meaning by David Lewis~\parencite*{lewis_convention_1969} and later naturalized by Brian Skyrms~\parencite*{skyrms_evolution_1996}, permits us to do exactly that while reaping the benefits of a formalism that enables the in-depth exploration of the implications of complex hypothetical language-games via mathematical analysis or computer simulation.
+%From this perspective, creating a signaling game model is like setting up a language-game, only using a formulation that improves perspicuity, and conducting mathematical analysis or computer simulations is like contemplating how the game would play out according to the assumptions built into the model.
 
 Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
-Before we start, we want to establish some vocabulary that will be useful later on.
-%
-%
-%\section{Facets of rationality}
-%\label{sec:facets-rationality}
-%
+But before we start, we need to establish some vocabulary to better frame our discussion.
 Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
@@
 We can start by considering what is deemed to be most important when characterizing the rationality of a given choice.
-The aforementioned classical picture of language and meaning is focused on dichotomies like true and false, meaningful and meaningless, correct and incorrect.
+The aforementioned logical picture of language and meaning is focused on dichotomies like true and false, meaningful and meaningless, correct and incorrect.
 A sentence can be true or false only if it is meaningful, and it is meaningful if it is constructed according to correct rules.
 Rationality is intimately connected with the ability to follow good procedures, not only of sentence production, but ultimately of sentence combination and reasoning (think of what is required for making a logically valid deductive inference).
-A \emph{procedural} account of rationality thus focuses on the means, rather than the ends.
+This is what we could call a \emph{procedural} account of rationality, one which focuses more on the means rather than the ends.
 One can also do the opposite and focus on the consequences instead.
 \emph{Instrumental} rationality, as it is typically called, characterizes an agent's choice as rational if it maximizes the possibility of achieving a desired goal, regardless of the means.
-The notion is linked to David Hume~\parencite*{hume_treatise_1738}, epitomized in his assertion that ``Reason is, and ought to only be the slave of the passions.''
+The notion is linked to David Hume~\parencite*{hume_treatise_1738}, epitomized in the following assertion: ``Reason is, and ought to only be the slave of the passions.''
 Instrumental rationality is close to a notion of rational choice that is used in economics and game theory%
@@
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
-Models working within the classical picture typically do not make a distinction between modeler and agent, thus there is no room to express any epistemic gaps.
+Models working within the logical picture typically do not make a distinction between modeler and agent, thus lacking room to express any epistemic gaps.
 It follows that agents can only be assumed to be omniscient of the details involved.
@@
 We can talk about \emph{bounded} rationality to characterize the choices of such agents.
-
+%
 % \todo[inline]{Following discussion is already drawing conclusions, and is too closely tied to notions that haven't been introduced yet}
@@
 % It is easier to gesture at what it is not: ignoring observational evidence; failing to draw simple and obvious conclusions; failing to make simple and obvious generalizations etc.
-
+%
 %We draw attention to two additional aspects of rationality which will be helpful in the following.
-Where the above definition of instrumental rationality is usually understood as requiring a single choice of act to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
+Where the above definition of instrumental rationality is usually understood as requiring a single choice to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
 A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
 When considering only a single situation, rationality can only be \emph{local}.
-If we take into account the possibility of the agent's choice or choice mechanism hinging on multiple situations, we can also talk about \emph{global} or \emph{ecological} rationality.
+If we take into account the possibility of the agent's choice or choice mechanism hinging on multiple situations, we can also talk about \emph{global} rationality.
 This can be important because, hypothetically, there could be choice mechanisms that are sub-optimal at a local level (for each situation) but are actually perfectly rational at a global level.
@@
 Note that we consider that, in theory, most combinations of these aspects are possible.
-Although we have been pinning logic to the classical picture, this is only with the most traditional logical systems in mind.
+Although we have been pinning procedural rationality to the logical picture of language, this is only with the most traditional logical systems in mind.
 We are not denying that advances in dynamic, epistemic, fuzzy, paraconsistent, and other types of logic could potentially enable one to capture procedural rationality with different characteristics.
-Game-theoretical models of language can, on the other hand, also assume local instrumental rationality with omniscient highly strategic agents.
+Game-theoretical models of language can, on the other hand, also combine local instrumental rationality with omniscient highly strategic agents.
 Our objective here is not to survey all the possibilities.
@@
 The vocabulary just introduced will, we hope, help inform the discussion that follows.
-\todo[inline]{Review}
 We proceed by introducing the framework of signaling games in Section~\ref{sec:signaling-games}.
 In Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a particular kind of signaling game.
-Section~\ref{sec:finite-experience} reflects on the implications of these explanations for considerations of rationality.
+Section~\ref{sec:ecology-of-vagueness} tries to generalize the considerations of these proposals to argue for a more ecological approach to vagueness.
+We propose and analyse the results of a novel multi-population model of imprecise imitation in Section~\ref{sec:multi-population-model}, and summarize our conclusions in Section~\ref{sec:conclusions}.
 %Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
@@
 More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.
-
-%So what about vagueness?
-In the context of game-theoretic approaches to language, vagueness presents a new challenge.
-Barton Lipman~\parencite*{lipman_why_2009} gives a detailed account of the problem.
-Very succinctly, it can be put as follows.
-In standard game-theoretic models of communication, vague signal use yields a lower expected utility than crisp use.
-Therefore, given that the dynamics (be it natural selection, cultural evolution, or rational choice) maximize utility, vagueness should be weeded out by these forces, giving rise to only precise languages.
-However, vagueness is pervasive in our natural languages, and there is no reason to believe it is going away.
-Lipman concludes that ``we cannot explain the prevalence of vague terms in natural language without a model of bounded rationality which is significantly different from anything in the existing literature''~\parencite*[1]{lipman_why_2009}.
-What proposals have been put forward to address this issue?
 
@@
 \section{The ecology of vagueness}
-\label{sec:finite-experience}
+\label{sec:ecology-of-vagueness}
 
@@
 $$
-In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma}$ and $\tilde{\rho}$ according to a parameter $\gamma$ defining the birth rate within the population.
-This can be described, for a certain time step $i+1$, in the following formulas:
+In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma^\alpha}$ and $\tilde{\rho^\alpha}$ (generated at each time step).
+The idea is that, at each time step, a certain percentage of each population will consist of ``newborn'' agents, \emph{i.e.}~agents that haven't yet had time to evolve their strategies.
+We define a parameter $\gamma$ that quantifies this percentage, or as we can also call it the birth rate, which we consider to be the same for every population.
+This mixing can be described, for a certain time step $i+1$, in the following formulas:
 $$
-\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}
+\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma_{i+1}^{\alpha}}
 $$
 $$
-\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}
+\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho_{i+1}^{\alpha}}
 $$
@@
 
-\section{Conclusion}
-\label{sec:conclusion}
+\section{Conclusions}
+\label{sec:conclusions}
 \todo[inline]{Extend/Rewrite}

### 2017-11-05 book_chapter/paper.tex
Small fixes to the text.

@@
 \usepackage{subcaption}
+\usepackage{rotating}
 
@@
 The expected utility $\text{EU}$ of a strategy can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
-Concretely, the expected utility of $\sigma$ given $\rho$ is $\text{EU}_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $\text{EU}_S(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
-The expected u %, \emph{i.e.}:
-% $$
-% \text{EU}(\sigma, \rho) = \sum_{t \in T} U(t, \rho(\sigma(t)))
-% $$
+Concretely, the expected utility of $\sigma$ given $\rho$ is $\text{EU}_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $\text{EU}_R(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
 As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, and the following utility matrix:
@@
 \end{tabular}
+%\captionof{table}{Example caption}
+%\label{tibble}
 \end{center}
@@
 
-The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
-However, we can already see how it attempts to address the study of meaning from a very different angle.
-There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific goals~\parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
-There is also no appeal to truth as a guide for our inquiry.
-One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
-We can merely focus on trying to see how agents can use signals to cope with the world and achieve their goals.
-On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
-In the examples discussed so far, agents are assumed to possess and exercise rationality.
-Unlike in the classical picture, it is an instrumental, rather than procedural, rationality, but one where agents are still omniscient and highly strategic.
+%The approach adumbrated so far is not, nor does it attempt to be, a full-blown theory of meaning like one would have in the classical picture.
+%However, we can already see how it attempts to address the study of meaning from a very different angle.
+%There is a focus, not on meaning as a kind of correspondence, but rather on the use of signals for specific goals~\parencite[but see][]{Hutteger:2007_Evol_Indicatives_Imperatives,Harms2010:Determining-tru,skyrms_signals_2010,Franke2013:An-adaptationis}.
+%There is also no appeal to truth as a guide for our inquiry.
+%One advantage of such a paradigm shift is that we are no longer tied to the need of explaining how language hooks on to the world; this question is no longer of primary relevance.
+%We can merely focus on trying to see how agents can use signals to cope with the world and achieve their goals.
+%On the way, we will better understand meaning, not by trying to say what meaning is, but by trying to better understand how communication works.
+%In the examples discussed so far, agents are assumed to possess and exercise rationality.
+%Unlike in the classical picture, it is an instrumental, rather than procedural, rationality, but one where agents are still omniscient and highly strategic.
 
@@
 
-The characterization of signaling problems in terms of evolutionary game theory allow us to explain why certain equilibria come to be and how.
+The characterization of signaling problems in terms of evolutionary game theory allows us to explain why certain equilibria come to be and how.
 A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
@@
 More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.
+In the following section, we give an overview of what vagueness looks like in a particular type of signaling games: sim-max games.
 
@@
 Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
-In terms of signaling games, we can take the state space to be constituted of values of the underlying dimension, and the message space as constituted by the terms in question.
+In terms of signaling games, we can model this by using a state space constituted of values of the underlying dimension, and a message space as constituted by the terms in question.
 Because of the difference in granularity, we will typically be interested in cases where the state space is much larger than the message space.
@@
 The simplest type of game to study in this scenario is one where the state space and the action space are the same. 
-We can imagine this as a game of guessing the state of affairs: the sender has knowledge of it, sends a message to the receiver, who in turn has to guess it; their payoff, as discussed above, is proportional to how close he got.
+We can imagine this as a game of guessing states of affairs: the sender has knowledge of a particular state, sends a message to the receiver, who in turn has to guess it; their payoff, as discussed above, is proportional to how close the guess got to the original state.
 These games, called similarity-maximization or sim-max games for short, were first introduced by Gerhard J{\"a}ger and Robert van~Rooij~\parencite{jager_language_2007,Jager2007} and further studied by~\textcite{jager_voronoi_2011}.
@@
 In general, at which point the sender switches the use of messages and which guesses of the receiver are optimal or rational critically hinges on the priors over states and the utility function.
-This will be important later.
+%This will be important later.
 Still, confirming Lipman's argument, there is no vagueness in any such optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, univocally associating one guess with each message.
@@
 In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space.
-The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where use of one message over the other by the sender is characterized by a smooth transition, and the receiver strategy assigns for each message a positive probability to more than one state.
+The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where the sender's choice of message to use is characterized by a smooth transition, and the receiver strategy assigns a positive probability to more than one state for each message.
 The particular shapes of these strategies are not important to characterize a vague signal use.
@@
 She studies sim-max games driven not by rational choice dynamics, but by generalized reinforcement learning (GRL), a variant of Herrnstein reinforcement learning (HRL)~\parencite{roth_learning_1995}.
-In HRL, agents learn to play a signaling game by strengthening particular choices (of messages for the sender, of responses for the receiver) proportionally to how successful that choice proves to be in an interaction.
+In HRL, agents learn to play a signaling game by strengthening particular choices (of messages for the sender, of responses for the receiver) proportionally to how successful those choices prove to be in an interaction.
 O'Connor's proposal is to model generalization as the propagation of reinforcement to nearby states, where ``nearby'' is defined in terms of distance in state space.
@@
 O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though less obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
-Considerations of optimality of strategies, for example by considering the evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
+Considerations of optimality of strategies, for example by analysing evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
 However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a shorter period of time.
@@
 Furthermore, it seems reasonable to assume that this state confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is that they will be mistaken for each other.
-Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of that dynamic that also induces vague signal use of the kind we expect here.
+Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of the dynamic that also induces vague signal use of the kind we expect here.
 The consequence is very similar to that of the GRL model discussed above, in that the way the behavior for a given state is updated takes into account the behavior of similar states, proportionally to their similarity.
@@
 A vague strategy is not claimed to have higher expected utility than a crisp one.
-However, the authors observe an effect similar to that pointed out by~\citeauthor{oconnor_evolution_2014}: signaling converges faster in scenarios where there is some degree of state confusability.
+However, the authors observe an effect similar to that pointed out by~\citeauthor{oconnor_evolution_2014}: signaling converges faster and more often in scenarios where there is some degree of state confusability.
 Furthermore, they observe one additional potentially beneficial property.
@@
 Thus, some amount of uncertainty seems to promote more homogeneous populations of signalers that are better at achieving cooperation within a group.
-In this picture, vagueness is the natural and unavoidable consequence of perceptual limitations, but may still contribute to the overall boundedly rational character of language if we take multiple contexts of learning into account.
+%In this picture, vagueness is the natural and unavoidable consequence of perceptual limitations, but may still contribute to the overall boundedly rational character of language if we take multiple contexts of learning into account.
 
 % \subsection{Aggregation from multiple senders}
-\textcite{lawry_vagueness_2017} explore the potential benefits of vagueness in variations of a sim-max game with multiple senders and one receiver.
+Finally, \textcite{lawry_vagueness_2017} explore the potential benefits of vagueness in variations of a sim-max game with multiple senders and one receiver.
 Although the models considered are not explicitly identified as sim-max games, they are equivalent in the relevant aspects.
@@
 The authors analyze a number of variations of this model.
-For a model with one sender, results are consistent with Lipman's~\parencite{lipman_why_2009} in that strategies without uncertainty lead to more accurate communication.
+For a model with one sender, results are consistent with Lipman's~\parencite*{lipman_why_2009} in that strategies without uncertainty lead to more accurate communication.
 For models with two messages and more than one sender, the expected error is a strictly decreasing function of the number of senders, and above a certain number of senders, strategies with uncertainty actually achieve lower expected error than crisp ones.
@@
 A combination of both would hypothetically lead to a better outcome in the scenario of \citeauthor{franke_vagueness_2017}.
-However, \citeauthor{lawry_vagueness_2017} give us something that holds even if the idealist is unwilling to acknowledge the constraints that pervade our finite existence.
-Whether the scenario of multiple senders is representative of the language games where vagueness comes about, and whether the strategies proposed by the authors would be the ones to evolve\footnote{The authors' analysis is a static one.
+\citeauthor{lawry_vagueness_2017} give us something that holds even if the idealist is unwilling to acknowledge the constraints that pervade our finite existence.
+Whether the strategies proposed by the authors would be the ones to evolve\footnote{The authors' analysis is a static one.
 Despite the advantage of the proposed vague strategies against strict ones in certain settings, one cannot know whether other strategies of higher accuracy would not be possible in those same settings without studying some dynamics in an evolutionary context.}, is something that is, however, open for debate.
@@
 % Another possible interpretation is that of a population of agents learning by imitation.
-Generalized reinforcement learning~\parencite{oconnor_evolution_2014}, and the work of~\textcite{franke_vagueness_2017}, both paint a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
+GRL~\parencite{oconnor_evolution_2014} and the work of~\textcite{franke_vagueness_2017} both assume a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
 %Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
@@
 But a crisp language would still have a higher expected utility than the evolved strategies.
-Agents in those models seem to be rational only to the extent of their capabilities.
-Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), this feels like a bittersweet conclusion because of the hypothetical possibility of an ideal strategy, seemingly barred from the agents in an artificial manner.
+Agents in those models seem to be only as rational as the modelers allow them to be.
+Despite the plausibility of the mechanisms proposed (limited memory, imprecise calculation of expected utilities, generalization, state confusability), the results of these models feel somewhat bittersweet because of the hypothetical possibility of an ideal strategy, seemingly barred from the agents in an artificial manner.
 Couldn't a more rational agent evolve and drive the system into crisp signal use?
@@
 Perhaps a deeper understanding of vagueness and the reasons for its pervasiveness in natural language are to be found only when we broaden the scope of the models employed.
-We call this the \emph{ecology} of vagueness.
+%We call this the \emph{ecology} of vagueness.
 All the models discussed so far explore evolutionary dynamics for one homogeneous population playing one game.
-Different types of agents and different game setups are considered, but each of these different possible scenarios has so far been tested separately.
+Different types of agents and different game setups are considered, but each of these different possible scenarios is always tested separately.
 We see at least two ways in which one could have a more ecological perspective.
@@
 These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[\S 23]{wittgenstein_philosophical_1953}.
-We could look for meaning at several levels in this pluralistic picture.
+We could look for rationality at several levels in this pluralistic picture.
 Firstly, as before, there is the actual behavior of a single agent in each actualized language game.
@@
 The argument goes that these side-effects, by temporarily enabling a higher expected utility, could allow a population using some generalization (or affected by some imprecision) to take over.
-However, despite the intuitiveness of the argument, it is based on comparing isolated runs of different dynamics.
+However, despite its intuitiveness the argument is based on comparing isolated runs of different dynamics.
 The models do not allow the hypothesis to be tested, because they do not accommodate different populations evolving together.
@@
 \label{sec:multi-population-model}
-We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations, with different imprecision parameters, evolving together.
+We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations with different imprecision values evolving together.
 In order to do this, first we consider that each population evolves according to the imprecise imitation dynamics.
@@
 Second, in order to see the impact of the success of different imprecision levels, we need to add another process of selection.
-The motivation is as follows: agents are born with varying perceptual abilities (modeled as populations with different imprecision values), they learn a signaling strategy by imitation of other agents (either within or across populations), depending on the success of the strategies they develop, they will be more or less likely to survive and reproduce, thus the proportion of each population changes depending on its fitness compared to other populations.
+The motivation is as follows: imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
+In such setup, we change not only each populations particular strategies, but also the proportion of each population depending on its fitness compared to other populations.
 
-Formally, let $A$ be the set of imprecision values considered.
+Formally, let's define $A$ to be the set of imprecision values considered.
 For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$, such that $\sum_{\alpha \in A} P(\alpha) = 1$.
@@
 These formulations are merely parameterized versions of the single-population model.
-They encapsulate calculations that we can then use to compute expected utilities and update probabilities for each type.
-The latter depend on the types of interaction that we imagine occurring between populations.
-In the following sections, we consider two different types.
+They encapsulate calculations that we can use to compute expected utilities and update probabilities for each type.
+The latter, however, depend on the types of interaction that we imagine occurring between populations.
+In the following sections, we consider two different possibilities.
 
@@
 Another important value to calculate has to do with the types which agents observe and imitate.
-In a model with tight interaction, we imagine that this occurring across populations.
+In a model with tight interaction, we imagine this occurring across populations.
 Therefore, we can define the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
@@
 $$
-Thus, these calculation also incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
-
-Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ are given by:
+Again, these calculation incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
+%
+Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ can be given by:
 $$
@@
 $$
-In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma^\alpha}$ and $\tilde{\rho^\alpha}$ (generated at each time step).
+In order to additionally account for the fact that strategies are not passed on to new generations, we mix the evolved strategies of a certain type $\sigma^\alpha$ and $\rho^\alpha$ with new random strategies $\tilde{\sigma}^\alpha$ and $\tilde{\rho}^\alpha$ (generated at each time step).
 The idea is that, at each time step, a certain percentage of each population will consist of ``newborn'' agents, \emph{i.e.}~agents that haven't yet had time to evolve their strategies.
@@
 $$
-\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma_{i+1}^{\alpha}}
+\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}_{i+1}^{\alpha}
 $$
 $$
-\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho_{i+1}^{\alpha}}
+\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}_{i+1}^{\alpha}
 $$
 
-We performed 25 simulation runs of this model for each of three population scenarios: only one population with $\alpha = 0$ (for reference), two populations with $\alpha \in \{0, 0.05\}$, and three populations with $\alpha \in \{0, 0.05\}$.
+We performed 25 simulation runs of this model for each of three population scenarios: only one population with $\alpha = 0$ (for reference), two populations with $\alpha \in \{0, 0.05\}$, and three populations with $\alpha \in \{0, 0.05, 0.1\}$.
 For each scenario, starting proportions $P(\alpha)$ were equal for each value of $\alpha$.
@@
 We also used a fixed uniform distribution for the priors and a message space with 2 messages.
-Regarding the duration of the simulations, due to the mixing in of random strategies at each step in order to simulate new individuals being introduced into the population (birth rate was fixed at $\gamma = 0.05$), the convergence criteria is not applicable.
+Regarding the duration of the simulations, due to the mixing in of random strategies at each step in order to simulate new individuals being introduced into the population (birth rate was fixed at $\gamma = 0.05$), the convergence criteria is no longer applicable.
 Therefore, all simulation runs were stopped after 200 iterations.
 
-The first thing to observe from the simulation results is that the population with no imprecision ($\alpha = 0$) always dominated the other populations in every run.
+The first thing to observe from the simulation results is that the population with no imprecision ($\alpha = 0$) dominated the other populations in every run.
+%\begin{sidewaysfigure}
 \begin{figure}
@@
 \end{figure}
+%\end{sidewaysfigure}
 In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for all trials.
-As the plot shows, for every trial, the population with no imprecision steadily increases its proportion against the others.
-In the three population scenario, the proportion of the population with $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
-
+As the plot shows, for every trial, the population with no imprecision steadily increased its proportion against the others.
+In the three population scenario, the proportion of type $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
 These observations go against the expectation of~\textcite{franke_vagueness_2017} that faster convergence to a convex strategy by populations with a certain level of imprecision could give them a temporary advantage to take over and eliminate other types.
@@
 What happens is that, because of the tight interaction between populations, the strategies of each type evolve in close tandem with each other.
-One of the consequences of this is that the population with no imprecision reaches convexity faster and develops more homogeneous strategies because of the interaction with the populations with imprecision.
-We can see the effect on convexity by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: population with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
+One of the consequences of this is that the population with no imprecision reaches convexity faster than it would on its own because of the interaction with the populations with imprecision.
+We can see this effect by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: population with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
 \begin{figure}
@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-percentage-all-strong.png}
-    \caption{Percentage of convex sender strategies.}
+    \caption{Percentage of convex sender strategies for $\alpha = 0$.}
     \label{fig:convexity-tight-interaction}
@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-sender-all-strong.png}
-    \caption{Mean entropy of sender strategies.}
+    \caption{Mean entropy of sender strategies for $\alpha = 0$.}
     \label{fig:entropy-sender-tight-interaction}
@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/entropy-receiver-all-strong.png}
-    \caption{Mean entropy of receiver strategies.}
+    \caption{Mean entropy of receiver strategies for $\alpha = 0$.}
     \label{fig:entropy-receiver-tight-interaction}
@@
 The flipside of this tight connection between populations is that strategies evolved by populations with no imprecision are also more vague (in the sense defined in Section~\ref{sec:sim-max-vagueness}).
-This is visible by looking at entropy values, namely sender strategy entropy, plotted in Figure~\ref{fig:entropy-sender-tight-interaction}, and receiver strategy entropy, plotted in Figure~\ref{fig:entropy-receiver-tight-interaction}.
+This is visible by looking at mean entropy values, namely sender strategy entropy, plotted in Figure~\ref{fig:entropy-sender-tight-interaction}, and receiver strategy entropy, plotted in Figure~\ref{fig:entropy-receiver-tight-interaction}.
 The values for the population with no imprecision are clearly higher in the scenarios where it evolves together with populations with imprecision than in the scenario where it evolves on its own.
@@
 
-Given the tight interaction between populations precluding the manifestation of temporary advantages of vagueness, we decided to test a variation of the model where populations interact more loosely.
-
 \subsection{Loose population interaction}
 \label{sec:loose-interaction-model}
-In order to model a different type of interaction between populations in our multi-population model, we need to make changes to some formulations.
-Any formula that is not redefined in this section should be assumed to be the same as used for the model in Section~\ref{sec:tight-interaction-model}.
-The dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of the same type.
+Given the tight interaction between populations precluding the manifestation of temporary advantages of vagueness, we decided to test a variation of the model where populations interact more loosely.
+In order to do this, we need to make changes to some calculations.
+However, any formula that is not redefined in this section should be assumed to remain the same.
+The dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of that same type.
 First, if agents learn only within their population, the imitation dynamic needs to consider only the expected utility against agents of that population.
@@
 $$
-The main difference is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type.
+The main difference with the previous model is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type ($\sigma^\alpha$ and $\rho^\alpha$).
+This also implies that population proportions do not play a role.
 
-We can redefine the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
+Regarding the imitation process, we can redefine the probability that a sender of type $\alpha$ observes a randomly sampled agent play message $m$ for observed state $t_o$ as:
 $$
@@
 $$
-Again, the main difference is that agents make observations within their own population, so to model a randomly sampled agent one needs only to take into account the same type.
-
-Update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ is then given by:
+Again, the main difference is that agents make observations within their own population, so to model a randomly sampled agent one needs only to take into account agents of the same type.
+Population proportions again do not play a role in these calculations.
+Based on these formulas, we can define the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ as:
 $$
@@
 $$
-and similarly for a receiver strategy of type $\alpha$ by:
+and similarly for a receiver strategy of type $\alpha$ as:
 $$
@@
 $$
+Note that, because imitation and learning occur only within populations, these formulations are essentially the same as for the single-population model of~\textcite{franke_vagueness_2017}, only parameterized by type $\alpha$.
 
-The selection process between different populations still follows the same motivation: agents of a certain type that evolve successful strategies (with respect to other types) will be more likely to survive and reproduce, benifitting the proportion of their population.
+The selection process between different populations still follows the same motivation as before: agents of a certain type that evolve successful strategies (with respect to other types) will be more likely to survive and reproduce, benifiting the proportion of their population.
 The formulation of the dynamic of the proportion of each type $P(\alpha)$ thus stays the same.
-In order to avoid confusion, we want to stress that this means that these calculations thus still rely on the calculation of expected utilities across populations (\emph{i.e.}~$\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})$) and not the newly introduced $\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\alpha})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\alpha})$.
+In order to avoid confusion, we want to stress that this means that these calculations rely on the definitions of expected utility across populations (\emph{i.e.}~$\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})$) and not the newly introduced $\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\alpha})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\alpha})$.
 
 We ran the same number of simulation trials under the same conditions as for the model with tight population interaction.
+%\begin{sidewaysfigure}
 \begin{figure}
@@
 \end{figure}
+%\end{sidewaysfigure}
 In Figure~\ref{fig:proportion-cases-loose-interaction} we plot the evolution of population proportions for all trials of the multi-population scenarios.
 The first thing to observe is that proportions evolve faster than in the model with tight interaction.
-Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in the one population fully dominating the others.
+Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in one population fully dominating the others.
 In the two-population scenario, some population reached 99\% proportion in 24 out of 25 trials.
 In the three-population scenario, this happened in 18 out of 25 trials.
-Given that strategies no longer develop in close tandem with each other, differences in expected utility can be more salient resulting in some population reaching a more clear advantage than the other.
+Given that strategies no longer develop in close tandem with each other, differences in expected utility can become more salient resulting in some type reaching a more clear advantage than the other.
 
@@
 In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
-In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then start losing ground to $\alpha = 0$.
+In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadly lose ground to $\alpha = 0$.
 For the three population scenario, $\alpha = 0.05$ fully reached 100\% proportion in only 1 trial and $\alpha = 0.1$ never reached more than 59\%.

### 2017-11-07 book_chapter/paper.tex
Added paragraph on relation with kin and multi-level selection theories.

@@
 We see at least two ways in which one could have a more ecological perspective.
-The first is to think about meaning and vagueness in a universe of multiple language games.
-We do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
-These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[\S 23]{wittgenstein_philosophical_1953}.
+The first is to think about meaning and vagueness with a more Wittgensteinian perspective.
+We can see each signaling game as embodying a particular language game.
+In Wittgenstein's picture of language, however, we do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
+These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[see][and in particular \S 23]{wittgenstein_philosophical_1953}.
 We could look for rationality at several levels in this pluralistic picture.
@@
 Second, in order to see the impact of the success of different imprecision levels, we need to add another process of selection.
-The motivation is as follows: imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
+A possible story to motivate these ideas goes as follows: imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
 In such setup, we change not only each populations particular strategies, but also the proportion of each population depending on its fitness compared to other populations.
 
-Formally, let's define $A$ to be the set of imprecision values considered.
+The motivations behind this model are closely related to the ideas behind the theories of kin selection~\parencite{hamilton_genetical_1964} and multi-level (or group) selection~\parencite{wilson_theory_1975}.
+These theories build upon the hypothesis that selection acts not only to directly favor genes that result in behaviors that benefit individuals, but also to indirectly favor genes that lead to behaviors that benefit either genotypically (kin) or socially (group) related individuals~\parencite[see][for more details]{okasha_evolution_2006}.
+%There are thus two aspects to be considered: for kin selection that would be the direct and indirect effects, and for multi-level selection the within-group and between-group components~\parencite[28--29]{birch_kin_2014}
+In our model there is also a similar structure: there is a process of selection of behavior within each population, levels of imprecision are selected across populations based on the strategies they give rise to.
+We believe there is an important distinction in that the two selection processes in our model act on different entities, the inner shaping signaling behavior, and the outer selecting levels of imprecision.
+In any case, we intend our model to be descriptive rather than causal or explanatory.
+That means that we do not want to commit to seeing populations either as kins or as social groups, and use them as merely descriptive abstractions that allow us to capture the hypothetical impact of indirect selection processes%
+\footnote{We make this note because of the heated debate between the two theories. See~\textcite{kohn_darwin_2008} and~\textcite{kramer_kin_2016} for more details on that.}%
+.
+
+In the following, we lay down the formal details of our model.
+Let's start by defining $A$ to be the set of imprecision values considered.
 For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$, such that $\sum_{\alpha \in A} P(\alpha) = 1$.

### 2017-11-07 book_chapter/simulation/analysis.R
Ran simulations for one scenario (weakest interaction, two populations with 0.0 and 0.05 impairment) with different initial proportions.

@@ -8,6 +8,7 @@ read.data <- function() {
         x$receiver.converged <- parse_logical(toupper(x$receiver.converged))
         x$file <- fname
         x$populations <- as.character(summarise(x, paste(unique(impairment), collapse='/')))
+        x$initial.proportions <- as.character(summarise(filter(x, iteration == 0), paste(proportion, collapse='/')))
         x$number.of.populations <- as.integer(summarise(x, length(unique(impairment))))
         measurement_data <<- rbind(measurement_data, x)
     }
@@ -64,7 +65,11 @@ plot.total.entropy <- function(interaction.type) {
     return(p)
 }
 
-plot.language <- function(simulation.id, impairment) {
+plot.language <- function(sim, impairment) {
+    fname <- unique(filter(measurement_data, sim.id == sim)$file)
+    fname.parts <- strsplit(fname, '-')
+    simulation.id <- paste(fname.parts[[1]][1:3], collapse='-')
+
     message.names <- c('m1','m2')
     sigma <- read_csv(paste0('strategies/', simulation.id, '-speaker-', impairment, '.csv'), col_names = message.names)
     nstates <- nrow(sigma)
@@ -108,3 +113,11 @@ plot.eu.mean <- function(interaction.type, population.scenario=NULL) {
     return(p)
 }
 
+do.stuff <- function() {
+    for (fname in unique(filter(measurement_data, interaction == 'weakest', populations == '0/0.05')$file)) {
+        parts <- strsplit(fname, '-')
+        sim.id <- paste(parts[[1]][1:3], collapse='-')
+        print(plot.language(sim.id, '0.0'))
+        print(plot.language(sim.id, '0.05'))
+    }
+}

### 2017-11-07 book_chapter/simulation/run-batch.sh

@@ -1,29 +1,14 @@
 #!/bin/bash
 
-for n in {1..25}
+for a in "0.0 0.1 0.05 0.9" "0.0 0.2 0.05 0.8" "0.0 0.3 0.05 0.7" "0.0 0.4 0.05 0.6" "0.0 0.5 0.05 0.5" "0.0 0.6 0.05 0.4" "0.0 0.7 0.05 0.3" "0.0 0.8 0.05 0.2" "0.0 0.9 0.05 0.1"
 do
-    echo -n "  $n "
-    python vagueness-games.py --batch 30 1 0.05 weakest 0.0
-    if [ $? -ne 0 ]; then
-        exit
-    fi
-done
-echo
-
-for a in "0.0 0.05" "0.0 0.05 0.1"
-do
-    echo $a
-    for b in "strong" "weak" "weakest"
+    for n in {1..25}
     do
-        echo " $b"
-        for n in {1..25}
-        do
-            echo -n "  $n "
-            python vagueness-games.py --batch 30 1 0.05 $b $a
-            if [ $? -ne 0 ]; then
-                exit
-            fi
-        done
-        echo
+        echo -n "  $n "
+        python vagueness-games.py --batch 30 1 0.05 weakest $a
+        if [ $? -ne 0 ]; then
+            exit
+        fi
     done
 done
+echo

### 2017-11-07 book_chapter/simulation/vagueness-games.py

@@ -193,13 +193,14 @@ NStates = 30
 PriorDistributionType = 'uniform'
 NMessages = 2
 Dynamics = 'replicator dynamics'
-Impairments = [0.0, 0.05, 0.1]
-NewPopulationRate = 0.0
+Impairments = [0.0, 0.05]
+PopulationProportions = makePDF(np.ones(len(Impairments)))
+NewPopulationRate = 0.05
 Tolerance = 0.1
 convThreshold = 0.001
 rounds = 200
 MetaDynamicsStep = 1
-SubPopulationInteraction = 'weak'
+SubPopulationInteraction = 'weakest'
 
 # # Batch mode
 
@@ -208,7 +209,7 @@ BatchMode = len(sys.argv) > 1
 if BatchMode:
     if len(sys.argv) < 5:
         print "Usage: python", sys.argv[
-            0], "--batch <number of states> <meta dynamics step> <birth rate> <sub-population interaction type> <impairments...>"
+            0], "--batch <number of states> <meta dynamics step> <birth rate> <sub-population interaction type> <<impairment proportion>...>"
         sys.exit(1)
     else:
         NStates = int(sys.argv[2])
@@ -216,13 +217,15 @@ if BatchMode:
         NewPopulationRate = float(sys.argv[4])
         SubPopulationInteraction = sys.argv[5]
         Impairments = []
-        for i in xrange(6, len(sys.argv)):
+        PopulationProportions = []
+        for i in xrange(6, len(sys.argv), 2):
             Impairments.append(float(sys.argv[i]))
+            PopulationProportions.append(float(sys.argv[i+1]))
 
 # # Initialization
 NPopulations = len(Impairments)
 
-PopulationProportions = makePDF(np.ones(NPopulations))
+PopulationProportions = makePDF(PopulationProportions)
 
 PerceptualSpace = np.linspace(0, 1, NStates, endpoint=True)
 if PriorDistributionType == 'uniform':

### 2017-11-10 book_chapter/simulation/analysis.R
Ran new round of simulations where random initializations (initial and newborns) are shared across populations.

@@ -1,3 +1,4 @@
+library(likert)
 library(tidyverse)
 
 read.data <- function() {
@@ -15,8 +16,10 @@ read.data <- function() {
     measurement_data$sim.id <<- as.integer(factor(measurement_data$file))
 }
 
-plot.proportion <- function(population.scenario, interaction.type, simulation.id=NULL) {
-    dt <- filter(measurement_data, populations == population.scenario, interaction == interaction.type)
+plot.proportion <- function(interaction.type, population.scenario, simulation.id=NULL) {
+    dt <- filter(measurement_data, 
+                 populations == population.scenario, 
+                 interaction == interaction.type)
     if (!is.null(simulation.id)) {
         dt <- filter(dt, sim.id == simulation.id)
     }
@@ -35,6 +38,16 @@ plot.convex.percentage <- function(interaction.type) {
     p <- ggplot(dt, aes(x=iteration)) +
         stat_summary(fun.y='mean', geom='line', aes(y=sender.convex, linetype=factor(populations))) +
         labs(x=expression(i), y='Convex percentage', linetype='Scenario') + theme_bw()
+    return(p)
+}
+
+plot.convex.cases <- function(interaction.type, population.scenario) {
+    dt <- filter(measurement_data, interaction == interaction.type, populations == population.scenario) %>% 
+        group_by(impairment, sim.id) %>% filter(sender.convex == 1) %>% 
+        summarise(convex.it = first(iteration))
+    p <- ggplot(dt) + geom_point(aes(x=convex.it, y=reverse.levels(factor(sim.id)), shape=factor(impairment))) +
+        scale_shape(solid = FALSE) +
+        labs(x=expression(i), y='Simulation', shape=expression(alpha))
     return(p)
 }
 
@@ -91,11 +104,25 @@ plot.language <- function(sim, impairment) {
         theme_bw() + theme(legend.position="none") + labs(y=NULL)
 }
 
+plot.eu <- function(interaction.type, population.scenario, sim.identificator) {
+    dt <- filter(measurement_data, 
+                 interaction == interaction.type, 
+                 populations == population.scenario,
+                 sim.id == sim.identificator)
+    dt2 <- filter(dt, sender.convex == 1) %>% group_by(impairment) %>% summarise(convex.it = first(iteration))
+    p <- ggplot(dt, aes(x=iteration)) +
+        geom_line(aes(y=sender.eu, linetype=factor(impairment)), color='red') +
[…]

### 2017-11-10 book_chapter/simulation/run-batch.sh

@@ -1,14 +1,29 @@
 #!/bin/bash
 
-for a in "0.0 0.1 0.05 0.9" "0.0 0.2 0.05 0.8" "0.0 0.3 0.05 0.7" "0.0 0.4 0.05 0.6" "0.0 0.5 0.05 0.5" "0.0 0.6 0.05 0.4" "0.0 0.7 0.05 0.3" "0.0 0.8 0.05 0.2" "0.0 0.9 0.05 0.1"
+for n in {1..25}
 do
-    for n in {1..25}
-    do
-        echo -n "  $n "
-        python vagueness-games.py --batch 30 1 0.05 weakest $a
-        if [ $? -ne 0 ]; then
-            exit
-        fi
-    done
+    echo -n "  $n "
+    python vagueness-games.py --batch 30 1 0.05 weakest 0.0 1
+    if [ $? -ne 0 ]; then
+        exit
+    fi
 done
 echo
+
+for a in "0.0 1 0.05 1" "0.0 1 0.05 1 0.1 1"
+do
+    echo $a
+    for b in "strong" "weak" "weakest"
+    do
+        echo " $b"
+        for n in {1..25}
+        do
+            echo -n "  $n "
+            python vagueness-games.py --batch 30 1 0.05 $b $a
+            if [ $? -ne 0 ]; then
+                exit
+            fi
+        done
+        echo
+    done
+done

### 2017-11-10 book_chapter/simulation/vagueness-games.py

@@ -246,8 +246,8 @@ ReceiverConfusions = [
     for x in Impairments
 ]
 
-Speakers = [random.dirichlet([1] * NMessages, NStates) for _ in xrange(NPopulations)]
-Hearers = [random.dirichlet([1] * NStates, NMessages) for _ in xrange(NPopulations)]
+Speakers = [random.dirichlet([1] * NMessages, NStates)] * NPopulations
+Hearers = [random.dirichlet([1] * NStates, NMessages)] * NPopulations
 
 ExpectedUtilitiesSenders = [ExpectedUtilitySpeaker(Speakers[j], Hearers, PopulationProportions, Utility)
                             for j in xrange(NPopulations)]
@@ -379,8 +379,8 @@ for i in xrange(1, rounds):
                                        *(ExpectedUtilitiesSenders[j] + ExpectedUtilitiesHearers[j]) \
                                        / (sum(ExpectedUtilitiesSenders) + sum(ExpectedUtilitiesHearers))
         PopulationProportions = makePDF(PopulationProportions)
-        NewSpeakers = [random.dirichlet([1] * NMessages, NStates) for _ in xrange(NPopulations)]
-        NewHearers = [random.dirichlet([1] * NStates, NMessages) for _ in xrange(NPopulations)]
+        NewSpeakers = [random.dirichlet([1] * NMessages, NStates)] * NPopulations
+        NewHearers = [random.dirichlet([1] * NStates, NMessages)] * NPopulations
         for j in xrange(NPopulations):
             Speakers[j] = makePDFPerRow((1 - NewPopulationRate) * Speakers[j] +
                                         NewPopulationRate * NewSpeakers[j])

### 2017-11-11 book_chapter/paper.tex
Wrote some more on the results of the loose interaction model.

@@
 We also used a fixed uniform distribution for the priors and a message space with 2 messages.
-Regarding the duration of the simulations, due to the mixing in of random strategies at each step in order to simulate new individuals being introduced into the population (birth rate was fixed at $\gamma = 0.05$), the convergence criteria is no longer applicable.
+Each type started with its own randomly generated strategy.
+Regarding the duration of the simulations, due to the mixing in of new individuals into the population (birth rate was fixed at $\gamma = 0.05$), the convergence criteria is no longer applicable because each strategy is randomly perturbed at each time step.
 Therefore, all simulation runs were stopped after 200 iterations.
@@
 In the three-population scenario, this happened in 18 out of 25 trials.
-Given that strategies no longer develop in close tandem with each other, differences in expected utility can become more salient resulting in some type reaching a more clear advantage than the other.
-
+%Given that strategies no longer develop in close tandem with each other, differences in expected utility can become more salient resulting in some type reaching a more clear advantage than the other.
 More interestingly for our investigation, some trials actually resulted in a population with some degree of imprecision dominating the others.
-For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials, a point from which it is technically impossible for the other population to recover.
+For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials (79, 86, 88, 89, 91, 93, 96, 98), a point from which it is technically impossible for the other population to recover.
 In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
 In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadly lose ground to $\alpha = 0$.
-For the three population scenario, $\alpha = 0.05$ fully reached 100\% proportion in only 1 trial and $\alpha = 0.1$ never reached more than 59\%.
-In 5 other trials we see a positive trend for $\alpha = 0.05$, but given the reversals that sometimes occur (\emph{e.g.}~trial 171), one cannot be sure whether full dominance would be achieved if the simulation was left to run for longer time.
+For the three population scenario, $\alpha = 0$ was reduced to 0\% proportion in 4 trials (152, 157, 159, and 175).
+In all of those cases, $\alpha = 0.05$ clearly has the upper hand over $\alpha = 0.1$, despite a temporary advantage of the latter in some trials.
+%In 5 other trials we see a positive trend for $\alpha = 0.05$, but given the reversals that sometimes occur (\emph{e.g.}~trial 171), one cannot be sure whether full dominance would be achieved if the simulation was left to run for longer time.
+
+Because of the loose interaction between populations, different types can now evolve separately.
+One consequence of this is that populations would a certain level of imprecision again reach convexity faster than those without imprecision.
+\begin{figure}
+  \centering
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-cases-0-005-weakest.png}
+    \caption{Two populations scenario.}
+    \label{fig:convex-cases-two-loose-interaction}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/convex-cases-0-005-01-weakest.png}
+    \caption{Three populations scenario.}
+    \label{fig:convex-cases-three-loose-interaction}
+  \end{subfigure}
+  \caption{First iteration with convex sender strategies for each simulation trial of the loose interaction model.}
+  \label{fig:convex-cases-loose-interaction}
+\end{figure}
+In Figure~\ref{fig:convex-cases-loose-interaction} we plot, for each trial, the first iteration when each type reached convexity.
+What we see is that, even though that is true on average, it is not always the case.
+This has certainly to do with the initial conditions of each trial, since the randomly generated strategies can simply by chance be more favourable to reaching convexity.
+More importantly, we also see that reaching convexity sooner is not a suficient condition for achieving population dominance .
+There were many trials where another type reached convexity much sooner than $\alpha = 0$ but was completely dominated nevertheless (\emph{e.g.}~76, 84, 92, and 99 in the two-population scenario; 156, 165, 168, and 172 in the three-population scenario).
+It also does not seem to be fully necessary, given a few examples where $\alpha = 0$ reached convexity early on and another type ended up dominating (89 in the two-population scenario and 152 in the three-population scenario).
+
+Another consequence of the two populations evolving separately is that they do not necessarily evolve towards the same equilibrium.
+In a sim-max game such as the one set up there are only two stable equilibria.
+These are two Voronoi systems of the kind shown in Figure~\ref{fig:example_strats}, one where the first message is used for the first half of the state space, and the other where the second message is used for this region.
+In the multi-population model with loose interaction, each population evolves independently towards one of these two equilibria.
+This is importantly conditioned by the initial strategy the process of selection initiates from.
+The populations are, however, not fully independent, since the process of selection of the level of precision relies on the expected utility of one population playing against itself and the others.
+And this is important because strategies in one equilibrium get the lowest payoff possible against strategies in the other equilibrium.
+Now, if two populations evolve towards different equilibria, whatever advantage one population has playing against itself could trigger a runaway effect by causing an increase in its proportion, which in turn will increase the population's relative expected utility, potentially increasing their proportion further in the next round, and so forth.
+In this case, one would expect that faster convergence towards convexity would be especially important for success.
+
+In the two-population scenario, of the 8 trials where $\alpha = 0$ was reduced to 0\%, all but two (88, 98) ended with the two populations each close to a different equilibria.
+In the two-population scenario, the population with $\alpha = 0$ evolved towards the same equilibrium as the other two in 1 of the trials where it was eliminated.
+In the other 3 trials, in 2 it evolved to a different equilibrium than the other two, and in 1 it evolved towards the same equilibrium as $\alpha = 0.05$ (but different to the one $\alpha = 0.1$ evolved towards).
+Again, we do not find a clear case for this being a decisive factor in the success of populations with impairment.
+That same goes for the impact of an initial advantage in expected utility creating the runaway effect we just mentioned.
+Despite none of these three factors (including reaching convexity sooner) sentencing the demise of $\alpha = 0$ with certainty, they do seem to conspire together to bring it about.
+In most trials where the type was eliminated, in both the two- and three-population scenarios, $\alpha = 0$ ended up evolving towards a different equilibrium than the other types, and either had an initial disadvantage in expected utility, or reached convexity much later.
+These factors can thus be seen as indicators at best, but the story of how a population with imprecision ends up dominating one without seems to be more complicated to tell.
+
+This is not surprising since we are facing a complex dynamical system with various interacting components (sender and receiver strategies and multiple populations).
+\begin{figure}
+  \centering
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/sender-eu-trial-88-0-005-weakest.png}
+    \caption{Sender expected utility.}
+    \label{fig:sender-eu-trial-88-loose-interaction}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[]{0.45\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/receiver-eu-trial-88-0-005-weakest.png}
+    \caption{Receiver expected utility.}
+    \label{fig:receiver-eu-trial-88-loose-interaction}
+  \end{subfigure}
+  \caption{Expected utilities for trial 88 of the loose interaction model. Vertical lines demarcate, for each type, the iteration where convexity was achieved.}
+  \label{fig:eu-trial-88-loose-interaction}
+\end{figure}
+Just as an example of this, in Figure~\ref{fig:eu-trial-88-loose-interaction} we plot the expected utilities of sender and receiver strategies for both populations of trial 88.
+The curves up to where $\alpha = 0.05$ achieved convexity seem to suggest that the sender strategy of that type was evolving in tandem with the receiver strategy of the other.
+This is possible in our model, although we expect it to be highly unlikely.
+The moment where the population reaches convexity marks the point where the sender strategy of $\alpha = 0.05$ aligns with its type's receiver strategy and this coincides with the moment where the population seems to gain real traction over the other type (see again the plot for trial 88 in Figure~\ref{fig:convex-cases-loose-interaction}).
 

### 2017-11-11 book_chapter/simulation/analysis.R

@@ -111,8 +111,8 @@ plot.eu <- function(interaction.type, population.scenario, sim.identificator) {
                  sim.id == sim.identificator)
     dt2 <- filter(dt, sender.convex == 1) %>% group_by(impairment) %>% summarise(convex.it = first(iteration))
     p <- ggplot(dt, aes(x=iteration)) +
-        geom_line(aes(y=sender.eu, linetype=factor(impairment)), color='red') +
-        geom_line(aes(y=receiver.eu, linetype=factor(impairment)), color='green') +
+        geom_line(aes(y=sender.eu, linetype=factor(impairment)), color = 'red') +
+        geom_line(aes(y=receiver.eu, linetype=factor(impairment)), color = 'green') +
         geom_vline(data = dt2, mapping = aes(xintercept = convex.it, linetype = factor(impairment))) +
         labs(x=expression(i), y=expression(EU), linetype=expression(alpha))
     return(p)
@@ -159,7 +159,7 @@ final.proportions <- function(interaction.type, population.scenario) {
 
 initial.eu <- function(interaction.type, population.scenario) {
     dt <- filter(measurement_data, interaction == interaction.type, populations == population.scenario) %>%
-        group_by(sim.id, impairment) %>% summarise(initial.eu = nth(sender.eu, 2)+nth(sender.eu, 2))
+        group_by(sim.id, impairment) %>% summarise(initial.eu = nth(sender.eu, 2) + nth(receiver.eu, 2))
     return(dt)
 }
 

### 2017-11-11 book_chapter/paper.tex
Re-wrote conclusions.

@@
 \label{sec:conclusions}
-\todo[inline]{Extend/Rewrite}
-We tried to show that, if we pragmatically acknowledge the complex and dynamic environment that surrounds language learning and use, we find many plausible sources of uncertainty that can lead to vague signal use.
-We supported this claim by giving an overview of recent research that tries to accommodate for vagueness in the context of sim-max games, and by discussing, in more general terms, how these considerations are relevant for reference games, and in the broader context of language as a multiplicity of language-games.
-Furthermore, we argued that if we broaden our notion of rationality beyond mere maximization of expected utility, we also find reasons to welcome the positive aspects of these processes and to see their acceptance as a perfectly rational choice.
-Irrational would be to ignore the finite nature of our existence and the constraints that condition language learning and use.
-Our challenge as researchers is of course not only to draw attention to the possibilities, but also to characterize which particular constraints give rise to vague signal use and how.
-
-The investigation doesn't end here, then.
-The research discussed here suggests that although vagueness can be beneficial, that is the case only in moderation: too much can lead to babbling equilibria where no information is transmitted.
-Thus, for agents capable of adjusting the level of vagueness in their strategies, questions remain regarding how this can be modeled and what factors influence the amount of vagueness that is optimal for a given situation.
-Other challenges include how to incorporate into our models the intuitions about the importance of speed of convergence, adaptability, and the promotion of group homogeneity.
-Finally, perhaps the biggest of them all is how to be able to focus on one language-game without losing sight of the ecosystem of games that need to be taken into account as well.
-Whatever solutions we find to these challenges, we believe that a notion of rationality cannot be strongly tied to mere utility maximization in a particular game, but has to consider an agents' need to optimize his behavior to navigate the plurality of uncertainties that besiege him.
-We need a clear notion of ecological rationality in order to see vagueness in the right light.
+Vagueness presents a challenge to both procedural and instrumental pictures of language and rationality alike.
+In the context of game-theoretical models of language, this takes the form of a question about evolution: how can vagueness persist if leads to demonstrably less efficient communication?
+Most existing proposals in the literature attempt to explain this by considering agents with some degree of bounded rationality.
+Despite the many ways in which our rationality is inevitably limited, the argument for the pervasiveness of vagueness in natural language would be much strengthened if there were also associated advantages.
+In this paper, we argued that finding those might require us to go beyond a local notion of rationality, as two of the proposals reviewed here~\parencite{oconnor_evolution_2014, franke_vagueness_2017} suggest.
+We advocate moving towards an ecological approach, studying vagueness in more heterogenenous ecosystems.
+Language is part of a very complex system~\parencite{beckner_language_2009} involving many components that interact in often unpredictable ways.
+An ecology of vagueness would involve considering models where different populations can evolve and interact with each other, or where different language games can be played between the individuals, or where the environment is uncertain and changing, all of the above, or anything else that more closely approximates the real context of language evolution.
+
+In light of this picture, we proposed two variants of a concrete multi-population signaling model to test the hypothesis~\parencite{franke_vagueness_2017} that certain features of imprecise imitation, like promoting faster convergence and regularity, could prove beneficial against full precision.
+Analysing these models did not provide us with a clear-cut answer, and revealed that the story is much more nuanced than initially expected.
+In a variant where populations with and without imprecision interact tightly, although precision always has the upper hand, vagueness seems to take a long time to be weeded out.
+When we let populations interact more loosely, we see a more complex pattern of outcomes.
+These include scenarios where imprecise imitation dominates over full precision, showing that strategies with vagueness can actually, under certain circumstances, be more successful.
+Bringing several populations together in a more complex ecosystem thus allowed us to not only spell out and test the original intuition, but also learn about new unforeseen effects.
+These models thus serve as an example of how moving to a more global perspective on rationality can allow us to achieve a more detailed awareness of the complex interactions that might be involved in sustaining vagueness in our language.
+They can be seen, we believe, as a first step towards an ecology of vagueness.
 

### 2017-11-11 book_chapter/paper.tex
Addressing Michael's comments.

@@
 For each value $\alpha \in A$, the proportion of its population is given by $P(\alpha)$, such that $\sum_{\alpha \in A} P(\alpha) = 1$.
+Each population has its own sender and receiver strategies, represented as $\sigma^\alpha$ and $\rho^\alpha$.
 The probability that a given agent with imprecision $\alpha$ (or of type $\alpha$) observes $t_o$ when the actual state is $t_a$ is given by $P_{o}^{\alpha}(t_{o}|t_{a})$.
@@
 These formulations are merely parameterized versions of the single-population model.
-They encapsulate calculations that we can use to compute expected utilities and update probabilities for each type.
+They encapsulate calculations that we can use to compute expected utilities and strategy update steps for each type.
 The latter, however, depend on the types of interaction that we imagine occurring between populations.
@@
 In a multi-population model with tight interaction between populations, each agent plays with, observes, and potentially imitates any other agent, regardless of their type.
-This has an impact on the expected utilities of sender and receiver strategies of each type, and on the update probabilities for those strategies.
+This has an impact on the expected utilities of sender and receiver strategies of each type, and on the update steps for those strategies.
 Let's start with the expected utilities.
@@
 %
-Finally, the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ can be given by:
-$$
-\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
+Finally, the update step for a sender strategy of type $\alpha$ at time instant $i+1$ is given by:
+$$
+\check{\sigma}_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\star})
 $$
@@
 $$
-\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
-$$
+\check{\rho}_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\star})
+$$
+We here use $\check{\sigma}$ and $\check{\rho}$ since there is still an additional adjustment to these values to be calculated before we get the final strategies $\sigma$ and $\rho$.
 
@@
 We define a parameter $\gamma$ that quantifies this percentage, or as we can also call it the birth rate, which we consider to be the same for every population.
-This mixing can be described, for a certain time step $i+1$, in the following formulas:
-$$
-\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\sigma_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}_{i+1}^{\alpha}
-$$
-$$
-\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\rho_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}_{i+1}^{\alpha}
+This mixing finally defines the strategies for time step $i+1$ and can be described in the following formulas:
+$$
+\sigma_{i+1}^{\alpha}(m|t)=(1 - \gamma)\check{\sigma}_{i+1}^{\alpha}(m|t) + \gamma \tilde{\sigma}_{i+1}^{\alpha}(m|t)
+$$
+$$
+\rho_{i+1}^{\alpha}(t|m)=(1 - \gamma)\check{\rho}_{i+1}^{\alpha}(t|m) + \gamma \tilde{\rho}_{i+1}^{\alpha}(t|m)
 $$
@@
   \centering
-  \begin{subfigure}[]{\textwidth}
-    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-strong.png}
+  \begin{subfigure}[]{.9\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-strong-square.png}
     \caption{Two populations scenario.}
@@
   \hfill
-  \begin{subfigure}[]{\textwidth}
-    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-strong.png}
+  \begin{subfigure}[]{.9\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-strong-square.png}
     \caption{Three populations scenario.}
@@
   \end{subfigure}
-  \caption{Evolution of population proportions through time for each simulation trial of the tight interaction model.}
+  \caption{Evolution of population proportions through time for each simulation trial of the tight interaction model. Numbers on top of each plot identify each trial.}
   \label{fig:proportion-cases-tight-interaction}
@@
 Population proportions again do not play a role in these calculations.
-Based on these formulas, we can define the update probabilities for a sender strategy of type $\alpha$ at time instant $i+1$ as:
-$$
-\sigma_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\alpha})
+Based on these formulas, we can define the update step for a sender strategy of type $\alpha$ at time instant $i+1$ as:
+$$
+\check{\sigma}_{i+1}^{\alpha}(m|t) \propto P_{o}^{\alpha}(m|t)\text{EU}_{\sigma_{i}}^{\alpha}(m,t,\rho_{i}^{\alpha})
 $$
@@
 $$
-\rho_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\alpha})
+\check{\rho}_{i+1}^{\alpha}(t|m) \propto P_{o}^{\alpha}(t|m)\text{EU}_{\rho_{i}}^{\alpha}(t,m,\sigma_{i}^{\alpha})
 $$
@@
   \centering
-  \begin{subfigure}[]{\textwidth}
-    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-weakest.png}
+  \begin{subfigure}[]{.9\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-weakest-square.png}
     \caption{Two populations scenario.}
@@
   \hfill
-  \begin{subfigure}[]{\textwidth}
-    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-weakest.png}
+  \begin{subfigure}[]{.9\textwidth}
+    \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-weakest-square.png}
     \caption{Three populations scenario.}
@@
   \end{subfigure}
-  \caption{Evolution of population proportions through time for each simulation trial of the loose interaction model.}
+  \caption{Evolution of population proportions through time for each simulation trial of the loose interaction model. Numbers on top of each plot identify each trial.}
   \label{fig:proportion-cases-loose-interaction}

### 2017-11-11 book_chapter/simulation/analysis.R

@@ -27,7 +27,7 @@ plot.proportion <- function(interaction.type, population.scenario, simulation.id
         geom_line(aes(linetype=factor(impairment))) +
         labs(x=expression(i), y=expression(P(alpha)), linetype=expression(alpha))
     if (is.null(simulation.id)) {
-        p <- p + facet_grid(. ~ sim.id) +
+        p <- p + facet_wrap(~ sim.id, nrow=5) +
             scale_x_continuous(breaks=c(100,200)) + theme(axis.text.x = element_text(angle=45))
     }
     return(p)

### 2017-11-11 book_chapter/paper.tex
Final changes.

@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-strong-square.png}
-    \caption{Two populations scenario.}
+    \caption{Two-population scenario.}
     \label{fig:proportion-cases-two-tight-interaction}
@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-strong-square.png}
-    \caption{Three populations scenario.}
+    \caption{Three-population scenario.}
     \label{fig:proportion-cases-three-tight-interaction}
@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-weakest-square.png}
-    \caption{Two populations scenario.}
+    \caption{Two-population scenario.}
     \label{fig:proportion-cases-two-loose-interaction}
@@
     \includegraphics[width=\textwidth]{simulation/results/round-3/plots/proportion-cases-0-005-01-weakest-square.png}
-    \caption{Three populations scenario.}
+    \caption{Three-population scenario.}
     \label{fig:proportion-cases-three-loose-interaction}
@@
 %Given that strategies no longer develop in close tandem with each other, differences in expected utility can become more salient resulting in some type reaching a more clear advantage than the other.
-More interestingly for our investigation, some trials actually resulted in a population with some degree of imprecision dominating the others.
+More interestingly for our investigation, some trials actually resulted in the population with $\alpha = 0$ being dominated.
 For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials (79, 86, 88, 89, 91, 93, 96, 98), a point from which it is technically impossible for the other population to recover.
@@
 Because of the loose interaction between populations, different types can now evolve separately.
-One consequence of this is that populations would a certain level of imprecision again reach convexity faster than those without imprecision.
+One consequence of this is that populations with a certain level of imprecision again reach convexity faster than those without imprecision.
 \begin{figure}
@@
 In Figure~\ref{fig:convex-cases-loose-interaction} we plot, for each trial, the first iteration when each type reached convexity.
-What we see is that, even though that is true on average, it is not always the case.
+What we see is that, even though $\alpha = 0$ usually reaches convexity later, this is not always the case.
 This has certainly to do with the initial conditions of each trial, since the randomly generated strategies can simply by chance be more favourable to reaching convexity.
 More importantly, we also see that reaching convexity sooner is not a suficient condition for achieving population dominance .
-There were many trials where another type reached convexity much sooner than $\alpha = 0$ but was completely dominated nevertheless (\emph{e.g.}~76, 84, 92, and 99 in the two-population scenario; 156, 165, 168, and 172 in the three-population scenario).
+There were many trials where another type reached convexity much sooner than $\alpha = 0$ but its population was completely dominated nevertheless (\emph{e.g.}~76, 84, 92, and 99 in the two-population scenario; 156, 165, 168, and 172 in the three-population scenario).
 It also does not seem to be fully necessary, given a few examples where $\alpha = 0$ reached convexity early on and another type ended up dominating (89 in the two-population scenario and 152 in the three-population scenario).
@@
 Another consequence of the two populations evolving separately is that they do not necessarily evolve towards the same equilibrium.
-In a sim-max game such as the one set up there are only two stable equilibria.
+In a sim-max game such as the one set up here, there are only two stable equilibria.
 These are two Voronoi systems of the kind shown in Figure~\ref{fig:example_strats}, one where the first message is used for the first half of the state space, and the other where the second message is used for this region.
@@
 Now, if two populations evolve towards different equilibria, whatever advantage one population has playing against itself could trigger a runaway effect by causing an increase in its proportion, which in turn will increase the population's relative expected utility, potentially increasing their proportion further in the next round, and so forth.
-In this case, one would expect that faster convergence towards convexity would be especially important for success.
+In this case, one would expect that faster convergence towards convexity would be especially important for a population's success.
 
 In the two-population scenario, of the 8 trials where $\alpha = 0$ was reduced to 0\%, all but two (88, 98) ended with the two populations each close to a different equilibria.
-In the two-population scenario, the population with $\alpha = 0$ evolved towards the same equilibrium as the other two in 1 of the trials where it was eliminated.
+In the three-population scenario, the population with $\alpha = 0$ evolved towards the same equilibrium as the other two in 1 of the trials where it was eliminated.
 In the other 3 trials, in 2 it evolved to a different equilibrium than the other two, and in 1 it evolved towards the same equilibrium as $\alpha = 0.05$ (but different to the one $\alpha = 0.1$ evolved towards).
-Again, we do not find a clear case for this being a decisive factor in the success of populations with impairment.
+Again, we do not find a clear case for this being a decisive factor in the success of populations with some imprecision.
 That same goes for the impact of an initial advantage in expected utility creating the runaway effect we just mentioned.
@@
 Just as an example of this, in Figure~\ref{fig:eu-trial-88-loose-interaction} we plot the expected utilities of sender and receiver strategies for both populations of trial 88.
-The curves up to where $\alpha = 0.05$ achieved convexity seem to suggest that the sender strategy of that type was evolving in tandem with the receiver strategy of the other.
-This is possible in our model, although we expect it to be highly unlikely.
+The curves up to where $\alpha = 0.05$ achieved convexity seem to suggest that the sender strategy of that type was evolving towards the same equilibrium as the receiver strategy of the other.
 The moment where the population reaches convexity marks the point where the sender strategy of $\alpha = 0.05$ aligns with its type's receiver strategy and this coincides with the moment where the population seems to gain real traction over the other type (see again the plot for trial 88 in Figure~\ref{fig:convex-cases-loose-interaction}).
+In this particular case, reaching convexity seems to have made an important difference.
 
@@
 \label{sec:conclusions}
-Vagueness presents a challenge to both procedural and instrumental pictures of language and rationality alike.
-In the context of game-theoretical models of language, this takes the form of a question about evolution: how can vagueness persist if leads to demonstrably less efficient communication?
+Vagueness presents a challenge to both procedural and instrumental pictures of rationality alike.
+In the context of game-theoretical models of language, this takes the form of a question about evolution: how can vagueness persist if it characterizes demonstrably less efficient communication?
 Most existing proposals in the literature attempt to explain this by considering agents with some degree of bounded rationality.
-Despite the many ways in which our rationality is inevitably limited, the argument for the pervasiveness of vagueness in natural language would be much strengthened if there were also associated advantages.
+Despite the many ways in which our rationality is inevitably limited, the argument for the pervasiveness of vagueness in natural language would be much stronger if one could also find associated advantages.
 In this paper, we argued that finding those might require us to go beyond a local notion of rationality, as two of the proposals reviewed here~\parencite{oconnor_evolution_2014, franke_vagueness_2017} suggest.
 We advocate moving towards an ecological approach, studying vagueness in more heterogenenous ecosystems.
-Language is part of a very complex system~\parencite{beckner_language_2009} involving many components that interact in often unpredictable ways.
-An ecology of vagueness would involve considering models where different populations can evolve and interact with each other, or where different language games can be played between the individuals, or where the environment is uncertain and changing, all of the above, or anything else that more closely approximates the real context of language evolution.
-
-In light of this picture, we proposed two variants of a concrete multi-population signaling model to test the hypothesis~\parencite{franke_vagueness_2017} that certain features of imprecise imitation, like promoting faster convergence and regularity, could prove beneficial against full precision.
+Language is part of a very complex system~\parencite{beckner_language_2009} that involves many components that interact in often unpredictable ways.
+An ecology of vagueness would involve studying models where different populations can evolve and interact with each other, where different language games can be played between the individuals, where the environment is uncertain and changing, or anything else that more closely approximates the real context of language evolution.
+
+In light of this picture, we proposed two variants of a concrete multi-population signaling model to test the hypothesis~\parencite{franke_vagueness_2017} that certain features of imprecise imitation, like promoting faster convergence and regularity, could prove beneficial in contrast with full precision.
 Analysing these models did not provide us with a clear-cut answer, and revealed that the story is much more nuanced than initially expected.
@@
 These include scenarios where imprecise imitation dominates over full precision, showing that strategies with vagueness can actually, under certain circumstances, be more successful.
-Bringing several populations together in a more complex ecosystem thus allowed us to not only spell out and test the original intuition, but also learn about new unforeseen effects.
-These models thus serve as an example of how moving to a more global perspective on rationality can allow us to achieve a more detailed awareness of the complex interactions that might be involved in sustaining vagueness in our language.
+Bringing several populations together in a more complex ecosystem thus allowed us to not only spell out and test the original intuition, but also learn about unforeseen effects.
+These models thus serve as an example of how moving to a more global perspective on rationality can allow us to achieve a more detailed awareness of the complex interactions that might be involved in sustaining vagueness in natural language.
 They can be seen, we believe, as a first step towards an ecology of vagueness.

### 2017-11-13 book_chapter/paper.tex
some typos, massages and some shortening in section 3

@@
 Despite the plausibility of the proposals, we argue that a deeper understanding of the benefits of vagueness needs a more ecological perspective, namely one that goes beyond local optimization of signaling strategies of a homogeneous population.
-As an example of one possible way to expand upon our current models, we propose two variants of a novel multi-population dynamic of imprecise imitation where, under certain conditions, populations with a certain level of imprecision dominate over precise ones.
+As an example of one possible way to expand upon our current models, we propose two variants of
+a novel multi-population dynamic of imprecise imitation where, under certain conditions,
+populations with vague language use dominate over populations with precise language use.
 \end{abstract}
@@
 Section~\ref{sec:ecology-of-vagueness} tries to generalize the considerations of these proposals to argue for a more ecological approach to vagueness.
+\mf{We should make clear what ``ecological'' means.}
 We propose and analyse the results of a novel multi-population model of imprecise imitation in Section~\ref{sec:multi-population-model}, and summarize our conclusions in Section~\ref{sec:conclusions}.
@@
 
-In order to work, the sorites paradox requires us to assume a relation between the vague terms and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
+The sorites paradox requires us to assume a relation between the vague terms and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
 Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
@@
 
-In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space.
+In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space which assigns all probability mass to a single state.
 The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where the sender's choice of message to use is characterized by a smooth transition, and the receiver strategy assigns a positive probability to more than one state for each message.
-The particular shapes of these strategies are not important to characterize a vague signal use.
-What is important is that the transition is smooth and monotonous for both players.
+The particular shapes of these strategies are less important for diagnosing vagueness.
+What is more important is a smooth and monotonous transition between areas where a message clearly applies and where it clearly does not.
 This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is as good as certain.
@@
 A third option is to see probabilities as capturing relative numbers in a population of agents.
-For example, if the sender strategy assigns a probability of $0.4$ of message $m$ being sent for state $t$, this would mean that $40\%$ of the population uses that message for that state, while $60\%$ uses the other.
+For example, if the sender strategy assigns a probability of $0.4$ to the event of message $m$ being sent for state $t$, this would mean that $40\%$ of the population uses that message for that state.
 This latter option leaves open the possible interpretation that each agent commands a crisp language and vagueness is only a population-level effect.
@@
 % \subsection{Bounded rationality}
-\textcite{franke_vagueness_2011} make two suggestions of how vagueness in a signaling game can be boundedly rational, \emph{i.e.}~how vagueness could arise as a consequence of cost-saving limitations in the cognitive capacities of instrumentally rational agents.
+\textcite{franke_vagueness_2011} make two suggestions as to how vagueness in a signaling game can be boundedly rational, \emph{i.e.}~how vagueness could arise as a consequence of cost-saving limitations in the cognitive capacities of instrumentally rational agents.
 The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
@@
 What they observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
+This is because each agent is exposed to a different history of previous play and so holds slightly different beliefs about language use.
 However, each agent still commands a crisp language, which is inadequate if the intention is to capture how vagueness presents itself in human languages.
@@
 This, in turn, may actually be boundedly rational since the calculation of expected utilities relies on assessing stochastic uncertainty, which in turn may be costly to calculate precisely. 
-Choice based on a few samples from a distribution can be optimal if taking more samples or other means of better approximating degrees of uncertainty is resource costly \parencite[\emph{e.g.}][]{VulGoodman2014:One-and-Done-Op,SanbornChater2016:Bayesian-Brains}.
+Choice based on a few samples from a distribution can be optimal if taking more samples or other means of better approximating probabilistic beliefs is resource costly \parencite[\emph{e.g.}][]{VulGoodman2014:One-and-Done-Op,SanbornChater2016:Bayesian-Brains}.
 The degree to which agents tremble in the calculation of expected utilities and therefore deviate from the instrumentally rational behavior can be characterized by a parameter.
@@
 This is done to a degree that is proportional to the similarity between $t$ and other states.
-The dynamics also gives rise to vague signaling of the kind we are looking for.
-
-But why should agents evolve to generalize?
-%First, it is not argued that vagueness in itself is beneficial, but that it is a by-product of a learning mechanism.
-O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages, though less obvious ones (see \textcite{oconnor_evolving_2015} for the detailed argument).
-Considerations of optimality of strategies, for example by analysing evolutionary stable states, are typically made in terms of hypothetical atemporal comparisons of expected utility.
-However, GRL has an interesting property in comparison with more strict learning mechanisms: it achieves higher payoffs in a shorter period of time.
-In a naturalized evolutionary setting, where we care about bounded rationality in a multi-context environment, learning speed is an advantage which should also be taken into account.
+The dynamics gives rise to vague signaling of the kind we are looking for.
+
+% \subsection{Noisy perception}
+Although there is a close relationship between reinforcement learning and population-level dynamics \citep{Borgers1997,Beggs2005}, O'Connor's GRL is, on the face of it, an account of learning between single players.
+Also, we need to ask for further justification to link generalization in reinforcement closely to the underlying payoff function of the sim-max game.
+Why should agents evolve to generalize in exactly the right way?
+O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages: it achieves higher payoffs in a shorter period of time.
+From a more encompassing, ecological point of view, learning speed can be an advantage.
 Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.
@@
 
-But would such a language be stable?
-In terms of evolutionary game theory, one could argue that a small enough group of mutants playing a precise signaling system could later invade a population of vague signalers.
-There is, however, an additional advantage to the generalized learning that is relevant here, namely the ability to adapt to a changing environment.
-Analysis of a game in terms of evolutionary stable states works well if we assume that the environment where the agents interact remains static.
-However, in a more realistic setting this is not the case.
-In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utility function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of its short-term gains.
-
-% \subsection{Noisy perception}
-\textcite{franke_vagueness_2017} study a variant of the replicator dynamics that is motivated by perceptual limitations.
+An alternative is offered by \textcite{franke_vagueness_2017} who study a variant of the replicator dynamics, \emph{i.e.}~a population-level dynamics, in which individual agents do not have the ability to generalize but simply make perceptual mistakes.
 Assuming that agents do not have perfect perception, there will always be a possibility that senders confuse states and receivers mix up responses.
@@
 %In this picture, vagueness is the natural and unavoidable consequence of perceptual limitations, but may still contribute to the overall boundedly rational character of language if we take multiple contexts of learning into account.
-
-% \subsection{Aggregation from multiple senders}
-Finally, \textcite{lawry_vagueness_2017} explore the potential benefits of vagueness in variations of a sim-max game with multiple senders and one receiver.
-Although the models considered are not explicitly identified as sim-max games, they are equivalent in the relevant aspects.
-They are signaling games where senders have private knowledge of a given state $x$, each sends a message to the receiver, which then selects an estimate $y$ of the original state.
-Although the authors do not define a utility function, their analysis evaluates the models in terms of the expected error between the estimate and the original state, defined as $(x-y)^2$.
-Better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse.
-The game could then be seen as a sim-max game with multiple senders and negative squared difference as utility function.
-
-Unlike the models discussed so far that introduce minimal constraints to a dynamic and study the strategies that emerge out of an evolutionary process, \citeauthor{lawry_vagueness_2017}' approach is mostly static and analytical.
-They consider several possible sender strategies, with and without uncertainty, and calculate if and when sender and receiver achieve better transmission accuracy, \emph{i.e.}~smaller expected error between original state and receiver estimate.
-Senders without uncertainty have a strategy that boils down to what is represented in Figure~\ref{fig:example_stratsA}: they split the state space according to a fixed threshold, switching from using one message to another abruptly.
-The model for the other type of sender is similar, but assumes that there is uncertainty regarding the value of the threshold, representing it as a uniformly distributed random variable.
-What results from this in terms of sender strategy is also a smooth transition between the use of two messages.%
-\footnote{Unlike what we see in Figure~\ref{fig:example_stratsB}, it is a linear transition.
+What is left to show, like for O'Connor's GRL approach, is whether the potential payoff advantages which were observed in simulations actually suffice to promote vague language use in an encompassing model of multi-level selection.
+The following section motivates an ecological approach to the evolution of vagueness in more detail, before Section~\ref{sec:multi-population-model} gives a concrete model.
+
+% In terms of evolutionary game theory, one could argue that a small enough group of mutants playing a precise signaling system could later invade a population of vague signalers.
+% There is, however, an additional advantage to the generalized learning that is relevant here, namely the ability to adapt to a changing environment.
+% Analysis of a game in terms of evolutionary stable states works well if we assume that the environment where the agents interact remains static.
+% However, in a more realistic setting this is not the case.
+% In a dynamic game with a variable environment, an ability to converge faster could give an evolutionary advantage over strict learning rules: as soon as the state space or the utility function change, GRL would quickly adapt and again potentially drive more strict learning mechanisms to extinction because of its short-term gains.
+
+% These results are interesting since they reveal scenarios where vague strategies have an advantage over strict ones.
+% The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the idealist\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
+% In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (better memory, or more reliable computation of expected utilities) would lead to more precision.
+% \citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{gilboa_social_1991}.
+% A combination of both would hypothetically lead to a better outcome in the scenario of \citeauthor{franke_vagueness_2017}.
+
+% % \subsection{Aggregation from multiple senders}
+% Finally, \textcite{lawry_vagueness_2017} explore the potential benefits of vagueness in variations of a sim-max game with multiple senders and one receiver.
+% Although the models considered are not explicitly identified as sim-max games, they are equivalent in the relevant aspects.
+% They are signaling games where senders have private knowledge of a given state $x$, each sends a message to the receiver, which then selects an estimate $y$ of the original state.
+% Although the authors do not define a utility function, their analysis evaluates the models in terms of the expected error between the estimate and the original state, defined as $(x-y)^2$.
+% Better models have lower expected error, thus one could say that the objective is to minimize this function, which would be equivalent to maximizing its inverse.
+% The game could then be seen as a sim-max game with multiple senders and negative squared difference as utility function.
+
+% Unlike the models discussed so far that introduce minimal constraints to a dynamic and study the strategies that emerge out of an evolutionary process, \citeauthor{lawry_vagueness_2017}' approach is mostly static and analytical.
+% They consider several possible sender strategies, with and without uncertainty, and calculate if and when sender and receiver achieve better transmission accuracy, \emph{i.e.}~smaller expected error between original state and receiver estimate.
+% Senders without uncertainty have a strategy that boils down to what is represented in Figure~\ref{fig:example_stratsA}: they split the state space according to a fixed threshold, switching from using one message to another abruptly.
+% The model for the other type of sender is similar, but assumes that there is uncertainty regarding the value of the threshold, representing it as a uniformly distributed random variable.
+% What results from this in terms of sender strategy is also a smooth transition between the use of two messages.%
+% \footnote{Unlike what we see in Figure~\ref{fig:example_stratsB}, it is a linear transition.
 This is certainly connected with the choice of modeling the uncertain threshold as uniformly distributed.
 We speculate that another choice of distribution, such as for example a normal distribution, could yield something similar to Figure~\ref{fig:example_stratsB}.}%
-The receiver strategy is not inherently vague for either type of sender, it merely aggregates the messages received to provide an estimate of the original state using a simple arithmetic mean.
-
-The authors analyze a number of variations of this model.
-For a model with one sender, results are consistent with Lipman's~\parencite*{lipman_why_2009} in that strategies without uncertainty lead to more accurate communication.
-For models with two messages and more than one sender, the expected error is a strictly decreasing function of the number of senders, and above a certain number of senders, strategies with uncertainty actually achieve lower expected error than crisp ones.
-This minimal number of senders from which vagueness surpasses sharpness is a decreasing function of the number of messages, \emph{i.e.}~whereas at least 8 senders are needed if the number of messages available is 2, this number lowers to 5 for 3 messages, 4 for and 5 messages, and 3 for more than 6 messages.
-If random errors are introduced between the choice of the sender and the message that reaches the receiver, increasing the number of senders can compensate for the noise, to a certain extent.
-Results are mixed when dropping the assumption of uniform priors for the state space.
-
-These results are interesting since they reveal scenarios where vague strategies have an advantage over strict ones, even in a setting of game-omniscient strategic agents.
-The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the idealist\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
-In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (better memory, or more reliable computation of expected utilities) would lead to more precision.
-\citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{gilboa_social_1991}.
-A combination of both would hypothetically lead to a better outcome in the scenario of \citeauthor{franke_vagueness_2017}.
-\citeauthor{lawry_vagueness_2017} give us something that holds even if the idealist is unwilling to acknowledge the constraints that pervade our finite existence.
-Whether the strategies proposed by the authors would be the ones to evolve\footnote{The authors' analysis is a static one.
-Despite the advantage of the proposed vague strategies against strict ones in certain settings, one cannot know whether other strategies of higher accuracy would not be possible in those same settings without studying some dynamics in an evolutionary context.}, is something that is, however, open for debate.
+% The receiver strategy is not inherently vague for either type of sender, it merely aggregates the messages received to provide an estimate of the original state using a simple arithmetic mean.
+
+% The authors analyze a number of variations of this model.
+% For a model with one sender, results are consistent with Lipman's~\parencite*{lipman_why_2009} in that strategies without uncertainty lead to more accurate communication.
+% For models with two messages and more than one sender, the expected error is a strictly decreasing function of the number of senders, and above a certain number of senders, strategies with uncertainty actually achieve lower expected error than crisp ones.
+% This minimal number of senders from which vagueness surpasses sharpness is a decreasing function of the number of messages, \emph{i.e.}~whereas at least 8 senders are needed if the number of messages available is 2, this number lowers to 5 for 3 messages, 4 for and 5 messages, and 3 for more than 6 messages.
+% If random errors are introduced between the choice of the sender and the message that reaches the receiver, increasing the number of senders can compensate for the noise, to a certain extent.
+% Results are mixed when dropping the assumption of uniform priors for the state space.
 
@@
 In sum, both at the level of behavior and at the level of beliefs about use or meaning, we should expect to find vagueness.
-Still, despite the natural vagueness, there does not seem to be anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent beliefs about language, use and meaning.
+Still, despite the natural vagueness, there does not seem to be anything fundamentally missing or conceptually incoherent in a naturalistic, rationality- or optimality-driven explanation of what each agent is doing or what each agent believes about language, use and meaning.
 
@@
 We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations with different imprecision values evolving together.
-In order to do this, first we consider that each population evolves according to the imprecise imitation dynamics.
-However, we want to model populations that are able to interact with each other.
-Therefore, the dynamic needs to be adapted in order to reflect the fact that agents can hypothetically imitate and play against agents with different imprecision levels.
-Second, in order to see the impact of the success of different imprecision levels, we need to add another process of selection.
-A possible story to motivate these ideas goes as follows: imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
-In such setup, we change not only each populations particular strategies, but also the proportion of each population depending on its fitness compared to other populations.
-
-The motivations behind this model are closely related to the ideas behind the theories of kin selection~\parencite{hamilton_genetical_1964} and multi-level (or group) selection~\parencite{wilson_theory_1975}.
+The resulting evolutionary dynamic has two layers: for one, signaling strategies change by imprecise imitation following the model of~\textcite{franke_vagueness_2017}; for another, there is higher-order evolutionary selection of the level of precision used by agents in their imitation of others' strategies.
+More concretely, imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
+In such setup, evolutionary processes will not only change the distribution of strategies within a population, where a population is identified by a shared level of imprecision in imitation.
+It will also promote those levels of imprecision which result in communicatively successful signal use.
+
+This is closely related to the ideas behind theories of kin selection~\parencite{hamilton_genetical_1964} and multi-level (or group) selection~\parencite{wilson_theory_1975}.
 These theories build upon the hypothesis that selection acts not only to directly favor genes that result in behaviors that benefit individuals, but also to indirectly favor genes that lead to behaviors that benefit either genotypically (kin) or socially (group) related individuals~\parencite[see][for more details]{okasha_evolution_2006}.
@@
 In any case, we intend our model to be descriptive rather than causal or explanatory.
-That means that we do not want to commit to seeing populations either as kins or as social groups, and use them as merely descriptive abstractions that allow us to capture the hypothetical impact of indirect selection processes%
+That means that we do not want to commit to seeing populations either as kins or as social groups, and use them as merely descriptive abstractions that allow us to capture the hypothetical impact of indirect selection processes.%
 \footnote{We make this note because of the heated debate between the two theories. See~\textcite{kohn_darwin_2008} and~\textcite{kramer_kin_2016} for more details on that.}%
-.
 
@@
 $$
-and, for a receiver of type $\alpha$, the expected utility of its strategy $\rho^\alpha$ against all other senderstrategies $\sigma^\star$, is given by:
+and, for a receiver of type $\alpha$, the expected utility of its strategy $\rho^\alpha$ against all other sender strategies $\sigma^\star$, is given by:
 $$
@@
 These formulations cover the evolution of the particular strategies of each type.
-We can think of this as the level of cultural evolution: agents are born with a certain level of imprecision and learn strategies based on the behavior of others.
+We can think of this as the level of cultural evolution: agents are born with a certain level of imprecision and adopt strategies based on the behavior of others.
 Alongside this process, we can imagine another level of selection, where agents die and new agents are born.
@@
 One of the consequences of this is that the population with no imprecision reaches convexity faster than it would on its own because of the interaction with the populations with imprecision.
-We can see this effect by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: population with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
+We can see this effect by looking into the percentage of trials with convex sender strategies at a given iteration, for each scenario, and comparing the three scenarios: populations with no imprecision evolving alone, two populations ($\alpha \in \{0, 0.05\}$), and three populations ($\alpha \in \{0, 0.05, 0.1\}$).
 \begin{figure}
@@
 \label{sec:loose-interaction-model}
-Given the tight interaction between populations precluding the manifestation of temporary advantages of vagueness, we decided to test a variation of the model where populations interact more loosely.
-In order to do this, we need to make changes to some calculations.
-However, any formula that is not redefined in this section should be assumed to remain the same.
-The dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of that same type.
+If we turn to a scenario where populations interact more loosely, we need to make changes to some calculations.
+Any formula that is not redefined in this section should be assumed to remain the same.
+The dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of that same precision type.
 First, if agents learn only within their population, the imitation dynamic needs to consider only the expected utility against agents of that population.
@@
 
-The selection process between different populations still follows the same motivation as before: agents of a certain type that evolve successful strategies (with respect to other types) will be more likely to survive and reproduce, benifiting the proportion of their population.
+The selection process between different populations still follows the same motivation as before: agents of a certain type that evolve successful strategies (with respect to other types) will be more likely to survive and reproduce, benefiting the proportion of their population.
 The formulation of the dynamic of the proportion of each type $P(\alpha)$ thus stays the same.
@@
 In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
-In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadly lose ground to $\alpha = 0$.
+In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadily lose ground to $\alpha = 0$.
 For the three population scenario, $\alpha = 0$ was reduced to 0\% proportion in 4 trials (152, 157, 159, and 175).
@@
 In the multi-population model with loose interaction, each population evolves independently towards one of these two equilibria.
-This is importantly conditioned by the initial strategy the process of selection initiates from.
+An important factor determining which strategy to converge to is the initial random population configuration.
 The populations are, however, not fully independent, since the process of selection of the level of precision relies on the expected utility of one population playing against itself and the others.
@@
 
-In the two-population scenario, of the 8 trials where $\alpha = 0$ was reduced to 0\%, all but two (88, 98) ended with the two populations each close to a different equilibria.
+In the two-population scenario, of the 8 trials where $\alpha = 0$ was reduced to 0\%, all but two (88, 98) ended with the two populations each close to different equilibria.
 In the three-population scenario, the population with $\alpha = 0$ evolved towards the same equilibrium as the other two in 1 of the trials where it was eliminated.

### 2017-11-13 book_chapter/paper.tex
Small changes to text. New round results with EU for imprecision selection calculated only within each population.

@@
 
-\section{Vagueness and the classical picture}
+\section{Vagueness and rationality}
 \label{sec:vagueness}
@@
 In Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a particular kind of signaling game.
-Section~\ref{sec:ecology-of-vagueness} tries to generalize the considerations of these proposals to argue for a more ecological approach to vagueness.
-\mf{We should make clear what ``ecological'' means.}
+Section~\ref{sec:ecology-of-vagueness} tries to generalize the considerations of these proposals to argue for an approach to vagueness anchored in a more global notion of rationality.
 We propose and analyse the results of a novel multi-population model of imprecise imitation in Section~\ref{sec:multi-population-model}, and summarize our conclusions in Section~\ref{sec:conclusions}.
@@
 The particular shapes of these strategies are less important for diagnosing vagueness.
-What is more important is a smooth and monotonous transition between areas where a message clearly applies and where it clearly does not.
+What is more important is a smooth and monotonous transition between areas where a message is clearly used and where it is clearly not.
 This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is as good as certain.
@@
 O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages: it achieves higher payoffs in a shorter period of time.
-From a more encompassing, ecological point of view, learning speed can be an advantage.
+From a more global point of view, learning speed can be an advantage.
 Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.

### 2017-11-13 book_chapter/simulation/run-batch.sh

@@ -1,19 +1,9 @@
 #!/bin/bash
 
-for n in {1..25}
-do
-    echo -n "  $n "
-    python vagueness-games.py --batch 30 1 0.05 weakest 0.0 1
-    if [ $? -ne 0 ]; then
-        exit
-    fi
-done
-echo
-
 for a in "0.0 1 0.05 1" "0.0 1 0.05 1 0.1 1"
 do
     echo $a
-    for b in "strong" "weak" "weakest"
+    for b in "weakest"
     do
         echo " $b"
         for n in {1..25}

### 2017-11-13 book_chapter/simulation/vagueness-games.py

@@ -246,12 +246,12 @@ ReceiverConfusions = [
     for x in Impairments
 ]
 
-Speakers = [random.dirichlet([1] * NMessages, NStates)] * NPopulations
-Hearers = [random.dirichlet([1] * NStates, NMessages)] * NPopulations
+Speakers = [random.dirichlet([1] * NMessages, NStates) for _ in xrange(NPopulations)]
+Hearers = [random.dirichlet([1] * NStates, NMessages) for _ in xrange(NPopulations)]
 
-ExpectedUtilitiesSenders = [ExpectedUtilitySpeaker(Speakers[j], Hearers, PopulationProportions, Utility)
+ExpectedUtilitiesSenders = [ExpectedUtility(Speakers[j], Hearers[j], Utility)
                             for j in xrange(NPopulations)]
-ExpectedUtilitiesHearers = [ExpectedUtilityHearer(Speakers, PopulationProportions, Hearers[j], Utility)
+ExpectedUtilitiesHearers = [ExpectedUtility(Speakers[j], Hearers[j], Utility)
                             for j in xrange(NPopulations)]
 
 SubPopulationMeasurements = []
@@ -266,8 +266,8 @@ for j in xrange(NPopulations):
         'sender.converged': False,
         'receiver.converged': False,
         'proportion': PopulationProportions[j],
-        'sender.eu': ExpectedUtilitySpeaker(Speakers[j], Hearers, PopulationProportions, Utility),
-        'receiver.eu': ExpectedUtilityHearer(Speakers, PopulationProportions, Hearers[j], Utility),
+        'sender.eu': ExpectedUtilitiesSenders[j],
+        'receiver.eu': ExpectedUtilitiesHearers[j],
         'sender.entropy': NormalizedEntropy(Speakers[j]),
         'receiver.entropy': NormalizedEntropy(Hearers[j]),
         'sender.convexity': Convexity(Speakers[j]),
@@ -365,8 +365,8 @@ for i in xrange(1, rounds):
                 % SubPopulationInteraction)
 
     for j in xrange(NPopulations):
-        ExpectedUtilitiesSenders[j] = ExpectedUtilitySpeaker(Speakers[j], Hearers, PopulationProportions, Utility)
-        ExpectedUtilitiesHearers[j] = ExpectedUtilityHearer(Speakers, PopulationProportions, Hearers[j], Utility)
+        ExpectedUtilitiesSenders[j] = ExpectedUtility(Speakers[j], Hearers[j], Utility)
[…]

### 2017-11-13 book_chapter/simulation/run-batch.sh
Separated new changes to be able to run them alongside the previous version of the model.

@@ -3,7 +3,7 @@
 for a in "0.0 1 0.05 1" "0.0 1 0.05 1 0.1 1"
 do
     echo $a
-    for b in "weakest"
+    for b in "detached"
     do
         echo " $b"
         for n in {1..25}

### 2017-11-13 book_chapter/simulation/vagueness-games.py

@@ -249,10 +249,16 @@ ReceiverConfusions = [
 Speakers = [random.dirichlet([1] * NMessages, NStates) for _ in xrange(NPopulations)]
 Hearers = [random.dirichlet([1] * NStates, NMessages) for _ in xrange(NPopulations)]
 
-ExpectedUtilitiesSenders = [ExpectedUtility(Speakers[j], Hearers[j], Utility)
-                            for j in xrange(NPopulations)]
-ExpectedUtilitiesHearers = [ExpectedUtility(Speakers[j], Hearers[j], Utility)
-                            for j in xrange(NPopulations)]
+if SubPopulationInteraction == 'detached':
+    ExpectedUtilitiesSenders = [ExpectedUtility(Speakers[j], Hearers[j], Utility)
+                                for j in xrange(NPopulations)]
+    ExpectedUtilitiesHearers = [ExpectedUtility(Speakers[j], Hearers[j], Utility)
+                                for j in xrange(NPopulations)]
+else:
+    ExpectedUtilitiesSenders = [ExpectedUtilitySpeaker(Speakers[j], Hearers, PopulationProportions, Utility)
+                                for j in xrange(NPopulations)]
+    ExpectedUtilitiesHearers = [ExpectedUtilityHearer(Speakers, PopulationProportions, Hearers[j], Utility)
+                                for j in xrange(NPopulations)]
 
 SubPopulationMeasurements = []
 for j in xrange(NPopulations):
@@ -340,7 +346,7 @@ for i in xrange(1, rounds):
                     for ti in xrange(NStates)]
                 for m in xrange(NMessages)])
             Hearers[j] = makePDFPerRow(PoReceiver * ExpUR)
-        elif SubPopulationInteraction == 'weakest':
+        elif SubPopulationInteraction == 'weakest' or SubPopulationInteraction == 'detached':
             PoSender = np.dot(PoT[j], PSigma[j])  # P_o(m|t_o)
             ExpUS = np.array([
                 [np.sum([PoT[j][to, ta] * np.sum(PRho[j][m, tr] * Utility[ta, tr]
@@ -361,12 +367,16 @@ for i in xrange(1, rounds):
             Hearers[j] = makePDFPerRow(PoReceiver * ExpUR)
         else:
             raise ValueError(
-                'Valid sub-population interaction types are \'strong\', \'weak\', and \'weakest\'. Unknown value \'%s\''
[…]

### 2017-11-14 book_chapter/paper.tex
Final changes before first submission.

@@
 \usepackage{rotating}
+\usepackage{graphicx}
 
@@
 \title{Towards an ecology of vagueness}
-\author{Jos\'e Pedro Correia \and Michael Franke}
+%\author{Jos\'e Pedro Correia \and Michael Franke}
 \date{}
@@
 
-Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and reflect on their implications for how one thinks of rationality in the context of language use.
-But before we start, we need to establish some vocabulary to better frame our discussion.
+%Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and suggest.
+%But before we start, we need to establish some vocabulary to better frame our discussion.
 Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
@@
 A sentence can be true or false only if it is meaningful, and it is meaningful if it is constructed according to correct rules.
-Rationality is intimately connected with the ability to follow good procedures, not only of sentence production, but ultimately of sentence combination and reasoning (think of what is required for making a logically valid deductive inference).
+Rationality is intimately connected with the ability to follow certain procedures, not only of sentence production, but ultimately of sentence combination and reasoning (think of what is required for making a logically valid deductive inference).
 This is what we could call a \emph{procedural} account of rationality, one which focuses more on the means rather than the ends.
@@
 When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between agents and their environment can come to bear on considerations of rationality.
-Depending on how accurate and complete an agent's knowledge of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, etc, the verdict over the rationality of certain choices can vary.
+The verdict over the rationality of certain choices can vary depending on how accurate and complete an agent's knowledge is of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, and so forth.
 An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
 Models working within the logical picture typically do not make a distinction between modeler and agent, thus lacking room to express any epistemic gaps.
-It follows that agents can only be assumed to be omniscient of the details involved.
+%It follows that agents can only be assumed to be omniscient of the details involved.
 By abstracting away from language users, these models cannot represent potential interactions between them, let alone repeated interaction and language change.
@@
 % Brian Skyrms~\parencite*[80--104]{skyrms_evolution_1996} identifies some problems with the story so far.
-Lewis' account of the stability of conventions rests on what could be considered strong demands for there to be a certain degree of required common knowledge between the players.
+Lewis' account of the stability of conventions rests on what could be considered strong demands. % for there to be a certain degree of required common knowledge between the players.
 Namely, there needs to be a state of affairs that indicates to everyone involved that a certain regularity will hold, as well as ``mutual ascription of some common inductive standards and background information, strategic rationality, mutual ascription of strategic rationality, and so on''~\parencite*[56--57]{lewis_convention_1969}.
-Agents are thus envisioned as omniscient of the game and highly strategic.
+Agents are thus envisioned as omniscient and highly strategic.
 These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
-The models were introduced in order to help explain how language could get off the ground as a conventional system without any sort of prior agreement.
+The models were introduced to help explain how language could have gotten off the ground as a conventional system without any sort of prior agreement.
 However, if we consider the origins of language from a historical perspective, it seems implausible that the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages possessed such advanced rationality.
@@
 Roughly, these are situations where the sender uses messages in a way that can be seen as partitioning the state space into convex regions and the receiver responds with the central element of those regions.
-For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280.%
-\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}
-Given two possible messages, the optimal sender strategy could be to use one message for all values from 40 up to 160, and the other for all values from 160 up to 280; the optimal receiver strategy could be to guess 100 if given the first message and 220 given the second message.
-These precise numerical values give mutually optimal behavior for a case where the prior probability is the same for each state and utility is a linear or quadratic function of the distance between the actual state and the receiver's guess.
-In general, at which point the sender switches the use of messages and which guesses of the receiver are optimal or rational critically hinges on the priors over states and the utility function.
+%For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280.%
+%\footnote{Giving some slack to the values of the shortest and tallest people ever as recorded by Guinness World Records, respectively at 54.6 and 272.}
+%Given two possible messages, the optimal sender strategy could be to use one message for all values from 40 up to 160, and the other for all values from 160 up to 280; the optimal receiver strategy could be to guess 100 if given the first message and 220 given the second message.
+%These precise numerical values give mutually optimal behavior for a case where the prior probability is the same for each state and utility is a linear or quadratic function of the distance between the actual state and the receiver's guess.
+%In general, at which point the sender switches the use of messages and which guesses of the receiver are optimal or rational critically hinges on the priors over states and the utility function.
 %This will be important later.
-Still, confirming Lipman's argument, there is no vagueness in any such optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, univocally associating one guess with each message.
+%Still, confirming Lipman's argument, there is no vagueness in any such optimal language: when given a height of 159 the sender will always use one message and given a height of 161 will always use the other; correspondingly, the receiver's response is also crisp, univocally associating one guess with each message.
 
@@
 \end{figure}
-
 In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space which assigns all probability mass to a single state.
+These strategies give mutually optimal behavior for a case where the prior probability is the same for each state and utility is a linear or quadratic function of the distance between the actual state and the receiver's guess.
+In general, at which point the sender switches the use of messages and which guesses of the receiver are optimal critically hinges on the priors over states and the utility function.
+%This will be important later.
+Still, confirming Lipman's argument, there is no vagueness in any such optimal language.
+
 The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where the sender's choice of message to use is characterized by a smooth transition, and the receiver strategy assigns a positive probability to more than one state for each message.
@@
 % These results are interesting since they reveal scenarios where vague strategies have an advantage over strict ones.
-% The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the idealist\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
-% In the models of \citeauthor{franke_vagueness_2011}, the idealist could argue that higher rationality (better memory, or more reliable computation of expected utilities) would lead to more precision.
+% The approaches discussed so far argue that vagueness is the inevitable byproduct of either cognitive limitations or a particular learning mechanism in the context of other learning mechanisms, which still leaves the skeptic\footnote{In an ordinary, non-philosophical, sense.} some room for speculation.
+% In the models of \citeauthor{franke_vagueness_2011}, the skeptic could argue that higher rationality (better memory, or more reliable computation of expected utilities) would lead to more precision.
 % \citeauthor{oconnor_evolving_2015}'s learning strategy could theoretically be dominated by a process with higher strategic rationality, like best response dynamics~\parencite{gilboa_social_1991}.
@@
 \item A corollary of the former is that there will always be heterogeneity in a population of language learners, at the very least in their prior experience, since each agent will have relied on a different set of observations.
-Furthermore, this information is not directly or fully accessible to others;
-\item Given that an agent is almost always learning and using language in a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice.
+Furthermore, this information is not directly or fully accessible to others.
+%\item Given that an agent is almost always learning and using language in a population of other agents, there are also various potential sources of linguistic input the agent is constantly integrating in his practice.
 \end{enumerate}
@@
 We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism thats leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
-\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).
+%\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).
 %Trying to answer the question of ``Why is language vague?'', is important to force us to recognize those aspects of language and the world that are easily forgotten when idealizing the subject matter.
@@
 These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[see][and in particular \S 23]{wittgenstein_philosophical_1953}.
-We could look for rationality at several levels in this pluralistic picture.
+
+We can look for rationality at several levels in this pluralistic picture.
 Firstly, as before, there is the actual behavior of a single agent in each actualized language game.
@@
 We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations with different imprecision values evolving together.
-The resulting evolutionary dynamic has two layers: for one, signaling strategies change by imprecise imitation following the model of~\textcite{franke_vagueness_2017}; for another, there is higher-order evolutionary selection of the level of precision used by agents in their imitation of others' strategies.
+The resulting evolutionary dynamic has two layers: first, signaling strategies change by imprecise imitation following the model of~\textcite{franke_vagueness_2017}; second, there is higher-order evolutionary selection of the level of precision used by agents in their imitation of others' strategies.
 More concretely, imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
@@
 
+\paragraph{Simulation results}
 We performed 25 simulation runs of this model for each of three population scenarios: only one population with $\alpha = 0$ (for reference), two populations with $\alpha \in \{0, 0.05\}$, and three populations with $\alpha \in \{0, 0.05, 0.1\}$.
@@
 \label{sec:loose-interaction-model}
-If we turn to a scenario where populations interact more loosely, we need to make changes to some calculations.
+We can also imagine a scenario where populations interact more loosely.
+Namely, the dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of the same type, but nevertheless use their signaling strategies with agents from all types.
+In order to capture this, we need to make changes to some calculations.
 Any formula that is not redefined in this section should be assumed to remain the same.
-The dynamic we want to model here is one where agents of a certain type $\alpha$ imitate and learn only from other agents of that same precision type.
 First, if agents learn only within their population, the imitation dynamic needs to consider only the expected utility against agents of that population.
@@
 The main difference with the previous model is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type ($\sigma^\alpha$ and $\rho^\alpha$).
-This also implies that population proportions do not play a role.
+This also implies that population proportions do not play a role in this level of selection.
 
@@
 The selection process between different populations still follows the same motivation as before: agents of a certain type that evolve successful strategies (with respect to other types) will be more likely to survive and reproduce, benefiting the proportion of their population.
-The formulation of the dynamic of the proportion of each type $P(\alpha)$ thus stays the same.
+The formulation of the dynamic for the proportion of each type $P(\alpha)$ thus stays the same.
 In order to avoid confusion, we want to stress that this means that these calculations rely on the definitions of expected utility across populations (\emph{i.e.}~$\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\star})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\star})$) and not the newly introduced $\text{EU}_{\sigma}^{\alpha}(m,t_{o},\rho^{\alpha})$ and $\text{EU}_{\rho}^{\alpha}(t_{i},m,\sigma^{\alpha})$.
 
+\paragraph{Simulation results}
 We ran the same number of simulation trials under the same conditions as for the model with tight population interaction.

### 2018-08-29 book_chapter/paper.tex
Fixing biblatex comma problem.
Fixed small language issues

@@
 
+\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
+
 \begin{document}
@@
 \title{Towards an ecology of vagueness}
-%\author{Jos\'e Pedro Correia \and Michael Franke}
+\author{Jos\'e Pedro Correia \and Michael Franke}
 \date{}
@@
 A vexing puzzle about vagueness, rationality and evolution runs, in crude abbreviation, as follows: vague language use is demonstrably suboptimal if the goal is efficient precise and cooperative information transmission; hence rational deliberation or evolutionary selection should, under this assumed goal, eradicate vagueness from language use.
-Since vagueness is persistent in all human languages, something has to give.
+Since vagueness is pervasive and entrenched in all human languages, something has to give.
 In this paper, we focus on this problem in the context of signaling games.
-We provide an overview of a number of proposed reasons why and mechanisms how vagueness may come into the picture in formal models of rational or evolutionary signaling.
+We provide an overview of a number of proposed reasons why and mechanisms how vagueness may come into the picture in formal models of rational and evolutionary signaling.
 Most of them argue that vague signal use is simply the best we can get, given certain factors.
-Despite the plausibility of the proposals, we argue that a deeper understanding of the benefits of vagueness needs a more ecological perspective, namely one that goes beyond local optimization of signaling strategies of a homogeneous population.
+Despite the plausibility of the proposals, we argue that a deeper understanding of the benefits of vagueness needs a more ecological perspective, namely one that goes beyond local optimization of signaling strategies in a homogeneous population.
 As an example of one possible way to expand upon our current models, we propose two variants of
@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
-The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, we can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, we would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
+The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, one can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, one would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
 Neither choice is, however, intuitively satisfying.
@@
 Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
-If we subscribe to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
+If one subscribes to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
 Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
-If we think of language as governed by logical rules, and of rationality as the ability to follow those rules, the sorites paradox seems \emph{prima facie} to show that vagueness in our language can undermine our rationality.
-
-But we can think of language in different terms.
+
+If one thinks of language as governed by logical rules, and of rationality as including the ability to follow those rules, the sorites paradox seems \emph{prima facie} to demonstrate that vagueness and rationality are incompatible.
+But one can think of language in different terms.
 A possibility is to think of signs as tools agents use to coordinate actions.
-An example of a way to formally study language along those lines is by using the framework of signaling games~\parencite{lewis_convention_1969}.
-In such a context, rationality becomes the ability to choose the use of signs that is optimal to achieve the desired coordination.
+An example of a way to formally study language along those lines is by using the game-theoretic framework of signaling games~\parencite{lewis_convention_1969}.
+Within such an approach, rationality can be seen as the ability to choose the use of signs that is optimal to achieve some form of coordination.
+The existence of vagueness in such models would not be at odds with rationality as long as vague languages turn out to be optimal for the purposes at hand.
 %So what about vagueness?
-However, in the context of these game-theoretic approaches to language, vagueness presents a new challenge.
-Barton Lipman~\parencite*{lipman_why_2009} gives a detailed account of the problem.
-Very succinctly, it can be put as follows.
+However, as Barton Lipman~\parencite*{lipman_why_2009} argues in detail, this is typically not the case.
+The problem can be put very succinctly as follows.
 In standard game-theoretic models of communication, vague signal use yields a lower expected utility than crisp use.
 Therefore, given that the dynamics (be it natural selection, cultural evolution, or rational choice) maximize utility, vagueness should be weeded out by these forces, giving rise to only precise languages.
-Thus it would be irrational to stick to vague language use when we could (theoretically) switch to a better system.
-But as we said before, vagueness is pervasive in our natural languages, and there is no reason to believe it is going away.
+Thus it would be irrational to stick to vague language use when one could (theoretically) switch to a better system.
+But vagueness is pervasive in natural language, and there is no reason to believe it is going away.
 The problem seems to be a theoretically serious one, but it does not obviously undermine our everyday linguistic practices.
@@
 
-%Following this idea, we here survey proposals for addressing vagueness in the context of signaling games and suggest.
-%But before we start, we need to establish some vocabulary to better frame our discussion.
+We here survey proposals for addressing vagueness in the context of signaling games, but before we start we want to establish some vocabulary to better frame the upcoming discussion.
 Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
-Discussions around it, although touching on different aspects of the concept, often fail to clearly demarcate them.
-We can start by considering what is deemed to be most important when characterizing the rationality of a given choice.
+Discussions around it usually touch on different aspects of the concept which we here want to try to clearly demarcate.
 The aforementioned logical picture of language and meaning is focused on dichotomies like true and false, meaningful and meaningless, correct and incorrect.
@@
 Rationality is intimately connected with the ability to follow certain procedures, not only of sentence production, but ultimately of sentence combination and reasoning (think of what is required for making a logically valid deductive inference).
-This is what we could call a \emph{procedural} account of rationality, one which focuses more on the means rather than the ends.
+This is an example of what we will call a \emph{procedural} account of rationality, one which focuses more on the means rather than the ends.
 One can also do the opposite and focus on the consequences instead.
@@
 When developing models where rationality is relevant, be it constructing a logical system or setting up a formal game, the assumed epistemic relation between agents and their environment can come to bear on considerations of rationality.
-The verdict over the rationality of certain choices can vary depending on how accurate and complete an agent's knowledge is of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, and so forth.
-An \emph{omniscient} agent is one that is in possession of the same information as the modeler, whereas an agent with less than that is said to only have \emph{limited} awareness of the relevant aspects of the model.
-Models working within the logical picture typically do not make a distinction between modeler and agent, thus lacking room to express any epistemic gaps.
+The verdict over how rational certain choices are can vary depending on how accurate and complete an agent's knowledge is of its environment, the goals to be achieved, the choices or rules available, the relation between those choices and the objectives, and so forth.
+In this respect, we will call an agent \emph{omniscient} if it is in possession of the same information as the modeler, whereas of an agent with less than that we will say that it only has \emph{limited} awareness of the relevant aspects of the model.
+Models working within the logical picture typically do not make a distinction between modeler and agent, and thus lack room to express these epistemic gaps.
 %It follows that agents can only be assumed to be omniscient of the details involved.
-By abstracting away from language users, these models cannot represent potential interactions between them, let alone repeated interaction and language change.
-In other types of models, however, a further aspect of rational belief formation needs to be considered, namely the ability of agents to make accurate predictions about how other agents behave.
+By abstracting away from language users, these models also typically do not represent potential interactions between them, let alone allow for repeated interaction and language change.
+In other types of models, however, a further aspect of rational choice needs to be considered, namely the ability of agents to make accurate predictions about how other agents behave.
 We can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
 %Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.\todo{too soon}
-Lack of perfect knowledge of the situation or lack of strategicness can be caused by many possible factors.
-These include, but are not limited to, limitations in handling information (receiving, storing, retrieving, transmitting) and limited computational resources to solve complex problems.
+Lack of perfect knowledge of the situation or lack of ability to choose strategically can be caused by many possible factors.
+These include, among other things, limitations in handling information (receiving, storing, retrieving, transmitting) and limited computational resources to solve complex problems.
 We can talk about \emph{bounded} rationality to characterize the choices of such agents.
@@
 %We draw attention to two additional aspects of rationality which will be helpful in the following.
-Where the above definition of instrumental rationality is usually understood as requiring a single choice to maximize the expected utility in a single concrete decision situation, we might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
+Where the above definition of instrumental rationality is usually understood as requiring a single choice to maximize the expected utility in a single concrete decision situation, one might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
 A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
@@
 % In order to extend his notion of convention to \emph{linguistic} exchanges
-In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where agents' choices involve sending and receiving signals or messages.
-Thus, we could think of two players with different roles.
+In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where agents' choices involve sending and receiving signals or messages\footnote{These terms will be used interchangeably.}.
+One can think of two players with different roles.
 The first player, the sender, has knowledge about which of a number of possible states of affairs obtains and, depending on this information, chooses a signal to send.
@@
 A preference relation exists between responses and states of affairs, and a payoff is attributed to each player based on the choices of both.
-Note that Lewis assumes that no player has any preference regarding the particular signal that is used, provided that it enables coordination.
-Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a probability measure $P$ such that $P(t)$ is the probability of frequency with which $t \in T$ occurs, a set of available signals or messages $M$, a set of responses or actions $A$, and a pair of utility functions $U_{S,R} : T \times A \rightarrow \mathbb{R}$, one for the sender and one for the receiver, each of which yields a payoff value for each possible pairing of state and action.
+Note that Lewis assumes that no player has any preference regarding the particular signal that is used for a given state, provided that it enables advantageous coordination with responses.
+Formally, in order to describe the setup all we need is to specify a set of possible states of affairs $T$, a probability measure $P$ such that $P(t)$ is the probability or frequency with which $t \in T$ occurs, a set of available signals or messages $M$, a set of responses or actions $A$, and a pair of utility functions $U_{S,R} : T \times A \rightarrow \mathbb{R}$, one for the sender and one for the receiver, each of which yields a payoff value for each possible pairing of state and action.
 These so-called signaling problems can be seen as particular cases of coordination problems if we consider the players' choices to be of contingency plans or strategies.
@@
 A receiver strategy analogously specifies a choice of action for each possible message.
-Thus, formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
-The expected utility $\text{EU}$ of a strategy can be calculated using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
+Formally, what the sender chooses is a function $\sigma : T \rightarrow M$ and the receiver a function $\rho : M \rightarrow A$.
+The expected utility $\text{EU}$ of a strategy can be calculated by using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
 Concretely, the expected utility of $\sigma$ given $\rho$ is $\text{EU}_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $\text{EU}_R(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
-As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, and the following utility matrix:
+As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, $P(t_1) = P(t_2) = 0.5$ and the following utility matrix:
 \begin{center}
@@
 \end{center}
-Possible sender and receiver strategies are, for example, $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
-These would have an expected utility of $1$ for both sender and receiver, since when $t_1$ obtains with probability $P(t_1)$ the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, and when $t_2$ obtains with probability $P(t_2) = 1-P(t_1)$ the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
+Consider sender strategy $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and receiver strategy $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
+These would have an expected utility of $1$ for both sender and receiver, since when $t_1$ obtains with probability $0.5$ the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, and when $t_2$ obtains with probability $0.5$ the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
 They also represent one of the two stable conventions in this game, the other being the pair of strategies $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
@@
 The models were introduced to help explain how language could have gotten off the ground as a conventional system without any sort of prior agreement.
-However, if we consider the origins of language from a historical perspective, it seems implausible that the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages possessed such advanced rationality.
+However, if one considers the circumstances of the origins of human language, it seems implausible that the agents that started making use of primordial signaling systems which (hypothetically) evolved into languages possessed such advanced rationality.
 Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
-A plausible account of the origin of language should first explain how signaling systems like those could get started, without requiring high standards of rationality of the agents involved.
-
-In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes to study signaling problems in evolutionary terms.
-Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, we can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
+A plausible account of the origin of language should first explain how signaling systems could get started, without requiring high standards of rationality of the agents involved.
+
+In order to address this problem, Brian Skyrms~\parencite*{skyrms_evolution_1996} proposes studying signaling problems in evolutionary terms.
+Rather than imagining, as Lewis does, rational agents making conscious decisions in possession of knowledge of the game and expectations of the behavior of other agents, one can imagine a simpler scenario inspired by biological evolution: there is a population of agents with biologically hardwired behaviors for engaging in interactions characteristic of a signaling problem; utility does not represent preference, but rather fitness for survival and reproduction; the make-up of the population evolves based on the relative fitness of the strategies represented in the population.
 Such a setup attempts to capture the main features of natural selection: in a diverse population, agents with more successful strategies thrive, while agents with less fit strategies die off.
@@
 The principles can be captured in formal models that abstract away from details of single interactions and behavior of individual agents, for example in the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
-The only thing relevant to the replicator dynamics are the relative proportions of strategies in a given population and the utility function.
+The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
 Using it, one can compute which strategies evolve under which conditions.
@@
 The characterization of signaling problems in terms of evolutionary game theory allows us to explain why certain equilibria come to be and how.
-A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under certain initial conditions and standard evolutionary pressures, and where it remains after small disturbances.
-With these tools, we can better understand why signaling systems are stable even without any strong assumptions of rationality.
-We can also map out which initial conditions drive the system towards which equilibria and which do not.
+A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under standard evolutionary pressures, and where it comes back to if slightly disturbed.
+With these tools, one can better understand why signaling systems are stable even without any strong assumptions of rationality.
+One can also map out which initial conditions drive the system towards which equilibria and which do not.
 In a simple case like the example discussed above, an evolutionary process of the kind described always drives the population into a state where one signaling system takes over completely.
@@
 Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few.
-More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \citet{FrankeWagner2014:Game-Theory-and}.
-In the following section, we give an overview of what vagueness looks like in a particular type of signaling games: sim-max games.
+More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \textcite{FrankeWagner2014:Game-Theory-and}.
+In the following section, we discuss how a particular type of signaling game has been used to address the problem of vagueness.
 
@@
 Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
-In terms of signaling games, we can model this by using a state space constituted of values of the underlying dimension, and a message space as constituted by the terms in question.
+In terms of signaling games, we can model this by using a state space constituted of values of the underlying dimension, and a message space constituted by the terms in question.
 Because of the difference in granularity, we will typically be interested in cases where the state space is much larger than the message space.
 We can model the structure of the state space by defining a distance or similarity function between every value, effectively making it a metric space.
-Another important ingredient of the paradox is the acknowledgment of a certain degree of tolerance with respect to whether a certain term applies or not.
+Another important ingredient of the sorites paradox is the acknowledgment of a certain degree of tolerance with respect to whether a certain term applies or not.
 This tolerance decreases with distance in state space: assuming a 180cm person is tall, one would easily tolerate the use of the term for a person measuring 179cm, less so for someone who is 170cm, and much less so for 160cm.
@@
 \end{figure}
-In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space which assigns all probability mass to a single state.
+
+In an abstract setup, using 50 states uniformly distributed over the unit interval and two possible messages, such an optimal language looks like what we see in Figure~\ref{fig:example_stratsA}: at a specific point, the probability that the sender uses one message decays sharply from 1 to 0, and increases sharply from 0 to 1 for the other message; the response of the receiver for each message is a degenerate distribution over the state space which assigns all probability mass to a single state.
 These strategies give mutually optimal behavior for a case where the prior probability is the same for each state and utility is a linear or quadratic function of the distance between the actual state and the receiver's guess.
@@
 %This will be important later.
-Still, confirming Lipman's argument, there is no vagueness in any such optimal language.
-
-The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where the sender's choice of message to use is characterized by a smooth transition, and the receiver strategy assigns a positive probability to more than one state for each message.
-The particular shapes of these strategies are less important for diagnosing vagueness.
-What is more important is a smooth and monotonous transition between areas where a message is clearly used and where it is clearly not.
-This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used, whereas for states in the extremes this is as good as certain.
+Still, confirming Lipman's argument, there is no vagueness in such optimal languages.
+
+The sender/receiver strategy pairs that we are looking for look more like Figure~\ref{fig:example_stratsB}, where the sender's probability of choosing a certain message gradually changes, and the receiver strategy assigns a positive probability to more than one state for each message.
+%The particular shapes of these strategies are less important for diagnosing vagueness.
+What characterizes vagueness in these models is thus a smooth and monotonous transition between parts of the state space where a message is clearly used and where it is clearly not.
+This means that, for some states in the middle of the state space, there is uncertainty as to which message will be used by the sender, whereas for states in the extremes this is as good as certain.
+For the receiver part, there is uncertainty as to which state will be picked for each message.
 
@@
 However, given the level of abstraction of the description so far, none of this is necessarily implied by the model.
-Our question is thus what additional modifications to sim-max games are sufficient for the optimal languages to be more like Figure~\ref{fig:example_stratsB}, rather than like Figure~\ref{fig:example_stratsA}.
+Our preferred stance is to see the model as descriptive, but we do not wish to focus on that debate here.
+The question we want to address is: what additional modifications to sim-max games are sufficient for optimal languages to be more like Figure~\ref{fig:example_stratsB}, rather than like Figure~\ref{fig:example_stratsA}?
 %Let's look at some proposals in the literature.
@@
 The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
-For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy.
-Each agent plays an instrumentally rational best response (given the utility function) to their estimate of the other's strategy.
+For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy, and plays an instrumentally rational best response to that strategy.
 In order to study the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
-What they observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the characteristics of a vague language as characterized before.
-This is because each agent is exposed to a different history of previous play and so holds slightly different beliefs about language use.
-However, each agent still commands a crisp language, which is inadequate if the intention is to capture how vagueness presents itself in human languages.
+What they observe is the emergence of vague signaling at the level of the population, \emph{i.e.}~population averages of individual strategies exhibit the hallmarks of a vague language.
+This is because each agent recalls a different history of previous play and so holds slightly different beliefs about language use.
+However, each agent still commands a crisp language, which is an inadequate feature of the model if the intention is to capture how vagueness presents itself in human languages.
 
@@
 % \subsection{Noisy perception}
-Although there is a close relationship between reinforcement learning and population-level dynamics \citep{Borgers1997,Beggs2005}, O'Connor's GRL is, on the face of it, an account of learning between single players.
+Although there is a close relationship between reinforcement learning and population-level dynamics \citep{Borgers1997,Beggs2005}, O'Connor's GRL is, on the face of it, an account of learning between individual players.
 Also, we need to ask for further justification to link generalization in reinforcement closely to the underlying payoff function of the sim-max game.
@@
 O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages: it achieves higher payoffs in a shorter period of time.
-From a more global point of view, learning speed can be an advantage.
-Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL to adapt to each other.
-Although the latter type of agent can hypothetically develop a precise and more efficient signaling system, agents using GRL would coordinate on vague signaling strategies with high (though not optimal) expected utility sooner than agents using HRL.
+From a global point of view, learning speed can be an advantage.
+Imagine an initial population of agents with random strategies, some using GRL and others using classical HRL.
+Although the latter type of agent can hypothetically develop a precise and more efficient signaling system, agents using GRL could coordinate on vague signaling strategies with high (though not optimal) expected utility sooner than agents using HRL.
 In such a scenario, they could drive the other agents to extinction before the latter had time to achieve coordination and reap the benefits of a more precise signaling system.
 
-An alternative is offered by \textcite{franke_vagueness_2017} who study a variant of the replicator dynamics, \emph{i.e.}~a population-level dynamics, in which individual agents do not have the ability to generalize but simply make perceptual mistakes.
-Assuming that agents do not have perfect perception, there will always be a possibility that senders confuse states and receivers mix up responses.
-Furthermore, it seems reasonable to assume that this state confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is that they will be mistaken for each other.
+A similar finding is made by \textcite{franke_vagueness_2017} when studying a variant of the replicator dynamics in which individual agents do not have the ability to generalize but simply make perceptual mistakes.
+In a scenario where agents learn by imitation, if one assumes that they do not have perfect perception, there will always be a possibility that senders confuse states and thus learn associations with messages that are different than the ones observed, and receivers similarly mix up responses.
+Furthermore, it seems reasonable to assume that this confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is that they will be mistaken for each other.
 Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of the dynamic that also induces vague signal use of the kind we expect here.
@@
 Given the known relation between reinforcement learning and the replicator dynamics~\parencite{Beggs2005}, it is actually quite plausible that the two are tightly related (although this would need to be formally demonstrated).
-The account is, furthermore, interpretable at a lower level of rationality, given that the replicator equation is suitable to represent biological processes of natural selection.
+The account is, furthermore, interpretable at a lower level of rationality.
+%, given that the replicator equation is suitable to represent biological processes of natural selection.
 
@@
 Furthermore, they observe one additional potentially beneficial property.
-By running several rounds of simulation for each parameter set, they can measure for each group of results how close resulting strategies are to each other, and how they would fare playing against one another.
+Running several rounds of simulation for each parameter set, they measure for each group of results how close resulting strategies are to each other, and how they would fare playing against one another.
 The results show that the within group distance between strategies becomes smaller with growing confusability, and the within group expected utility is actually higher for strategies evolving under a certain degree of state confusion.
@@
 What is left to show, like for O'Connor's GRL approach, is whether the potential payoff advantages which were observed in simulations actually suffice to promote vague language use in an encompassing model of multi-level selection.
-The following section motivates an ecological approach to the evolution of vagueness in more detail, before Section~\ref{sec:multi-population-model} gives a concrete model.
+The following section motivates the need for taking a more ecological approach to the evolution of vagueness, before Section~\ref{sec:multi-population-model} gives a concrete model.
 
@@
 \end{enumerate}
-All of these observations support the weakening of the modeling assumptions.
+All of these observations support the weakening of modeling assumptions.
 % If we take these points seriously, we have to concede that it is almost inevitable that uncertainty gets introduced at almost all levels of language learning and language use.
@@
 \citeauthor{oconnor_evolving_2015} also showed how generalization, an invaluable feature of any procedure for learning from a limited number of observations (point 2), leads to vagueness.
-We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism thats leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
+We also learned from \citeauthor{franke_vagueness_2017} that state confusability, a mechanism that leads to vague signal use, can have a homogenizing effect on vocabularies, potentially compensating for the heterogeneity of agents' experiences (point 3).
 %\citeauthor{lawry_vagueness_2017} demonstrated the benefits of strategies that incorporate uncertainty, leading to vagueness, in language games where agents need to aggregate information from various sources (point 4).
@@
 All the models discussed so far explore evolutionary dynamics for one homogeneous population playing one game.
-Different types of agents and different game setups are considered, but each of these different possible scenarios is always tested separately.
-We see at least two ways in which one could have a more ecological perspective.
-The first is to think about meaning and vagueness with a more Wittgensteinian perspective.
-We can see each signaling game as embodying a particular language game.
+Different types of agents and different game setups are considered, but each of these different possible scenarios is tested separately.
+We see at least two ways in which one could embrace a more ecological perspective.
+The first is to think about meaning and vagueness from a more Wittgensteinian perspective.
+One can see each signaling game as embodying a particular language game.
 In Wittgenstein's picture of language, however, we do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
@@
 The argument goes that these side-effects, by temporarily enabling a higher expected utility, could allow a population using some generalization (or affected by some imprecision) to take over.
-However, despite its intuitiveness the argument is based on comparing isolated runs of different dynamics.
+However, despite its intuitiveness, the argument is based on comparing isolated runs of different dynamics.
 The models do not allow the hypothesis to be tested, because they do not accommodate different populations evolving together.
@@
 %There are many ways in which this can be done.
-In the remainder of the paper we propose a way to do this for the model of~\textcite{franke_vagueness_2017}.
-We introduce two variations of a multi-population model of the imprecise imitation dynamics, where populations characterized by different imprecision values interact and evolve together.
+In the remainder of the paper we propose a way to do this based on the model of~\textcite{franke_vagueness_2017}.
+We introduce two variants of a multi-population model of the imprecise imitation dynamics, where populations characterized by different imprecision values interact and evolve together.
 Using this model, we can better see under which conditions the hypothesized advantages of some imprecision can lead to the evolution of vagueness.
@@
 We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations with different imprecision values evolving together.
-The resulting evolutionary dynamic has two layers: first, signaling strategies change by imprecise imitation following the model of~\textcite{franke_vagueness_2017}; second, there is higher-order evolutionary selection of the level of precision used by agents in their imitation of others' strategies.
+The resulting evolutionary dynamic has two layers: first, signaling strategies change by imprecise imitation following the model of~\textcite{franke_vagueness_2017}; second, there is evolutionary selection at the level of the imprecision values themselves.
 More concretely, imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
-In such setup, evolutionary processes will not only change the distribution of strategies within a population, where a population is identified by a shared level of imprecision in imitation.
-It will also promote those levels of imprecision which result in communicatively successful signal use.
+In such a setup, evolutionary processes will not only change the distribution of strategies within a population, where a population is identified by a shared level of imprecision in imitation, they will also promote those levels of imprecision which result in the development of more successful signal use.
 
@@
 %There are thus two aspects to be considered: for kin selection that would be the direct and indirect effects, and for multi-level selection the within-group and between-group components~\parencite[28--29]{birch_kin_2014}
-In our model there is also a similar structure: there is a process of selection of behavior within each population, levels of imprecision are selected across populations based on the strategies they give rise to.
-We believe there is an important distinction in that the two selection processes in our model act on different entities, the inner shaping signaling behavior, and the outer selecting levels of imprecision.
+In our model there is also a similar structure: there is a process of selection of behavior within each population, and levels of imprecision are selected across populations based on the strategies they give rise to.
+We believe there is an important difference with group and kin selection models in that the two selection processes in our model act on different entities, the inner shaping signaling behavior, and the outer selecting levels of imprecision.
 In any case, we intend our model to be descriptive rather than causal or explanatory.
@@
 These formulations are merely parameterized versions of the single-population model.
-They encapsulate calculations that we can use to compute expected utilities and strategy update steps for each type.
+They encapsulate calculations that one can use to compute expected utilities and strategy update steps for each type.
 The latter, however, depend on the types of interaction that we imagine occurring between populations.
@@
 In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for all trials.
-As the plot shows, for every trial, the population with no imprecision steadily increased its proportion against the others.
+As the plot shows, the population with no imprecision steadily increased its proportion against the others in every trial.
 In the three population scenario, the proportion of type $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
@@
 This is plotted in Figure~\ref{fig:convexity-tight-interaction}.
-What we see is more trials reaching convexity earlier for the multi-population scenarios when compared with the single-poulation scenario.
-This effect precludes the hypothesized temporary advantage of imprecision to even manifest itself, but it can be seen as a positive influence on the population with no imprecision.
-
-The flipside of this tight connection between populations is that strategies evolved by populations with no imprecision are also more vague (in the sense defined in Section~\ref{sec:sim-max-vagueness}).
+What we see is more trials reaching convexity earlier for the multi-population scenarios when compared with the single-population scenario.
+This effect precludes the hypothesized temporary advantage of imprecision to even manifest itself, but it can be seen as a positive influence of the population(s) with imprecision on the population with no imprecision.
+
+The flip side of this tight connection between populations is that strategies evolved by populations with no imprecision are also more vague (in the sense defined in Section~\ref{sec:sim-max-vagueness}).
 This is visible by looking at mean entropy values, namely sender strategy entropy, plotted in Figure~\ref{fig:entropy-sender-tight-interaction}, and receiver strategy entropy, plotted in Figure~\ref{fig:entropy-receiver-tight-interaction}.
 The values for the population with no imprecision are clearly higher in the scenarios where it evolves together with populations with imprecision than in the scenario where it evolves on its own.
-Given the trends in population proportions, one expects this to eventually be eliminated when the population with no imprecision finally takes over the others, but it is interesting to observe that while populations with vague strategies persist, the population with no imprecision takes much longer to evolve a crisp strategy.
+Given the trends in population proportions, one expects this to eventually be eliminated when the population with no imprecision finally takes over the others, but it is interesting to observe that while populations with vague strategies persist, the population with no imprecision takes much longer to evolve a fully crisp strategy.
 
@@
 The main difference with the previous model is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type ($\sigma^\alpha$ and $\rho^\alpha$).
-This also implies that population proportions do not play a role in this level of selection.
+This also implies that population proportions do not play a role at this level of selection.
 
@@
 Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in one population fully dominating the others.
-In the two-population scenario, some population reached 99\% proportion in 24 out of 25 trials.
+In the two-population scenario, some population reached at least 99\% proportion in 24 out of 25 trials.
 In the three-population scenario, this happened in 18 out of 25 trials.
@@
 In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
-In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadily lose ground to $\alpha = 0$.
+In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadily start losing ground to $\alpha = 0$.
 For the three population scenario, $\alpha = 0$ was reduced to 0\% proportion in 4 trials (152, 157, 159, and 175).
@@
 This has certainly to do with the initial conditions of each trial, since the randomly generated strategies can simply by chance be more favourable to reaching convexity.
-More importantly, we also see that reaching convexity sooner is not a suficient condition for achieving population dominance .
+More importantly, we also see that reaching convexity sooner is not a sufficient condition for achieving population dominance .
 There were many trials where another type reached convexity much sooner than $\alpha = 0$ but its population was completely dominated nevertheless (\emph{e.g.}~76, 84, 92, and 99 in the two-population scenario; 156, 165, 168, and 172 in the three-population scenario).
@@
 
-Another consequence of the two populations evolving separately is that they do not necessarily evolve towards the same equilibrium.
+Another consequence of the populations evolving separately in this model is that they do not necessarily evolve towards the same equilibrium.
 In a sim-max game such as the one set up here, there are only two stable equilibria.
-These are two Voronoi systems of the kind shown in Figure~\ref{fig:example_strats}, one where the first message is used for the first half of the state space, and the other where the second message is used for this region.
+These are the two Voronoi languages of the kind shown in Figure~\ref{fig:example_strats}, one where the first message is used for the first half of the state space, and the other where the second message is used for this region.
 In the multi-population model with loose interaction, each population evolves independently towards one of these two equilibria.
-An important factor determining which strategy to converge to is the initial random population configuration.
+An important factor determining which language strategies converge to is the random initial population configuration.
 The populations are, however, not fully independent, since the process of selection of the level of precision relies on the expected utility of one population playing against itself and the others.
@@
 Again, we do not find a clear case for this being a decisive factor in the success of populations with some imprecision.
-That same goes for the impact of an initial advantage in expected utility creating the runaway effect we just mentioned.
+The same goes for the impact of an initial advantage in expected utility creating the runaway effect we just mentioned.
 Despite none of these three factors (including reaching convexity sooner) sentencing the demise of $\alpha = 0$ with certainty, they do seem to conspire together to bring it about.
@@
 In this paper, we argued that finding those might require us to go beyond a local notion of rationality, as two of the proposals reviewed here~\parencite{oconnor_evolution_2014, franke_vagueness_2017} suggest.
-We advocate moving towards an ecological approach, studying vagueness in more heterogenenous ecosystems.
+We advocate moving towards an ecological approach, studying vagueness in more heterogeneous ecosystems.
 Language is part of a very complex system~\parencite{beckner_language_2009} that involves many components that interact in often unpredictable ways.

### 2018-08-31 book_chapter/paper.tex
minor edits book chapter

@@
 
-\usepackage[backend=biber,natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
+\usepackage[backend=bibtex,natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
@@
 
-\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
+% \DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
 
@@
 In this paper, we focus on this problem in the context of signaling games.
-We provide an overview of a number of proposed reasons why and mechanisms how vagueness may come into the picture in formal models of rational and evolutionary signaling.
+We provide an overview of a number of proposed ways of how vagueness may come into the picture in formal models of rational and evolutionary signaling.
 Most of them argue that vague signal use is simply the best we can get, given certain factors.
@@
 $$
-Again, these calculation incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
+Again, these calculations incorporate the probabilities that the imitating agent might observe the behavior of an agent of another type $\alpha^\prime$, weighed by its relative proportion.
 %

### 2018-10-25 book_chapter/paper.tex
Addressed proofreader's suggestions

@@
 
-\usepackage[backend=bibtex,natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
+\usepackage[backend=biber,natbib=true,style=authoryear-comp,doi=false,url=false,isbn=false]{biblatex}
 \bibliography{paper}
@@
 \title{Towards an ecology of vagueness}
-\author{Jos\'e Pedro Correia \and Michael Franke}
+\author{Jos\'e Pedro Correia\footnote{Holder of a PhD studentship by the \emph{Fundação para a Ciência e a Tecnologia} (SFRH/BD/100437/2014).}
+\and Michael Franke}
 \date{}
@@
 \begin{abstract}
-A vexing puzzle about vagueness, rationality and evolution runs, in crude abbreviation, as follows: vague language use is demonstrably suboptimal if the goal is efficient precise and cooperative information transmission; hence rational deliberation or evolutionary selection should, under this assumed goal, eradicate vagueness from language use.
+A vexing puzzle about vagueness, rationality, and evolution runs, in crude abbreviation, as follows: vague language use is demonstrably suboptimal if the goal is efficient, precise and cooperative information transmission; hence rational deliberation or evolutionary selection should, under this assumed goal, eradicate vagueness from language use.
 Since vagueness is pervasive and entrenched in all human languages, something has to give.
 In this paper, we focus on this problem in the context of signaling games.
-We provide an overview of a number of proposed ways of how vagueness may come into the picture in formal models of rational and evolutionary signaling.
-Most of them argue that vague signal use is simply the best we can get, given certain factors.
-Despite the plausibility of the proposals, we argue that a deeper understanding of the benefits of vagueness needs a more ecological perspective, namely one that goes beyond local optimization of signaling strategies in a homogeneous population.
+We provide an overview of a number of proposed ways in which vagueness may come into the picture in formal models of rational and evolutionary signaling.
+Most argue that vague signal use is simply the best we can get, given certain factors.
+Despite the plausibility of the proposals, we argue that a deeper understanding of the benefits of vagueness needs a more ecological perspective, namely one that goes beyond the local optimization of signaling strategies in a homogeneous population.
 As an example of one possible way to expand upon our current models, we propose two variants of
@@
 The classical philosophical problem of vagueness is most starkly embodied by the sorites paradox.
-The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, one can repeatedly remove all but one grain of sand of something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, one would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
+The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\parencite{sorensen_sorites_2009}, and uses the example of a heap of sand: if no removal of one grain of sand can make a heap into a non-heap, one can repeatedly remove all but one grain of sand from something that is clearly a heap and be forced to acknowledge that the remaining single grain of sand is still a heap; otherwise, it seems, one would have to accept that there is a determinate number of grains that forms a heap, and anything under it is not a heap.
 Neither choice is, however, intuitively satisfying.
@@
 It is easy to find more examples of predicates based on more finely grained properties---as `tall' is intuitively based on height---for which constructing a sorites paradox would be easy.
-Mereological nihilists argue that instances of the sorites can be designed for any material object that can be decomposed into small enough parts.
+Mereological nihilists argue that instances of the sorites paradox can be designed for any material object that can be decomposed into small enough parts.
 If one subscribes to the scientific picture of matter as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\parencite{Unger1979}.
-Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic'', are vague when used by human beings.
+Bertrand Russell famously argued~\parencite*{russell_vagueness_1923} that all words, including ``the words of pure logic,'' are vague when used by human beings.
 
@@
 But one can think of language in different terms.
-A possibility is to think of signs as tools agents use to coordinate actions.
-An example of a way to formally study language along those lines is by using the game-theoretic framework of signaling games~\parencite{lewis_convention_1969}.
+One possibility is to think of signs as tools agents use to coordinate actions.
+An example of a way to formally study language along those lines is the game-theoretic framework of signaling games~\parencite{lewis_convention_1969}.
 Within such an approach, rationality can be seen as the ability to choose the use of signs that is optimal to achieve some form of coordination.
@@
 
-We here survey proposals for addressing vagueness in the context of signaling games, but before we start we want to establish some vocabulary to better frame the upcoming discussion.
-Rationality is an elusive notion which is frequently debated in areas like philosophy, psychology, and economics.
-Discussions around it usually touch on different aspects of the concept which we here want to try to clearly demarcate.
+Here we survey proposals for addressing vagueness in the context of signaling games, but before we start we want to establish some vocabulary to better frame the upcoming discussion.
+Rationality is an elusive notion that is frequently debated in areas like philosophy, psychology, and economics.
+Discussions around it usually touch on different aspects of the concept without always clearly demarcating them; this is what we want to do before going further.
 The aforementioned logical picture of language and meaning is focused on dichotomies like true and false, meaningful and meaningless, correct and incorrect.
@@
 The notion is linked to David Hume~\parencite*{hume_treatise_1738}, epitomized in the following assertion: ``Reason is, and ought to only be the slave of the passions.''
-Instrumental rationality is close to a notion of rational choice that is used in economics and game theory%
-\footnote{In fact, it has been argued~\parencite{vanderschraaf_informal_1998} that Hume's whole account of convention is very closely in line with modern game theory.}%
-: agents are rational if and only if they make decisions that maximize their expected utility.
-A more in-depth discussion of the opposition in the context of theoretical economics can be found, for example, in the work of Herbert A.~Simon~\parencite*[\emph{e.g.}][]{simon_rationality_1986}%
-\footnote{Simon uses the term substantive instead of instrumental rationality, but the characterization is basically the same~\parencite[210-212]{simon_rationality_1986}.}%
-.
+Instrumental rationality is close to a notion of rational choice that is used in economics and game theory:%
+\footnote{In fact, it has been argued~\parencite{vanderschraaf_informal_1998} that Hume's whole account of convention is very closely in line with modern game theory.}
+agents are rational if and only if they make decisions that maximize their expected utility.
+A more in-depth discussion of the opposition in the context of theoretical economics can be found, for example, in the work of Herbert A.~Simon~\parencite*[\emph{e.g.}][]{simon_rationality_1986}.%
+\footnote{Simon uses the term substantive instead of instrumental rationality, but the characterization is basically the same~\parencite[210-212]{simon_rationality_1986}.}
+
 %Concretely, for a (finite) set of acts $A$, a (finite) set of world states $T$ with prior probability $P \in \Delta(T)$ and a utility function $u \colon A \times T \rightarrow \mathbb{R}$, choice of action $a$ is rational only if the expected utility $\text{EU}(a) = \sum_{t\in T} P(t) \, u(a,t)$ of $a$ is at least as high as that of any other $a' \in A$.\todo{Brings in a particular model too soon}
@@
 In other types of models, however, a further aspect of rational choice needs to be considered, namely the ability of agents to make accurate predictions about how other agents behave.
-We can say that an agent is more or less \emph{strategic} depending on the extent to which he is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
+We can say that an agent is more or less \emph{strategic} depending on the extent to which she is able to anticipate the actions and beliefs of other agents, and to predict medium/long term gains from repeated play.
 %Strategic rationality is important for language use, because it is natural to assume that an agent's beliefs about how an expression is used normally or how it ought to be used properly will depend, at least in part, on beliefs about how others use said expression.\todo{too soon}
@@
 %We draw attention to two additional aspects of rationality which will be helpful in the following.
-Where the above definition of instrumental rationality is usually understood as requiring a single choice to maximize the expected utility in a single concrete decision situation, one might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2016:Smart-Transform}.
+While the above definition of instrumental rationality is usually understood as requiring a single choice to maximize the expected utility in a single concrete decision situation, one might also be interested in more general choice mechanisms~\parencite[\emph{e.g.}][]{ZollmanSmead2010:Plasticity-and-,HagenChater2012:Decision-Making,FawcettHamblin2013:Exposing-the-be,GaleazziFranke2017:Smart-Represent}.
 A choice mechanism is a general way of behaving for an agent involved in a variety of decision situations.
@@
 The vocabulary just introduced will, we hope, help inform the discussion that follows.
-We proceed by introducing the framework of signaling games in Section~\ref{sec:signaling-games}.
-In Section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a particular kind of signaling game.
+We proceed by introducing the framework of signaling games in section~\ref{sec:signaling-games}.
+In section~\ref{sec:sim-max-vagueness} we look into explanations of vagueness in a particular kind of signaling game.
 Section~\ref{sec:ecology-of-vagueness} tries to generalize the considerations of these proposals to argue for an approach to vagueness anchored in a more global notion of rationality.
-We propose and analyse the results of a novel multi-population model of imprecise imitation in Section~\ref{sec:multi-population-model}, and summarize our conclusions in Section~\ref{sec:conclusions}.
+We propose and analyse the results of a novel multi-population model of imprecise imitation in section~\ref{sec:multi-population-model}, and summarize our conclusions in section~\ref{sec:conclusions}.
 %Section~\ref{sec:referential-vagueness} generalizes our discussion to the natural case where agents play more than one game.
@@
 % In order to extend his notion of convention to \emph{linguistic} exchanges
-In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where agents' choices involve sending and receiving signals or messages\footnote{These terms will be used interchangeably.}.
+In order to support the idea that linguistic conventions can arise without any prior conventional activity, Lewis considers situations where agents' choices involve sending and receiving signals or messages.\footnote{These terms will be used interchangeably.}
 One can think of two players with different roles.
@@
 The expected utility $\text{EU}$ of a strategy can be calculated by using the utility function and aggregating payoffs for all pairings of states of affairs and actions, weighted by the probability of each state.
-Concretely, the expected utility of $\sigma$ given $\rho$ is $\text{EU}_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$ and the expected utility of $\rho$ given $\sigma$ is $\text{EU}_R(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
+Concretely, the expected utility of $\sigma$ given $\rho$ is $\text{EU}_S(\sigma \mid \rho) = \sum_{t \in T} P(t) \ U_S(t, \rho(\sigma(t)))$, and the expected utility of $\rho$ given $\sigma$ is $\text{EU}_R(\rho \mid \sigma) = \sum_{t \in T} P(t) \ U_R(t, \rho(\sigma(t)))$.
 As an example, consider a game with $T = \lbrace t_1, t_2 \rbrace$, $M = \lbrace m_1, m_2 \rbrace$, $A = \lbrace a_1, a_2 \rbrace$, $P(t_1) = P(t_2) = 0.5$ and the following utility matrix:
@@
 Consider sender strategy $\sigma = \lbrace t_1 \mapsto m_2, t_2 \mapsto m_1 \rbrace$ and receiver strategy $\rho = \lbrace m_1 \mapsto a_2, m_2 \mapsto a_1 \rbrace$.
-These would have an expected utility of $1$ for both sender and receiver, since when $t_1$ obtains with probability $0.5$ the sender will use $m_2$ and to this message the receiver will respond with $a_1$ which achieves a payoff of $1$, and when $t_2$ obtains with probability $0.5$ the sender will use $m_1$ and to this message the receiver will respond with $a_2$ which also achieves a payoff of $1$.
+These would have an expected utility of $1$ for both sender and receiver, since when $t_1$ obtains with probability $0.5$ the sender will use $m_2$, and to this message the receiver will respond with $a_1$, which achieves a payoff of $1$, and when $t_2$ obtains with probability $0.5$ the sender will use $m_1$, and to this message the receiver will respond with $a_2$, which also achieves a payoff of $1$.
 They also represent one of the two stable conventions in this game, the other being the pair of strategies $\sigma = \lbrace t_1 \mapsto m_1, t_2 \mapsto m_2 \rbrace$ and $\rho = \lbrace m_1 \mapsto a_1, m_2 \mapsto a_2 \rbrace$.
@@
 Agents are thus envisioned as omniscient and highly strategic.
-These requirements can seem excessive, even more so if we consider how simple signaling systems are when compared to human languages.
+These requirements can seem excessive, and even more so if we consider how simple signaling systems are when compared to human languages.
 The models were introduced to help explain how language could have gotten off the ground as a conventional system without any sort of prior agreement.
@@
 Furthermore, communication through simple message exchange is something that almost all animals do: monkeys use calls, birds use singing, bees use dances, ants use pheromone trails, and so on.
-A plausible account of the origin of language should first explain how signaling systems could get started, without requiring high standards of rationality of the agents involved.
+A plausible account of the origin of language should first explain how signaling systems could get started, without requiring high standards of rationality from the agents involved.
 
@@
 The principles can be captured in formal models that abstract away from details of single interactions and behavior of individual agents, for example in the replicator dynamics~\parencite{TaylorJonker1978:Evolutionary-St}.
-The only thing relevant to this equation are the relative proportions of strategies in a given population and the utility function.
+The only things relevant to this equation are the relative proportions of strategies in a given population and the utility function.
 Using it, one can compute which strategies evolve under which conditions.
 
-Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, but it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
+Skyrms' evolutionary game theory approach to signaling games not only gives more plausible grounds to support Lewis' discussion of convention, it also accomplishes an important conceptual change: it moves most of the theory and mathematical formalism to the descriptive side of the investigation.
 Utility represents how the modeler views the signaling problem and understands the relative advantages or disadvantages of different possible strategy combinations.
 Dynamics describe how strategies can evolve when driven by mechanisms of utility maximization.
-The shift in perspective allows interpretations that accommodate for limited non-strategic agents.
+The shift in perspective allows interpretations that accommodate limited non-strategic agents.
 % Focus is put on understanding how various ingredients of the model interact, and which results they produce, not on metaphysical concerns.
 While the general framework manages to abstract quite some details away from the formalization, it nevertheless leaves room for them, especially when it comes to the dynamics.
-We already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
+We have already mentioned the replicator equation that can be seen as representing biological or cultural evolution, but one can also use dynamics inspired by learning mechanisms~\parencite[\emph{e.g.}][]{roth_learning_1995}, or even ones assuming a high degree of knowledge of the game and other players~\parencite[\emph{e.g.}][]{gilboa_social_1991,Muhlenbernd2011:Learning-with-N,SpikeStadler2016:Minimal-Require}.
 This range of options goes hand in hand with a range of pictures of rationality, from nothing more than survival of the fittest in a biologically-inspired setting, to a certain degree of instrumental but limited and non-strategic rationality in the case of learning dynamics, to higher levels of rationality and even recursive strategic reasoning about the co-players' beliefs and choices.
@@
 The characterization of signaling problems in terms of evolutionary game theory allows us to explain why certain equilibria come to be and how.
-A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under standard evolutionary pressures, and where it comes back to if slightly disturbed.
+A core notion in this context is that of an \emph{evolutionary stable state}~\parencite{maynard_smith_evolution_1982}: an equilibrium situation that a population tends to under standard evolutionary pressures, and to which it returns if slightly disturbed.
 With these tools, one can better understand why signaling systems are stable even without any strong assumptions of rationality.
@@
 Skyrms~\parencite*{skyrms_signals_2010} gives an overview of different topics studied using signaling games, including expansions of the framework itself (for example, considering other dynamics beyond the replicator equation), exploration of other factors that impact the evolution of signaling (for example, how agents are interconnected), or variations on the signaling problem and its basic assumptions (for example, loosening the alignment of interests in order to provide accounts of deceptive signal use).
-Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, just to name a few.
+Other uses of signaling games include discussions of categorization~\parencite[\emph{e.g.}][]{jager_language_2007}, compositionality~\parencite[\emph{e.g.}][]{barrett_evolution_2009}, incommensurability~\parencite[\emph{e.g.}][]{barrett_faithful_2010}, to name a few.
 More recent overviews are given by \textcites{huttegger_how_2014}, \textcite{huttegger_dynamics_2014}, and \textcite{FrankeWagner2014:Game-Theory-and}.
@@
 The sorites paradox requires us to assume a relation between the vague terms and a more precise underlying dimension (height for tallness, number of hairs for baldness, number of grains of sand for ``heapness'', and so on).
-Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure, in the sense that there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), but usually even a degree of how far apart these elements are from each other.
-In terms of signaling games, we can model this by using a state space constituted of values of the underlying dimension, and a message space constituted by the terms in question.
+Not only does this property need to be much more fine-grained than the vague term, but it also needs to have some structure: there is at least an order between the elements in it (thinking of height in centimeters, $180 > 179 > \ldots > 120$), and usually even a degree of how far apart these elements are from each other.
+In terms of signaling games, one can model this using a state space constituted of values of the underlying dimension, and a message space constituted by the terms in question.
 Because of the difference in granularity, we will typically be interested in cases where the state space is much larger than the message space.
@@
 This tolerance decreases with distance in state space: assuming a 180cm person is tall, one would easily tolerate the use of the term for a person measuring 179cm, less so for someone who is 170cm, and much less so for 160cm.
-This can be modeled by using a utility function that is continuous rather than discrete and that is monotonously decreasing with distance, \emph{i.e.}~success is not a matter of black and white, right or wrong, but a matter of degree, of how close the receiver got to the optimal response to the sender's perceived state.
+This can be modeled using a utility function that is continuous rather than discrete and that monotonously decreases with distance, \emph{i.e.}~success is not a matter of black and white, right or wrong, but a matter of degree, of how close the receiver got to the optimal response to the sender's perceived state.
 
@@
 What these authors find about this setup is that the evolutionary stable states are what they call Voronoi languages.
-Roughly, these are situations where the sender uses messages in a way that can be seen as partitioning the state space into convex regions and the receiver responds with the central element of those regions.
+Roughly, these are situations where the sender uses messages in a way that can be seen as partitioning the state space into convex regions, and the receiver responds with the central element of those regions.
 %For example, imagine a state space consisting of values of height in centimeters ranging from 40 to 280.%
@@
 The first proposal is called limited memory fictitious play (LMF) and models agents playing a sim-max game where their ability to recall past interactions with others is limited to a certain number.
-For a given interaction, each agent uses his limited memory of the other agent's past behavior to estimate the other's strategy, and plays an instrumentally rational best response to that strategy.
+For a given interaction, each agent uses her limited memory of the other agent's past behavior to estimate the other's strategy, and plays an instrumentally rational best response to that strategy.
 In order to study the evolution of strategies in repeated interaction, the authors model several individual agents in actual play.
@@
 \citeauthor{franke_vagueness_2011} find that for low values of this parameter, only babbling equilibria are possible, where sender and receiver simply randomize, respectively, message and interpretation choice uniformly.
-Above a certain value of the parameter, other equilibria of the kind described in the beginning of this section arise, where agents communicate successfully, though not perfectly, using fuzzy strategies similar to the ones depicted in Figure~\ref{fig:example_stratsB}.
+Above a certain value of the parameter, other equilibria of the kind described in the beginning of this section arise, where agents communicate successfully, though not perfectly, using fuzzy strategies similar to those depicted in Figure~\ref{fig:example_stratsB}.
 However, it is not clear whether soft-max choices capture the right stochastic trembles in decision making as they would arise under natural sources of uncertainty about the context \parencite[see][]{franke_vagueness_2017}.
@@
 O'Connor's proposal is to model generalization as the propagation of reinforcement to nearby states, where ``nearby'' is defined in terms of distance in state space.
-For example, if a sender was successful in using message $m$ for state $t$, he will not only positively reinforce that choice of message for $t$, but also for states similar to $t$.
+For example, if a sender was successful in using message $m$ for state $t$, she will not only positively reinforce that choice of message for $t$, but also for states similar to $t$.
 This is done to a degree that is proportional to the similarity between $t$ and other states.
@@
 Although there is a close relationship between reinforcement learning and population-level dynamics \citep{Borgers1997,Beggs2005}, O'Connor's GRL is, on the face of it, an account of learning between individual players.
-Also, we need to ask for further justification to link generalization in reinforcement closely to the underlying payoff function of the sim-max game.
+Also, we need further justification for linking generalization in reinforcement so closely to the underlying payoff function of the sim-max game.
 Why should agents evolve to generalize in exactly the right way?
-O'Connor suggests that, despite a vague language having lower expected utility than a precise one, the learning mechanism that induces vagueness does have evolutionary advantages: it achieves higher payoffs in a shorter period of time.
+O'Connor suggests that, despite a language with vagueness having lower expected utility than a language without it, the learning mechanism that induces vagueness does have evolutionary advantages: it achieves higher payoffs in a shorter period of time.
 From a global point of view, learning speed can be an advantage.
@@
 A similar finding is made by \textcite{franke_vagueness_2017} when studying a variant of the replicator dynamics in which individual agents do not have the ability to generalize but simply make perceptual mistakes.
-In a scenario where agents learn by imitation, if one assumes that they do not have perfect perception, there will always be a possibility that senders confuse states and thus learn associations with messages that are different than the ones observed, and receivers similarly mix up responses.
+In a scenario where agents learn by imitation, if one assumes that they do not have perfect perception, there will always be the possibility for senders to confuse states and thus learn associations with messages that are different than the ones observed, and for receivers to similarly mix up responses.
 Furthermore, it seems reasonable to assume that this confusability is proportional to state similarity, \emph{i.e.}~that the more similar two states are, the more likely it is that they will be mistaken for each other.
-Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, they develop a variant of the dynamic that also induces vague signal use of the kind we expect here.
+Incorporating these considerations into a derivation of the replicator dynamics based on imitation processes, Franke and Correia develop a variant of the dynamic that also induces vague signal use of the kind we expect here.
 The consequence is very similar to that of the GRL model discussed above, in that the way the behavior for a given state is updated takes into account the behavior of similar states, proportionally to their similarity.
@@
 %In this picture, vagueness is the natural and unavoidable consequence of perceptual limitations, but may still contribute to the overall boundedly rational character of language if we take multiple contexts of learning into account.
-What is left to show, like for O'Connor's GRL approach, is whether the potential payoff advantages which were observed in simulations actually suffice to promote vague language use in an encompassing model of multi-level selection.
-The following section motivates the need for taking a more ecological approach to the evolution of vagueness, before Section~\ref{sec:multi-population-model} gives a concrete model.
+What is left to show, as it was with O'Connor's GRL approach, is whether the potential payoff advantages that were observed in simulations actually suffice to promote vague language use in an encompassing model of multi-level selection.
+The following section motivates the need for taking a more ecological approach to the evolution of vagueness, before section~\ref{sec:multi-population-model} gives a concrete model.
 
@@
 
-Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that vague signal use is suboptimal when compared to a crisp one, work under a number of various assumptions.
+Arguments of the kind presented by Lipman~\parencite*{lipman_why_2009}, that vague signal use is suboptimal when compared to crisp use, work under various assumptions.
 Part of the picture formed by these assumptions is a highly idealized conception of the agents involved and of the context in which they develop and use signals.
@@
 \item Language learning through experience has to rely on a limited number of observations.
-Not only is the state space typically much larger than what one can survey in sufficient time, it is even potentially infinite and constantly changing;
+Not only is the state space typically much larger than one can survey in sufficient time, it is even potentially infinite and constantly changing;
 \item A corollary of the former is that there will always be heterogeneity in a population of language learners, at the very least in their prior experience, since each agent will have relied on a different set of observations.
@@
 % Another possible interpretation is that of a population of agents learning by imitation.
-GRL~\parencite{oconnor_evolution_2014} and the work of~\textcite{franke_vagueness_2017} both assume a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
+GRL~\parencite{oconnor_evolution_2014} and the work of~\textcite{franke_vagueness_2017} both assume a picture of agents with a basic level of instrumental rationality, possibly limited awareness of the game, and a lack of strategic capabilities, adapting their behavior with only short-term gains in sight.
 %Under the standard assumptions, even agents with such limited rationality would evolve crisp signaling in sim-max games.
@@
 One can see each signaling game as embodying a particular language game.
-In Wittgenstein's picture of language, however, we do not play only one language game in our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might take out of it by choosing a certain behavior over another.
+In Wittgenstein's picture of language, however, we do not play only one language game over the course of our existence; there is a plurality of them and which one an agent is engaged in at a particular moment is never clearly identified, neither are the exact benefits one might gain by choosing a certain behavior over another.
 These are furthermore not fixed in time; old language games fall out of fashion or stop being useful, and new ones emerge all the time~\parencite[see][and in particular \S 23]{wittgenstein_philosophical_1953}.
@@
 We can look for rationality at several levels in this pluralistic picture.
-Firstly, as before, there is the actual behavior of a single agent in each actualized language game.
+First, as before, there is the actual behavior of a single agent in each actualized language game.
 As mentioned above in connection the soft-max choice function used by \textcite{franke_vagueness_2011}, behavior that strictly maximizes expected utility under uncertainty may be resource heavy, so it might be compatible with local strategic rationality that agents' production choices are stochastic.
 %% \mf{mention ``bounded rationality'' here}
-Secondly, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
+Second, if we look at behavior across many game types and contexts, there is also the level of an agent's internal theory of how words and phrases are likely to be used (or even normatively: how messages should be used), conditional on a given context.
 Notice that a single agent's rational beliefs about linguistic practices or linguistic meaning may well have to reflect the actual stochasticity: under natural assumptions about information loss, the best belief for prediction matches the actual distribution in the real world \parencite[\emph{e.g.}][]{VehtariOjanen2012:A-survey-of-Bay}.
@@
 %Is it rational for a sender to use a message in a crisp way, splitting the state space as if a strict threshold existed between the use of one term and the other?
-%In a scenario where the agent's ability to make the optimal choice is bounded, such as in the models of~\textcite{franke_vagueness_2011}, the agent is again doing the best he can given the constraints imposed on him.
+%In a scenario where the agent's ability to make the optimal choice is bounded, such as in the models of~\textcite{franke_vagueness_2011}, the agent is again doing the best she can given the constraints imposed on him.
 %\textcite{lawry_vagueness_2017} draw attention to another option we need to consider: such agents could opt to ignore their limitations and external sources of uncertainty and use a crisp strategy regardless.
@@
 \label{sec:multi-population-model}
-We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics the support for multiple populations with different imprecision values evolving together.
+We build upon the model of~\textcite{franke_vagueness_2017}, adding to the imprecise imitation dynamics support for multiple populations with different imprecision values evolving together.
 The resulting evolutionary dynamic has two layers: first, signaling strategies change by imprecise imitation following the model of~\textcite{franke_vagueness_2017}; second, there is evolutionary selection at the level of the imprecision values themselves.
-More concretely, imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop are more or less likely to survive and reproduce.
+More concretely, imagine agents who are born with varying perceptual abilities, subsequently learn a signaling strategy by imitation of other agents (either within or across populations), and depending on the success of the strategies they develop, are more or less likely to survive and reproduce.
 In such a setup, evolutionary processes will not only change the distribution of strategies within a population, where a population is identified by a shared level of imprecision in imitation, they will also promote those levels of imprecision which result in the development of more successful signal use.
@@
 %There are thus two aspects to be considered: for kin selection that would be the direct and indirect effects, and for multi-level selection the within-group and between-group components~\parencite[28--29]{birch_kin_2014}
-In our model there is also a similar structure: there is a process of selection of behavior within each population, and levels of imprecision are selected across populations based on the strategies they give rise to.
-We believe there is an important difference with group and kin selection models in that the two selection processes in our model act on different entities, the inner shaping signaling behavior, and the outer selecting levels of imprecision.
+In our model there is a similar structure: there is a process of selection of behavior within each population, and levels of imprecision are selected across populations based on the strategies they give rise to.
+We believe there is an important difference with group and kin selection models in that the two selection processes in our model act on different entities: the inner shaping signaling behavior, and the outer selecting levels of imprecision.
 In any case, we intend our model to be descriptive rather than causal or explanatory.
-That means that we do not want to commit to seeing populations either as kins or as social groups, and use them as merely descriptive abstractions that allow us to capture the hypothetical impact of indirect selection processes.%
-\footnote{We make this note because of the heated debate between the two theories. See~\textcite{kohn_darwin_2008} and~\textcite{kramer_kin_2016} for more details on that.}%
+That means that we do not want to commit to seeing populations either as kin or as social groups, and use them as merely descriptive abstractions that allow us to capture the hypothetical impact of indirect selection processes.%
+\footnote{We make this note because of the heated debate between the two theories. See~\textcite{kohn_darwin_2008}, and~\textcite{kramer_kin_2016} for more details.}
 
@@
 The probability that a given agent with imprecision $\alpha$ (or of type $\alpha$) observes $t_o$ when the actual state is $t_a$ is given by $P_{o}^{\alpha}(t_{o}|t_{a})$.
-If the same agent intends to realize $t_i$, the probability that he actually realizes $t_r$ instead is given by $P_{r}^{\alpha}(t_{r}|t_{i})$.
+If the same agent intends to realize $t_i$, the probability that she actually realizes $t_r$ instead is given by $P_{r}^{\alpha}(t_{r}|t_{i})$.
 Following~\textcite[26]{franke_vagueness_2017}, we can then define the following values.
+
 Probability that $t_a$ is actual if $t_o$ is observed by an agent of type $\alpha$:
@@
 $$
+
 Probability that a random sender of type $\alpha$ produces $m$ when the actual state is $t_a$:
@@
 $$
+
 Probability that the actual state is $t_a$ if a random sender of type $\alpha$ produced $m$:
@@
 $$
+
 Probability that $t_r$ is realized by a random receiver of type $\alpha$ in response to message $m$:
@@
 $$
+
 These formulations are merely parameterized versions of the single-population model.
@@
 
-Another important value to calculate has to do with the types which agents observe and imitate.
+Another important value to calculate has to do with the types that agents observe and imitate.
 In a model with tight interaction, we imagine this occurring across populations.
@@
 A population of type $\alpha$ consists of agents employing both a sender and a receiver strategy, thus the overall fitness of the population must include the expected utilities of both.
-We could imagine this process happening at a different speed than the cultural process, in which case we could have different time scales.
+We could imagine this process happening at a different speed to the cultural process, in which case we could have different time scales.
 For the sake of simplicity, we choose to have them both happen at each time step, but calculate the changes occurring at this level of selection after the calculation for the other level.
@@
 %\end{sidewaysfigure}
-In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-populations scenario for all trials.
+In Figure~\ref{fig:proportion-cases-tight-interaction} we plot the evolution of the proportion of each population in the two-population and three-population scenarios for all trials.
 As the plot shows, the population with no imprecision steadily increased its proportion against the others in every trial.
-In the three population scenario, the proportion of type $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
+In the three-population scenario, the proportion of type $\alpha = 0.05$ sees a slight increase in the beginning of the simulation, but inexorably starts a downward trend.
 These observations go against the expectation of~\textcite{franke_vagueness_2017} that faster convergence to a convex strategy by populations with a certain level of imprecision could give them a temporary advantage to take over and eliminate other types.
@@
 What we see is more trials reaching convexity earlier for the multi-population scenarios when compared with the single-population scenario.
-This effect precludes the hypothesized temporary advantage of imprecision to even manifest itself, but it can be seen as a positive influence of the population(s) with imprecision on the population with no imprecision.
-
-The flip side of this tight connection between populations is that strategies evolved by populations with no imprecision are also more vague (in the sense defined in Section~\ref{sec:sim-max-vagueness}).
+This effect precludes the hypothesized temporary advantage of imprecision manifesting itself, but it can be seen as a positive influence of the population(s) with imprecision on the population with no imprecision.
+
+The flip side of this tight connection between populations is that strategies evolved by populations with no imprecision are also more vague (in the sense defined in section~\ref{sec:sim-max-vagueness}).
 This is visible by looking at mean entropy values, namely sender strategy entropy, plotted in Figure~\ref{fig:entropy-sender-tight-interaction}, and receiver strategy entropy, plotted in Figure~\ref{fig:entropy-receiver-tight-interaction}.
@@
 $$
-The main difference with the previous model is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type ($\sigma^\alpha$ and $\rho^\alpha$).
+The main difference from the previous model is that these expected utilities are not calculated against all populations ($\sigma^\star$ and $\rho^\star$) but only against the agent's own type ($\sigma^\alpha$ and $\rho^\alpha$).
 This also implies that population proportions do not play a role at this level of selection.
@@
 Whereas in the latter no given population ever reached much more than 70\% proportion, in this model we see that many trials resulted in one population fully dominating the others.
-In the two-population scenario, some population reached at least 99\% proportion in 24 out of 25 trials.
+In the two-population scenario, some population reached at least 99\% in 24 out of 25 trials.
 In the three-population scenario, this happened in 18 out of 25 trials.
@@
 More interestingly for our investigation, some trials actually resulted in the population with $\alpha = 0$ being dominated.
-For the two-population scenario, $\alpha = 0.05$ fully reached 100\% proportion in 8 trials (79, 86, 88, 89, 91, 93, 96, 98), a point from which it is technically impossible for the other population to recover.
+For the two-population scenario, $\alpha = 0.05$ fully reached 100\% in 8 trials (79, 86, 88, 89, 91, 93, 96, 98), a point from which it is technically impossible for the other population to recover.
 In most cases the dominating population gains its ground from the start, but in 5 cases we see a temporary advantage of $\alpha = 0.05$ that is then lost to the other population.
 In one interesting trial (85), $\alpha = 0.05$ reached 99.87\% only to then steadily start losing ground to $\alpha = 0$.
-For the three population scenario, $\alpha = 0$ was reduced to 0\% proportion in 4 trials (152, 157, 159, and 175).
+For the three-population scenario, $\alpha = 0$ was reduced to 0\% in 4 trials (152, 157, 159, and 175).
 In all of those cases, $\alpha = 0.05$ clearly has the upper hand over $\alpha = 0.1$, despite a temporary advantage of the latter in some trials.
@@
 This has certainly to do with the initial conditions of each trial, since the randomly generated strategies can simply by chance be more favourable to reaching convexity.
-More importantly, we also see that reaching convexity sooner is not a sufficient condition for achieving population dominance .
+More importantly, we also see that reaching convexity sooner is not a sufficient condition for achieving population dominance.
 There were many trials where another type reached convexity much sooner than $\alpha = 0$ but its population was completely dominated nevertheless (\emph{e.g.}~76, 84, 92, and 99 in the two-population scenario; 156, 165, 168, and 172 in the three-population scenario).
@@
 In a sim-max game such as the one set up here, there are only two stable equilibria.
-These are the two Voronoi languages of the kind shown in Figure~\ref{fig:example_strats}, one where the first message is used for the first half of the state space, and the other where the second message is used for this region.
+These are the two Voronoi languages of the kind shown in Figure~\ref{fig:example_strats}: one where the first message is used for the first half of the state space, and the other where the second message is used for this region.
 In the multi-population model with loose interaction, each population evolves independently towards one of these two equilibria.
@@
 In most trials where the type was eliminated, in both the two- and three-population scenarios, $\alpha = 0$ ended up evolving towards a different equilibrium than the other types, and either had an initial disadvantage in expected utility, or reached convexity much later.
-These factors can thus be seen as indicators at best, but the story of how a population with imprecision ends up dominating one without seems to be more complicated to tell.
-
-This is not surprising since we are facing a complex dynamical system with various interacting components (sender and receiver strategies and multiple populations).
+These factors can thus be seen as indicators at best, and the story of how a population with imprecision ends up dominating one without it seems to be more complicated than expected.
+
+This is not surprising, since we are facing a complex dynamical system with various interacting components (sender and receiver strategies and multiple populations).
 \begin{figure}
@@
 The curves up to where $\alpha = 0.05$ achieved convexity seem to suggest that the sender strategy of that type was evolving towards the same equilibrium as the receiver strategy of the other.
-The moment where the population reaches convexity marks the point where the sender strategy of $\alpha = 0.05$ aligns with its type's receiver strategy and this coincides with the moment where the population seems to gain real traction over the other type (see again the plot for trial 88 in Figure~\ref{fig:convex-cases-loose-interaction}).
+The moment where the population reaches convexity marks the point where the sender strategy of $\alpha = 0.05$ aligns with its type's receiver strategy and this coincides with the moment when the population seems to gain real traction over the other type (see again the plot for trial 88 in Figure~\ref{fig:convex-cases-loose-interaction}).
 In this particular case, reaching convexity seems to have made an important difference.
@@
 We advocate moving towards an ecological approach, studying vagueness in more heterogeneous ecosystems.
-Language is part of a very complex system~\parencite{beckner_language_2009} that involves many components that interact in often unpredictable ways.
-An ecology of vagueness would involve studying models where different populations can evolve and interact with each other, where different language games can be played between the individuals, where the environment is uncertain and changing, or anything else that more closely approximates the real context of language evolution.
-
-In light of this picture, we proposed two variants of a concrete multi-population signaling model to test the hypothesis~\parencite{franke_vagueness_2017} that certain features of imprecise imitation, like promoting faster convergence and regularity, could prove beneficial in contrast with full precision.
+Language is part of a very complex system~\parencite{beckner_language_2009} that involves many components interacting in often unpredictable ways.
+An ecology of vagueness would involve studying models where different populations can evolve and interact with each other, where different language games can be played between individuals, where the environment is uncertain and changing, or anything else that more closely approximates the real context of language evolution.
+
+In light of this picture, we have proposed two variants of a concrete multi-population signaling model to test the hypothesis~\parencite{franke_vagueness_2017} that certain features of imprecise imitation, like promoting faster convergence and regularity, could prove beneficial in contrast with full precision.
 Analysing these models did not provide us with a clear-cut answer, and revealed that the story is much more nuanced than initially expected.
